### 2010-09-18 doc/jInfer/Architecture.tex
Documentation: Architecture: work started.

\documentclass[10pt,a4paper,notitlepage]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\begin{document}

\title{jInfer Architecture} 
\author{Matej Vitásek} 
\maketitle

The description of jInfer architecture will commence by describing the data structures, namely representations of regular expressions and XML elements, attributes and simple data.\\

Afterwards the interfaces of basic inference modules - Initial Grammar Generator, Simplifier and Schema Generator - will be explained.\\

Finally, the process of inference will be described.

\subsection{Package naming conventions}

All packages start with \texttt{cz.cuni.mff.ksi.jinfer}. Afterwards is the short, normalized name of the module (e.g. \texttt{base}) and finally the package structure in this module (e.g. \texttt{objects.utils}). All in all, a package in the Base module could look like 
\begin{verbatim}
cz.cuni.mff.ksi.jinfer.base.objects.utils
\end{verbatim}

\section{Data structures}

\subsection{Regular expressions}

For general information on regular expressions, please refer for example to Wikipedia: $http://en.wikipedia.org/wiki/Regular\_expression$ .\\

All classes pertaining to regular expressions can be found in the package \texttt{cz.cuni.mff.ksi.jinfer.base.regexp}.\\

In jInfer, the following types of regular expressions are represented:
\begin{itemize}
	\item Token - a letter of the language described by this regular expression. In this case, an \texttt{AbstractNode} (see \ref{xml-rep}). Eg. \textit{"a"}.
	\item Concatenation - one or more regular expression in an ordered sequence. Eg. \textit{"(a, b, c, d)"}.
	\item Alternation - a choice between one or more regular expressions. Eg. \textit{"(a $|$ b $|$ c $|$ d)"}.
	\item Kleene star ("*") - a regular expression that can occur any number times (zero to infinity) in a row. Eg. \textit{"a*"}.
\end{itemize}

These types are defined in the \texttt{RegexpType} enum as \texttt{TOKEN}, \texttt{CONCATENATION}, \texttt{ALTERNATION} and \texttt{KLEENE}.\\

The class representing a regular expression is \texttt{Regexp}. A \texttt{Regexp} is a generic class (let's denominate its type variable as \textit{T}) that has three main members:
\begin{itemize}
	\item \texttt{type} - a \texttt{RegexpType} instance, set in the constructor and immutable afterwards. Can be queried via the method \texttt{getType()} or shorthand boolean methods
	\begin{itemize}
		\item \texttt{isToken()}
		\item \texttt{isConcatenation()}
		\item \texttt{isAlternation()}
		\item \texttt{isKleene()}
	\end{itemize}

	\item \texttt{content} - a type \textit{T} member. Must be set iff the \texttt{type} is \texttt{TOKEN}, \texttt{null} otherwise. Means "the token represented by this regular expression". % TODO how to query

	\item \texttt{children} - a list of \texttt{Regexp}s of the type \textit{T}. Must be set iff the \texttt{type} is \textbf{not} \texttt{TOKEN}, \texttt{null} otherwise. If the \texttt{type} is \texttt{KLEENE}, it must contain exactly one element, although this rule is not enforced by the code. Otherwise the number of elements in this list is unconstrained. The meaning of this member is "the list of regular sub-expressions in this concatenation / alternation / Kleene star". % TODO how to query
\end{itemize}

\textit{Example:} regular expression \textit{"(a, b, (c $|$ d))*"} would be represented as a \texttt{Regexp} of type \texttt{KLEENE} with children of size 1, containing a \texttt{CONCATENATION} with 3 children. Of these three, the first two would be \texttt{TOKEN}s with content \textit{"a"} and \textit{"b"} respectively. The last child would be an \texttt{ALTERNATION} containing two \texttt{TOKEN}s \textit{"c"} and \textit{"d"}. It is obvious that this representation of regular expressions creates tree structures.

\subsection{XML representation}
\label{xml-rep}

The classes that represent XML elements, attributes and simple text data are summarized in figure \ref{img-arch-nodes}.

\begin{figure}
	\includegraphics[scale=0.3]{Architecture-images/nodes.eps} 
	\caption{UML class diagram of classes representing parts of XML}
	\label{img-arch-nodes}
\end{figure}

It is important to note that attributes in jInfer's data structures are nodes, just like elements and simple data.\\

Description of all classes from the schema follows.

\subsubsection{AbstractNode}

\subsubsection{Element}

\subsubsection{AbstractContentNode}

\subsubsection{Attribute}

\subsubsection{SimpleData}

\section{Interfaces}

\section{Inference process}

\end{document}

### 2010-09-18 doc/jInfer/Architecture.tex
Documentation: Architecture: work in progress.

@@
 
-The classes that represent XML elements, attributes and simple text data are summarized in figure \ref{img-arch-nodes}.
+The classes that represent XML elements, attributes and simple text data (collectively known as \textit{XML nodes}) are summarized in figure \ref{img-arch-nodes}.
 
@@
 	\includegraphics[scale=0.3]{Architecture-images/nodes.eps} 
-	\caption{UML class diagram of classes representing parts of XML}
+	\caption{UML class diagram of classes representing XML nodes}
 	\label{img-arch-nodes}
@@
 
-It is important to note that attributes in jInfer's data structures are nodes, just like elements and simple data.\\
+It is important to note that attributes in jInfer's data structures are nodes too, just like elements and simple data.\\
 
-Description of all classes from the schema follows.
+Description of all classes from the schema follows. All classes can be found in \texttt{cz.cuni.mff.ksi.jinfer.base.objects} package.
 
-\subsubsection{AbstractNode}
+\subsubsection{\texttt{NodeType}}
 
-\subsubsection{Element}
+NodeType is an enum of all non-abstract node types, i.e. ATTRIBUTE, ELEMENT, SIMPLE\_DATA.
 
-\subsubsection{AbstractContentNode}
+\subsubsection{\texttt{AbstractNode}}
 
-\subsubsection{Attribute}
+A common ancestor of all XML nodes. Contains the following members:
+\begin{itemize}
+	\item \texttt{context} - a list of \texttt{String}s, element names, marking the path from this document's root element to the current node. For example, in a document like this
+\begin{verbatim}
+<people>
+<person>
+<name first="John" last="Doe">
+</person>
+</people>
+\end{verbatim} % TODO tabs
+the element \texttt{$<$name$>$} would have context \texttt{\{"people", "person"\}} while the element \texttt{$<$people$>$} would have an empty list for context.
+	\item \texttt{name} - a String representing this node's name.
+For \texttt{ELEMENT}s and \texttt{ATTRIBUTES}s this is their name.
+For \texttt{SIMPLE\_DATA}, this is the string representation of the textual data in this node.
+	\item \texttt{metadata} - a map indexed by \texttt{String}s, containing general \texttt{Object}s containing various metadata related to this node.
+At this moment, three types of metadata are recognized:
+	\begin{itemize}
+		\item required - used in \texttt{ATTRIBUTE}s.
+If there is a \texttt{Boolean.TRUE} value under this key in metadata, it means that the attribute represented by this node is considered to be required (for example, in DTD this would be marked as \texttt{\#REQUIRED}).
+		\item \texttt{from.schema} - a \texttt{Boolean.TRUE} value under this key means that this node was obtained while processing an XML schema.
+		\item \texttt{from.query} - a \texttt{Boolean.TRUE} value under this key means that this node was obtained while processing a query.
+	\end{itemize}
+	jInfer module are free to ignore or process these metadata.
+\end{itemize}
+Method \texttt{getType()} returning this node's \texttt{NodeType} is declared as abstract and overriden in corresponding classes.
 
-\subsubsection{SimpleData}
+\subsubsection{\texttt{Element}}
+
+Represents an actual XML element.
+Apart from the members mentioned before, a new member \texttt{subnodes} is introduced.
+\texttt{subnodes} is of type \texttt{Regexp$<$AbstractNodes$>$}.
+This regular expression describes the allowed content of this node (i.e.
+the nodes that can be within, hence \textit{subnodes}).
+\texttt{subnodes} may be \texttt{null}.
+
+\subsubsection{\texttt{AbstractContentNode}}
+
+A common ancestor for nodes that have some (usually textual) content.
+This class is generic, however, this document assumes it to be parametrized as \texttt{AbstractContentNode$<$String$>$} for the sake of simplicity.
+Contains the following members:
+\begin{itemize}
+	\item \texttt{contentType} - a \texttt{String} describing the type of the content in this node. Note that the parameter is currently ignored and will probably be deleted or changed.
+	\item \texttt{content} - a list of \texttt{String}s holding the actual content of this node.
+Note that a single node has only one string as its content, however, as the nodes are routinely aggregated, this list might hold all the contents of the original aggregated nodes.
+\end{itemize}
+
+\subsubsection{\texttt{Attribute}}
+
+This is just a marker implementation of an \texttt{AbstractContentNode$<$String$>$}. Represents an XML attribute.
+
+\subsubsection{\texttt{SimpleData}}
+
+This is just a marker implementation of an \texttt{AbstractContentNode$<$String$>$}. Represents XML simple text data.
 

### 2010-10-26 doc/jInfer/full/main.tex
documentation framework baby

\documentclass[a4paper,10pt,oneside,twocolumn]{article}
\usepackage{epsfig}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{import}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{26cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer - Framework for Automatic XML Schema Inference}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle

\input{00abstract}

\input{01introduction}

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-09 doc/jInfer/full/architecture.tex
some docs

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{epsfig}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{import}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{26cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer Architecture}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
Target audience: developers willing to extend jInfer.\\

\textit{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\

The description of jInfer architecture will commence by describing the data
structures, namely representations of regular expressions and XML elements,
attributes and simple data.

Afterwards the interfaces of basic inference modules - Initial Grammar Gen-
erator, Simplifier and Schema Generator - will be explained.

Finally, the process of inference will be described.

\section{Package naming conventions}
All packages start with \code{cz.cuni.mff.ksi.jinfer}. Afterwards is the short, normalized name of the module (e.g. \code{base}) and finally the package structure in this module (e.g. \texttt{objects.utils}). All in all, a package in the Base module could look like 
\code{cz.cuni.mff.ksi.jinfer.base.objects.utils}

\section{Data structures}
\subsection{Regular expressions}

For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{Hopcroft:2006:IAT:1177300}.

All classes pertaining to regular expressions can be found in the package \code{cz.cuni.mff.ksi.jinfer.base.regexp}.

In theory, regular expression is defined recursively:
\begin{itemize}
	\item Every symbol $x$ from alphabet is regular expression representing language ${x}$.
	\item If $R_1, R_2, \dots, R_n$ are regular expressions representing languages $L_1, L_2, \dots, L_n$, then $R_1 | R_2 | \dots | R_n$
	is regular expression representing language $\bigcup_{i = 1}^{n}{L_i}$. Operation $|$ is called alternation.
	\item If $R_1, R_2, \dots, R_n$ are regular expressions representing languages $L_1, L_2, \dots, L_n$, then $R_1, R_2 , \dots , R_n$
	is regular expression representing language $L_1 \cdot L_2 \cdot \ldots \cdot L_n$. Where $\cdot$ is languague concatenation,
	and operation $,$ is called concatenation as well.
	\item If $R$ is regular expression representing language $L$, then $R*$ 
	is regular expression representing language $L*$. In regular expressions, operator $*$ is called Kleene star.
\end{itemize}

In common practice, extended regular expressions are commonly used. Although they are not extending set of
languages representable by regular expressions, they give us nicer syntax. Extended regular expression adds syntax:
\begin{itemize}
	\item $R{m,n}$, where $m,n \in N$ with meaning of $R$ at least $m$-times, at most $n$-times,
	\item $R{m,}$ with meaning at least $m$-times,
	\item $R?$ with meaning $R{0,1}$,
	\item $R+$ with meaning $R{1,}$.
\end{itemize}

By this syntax, Kleene star is simply interval ${0,}$. This is just syntactic shortcut for long expresssions like $R, R, R, R?, R?$, which is just $R{3,5}$.

In jInfer, we implement regular expresssions as class \code{Regexp} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
Each \code{Regexp} instance has one of the \code{RegexpType} type:
\begin{itemize}
	\item Token - a letter of the language.
	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$.
	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$.
	\item Permutation - shortcut for all possible permutations of regular expressions (to work with \code{xs:all} in future).
	Our syntax to write down permutation is $(a  b c  d)$.
\end{itemize}

Each \code{Regexp} instance has interval associated with it (as a member). Class \code{RegexpInterval} represents all kinds of intervals
${m,n}; {m,}$.

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-09 doc/jInfer/full/main.tex
empty skulp for vitooo
some docs

@@
 
-\input{00abstract}
+\section*{Abstract}
 
-\input{01introduction}
+jInfer is a framework for XML schema inference based on NetBeans platform.
+Using jInfer, you can create DTD or XSD schema from your existing XML documents, schemas and queries.
+With our modules, you can fine-tune the inference process and see it unfolding.
+With your modules, you can do even more: export to exotic languages, try cutting-edge algorithms and more.
+
+\paragraph{keywords:} XML, schema
+
+\section*{Introduction}
+
+\section*{Tutorial}
 

### 2011-01-10 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - little folder #2

--- 2011-01-09 doc/jInfer/full/architecture.tex
@@
 
-For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{Hopcroft:2006:IAT:1177300}.
+For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{automatatheory}.
+All classes pertaining to regular expressions can be found in the package \code{cz.cuni.mff.ksi.jinfer.base.regexp}.
+In jInfer, we use extended regular expressions as they give us nicer syntax (and easier programming).
 
-All classes pertaining to regular expressions can be found in the package \code{cz.cuni.mff.ksi.jinfer.base.regexp}.
-
-In theory, regular expression is defined recursively:
+Regular expression is implemented as class \code{Regexp} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
+Each \code{Regexp} instance has one of the enum \code{RegexpType} type:
 \begin{itemize}
-	\item Every symbol $x$ from alphabet is regular expression representing language ${x}$.
-	\item If $R_1, R_2, \dots, R_n$ are regular expressions representing languages $L_1, L_2, \dots, L_n$, then $R_1 | R_2 | \dots | R_n$
-	is regular expression representing language $\bigcup_{i = 1}^{n}{L_i}$. Operation $|$ is called alternation.
-	\item If $R_1, R_2, \dots, R_n$ are regular expressions representing languages $L_1, L_2, \dots, L_n$, then $R_1, R_2 , \dots , R_n$
-	is regular expression representing language $L_1 \cdot L_2 \cdot \ldots \cdot L_n$. Where $\cdot$ is languague concatenation,
-	and operation $,$ is called concatenation as well.
-	\item If $R$ is regular expression representing language $L$, then $R*$ 
-	is regular expression representing language $L*$. In regular expressions, operator $*$ is called Kleene star.
+	\item Lambda - empty string (also called $\epsilon$ in literature),
+	\item Token - a letter of the alphabet,
+	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$,
+	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$,
+	\item Permutation - shortcut for all possible permutations of regular expressions. Our syntax to write down permutation is $(a\& b\& c\& d)$.
 \end{itemize}
 
-In common practice, extended regular expressions are commonly used. Although they are not extending set of
-languages representable by regular expressions, they give us nicer syntax. Extended regular expression adds syntax:
+Each \code{Regexp} instance has one instance of \code{RegexpInterval} as member.
+Class \code{RegexpInterval} represents POSIX-like intervals for expression:
 \begin{itemize}
-	\item $R{m,n}$, where $m,n \in N$ with meaning of $R$ at least $m$-times, at most $n$-times,
-	\item $R{m,}$ with meaning at least $m$-times,
-	\item $R?$ with meaning $R{0,1}$,
-	\item $R+$ with meaning $R{1,}$.
+	\item $a\{m,n\}$ means $a$ at least $m$-times, at most $n$-times,
+	\item $a\{m,\}$ means at least $m$-times (unbounded).
 \end{itemize}
 
-By this syntax, Kleene star is simply interval ${0,}$. This is just syntactic shortcut for long expresssions like $R, R, R, R?, R?$, which is just $R{3,5}$.
-
-In jInfer, we implement regular expresssions as class \code{Regexp} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
-Each \code{Regexp} instance has one of the \code{RegexpType} type:
-\begin{itemize}
-	\item Token - a letter of the language.
-	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$.
-	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$.
-	\item Permutation - shortcut for all possible permutations of regular expressions (to work with \code{xs:all} in future).
-	Our syntax to write down permutation is $(a  b c  d)$.
-\end{itemize}
-
-Each \code{Regexp} instance has interval associated with it (as a member). Class \code{RegexpInterval} represents all kinds of intervals
-${m,n}; {m,}$.
+Class \code{Regexp} can represent regular expression over any alphabet.
+This is done by using java generics, \code{Regexp} is implemented as \code{Regexp<T>}.
+Regular expression is in fact $n$-ary tree, for example expression $(a, b, (c | d))$ can be viewed as in figure \ref{reg:tree}.
+\begin{figure}
+\caption{Example tree for regular expression $(a, b, (c | d))$} \label{reg:tree}
+\includegraphics[scale=1]{reg_tree}
+\end{figure}
 

### 2011-01-10 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - picture
3154619 Create documentation for jInfer architecture - regexp draft completed

@@
 \usepackage{import}
+\usepackage{listings}
+\usepackage{color}
+\usepackage{textcomp}
+\usepackage{upquote}
+
+\definecolor{listinggray}{gray}{0.9}
+\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
+\lstset{
+%	backgroundcolor=\color{lbcolor},
+	tabsize=2,
+	rulecolor=,
+	language=java,
+        basicstyle=\small\ttfamily,
+        basewidth=0.51em,
+        upquote=true,
+        aboveskip={1.5\baselineskip},
+        columns=fixed,
+        showstringspaces=false,
+        extendedchars=true,
+        breaklines=true,
+%        prebreak =\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
+ %       frame=single,
+        showtabs=false,
+        showspaces=false,
+        showstringspaces=false,
+        identifierstyle=\ttfamily,
+%        keywordstyle=\color[rgb]{0,0,1},
+%        commentstyle=\color[rgb]{0.133,0.545,0.133},
+%        stringstyle=\color[rgb]{0.627,0.126,0.941},
+}
 
@@
 
+Type of regexp is held in \code{type} member in class \code{Regexp} and can be tested by calling methods \code{isLambda()}, \code{isToken()} etc.
+
 Each \code{Regexp} instance has one instance of \code{RegexpInterval} as member.
@@
 \end{itemize}
+Interval is either bounded (you have to set both lower and upper bound integers), or unbounded (you have to set only lower bound).
+Testing interval commonly follows routine:
+\begin{lstlisting}
+	RegexpInterval i = r.getInterval();
+	if (i.isUnbounded()) {
+		print(i.getMin());
+	} else {
+		print(i.getMin(), i.getMax());
+	}
+\end{lstlisting}
+That is, first check interval for being unbounded, only if it is bounded, you can ask for maximum.
 
@@
 This is done by using java generics, \code{Regexp} is implemented as \code{Regexp<T>}.
-Regular expression is in fact $n$-ary tree, for example expression $(a, b, (c | d))$ can be viewed as in figure \ref{reg:tree}.
+Only token regexps hold instance of type \code{T} in member \code{content}.
+
+Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in figure \ref{reg:tree}.
 \begin{figure}
-\caption{Example tree for regular expression $(a, b, (c | d))$} \label{reg:tree}
-\includegraphics[scale=1]{reg_tree}
+\caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg:tree}
+\centering\includegraphics[scale=0.5]{reg_tree.eps}
 \end{figure}
+We implement this tree by member of \code{Regexp} class called \code{children}, which is of type \code{List<Regexp<T>>}.
+List contains children of regexp in means of regexp tree.
+
+Regexp has to obey constraits:
+\begin{itemize}
+	\item type, children and interval have to be non-null references,
+	\item when type is lamba, content and interval has to be null,
+	\item when type is token, content has to be non-null,
+	\item when type concatenation, alternation or permutation, content has to be null.
+\end{itemize}
+These constraits are checked by constructors, so the best way to construct new regexps is by using 
+methods \code{getToken(), getConcatenation()} etc.
+
+Regexp instance is by default created as immutable, that is, once instantiated, you cannot add more children to list of children, cannot change type, content etc.
+It is to prevent missuse.
+In special circumstances, one does not know future children of regexp in time of creation.
+This occurs mainly in input modules, where by parsing XML data sequentially, one does not know contents of element in time of handling start element event.
+For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of members set.
+One has to fill in all properties carefully and call \code{setImmutable()} aftewards.
+Proper usage should be one of following:
+\begin{lstlisting}
+	Regexp<T> r = Regexp.<T>getMutable();
+	r.setInterval(...);
+	r.setType(RegexpType.LAMBDA);
+	r.setImmutable();
+\end{lstlisting}
+\begin{lstlisting}
+	Regexp<T> r = Regexp.<T>getMutable();
+	r.setInterval(...);
+	r.setType(RegexpType.TOKEN);
+	r.setContent(...)
+	r.setImmutable();
+\end{lstlisting}
+\begin{lstlisting}
+	Regexp<T> r = Regexp.<T>getMutable();
+	r.setInterval(...);
+	r.setType(RegexpType.CONCATENATION);
+	r.addChild(...);
+	r.addChild(...);
+	r.addChild(...);
+	r.setImmutable();
+\end{lstlisting}
+
+Finally, regexp contain one useful method for obtaining all leaves in the regexp tree, it is called \code{getTokens()} and it recursively traverses tree returning list of leaves (token type regexps).
+
+\subsection{XML representation}
 

### 2011-01-11 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - xml representation draft done

@@
 Class \code{Regexp} can represent regular expression over any alphabet.
-This is done by using java generics, \code{Regexp} is implemented as \code{Regexp<T>}.
+This is done by using java generics, \code{Regexp} evinceis implemented as \code{Regexp<T>}.
 Only token regexps hold instance of type \code{T} in member \code{content}.
 
-Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in figure \ref{reg:tree}.
+Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in figure \ref{reg_tree}.
 \begin{figure}
-\caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg:tree}
+\caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg_tree}
 \centering\includegraphics[scale=0.5]{reg_tree.eps}
@@
 \subsection{XML representation}
+XML data basically contains elements, text nodes (characters inside elements) and attributes.
+For maximum generality, we decided to break apart theese objects.
+We define three basic interfaces: \code{NamedNode}, \code{StructuralNode} and \code{ContentNode} (see package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes}).
+
+The first stands for bare node in XML document tree, it has its name and context withing the tree (path from root).
+The latter two extends \code{NamedNode} interface.
+\code{StructuralNode} is for nodes, which form structure of XML document tree: elements and text nodes.
+\code{ContentNode} is for nodes, that have content in XML documents: text nodes and attributes.
+We have three classes: \code{Element} for elements, \code{SimpleData} for text nodes, \code{Attribute} for attributes (see package \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes}).
+In theory, the classes and interfaces would be layed out as on figure \ref{interfaces_nodes}
+\begin{figure}
+\caption{How should interfaces and classes for XML representation look like in theory} \label{interfaces_nodes}
+\centering\includegraphics[scale=0.5]{interfaces_nodes.eps}
+\end{figure}
+\begin{figure}
+\caption{How are interfaces and classes for XML representation arranged in practice} \label{nodes}
+\centering\includegraphics[scale=0.5]{nodes.eps}
+\end{figure}
+\begin{figure}
+\caption{XML representing interfaces and classes in detail} \label{nodes_full}
+\centering\includegraphics[scale=0.3]{nodes_full.eps}
+\end{figure}
+
+For even more generality in design, we decided to implement abstract classes in midlevel:
+\begin{itemize}
+	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later),
+	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
+\end{itemize}
+As practice showed, for methods handling and infering structural properties, it is important to recognize whether structural node on input is element or text node.
+However methods for content devising don't need to know, if they are working on infering model for content of attribute or text node.
+
+Finally, our interface/class model for representing XML nodes is drafted on figure \ref{nodes}. Those, who are brave enough, can look on figure \ref{nodes_full}.
+
+In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{AbstractStructuralNode} variables.
+And \code{SimpleData} and \code{Attribute} have methods \code{getContentType()} and \code{getContent()} to work with content model.
+
+Class \code{Element} has two important members of course:
+\begin{itemize}
+	\item \code{Regexp<AbstractStructuralNode> subnodes} - for representing right side of grammar rule in resulting infered schema,
+	\item \code{List<Attribute> attributes} - for representing all attributes in resulting infered schema.
+\end{itemize}
+Theese two are filled by import modules, processed further by infering (simplifying) modules and finally exported by exporter modules. We will look at proper interfaces later.
+
+As in regular expressions, classes pertaining XML nodes are by default immutable.
+For elements, it means no adding of attributes and changing regexp reference (regexp instance itself is immutable).
+Same \code{getMutable()} principles and good usage practises hold for theese classes.
+
+\subsection{Module interfaces}
 

### 2011-01-11 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - font and minors

@@
 \usepackage{amssymb}
-\usepackage{import}
+\usepackage{palatino}
 \usepackage{listings}
-\usepackage{color}
 \usepackage{textcomp}
@@
 
-Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in figure \ref{reg_tree}.
+Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in fig. \ref{reg_tree}.
 \begin{figure}
@@
 We have three classes: \code{Element} for elements, \code{SimpleData} for text nodes, \code{Attribute} for attributes (see package \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes}).
-In theory, the classes and interfaces would be layed out as on figure \ref{interfaces_nodes}
+In theory, the classes and interfaces would be layed out as on fig. \ref{interfaces_nodes}
 \begin{figure}
@@
 
-Finally, our interface/class model for representing XML nodes is drafted on figure \ref{nodes}. Those, who are brave enough, can look on figure \ref{nodes_full}.
+Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full}.
 

### 2011-01-11 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - desired section
3154619 Create documentation for jInfer architecture - minors

@@
 \usepackage{amssymb}
-\usepackage{palatino}
+%\usepackage{palatino}
+\usepackage[T1]{fontenc}
+\usepackage[sc]{mathpazo}
+\linespread{1.05}
+
 \usepackage{listings}
@@
         upquote=true,
-        aboveskip={1.5\baselineskip},
+%        aboveskip={1.5\baselineskip},
         columns=fixed,
@@
         showstringspaces=false,
-        identifierstyle=\ttfamily,
+%        identifierstyle=\ttfamily,
 %        keywordstyle=\color[rgb]{0,0,1},
@@
 \end{figure}
-We implement this tree by member of \code{Regexp} class called \code{children}, which is of type \code{List<Regexp<T>>}.
+We implement this tree by member of \code{Regexp} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
 List contains children of regexp in means of regexp tree.
@@
 
-\subsection{Module interfaces}
+\subsection{Something that will vektor declare priority}
+
+\subsection{(Nondeterministic) Finite Automaton}
 

### 2011-01-11 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - eps and listing strip

@@
 \documentclass[a4paper,10pt,oneside]{article}
-\usepackage{epsfig}
+\usepackage{graphicx}
 \usepackage{color}
@@
 \usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-%\usepackage{palatino}
-\usepackage[T1]{fontenc}
-\usepackage[sc]{mathpazo}
-\linespread{1.05}
-
-\usepackage{listings}
-\usepackage{textcomp}
-\usepackage{upquote}
-
-\definecolor{listinggray}{gray}{0.9}
-\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
-\lstset{
-%	backgroundcolor=\color{lbcolor},
-	tabsize=2,
-	rulecolor=,
-	language=java,
-        basicstyle=\small\ttfamily,
-        basewidth=0.51em,
-        upquote=true,
-%        aboveskip={1.5\baselineskip},
-        columns=fixed,
-        showstringspaces=false,
-        extendedchars=true,
-        breaklines=true,
-%        prebreak =\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
- %       frame=single,
-        showtabs=false,
-        showspaces=false,
-        showstringspaces=false,
-%        identifierstyle=\ttfamily,
-%        keywordstyle=\color[rgb]{0,0,1},
-%        commentstyle=\color[rgb]{0.133,0.545,0.133},
-%        stringstyle=\color[rgb]{0.627,0.126,0.941},
-}
 
@@
 \setlength{\topmargin}{1.5cm}
-\setlength{\textheight}{26cm}
+\setlength{\textheight}{24cm}
 \setlength{\textwidth}{18cm}
@@
 Testing interval commonly follows routine:
-\begin{lstlisting}
+\begin{verbatim}
 RegexpInterval i = r.getInterval();
@@
 }
-\end{lstlisting}
+\end{verbatim}
 That is, first check interval for being unbounded, only if it is bounded, you can ask for maximum.
@@
 \caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg_tree}
-\centering\includegraphics[scale=0.5]{reg_tree.eps}
+\centering\includegraphics[scale=0.5]{reg_tree}
 \end{figure}
@@
 Proper usage should be one of following:
-\begin{lstlisting}
+\begin{verbatim}
 	Regexp<T> r = Regexp.<T>getMutable();
@@
 	r.setImmutable();
-\end{lstlisting}
-\begin{lstlisting}
+\end{verbatim}
+\begin{verbatim}
 	Regexp<T> r = Regexp.<T>getMutable();
@@
 	r.setImmutable();
-\end{lstlisting}
+\end{verbatim}
 
@@
 \caption{How should interfaces and classes for XML representation look like in theory} \label{interfaces_nodes}
-\centering\includegraphics[scale=0.5]{interfaces_nodes.eps}
+\centering\includegraphics[scale=0.5]{interfaces_nodes}
 \end{figure}
@@
 \caption{How are interfaces and classes for XML representation arranged in practice} \label{nodes}
-\centering\includegraphics[scale=0.5]{nodes.eps}
+\centering\includegraphics[scale=0.5]{nodes}
 \end{figure}
@@
 \caption{XML representing interfaces and classes in detail} \label{nodes_full}
-\centering\includegraphics[scale=0.3]{nodes_full.eps}
+\centering\includegraphics[scale=0.3]{nodes_full}
 \end{figure}

### 2011-01-12 doc/jInfer/full/architecture/architecture.tex
Documentation: Architecture: Inference Process

@@
 
-\subsection{Something that will vektor declare priority}
+\subsection{(Nondeterministic) Finite Automaton}
 
-\subsection{(Nondeterministic) Finite Automaton}
+% TODO explain what IG (or a grammar, for that matter) is
+% TODO explain metadata and those that we already use
+% TODO section about NBMs
+
+\section{Inference process}
+\begin{figure}
+	\caption{High-level view of the inference process} \label{inference_process}
+	\centering\includegraphics[scale=0.5]{inference_process}
+\end{figure}
+
+% TODO anti: is it really necessary to write fig. by hand? Shouldn't it be automatic?
+
+The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig.
+\ref{inference_process}.
+From the high-level viewpoint, it consists of three consecutive steps carried out by three different modules:
+\begin{enumerate}
+	\item Initial Grammar (IG) generation: done by the \textit{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation.
+All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
+	% TODO anti: how about the whole examples in italic?
+	For example, a trivial XML document
+	% TODO anti: please format the indentation	
+	\begin{verbatim}
+		<a>
+			<b>
+				<c/>
+			</b>
+			<d/>
+		</a>
+	\end{verbatim}
+	would translate into the following IG rules
+	\begin{verbatim}
+		a -> b, d
+		b -> c
+	\end{verbatim}	
+	\item Simplification: done by the \textit{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules.
+User interaction might be used in this step to help achieve better simplification.
+At the end of this step, all rules are sent to the last step.\\
+	For example, rules
+	\begin{verbatim}
+		a -> b
+		a -> c, d, d, d
+	\end{verbatim}
+	could be simplified to a single rule
+	\begin{verbatim}
+		a -> b | (c, d*)
+	\end{verbatim}
+	\item Schema export: done by the \textit{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema from the simplified rules.
+Result of this step is a string representation of the schema, which is sent back to the framework (and later displayed, saved, etc).\\
+	% TODO vektor Example
+\end{enumerate}
+Important thing to note here is that all these steps are executed consecutively.
+That means, \textit{Simplifier} is only started \textit{after} the \textit{IGG} completely finished its work and returned IG to be simplified.
+Similarly, \textit{SchemaGen} gets all the rules to export at once, in one list.\\
+
+This modular architecture means that (at least theoretically) it is possible to replace any and all of these modules with \textit{something} else that does similar job.
+In practice, it might be useful to use the logic already implemented in \textit{BasicIGG} and just extend it to handle new input type.
+
+% TODO vektor Describe requirements on what gets passed on between modules
+
+\subsection{Programmatic view}
+
+From developer's point of view, inference modules are just properly annotated classes implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
+
+\begin{itemize}
+	\item \code{IGGenerator}
+	\item \code{Simplifier}
+	\item \code{SchemaGenerator}
+\end{itemize}
+
+A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{SimplifierImpl}.\\
+
+% TODO vektor Connection to NBMs
+
+Annotation required for the framework to recognize such a class as an inference module is the following
+\begin{verbatim}
+	@ServiceProvider(service = <interface>.class)
+\end{verbatim}
+for example
+\begin{verbatim}
+	@ServiceProvider(service = Simplifier.class)
+\end{verbatim}
+
+The most important method in each module is \code{start}, defined in each of the interfaces.
+This method is called by the framework when the respective step of inference is being executed.
+It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished.
+We will look at both parameter now in more detail.
+
+\subsubsection{Module input}
+Each inference module takes the actual input data as the first parameter of its \code{start} method. The type of the argument differs based on the inference module.\\
+
+\textit{Initial Grammar Generator} takes an object of type \textit{Input}.
+This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}.
+Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.\\
+
+\textit{Simplifier} and \textit{Schema Generator} take grammar, % TODO vektor Link to grammar definition
+in other words a list of rules as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
+
+\subsubsection{Module output}
+Second parameter of each inference module's \code{start} method is a callback object. There are 3 callback interfaces defined in the \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package
+
+\begin{itemize}
+	\item \code{IGGeneratorCallback}
+	\item \code{SimplifierCallback}
+	\item \code{SchemaGeneratorCallback}
+\end{itemize}
+
+Each callback interface naturally belongs to the similarly named module interface.
+As their respective interfaces, also callbacks define one crucial method: \textit{finished}.
+Each inference module is responsible for invoking this method on the callback it got as a parameter after it has finished its work and has results to be passed on.
+Again, these 3 finished methods have different arguments based on the inference module.\\
+
+\code{IGGeneratorCallback.finished()} and \code{SimplifierCallback.finished()} have a grammar (Initial Grammar in case of IGG) as their only argument.\\
+
+\code{SchemaGeneratorCallback.finished} has two \code{String}s as arguments: \code{schema} is the actual string representation of the resulting schema, \code{extension} is a file extension of the result (such as "dtd" or "xsd") which the framework will use when saving the result in a file.
+
+\subsubsection{Error handling}
+
+Because the run of each inference module is encapsulated in a \code{try-catch} block by the framework, it is safe to throw any exception out of the \code{start} method: it will get logged, presented to the user and inference will stop.
+However, if the module uses threads that could throw an exception, it is responsible for catching these exceptions and possibly re-throwing them in the thread where \code{start} runs.
+
+\subsubsection{Interruptions}
+
+User running the inference might change his mind and try to stop this. For this reason, modules have to check for this case in every time-consuming place such as long loops with the following code
+% TODO anti Please indent
+\begin{verbatim}
+	for (forever) {
+		if (Thread.interrupted()) {
+			throw new InterruptedException();
+		}
+		doStuff();
+	}
+\end{verbatim}
+
+\subsubsection{Runner}
+
+The part of framework responsible for actually gathering user input, running all modules one after another and presenting the results is the \code{Runner} class  in \code{cz.cuni.mff.ksi.jinfer.runner} package.\\
+
+A new instance of \code{Runner} is constructed for each inference run.
+While being created, \code{Runner}  loads the preferences for current project and looks up user-selected inference modules.
+Also, callback objects pointing back to methods in \code{Runner} are created.
+The inference process itself is then as follows
+
+% TODO vektor Picture!!!
+\begin{enumerate}
+	\item Selected IGG's \code{start} is encapsulated with error/interruption handling and executed, passing \code{Input} and first callback as parameters.
+	\item When IGG finishes, it invokes callback's \code{finish} method, passing the IG as parameter.
+	\item This in turn causes \code{Runner} to encapsulate and execute Simplifier's \code{start}, passing IG from the first callback and the second callback as parameters.
+	\item When Simplifier finishes, it invokes callback's \code{finish} method, passing the simplified grammar as parameter.
+	\item This again causes \code{Runner} to encapsulate and execute SchemaGen's \code{start}, passing the simplified grammar from the second callback and the third callback as parameters.
+	\item SchemaGen finishes and invokes the last callback's \code{finish}, passing the resulting schema and its extension as parameters.
+	\item \code{Runner} receives the resulting schema and based on preferences, saves it to a file, displays it, etc.
+\end{enumerate}
+
+% TODO vektor Talk about Capabilities
 

### 2011-01-12 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - better fonts and pictures
3154619 Create documentation for jInfer architecture - minors
3154619 Create documentation for jInfer architecture - verbatim with spaces instead tabs
Documentation: Architecture: Inference Process

@@
 \usepackage[T1]{fontenc}
+\usepackage{tgpagella}
+%\usepackage[scale=0.9]{tgcursor}
+%\usepackage[scale=0.9]{tgheros}
 
@@
 \caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg_tree}
-\centering\includegraphics[scale=0.5]{reg_tree}
+\centering\includegraphics[scale=\myscale]{reg_tree}
 \end{figure}
@@
 \caption{How should interfaces and classes for XML representation look like in theory} \label{interfaces_nodes}
-\centering\includegraphics[scale=0.5]{interfaces_nodes}
+\centering\includegraphics[scale=\myscale]{interfaces_nodes}
 \end{figure}
@@
 \caption{How are interfaces and classes for XML representation arranged in practice} \label{nodes}
-\centering\includegraphics[scale=0.5]{nodes}
+\centering\includegraphics[scale=\myscale]{nodes}
 \end{figure}
@@
 
-% TODO explain what IG (or a grammar, for that matter) is
+\subsection{Rules, Grammars, etc}
+
+jInfer and its documentation uses extended context-free grammars. % TODO vektor Cite, for example http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf
+\textit{Rules} in such grammar are in the form
+% TODO anti Please make a proper arrow
+\[  Left Hand Side (LHS)  -> Right Hand Side (RHS) \]
+where LHS is a letter of the alphabet (token), RHS is a regular expression over this alphabet. Example would be
+\begin{verbatim}
+  a -> b, (c | d)*
+\end{verbatim}
+
+In jInfer each such rule is represented with an \code{Element} instance. In this representation, the \code{Element} itself is the LHS, its \code{subnodes} are the RHS.\\
+
+Another important notion is a \textit{grammar}.
+A grammar consists of its rules, so in jInfer a grammar is just a collection of \code{Element}s.
+Closely related term is \textit{Initial Grammar}, which for us is a grammar consisting of rules with \textit{simple} right hand sides, i.e.
+just concatenations of tokens.
+Initial Grammar is produced by Initial Grammar Generator.
+% TODO vektor Link
+
 % TODO explain metadata and those that we already use
@@
 	\caption{High-level view of the inference process} \label{inference_process}
-	\centering\includegraphics[scale=0.5]{inference_process}
+	\centering\includegraphics[scale=\myscale]{inference_process}
 \end{figure}
@@
 	For example, a trivial XML document
-	% TODO anti: please format the indentation	
 	\begin{verbatim}

### 2011-01-12 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - consistent xml representation looking, some points for vektor and xml example of tree

@@
 
-\textit{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\
+\emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\
 
@@
 \subsection{Regular expressions}
-
+% TODO anti lambda, epsilon notation
 For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{automatatheory}.
@@
 	r.setType(RegexpType.LAMBDA);
-	r.setImmutable();
+	r.setImmutable();cz.cuni.mff.ksi.jinfer.base
 \end{verbatim}
@@
 
+\begin{figure}
+\caption{XML document representation} \label{xml_example}
+\centering\includegraphics[scale=\myscale]{xml_example}
+\end{figure}
+Let's take an example, the following XML document would be represented as on fig. \ref{xml_example}.
+\begin{verbatim}
+  <person name="john" surname="smith">
+    <info>
+      Some text
+      <note/>
+    </info>
+    <more/>
+  </person>
+\end{verbatim}
+
 As in regular expressions, classes pertaining XML nodes are by default immutable.
@@
 
-\subsection{(Nondeterministic) Finite Automaton}
-
 \subsection{Rules, Grammars, etc}
 
-jInfer and its documentation uses extended context-free grammars. % TODO vektor Cite, for example http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf
-\textit{Rules} in such grammar are in the form
-% TODO anti Please make a proper arrow
-\[  Left Hand Side (LHS)  -> Right Hand Side (RHS) \]
+jInfer and its documentation uses extended context-free grammars\cite{extendedcfg}. % TODO vektor seee literature.bib and repair citation http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf
+\emph{Rules} in such grammar are in the form
+% TODO vektor read "the not so short introduction to latex" and math mode section
+$$
+  \textnormal{Left Hand Side (LHS)} \to \textnormal{Right Hand Side (RHS)}
+$$
 where LHS is a letter of the alphabet (token), RHS is a regular expression over this alphabet. Example would be
-\begin{verbatim}
-  a -> b, (c | d)*
-\end{verbatim}
-
+$$
+  a \to b, (c | d)\ast
+$$
 In jInfer each such rule is represented with an \code{Element} instance. In this representation, the \code{Element} itself is the LHS, its \code{subnodes} are the RHS.\\
 
-Another important notion is a \textit{grammar}.
+Another important notion is a \emph{grammar}.
 A grammar consists of its rules, so in jInfer a grammar is just a collection of \code{Element}s.
-Closely related term is \textit{Initial Grammar}, which for us is a grammar consisting of rules with \textit{simple} right hand sides, i.e.
+Closely related term is \emph{Initial Grammar}, which for us is a grammar consisting of rules with \emph{simple} right hand sides, i.e.
 just concatenations of tokens.
@@
 % TODO vektor Link
+% TODO vektor don't forget about advanced interface by capabilities.... don't know if to be mentioned right here, but don't forget
+% TODO explain metadata and those that we already use
+% TODO section about NBMs (anti: but i suppose it on the end of document)
 
-% TODO explain metadata and those that we already use
-% TODO section about NBMs
+\subsection{Nondeterministic Finite Automaton}
+For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
+Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
+Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.
+
+Let's begin by smalles of the classes, the \code{State<T>}.
+It represents automaton state, it has its \code{name} member (integer)
 
@@
 
-% TODO anti: is it really necessary to write fig. by hand? Shouldn't it be automatic?
+% TODO anti: is it really necessary to write fig. by hand? Shouldn't it be automatic? anti: don't know, who cares... once it figure, once picture and maybe painting :D
 
@@
 \begin{enumerate}
-	\item Initial Grammar (IG) generation: done by the \textit{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation.
+	\item Initial Grammar (IG) generation: done by the \emph{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation.
 All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
-	% TODO anti: how about the whole examples in italic?
+	% TODO anti: how about the whole examples in italic? anti: bleee
 	For example, a trivial XML document
@@
 	would translate into the following IG rules
-	\begin{verbatim}
-		a -> b, d
-		b -> c
-	\end{verbatim}	
-	\item Simplification: done by the \textit{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules.
+	\begin{eqnarray*}
+		a & \to &  b, d \\
+		b & \to & c \\
+		c & \to & \lambda \\
+		d & \to & \lambda 
+	\end{eqnarray*}
+	\item Simplification: done by the \emph{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules.
 User interaction might be used in this step to help achieve better simplification.
@@
 	For example, rules
-	\begin{verbatim}
-		a -> b
-		a -> c, d, d, d
-	\end{verbatim}
+	\begin{eqnarray*}
+		a & \to & b \\
+		a & \to & c, d, d, d
+	\end{eqnarray*}
 	could be simplified to a single rule
-	\begin{verbatim}
-		a -> b | (c, d*)
-	\end{verbatim}
-	\item Schema export: done by the \textit{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema from the simplified rules.
+	\begin{eqnarray*}
+		a & \to & b | (c, d\ast)
+	\end{eqnarray*}
+	\item Schema export: done by the \emph{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema from the simplified rules.
 Result of this step is a string representation of the schema, which is sent back to the framework (and later displayed, saved, etc).\\
@@
 Important thing to note here is that all these steps are executed consecutively.
-That means, \textit{Simplifier} is only started \textit{after} the \textit{IGG} completely finished its work and returned IG to be simplified.
-Similarly, \textit{SchemaGen} gets all the rules to export at once, in one list.\\
+That means, \emph{Simplifier} is only started \emph{after} the \emph{IGG} completely finished its work and returned IG to be simplified.
+Similarly, \emph{SchemaGen} gets all the rules to export at once, in one list.\\
 
-This modular architecture means that (at least theoretically) it is possible to replace any and all of these modules with \textit{something} else that does similar job.
-In practice, it might be useful to use the logic already implemented in \textit{BasicIGG} and just extend it to handle new input type.
+This modular architecture means that (at least theoretically) it is possible to replace any and all of these modules with \emph{something} else that does similar job.
+In practice, it might be useful to use the logic already implemented in \emph{BasicIGG} and just extend it to handle new input type.
 
@@
 
-\textit{Initial Grammar Generator} takes an object of type \textit{Input}.
+\emph{Initial Grammar Generator} takes an object of type \emph{Input}.
 This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}.
@@
 
-\textit{Simplifier} and \textit{Schema Generator} take grammar, % TODO vektor Link to grammar definition
+\emph{Simplifier} and \emph{Schema Generator} take grammar, % TODO vektor Link to grammar definition
 in other words a list of rules as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
@@
 Each callback interface naturally belongs to the similarly named module interface.
-As their respective interfaces, also callbacks define one crucial method: \textit{finished}.
+As their respective interfaces, also callbacks define one crucial method: \emph{finished}.
 Each inference module is responsible for invoking this method on the callback it got as a parameter after it has finished its work and has results to be passed on.
@@
 User running the inference might change his mind and try to stop this. For this reason, modules have to check for this case in every time-consuming place such as long loops with the following code
-% TODO anti Please indent
 \begin{verbatim}

### 2011-01-12 doc/jInfer/full/architecture/architecture.tex
Documentation: Architecture

@@
 \begin{itemize}
-	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later),
+	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later), % TODO anti Link to proper subsection
 	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
@@
 
+TODO anti
+
 \subsection{Rules, Grammars, etc}
@@
 % TODO vektor Link
-% TODO vektor don't forget about advanced interface by capabilities.... don't know if to be mentioned right here, but don't forget
-% TODO explain metadata and those that we already use
-% TODO section about NBMs (anti: but i suppose it on the end of document)
+
+\subsection{Metadata}
+To allow simple extensibility of our rules, each descendant of \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
+
+At the present, jInfer modules use the following metadata, all defined in \textit{IGGUtils} class:
+\begin{itemize}
+	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
+	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
+	\item \code{is.sentinel}, filled in by XSDImport module: TODO anti Explain
+	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
+\end{itemize}
+
+\section{NetBeans Modules}
+
+TODO vektor
 
@@
 
-% TODO anti: is it really necessary to write fig. by hand? Shouldn't it be automatic? anti: don't know, who cares... once it figure, once picture and maybe painting :D
-
 The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig.
@@
 All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
-	% TODO anti: how about the whole examples in italic? anti: bleee
 	For example, a trivial XML document

### 2011-01-13 doc/jInfer/full/architecture/architecture.tex
little editing

@@
 \newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\jmodule}[1]{\emph{#1}}
 
@@
 
-Afterwards the interfaces of basic inference modules - Initial Grammar Gen-
-erator, Simplifier and Schema Generator - will be explained.
+Afterwards the interfaces of basic inference modules - \jmodule{Initial Grammar Generator}, \jmodule{Simplifier} and \jmodule{Schema Generator} - will be explained.
 
@@
 \texttt{objects.utils}).
-All in all, a package in the Base module could look like
-\code{cz.cuni.mff.ksi.jinfer.base.objects.utils}
+All in all, a package in the \jmodule{Base} module could look like \code{cz.cuni.mff.ksi.jinfer.base.objects.utils}.
 
@@
 \subsection{Regular expressions}
-% TODO anti lambda, epsilon notation
 For general information on regular expressions, please refer to \cite{wikiregexp}, \cite{automatatheory}.
@@
 
-Regular expression is implemented as class \code{Regexp} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
-Each \code{Regexp} instance has one of the enum \code{RegexpType} type:
+Regular expression is implemented as class \code{Regexp<T>} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
+Each \code{Regexp<T>} instance has one of the enum \code{RegexpType} type:
 \begin{itemize}
-	\item Lambda - empty string (also called $\epsilon$ in literature),
+	\item Lambda ($\lambda$) - empty string (also called $\epsilon$ in literature),
 	\item Token - a letter of the alphabet,
@@
 
-Type of regexp is held in \code{type} member in class \code{Regexp} and can be tested by calling methods \code{isLambda()}, \code{isToken()} etc.
+Type of regexp is held in \code{type} member in class \code{Regexp<T>} and can be tested by calling methods \code{isLambda()}, \code{isToken()} etc.
 
-Each \code{Regexp} instance has one instance of \code{RegexpInterval} as member.
+Each \code{Regexp<T>} instance has one instance of \code{RegexpInterval} as member.
 Class \code{RegexpInterval} represents POSIX-like intervals for expression:
@@
 \end{itemize}
-Interval is either bounded (you have to set both lower and upper bound integers), or unbounded (you have to set only lower bound).
-Testing interval commonly follows routine:
+Interval can be either bounded (you have to set both lower and upper bound integers), or unbounded (you have to set only lower bound).
+Testing interval value commonly follows routine:
 \begin{verbatim}
@@
 
-Class \code{Regexp} can represent regular expression over any alphabet.
-This is done by using java generics, \code{Regexp} evinceis implemented as \code{Regexp<T>}.
+Class \code{Regexp<T>} can represent regular expression over any alphabet.
+This is done by using java generics.
 Only token regexps hold instance of type \code{T} in member \code{content}.
@@
 \end{figure}
-We implement this tree by member of \code{Regexp} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
+We implement this tree by member of \code{Regexp<T>} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
 List contains children of regexp in means of regexp tree.
@@
 \begin{itemize}
-	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later), % TODO anti Link to proper subsection
+	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metada}),
 	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
@@
 \end{itemize}
-Theese two are filled by import modules, processed further by infering (simplifying) modules and finally exported by exporter modules. We will look at proper interfaces later.
+Theese two are filled by import modules, processed further by infering (simplifying) modules and finally exported by exporter modules.
+
+As in regular expressions, classes pertaining XML nodes are by default immutable.
+For elements, it means no adding of attributes and changing regexp reference (regexp instance itself is immutable as well).
+Same \code{getMutable()} principles and good usage practises as for regexps, hold for theese classes.
 
@@
 \end{figure}
-Let's take an example, the following XML document would be represented as on fig. \ref{xml_example}.
+Let's take an example, the following XML document would be represented as tree on fig. \ref{xml_example}.
 \begin{verbatim}
@@
 \end{verbatim}
-
-As in regular expressions, classes pertaining XML nodes are by default immutable.
-For elements, it means no adding of attributes and changing regexp reference (regexp instance itself is immutable).
-Same \code{getMutable()} principles and good usage practises hold for theese classes.
+Although in example we present whole document tree, input modules produce slightly different format (consisting of rules).
 
 TODO anti
+% hey vektor, what to DO here?
 
@@
 
-jInfer and its documentation uses extended context-free grammars\cite{extendedcfg}. % TODO vektor seee literature.bib and repair citation http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf
+jInfer and its documentation uses extended context-free grammars\cite{extendedcfg}.
+% TODO vektor check citation of http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf in literature.bib
 \emph{Rules} in such grammar are in the form
-% TODO vektor read "the not so short introduction to latex" and math mode section
 $$
@@
 Closely related term is \emph{Initial Grammar}, which for us is a grammar consisting of rules with \emph{simple} right hand sides, i.e.
-just concatenations of tokens.
-Initial Grammar is produced by Initial Grammar Generator.
+just concatenations of tokens (even with no children, see fig.
+\ref{xml_example} again).
+Initial Grammar is produced by \jmodule{Initial Grammar Generator}.
 % TODO vektor Link
-
-\subsection{Metadata}
-To allow simple extensibility of our rules, each descendant of \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
-
-At the present, jInfer modules use the following metadata, all defined in \textit{IGGUtils} class:
-\begin{itemize}
-	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
-	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
-	\item \code{is.sentinel}, filled in by XSDImport module: TODO anti Explain
-	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
-\end{itemize}
 
@@
 % TODO vektor Describe requirements on what gets passed on between modules
+
+% anti : should be after IGG explained, to fulfil is.sentinel explanation i will refer back.
+\subsection{Node metadata} \label{section_metadata}
+To allow simple extensibility of rules, class \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
+
+At the present, jInfer modules use the following metadata, all defined in \code{IGGUtils} class:
+\begin{itemize}
+	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
+	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
+	\item \code{is.sentinel}, filled in by  BasicIGG and XSDImport module: when importing XML document, one can build whole tree in memory (as in fig.
+\ref{xml_example}) and then construct list of rules from it (thus saving memory).
+With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import) but right side can be defined anywhere in source file.
+To save complicated loading of whole schema, searching and pairing elements thorough rules, nodes on right side of rule are created empty - holding only name of node.
+Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling node as sentinel.
+	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
+\end{itemize}
 

### 2011-01-13 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - section about automatons complete

@@
 \usepackage{url}
+\usepackage{subfigure}
 \usepackage[utf8]{inputenc}
@@
 \begin{figure}
+\centering\includegraphics[scale=\myscale]{reg_tree}
 \caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg_tree}
-\centering\includegraphics[scale=\myscale]{reg_tree}
 \end{figure}
@@
 \begin{figure}
+\centering\includegraphics[scale=\myscale]{interfaces_nodes}
 \caption{How should interfaces and classes for XML representation look like in theory} \label{interfaces_nodes}
-\centering\includegraphics[scale=\myscale]{interfaces_nodes}
 \end{figure}
 \begin{figure}
+\centering\includegraphics[scale=\myscale]{nodes}
 \caption{How are interfaces and classes for XML representation arranged in practice} \label{nodes}
-\centering\includegraphics[scale=\myscale]{nodes}
 \end{figure}
 \begin{figure}
+\centering\includegraphics[scale=0.3]{nodes_full}
 \caption{XML representing interfaces and classes in detail} \label{nodes_full}
-\centering\includegraphics[scale=0.3]{nodes_full}
 \end{figure}
@@
 \begin{figure}
+\centering\includegraphics[scale=\myscale]{xml_example}
 \caption{XML document representation} \label{xml_example}
-\centering\includegraphics[scale=\myscale]{xml_example}
 \end{figure}
@@
 
-\subsection{Nondeterministic Finite Automaton}
-For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
-Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
-Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.
-
-Let's begin by smalles of the classes, the \code{State<T>}.
-It represents automaton state, it has its \code{name} member (integer)
-
 \section{Inference process}
 \begin{figure}
+	\centering\includegraphics[scale=\myscale]{inference_process}
 	\caption{High-level view of the inference process} \label{inference_process}
-	\centering\includegraphics[scale=\myscale]{inference_process}
 \end{figure}
@@
 \begin{enumerate}
-	\item Initial Grammar (IG) generation: done by the \emph{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation.
+	\item Initial Grammar (IG) generation: done by the \jmodule{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation.
 All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
@@
 	\end{eqnarray*}
-	\item Simplification: done by the \emph{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules.
+	\item Simplification: done by the \jmodule{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules (exactly one rule for each element).
 User interaction might be used in this step to help achieve better simplification.
-At the end of this step, all rules are sent to the last step.\\
+At the end of this step, all rules are sent to the export step.\\
 	For example, rules
@@
 	\end{eqnarray*}
-	\item Schema export: done by the \emph{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema from the simplified rules.
+	\item Schema export: done by the \jmodule{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema file from the simplified rules.
 Result of this step is a string representation of the schema, which is sent back to the framework (and later displayed, saved, etc).\\
@@
 Important thing to note here is that all these steps are executed consecutively.
-That means, \emph{Simplifier} is only started \emph{after} the \emph{IGG} completely finished its work and returned IG to be simplified.
-Similarly, \emph{SchemaGen} gets all the rules to export at once, in one list.\\
+That means, \jmodule{Simplifier} is only started \emph{after} the \jmodule{IGG} completely finished its work and returned IG to be simplified.
+Similarly, \jmodule{SchemaGen} gets all the rules to export at once, in one list.\\
 
-This modular architecture means that (at least theoretically) it is possible to replace any and all of these modules with \emph{something} else that does similar job.
-In practice, it might be useful to use the logic already implemented in \emph{BasicIGG} and just extend it to handle new input type.
-
-% TODO vektor Describe requirements on what gets passed on between modules
+This modular architecture means that it is possible to replace any and/or all of these modules with \emph{something} else that does similar job.
+In practice, scientists will probably implement new \jmodule{Simplifier} modules, while using our implementation of BasicIGG and export modules.
+This modular scheme is used inside our modules to divide them into submodules.
+It is possible to add new file format processing to import/export.
+One can extend automaton merging state algorithm of our implementation of \cite{ahonen} in TwoStepSimplifier module by replacing submodules.
 
@@
 \ref{xml_example}) and then construct list of rules from it (thus saving memory).
-With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import) but right side can be defined anywhere in source file.
+With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import), but right side can be defined anywhere in source file.
 To save complicated loading of whole schema, searching and pairing elements thorough rules, nodes on right side of rule are created empty - holding only name of node.
-Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling node as sentinel.
+Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
 	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
+	All of these metadata are of set/not set character (using \code{Boolean.TRUE}).
 \end{itemize}
+
+% TODO vektor Describe requirements on what gets passed on between modules
+
+\subsection{Nondeterministic Finite Automaton}
+For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
+Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
+Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.
+
+Let's begin by smallest of the classes, the \code{State<T>}.
+It represents automaton state, it has two integer members: \code{name} and \code{finalCount}.
+Name clearly serves as name of state in visualization and to string conversion.
+Final count is for representing whether the state is final in automaton.
+The field is not true/false but integral to help algorithms use stastistics over automatons (how many times in XML input is this state final?).
+
+Next we have \code{Step<T>} which stands for automaton transition.
+It has its \code{source} and \code{destination} states references.
+Symbol accepted by using this transition is store in \code{acceptsSymbol} member of generic type \code{T}.
+And finally member \code{useCount}, which is integer stating how many times the transition was used when constructing prefix tree automaton from input data.
+Simplifying algorithm can use this number for statistic purposes.
+
+\code{Automaton<T>} class puts theese together into nondeterministic finite automaton.
+It has reference to \code{initialState}, it has \code{newStateName} integer value to assure unique state names inside one automaton (incrementing every time new state creates).
+We use two maps to implement transition function (delta($\delta$) function  in most literature).
+One map of type \code{Map<State<T>, Set<Step<T>>>} called \code{delta} represents mapping from state into set of all outgoing transitions from state.
+Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better perfomance).
+There exist only one instance of each step from one state to another.
+That instance is referenced in delta map (on place of source state), and in reverse delta map (on place of destination state).
+Loops are no speciality, just source = destination.
+
+Automaton supports creating automaton as a copy of another one (not reference copy, but true copy), this can be used when searching solution space to create more versions of automaton to edit.
+We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
+Create empty automaton and then call this method for every positive example input string of language.
+You will get PTA with useCounts and finalCounts set properly on steps/states.
+
+\begin{figure}
+	\centering
+	\subfigure[Before merge.]
+	{
+		\includegraphics[scale=\myscale]{automaton_merge1}
+		\label{automaton_merge1}
+	}
+	\subfigure[After states 3 and 4 merged.]
+	{
+		\centering\includegraphics[scale=\myscale]{automaton_merge2}
+		\label{automaton_merge2}
+	}
+	\subfigure[After states 2 and 4 merged.]
+	{
+		\centering\includegraphics[scale=\myscale]{automaton_merge3}
+		\label{automaton_merge3}
+	}
+	\caption{Sample automaton.} \label{automaton_merge}
+\end{figure}
+
+The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates()} method.
+Method merges second state given into first one (or all states in list into first one in list).
+All \{in | out\}-transitions are redirected properly (or discarted as needed).
+Use count and final count are updated as sums of values from merging transitions/states.
+One can ask for merging states, that are merged out from automaton already.
+Take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
+One asks to merge states 3 and 4.
+State 3 then becomes final state with loop and state 4 disappears (fig. \ref{automaton_merge2}).
+If then one ask to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
+This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).
 

### 2011-01-13 doc/jInfer/full/architecture/architecture.tex
minors

@@
 \begin{itemize}
-	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metada}),
+	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metadata}),
 	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
@@
 
-TODO anti
-% hey vektor, what to DO here?
-
-\subsection{Rules, Grammars, etc}
+\subsection{Rules and grammars}
 
@@
 
-\section{NetBeans Modules}
+\subsection{Nondeterministic Finite Automaton}
+We recommend skipping this section in first read as it describes advanced features.
+For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
+Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
+Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.
 
-TODO vektor
+\paragraph{State}
+Let's begin by smallest of the classes, the \code{State<T>}.
+It represents automaton state, it has two integer members: \code{name} and \code{finalCount}.
+Name clearly serves as name of state in visualization and to string conversion.
+Final count is for representing whether the state is final in automaton.
+The field is not true/false but integral to help algorithms use stastistics over automatons (how many times in XML input is this state final?).
+
+\paragraph{Step}
+Next we have \code{Step<T>} which stands for automaton transition.
+It has its \code{source} and \code{destination} states references.
+Symbol accepted by using this transition is store in \code{acceptsSymbol} member of generic type \code{T}.
+And finally member \code{useCount}, which is integer stating how many times the transition was used when constructing prefix tree automaton from input data.
+Simplifying algorithm can use this number for statistic purposes.
+
+\paragraph{Automaton}
+\code{Automaton<T>} class puts theese together into nondeterministic finite automaton.
+It has reference to \code{initialState}, it has \code{newStateName} integer value to assure unique state names inside one automaton (incrementing every time new state is created).
+We use two maps to implement transition function ($\delta$-function).
+One map of type \code{Map<State<T>, Set<Step<T>{}>{}>} called \code{delta} represents mapping from state into set of all outgoing transitions from state.
+Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better perfomance).
+There exist only one instance of each step from one state to another.
+That instance is referenced in delta map (on place of source state), and in reverse delta map (on place of destination state).
+Loops are no speciality, just source = destination.
+
+Automaton supports creating automaton as a copy of another one (not reference copy, but true copy), this can be used when searching solution space to create more versions of automaton to edit.
+We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
+Create empty automaton and then call this method for every input string of language.
+You will get PTA with useCounts and finalCounts set properly on steps/states.
+
+\begin{figure}
+	\centering
+	\subfigure[Before merge.]
+	{
+		\includegraphics[scale=\myscale]{automaton_merge1}
+		\label{automaton_merge1}
+	}
+	\subfigure[After states 3 and 4 merged.]
+	{
+		\centering\includegraphics[scale=\myscale]{automaton_merge2}
+		\label{automaton_merge2}
+	}
+	\subfigure[After states 2 and 4 merged.]
+	{
+		\centering\includegraphics[scale=\myscale]{automaton_merge3}
+		\label{automaton_merge3}
+	}
+	\caption{Sample automaton.} \label{automaton_merge}
+\end{figure}
+
+The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates(state1, state2)} method.
+Method merges second state given into first one (or an overloaded version - all states in list into first one in list).
+All \{in|out\}-transitions are redirected properly (or discarted as needed).
+Variables useCount and finalCount are updated to sums of values from merged transitions/states.
+One can ask for merging states, that are merged out from automaton already.
+Lets take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
+One asks to merge states 3 and 4.
+State 3 then becomes final state with loop and state 4 disappears (fig. \ref{automaton_merge2}).
+If then one ask to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
+This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).
 
@@
 	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
+\end{itemize}
 All of these metadata are of set/not set character (using \code{Boolean.TRUE}).
-\end{itemize}
 
 % TODO vektor Describe requirements on what gets passed on between modules
-
-\subsection{Nondeterministic Finite Automaton}
-For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
-Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
-Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.
-
-Let's begin by smallest of the classes, the \code{State<T>}.
-It represents automaton state, it has two integer members: \code{name} and \code{finalCount}.
-Name clearly serves as name of state in visualization and to string conversion.
-Final count is for representing whether the state is final in automaton.
-The field is not true/false but integral to help algorithms use stastistics over automatons (how many times in XML input is this state final?).
-
-Next we have \code{Step<T>} which stands for automaton transition.
-It has its \code{source} and \code{destination} states references.
-Symbol accepted by using this transition is store in \code{acceptsSymbol} member of generic type \code{T}.
-And finally member \code{useCount}, which is integer stating how many times the transition was used when constructing prefix tree automaton from input data.
-Simplifying algorithm can use this number for statistic purposes.
-
-\code{Automaton<T>} class puts theese together into nondeterministic finite automaton.
-It has reference to \code{initialState}, it has \code{newStateName} integer value to assure unique state names inside one automaton (incrementing every time new state creates).
-We use two maps to implement transition function (delta($\delta$) function  in most literature).
-One map of type \code{Map<State<T>, Set<Step<T>>>} called \code{delta} represents mapping from state into set of all outgoing transitions from state.
-Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better perfomance).
-There exist only one instance of each step from one state to another.
-That instance is referenced in delta map (on place of source state), and in reverse delta map (on place of destination state).
-Loops are no speciality, just source = destination.
-
-Automaton supports creating automaton as a copy of another one (not reference copy, but true copy), this can be used when searching solution space to create more versions of automaton to edit.
-We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
-Create empty automaton and then call this method for every positive example input string of language.
-You will get PTA with useCounts and finalCounts set properly on steps/states.
-
-\begin{figure}
-	\centering
-	\subfigure[Before merge.]
-	{
-		\includegraphics[scale=\myscale]{automaton_merge1}
-		\label{automaton_merge1}
-	}
-	\subfigure[After states 3 and 4 merged.]
-	{
-		\centering\includegraphics[scale=\myscale]{automaton_merge2}
-		\label{automaton_merge2}
-	}
-	\subfigure[After states 2 and 4 merged.]
-	{
-		\centering\includegraphics[scale=\myscale]{automaton_merge3}
-		\label{automaton_merge3}
-	}
-	\caption{Sample automaton.} \label{automaton_merge}
-\end{figure}
-
-The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates()} method.
-Method merges second state given into first one (or all states in list into first one in list).
-All \{in | out\}-transitions are redirected properly (or discarted as needed).
-Use count and final count are updated as sums of values from merging transitions/states.
-One can ask for merging states, that are merged out from automaton already.
-Take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
-One asks to merge states 3 and 4.
-State 3 then becomes final state with loop and state 4 disappears (fig. \ref{automaton_merge2}).
-If then one ask to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
-This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).
 
@@
 The most important method in each module is \code{start}, defined in each of the interfaces.
-This method is called by the framework when the respective step of inference is being executed.
+This method is called by the framework when the respective step of inference is to be executed.
 It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished.
-We will look at both parameter now in more detail.
+We will look at both parameters now in more detail.
 
@@
 
-\emph{Initial Grammar Generator} takes an object of type \emph{Input}.
+\jmodule{Initial Grammar Generator} takes an object of type \emph{Input}.
 This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}.
@@
 % TODO vektor Talk about Capabilities
+\section{NetBeans Modules}
+
+TODO vektor
 

### 2011-01-13 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - picture for vektor runner

@@
 
-\jmodule{Initial Grammar Generator} takes an object of type \emph{Input}.
+\jmodule{Initial Grammar Generator} takes an object of type \code{Input}.
 This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}.
@@
 
-\emph{Simplifier} and \emph{Schema Generator} take grammar, % TODO vektor Link to grammar definition
-in other words a list of rules as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
+\jmodule{Simplifier} and \jmodule{Schema Generator} take grammar, % TODO vektor Link to grammar definition
+in other words a list of \code{Element}s as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
 
@@
 As their respective interfaces, also callbacks define one crucial method: \emph{finished}.
-Each inference module is responsible for invoking this method on the callback it got as a parameter after it has finished its work and has results to be passed on.
+Each inference module is responsible for invoking this method on the callback it got as a parameter, after it has finished its work and has results to be passed on.
 Again, these 3 finished methods have different arguments based on the inference module.\\
@@
 
-\code{SchemaGeneratorCallback.finished} has two \code{String}s as arguments: \code{schema} is the actual string representation of the resulting schema, \code{extension} is a file extension of the result (such as "dtd" or "xsd") which the framework will use when saving the result in a file.
+\code{SchemaGeneratorCallback.finished()} has two \code{String}s as arguments: \code{schema} is the actual string representation of the resulting schema, \code{extension} is a file extension of the result (such as "dtd" or "xsd") which the framework will use when saving the result in a file.
 
@@
 
-User running the inference might change his mind and try to stop this. For this reason, modules have to check for this case in every time-consuming place such as long loops with the following code
+User running the inference might change his mind and try to stop this. For this reason, modules have to check for this case in every time-consuming place such as long loops with the following code:
 \begin{verbatim}
@@
 % TODO vektor Picture!!!
+\begin{figure}
+\centering\includegraphics[scale=\myscale]{runner}
+\caption{Runner} \label{runner}
+\end{figure}
+
 \begin{enumerate}

### 2011-01-14 doc/jInfer/building-jinfer.tex
reorganizing documentation

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{epsfig}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{import}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{26cm}
\setlength{\textwidth}{18cm}

\def\mftitle{Building jInfer}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle

Target audience: jInfer power users. Anyone who needs to build jInfer from sources.\\

\textit{This document assumes that your computer is set to use jInfer, meaning you have correct version of NetBeans installed and you can run jInfer.}

\section{Getting jInfer sources}
There are two main ways of obtaining jInfer source files: from a version distribution, or from the Subversion repository. The first way might be more convenient and secure (version distributions are usually more stable than the bleeding edge code in repository), however you might prefer to build the latest code.\\

The first way, obtaining sources from version distribution:
\begin{enumerate}
	\item Go to https://sourceforge.net/projects/jinfer/files/
	\item Download the latest jInfer sources (jInfer-X.Y-src.zip).
	\item Unpack the ZIP, you’ll get a few directories. One of them will be \texttt{src}.
	\item Continue by building jInfer.
\end{enumerate}

Second way, checking out Subversion repository: you will need a Subversion client, such as the command line client or TortoiseSVN. In any case, http://subversion.apache.org/ might be a good place to start. To check out jInfer, use the following URL:\\

\texttt{https://jinfer.svn.sourceforge.net/svnroot/jinfer}

After the checkout, continue by building jInfer.

\section{Building from NetBeans}
\textit{This is perhaps the easier, more user friendly way of building jInfer. If you however wish to automate the building process, you might want to try using Ant - in that case, see the next section.}

\begin{enumerate}
	\item Start NetBeans, select \textit{File} > \textit{Open Project...} from the main menu.
	\item In the dialog window, navigate to the root folder with jInfer sources. Its icon should look like two brown puzzle pieces.
	\item You should see the name "jInfer" in the dialog window.
	\item Make sure the \textit{Open Required Projects} check box is checked.
	\item You might want checking the \textit{Open as Main Project} checkbox.
	\item Click \textit{Open Project}. After a while, all jInfer projects should open and be visible in the \textit{Projects} window.
\end{enumerate}

At this moment, you might want to look around, play with the code a bit or apply some patches. If you want to try modified jInfer in a running NetBeans, select \textit{Run} from project jInfer (two brown puzzle pieces icon) context menu. If jInfer is the main project, clicking the green "Play" icon in the toolbar should work too.\\

If you want to build jInfer, for example to test if your changes didn't break something, select \textit{Build} from the context menu. Sometimes it might be necessary to run \textit{Clean \& Build}, try this if it "stops working". If you want to run JUnit test, select \textit{Test} from the same context menu.\\

If you now want to create your own distribution (= NBM modules), select \textit{Create NBMs} from the context menu. They will be created in \textit{build/updates} directory.

\section{Building with Ant}

Each jInfer module and jInfer suite itself has its own \texttt{build.xml} Ant build file. The most important one, for the whole jInfer suite, is located in the \texttt{src} folder.\\

It is possible to use vanilla Ant from Apache, but perhaps safer to resort to the Ant bundled with NetBeans. This one should be located in \textit{NetBeans install directory} > \texttt{java/ant/bin/ant}.\\

Whichever Ant is chosen, to build jInfer do the following on command line:
\begin{enumerate}
	\item Change directory to jInfer's \texttt{src}.
	\item Run \texttt{ant [chosen target]}. Interesting targets include:
	\begin{itemize}
		\item \texttt{clean}: removes any build artefacts.
		\item \texttt{build}: compiles and builds jInfer.
		\item \texttt{clean build}: performs a clean build, useful if something "stops working".
		\item \texttt{nbms}: creates a distribution by packaging jInfer modules in NBM files.
	\end{itemize}
\end{enumerate}

Chosen target may be empty, which should default to \texttt{build}.

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-14 doc/jInfer/user-tutorial.tex
reorganizing documentation

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{epsfig}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{import}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{26cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer User Tutorial}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle

Target audience: jInfer users. Anyone who needs to use jInfer to create DTD, XSD or another schema from existing XML documents, schemas (again: DTD, XSD, ...) or queries (such as XPath).\\

\textit{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\

\section{Overview}
\begin{enumerate}
	\item Get \& install NetBeans.
	\item Get \& install jInfer.
	\item Infer schema.
\end{enumerate}

\section{NetBeans}
\textit{Do this if you don’t have any NetBeans installed. If you have an older version than 6.9, jInfer will not run - please update it first.}
\begin{enumerate}
	\item Go to http://netbeans.org/
	\item Download NetBeans 6.9+, whichever bundle will do.
	\item Follow installation instructions.
\end{enumerate}

\section{jInfer}
\textit{Do this if you have the correct version of NetBeans installed. Note that jInfer features automatic update - you don’t need to follow these steps each time new version comes out.}

\begin{enumerate}
	\item Go to https://sourceforge.net/projects/jinfer/files/
	\item Download the latest jInfer binaries (jInfer-X.Y-bin.zip).
	\item Unpack the ZIP, you’ll get a few NBM files.
	\item Start NetBeans, open \textit{Tools} > \textit{Plugins} from the main menu.
	\item Switch to \textit{Downloaded} tab.
	\item Click \textit{Add Plugins...}.
	\item Select all the unpacked NBM files.
	\item Click \textit{Install} and follow the instructions.
	\item Restart NetBeans.
\end{enumerate}

\section{Basic Inference}
\textit{This section assumes you have NetBeans and jInfer installed. After the first run, a jInfer Welcome window will appear, guiding you through basically the same steps that will be covered here. You might want to follow them instead.}\\

Let’s assume that you have one or more XML files, and you want to create XSD schema from them.\\

\begin{enumerate}
	\item Create a new jInfer project.
	\begin{enumerate}
		\item \textit{File} > \textit{New Project} \textbf{or} \textit{New Project} icon from the toolbar \textbf{or} right-click the area in \textit{Projects} window > \textit{New Project}.
		\item Select \textit{jInfer} category, \textit{jInfer Project}.
		\item Click \textit{Next}.
		\item \label{project-location} Choose a name (for example MyInference) and location of your project.
		\item Click \textit{Next}.
		\item In the \textit{Schema Generator} combo box select \textit{Basic XSD Exporter}.
		\item Click \textit{Finish}.
	\end{enumerate}	
	\item Add input XML files to this project.
	\begin{enumerate}
		\item Right click the newly created project in \textit{Projects} window.
		\item Select \textit{Add files}.
		\item Select XML files you want to use for schema inference.
		\item If you now open the \textit{XML} folder in your project, you will see the added files.
	\end{enumerate}	
	\item Run inference.
	\begin{enumerate}
		\item \textit{Run Project} in the toolbar (green play icon) \textbf{or} right click the project, select \textit{Run}.
		\item Generated schema will open in a new \textit{Editor} window. It will also appear in the \textit{Output} folder.
		\item You will find the file with the schema in your project (location from step \ref{project-location}, folder \textit{output}).
	\end{enumerate}
\end{enumerate}

\section{Project files}
\textit{This section assumes you have successfully inferred at least one schema.}
\begin{itemize}
	\item To add input files to a specific folder (XML documents, schemas, queries) without relying on jInfer to guess the folder from the extension, right click the folder and select \textit{Add XYZ files}.
	\item Note that input folders are virtual - files in them still reside at their original locations. You can’t actually delete them from disk from within jInfer, just remove them from these virtual folders by selecting \textit{Delete} from their context menu.
	\item \textit{Output} folder is, on the other hand, a faithful representation of the folder \textit{output} inside your jInfer project (step \ref{project-location}). If you delete a schema here, it will be deleted from the disk too.
	\item You might want to see a diff between two input documents, or old and inferred schema - just select both files at once, right-click and select \textit{Tools} > \textit{Diff} from their context menu.
\end{itemize}

\section{Setting up jInfer}
\textit{This section assumes that you have successfully inferred at least one schema and know how to manipulate input and output files in a project.}\\

To change jInfer settings, you have to first realize whether it is a change affecting a single project at a time (for example, changing the output language from XSD to DTD) or whole jInfer at once (whether schema should open in Editor window after the inference finishes).
\begin{enumerate}
	\item First type of settings is found in so called Project Preferences. You can access them from the respective jInfer project’s context menu: right click the project and select \textit{Properties}.
	\item Second type is called jInfer Options and can be found among NetBeans’ options: open them (Windows, Linux: \textit{Tools} > \textit{Preferences...}, Mac: \textit{NetBeans} > \textit{Preferences...}) and switch to \textit{jInfer} tab.
\end{enumerate}

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-14 doc/jInfer/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - automaton cloner

--- 2011-01-13 doc/jInfer/full/architecture/architecture.tex
@@
 \begin{figure}
-\centering\includegraphics[scale=0.3]{nodes_full}
+\centering\includegraphics[scale=\myscale]{nodes_full}
 \caption{XML representing interfaces and classes in detail} \label{nodes_full}
@@
 Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
-Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.
+Whole implementation uses java generics for representation of symbol of alphabet.
+We denote \code{T} the java type of symbol.
+Take care of symbol class and its \code{equals()} implementation.
+Automaton uses \code{equals()} to compare symbols on transitions (when building prefix-tree automaton and when merging states).
+If you are using strings, you're just fine, but with complicated objects, either take care, that equivalent objects are properly tested in \code{equals()}.
+Or (maybe faster) solution is to cluster objects into classes of equivalence before inserting them into automaton.
+Then give automaton only cluster representant object (which will use java \code{Object.equals()} with reference comparison) for each object encountered.
 
@@
 It has its \code{source} and \code{destination} states references.
-Symbol accepted by using this transition is store in \code{acceptsSymbol} member of generic type \code{T}.
+Symbol accepted by using this transition is stored in \code{acceptsSymbol} member of generic type \code{T}.
 And finally member \code{useCount}, which is integer stating how many times the transition was used when constructing prefix tree automaton from input data.
@@
 
-Automaton supports creating automaton as a copy of another one (not reference copy, but true copy), this can be used when searching solution space to create more versions of automaton to edit.
+Automaton supports creating of automaton as a copy of another one (not reference copy, but deep copy expect of symbols), this can be used when searching solution space to create more versions of automaton to edit.
 We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
@@
 \end{figure}
-
 The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates(state1, state2)} method.
@@
 This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).
+
+\paragraph{AutomatonCloner}
+Class \code{AutomatonCloner<A, B>} has one overloaded method \code{convertAutomaton()}.
+First version accepts automaton and class implementing \code{AutomatonClonerSymbolConverter<A, B>} interface, and returns new automaton with same structure, but with symbols on transitions from alphabet of java type \code{B}.
+Second version takes two automatons, second one has to be empty (without initial state created), and symbol converter.
+It fills in second automaton to have same structure as first one, but with symbols of type \code{B}.
+
+To fulfill symbol conversion, one have to provide imlementation of \code{AutomatonClonerSymbolConverter<A, B>} interface.
+Interface has one method:
+\begin{verbatim}
+	B convertSymbol(final A symbol);
+\end{verbatim}
+Its purpose is to give mapping from symbol over one domain to new symbols.
+Implementation have to be \code{equals} consistent, if in first automaton are two transitions with symbols \code{a}, \code{b} such that
+\code{a.equals(b) == true}, convertor has to produce new symbols that are equal.
 

### 2011-01-14 doc/jInfer/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - example of inference proces

@@
 	For example, a trivial XML document
+	\label{example_inference_xml}
 	\begin{verbatim}
-  <a>
-    <b>
-      <c/>
-    </b>
-    <d/>
-  </a>
+  <person name="john" surname="smith">
+    <info>
+      Some text
+      <note/>
+    </info>
+    <more/>
+  </person>
+  <person>
+    <more/>
+    <more/>
+    <more/>
+  </person>
 	\end{verbatim}
-	would translate into the following IG rules
+	will translate into the following IG rules
 	\begin{eqnarray*}
-		a & \to &  b, d \\
-		b & \to & c \\
-		c & \to & \lambda \\
-		d & \to & \lambda 
+		person & \to & \mathit{info}, more, more, more, more \\
+		\mathit{info} & \to & simple\_data, note \\
+		note & \to & empty\_concatenation \\
+		more & \to & empty\_concatenation \\
+		person & \to & more, more, more \\
+		more & \to & empty\_concatenation \\
+		more & \to & empty\_concatenation \\
+		more & \to & empty\_concatenation \\
 	\end{eqnarray*}
@@
 At the end of this step, all rules are sent to the export step.\\
-	For example, rules
+	For example, previous rules for element person could be simplified to a single rule
 	\begin{eqnarray*}
-		a & \to & b \\
-		a & \to & c, d, d, d
+		person & \to & \mathit{info}?, more\{1, 3\}
 	\end{eqnarray*}
-	could be simplified to a single rule
+	Rules for elements info, more and note after simplification will be:
 	\begin{eqnarray*}
-		a & \to & b | (c, d\ast)
+		\mathit{info} & \to & simple\_data, note \\
+		note & \to & \lambda \\
+		more & \to & \lambda
 	\end{eqnarray*}
+	Note the lambda regular expressions for note and more.
+In Initial Grammar, all regexps are concatenations (even empty), but in simplified grammar, if element have to be empty in schema, it has to have lambda regular expresion as subnodes.
 	\item Schema export: done by the \jmodule{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema file from the simplified rules.
 Result of this step is a string representation of the schema, which is sent back to the framework (and later displayed, saved, etc).\\
-	% TODO vektor Example
+	For previous simplifier rules, the resulting DTD would be:
+	\begin{verbatim}
+	<!ELEMENT person (info?, more, more?, more?)>
+	<!ELEMENT info (#PCDATA | note)*>
+	<!ELEMENT note EMPTY>
+	<!ELEMENT more EMPTY>
+	+ attributes
+	\end{verbatim}
+	For element person, even when simplified grammar specifies its occurence to at least once, at most 3 times, as DTD has no such construct, export module have to do some magic.
+Situation is even worse for elements that contain simple data inside simplified rule.
+Only way to express mixed content in DTD is to use \verb.(#PCDATA | note)*.
+construct.
+Even if regular expression is complicated, export module has to do this "flattening".
+	Lets look on XSD output of same rules:
+	% TODO rio review this example!
+	\begin{verbatim}
+	<xs:element name="person">
+	  <xs:complexType>
+	    <xs:sequence>
+	      <xs:element name="info" minOccurs="0" maxOccurs="1" type="xs:string"/>
+	      <xs:element name="more" minOccurs="1" maxOccurs="3" type="xs:string"/>
+	    </xs:sequence>
+	  </xs:complexType>
+	</xs:element>
+	<xs:element name="info">
+	  <xs:complexType mixed="true">
+	    <xs:sequence>
+	      <xs:element name="note" minOccurs="1" maxOccurs="1" type="xs:string"/>
+	    </xs:sequence>
+	  </xs:complexType>
+	</xs:element>
+	<xs:element name="note">
+	  <xs:complexType>
+	  </xs:complexType>
+	</xs:element>
+	<xs:element name="more">
+	  <xs:complexType>
+	  </xs:complexType>
+	</xs:element>
+	+attributes
+	\end{verbatim}
 \end{enumerate}
@@
 
-% anti : should be after IGG explained, to fulfil is.sentinel explanation i will refer back.
 \subsection{Node metadata} \label{section_metadata}
@@
 
-% TODO vektor Picture!!!
 \begin{figure}

### 2011-01-15 doc/jInfer/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - example of inference process - corrected as rio said

@@
 	    <xs:sequence>
-	      <xs:element name="info" minOccurs="0" maxOccurs="1" type="xs:string"/>
-	      <xs:element name="more" minOccurs="1" maxOccurs="3" type="xs:string"/>
+	      <xs:element name="info" minOccurs="0" maxOccurs="1">
+	        <xs:complexType mixed="true">
+	          <xs:sequence>
+	            <xs:element name="note" minOccurs="1" maxOccurs="1">
+	              <xs:complexType>
+	              </xs:complexType>
+	            </xs:element>
 	          </xs:sequence>
@@
 	      </xs:element>
-	<xs:element name="info">
-	  <xs:complexType mixed="true">
-	    <xs:sequence>
-	      <xs:element name="note" minOccurs="1" maxOccurs="1" type="xs:string"/>
-	    </xs:sequence>
-	  </xs:complexType>
-	</xs:element>
-	<xs:element name="note">
+	      <xs:element name="more" minOccurs="1" maxOccurs="3"/>
 	        <xs:complexType>
@@
 	      </xs:element>
-	<xs:element name="more">
-	  <xs:complexType>
+	    </xs:sequence>
 	  </xs:complexType>

### 2011-01-15 doc/jInfer/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - talked about grammar interfaces - igg > simplifier, two formats determined by capability and simplifier > export, that is just one sentence mention:). but i think, this document is nearly completed

@@
 \begin{itemize}
+	\item \code{from.xml}, filled in by BasicIGG module: means that this rule was created (originates from) from a XML document.
 	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
@@
 \end{itemize}
-All of these metadata are of set/not set character (using \code{Boolean.TRUE}).
+All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively mean true).
 
@@
 
+\jmodule{Simplifier} and \jmodule{Schema Generator} take grammar, % TODO vektor Link to grammar definition
+in other words a list of \code{Element}s as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
+
 \jmodule{Initial Grammar Generator} takes an object of type \code{Input}.
@@
 Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.\\
+\paragraph{Initial Grammar} has two formats:
+\begin{itemize}
+	\item Only concatenations. This is default format. All regular expressions coming from IGG have to be concatenations, even empty.
+	When importing XML documents, this is just fine, as document is positive example from language we are trying to identify.
+	But even when importing schemas, queries or other stuff that can contain more information, one has to convert it somehow to concatenations.
+	For example, when importing DTD:
+	\begin{verbatim}
+	<!ELEMENT person (info?, more, more?, more?)>
+	<!ELEMENT info (#PCDATA | note)*>
+	<!ELEMENT note EMPTY>
+	<!ELEMENT more EMPTY>
+	\end{verbatim}
+	Elements note and more are empty, but we cannot use lambda regexp. Import will use empty concatenation instead.
+	For element person, import module have to generate some positive examples of language generated by regular expression \verb+(info?, more, more?, more?)+.
+This generating is basicly done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{expand()}.
+One can implement own intelligent expander, or use our provided \code{cz.cuni.mff.ksi.jinfer.basicigg.expansion.ExpanderImpl}.
 
-\jmodule{Simplifier} and \jmodule{Schema Generator} take grammar, % TODO vektor Link to grammar definition
-in other words a list of \code{Element}s as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
+	\item Complex regular expressions. Simplifier module running in current inference can indicate support for this format by having capability called \code{can.handle.complex.regexps}.
+	Even then, import module is not obliged to use it of course. But it is reasonable to do so, if imported file type gives more information than simple concatenations (when importing schemas).
+	Initial Grammar can contain all regular expressions.
+If schema defines empty element, make its subnodes lambda regexp.
+If schema defines element content as \verb.(a, b){2,7}(c | d)., just construct that regexp.
+Empty concatenations are also allowed - as they are positive examples in XML document input.
+Its simplifier business to divide rules by import file formats (by using metadata described in \ref{section_metada}) and to handle empty concatenations from XML file import with complex regular expressions from other formats.
+\end{itemize}
+If you are writing import module, check if you can provide more information to simplifier from input file type by using complex format.
+This interface division allows us to plug in old methods (accepting only XML input) to the chain of inference, even with schema input files.
+It may be useful for method benchmarking too.
+
+\paragraph{Simplified grammar} has only one format and that is strict regular expression format.
+Empty concatenations, alternations or permutations are not allowed.
+If element should be exported as empty, use lambda regexp as its subnodes.
 
@@
 
-TODO vektor
+TODO vektor capabilities too
+TODO vektor anti thinks, we should talk about properties windows and plugability of autoeditor, or thats part of some other developer doc?
 

### 2011-01-16 doc/jInfer/architecture/architecture.tex
Documentation: Capabilities

@@
 
-% TODO vektor Talk about Capabilities
+\subsection{Capabilities} \label{section_capabilities}
+From time to time, modules need a flow of certain information in completely opposite way to the usual inference: they need to know what the following module can or cannot do.\\
+
+For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCapabilities()}.
+This method returns a list of strings representing the abilities, or "capabilities" of this module.\\
+
+An inference module can query the capabilities of the \textit{next} module in the inference chain by invoking the \code{RunningProject.getNextModuleCaps()} method to obtain the next module encapsulated in the \code{Capabilities} interface.\\
+
+At the moment, only one capability is defined to be communicated across modules, more specifically between \jmodule{IGGenerator} and \jmodule{Simplifier}.
+This is \code{Capabilities.CAN\_HANDLE\_COMPLEX\_REGEXPS} meaning whether the \code{Simplifier} accepts more complicated regular expressions than simple concatenations of tokens on its input.\\
+% TODO vektor check later - didn't this change?!?
+
+Example of Capabilities usage taken from \jmodule{XSDImporter} module:
+\begin{verbatim}
+  // if the next module cannot handle complex regexps, help it by expanding our result
+  if (!RunningProject.getNextModuleCaps().getCapabilities()
+               .contains(Capabilities.CAN_HANDLE_COMPLEX_REGEXPS)) { 
+    // lookup expander
+    final Expander expander = Lookup.getDefault().lookup(Expander.class);
+    // return expanded
+    return expander.expand(parser.getRules());
+  }
+  // return not expanded rules
+  return parser.getRules();
+\end{verbatim}
+
 \section{NetBeans Modules}
 
-TODO vektor capabilities too
-TODO vektor anti thinks, we should talk about properties windows and plugability of autoeditor, or thats part of some other developer doc?
+% TODO vektor anti thinks, we should talk about properties windows and plugability of autoeditor, or thats part of some other developer doc?
+% property windows, autoeditor and ruledisplayer will be handled in dev tutorial -- vektor
 

### 2011-01-16 doc/jInfer/architecture/architecture.tex
Documentation: NB modules.

@@
 All packages start with \code{cz.cuni.mff.ksi.jinfer}.
-Afterwards is the short, normalized name of the module (e.g.
+Afterwards is the short, normalized name of the NetBeans module (e.g.
 \code{base}) and finally the package structure in this module (e.g.
@@
 
-% TODO vektor Connection to NBMs
-
 Annotation required for the framework to recognize such a class as an inference module is the following
@@
 
-% TODO vektor anti thinks, we should talk about properties windows and plugability of autoeditor, or thats part of some other developer doc?
-% property windows, autoeditor and ruledisplayer will be handled in dev tutorial -- vektor
+jInfer, being a NetBeans Platform application, is divided into multiple \textit{NetBeans Modules}.
+It is very important to understand that this division is in principle orthogonal to that naturally sketched in the previous chapters - theoretically, the whole jInfer could be contained in a single NB module, because the framework uses Lookups to locate logical modules.
+However, for the sake of organization, jInfer is split into several modules and most of the time there is a strong correlation between logical units of jInfer and its NB modules.\\
+
+A short overview of these modules and their functions follow. To get a deeper understanding of any of them, please refer to its documentation.\\
+
+Base modules:
+
+\begin{itemize}
+	\item \jmodule{jInfer}: not a NBP module in fact, but rather a module suite - the ,,root" NB project for the whole framework.
+	\item \jmodule{Base}: contains common data structures, interfaces and utility logic shared across all other modules.
+\end{itemize}
+
+Inference modules:
+
+\begin{itemize}
+	\item \jmodule{Runner}: as described in the section TODO anti link, this module encapsulates the inference process (including the fact that inference consists of 3 phases as described before).
+Most important class: \code{Runner}.
+	\item \jmodule{BasicIGG}: this module contains an extensible implementation of \jmodule{IGGenerator}.
+Basic version handles all XML-typed input documents, DTD schemas and XPath queries.
+To find out more about the extensibility to handle other outputs, refer to \jmodule{BasicIGG}'s documentation.
+	\item \jmodule{XSDImport}: demonstrating \jmodule{BasicIGG}'s extensibility, this is a base module for XSD input.
+	\item \jmodule{XSDImportSAX}: XSD input handling using SAX approach.
+	\item \jmodule{XSDImportDOM}: XSD input handling using DOM approach.
+	\item \jmodule{TwoStepSimplifier}: a \jmodule{Simplifier} implementation. TODO anti
+	\item \jmodule{BasicDTDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to DTD.
+	\item \jmodule{BasicXSDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to XSD.
+\end{itemize}
+
+Utility modules:
+
+\begin{itemize}
+	\item \jmodule{RuleDisplayer}: module capable of displaying a grammar in a graphical way.
+	\item \jmodule{AdvancedRuleDisplayer}: more advanced version of the former using \jmodule{JUNG}'s graphing capabilities.
+	\item \jmodule{AutoEditor}: encapsulation of an interactive finite state automaton editor.
+	\item \jmodule{JUNG}: wrapper module for the JUNG TODO anti link graphing library.
+	\item \jmodule{Options}: NBP specific: this module adds a jInfer category in NetBeans Options window.
+	\item \jmodule{ProjectType}: NBP specific: this module defines a jInfer project type, along with its input and output files, settings etc.
+	\item \jmodule{XPathFileType}: NBP specific: this module teaches NetBeans to recognize a \code{.xpath} file extension.
+\end{itemize}
+
+TODO vektor Lookups
+
+% TODO module dependency hierarchy
+% TODO picture showing out inference modules in a humble graph
+
+To get deeper understanding of NBP modules, their versions, dependencies etc, refer to http://bits.netbeans.org/dev/javadoc/org-openide-modules/org/openide/modules/doc-files/api.html . TODO anti Cite
 

### 2011-01-16 doc/jInfer/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - suggesting changes

@@
 construct.
-Even if regular expression is complicated, export module has to do this "flattening".
+Even if regular expression is complicated, export module has to do this "flattening".texmaker
 	Lets look on XSD output of same rules:
@@
 
-\subsubsection{Runner}
+\subsubsection{Runner} \label{section_runner}
 
@@
 This method returns a list of strings representing the abilities, or "capabilities" of this module.\\
-
-An inference module can query the capabilities of the \textit{next} module in the inference chain by invoking the \code{RunningProject.getNextModuleCaps()} method to obtain the next module encapsulated in the \code{Capabilities} interface.\\
+An inference module can query the capabilities of the \emph{next} module in the inference chain by invoking the 
+\begin{verbatim}
+RunningProject.getNextModuleCaps()
+\end{verbatim}
+method to obtain the next module encapsulated in the \code{Capabilities} interface.\\
 
@@
 
-jInfer, being a NetBeans Platform application, is divided into multiple \textit{NetBeans Modules}.
+jInfer, being a NetBeans Platform application, is divided into multiple \emph{NetBeans Modules}.
 It is very important to understand that this division is in principle orthogonal to that naturally sketched in the previous chapters - theoretically, the whole jInfer could be contained in a single NB module, because the framework uses Lookups to locate logical modules.
@@
 \begin{itemize}
-	\item \jmodule{Runner}: as described in the section TODO anti link, this module encapsulates the inference process (including the fact that inference consists of 3 phases as described before).
+	\item \jmodule{Runner}: as described in the section \ref{section_runner}, this module encapsulates the inference process (including the fact that inference consists of 3 phases as described before).
 Most important class: \code{Runner}.
@@
 	\item \jmodule{XSDImport}: demonstrating \jmodule{BasicIGG}'s extensibility, this is a base module for XSD input.
-	\item \jmodule{XSDImportSAX}: XSD input handling using SAX approach.
-	\item \jmodule{XSDImportDOM}: XSD input handling using DOM approach.
-	\item \jmodule{TwoStepSimplifier}: a \jmodule{Simplifier} implementation. TODO anti
+	\item \jmodule{XSDImportSAX}: XSD input handling using SAX parser.
+	\item \jmodule{XSDImportDOM}: XSD input handling using DOM parser.
+	\item \jmodule{TwoStepSimplifier}: a \jmodule{Simplifier} implementation.
+Implements \cite{ahonen} merging state algorithm with basic handling of attributes.
+Read the separated documentation of module.
 	\item \jmodule{BasicDTDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to DTD.
@@
 	\item \jmodule{AutoEditor}: encapsulation of an interactive finite state automaton editor.
-	\item \jmodule{JUNG}: wrapper module for the JUNG TODO anti link graphing library.
+	\item \jmodule{JUNG}: wrapper module for the JUNG (see \cite{jung}).
 	\item \jmodule{Options}: NBP specific: this module adds a jInfer category in NetBeans Options window.
@@
 
-To get deeper understanding of NBP modules, their versions, dependencies etc, refer to http://bits.netbeans.org/dev/javadoc/org-openide-modules/org/openide/modules/doc-files/api.html . TODO anti Cite
+To get deeper understanding of NBP modules, their versions, dependencies etc, refer to \cite{modules_api}.
 

### 2011-01-19 doc/jInfer/architecture/architecture.tex
Documentation: some more work on Architecture.

@@
 
-The description of jInfer architecture will commence by describing the data
-structures, namely representations of regular expressions and XML elements,
-attributes and simple data.
+The description of jInfer architecture will be divided into 3 main views: the data view, the process view and the plaform view.\\
 
-Afterwards the interfaces of basic inference modules - \jmodule{Initial Grammar Generator}, \jmodule{Simplifier} and \jmodule{Schema Generator} - will be explained.
+The data view will commence by describing the data structures, namely representations of regular expressions and XML elements, attributes and simple data.
+The notions of a \emph{rule} and a \emph{grammar} will be explained, along with the finite state automata representation we use.\\
 
-Finally, the process of inference will be described.
+Afterwards the process of inference will be described, both from high-level and programmatic point of view.\\
+
+Finally, we will look at jInfer as at an NetBeans Platform application consisting of a number of modules performing their tasks and communicating together through a set of well-defined interfaces.\\
 
@@
 
-\section{Data structures}
+\section{Data view}
 \subsection{Regular expressions}
@@
 
-\subsection{Rules and grammars}
+\subsection{Rules and grammars} \label{rules_grammar}
 
@@
 
-\section{Inference process}
+\section{Process view} \label{section_inference_process}
 \begin{figure}
@@
 
-\subsection{Node metadata} \label{section_metadata}
-To allow simple extensibility of rules, class \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
-
-At the present, jInfer modules use the following metadata, all defined in \code{IGGUtils} class:
-\begin{itemize}
-	\item \code{from.xml}, filled in by BasicIGG module: means that this rule was created (originates from) from a XML document.
-	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
-	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
-	\item \code{is.sentinel}, filled in by  BasicIGG and XSDImport module: when importing XML document, one can build whole tree in memory (as in fig.
-\ref{xml_example}) and then construct list of rules from it (thus saving memory).
-With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import), but right side can be defined anywhere in source file.
-To save complicated loading of whole schema, searching and pairing elements thorough rules, nodes on right side of rule are created empty - holding only name of node.
-Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
-	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
-\end{itemize}
-All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively mean true).
-
-% TODO vektor Describe requirements on what gets passed on between modules
-
 \subsection{Programmatic view}
 
-From developer's point of view, inference modules are just properly annotated classes implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
+From developer's point of view, inference modules are just properly annotated classes (see \ref{section_lookups}) implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
 
@@
 
-\jmodule{Simplifier} and \jmodule{Schema Generator} take grammar, % TODO vektor Link to grammar definition
-in other words a list of \code{Element}s as input. In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
+\jmodule{Simplifier} and \jmodule{Schema Generator} take grammar (see \ref{rules_grammar}), in other words a list of \code{Element}s as input.
+In the first case, this grammar is the Initial Grammar, in the second case it is the simplified grammar.
 
@@
 Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.\\
+
 \paragraph{Initial Grammar} has two formats:
@@
 
+\subsection{Node metadata} \label{section_metadata} % TODO vektor Move after Programmatic view
+To allow simple extensibility of rules, class \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
+
+At the present, jInfer modules use the following metadata, all defined in \code{IGGUtils} class:
+\begin{itemize}
+	\item \code{from.xml}, filled in by BasicIGG module: means that this rule was created (originates from) from a XML document.
+	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
+	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
+	\item \code{is.sentinel}, filled in by  BasicIGG and XSDImport module: when importing XML document, one can build whole tree in memory (as in fig.
+\ref{xml_example}) and then construct list of rules from it (thus saving memory).
+With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import), but right side can be defined anywhere in source file.
+To save complicated loading of whole schema, searching and pairing elements thorough rules, nodes on right side of rule are created empty - holding only name of node.
+Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
+	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
+\end{itemize}
+All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively mean true).
+
 \subsection{Capabilities} \label{section_capabilities}
@@
 
-\section{NetBeans Modules}
+\section{Platform view}
 
@@
 
-A short overview of these modules and their functions follow. To get a deeper understanding of any of them, please refer to its documentation.\\
+A short overview of these modules and their functions follows.
+To get a deeper understanding of any of them, please refer to its documentation.
+To get deeper understanding of how NBP modules work, their versions, dependencies etc, refer to \cite{modules_api}.\\
 
@@
 \begin{itemize}
-	\item \jmodule{jInfer}: not a NBP module in fact, but rather a module suite - the ,,root" NB project for the whole framework.
+	\item \jmodule{jInfer}: not a NBP module in fact, but rather a module suite - the ``root" NB project for the whole framework.
 	\item \jmodule{Base}: contains common data structures, interfaces and utility logic shared across all other modules.
@@
 \begin{itemize}
-	\item \jmodule{RuleDisplayer}: module capable of displaying a grammar in a graphical way.
-	\item \jmodule{AdvancedRuleDisplayer}: more advanced version of the former using \jmodule{JUNG}'s graphing capabilities.
+	\item \jmodule{BasicRuleDisplayer}: module capable of displaying a grammar in a graphical way.
+	\item \jmodule{TreeRuleDisplayer}: more advanced version of the former using \jmodule{JUNG}'s graphing capabilities.
 	\item \jmodule{AutoEditor}: encapsulation of an interactive finite state automaton editor.
@@
 
-TODO vektor Lookups
+The dependencies between all the modules are summarized in figure \ref{module_deps}.
+Figure \ref{modules_inference} shows modules implementing the 3 main inference interfaces described in \ref{section_inference_process}.\\
 
-% TODO module dependency hierarchy
-% TODO picture showing out inference modules in a humble graph
+\begin{figure}
+	\centering\includegraphics[scale=0.4]{module_deps}
+	\caption{Module dependencies} \label{module_deps}
+\end{figure}
 
-To get deeper understanding of NBP modules, their versions, dependencies etc, refer to \cite{modules_api}.
+\begin{figure}
+	\centering\includegraphics[scale=0.4]{modules_inference}
+	\caption{Inference modules in default jInfer} \label{modules_inference}
+\end{figure}
+
+\subsection{Lookups} \label{section_lookups}
+
+\emph{Lookups} are a part of NBP API, a way to achieve late-binding between code based on this platform.
+An important notion is the one of a \emph{service provider}.
+This is a class implementing an interface, annotated in the following way (shown on an example from \jmodule{BasicDTDExporter}).
+\begin{verbatim}
+  @ServiceProvider(service = SchemaGenerator.class)
+  public class SchemaGeneratorImpl implements SchemaGenerator {
+    ...
+\end{verbatim}
+The \code{@ServiceProvider} annotation tells NBP that \code{SchemaGeneratorImpl} is to be registered among all providers of the \code{SchemaGenerator} service (or interface).\\
+
+To look up a service provider, or all of them, one can use the \jmodule{Lookups} API in the following way.
+\begin{verbatim}
+  Collection<SchemaGenerator> generators = Lookup.getDefault().lookupAll(SchemaGenerator.class);
+
+  for (SchemaGenerator g : generators) {
+    System.out.println(g.getName());
+  }
+\end{verbatim}
+Call to \code{Lookup.getDefault().lookupAll()} will return a collection of all service providers registered to be providing the \code{SchemaGenerator} interface.
+This collection can be walked in some way, and on one or more of service providers methods can be called.
+We didn't have to actually bind providers to logic that uses their services, NB Platform took care of this for us.\\
+
+At this point it is important to understand that when using lookups this way, all service providers are kept as singletons.
+In case a provider needs to change its state during providing the service (and needs to be ``fresh" each time it is retrieved), it is best to convert this to a factory design pattern: the service provider registered is just a factory returning instances that do the actual work.\\
+
+TODO anti Link to http://bits.netbeans.org/dev/javadoc/org-openide-util-lookup/org/openide/util/Lookup.html as the official documentation to lookup API.
+
+\subsection{Tokens}
+
+TODO vektor
 

### 2011-01-19 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\emph{#1}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer TwoStep simplifier design and implementation}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
Target audience: developers willing to extend jInfer, scientist willing to implement own inference methods\\

\emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}

\section{Using of factory pattern}
TwoStep simplifier is divided into two submodules, which are then divided into submodules and so on.
Each submodule is simply class, properly anotated.
User selects proper classes - submodules in chain to work on inference.
Classes are then, in runtime looked up by using NetBeans lookup mechanism.
But NetBeans holds one instance of each class, that may be looked upon.
When user runs inference first time, classes are looked up and used.
Maybe class members are initialised, instances are in some state after inference completes.
Then, user clicks run button again, and same instances of classes are returned by lookups.
Ooops, theese are not freshly created instances, using them as they are may cause harm.
And problem arises - force each class (submodule) that is being looked up in inference process, to implement
some sort of cleanup method, that would restart it into fresh state and make it ready to use by another 
inference run? Or create new instances of theese classes in each inference run and make singleton classes (those which NetBeans return by lookups) their factory classes?

We decided to use second approach. 
It enables us not only to produce fresh instance in each inference run, but also factory classes implement obligatory
module methods such as \code{getName}, \code{getDescription} and so on, which are really same in each inference run.
Each submodule is then defined by (at least) two interfaces.
One is the factory interface, like this one:
\begin{verbatim}
public interface AutomatonSimplifierFactory extends NamedModule, Capabilities, UserModuleDescription {
  <T> AutomatonSimplifier<T> create();
}
\end{verbatim}
It extends \code{NamedModule} with all module methods.
It extends \code{Capabilities}, thus each module have to answer, if it has some capabilities.
And, in TwoStep we often extend \code{UserModuleDescription}, which defines method for obtaining description of the module, comprehensive to user (it is displayed in properties panels).
In some circumstances, it is useful to have method \code{create} generic.
The automaton simplifier works with automaton, which itself is generic too.
But simplifying does not depend on type of symbol of automaton, so interface \code{AutomatonSimplifier<T>}
is also generic as simplifier can simplify automaton of any symbol java type.
Factory interface deals with this by defining create method generic too.

Second interface is the one, which is returned by \code{create} method in factory.
That is the interface, which defines real work cycle with submodule (in example it is \code{AutomatonSimplifier<T>}).
We call this interface the worker interface of submodule.
Usage of this factory pattern follows the routine:
\begin{verbatim}
final Properties p = RunningProject.getActiveProjectProps(getName());

AutomatonSimplifierFactory f = ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
   p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER));

AutomatonSimplifier<AbstractStructuralNode> autSmp = f.<AbstractStructuralNode>create();
...
give some work to autSmp
\end{verbatim}
If our module has some submodules, we often implement lookups for submodules implementations in our own factory create method.
Worker class receives factories of all submodules it needs as a constructor parameters.

Lets look on \code{AutomatonMergingStateFactory}, that is factory of module, which has AutomatonSimplifier as submodule.
Its create method looks like this (shortened): 
\begin{verbatim}
@Override
public ClusterProcessor<AbstractStructuralNode> create() {
  LOG.debug("Creating new ClusterProcessorAutomatonMergingState.");
  return new AutomatonMergingState(getAutomatonSimplifierFactory(), getRegexpAutomatonSimplifierFactory());
}
\end{verbatim}
Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical, we show you the former one:
\begin{verbatim}
private AutomatonSimplifierFactory getAutomatonSimplifierFactory() {
  final Properties p = RunningProject.getActiveProjectProps(getName());

  return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER));
}
\end{verbatim}
Cluster processor AutomatonMergingState then receives factories for AutomatonSimplifier and RegexpAutomatonSipmlifier submodules in its constructor.
Cluster processor then may create as many instances of submodule classes as it needs (maybe simplifying more than one automaton).
Thorough this document, we will mention only worker interface when describing submodules, since all factory interfaces are designed same way just described.

\section{TwoStep module}
TwoStep simplifier is inspired by \cite{1802522} design. Inference proceeds in two steps:
\begin{enumerate}
	\item clustering of element instances into clusters of (probably) same elements
	\item infering regular expression for each element from examples of element contents taken from all elements in cluster
\end{enumerate}
Main routine of TwoStep simplifer does basically this:
\begin{verbatim}
    // 1. cluster elements
    final Clusterer<AbstractStructuralNode> clusterer= clustererFactory.create();
    clusterer.addAll(initialGrammar);
    clusterer.cluster();

    // 2. prepare emtpy final grammar
    final List<Element> finalGrammar= new LinkedList<Element>();

    // 3. process rules
    final ClusterProcessor<AbstractStructuralNode> processor= clusterProcessorFactory.create();
    for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
      final AbstractStructuralNode node =  processor.processCluster(clusterer, cluster.getMembers() - without sentinels);
      // 4. add to rules
      finalGrammar.add(
              new Element(node.getContext(),
              node.getName(),
              node.getMetadata(),
              ((Element) node).getSubnodes(),
              attList);
    }
\end{verbatim}
Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
Modules currently in jInfer are drawn on fig. \ref{twostep_modules}.
\begin{figure}
	\centering\includegraphics[scale=\myscale]{twostep_modules}
	\caption{Submodules of TwoStep simplifier} \label{twostep_modules}
\end{figure}
We provide one \jmodule{Clusterer} and four \jmodule{ClusterProcessor} implementations.
Each of those will be explained further in this document.

TODO sentinel processing

\subsection{Clusterer submodule}
Lets start by example, let input document(s) contain XML:
\begin{verbatim}
  <person name="john" surname="smith">
    <info>
      Some text
      <note/>
    </info>
  </person>
  <inform>
    Another text
    <note/>
  </inform>
  <person>
      <information>
        Some text
        <note/>
      </information>
  </person>
\end{verbatim}
We examine info-like named elements.
If we cluster elements sipmly by their name, we get one cluster with info element, another cluster with inform element and another one with information element.
As \cite{1802522} suggests, we should consider element content in clustering process.
If we take care of element content, elements info, inform and information would look same (in means of node tree inside them - text node and element note).
Elements information and info have even same context (inside person element).

We implement clustering in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
One cluster is represented by class \code{Cluster<T>} with \code{T} as type of clustered items.
This class simply holds java set of member of cluster and one of the references is held also in \code{representant} member.

We provide \code{Clusterer} interface for classes to implement.
Its purpose is to cluster bunch of elements (rules) on input, into  bunch of cluster class instances  (cluters).
It has methods \code{add()} and \code{addAll()} for adding items for clustering.
Center is method \code{cluster()}, which does the clustering itself.
As it may be time-consuming operation, method throws \code{InterruptedException}, implementation should take care 
of checking whether thread is user interrupted (see \cite{arch}).
After clustering, implementation should hold clusters in member, as it it will be further asked by calling method \code{T getRepresentantForItem(T item)}.
Given item to this method, one can ask for representant of cluster, to which the item belongs.
If no such cluster exists (item was not added for clustering before), we recommend you throwing an exception rather than returning \code{null}.
Missing will probably  indicate error in algorithm rather than normal workflow.
One can pull clusters from clusterer by calling \code{getClusters()} method.

Basic work usage of clusterer is:
\begin{verbatim}
Clusterer <T> c = new MyContextClusterer<T>();
c.addAll(initialGrammar);
c.cluster();
...
c.getClusters();
or
c.getRepresentantForItem(x);
\end{verbatim}

\subsubsection{ClustererWithAttributes extended interface}
Maybe you noticed that whole clusterer interface and cluster class are generic.
They may be used as design pattern not only for clustering elements in inference process.
To address clustering of elements in more detail, we created \code{ClustererWithAttributes<T, S>} interface, which extends \code{Clusterer<T>} interface.
It add method \code{List<Cluster<S>> getAttributeClusters(T representant)}, implying that each representant of
type \code{T} (that is representant of some main cluster) has some "attribute" clusters associated with it.
Theese clusters are obtained by calling \code{getAttributeClusters(x)}.
Attribute clusters are of type \code{S}.

Finally, we use this scheme to implement clusterer \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
For each cluster of elements, it creates list of clusters of attributes observed with all elements in main cluster.

All simple data are considered equal and given one dedicated cluster.
But in future implementations, extensibility is ensured to cluster simple data to obtain content models of elements.

All sentinel elements are clustered too.
TODO sentinels
TODO attributes clustering

\subsection{ClusterProcessor submodule}
Cluster processor takes rules of one cluster of element and somehow obtains regular expression for that set of elements. It returns rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for  resulting element) and with subnodes set to regular expression infered.
It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to resulting element.

Worker interface itself is defined as follows:
\begin{verbatim}
public interface ClusterProcessor<T> {
  T processCluster(
          final Clusterer<T> clusterer,
          final List<T> rules
          ) throws InterruptedException;
}
\end{verbatim}
Maybe you are asking, why cluster processor is given the clusterer instance.
Rules themself contain information about which elements to process, but clusterer has more information
about the topic.
Clusterer can tell you representant for any element in whole input (not only those elements in rules, but also those that may be on right side of rules), clusterer (if it is with attributes) has information about attributes <of each cluster.

We will now shortly describe each cluster processor implementation we've got.

\subsubsection{PassRepresentant}
Simple example to read. For each cluster return its representant as a rule to be in schema.
This has nothing to do with infering grammar, it is just proof of submodules concept.
Input documents are not valid against this odd grammar.
Do not use this in practice, just read the code to understand the bare minimum needed to implement submodule.

\subsubsection{Alternations}
This processor simply gets all right sides from elements in cluster, puts them in one big list and creates alternation regular expression with this list as children.
That is, it creates one big rule with alternation of every positive example observed.
No generalization is done at all.

\subsubsection{Trie}
TODO vektor describe trie cluster processor

\subsubsection{AutomatonMergingState}
AutomatonMergingState is implementation of merging state algorithm on nondeterministic finite automaton.
It creates prefix-tree automaton (PTA) from positive examples - right sides of rules given.
Then it calls its submodule called \jmodule{AutomatonSimplifier} to  modify PTA to some generalized 
automaton by merging states.

Simplified automaton is then converted to an instance of \code{RegexpAutomaton} by using clone constructor in regexp automaton.
Regexp automaton is automaton with regular expression as symbol on transitions.
In automata theory, such automaton is called extended NFA.
Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as content.

Regexp automaton is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
Its job is to derive regular expression from automaton, such that automaton and regular expression represents same language.

AutomatonMergingState has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by AutomatonMergingState.
It is at disposal for implementations of AutomatonSimplifier interface for testing, whether two states in automaton are equivalent and should be merged into one state.

AutomatonSimplifier implements solution searching logic in simplifying automaton by merging states.
We implement greedy strategy in class \code{Greedy}.
It simply asks given MergeConditionTester if it can merge any of automaton states and merges states until there are no states to be merged (on every pair of states asked, MergeConditionTester answers, they cannot be merged).
One can implement ACO heuristics or MDL principle heuristics as AutomatonSimplifier submodule using MergeConditionTesters provided.

Whole submodule structure of AutomatonMergingState cluster processor is drawn on fig. \ref{automaton_merging_state_submodules}.
\begin{figure}
	\centering\includegraphics[scale=\myscale]{automaton_merging_state_submodules}
	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_submodules}
\end{figure}
We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext}.
We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.

\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-20 doc/jInfer/architecture/architecture.tex
Documentation: finishing Architecture.

@@
 % TODO vektor check citation of http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf in literature.bib
+% TODO anti what do you mean, "check citation"?
 \emph{Rules} in such grammar are in the form
@@
 Initial Grammar is produced by \jmodule{Initial Grammar Generator}.
-% TODO vektor Link
 
@@
 
-\subsection{Node metadata} \label{section_metadata} % TODO vektor Move after Programmatic view
+\subsection{Node metadata} \label{section_metadata}
 To allow simple extensibility of rules, class \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
@@
 
-TODO vektor
+An important topic related to NB modules and their dependencies is a notion of \emph{tokens}.
+Imagine a situation where module \emph{M} needs \emph{someone} else to provide a certain service \emph{S}, but doesn't care who will that be.
+It cannot declare a direct dependency on a module containing this service provider, because it might not know such a module.
+Instead, \emph{M} declares a \emph{required token}.
+Any module providing \emph{S} will declare a \emph{provided token}, which is a simple string (usually fully qualified Java name of the interface representing \emph{S}).
+Later, NBP will make sure that \emph{M} can be installed in the platform only if there is at least one module providing that specific token.\\
+
+jInfer uses tokens in the following way: \jmodule{Runner}, as the module encapsulating the inference process, and \jmodule{ProjectType} as an encapsulation of jInfer projects both require the following 3 tokens.
+
+\begin{itemize}
+	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator}
+	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator}
+	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier}
+\end{itemize}
+
+All inference modules declare their respective token as provided.
+This way it is ensured that jInfer will run only if there is at least one implementation of each of the inference modules, regardless of the implementation itself.
 

### 2011-01-21 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - better modules images, reflecting cleaners and better layouts
ignores

@@
 \section{Using of factory pattern}
-TwoStep simplifier is divided into two submodules, which are then divided into submodules and so on.
+TwoStep simplifier is divided into three submodules, which are then divided into submodules and so on.
 Each submodule is simply class, properly anotated.
@@
 We will examine both of them.
+There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
 Modules currently in jInfer are drawn on fig. \ref{twostep_modules}.
@@
 \end{figure}
-We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext}.
+We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext}, which is used by \class{Greedy}
+to test mergability of states by default configuration.
+
+\subsubsubsection{StateRemoval}
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
+This is implemented in \class{StateRemoval} class.
+We defined one submodule of this class with interface called \code{Orderer}.
+It has only one method to implement: \code{getStateToRemove}.
+Given automaton it has to return reference to one state which should be removed from automaton at first.
+State removal calls this submodule and removes states given until there are only two states in automaton - superInitial and superFinal states,
+with exactly one transition.
+That transition has final regular expression on it as symbol, it is read and returned to AutomatonMergingState.
+We implement one orderer, called \code{Weighted}.
+It is simple heuristic - weights all states (weight = sum of {in | out | loop}-transition regular expression lengths) and returns state with lowest weight.
+
+\subsection{RegularExpressionCleaner module}
+Last we examine \code{RegularExpressionCleaner} interface.
+TODO anti cleaners
+
+Whole TwoStep submodules structure is on fig. \ref{twostep_modules_complete}.
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{twostep_modules_complete}
+	\caption{Modules of TwoStep simplifier and their submodules. Filled classes are default selection (best of).} \label{twostep_modules_complete}
+\end{figure}
 

### 2011-01-21 doc/jInfer/twostep/twostep.tex
Documentation: minor repairs in Architecture and TwoStep.

@@
 \subsubsection{Trie}
-TODO vektor describe trie cluster processor
+This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie") of them.
+More precisely, it takes the first rule and delares it to be a long branch (concatenation of tokens) in a newly created tree.
+After that, it adds the remaining rules one by one as branches like this: as long as it can follow an existing branch, it follows it.
+As soon as the newly added branch starts to differ, it ``branches off" (creates an alternation at that point) the existing tree and hangs the rest of the newly added rule there.
+Repeating this process creates a prefix tree describing all the rules in the cluster.
 

### 2011-01-21 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - middlework, added image for factory pattern

@@
 \newcommand{\code}[1]{\texttt{#1}}
-\newcommand{\jmodule}[1]{\emph{#1}}
+\newcommand{\jmodule}[1]{\texttt{\textsl{#1}}}
 
@@
 \maketitle
-Target audience: developers willing to extend jInfer, scientist willing to implement own inference methods\\
+\noindent Target audience: developers willing to extend jInfer, scientist willing to implement own inference methods
 
-\emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}
+\noindent \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}
 
 \section{Using of factory pattern}
-TwoStep simplifier is divided into three submodules, which are then divided into submodules and so on.
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{factory_pattern1}
+	\caption{Same instance of class returned by NetBeans in two successive inference runs} \label{factory_pattern1}
+\end{figure}
+\jmodule{TwoStepSimplifier} is divided into three submodules, which are then divided into submodules and so on.
 Each submodule is simply class, properly anotated.
-User selects proper classes - submodules in chain to work on inference.
-Classes are then, in runtime looked up by using NetBeans lookup mechanism.
+User selects proper classes - submodules to work in chain on inference.
+Classes are then, in runtime looked up by using NetBeans lookup mechanism (see \cite[p. 16]{arch}).
 But NetBeans holds one instance of each class, that may be looked upon.
@@
 Then, user clicks run button again, and same instances of classes are returned by lookups.
-Ooops, theese are not freshly created instances, using them as they are may cause harm.
-And problem arises - force each class (submodule) that is being looked up in inference process, to implement
+Oops, theese are not freshly created instances, using them as they are may cause harm.
+The problem is ilustrated on fig. \ref{factory_pattern1}.
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{factory_pattern2}
+	\caption{Factory solution of singleton lookup class problem} \label{factory_pattern2}
+\end{figure}
+
+There are basically two solutions:
+\begin{itemize}
+	\item force each class (submodule) that is being looked up in inference process, to implement
 some sort of cleanup method, that would restart it into fresh state and make it ready to use by another 
-inference run? Or create new instances of theese classes in each inference run and make singleton classes (those which NetBeans return by lookups) their factory classes?
-
-We decided to use second approach. 
-It enables us not only to produce fresh instance in each inference run, but also factory classes implement obligatory
+inference run,
+	\item or create new instances of theese classes in each inference run and make singleton classes (those which NetBeans return by lookups) their factory classes.
+\end{itemize}
+We decided for the latter approach (as ilustrated on fig. \ref{factory_pattern2}
+It enables us not only to produce fresh class instance in each inference run, but also factory classes implement obligatory
 module methods such as \code{getName}, \code{getDescription} and so on, which are really same in each inference run.
@@
 \begin{verbatim}
-public interface AutomatonSimplifierFactory extends NamedModule, Capabilities, UserModuleDescription {
+public interface AutomatonSimplifierFactory extends 
+                     NamedModule, Capabilities, UserModuleDescription {
   <T> AutomatonSimplifier<T> create();
@@
 \end{verbatim}
-It extends \code{NamedModule} with all module methods.
+It extends \code{NamedModule} with its modules describing methods:
+\begin{verbatim}
+String getName();
+String getDisplayName();
+String getModuleDescription();
+\end{verbatim}
 It extends \code{Capabilities}, thus each module have to answer, if it has some capabilities.
-And, in TwoStep we often extend \code{UserModuleDescription}, which defines method for obtaining description of the module, comprehensive to user (it is displayed in properties panels).
-In some circumstances, it is useful to have method \code{create} generic.
-The automaton simplifier works with automaton, which itself is generic too.
+And, in TwoStep we often extend \code{UserModuleDescription}, which defines method:
+\begin{verbatim}
+String getUserModuleDescription();
+\end{verbatim}
+It returns description of the module, comprehensive to user, which is then displayed in properties panels.
+
+In some circumstances, it is useful to have method \code{create()} generic.
+The \jmodule{AutomatonSimplifier} works with \code{Automaton}, which itself is generic too.
 But simplifying does not depend on type of symbol of automaton, so interface \code{AutomatonSimplifier<T>}
-is also generic as simplifier can simplify automaton of any symbol java type.
-Factory interface deals with this by defining create method generic too.
+is also generic as simplifier can simplify automaton of any java type symbol.
+Factory interface deals with this by defining \code{create()} method generic too.
 
-Second interface is the one, which is returned by \code{create} method in factory.
+Second interface is the one, which is returned by \code{create()} method.
 That is the interface, which defines real work cycle with submodule (in example it is \code{AutomatonSimplifier<T>}).
@@
 \end{verbatim}
-Cluster processor AutomatonMergingState then receives factories for AutomatonSimplifier and RegexpAutomatonSipmlifier submodules in its constructor.
+Cluster processor \code{AutomatonMergingState} then receives factories of \code{AutomatonSimplifier} and \code{RegexpAutomatonSimplifier} submodules in its constructor.
 Cluster processor then may create as many instances of submodule classes as it needs (maybe simplifying more than one automaton).
@@
   // 4. add to rules
-      finalGrammar.add(
-              new Element(node.getContext(),
+  finalGrammar.add(new Element(node.getContext(),
                    node.getName(),
@@
 We provide \code{Clusterer} interface for classes to implement.
-Its purpose is to cluster bunch of elements (rules) on input, into  bunch of cluster class instances  (cluters).
+Its purpose is to cluster bunch of elements (rules) on input, into  bunch of cluster class instances (cluters) on output.
 It has methods \code{add()} and \code{addAll()} for adding items for clustering.
 Center is method \code{cluster()}, which does the clustering itself.
-As it may be time-consuming operation, method throws \code{InterruptedException}, implementation should take care 
-of checking whether thread is user interrupted (see \cite{arch}).
+As it may be time-consuming operation, method throws \code{InterruptedException}.
+Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{arch}).
 After clustering, implementation should hold clusters in member, as it it will be further asked by calling method \code{T getRepresentantForItem(T item)}.
@@
 If no such cluster exists (item was not added for clustering before), we recommend you throwing an exception rather than returning \code{null}.
-Missing will probably  indicate error in algorithm rather than normal workflow.
+Missing item will probably  indicate error in algorithm rather than normal workflow.
 One can pull clusters from clusterer by calling \code{getClusters()} method.
@@
 To address clustering of elements in more detail, we created \code{ClustererWithAttributes<T, S>} interface, which extends \code{Clusterer<T>} interface.
-It add method \code{List<Cluster<S>> getAttributeClusters(T representant)}, implying that each representant of
+It adds method \code{List<Cluster<S>{}> getAttributeClusters(T representant)}, implying that each representant of
 type \code{T} (that is representant of some main cluster) has some "attribute" clusters associated with it.
-Theese clusters are obtained by calling \code{getAttributeClusters(x)}.
-Attribute clusters are of type \code{S}.
+Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.
 
@@
 \end{figure}
-We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext}, which is used by \class{Greedy}
+We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext}, which is used by \code{Greedy}
 to test mergability of states by default configuration.
 
-\subsubsubsection{StateRemoval}
+\subsubsection{StateRemoval}
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
-This is implemented in \class{StateRemoval} class.
+This is implemented in \code{StateRemoval} class.
 We defined one submodule of this class with interface called \code{Orderer}.

### 2011-01-21 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - state removal modules picture

@@
 
-Whole submodule structure of AutomatonMergingState cluster processor is drawn on fig. \ref{automaton_merging_state_submodules}.
+Whole submodule structure of AutomatonMergingState cluster processor is drawn on fig. \ref{automaton_merging_state_modules}.
 \begin{figure}
-	\centering\includegraphics[scale=\myscale]{automaton_merging_state_submodules}
-	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_submodules}
+	\centering\includegraphics[scale=\myscale]{automaton_merging_state_modules}
+	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_modules}
 \end{figure}
@@
 \subsubsection{StateRemoval}
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{state_removal_modules}
+	\caption{Modules of StateRemoval.} \label{state_removal_modules}
+\end{figure}
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
-This is implemented in \code{StateRemoval} class.
+This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}).
 We defined one submodule of this class with interface called \code{Orderer}.

### 2011-01-22 doc/jInfer/architecture/architecture.tex
removing "schema.data"
3154619 Create documentation for jInfer architecture - better pictures for modules_inference and module_deps (stretching to textwidth with bigger font)
3154619 Create documentation for jInfer architecture - resolved my todos, given to reseto
3159165 Create documentation fof TwoStep simplifier - experimenting
formatting paragraphs and verbatims in architecture

@@
 \maketitle
-Target audience: developers willing to extend jInfer.\\
+\noindent Target audience: developers willing to extend jInfer.
 
-\emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\
+\noindent \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\
 
-The description of jInfer architecture will be divided into 3 main views: the data view, the process view and the plaform view.\\
+The description of jInfer architecture will be divided into 3 main views: the data view, the process view and the plaform view.
+The data view will commence by describing the data structures, namely representations of regular expressions and XML elements, attributes and simple data.
+The notions of a \emph{rule} and a \emph{grammar} will be explained, along with the finite state automata representation we use.
 
-The data view will commence by describing the data structures, namely representations of regular expressions and XML elements, attributes and simple data.
-The notions of a \emph{rule} and a \emph{grammar} will be explained, along with the finite state automata representation we use.\\
-
-Afterwards the process of inference will be described, both from high-level and programmatic point of view.\\
-
-Finally, we will look at jInfer as at an NetBeans Platform application consisting of a number of modules performing their tasks and communicating together through a set of well-defined interfaces.\\
+Afterwards the process of inference will be described, both from high-level and programmatic point of view.
+Finally, we will look at jInfer as at an NetBeans Platform application consisting of a number of modules performing their tasks and communicating together through a set of well-defined interfaces.
 
@@
 \end{itemize}
-
 Type of regexp is held in \code{type} member in class \code{Regexp<T>} and can be tested by calling methods \code{isLambda()}, \code{isToken()} etc.
@@
 
+For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of members set.
+One has to fill in all properties carefully and call \code{setImmutable()} aftewards.
+Proper usage should be one of following:
 Regexp instance is by default created as immutable, that is, once instantiated, you cannot add more children to list of children, cannot change type, content etc.
@@
 This occurs mainly in input modules, where by parsing XML data sequentially, one does not know contents of element in time of handling start element event.
-For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of members set.
-One has to fill in all properties carefully and call \code{setImmutable()} aftewards.
-Proper usage should be one of following:
 \begin{verbatim}
@@
 Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full}.
-
 In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{AbstractStructuralNode} variables.
@@
 \subsection{Rules and grammars} \label{rules_grammar}
-
 jInfer and its documentation uses extended context-free grammars\cite{extendedcfg}.
-% TODO vektor check citation of http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf in literature.bib
-% TODO anti what do you mean, "check citation"?
 \emph{Rules} in such grammar are in the form
@@
 $$
-In jInfer each such rule is represented with an \code{Element} instance. In this representation, the \code{Element} itself is the LHS, its \code{subnodes} are the RHS.\\
+In jInfer each such rule is represented with an \code{Element} instance. In this representation, the \code{Element} itself is the LHS, its \code{subnodes} are the RHS.
 
@@
 We recommend skipping this section in first read as it describes advanced features.
+
 For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
-Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
+Implementation consists of 4 classes (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}): 
+\begin{itemize}
+  \item \code{Automaton<T>},
+  \item \code{Step<T>},
+  \item \code{State<T>},
+  \item \code{AutomatonCloner<A, B>}.
+\end{itemize}
 Whole implementation uses java generics for representation of symbol of alphabet.
 We denote \code{T} the java type of symbol.
+
 Take care of symbol class and its \code{equals()} implementation.
@@
 You will get PTA with useCounts and finalCounts set properly on steps/states.
-
 \begin{figure}
@@
 All \{in|out\}-transitions are redirected properly (or discarted as needed).
-Variables useCount and finalCount are updated to sums of values from merged transitions/states.
+Variables \code{useCount} and \code{finalCount} are updated to sums of values from merged transitions/states.
+
 One can ask for merging states, that are merged out from automaton already.
@@
 \end{figure}
-
 The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig.
@@
   </info>
+</person>
 <more/>
-  </person>
 <person>
@@
 	Lets look on XSD output of same rules:
-	% TODO rio review this example!
 	\begin{verbatim}
@@
 That means, \jmodule{Simplifier} is only started \emph{after} the \jmodule{IGG} completely finished its work and returned IG to be simplified.
-Similarly, \jmodule{SchemaGen} gets all the rules to export at once, in one list.\\
+Similarly, \jmodule{SchemaGen} gets all the rules to export at once, in one list.
 
@@
 \subsection{Programmatic view}
-
 From developer's point of view, inference modules are just properly annotated classes (see \ref{section_lookups}) implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
-
 \begin{itemize}
@@
 \end{itemize}
-
-A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{SimplifierImpl}.\\
-
+A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{SimplifierImpl}.
 Annotation required for the framework to recognize such a class as an inference module is the following
@@
 \end{verbatim}
-
 The most important method in each module is \code{start}, defined in each of the interfaces.
@@
 \subsubsection{Module input}
-Each inference module takes the actual input data as the first parameter of its \code{start} method. The type of the argument differs based on the inference module.\\
+Each inference module takes the actual input data as the first parameter of its \code{start} method. The type of the argument differs based on the inference module.
 
@@
 This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}.
-Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.\\
+Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.
 
@@
 	For element person, import module have to generate some positive examples of language generated by regular expression \verb+(info?, more, more?, more?)+.
-This generating is basicly done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{expand()}.
-One can implement own intelligent expander, or use our provided \code{cz.cuni.mff.ksi.jinfer.basicigg.expansion.ExpanderImpl}.
+
+	This generating is basicaly done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{expand()}.
+One can implement own intelligent expander, or use our provided\\ \code{cz.cuni.mff.ksi.jinfer.basicigg.expansion.ExpanderImpl}.
 
@@
 	Even then, import module is not obliged to use it of course. But it is reasonable to do so, if imported file type gives more information than simple concatenations (when importing schemas).
+
 	Initial Grammar can contain all regular expressions.
@@
 Empty concatenations are also allowed - as they are positive examples in XML document input.
-Its simplifier business to divide rules by import file formats (by using metadata described in \ref{section_metada}) and to handle empty concatenations from XML file import with complex regular expressions from other formats.
+Its simplifier business to divide rules by import file formats (by using metadata described in \ref{section_metadata}) and to handle empty concatenations from XML file import with complex regular expressions from other formats.
 \end{itemize}
@@
 Second parameter of each inference module's \code{start} method is a callback object. There are 3 callback interfaces defined in the \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package
-
 \begin{itemize}
@@
 \end{itemize}
-
 Each callback interface naturally belongs to the similarly named module interface.
@@
 Each inference module is responsible for invoking this method on the callback it got as a parameter, after it has finished its work and has results to be passed on.
-Again, these 3 finished methods have different arguments based on the inference module.\\
+Again, these 3 finished methods have different arguments based on the inference module.
 
-\code{IGGeneratorCallback.finished()} and \code{SimplifierCallback.finished()} have a grammar (Initial Grammar in case of IGG) as their only argument.\\
+\code{IGGeneratorCallback.finished()} and \code{SimplifierCallback.finished()} have a grammar (Initial Grammar in case of IGG) as their only argument.
 
@@
 \subsubsection{Error handling}
-
 Because the run of each inference module is encapsulated in a \code{try-catch} block by the framework, it is safe to throw any exception out of the \code{start} method: it will get logged, presented to the user and inference will stop.
@@
 \subsubsection{Interruptions}
-
 User running the inference might change his mind and try to stop this. For this reason, modules have to check for this case in every time-consuming place such as long loops with the following code:
@@
 \subsubsection{Runner} \label{section_runner}
-
-The part of framework responsible for actually gathering user input, running all modules one after another and presenting the results is the \code{Runner} class  in \code{cz.cuni.mff.ksi.jinfer.runner} package.\\
+The part of framework responsible for actually gathering user input, running all modules one after another and presenting the results is the \code{Runner} class  in \code{cz.cuni.mff.ksi.jinfer.runner} package.
 
@@
 The inference process itself is then as follows
-
 \begin{figure}
@@
 \end{figure}
-
 \begin{enumerate}
@@
 \subsection{Node metadata} \label{section_metadata}
-To allow simple extensibility of rules, class \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.\\
+To allow simple extensibility of rules, class \code{AbstractNamedNode} contains a string-addressed map called \code{metadata} that can contain arbitrary object values.
 
@@
 Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
-	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
 \end{itemize}
@@
 \subsection{Capabilities} \label{section_capabilities}
-From time to time, modules need a flow of certain information in completely opposite way to the usual inference: they need to know what the following module can or cannot do.\\
+From time to time, modules need a flow of certain information in completely opposite way to the usual inference: they need to know what the following module can or cannot do.
 
 For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCapabilities()}.
-This method returns a list of strings representing the abilities, or "capabilities" of this module.\\
+This method returns a list of strings representing the abilities, or "capabilities" of this module.
 An inference module can query the capabilities of the \emph{next} module in the inference chain by invoking the 
@@
 \end{verbatim}
-method to obtain the next module encapsulated in the \code{Capabilities} interface.\\
+method to obtain the next module encapsulated in the \code{Capabilities} interface.
 
 At the moment, only one capability is defined to be communicated across modules, more specifically between \jmodule{IGGenerator} and \jmodule{Simplifier}.
-This is \code{Capabilities.CAN\_HANDLE\_COMPLEX\_REGEXPS} meaning whether the \code{Simplifier} accepts more complicated regular expressions than simple concatenations of tokens on its input.\\
+This is \code{Capabilities.CAN\_HANDLE\_COMPLEX\_REGEXPS} meaning whether the \code{Simplifier} accepts more complicated regular expressions than simple concatenations of tokens on its input.
 % TODO vektor check later - didn't this change?!?
@@
 \section{Platform view}
-
 jInfer, being a NetBeans Platform application, is divided into multiple \emph{NetBeans Modules}.
 It is very important to understand that this division is in principle orthogonal to that naturally sketched in the previous chapters - theoretically, the whole jInfer could be contained in a single NB module, because the framework uses Lookups to locate logical modules.
-However, for the sake of organization, jInfer is split into several modules and most of the time there is a strong correlation between logical units of jInfer and its NB modules.\\
+However, for the sake of organization, jInfer is split into several modules and most of the time there is a strong correlation between logical units of jInfer and its NB modules.
 
@@
 To get a deeper understanding of any of them, please refer to its documentation.
-To get deeper understanding of how NBP modules work, their versions, dependencies etc, refer to \cite{modules_api}.\\
+To get deeper understanding of how NBP modules work, their versions, dependencies etc, refer to \cite{modules_api}.
 
-Base modules:
-
+\noindent Base modules:
 \begin{itemize}
@@
 
-Inference modules:
-
+\noindent Inference modules:
 \begin{itemize}
@@
 
-Utility modules:
-
+\noindent Utility modules:
 \begin{itemize}
@@
 \end{itemize}
-
 The dependencies between all the modules are summarized in figure \ref{module_deps}.
-Figure \ref{modules_inference} shows modules implementing the 3 main inference interfaces described in \ref{section_inference_process}.\\
+Figure \ref{modules_inference} shows modules implementing the 3 main inference interfaces described in \ref{section_inference_process}.
 
 \begin{figure}
-	\centering\includegraphics[scale=0.4]{module_deps}
+	\centering\includegraphics[scale=1]{module_deps}
 	\caption{Module dependencies} \label{module_deps}
@@
 \begin{figure}
-	\centering\includegraphics[scale=0.4]{modules_inference}
+	\centering\includegraphics[scale=\myscale]{modules_inference}
 	\caption{Inference modules in default jInfer} \label{modules_inference}
@@
 \subsection{Lookups} \label{section_lookups}
-
 \emph{Lookups} are a part of NBP API, a way to achieve late-binding between code based on this platform.
@@
 \end{verbatim}
-The \code{@ServiceProvider} annotation tells NBP that \code{SchemaGeneratorImpl} is to be registered among all providers of the \code{SchemaGenerator} service (or interface).\\
+The \code{@ServiceProvider} annotation tells NBP that \code{SchemaGeneratorImpl} is to be registered among all providers of the \code{SchemaGenerator} service (or interface).
 
@@
 This collection can be walked in some way, and on one or more of service providers methods can be called.
-We didn't have to actually bind providers to logic that uses their services, NB Platform took care of this for us.\\
+We didn't have to actually bind providers to logic that uses their services, NB Platform took care of this for us.
 
 At this point it is important to understand that when using lookups this way, all service providers are kept as singletons.
-In case a provider needs to change its state during providing the service (and needs to be ``fresh" each time it is retrieved), it is best to convert this to a factory design pattern: the service provider registered is just a factory returning instances that do the actual work.\\
+In case a provider needs to change its state during providing the service (and needs to be ``fresh" each time it is retrieved), it is best to convert this to a factory design pattern: the service provider registered is just a factory returning instances that do the actual work.
 
-TODO anti Link to http://bits.netbeans.org/dev/javadoc/org-openide-util-lookup/org/openide/util/Lookup.html as the official documentation to lookup API.
+See \cite{lookup_api} as the official documentation to lookup API.
 
 \subsection{Tokens}
-
 An important topic related to NB modules and their dependencies is a notion of \emph{tokens}.
@@
 Any module providing \emph{S} will declare a \emph{provided token}, which is a simple string (usually fully qualified Java name of the interface representing \emph{S}).
-Later, NBP will make sure that \emph{M} can be installed in the platform only if there is at least one module providing that specific token.\\
+Later, NBP will make sure that \emph{M} can be installed in the platform only if there is at least one module providing that specific token.
 
 jInfer uses tokens in the following way: \jmodule{Runner}, as the module encapsulating the inference process, and \jmodule{ProjectType} as an encapsulation of jInfer projects both require the following 3 tokens.
-
 \begin{itemize}
@@
 \end{itemize}
-
 All inference modules declare their respective token as provided.
@@
 
-\nocite{*}
+%\nocite{*}
+\newpage
 \bibliographystyle{alpha}

### 2011-01-22 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - removing unneeded pdfs
formatting paragraphs and verbatims in architecture
3159165 Create documentation fof TwoStep simplifier - more text on twostep module, subfigures

@@
 \newcommand{\code}[1]{\texttt{#1}}
-\newcommand{\jmodule}[1]{\texttt{\textsl{#1}}}
+\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
 
@@
 
+\noindent \begin{tabular}{|l|l|} \hline
+Responsible developer: & anti \\ \hline
+Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
+Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\ \hline
+Module dependencies:   & AutoEditor, Base, JUNG, Lookup \\ \hline
+Public packages:       & \\ \hline
+\end{tabular}
+
 \section{Using of factory pattern}
 \begin{figure}
+  \centering
+	\subfigure[Same instance of class returned by NetBeans in two successive inference runs]
+	{
 	  \centering\includegraphics[scale=\myscale]{factory_pattern1}
-	\caption{Same instance of class returned by NetBeans in two successive inference runs} \label{factory_pattern1}
+    \label{factory_pattern1}
+  }
+	\subfigure[Factory solution of singleton lookup class problem]
+	{
+	  \centering\includegraphics[scale=\myscale]{factory_pattern2}
+    \label{factory_pattern2}
+  }
 \end{figure}
@@
 User selects proper classes - submodules to work in chain on inference.
-Classes are then, in runtime looked up by using NetBeans lookup mechanism (see \cite[p. 16]{arch}).
+Classes are then, in runtime looked up by using NetBeans lookup mechanism (see \cite[p. 16]{archdoc}).
 But NetBeans holds one instance of each class, that may be looked upon.
@@
 The problem is ilustrated on fig. \ref{factory_pattern1}.
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{factory_pattern2}
-	\caption{Factory solution of singleton lookup class problem} \label{factory_pattern2}
-\end{figure}
 
@@
 \section{TwoStep module}
-TwoStep simplifier is inspired by \cite{1802522} design. Inference proceeds in two steps:
+\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
 \begin{enumerate}
@@
 \end{enumerate}
-Main routine of TwoStep simplifer does basically this:
+Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule.
+We will examine both of them.
+There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
+Modules are drawn on fig. \ref{twostep_modules}.
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{twostep_modules}
+	\caption{Submodules of TwoStep simplifier} \label{twostep_modules}
+\end{figure}
+We provide one \jmodule{Clusterer} and 4 \jmodule{ClusterProcessor} implementations.
+Each of those will be explained further in this document.
+
+\jmodule{TwoStepSimplfier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
+Module is implemented in two classes: \code{TwoStepSimplifier} (main logic), \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
+Factory class implements \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
+Thus has main method called \code{start} which receives Initial Grammar in form of
+\begin{verbatim}
+final List<Element> grammar
+\end{verbatim}
+Each grammar rule is represented as class \code{Element}, where element is left side of rule and it's \code{getSubnodes()} method
+returns regular expression, which is right side of rule.
+Second parameter is 
+\begin{verbatim}
+final SimplifierCallback callback
+\end{verbatim}
+Callback which should be called when work is done.
+On output, simplifier provides simplified grammar as a parameter into callback function:
+\begin{verbatim}
+void finished(final List<Element> grammar);
+\end{verbatim}
+Thus simplifier receives initial grammar and returns simplified grammar, both represented same way - as elements.
+
+\code{TwoStepSimplifierFactory} in method \code{start()} created new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
+Then it calls \code{simplify(initialGrammar)} method on this instance and its result passes as parameter to callback function.
+Thats all, the magic comes in \code{TwoStepSimplifier} class.
+
+Method \code{TwoStepSimplifier.simplify()} does basically this:
 \begin{verbatim}
@@
 // 3. process rules
-final ClusterProcessor<AbstractStructuralNode> processor= clusterProcessorFactory.create();
+final ClusterProcessor<AbstractStructuralNode> processor =
+             clusterProcessorFactory.create();
+
 for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
-  final AbstractStructuralNode node =  processor.processCluster(clusterer, cluster.getMembers() - without sentinels);
+  final AbstractStructuralNode node =  
+                          processor.processCluster(clusterer, cluster.getMembers());
+  final RegularExpressionCleaner<AbstractStructuralNode> cleaner = 
+                          regularExpressionCleanerFactory.<AbstractStructuralNode>create();
   // 4. add to rules
-  finalGrammar.add(new Element(node.getContext(),
+  finalGrammar.add(
+              new Element(node.getContext(),
               node.getName(),
               node.getMetadata(),
-                   ((Element) node).getSubnodes(),
-                   attList);
+              cleaner.cleanRegularExpression(((Element) node).getSubnodes()),
+              attList));
 }
+
+return finalGrammar;
 \end{verbatim}
-Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule.
-We will examine both of them.
-There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
-Modules currently in jInfer are drawn on fig. \ref{twostep_modules}.
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{twostep_modules}
-	\caption{Submodules of TwoStep simplifier} \label{twostep_modules}
-\end{figure}
-We provide one \jmodule{Clusterer} and four \jmodule{ClusterProcessor} implementations.
-Each of those will be explained further in this document.
+It creates clusterer, gives it all rules, orders it to cluster elements.
+Then, empty list of elements is created as final (simplified) grammar.
+For each input rule in initialGrammar, submodule \jmodule{ClusterProcessor} is called to do infering of regular expression of that element.
+Finally, regular expression cleaning is done in submodule and new \code{Element} instance is created as a copy of processed node, but with cleaned regexp.
 
+Now we will examine submodules for clustering, processing and cleaning.
+
+TODO sentinel cant be on left side
 TODO sentinel processing
 
-\subsection{Clusterer submodule}
+\subsection{Clusterer module}
 Lets start by example, let input document(s) contain XML:
@@
 As it may be time-consuming operation, method throws \code{InterruptedException}.
-Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{arch}).
+Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
 After clustering, implementation should hold clusters in member, as it it will be further asked by calling method \code{T getRepresentantForItem(T item)}.

### 2011-01-22 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - more text on twostep module, subfigures

@@
 Finally, we use this scheme to implement clusterer \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
-For each cluster of elements, it creates list of clusters of attributes observed with all elements in main cluster.
-
-All simple data are considered equal and given one dedicated cluster.
-But in future implementations, extensibility is ensured to cluster simple data to obtain content models of elements.
-
-All sentinel elements are clustered too.
-TODO sentinels
-TODO attributes clustering
+Clustering is done based on elements \code{getName} equality test (ignore case).
+For each rule (element), it is first clustered by finding cluster where representant has same name - or create new cluster with this element as representant.
+Then we process right side of this element rule (that are nodes from \code{getSubnodes}).
+Since right side of rule is always concatenation, we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
+Each node on right side, is examined:
+\begin{itemize}
+  \item if it is simple data throw it to \code{SimpleDataClusterer} class, which does nothing more, than holds all of simple data in one cluster.
+But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
+  \item if it is element and it is tagged as sentinel by metada - search main clusters to find cluster with representant of same \code{getName()}, or create new cluster with this sentinel as representant.
+From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future).
+So there can't be cluster with only one sentinel element in it.
+  \item do nothing otherwise, since it is element, that has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in main initial grammar list, thus it is already processed or will be on schedule in clustering.
+\end{itemize}
+Attributes of element are processed through helper attribute clusterer, which is created for each element cluster.
+We have bunch of elements of same name in cluster and one attribute clusterer associated with this bunch.
+This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
+It clusters them by name (case insensitive).
 

### 2011-01-22 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - and writing and writing

@@
 \section{Using of factory pattern}
+TODO anti/vektor move to architecture as described in twostepTODO.pdf
 \begin{figure}
@@
 
-\section{TwoStep module}
+\section{TwoStepSimplifier module}
 \jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
@@
 It has methods \code{add()} and \code{addAll()} for adding items for clustering.
-Center is method \code{cluster()}, which does the clustering itself.
+Centerpart is method \code{cluster()}, which does the clustering itself.
 As it may be time-consuming operation, method throws \code{InterruptedException}.
@@
 Missing item will probably  indicate error in algorithm rather than normal workflow.
-One can pull clusters from clusterer by calling \code{getClusters()} method.
+One can pull clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
 
@@
 \begin{itemize}
-  \item if it is simple data throw it to \code{SimpleDataClusterer} class, which does nothing more, than holds all of simple data in one cluster.
+  \item if it is simple data throw it to \code{SimpleDataClusterer} class which is associated with main element (of which this is right side), which does nothing more, than holds all of simple data in one cluster.
 But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
@@
 
+To rehearse, there are main clusters for elements, for each main cluster there are two helper clusterer - one for attribute clusters and one for simple data clusters.
+
 \subsection{ClusterProcessor submodule}
 Cluster processor takes rules of one cluster of element and somehow obtains regular expression for that set of elements.
-It returns rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for  resulting element) and with subnodes set to regular expression infered.
+It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for the resulting element) and with subnodes set to regular expression infered.
 It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to resulting element.
@@
 
-\subsubsection{PassRepresentant}
+\subsubsection{PassRepresentant cluster processor}
 Simple example to read. For each cluster return its representant as a rule to be in schema.
@@
 
-\subsubsection{Alternations}
+\subsubsection{Alternations cluster processor}
 This processor simply gets all right sides from elements in cluster, puts them in one big list and creates alternation regular expression with this list as children.
@@
 
-\subsubsection{Trie}
+\subsubsection{Trie cluster processor}
 This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie") of them.
@@
 Repeating this process creates a prefix tree describing all the rules in the cluster.
+TODO vektor and then what? how do you get regexp from that.
 
-\subsubsection{AutomatonMergingState}
-AutomatonMergingState is implementation of merging state algorithm on nondeterministic finite automaton.
+\subsubsection{AutomatonMergingState cluster processor}
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{automaton_merging_state_modules}
+	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_modules}
+\end{figure}
+\jmodule{AutomatonMergingState} is implementation of merging state algorithm on nondeterministic finite automaton (see package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate}).
 It creates prefix-tree automaton (PTA) from positive examples - right sides of rules given.
@@
 
-Simplified automaton is then converted to an instance of \code{RegexpAutomaton} by using clone constructor in regexp automaton.
+Simplified automaton is then converted to an instance of \code{RegexpAutomaton} by using clone constructor.
 Regexp automaton is automaton with regular expression as symbol on transitions.
 In automata theory, such automaton is called extended NFA.
-Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as content.
+Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
 
@@
 
-AutomatonMergingState has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by AutomatonMergingState.
-It is at disposal for implementations of AutomatonSimplifier interface for testing, whether two states in automaton are equivalent and should be merged into one state.
+\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{AutomatonMergingState}.
+It is at disposal for implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
 
-AutomatonSimplifier implements solution searching logic in simplifying automaton by merging states.
-We implement greedy strategy in class \code{Greedy}.
-It simply asks given MergeConditionTester if it can merge any of automaton states and merges states until there are no states to be merged (on every pair of states asked, MergeConditionTester answers, they cannot be merged).
-One can implement ACO heuristics or MDL principle heuristics as AutomatonSimplifier submodule using MergeConditionTesters provided.
+Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states.
+One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.
 
-Whole submodule structure of AutomatonMergingState cluster processor is drawn on fig. \ref{automaton_merging_state_modules}.
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{automaton_merging_state_modules}
-	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_modules}
-\end{figure}
-We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext}, which is used by \code{Greedy}
-to test mergability of states by default configuration.
+Shortened code of cluster processor operation:
+\begin{verbatim}
+// 1. Construct PTA
+final Automaton<AbstractStructuralNode> automaton = new Automaton<AbstractStructuralNode>(true);
 
-\subsubsection{StateRemoval}
+// Take each rule in cluster and pass right side to automaton to create PTA
+for (AbstractStructuralNode instance : rules) {
+  final Element element = (Element) instance;
+  final Regexp<AbstractStructuralNode> rightSide = element.getSubnodes();
+
+  final List<AbstractStructuralNode> rightSideTokens = rightSide.getTokens();
+
+  final List<AbstractStructuralNode> symbolString = new LinkedList<AbstractStructuralNode>();
+  for (AbstractStructuralNode token : rightSideTokens) {
+    symbolString.add(clusterer.getRepresentantForItem(token));
+  }
+  automaton.buildPTAOnSymbol(symbolString);
+}
+
+// 2. Simplify automaton by merging states using automatonSimplifier
+final Automaton<AbstractStructuralNode> simplifiedAutomaton =
+    automatonSimplifier.simplify(automaton, elementSymbolToString);
+
+// 3. Convert Automaton<AbstractStructuralNode> to RegexpAutomaton<AbstractStructuralNode>
+final RegexpAutomaton<AbstractStructuralNode> regexpAutomaton =
+    new RegexpAutomaton<AbstractStructuralNode>(simplifiedAutomaton);
+// 4. Call regexpAutomatonSimplifier to obtain regular expression from regexpAutomaton
+final Regexp<AbstractStructuralNode> regexp =
+    regexpAutomatonSimplifier.simplify(regexpAutomaton, regexpAbstractToString);
+
+// 5. Return element with regexp
+return new Element(
+      new ArrayList<String>(),
+      rules.get(0).getName(),
+      new HashMap<String, Object>(),
+      regexp,
+      new ArrayList<Attribute>()
+      );
+\end{verbatim}
+Whole submodule structure of \jmodule{AutomatonMergingState} cluster processor is drawn on fig. \ref{automaton_merging_state_modules}.
+
+\subsubsection{AutomatonSimplifier submodule}
+\jmodule{AutomatonSimplifier} module has its worker interface defined as:
+\begin{verbatim}
+public interface AutomatonSimplifier<T> {
+  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
+                        final SymbolToString<T> symbolToString) throws InterruptedException;
+}
+\end{verbatim}
+Given input automaton it returns automaton, there is nothing magical on it.
+Interface is defined as generic, implementation don't have to be, but there is no reason to not make them generic.
+When implementation needs to present automaton to user, however, it needs some string representation of symbols to display on automaton transitions.
+For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
+It is responsible for converting symbol to string.
+To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplfifier} submodule.
+
+By this genericity, simplifier can be used to perfom same algorithms for simplifying not only to infer rules for elements, but maybe also to infer patterns strings for content model of attributes.
+The actual implementation of this still waits for its developer to come.
+
+TODO how automaton works?
+
+Lets go on to implementations of \jmodule{AutomatonSimplifier}.
+
+\paragraph{Greedy automaton simplifier} We implement greedy strategy of automaton simplifying in class \code{Greedy}.
+It simply asks given \code{MergeConditionTester} if it can merge any of automaton states and merges states until there are no states to be merged:
+\begin{verbatim}
+while (there exist pair of states that are equivalent) {
+  merge them
+}
+\end{verbatim}
+We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext} (implements \code{MergeConditionTester} interface), which is used by \code{Greedy} to test mergability of states by default configuration.
+
+\paragraph{UserInteractive automaton simplifier}
+We have created automaton simplifier which displays user input automaton asking him to select some states to merge.
+Then it merges them and asks user again:
+\begin{verbatim}
+repeat {
+  draw automaton
+  if (user selects more than one state to merge) {
+    merge them
+  }
+} until (user selected at least one state in last trial)
+\end{verbatim}
+We user \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
+To make reader more happy, we include one screenshot of automaton visualization on picture \ref{screenshot_viz}.
+TODO anti screenshot for happy reader
+
+\subsubsection{RegexpAutomatonSimplifier submodule}
+TODO anti interface and flow of regexpautomatonsimplifier
+
 \begin{figure}
@@
 \end{figure}
+\paragraph{StateRemoval regexp automaton simplifier}
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
-This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}).
+This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
 We defined one submodule of this class with interface called \code{Orderer}.

### 2011-01-24 doc/jInfer/base/base.tex
Documentation: Base module.

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer Base Module Description}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
\noindent Target audience: developers willing to extend jInfer.

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & vektor \\ \hline
Required tokens:       & none \\ \hline
Provided tokens:       & none \\ \hline
Module dependencies:   & none \\ \hline
Public packages:       & cz.cuni.mff.ksi.jinfer.base.automaton \newline
cz.cuni.mff.ksi.jinfer.base.interfaces \newline
cz.cuni.mff.ksi.jinfer.base.interfaces.inference \newline
cz.cuni.mff.ksi.jinfer.base.interfaces.nodes \newline
cz.cuni.mff.ksi.jinfer.base.objects \newline
cz.cuni.mff.ksi.jinfer.base.objects.nodes \newline
cz.cuni.mff.ksi.jinfer.base.regexp \newline
cz.cuni.mff.ksi.jinfer.base.utils \newline
org.apache.log4j.* \\ \hline
\end{tabular}

\section{Introduction}

This is the module containing data structures, interfaces and logic shared across the whole jInfer framework. Virtually every other module containing logic should in theory depend on \jmodule{Base}.

\section{Structure}

Description of \jmodule{Base} structure will partially mirror its JavaDoc documentation. For more detailed information, refer to it directly.\\

\jmodule{Base} contains logic for Log4j initialization in \code{Installer} class. Configuration of the overall logging granularity level (NetBeans options integration) is contained in \code{cz.cuni.mff.ksi.jinfer.base.options} package. Shared jInfer graphics is contained in the \code{cz.cuni.mff.ksi.jinfer.base.graphics} and \code{cz.cuni.mff.ksi.jinfer.base.graphics.icons} packages.

\subsection{Data structures}

Regular expression representation is contained in \code{cz.cuni.mff.ksi.jinfer.base.regexp} package. Most important class is of course \code{Regexp}, assisted by an enum of its type \code{RegexpType} and representation of its interval \code{RegexpInterval}.\\
XML node representation is described by interfaces in \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes} package and more-less concrete implementations in \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes} package.\\
Finite state automata representation is contained in the \cite{cz.cuni.mff.ksi.jinfer.base.automaton} package.\\
Refer to \ref{archdoc} to get an accurate description of all these representations.\\

Miscellaneous shared classes are contained in \code{cz.cuni.mff.ksi.jinfer.base.objects}. For example, \code{Input} is used to provide the \jmodule{Initial Grammar Generator} with input data. \code{Pair} is a generic class binding two object together in a \emph{pair}.

\subsection{Interfaces}

Apart from interfaces contained in already discussed \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes} package, there is an important group of interfaces contained in \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package: the \emph{inference} interfaces. They come in two flavours: the actual inference interface, and its callback. For a comprehensive description of them and their interaction refer again to \ref{archdoc}.\\

There is one more package containing interfaces: \code{cz.cuni.mff.ksi.jinfer.base.interfaces}. There are a few groups of them.
\begin{itemize}
	\item Module lookup support: \code{NamedModule} and \code{UserModuleDescription}.
	\item Inference support: \code{Capabilities}.
	\item Service provider definitions: \code{Expander}, \code{Processor} and \code{RuleDisplayer}.
\end{itemize}

\subsection{Utility logic}

Useful logic for the whole framework is focused in the \code{cz.cuni.mff.ksi.jinfer.base.utils} package. A few highlights from here follow.
\begin{itemize}
	\item Testing whether a collection is empty: \code{BaseUtils.isEmpty()} - handles \code{null} and zero elements.
	\item Filtering a list based on a predicate: \code{BaseUtils.filter()}.
	\item Cloning a list N times in a row: \code{baseUtils.cloneList()} - e.g. from \emph{abc, 3 times} creates \emph{abcabcabc}.
	\item Deep cloning a grammar: \code{CloneHelper.cloneGrammar()}.
	\item Writing out a list separated by specified separator: \code{CollectionToString.colToString()} - accepts the list, operation to perform on each element and separator character. Separator is placed smartly, i.e. only between elements.
	\item XML element comparison while ignoring specified members: \code{EqualityUtils}.
	\item Module lookup/selection: \code{ModuleSelectionHelper}.
	\item Singleton class reporting on the currently running inference: \code{RunningProject}.
	\item Various utilities for JUnit tests: \code{TestUtils}.
	\item Topological sorting of grammar: \code{TopologicalSort}. 
\end{itemize}

\section{Data flow}

This is not an inference module and there is no real data flow in it. Refer to \ref{archdoc} to understand the inference process described by interfaces from \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package.

\section{Extensibility}

From one point of view, extensibility of \jmodule{Base} means creating service providers implementing inference or other interfaces. This is described in the documentation for respective modules that use these interfaces.\\
On the other hand, if there is a need to share data structures, interfaces or logic across multiple developed modules (for example between custom schema generator and simplifier), it is advised to create a module similar to \jmodule{Base} instead of changing it directly.

%\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-24 doc/jInfer/basicigg/basicigg.tex
Documentation: BasicIGG module.

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer BasicIGG Module Description}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
\noindent Target audience: developers willing to extend jInfer, specifically modify the way \jmodule{BasicIGG} creates initial grammar from input files, for example by adding support for a new schema language.

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & Matej Vitásek \\ \hline
Required tokens:       & none \\ \hline
Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\ \hline
Module dependencies:   & Base \\ \hline
Public packages:       & cz.cuni.mff.ksi.jinfer.basicigg.properties \\ \hline
\end{tabular}

\section{Introduction}

This is an extensible implementation of the \jmodule{IGGenerator} inference interface. It is the only IG generator officialy shipped with jInfer.\\
Make sure you understand the difference between the two types of initial grammar as described in \ref{archdoc}.

\section{Structure}

The main class implementing \code{IGGenerator} inference interface and simultaneously registered as its service provider is \code{IGGeneratorImpl}. In its \code{start} method it enumerates all files in the input parameter (of type \code{Input}). For each file, based on its extension the correct \emph{processor} is selected, executed and returned rules are aggregated. After each file has been processed, the resulting grammar is returned by invoking the \code{finished} method of the callback argument.

\section{Processors}

A \emph{processor} is a class capable of extracting IG from an arbitrary \code{InputStream} (usually encapsulating a file). Various processors may handle generic XML, schemas like DTD, XSD or Schematron, query languages such as XPath, and so on.\\
A processor has to be registered as a service provider of the \code{Processor} interface from \jmodule{Base} module. Due to the nature of NBP lookups, each processor is internally kept as a singleton, and should not use its inner state (refer to the chapter Lookups in \ref{archdoc}). Note that the factory pattern is not used here.\\ % TODO anti reference to a chapter?
Each processor declares the class of inputs (documents, schemas, queries) and file extensions it is able to handle by implementing methods \code{getFolder()}, \code{getExtension()} and \code{processUndefined()}. Refer to JavaDoc of these methods for further details.\\

\jmodule{BasicIGG} comes bundled with 3 processors: for generic XML documents, DTD schemas and XPath queries. Support for XSD queries is implemented in \jmodule{XSDImporter} to demonstrate \jmodule{BasicIGG}'s extension capabilities.

\subsection{XML processor}

XML processor registers itself into \emph{document} input folder, \emph{xml} file extension and declares that it can process other arbitrary file extensions.\\
SAX traversal is used to collect the rules; the relevant \code{ContentHandler} is the class \code{TrivialHandler}. The way it works is following: every time a start of an element is encountered, a new \code{Element} is created along with its attributes. This new element representation is then placed on the top of a stack. Every other element or simple data found until its ending tag is created and attached to its subnodes. When an end of an element is encountered, the element currently on the top of the stack is closed and declared to be a rule of the IG.\\
After reaching the end of the document, all IG rules are returned. Note that this approach creates "simple" initial grammar.

\subsection{DTD processor}

DTD processor registers itself into \emph{schema} input folder, \emph{dtd} file extension and declares it cannot handle other file extensions.\\
To parse DTD files a 3rd party library is used: \jmodule{dtdparser}. % TODO anti Reference http://www.rpbourret.com/dtdparser/index.htm
Translation from the object model in this library to our own is handled mostly by \code{DTD2RETranslator} class.\\
Note that the initial grammar generated in this way is complex. Therefore, this processor checks for the \code{CAN\_HANDLE\_COMPLEX\_REGEXPS} capability of the simplifier following this IG generator: in positive case it simply returns the grammar, if the following module cannot handle complex regular expressions, DTD processor first invokes an \emph{Expander} (see \ref{section-expansion}).

\subsection{XPath processor}

This is a rather na\"{i}ve implementation of an XPath processor. It registers itself into \emph{query} input folder, \emph{xpath} file extension (text file containing one XPath query per line) and declares it cannot handle other file extensions.\\
To parse XPath queries, standard support present in JDK is used. The relevant \code{XPathHandler} is the \code{XPathHandler} class.\\
TODO vektor Explain what it does...

\section{Expansion} \label{section-expansion}

Expansion is used to convert complex regular expressions in initial grammar to simple ones (concatenations of tokens) for simplifiers that can handle only the simple form. Relevant interface encapsulating any implementation of such an expander is \code{Expander} in \jmodule{Base}. Reference implementation is \code{ExpanderImpl} in this module. Note that even though this implementation is retrieved using lookups, jInfer bundles only one such implementation and does not support choice among more of them. Anyone wishing to implement his own expander will thus have to either remove jInfer's implementation, or implement the usual module selection. % TODO vektor Is this documented somewhere?

\subsection{ExpanderImpl internals}

TODO vektor

\section{Data flow}

Flow of data in this module is following.
\begin{enumerate}
	\item IGGeneratorImpl walks over input files in a loop.
	\item Each file gets processed by a processor based on its folder and extension.
	\item Rules from all files are gathered in a single list (IG) and returned via a callback.
\end{enumerate}

\section{Extensibility}

\jmodule{BasicIGG} can be easily extended to support a new input type: just create a class implementing \code{Processor}, annotate it as a service provider and implement any logic needed. \jmodule{XSDImporter} is an example of this.\\

It is possible to replace the default \code{Expander} implementation: but as mentioned in \ref{section-expansion}, either the old implementation must be removed or module selection must be introduced.

%\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-24 doc/jInfer/base/base.tex
Documentation: BasicIGG module.

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & vektor \\ \hline
+Responsible developer: & Matej Vitásek \\ \hline
 Required tokens:       & none \\ \hline

### 2011-01-25 doc/jInfer/twostep/twostep.tex
Changed type of arrows to empty for uml-like diagrams

@@
 
-\def\mftitle{jInfer TwoStep simplifier design and implementation}
+\def\mftitle{jInfer TwoStepSimplifier design and implementation}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & anti \\ \hline
+Responsible developer: & Michal Klempa \\ \hline
 Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
@@
 \subsubsection{RegexpAutomatonSimplifier submodule}
-TODO anti interface and flow of regexpautomatonsimplifier
+Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain regular expression from given automaton.
+Interface for doing this is thus simple:
+\begin{verbatim}
+public interface RegexpAutomatonSimplifier<T> {
+  Regexp<T> simplify(final RegexpAutomaton<T> inputAutomaton, 
+                     final SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
+}
+\end{verbatim}
+Once again we are encountering \code{symbolToString} with same purpose as before, if submodule would need to present automaton to user
+it has to be able to convert symbols of (at runtime) unknown type to strings.
 
@@
 This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
-We defined one submodule of this class with interface called \code{Orderer}.
+State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - superInitial and superFinal.
+Both are added before algorithm starts. Former one with $\lambda$-transition to initial state, and from all final states $\lambda$ transition to superFinal state is added.
+After removing all states, there is only one transition from superInitial to superFinal state.
+That transition has final regular expression on it as symbol, it is read and returned to \jmodule{AutomatonMergingState}.
+
+We defined one submodule of this class with \jmodule{Orderer}.
 It has only one method to implement: \code{getStateToRemove}.
-Given automaton it has to return reference to one state which should be removed from automaton at first.
-State removal calls this submodule and removes states given until there are only two states in automaton - superInitial and superFinal states,
-with exactly one transition.
-That transition has final regular expression on it as symbol, it is read and returned to AutomatonMergingState.
-We implement one orderer, called \code{Weighted}.
+Given automaton, it has to return reference to one state which should be removed from automaton at first.
+State removal calls this submodule and removes state returned.
+
+\paragraph{Weighted} We implement one orderer, called \code{Weighted}.
 It is simple heuristic - weights all states (weight = sum of {in | out | loop}-transition regular expression lengths) and returns state with lowest weight.
@@
 Last we examine \code{RegularExpressionCleaner} interface.
-TODO anti cleaners
+Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
+For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such like $(name, (person, id))$.
+To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
+Interface is definition is straightforward:
+\begin{verbatim}
+public interface RegularExpressionCleaner<T> {
+  Regexp<T> cleanRegularExpression(final Regexp<T> regexp);
+}
+\end{verbatim}
+Given regular expression, return regular expression.
+Converting of regular expression is commonly a recursive task and all our implementations work this way.
+Lets look at them.
+
+\paragraph{NestedConcatention}
+This cleaner does exactly what we have just described.
+It converts nested concatenations into flat ones.
+
+\paragraph{Null}
+This cleaner just returns regexp it receives and does nothing.
+Plug this one into chain, if you want to omit this step of inference.
+
+\paragraph{EmptyChildren}
+Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
+For example some pitty inferring method produces unnecessary empty regexps in regexp tree as:
+$$
+(name, (), (person, id))
+$$
+There is one empty concatenation/alternation/permutation with no sense.
+This is wiped out by this cleaner.
+
+\paragraph{Chained}
+Enables user to chain more cleaners with output from first one plugged as input into second one and so one.
+Thus regular expression from inference can pass through \jmodule{EmptyChildren} and them through \jmodule{NestedConcatenation} cleaners and then it is returned.
 

### 2011-01-25 doc/jInfer/base/base.tex
Documentation: Base and BasicIGG modules.

--- 2011-01-24 doc/jInfer/base/base.tex
@@
 Module dependencies:   & none \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.base.automaton \newline
-cz.cuni.mff.ksi.jinfer.base.interfaces \newline
-cz.cuni.mff.ksi.jinfer.base.interfaces.inference \newline
-cz.cuni.mff.ksi.jinfer.base.interfaces.nodes \newline
-cz.cuni.mff.ksi.jinfer.base.objects \newline
-cz.cuni.mff.ksi.jinfer.base.objects.nodes \newline
-cz.cuni.mff.ksi.jinfer.base.regexp \newline
-cz.cuni.mff.ksi.jinfer.base.utils \newline
-org.apache.log4j.* \\ \hline
+Public packages:       & cz.cuni.mff.ksi.jinfer.base.automaton \\ 
+& cz.cuni.mff.ksi.jinfer.base.interfaces \\ 
+& cz.cuni.mff.ksi.jinfer.base.interfaces.inference \\ 
+& cz.cuni.mff.ksi.jinfer.base.interfaces.nodes \\
+& cz.cuni.mff.ksi.jinfer.base.objects \\ 
+& cz.cuni.mff.ksi.jinfer.base.objects.nodes \\ 
+& cz.cuni.mff.ksi.jinfer.base.regexp \\ 
+& cz.cuni.mff.ksi.jinfer.base.utils \\
+& org.apache.log4j.* \\ \hline
 \end{tabular}

### 2011-01-25 doc/jInfer/basicigg/basicigg.tex
Documentation: Base and BasicIGG modules.

@@
 Expansion is used to convert complex regular expressions in initial grammar to simple ones (concatenations of tokens) for simplifiers that can handle only the simple form.
+The process can be seen as breaking down a regular expression into a set of words (positive examples) that are described by this regexp.
 Relevant interface encapsulating any implementation of such an expander is \code{Expander} in \jmodule{Base}.
@@
 
-TODO vektor
+\code{ExpanderImpl} works recursively, and it will probably be the best to explain it from bottom to up.\\
+
+At this point, let's denote a \emph{word} over an \emph{alphabet} typed (in Java generics sense) \emph{T} as a \code{List<T>}.
+Following the same logic, let's denote a set of such words as \code{List<List<T>{}>}.
+Our aim while expanding will be to convert a regular expression representing some node's subnodes to a list of words that are described by this regexp.
+In this case, \emph{T} will be \code{AbstractStructuralNode}.\\
+
+The first interesting function is \code{ExpansionHelper.applyInterval()}.
+It takes a list of words extracted from a regexp (thus in the List<List<AbstractStructuralNode>{}> form) and a \code{RegexpInterval} and creates a new list of words: applying the interval on each of the words in a heuristic sense.
+The behaviour depends on the type of the interval.
+\begin{itemize}
+	\item \emph{Once}: the resulting set of words is exactly the original set of words.
+	\item \emph{Optional}: same as above, but an empty word is added to the resulting set (each word from input \emph{may} or \emph{may not} occur - but one empty word is enough).
+	\item \emph{General interval}: has a minimum and a maximum number of occurences.
+Theoretically, all counts in interval $ [min, max] $ should be considered, but for practical reasons the following heuristic is employed: each word from the input is duplicated\footnote{Duplicating a word $abc$ $N$ times results in a word $\underbrace{abcabc\ldots abc}_{N\times}$.} first $min$, then $max$ times and added to the result.
+	\item \emph{Kleene star}: similar to the general interval and \emph{optional}, the result contains an empty word.
+Then for each word on input: this word and its 3$\times$ duplication is added to the result.
+Obviously, this is a heuristic approach again - theoretically each word should be duplicated between 0 and $\infty$ times in the result.
+	\item \emph{Kleene cross}: except for not including empty word, this is the same case as \emph{Kleene star}: each input word is duplicated 1 and 3 times in the result.
+\end{itemize}
+
+Understanding how intervals affect the expansion, it is time to move a level up: how a regular expression is expanded.
+This is done in \code{ExpanderImpl.unpackRE()} performing a recursive descent into the regexp tree and returning a set of words.
+Based on the regexp type, one of the following is performed.
+\begin{itemize}
+	\item \emph{LAMBDA}: resulting word set contains one empty word.
+	\item \emph{TOKEN}: resulting set contains a word with one letter: the content (an \code{AbstractStructuralNode}) of this token; the interval of the regexp is applied on this word before returning it.
+	\item \emph{ALTERNATION}: first, all children of this alternation are recursively evaluated.
+Their respective words are appended in a list, this list has the regexp interval applied and is returned.
+Reasoning is the following: we are trying to get all the words a regexp can produce.
+If the regexp is an alternation, any word that can be produced by one of its children can be a word produced by the regexp itself.
+	\item \emph{PERMUTATION}: a heuristic is applied again: because there are $n!$ ways to order children of a permutation, only two are actually picked: the original order in which they appear, and reversal of this order.
+In both cases the children are treated as if they we concatenation (and thus the following case is invoked) and appended in a list.
+On this list the regexp interval is applied and the list is returned.
+	\item \emph{CONCATENATION}: TODO vektor
+\end{itemize}
 

### 2011-01-25 doc/jInfer/basicigg/basicigg.tex
Documentation: BasicIGG - expansion description.

@@
 To parse XPath queries, standard support present in JDK is used. The relevant \code{XPathHandler} is the \code{XPathHandler} class.\\
-TODO vektor Explain what it does...
+% TODO vektor Explain how it works...
 
@@
 On this list the regexp interval is applied and the list is returned.
-	\item \emph{CONCATENATION}: TODO vektor
+	\item \emph{CONCATENATION}: this is by far the most complicated case.
+Let's denote word sets produced by the $N$ children of the concatenation $ w_1 \ldots w_N $.
+Now, all words that could be produced by this concatenations correspond to concatenations of form $w_{a_1} w_{a_2} \ldots w_{a_N}$ for all possible combinations of indices $a_i, i  \in \{1 \ldots N\}$.
+Number of such words grows exponentially with $N$, therefore a heuristic approach has to be employed again.\newline
+In this case, consider $ m = \max_{i \in 1 \ldots N}{|w_i|} $.
+Now, $ m $ words will be constructed in the following way: in a for-loop enumerating $ i \in \{1 \ldots m\}$ the $j$-th part of this $i$-th word will be $w_{i mod |w_j|}$.
+The method \code{ExpanderImp.unpackConcat()} illustrates this.
 \end{itemize}
+
+Now for the next part, description of \code{Element} expansion.
+This is the content of \code{ExpanderImpl.expandElement()} method.
+First, element's subnodes are expanded as a regexp.
+For each word from this expansion a new \code{Element} is created: letters of this word are encapsulated in tokens, these tokens are put in a concatenation that constitutes subnodes of this new element.
+If the letter is an element itself, it is not recursively expanded - it is just declared to be a sentinel.\\ % TODO anti Link to architecture
+
+Finally, expanding a grammar means expanding each of its rules (elements) and returning the gathered rules (elements).
 

### 2011-01-25 doc/jInfer/basicruledisplayer/basicruledisplayer.tex
Documentation: BasicDTDExporter and BasicRuleDisplayer.

--- 2011-01-24 doc/jInfer/basicigg/basicigg.tex
@@
 
-\def\mftitle{jInfer BasicIGG Module Description}
+\def\mftitle{jInfer BasicRuleDisplayer Module Description}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \maketitle
-\noindent Target audience: developers willing to extend jInfer, specifically modify the way \jmodule{BasicIGG} creates initial grammar from input files, for example by adding support for a new schema language.
+\noindent Target audience: developers willing to extend jInfer, looking for ways to visualize grammars.
 
@@
 Required tokens:       & none \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\ \hline
+Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
 Module dependencies:   & Base \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.basicigg.properties \\ \hline
+Public packages:       & none \\ \hline
 \end{tabular}
@@
 
-This is an extensible implementation of the \jmodule{IGGenerator} inference interface. It is the only IG generator officialy shipped with jInfer.\\
-Make sure you understand the difference between the two types of initial grammar as described in \ref{archdoc}.
+This is fairly basic rule displayer draws rules as series of nested coloured rectangles.
 
@@
 
-The main class implementing \code{IGGenerator} inference interface and simultaneously registered as its service provider is \code{IGGeneratorImpl}.
-In its \code{start} method it enumerates all files in the input parameter (of type \code{Input}).
-For each file, based on its extension the correct \emph{processor} is selected, executed and returned rules are aggregated.
-After each file has been processed, the resulting grammar is returned by invoking the \code{finished} method of the callback argument.
-
-\section{Processors}
-
-A \emph{processor} is a class capable of extracting IG from an arbitrary \code{InputStream} (usually encapsulating a file).
-Various processors may handle generic XML, schemas like DTD, XSD or Schematron, query languages such as XPath, and so on.\\
-A processor has to be registered as a service provider of the \code{Processor} interface from \jmodule{Base} module.
-Due to the nature of NBP lookups, each processor is internally kept as a singleton, and should not use its inner state (refer to the chapter Lookups in \ref{archdoc}).
-Note that the factory pattern is not used here.\\ % TODO anti reference to a chapter?
-Each processor declares the class of inputs (documents, schemas, queries) and file extensions it is able to handle by implementing methods \code{getFolder()}, \code{getExtension()} and \code{processUndefined()}.
-Refer to JavaDoc of these methods for further details.\\
-
-\jmodule{BasicIGG} comes bundled with 3 processors: for generic XML documents, DTD schemas and XPath queries.
-Support for XSD queries is implemented in \jmodule{XSDImporter} to demonstrate \jmodule{BasicIGG}'s extension capabilities.
-
-\subsection{XML processor}
-
-XML processor registers itself into \emph{document} input folder, \emph{xml} file extension and declares that it can process other arbitrary file extensions.\\
-SAX traversal is used to collect the rules; the relevant \code{ContentHandler} is the class \code{TrivialHandler}.
-The way it works is following: every time a start of an element is encountered, a new \code{Element} is created along with its attributes.
-This new element representation is then placed on the top of a stack.
-Every other element or simple data found until its ending tag is created and attached to its subnodes.
-When an end of an element is encountered, the element currently on the top of the stack is closed and declared to be a rule of the IG.\\
-After reaching the end of the document, all IG rules are returned. Note that this approach creates "simple" initial grammar.
-
-\subsection{DTD processor}
-
-DTD processor registers itself into \emph{schema} input folder, \emph{dtd} file extension and declares it cannot handle other file extensions.\\
-To parse DTD files a 3rd party library is used: \jmodule{dtdparser}. % TODO anti Reference http://www.rpbourret.com/dtdparser/index.htm
-Translation from the object model in this library to our own is handled mostly by \code{DTD2RETranslator} class.\\
-Note that the initial grammar generated in this way is complex.
-Therefore, this processor checks for the \code{CAN\_HANDLE\_COMPLEX\_REGEXPS} capability of the simplifier following this IG generator: in positive case it simply returns the grammar, if the following module cannot handle complex regular expressions, DTD processor first invokes an \emph{Expander} (see \ref{section-expansion}).
-
-\subsection{XPath processor}
-
-This is a rather na\"{i}ve implementation of an XPath processor.
-It registers itself into \emph{query} input folder, \emph{xpath} file extension (text file containing one XPath query per line) and declares it cannot handle other file extensions.\\
-To parse XPath queries, standard support present in JDK is used. The relevant \code{XPathHandler} is the \code{XPathHandler} class.\\
-TODO vektor Explain what it does...
-
-\section{Expansion} \label{section-expansion}
-
-Expansion is used to convert complex regular expressions in initial grammar to simple ones (concatenations of tokens) for simplifiers that can handle only the simple form.
-Relevant interface encapsulating any implementation of such an expander is \code{Expander} in \jmodule{Base}.
-Reference implementation is \code{ExpanderImpl} in this module.
-Note that even though this implementation is retrieved using lookups, jInfer bundles only one such implementation and does not support choice among more of them.
-Anyone wishing to implement his own expander will thus have to either remove jInfer's implementation, or implement the usual module selection.
-% TODO vektor Is this documented somewhere?
-
-\subsection{ExpanderImpl internals}
-
-TODO vektor
-
-\section{Data flow}
-
-Flow of data in this module is following.
-\begin{enumerate}
-	\item IGGeneratorImpl walks over input files in a loop.
-	\item Each file gets processed by a processor based on its folder and extension.
-	\item Rules from all files are gathered in a single list (IG) and returned via a callback.
-\end{enumerate}
-
-\section{Extensibility}
-
-\jmodule{BasicIGG} can be easily extended to support a new input type: just create a class implementing \code{Processor}, annotate it as a service provider and implement any logic needed.
-\jmodule{XSDImporter} is an example of this.\\
-
-It is possible to replace the default \code{Expander} implementation: but as mentioned in \ref{section-expansion}, either the old implementation must be removed or module selection must be introduced.
+The main class implementing \code{RuleDisplayer} inference interface and simultaneously registered as its service provider is \code{BasicRuleDisplayer}.
 

### 2011-01-25 doc/jInfer/basicdtd/basicdtd.tex
Documentation: BasicDTDExporter and BasicRuleDisplayer.

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer BasicDTDExporter Module Description}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
\noindent Target audience: developers willing to extend jInfer, specifically hack the DTD export.

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & Matej Vitásek \\ \hline
Required tokens:       & none \\ \hline
Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\ \hline
Module dependencies:   & Base \\ \hline
Public packages:       & none \\ \hline
\end{tabular}

\section{Introduction}

This is a relatively simple implementation of a \jmodule{SchemaGenerator} exporting the inferred schema to DTD.

\section{Structure}

The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl}. Its \code{start} method first topologically sorts all rules (elements) it got in the simplified grammar on input. This toposorting is necessary to avoid using anything not yet defined in the resulting schema. Afterwards, it creates their DTD string represenation.\\

Export of a single element is handled in the \code{elementToString()} method. First the actual \code{<!ELEMENT \ldots>} tag is exported, after that its attributes in a \code{<!ATTLIST \ldots>} tag (if there are any).

\subsection{Element content export}

TODO anti

\subsection{Attribute export}

Code exporting attributes is in \code{attributeToString()}. First thing this method does is to assess the domain of a particular atribute: this is a map indexed by attribute values containing number of occurences for each such attribute. Type definition of an attribute is generated in the \code{DomainUtils.getAttributeType()} method. Based on a user setting, this might decide to enumerate all possible values of this attribute using the \code{(a|b|c)} notation, otherwise it just returns \code{\#CDATA}.\\
Attribute requiredness is assessed based on \code{required} metadata presence. If an attribute is not deemed required, it might have a default value: if a certain value is prominent in the attribute domain (based on user setting again), it is declared default.

\section{Data flow}

Flow of data in this module is following.
\begin{enumerate}
	\item \code{SchemaGeneratorImpl} topologically sorts elements (rules) it got on input.
	\item For each element, relevant portion of DTD schema is generated.
	\item String representation of the schema is returned along with the information that file extension should be "dtd".
\end{enumerate}

%\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-25 doc/jInfer/basicruledisplayer/basicruledisplayer.tex
Documentation.

@@
 This is fairly basic rule displayer draws rules as series of nested coloured rectangles.
+It does so by recursively descending down the regexp tree to draw the innermost nodes first, composing the outer nodes from them.
+User interface-wise, it creates a component with multiple panels: each one for a new grammar to display.
 
@@
 The main class implementing \code{RuleDisplayer} inference interface and simultaneously registered as its service provider is \code{BasicRuleDisplayer}.
+Main method of this class is \code{createDisplayer()}, which looks up the component, adds a new panel to it and renders the specified grammar in it.\\
+
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{class-diagram}
+	\caption{BasicRuleDisplayer class diagram} \label{figure-class-diagram}
+\end{figure}
+
+The diagram of classes involved in rule display is in figure \ref{figure-class-diagram}.
+The top component of the rule displayer has a few panels in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar once into an internal \code{Image}, then keeping it and drawing it onto panel's \code{Canvas} each time repainting is needed.\\
+
+Grammar is rendered in the \code{drawRules()} method by recursively rendering its rules and stacking their image representations into the image.
+Rule painting is handled in the \code{NodePainter} class, with most important method \code{drawNode()}.
+While the code itself is a nice recursion programming exercise, there is nothing of a special interest in it.\\
+
+All graphics used by \jmodule{BasicRuleDisplayer} is contained in the \code{cz.cuni.mff.ksi.jinfer.basicruledisplayer.graphics} package.
+
+\subsection{Settings}
+
+All settings provided by \jmodule{BasicRuleDisplayer} are NetBeans-wide.
+The options panel along with all the logic is in the \code{cz.cuni.mff.ksi.jinfer.basicruledisplayer.options} package.
+Available options include setting the maximum number of panels concurrently displayed, rules drawn, as well as maximum nesting level.
+Everything beyond these limits will be rendered as an icon suggesting omitted information.
+Furthermore, colours of various nodes can be set here.
 

### 2011-01-25 doc/jInfer/basicdtd/basicdtd.tex
Documentation.

@@
 
+\subsection{Preferences}
+
+All settings provided by \jmodule{BasicDTDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicdtd.properties} package.
+As mentioned before, it is possible to set the following.
+\begin{itemize}
+	\item Maximum attribute domain size which is exported as a list of all values (\code{(a|b|c)} notation).
+	\item Minimal ratio an attribute value in the domain needs to have in order to be declared default.
+\end{itemize}
+
 \section{Data flow}

### 2011-01-25 doc/jInfer/basicigg/basicigg.tex
Documentation: BasicIGG - process view schema.
Documentation.

@@
 After each file has been processed, the resulting grammar is returned by invoking the \code{finished} method of the callback argument.
+This process is illustrated in fig.
+\ref{figure-process}.
+
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{process-view}
+	\caption{Process view of BasicIGG} \label{figure-process}
+\end{figure}
 
@@
 
+\section{Preferences}
+
+All settings provided by \jmodule{BasicIGG} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicigg.properties} package. It is possible to set the following flags. 
+\begin{itemize}
+	\item Flag to keep values of attributes found in XML documents - unchecking this saves memory during the inference process, but loses all information about the attribute domain, which can lead to less expressive schema.
+	\item Flag to keep simple data values from XML documents - again, unchecking saves memory (potentially a lot, with input files containing a lot of text).
+Drawback might again be losing information that could be useful in schema generation.
+	\item Flag to stop on errors: if this flag is set, \jmodule{BasicIGG} will interrupt the inference if it comes across an error while extracting IG.
+Unsetting this flag allows \jmodule{BasicIGG} to ignore these errors and retrieve as much IG as possible.
+\end{itemize}
+
 \section{Data flow}

### 2011-01-25 doc/jInfer/architecture/architecture.tex
UML like adaptation, not done
Changed type of arrows to empty for uml-like diagrams
Documentation: Architecture: adding required metadata reference.

@@
 \end{figure}
+% This figure has been replaced by UML-like
+%\begin{figure}
+%\centering\includegraphics[scale=\myscale]{nodes_full}
+%\caption{XML representing interfaces and classes in detail} \label{nodes_full}
+%\end{figure}
 \begin{figure}
-\centering\includegraphics[scale=\myscale]{nodes_full}
-\caption{XML representing interfaces and classes in detail} \label{nodes_full}
+\centering\includegraphics[scale=\myscale]{nodes_full2}
+\caption{Class diagram for XML representing interfaces and classes} \label{nodes_full2}
 \end{figure}
@@
 \begin{itemize}
-	\item \code{from.xml}, filled in by BasicIGG module: means that this rule was created (originates from) from a XML document.
-	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
-	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
-	\item \code{is.sentinel}, filled in by  BasicIGG and XSDImport module: when importing XML document, one can build whole tree in memory (as in fig.
+	\item \code{from.xml}, filled in by \jmodule{BasicIGG} module: means that this rule was created (originates from) from a XML document.
+	\item \code{from.schema}, filled in by \jmodule{BasicIGG} and \jmodule{XSDImporter} modules: means that this rule was created (originates from) from a schema.
+	\item \code{from.query}, filled in by \jmodule{BasicIGG} and \jmodule{XSDImporter} modules: means that this rule was created from a query.
+	\item \code{is.sentinel}, filled in by \jmodule{BasicIGG} and \jmodule{XSDImporter} modules: when importing XML document, one can build whole tree in memory (as in fig.
 \ref{xml_example}) and then construct list of rules from it (thus saving memory).
@@
 Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
+	\item \code{required}, filled in by \jmodule{BasicIGG} and \jmodule{XSDImporter} modules: indicates that the attribute is required.
 \end{itemize}
-All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively mean true).
+All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively means "true").
 

### 2011-01-26 doc/jInfer/basicxsd/basicxsd.tex
Part of BasicXSDExporter documentation.

--- 2011-01-25 doc/jInfer/basicdtd/basicdtd.tex
@@
 
-\def\mftitle{jInfer BasicDTDExporter Module Description}
+\def\mftitle{jInfer BasicXSDExporter Module Description}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \maketitle
-\noindent Target audience: developers willing to extend jInfer, specifically hack the DTD export.
+\noindent Target audience: developers willing to extend jInfer, specifically hack the XSD export.
 
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
+Responsible developer: & Mário Mikula \\ \hline
 Required tokens:       & none \\ \hline
@@
 
-This is a relatively simple implementation of a \jmodule{SchemaGenerator} exporting the inferred schema to DTD.
+This is an implementation of a \jmodule{SchemaGenerator} exporting the inferred schema to XSD, providing basic features of the language.
 
@@
 The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl}.
-Its \code{start} method first topologically sorts all rules (elements) it got in the simplified grammar on input.
-This toposorting is necessary to avoid using anything not yet defined in the resulting schema.
-Afterwards, it creates their DTD string represenation.\\
+Process of export consists of two phases described in detail in later sections:
+\begin{enumerate}
+	\item Preprocessing
+	\item Own export to string representation of XSD
+\end{enumerate}
+Method \code{start} first creates instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
+Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such which elements should be globally defined and which element is the top level element (TODO rio how is this element called in XML??).\\
 
-Export of a single element is handled in the \code{elementToString()} method.
-First the actual \code{<!ELEMENT \ldots>} tag is exported, after that its attributes in a \code{<!ATTLIST \ldots>} tag (if there are any).
+Afterwards, it recursively traverses global elements followed by other elements starting at the top level element and for each it creates element's XSD string representation.\\
 
-\subsection{Element content export}
+\subsection{Preprocessing}
 
-TODO anti
+TODO rio
 
-\subsection{Attribute export}
+\subsection{Own export}
+
+TODO rio\\
 
@@
 \end{enumerate}
-
+\cite{ahonen}
 %\nocite{*}

### 2011-01-26 doc/jInfer/base/base.tex
Some cite{} corrections to base

@@
 
-Description of \jmodule{Base} structure will partially mirror its JavaDoc documentation. For more detailed information, refer to it directly.\\
+Description of \jmodule{Base} structure will partially mirror its JavaDoc documentation (\cite{javadoc}). For more detailed information, refer to it directly.\\
 
-\jmodule{Base} contains logic for Log4j initialization in \code{Installer} class.
+\jmodule{Base} contains logic for Log4j (\cite{log4j}) initialization in \code{Installer} class.
 Configuration of the overall logging granularity level (NetBeans options integration) is contained in \code{cz.cuni.mff.ksi.jinfer.base.options} package.
@@
 XML node representation is described by interfaces in \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes} package and more-less concrete implementations in \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes} package.\\
-Finite state automata representation is contained in the \cite{cz.cuni.mff.ksi.jinfer.base.automaton} package.\\
-Refer to \ref{archdoc} to get an accurate description of all these representations.\\
+Finite state automata representation is contained in the \code{cz.cuni.mff.ksi.jinfer.base.automaton} package.\\
+Refer to \cite{archdoc} to get an accurate description of all these representations.\\
 
@@
 They come in two flavours: the actual inference interface, and its callback.
-For a comprehensive description of them and their interaction refer again to \ref{archdoc}.\\
+For a comprehensive description of them and their interaction refer again to \cite{archdoc}.\\
 
@@
 This is not an inference module and there is no real data flow in it.
-Refer to \ref{archdoc} to understand the inference process described by interfaces from \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package.
+Refer to \cite{archdoc} to understand the inference process described by interfaces from \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package.
 
@@
 
-%\nocite{*}
+\nocite{*}
 \newpage

### 2011-01-26 doc/jInfer/basicigg/basicigg.tex
regenerating, ahonen cite to make bibtex happy
some \cite corrections to basicigg, only one note from to this document

@@
 This is an extensible implementation of the \jmodule{IGGenerator} inference interface. It is the only IG generator officialy shipped with jInfer.\\
-Make sure you understand the difference between the two types of initial grammar as described in \ref{archdoc}.
+Make sure you understand the difference between the two types of initial grammar as described in \cite{archdoc}.
 
@@
 A processor has to be registered as a service provider of the \code{Processor} interface from \jmodule{Base} module.
-Due to the nature of NBP lookups, each processor is internally kept as a singleton, and should not use its inner state (refer to the chapter Lookups in \ref{archdoc}).
-Note that the factory pattern is not used here.\\ % TODO anti reference to a chapter?
+Due to the nature of NBP lookups, each processor is internally kept as a singleton, and should not use its inner state (refer to the chapter Lookups in \cite{archdoc}).
+Note that the factory pattern is not used here (see \cite[Factory]{archdoc}).\\ % TODO anti reference to a chapter number when factory pattern is moved to arch
 Each processor declares the class of inputs (documents, schemas, queries) and file extensions it is able to handle by implementing methods \code{getFolder()}, \code{getExtension()} and \code{processUndefined()}.
@@
 DTD processor registers itself into \emph{schema} input folder, \emph{dtd} file extension and declares it cannot handle other file extensions.\\
-To parse DTD files a 3rd party library is used: \jmodule{dtdparser}. % TODO anti Reference http://www.rpbourret.com/dtdparser/index.htm
+To parse DTD files a 3rd party library is used: \jmodule{dtdparser} (\cite{dtdparser}).
 Translation from the object model in this library to our own is handled mostly by \code{DTD2RETranslator} class.\\
@@
 % TODO vektor Is this documented somewhere?
+anti: NO, should be in developer doc with preferences panel and so on!!!
 
@@
 For each word from this expansion a new \code{Element} is created: letters of this word are encapsulated in tokens, these tokens are put in a concatenation that constitutes subnodes of this new element.
-If the letter is an element itself, it is not recursively expanded - it is just declared to be a sentinel.\\ % TODO anti Link to architecture
+If the letter is an element itself, it is not recursively expanded - it is just declared to be a sentinel (see \cite[section 3.3]{archdoc}).\\
 

### 2011-01-26 doc/jInfer/runner/runner.tex
Documentation: Runner module.

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer Runner Module Description}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
\noindent Target audience: developers willing to extend jInfer, specifically hack (or expand) the inference process.

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & Matej Vitásek \\ \hline
Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\
 & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\
 & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\ 
 & org.openide.windows.IOProvider \\ \hline
Provided tokens:       & none \\ \hline
Module dependencies:   & Base \\ \hline
Public packages:       & cz.cuni.mff.ksi.jinfer.runner \\
  & cz.cuni.mff.ksi.jinfer.runner.properties \\ \hline
\end{tabular}

\section{Introduction}

\jmodule{Runner} it the module responsible for inference process. The fact that the inference consists of exactly 3 steps as described in TODO link architecture is hard-wired in jInfer via this module.

\section{Structure and operation}

The main class responsible for the inference run is the \code{Runner} in \code{cz.cuni.mff.ksi.jinfer.runner} package. During its construction it loads the properties of the currently running project (via the \code{RunningProject} class) to find out which modules are selected for the inference. These modules are looked up and remembered - each new inference run should therefore use a new instance of runner.\\

The only public method in \code{Runner} is, unsurprisingly, \code{run()}. This method will start the first step of the inference process by invoking the selected \jmodule{IGGenerator}'s \code{start()} method, as described in TODO link inference process in architecture. Callback methods \code{finishedIGGenerator()}, \code{finishedSimplifier()} and \code{finishedSchemaGenerator()} are responsible for invoking the following stages of inference, or in the latter case for presenting the resulting schema to the user and terminating the inference process.\\

Invocation of every step in the process is encapsulated in a NetBeans \emph{task}: this is the responsibility of \code{runAsync()} method. First of all, this means that all the work is done in an asynchronous thread independent from the GUI. Second, NBP presents each such task as a progressbar in the bottom right corner of the window, and allows the user to cancel it. Should this happen, the currently running module detects this  by checking for \code{Thread.interrupted()} and responds by throwing an \code{InterruptedException}. \code{Runner} catches this exception, terminates the inference and informs the user (\code{interrupted()} method).\\

Furthermore, should any unexpected exception occur while running one of the modules, this will get caught in \code{Runner} again. Inference will be interrupted and user will be notified - this is the responsibility of \code{unexpected()} method.\\

Finally, the generated schema is annotated in the end with a comment stating the current date and time, and the modules used in the process. This is the responsibility of \code{getCommentedSchema()} method.

\subsection{Settings}

\jmodule{Runner} has NetBeans-wide settings determining what should happen after the schema is inferred and which rule displayer should be used. The options panel along with all the logic is in the \code{cz.cuni.mff.ksi.jinfer.runner.options} package.

\subsection{Preferences}

\jmodule{Runner} naturally has project-wide preferences for selection of inference modules. The panel and its logic is in \code{cz.cuni.mff.ksi.jinfer.runner.properties} package.

\section{Extensibility}

For certain inference algorithms it might be necessary to completely change the number or order of inference modules. For example, it might be necessary to have a dynamic inference with arbitrary number of iterations over the same module (modules). In this case, \jmodule{Runner} can serve as a template: adding e.g. a cleaner between \jmodule{Simplifier} and \jmodule{SchemaGenerator} would mean just copy-pasting members and methods currently associated to \jmodule{Simplifier}.\\

%\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-01-26 doc/jInfer/basicigg/basicigg.tex
fixed overfull hboxes by package names

@@
 %\usepackage[scale=0.9]{tgheros}
+\usepackage{xstring}
 
@@
 \newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\punts}[1]{\StrSubstitute{#1}{.}{.\.}}
+\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
+\def\.{\discretionary{}{}{}}
 \newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
@@
 Note that the initial grammar generated in this way is complex.
-Therefore, this processor checks for the \code{CAN\_HANDLE\_COMPLEX\_REGEXPS} capability of the simplifier following this IG generator: in positive case it simply returns the grammar, if the following module cannot handle complex regular expressions, DTD processor first invokes an \emph{Expander} (see \ref{section-expansion}).
+Therefore, this processor checks for the \code{CAN\_\.HANDLE\_\.COMPLEX\_\.REGEXPS} capability of the simplifier following this IG generator: in positive case it simply returns the grammar, if the following module cannot handle complex regular expressions, DTD processor first invokes an \emph{Expander} (see \ref{section-expansion}).
 

### 2011-01-26 doc/jInfer/basicxsd/basicxsd.tex
just cite repair
fixed overfull hboxes by package names
Documenting BasicXSDExporter.

@@
 \newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
+\def\.{\discretionary{}{}{}}
 \newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
@@
 Method \code{start} first creates instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
-Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such which elements should be globally defined and which element is the top level element (TODO rio how is this element called in XML??).\\
+Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such which elements should be globally defined and which element is the top level element (TODO rio how is this element called in XML??).
+This instance is kept as a member variable of module to be used during the whole export process.\\
 
-Afterwards, it recursively traverses global elements followed by other elements starting at the top level element and for each it creates element's XSD string representation.\\
+Afterwards, \code{start} method recursively traverses global elements followed by other elements starting at the top level element and for each it creates element's XSD string representation.\\
 
@@
 
-TODO rio
+As mentioned before, preprocessing is implemented in \code{Preprocessor} class and its functions are following.
+\begin{itemize}
+	\item Decide which elements should be defined globally.
+	\item Remove unused elements.
+	\item Find the top level element.
+	\item Find an instance of element by its name.
+\end{itemize}
+
+Constructor of \code{Preprocessor} class gets elements and a number, defining minimal number of occurrences of an element to be defined.
+It first topologically sorts input elements to decide which one is the top level element.
+Afterwards, it counts occurrences of the elements and removes unused ones (those which did not occurred).
+Finally, for each element it decides whether mark it as a global one or not.
+An element is considered global if its occurrence count is greater than or equal number of occurrences supplied on input.\\
+
+Results of preprocessing are provided by public methods of \code{Preprocessor} class. For details see their JavaDoc.\\
 
@@
 
-TODO rio\\
+Own XSD export is performed in module's \code{start} function right after the preprocessing.\\
+
+Useful helper class to handle indentation of text in a resulting XSD is named \code{Indentator}.
+Instance of this class is a member variable of the module (alike instance of \code{Preprocessor}), it holds text appended to it and keeps indentation level state.
+Text can be appended without indentation (method \code{append}) or indented (method \code{indent}).
+Level of indentation can be incremented or decremented by methods \code{increaseIndentation} and \code{decreaseIndentation}.
+At the end of export, when textual representation of each element has been appended to the \code{Indentator}, \code{Indentator}'s method \code{toString} will return string representation of resulting XSD.\\
+
+First, global elements are exported. For each element, its type is defined as a global type. TODO rio example
+
+After global elements, others are exported.
 
@@
 \end{enumerate}
-\cite{ahonen}
-%\nocite{*}
+
+\nocite{*}
 \newpage

### 2011-01-27 doc/jInfer/architecture/architecture.tex
fixed overfull hboxes by package names
moved factory pattern to arch from twostep

@@
 \newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
+\def\.{\discretionary{}{}{}}
 \newcommand{\jmodule}[1]{\emph{#1}}
@@
 
-\noindent \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\
+\noindent\emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}\\
 
@@
 
-Regular expression is implemented as class \code{Regexp<T>} with supplementing classes \code{RegexpInterval} and \code{RegexpType}.
+Regular expression is implemented as class \code{Regexp<T>} with supplementing classes \code{RegexpInterval} and \code{Reg\.exp\.Type}.
 Each \code{Regexp<T>} instance has one of the enum \code{RegexpType} type:
@@
 \end{itemize}
-Type of regexp is held in \code{type} member in class \code{Regexp<T>} and can be tested by calling methods \code{isLambda()}, \code{isToken()} etc.
+Type of regexp is in \code{type} member in class \code{Regexp<T>} and can be tested by calling methods \code{isLa\.mb\.da()}, \code{isTo\.ken()}, \code{isCon\.ca\.te\.na\.tion()}, \code{isAlt\.er\.na\.tion()}, \code{isPer\.mu\.ta\.tion()}.
 
@@
 These constraits are checked by constructors, so the best way to construct new regexps is by using 
-methods \code{getToken(), getConcatenation()} etc.
+methods \code{get\.To\.ken(), get\.Con\.ca\.te\.na\.tion()} etc.
 
@@
 
-Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full}.
-In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{AbstractStructuralNode} variables.
+Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full2}.
+In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{Abstr\.act\.Stru\.ctu\.ral\.Node} variables.
 And \code{SimpleData} and \code{Attribute} have methods \code{getContentType()} and \code{getContent()} to work with content model.
@@
 \end{itemize}
-A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{SimplifierImpl}.
+A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{Sim\.pli\.fier\.Impl}.
 Annotation required for the framework to recognize such a class as an inference module is the following
@@
 
-For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCapabilities()}.
+For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCa\.pa\.bi\.li\.ties()}.
 This method returns a list of strings representing the abilities, or "capabilities" of this module.
@@
 
+See \cite{lookup_api} as the official documentation to lookup API.
+
+\subsection{Factory pattern}
+\begin{figure}
+  \centering
+	\subfigure[Same instance of class returned by NetBeans in two successive inference runs]
+	{
+	  \centering\includegraphics[scale=\myscale]{factory_pattern1}
+    \label{factory_pattern1}
+  }
+	\subfigure[Factory solution of singleton lookup class problem]
+	{
+	  \centering\includegraphics[scale=\myscale]{factory_pattern2}
+    \label{factory_pattern2}
+  }
+\end{figure}
 At this point it is important to understand that when using lookups this way, all service providers are kept as singletons. 
-In case a provider needs to change its state during providing the service (and needs to be ``fresh" each time it is retrieved), it is best to convert this to a factory design pattern: the service provider registered is just a factory returning instances that do the actual work.
+We ilustrate this behaviour on \ref{factory_pattern1}, user runs inference first time, classes are looked up and used.
+When user clicks run button again, and same instances of classes are returned by lookups.
 
-See \cite{lookup_api} as the official documentation to lookup API.
+In case a provider needs to change its state during providing the service (and needs to be ``fresh'' each time it is retrieved), there are two solutions:
+\begin{itemize}
+	\item force each service providing class (submodule), to implement
+some sort of cleanup method, that would restart it into fresh state and make it ready to use by another 
+inference run,
+	\item or create new instances of stateful classes in each inference run and make service providing classes (those which NetBeans return by lookups) factories.
+\end{itemize}
+We decided for the latter approach (as ilustrated on fig. \ref{factory_pattern2}
+It enables us not only to produce fresh class instance in each inference run, but also factory classes implement obligatory
+module methods such as \code{getName}, \code{getDescription} and so on, which are really same in each inference run.
+Each submodule is then defined by (at least) two interfaces.
+One is the factory interface, like this one (example from \jmodule{TwoStepSimplifier}:
+\begin{verbatim}
+public interface AutomatonSimplifierFactory extends 
+                     NamedModule, Capabilities, UserModuleDescription {
+  <T> AutomatonSimplifier<T> create();
+}
+\end{verbatim}
+Second is worker interface, in the example it is \code{AutomatonSimplifier<T>} interface.
+
+This pattern is used thorough jInfer modules to divide module into submodules.
 

### 2011-01-27 doc/jInfer/twostep/twostep.tex
fixed overfull hboxes by package names
moved factory pattern to arch from twostep

@@
 
+\section{Introduction}
+With \jmodule{TwoStepSimplifier} we provide user complex solution of infering problem.
+This is the only one simplifier shipped with jInfer, but it has many replaceable submodules of which we implement more alternatives in basic bundle.
+The best we implement is \cite{ahonen} $k,h$-context method od infering grammar rules.
+This process is split into submodules with generic automaton implementation, so it is easy to plug-in any other automaton merging state algorithm.
+In this document, we will walk over all modules and submodules in simplifier and describe them briefly.
+
 \section{Using of factory pattern}
-TODO anti/vektor move to architecture as described in twostepTODO.pdf
-\begin{figure}
-  \centering
-	\subfigure[Same instance of class returned by NetBeans in two successive inference runs]
-	{
-	  \centering\includegraphics[scale=\myscale]{factory_pattern1}
-    \label{factory_pattern1}
-  }
-	\subfigure[Factory solution of singleton lookup class problem]
-	{
-	  \centering\includegraphics[scale=\myscale]{factory_pattern2}
-    \label{factory_pattern2}
-  }
-\end{figure}
-\jmodule{TwoStepSimplifier} is divided into three submodules, which are then divided into submodules and so on.
-Each submodule is simply class, properly anotated.
-User selects proper classes - submodules to work in chain on inference.
-Classes are then, in runtime looked up by using NetBeans lookup mechanism (see \cite[p. 16]{archdoc}).
-But NetBeans holds one instance of each class, that may be looked upon.
-When user runs inference first time, classes are looked up and used.
-Maybe class members are initialised, instances are in some state after inference completes.
-Then, user clicks run button again, and same instances of classes are returned by lookups.
-Oops, theese are not freshly created instances, using them as they are may cause harm.
-The problem is ilustrated on fig. \ref{factory_pattern1}.
-
-There are basically two solutions:
-\begin{itemize}
-	\item force each class (submodule) that is being looked up in inference process, to implement
-some sort of cleanup method, that would restart it into fresh state and make it ready to use by another 
-inference run,
-	\item or create new instances of theese classes in each inference run and make singleton classes (those which NetBeans return by lookups) their factory classes.
-\end{itemize}
-We decided for the latter approach (as ilustrated on fig. \ref{factory_pattern2}
-It enables us not only to produce fresh class instance in each inference run, but also factory classes implement obligatory
-module methods such as \code{getName}, \code{getDescription} and so on, which are really same in each inference run.
-Each submodule is then defined by (at least) two interfaces.
-One is the factory interface, like this one:
+But in first, we will stay a while at factory pattern usage.
+In \jmodule{TwoStepSimplifier} we use factory pattern (as described in \cite{archdoc}) to divide module into submodules.
+As service providing classes are kept singleton, we make them factories:
 \begin{verbatim}
@@
 \end{verbatim}
-It extends \code{NamedModule} with its modules describing methods:
+The real submodule interface (called familiarly worker interface) is the \code{AutomatonSimplifier<T>} interface (of wich instance is returned by factory).
+It is now not important how worker interface looks like, lets examine the factory.
+
+In \code{TwoStepSimplifier} we design service providing factory interfaces so that they extends extends \code{NamedModule} with module describing methods:
 \begin{verbatim}
@@
 \end{verbatim}
-It extends \code{Capabilities}, thus each module have to answer, if it has some capabilities.
-And, in TwoStep we often extend \code{UserModuleDescription}, which defines method:
+And to extend \code{Capabilities}, thus each module have to answer, if it has some capabilities.
+And to extend \code{UserModuleDescription} interface, which defines method:
 \begin{verbatim}
@@
 
-Second interface is the one, which is returned by \code{create()} method.
-That is the interface, which defines real work cycle with submodule (in example it is \code{AutomatonSimplifier<T>}).
-We call this interface the worker interface of submodule.
-Usage of this factory pattern follows the routine:
+Our usage of this factory pattern follows the routine:
 \begin{verbatim}
@@
 ...
-give some work to autSmp
+result= autSmp.work(something_to_process);
 \end{verbatim}
-If our module has some submodules, we often implement lookups for submodules implementations in our own factory create method.
-Worker class receives factories of all submodules it needs as a constructor parameters.
+If one module has some submodules, we implement lookups for submodule implementations in our own factory \code{create} method.
+Worker class receives factories of all of its submodules as a constructor parameters.
 
-Lets look on \code{AutomatonMergingStateFactory}, that is factory of module, which has AutomatonSimplifier as submodule.
+Lets look on \code{AutomatonMergingStateFactory}, that is factory of module, which has \jmodule{AutomatonSimplifier} as submodule.
 Its create method looks like this (shortened): 
@@
 
-\section{TwoStepSimplifier module}
-\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
-\begin{enumerate}
-	\item clustering of element instances into clusters of (probably) same elements
-	\item infering regular expression for each element from examples of element contents taken from all elements in cluster
-\end{enumerate}
-Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule.
-We will examine both of them.
-There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
-Modules are drawn on fig. \ref{twostep_modules}.
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{twostep_modules}
-	\caption{Submodules of TwoStep simplifier} \label{twostep_modules}
-\end{figure}
-We provide one \jmodule{Clusterer} and 4 \jmodule{ClusterProcessor} implementations.
-Each of those will be explained further in this document.
-
+\section{Structure}
 \jmodule{TwoStepSimplfier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
 Module is implemented in two classes: \code{TwoStepSimplifier} (main logic), \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
-Factory class implements \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
-Thus has main method called \code{start} which receives Initial Grammar in form of
+Latter one is registered as service provider (and implements) of \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
+
+It has main method called \code{start} which receives Initial Grammar in form of
 \begin{verbatim}
@@
 
-\code{TwoStepSimplifierFactory} in method \code{start()} created new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
+\code{TwoStepSimplifierFactory} in method \code{start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
 Then it calls \code{simplify(initialGrammar)} method on this instance and its result passes as parameter to callback function.
 Thats all, the magic comes in \code{TwoStepSimplifier} class.
+
+\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
+\begin{enumerate}
+	\item clustering of element instances into clusters of (probably) same elements
+	\item infering regular expression for each element from examples of element contents taken from all elements in cluster
+\end{enumerate}
+Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule.
+We will examine both of them.
+There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
+Modules are drawn on fig. \ref{twostep_modules}.
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{twostep_modules}
+	\caption{Submodules of TwoStep simplifier with their implementations} \label{twostep_modules}
+\end{figure}
+We provide one \jmodule{Clusterer} and 4 \jmodule{ClusterProcessor} implementations.
+Each of those will be explained further in this document.
 
@@
 
-TODO sentinel cant be on left side
-TODO sentinel processing
+TODO anti sentinel cant be on left side
+TODO anti sentinel processing
 
@@
 
-\subsection{ClusterProcessor submodule}
+\subsection{ClusterProcessor module}
 Cluster processor takes rules of one cluster of element and somehow obtains regular expression for that set of elements.

### 2011-01-27 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - flooooooooooooow

@@
 
-\subsection{Clusterer module}
+\section{Preferences}
+All settings provided by \jmodule{TwoStepSimplifier} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.twostep.properties} package.
+It is possible to:
+\begin{itemize}
+	\item select \jmodule{Clusterer} submodule implementation from those installed
+	\item select \jmodule{ClusterProcessor} submodule implementation from those installed
+	\item select \jmodule{RegularExpressionCleaner} submodule implementation from those installed
+\end{itemize}
+
+\section{Data flow}
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{twostep_flow}
+	\caption{TwoStep data flow} \label{twostep_flow}
+\end{figure}
+\begin{enumerate}
+	\item On input there is Initial Grammar, which is ruled into \jmodule{Clusterer}.
+	\item Then, each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
+	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
+	\item From each cluster processing, regular expression is added to the list of simplified grammar rules.
+	\item List of simplifier rules is returned (simplified grammar).
+\end{enumerate}
+Process is illustrated on fig. \ref{twostep_flow}
+
+\section{Extensibility}
+If you are willing to, you can replaced each of three submodules of \jmodule{TwoStepSimplifier}.
+It may be useful to replace default \jmodule{ClustererWithAttributes} implementation to clever one.
+And regular expression cleaning will be probably target of adding new implementations, too.
+We recommend replacing other parts at lower levels, however.
+You will need to replace \jmodule{ClusterProcessor} only if you want to write infering method that is not based on merging state algorithm (or NFA's at all).
+If this is not your case, read the rest of this document and find proper place where to place your logic.
+
+\section{Clusterer module}
 Lets start by example, let input document(s) contain XML:

### 2011-01-27 doc/jInfer/twostep/twostep.tex
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa, kkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkk

@@
 
-\def\mftitle{jInfer TwoStepSimplifier design and implementation}
+\def\mftitle{jInfer TwoStepSimplifier Module Description}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 
+%#### INTRO ############################################
 \section{Introduction}
@@
 
-\section{Using of factory pattern}
+%#### FACTORY ############################################
+\section{Using of factory pattern} \label{section_factory}
 But in first, we will stay a while at factory pattern usage.
@@
 
+%#### STRUCTURE ############################################
 \section{Structure}
@@
 
-\section{Preferences}
-All settings provided by \jmodule{TwoStepSimplifier} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.twostep.properties} package.
-It is possible to:
-\begin{itemize}
-	\item select \jmodule{Clusterer} submodule implementation from those installed
-	\item select \jmodule{ClusterProcessor} submodule implementation from those installed
-	\item select \jmodule{RegularExpressionCleaner} submodule implementation from those installed
-\end{itemize}
-
-\section{Data flow}
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{twostep_flow}
-	\caption{TwoStep data flow} \label{twostep_flow}
-\end{figure}
-\begin{enumerate}
-	\item On input there is Initial Grammar, which is ruled into \jmodule{Clusterer}.
-	\item Then, each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
-	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
-	\item From each cluster processing, regular expression is added to the list of simplified grammar rules.
-	\item List of simplifier rules is returned (simplified grammar).
-\end{enumerate}
-Process is illustrated on fig. \ref{twostep_flow}
-
-\section{Extensibility}
-If you are willing to, you can replaced each of three submodules of \jmodule{TwoStepSimplifier}.
-It may be useful to replace default \jmodule{ClustererWithAttributes} implementation to clever one.
-And regular expression cleaning will be probably target of adding new implementations, too.
-We recommend replacing other parts at lower levels, however.
-You will need to replace \jmodule{ClusterProcessor} only if you want to write infering method that is not based on merging state algorithm (or NFA's at all).
-If this is not your case, read the rest of this document and find proper place where to place your logic.
-
-\section{Clusterer module}
-Lets start by example, let input document(s) contain XML:
-\begin{verbatim}
-<person name="john" surname="smith">
-  <info>
-    Some text
-    <note/>
-  </info>
-</person>
-<inform>
-  Another text
-  <note/>
-</inform>
-<person>
-    <information>
-      Some text
-      <note/>
-    </information>
-</person>
-\end{verbatim}
-We examine info-like named elements.
-If we cluster elements sipmly by their name, we get one cluster with info element, another cluster with inform element and another one with information element.
-As \cite{1802522} suggests, we should consider element content in clustering process.
-If we take care of element content, elements info, inform and information would look same (in means of node tree inside them - text node and element note).
-Elements information and info have even same context (inside person element).
-
+%#### STRUCTURE->Clusterer ############################################
+\subsection{Clusterer module}
 We implement clustering in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
@@
 
-Basic work usage of clusterer is:
+Basic work usage with clusterer is:
 \begin{verbatim}
@@
 
+%#### STRUCTURE->Clusterer->WithAttributes ############################################
 \subsubsection{ClustererWithAttributes extended interface}
@@
 
-Finally, we use this scheme to implement clusterer \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
+We provide one implementation of this interface.
+It is described in \ref{section_Iname}.
+
+%#### STRUCTURE->ClusterProcessor ############################################
+\subsection{ClusterProcessor module}
+Cluster processor takes rules of one cluster of element and somehow obtains regular expression for that set of elements.
+It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for the resulting element) and with subnodes set to regular expression infered.
+It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to resulting element.
+
+Worker interface itself is defined as follows:
+\begin{verbatim}
+public interface ClusterProcessor<T> {
+  T processCluster(
+          final Clusterer<T> clusterer,
+          final List<T> rules
+          ) throws InterruptedException;
+}
+\end{verbatim}
+Maybe you are asking, why cluster processor is given the clusterer instance.
+Rules themself contain information about which elements to process, but clusterer has more information
+about the topic.
+Clusterer can tell you representant for any element in whole input (not only those elements in rules, but also those that may be on right side of rules), clusterer (if it is with attributes) has information about attributes <of each cluster.
+
+We describe each cluster processor implementation we've got in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.
+
+%#### STRUCTURE->RegularExpressionCleaner ############################################
+\subsection{RegularExpressionCleaner module}
+Last we examine \code{RegularExpressionCleaner} interface.
+Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
+For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such like $(name, (person, id))$.
+To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
+Interface is definition is straightforward:
+\begin{verbatim}
+public interface RegularExpressionCleaner<T> {
+  Regexp<T> cleanRegularExpression(final Regexp<T> regexp);
+}
+\end{verbatim}
+Given regular expression, return regular expression.
+Converting of regular expression is commonly a recursive task and all our implementations work this way.
+
+%#### PREFERENCES ############################################
+\section{Preferences}
+All settings provided by \jmodule{TwoStepSimplifier} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.twostep.properties} package.
+It is possible to:
+\begin{itemize}
+	\item select \jmodule{Clusterer} submodule implementation from those installed
+	\item select \jmodule{ClusterProcessor} submodule implementation from those installed
+	\item select \jmodule{RegularExpressionCleaner} submodule implementation from those installed
+\end{itemize}
+
+%#### DATA FLOW ############################################
+\section{Data flow}
+\begin{figure}
+	\centering\includegraphics[scale=1]{twostep_flow}
+	\caption{TwoStep data flow} \label{twostep_flow}
+\end{figure}
+\begin{enumerate}
+	\item On input there is Initial Grammar, which is ruled into \jmodule{Clusterer}.
+	\item Then, each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
+	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
+	\item From each cluster processing, regular expression is added to the list of simplified grammar rules.
+	\item List of simplifier rules is returned (simplified grammar).
+\end{enumerate}
+Process is illustrated on fig. \ref{twostep_flow}
+
+%#### EXTENSIBILITY ############################################
+\section{Extensibility}
+If you are willing to, you can replaced each of three submodules of \jmodule{TwoStepSimplifier}.
+It may be useful to replace default \jmodule{ClustererWithAttributes} implementation to clever one.
+And regular expression cleaning will be probably target of adding new implementations, too.
+We recommend replacing other parts at lower levels, however.
+You will need to replace \jmodule{ClusterProcessor} only if you want to write infering method that is not based on merging state algorithm (or NFA's at all).
+If this is not your case, read the rest of this document and find proper place where to place your logic.
+
+%#### SUBMODULES ############################################
+\section{Submodules implementations}
+%#### SUBMODULES->Iname ############################################
+\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
+We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
 Clustering is done based on elements \code{getName} equality test (ignore case).
@@
 
-\subsection{ClusterProcessor module}
-Cluster processor takes rules of one cluster of element and somehow obtains regular expression for that set of elements.
-It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for the resulting element) and with subnodes set to regular expression infered.
-It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to resulting element.
-
-Worker interface itself is defined as follows:
-\begin{verbatim}
-public interface ClusterProcessor<T> {
-  T processCluster(
-          final Clusterer<T> clusterer,
-          final List<T> rules
-          ) throws InterruptedException;
-}
-\end{verbatim}
-Maybe you are asking, why cluster processor is given the clusterer instance.
-Rules themself contain information about which elements to process, but clusterer has more information
-about the topic.
-Clusterer can tell you representant for any element in whole input (not only those elements in rules, but also those that may be on right side of rules), clusterer (if it is with attributes) has information about attributes <of each cluster.
-
-We will now shortly describe each cluster processor implementation we've got.
-
-\subsubsection{PassRepresentant cluster processor}
+%#### SUBMODULES->PassRepresentant ############################################
+\subsection{PassRepresentant (ClusterProcessor)} \label{section_PassRepresentant}
 Simple example to read. For each cluster return its representant as a rule to be in schema.
@@
 
-\subsubsection{Alternations cluster processor}
+%#### SUBMODULES->Alternations ############################################
+\subsection{Alternations (ClusterProcessor)} \label{section_Alternations}
 This processor simply gets all right sides from elements in cluster, puts them in one big list and creates alternation regular expression with this list as children.
@@
 
-\subsubsection{Trie cluster processor}
+%#### SUBMODULES->Trie ############################################
+\subsection{Trie (ClusterProcessor)} \label{section_Trie}
 This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie") of them.
@@
 
-\subsubsection{AutomatonMergingState cluster processor}
+%#### SUBMODULES->AutomatonMergingState ############################################
+\subsection{AutomatonMergingState (ClusterProcessor)} \label{section_AutomatonMergingState}
 \begin{figure}
@@
 	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_modules}
+\end{figure}
+\begin{figure}
+	\centering\includegraphics[scale=1]{automaton_merging_state_flow}
+	\caption{Data flow of AutomatonMergingState cluster processor} \label{automaton_merging_state_flow}
 \end{figure}
@@
 Its job is to derive regular expression from automaton, such that automaton and regular expression represents same language.
+The data flow is on fig. \ref{automaton_merging_state_flow}.
+
+\subsubsection{AutomatonSimplifier module}
+\jmodule{AutomatonSimplifier} module has its worker interface defined as:
+\begin{verbatim}
+public interface AutomatonSimplifier<T> {
+  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
+                        final SymbolToString<T> symbolToString) throws InterruptedException;
+}
+\end{verbatim}
+Given input automaton it returns automaton, there is nothing magical on it.
+Interface is defined as generic, implementation don't have to be, but there is no reason to not make them generic.
+When implementation needs to present automaton to user, however, it needs some string representation of symbols to display on automaton transitions.
+For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
+It is responsible for converting symbol to string.
+To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplfifier} submodule.
+
+By this genericity, simplifier can be used to perfom same algorithms for simplifying not only to infer rules for elements, but maybe also to infer patterns strings for content model of attributes.
+The actual implementation of this still waits for its developer to come.
+
+TODO how automaton works?
+
+Lets go on to implementations of \jmodule{AutomatonSimplifier}.
+
+\subsubsection{RegexpAutomatonSimplifier submodule}
+Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain regular expression from given automaton.
+Interface for doing this is thus simple:
+\begin{verbatim}
+public interface RegexpAutomatonSimplifier<T> {
+  Regexp<T> simplify(final RegexpAutomaton<T> inputAutomaton, 
+                     final SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
+}
+\end{verbatim}
+Once again we are encountering \code{symbolToString} with same purpose as before, if submodule would need to present automaton to user
+it has to be able to convert symbols of (at runtime) unknown type to strings.
+
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{state_removal_modules}
+	\caption{Modules of StateRemoval.} \label{state_removal_modules}
+\end{figure}
 
@@
 
-\subsubsection{AutomatonSimplifier submodule}
-\jmodule{AutomatonSimplifier} module has its worker interface defined as:
-\begin{verbatim}
-public interface AutomatonSimplifier<T> {
-  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
-                        final SymbolToString<T> symbolToString) throws InterruptedException;
-}
-\end{verbatim}
-Given input automaton it returns automaton, there is nothing magical on it.
-Interface is defined as generic, implementation don't have to be, but there is no reason to not make them generic.
-When implementation needs to present automaton to user, however, it needs some string representation of symbols to display on automaton transitions.
-For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
-It is responsible for converting symbol to string.
-To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplfifier} submodule.
+\subsection{NestedConcatention (RegularExpressionCleaner)}
+This cleaner does exactly what we have just described.
+It converts nested concatenations into flat ones.
 
-By this genericity, simplifier can be used to perfom same algorithms for simplifying not only to infer rules for elements, but maybe also to infer patterns strings for content model of attributes.
-The actual implementation of this still waits for its developer to come.
+\subsection{Null (RegularExpressionCleaner)}
+This cleaner just returns regexp it receives and does nothing.
+Plug this one into chain, if you want to omit this step of inference.
 
-TODO how automaton works?
+\subsection{EmptyChildren (RegularExpressionCleaner)}
+Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
+For example some pitty inferring method produces unnecessary empty regexps in regexp tree as:
+$$
+(name, (), (person, id))
+$$
+There is one empty concatenation/alternation/permutation with no sense.
+This is wiped out by this cleaner.
 
-Lets go on to implementations of \jmodule{AutomatonSimplifier}.
+\subsection{Chained (RegularExpressionCleaner)}
+Enables user to chain more cleaners with output from first one plugged as input into second one and so one.
+Thus regular expression from inference can pass through \jmodule{EmptyChildren} and them through \jmodule{NestedConcatenation} cleaners and then it is returned.
 
@@
 
-\subsubsection{RegexpAutomatonSimplifier submodule}
-Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain regular expression from given automaton.
-Interface for doing this is thus simple:
-\begin{verbatim}
-public interface RegexpAutomatonSimplifier<T> {
-  Regexp<T> simplify(final RegexpAutomaton<T> inputAutomaton, 
-                     final SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
-}
-\end{verbatim}
-Once again we are encountering \code{symbolToString} with same purpose as before, if submodule would need to present automaton to user
-it has to be able to convert symbols of (at runtime) unknown type to strings.
-
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{state_removal_modules}
-	\caption{Modules of StateRemoval.} \label{state_removal_modules}
-\end{figure}
 \paragraph{StateRemoval regexp automaton simplifier}
@@
 
-\subsection{RegularExpressionCleaner module}
-Last we examine \code{RegularExpressionCleaner} interface.
-Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
-For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such like $(name, (person, id))$.
-To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
-Interface is definition is straightforward:
-\begin{verbatim}
-public interface RegularExpressionCleaner<T> {
-  Regexp<T> cleanRegularExpression(final Regexp<T> regexp);
-}
-\end{verbatim}
-Given regular expression, return regular expression.
-Converting of regular expression is commonly a recursive task and all our implementations work this way.
-Lets look at them.
-
-\paragraph{NestedConcatention}
-This cleaner does exactly what we have just described.
-It converts nested concatenations into flat ones.
-
-\paragraph{Null}
-This cleaner just returns regexp it receives and does nothing.
-Plug this one into chain, if you want to omit this step of inference.
-
-\paragraph{EmptyChildren}
-Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
-For example some pitty inferring method produces unnecessary empty regexps in regexp tree as:
-$$
-(name, (), (person, id))
-$$
-There is one empty concatenation/alternation/permutation with no sense.
-This is wiped out by this cleaner.
-
-\paragraph{Chained}
-Enables user to chain more cleaners with output from first one plugged as input into second one and so one.
-Thus regular expression from inference can pass through \jmodule{EmptyChildren} and them through \jmodule{NestedConcatenation} cleaners and then it is returned.
-
 Whole TwoStep submodules structure is on fig. \ref{twostep_modules_complete}.

### 2011-01-27 doc/jInfer/twostep/twostep.tex
aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa, kkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkkk

@@
 
-Lets go on to implementations of \jmodule{AutomatonSimplifier}.
+Implementations of \jmodule{AutomatonSimplifier} are discussed in \ref{section_Greedy} and \ref{section_UserInteractive}.
 
@@
 it has to be able to convert symbols of (at runtime) unknown type to strings.
+Our implementation of \jmodule{RegexpAutomatonSimplifier} is discussed in \ref{section_StateRemoval}
 
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{state_removal_modules}
-	\caption{Modules of StateRemoval.} \label{state_removal_modules}
-\end{figure}
-
-\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{AutomatonMergingState}.
+\subsubsection{MergeConditionTester submodule}
+\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{Au\.to\.ma\.ton\.Merg\.ing\.Sta\.te}.
 It is at disposal for implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
@@
 Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states.
-One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.
+One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.\\
 
@@
 
-\subsection{NestedConcatention (RegularExpressionCleaner)}
+\subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
 This cleaner does exactly what we have just described.
@@
 
-\subsection{Null (RegularExpressionCleaner)}
+\subsection{Null (RegularExpressionCleaner)} \label{section_Null}
 This cleaner just returns regexp it receives and does nothing.
@@
 
-\subsection{EmptyChildren (RegularExpressionCleaner)}
+\subsection{EmptyChildren (RegularExpressionCleaner)} \label{section_EmptyChildren}
 Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
@@
 
-\subsection{Chained (RegularExpressionCleaner)}
+\subsection{Chained (RegularExpressionCleaner)} \label{section_Chained}
 Enables user to chain more cleaners with output from first one plugged as input into second one and so one.
@@
 
-\paragraph{Greedy automaton simplifier} We implement greedy strategy of automaton simplifying in class \code{Greedy}.
+\subsection{Greedy (AutomatonSimplifier)} \label{section_Greedy}
+We implement greedy strategy of automaton simplifying in class \code{Greedy}.
 It simply asks given \code{MergeConditionTester} if it can merge any of automaton states and merges states until there are no states to be merged:
@@
 
-\paragraph{UserInteractive automaton simplifier}
+\subsection{UserInteractive (AutomatonSimplifier)} \label{section_UserInteractive}
 We have created automaton simplifier which displays user input automaton asking him to select some states to merge.
@@
 
-\paragraph{StateRemoval regexp automaton simplifier}
+\subsection{StateRemoval (RegexpAutomatonSimplifier)} \label{section_StateRemoval}
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{state_removal_modules}
+	\caption{Modules of StateRemoval.} \label{state_removal_modules}
+\end{figure}
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
@@
 
-\paragraph{Weighted} We implement one orderer, called \code{Weighted}.
+\subsection{Weighted (Orderer)} \label{section_Weighted}
+We implement one orderer, called \code{Weighted}.
 It is simple heuristic - weights all states (weight = sum of {in | out | loop}-transition regular expression lengths) and returns state with lowest weight.

### 2011-01-28 doc/jInfer/twostep/twostep.tex
miniiiiiiiiiiiiiior
3159165 Create documentation fof TwoStep simplifier - DONE svn stsvn stsvn stsvn stsvn stsvn stsvn stsvn stsvn st

@@
 \section{Introduction}
-With \jmodule{TwoStepSimplifier} we provide user complex solution of infering problem.
-This is the only one simplifier shipped with jInfer, but it has many replaceable submodules of which we implement more alternatives in basic bundle.
-The best we implement is \cite{ahonen} $k,h$-context method od infering grammar rules.
+With \jmodule{TwoStepSimplifier} we provide user a complex solution of grammar inference problem.
+This is the only one simplifier shipped with jInfer, but it has many features and replaceable submodules of which we implement alternatives in basic bundle.
+We implement is $k,h$-context method of infering grammar rules (see \cite{ahonen}).
 This process is split into submodules with generic automaton implementation, so it is easy to plug-in any other automaton merging state algorithm.
@@
 \section{Using of factory pattern} \label{section_factory}
-But in first, we will stay a while at factory pattern usage.
-In \jmodule{TwoStepSimplifier} we use factory pattern (as described in \cite{archdoc}) to divide module into submodules.
-As service providing classes are kept singleton, we make them factories:
+Lets stay a while at factory pattern usage.
+In \jmodule{TwoStepSimplifier} we employ factory pattern (as described in \cite[section 4.2]{archdoc}) to divide module into submodules.
+Since service providing classes kept singletons, we make them factories:
+For example:
 \begin{verbatim}
+@ServiceProvider(service = AutomatonSimplifierFactory.class)
 public interface AutomatonSimplifierFactory extends 
@@
 \end{verbatim}
-The real submodule interface (called familiarly worker interface) is the \code{AutomatonSimplifier<T>} interface (of wich instance is returned by factory).
+The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (of which instance is returned by factory).
 It is now not important how worker interface looks like, lets examine the factory.
 
-In \code{TwoStepSimplifier} we design service providing factory interfaces so that they extends extends \code{NamedModule} with module describing methods:
+In \jmodule{TwoStepSimplifier} we design service providing factory interfaces so that they extend \code{NamedModule}, \code{Capabilities} and \code{UserModuleDescription}.
+That means, modules must implement methods:
 \begin{verbatim}
@@
 String getModuleDescription();
-\end{verbatim}
-And to extend \code{Capabilities}, thus each module have to answer, if it has some capabilities.
-And to extend \code{UserModuleDescription} interface, which defines method:
-\begin{verbatim}
+List<String> getCapabilities();
 String getUserModuleDescription();
 \end{verbatim}
+Nothing non-standard, but \code{getUserModuleDescription}.
 It returns description of the module, comprehensive to user, which is then displayed in properties panels.
 
-In some circumstances, it is useful to have method \code{create()} generic.
-The \jmodule{AutomatonSimplifier} works with \code{Automaton}, which itself is generic too.
-But simplifying does not depend on type of symbol of automaton, so interface \code{AutomatonSimplifier<T>}
-is also generic as simplifier can simplify automaton of any java type symbol.
-Factory interface deals with this by defining \code{create()} method generic too.
+In some circumstances, it is useful to declare method \code{create()} generic.
+The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which themself are generic, too.
+But simplifying of automaton does not depend on type of symbol of automaton transitions, so interface \code{AutomatonSimplifier<T>}
+is also generic.
+Factory interface deals with this by defining \code{create()} method generic, too.
 
@@
 ...
-result= autSmp.work(something_to_process);
+result= autSmp.simplify(something_to_process);
 \end{verbatim}
@@
 
-Lets look on \code{AutomatonMergingStateFactory}, that is factory of module, which has \jmodule{AutomatonSimplifier} as submodule.
-Its create method looks like this (shortened): 
+Lets look on \code{AutomatonMergingStateFactory}, that is factory of module, which has \jmodule{AutomatonSimplifier} as a submodule.
+Its create method is (shortened): 
 \begin{verbatim}
@@
   LOG.debug("Creating new ClusterProcessorAutomatonMergingState.");
-  return new AutomatonMergingState(getAutomatonSimplifierFactory(), getRegexpAutomatonSimplifierFactory());
+  return new AutomatonMergingState(getAutomatonSimplifierFactory(),
+                                   getRegexpAutomatonSimplifierFactory());
 }
@@
   return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
-            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER));
+            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER, PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
 }
 \end{verbatim}
-Cluster processor \code{AutomatonMergingState} then receives factories of \code{AutomatonSimplifier} and \code{RegexpAutomatonSimplifier} submodules in its constructor.
+Cluster processor \code{AutomatonMergingState} then receives factories of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
 Cluster processor then may create as many instances of submodule classes as it needs (maybe simplifying more than one automaton).
-Thorough this document, we will mention only worker interface when describing submodules, since all factory interfaces are designed same way just described.
+Thorough this document, we will mention only worker interface when describing submodule, since all factory interfaces are designed same way, we've just described.
 
@@
 \section{Structure}
-\jmodule{TwoStepSimplfier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
-Module is implemented in two classes: \code{TwoStepSimplifier} (main logic), \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
+\jmodule{TwoStepSimplifier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
+Module is divided into two classes: \code{TwoStepSimplifier} (main logic), \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
 Latter one is registered as service provider (and implements) of \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
@@
 Each grammar rule is represented as class \code{Element}, where element is left side of rule and it's \code{getSubnodes()} method
-returns regular expression, which is right side of rule.
+returns regular expression, which is right side of rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
 Second parameter is 
@@
 Callback which should be called when work is done.
-On output, simplifier provides simplified grammar as a parameter into callback function:
+On output, simplifier provides Simplified grammar as a parameter into callback function:
 \begin{verbatim}
@@
 \end{verbatim}
-Thus simplifier receives initial grammar and returns simplified grammar, both represented same way - as elements.
+Thus simplifier receives Initial Grammar and returns Simplified grammar, both represented same way - list of elements.
 
-\code{TwoStepSimplifierFactory} in method \code{start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
-Then it calls \code{simplify(initialGrammar)} method on this instance and its result passes as parameter to callback function.
-Thats all, the magic comes in \code{TwoStepSimplifier} class.
+\code{TwoStepSimplifierFactory.start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
+Then it calls \code{simplify(initialGrammar)} method of this instance and its result passes as parameter to callback function.
+Thats all, the magic comes in \code{TwoStepSimplifierFactory} class.
 
+\section{Modular design}
 \jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
@@
 We will examine both of them.
-There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beatify output regular expressions, no inference logic is implemented there. 
+There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beautify
+ output regular expressions, no inference logic is implemented there. 
 Modules are drawn on fig. \ref{twostep_modules}.
@@
 \end{figure}
-We provide one \jmodule{Clusterer} and 4 \jmodule{ClusterProcessor} implementations.
+We provide one \jmodule{Clusterer}, 4 \jmodule{ClusterProcessor} and 4 \jmodule{RegularExpressionCleaner} implementations.
 Each of those will be explained further in this document.
@@
 Then, empty list of elements is created as final (simplified) grammar.
-For each input rule in initialGrammar, submodule \jmodule{ClusterProcessor} is called to do infering of regular expression of that element.
-Finally, regular expression cleaning is done in submodule and new \code{Element} instance is created as a copy of processed node, but with cleaned regexp.
+For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is called to do infering of regular expression of that element.
+Finally, regular expression cleaning is done in submodule and new \code{Element} instance is created as a copy of processed node, but with infered and cleaned regexp.
+
+\paragraph{Sentinel processing}
+There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
+Sentinels can be only on right side of rules and they have to be send to clusterer, as someone further in the chain can ask for cluster of sentinel element - which appears on right side.
+But sentinels have no content, they must not be taken into account when infering attribute properties (whether it is required or optional) - since sentinel have no attributes at all.
 
@@
 
-TODO anti sentinel cant be on left side
-TODO anti sentinel processing
-
 %#### STRUCTURE->Clusterer ############################################
-\subsection{Clusterer module}
+\subsection{Clusterer submodule}
 We implement clustering in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
 One cluster is represented by class \code{Cluster<T>} with \code{T} as type of clustered items.
-This class simply holds java set of member of cluster and one of the references is held also in \code{representant} member.
+This class simply holds java set of members of cluster and one of the references is held also in \code{representant} member.
 
 We provide \code{Clusterer} interface for classes to implement.
-Its purpose is to cluster bunch of elements (rules) on input, into  bunch of cluster class instances (cluters) on output.
+Its purpose is to cluster bunch of elements (rules) on input, into  bunch of \code{Cluster} class instances (cluters) on output.
 It has methods \code{add()} and \code{addAll()} for adding items for clustering.
 Centerpart is method \code{cluster()}, which does the clustering itself.
+
 As it may be time-consuming operation, method throws \code{InterruptedException}.
 Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
-After clustering, implementation should hold clusters in member, as it it will be further asked by calling method \code{T getRepresentantForItem(T item)}.
+After clustering, implementation should hold clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
 Given item to this method, one can ask for representant of cluster, to which the item belongs.
@@
 One can pull clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
-
 Basic work usage with clusterer is:
@@
 \subsubsection{ClustererWithAttributes extended interface}
-Maybe you noticed that whole clusterer interface and cluster class are generic.
+You noticed that whole \jmodule{Clusterer} interface and \code{Cluster} class are generic.
 They may be used as design pattern not only for clustering elements in inference process.
@@
 
-We provide one implementation of this interface.
-It is described in \ref{section_Iname}.
+We provide one implementation of this interface described in \ref{section_Iname}.
 
@@
 \subsection{ClusterProcessor module}
-Cluster processor takes rules of one cluster of element and somehow obtains regular expression for that set of elements.
+\jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression for that set of elements.
 It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for the resulting element) and with subnodes set to regular expression infered.
-It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to resulting element.
+It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to the resulting element.
 
@@
 \end{verbatim}
-Maybe you are asking, why cluster processor is given the clusterer instance.
+Why cluster processor is given the clusterer instance?
 Rules themself contain information about which elements to process, but clusterer has more information
 about the topic.
-Clusterer can tell you representant for any element in whole input (not only those elements in rules, but also those that may be on right side of rules), clusterer (if it is with attributes) has information about attributes <of each cluster.
+Clusterer can tell you representant for any element in whole input (not only those elements in rules, but also those that may be on right side of rules), clusterer (if it is with attributes) has information about attributes of each cluster.
 
-We describe each cluster processor implementation we've got in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.
+We describe each \jmodule{ClusterProcessor} implementation we've got in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.
 
@@
 \subsection{RegularExpressionCleaner module}
-Last we examine \code{RegularExpressionCleaner} interface.
+Last we examine \jmodule{RegularExpressionCleaner} interface.
 Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
@@
 \end{figure}
+Process illustrated on fig. \ref{twostep_flow}:
 \begin{enumerate}
@@
 	\item From each cluster processing, regular expression is added to the list of simplified grammar rules.
-	\item List of simplifier rules is returned (simplified grammar).
+	\item List of simplified rules is returned (Simplified grammar).
 \end{enumerate}
-Process is illustrated on fig. \ref{twostep_flow}
-
-%#### EXTENSIBILITY ############################################
-\section{Extensibility}
-If you are willing to, you can replaced each of three submodules of \jmodule{TwoStepSimplifier}.
-It may be useful to replace default \jmodule{ClustererWithAttributes} implementation to clever one.
-And regular expression cleaning will be probably target of adding new implementations, too.
-We recommend replacing other parts at lower levels, however.
-You will need to replace \jmodule{ClusterProcessor} only if you want to write infering method that is not based on merging state algorithm (or NFA's at all).
-If this is not your case, read the rest of this document and find proper place where to place your logic.
 
@@
 \begin{itemize}
-  \item if it is simple data throw it to \code{SimpleDataClusterer} class which is associated with main element (of which this is right side), which does nothing more, than holds all of simple data in one cluster.
+  \item if it is simple data throw it to \code{SimpleDataClusterer} class which is associated with main element cluster.
+It does nothing more, than holds all \code{SimpleData} instances in one cluster.
 But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
@@
 So there can't be cluster with only one sentinel element in it.
-  \item do nothing otherwise, since it is element, that has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in main initial grammar list, thus it is already processed or will be on schedule in clustering.
+  \item do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammr, thus it is already processed or is on schedule.
 \end{itemize}
@@
 It clusters them by name (case insensitive).
+Each main cluster also has \code{SimpleData} clusterer associated with it.
+Whole scheme is on fig. \ref{clusterer_structure}.
+\begin{figure}
+	\centering\includegraphics[scale=1]{clusterer_structure}
+	\caption{Iname clusterer structure} \label{clusterer_structure}
+\end{figure}
 
@@
 This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie") of them.
-More precisely, it takes the first rule and delares it to be a long branch (concatenation of tokens) in a newly created tree.
+More precisely, it takes the first rule and declares it to be a long branch (concatenation of tokens) in a newly created tree.
 After that, it adds the remaining rules one by one as branches like this: as long as it can follow an existing branch, it follows it.
@@
 Repeating this process creates a prefix tree describing all the rules in the cluster.
-TODO vektor and then what? how do you get regexp from that.
+The tree is the final regular expression.
 
@@
 \jmodule{AutomatonMergingState} is implementation of merging state algorithm on nondeterministic finite automaton (see package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate}).
-It creates prefix-tree automaton (PTA) from positive examples - right sides of rules given.
-Then it calls its submodule called \jmodule{AutomatonSimplifier} to  modify PTA to some generalized 
-automaton by merging states.
-
-Simplified automaton is then converted to an instance of \code{RegexpAutomaton} by using clone constructor.
-Regexp automaton is automaton with regular expression as symbol on transitions.
-In automata theory, such automaton is called extended NFA.
-Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
-
-Regexp automaton is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
-Its job is to derive regular expression from automaton, such that automaton and regular expression represents same language.
-The data flow is on fig. \ref{automaton_merging_state_flow}.
-
-\subsubsection{AutomatonSimplifier module}
-\jmodule{AutomatonSimplifier} module has its worker interface defined as:
-\begin{verbatim}
-public interface AutomatonSimplifier<T> {
-  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
-                        final SymbolToString<T> symbolToString) throws InterruptedException;
-}
-\end{verbatim}
-Given input automaton it returns automaton, there is nothing magical on it.
-Interface is defined as generic, implementation don't have to be, but there is no reason to not make them generic.
-When implementation needs to present automaton to user, however, it needs some string representation of symbols to display on automaton transitions.
-For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
-It is responsible for converting symbol to string.
-To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplfifier} submodule.
-
-By this genericity, simplifier can be used to perfom same algorithms for simplifying not only to infer rules for elements, but maybe also to infer patterns strings for content model of attributes.
-The actual implementation of this still waits for its developer to come.
-
-TODO how automaton works?
-
-Implementations of \jmodule{AutomatonSimplifier} are discussed in \ref{section_Greedy} and \ref{section_UserInteractive}.
-
-\subsubsection{RegexpAutomatonSimplifier submodule}
-Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain regular expression from given automaton.
-Interface for doing this is thus simple:
-\begin{verbatim}
-public interface RegexpAutomatonSimplifier<T> {
-  Regexp<T> simplify(final RegexpAutomaton<T> inputAutomaton, 
-                     final SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
-}
-\end{verbatim}
-Once again we are encountering \code{symbolToString} with same purpose as before, if submodule would need to present automaton to user
-it has to be able to convert symbols of (at runtime) unknown type to strings.
-Our implementation of \jmodule{RegexpAutomatonSimplifier} is discussed in \ref{section_StateRemoval}
-
-\subsubsection{MergeConditionTester submodule}
-\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{Au\.to\.ma\.ton\.Merg\.ing\.Sta\.te}.
-It is at disposal for implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
-
-Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states.
-One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.\\
-
+\subsubsection{Structure} \label{section_AutomatonMergingState_structure}
 Shortened code of cluster processor operation:
@@
 \end{verbatim}
-Whole submodule structure of \jmodule{AutomatonMergingState} cluster processor is drawn on fig. \ref{automaton_merging_state_modules}.
+The whole submodule structure is drawn on fig. \ref{automaton_merging_state_modules}.
+
+\subsubsection{Data flow}
+It works in steps:
+\begin{enumerate}
+	\item It creates prefix-tree automaton (PTA) from positive examples - right sides of rules given.
+Then it calls its submodule called \jmodule{AutomatonSimplifier} to  modify PTA to some generalized 
+automaton by merging states.
+	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.
+Regexp automaton is automaton with regular expression as symbol on transitions.
+In automata theory, such automaton is called extended NFA.
+Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
+	\item \code{RegexpAutomaton} is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
+Its job is to derive regular expression from automaton, such that automaton and regular expression represents same language.
+\end{enumerate}
+The data flow is on fig. \ref{automaton_merging_state_flow}.
+
+\subsubsection{Data structures - Automaton} \label{section_Automaton}
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{pta}
+	\caption{PTA created from input strings: $aa, aaaaa, abaa, bab, baa$} \label{pta}
+\end{figure}
+Class \code{Automaton} is shortly described in \cite{archdoc}.
+In this place, it is interestesting to look on method \code{buildPTAOnSymbol}.
+Given \code{List<T> symbolString} it traverses automaton - comparing transition symbols with symbols in string.
+It follows transitions in automaton until first difference with string symbol is found (like in prefix trees).
+On different symbol, it creates new branch and all states and transitions on that branch.
+Last state visited got's incremented its \code{finalCount} value, as one more input string ended in this state.
+Lets see code:
+\begin{verbatim}
+public void buildPTAOnSymbol(final List<T> symbolString) {
+  State<T> xState= this.getInitialState();
+  for (T symbol : symbolString) {
+    final Step<T> xStep= this.getOutStepOnSymbol(xState, symbol);
+    if (xStep != null) {
+      xStep.incUseCount();
+      xState= xStep.getDestination();
+    } else {
+      final State<T> newState= this.createNewState();
+      final Step<T> newStep= this.createNewStep(symbol, xState, newState);
+      assert newStep.getDestination().equals(newState);
+      xState= newStep.getDestination();
+    }
+  }
+  xState.incFinalCount();
+}
+\end{verbatim}
+Start from initial state traverse automaton by asking for \code{getOutStepOnSymbol}.
+If there is such step (transition), follow it to its destination state and move the position in \code{symbolString}.
+If there is no such step, create one and create also the destination state, then follow this newly created step to new state.
+
+As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
+By the way, that is why we cluster all simple data and all sentinels on right sides of rules.
+To automaton, instead of instances on right side of rules, their cluster representants are inserted as one can see in section \ref{section_AutomatonMergingState_structure}.
+This assures, that comparisons in automaton are made by using java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
+That results in cluster-equal behaviour.
+Example of prefix-tree automaton is on fig. \ref{pta}.
+
+\subsubsection{Data structures - RegexpAutomaton} \label{section_RegexpAutomaton}
+Class \code{RegexpAutomaton} is extending class \code{Automaton} with generic type \code{T} set to \code{Regexp<T>}.
+That is, you have to instantiate it with same symbol type (e.g. \code{AbstractStructuralNode}) as automaton.
+Implementation can be found in package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate.regexping}, that is same, as where \jmodule{RegexpAutomatonSimplifier} interface is defined.
+Implementation is nothing special, it has only copy constructor which creates regular expression automaton from ordinary automaton by enclosing each symbol on transition of original automaton by \code{Regexp.TOKEN}.
+
+\subsubsection{AutomatonSimplifier module}
+\jmodule{AutomatonSimplifier} module has its worker interface defined as:
+\begin{verbatim}
+public interface AutomatonSimplifier<T> {
+  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
+                        final SymbolToString<T> symbolToString) throws InterruptedException;
+  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
+                        final SymbolToString<T> symbolToString,
+                        final String elementName) throws InterruptedException;
+}
+\end{verbatim}
+Given input automaton it returns automaton, there is nothing magical on it.
+Interface is defined as generic, implementation don't have to be, but there is no reason to not make them generic.
+When implementation needs to present automaton to user, however, it needs some string representation of symbols to display on automaton transitions.
+For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
+It is responsible for converting symbol to string.
+To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule.
+Overloaded version is to inform user about name of processed element when presenting automaton by \jmodule{AutoEditor}.
+This version of interface will probably change in future.
+
+By this genericity, simplifier can be used to perfom same algorithms for simplifying not only to infer rules for elements, but maybe also to infer patterns strings for content model of attributes.
+The actual implementation of this still waits for its developer to come.
+Implementations of \jmodule{AutomatonSimplifier} are discussed in \ref{section_Greedy} and \ref{section_UserInteractive}.
+
+\subsubsection{RegexpAutomatonSimplifier submodule}
+Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain regular expression from given automaton.
+Interface for doing this is thus simple:
+\begin{verbatim}
+public interface RegexpAutomatonSimplifier<T> {
+  Regexp<T> simplify(final RegexpAutomaton<T> inputAutomaton, 
+                     final SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
+}
+\end{verbatim}
+Once again we are encountering \code{symbolToString} with same purpose as before, if submodule would need to present automaton to user
+it has to be able to convert symbols of (at runtime) unknown type to strings.
+Our implementation of \jmodule{RegexpAutomatonSimplifier} is discussed in \ref{section_StateRemoval}
+
+\subsubsection{MergeConditionTester submodule}
+\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{Au\.to\.ma\.ton\.Merg\.ing\.Sta\.te}.
+It is at disposal for implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
+
+Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states.
+One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.
 
@@
 It converts nested concatenations into flat ones.
+Example: $(name, (person, id))$ is converted to $(name, person, id)$.
 
@@
 This cleaner just returns regexp it receives and does nothing.
-Plug this one into chain, if you want to omit this step of inference.
+Plug this one into chain, if you want to omit this step in inference.
 
@@
 There is one empty concatenation/alternation/permutation with no sense.
-This is wiped out by this cleaner.
+This is wiped out by this cleaner to produce 
+$$
+(name, (person, id))
+$$
 
@@
 \end{verbatim}
-We have implemented \emph{k,h-context} (see \cite{ahonen}) state equivalence in class \code{KHContext} (implements \code{MergeConditionTester} interface), which is used by \code{Greedy} to test mergability of states by default configuration.
+We provide one \jmodule{MergeConditionTester} implementation, see \ref{section_KHContext}.
 
@@
 \end{verbatim}
-We user \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
-To make reader more happy, we include one screenshot of automaton visualization on picture \ref{screenshot_viz}.
-TODO anti screenshot for happy reader
+We use \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
+To make reader more happy, we include one screenshot of automaton visualization on picture \ref{screenshot}.
+\begin{figure}
+	\centering\includegraphics[scale=0.5]{screenshot}
+	\caption{Screenshot of UserInteractive (AutomatonSimplifier).} \label{screenshot}
+\end{figure}
+
+\subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
+We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{MergeConditionTester} interface), which is used by \code{Greedy} to test mergability of states by default configuration.
+TODO vektor : mention this? (we have stupid algorithm:)
+We don't use $k-grams$ algorithm from \cite{ahonen}, but a simple DFS.
 
@@
 This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
+before algorithm starts, states superInitial and superFinal are added to automaton, former one with $\lambda$-transition to initial state.
+From all final states $\lambda$ transition to superFinal state is added.
 State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - superInitial and superFinal.
-Both are added before algorithm starts. Former one with $\lambda$-transition to initial state, and from all final states $\lambda$ transition to superFinal state is added.
 After removing all states, there is only one transition from superInitial to superFinal state.
-That transition has final regular expression on it as symbol, it is read and returned to \jmodule{AutomatonMergingState}.
+That transition has final regular expression on it as symbol, it is read and returned.
 
@@
 
+\subsubsection{Data Structures - StateRemovalRegexpAutomaton}
+We extend \code{RegexpAutomaton} to obtain automaton that fits state removing procedure.
+This implementation has methods \code{createSuperInitialState} and \code{createSuperInitialState}.
+They does exactly what their names stands for.
+SuperInitial state has only one $\lambda$-out-transition pointing to original initial state.
+SuperFinal state has as many $\lambda$-in-transitions, as there are original final states - from each one, it gets one.
+Without going further to many technical details, we just mention method \code{removeState}, which removes given state from automaton.
+Transitions are handled this way:
+\begin{enumerate}
+	\item Collapse all loop transition of removed state to one loop transition with new \code{Regex.ALTERNATION} of all regexps on original loops (with Kleene interval set).
+	\item Collapse all {in | out}-transitions, which same {source | destination} state to one transition with \code{Regexp.ALTERNATION} of original regexps.
+	\item For each in-transition, loop, and out-transition, add transition from in-transition.getSource to out-transition.getDestination with symbol of \code{Regexp.CONCATENATION} of in-transition regexp, loop regexp and out-transition regexp.
+Thus by-passing the state completely.
+	\item Remove the state and transitions associated with it.
+\end{enumerate}
+Whole process is step-by-step illustrated on fig. \ref{regaut}.
+\begin{figure}
+	\centering
+	\subfigure[RegexpAutomaton after conversion from ordinary automaton.]
+	{
+		\includegraphics[scale=\myscale]{regaut1}
+		\label{regaut1}
+	}
+	\subfigure[First, collapse loop transitions to alternation with Kleene interval set.]
+	{
+		\includegraphics[scale=\myscale]{regaut2}
+		\label{regaut2}
+	}
+	\subfigure[Collapse {in | out}-transitions {from | to} same states.]
+	{
+		\includegraphics[scale=\myscale]{regaut3}
+		\label{regaut3}
+	}
+	\subfigure[Bypass state 3 by adding concatenation transitions from states 1 and 2 to state 4.]
+	{
+		\includegraphics[scale=\myscale]{regaut4}
+		\label{regaut4}
+	}
+	\subfigure[Finally remove state 3 and associated transitions.]
+	{
+		\includegraphics[scale=\myscale]{regaut5}
+		\label{regaut5}
+	}
+	\caption{RegexpAutomaton and removing state 3 from it.} \label{regaut}
+\end{figure}
+
 \subsection{Weighted (Orderer)} \label{section_Weighted}
@@
 It is simple heuristic - weights all states (weight = sum of {in | out | loop}-transition regular expression lengths) and returns state with lowest weight.
+
+%#### EXTENSIBILITY ############################################
+\section{Extensibility}
+If you are willing to, you can replace each of three submodules of \jmodule{TwoStepSimplifier}.
+It may be useful to replace default \jmodule{ClustererWithAttributes} implementation to a clever one.
+And regular expression cleaning will be probably target of adding new implementations, too.
+We recommend replacing other parts at lower levels, however.
+You will need to replace \jmodule{ClusterProcessor} only if you want to write infering method that is not based on merging state algorithm (or NFA's at all).
+
+If this is not your case, you can replace module \jmodule{AutomatonSimplifier} to implement ACO heuristics or MDL principle in searching solutions of merging states.
+
+Of course, merge criterion is propably the point to add implementation, implement \jmodule{MergeConditionTester} interface to obtain for example $s,k$-string merge criterion.
+
+Changing ordering of states to remove in \jmodule{StateRemoval} module is easy by replacing \jmodule{Orderer} implementation, for example to one of suggested in \cite{1224321} may lead to shorter resulting regular expressions.
 

### 2011-01-28 doc/jInfer/basicxsd/basicxsd.tex
Documenting BasicXSDExporter.

@@
 
-First, global elements are exported. For each element, its type is defined as a global type. TODO rio example
+First, global elements are exported. For each element, its type is defined as a global type. TODO rio example. This is done by passing global elements to \code{processGlobalElement} method.
 
 After global elements, others are exported.
+To sa deje rekurzivnym priechodom so startom v korenovom elemente.
+Elementy, ktorych typy su definovane globalne a elementy so vstavanymi typmi, su exportovane jednoducho, uvedenim ich mena a typu.
+Ostatne elementy su definovane na mieste.
+TODO rio priklady.
+Na toto je potrebne zavolat metodu \code{processElement} s korenovym elementov, ako argumentom.
+
+\subsubsection{Content of elements}
+
+Spracovava sa vo funkcii \code{processElementContent}.
+Obsah elementu je definovany pomocou elementu element a bud pomocou jedneho zo vstavanych typov ako napriklad string, integer, positiveInteger, etc alebo je definovany vlastny.
+
+Vlastny obsah elementu je definovany bud ako simpleType alebo complexType.
+TODO rio priklady.
+Actually exporter nepodporuje konstrukcie XSD, ktore sa definuje cez simpleType, takze tento typ sa vo vysledku neobjavi.
+complexType moze obsahovat konstrukcie sequence a choice, v ktorych sa mozu vyskytovat dalsie elementy alebo zas tieto konstrukcie.
+complexType moze na konci obsahovat definicie atributov (popisane nizsie).
+Prazdny element je vyjadreny prazdnym complexType.
+
+Konstrukcie element, sequence a choice moze obsahovat atributy minOccurs a maxOccurs, ktore udavaju interval poctu instancii danej konstruckie v mieste definovania.
+TODO priklad.
+Moze byt zadane aj jednotlivo.
+Pri nezadani plati podla pravidiel XSD default hodnota 1.
+Tieto hodnoty sa ziskavaju z metadat pomocou funkcie getInterval definovaj na triede regexp (TODO odkaz niekam).
+
+Exporter podporuje zmiesany obsah elementov definovany atributom mixed="true". TODO priklad.
+
+\subsection{Attributes}
+
+Atributy elementu sa definuju pomocou elementu attribute and takes place at the end of complexType element. Spracovavaju sa vo funkcii \code{processElementAttributes}.
+\begin{itemize}
+	\item Built-in types like \code{string}, \code{integer}, \code{positiveInteger}, etc.
+	\item Attribute requiredness defined by attribute \code{use="required"}.
+\end{itemize}
 

### 2011-01-28 doc/jInfer/basicxsd/basicxsd.tex
Documenting BasicXSDExporter.

@@
 	\item Preprocessing
-	\item Own export to string representation of XSD
+	\item Own export to a string representation of XSD
 \end{enumerate}
 Method \code{start} first creates instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
-Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such which elements should be globally defined and which element is the top level element (TODO rio how is this element called in XML??).
+Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such which elements should be globally defined and which element is the root element.
 This instance is kept as a member variable of module to be used during the whole export process.\\
 
-Afterwards, \code{start} method recursively traverses global elements followed by other elements starting at the top level element and for each it creates element's XSD string representation.\\
+Afterwards, \code{start} method recursively traverses global elements followed by other elements starting at the root element and for each it creates element's XSD string representation.\\
 
@@
 Constructor of \code{Preprocessor} class gets elements and a number, defining minimal number of occurrences of an element to be defined.
-It first topologically sorts input elements to decide which one is the top level element.
+It first topologically sorts input elements to decide which one is the root element.
 Afterwards, it counts occurrences of the elements and removes unused ones (those which did not occurred).
@@
 
+Before describing export of elements, let's take a look on how we define elements and their attributes using XSD language.
+
+\subsubsection{Definition of elements}
+
+Element is defined by XSD element \code{element}, specifying its name and type. Let \code{xs} be XMLSchema namespace.
+
+\begin{verbatim}
+<xs:element name="Person" type="...
+\end{verbatim}
+
+Type of an element is one of following.
+
+\begin{itemize}
+	\item XSD built-in type. One of types like \code{xs:string}, \code{xs:integer}, \code{xs:positiveInteger}, etc.
+	\begin{verbatim}
+	<xs:element name="Person" type="xs:string"/>
+	\end{verbatim}
+
+	\item \code{simpleType}. Actually, exporter does not support any XSD features, which are defined in \code{simpleType}. That means, this types will not occur in result XSDs.
+	\begin{verbatim}
+	<xs:element name="Person">
+	  <xs:simpleType>
+	    ...
+	  </xs:simpleType>
+	</xs:element>
+	\end{verbatim}
+
+	\item \code{complexType}.
+This type can contain XSD element \code{xs:sequence} or \code{xs:choice}.
+Each of these elements can contain definitions of elements, \code{xs:sequence}s and \code{xs:choice}s again.
+	\begin{verbatim}
+	<xs:element name="Person">
+	  <xs:complexType>
+	    <xs:sequence>
+	      <xs:element name="Name" type="xs:string"/>
+	      <xs:element name="Surname" type="xs:string"/>
+	      <xs:choice>
+	      ...
+	      </xs:choice>
+	    </xs:sequence>
+	  </xs:complexType>
+	</xs:element>
+	\end{verbatim}
+	An empty element is defined as empty \code{complexType}
+	\begin{verbatim}
+	<xs:element name="EmptyElement">
+	  <xs:complexType>
+	  </xs:complexType>
+	</xs:element>
+	\end{verbatim}
+\end{itemize}
+
+Named type of an element is defined by XSD element \code{simpleType}(lack of features mentioned above) or \code{complexType} with specified attribute \code{name}.
+Its content is exactly the same as described above.
+There is, of course, no need to define built-in types.
+
+\begin{verbatim}
+<xs:complexType name="PersonType">
+...
+</xs:complexType>
+\end{verbatim}
+
+Element of this type can be then defined by specifying name of the type.
+
+\begin{verbatim}
+<xs:element name="Person" type="PersonType"/>
+\end{verbatim}
+
+XSD elements \code{xs:element}, \code{xs:sequence} and \code{xs:choice} can has attributes \code{minOccurs} and \code{maxOccurs}.
+These attributes defines interval of number of instances of a particular element.
+Legal values of these attributes are non-negative integers.
+
+\begin{verbatim}
+<xs:element name="Person" type="PersonType" minOccurs="1" maxOccurs="3"/>
+\end{verbatim}
+
+Default values for \code{minOccurs} and \code{maxOccurs} attributes are "1", if they are not specified. So the example above has the same meaning as following.
+
+\begin{verbatim}
+<xs:element name="Person" type="PersonType" maxOccurs="3"/>
+\end{verbatim}
+
 First, global elements are exported. For each element, its type is defined as a global type. TODO rio example. This is done by passing global elements to \code{processGlobalElement} method.
 
-After global elements, others are exported.
+After global elements, others are exported by calling of method \code{processElement}, supplied by the top level element as its argument.
+Reference to a global element type is done simply by declaring element's name and type.
+
+\begin{verbatim}
+<xs:element name="Person" type="PersonType"/>
+\end{verbatim}
+
 To sa deje rekurzivnym priechodom so startom v korenovom elemente.
@@
 
-\subsubsection{Content of elements}
+\subsubsection{Definition of elements}
 

### 2011-01-28 doc/jInfer/basicxsd/basicxsd.tex
Documenting BasicXSDExporter.

@@
 
+Exporter supports types of mixed elements. Mixed element is an element that contains other elements as well as some text.
+
+\begin{verbatim}
+<mixedElement>
+  some text
+  <anotherElement/>
+  another text
+</mixedElement>
+\end{verbatim}
+
+Mixed element type is defined as \code{complexType} with attribute \code{mixed="true"}. Definition of the element from the last example may be as following.
+
+\begin{verbatim}
+<xs:element name="mixedElement">
+  <xs:complexType mixed="true">
+    <xs:sequence>
+      <xs:element name="anotherElement" type="..."/>
+    </xs:sequence>
+  </xs:complexType>
+</xs:element>
+\end{verbatim}
+
+\subsection{Definition of attributes}
+
+Attributes are defined by XSD element \code{xs:attribute} with attributes \code{name}, \code{type} and optional \code{use}.
+Elements \code{xs:attributes} have to be placed at the end of a \code{complexType} definition.
+
+\begin{verbatim}
+<xs:element name="Person">
+  <xs:complexType>
+  ...
+  </xs:complexType>
+  <xs:attribute name="age" type="xs:positiveInteger"/>
+  <xs:attribute name="id" type="xs:string" use="required"/>
+</xs:element>
+\end{verbatim}
+
+Attribute \code{type} is one of a built-in types. If an attribute is obligatory, this is defined by specifying \code{use="required"}.
+
+--------------------------------------------------------------------------------------
+
 First, global elements are exported. For each element, its type is defined as a global type. TODO rio example. This is done by passing global elements to \code{processGlobalElement} method.
@@
 Na toto je potrebne zavolat metodu \code{processElement} s korenovym elementov, ako argumentom.
-
-\subsubsection{Definition of elements}
-
-Spracovava sa vo funkcii \code{processElementContent}.
-Obsah elementu je definovany pomocou elementu element a bud pomocou jedneho zo vstavanych typov ako napriklad string, integer, positiveInteger, etc alebo je definovany vlastny.
-
-Vlastny obsah elementu je definovany bud ako simpleType alebo complexType.
-TODO rio priklady.
-Actually exporter nepodporuje konstrukcie XSD, ktore sa definuje cez simpleType, takze tento typ sa vo vysledku neobjavi.
-complexType moze obsahovat konstrukcie sequence a choice, v ktorych sa mozu vyskytovat dalsie elementy alebo zas tieto konstrukcie.
-complexType moze na konci obsahovat definicie atributov (popisane nizsie).
-Prazdny element je vyjadreny prazdnym complexType.
-
-Konstrukcie element, sequence a choice moze obsahovat atributy minOccurs a maxOccurs, ktore udavaju interval poctu instancii danej konstruckie v mieste definovania.
-TODO priklad.
-Moze byt zadane aj jednotlivo.
-Pri nezadani plati podla pravidiel XSD default hodnota 1.
-Tieto hodnoty sa ziskavaju z metadat pomocou funkcie getInterval definovaj na triede regexp (TODO odkaz niekam).
-
-Exporter podporuje zmiesany obsah elementov definovany atributom mixed="true". TODO priklad.
-
-\subsection{Attributes}
-
-Atributy elementu sa definuju pomocou elementu attribute and takes place at the end of complexType element. Spracovavaju sa vo funkcii \code{processElementAttributes}.
-\begin{itemize}
-	\item Built-in types like \code{string}, \code{integer}, \code{positiveInteger}, etc.
-	\item Attribute requiredness defined by attribute \code{use="required"}.
-\end{itemize}
 

### 2011-01-29 doc/jInfer/twostep/twostep.tex
3159165 Create documentation fof TwoStep simplifier - magic number mangling on figures to get it on 15 pages, also wraped figure used.

@@
 \usepackage{xstring}
+\usepackage{wrapfig}
 
@@
 
-In \jmodule{TwoStepSimplifier} we design service providing factory interfaces so that they extend \code{NamedModule}, \code{Capabilities} and \code{UserModuleDescription}.
+In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
 That means, modules must implement methods:
@@
   return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
-            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER, PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
+            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER,
+                          PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
 }
@@
 Modules are drawn on fig. \ref{twostep_modules}.
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=\myscale]{twostep_modules}
@@
 They may be used as design pattern not only for clustering elements in inference process.
-To address clustering of elements in more detail, we created \code{ClustererWithAttributes<T, S>} interface, which extends \code{Clusterer<T>} interface.
-It adds method \code{List<Cluster<S>{}> getAttributeClusters(T representant)}, implying that each representant of
+To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
+It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representant of
 type \code{T} (that is representant of some main cluster) has some "attribute" clusters associated with it.
@@
 \section{Data flow}
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=1]{twostep_flow}
@@
 Whole scheme is on fig. \ref{clusterer_structure}.
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=1]{clusterer_structure}
@@
 \subsection{AutomatonMergingState (ClusterProcessor)} \label{section_AutomatonMergingState}
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=\myscale]{automaton_merging_state_modules}
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=1]{automaton_merging_state_flow}
@@
 \subsubsection{Data structures - Automaton} \label{section_Automaton}
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=\myscale]{pta}
@@
 Class \code{Automaton} is shortly described in \cite{archdoc}.
-In this place, it is interestesting to look on method \code{buildPTAOnSymbol}.
+In this place, it is interestesting to look on method \code{build\.PTA\.On\.Sym\.bol}.
 Given \code{List<T> symbolString} it traverses automaton - comparing transition symbols with symbols in string.
@@
 To make reader more happy, we include one screenshot of automaton visualization on picture \ref{screenshot}.
-\begin{figure}
+\begin{figure}[tpb]
 	\centering\includegraphics[scale=0.5]{screenshot}
@@
 \subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
-We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{MergeConditionTester} interface), which is used by \code{Greedy} to test mergability of states by default configuration.
+We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergability of states by default configuration.
 TODO vektor : mention this? (we have stupid algorithm:)
@@
 \subsection{StateRemoval (RegexpAutomatonSimplifier)} \label{section_StateRemoval}
-\begin{figure}
+\begin{wrapfigure}{R}{40mm}
 	\centering\includegraphics[scale=\myscale]{state_removal_modules}
 	\caption{Modules of StateRemoval.} \label{state_removal_modules}
-\end{figure}
+	\vspace{10mm}
+\end{wrapfigure}
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
@@
 We extend \code{RegexpAutomaton} to obtain automaton that fits state removing procedure.
-This implementation has methods \code{createSuperInitialState} and \code{createSuperInitialState}.
+This implementation has methods \code{crea\.te\.Su\.per\.In\.it\.ial\.Sta\.te} and \code{crea\.te\.Su\.per\.In\.it\.ial\.Sta\.te}.
 They does exactly what their names stands for.
@@
 	\item Collapse all loop transition of removed state to one loop transition with new \code{Regex.ALTERNATION} of all regexps on original loops (with Kleene interval set).
-	\item Collapse all {in | out}-transitions, which same {source | destination} state to one transition with \code{Regexp.ALTERNATION} of original regexps.
-	\item For each in-transition, loop, and out-transition, add transition from in-transition.getSource to out-transition.getDestination with symbol of \code{Regexp.CONCATENATION} of in-transition regexp, loop regexp and out-transition regexp.
+	\item Collapse all {in | out}-transitions, which same {source | destination} state to one transition with \code{Reg\.exp.AL\.TER\.NA\.TION} of original regexps.
+	\item For each in-transition, loop, and out-transition, add transition from in-transition source to out-transition destination with symbol of \code{Regexp.CONCATENATION} of in-transition regexp, loop regexp and out-transition regexp.
 Thus by-passing the state completely.
@@
 Whole process is step-by-step illustrated on fig. \ref{regaut}.
-\begin{figure}
+\begin{figure}[pb]
 	\centering
@@
 	}
-	\subfigure[Bypass state 3 by adding concatenation transitions from states 1 and 2 to state 4.]
+	\subfigure[Bypass state 3 by adding concatenation transitions from states 1 and 2 to state 4. Remove state 3 and associated transitions.]
 	{
@@
 	}
-	\subfigure[Finally remove state 3 and associated transitions.]
-	{
-		\includegraphics[scale=\myscale]{regaut5}
-		\label{regaut5}
-	}
+%	
+%	\subfigure[Finally remove state 3 and associated transitions.]
+%	{
+%		\includegraphics[scale=\myscale]{regaut5}
+%		\label{regaut5}
+%	}
 	\caption{RegexpAutomaton and removing state 3 from it.} \label{regaut}
@@
 Whole TwoStep submodules structure is on fig. \ref{twostep_modules_complete}.
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{twostep_modules_complete}
+\begin{figure}[tpb]
+	\centering\includegraphics[scale=0.37]{twostep_modules_complete}
 	\caption{Modules of TwoStep simplifier and their submodules. Filled classes are default selection (best of).} \label{twostep_modules_complete}

### 2011-01-30 doc/jInfer/twostep/twostep.tex
Documentation: TwoStep review.

@@
 \begin{document}
+
+% TODO anti Representant -> Representative!
+
 \maketitle
-\noindent Target audience: developers willing to extend jInfer, scientist willing to implement own inference methods
+\noindent Target audience: developers willing to extend jInfer, researchers willing to implement own inference methods.
 
@@
 Module dependencies:   & AutoEditor, Base, JUNG, Lookup \\ \hline
-Public packages:       & \\ \hline
+Public packages:       & none \\ \hline
 \end{tabular}
@@
 \section{Introduction}
-With \jmodule{TwoStepSimplifier} we provide user a complex solution of grammar inference problem.
-This is the only one simplifier shipped with jInfer, but it has many features and replaceable submodules of which we implement alternatives in basic bundle.
+In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
+% TODO anti Sorry, I don't understand the following 3 sentences, so I cannot correct them
+This is the only one simplifier shipped with jInfer, but it has many features and exchangeable submodules of which we implement alternatives in basic bundle.
 We implement is $k,h$-context method of infering grammar rules (see \cite{ahonen}).
@@
 %#### FACTORY ############################################
-\section{Using of factory pattern} \label{section_factory}
+\section{Factory pattern usage} \label{section_factory}
 Lets stay a while at factory pattern usage.
 In \jmodule{TwoStepSimplifier} we employ factory pattern (as described in \cite[section 4.2]{archdoc}) to divide module into submodules.
-Since service providing classes kept singletons, we make them factories:
+Since service providing classes are being kept as singletons in the NB platform, we use them as factories:
 For example:
@@
 The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (of which instance is returned by factory).
-It is now not important how worker interface looks like, lets examine the factory.
+For now, it is not important how the worker interface looks like, lets just examine the factory.
 
@@
 \end{verbatim}
-Nothing non-standard, but \code{getUserModuleDescription}.
-It returns description of the module, comprehensive to user, which is then displayed in properties panels.
+Nothing non-standard apart from \code{getUserModuleDescription}.
+It returns user-friendly description of the module which is then displayed in properties panels.
 
-In some circumstances, it is useful to declare method \code{create()} generic.
-The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which themself are generic, too.
-But simplifying of automaton does not depend on type of symbol of automaton transitions, so interface \code{AutomatonSimplifier<T>}
-is also generic.
-Factory interface deals with this by defining \code{create()} method generic, too.
+% TODO anti Rewrite so that these sentences don't end with "too" all the time
+Under certain circumstances it is useful to declare method \code{create()} generic.
+The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic too.
+But automaton simplification does not depend on type of symbol of automaton transitions, so interface \code{AutomatonSimplifier<T>}
+generic too.
+Factory interface deals with this by defining the \code{create()} method generic.
 
@@
 \begin{verbatim}
-final Properties p = RunningProject.getActiveProjectProps(getName());
+Properties p = RunningProject.getActiveProjectProps(getName());
 
@@
 \end{verbatim}
-If one module has some submodules, we implement lookups for submodule implementations in our own factory \code{create} method.
+If a module has submodules, we implement lookups for submodule implementations in our own factory \code{create()} method.
 Worker class receives factories of all of its submodules as a constructor parameters.
 
-Lets look on \code{AutomatonMergingStateFactory}, that is factory of module, which has \jmodule{AutomatonSimplifier} as a submodule.
-Its create method is (shortened): 
+Lets look at \code{AutomatonMergingStateFactory}. This is factory of module which has \jmodule{AutomatonSimplifier} as a submodule.
+Its create method (shortened): 
 \begin{verbatim}
@@
 \end{verbatim}
-Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical, we show you the former one:
+Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical. Here is the former:
 \begin{verbatim}
 private AutomatonSimplifierFactory getAutomatonSimplifierFactory() {
-  final Properties p = RunningProject.getActiveProjectProps(getName());
+  Properties p = RunningProject.getActiveProjectProps(getName());
 
@@
 Cluster processor then may create as many instances of submodule classes as it needs (maybe simplifying more than one automaton).
-Thorough this document, we will mention only worker interface when describing submodule, since all factory interfaces are designed same way, we've just described.
+Thorough this document, we will mention only worker interface when describing submodule, since all factory interfaces are designed the same way as was just described.
 
@@
 \jmodule{TwoStepSimplifier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
-Module is divided into two classes: \code{TwoStepSimplifier} (main logic), \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
-Latter one is registered as service provider (and implements) of \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
+Module is divided into two classes: \code{TwoStepSimplifier} (main logic) and \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
+The latter is registered as service provider (and implements) of \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
 
-It has main method called \code{start} which receives Initial Grammar in form of
+Its main method is \code{start()} which receives Initial Grammar in form of
 \begin{verbatim}
-final List<Element> grammar
+List<Element> grammar
 \end{verbatim}
-Each grammar rule is represented as class \code{Element}, where element is left side of rule and it's \code{getSubnodes()} method
-returns regular expression, which is right side of rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
+Each grammar rule is represented as class \code{Element}, where element is left side of rule and its \code{getSubnodes()} method
+returns regular expression representing right side of rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
 Second parameter is 
 \begin{verbatim}
-final SimplifierCallback callback
+SimplifierCallback callback
 \end{verbatim}
-Callback which should be called when work is done.
-On output, simplifier provides Simplified grammar as a parameter into callback function:
+Callback which should be invoked when the simplification is done.
+On output, simplifier provides Simplified grammar as a parameter of callback function:
 \begin{verbatim}
-void finished(final List<Element> grammar);
+void finished(List<Element> grammar);
 \end{verbatim}
-Thus simplifier receives Initial Grammar and returns Simplified grammar, both represented same way - list of elements.
+Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented same way - list of elements.
 
 \code{TwoStepSimplifierFactory.start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
-Then it calls \code{simplify(initialGrammar)} method of this instance and its result passes as parameter to callback function.
-Thats all, the magic comes in \code{TwoStepSimplifierFactory} class.
+Then it calls \code{simplify(initialGrammar)} method of this instance and its result is passed as parameter to callback function.
+That's all, the magic is in \code{TwoStepSimplifierFactory} class.
 
@@
 \begin{enumerate}
-	\item clustering of element instances into clusters of (probably) same elements
-	\item infering regular expression for each element from examples of element contents taken from all elements in cluster
+	\item Clustering the element instances into clusters of (probably) same elements.
+	\item Inferring regular expression for each element from examples of element contents taken from all elements in a cluster.
 \end{enumerate}
-Task of clustering is dedicated to \jmodule{Clusterer} submodule, and task of infering regular expression for each cluster is dedicated to \jmodule{ClusterProcessor} submodule.
-We will examine both of them.
+Clustering is delegated to \jmodule{Clusterer} submodule, and task of regular expression inference for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
 There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beautify
@@
 
-Method \code{TwoStepSimplifier.simplify()} does basically this:
+Method \code{TwoStepSimplifier.simplify()} basically does the following.
+% TODO anti Rename finalGrammar to resultGrammar or something like that
 \begin{verbatim}
 // 1. cluster elements
-final Clusterer<AbstractStructuralNode> clusterer= clustererFactory.create();
+Clusterer<AbstractStructuralNode> clusterer = clustererFactory.create();
 clusterer.addAll(initialGrammar);
@@
 // 2. prepare emtpy final grammar
-final List<Element> finalGrammar= new LinkedList<Element>();
+List<Element> finalGrammar= new LinkedList<Element>();
 
 // 3. process rules
-final ClusterProcessor<AbstractStructuralNode> processor =
+ClusterProcessor<AbstractStructuralNode> processor =
              clusterProcessorFactory.create();
@@
 for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
-  final AbstractStructuralNode node =  
+  AbstractStructuralNode node =  
                           processor.processCluster(clusterer, cluster.getMembers());
-  final RegularExpressionCleaner<AbstractStructuralNode> cleaner = 
+  RegularExpressionCleaner<AbstractStructuralNode> cleaner = 
                           regularExpressionCleanerFactory.<AbstractStructuralNode>create();
@@
 \end{verbatim}
-It creates clusterer, gives it all rules, orders it to cluster elements.
-Then, empty list of elements is created as final (simplified) grammar.
-For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is called to do infering of regular expression of that element.
-Finally, regular expression cleaning is done in submodule and new \code{Element} instance is created as a copy of processed node, but with infered and cleaned regexp.
+First, it creates a clusterer, gives it all the rules and performs the clustering.
+Then, empty list of elements is created as the final (simplified) grammar.
+For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
+Finally, regular expression cleaning is done in a submodule and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.
 
@@
 There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
-Sentinels can be only on right side of rules and they have to be send to clusterer, as someone further in the chain can ask for cluster of sentinel element - which appears on right side.
-But sentinels have no content, they must not be taken into account when infering attribute properties (whether it is required or optional) - since sentinel have no attributes at all.
+Sentinels can be only on right side of rules and they have to be sent to clusterer, as someone further in the chain can ask for cluster of sentinel element - which appears on right side.
+But sentinels have no content, they must not be taken into account when inferring attribute properties (whether it is required or optional) - since sentinels have no attributes at all.
 
@@
 \subsection{Clusterer submodule}
-We implement clustering in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
-One cluster is represented by class \code{Cluster<T>} with \code{T} as type of clustered items.
-This class simply holds java set of members of cluster and one of the references is held also in \code{representant} member.
+Clustering is implemented in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
+One cluster is represented by the \code{Cluster<T>} class with \code{T} as type parameter of clustered items.
+This class simply holds a \code{Set} of members of this cluster and one of the references as \code{representant} member.
 
-We provide \code{Clusterer} interface for classes to implement.
-Its purpose is to cluster bunch of elements (rules) on input, into  bunch of \code{Cluster} class instances (cluters) on output.
+Any clusterer has to comply to \code{Clusterer} interface.
+Its purpose is to cluster bunch of elements (rules) on input, into bunch of \code{Cluster} class instances (clusters) on output.
 It has methods \code{add()} and \code{addAll()} for adding items for clustering.
-Centerpart is method \code{cluster()}, which does the clustering itself.
+Main part is the method \code{cluster()}, which does the clustering itself.
 
-As it may be time-consuming operation, method throws \code{InterruptedException}.
+As this may be a time-consuming operation, method throws \code{InterruptedException}.
 Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
+% TODO anti Don't understand the following two sentences, please rephrase
 After clustering, implementation should hold clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
 Given item to this method, one can ask for representant of cluster, to which the item belongs.
-If no such cluster exists (item was not added for clustering before), we recommend you throwing an exception rather than returning \code{null}.
-Missing item will probably  indicate error in algorithm rather than normal workflow.
-One can pull clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
-Basic work usage with clusterer is:
+If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
+Missing an item probably indicates error in algorithm rather than normal workflow.
+One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
+Basic usage of the clusterer is:
 \begin{verbatim}
@@
 \subsubsection{ClustererWithAttributes extended interface}
-You noticed that whole \jmodule{Clusterer} interface and \code{Cluster} class are generic.
+You might have noticed that the \jmodule{Clusterer} interface and \code{Cluster} class are generic.
 They may be used as design pattern not only for clustering elements in inference process.
@@
 \subsection{ClusterProcessor module}
-\jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression for that set of elements.
-It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose right name for the resulting element) and with subnodes set to regular expression infered.
-It process attributes of all elements in cluster to obtain meaningful schema attribute specification and theese attributes has to attach to the resulting element.
+\jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression representing that set of elements.
+It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
+It processes attributes of all elements in cluster to obtain meaningful schema attribute specification and these attributes have to be attached to the resulting element.
 
-Worker interface itself is defined as follows:
+% TODO anti Not clear what "worker" is
+Worker interface itself is defined as follows.
 \begin{verbatim}
@@
   T processCluster(
-          final Clusterer<T> clusterer,
-          final List<T> rules
+          Clusterer<T> clusterer,
+          List<T> rules
           ) throws InterruptedException;
@@
 \end{verbatim}
-Why cluster processor is given the clusterer instance?
-Rules themself contain information about which elements to process, but clusterer has more information
-about the topic.
-Clusterer can tell you representant for any element in whole input (not only those elements in rules, but also those that may be on right side of rules), clusterer (if it is with attributes) has information about attributes of each cluster.
+Why is the cluster processor given a clusterer instance?
+Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
+Clusterer can tell you the representant for any element in the whole input (not only those elements in rules, but also those that may be on right side of rules).
+Also, clusterer (if it is with attributes) has information about attributes in each cluster.
 
-We describe each \jmodule{ClusterProcessor} implementation we've got in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.
+We describe each \jmodule{ClusterProcessor} implementation shipped in jInfer in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.
 
@@
 public interface RegularExpressionCleaner<T> {
-  Regexp<T> cleanRegularExpression(final Regexp<T> regexp);
+  Regexp<T> cleanRegularExpression(Regexp<T> regexp);
 }
@@
 // 1. Construct PTA
-final Automaton<AbstractStructuralNode> automaton = new Automaton<AbstractStructuralNode>(true);
+Automaton<AbstractStructuralNode> automaton = new Automaton<AbstractStructuralNode>(true);
 
@@
 for (AbstractStructuralNode instance : rules) {
-  final Element element = (Element) instance;
-  final Regexp<AbstractStructuralNode> rightSide = element.getSubnodes();
+  Element element = (Element) instance;
+  Regexp<AbstractStructuralNode> rightSide = element.getSubnodes();
 
-  final List<AbstractStructuralNode> rightSideTokens = rightSide.getTokens();
+  List<AbstractStructuralNode> rightSideTokens = rightSide.getTokens();
 
-  final List<AbstractStructuralNode> symbolString = new LinkedList<AbstractStructuralNode>();
+  List<AbstractStructuralNode> symbolString = new LinkedList<AbstractStructuralNode>();
   for (AbstractStructuralNode token : rightSideTokens) {
@@
 // 2. Simplify automaton by merging states using automatonSimplifier
-final Automaton<AbstractStructuralNode> simplifiedAutomaton =
+Automaton<AbstractStructuralNode> simplifiedAutomaton =
     automatonSimplifier.simplify(automaton, elementSymbolToString);
@@
 // 3. Convert Automaton<AbstractStructuralNode> to RegexpAutomaton<AbstractStructuralNode>
-final RegexpAutomaton<AbstractStructuralNode> regexpAutomaton =
+RegexpAutomaton<AbstractStructuralNode> regexpAutomaton =
     new RegexpAutomaton<AbstractStructuralNode>(simplifiedAutomaton);
 // 4. Call regexpAutomatonSimplifier to obtain regular expression from regexpAutomaton
-final Regexp<AbstractStructuralNode> regexp =
+Regexp<AbstractStructuralNode> regexp =
     regexpAutomatonSimplifier.simplify(regexpAutomaton, regexpAbstractToString);
@@
 \begin{verbatim}
-public void buildPTAOnSymbol(final List<T> symbolString) {
+public void buildPTAOnSymbol(List<T> symbolString) {
   State<T> xState= this.getInitialState();
   for (T symbol : symbolString) {
-    final Step<T> xStep= this.getOutStepOnSymbol(xState, symbol);
+    Step<T> xStep= this.getOutStepOnSymbol(xState, symbol);
     if (xStep != null) {
@@
     } else {
-      final State<T> newState= this.createNewState();
-      final Step<T> newStep= this.createNewStep(symbol, xState, newState);
+      State<T> newState= this.createNewState();
+      Step<T> newStep= this.createNewStep(symbol, xState, newState);
       assert newStep.getDestination().equals(newState);
@@
 public interface AutomatonSimplifier<T> {
-  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
-                        final SymbolToString<T> symbolToString) throws InterruptedException;
-  Automaton<T> simplify(final Automaton<T> inputAutomaton, 
-                        final SymbolToString<T> symbolToString,
-                        final String elementName) throws InterruptedException;
+  Automaton<T> simplify(Automaton<T> inputAutomaton, 
+                        SymbolToString<T> symbolToString) throws InterruptedException;
+  Automaton<T> simplify(Automaton<T> inputAutomaton, 
+                        SymbolToString<T> symbolToString,
+                        String elementName) throws InterruptedException;
 }
@@
 public interface RegexpAutomatonSimplifier<T> {
-  Regexp<T> simplify(final RegexpAutomaton<T> inputAutomaton, 
-                     final SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
+  Regexp<T> simplify(RegexpAutomaton<T> inputAutomaton, 
+                     SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
 }

### 2011-01-30 doc/jInfer/basicdtd/basicdtd.tex
regenerating, ahonen cite to make bibtex happy
Some quote and cite corrections to basictdt
fixed overfull hboxes by package names
written basicdtd element processing and better screenshot in twostep

@@
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl}.
+The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provi\-der is \code{Sche\.ma\.Ge\.ne\.ra\.tor\.Impl}.
 Its \code{start} method first topologically sorts all rules (elements) it got in the simplified grammar on input.
@@
 \subsection{Element content export}
+Elements are processed in method \code{elementToString()}, but the real work is done in method \code{regexpToString()}, which takes regexp and recursively converts it to string representation.
+But before, whole element is sent to method \code{expandIntervalsElement} of class \code{IntervalExpander}.
+Its purpose is to convert intervals on regexp and its children to those, that are representable in DTD.
+For example regular expression $(a\{2,5\}, b\{0,2\})$ would be transformed to $(a,a,a?,a?,a?,b?,b?)$.
+Class \code{IntervalExpander} works recursively.
+First, element is passed to \code{expand\.In\.ter\.vals\.E\.le\.me\.nt}, which does (shortened):
+\begin{verbatim}
+public Element expandIntervalsElement(final Element treeBase) {
+  return new Element(
+    treeBase.getContext(),
+    treeBase.getName(),
+    treeBase.getMetadata(),
+    expandIntervalsRegexp(treeBase.getSubnodes()),
+    treeBase.getAttributes());
+}
+\end{verbatim}
+It calls private method \code{expandIntervalsRegexp} to handle regexp in that element.
+This private method makes big switch according to regexp type.
+For $\lambda$, it returns $\lambda$.
+Otherwise, it examines interval of regexp in method \code{is\.Sa\.fe\.In\.ter\.val}.
+In DTD one can represent $+, ?, \ast$, so safe intervals are $\{1, \infty\}, \{0, 1\}, \{0, \infty\}$ respectively.
+If interval is not safe, it has to be expanded.
+It is easy to do so, first output $min$-times the regexp itself - that is the minimum occurences, with interval set to $\{1, 1\}$.
+Then, if interval is bounded, output $max - min$-times the regexp itself with interval $\{0, 1\}$ - that is optional part.
+If it is unbounded, attach the regexp once, with interval $\{0, \infty\}$.
 
-TODO anti
+After intervals are expanded, further processing in \code{regexpToString} is divided by big \code{switch} statement of type of regexp.
+For $\lambda$, it simply returns \code{EMPY} as string.
+Tokens are first examined if they are \code{SimpleData}, if so, string \code{\#PCDATA} is returned.
+If not, element name is returned.
+If interval of this regexp is different from $\{1, 1\}$, the interval \code{toString()} representation is appended.
+
+Little complication is with complex regexps, that contain \code{SimpleData} somewhere inside tree.
+They are processed in \code{comboToString} method, it first checks, if there are no simple data in whole tree.
+If not, regexp can be outputted just as list, e.g. $(a, b, c)$ or $(a | b | c)$ or $(a \& b \& c)$.
+If there is at least one simple data, flattening is applied.
+That means, all elements from regexp are collected into one flat list.
+All simple datas are trashed away.
+On output is string \code{(\#PCDATA, a, b, c, d)*}, as this is the only way to represent mixed content in DTDs.
 
@@
 	\item For each element, relevant portion of DTD schema is generated.
-	\item String representation of the schema is returned along with the information that file extension should be "dtd".
+	\item String representation of the schema is returned along with the information that file extension should be ``dtd''.
 \end{enumerate}
 
-%\nocite{*}
+\nocite{*}
 \newpage

### 2011-01-30 doc/jInfer/basicigg/basicigg.tex
written basicdtd element processing and better screenshot in twostep

@@
 Due to the nature of NBP lookups, each processor is internally kept as a singleton, and should not use its inner state (refer to the chapter Lookups in \cite{archdoc}).
-Note that the factory pattern is not used here (see \cite[Factory]{archdoc}).\\ % TODO anti reference to a chapter number when factory pattern is moved to arch
+Note that the factory pattern is not used here (see \cite[section 4.2]{archdoc}).\\
 Each processor declares the class of inputs (documents, schemas, queries) and file extensions it is able to handle by implementing methods \code{getFolder()}, \code{getExtension()} and \code{processUndefined()}.

### 2011-01-30 doc/jInfer/twostep/twostep.tex
Documentation: TwoStep review.

@@
 \subsection{RegularExpressionCleaner module}
-Last we examine \jmodule{RegularExpressionCleaner} interface.
+Finally we examine the \jmodule{RegularExpressionCleaner} interface.
 Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
-For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such like $(name, (person, id))$.
+For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
 To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
-Interface is definition is straightforward:
+Interface definition is straightforward here.
 \begin{verbatim}
@@
 Given regular expression, return regular expression.
-Converting of regular expression is commonly a recursive task and all our implementations work this way.
+Conversion of regular expression is commonly a recursive task and all our implementations work this way.
 
@@
 All settings provided by \jmodule{TwoStepSimplifier} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.twostep.properties} package.
-It is possible to:
+It is possible to set the following.
 \begin{itemize}
-	\item select \jmodule{Clusterer} submodule implementation from those installed
-	\item select \jmodule{ClusterProcessor} submodule implementation from those installed
-	\item select \jmodule{RegularExpressionCleaner} submodule implementation from those installed
+	\item Select \jmodule{Clusterer} submodule implementation from those installed.
+	\item Select \jmodule{ClusterProcessor} submodule implementation from those installed.
+	\item Select \jmodule{RegularExpressionCleaner} submodule implementation from those installed.
 \end{itemize}
@@
 \end{figure}
-Process illustrated on fig. \ref{twostep_flow}:
+Process illustrated in fig. \ref{twostep_flow} is as follows.
 \begin{enumerate}
-	\item On input there is Initial Grammar, which is ruled into \jmodule{Clusterer}.
-	\item Then, each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
+	\item Initial Grammar on input, which is ruled into \jmodule{Clusterer}. % TODO anti WTF is "ruled into"?
+	\item Each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
 	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
-	\item From each cluster processing, regular expression is added to the list of simplified grammar rules.
+	\item Regular expressions representing all the processed clusters are added to the list of simplified grammar rules.
 	\item List of simplified rules is returned (Simplified grammar).
@@
 %#### SUBMODULES ############################################
-\section{Submodules implementations}
+\section{Submodule implementations}
 %#### SUBMODULES->Iname ############################################
@@
 We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
-Clustering is done based on elements \code{getName} equality test (ignore case).
-For each rule (element), it is first clustered by finding cluster where representant has same name - or create new cluster with this element as representant.
-Then we process right side of this element rule (that are nodes from \code{getSubnodes}).
-Since right side of rule is always concatenation, we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
-Each node on right side, is examined:
+% TODO anti I don't understand the sentence, rewrite.
+Clustering is done based on element's \code{getName()} equality (ignoring case).
+For each rule (element), it is first clustered by finding the cluster which representant has the same name - or create new cluster with this element as representant.
+Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
+Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
+Each node on right side, is examined in the following way.
 \begin{itemize}
-  \item if it is simple data throw it to \code{SimpleDataClusterer} class which is associated with main element cluster.
+  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster.
 It does nothing more, than holds all \code{SimpleData} instances in one cluster.
 But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
-  \item if it is element and it is tagged as sentinel by metada - search main clusters to find cluster with representant of same \code{getName()}, or create new cluster with this sentinel as representant.
+  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representant of same \code{getName()}, or create new cluster with this sentinel as representant.
 From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future).
 So there can't be cluster with only one sentinel element in it.
-  \item do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammr, thus it is already processed or is on schedule.
+% TODO anti Please rewrite the really long sentence.
+  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule.
+% TODO anti Totally unclear, rewrite
 \end{itemize}
-Attributes of element are processed through helper attribute clusterer, which is created for each element cluster.
-We have bunch of elements of same name in cluster and one attribute clusterer associated with this bunch.
+Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
+We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
 This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
-It clusters them by name (case insensitive).
+It clusters them by name, case insensitive.
 Each main cluster also has \code{SimpleData} clusterer associated with it.
@@
 
-To rehearse, there are main clusters for elements, for each main cluster there are two helper clusterer - one for attribute clusters and one for simple data clusters.
+Once again, there are main clusters for elements, for each main cluster there are two helper clusterers - one for attribute clusters and one for simple data clusters.
 

### 2011-01-30 doc/jInfer/twostep/twostep.tex
Documentation: TwoStep review.

@@
 \subsection{PassRepresentant (ClusterProcessor)} \label{section_PassRepresentant}
-Simple example to read. For each cluster return its representant as a rule to be in schema.
-This has nothing to do with infering grammar, it is just proof of submodules concept.
+Trivial ClusterProcessor implementation. For each cluster return its representant as the rule to be in resulting schema.
+This has nothing to do with grammar simplification, it is just a proof-of-concept submodule.
 Input documents are not valid against this odd grammar.
-Do not use this in practice, just read the code to understand the bare minimum needed to implement submodule.
+Do not use this in practice, just read the code to understand the bare minimum needed to implement a submodule.
 
@@
 \end{figure}
-\jmodule{AutomatonMergingState} is implementation of merging state algorithm on nondeterministic finite automaton (see package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate}).
+\jmodule{AutomatonMergingState} is an implementation of merging state algorithm on nondeterministic finite automaton (see package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate}).
 \subsubsection{Structure} \label{section_AutomatonMergingState_structure}
-Shortened code of cluster processor operation:
+Abridged code of cluster processor operation follows.
 \begin{verbatim}
@@
     new RegexpAutomaton<AbstractStructuralNode>(simplifiedAutomaton);
+
 // 4. Call regexpAutomatonSimplifier to obtain regular expression from regexpAutomaton
@@
 \end{verbatim}
-The whole submodule structure is drawn on fig. \ref{automaton_merging_state_modules}.
+The whole submodule structure is in fig. \ref{automaton_merging_state_modules}.
 
 \subsubsection{Data flow}
-It works in steps:
+\jmodule{AutomatonMergingState} works in following steps.
 \begin{enumerate}
-	\item It creates prefix-tree automaton (PTA) from positive examples - right sides of rules given.
-Then it calls its submodule called \jmodule{AutomatonSimplifier} to  modify PTA to some generalized 
-automaton by merging states.
+	\item Create prefix-tree automaton (PTA) from positive examples - right sides of rules.
+Call its submodule \jmodule{AutomatonSimplifier} to modify PTA to some generalized automaton by merging states.
 	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.
@@
 	\item \code{RegexpAutomaton} is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
-Its job is to derive regular expression from automaton, such that automaton and regular expression represents same language.
+Its job is to derive regular expression from automaton, such that automaton and regular expression represent the same language.
 \end{enumerate}
-The data flow is on fig. \ref{automaton_merging_state_flow}.
+The data flow is in fig. \ref{automaton_merging_state_flow}.
 

### 2011-01-30 doc/jInfer/basicdtd/basicdtd.tex
updated documentation to new type of interval expander in dtd

@@
 \subsection{Element content export}
-Elements are processed in method \code{elementToString()}, but the real work is done in method \code{regexpToString()}, which takes regexp and recursively converts it to string representation.
-But before, whole element is sent to method \code{expandIntervalsElement} of class \code{IntervalExpander}.
+Elements are processed in method \code{elementToString()}.
+Whole element regexp is sent to method \code{expandIn\.ter\.vals\.Reg\.exp} of class \code{IntervalExpander}.
 Its purpose is to convert intervals on regexp and its children to those, that are representable in DTD.
@@
 Class \code{IntervalExpander} works recursively.
-First, element is passed to \code{expand\.In\.ter\.vals\.E\.le\.me\.nt}, which does (shortened):
-\begin{verbatim}
-public Element expandIntervalsElement(final Element treeBase) {
-  return new Element(
-    treeBase.getContext(),
-    treeBase.getName(),
-    treeBase.getMetadata(),
-    expandIntervalsRegexp(treeBase.getSubnodes()),
-    treeBase.getAttributes());
-}
-\end{verbatim}
-It calls private method \code{expandIntervalsRegexp} to handle regexp in that element.
-This private method makes big switch according to regexp type.
+Regexp is passed to method \code{expandIntervalsRegexp} to handle regexp in that element.
+This method makes big switch according to regexp type.
 For $\lambda$, it returns $\lambda$.
@@
 Then, if interval is bounded, output $max - min$-times the regexp itself with interval $\{0, 1\}$ - that is optional part.
-If it is unbounded, attach the regexp once, with interval $\{0, \infty\}$.
+If it is unbounded and $min$ is zero, attach the regexp once, with interval $\{0, \infty\}$.
+If $min$ is nonzero, last regexp in required string gets interval $\{1, \infty\}$.
 
-After intervals are expanded, further processing in \code{regexpToString} is divided by big \code{switch} statement of type of regexp.
+After intervals are expanded, returned regexp is passed to \code{regexpToString()}, which is divided by big \code{switch} statement of type of regexp.
 For $\lambda$, it simply returns \code{EMPY} as string.

### 2011-01-30 doc/jInfer/twostep/twostep.tex
Documentation: TwoStep review.

@@
 Class \code{Automaton} is shortly described in \cite{archdoc}.
-In this place, it is interestesting to look on method \code{build\.PTA\.On\.Sym\.bol}.
+Let's take a look at method \code{build\.PTA\.On\.Sym\.bol} here.
 Given \code{List<T> symbolString} it traverses automaton - comparing transition symbols with symbols in string.
@@
 On different symbol, it creates new branch and all states and transitions on that branch.
-Last state visited got's incremented its \code{finalCount} value, as one more input string ended in this state.
-Lets see code:
+Last state visited gets its \code{finalCount} value incremented, as one more input string ended in this state.
+The code is as follows.
 \begin{verbatim}
@@
 \end{verbatim}
-Start from initial state traverse automaton by asking for \code{getOutStepOnSymbol}.
+Starting from initial state, traverse the automaton by asking for \code{getOutStepOnSymbol()}.
 If there is such step (transition), follow it to its destination state and move the position in \code{symbolString}.
@@
 As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
-By the way, that is why we cluster all simple data and all sentinels on right sides of rules.
+By the way, this is why we cluster all simple data and all sentinels on right sides of rules.
 To automaton, instead of instances on right side of rules, their cluster representants are inserted as one can see in section \ref{section_AutomatonMergingState_structure}.
-This assures, that comparisons in automaton are made by using java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
+% TODO anti Don't understand the sentence
+This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
 That results in cluster-equal behaviour.
-Example of prefix-tree automaton is on fig. \ref{pta}.
+Example of prefix-tree automaton is in fig. \ref{pta}.
 
 \subsubsection{Data structures - RegexpAutomaton} \label{section_RegexpAutomaton}
-Class \code{RegexpAutomaton} is extending class \code{Automaton} with generic type \code{T} set to \code{Regexp<T>}.
+Class \code{RegexpAutomaton} extends class \code{Automaton} with generic type \code{T} set to \code{Regexp<T>}.
 That is, you have to instantiate it with same symbol type (e.g. \code{AbstractStructuralNode}) as automaton.
-Implementation can be found in package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate.regexping}, that is same, as where \jmodule{RegexpAutomatonSimplifier} interface is defined.
-Implementation is nothing special, it has only copy constructor which creates regular expression automaton from ordinary automaton by enclosing each symbol on transition of original automaton by \code{Regexp.TOKEN}.
+Implementation is in package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate.regexping}. This is the same package where \jmodule{RegexpAutomatonSimplifier} interface is defined.
+Implementation is nothing special - there is a copy constructor which creates regular expression automaton from ordinary automaton by enclosing each symbol on transition of original automaton in a \code{Regexp.TOKEN}.
 
 \subsubsection{AutomatonSimplifier module}
-\jmodule{AutomatonSimplifier} module has its worker interface defined as:
+\jmodule{AutomatonSimplifier} module has its worker interface defined as following.
 \begin{verbatim}
@@
 \end{verbatim}
-Given input automaton it returns automaton, there is nothing magical on it.
-Interface is defined as generic, implementation don't have to be, but there is no reason to not make them generic.
-When implementation needs to present automaton to user, however, it needs some string representation of symbols to display on automaton transitions.
+Given input automaton it returns automaton, there is nothing magical to it.
+Interface is defined as generic, implementations don't have to be. However, there is no reason not to make them generic.
+When implementation needs to present an automaton to user, it needs string representation of symbols to display on automaton transitions.
 For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
@@
 To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule.
-Overloaded version is to inform user about name of processed element when presenting automaton by \jmodule{AutoEditor}.
+% TODO anti Unreadable....
+Overloaded version is to inform the user about name of processed element when presenting automaton in \jmodule{AutoEditor}.
 This version of interface will probably change in future.
 
-By this genericity, simplifier can be used to perfom same algorithms for simplifying not only to infer rules for elements, but maybe also to infer patterns strings for content model of attributes.
+This generality means that simplifier can not only be used to infer rules for elements, but possibly also to infer patterns strings for content model of attributes using the same algorithms.
 The actual implementation of this still waits for its developer to come.
@@
 \subsubsection{RegexpAutomatonSimplifier submodule}
-Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain regular expression from given automaton.
-Interface for doing this is thus simple:
+Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain a regular expression from a given automaton.
+Interface for doing is is thus simple.
 \begin{verbatim}
@@
 \end{verbatim}
-Once again we are encountering \code{symbolToString} with same purpose as before, if submodule would need to present automaton to user
-it has to be able to convert symbols of (at runtime) unknown type to strings.
+Once again we encounter \code{symbolToString} with the same purpose as before, if submodule would need to present the automaton to user, it has to be able to convert symbols of (at runtime) unknown type to strings.
 Our implementation of \jmodule{RegexpAutomatonSimplifier} is discussed in \ref{section_StateRemoval}
@@
 \jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{Au\.to\.ma\.ton\.Merg\.ing\.Sta\.te}.
-It is at disposal for implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
+It is at disposal of implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
 
-Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states.
+Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states. % TODO anti Don't understand
 One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.
@@
 \subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
-This cleaner does exactly what we have just described.
+This cleaner does exactly what we have just described. % TODO anti Doesn't make sense
 It converts nested concatenations into flat ones.
@@
 Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
-For example some pitty inferring method produces unnecessary empty regexps in regexp tree as:
+If for example an inference method produces unnecessary empty regexps in regexp tree such as
 $$
@@
 \subsection{Chained (RegularExpressionCleaner)} \label{section_Chained}
-Enables user to chain more cleaners with output from first one plugged as input into second one and so one.
-Thus regular expression from inference can pass through \jmodule{EmptyChildren} and them through \jmodule{NestedConcatenation} cleaners and then it is returned.
+Allows the user to chain more cleaners with output from first one plugged as input into second one and so one.
+Thus regular expression from inference can pass through \jmodule{EmptyChildren} and then through \jmodule{NestedConcatenation} cleaners and then it is returned.
 
@@
 We implement greedy strategy of automaton simplifying in class \code{Greedy}.
-It simply asks given \code{MergeConditionTester} if it can merge any of automaton states and merges states until there are no states to be merged:
+It simply asks given \code{MergeConditionTester} if it can merge any of automaton states and merges states until there are no states to be merged.
 \begin{verbatim}
@@
 \subsection{UserInteractive (AutomatonSimplifier)} \label{section_UserInteractive}
-We have created automaton simplifier which displays user input automaton asking him to select some states to merge.
-Then it merges them and asks user again:
+We have created automaton simplifier which shows the user the input automaton, asking him to select states to merge.
+Then it merges them and asks user again.
 \begin{verbatim}
@@
   }
-} until (user selected at least one state in last trial)
+} until (user selected at least one state in last trial) TODO anti Trial -> Attempt?
 \end{verbatim} 
 We use \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
-To make reader more happy, we include one screenshot of automaton visualization on picture \ref{screenshot}.
+A screenshot of automaton visualization is in picture \ref{screenshot}.
 \begin{figure}[tpb]
@@
 \subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
-We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergability of states by default configuration.
-TODO vektor : mention this? (we have stupid algorithm:)
-We don't use $k-grams$ algorithm from \cite{ahonen}, but a simple DFS.
+We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergeability of states by default configuration.
+% TODO anti Sentence unclear
+% TODO vektor : mention this? (we have stupid algorithm:) TODO anti Of course;)
+We don't use $k-grams$ algorithm from \cite{ahonen}, just a simple DFS.
 
@@
 This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
-before algorithm starts, states superInitial and superFinal are added to automaton, former one with $\lambda$-transition to initial state.
-From all final states $\lambda$ transition to superFinal state is added.
+% TODO anti Emphasise the states in following sentences
+Before the algorithm starts, states superInitial and superFinal are added to automaton, former with $\lambda$-transition to initial state.
+From all final states a $\lambda$ transition to superFinal state is added.
 State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - superInitial and superFinal.
 After removing all states, there is only one transition from superInitial to superFinal state.
-That transition has final regular expression on it as symbol, it is read and returned.
+That transition has the final regular expression as symbol on it. This one is returned.
 
@@
 \subsubsection{Data Structures - StateRemovalRegexpAutomaton}
-We extend \code{RegexpAutomaton} to obtain automaton that fits state removing procedure.
+We extend \code{RegexpAutomaton} to obtain automaton that fits the state removal procedure.
 This implementation has methods \code{crea\.te\.Su\.per\.In\.it\.ial\.Sta\.te} and \code{crea\.te\.Su\.per\.In\.it\.ial\.Sta\.te}.
 They does exactly what their names stands for.
+% TODO anti Emphasise the states again
 SuperInitial state has only one $\lambda$-out-transition pointing to original initial state.
-SuperFinal state has as many $\lambda$-in-transitions, as there are original final states - from each one, it gets one.
-Without going further to many technical details, we just mention method \code{removeState}, which removes given state from automaton.
-Transitions are handled this way:
+SuperFinal state has as many $\lambda$-in-transitions, as there are original final states - from each one it gets one.
+Without going further into many technical details, we just mention method \code{removeState}, which removes given state from automaton.
+Transitions are handled in the following way.
 \begin{enumerate}
-	\item Collapse all loop transition of removed state to one loop transition with new \code{Regex.ALTERNATION} of all regexps on original loops (with Kleene interval set).
+	\item Collapse all loop transitions of removed state to one loop transition with new \code{Regex.ALTERNATION} of all regexps on original loops (with Kleene interval set).
 	\item Collapse all {in | out}-transitions, which same {source | destination} state to one transition with \code{Reg\.exp.AL\.TER\.NA\.TION} of original regexps.
@@
 \end{enumerate}
-Whole process is step-by-step illustrated on fig. \ref{regaut}.
+The whole process is step-by-step illustrated in fig. \ref{regaut}.
 \begin{figure}[pb]
@@
 If you are willing to, you can replace each of three submodules of \jmodule{TwoStepSimplifier}.
-It may be useful to replace default \jmodule{ClustererWithAttributes} implementation to a clever one.
-And regular expression cleaning will be probably target of adding new implementations, too.
+It may be useful to replace the default \jmodule{ClustererWithAttributes} implementation with a clever one.
+And regular expression cleaning will be probably the aim of new implementations, too.
 We recommend replacing other parts at lower levels, however.
-You will need to replace \jmodule{ClusterProcessor} only if you want to write infering method that is not based on merging state algorithm (or NFA's at all).
+You will need to replace \jmodule{ClusterProcessor} only if you want to write inference method that is not based on merging state algorithm (or NFA's at all).
 
@@
 
-Of course, merge criterion is propably the point to add implementation, implement \jmodule{MergeConditionTester} interface to obtain for example $s,k$-string merge criterion.
+Of course, merge criterion is probably the point to add implementation, implement \jmodule{MergeConditionTester} interface to obtain for example $s,k$-string merge criterion.
 
-Changing ordering of states to remove in \jmodule{StateRemoval} module is easy by replacing \jmodule{Orderer} implementation, for example to one of suggested in \cite{1224321} may lead to shorter resulting regular expressions.
+Changing ordering of states to remove in \jmodule{StateRemoval} module is easy by replacing \jmodule{Orderer} implementation, for example to one suggested in \cite{1224321} may lead to shorter resulting regular expressions.
 
-Whole TwoStep submodules structure is on fig. \ref{twostep_modules_complete}.
+Whole TwoStep submodules structure is in fig. \ref{twostep_modules_complete}.
 \begin{figure}[tpb]

### 2011-01-30 doc/jInfer/basicdtd/basicdtd.tex
Documentation: BasicDTD review.

@@
 \subsection{Element content export}
-Elements are processed in method \code{elementToString()}.
+Elements are processed \code{elementToString()} method.
 Whole element regexp is sent to method \code{expandIn\.ter\.vals\.Reg\.exp} of class \code{IntervalExpander}.
-Its purpose is to convert intervals on regexp and its children to those, that are representable in DTD.
+Its purpose is to convert intervals on regexp and its children to those representable in DTD.
 For example regular expression $(a\{2,5\}, b\{0,2\})$ would be transformed to $(a,a,a?,a?,a?,b?,b?)$.
@@
 Regexp is passed to method \code{expandIntervalsRegexp} to handle regexp in that element.
-This method makes big switch according to regexp type.
+This method contains a big\code{switch} based on the regexp type.
 For $\lambda$, it returns $\lambda$.
@@
 In DTD one can represent $+, ?, \ast$, so safe intervals are $\{1, \infty\}, \{0, 1\}, \{0, \infty\}$ respectively.
-If interval is not safe, it has to be expanded.
+If an interval is not safe, it has to be expanded.
 It is easy to do so, first output $min$-times the regexp itself - that is the minimum occurences, with interval set to $\{1, 1\}$.
@@
 If it is unbounded and $min$ is zero, attach the regexp once, with interval $\{0, \infty\}$.
-If $min$ is nonzero, last regexp in required string gets interval $\{1, \infty\}$.
+If $min$ is non-zero, last regexp in required string gets interval $\{1, \infty\}$.
 
-After intervals are expanded, returned regexp is passed to \code{regexpToString()}, which is divided by big \code{switch} statement of type of regexp.
+After intervals are expanded, resulting regexp is passed to \code{regexpToString()}, which contains a big \code{switch} statement based on the type of regexp.
 For $\lambda$, it simply returns \code{EMPY} as string.
@@
 
-Little complication is with complex regexps, that contain \code{SimpleData} somewhere inside tree.
-They are processed in \code{comboToString} method, it first checks, if there are no simple data in whole tree.
+Situation is a bit complicated with complex regexps that contain \code{SimpleData} somewhere inside the tree.
+They are processed in \code{comboToString} method which first checks if there are no simple data in whole tree.
 If not, regexp can be outputted just as list, e.g. $(a, b, c)$ or $(a | b | c)$ or $(a \& b \& c)$.

### 2011-01-30 doc/jInfer/architecture/architecture.tex
Documentation: Architecture review.

@@
 
-Regular expression is implemented as class \code{Regexp<T>} with supplementing classes \code{RegexpInterval} and \code{Reg\.exp\.Type}.
-Each \code{Regexp<T>} instance has one of the enum \code{RegexpType} type:
+Regular expression is implemented as class \code{Regexp<T>} with supporting classes \code{RegexpInterval} and \code{Reg\.exp\.Type}.
+Each \code{Regexp<T>} instance has a type of \code{RegexpType} enum.
 \begin{itemize}
-	\item Lambda ($\lambda$) - empty string (also called $\epsilon$ in literature),
-	\item Token - a letter of the alphabet,
-	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$,
-	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$,
-	\item Permutation - shortcut for all possible permutations of regular expressions. Our syntax to write down permutation is $(a\& b\& c\& d)$.
+	\item Lambda ($\lambda$) - empty string (also called $\epsilon$ in literature).
+	\item Token - a letter of the alphabet.
+	\item Concatenation - one or more regular expression in an ordered sequence. Eg. $(a, b, c, d)$.
+	\item Alternation - a choice between one or more regular expressions. Eg. $(a | b | c | d)$.
+	\item Permutation - shortcut for all possible permutations of regular expressions. Our notation is $(a\& b\& c\& d)$.
 \end{itemize}
@@
 Each \code{Regexp<T>} instance has one instance of \code{RegexpInterval} as member.
-Class \code{RegexpInterval} represents POSIX-like intervals for expression:
+Class \code{RegexpInterval} represents POSIX-like intervals for expression.
 \begin{itemize}
-	\item $a\{m,n\}$ means $a$ at least $m$-times, at most $n$-times,
+	\item $a\{m,n\}$ means $a$ at least $m$-times, at most $n$-times.
 	\item $a\{m,\}$ means at least $m$-times (unbounded).
@@
 Interval can be either bounded (you have to set both lower and upper bound integers), or unbounded (you have to set only lower bound).
-Testing interval value commonly follows routine:
+Testing interval value commonly follows this routine.
 \begin{verbatim}
@@
 
-Class \code{Regexp<T>} can represent regular expression over any alphabet.
-This is done by using java generics.
-Only token regexps hold instance of type \code{T} in member \code{content}.
+Using Java generics, \code{Regexp<T>} can represent regular expression over any alphabet.
+Only token regexps actually hold instance of type \code{T} in member \code{content}.
 
@@
 \end{figure}
-We implement this tree by member of \code{Regexp<T>} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
-List contains children of regexp in means of regexp tree.
+We implement this tree in a member of \code{Regexp<T>} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
+List contains children of regexp: a node contains a list of child nodes in the tree view.
 
-Regexp has to obey constraits:
+Regexp has to obey the following constraits.
 \begin{itemize}
-	\item type, children and interval have to be non-null references,
-	\item when type is lamba, content and interval has to be null,
-	\item when type is token, content has to be non-null,
-	\item when type concatenation, alternation or permutation, content has to be null.
+	\item Type, children and interval have to be non-\code{null} references.
+	\item When type is lamba, content and interval has to be \code{null}.
+	\item When type is token, content has to be non-\code{null}.
+	\item When type concatenation, alternation or permutation, content has to be \code{null}.
 \end{itemize}
-These constraits are checked by constructors, so the best way to construct new regexps is by using 
-methods \code{get\.To\.ken(), get\.Con\.ca\.te\.na\.tion()} etc.
+These constraits are checked by constructors, therefore the best way to construct new regexps is by using methods \code{get\.To\.ken(), get\.Con\.ca\.te\.na\.tion()} etc.
 
-For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of members set.
+For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of the members set.
 One has to fill in all properties carefully and call \code{setImmutable()} aftewards.
-Proper usage should be one of following:
+Proper usage should be one of following: % TODO anti WTF?
 Regexp instance is by default created as immutable, that is, once instantiated, you cannot add more children to list of children, cannot change type, content etc.
-It is to prevent missuse.
-In special circumstances, one does not know future children of regexp in time of creation.
-This occurs mainly in input modules, where by parsing XML data sequentially, one does not know contents of element in time of handling start element event.
+In special circumstances, one does not know future children of regexp at the time of creation.
+This occurs mainly in input modules: while parsing XML data sequentially, one does not know contents of element in time of handling start element event.
 \begin{verbatim}
@@
 
-Finally, regexp contain one useful method for obtaining all leaves in the regexp tree, it is called \code{getTokens()} and it recursively traverses tree returning list of leaves (token type regexps).
+Finally, regexp contains a useful method to obtain all leaves in the regexp tree. It is called \code{getTokens()} and it recursively traverses tree returning list of leaves (token type regexps).
 
 \subsection{XML representation}
-XML data basically contains elements, text nodes (characters inside elements) and attributes.
-For maximum generality, we decided to break apart theese objects.
+XML data basically encompasses elements, text nodes (characters inside elements) and attributes.
+For maximum generality, we decided to break apart these objects.
 We define three basic interfaces: \code{NamedNode}, \code{StructuralNode} and \code{ContentNode} (see package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes}).
 
-The first stands for bare node in XML document tree, it has its name and context withing the tree (path from root).
-The latter two extends \code{NamedNode} interface.
-\code{StructuralNode} is for nodes, which form structure of XML document tree: elements and text nodes.
-\code{ContentNode} is for nodes, that have content in XML documents: text nodes and attributes.
+The first stands for a bare node in XML document tree, it has its name and context within the tree (path from root).
+The latter two extend \code{NamedNode} interface.
+\code{StructuralNode} is for nodes which form structure of XML document tree: elements and text nodes.
+\code{ContentNode} is for nodes that have content in XML documents: text nodes and attributes.
 We have three classes: \code{Element} for elements, \code{SimpleData} for text nodes, \code{Attribute} for attributes (see package \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes}).
-In theory, the classes and interfaces would be layed out as on fig. \ref{interfaces_nodes}
+In theory, the classes and interfaces would be layed out as in fig. \ref{interfaces_nodes}
 \begin{figure}
@@
 \begin{itemize}
-	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metadata}),
-	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
+	\item \code{AbstractNamedNode} which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metadata}),
+	\item \code{AbstractStructuralNode} which implements only the task of deciding if instance is \code{Element} or \code{SimpleData}.
 \end{itemize}
-As practice showed, for methods handling and infering structural properties, it is important to recognize whether structural node on input is element or text node.
-However methods for content devising don't need to know, if they are working on infering model for content of attribute or text node.
+As practice showed, for methods handling and infering structural properties, it is important to recognize whether structural node on input is element or text node. % TODO anti Unclear, rewrite
+However methods for content devising don't need to know if they are working on infering model for content of attribute or text node. % TODO anti Unclear, rewrite
 
-Finally, our interface/class model for representing XML nodes is drafted on fig. \ref{nodes}. Those, who are brave enough, can look on fig. \ref{nodes_full2}.
-In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{Abstr\.act\.Stru\.ctu\.ral\.Node} variables.
+Finally, our interface/class model for representing XML nodes is drafted in fig. \ref{nodes}. Those who are brave enough can look at fig. \ref{nodes_full2}.
+In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{Abstr\.act\.Stru\.ctu\.ral\.Node} variables. % TODO anti What is "devise"?
 And \code{SimpleData} and \code{Attribute} have methods \code{getContentType()} and \code{getContent()} to work with content model.
+% TODO anti Last 2 sentences make no sense to me...
 
-Class \code{Element} has two important members of course:
+Class \code{Element} has two important members.
 \begin{itemize}
-	\item \code{Regexp<AbstractStructuralNode> subnodes} - for representing right side of grammar rule in resulting infered schema,
-	\item \code{List<Attribute> attributes} - for representing all attributes in resulting infered schema.
+	\item \code{Regexp<AbstractStructuralNode> subnodes} - representing the right side of a grammar rule,
+	\item \code{List<Attribute> attributes} - representing all attributes in this element.
 \end{itemize}
-Theese two are filled by import modules, processed further by infering (simplifying) modules and finally exported by exporter modules.
+These two are filled in by import modules, processed further by inference (simplification) modules and finally exported by schema generator modules.
 
-As in regular expressions, classes pertaining XML nodes are by default immutable.
+As in regular expressions, classes pertaining to XML nodes are by default immutable.
 For elements, it means no adding of attributes and changing regexp reference (regexp instance itself is immutable as well).
-Same \code{getMutable()} principles and good usage practises as for regexps, hold for theese classes.
+The same \code{getMutable()} principles and good usage practises hold for these classes.
 
@@
 \end{figure}
-Let's take an example, the following XML document would be represented as tree on fig. \ref{xml_example}.
+Let's have an example: the following XML document would be represented as tree in fig. \ref{xml_example}.
 \begin{verbatim}
@@
 \end{verbatim} 
-Although in example we present whole document tree, input modules produce slightly different format (consisting of rules).
+Although in example we present the whole document tree, input modules produce slightly different format (consisting of rules).
 
 \subsection{Rules and grammars} \label{rules_grammar}
-jInfer and its documentation uses extended context-free grammars\cite{extendedcfg}.
+jInfer and its documentation use extended context-free grammars\cite{extendedcfg}.
 \emph{Rules} in such grammar are in the form

### 2011-01-30 doc/jInfer/architecture/architecture.tex
Documentation: Architecture review - finished.

@@
 \subsection{Nondeterministic Finite Automaton}
-We recommend skipping this section in first read as it describes advanced features.
+We recommend skipping this section during the first reading as it describes advanced features.
 
 For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
-Implementation consists of 4 classes (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}): 
+Implementation consists of 4 classes (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
 \begin{itemize}
-  \item \code{Automaton<T>},
-  \item \code{Step<T>},
-  \item \code{State<T>},
-  \item \code{AutomatonCloner<A, B>}.
+  \item \code{Automaton<T>}
+  \item \code{Step<T>}
+  \item \code{State<T>}
+  \item \code{AutomatonCloner<A, B>}
 \end{itemize}
-Whole implementation uses java generics for representation of symbol of alphabet.
-We denote \code{T} the java type of symbol.
+The implementation uses Java generics to represent symbol of alphabet.
+We denote \code{T} the Java type of that symbol.
 
-Take care of symbol class and its \code{equals()} implementation.
+Take care of symbol class and its \code{equals()} implementation. % TODO anti What does it mean, to take care?
 Automaton uses \code{equals()} to compare symbols on transitions (when building prefix-tree automaton and when merging states).
-If you are using strings, you're just fine, but with complicated objects, either take care, that equivalent objects are properly tested in \code{equals()}.
-Or (maybe faster) solution is to cluster objects into classes of equivalence before inserting them into automaton.
+If you are using strings, you're just fine. With complicated objects, the first option is to make sure that equivalent objects are properly tested in \code{equals()}.
+Second, maybe faster solution is to cluster objects into classes of equivalence before inserting them into automaton.
 Then give automaton only cluster representant object (which will use java \code{Object.equals()} with reference comparison) for each object encountered.
@@
 \paragraph{State}
-Let's begin by smallest of the classes, the \code{State<T>}.
+Let's begin with the smallest of classes, the \code{State<T>}.
 It represents automaton state, it has two integer members: \code{name} and \code{finalCount}.
-Name clearly serves as name of state in visualization and to string conversion.
+Name clearly serves as name of state in visualization and to-string conversion.
 Final count is for representing whether the state is final in automaton.
-The field is not true/false but integral to help algorithms use stastistics over automatons (how many times in XML input is this state final?).
+The field is not true/false but integral to help algorithms use statistics over automatons (how many times in XML input is this state final?).
 
@@
 \paragraph{Automaton}
-\code{Automaton<T>} class puts theese together into nondeterministic finite automaton.
+\code{Automaton<T>} class puts these together to form a nondeterministic finite automaton.
 It has reference to \code{initialState}, it has \code{newStateName} integer value to assure unique state names inside one automaton (incrementing every time new state is created).
@@
 One map of type \code{Map<State<T>, Set<Step<T>{}>{}>} called \code{delta} represents mapping from state into set of all outgoing transitions from state.
-Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better perfomance).
-There exist only one instance of each step from one state to another.
-That instance is referenced in delta map (on place of source state), and in reverse delta map (on place of destination state).
-Loops are no speciality, just source = destination.
+Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better performance).
+There is only one instance of each step from one state to another.
+That instance is referenced in delta map (on place of source state) and in reverse delta map (on place of destination state).
+Loops are no speciality, only source = destination.
 
-Automaton supports creating of automaton as a copy of another one (not reference copy, but deep copy expect of symbols), this can be used when searching solution space to create more versions of automaton to edit.
+Automaton supports creation as a copy of another one (not reference copy, but deep copy expect of symbols), this can be used when searching solution space to create more versions of automaton to edit.
 We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
@@
 \end{figure}
-The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates(state1, state2)} method.
+The biggest thing we offer to researchers is state merging by simply calling \code{mergeStates(state1, state2)} method.
 Method merges second state given into first one (or an overloaded version - all states in list into first one in list).
@@
 
-One can ask for merging states, that are merged out from automaton already.
+One can ask for merging states, that are merged out from automaton already. % TODO anti What does it mean?
 Lets take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
@@
 State 3 then becomes final state with loop and state 4 disappears (fig. \ref{automaton_merge2}).
-If then one ask to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
+If then one asks to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
 This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).
@@
 Class \code{AutomatonCloner<A, B>} has one overloaded method \code{convertAutomaton()}.
-First version accepts automaton and class implementing \code{AutomatonClonerSymbolConverter<A, B>} interface, and returns new automaton with same structure, but with symbols on transitions from alphabet of java type \code{B}.
+First version accepts automaton and class implementing \code{AutomatonClonerSymbolConverter<A, B>} interface, and returns new automaton with same structure, but with symbols on transitions from alphabet of Java type \code{B}.
 Second version takes two automatons, second one has to be empty (without initial state created), and symbol converter.
@@
 
-To fulfill symbol conversion, one have to provide imlementation of \code{AutomatonClonerSymbolConverter<A, B>} interface.
-Interface has one method:
+To fulfil symbol conversion, one have to provide implementation of \code{AutomatonClonerSymbolConverter<A, B>} interface.
+Interface has one method.
 \begin{verbatim}
@@
 \end{verbatim}
-Its purpose is to give mapping from symbol over one domain to new symbols.
-Implementation have to be \code{equals} consistent, if in first automaton are two transitions with symbols \code{a}, \code{b} such that
-\code{a.equals(b) == true}, convertor has to produce new symbols that are equal.
+Its purpose is to provide mapping from symbol over one domain to new symbols.
+Implementation has to be \code{equals()} consistent, if in first automaton are two transitions with symbols \code{a}, \code{b} such that
+\code{a.equals(b) == true}, converter has to produce new symbols that are equal.
 
@@
 This modular architecture means that it is possible to replace any and/or all of these modules with \emph{something} else that does similar job.
-In practice, scientists will probably implement new \jmodule{Simplifier} modules, while using our implementation of BasicIGG and export modules.
-This modular scheme is used inside our modules to divide them into submodules.
+In practice, researchers will probably implement new \jmodule{Simplifier} modules, while using our implementation of BasicIGG and export modules.
+This modular scheme is also used inside our modules to divide them into submodules.
 It is possible to add new file format processing to import/export.
 One can extend automaton merging state algorithm of our implementation of \cite{ahonen} in TwoStepSimplifier module by replacing submodules.
+% TODO anti This is a terrible paragraph with no clear path to follow, please reorganize (sentences on their own are OK)
 
@@
 \begin{itemize}
-	\item Only concatenations. This is default format. All regular expressions coming from IGG have to be concatenations, even empty.
-	When importing XML documents, this is just fine, as document is positive example from language we are trying to identify.
-	But even when importing schemas, queries or other stuff that can contain more information, one has to convert it somehow to concatenations.
+	\item \textbf{Only concatenations.} This is the default format. All regular expressions coming from IGG have to be concatenations, even if they are empty.
+	When importing XML documents this is just fine, because a document is a positive example of language we are trying to identify.
+	But while importing schemas, queries or other stuff that can contain more information, it might be necessary to convert it somehow to concatenations.
 	For example, when importing DTD:
@@
 	\end{verbatim}
+	% TODO anti Emphasise elements in the following paragraph
 	Elements note and more are empty, but we cannot use lambda regexp. Import will use empty concatenation instead.
-	For element person, import module have to generate some positive examples of language generated by regular expression \verb+(info?, more, more?, more?)+.
+	For element person, import module has to generate some positive examples of language generated by regular expression \verb+(info?, more, more?, more?)+.
 
-	This generating is basicaly done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{expand()}.
+	This is basically done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{expand()}.
 One can implement own intelligent expander, or use our provided\\ \code{cz.cuni.mff.ksi.jinfer.basicigg.expansion.ExpanderImpl}.
 
-	\item Complex regular expressions. Simplifier module running in current inference can indicate support for this format by having capability called \code{can.handle.complex.regexps}.
+	\item \textbf{Complex regular expressions.} Simplifier module running in current inference can indicate support for this format by declaring capability \code{Capabilities.CAN\_HANDLE\_COMPLEX\_REGEXPS}.
 	Even then, import module is not obliged to use it of course. But it is reasonable to do so, if imported file type gives more information than simple concatenations (when importing schemas).
@@
 Empty concatenations are also allowed - as they are positive examples in XML document input.
-Its simplifier business to divide rules by import file formats (by using metadata described in \ref{section_metadata}) and to handle empty concatenations from XML file import with complex regular expressions from other formats.
+It is simplifier's responsibility to divide rules by import file formats (by using metadata described in \ref{section_metadata}) and to handle empty concatenations from XML file import with complex regular expressions from other formats.
 \end{itemize}
@@
 
-\paragraph{Simplified grammar} has only one format and that is strict regular expression format.
+\paragraph{Simplified grammar} has only one format and that is \textbf{strict regular expression format}.
 Empty concatenations, alternations or permutations are not allowed.

### 2011-01-31 doc/jInfer/basicxsd/basicxsd.tex
Documentation: review of finished part of BasicXSDExporter.

@@
 
-This is an implementation of a \jmodule{SchemaGenerator} exporting the inferred schema to XSD, providing basic features of the language.
+This is an implementation of a \jmodule{SchemaGenerator} exporting the inferred schema to XSD, supporting basic features of the language.
 
@@
 The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl}.
-Process of export consists of two phases described in detail in later sections:
+Process of export consists of two phases described in detail in later sections.
 \begin{enumerate}
-	\item Preprocessing
-	\item Own export to a string representation of XSD
+	\item Preprocessing.
+	\item The export to a string representation itself.
 \end{enumerate}
-Method \code{start} first creates instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
-Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such which elements should be globally defined and which element is the root element.
+Method \code{start()} first creates an instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
+Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such as which elements should be globally defined and which element is the root element.
 This instance is kept as a member variable of module to be used during the whole export process.\\
 
-Afterwards, \code{start} method recursively traverses global elements followed by other elements starting at the root element and for each it creates element's XSD string representation.\\
+Afterwards, \code{start()} method recursively traverses global elements followed by other elements starting at the root element and for each it creates element's XSD string representation.\\
 
@@
 
-As mentioned before, preprocessing is implemented in \code{Preprocessor} class and its functions are following.
+As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its functions are following.
 \begin{itemize}
@@
 
-Constructor of \code{Preprocessor} class gets elements and a number, defining minimal number of occurrences of an element to be defined.
-It first topologically sorts input elements to decide which one is the root element.
-Afterwards, it counts occurrences of the elements and removes unused ones (those which did not occurred).
-Finally, for each element it decides whether mark it as a global one or not.
-An element is considered global if its occurrence count is greater than or equal number of occurrences supplied on input.\\
+Constructor of \code{Preprocessor} class gets a list of elements and a number, defining minimal number of occurrences of an element to be defined globally.
+It first topologically sorts input elements to decide which of the elements is the root element.
+Afterwards, it counts occurrences of the elements and removes unused ones (those which did not occur).
+Finally, for each element it decides whether to mark it as a global one or not.
+An element is considered global if its occurrence count is greater than or equal to the number of occurrences provided on input.\\
 
@@
 
-\subsection{Own export}
+\subsection{Export}
 
-Own XSD export is performed in module's \code{start} function right after the preprocessing.\\
+XSD export itself is performed in module's \code{start()} function right after the preprocessing.\\
 
-Useful helper class to handle indentation of text in a resulting XSD is named \code{Indentator}.
+Useful helper class to handle indentation of text in a resulting XSD is called \code{Indentator}.
 Instance of this class is a member variable of the module (alike instance of \code{Preprocessor}), it holds text appended to it and keeps indentation level state.
-Text can be appended without indentation (method \code{append}) or indented (method \code{indent}).
-Level of indentation can be incremented or decremented by methods \code{increaseIndentation} and \code{decreaseIndentation}.
-At the end of export, when textual representation of each element has been appended to the \code{Indentator}, \code{Indentator}'s method \code{toString} will return string representation of resulting XSD.\\
+Text can be appended without indentation (method \code{append()}) or indented (method \code{indent()}).
+Level of indentation can be incremented or decremented by methods \code{increaseIndentation()} and \code{decreaseIndentation()}.
+At the end of export, when textual representation of each element has been appended to the \code{Indentator}, \code{Indentator}'s method \code{toString()} will return string representation of the resulting XSD.\\
 
-Before describing export of elements, let's take a look on how we define elements and their attributes using XSD language.
+Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language.
 
@@
 
-	\item \code{simpleType}. Actually, exporter does not support any XSD features, which are defined in \code{simpleType}. That means, this types will not occur in result XSDs.
+	\item \code{simpleType}. Actually, exporter does not support any XSD features which are defined in \code{simpleType}. This means, that these types will not occur in result XSDs.
 	\begin{verbatim}
@@
 	\end{verbatim}
-	An empty element is defined as empty \code{complexType}
+	An empty element is defined as empty \code{complexType}.
 	\begin{verbatim}
@@
 
-Named type of an element is defined by XSD element \code{simpleType}(lack of features mentioned above) or \code{complexType} with specified attribute \code{name}.
+Named type of an element is defined by XSD element \code{simpleType} (lack of features mentioned above) or \code{complexType} with specified attribute \code{name}.
 Its content is exactly the same as described above.
-There is, of course, no need to define built-in types.
+There is of course no need to define built-in types.
 
@@
 
-XSD elements \code{xs:element}, \code{xs:sequence} and \code{xs:choice} can has attributes \code{minOccurs} and \code{maxOccurs}.
+XSD elements \code{xs:element}, \code{xs:sequence} and \code{xs:choice} can have attributes \code{minOccurs} and \code{maxOccurs}.
 These attributes defines interval of number of instances of a particular element.
@@
 
-Default values for \code{minOccurs} and \code{maxOccurs} attributes are "1", if they are not specified. So the example above has the same meaning as following.
+Default values for \code{minOccurs} and \code{maxOccurs} attributes are "1", if they are not specified. So the example above has the same meaning as the following one.
 

### 2011-01-31 doc/jInfer/twostep/twostep.tex
Documentation: TODOs in progress.

@@
 In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
-% TODO anti Sorry, I don't understand the following 3 sentences, so I cannot correct them
+% TODO vito Sorry, I don't understand the following 3 sentences, so I cannot correct them
 This is the only one simplifier shipped with jInfer, but it has many features and exchangeable submodules of which we implement alternatives in basic bundle.
@@
 Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
-% TODO anti Don't understand the following two sentences, please rephrase
-After clustering, implementation should hold clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
+% TODO vito Don't understand the following two sentences, please rephrase
+After clustering, implementation should keep the clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
 Given item to this method, one can ask for representant of cluster, to which the item belongs.
+
 If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
@@
 \subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
-We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class, which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
+We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class (Iname je clusterer!), which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
 % TODO anti I don't understand the sentence, rewrite.
@@
 To automaton, instead of instances on right side of rules, their cluster representants are inserted as one can see in section \ref{section_AutomatonMergingState_structure}.
-% TODO anti Don't understand the sentence
+% TODO vito Don't understand the sentence
+% Do automatu sypeme concat, ale nie jeho prvky, ale ich reprezentantov
 This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
@@
 It is responsible for converting symbol to string.
-To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule.
+% To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule.
 % TODO anti Unreadable....
@@
 
-Module \jmodule{AutomatonSimplifier} implements solution searching logic in simplifying automaton by merging states. % TODO anti Don't understand
 One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.
@@
 \subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
-This cleaner does exactly what we have just described. % TODO anti Doesn't make sense
 It converts nested concatenations into flat ones.
@@
   }
-} until (user selected at least one state in last trial) TODO anti Trial -> Attempt?
+} until (user selected at least one state in last attempt)
 \end{verbatim} 
@@
 \subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
-We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergeability of states by default configuration.
-% TODO anti Sentence unclear
-% TODO vektor : mention this? (we have stupid algorithm:) TODO anti Of course;)
+We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergeability of states.
 We don't use $k-grams$ algorithm from \cite{ahonen}, just a simple DFS.

### 2011-01-31 doc/jInfer/architecture/architecture.tex
Documentation: TODOs in progress.

@@
 \end{itemize}
-As practice showed, for methods handling and infering structural properties, it is important to recognize whether structural node on input is element or text node. % TODO anti Unclear, rewrite
-However methods for content devising don't need to know if they are working on infering model for content of attribute or text node. % TODO anti Unclear, rewrite
+% TODO vektor Na jednej strane nas zaujima struktura, na to tam mame isElement() isSimpleData(), na druhej strane nas zaujimaju textove data, na to tam mame getContent() v SD aj atribute.
 
 Finally, our interface/class model for representing XML nodes is drafted in fig. \ref{nodes}. Those who are brave enough can look at fig. \ref{nodes_full2}.
-In result, \code{Element} and \code{SimpleData} have method \code{getType()} to devise type of \code{Abstr\.act\.Stru\.ctu\.ral\.Node} variables. % TODO anti What is "devise"?
-And \code{SimpleData} and \code{Attribute} have methods \code{getContentType()} and \code{getContent()} to work with content model.
-% TODO anti Last 2 sentences make no sense to me...
 
@@
 
-Take care of symbol class and its \code{equals()} implementation. % TODO anti What does it mean, to take care?
 Automaton uses \code{equals()} to compare symbols on transitions (when building prefix-tree automaton and when merging states).
@@
 
-One can ask for merging states, that are merged out from automaton already. % TODO anti What does it mean?
+% TODO vito Aby bolo mozne odkazovat sa na uz zmergovane stavy, ktore sa uz volaju inak, pouzivame mapu premenovani.
 Lets take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).

### 2011-01-31 doc/jInfer/architecture/architecture.tex
Documentation: Architecture: fixing TODOs.

@@
 \end{itemize}
-% TODO vektor Na jednej strane nas zaujima struktura, na to tam mame isElement() isSimpleData(), na druhej strane nas zaujimaju textove data, na to tam mame getContent() v SD aj atribute.
+On the one hand, we are interested in the XML structure - for this reason we use \code{isElement()} and \code{isSimpleData()} of \code{AbstractStructuralNode}.
+On the other hand, we want to access textual content and its type, for this reason we use \code{getContent()} and \code{getContentType()} of \code{ContentNode}.
 
@@
 
-% TODO vito Aby bolo mozne odkazovat sa na uz zmergovane stavy, ktore sa uz volaju inak, pouzivame mapu premenovani.
+To be able to refer to states that have already been merged (and thus have a different name), we use a map of renamings.
 Lets take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
@@
 This is \code{Capabilities.CAN\_HANDLE\_COMPLEX\_REGEXPS} meaning whether the \code{Simplifier} accepts more complicated regular expressions than simple concatenations of tokens on its input.
-% TODO vektor check later - didn't this change?!?
 

### 2011-01-31 doc/jInfer/twostep/twostep.tex
Documentation: TwoStep: fixing TODOs.

@@
 In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
-% TODO vito Sorry, I don't understand the following 3 sentences, so I cannot correct them
-This is the only one simplifier shipped with jInfer, but it has many features and exchangeable submodules of which we implement alternatives in basic bundle.
-We implement is $k,h$-context method of infering grammar rules (see \cite{ahonen}).
-This process is split into submodules with generic automaton implementation, so it is easy to plug-in any other automaton merging state algorithm.
+While this is the only simplifier bundled with jInfer, its strong modularity ensures that it can be extended easily.
+For most of the points of extension we indeed provide multiple implementations.
+These vary from trivial proof-of-concept to for example $k,h$-context method for grammar rules inference (see \cite{ahonen}).
+Strong emphasis is put on generic automata, therefore it is easy to plug-in any automaton merging state algorithm.
 In this document, we will walk over all modules and submodules in simplifier and describe them briefly.
@@
 Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
-% TODO vito Don't understand the following two sentences, please rephrase
-After clustering, implementation should keep the clusters in member, as it will be later requested by calling method \code{T getRepresentantForItem(T item)}.
-Given item to this method, one can ask for representant of cluster, to which the item belongs.
+To be able to respond to calls to \code{T getRepresentantForItem(T item)} method, every clusterer implementation has to keep all its clusters even after the clustering is finished.
+Using this method, given an item the correct cluster representative for this item has to be returned.
 
@@
 By the way, this is why we cluster all simple data and all sentinels on right sides of rules.	
-To automaton, instead of instances on right side of rules, their cluster representants are inserted as one can see in section \ref{section_AutomatonMergingState_structure}.
-% TODO vito Don't understand the sentence
-% Do automatu sypeme concat, ale nie jeho prvky, ale ich reprezentantov
+While adding a concatenation to the automaton, instead of the instances themselves their \emph{representatives} are used. See \ref{section_AutomatonMergingState_structure} for further reference.
 This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.

### 2011-01-31 doc/jInfer/basicigg/basicigg.tex
Documentation: BasicIGG: fixing TODOs.

@@
 Responsible developer: & Matej Vitásek \\ \hline
-Required tokens:       & none \\ \hline
+Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
 Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\ \hline
@@
 Anyone wishing to implement his own expander will thus have to either remove jInfer's implementation, or implement the usual module selection.
-% TODO vektor Is this documented somewhere?
-anti: NO, should be in developer doc with preferences panel and so on!!!
+\\
+
+One interesting use case for expansion is simple adaptation of legacy inference methods. "Old" methods working only with simple grammar can be used unchanged with "new" inputs - schemas.
 
@@
 Now, $ m $ words will be constructed in the following way: in a for-loop enumerating $ i \in \{1 \ldots m\}$ the $j$-th part of this $i$-th word will be $w_{i mod |w_j|}$.
-The method \code{ExpanderImp.unpackConcat()} illustrates this.
+The method \code{ExpanderImpl.unpackConcat()} illustrates this.
 \end{itemize}

### 2011-02-01 doc/jInfer/basicigg/basicigg.tex
Documentation: BasicIGG

@@
 It registers itself into \emph{query} input folder, \emph{xpath} file extension (text file containing one XPath query per line) and declares it cannot handle other file extensions.\\
-To parse XPath queries, standard support present in JDK is used. The relevant \code{XPathHandler} is the \code{XPathHandler} class.\\
-% TODO vektor Explain how it works...
+To parse XPath queries, standard support present in JDK is used. The relevant \code{XPathHandler} is the \code{XPathHandlerImpl} class.\\
+While the exact process by which \code{XPathHandlerImpl} creates the IG from a series of XPath events is an interesting coding exercise, its description is beyond the scope and requirements of this document.
 

### 2011-02-01 doc/jInfer/architecture/architecture.tex
resolving todos in arch and twostep, and fresh rebuild of all pdfs

@@
 
-For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of the members set.
-One has to fill in all properties carefully and call \code{setImmutable()} aftewards.
-Proper usage should be one of following: % TODO anti WTF?
 Regexp instance is by default created as immutable, that is, once instantiated, you cannot add more children to list of children, cannot change type, content etc.
@@
 This occurs mainly in input modules: while parsing XML data sequentially, one does not know contents of element in time of handling start element event.
+
+For these cases, special \code{getMutable()} method is implemented to obtain regexp with none of the members set.
+One has to fill in all properties carefully and call \code{setImmutable()} aftewards.
+Proper usage should be one of following:
 \begin{verbatim}
@@
 
-This modular architecture means that it is possible to replace any and/or all of these modules with \emph{something} else that does similar job.
-In practice, researchers will probably implement new \jmodule{Simplifier} modules, while using our implementation of BasicIGG and export modules.
-This modular scheme is also used inside our modules to divide them into submodules.
-It is possible to add new file format processing to import/export.
-One can extend automaton merging state algorithm of our implementation of \cite{ahonen} in TwoStepSimplifier module by replacing submodules.
-% TODO anti This is a terrible paragraph with no clear path to follow, please reorganize (sentences on their own are OK)
-
 \subsection{Programmatic view}
@@
 	\end{verbatim}
-	% TODO anti Emphasise elements in the following paragraph
-	Elements note and more are empty, but we cannot use lambda regexp. Import will use empty concatenation instead.
-	For element person, import module has to generate some positive examples of language generated by regular expression \verb+(info?, more, more?, more?)+.
+	\emph{Elements} note and more are empty, but we cannot use lambda regexp. Import will use empty concatenation instead.
+	For \emph{element} person, import module has to generate some positive examples of language generated by regular expression \verb+(info?, more, more?, more?)+.
 

### 2011-02-01 doc/jInfer/twostep/twostep.tex
resolving todos in arch and twostep, and fresh rebuild of all pdfs

@@
 \begin{document}
-
-% TODO anti Representant -> Representative!
 
@@
 
-% TODO anti Rewrite so that these sentences don't end with "too" all the time
 Under certain circumstances it is useful to declare method \code{create()} generic.
-The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic too.
+The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic.
 But automaton simplification does not depend on type of symbol of automaton transitions, so interface \code{AutomatonSimplifier<T>}
@@
 Method \code{TwoStepSimplifier.simplify()} basically does the following.
-% TODO anti Rename finalGrammar to resultGrammar or something like that
 \begin{verbatim}
@@
 
-// 2. prepare emtpy final grammar
-List<Element> finalGrammar= new LinkedList<Element>();
+// 2. prepare emtpy simplified grammar
+List<Element> simplifiedGrammar= new LinkedList<Element>();
 
@@
   // 4. add to rules
-  finalGrammar.add(
+  simplifiedGrammar.add(
               new Element(node.getContext(),
@@
 
-return finalGrammar;
+return simplifiedGrammar;
 \end{verbatim}
 First, it creates a clusterer, gives it all the rules and performs the clustering.
-Then, empty list of elements is created as the final (simplified) grammar.
+Then, empty list of elements is created as the simplified grammar.
 For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
@@
 To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
-It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representant of
-type \code{T} (that is representant of some main cluster) has some "attribute" clusters associated with it.
+It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
+type \code{T} (that is representative of some main cluster) has some "attribute" clusters associated with it.
 Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.
@@
 
-% TODO anti Not clear what "worker" is
-Worker interface itself is defined as follows.
+Interface of submodule is defined as follows.
 \begin{verbatim}
@@
 Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
-Clusterer can tell you the representant for any element in the whole input (not only those elements in rules, but also those that may be on right side of rules).
+Clusterer can tell you the representative for any element in the whole input (not only those elements in rules, but also those that may be on right side of rules).
 Also, clusterer (if it is with attributes) has information about attributes in each cluster.
@@
 \begin{enumerate}
-	\item Initial Grammar on input, which is ruled into \jmodule{Clusterer}. % TODO anti WTF is "ruled into"?
+	\item Initial Grammar on input, which is sent into \jmodule{Clusterer}.
 	\item Each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
@@
 Clustering is done based on element's \code{getName()} equality (ignoring case).
-For each rule (element), it is first clustered by finding the cluster which representant has the same name - or create new cluster with this element as representant.
+For each rule (element), it is first clustered by finding the cluster which representative has the same name - or create new cluster with this element as representative.
 Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
@@
 But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
-  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representant of same \code{getName()}, or create new cluster with this sentinel as representant.
+  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representative of same \code{getName()}, or create new cluster with this sentinel as representative.
 From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future).
@@
 \subsection{PassRepresentant (ClusterProcessor)} \label{section_PassRepresentant}
-Trivial ClusterProcessor implementation. For each cluster return its representant as the rule to be in resulting schema.
+Trivial ClusterProcessor implementation. For each cluster return its representative as the rule to be in resulting schema.
 This has nothing to do with grammar simplification, it is just a proof-of-concept submodule.
@@
 While adding a concatenation to the automaton, instead of the instances themselves their \emph{representatives} are used. See \ref{section_AutomatonMergingState_structure} for further reference.
-This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representant is only present in automaton.
+This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representative is only present in automaton.
 That results in cluster-equal behaviour.
@@
 It is responsible for converting symbol to string.
-% To \jmodule{AutomatonSimplifier} it is passed from \jmodule{AutomatonMergingState} cluster processor, where we know that symbols are of type \code{AbstractStructuralNode}, so we implement \code{SymbolToString} by using \code{getName()} method on nodes and pass it down to generic \code{AutomatonSimplififier} submodule.
-% TODO anti Unreadable....
 Overloaded version is to inform the user about name of processed element when presenting automaton in \jmodule{AutoEditor}.
@@
 This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
-% TODO anti Emphasise the states in following sentences
-Before the algorithm starts, states superInitial and superFinal are added to automaton, former with $\lambda$-transition to initial state.
-From all final states a $\lambda$ transition to superFinal state is added.
-State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - superInitial and superFinal.
-After removing all states, there is only one transition from superInitial to superFinal state.
+Before the algorithm starts, states \emph{superInitial} and \emph{superFinal} are added to automaton, former with $\lambda$-transition to initial state.
+From all final states a $\lambda$ transition to \emph{superFinal} state is added.
+State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - \emph{superInitial} and \emph{superFinal}.
+After removing all states, there is only one transition from \emph{superInitial} to \emph{superFinal} state.
 That transition has the final regular expression as symbol on it. This one is returned.
@@
 They does exactly what their names stands for.
-% TODO anti Emphasise the states again
-SuperInitial state has only one $\lambda$-out-transition pointing to original initial state.
-SuperFinal state has as many $\lambda$-in-transitions, as there are original final states - from each one it gets one.
+\emph{SuperInitial} state has only one $\lambda$-out-transition pointing to original initial state.
+\emph{SuperFinal} state has as many $\lambda$-in-transitions, as there are original final states - from each one it gets one.
 Without going further into many technical details, we just mention method \code{removeState}, which removes given state from automaton.

### 2011-02-03 doc/jInfer/architecture/architecture.tex
todo vektor ruledisplayer
Documentation: Architecture: RuleDisplayer token.

@@
 All inference modules declare their respective token as provided.
-This way it is ensured that jInfer will run only if there is at least one implementation of each of the inference modules, regardless of the implementation itself.
+This way it is ensured that jInfer will run only if there is at least one implementation of each of the inference modules, regardless of the implementation itself.\\
+
+Another use of tokens is \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} representing a module providing a rule displayer.
+This token is provided by \jmodule{BasicRuleDisplayer} and \jmodule{TreeRuleDisplayer}, requested by \jmodule{BasicIGG}, \jmodule{XSDImporter} and \jmodule{TwoStepSimplifier}.
 

### 2011-02-04 doc/jInfer/basicigg/basicigg.tex
Documentation: Modifying BasicIGG in accordance with #3172486.

@@
 For each file, based on its extension the correct \emph{processor} is selected, executed and returned rules are aggregated.
-After each file has been processed, the resulting grammar is returned by invoking the \code{finished} method of the callback argument.
+Rules coming from XML documents are verified to be \emph{simple} grammar representation.
+After each file has been processed, the resulting grammar is expanded if necessary and returned by invoking the \code{finished} method of the callback argument.
 This process is illustrated in fig.
@@
 When an end of an element is encountered, the element currently on the top of the stack is closed and declared to be a rule of the IG.\\
-After reaching the end of the document, all IG rules are returned. Note that this approach creates "simple" initial grammar.
+After reaching the end of the document, all IG rules are returned. Note that this approach creates \emph{simple} initial grammar.
 
@@
 Translation from the object model in this library to our own is handled mostly by \code{DTD2RETranslator} class.\\
-Note that the initial grammar generated in this way is complex.
-Therefore, this processor checks for the \code{CAN\_\.HANDLE\_\.COMPLEX\_\.REGEXPS} capability of the simplifier following this IG generator: in positive case it simply returns the grammar, if the following module cannot handle complex regular expressions, DTD processor first invokes an \emph{Expander} (see \ref{section-expansion}).
+Note that the initial grammar generated in this way is \emph{complex}.
 
@@
 Note that even though this implementation is retrieved using lookups, jInfer bundles only one such implementation and does not support choice among more of them.
-Anyone wishing to implement his own expander will thus have to either remove jInfer's implementation, or implement the usual module selection.
-\\
+Anyone wishing to implement his own expander will thus have to either remove jInfer's implementation, or implement the usual module selection.\\
 
-One interesting use case for expansion is simple adaptation of legacy inference methods. "Old" methods working only with simple grammar can be used unchanged with "new" inputs - schemas.
+One interesting use case for expansion is simple adaptation of legacy inference methods. "Old" methods working only with \emph{simple} grammar can be used unchanged with "new" inputs - schemas.\\
+
+In \jmodule{BasicIGG}, only grammar coming from input schemas and queries is expanded, and this happens only if the following module lacks the \code{CAN\_HANDLE\_COMPLEX\_REGEXPS}.
 
@@
 	\item Each file gets processed by a processor based on its folder and extension.
-	\item Rules from all files are gathered in a single list (IG) and returned via a callback.
+	\item Grammar coming from XML documents is checked to be \emph{simple}.
+	\item If needed, grammar coming from schemas and queries is expanded.
+	\item Grammar (from now on called IG) is returned via a callback.
 \end{enumerate}

### 2011-02-09 doc/jInfer/treeruledisplayer/treeruledisplayer.tex
Added Documentation for Tree rule displayer

--- 2011-01-25 doc/jInfer/basicruledisplayer/basicruledisplayer.tex
@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
+Responsible developer: & Michal Švirec \\ \hline
 Required tokens:       & none \\ \hline
 Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
-Module dependencies:   & Base \\ \hline
+Module dependencies:   & Base \\
+	& JUNG \\ \hline
 Public packages:       & none \\ \hline
@@
 
-This is fairly basic rule displayer draws rules as series of nested coloured rectangles.
+This rule displayer uses JUNG graph implementation to draw rules as forest of trees, where each tree represents one rule of whole set.
+In each tree, root vertex represents left side of the rule, inner vertices represents concatenation, alternation or permutation and leafs are elements, attributes or simple data.
+As in basic rule displayer it creates a component with multiple tabs: each one for a new grammar to display.
 
@@
 
-The main class implementing \code{RuleDisplayer} inference interface and simultaneously registered as its service provider is \code{BasicRuleDisplayer}.
+The main class implementing \code{RuleDisplayer} inference interface and simultaneously registered as its service provider is \code{TreeRuleDisplayer}.
+Main method of this class is \code{createDisplayer()}, which looks up the component, adds a new panel to it and renders the specified grammar in it.\\
 
-%\nocite{*}
+Main window of rule displayer may have multiple tabs in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar using \jmodule{JUNG} library.\\
+
+All the trees representing rules are created in the \code{GraphBuilder} class with method \code{buildGraphPanel()}.
+While the code itself is a nice recursion programming exercise, there is nothing of a special interest in it.\\
+
+All graphics used by \jmodule{TreeRuleDisplayer} is contained in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.graphics} package.
+
+\subsection{Settings}
+
+All settings provided by \jmodule{TreeRuleDisplayer} are NetBeans-wide.
+The options panel along with all the logic is in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.options} package.
+Available options include setting the color of the background, horizontal and vertical distance between vertices.
+Also for each type of vertex is possible to set the size, the shape and the color.
+
+\nocite{*}
 \newpage

### 2011-02-13 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl}.
+The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
 Process of export consists of two phases described in detail in later sections.
@@
 Method \code{start()} first creates an instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
-Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such as which elements should be globally defined and which element is the root element.
-This instance is kept as a member variable of module to be used during the whole export process.\\
+Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such as which elements should be globally defined and which element is the root element.\\
 
-Afterwards, \code{start()} method recursively traverses global elements followed by other elements starting at the root element and for each it creates element's XSD string representation.\\
+Afterwards, \code{start()} method uses instances of classes derived from \code{AbstractElementsProcessor} class to export elements of input grammar.
 
 \subsection{Preprocessing}
+
+Code to handle preprocessing resides in package \code{cz.cuni.mff.ksi.jinfer.basicxsd.preprocessing}.
+
+\subsubsection{Purpose}
 
@@
 
+\subsubsection{TODO spravny vyraz sem}
+
 Constructor of \code{Preprocessor} class gets a list of elements and a number, defining minimal number of occurrences of an element to be defined globally.
@@
 
-Results of preprocessing are provided by public methods of \code{Preprocessor} class. For details see their JavaDoc.\\
+\subsubsection{Running preprocessor and obtaining its result}
+
+As described above, preprocessing is performed by creating an instance of \code{Preprocessor} class.\\
+
+Information discovered by the preprocessing can be obtained by calling \code{Preprocessor}'s \code{getResult()} method.
+This method returns an instance of \code{PreprocessingResult} class.
+Purpose of this class is to provide of what the preprocessor has discovered and to provide an easy way to search the input grammar for an element by its name.
+For details see JavaDoc of \code{PreprocessingResult}'s public methods.
 
@@
 
-XSD export itself is performed in module's \code{start()} function right after the preprocessing.\\
+XSD export itself is performed using classes derived from \code{AbstractElementsProcessor} class and a helper class named \code{Indentator}.
 
-Useful helper class to handle indentation of text in a resulting XSD is called \code{Indentator}.
-Instance of this class is a member variable of the module (alike instance of \code{Preprocessor}), it holds text appended to it and keeps indentation level state.
+\subsubsection{Indentation}
+
+To generate a human readable XSD output, it is necessary to apply correct indentation of XSD elements and their content.
+This is handled by \code{Indentator} class.
+This classes also serves as a buffer for string representation of XSD that the exporter is creating.\\
+
+Instance of this class is a member variable of the module, it holds text appended to it and keeps indentation level state.
 Text can be appended without indentation (method \code{append()}) or indented (method \code{indent()}).
@@
 
-Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language.
+\subsubsection{Definition of elements}
 
-\subsubsection{Definition of elements}
+Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language.\\
 
@@
 \begin{itemize}
-	\item XSD built-in type. One of types like \code{xs:string}, \code{xs:integer}, \code{xs:positiveInteger}, etc.
+	\item XSD \emph{built-in type}. One of types like \code{xs:string}, \code{xs:integer}, \code{xs:positiveInteger}, etc.
 	\begin{verbatim}
@@
 	\end{verbatim}
-	An empty element is defined as empty \code{complexType}.
+	An \emph{empty element} is defined as empty \code{complexType}.
 	\begin{verbatim}
@@
 Its content is exactly the same as described above.
-There is of course no need to define built-in types.
+There is of course no need to define \emph{built-in types}.
 
@@
 
-Exporter supports types of mixed elements. Mixed element is an element that contains other elements as well as some text.
+Exporter supports types of \emph{mixed elements}. \emph{Mixed element} is an element that contains other elements as well as some text.
 
@@
 
-Mixed element type is defined as \code{complexType} with attribute \code{mixed="true"}. Definition of the element from the last example may be as following.
+\emph{Mixed element} type is defined as \code{complexType} with attribute \code{mixed="true"}. Definition of the element from the last example may be as following.
 
@@
 
-\subsection{Definition of attributes}
+\subsubsection{Definition of attributes}
 
@@
 Attribute \code{type} is one of a built-in types. If an attribute is obligatory, this is defined by specifying \code{use="required"}.
+
+\subsubsection{Export of elements}
+
+TODO
 

### 2011-02-13 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 
-Instance of this class is a member variable of the module, it holds text appended to it and keeps indentation level state.
+Instance of this class holds text appended to it and keeps indentation level state.
 Text can be appended without indentation (method \code{append()}) or indented (method \code{indent()}).
@@
 
-TODO
+Classes to handle export of elements are in package \code{cz.cuni.mff.ksi.jinfer.basicxsd.elementsexporters}.
+Basic common logic is implemented in \code{AbstractElementsExporter} class.
+This class is abstract and is supposed to be extended by classes with a particular purpose.
+Its constructor signature is defined as following.
+
+\begin{verbatim}
+public AbstractElementsExporter(final PreprocessingResult preprocessingResult,
+                                final Indentator indentator)
+\end{verbatim}
+
+Names of parameters are self explanatory.
+Parameter \code{preprocessingResult} is a result of preprocessing.
+Parameter \code{indentator} is an instance of \code{Indentator} class to be used to buffer and indent output of exporter.
+This instance has not to be empty.
+Output of exporter is appended at the end of text held by the \code{indentator}.
+This behaviour is convenient when chaining output of several elements exporters.\\
+
+There are two classes extending \code{AbstractElementsExporter}.
+\code{GlobalElementsExporter} and \code{RootElementExporter}.
+Their constructors has the same signature as constructor of \code{AbstractElementsExporter} and both have \code{run()} method to perform their function.\\
+
+\code{GlobalElementsExporter} retrieves global elements from the result of preprocessing and creates global definition of their types.
+These definitions are appended to the \code{indentator}.
+The \code{indentator} should be set to a level of indentation at which it is desired to append the global type definitions (typically, no indentation).
+After return from a \code{run()} method call, output of exporter is appended to the \code{indentator}.
+TODO exmaple.
+TODO names and number of occurences, see properties.
+
+\code{RootElementExporter} retrieves the root element from the result of preprocessing and creates its definition.
+This definition is also appended to the \code{indentator} (supplied to constructor) at a level of indentation, the \code{indentator} is set.
+Way of handling the root element's subelements depends on if these are global elements or not.\\
+
+Global elements (retrieves from the result of preprocessing) are defined by referencing its type, which is supposed to be defined globally. TODO example.
+
+Elements which are not global are defined recursively. TODO example.
 

### 2011-02-13 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 TODO exmaple.
-TODO names and number of occurences, see properties.
+TODO names and number of occurences, see properties.\\
 
@@
 
-Global elements (retrieves from the result of preprocessing) are defined by referencing its type, which is supposed to be defined globally. TODO example.
+Global elements (retrieves from the result of preprocessing) are defined by referencing its type, which is supposed to be defined globally.
+TODO example.
+This applies also to elements which are not global and their type is one of a built-in types.
+TODO example.
 
-Elements which are not global are defined recursively. TODO example.
+Elements which are not global and are not of a build-in type are defined recursively, at the place of their occurrence. TODO example.
+
+\subsubsection{Export of attributes}
+
+TODO translate. Export elementov zahrna taktiez export ich atributov, kedze atributy sa exportuje ako element <xs:attribute> v elemente <xs:complexType>.
+
+\subsection{Preferences}
+
+All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package. It is possible to set the following. 
+\begin{itemize}
+	\item Turn off generation of global element types.
+	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
+	\item Number of spaces in output per one level of indentation.
+	\item Global type name prefix.
+It is a string which will be inserted before a name of a type, which is derived from element's name.
+Can be also an empty string.
+(Only if generation of global elements is active.)
+	\item Global type name suffix.
+It is a string which will be appended after a name of a type, which is derived from element's name.
+Can be also an empty string.
+(Only if generation of global elements is active.)
+\end{itemize}
+
+\section{Known problems and limitations}
+
+TODO namespaces
+TODO problem s vypnutym generovanim global elements
+TODO problem s rovnakymi zaciatkami sekvencii
 

### 2011-02-14 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 
-\subsubsection{TODO spravny vyraz sem}
+\subsubsection{How does it work}
 
@@
 An element is considered global if its occurrence count is greater than or equal to the number of occurrences provided on input.\\
+
+If it is desired to not generate global types, this feature can be turned off in preferences.
+However, turning it off is not recommended and in some cases it may cause invalidity of resulting XSD output.
+For more information, see \ref{section:problems}.\\
+
+Minimal number of occurrences of an element to define its type globally can be altered if preferences, as well. See \ref{subsection:preferences}.
 
@@
 At the end of export, when textual representation of each element has been appended to the \code{Indentator}, \code{Indentator}'s method \code{toString()} will return string representation of the resulting XSD.\\
+
+Number of spaces characters per one level of indentation can be altered in project properties. See \ref{subsection:preferences}.
 
@@
 The \code{indentator} should be set to a level of indentation at which it is desired to append the global type definitions (typically, no indentation).
-After return from a \code{run()} method call, output of exporter is appended to the \code{indentator}.
-TODO exmaple.
-TODO names and number of occurences, see properties.\\
+After return from a \code{run()} method call, output of exporter is appended to the \code{indentator}.\\
+
+If a global type contains another elements, these are processed as following.
+
+\begin{itemize}
+	\item If a contained element is global or of a built-in type, its type is simply referenced.
+	\item If a contained element is not global nor of a build-in type, it is fully defined in a place of its occurrence inside the global type.
+\end{itemize}
+
+Name of a global type is derived from name of a corresponding global element by prefixing and suffixing it.
+Default prefix is "T" and default suffix is empty.
+These values can be changed in project properties.
+
+Examples of definition of global types.
+
+Type generated from element named "GlobalElement", containing elements "Text", which is string, and element "AnotherGlobalElement", which is another global element.
+
+\begin{verbatim}
+<xs:complexType name="TGlobalElement">
+  <xs:sequence>
+    <xs:element name="Text" type="xs:string"/>
+    <xs:element name="AnotherGlobalElement" type="TAnotherGlobalElement"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="TAnotherGlobalElement">
+  ...
+</xs:complexType>
+\end{verbatim}
+
+Type generated from element named "X", containing element "Y", which is not global element and contains elements "Text" and "GlobalElement" from the previous example.
+Element "X" has one mandatory string attribute named "id".
+
+\begin{verbatim}
+<xs:complexType name="TX">
+  <xs:sequence>
+    <xs:element name="Y">
+      <xs:complexType>
+        <xs:element name="Text" type="xs:string"/>
+        <xs:element name="GlobalElement" type="TGlobalElement"/>
+      </xs:complexType>
+    </xs:element>
+  </xs:sequence>
+  <xs:attribute name="id" type="xs:string"/>
+</xs:complexType>
+\end{verbatim}
 
@@
 Global elements (retrieves from the result of preprocessing) are defined by referencing its type, which is supposed to be defined globally.
-TODO example.
 This applies also to elements which are not global and their type is one of a built-in types.
-TODO example.
+For example, root element "Root" contains one the following two elements.
+Global element "A" and string element "B".
 
-Elements which are not global and are not of a build-in type are defined recursively, at the place of their occurrence. TODO example.
+\begin{verbatim}
+<xs:complexType name="TA">
+  ...
+</xs:complexType>
+
+<xs:element name="Root">
+  <xs:complexType>
+    <xs:choice>
+      <xs:element name="A" type="TA"/>
+      <xs:element name="B" type="xs:string"/>
+    </xs:choice>
+  </xs:complexType>
+</xs:element>
+\end{verbatim}
+
+Elements which are not global and are not of a build-in type are defined recursively, at the place of their occurrence.
+For example, root element "Root" contains element "A", which contains two another string elements.
+
+\begin{verbatim}
+<xs:element name="Root">
+  <xs:complexType>
+    <xs:sequence>
+      <xs:element name="A">
+        <xs:complexType>
+          <xs:sequence>
+            <xs:element name="text1" type="xs:string"/>
+            <xs:element name="text2" type="xs:string"/>
+          </xs:sequence>          
+        </xs:complexType>
+      <xs:element>
+    </xs:sequence>
+  </xs:complexType>
+</xs:element>
+\end{verbatim}
 
@@
 
-TODO translate. Export elementov zahrna taktiez export ich atributov, kedze atributy sa exportuje ako element <xs:attribute> v elemente <xs:complexType>.
+Export of attributes is done along with export of elements, as attributes are exported as XSD element \code{xs:attribute} in XSD element \code{xs:complexType}.
 
-\subsection{Preferences}
+\begin{verbatim}
+<xs:complexType name="...">
+  <xs:sequence>
+    ...
+  </xs:sequence>
+  <xs:attribute name="id" type="xs:string"/>
+</xs:complexType>
+\end{verbatim}
 
-All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package. It is possible to set the following. 
+\subsection{Preferences} \label{subsection:preferences}
+
+All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package.
+As mentioned above, it is possible to set the following.
 \begin{itemize}
-	\item Turn off generation of global element types.
+	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
 	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
@@
 
-\section{Known problems and limitations}
+\section{Known problems and limitations} \label{section:problems}
+
+This section describes some limitations of XSD export and some known problems, which occur in some specific cases of input grammar.
 

### 2011-02-14 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 As mentioned above, it is possible to set the following.
+
 \begin{itemize}
@@
 
-TODO namespaces
-TODO problem s vypnutym generovanim global elements
-TODO problem s rovnakymi zaciatkami sekvencii
+\subsection{Namespaces}
 
---------------------------------------------------------------------------------------
+\subsubsection{Description}
 
-First, global elements are exported. For each element, its type is defined as a global type. TODO rio example. This is done by passing global elements to \code{processGlobalElement} method.
+This basic implementation of XSD export does not support and does not handle namespaces.
+That means that no namespace definitions are generated in a result, namespace definitions in input are processed as regular elements and attributes and namespace usages (namespace:element) are considered as a part of name of an element.
 
-After global elements, others are exported by calling of method \code{processElement}, supplied by the top level element as its argument.
-Reference to a global element type is done simply by declaring element's name and type.
+\subsubsection{Workaround}
+
+To generate a valid XSD output, it is necessary to remove all namespace definitions and usages from input.
+If a presence of namespaces in a resulting XSD is necessary, they need to be inserted there manually.
+
+\subsection{XSD invalidity if generation of global element types is disabled}
+
+\subsubsection{Error conditions}
+
+Following conditions have to be met to cause a problem.
+
+\begin{itemize}
+	\item Generation of global element types if turned off (in project properties).
+	\item Input grammar contains a concatenation or an alternation regexp with several (two or more) same elements.
+\end{itemize}
+
+\subsubsection{Description}
+
+Example of a part of generated XSD.
 
 \begin{verbatim}
-<xs:element name="Person" type="PersonType"/>
+...
+  <xs:sequence>
+    <xs:element name="A">
+      ...
+    </xs:element>
+    <xs:element name="A">
+      ...
+    </xs:element>
+  <xs:sequence>
+...
 \end{verbatim}
 
-To sa deje rekurzivnym priechodom so startom v korenovom elemente.
-Elementy, ktorych typy su definovane globalne a elementy so vstavanymi typmi, su exportovane jednoducho, uvedenim ich mena a typu.
-Ostatne elementy su definovane na mieste.
-TODO rio priklady.
-Na toto je potrebne zavolat metodu \code{processElement} s korenovym elementov, ako argumentom.
+In XSD, it is not allowed to define elements with the same name in one sequence (and other XSD constructs).
 
-Code exporting attributes is in \code{attributeToString()}.
-First thing this method does is to assess the domain of a particular atribute: this is a map indexed by attribute values containing number of occurences for each such attribute.
-Type definition of an attribute is generated in the \code{DomainUtils.getAttributeType()} method.
-Based on a user setting, this might decide to enumerate all possible values of this attribute using the \code{(a|b|c)} notation, otherwise it just returns \code{\#CDATA}.\\
-Attribute requiredness is assessed based on \code{required} metadata presence.
-If an attribute is not deemed required, it might have a default value: if a certain value is prominent in the attribute domain (based on user setting again), it is declared default.
+\subsubsection{Workaround}
 
-\subsection{Preferences}
+Do not turn the generation of global element types off.
 
-All settings provided by \jmodule{BasicDTDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicdtd.properties} package.
-As mentioned before, it is possible to set the following.
-\begin{itemize}
-	\item Maximum attribute domain size which is exported as a list of all values (\code{(a|b|c)} notation).
-	\item Minimal ratio an attribute value in the domain needs to have in order to be declared default.
-\end{itemize}
+\subsection{"Unique Particle Attribution" problem}
+
+\subsubsection{Error conditions}
+
+This problem appears if input grammar contains an alternation of two or more concatenations, which begin with at least one same element. Part of a problematic XSD is for example following.
+
+\begin{verbatim}
+...
+  <xs:choice>
+    <xs:sequence>
+      <xs:element name="A" type="TA"/>
+      <xs:element name="B" type="TB"/>
+    </xs:sequence>
+    <xs:sequence>
+      <xs:element name="A" type="TA"/>
+      <xs:element name="C" type="TC"/>
+    </xs:sequence>
+  </choice>
+...
+\end{verbatim}
+
+\subsubsection{Description}
+
+If error conditions are met, input XMLs are not valid against a generated XSD.
 
@@
 \begin{enumerate}
-	\item \code{SchemaGeneratorImpl} topologically sorts elements (rules) it got on input.
-	\item For each element, relevant portion of DTD schema is generated.
-	\item String representation of the schema is returned along with the information that file extension should be "dtd".
+	\item \code{Preprocessor} topologically sorts elements (rules) it got on input.
+	\item For each element, its occurrence count is computed.
+	\item Unused elements (occurrence count equals 0) are removed.
+	\item For each element, it is determined if it type will be defined as a global type or not.
+	\item For each global element, its type is exported to a XSD representation by \code{GlobalElementsExporter}.
+	\item Root element and recursively all remaining elements are exported to a XSD representation by \code{RootElementExporter}.
+	\item String representation of the schema is returned along with the information that file extension should be "xsd".
 \end{enumerate}

### 2011-02-14 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor skeleton created.

--- 2011-02-13 doc/jInfer/basicxsd/basicxsd.tex
@@
 
-\def\mftitle{jInfer BasicXSDExporter Module Description}
+\def\mftitle{jInfer AutoEditor Module Description}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \maketitle
-\noindent Target audience: developers willing to extend jInfer, specifically hack the XSD export.
+\noindent Target audience: developers willing to extend jInfer, specifically alter displaying of automata .
 
@@
 Responsible developer: & Mário Mikula \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\ \hline
-Module dependencies:   & Base \\ \hline
-Public packages:       & none \\ \hline
+Required tokens:       & org.openide.windows.WindowManager \\ \hline
+Provided tokens:       & none \\ \hline
+Module dependencies:   & Base \\ 
+					   & JUNG \\ \hline
+Public packages:       & cz.cuni.mff.ksi.jinfer.autoeditor \\ 
+					   & cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer \\
+   					   & cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts \\
+   					   & cz.cuni.mff.ksi.jinfer.autoeditor.gui.component \\ \hline
 \end{tabular}
@@
 
-This is an implementation of a \jmodule{SchemaGenerator} exporting the inferred schema to XSD, supporting basic features of the language.
+This is an implementation of a \jmodule{AutoEditor} TODO
 
 \section{Structure}
+
+TODO TODO TODO
 
@@
 
-\subsection{Preprocessing}
+\subsection{Preferences}
 
-Code to handle preprocessing resides in package \code{cz.cuni.mff.ksi.jinfer.basicxsd.preprocessing}.
+TODO
 
-\subsubsection{Purpose}
-
-As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its functions are following.
-\begin{itemize}
-	\item Decide which elements should be defined globally.
-	\item Remove unused elements.
-	\item Find the top level element.
-	\item Find an instance of element by its name.
-\end{itemize}
-
-\subsubsection{TODO spravny vyraz sem}
-
-Constructor of \code{Preprocessor} class gets a list of elements and a number, defining minimal number of occurrences of an element to be defined globally.
-It first topologically sorts input elements to decide which of the elements is the root element.
-Afterwards, it counts occurrences of the elements and removes unused ones (those which did not occur).
-Finally, for each element it decides whether to mark it as a global one or not.
-An element is considered global if its occurrence count is greater than or equal to the number of occurrences provided on input.\\
-
-\subsubsection{Running preprocessor and obtaining its result}
-
-As described above, preprocessing is performed by creating an instance of \code{Preprocessor} class.\\
-
-Information discovered by the preprocessing can be obtained by calling \code{Preprocessor}'s \code{getResult()} method.
-This method returns an instance of \code{PreprocessingResult} class.
-Purpose of this class is to provide of what the preprocessor has discovered and to provide an easy way to search the input grammar for an element by its name.
-For details see JavaDoc of \code{PreprocessingResult}'s public methods.
-
-\subsection{Export}
-
-XSD export itself is performed using classes derived from \code{AbstractElementsProcessor} class and a helper class named \code{Indentator}.
-
-\subsubsection{Indentation}
-
-To generate a human readable XSD output, it is necessary to apply correct indentation of XSD elements and their content.
-This is handled by \code{Indentator} class.
-This classes also serves as a buffer for string representation of XSD that the exporter is creating.\\
-
-Instance of this class holds text appended to it and keeps indentation level state.
-Text can be appended without indentation (method \code{append()}) or indented (method \code{indent()}).
-Level of indentation can be incremented or decremented by methods \code{increaseIndentation()} and \code{decreaseIndentation()}.
-At the end of export, when textual representation of each element has been appended to the \code{Indentator}, \code{Indentator}'s method \code{toString()} will return string representation of the resulting XSD.\\
-
-\subsubsection{Definition of elements}
-
-Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language.\\
-
-Element is defined by XSD element \code{element}, specifying its name and type. Let \code{xs} be XMLSchema namespace.
-
-\begin{verbatim}
-<xs:element name="Person" type="...
-\end{verbatim}
-
-Type of an element is one of following.
+All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package.
+As mentioned above, it is possible to set the following.
 
 \begin{itemize}
-	\item XSD \emph{built-in type}. One of types like \code{xs:string}, \code{xs:integer}, \code{xs:positiveInteger}, etc.
-	\begin{verbatim}
-	<xs:element name="Person" type="xs:string"/>
-	\end{verbatim}
-
-	\item \code{simpleType}. Actually, exporter does not support any XSD features which are defined in \code{simpleType}. This means, that these types will not occur in result XSDs.
-	\begin{verbatim}
-	<xs:element name="Person">
-	  <xs:simpleType>
-	    ...
-	  </xs:simpleType>
-	</xs:element>
-	\end{verbatim}
-
-	\item \code{complexType}.
-This type can contain XSD element \code{xs:sequence} or \code{xs:choice}.
-Each of these elements can contain definitions of elements, \code{xs:sequence}s and \code{xs:choice}s again.
-	\begin{verbatim}
-	<xs:element name="Person">
-	  <xs:complexType>
-	    <xs:sequence>
-	      <xs:element name="Name" type="xs:string"/>
-	      <xs:element name="Surname" type="xs:string"/>
-	      <xs:choice>
-	      ...
-	      </xs:choice>
-	    </xs:sequence>
-	  </xs:complexType>
-	</xs:element>
-	\end{verbatim}
-	An \emph{empty element} is defined as empty \code{complexType}.
-	\begin{verbatim}
-	<xs:element name="EmptyElement">
-	  <xs:complexType>
-	  </xs:complexType>
-	</xs:element>
-	\end{verbatim}
-\end{itemize}
-
-Named type of an element is defined by XSD element \code{simpleType} (lack of features mentioned above) or \code{complexType} with specified attribute \code{name}.
-Its content is exactly the same as described above.
-There is of course no need to define \emph{built-in types}.
-
-\begin{verbatim}
-<xs:complexType name="PersonType">
-...
-</xs:complexType>
-\end{verbatim}
-
-Element of this type can be then defined by specifying name of the type.
-
-\begin{verbatim}
-<xs:element name="Person" type="PersonType"/>
-\end{verbatim}
-
-XSD elements \code{xs:element}, \code{xs:sequence} and \code{xs:choice} can have attributes \code{minOccurs} and \code{maxOccurs}.
-These attributes defines interval of number of instances of a particular element.
-Legal values of these attributes are non-negative integers.
-
-\begin{verbatim}
-<xs:element name="Person" type="PersonType" minOccurs="1" maxOccurs="3"/>
-\end{verbatim}
-
-Default values for \code{minOccurs} and \code{maxOccurs} attributes are "1", if they are not specified. So the example above has the same meaning as the following one.
-
-\begin{verbatim}
-<xs:element name="Person" type="PersonType" maxOccurs="3"/>
-\end{verbatim}
-
-Exporter supports types of \emph{mixed elements}. \emph{Mixed element} is an element that contains other elements as well as some text.
-
-\begin{verbatim}
-<mixedElement>
-  some text
-  <anotherElement/>
-  another text
-</mixedElement>
-\end{verbatim}
-
-\emph{Mixed element} type is defined as \code{complexType} with attribute \code{mixed="true"}. Definition of the element from the last example may be as following.
-
-\begin{verbatim}
-<xs:element name="mixedElement">
-  <xs:complexType mixed="true">
-    <xs:sequence>
-      <xs:element name="anotherElement" type="..."/>
-    </xs:sequence>
-  </xs:complexType>
-</xs:element>
-\end{verbatim}
-
-\subsubsection{Definition of attributes}
-
-Attributes are defined by XSD element \code{xs:attribute} with attributes \code{name}, \code{type} and optional \code{use}.
-Elements \code{xs:attributes} have to be placed at the end of a \code{complexType} definition.
-
-\begin{verbatim}
-<xs:element name="Person">
-  <xs:complexType>
-  ...
-  </xs:complexType>
-  <xs:attribute name="age" type="xs:positiveInteger"/>
-  <xs:attribute name="id" type="xs:string" use="required"/>
-</xs:element>
-\end{verbatim}
-
-Attribute \code{type} is one of a built-in types. If an attribute is obligatory, this is defined by specifying \code{use="required"}.
-
-\subsubsection{Export of elements}
-
-Classes to handle export of elements are in package \code{cz.cuni.mff.ksi.jinfer.basicxsd.elementsexporters}.
-Basic common logic is implemented in \code{AbstractElementsExporter} class.
-This class is abstract and is supposed to be extended by classes with a particular purpose.
-Its constructor signature is defined as following.
-
-\begin{verbatim}
-public AbstractElementsExporter(final PreprocessingResult preprocessingResult,
-                                final Indentator indentator)
-\end{verbatim}
-
-Names of parameters are self explanatory.
-Parameter \code{preprocessingResult} is a result of preprocessing.
-Parameter \code{indentator} is an instance of \code{Indentator} class to be used to buffer and indent output of exporter.
-This instance has not to be empty.
-Output of exporter is appended at the end of text held by the \code{indentator}.
-This behaviour is convenient when chaining output of several elements exporters.\\
-
-There are two classes extending \code{AbstractElementsExporter}.
-\code{GlobalElementsExporter} and \code{RootElementExporter}.
-Their constructors has the same signature as constructor of \code{AbstractElementsExporter} and both have \code{run()} method to perform their function.\\
-
-\code{GlobalElementsExporter} retrieves global elements from the result of preprocessing and creates global definition of their types.
-These definitions are appended to the \code{indentator}.
-The \code{indentator} should be set to a level of indentation at which it is desired to append the global type definitions (typically, no indentation).
-After return from a \code{run()} method call, output of exporter is appended to the \code{indentator}.
-TODO exmaple.
-TODO names and number of occurences, see properties.\\
-
-\code{RootElementExporter} retrieves the root element from the result of preprocessing and creates its definition.
-This definition is also appended to the \code{indentator} (supplied to constructor) at a level of indentation, the \code{indentator} is set.
-Way of handling the root element's subelements depends on if these are global elements or not.\\
-
-Global elements (retrieves from the result of preprocessing) are defined by referencing its type, which is supposed to be defined globally.
-TODO example.
-This applies also to elements which are not global and their type is one of a built-in types.
-TODO example.
-
-Elements which are not global and are not of a build-in type are defined recursively, at the place of their occurrence. TODO example.
-
-\subsubsection{Export of attributes}
-
-TODO translate. Export elementov zahrna taktiez export ich atributov, kedze atributy sa exportuje ako element <xs:attribute> v elemente <xs:complexType>.
-
-\subsection{Preferences}
-
-All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package. It is possible to set the following. 
-\begin{itemize}
-	\item Turn off generation of global element types.
+	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
 	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
@@
 
-\section{Known problems and limitations}
-
-TODO namespaces
-TODO problem s vypnutym generovanim global elements
-TODO problem s rovnakymi zaciatkami sekvencii
-
---------------------------------------------------------------------------------------
-
-First, global elements are exported. For each element, its type is defined as a global type. TODO rio example. This is done by passing global elements to \code{processGlobalElement} method.
-
-After global elements, others are exported by calling of method \code{processElement}, supplied by the top level element as its argument.
-Reference to a global element type is done simply by declaring element's name and type.
-
-\begin{verbatim}
-<xs:element name="Person" type="PersonType"/>
-\end{verbatim}
-
-To sa deje rekurzivnym priechodom so startom v korenovom elemente.
-Elementy, ktorych typy su definovane globalne a elementy so vstavanymi typmi, su exportovane jednoducho, uvedenim ich mena a typu.
-Ostatne elementy su definovane na mieste.
-TODO rio priklady.
-Na toto je potrebne zavolat metodu \code{processElement} s korenovym elementov, ako argumentom.
-
-Code exporting attributes is in \code{attributeToString()}.
-First thing this method does is to assess the domain of a particular atribute: this is a map indexed by attribute values containing number of occurences for each such attribute.
-Type definition of an attribute is generated in the \code{DomainUtils.getAttributeType()} method.
-Based on a user setting, this might decide to enumerate all possible values of this attribute using the \code{(a|b|c)} notation, otherwise it just returns \code{\#CDATA}.\\
-Attribute requiredness is assessed based on \code{required} metadata presence.
-If an attribute is not deemed required, it might have a default value: if a certain value is prominent in the attribute domain (based on user setting again), it is declared default.
-
-\subsection{Preferences}
-
-All settings provided by \jmodule{BasicDTDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicdtd.properties} package.
-As mentioned before, it is possible to set the following.
-\begin{itemize}
-	\item Maximum attribute domain size which is exported as a list of all values (\code{(a|b|c)} notation).
-	\item Minimal ratio an attribute value in the domain needs to have in order to be declared default.
-\end{itemize}
-
-\section{Data flow}
-
-Flow of data in this module is following.
-\begin{enumerate}
-	\item \code{SchemaGeneratorImpl} topologically sorts elements (rules) it got on input.
-	\item For each element, relevant portion of DTD schema is generated.
-	\item String representation of the schema is returned along with the information that file extension should be "dtd".
-\end{enumerate}
-
 \nocite{*}

### 2011-02-15 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor concept.

@@
 
-This is an implementation of a \jmodule{AutoEditor} TODO
+This is an implementation of a \jmodule{AutoEditor}.
+Using JUNG library, it provides an API to display and user interactively modify automata, so the process of inference can be easily made user interactive.
 
@@
 
-TODO TODO TODO
+Structure of \jmodule{AutoEditor} can be divided into following 3 main parts.
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
-Process of export consists of two phases described in detail in later sections.
-\begin{enumerate}
-	\item Preprocessing.
-	\item The export to a string representation itself.
-\end{enumerate}
-Method \code{start()} first creates an instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
-Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such as which elements should be globally defined and which element is the root element.\\
+\begin{itemize}
+	\item API - API to display automaton in GUI.
+	\item Base classes - Classes providing basic functionality that can be extended and combined to achieve desired visualization of an automaton.
+	\item Derived classes - Classes derived from the base classes that are used in existing modules and simultaneously serve as examples.
+\end{itemize}
 
-Afterwards, \code{start()} method uses instances of classes derived from \code{AbstractElementsProcessor} class to export elements of input grammar.
+First, base classes and a creation of automaton visualization will be described.
+
+\subsection{Base classes}
+
+Main two classes representing visualization of automaton are \code{Visualizer} and \code{AbstractComponent}.
+\code{Visualizer} is a graphical representation of automaton and \code{AbstractComponent} is a panel (extends \code{JPanel}) containing the \code{Visualizer} which will be displayed in GUI.
+TODO obrazok ako AC dedi od JPanelu a obsahuje Visualizer.
+
+\subsubsection{Visualizer}
+
+TODO translate
+
+Trieda \code{Visualizer} dedi od JUNGoveho \code{VisualizationViewer}, takze poskytuje vsetky jeho metody a navyse podporu pre ulozenie automatu do obrazku - metody \code{saveImage()} a \code{getSupportedImageFormatNames()}.
+Pre ulozenie obrazku vsak tieto metody nie je nutne volat, pretoze AutoEditor GUI obsahuje tlacitko na ulozenie prave vykresleneho automatu do obrazku (viz dalej).
+
+Constructor method bere ako argument instanciu triedy \code{Layout}. Viac informacii o tejto triede a jej pouzitie v kapitole TODO ref.
+
+TODO obrazok ako Visualizer dedi od VisualizationVieweru a obsahuje Layout.
+
+\subsubsection{AbstractComponent}
+
+TODO translate
+
+Trieda \code{AbstractComponent} je panel v ktorom bude vykresleny automat, presnejsie Visualizer reprezentujuci nejaky automat.
+Dedi od triedy \code{JPanel}, takze poskytuje vsetky jej metody a spravanie.
+Navyse poskytuje metody
+
+setVisualizer()
+getVisualizer()
+waitForGuiDone()
+guiDone()
+guiInterrupt()
+guiInterrupted()
+
+a abstract metodu
+getAutomatonDrawPanel()
+
+Purpose tejto triedy je rozsirit ju a poskladat si panel aky sa hodi (tlacitka, napisy, ...) s tym, ze musi obsahovat aspon jeden JPanel, v ktorom bude vykresleny nastaveny Visualizer.
+Ucel metody getAutomatonDrawPanel() je vratit tento JPanel, aby AutoEditor vedel, kam ma ten Visualizer vykreslit.
+
+Ak je ziadany user interaktivita, je nutne si podporu pre nu zahrnut prave do tejto triedy. Pre viac informacii viz TODO ref.
+
+Visualizer sa nenastavuje v konstruktore z toho dovodu, ze casto je ziaduce, aby sa na rovnakom paneli kreslilo postupne viac roznych automatov.
+Na to nie je nutne vyrabat novu instanciu, ale staci na jednej instancii volat setVisualizer().
+
+\subsection{API}
+
+API AE je velmi jednoduche. Trieda \code{AutoEditor} poskytuje tieto 3 staticke metody.
+
+drawComponentAsync()
+drawComponentAndWaitForGUI()
+closeTab()
+
+\subsection{Derived classes}
+
+StatePickingVisualizer
+StatesPickingVisualizer
+
+\subsection{GUI}
+
+TODO
+
+tlacitka
 

### 2011-02-15 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor concept.

@@
 
-Structure of \jmodule{AutoEditor} can be divided into following 3 main parts.
+Structure of \jmodule{AutoEditor} can be divided into following four main parts.
 
@@
 	\item Derived classes - Classes derived from the base classes that are used in existing modules and simultaneously serve as examples.
+	\item Layout factory - TODO
 \end{itemize}
@@
 TODO obrazok ako Visualizer dedi od VisualizationVieweru a obsahuje Layout.
+
+\subsubsection{PluggableVisualizer}
+
+Tried \code{PluggableVisualizer} je rozsirenim triedy \code{Visualizer}, ktora poskytuje navyse jednoduchy sposob ako zapajat mouse pluginy.
+
+Mouse plugin je trieda JUNGu, ktore rozsiruje moznosti Visualizera o pracu s mysou (vyberanie stavov, zoom, ...). Spomenut uz existujuce pluginy a ref.
+
+TODO obrazok ako PluggableVisualizer dedi od Visualizeru
+
+Metody:
+
+addGraphMousePlugin()
+setVertexLabelTransformer()
+replaceVertexLabelTransformer()
+setEdgeLabelTransformer()
+replaceEdgeLabelTransformer()
+
+By default obsahuje 2 pluginy, jeden pre zoom a jeden pre posuvanie canvasu. V pripade potreby je pozne ich odstanit pomocou metod VisualizationVieweru. TODO dopisat.
 
@@
 
+Popis tried pouzitych v inych moduloch, ktore sluzia zaroven ako priklad.
+
 StatePickingVisualizer
 StatesPickingVisualizer
+
+\subsection{Layout factory}
+
+TODO
 

### 2011-02-16 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 
-Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language.\\
+Before we describe the export of elements, let's take a look on how we define elements and their attributes using XSD language.
+This subsection covers only XSD features that the exporter supports.
+For list of supported features, see TODO ref.\\
 
@@
 
-\section{Known problems and limitations} \label{section:problems}
+\section{Known issues and limitations} \label{section:problems}
 
-This section describes some limitations of XSD export and some known problems, which occur in some specific cases of input grammar.
+This section describes some limitations of XSD export and some known issues, which occur in some specific cases of input grammar.
+
+\subsection{Supported XSD features} \label{subsection:supportedXSDFeatures}
+
+This module is just basic implementation so many XSD features are not supported. List of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.
+
+\begin{itemize}
+	\item \code{xs:element} - \code{name}, \code{type}, \code{minOccurs}, \code{maxOccurs}
+	\item \code{xs:attribute} - \code{name}, \code{type}, \code{use}
+	\item \code{xs:complexType} - \code{name}, \code{mixed}
+	\item \code{xs:sequence} - \code{minOccurs}, \code{maxOccurs}
+	\item \code{xs:choice} - \code{minOccurs}, \code{maxOccurs}
+\end{itemize}
 

### 2011-02-16 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 	\item Derived classes - Classes derived from the base classes that are used in existing modules and simultaneously serve as examples.
-	\item Layout factory - TODO
+	\item Layout creation - System of creating \code{Layout}s.
 \end{itemize}
 
-First, base classes and a creation of automaton visualization will be described.
+First, \code{Layout}s and use of base classes to create a visualization of automaton will be described.
+
+\subsection{Layout}
+
+\code{Layout} is a JUNG interface responsible primarily of representation of automaton and positions of its states.
+JUNG library provides several implementation of \code{Layout} interface.
+However, none of them is convenient for automatic automaton displaying, \jmodule{AutoEditor} provides two additional implementations.
+\code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.
+TODO odkazy v predchadzajucej vete.
+
+Class providing creation of \code{Layout} instances is named \code{LayoutHelperFactory}.
+
+\subsubsection{Vyhnanovska Layout}
+
+As mentioned above, this Layout was implemented by \emph{Julie Vyhnanovska} as a part of her master thesis.
+TODO link?
+It positions automaton states to a square grid.
+This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confused.
+
+Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.vyhnanovska}.
+
+\subsubsection{Graphviz Layout}
+
+\code{Graphviz Layout} uses \emph{Graphviz}, third-party graph visualization software, to create positions of automaton states.
+TODO link?
+To use this \code{Layout}, \emph{Graphviz} has to be installed and path to \emph{dot} binary has to be set in options.
+This \code{Layout} gives nice results even on large automata.
+
+Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.
+
+\subsubsection{LayoutHelperFactory}
 
@@
 
-TODO translate
+\code {Visualizer} class extends JUNG \code{VisualizationViewer} class, so it provides all its methods and adds support for saving contained automaton to an image file.
+Responsible methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
+However, to save an image of automaton it is not necessary to call this methods directly.
+\jmodule{AutoEditor} GUI contains button to save a save an image of displayed automaton.
+For information on how to to this, see TODO ref.
 
-Trieda \code{Visualizer} dedi od JUNGoveho \code{VisualizationViewer}, takze poskytuje vsetky jeho metody a navyse podporu pre ulozenie automatu do obrazku - metody \code{saveImage()} a \code{getSupportedImageFormatNames()}.
-Pre ulozenie obrazku vsak tieto metody nie je nutne volat, pretoze AutoEditor GUI obsahuje tlacitko na ulozenie prave vykresleneho automatu do obrazku (viz dalej).
-
-Constructor method bere ako argument instanciu triedy \code{Layout}. Viac informacii o tejto triede a jej pouzitie v kapitole TODO ref.
+Constructor has one argument, instance of \code{Layout} class. \code{Layout} class is described in TODO ref.
 

### 2011-02-16 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 \code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.
-TODO odkazy v predchadzajucej vete.
+TODO odkazy v predchadzajucej vete.\\
 
@@
 
-\subsubsection{Vyhnanovska Layout}
+\subsubsection{Vyhnanovska Layout} \label{subsubsection:vyhnanovskaLayout}
 
@@
 It positions automaton states to a square grid.
-This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confused.
+This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confused.\\
 
@@
 
-\subsubsection{Graphviz Layout}
+\subsubsection{Graphviz Layout} \label{subsubsection:graphvizLayout}
 
@@
 To use this \code{Layout}, \emph{Graphviz} has to be installed and path to \emph{dot} binary has to be set in options.
-This \code{Layout} gives nice results even on large automata.
+This \code{Layout} gives nice results even on large automata.\\
+
+TODO ako presne ziskava pozicie z graphvizu
 
@@
 
-\subsubsection{LayoutHelperFactory}
+\subsubsection{LayoutHelperFactory} \label{subsubsection:LayoutHelperFactory}
+
+In project properties, it is possible to select a \code{Layout} to be used to display automata.
+\code{LayoutHelperFactory} is class, providing just one static method, responsible for creating instances of \code{Layout}s according to a selection in project properties.\\
+
+This method has following signature.
+
+\begin{verbatim}
+public static <T> Layout<State<T>, Step<T>> createUserLayout(final Automaton<T> automaton, final Transformer<Step<T>, String> edgeLabelTransformer) throws InterruptedException;
+\end{verbatim}
+
+The first argument is a automaton to create a layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.
+
+Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}.
+
+\subsubsection{How to create a new Layout}
+
+TODO dopisat, translate, odkaz niekam na popis modulov
+
+\code{Layout}s su implementovane pomocou modularneho systemu. Pre vyrobu noveho je treba vyrobit triedu implementujucu LayoutFactory a spravne ju oanotovat.
 
@@
 However, to save an image of automaton it is not necessary to call this methods directly.
-\jmodule{AutoEditor} GUI contains button to save a save an image of displayed automaton.
-For information on how to to this, see TODO ref.
+\jmodule{AutoEditor} GUI contains button to save an image of displayed automaton.
+For information on how to to this, see TODO ref.\\
 
-Constructor has one argument, instance of \code{Layout} class. \code{Layout} class is described in TODO ref.
+Constructor has one argument, instance of \code{Layout} class created from an automaton, typically by \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
 

### 2011-02-16 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 
-TODO dopisat, translate, odkaz niekam na popis modulov
+\code{Layout}s can be implemented using the modular system.
+To create a new implementation of \code{Layout} interface, it is needed to create a new class implementing \code{LayoutFactory} interface (package \code{package cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.
 
-\code{Layout}s su implementovane pomocou modularneho systemu. Pre vyrobu noveho je treba vyrobit triedu implementujucu LayoutFactory a spravne ju oanotovat.
+\begin{verbatim}
+@ServiceProvider(service = LayoutFactory.class)
+\end{verbatim}
+
+Created implementation will be shown in project properties in the \code{Layout} selection.\\
+
+System of modules is described in detail in TODO ref.
 
 \subsection{Base classes}
+
+This section describes classes implementing basic common functionality that are supposed to be extended to create a new suitable visualization of automata for a particular method of inference.
+The new visualization may involve a brand new GUI panel with buttons with various functions, user interaction like selecting states or edges and others.\\
 
@@
 
-Constructor has one argument, instance of \code{Layout} class created from an automaton, typically by \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
+Constructor has one argument, instance of \code{Layout} interface created from an automaton, typically by \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
 
@@
 
-Tried \code{PluggableVisualizer} je rozsirenim triedy \code{Visualizer}, ktora poskytuje navyse jednoduchy sposob ako zapajat mouse pluginy.
+\code{PluggableVisualizer} class is extension of \code{Visualizer} class, which primarily provides an easy way to plug \emph{graph mouse plugins}.\\
 
-Mouse plugin je trieda JUNGu, ktore rozsiruje moznosti Visualizera o pracu s mysou (vyberanie stavov, zoom, ...). Spomenut uz existujuce pluginy a ref.
+\emph{Graph mouse plugin}s are classes implementing JUNG \code{GraphMousePlugin} interface and their purpose is to enhance \code{Visualizer} with mouse support.\\
+
+By default, instance of \code{PluggableVisualizer} is constructed with two plugins enabled.
+They are \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
+In the most cases, these plugins are useful but if they are not wanted they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
 
@@
 
-Metody:
+Public (not inherited) methods of \code{PluggableVisualizer}.
 
-addGraphMousePlugin()
-setVertexLabelTransformer()
-replaceVertexLabelTransformer()
-setEdgeLabelTransformer()
-replaceEdgeLabelTransformer()
+TODO
+\begin{itemize}
+	\item addGraphMousePlugin()
+	\item removeGraphMousePlugin()
+	\item getGraphMousePlugins()
+	\item setVertexLabelTransformer()
+	\item replaceVertexLabelTransformer()
+	\item setEdgeLabelTransformer()
+	\item replaceEdgeLabelTransformer()
+\end{itemize}
 

### 2011-02-16 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 This is an implementation of a \jmodule{AutoEditor}.
-Using JUNG library, it provides an API to display and user interactively modify automata, so the process of inference can be easily made user interactive.
+Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.
 
@@
 
-\code {Visualizer} class extends JUNG \code{VisualizationViewer} class, so it provides all its methods and adds support for saving contained automaton to an image file.
+\code {Visualizer} class extends JUNG \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
 Responsible methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
@@
 
-Public (not inherited) methods of \code{PluggableVisualizer}.
+Public (not inherited) methods of \code{PluggableVisualizer} are the following. Their purpose is clear from their names, for details see their JavaDoc.
 
-TODO
 \begin{itemize}
-	\item addGraphMousePlugin()
-	\item removeGraphMousePlugin()
-	\item getGraphMousePlugins()
-	\item setVertexLabelTransformer()
-	\item replaceVertexLabelTransformer()
-	\item setEdgeLabelTransformer()
-	\item replaceEdgeLabelTransformer()
+	\item \code{addGraphMousePlugin()}
+	\item \code{removeGraphMousePlugin()}
+	\item \code{getGraphMousePlugins()}
+	\item \code{setVertexLabelTransformer()}
+	\item \code{replaceVertexLabelTransformer()}
+	\item \code{setEdgeLabelTransformer()}
+	\item \code{replaceEdgeLabelTransformer()}
 \end{itemize}
-
-By default obsahuje 2 pluginy, jeden pre zoom a jeden pre posuvanie canvasu. V pripade potreby je pozne ich odstanit pomocou metod VisualizationVieweru. TODO dopisat.
 
@@
 
-TODO translate
+\code{AbstractComponent} class is a representation of GUI panel containing an instance of \code{Visualizer} class for some automaton.
+It is inherited from \code{JPanel} class thus provides JPanel's method and behaviour.
+In addition, it provides the following methods.
 
-Trieda \code{AbstractComponent} je panel v ktorom bude vykresleny automat, presnejsie Visualizer reprezentujuci nejaky automat.
-Dedi od triedy \code{JPanel}, takze poskytuje vsetky jej metody a spravanie.
-Navyse poskytuje metody
+\begin{itemize}
+	\item \code{setVisualizer()} - Setter of \code{Visualizer}.
+	\item \code{getVisualizer()} - Getter of \code{Visualizer}.
+	\item \code{waitForGuiDone()} - Suspends its thread until method \code{guiDone} is called on this instance.
+Do not call this method directly, it is called by \jmodule{AutoEditor}.
+For more information, see \ref{subsubsection:AbstractComponent user-interactivity support}.
+	\item \code{guiDone()} - Wakes up this instance from a suspended state. For detailed description, see \ref{subsubsection:AbstractComponent user-interactivity support}.
+	\item \code{guiInterrupt()} - Called when \jmodule{AutoEditor}'s tab is closed to propagate information about terminating of inference to a caller of \jmodule{AutoEditor}.
+There is no need to called this method directly.
+	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by interrupt or regularly.
+Also, there is no need to call this method directly, it is called by \jmodule{AutoEditor}.
+For details, see TODO ref GUI.
+\end{itemize}
 
-setVisualizer()
-getVisualizer()
-waitForGuiDone()
-guiDone()
-guiInterrupt()
-guiInterrupted()
+Besides those methods, \code{AbstractComponent} has one abstract method, named \code{getAutomatonDrawPanel()}.\\
 
-a abstract metodu
-getAutomatonDrawPanel()
+Purpose of this class is to be extended to create own GUI panel, which displays some automaton using a supplied instance of \code{Visualizer}.
+Method \code{getAutomatonDrawPanel()} is meant to be overridden to returns an instance of \code{JPanel}, in which the \code{Visualizer} is to be drawn.\\
 
-Purpose tejto triedy je rozsirit ju a poskladat si panel aky sa hodi (tlacitka, napisy, ...) s tym, ze musi obsahovat aspon jeden JPanel, v ktorom bude vykresleny nastaveny Visualizer.
-Ucel metody getAutomatonDrawPanel() je vratit tento JPanel, aby AutoEditor vedel, kam ma ten Visualizer vykreslit.
+Programmer implementing an extension of \code{AbstractComponent} is not forced to place the \code{Visualizer} on its own.
+It is just needed to create \code{JPanel} and define \code{getAutomatonDrawPanel()} method to return this \code{JPanel}.
+\jmodule{AutoEditor} will take care of placing and displaying the \code{Visualizer} in the \code{JPanel}.\\
+
+\code{Visualizer} is not set in constructor, because it is often desired to subsequently display several different automata (\code{Visualizer}s) in the same panel.
+In this case it is not needed to create new instance of \code{AbstractComponent} for each \code{Visualizer}, but subsequently call \code{setVisualizer()} method using one instance of \code{AbstractComponent}.
+
+\subsubsection{AbstractComponent user-interactivity support} \label{subsubsection:AbstractComponent user-interactivity support}
 
@@
 
-Visualizer sa nenastavuje v konstruktore z toho dovodu, ze casto je ziaduce, aby sa na rovnakom paneli kreslilo postupne viac roznych automatov.
-Na to nie je nutne vyrabat novu instanciu, ale staci na jednej instancii volat setVisualizer().
+metody
+waitForGuiDone
+guiDone
 
@@
 StatesPickingVisualizer
-
-\subsection{Layout factory}
-
-TODO
 

### 2011-02-16 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.
+
+TODO spomenut vsade packages
 
@@
 
-\subsubsection{AbstractComponent user-interactivity support} \label{subsubsection:AbstractComponent user-interactivity support}
+\subsubsection{AbstractComponent user interactivity support} \label{subsubsection:AbstractComponent user-interactivity support}
 
-Ak je ziadany user interaktivita, je nutne si podporu pre nu zahrnut prave do tejto triedy. Pre viac informacii viz TODO ref.
+If some kind of user interactivity is desired, \code{AbstractComponent} is a right place to implement it.\\
 
-metody
-waitForGuiDone
-guiDone
+To display the component in GUI and wait for some user action, method \code{waitForGuiDone()} is used.
+After displaying the component, calling of this method suspend running thread thus code execution of a caller module is stopped at the place of this call.
+However, GUI is ran in another thread, user is able to interact with the panel (component).\\*
+Do not call method \code{waitForGuiDone()} directly.
+It is called by \jmodule{AutoEditor} when displaying the component by \jmodule{AutoEditor} \emph{API} method named \code{drawComponentAndWaitForGUI()}.
+For more information on \jmodule{AutoEditor} \emph{API}, see TODO ref.\\
+
+Method important for a programmer extending \code{AbstractComponent} is named \code{guiDone()}.
+This method wakes up the thread suspended in \code{waitForGuiDone()} method and the programmer is responsible for calling it.
+Typically, it is called upon some user action like button click, vertex pick or other GUI event.\\*
+After calling of \code{guiDone()} method, code execution of the caller module is resumed and holding instance of the \code{AbstractComponent} it is able to retrieve results of user interaction, saved in its state.\\
+
+For examples of user-interactive component, see TODO ref.
 
@@
 
-API AE je velmi jednoduche. Trieda \code{AutoEditor} poskytuje tieto 3 staticke metody.
+\jmodule{AutoEditor} \emph{API} is pretty simple. Package \code{cz.cuni.mff.ksi.jinfer.autoeditor} contains class \code{AutoEditor} with three public static methods.
 
-drawComponentAsync()
-drawComponentAndWaitForGUI()
-closeTab()
+\begin{itemize}
+	\item \code{drawComponentAsync()} - Displays given \code{AbstractComponent} asynchronously in a GUI thread and immediately returns.
+Use this method to just display automaton, without any user interaction and without waiting for any external event.
+This method does not support these.
+	\item \code{drawComponentAndWaitForGUI()} - Displays given \code{AbstractComponent} in a GUI thread and a caller thread is suspended until  \code{guiDone()} method of \code{AbstractComponent()} is called.
+This method can wait for GUI events thus is convenient for user interaction.
+How to achieve it is described in detail in \ref{subsubsection:AbstractComponent user-interactivity support}.
+	\item \code{closeTab()} - Closes \jmodule{AutoEditor}'s GUI tab and interrupts inference, if running.
+\end{itemize}
+
+For examples of API usage, see TODO ref.
 

### 2011-02-17 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 
+For the whole document, let \code{xs} be XMLSchema namespace.
+
 \section{Structure}
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provider is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
+The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provid\-er is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
 Process of export consists of two phases described in detail in later sections.
@@
 \end{enumerate}
-Method \code{start()} first creates an instance of \code{Preprocessor} class supplied by rules (elements) it got in the simplified grammar on input.
-Phase of preprocessing is done by creating that instance (calling its constructor) and its purpose is to discover information such as which elements should be globally defined and which element is the root element.\\
+Method \code{start()} first creates an instance of \code{Preprocessor} class supplying the rules (elements) it got in the simplified grammar on input.
+Preprocessing itself is done by creating that instance (calling its constructor) and its results can be then retrieved by calling \code{getResult()} method.
+Purpose of preprocessing is to discover information necessary for export of elements.
+For example, determine elements which should be globally defined or which element is the root element.\\
 
@@
 
-As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its functions are following.
+As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its is the following following.
 \begin{itemize}
@@
 
-Minimal number of occurrences of an element to define its type globally can be altered if preferences, as well. See \ref{subsection:preferences}.
+Minimal number of occurrences of an element to define its type globally can be altered if preferences as well. See \ref{subsection:preferences}.
 
@@
 This subsection covers only XSD features that the exporter supports.
-For list of supported features, see TODO ref.\\
+For list of supported features, see \ref{subsection:supportedXSDFeatures}\\
 
-Element is defined by XSD element \code{element}, specifying its name and type. Let \code{xs} be XMLSchema namespace.
+Element is defined by XSD element \code{element}, specifying its name and type.
 
@@
 	\end{verbatim}
-	An \emph{empty element} is defined as empty \code{complexType}.
+	An \emph{empty element} is defined as an empty \code{complexType}.
 	\begin{verbatim}
@@
 
-Attribute \code{type} is one of a built-in types. If an attribute is obligatory, this is defined by specifying \code{use="required"}.
+Attribute \code{type} is one of a built-in types.
+If an attribute is obligatory, this is defined by specifying \code{use="required"}.
+Whether an element is obligatory can determined from its metadata by the following code, assuming \code{attribute} is an instance of \code{Attribute} class.
+
+\begin{verbatim}
+if (attribute.getMetadata().containsKey(IGGUtils.REQUIRED)) { ...
+\end{verbatim}
 
@@
 \begin{verbatim}
-public AbstractElementsExporter(final PreprocessingResult preprocessingResult,
-                                final Indentator indentator)
+public AbstractElementsExporter(PreprocessingResult preprocessingResult,
+                                Indentator indentator)
 \end{verbatim}
@@
 Parameter \code{indentator} is an instance of \code{Indentator} class to be used to buffer and indent output of exporter.
-This instance has not to be empty.
+This instance doesn't need to be empty .
 Output of exporter is appended at the end of text held by the \code{indentator}.
@@
 
+\paragraph{GlobalElementsExporter}
 \code{GlobalElementsExporter} retrieves global elements from the result of preprocessing and creates global definition of their types.
@@
 
-If a global type contains another elements, these are processed as following.
+If a global type contains other elements, these are processed as following.
 
@@
 	\item If a contained element is global or of a built-in type, its type is simply referenced.
-	\item If a contained element is not global nor of a build-in type, it is fully defined in a place of its occurrence inside the global type.
+	\item If a contained element is not global nor of a built-in type, it is fully defined in a place of its occurrence inside the global type.
 \end{itemize}
@@
 
-Examples of definition of global types.
-
+\paragraph{Examples of definition of global types}
 Type generated from element named "GlobalElement", containing elements "Text", which is string, and element "AnotherGlobalElement", which is another global element.
@@
 
+\paragraph{RootElementExporter}
 \code{RootElementExporter} retrieves the root element from the result of preprocessing and creates its definition.
@@
 
-Elements which are not global and are not of a build-in type are defined recursively, at the place of their occurrence.
-For example, root element "Root" contains element "A", which contains two another string elements.
+Elements which are not global and are not of a built-in type are defined recursively, at the place of their occurrence.
+For example, root element "Root" contains element "A", which contains two other string elements.
 
@@
 
-Export of attributes is done along with export of elements, as attributes are exported as XSD element \code{xs:attribute} in XSD element \code{xs:complexType}.
+Export of attributes is done alongside with export of elements, as attributes are exported as XSD element \code{xs:attribute} in XSD element \code{xs:complexType}.
 
@@
 \begin{itemize}
-	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
-	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
+	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of the resulting XSD. See \ref{section:problems}.
+	\item Minimal number of occurrences of element to define its type globally. (Applies only if generation of global elements is active.)
 	\item Number of spaces in output per one level of indentation.
@@
 Can be also an empty string.
-(Only if generation of global elements is active.)
+(Applies only if generation of global elements is active.)
 	\item Global type name suffix.
@@
 Can be also an empty string.
-(Only if generation of global elements is active.)
+(Applies only if generation of global elements is active.)
 \end{itemize}
@@
 
-This module is just basic implementation so many XSD features are not supported. List of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.
+This module is just basic implementation, thus many XSD features are not supported. List of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.
 
@@
 
-This problem appears if input grammar contains an alternation of two or more concatenations, which begin with at least one same element. Part of a problematic XSD is for example following.
+This problem appears if input grammar contains an alternation of two or more concatenations, which share the same prefix (begin with at least one same element). An example of offending XSD follows.
 
@@
       <xs:element name="B" type="TB"/>
+      <xs:element name="C" type="TC"/>
     </xs:sequence>
@@
       <xs:element name="A" type="TA"/>
-      <xs:element name="C" type="TC"/>
+      <xs:element name="B" type="TB"/>
+      <xs:element name="D" type="TD"/>
+      <xs:element name="E" type="TE"/>
     </xs:sequence>
@@
 
+\subsubsection{Workaround}
+
+To make the result valid, manual modification is needed.
+The same prefix elements have to be removed from both sequences and inserted into a new \code{<xs:sequence>} element before the \code{<xs:choice>} element.\\
+
+Adjusted XSD from the example can look like the following.
+\begin{verbatim}
+...
+  <xs:sequence>
+    <xs:element name="A" type="TA"/>
+    <xs:element name="B" type="TB"/>
+  <xs:sequence>
+  <xs:choice>
+    <xs:element name="C" type="TC"/>
+    <xs:sequence>
+      <xs:element name="D" type="TD"/>
+      <xs:element name="E" type="TE"/>
+    </xs:sequence>
+  </choice>
+...
+\end{verbatim}
+
 \section{Data flow}
 
-Flow of data in this module is following.
+Flow of data in this module is the following.
 \begin{enumerate}

### 2011-02-18 doc/jInfer/basicxsd/basicxsd.tex
Documentation: BasicXSDExporter.

@@
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provid\-er is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
+The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provi\-der is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
 Process of export consists of two phases described in detail in later sections.
@@
 
-As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its is the following following.
+As mentioned above, preprocessing is implemented in \code{Preprocessor} class and its purpose is the following.
 \begin{itemize}
@@
 
-Number of spaces characters per one level of indentation can be altered in project properties. See \ref{subsection:preferences}.
+Number of spaces characters per one level of indentation can be altered in project properties, see \ref{subsection:preferences}.
 

### 2011-02-19 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 TODO spomenut vsade packages
+
+TODO UML diagramy su zjednodusene z dovodu prehladnosti
 
@@
 
-TODO obrazok ako Visualizer dedi od VisualizationVieweru a obsahuje Layout.
-
 \subsubsection{PluggableVisualizer}
@@
 	\item \code{getGraphMousePlugins()}
-	\item \code{setVertexLabelTransformer()}
 	\item \code{replaceVertexLabelTransformer()}
-	\item \code{setEdgeLabelTransformer()}
 	\item \code{replaceEdgeLabelTransformer()}
@@
 In this case it is not needed to create new instance of \code{AbstractComponent} for each \code{Visualizer}, but subsequently call \code{setVisualizer()} method using one instance of \code{AbstractComponent}.
+
+\begin{figure}
+\centering\includegraphics[scale=\myscale]{uml_basic_full}
+\caption{Class diagram for the base classes.} \label{uml_basic_full}
+\end{figure}
 
@@
 
-Popis tried pouzitych v inych moduloch, ktore sluzia zaroven ako priklad.
+\begin{figure}
+\centering\includegraphics[scale=\myscale]{uml_derived_classes}
+\caption{Class diagram for the derived classes.} \label{uml_derived_classes}
+\end{figure}
+
+This section describes classes derived from the base classes used in \jmodule{Two Step Simplifier} module to display automata. This classes may also serve as examples of extending the base classes.
+
+\subsubsection{StatePickingComponent}
+\code{StatePickingComponent} (extensions of \code{abstractComponent}) alongside with \code{StatePickingVisualizer} (extension of \code{PluggableVisualizer}) provides possibility to pick one automaton state in GUI and immediately return to the calling code, which can retrieve the picked state.
+
+\code{StatePickingVisualizer} is trivial extension of \code{PluggableVisualizer}.
+It has no additional methods.
+Its constructor has additional two arguments, superinitial and superfinal states as we want to distinguish these states in displayed automaton.\
+And upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \code{PluggableVisualizer}.
+Purpose of this plugin is to allow user to pick some state of automaton and then call \code{guiDone()} method on instance of \code{AbstractComponent}.
+For description of \code{guiDone()} method, see TODO ref.
+
+\code{StatePickingComponent} provides the following additional methods.
+\begin{itemize}
+	\item \code{getPickedState()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see TODO ref), this method retrieves the user picked automaton state.
+	\item \code{setLabel()} - Sets text of a component label. The label can be used to communicate some information to user, for example instructions.
+\end{itemize}
+
+Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \code{VertexPickingGraphMousePlugin}.
+
+Example of usage follows.
+\begin{verbatim}
+final Transformer<Step<Regexp<T>>, String> transformer = new Transformer<Step<Regexp<T>>, String>() {
+
+      @Override
+      public String transform(final Step<Regexp<T>> step) {
+        StringBuilder sb = new StringBuilder();
+        sb.append("{");
+        sb.append(symbolToString.toString(step.getAcceptSymbol()));
+        sb.append("|");
+        sb.append(String.valueOf(step.getUseCount()));
+        sb.append("}");
+        return sb.toString();
+      }
+    };
+
+    State<Regexp<T>> removeState;
+    final StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
+    final StatePickingVisualizer<Regexp<T>> visualizer = new StatePickingVisualizer<Regexp<T>>(LayoutHelperFactory.createUserLayout(automaton, transformer), transformer, component, automaton.getSuperInitialState(), automaton.getSuperFinalState());
+    component.setVisualizer(visualizer);
+
+    do {
+      AutoEditor.drawComponentAndWaitForGUI(component);
+      removeState = component.getPickedState();
+
+      if ((removeState.equals(automaton.getSuperFinalState())) || (removeState.equals(automaton.getSuperInitialState()))) {
+        component.setLabel("Do not select superInitial and superFinal states.");
+        continue;
+      }
+      LOG.debug("AUTO EDITOR selected: " + removeState.toString());
+      return removeState;
+    } while (true);
+\end{verbatim}
 

### 2011-02-19 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 \subsubsection{StatePickingComponent}
-\code{StatePickingComponent} (extensions of \code{abstractComponent}) alongside with \code{StatePickingVisualizer} (extension of \code{PluggableVisualizer}) provides possibility to pick one automaton state in GUI and immediately return to the calling code, which can retrieve the picked state.
+\code{StatePickingComponent} (extensions of \code{abstractComponent}) alongside with \code{StatePickingVisualizer} (extension of \code{PluggableVisualizer}) provides possibility to pick one automaton state in GUI and immediately return to the calling code, which can retrieve the picked state.\\
 
@@
 Purpose of this plugin is to allow user to pick some state of automaton and then call \code{guiDone()} method on instance of \code{AbstractComponent}.
-For description of \code{guiDone()} method, see TODO ref.
+For description of \code{guiDone()} method, see TODO ref.\\
 
@@
 
-Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \code{VertexPickingGraphMousePlugin}.
+Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \code{VertexPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
 
@@
       }
-      LOG.debug("AUTO EDITOR selected: " + removeState.toString());
       return removeState;
@@
 
-StatePickingVisualizer
-StatesPickingVisualizer
+\subsubsection{StatesPickingComponent}
+Classes \code{StatePickingComponent} and \code{StatesPickingVisualizer} are similar to the classes described in the previous section.
+Purpose of these is to provide picking of multiple automaton states.\\
+
+\code{StatesPickingVisualizer} is extension of \code{PluggableVisualizer} and upon its construction it adds \code{VerticesPickingGraphMousePlugin} to the graph mouse plugins.
+With this plugin used, user can pick and unpick automaton states separately or pick several states at once by dragging rectangular selection box over states.\
+Main difference compared to the \code{VertexPickingGraphMousePlugin} is \code{VerticesPickingGraphMousePlugin} does not call component's \code{guiDone()} method.\\
+
+\code{StatesPickingComponent} is almost the same as \code{StatePickingComponent} but it contains two extra GUI controls.
+'Continue' button and 'Don't ask me anymore to select states to be merger' checkbox.\
+The button is supposed to be pushed when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see TODO ref).\
+State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.
+
+Methods of \code{StatesPickingComponent} are the following.
+\begin{itemize}
+	\item \code{shallAskUser()} - Retrieves state of the checkbox.
+	\item \code{getPickedStates()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see TODO ref), this method retrieves the list of user picked automaton states.
+In case that none of automaton states was picked, it returns an empty list.
+	\item \code{setLabel()} - Sets text of a component label. The label can be used to communicate some information to user, for example instructions.
+\end{itemize}
+
+Source codes of \code{StatesPickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \code{VerticesPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
+
+Example of usage follows.
+\begin{verbatim}
+  Automaton<T> simplify(Automaton<T> inputAutomaton, final SymbolToString<T> symbolToString, final String elementName) {
+    if (!askUser) {
+      return inputAutomaton;
+    }
+
+    List<State<T>> mergeLst;
+    final Transformer<Step<T>, String> transformer = new Transformer<Step<T>, String>() {
+
+      @Override
+      public String transform(Step<T> step) {
+        StringBuilder sb = new StringBuilder();
+        sb.append("{");
+        sb.append(symbolToString.toString(step.getAcceptSymbol()));
+        sb.append("|");
+        sb.append(String.valueOf(step.getUseCount()));
+        sb.append("}");
+        return sb.toString();
+      }
+    };
+
+    Boolean selectTwo = false;
+    do {
+      final StatesPickingVisualizer<T> visualizer = new StatesPickingVisualizer<T>(LayoutHelperFactory.createUserLayout(inputAutomaton, transformer), transformer);
+
+      final StatesPickingComponent<T> panel = new StatesPickingComponent<T>();
+      panel.setVisualizer(visualizer);
+      if (selectTwo) {
+        panel.setLabel("Automaton for element <" + elementName + ">. Please select states to be merged and click continue. Select at least 2 states.");
+      } else {
+        panel.setLabel("Automaton for element <" + elementName + ">. Please select states to be merged and click continue.");
+      }
+
+      AutoEditor.drawComponentAndWaitForGUI(panel);
+      mergeLst = panel.getPickedStates();
+
+      if ((!BaseUtils.isEmpty(mergeLst)) && (mergeLst.size() >= 2)) {
+        inputAutomaton.mergeStates(mergeLst);
+        selectTwo = false;
+      } else if (mergeLst.size() < 2) {
+        selectTwo = true;
+      }
+
+      if (!panel.shallAskUser()) {
+        askUser = false;
+        break;
+      }
+    } while (!BaseUtils.isEmpty(mergeLst));
+    return inputAutomaton;
+  }
+\end{verbatim}
 

### 2011-02-21 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 \begin{figure}
-\centering\includegraphics[scale=\myscale]{uml_basic_full}
-\caption{Class diagram for the base classes.} \label{uml_basic_full}
+\centering\includegraphics[scale=\myscale]{uml_base_classes}
+\caption{Class diagram for the base classes.} \label{uml_base_classes}
 \end{figure}
@@
 
-TODO
+\begin{figure}
+\centering\includegraphics[scale=0.6]{screenshot}
+\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
+\end{figure}
 
-tlacitka
+As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the main NetBeans window.
+It consists of two buttons, horizontal line below them and below the line, there is panel to place an extension of \code{AbstractComponent} (see TODO ref).\\
 
-\subsection{Preferences}
+Buttons have labels 'Save as image' and 'Show settings'.
+Pushing the first one will raise a dialog box to save currently displayed automaton to an image file.
+Set of supported image formats depends on installed JRE.\
+The second one will open \jmodule{AutoEditor} tab in NetBeans options. For description of the settings, see \ref{subsection:Settings}.
 
-TODO
+\subsection{Settings}
 
-All settings provided by \jmodule{BasicXSDExporter} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.basicxsd.properties} package.
-As mentioned above, it is possible to set the following.
+All settings provided by \jmodule{AutoEditor} are NetBeans-wide.
+The options panel along with all the logic is in the \code{cz.cuni.mff.ksi.jinfer.autoeditor.options} package.
+Available options include setting the color of the background, colors and shapes of some special types of automaton states.\\
 
-\begin{itemize}
-	\item Turn off generation of global element types. Turning off this feature is not recommended as it may cause certain problems with validity of resulting XSD. See \ref{section:problems}.
-	\item Minimal number of occurrences of element to define its type globally. (Only if generation of global elements is active.)
-	\item Number of spaces in output per one level of indentation.
-	\item Global type name prefix.
-It is a string which will be inserted before a name of a type, which is derived from element's name.
-Can be also an empty string.
-(Only if generation of global elements is active.)
-	\item Global type name suffix.
-It is a string which will be appended after a name of a type, which is derived from element's name.
-Can be also an empty string.
-(Only if generation of global elements is active.)
-\end{itemize}
+To have some effect, these settings need to be implemented by extensions of \code{Visualizer} class.
+For examples, see source codes of \code{StatePickingVisualizer} and \code{StatesPickingVisualizer} classes.\\
 

### 2011-02-21 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 This is an implementation of a \jmodule{AutoEditor}.
-Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.
+Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.\\
 
-TODO spomenut vsade packages
-
-TODO UML diagramy su zjednodusene z dovodu prehladnosti
+UML diagrams shown in this document are simplified to keep them synoptical.
+Simplification involves removing not important class members which are not mentioned in the text, removing members of not jInfer classes and truncating string of inheritance which are not important for this document.\ For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extends \code{JPanel} class directly, though in an UML diagram it has not any methods and extends \code{JPanel} directly, because it is sufficient for understanding of this document.
 
@@
 \code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.
-TODO odkazy v predchadzajucej vete.\\
+TODO odkazy na juliu a JUNG v predchadzajucej vete.\\
 
@@
 
-TODO ako presne ziskava pozicie z graphvizu
+Automaton is transformed to the \emph{dot} representation and the \emph{dot} binary is ran with this representation on input.
+It processes the automaton and writes positions of its states to the standard output.
+Than the output is parsed and an instances of \code{Layout} is created using these positions.\\
 
@@
 \begin{verbatim}
-public static <T> Layout<State<T>, Step<T>> createUserLayout(final Automaton<T> automaton, final Transformer<Step<T>, String> edgeLabelTransformer) throws InterruptedException;
+public static <T> Layout<State<T>, Step<T>> createUserLayout(Automaton<T> automaton,
+                                                             Transformer<Step<T>, String> edgeLabelTransformer)
 \end{verbatim}
@@
 
-System of modules is described in detail in TODO ref.
+System of modules is described in documentation of Architecture, TODO ref na bib.
 
@@
 Main two classes representing visualization of automaton are \code{Visualizer} and \code{AbstractComponent}.
-\code{Visualizer} is a graphical representation of automaton and \code{AbstractComponent} is a panel (extends \code{JPanel}) containing the \code{Visualizer} which will be displayed in GUI.
-TODO obrazok ako AC dedi od JPanelu a obsahuje Visualizer.
+\code{Visualizer} is a graphical representation of automaton and \code{AbstractComponent} is a panel (extends \code{JPanel}) containing the \code{Visualizer} which will be displayed in GUI.\\
 
@@
 
-\code {Visualizer} class extends JUNG \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
+\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
 Responsible methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
@@
 \jmodule{AutoEditor} GUI contains button to save an image of displayed automaton.
-For information on how to to this, see TODO ref.\\
+For information on how to to this, see \ref{subsection:GUI}.\\
 
@@
 
-\code{PluggableVisualizer} class is extension of \code{Visualizer} class, which primarily provides an easy way to plug \emph{graph mouse plugins}.\\
+\code{PluggableVisualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package is extension of \code{Visualizer} class, which primarily provides an easy way to plug \emph{graph mouse plugins}.\\
 
@@
 In the most cases, these plugins are useful but if they are not wanted they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
-
-TODO obrazok ako PluggableVisualizer dedi od Visualizeru
 
@@
 
-\subsubsection{AbstractComponent}
+\subsubsection{AbstractComponent} \label{subsubsection:AbstractComponent}
 
-\code{AbstractComponent} class is a representation of GUI panel containing an instance of \code{Visualizer} class for some automaton.
+\code{AbstractComponent} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component} package is a representation of GUI panel containing an instance of \code{Visualizer} class for some automaton.
 It is inherited from \code{JPanel} class thus provides JPanel's method and behaviour.
@@
 Also, there is no need to call this method directly, it is called by \jmodule{AutoEditor}.
-For details, see TODO ref GUI.
+For details, see \ref{subsection:GUI}.
 \end{itemize}
@@
 It is called by \jmodule{AutoEditor} when displaying the component by \jmodule{AutoEditor} \emph{API} method named \code{drawComponentAndWaitForGUI()}.
-For more information on \jmodule{AutoEditor} \emph{API}, see TODO ref.\\
+For more information on \jmodule{AutoEditor} \emph{API}, see \ref{subsection:API}.\\
 
@@
 
-For examples of user-interactive component, see TODO ref.
+For examples of user-interactive component, see \ref{subsubsection:StatePickingComponent} and \ref{subsubsection:StatesPickingComponent}.
 
-\subsection{API}
+\subsection{API} \label{subsection:API}
 
@@
 
-For examples of API usage, see TODO ref.
+For examples of API usage, see \ref{subsubsection:StatePickingComponent} and \ref{subsubsection:StatesPickingComponent}.
 
@@
 
-\subsubsection{StatePickingComponent}
+\subsubsection{StatePickingComponent} \label{subsubsection:StatePickingComponent}
 \code{StatePickingComponent} (extensions of \code{abstractComponent}) alongside with \code{StatePickingVisualizer} (extension of \code{PluggableVisualizer}) provides possibility to pick one automaton state in GUI and immediately return to the calling code, which can retrieve the picked state.\\
@@
 Purpose of this plugin is to allow user to pick some state of automaton and then call \code{guiDone()} method on instance of \code{AbstractComponent}.
-For description of \code{guiDone()} method, see TODO ref.\\
+For description of \code{guiDone()} method, see \ref{subsubsection:AbstractComponent}.\\
 
@@
 \begin{itemize}
-	\item \code{getPickedState()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see TODO ref), this method retrieves the user picked automaton state.
+	\item \code{getPickedState()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see \ref{subsection:API}), this method retrieves the user picked automaton state.
 	\item \code{setLabel()} - Sets text of a component label. The label can be used to communicate some information to user, for example instructions.
@@
 
-\subsubsection{StatesPickingComponent}
+\subsubsection{StatesPickingComponent} \label{subsubsection:StatesPickingComponent}
 Classes \code{StatePickingComponent} and \code{StatesPickingVisualizer} are similar to the classes described in the previous section.
@@
 'Continue' button and 'Don't ask me anymore to select states to be merger' checkbox.\
-The button is supposed to be pushed when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see TODO ref).\
+The button is supposed to be pushed when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
 State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.
@@
 	\item \code{shallAskUser()} - Retrieves state of the checkbox.
-	\item \code{getPickedStates()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see TODO ref), this method retrieves the list of user picked automaton states.
+	\item \code{getPickedStates()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see \ref{subsection:API}), this method retrieves the list of user picked automaton states.
 In case that none of automaton states was picked, it returns an empty list.
@@
 
-\subsection{GUI}
+\subsection{GUI} \label{subsection:GUI}
 
@@
 As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the main NetBeans window.
-It consists of two buttons, horizontal line below them and below the line, there is panel to place an extension of \code{AbstractComponent} (see TODO ref).\\
+It consists of two buttons, horizontal line below them and below the line, there is panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}).
+Class representing this panel is named \code{AutoEditorTopComponent} and resides in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.topcomponent} package.\\
 

### 2011-02-21 doc/jInfer/treeruledisplayer/treeruledisplayer.tex
Added Documentation for Tree rule displayer
#3187360 improved TreeRD documentation.

@@
 
-\def\mftitle{jInfer BasicRuleDisplayer Module Description}
+\def\mftitle{jInfer TreeRuleDisplayer Module Description}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 This rule displayer uses JUNG graph implementation to draw rules as forest of trees, where each tree represents one rule of whole set.
-In each tree, root vertex represents left side of the rule, inner vertices represents concatenation, alternation or permutation and leafs are elements, attributes or simple data.
+In each tree, root vertex represents left side of the rule, inner vertices represents concatenation, alternation or permutation and leafs are elements, simple data or attribute.
 As in basic rule displayer it creates a component with multiple tabs: each one for a new grammar to display.
@@
 
-Main window of rule displayer may have multiple tabs in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar using \jmodule{JUNG} library.\\
+Main window of rule displayer may have multiple tabs in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar using \jmodule{JUNG} library.
+Each tab in rule displayer window is represented by instance of \code{RulePanel} which contains button linking to options and the main panel(canvas) with rendered rules.
+Canvas is made by \code{VisualizationViewer} instance with \code{TreeLayout} layout which are both classes from JUNG library.
+For \code{VisualizationViewer} is set \code{GraphMouse} with transforming mode allowing zooming and moving of graph in canvas.
+To allow user set custom appearance of rules, we set various \code{Transformer}s in \code{RenderContext} of \code{VisualizationViewer}.
+These transormers change color, shape and size vertices in each rule tree according to properties set in options.\\
 
 All the trees representing rules are created in the \code{GraphBuilder} class with method \code{buildGraphPanel()}.
-While the code itself is a nice recursion programming exercise, there is nothing of a special interest in it.\\
+This method creates \code{DelegateForrest} of \code{DelegateTree}s where each tree of the forrest represents one rule.
+Each vertex of the tree is represented by \code{Regexp<?
+extends AbstractNamedNode>} and edge by \code{RegexpInterval}.
+While the code building tree structure is a nice recursion programming exercise, there is nothing of a special interest in it.
+Finally the method appends forrest into layout and visualizationViewer which were described above.
+All code responsible for building rule trees is contained in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.logic}.
+For more details about JUNG library please see \cite{jung}.\\
 

### 2011-02-22 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 \section{Introduction}
-
 This is an implementation of a \jmodule{AutoEditor}.
-Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.\\
+Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.
+For more information on JUNG library, see \cite{jung}.\\
 
@@
 \section{Structure}
-
 Structure of \jmodule{AutoEditor} can be divided into following four main parts.
-
 \begin{itemize}
@@
 \end{itemize}
+First, \code{Layout}s and use of base classes to create a visualization of automaton will be described.\\
 
-First, \code{Layout}s and use of base classes to create a visualization of automaton will be described.
+TODO prelozit
+Ak je nejaka genericka trieda parametrizovana typom \code{T}, tak pri instanciovani tejto triedy je nutne tento typ zadat rovnaky, akym je instanciovany automat, pre ktory sa tato trieda vyraba.
+JUNG triedy su vacsinou parametrizovane typom pre stav a hranu.
+V takom pripade sa tieto triedy instanciuju typmi \code{State\textless T\textgreater} pre stav a \code{Step\textless T\textgreater} pre hranu.
+Refer to \cite{archdoc} for more information on these classes and see \ref{subsubsection:StatePickingComponent} and \ref{subsubsection:StatesPickingComponent} for source code examples.
 
 \subsection{Layout}
-
 \code{Layout} is a JUNG interface responsible primarily of representation of automaton and positions of its states.
@@
 \code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.
-TODO odkazy na juliu a JUNG v predchadzajucej vete.\\
+TODO odkazy na juliu, graphviz a JUNG v predchadzajucej vete.\\
 
@@
 \subsubsection{Vyhnanovska Layout} \label{subsubsection:vyhnanovskaLayout}
-
 As mentioned above, this Layout was implemented by \emph{Julie Vyhnanovska} as a part of her master thesis.
@@
 
-Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.vyhnanovska}.
+Source codes reside in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.vyhnanovska}.
 
 \subsubsection{Graphviz Layout} \label{subsubsection:graphvizLayout}
-
 \code{Graphviz Layout} uses \emph{Graphviz}, third-party graph visualization software, to create positions of automaton states.
@@
 
-Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.
+Source codes reside in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.
 
 \subsubsection{LayoutHelperFactory} \label{subsubsection:LayoutHelperFactory}
-
 In project properties, it is possible to select a \code{Layout} to be used to display automata.
@@
 This method has following signature.
-
 \begin{verbatim}
-public static <T> Layout<State<T>, Step<T>> createUserLayout(Automaton<T> automaton,
-                                                             Transformer<Step<T>, String> edgeLabelTransformer)
+public static <T> Layout<State<T>, Step<T>>
+createUserLayout(Automaton<T> automaton, Transformer<Step<T>, String> edgeLabelTransformer)
 \end{verbatim}
-
-The first argument is a automaton to create a layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.
+The first argument is a automaton to create a layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\
 
@@
 
-System of modules is described in documentation of Architecture, TODO ref na bib.
+System of modules is described in documentation of Architecture, see \cite{archdoc}.
 
@@
 Main two classes representing visualization of automaton are \code{Visualizer} and \code{AbstractComponent}.
-\code{Visualizer} is a graphical representation of automaton and \code{AbstractComponent} is a panel (extends \code{JPanel}) containing the \code{Visualizer} which will be displayed in GUI.\\
+\code{Visualizer} is a graphical representation of automaton and \code{AbstractComponent} is a panel (extends \code{JPanel}) containing \linebreak the \code{Visualizer} which will be displayed in GUI.\\
 
@@
 
-\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
+\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
 Responsible methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
@@
 
-Constructor has one argument, instance of \code{Layout} interface created from an automaton, typically by \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
+Constructor has one argument, instance of \code{Layout} interface created from an automaton, typically by \linebreak \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
 
@@
 By default, instance of \code{PluggableVisualizer} is constructed with two plugins enabled.
-They are \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
+They are \linebreak \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
 In the most cases, these plugins are useful but if they are not wanted they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
@@
 \begin{figure}
-\centering\includegraphics[scale=\myscale]{uml_derived_classes}
+\centering\includegraphics[scale=0.65]{uml_derived_classes}
 \caption{Class diagram for the derived classes.} \label{uml_derived_classes}
@@
 Its constructor has additional two arguments, superinitial and superfinal states as we want to distinguish these states in displayed automaton.\
-And upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \code{PluggableVisualizer}.
+And upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \linebreak \code{PluggableVisualizer}.
 Purpose of this plugin is to allow user to pick some state of automaton and then call \code{guiDone()} method on instance of \code{AbstractComponent}.
@@
 
-Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \code{VertexPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
+Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \linebreak \code{VertexPickingGraphMousePlugin} in \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
 
@@
 \begin{verbatim}
-    final Transformer<Step<Regexp<T>>, String> transformer = new Transformer<Step<Regexp<T>>, String>() {
+    final Transformer<Step<Regexp<T>>, String>
+      transformer = new Transformer<Step<Regexp<T>>, String>() {
 
@@
     final StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
-    final StatePickingVisualizer<Regexp<T>> visualizer = new StatePickingVisualizer<Regexp<T>>(LayoutHelperFactory.createUserLayout(automaton, transformer), transformer, component, automaton.getSuperInitialState(), automaton.getSuperFinalState());
+    final Layout<State<T>, Step<T>> layout = LayoutHelperFactory.createUserLayout(automaton,
+                                                                                  transformer);
+    final StatePickingVisualizer<Regexp<T>>
+      visualizer = new StatePickingVisualizer<Regexp<T>>(layout,
+                                                         transformer,
+                                                         component,
+                                                         automaton.getSuperInitialState(),
+                                                         automaton.getSuperFinalState());
     component.setVisualizer(visualizer);
@@
 
-      if ((removeState.equals(automaton.getSuperFinalState())) || (removeState.equals(automaton.getSuperInitialState()))) {
+      if ((removeState.equals(automaton.getSuperFinalState()))
+           || (removeState.equals(automaton.getSuperInitialState()))) {
         component.setLabel("Do not select superInitial and superFinal states.");
@@
 
-\code{StatesPickingVisualizer} is extension of \code{PluggableVisualizer} and upon its construction it adds \code{VerticesPickingGraphMousePlugin} to the graph mouse plugins.
+\code{StatesPickingVisualizer} is extension of \code{PluggableVisualizer} and upon its construction it adds \linebreak \code{VerticesPickingGraphMousePlugin} to the graph mouse plugins.
 With this plugin used, user can pick and unpick automaton states separately or pick several states at once by dragging rectangular selection box over states.\
@@
 
-Source codes of \code{StatesPickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \code{VerticesPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
+Source codes of \code{StatesPickingComponent} resides in package \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \newline \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \newline \code{VerticesPickingGraphMousePlugin} in \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
 
@@
 \begin{verbatim}
-  Automaton<T> simplify(Automaton<T> inputAutomaton, final SymbolToString<T> symbolToString, final String elementName) {
+  Automaton<T> simplify(Automaton<T> inputAutomaton,
+                        SymbolToString<T> symbolToString,
+                        String elementName) {
     if (!askUser) {
@@
     do {
-      final StatesPickingVisualizer<T> visualizer = new StatesPickingVisualizer<T>(LayoutHelperFactory.createUserLayout(inputAutomaton, transformer), transformer);
+      final Layout<State<T>, Step<T>>
+        layout = LayoutHelperFactory.createUserLayout(inputAutomaton, transformer);
+      final StatesPickingVisualizer<T>
+        visualizer = new StatesPickingVisualizer<T>(layout, transformer);
 
@@
       if (selectTwo) {
-        panel.setLabel("Automaton for element <" + elementName + ">. Please select states to be merged and click continue. Select at least 2 states.");
+        panel.setLabel("Automaton for element <" + elementName
+          + ">. Please select states to be merged and click continue. Select at least 2 states.");
       } else {
-        panel.setLabel("Automaton for element <" + elementName + ">. Please select states to be merged and click continue.");
+        panel.setLabel("Automaton for element <" + elementName
+          + ">. Please select states to be merged and click continue.");
       }
@@
 
-\subsection{Settings}
+\subsection{Settings} \label{subsection:Settings}
 

### 2011-02-22 doc/jInfer/treeruledisplayer/treeruledisplayer.tex
#3187360 fixed doc from review, added basic class diagram

@@
 
-This rule displayer uses JUNG graph implementation to draw rules as forest of trees, where each tree represents one rule of whole set.
+This rule displayer uses JUNG graph implementation to draw rules as a forest of trees, where each tree represents one rule from the whole set.
 In each tree, root vertex represents left side of the rule, inner vertices represents concatenation, alternation or permutation and leafs are elements, simple data or attribute.
-As in basic rule displayer it creates a component with multiple tabs: each one for a new grammar to display.
+As in basic rule displayer, tree rule displayer creates a component with multiple tabs: each one for a new grammar to display.
 
@@
 
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{class-diagram}
+	\caption{TreeRuleDisplayer class diagram} \label{figure-class-diagram}
+\end{figure}
+
+The diagram of classes involved in rule display is in figure \ref{figure-class-diagram}.
 Main window of rule displayer may have multiple tabs in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar using \jmodule{JUNG} library.
-Each tab in rule displayer window is represented by instance of \code{RulePanel} which contains button linking to options and the main panel(canvas) with rendered rules.
-Canvas is made by \code{VisualizationViewer} instance with \code{TreeLayout} layout which are both classes from JUNG library.
-For \code{VisualizationViewer} is set \code{GraphMouse} with transforming mode allowing zooming and moving of graph in canvas.
-To allow user set custom appearance of rules, we set various \code{Transformer}s in \code{RenderContext} of \code{VisualizationViewer}.
-These transormers change color, shape and size vertices in each rule tree according to properties set in options.\\
+Each tab in rule displayer window is represented by instance of \code{RulePanel} which contains button linking to options and the main panel (canvas) with rendered rules.
+Canvas is created by \code{VisualizationViewer} instance with \code{TreeLayout} layout (which are both classes from JUNG library).
+For \code{VisualizationViewer}, \code{GraphMouse} is set with transforming mode to allows zooming and moving of graph in canvas.
+To allow the user to set custom appearance of rules, we set various \code{Transformer}s in \code{RenderContext} of \code{VisualizationViewer}.
+These transormers change color, shape and vertex size in each rule tree according to properties set in options.\\
 
-All the trees representing rules are created in the \code{GraphBuilder} class with method \code{buildGraphPanel()}.
+All the trees representing rules are created in the \code{GraphBuilder} class using the \code{buildGraphPanel()} method.
 This method creates \code{DelegateForrest} of \code{DelegateTree}s where each tree of the forrest represents one rule.
-Each vertex of the tree is represented by \code{Regexp<?
-extends AbstractNamedNode>} and edge by \code{RegexpInterval}.
+Each vertex of the tree is represented by \code{Regexp<?\ extends AbstractNamedNode>} and edge by \code{RegexpInterval}.
 While the code building tree structure is a nice recursion programming exercise, there is nothing of a special interest in it.
-Finally the method appends forrest into layout and visualizationViewer which were described above.
+Finally the method appends forrest into layout and VisualizationViewer which were described above.
 All code responsible for building rule trees is contained in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.logic}.

### 2011-02-22 doc/jInfer/projecttype/projecttype.tex
#3187361 first commit of projecttype documentation

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}
\usepackage{xstring}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer ProjectType Module Description}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
\noindent Target audience: developers willing to extend jInfer, specifically extend jInfer project structure.

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & Michal Švirec \\ \hline
Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\
 & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\
 & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\
 & org.openide.windows.IOProvider \\ \hline
Provided tokens:       & none \\ \hline
Module dependencies:   & Base \\
	& Runner \\ \hline
Public packages:       & cz.cuni.mff.ksi.jinfer.projecttype.actions \\ \hline
\end{tabular}

\section{Introduction}

\jmodule{ProjectType} is the module responsible for creation of NBP project type which groups input/output files that belongs to one specific inference. Each jInfer project also allows user to set properties specific for inference.

\section{Structure}

Structure of \jmodule{ProjectType} can be divided into following five main parts.
\begin{itemize}
	\item Base classes - Classes providing main functionality like creation of project, defining operations like move, delete, copy etc. All the base classes are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype} package.
	\item Visualization classes - These classes create tree structure of project in NBP Projects view and are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype.nodes}
	\item Actions - Classes from \code{cz.cuni.mff.ksi.jinfer.projecttype.actions} package that provides actions allowing adding, removing input files into project, or running the project.
	\item Properties - Classes responsible for creation of project properties window with properties category tree. These are situated in  \code{cz.cuni.mff.ksi.jinfer.projecttype.properties} package.
	\item Project wizard - Classes representing creation of project through new project wizard from  \code{cz.cuni.mff.ksi.jinfer.projecttype.sample} package.
\end{itemize}

\subsection{Base classes}

\subsection{Visualization classes}

\subsection{Actions}

\subsection{Properties}

\subsection{Project wizard}

\subsection{Preferences}

\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-02-22 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{project_dir_structure}
+	\caption{Project directory structure} \label{dir-structure}
+\end{figure}
+
+This section describes classes needed for creation of the jInfer project and correct integration into NBP.
+With each project created in jInfer is also created specific directory structure on filesystem that describes jInfer project.
+This structure is described in figure \ref{dir-structure}, where rectangles represent folders and rounded rectangles files.\\
+
+Main two classes representing jInfer project type are \code{JInferProjectFactory} and \code{JInferProject}.
+\code{JInferProjectFactory} is a factory classes responsible for creation of jInfer project from disk directory and \code{JinferProject} is represent jInfer project in NBP.
+
 \subsection{Visualization classes}

### 2011-02-23 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor modified according to the review remarks. Not finished yet.

@@
 \section{Introduction}
-This is an implementation of a \jmodule{AutoEditor}.
-Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive.
+This is an implementation of a an automaton editor.
+Using JUNG library, it provides an API to display and modify automata in an interactive mode.
 For more information on JUNG library, see \cite{jung}.\\
 
-UML diagrams shown in this document are simplified to keep them synoptical.
-Simplification involves removing not important class members which are not mentioned in the text, removing members of not jInfer classes and truncating string of inheritance which are not important for this document.\ For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extends \code{JPanel} class directly, though in an UML diagram it has not any methods and extends \code{JPanel} directly, because it is sufficient for understanding of this document.
+UML diagrams shown in this document are simplified to keep them readable.
+Simplification involves removing not important class members which are not mentioned in the text, removing members of non-jInfer classes and truncating string of inheritance which are not important for this document.\ For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extends \code{JPanel} class directly, though in an UML diagram it has not any methods and extends \code{JPanel} directly, because it is sufficient for understanding of this document.
 
@@
 \subsection{Layout}
-\code{Layout} is a JUNG interface responsible primarily of representation of automaton and positions of its states.
+\code{Layout} is a JUNG interface responsible primarily of representation of an automaton and positions of its states.
 JUNG library provides several implementation of \code{Layout} interface.
-However, none of them is convenient for automatic automaton displaying, \jmodule{AutoEditor} provides two additional implementations.
+However, because none of them is convenient for automatic automaton displaying, \jmodule{AutoEditor} provides two additional implementations.
 \code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.
@@
 TODO link?
-It positions automaton states to a square grid.
-This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confused.\\
+It positions automaton states on a square grid.
+This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confusing.\\
 
-Source codes reside in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.vyhnanovska}.
+Source code resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.}\\*\code{vyhnanovska}.
 
@@
 
-Automaton is transformed to the \emph{dot} representation and the \emph{dot} binary is ran with this representation on input.
+Automaton is transformed to the \emph{dot} representation and the \emph{dot} binary is invoked with this representation on input.
 It processes the automaton and writes positions of its states to the standard output.
-Than the output is parsed and an instances of \code{Layout} is created using these positions.\\
+The output is then parsed and an instance of \code{Layout} is created using these positions.\\
 
-Source codes reside in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.
+Source code resides in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.
 
@@
 In project properties, it is possible to select a \code{Layout} to be used to display automata.
-\code{LayoutHelperFactory} is class, providing just one static method, responsible for creating instances of \code{Layout}s according to a selection in project properties.\\
+\code{LayoutHelperFactory} is a class providing just one static method responsible for creating instances of \code{Layout}s according to a selection in project properties.\\
 
-This method has following signature.
+This method has the following signature.
 \begin{verbatim}
@@
 \end{verbatim}
-The first argument is a automaton to create a layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\
+The first argument is an automaton to create layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\
 
-Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}.
+Source code resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}.
 
@@
 
-\code{Layout}s can be implemented using the modular system.
-To create a new implementation of \code{Layout} interface, it is needed to create a new class implementing \code{LayoutFactory} interface (package \code{package cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.
+New \code{Layout}s can be implemented using the modular system.
+For information on system of modules, see \cite{archdoc}.
+To create a new implementation of \code{Layout} interface, it is necessary to create a new class implementing \code{LayoutFactory} interface (package \code{package cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.
 
@@
 Created implementation will be shown in project properties in the \code{Layout} selection.\\
-
-System of modules is described in documentation of Architecture, see \cite{archdoc}.
 
@@
 \code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
-Responsible methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
+Relevant methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
 However, to save an image of automaton it is not necessary to call this methods directly.
@@
 
-Constructor has one argument, instance of \code{Layout} interface created from an automaton, typically by \linebreak \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
+Constructor has one argument, an instance of \code{Layout} interface created from an automaton, typically by \linebreak \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
 
@@
 They are \linebreak \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
-In the most cases, these plugins are useful but if they are not wanted they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
+In the most cases these plugins are useful, otherwise they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
 
@@
 \code{AbstractComponent} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component} package is a representation of GUI panel containing an instance of \code{Visualizer} class for some automaton.
-It is inherited from \code{JPanel} class thus provides JPanel's method and behaviour.
+It is inherited from \code{JPanel} class, thus provides JPanel's method and behaviour.
 In addition, it provides the following methods.
@@
 	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by interrupt or regularly.
-Also, there is no need to call this method directly, it is called by \jmodule{AutoEditor}.
+Again, there is no need to call this method directly, it is called by \jmodule{AutoEditor}.
 For details, see \ref{subsection:GUI}.
@@
 
-Programmer implementing an extension of \code{AbstractComponent} is not forced to place the \code{Visualizer} on its own.
+Programmer implementing an extension of \code{AbstractComponent} is not forced to place the \code{Visualizer} on his own.
 It is just needed to create \code{JPanel} and define \code{getAutomatonDrawPanel()} method to return this \code{JPanel}.
@@
 
-If some kind of user interactivity is desired, \code{AbstractComponent} is a right place to implement it.\\
+If some kind of user interactivity is desired, \code{AbstractComponent} is the right place to implement it.\\
 
 To display the component in GUI and wait for some user action, method \code{waitForGuiDone()} is used.
-After displaying the component, calling of this method suspend running thread thus code execution of a caller module is stopped at the place of this call.
+After displaying the component, calling this method suspends the running thread, thus code execution of a caller module is stopped at the place of this call.
 However, GUI is ran in another thread, user is able to interact with the panel (component).\\*
@@
 
-Method important for a programmer extending \code{AbstractComponent} is named \code{guiDone()}.
+Method important for a programmer extending \code{AbstractComponent} is called \code{guiDone()}.
 This method wakes up the thread suspended in \code{waitForGuiDone()} method and the programmer is responsible for calling it.
@@
 Use this method to just display automaton, without any user interaction and without waiting for any external event.
-This method does not support these.
-	\item \code{drawComponentAndWaitForGUI()} - Displays given \code{AbstractComponent} in a GUI thread and a caller thread is suspended until  \code{guiDone()} method of \code{AbstractComponent()} is called.
-This method can wait for GUI events thus is convenient for user interaction.
-How to achieve it is described in detail in \ref{subsubsection:AbstractComponent user-interactivity support}.
+This method does not support it.
+	\item \code{drawComponentAndWaitForGUI()} - Displays given \code{AbstractComponent} in a GUI thread, while the caller thread is suspended until  \code{guiDone()} method of \code{AbstractComponent()} is called.
+This method can wait for GUI events and thus is convenient for user interaction.
+How to achieve this is described in detail in \ref{subsubsection:AbstractComponent user-interactivity support}.
 	\item \code{closeTab()} - Closes \jmodule{AutoEditor}'s GUI tab and interrupts inference, if running.
@@
 
-This section describes classes derived from the base classes used in \jmodule{Two Step Simplifier} module to display automata. This classes may also serve as examples of extending the base classes.
+This section describes classes derived from the base classes used in \jmodule{Two Step Simplifier} module to display automata. This classes may also serve as examples of base classes extensions.
 
@@
 
-Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \linebreak \code{VertexPickingGraphMousePlugin} in \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
+Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.}\linebreak\code{component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and\linebreak \code{VertexPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.}\linebreak\code{graphmouseplugins}.\\
 
@@
 \begin{verbatim}
-    final Transformer<Step<Regexp<T>>, String>
+    Transformer<Step<Regexp<T>>, String>
       transformer = new Transformer<Step<Regexp<T>>, String>() {
@@
     State<Regexp<T>> removeState;
-    final StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
-    final Layout<State<T>, Step<T>> layout = LayoutHelperFactory.createUserLayout(automaton,
-                                                                                  transformer);
-    final StatePickingVisualizer<Regexp<T>>
+    StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
+    Layout<State<T>, Step<T>> layout = LayoutHelperFactory.createUserLayout(automaton, transformer);
+    StatePickingVisualizer<Regexp<T>>
       visualizer = new StatePickingVisualizer<Regexp<T>>(layout,
@@
 'Continue' button and 'Don't ask me anymore to select states to be merger' checkbox.\
-The button is supposed to be pushed when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
+The button is supposed to be clicked when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
 State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.
@@
 	\item \code{getPickedStates()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see \ref{subsection:API}), this method retrieves the list of user picked automaton states.
-In case that none of automaton states was picked, it returns an empty list.
-	\item \code{setLabel()} - Sets text of a component label. The label can be used to communicate some information to user, for example instructions.
+In the case that none of automaton states was picked, it returns an empty list.
+	\item \code{setLabel()} - Sets text of the component label. The label can be used to communicate some information to user, for example instructions.
 \end{itemize}
 
-Source codes of \code{StatesPickingComponent} resides in package \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \newline \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \newline \code{VerticesPickingGraphMousePlugin} in \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\
+Source codes of \code{StatesPickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.}\linebreak\code{component}, \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and\linebreak \code{VerticesPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.}\linebreak\code{graphmouseplugins}.\\
 
@@
     List<State<T>> mergeLst;
-    final Transformer<Step<T>, String> transformer = new Transformer<Step<T>, String>() {
+    Transformer<Step<T>, String> transformer = new Transformer<Step<T>, String>() {
 
@@
     do {
-      final Layout<State<T>, Step<T>>
+      Layout<State<T>, Step<T>>
         layout = LayoutHelperFactory.createUserLayout(inputAutomaton, transformer);
-      final StatesPickingVisualizer<T>
-        visualizer = new StatesPickingVisualizer<T>(layout, transformer);
+      StatesPickingVisualizer<T> visualizer = new StatesPickingVisualizer<T>(layout, transformer);
 
-      final StatesPickingComponent<T> panel = new StatesPickingComponent<T>();
+      StatesPickingComponent<T> panel = new StatesPickingComponent<T>();
       panel.setVisualizer(visualizer);
@@
 
-As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the main NetBeans window.
-It consists of two buttons, horizontal line below them and below the line, there is panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}).
+As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the editor window.
+It consists of two buttons, horizontal line below them and and a panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}).
 Class representing this panel is named \code{AutoEditorTopComponent} and resides in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.topcomponent} package.\\
 
-Buttons have labels 'Save as image' and 'Show settings'.
+Buttons have labels 'Save as Image' and 'Show Settings'.
 Pushing the first one will raise a dialog box to save currently displayed automaton to an image file.

### 2011-02-23 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor modified according to the review remarks.

@@
 However, because none of them is convenient for automatic automaton displaying, \jmodule{AutoEditor} provides two additional implementations.
-\code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.
-TODO odkazy na juliu, graphviz a JUNG v predchadzajucej vete.\\
+\code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software.\\
 
@@
 \subsubsection{Vyhnanovska Layout} \label{subsubsection:vyhnanovskaLayout}
-As mentioned above, this Layout was implemented by \emph{Julie Vyhnanovska} as a part of her master thesis.
-TODO link?
+As mentioned above, this Layout was implemented by \emph{Julie Vyhnanovska} as a part of her master thesis (see \cite{Vyhnanovska}).
 It positions automaton states on a square grid.
@@
 \subsubsection{Graphviz Layout} \label{subsubsection:graphvizLayout}
-\code{Graphviz Layout} uses \emph{Graphviz}, third-party graph visualization software, to create positions of automaton states.
-TODO link?
+\code{Graphviz Layout} uses \emph{Graphviz}, third-party graph visualization software (see \cite{graphviz}), to create positions of automaton states.
 To use this \code{Layout}, \emph{Graphviz} has to be installed and path to \emph{dot} binary has to be set in options.
@@
 \begin{verbatim}
-public static <T> Layout<State<T>, Step<T>>
-createUserLayout(Automaton<T> automaton, Transformer<Step<T>, String> edgeLabelTransformer)
+public static <T> Layout<State<T>, Step<T>> createUserLayout(
+  Automaton<T> automaton,
+  Transformer<Step<T>, String> edgeLabelTransformer)
 \end{verbatim}
@@
 
-\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file.
+\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it inherits all its methods and adds support for saving contained automaton to an image file.
 Relevant methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
@@
 There is no need to called this method directly.
-	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by interrupt or regularly.
+	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by \code{guiInterrupt()} method or regularly (\code{guiDone()} method or GUI was not waiting for user interaction).
 Again, there is no need to call this method directly, it is called by \jmodule{AutoEditor}.
@@
 \centering\includegraphics[scale=0.65]{uml_derived_classes}
-\caption{Class diagram for the derived classes.} \label{uml_derived_classes}
+\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.} \label{uml_derived_classes}
 \end{figure}

### 2011-02-23 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 \jmodule{ProjectType} is the module responsible for creation of NBP project type which groups input/output files that belongs to one specific inference.
-Each jInfer project also allows user to set properties specific for inference.
+Each jInfer project also allows user to set properties specific for inference.\\
+
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{project_dir_structure}
+	\caption{Project directory structure} \label{dir-structure}
+\end{figure}
+
+With each project created in jInfer is also created specific directory structure on filesystem that describes jInfer project.
+This structure is described in figure \ref{dir-structure}, where rectangles represent folders and rounded rectangles files.
+Folder in filesystem is considered as jInfer project folder, if contains \emph{jinferproject} subfolder.
+These folder contains two files, first is \emph{project.properties}, where all properties set for particular project are saved.
+Second file is \emph{input.files} which is xml file filled with paths to files inserted as input into particular project (structure of this file is described in section \ref{input.files}).
 
@@
 
-\begin{figure}
-	\centering\includegraphics[scale=\myscale]{project_dir_structure}
-	\caption{Project directory structure} \label{dir-structure}
-\end{figure}
+This section describes classes needed for creation of the jInfer project and correct integration into NBP.
+Main two classes representing jInfer project type are \code{JInferProjectFactory} and \code{JInferProject}.
 
-This section describes classes needed for creation of the jInfer project and correct integration into NBP.
-With each project created in jInfer is also created specific directory structure on filesystem that describes jInfer project.
-This structure is described in figure \ref{dir-structure}, where rectangles represent folders and rounded rectangles files.\\
+\subsubsection{JInferProjectFactory}
 
-Main two classes representing jInfer project type are \code{JInferProjectFactory} and \code{JInferProject}.
-\code{JInferProjectFactory} is a factory classes responsible for creation of jInfer project from disk directory and \code{JinferProject} is represent jInfer project in NBP.
+\code{JInferProjectFactory} is a factory class implementing \code{ProjectFactory} interface provided by NBP.
+This class is responsible for loading/saving of jInfer project from/to filesystem and also determining if folder in filesystem is type of jInfer project (contains \emph{jinferproject} subfolder).
+For this purpose, class has three methods: \code{isProject()}, \code{save()} and \code{load()}.
+When jInfer project is loaded from filesystem, \code{JInferProjectFactory} creates new instance of \code{JInferProject} passing path to project folder as a constructor parameter.
+When \code{save()} method is invoced, all the project properties are saved in \emph{projectproperties} file and paths to input files are saved in \emph{input.files}.
+
+\subsubsection{JInferProject}
+
+\code{JinferProject} class implements \code{Project} interface from NBP and represents in-memory representation of jInfer project.
+Inner structure of this class is very simple, interface provides only two methods: \code{getProjectDirectory()} and \code{getLookup()}.
+All the extensibility of project type is done by \emph{Lookup} functionality of NBP.
+Project lookup contains besides properties and \code{Input} class, which encapsulates input files, also class that creates output files, class that builds project tree in project window, etc.
+
+\subsubsection{InputFiles class}\label{input.files}
+
+\code{InputFiles} is utility class that stores/loads input file paths into/from input.files xml file.
+For this purpose, class uses \emph{JAXB} architecture to unmarshall/marshall xml file into/from java content objects.
+Xml structure of the input.files is very simple, \emph{jinferinput} is root element under which are \emph{xml}, \emph{schemas} and \emph{queries} elements.
+Each of this elements could have none or more \emph{file} elements with string attribute \emph{loc} that contains absolute path to input file.\\
+
+\noindent Example of the structure follows.
+\begin{verbatim}
+	<?xml version="1.0" encoding="UTF-8" standalone="yes"?>
+	<jinferinput>
+	    <xml>
+	        <file loc="/home/user/example.xml"/>
+	    </xml>
+	    <schemas>
+		<file loc="/home/user/example.dtd"/>
+	    </schemas>
+	    </queries/>
+	</jinferinput>
+\end{verbatim}
 

### 2011-02-23 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 Folder in filesystem is considered as jInfer project folder, if contains \emph{jinferproject} subfolder.
-These folder contains two files, first is \emph{project.properties}, where all properties set for particular project are saved.
+This folder contains two files, first is \emph{project.properties}, where all properties set for particular project are saved.
 Second file is \emph{input.files} which is xml file filled with paths to files inserted as input into particular project (structure of this file is described in section \ref{input.files}).
+\emph{Output} subfolder contains schema files which were generated by jInfer inference.
 
@@
 \code{InputFiles} is utility class that stores/loads input file paths into/from input.files xml file.
-For this purpose, class uses \emph{JAXB} architecture to unmarshall/marshall xml file into/from java content objects.
+For this purpose, class uses \emph{JAXB} \cite{JAXB} architecture to unmarshall/marshall xml file into/from java content objects.
 Xml structure of the input.files is very simple, \emph{jinferinput} is root element under which are \emph{xml}, \emph{schemas} and \emph{queries} elements.
@@
 
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{node_structure}
+	\caption{Project directory structure} \label{node-structure}
+\end{figure}
+
+Each project in NBP is after creation opened in project window and has some visual tree structure where each node in this tree is represented by NBP \code{Node} class.
+This structure is in jInfer project type provided by \code{JInferLogicalView} class that implements \code{LogicalViewProvider} interface from NBP and is available through \code{JInferProject} lookup.
+This class has one important method, \code{createLogicalView()} that returns instance of \code{RootNode} class of root node.
+Simple class structure is described in figure \ref{node-structure}.\\
+
+\code{RootNode} extends NBP \code{AbstractNode} class which is basic implementation of NBP \code{Node}.
+This node contains four children nodes: three for input files (Document, Schema and Query) and one for output files generated by inference (Output).
+Root node also offers actions for adding input files, run inference and standard project actions like delete or move/rename project.\\
+
+Each of the \emph{input} nodes is represented by \code{FolderNode} class and contains \emph{input file} nodes (represented by \code{FileNode} class) representing files added to the jInfer project.
+Notice that these \emph{input file} nodes represents files from filesystem and are not copied to project folder structure.
+Because of this, if you delete this nodes from project tree, files from filesystem are not deleted.\\
+
+\code{OutputNode} class represents \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} subfolder in project filesystem hierarchy.
+In other words, files contained in this folder are shown as subnodes in \emph{Output} node in project tree.
+This subnodes are represented by \code{OutputFileNode} class.
+Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is also deleted from filesystem.
+
 \subsection{Actions}

### 2011-02-24 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 With each project created in jInfer is also created specific directory structure on filesystem that describes jInfer project.
-This structure is described in figure \ref{dir-structure}, where rectangles represent folders and rounded rectangles files.
+This structure is described in figure \ref{dir-structure}.
 Folder in filesystem is considered as jInfer project folder, if contains \emph{jinferproject} subfolder.
 This folder contains two files, first is \emph{project.properties}, where all properties set for particular project are saved.
-Second file is \emph{input.files} which is xml file filled with paths to files inserted as input into particular project (structure of this file is described in section \ref{input.files}).
+Second file is \emph{input.files} which is XML file filled with paths to files inserted as input into particular project (structure of this file is described in section \ref{input.files}).
 \emph{Output} subfolder contains schema files which were generated by jInfer inference.
@@
 
-\code{InputFiles} is utility class that stores/loads input file paths into/from input.files xml file.
-For this purpose, class uses \emph{JAXB} \cite{JAXB} architecture to unmarshall/marshall xml file into/from java content objects.
-Xml structure of the input.files is very simple, \emph{jinferinput} is root element under which are \emph{xml}, \emph{schemas} and \emph{queries} elements.
+\code{InputFiles} is utility class that stores/loads input file paths into/from input.files XML file.
+For this purpose, class uses \emph{JAXB} \cite{JAXB} architecture to unmarshall/marshall XML file into/from java content objects.
+XML structure of the input.files is very simple, \emph{jinferinput} is root element under which are \emph{documents}, \emph{schemas} and \emph{queries} elements.
 Each of this elements could have none or more \emph{file} elements with string attribute \emph{loc} that contains absolute path to input file.\\
@@
 
+Each node in project tree hierarchy contains set of actions available from its context menu.
+Besides standard actions for root node or file node, jInfer project type introduce its own actions.
+These actions could be separated into three groups: actions for \emph{root} node, actions for \emph{input folder} nodes and one action for \emph{input/output file} nodes implemented by \code{ValidateAction} class.
+All classes implementing some jInfer project action extend standard java \code{javax.swing.AbstractAction} class besides \code{ValidateAction} class which is described in \ref{validate-action}.\\
+
+\subsubsection{Root node actions}
+
+This group contains two actions, first is action that allows adding input files into \emph{input folder} nodes (implemented by \code{FilesAddAction} class).
+Second one is slightly different, because unlike from the previous action, this one doesn't create new item in \emph{root} node context menu but implements functionality for item that already exists in context menu - Run action.\\
+
+First of the two actions, after it was invoked, open file choose dialog which has set file filter according to available classes that implement \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
+After selecting files in dialog, these files are inserted into particular \emph{input file} node according to mapping of file extension to folder type gained from \code{Processor} classes.\\
+
+Run action, which is available in root node context menu, is also present in NBP toolbar as a \emph{Run} button and in jInfer Welcome window.
+This action is implemented in \code{RunAction} class and simply invokes \jmodule{Runner}s \code{run()} method, if no other inference is already running, otherwise information dialog pops up informing about already running inference.
+This check is done with \code{setActiveProject()} from \code{RunningProject} class.
+
+\subsubsection{Input folder node actions}
+
+To this group of actions belong two actions: action for adding input files into \emph{input folder} node (\code{FileAddAction} class) and action for deleting all input files in \emph{input folder} node (\code{DeleteAllAction} class).
+\code{FileAddAction} class implements action which opens file choose dialog and selected files are added into particular \emph{file input}node.
+\code{DeleteAllAction} class simply removes all input files previuosly added into particular \emph{file input} node.
+
+\subsubsection{ValidateAction class}\label{validate-action}
+
+\code{ValidateAction} class implements action, which allows documents and schema files to be validated against each other.
+Because of this, action is presented in context menu of document and schema files (in actual implementation is available for XML, dtd and XML schema files).
+Unlike the other action classes, this extends NBP \code{NodeAction} class, which allows to enable/disable action in context menu according to node selection.
+Action is enabled only if exactly one schema file is selected and one or more XML files.
+If the selected XML files are valid against selected schema, information dialog is popped up, otherwise error dialog is popped up and in \emph{validate} output window is printed error message describing cause of non-validity XML files against schema.
+
 \subsection{Properties}
+
+Each jInfer project has associated properties window through which can user change project-wide properties.
+This window is accesible through \emph{properties} action in project \emph{root} node context menu.
+Properties window consists of category tree on a left side and properties pane on a right side.
+For each category in tree is associated one properties pane.
+Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
 

### 2011-02-25 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 
-\jmodule{ProjectType} is the module responsible for creation of NBP project type which groups input/output files that belongs to one specific inference.
+\jmodule{ProjectType} is the module representing a jInfer project in the NB platform, which groups input/output files that belongs to one specific inference.
 Each jInfer project also allows user to set properties specific for inference.\\
@@
 \begin{figure}
+	\subfigure[Legend: Rectangle is folder, Rounded rectangle is file] {
 		\centering\includegraphics[scale=\myscale]{project_dir_structure}
+	}
 	\caption{Project directory structure} \label{dir-structure}
@@
 All the base classes are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype} package.
-	\item Visualization classes - These classes create tree structure of project in NBP Projects view and are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype.nodes}
+	\item Visualization classes - These classes create tree structure of project in NBP Projects view and are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype.nodes}.
 	\item Actions - Classes from \code{cz.cuni.mff.ksi.jinfer.projecttype.actions} package that provides actions allowing adding, removing input files into project, or running the project.
@@
 	        <file loc="/home/user/example.xml"/>
+	        <file loc="/home/user/example2.xml"/>
 	    </xml>
@@
 
-Each project in NBP is after creation opened in project window and has some visual tree structure where each node in this tree is represented by NBP \code{Node} class.
+Each jInfer project type in NBP is after creation opened in project window and has some visual tree structure where each node in this tree is represented by NBP \code{Node} class.
 This structure is in jInfer project type provided by \code{JInferLogicalView} class that implements \code{LogicalViewProvider} interface from NBP and is available through \code{JInferProject} lookup.
@@
 
+TODO what kind of validation did we use. How to extend validation.\\
+
 \code{ValidateAction} class implements action, which allows documents and schema files to be validated against each other.
@@
 
-Each jInfer project has associated properties window through which can user change project-wide properties.
+Each jInfer project type has associated properties window through which can user change project-wide properties.
 This window is accesible through \emph{properties} action in project \emph{root} node context menu.
@@
 For each category in tree is associated one properties pane.
-Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
+Category in a properties window is represented by NB platform \code{ProjectCustomizer.Category} class and properties pane is represented by \code{JPanel} class.
+Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.\\
+
+\code{JInferComponentProvider} is very simple class that extends \code{ProjectCustomizer.CategoryComponentProvider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
+In the \code{JInferCustomizerProvider} class, which extends NBP \code{CustomizerProvider} class, is the core functionality of collecting categories with associated properties panes from all jInfer modules.
+In section \ref{properties_reg} is described how to create your own category with associated properties pane, which will be present in jInfer project type properties window.\\
+
+\subsubsection{Custom project properties}\label{properties_reg}
 

### 2011-02-25 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 
-TODO prelozit
-Ak je nejaka genericka trieda parametrizovana typom \code{T}, tak pri instanciovani tejto triedy je nutne tento typ zadat rovnaky, akym je instanciovany automat, pre ktory sa tato trieda vyraba.
-JUNG triedy su vacsinou parametrizovane typom pre stav a hranu.
-V takom pripade sa tieto triedy instanciuju typmi \code{State\textless T\textgreater} pre stav a \code{Step\textless T\textgreater} pre hranu.
-Refer to \cite{archdoc} for more information on these classes and see \ref{subsubsection:StatePickingComponent} and \ref{subsubsection:StatesPickingComponent} for source code examples.
+In case a generic class has a type parameter \code{T}, this has to be the same \code{T} that is used to parametrize jInfer automata.
+JUNG classes usually need a type parameter for a state and for an edge.
+These should be then \code{State\textless T\textgreater} (for a state) and \code{Step\textless T\textgreater} (for edge).
 
@@
 Purpose of this class is to be extended to create own GUI panel, which displays some automaton using a supplied instance of \code{Visualizer}.
-Method \code{getAutomatonDrawPanel()} is meant to be overridden to returns an instance of \code{JPanel}, in which the \code{Visualizer} is to be drawn.\\
+Method \code{getAutomatonDrawPanel()} is meant to be overridden to return an instance of \code{JPanel}, in which the \code{Visualizer} is to be drawn.\\
 
@@
 \subsubsection{StatesPickingComponent} \label{subsubsection:StatesPickingComponent}
+\begin{figure}
+\centering\includegraphics[scale=0.6]{screenshot}
+\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
+\end{figure}
+
 Classes \code{StatePickingComponent} and \code{StatesPickingVisualizer} are similar to the classes described in the previous section.
@@
 
-\begin{figure}
-\centering\includegraphics[scale=0.6]{screenshot}
-\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
-\end{figure}
-
 As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the editor window.
-It consists of two buttons, horizontal line below them and and a panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}).
-Class representing this panel is named \code{AutoEditorTopComponent} and resides in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.topcomponent} package.\\
+It consists of two buttons, horizontal line below them and a panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}).
+Class representing this panel is called \code{AutoEditorTopComponent} and resides in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.topcomponent} package.\\
 

### 2011-02-26 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 
-With each project created in jInfer is also created specific directory structure on filesystem that describes jInfer project.
+With each project created in jInfer is also created specific directory structure on file-system that describes jInfer project.
 This structure is described in figure \ref{dir-structure}.
-Folder in filesystem is considered as jInfer project folder, if contains \emph{jinferproject} subfolder.
+Folder in file-system is considered as jInfer project folder, if contains \emph{jinferproject} sub-folder.
 This folder contains two files, first is \emph{project.properties}, where all properties set for particular project are saved.
 Second file is \emph{input.files} which is XML file filled with paths to files inserted as input into particular project (structure of this file is described in section \ref{input.files}).
-\emph{Output} subfolder contains schema files which were generated by jInfer inference.
+\emph{Output} sub-folder contains schema files which were generated by jInfer inference.
 
 \section{Structure}
-
+PropertiesPanelProvider
 Structure of \jmodule{ProjectType} can be divided into following five main parts.
@@
 \code{JInferProjectFactory} is a factory class implementing \code{ProjectFactory} interface provided by NBP.
-This class is responsible for loading/saving of jInfer project from/to filesystem and also determining if folder in filesystem is type of jInfer project (contains \emph{jinferproject} subfolder).
+This class is responsible for loading/saving of jInfer project from/to file-system and also determining if folder in file-system is type of jInfer project (contains \emph{jinferproject} sub-folder).
 For this purpose, class has three methods: \code{isProject()}, \code{save()} and \code{load()}.
-When jInfer project is loaded from filesystem, \code{JInferProjectFactory} creates new instance of \code{JInferProject} passing path to project folder as a constructor parameter.
-When \code{save()} method is invoced, all the project properties are saved in \emph{projectproperties} file and paths to input files are saved in \emph{input.files}.
+When jInfer project is loaded from file-system, \code{JInferProjectFactory} creates new instance of \code{JInferProject} passing path to project folder as a constructor parameter.
+When \code{save()} method is invoked, all the project properties are saved in \emph{projectproperties} file and paths to input files are saved in \emph{input.files}.
 
@@
 Each of the \emph{input} nodes is represented by \code{FolderNode} class and contains \emph{input file} nodes (represented by \code{FileNode} class) representing files added to the jInfer project.
-Notice that these \emph{input file} nodes represents files from filesystem and are not copied to project folder structure.
-Because of this, if you delete this nodes from project tree, files from filesystem are not deleted.\\
+Notice that these \emph{input file} nodes represents files from file-system and are not copied to project folder structure.
+Because of this, if you delete this nodes from project tree, files from file-system are not deleted.\\
 
-\code{OutputNode} class represents \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} subfolder in project filesystem hierarchy.
-In other words, files contained in this folder are shown as subnodes in \emph{Output} node in project tree.
-This subnodes are represented by \code{OutputFileNode} class.
-Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is also deleted from filesystem.
+\code{OutputNode} class represents \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} sub-folder in project file-system hierarchy.
+In other words, files contained in this folder are shown as sub-nodes in \emph{Output} node in project tree.
+This sub-nodes are represented by \code{OutputFileNode} class.
+Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is also deleted from file-system.
 
@@
 \code{FileAddAction} class implements action which opens file choose dialog and selected files are added into particular \emph{file input}node.
-\code{DeleteAllAction} class simply removes all input files previuosly added into particular \emph{file input} node.
+\code{DeleteAllAction} class simply removes all input files previously added into particular \emph{file input} node.
 
@@
 Each jInfer project type has associated properties window through which can user change project-wide properties.
-This window is accesible through \emph{properties} action in project \emph{root} node context menu.
+This window is accessible through \emph{properties} action in context menu of project \emph{root} node.
 Properties window consists of category tree on a left side and properties pane on a right side.
-For each category in tree is associated one properties pane.
+With each category in a tree is associated one properties pane.
+Structure of the category tree consists of two type of categories: actual category provided by some module and \emph{virtual} category.
+Actual category has associated properties pane, which is displayed in a right side of the properties window.
+\emph{Virtual} category displays as a properties pane special information panel that provides information about count and name of modules which are children of this virtual category.
+Each actual category could have virtual categories as a children or another actual categories, but not both at a same time.
+On the other side \emph{virtual} category can have only actual categories as a children.
+If a \emph{virtual} category has no child, this category is not displayed in a category tree.\\
+
 Category in a properties window is represented by NB platform \code{ProjectCustomizer.Category} class and properties pane is represented by \code{JPanel} class.
-Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.\\
+Classes responsible for creation of this properties window are\\ \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
+From Jinfer module point of view is properties category represented by \code{PropertiesPanelProvider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
+How to use this interface and abstract class to create custom project properties category presented in jInfer project type properties windowis described in section \ref{properties_reg}.\\
 
 \code{JInferComponentProvider} is very simple class that extends \code{ProjectCustomizer.CategoryComponentProvider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
-In the \code{JInferCustomizerProvider} class, which extends NBP \code{CustomizerProvider} class, is the core functionality of collecting categories with associated properties panes from all jInfer modules.
-In section \ref{properties_reg} is described how to create your own category with associated properties pane, which will be present in jInfer project type properties window.\\
+In the \code{JInferCustomizerProvider} class, which extends NBP \code{CustomizerProvider} class, is the core functionality of collecting categories (\code{PropertiesPanelProvider} interface) with associated properties panes (\code{AbstractPropertiesPanel} class) from all jInfer modules.\\
 
-\subsubsection{Custom project properties}\label{properties_reg}
+\subsubsection{Custom project properties category}\label{properties_reg}
+
+As was written before, each jInfer project properties category is represented in jInfer modules by \code{PropertiesPanelProvider} interface.
+To create new category, this interface must be implemented and implementation must be annotated by the following code.
+\begin{verbatim}
+	@ServiceProvider(service = PropertiesPanelProvider.class)
+\end{verbatim}
+
+\code{PropertiesPanelProvider} interface provides the following methods:
+\begin{itemize}
+	\item getName() - Gets a programmatic name of category
+	\item getDisplayName() - Gets a name of category displayed in properties window.
+	\item getPriority() - Gets a priority of category, by which is sorted among other sibling categories in a tree.
+	\item getPanel() - Gets \code{AbstractPropertiesPanel} representing the category properties pane.
+	\item getSubCategories() - Gets a \code{List} of \code{VirtualCategoryPanel} representing virtual category child.
+\end{itemize}
+
+\begin{figure}
+	\centering\includegraphics[scale=\myscale]{uml_properties}
+	\caption{Project directory structure} \label{uml-properties}
+\end{figure}
 

### 2011-02-27 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 
-\code{PropertiesPanelProvider} interface provides the following methods:
+\noindent\code{PropertiesPanelProvider} interface provides the following methods:
 \begin{itemize}
@@
 	\item getPriority() - Gets a priority of category, by which is sorted among other sibling categories in a tree.
-	\item getPanel() - Gets \code{AbstractPropertiesPanel} representing the category properties pane.
+	\item getPanel(Properties) - Gets \code{AbstractPropertiesPanel} representing the category properties pane.
+The only parameter is instance of \code{Properties} class to which are saved properties which were set in \emph{properties pane}.
+	\item getParent() - getter for a programmatic name of category which is parent of this category in a tree hierarchy.
 	\item getSubCategories() - Gets a \code{List} of \code{VirtualCategoryPanel} representing virtual category child.
@@
 
+\noindent\code{VirtualCategoryPanel} extends \code{JPanel} and its constructor has three parameters:
+ \begin{itemize}
+	\item \code{categoryId} - programmatic name of the category or ID.
+	\item \code{categoryName} - human readable name of the category displayed in jInfer properties window.
+	\item \code{modules} - \code{List<? extends NamedModule>} of modules that will be children of this virtual category.
+\end{itemize}
+
+To create properties pane associated to custom category, class that extends \code{AbstractPropertiesPanel} class must be implemented.
+As you can see from a class diagram in the Figure \ref{uml-properties}, this abstract class has constructor with \code{Properties} parameter and two abstract methods, \code{store()} and \code{load()}, which must be implemented.
+Constructor gets a \code{Properties} instance as a parameter and saves it to \code{protected} field \code{properties}.
+This \code{Properties} instance is associated with jInfer project type \emph{lookup} and is used to store properties that are set in this panel.\\
+
+Purpose of the method \code{store()} can be easily guessed from its name, it's used to store values that are set in \emph{properties pane} elements (text field, checkbox etc.) to the provided \emph{properties} field.
+On the other side, \code{load()} method is used to populate \emph{properties pane} elements with already saved values in the \emph{properties}.
+
 \begin{figure}
 	\centering\includegraphics[scale=\myscale]{uml_properties}
-	\caption{Project directory structure} \label{uml-properties}
+	\caption{Class diagram for the custom jInfer project properties category.} \label{uml-properties}
 \end{figure}

### 2011-02-27 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 
-First of the two actions, after it was invoked, open file choose dialog which has set file filter according to available classes that implement \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
-After selecting files in dialog, these files are inserted into particular \emph{input file} node according to mapping of file extension to folder type gained from \code{Processor} classes.\\
+First of the two actions, after it was invoked, file choose dialog is opened which has set file filter according to available classes that implement \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
+After selecting files in dialog, these files are inserted into particular \emph{input file} node according to mapping of file extension to folder type gained from \code{Processor} classes.
+If no mapping for some extension exists, this file is inserted into folder which is defined in \emph{Miscellaneous} properties category (section \ref{pref}).\\
 
@@
 
+\noindent Example of the implemented \code{PropertiesPanelProvider}.
+
+\begin{verbatim}
+@ServiceProvider(service = PropertiesPanelProvider.class)
+public class PropertiesPanelProviderImpl implements PropertiesPanelProvider{
+
+  @Override
+  public AbstractPropertiesPanel getPanel(final Properties properties) {
+    return new ExamplePropertiesPanel(properties);
+  }
+
+  @Override
+  public String getName() {
+    return "exampleCategoryID";
+  }
+
+  @Override
+  public String getDisplayName() {
+    return "Example category";
+  }
+
+  @Override
+  public int getPriority() {
+    return 1000;
+  }
+
+  @Override
+  public String getParent() {
+    return null;
+  }
+
+  @Override
+  public List<VirtualCategoryPanel> getSubCategories() {
+    final List<VirtualCategoryPanel> result = new ArrayList<VirtualCategoryPanel>();
+    result.add(new VirtualCategoryPanel("virtual_categoryID", "example virtual category",
+            ModuleSelectionHelper.lookupImpls(IGGenerator.class)));
+
+    return result;
+  }
+\end{verbatim}
+
 \noindent\code{VirtualCategoryPanel} extends \code{JPanel} and its constructor has three parameters:
@@
 
+\noindent Example of the implemented \code{AbstractPropertiesPanel}.
+
+\begin{verbatim}
+public class ProjectPropertiesPanel extends AbstractPropertiesPanel {
+
+  private javax.swing.JCheckBox checkBox;
+
+  public ProjectPropertiesPanel(final Properties properties) {
+    super(properties);
+    init();
+  }
+
+  public void init() {
+    ...
+    initialize checkBox
+    ...
+  }
+
+  @Override
+  public void store() {
+    properties.setProperty("some_property", Boolean.toString(checkBox.isSelected()));
+  }
+
+  @Override
+  public void load() {
+    checkBox.setSelected(Boolean.parseBoolean(properties.getProperty("some_property", "true")));
+  }
+}
+\end{verbatim}
+
 \begin{figure}
@@
 
-\subsection{Preferences}
+This section describes ability of jInfer to create new jInfer project through \emph{new project} wizard.
+This wizard contains two steps, where first is standard one with project name and location selection, and the second is the selection of Initial grammar module, Simplifier and Schema generator module which will be used in inference for newly created module.
+In the new project creation process can be already finished in a first step, in that case, default modules will be used for inference.
+After finishing of the wizard, new jInfer project type is created with typed name in selected location.
+It is important to notice that package \code{cz.cuni.mff.ksi.jinfer.projecttype} contains \code{JInferTemplateProject.zip} file, in which is default folder structure of the jInfer project type, and the content of a zip is copied into location defined in the wizard.
+For more information, how to create new project wizard,  please look at \cite{project_sample}.\\
+
+All classes responsible for creation of this wizard are in package \code{cz.cuni.mff.ksi.jinfer.projecttype.sample}.
+It contains \code{JinferTemplateWizardIterator} class which defines number of steps in the wizard and creates panel for each step.
+Each step is  represented by a class implementing \code{WizardDescriptor.Panel} interface, which creates steps panel and validate its data inserted by a user.
+Finally there is class extending \code{JPanel} in which are components of the first step.
+
+\subsection{Preferences}\label{pref}
+
+Also jInfer project type has its own properties category in a project properties window and is called \emph{Miscellaneous}. In this category,
+two properties can be set.
+First property is default \emph{input file} folder to which is inserted input file selected from \emph{FilesAddAction}, that has extension, which is not defined by any \code{Processor} registered in jInfer.
+Second property is global \emph{LOG} level, which defines the minimum level of messages that will be displayed in jInfer output window.
+If some module has its own \emph{LOG} level setting, global settings are overridden by these.
 

### 2011-02-28 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 
+\begin{figure}
+\centering\includegraphics[scale=\myscale]{uml_base_classes}
+\caption{Class diagram for the base classes.} \label{uml_base_classes}
+\end{figure}
+
 This section describes classes implementing basic common functionality that are supposed to be extended to create a new suitable visualization of automata for a particular method of inference.
@@
 
-\begin{figure}
-\centering\includegraphics[scale=\myscale]{uml_base_classes}
-\caption{Class diagram for the base classes.} \label{uml_base_classes}
-\end{figure}
-
 \subsubsection{AbstractComponent user interactivity support} \label{subsubsection:AbstractComponent user-interactivity support}
@@
 \begin{figure}
-\centering\includegraphics[scale=0.65]{uml_derived_classes}
-\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.} \label{uml_derived_classes}
+\centering\includegraphics[scale=\myscale]{uml_derived_classes}
+\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.
+Also not showing signatures of constructors of StatePickingVisualizer and StatePickingVisualizer for their length.} \label{uml_derived_classes}
 \end{figure}
@@
 
-\code{StatePickingVisualizer} is trivial extension of \code{PluggableVisualizer}.
-It has no additional methods.
-Its constructor has additional two arguments, superinitial and superfinal states as we want to distinguish these states in displayed automaton.\
-And upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \linebreak \code{PluggableVisualizer}.
+\code{StatePickingVisualizer} is trivial extension of \code{PluggableVisualizer}. It has no additional methods. Its constructor has several additional arguments. Their description follows.
+\begin{verbatim}
+public StatePickingVisualizer(
+    Layout<State<T>, Step<T>> layout,
+    Transformer<Step<T>, String> edgeLabelTransformer,
+    AbstractComponent<T> component,
+    State<T> superinitialState,
+    State<T> superfinalState)
+\end{verbatim}
+\begin{itemize}
+	\item \code{layout} - Instance of \code{Layout} which will be provided to the constructor of parent \code{Visualizer} class.
+	\item \code{edgeLabelTransformer} - Transformer to convert automaton edges to their string representations.
+	\item \code{component} - Instance of \code{AbstractComponent} to call \code{guiDone()} method on, when a state is picked.
+	\item \code{superinitialState} and \code{superfinalState} - Superinitial and superfinal states as we want to distinguish these states in displayed automaton.
+\end{itemize}
+
+Upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \linebreak \code{PluggableVisualizer}.
 Purpose of this plugin is to allow user to pick some state of automaton and then call \code{guiDone()} method on instance of \code{AbstractComponent}.
@@
 Main difference compared to the \code{VertexPickingGraphMousePlugin} is \code{VerticesPickingGraphMousePlugin} does not call component's \code{guiDone()} method.\\
+
+Signature of constructor and description of arguments follow.
+\begin{verbatim}
+public StatesPickingVisualizer(Layout<State<T>, Step<T>> layout, Transformer<Step<T>, String> edgeLabelTransformer)
+\end{verbatim}
+\begin{itemize}
+	\item \code{layout} - Instance of \code{Layout} which will be provided to the constructor of parent \code{Visualizer} class.
+	\item \code{edgeLabelTransformer} - Transformer to convert automaton edges to their string representations.
+\end{itemize}
 

### 2011-02-28 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 \jmodule{ProjectType} is the module representing a jInfer project in the NB platform, which groups input/output files that belongs to one specific inference.
-Each jInfer project also allows user to set properties specific for inference.\\
+Each jInfer project also allows user to set properties specific for a inference.\\
 
@@
 \section{Structure}
-PropertiesPanelProvider
+
 Structure of \jmodule{ProjectType} can be divided into following five main parts.
@@
 	\item Visualization classes - These classes create tree structure of project in NBP Projects view and are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype.nodes}.
-	\item Actions - Classes from \code{cz.cuni.mff.ksi.jinfer.projecttype.actions} package that provides actions allowing adding, removing input files into project, or running the project.
+	\item Actions - Classes from \code{cz.cuni.mff.ksi.jinfer.projecttype.actions} package that provides actions allowing adding input files into project, removing input files, or running the project.
 	\item Properties - Classes responsible for creation of project properties window with properties category tree. These are situated in  \code{cz.cuni.mff.ksi.jinfer.projecttype.properties} package.
@@
 This action is implemented in \code{RunAction} class and simply invokes \jmodule{Runner}s \code{run()} method, if no other inference is already running, otherwise information dialog pops up informing about already running inference.
-This check is done with \code{setActiveProject()} from \code{RunningProject} class.
+This check is done with \code{setActiveProject()} from \code{RunningProject} class (section \ref{running-project}).
 
@@
 
+\subsection{RunningProject utility class}\label{running-project}
+
+This section describes \code{RunningProject} utility class, which is not part of the \jmodule{ProjectType} module, but is closely related to it.
+This class can be found in a \code{cz.cuni.mff.ksi.jinfer.base.utils} package in the \jmodule{Base} module.
+The main purpose of this class is to allow all jInfer \jmodule{module}s to access \code{JInferProject} instance of the actual running inference.
+Notice that in jInfer application can be only one running inference at a time.\\
+
+\noindent \code{RunningProject} class implements these methods:
+\begin{itemize}
+	\item \code{setActiveProject()} - Setter of the \code{JInferProject} instance of a running project, which set the \code{JInferProject} instance if no other inference is running (no other \code{JInferProject} instance is already set).
+Returns \code{boolean} value, according as the setting was successful or not.
+	\item \code{removeActiveProject()} - Removes \code{JInferProject} instance if some inference is running.
+	\item \code{getActiveProject()} - Getter of the \code{JInferProject} instance of a running project.
+	\item \code{isActiveProject()} - Checks if some inference is running (the \code{JInferProject} instance is set).
+	\item \code{getActiveProjectProps()} - Gets \code{Properties} instance of the actual running inference (\code{Properties} instance is part of the \code{JInferProject} instace).
+	\item \code{getNextModuleCaps()} - Getter of a capability of the next module in the inference chain.
+	\item \code{setNextModuleCaps()} - Setter of a capability of the next module in the inference chain.
+\end{itemize}
+
 \nocite{*}

### 2011-02-28 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 These should be then \code{State\textless T\textgreater} (for a state) and \code{Step\textless T\textgreater} (for edge).
+Refer to \cite{archdoc} for more information on these classes and see \ref{subsubsection:StatePickingComponent} and \ref{subsubsection:StatesPickingComponent} for source code examples.
 
@@
 For information on system of modules, see \cite{archdoc}.
-To create a new implementation of \code{Layout} interface, it is necessary to create a new class implementing \code{LayoutFactory} interface (package \code{package cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.
+To create a new implementation of \code{Layout} interface, it is necessary to create a new class implementing \code{LayoutFactory} interface (package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.
 
@@
 
+\begin{figure}
+\centering\includegraphics[scale=\myscale]{uml_derived_classes}
+\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.
+Also not showing signatures of constructors of StatePickingVisualizer and StatePickingVisualizer for their excessive length.} \label{uml_derived_classes}
+\end{figure}
+
 \subsection{API} \label{subsection:API}
@@
 \subsection{Derived classes}
-
-\begin{figure}
-\centering\includegraphics[scale=\myscale]{uml_derived_classes}
-\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.
-Also not showing signatures of constructors of StatePickingVisualizer and StatePickingVisualizer for their length.} \label{uml_derived_classes}
-\end{figure}
 
@@
 \subsubsection{StatesPickingComponent} \label{subsubsection:StatesPickingComponent}
-\begin{figure}
-\centering\includegraphics[scale=0.6]{screenshot}
-\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
-\end{figure}
-
 Classes \code{StatePickingComponent} and \code{StatesPickingVisualizer} are similar to the classes described in the previous section.
@@
 \begin{verbatim}
-public StatesPickingVisualizer(Layout<State<T>, Step<T>> layout, Transformer<Step<T>, String> edgeLabelTransformer)
+public StatesPickingVisualizer(
+  Layout<State<T>, Step<T>> layout,
+  Transformer<Step<T>, String> edgeLabelTransformer)
 \end{verbatim}
@@
 The button is supposed to be clicked when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
-State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.
+State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.\\
 
@@
 \end{itemize}
+
+\begin{figure}
+\centering\includegraphics[scale=0.6]{screenshot}
+\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
+\end{figure}
 

### 2011-02-28 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor.

@@
 JUNG classes usually need a type parameter for a state and for an edge.
-These should be then \code{State\textless T\textgreater} (for a state) and \code{Step\textless T\textgreater} (for edge).
+These should be then \code{State<T>} (for a state) and \code{Step<T>} (for edge).
 Refer to \cite{archdoc} for more information on these classes and see \ref{subsubsection:StatePickingComponent} and \ref{subsubsection:StatesPickingComponent} for source code examples.
@@
 
-\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it inherits all its methods and adds support for saving contained automaton to an image file.
+\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG\linebreak\code{Visualization\-Viewer} class thus it inherits all its methods and adds support for saving contained automaton to an image file.
 Relevant methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
@@
 
-Constructor has one argument, an instance of \code{Layout} interface created from an automaton, typically by \linebreak \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).
+Constructor has one argument, an instance of \code{Layout} interface created from an automaton, typically by \code{Layout\-Helper\-Factory} (see \ref{subsubsection:LayoutHelperFactory}).
 
@@
 By default, instance of \code{PluggableVisualizer} is constructed with two plugins enabled.
-They are \linebreak \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
+They are \code{Scaling\-Graph\-Mouse\-Plugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
 In the most cases these plugins are useful, otherwise they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
@@
 \begin{itemize}
-	\item \code{setVisualizer()} - Setter of \code{Visualizer}.
+	\item \code{setVisualizer()} - Protected setter of \code{Visualizer}.
+It is defined protected because extensions of this class may want to declare this setter with argument type that extends \code{Visualizer} class.
 	\item \code{getVisualizer()} - Getter of \code{Visualizer}.
@@
 There is no need to called this method directly.
-	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by \code{guiInterrupt()} method or regularly (\code{guiDone()} method or GUI was not waiting for user interaction).
+	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by \code{guiInterrupt()} method or regularly\linebreak(\code{guiDone()} method or GUI was not waiting for user interaction).
 Again, there is no need to call this method directly, it is called by \jmodule{AutoEditor}.
@@
 
-\begin{figure}
-\centering\includegraphics[scale=\myscale]{uml_derived_classes}
-\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.
-Also not showing signatures of constructors of StatePickingVisualizer and StatePickingVisualizer for their excessive length.} \label{uml_derived_classes}
-\end{figure}
-
 \subsection{API} \label{subsection:API}
@@
   Transformer<Step<T>, String> edgeLabelTransformer,
-  AbstractComponent<T> component,
+  StatePickingComponent<T> component,
   State<T> superinitialState,
@@
 	\item \code{edgeLabelTransformer} - Transformer to convert automaton edges to their string representations.
-	\item \code{component} - Instance of \code{AbstractComponent} to call \code{guiDone()} method on, when a state is picked.
+	\item \code{component} - Instance of \code{StatePickingComponent} to call \code{guiDone()} method on, when a state is picked.
 	\item \code{superinitialState} and \code{superfinalState} - Superinitial and superfinal states as we want to distinguish these states in displayed automaton.
@@
 
-Upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \linebreak \code{PluggableVisualizer}.
+Upon construction, it just adds \code{VertexPickingGraphMousePlugin} to the plugins of \code{Pluggable\-Visualizer}.
 Purpose of this plugin is to allow user to pick some state of automaton and then call \code{guiDone()} method on instance of \code{AbstractComponent}.
@@
 \begin{itemize}
+	\item \code{setVisualizer()} - Setter for \code{Visualizer} restricted to accept instances of \code{StatePickingVisualizer} class to prevent misuse.
 	\item \code{getPickedState()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see \ref{subsection:API}), this method retrieves the user picked automaton state.
@@
 
-Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.}\linebreak\code{component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and\linebreak \code{VertexPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.}\linebreak\code{graphmouseplugins}.\\
+Source codes of \code{State\.Picking\.Component} resides in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.component}, \code{State\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertex\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
 
@@
 
+\begin{figure}
+\centering\includegraphics[scale=0.7]{uml_derived_classes}
+\caption{Class diagram for the derived classes, not showing members of classes from figure \ref{uml_base_classes} to keep this diagram simple.
+Also not showing signatures of constructors of StatePickingVisualizer and StatePickingVisualizer for their excessive length.} \label{uml_derived_classes}
+\end{figure}
+
 \subsubsection{StatesPickingComponent} \label{subsubsection:StatesPickingComponent}
@@
 
-\code{StatesPickingVisualizer} is extension of \code{PluggableVisualizer} and upon its construction it adds \linebreak \code{VerticesPickingGraphMousePlugin} to the graph mouse plugins.
+\code{StatesPickingVisualizer} is extension of \code{PluggableVisualizer} and upon its construction it adds \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} to the graph mouse plugins.
 With this plugin used, user can pick and unpick automaton states separately or pick several states at once by dragging rectangular selection box over states.\
@@
 
+\begin{figure}
+\centering\includegraphics[scale=0.6]{screenshot}
+\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
+\end{figure}
+
 Methods of \code{StatesPickingComponent} are the following.
 \begin{itemize}
+	\item \code{setVisualizer()} - Setter for \code{Visualizer} restricted to accept instances of \code{StatesPickingVisualizer} class to prevent misuse.
 	\item \code{shallAskUser()} - Retrieves state of the checkbox.
@@
 
-\begin{figure}
-\centering\includegraphics[scale=0.6]{screenshot}
-\caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
-\end{figure}
-
-Source codes of \code{StatesPickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.}\linebreak\code{component}, \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and\linebreak \code{VerticesPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.}\linebreak\code{graphmouseplugins}.\\
+Source codes of \code{StatesPickingComponent} resides in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.component}, \code{States\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
 

### 2011-02-28 doc/jInfer/projecttype/projecttype.tex
#3187361 projecttype doc

@@
 When jInfer project is loaded from file-system, \code{JInferProjectFactory} creates new instance of \code{JInferProject} passing path to project folder as a constructor parameter.
-When \code{save()} method is invoked, all the project properties are saved in \emph{projectproperties} file and paths to input files are saved in \emph{input.files}.
+When \code{save()} method is invoked, all the project properties are saved in \emph{project.properties} file and paths to input files are saved into \emph{input.files}.
 
@@
 
-\code{InputFiles} is utility class that stores/loads input file paths into/from input.files XML file.
+\code{InputFiles} is utility class that stores/loads input file paths into/from \emph{input.files} XML file.
 For this purpose, class uses \emph{JAXB} \cite{JAXB} architecture to unmarshall/marshall XML file into/from java content objects.
@@
 Each of the \emph{input} nodes is represented by \code{FolderNode} class and contains \emph{input file} nodes (represented by \code{FileNode} class) representing files added to the jInfer project.
-Notice that these \emph{input file} nodes represents files from file-system and are not copied to project folder structure.
+Notice that these \emph{input file} nodes represents files from file-system and are not copied to the project folder structure.
 Because of this, if you delete this nodes from project tree, files from file-system are not deleted.\\
@@
 
-First of the two actions, after it was invoked, file choose dialog is opened which has set file filter according to available classes that implement \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
-After selecting files in dialog, these files are inserted into particular \emph{input file} node according to mapping of file extension to folder type gained from \code{Processor} classes.
+First of the two actions, after its invocation, file choose dialog is opened which has set file filter according to available classes that implement \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
+After selecting files in dialog, these files are inserted into particular \emph{input folder} node according to mapping of file extension to folder type gained from \code{Processor} classes.
 If no mapping for some extension exists, this file is inserted into folder which is defined in \emph{Miscellaneous} properties category (section \ref{pref}).\\
@@
 Run action, which is available in root node context menu, is also present in NBP toolbar as a \emph{Run} button and in jInfer Welcome window.
-This action is implemented in \code{RunAction} class and simply invokes \jmodule{Runner}s \code{run()} method, if no other inference is already running, otherwise information dialog pops up informing about already running inference.
+This action is implemented in \code{RunAction} class and simply invokes \jmodule{Runner}s \code{run()} method, if no other inference is already running, otherwise information dialog pops up, informing about already running inference.
 This check is done with \code{setActiveProject()} from \code{RunningProject} class (section \ref{running-project}).
@@
 To this group of actions belong two actions: action for adding input files into \emph{input folder} node (\code{FileAddAction} class) and action for deleting all input files in \emph{input folder} node (\code{DeleteAllAction} class).
-\code{FileAddAction} class implements action which opens file choose dialog and selected files are added into particular \emph{file input}node.
-\code{DeleteAllAction} class simply removes all input files previously added into particular \emph{file input} node.
+\code{FileAddAction} class implements action which opens file choose dialog and selected files are added into particular \emph{input folder} node.
+\code{DeleteAllAction} class simply removes all input files previously added into particular \emph{input folder} node.
 
 \subsubsection{ValidateAction class}\label{validate-action}
-
-TODO what kind of validation did we use. How to extend validation.\\
 
@@
 Action is enabled only if exactly one schema file is selected and one or more XML files.
-If the selected XML files are valid against selected schema, information dialog is popped up, otherwise error dialog is popped up and in \emph{validate} output window is printed error message describing cause of non-validity XML files against schema.
+If the selected XML files are valid against selected schema, information dialog is popped up, otherwise error dialog is popped up and in \emph{validate} output window is printed error message describing cause of non-validity XML files against schema.\\
+
+For the validation against XML Schema, we use standard Java validation API \cite{xml_validate}.
+Besides XML Schema, this API allows to  validate XML also against \emph{Relax NG} and \emph{Schematron} schema.
+If you want to extend the functionality of one of these schemas, you need to set \code{schemaLanguage} local variable to appropriate \code{XMLConstant}.
+If you need to extend validation to some other type of schema, you can use dtd validation as a sample.\\
+
+For a validation error handling is used \code{JInferErrorHander} class which extends \code{DefaultHandler} class and overrides \code{error()} and \code{fatalError()} methods from this class.
+In these methods is created error message, which is consequently printed in validation \emph{output} window.
 
@@
 From Jinfer module point of view is properties category represented by \code{PropertiesPanelProvider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
-How to use this interface and abstract class to create custom project properties category presented in jInfer project type properties windowis described in section \ref{properties_reg}.\\
+How to use this interface and abstract class to create custom project properties category presented in jInfer project type properties window is described in section \ref{properties_reg}.\\
 
@@
 \begin{itemize}
-	\item getName() - Gets a programmatic name of category
+	\item getName() - Gets a programmatic name of category.
 	\item getDisplayName() - Gets a name of category displayed in properties window.
 	\item getPriority() - Gets a priority of category, by which is sorted among other sibling categories in a tree.
-	\item getPanel(Properties) - Gets \code{AbstractPropertiesPanel} representing the category properties pane.
+	\item getPanel(Properties) - Gets the \code{AbstractPropertiesPanel} instance representing the category \emph{properties pane}.
 The only parameter is instance of \code{Properties} class to which are saved properties which were set in \emph{properties pane}.
-	\item getParent() - getter for a programmatic name of category which is parent of this category in a tree hierarchy.
-	\item getSubCategories() - Gets a \code{List} of \code{VirtualCategoryPanel} representing virtual category child.
+	\item getParent() - Getter for a programmatic name of category which is parent of this category in a tree hierarchy.
+	\item getSubCategories() - Gets a \code{List} of \code{VirtualCategoryPanel} instances representing virtual category children.
 \end{itemize}
@@
 	\item \code{categoryName} - human readable name of the category displayed in jInfer properties window.
-	\item \code{modules} - \code{List<? extends NamedModule>} of modules that will be children of this virtual category.
+	\item \code{modules} - \code{List<?\ extends NamedModule>} of modules that will be children of this virtual category.
 \end{itemize}
@@
 \begin{verbatim}
-public class ProjectPropertiesPanel extends AbstractPropertiesPanel {
+public class ExamplePropertiesPanel extends AbstractPropertiesPanel {
 
@@
 
-  public ProjectPropertiesPanel(final Properties properties) {
+  public ExamplePropertiesPanel(final Properties properties) {
     super(properties);
@@
 This wizard contains two steps, where first is standard one with project name and location selection, and the second is the selection of Initial grammar module, Simplifier and Schema generator module which will be used in inference for newly created module.
-In the new project creation process can be already finished in a first step, in that case, default modules will be used for inference.
+New project creation process can be already finished in a first step, in that case, default modules will be used for inference.
 After finishing of the wizard, new jInfer project type is created with typed name in selected location.
@@
 
-Also jInfer project type has its own properties category in a project properties window and is called \emph{Miscellaneous}. In this category,
-two properties can be set.
-First property is default \emph{input file} folder to which is inserted input file selected from \emph{FilesAddAction}, that has extension, which is not defined by any \code{Processor} registered in jInfer.
+Also jInfer project type has its own properties category in a project properties window called \emph{Miscellaneous}, and it is implemented by \code{ProjectPropertiesPanelProvider} and \code{ProjectPropertiesPanel} classes.
+In this category, two properties can be set.
+First property is default \emph{input file} folder to which is inserted input file selected from \emph{FilesAddAction}, that has extension which is not defined by any \code{Processor} registered in jInfer.
 Second property is global \emph{LOG} level, which defines the minimum level of messages that will be displayed in jInfer output window.

### 2011-03-01 doc/jInfer/projecttype/projecttype.tex
Documentation: ProjectType module: review round 1 part 1.

@@
 
-\jmodule{ProjectType} is the module representing a jInfer project in the NB platform, which groups input/output files that belongs to one specific inference.
-Each jInfer project also allows user to set properties specific for a inference.\\
+\jmodule{ProjectType} is the module representing a jInfer project in the NB platform, grouping input/output files that belong to one specific inference.
+Each jInfer project also allows the user to set properties specific for this inference.\\
 
@@
 
-With each project created in jInfer is also created specific directory structure on file-system that describes jInfer project.
+For each jInfer project a specific directory structure in the filesystem describing this project is created.
 This structure is described in figure \ref{dir-structure}.
-Folder in file-system is considered as jInfer project folder, if contains \emph{jinferproject} sub-folder.
-This folder contains two files, first is \emph{project.properties}, where all properties set for particular project are saved.
-Second file is \emph{input.files} which is XML file filled with paths to files inserted as input into particular project (structure of this file is described in section \ref{input.files}).
-\emph{Output} sub-folder contains schema files which were generated by jInfer inference.
+A folder in the filesystem is considered to be jInfer project folder iff it contains a \emph{jinferproject} subfolder.
+This folder contains two files: \emph{project.properties}, where all properties set for that particular project are stored.
+The second file is \emph{input.files} which is an XML file filled with paths to files selected particular project input (structure of this file is described in section \ref{input.files}).
+\emph{output} subfolder contains schema files which were generated by jInfer inference runs.
 
@@
 
-Structure of \jmodule{ProjectType} can be divided into following five main parts.
+Structure of \jmodule{ProjectType} can be divided into the following five main parts.
 \begin{itemize}
-	\item Base classes - Classes providing main functionality like creation of project, defining operations like move, delete, copy etc.
-All the base classes are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype} package.
-	\item Visualization classes - These classes create tree structure of project in NBP Projects view and are contained in the \code{cz.cuni.mff.ksi.jinfer.projecttype.nodes}.
-	\item Actions - Classes from \code{cz.cuni.mff.ksi.jinfer.projecttype.actions} package that provides actions allowing adding input files into project, removing input files, or running the project.
-	\item Properties - Classes responsible for creation of project properties window with properties category tree. These are situated in  \code{cz.cuni.mff.ksi.jinfer.projecttype.properties} package.
-	\item Project wizard - Classes representing creation of project through new project wizard from  \code{cz.cuni.mff.ksi.jinfer.projecttype.sample} package.
+	\item Base classes - classes providing core functionality like project creation and definitions of operations like move, delete, copy etc.
+All the base classes are contained in the \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype} package.
+	\item Visualization classes - these classes create tree structure of jInfer project in NBP Projects view and are contained in the \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.nodes}.
+	\item Actions - classes from \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.actions} package that provide actions allowing adding input files into the project, removing input files, or running the inference on the project.
+	\item Properties - classes responsible for creation of project properties window with properties category tree.
+These are located in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.properties} package.
+	\item Project wizard - classes representing new project creation through a project wizard from \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.sample} package.
 \end{itemize}
@@
 
-This section describes classes needed for creation of the jInfer project and correct integration into NBP.
-Main two classes representing jInfer project type are \code{JInferProjectFactory} and \code{JInferProject}.
+This section describes classes needed for jInfer project creation and its integration into the NBP.
+The two main classes representing a jInfer project type are \code{JInferProjectFactory} and \code{JInferProject}.
 
@@
 \code{JInferProjectFactory} is a factory class implementing \code{ProjectFactory} interface provided by NBP.
-This class is responsible for loading/saving of jInfer project from/to file-system and also determining if folder in file-system is type of jInfer project (contains \emph{jinferproject} sub-folder).
+This class is responsible for loading/saving the jInfer project from/to the filesystem and also determining if a folder in filesystem is of jInfer project type (= contains \emph{jinferproject} subfolder).
 For this purpose, class has three methods: \code{isProject()}, \code{save()} and \code{load()}.
-When jInfer project is loaded from file-system, \code{JInferProjectFactory} creates new instance of \code{JInferProject} passing path to project folder as a constructor parameter.
+When jInfer project is loaded from the filesystem, \code{JInferProjectFactory} creates new instance of \code{JInferProject} passing path to project folder as a constructor parameter.
 When \code{save()} method is invoked, all the project properties are saved in \emph{project.properties} file and paths to input files are saved into \emph{input.files}.
@@
 
-\code{JinferProject} class implements \code{Project} interface from NBP and represents in-memory representation of jInfer project.
-Inner structure of this class is very simple, interface provides only two methods: \code{getProjectDirectory()} and \code{getLookup()}.
-All the extensibility of project type is done by \emph{Lookup} functionality of NBP.
-Project lookup contains besides properties and \code{Input} class, which encapsulates input files, also class that creates output files, class that builds project tree in project window, etc.
+\code{JinferProject} class implements the \code{Project} interface from NBP and represents an in-memory representation of a jInfer project.
+Inner structure of this class is very simple, interface specifies only two methods: \code{getProjectDirectory()} and \code{getLookup()}.
+All the extensibility of a project type is done by the \emph{Lookup} functionality of NBP.
+Project lookup contains properties, \code{Input} class (encapsulating input files), class creating output files, class building project tree in the Projects view, etc.
 
-\subsubsection{InputFiles class}\label{input.files}
+\subsubsection{InputFiles}\label{input.files}
 
-\code{InputFiles} is utility class that stores/loads input file paths into/from \emph{input.files} XML file.
-For this purpose, class uses \emph{JAXB} \cite{JAXB} architecture to unmarshall/marshall XML file into/from java content objects.
-XML structure of the input.files is very simple, \emph{jinferinput} is root element under which are \emph{documents}, \emph{schemas} and \emph{queries} elements.
-Each of this elements could have none or more \emph{file} elements with string attribute \emph{loc} that contains absolute path to input file.\\
+\code{InputFiles} is a utility class that stores/loads input file paths into/from the \emph{input.files} XML file.
+For this purpose, \code{InputFiles} uses \emph{JAXB} \cite{JAXB} framework to unmarshall/marshall XML file into/from Java content objects.
+XML structure of \emph{input.files} is very simple, \emph{jinferinput} is the root element under which are \emph{documents}, \emph{schemas} and \emph{queries} elements.
+Each of this elements may have none or more \emph{file} elements with string attribute \emph{loc} that contains the absolute path to the input file.\\
 
-\noindent Example of the structure follows.
+\noindent Example of \emph{input.files} follows.
 \begin{verbatim}
@@
 	    <xml>
-	        <file loc="/home/user/example.xml"/>
+	        <file loc="/home/user/example1.xml"/>
 	        <file loc="/home/user/example2.xml"/>
@@
 Each of the \emph{input} nodes is represented by \code{FolderNode} class and contains \emph{input file} nodes (represented by \code{FileNode} class) representing files added to the jInfer project.
-Notice that these \emph{input file} nodes represents files from file-system and are not copied to the project folder structure.
-Because of this, if you delete this nodes from project tree, files from file-system are not deleted.\\
+Notice that these \emph{input file} nodes represents files from filesystem and are not copied to the project folder structure.
+Because of this, if you delete this nodes from project tree, files from filesystem are not deleted.\\
 
-\code{OutputNode} class represents \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} sub-folder in project file-system hierarchy.
+\code{OutputNode} class represents \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} subfolder in project filesystem hierarchy.
 In other words, files contained in this folder are shown as sub-nodes in \emph{Output} node in project tree.
 This sub-nodes are represented by \code{OutputFileNode} class.
-Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is also deleted from file-system.
+Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is also deleted from filesystem.
 

### 2011-03-01 doc/jInfer/projecttype/projecttype.tex
Documentation: ProjectType module: review round 1 part 2.

@@
 	\centering\includegraphics[scale=\myscale]{node_structure}
-	\caption{Project directory structure} \label{node-structure}
+	\caption{Project nodes structure} \label{node-structure}
 \end{figure}
 
-Each jInfer project type in NBP is after creation opened in project window and has some visual tree structure where each node in this tree is represented by NBP \code{Node} class.
-This structure is in jInfer project type provided by \code{JInferLogicalView} class that implements \code{LogicalViewProvider} interface from NBP and is available through \code{JInferProject} lookup.
-This class has one important method, \code{createLogicalView()} that returns instance of \code{RootNode} class of root node.
+Each jInfer project in NBP is open after its creation in the Projects view and has tree structure where each node is represented by a NBP \code{Node} class.
+This structure in jInfer project type is provided by the \code{JInferLogicalView} class implementing \code{LogicalViewProvider} interface from NBP and is available through the \code{JInferProject} lookup.
+This class has one important method: \code{createLogicalView()} returning an instance of \code{RootNode} class.
 Simple class structure is described in figure \ref{node-structure}.\\
 
-\code{RootNode} extends NBP \code{AbstractNode} class which is basic implementation of NBP \code{Node}.
-This node contains four children nodes: three for input files (Document, Schema and Query) and one for output files generated by inference (Output).
-Root node also offers actions for adding input files, run inference and standard project actions like delete or move/rename project.\\
+\code{RootNode} extends NBP \code{AbstractNode} class which is a basic implementation of NBP \code{Node}.
+This node contains four children nodes: three for input files (Document, Schema and Query) and one for output files generated by the inference (Output).
+Root node also offers actions for adding input files, inference running and standard project actions like delete or move/rename project.\\
 
-Each of the \emph{input} nodes is represented by \code{FolderNode} class and contains \emph{input file} nodes (represented by \code{FileNode} class) representing files added to the jInfer project.
-Notice that these \emph{input file} nodes represents files from filesystem and are not copied to the project folder structure.
-Because of this, if you delete this nodes from project tree, files from filesystem are not deleted.\\
+Each of the \emph{input} nodes is represented by a \code{FolderNode} class and contains \emph{input file} nodes (represented by a \code{FileNode} class) representing files added to the jInfer project.
+Notice that these \emph{input file} nodes represent existing files in the filesystem and are not copied to the project folder structure.
+Because of this, if you delete these nodes from project tree, files in the filesystem are not affected.\\
 
-\code{OutputNode} class represents \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} subfolder in project filesystem hierarchy.
-In other words, files contained in this folder are shown as sub-nodes in \emph{Output} node in project tree.
-This sub-nodes are represented by \code{OutputFileNode} class.
-Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is also deleted from filesystem.
+\code{OutputNode} class represents an \emph{Output} node in jInfer project type tree and is slightly different from \emph{input} nodes, because it refers to \emph{output} subfolder in project filesystem hierarchy.
+In other words, files contained in this folder are shown as subnodes in \emph{Output} node in project tree.
+These subnodes are represented by \code{OutputFileNode} class.
+Unlike \emph{input file} nodes, if you delete \emph{output file} node under the \emph{Output} node, file represented by this node is deleted from filesystem too.
 
@@
 
-Each node in project tree hierarchy contains set of actions available from its context menu.
-Besides standard actions for root node or file node, jInfer project type introduce its own actions.
-These actions could be separated into three groups: actions for \emph{root} node, actions for \emph{input folder} nodes and one action for \emph{input/output file} nodes implemented by \code{ValidateAction} class.
-All classes implementing some jInfer project action extend standard java \code{javax.swing.AbstractAction} class besides \code{ValidateAction} class which is described in \ref{validate-action}.\\
+Each node in jInfer project tree hierarchy contains a set of actions available from its context menu.
+Besides standard NBP actions for root node or file node, jInfer project type introduces its own actions.
+These actions can be divided into three groups: actions for \emph{root} node, actions for \emph{input folder} nodes and one action for \emph{input/output file} nodes implemented by \code{ValidateAction} class.
+All classes implementing a jInfer project action extend standard java \code{javax.\.swing.\.AbstractAction} class besides \code{ValidateAction} class which is described in \ref{validate-action}.\\
 
@@
 
-This group contains two actions, first is action that allows adding input files into \emph{input folder} nodes (implemented by \code{FilesAddAction} class).
-Second one is slightly different, because unlike from the previous action, this one doesn't create new item in \emph{root} node context menu but implements functionality for item that already exists in context menu - Run action.\\
+This group contains two actions: an action that allows adding input files into \emph{input folder} nodes (implemented by \code{FilesAddAction} class).
+Second one is slightly different, because it implements functionality for an item that already exists in context menu - Run action.\\
 
-First of the two actions, after its invocation, file choose dialog is opened which has set file filter according to available classes that implement \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
-After selecting files in dialog, these files are inserted into particular \emph{input folder} node according to mapping of file extension to folder type gained from \code{Processor} classes.
-If no mapping for some extension exists, this file is inserted into folder which is defined in \emph{Miscellaneous} properties category (section \ref{pref}).\\
+After invoking the first of the two actions, a file chooser dialog is open.
+Its file filter is set according to available classes implementing \code{Processor} interface registered in jInfer with \code{ServiceProvider}.
+After selecting file(s) in dialog, these files are inserted into the particular \emph{input folder} node according to mapping of their file extensions to folder type retrieved from \code{Processor} classes.
+If no mapping for some extension exists, this file is inserted into a default folder defined in \emph{Miscellaneous} properties category (section \ref{pref}).\\
 
-Run action, which is available in root node context menu, is also present in NBP toolbar as a \emph{Run} button and in jInfer Welcome window.
-This action is implemented in \code{RunAction} class and simply invokes \jmodule{Runner}s \code{run()} method, if no other inference is already running, otherwise information dialog pops up, informing about already running inference.
-This check is done with \code{setActiveProject()} from \code{RunningProject} class (section \ref{running-project}).
+Run action, available from root node context menu, is also present in NBP toolbar as a \emph{Run} button and in jInfer Welcome window.
+This action is implemented in \code{RunAction} class and simply invokes \jmodule{Runner}'s \code{run()} method, if no other inference is already running.
+Otherwise an information dialog pops up, informing about the already running inference.
+This check is done invoking the \code{setActiveProject()} from \code{RunningProject} class (section \ref{running-project}).
 
@@
 
-To this group of actions belong two actions: action for adding input files into \emph{input folder} node (\code{FileAddAction} class) and action for deleting all input files in \emph{input folder} node (\code{DeleteAllAction} class).
-\code{FileAddAction} class implements action which opens file choose dialog and selected files are added into particular \emph{input folder} node.
-\code{DeleteAllAction} class simply removes all input files previously added into particular \emph{input folder} node.
+Two actions belong to this group: action for adding input files into \emph{input folder} node (\code{FileAddAction} class) and action for deleting all input files in the \emph{input folder} node (\code{DeleteAllAction} class).
+\code{FileAddAction} class implements action which opens a file chooser dialog and adds the selected files into the particular \emph{input folder} node.
+\code{DeleteAllAction} class simply removes all input files previously added into that particular \emph{input folder} node.
 
-\subsubsection{ValidateAction class}\label{validate-action}
+\subsubsection{ValidateAction}\label{validate-action}
 
-\code{ValidateAction} class implements action, which allows documents and schema files to be validated against each other.
-Because of this, action is presented in context menu of document and schema files (in actual implementation is available for XML, dtd and XML schema files).
-Unlike the other action classes, this extends NBP \code{NodeAction} class, which allows to enable/disable action in context menu according to node selection.
-Action is enabled only if exactly one schema file is selected and one or more XML files.
-If the selected XML files are valid against selected schema, information dialog is popped up, otherwise error dialog is popped up and in \emph{validate} output window is printed error message describing cause of non-validity XML files against schema.\\
+\code{ValidateAction} class implements action that allows documents and schema files to be validated against each other.
+Because of this, action is presented in context menu of document and schema files (its actual implementation is available for XML, DTD and XML Schema files).
+Unlike the other action classes, this extends NBP \code{NodeAction} class, which allows to enable/disable the action in context menu according to node selection.
+Action is enabled only if exactly one schema file and one or more XML files are selected at the same time.
+If the selected XML files are valid against selected schema, information dialog pops up, otherwise error dialog pops up and an error message describing the cause of non-validity of XML files against this schema is printed in \emph{validate} output window.\\
 
@@
 If you want to extend the functionality of one of these schemas, you need to set \code{schemaLanguage} local variable to appropriate \code{XMLConstant}.
-If you need to extend validation to some other type of schema, you can use dtd validation as a sample.\\
+If you need to extend validation to some other type of schema, you can use DTD validation as an example.\\
 
-For a validation error handling is used \code{JInferErrorHander} class which extends \code{DefaultHandler} class and overrides \code{error()} and \code{fatalError()} methods from this class.
-In these methods is created error message, which is consequently printed in validation \emph{output} window.
+For validation error handling the \code{JInferErrorHander} class is used.
+It extends \code{DefaultHandler} class and overrides its \code{error()} and \code{fatalError()} methods.
+In these methods an error message is created, which is consequently printed in validation \emph{output} window.
 

### 2011-03-01 doc/jInfer/projecttype/projecttype.tex
Documentation: ProjectType module: review round 1 part 3.

@@
 
-Each jInfer project type has associated properties window through which can user change project-wide properties.
-This window is accessible through \emph{properties} action in context menu of project \emph{root} node.
-Properties window consists of category tree on a left side and properties pane on a right side.
-With each category in a tree is associated one properties pane.
-Structure of the category tree consists of two type of categories: actual category provided by some module and \emph{virtual} category.
-Actual category has associated properties pane, which is displayed in a right side of the properties window.
-\emph{Virtual} category displays as a properties pane special information panel that provides information about count and name of modules which are children of this virtual category.
-Each actual category could have virtual categories as a children or another actual categories, but not both at a same time.
-On the other side \emph{virtual} category can have only actual categories as a children.
-If a \emph{virtual} category has no child, this category is not displayed in a category tree.\\
+Each jInfer project has its associated properties window through which the user can change project-wide properties.
+This window is accessible through \emph{Properties} action in context menu of the project \emph{root} node.
+Properties window consists of category tree on the left side and properties pane on the right side.
+With each category in the tree, one properties pane is associated.
+Structure of the category tree consists of two types of categories: actual categories provided by some modules and \emph{virtual} categories.
+An actual category has its associated properties pane, which is displayed in the right side of the properties window.
+\emph{Virtual} category is displayed as a special information panel in the properties pane that provides information about count and names of modules which are children of this virtual category.
+Each actual category can have either virtual or actual categories as its children, but not both at a same time.
+On the other hand, a \emph{virtual} category can have only actual categories as its children.
+If a \emph{virtual} category has no children, this category is not displayed in the category tree.\\
 
-Category in a properties window is represented by NB platform \code{ProjectCustomizer.Category} class and properties pane is represented by \code{JPanel} class.
-Classes responsible for creation of this properties window are\\ \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
-From Jinfer module point of view is properties category represented by \code{PropertiesPanelProvider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
-How to use this interface and abstract class to create custom project properties category presented in jInfer project type properties window is described in section \ref{properties_reg}.\\
+A category in the properties window is represented by NBP \code{ProjectCustomizer.Category} class, properties pane is represented by \code{JPanel} class.
+Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
+From jInfer module point of view the properties category is represented by \code{PropertiesPanelProvider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
+How to use this interface and abstract class to create custom project properties category presented in jInfer project properties window is described in section \ref{properties_reg}.\\
 
-\code{JInferComponentProvider} is very simple class that extends \code{ProjectCustomizer.CategoryComponentProvider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
-In the \code{JInferCustomizerProvider} class, which extends NBP \code{CustomizerProvider} class, is the core functionality of collecting categories (\code{PropertiesPanelProvider} interface) with associated properties panes (\code{AbstractPropertiesPanel} class) from all jInfer modules.\\
+\code{JInferComponentProvider} is a very simple class that extends \code{ProjectCustomizer.CategoryComponentProvider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
+In the \code{JInferCustomizerProvider} class (which extends NBP \code{CustomizerProvider} class) is the core functionality of collecting categories (\code{PropertiesPanelProvider} interface) with associated properties panes (\code{AbstractPropertiesPanel} class) from all jInfer modules.\\
 
@@
 \begin{itemize}
-	\item getName() - Gets a programmatic name of category.
-	\item getDisplayName() - Gets a name of category displayed in properties window.
-	\item getPriority() - Gets a priority of category, by which is sorted among other sibling categories in a tree.
-	\item getPanel(Properties) - Gets the \code{AbstractPropertiesPanel} instance representing the category \emph{properties pane}.
-The only parameter is instance of \code{Properties} class to which are saved properties which were set in \emph{properties pane}.
-	\item getParent() - Getter for a programmatic name of category which is parent of this category in a tree hierarchy.
-	\item getSubCategories() - Gets a \code{List} of \code{VirtualCategoryPanel} instances representing virtual category children.
+	\item getName() - gets a programmatic name of this category.
+	\item getDisplayName() - gets a name of this category which will be displayed in properties window.
+	\item getPriority() - gets priority of this category, by which it is sorted among other sibling categories in the tree.
+	\item getPanel(Properties) - gets the \code{AbstractPropertiesPanel} instance representing the category \emph{properties pane}.
+The only parameter is an instance of \code{Properties} class holding stored properties for the \emph{properties pane}.
+	\item getParent() - getter for a programmatic name of category which is the parent of this category in the tree hierarchy.
+	\item getSubCategories() - gets a \code{List} of \code{VirtualCategoryPanel} instances representing virtual category children.
 \end{itemize}
@@
 
-\noindent\code{VirtualCategoryPanel} extends \code{JPanel} and its constructor has three parameters:
+\noindent\code{VirtualCategoryPanel} extends \code{JPanel}, its constructor has three parameters:
  \begin{itemize}
@@
 
-To create properties pane associated to custom category, class that extends \code{AbstractPropertiesPanel} class must be implemented.
-As you can see from a class diagram in the Figure \ref{uml-properties}, this abstract class has constructor with \code{Properties} parameter and two abstract methods, \code{store()} and \code{load()}, which must be implemented.
+To create properties pane associated with a custom category, class extending \code{AbstractPropertiesPanel} class must be implemented.
+As you can see from the class diagram in the Figure \ref{uml-properties}, this abstract class has a constructor with \code{Properties} parameter and two abstract methods, \code{store()} and \code{load()}, which must be implemented.
 Constructor gets a \code{Properties} instance as a parameter and saves it to \code{protected} field \code{properties}.
@@
 
-Purpose of the method \code{store()} can be easily guessed from its name, it's used to store values that are set in \emph{properties pane} elements (text field, checkbox etc.) to the provided \emph{properties} field.
-On the other side, \code{load()} method is used to populate \emph{properties pane} elements with already saved values in the \emph{properties}.
+Purpose of the method \code{store()} can be easily guessed from its name, it's used to store values that are set in \emph{properties pane} elements (text fields, checkboxes etc.) to the provided \emph{properties} field.
+On the other hand, \code{load()} method is used to populate \emph{properties pane} elements with already saved values in the \emph{properties}.
 
@@
 
-  private javax.swing.JCheckBox checkBox;
+  private JCheckBox checkBox;
 
@@
     ...
-    initialize checkBox
+    // initialize checkBox
     ...
@@
 
-This section describes ability of jInfer to create new jInfer project through \emph{new project} wizard.
-This wizard contains two steps, where first is standard one with project name and location selection, and the second is the selection of Initial grammar module, Simplifier and Schema generator module which will be used in inference for newly created module.
-New project creation process can be already finished in a first step, in that case, default modules will be used for inference.
-After finishing of the wizard, new jInfer project type is created with typed name in selected location.
-It is important to notice that package \code{cz.cuni.mff.ksi.jinfer.projecttype} contains \code{JInferTemplateProject.zip} file, in which is default folder structure of the jInfer project type, and the content of a zip is copied into location defined in the wizard.
-For more information, how to create new project wizard,  please look at \cite{project_sample}.\\
+This section describes the ability of jInfer to create a new jInfer project through a \emph{new project} wizard.
+This wizard contains two steps, where first is a standard one with project name and location selection, and the second is the selection of Initial Grammar Generator, Simplifier and Schema Generator modules which will be used in inference for this newly created project.
+New project creation process can be finished in the first step, in which case the default modules will be used for the inference.
+After finishing the wizard, new jInfer project is created with chosen name in the chosen location.
+It is important to notice that package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype} contains \code{JInferTemplateProject.zip} file, in which the default folder structure of a jInfer project is stored.
+The content of this zip is copied into location defined in the wizard.
+For more information on how to create a new project wizard, please refer to \cite{project_sample}.\\
 
-All classes responsible for creation of this wizard are in package \code{cz.cuni.mff.ksi.jinfer.projecttype.sample}.
-It contains \code{JinferTemplateWizardIterator} class which defines number of steps in the wizard and creates panel for each step.
-Each step is  represented by a class implementing \code{WizardDescriptor.Panel} interface, which creates steps panel and validate its data inserted by a user.
-Finally there is class extending \code{JPanel} in which are components of the first step.
+All classes responsible for creation of this wizard are in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.sample} package.
+It contains \code{JinferTemplateWizardIterator} class defining the number of steps in the wizard and creates a panel for each of these steps.
+Each step is represented by a class implementing \code{WizardDescriptor.Panel} interface, which creates a step panel and validates its data inserted by the user.
+Finally there is a class extending \code{JPanel} containing components of the first step.
 
@@
 
-Also jInfer project type has its own properties category in a project properties window called \emph{Miscellaneous}, and it is implemented by \code{ProjectPropertiesPanelProvider} and \code{ProjectPropertiesPanel} classes.
+A jInfer project has its own properties category in the project properties window called \emph{Miscellaneous}.
+This is implemented by \code{ProjectPropertiesPanelProvider} and \code{ProjectPropertiesPanel} classes.
 In this category, two properties can be set.
-First property is default \emph{input file} folder to which is inserted input file selected from \emph{FilesAddAction}, that has extension which is not defined by any \code{Processor} registered in jInfer.
-Second property is global \emph{LOG} level, which defines the minimum level of messages that will be displayed in jInfer output window.
-If some module has its own \emph{LOG} level setting, global settings are overridden by these.
+First property is default \emph{input file} folder to which an unknown input file is inserted (its extension is not defined by any \code{Processor} registered in jInfer).
+Second property is a global \emph{log} level, which defines the minimum level of messages that will be displayed in jInfer output window.
+If a module has its own \emph{log} level setting, global settings are overridden by it.
 
@@
 This section describes \code{RunningProject} utility class, which is not part of the \jmodule{ProjectType} module, but is closely related to it.
-This class can be found in a \code{cz.cuni.mff.ksi.jinfer.base.utils} package in the \jmodule{Base} module.
-The main purpose of this class is to allow all jInfer \jmodule{module}s to access \code{JInferProject} instance of the actual running inference.
-Notice that in jInfer application can be only one running inference at a time.\\
+This class can be found in a \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.base.\.utils} package in the \jmodule{Base} module.
+The main purpose of this class is to allow all jInfer \jmodule{module}s to access the \code{JInferProject} instance of the actual running inference.
+Notice that in jInfer application only one running inference is allowed at a time.\\
 
@@
 \begin{itemize}
-	\item \code{setActiveProject()} - Setter of the \code{JInferProject} instance of a running project, which set the \code{JInferProject} instance if no other inference is running (no other \code{JInferProject} instance is already set).
-Returns \code{boolean} value, according as the setting was successful or not.
-	\item \code{removeActiveProject()} - Removes \code{JInferProject} instance if some inference is running.
-	\item \code{getActiveProject()} - Getter of the \code{JInferProject} instance of a running project.
-	\item \code{isActiveProject()} - Checks if some inference is running (the \code{JInferProject} instance is set).
-	\item \code{getActiveProjectProps()} - Gets \code{Properties} instance of the actual running inference (\code{Properties} instance is part of the \code{JInferProject} instace).
-	\item \code{getNextModuleCaps()} - Getter of a capability of the next module in the inference chain.
-	\item \code{setNextModuleCaps()} - Setter of a capability of the next module in the inference chain.
+	\item \code{setActiveProject()} - setter of the \code{JInferProject} instance of a running project, which sets the \code{JInferProject} instance if no other inference is running (no other \code{JInferProject} instance is already set).
+Returns \code{boolean} value, based on whether the setting was successful.
+	\item \code{removeActiveProject()} - removes the \code{JInferProject} instance if some inference is running.
+	\item \code{getActiveProject()} - getter of the \code{JInferProject} instance of a running project.
+	\item \code{isActiveProject()} - checks if some inference is running (the \code{JInferProject} instance is set).
+	\item \code{getActiveProjectProps()} - gets the \code{Properties} instance from the actual running inference (\code{Properties} instance is part of the \code{JInferProject} instace).
+If no project is running at the moment, an empty \code{Properties} instance is returned - this allows the JUnit tests to run with default settings.
+	\item \code{getNextModuleCaps()} - getter of capabilities of the next module in the inference chain.
+	\item \code{setNextModuleCaps()} - setter of capabilities of the next module in the inference chain.
 \end{itemize}

### 2011-03-01 doc/jInfer/architecture/architecture.tex
architecture final touch

--- 2011-02-01 doc/jInfer/architecture/architecture.tex
@@
 \begin{figure}
-\centering\includegraphics[scale=\myscale]{nodes_full2}
+\centering\includegraphics[width=\textwidth]{nodes_full2}
 \caption{Class diagram for XML representing interfaces and classes} \label{nodes_full2}
@@
 For elements, it means no adding of attributes and changing regexp reference (regexp instance itself is immutable as well).
-The same \code{getMutable()} principles and good usage practises hold for these classes.
+The same \code{get\.Mu\.tab\.le()} principles and good usage practises hold for these classes.
 
@@
 \begin{verbatim}
-	B convertSymbol(final A symbol);
+	B convertSymbol(A symbol);
 \end{verbatim}
@@
 construct.
-Even if regular expression is complicated, export module has to do this "flattening".texmaker
+Even if regular expression is complicated, export module has to do this ``flattening''.texmaker
 	Lets look on XSD output of same rules:
@@
 
-	This is basically done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{expand()}.
-One can implement own intelligent expander, or use our provided\\ \code{cz.cuni.mff.ksi.jinfer.basicigg.expansion.ExpanderImpl}.
+This is basically done by calling \code{cz.cuni.mff.ksi.jinfer.base.interfaces.Expander} interface method \code{ex\.pand()}.
+One can implement own intelligent expander, or use our provided \code{cz.cuni.mff.ksi.jinfer.basic\.igg.ex\.pansion.ExpanderImpl}.
 
@@
 
-\code{SchemaGeneratorCallback.finished()} has two \code{String}s as arguments: \code{schema} is the actual string representation of the resulting schema, \code{extension} is a file extension of the result (such as "dtd" or "xsd") which the framework will use when saving the result in a file.
+\code{SchemaGeneratorCallback.finished()} has two \code{String}s as arguments: \code{schema} is the actual string representation of the resulting schema, \code{extension} is a file extension of the result (such as ``dtd'' or ``xsd'') which the framework will use when saving the result in a file.
 
@@
 \end{itemize}
-All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively means "true").
+All of these metadata are of set/not set character (using \code{Boolean.TRUE} as value, but their presence effectively means ``true'').
 
@@
 For this reason, inference modules extend the \code{Capabilities} interface defining a single method: \code{getCa\.pa\.bi\.li\.ties()}.
-This method returns a list of strings representing the abilities, or "capabilities" of this module.
+This method returns a list of strings representing the abilities, or ``capabilities'' of this module.
 An inference module can query the capabilities of the \emph{next} module in the inference chain by invoking the 
@@
   // lookup expander
-  final Expander expander = Lookup.getDefault().lookup(Expander.class);
+  Expander expander = Lookup.getDefault().lookup(Expander.class);
   // return expanded
@@
 \begin{itemize}
-	\item \jmodule{jInfer}: not a NBP module in fact, but rather a module suite - the ``root" NB project for the whole framework.
+	\item \jmodule{jInfer}: not a NBP module in fact, but rather a module suite - the ``root'' NB project for the whole framework.
 	\item \jmodule{Base}: contains common data structures, interfaces and utility logic shared across all other modules.
@@
 
-%\nocite{*}
+\paragraph{}
+Another use of tokens is \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} representing a module providing a rule displayer.
+This token is provided by \jmodule{BasicRuleDisplayer} and \jmodule{TreeRuleDisplayer}, requested by \jmodule{BasicIGG}, \jmodule{XSDImporter} and \jmodule{TwoStepSimplifier}.
+
+\nocite{*}
 \newpage

### 2011-03-01 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor module: review round 4.

@@
 
-\code{StatePickingVisualizer} is trivial extension of \code{PluggableVisualizer}. It has no additional methods. Its constructor has several additional arguments. Their description follows.
+\code{StatePickingVisualizer} is a trivial extension of \code{PluggableVisualizer}. It has no additional methods. Its constructor has several additional arguments. Their description follows.
 \begin{verbatim}
@@
 With this plugin used, user can pick and unpick automaton states separately or pick several states at once by dragging rectangular selection box over states.\
-Main difference compared to the \code{VertexPickingGraphMousePlugin} is \code{VerticesPickingGraphMousePlugin} does not call component's \code{guiDone()} method.\\
+Main difference compared to the \code{VertexPickingGraphMousePlugin} is that \code{VerticesPickingGraphMousePlugin} does not call component's \code{guiDone()} method.\\
 
@@
 
-Source codes of \code{StatesPickingComponent} resides in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.component}, \code{States\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
+Source codes of \code{StatesPickingComponent} reside in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.component}, \code{States\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
 

### 2011-03-01 doc/jInfer/projecttype/projecttype.tex
Documentation: ProjectType module: review round 1 part 4.

@@
 \begin{figure}
-	\subfigure[Legend: Rectangle is folder, Rounded rectangle is file] {
+	\subfigure[Legend: Rectangle is folder, Rounded rectangle is file.] {
 		\centering\includegraphics[scale=\myscale]{project_dir_structure}
 	}
-	\caption{Project directory structure} \label{dir-structure}
+	\caption{Project directory structure.} \label{dir-structure}
 \end{figure}
@@
 This structure is described in figure \ref{dir-structure}.
-A folder in the filesystem is considered to be jInfer project folder iff it contains a \emph{jinferproject} subfolder.
+A folder in the filesystem is considered to be jInfer project folder if it contains a \emph{jinferproject} subfolder.
 This folder contains two files: \emph{project.properties}, where all properties set for that particular project are stored.
-The second file is \emph{input.files} which is an XML file filled with paths to files selected particular project input (structure of this file is described in section \ref{input.files}).
+The second file is \emph{input.files} which is an XML file filled with paths to files selected as particular project input (structure of this file is described in section \ref{input.files}).
 \emph{output} subfolder contains schema files which were generated by jInfer inference runs.
@@
 \begin{itemize}
-	\item Base classes - classes providing core functionality like project creation and definitions of operations like move, delete, copy etc.
+	\item Base classes - Classes providing core functionality like project creation and definitions of operations like move, delete, copy etc.
 All the base classes are contained in the \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype} package.
-	\item Visualization classes - these classes create tree structure of jInfer project in NBP Projects view and are contained in the \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.nodes}.
-	\item Actions - classes from \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.actions} package that provide actions allowing adding input files into the project, removing input files, or running the inference on the project.
-	\item Properties - classes responsible for creation of project properties window with properties category tree.
+	\item Visualization classes - These classes create tree structure of jInfer project in NBP Projects view and are contained in the \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.nodes}.
+	\item Actions - Classes from \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.actions} package that provide actions allowing adding input files into the project, removing input files, or running the inference on the project.
+	\item Properties - Classes responsible for creation of project properties window with properties category tree.
 These are located in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.properties} package.
-	\item Project wizard - classes representing new project creation through a project wizard from \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.sample} package.
+	\item Project wizard - Classes representing new project creation through a project wizard from \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.projecttype.\.sample} package.
 \end{itemize}
@@
 	\centering\includegraphics[scale=\myscale]{node_structure}
-	\caption{Project nodes structure} \label{node-structure}
+	\caption{Project nodes structure.} \label{node-structure}
 \end{figure}
@@
 Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
-From jInfer module point of view the properties category is represented by \code{PropertiesPanelProvider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
+From jInfer module point of view the properties category is represented by \code{Properties\-Panel\-Provider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
 How to use this interface and abstract class to create custom project properties category presented in jInfer project properties window is described in section \ref{properties_reg}.\\
@@
 \code{JInferComponentProvider} is a very simple class that extends \code{ProjectCustomizer.CategoryComponentProvider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
-In the \code{JInferCustomizerProvider} class (which extends NBP \code{CustomizerProvider} class) is the core functionality of collecting categories (\code{PropertiesPanelProvider} interface) with associated properties panes (\code{AbstractPropertiesPanel} class) from all jInfer modules.\\
+In the \code{JInferCustomizer\-Provider} class (which extends NBP \code{CustomizerProvider} class) is the core functionality of collecting categories (\code{PropertiesPanelProvider} interface) with associated properties panes (\code{AbstractPropertiesPanel} class) from all jInfer modules.\\
 
@@
 \begin{itemize}
-	\item getName() - gets a programmatic name of this category.
-	\item getDisplayName() - gets a name of this category which will be displayed in properties window.
-	\item getPriority() - gets priority of this category, by which it is sorted among other sibling categories in the tree.
-	\item getPanel(Properties) - gets the \code{AbstractPropertiesPanel} instance representing the category \emph{properties pane}.
+	\item getName() - Gets a programmatic name of this category.
+	\item getDisplayName() - Gets a name of this category which will be displayed in properties window.
+	\item getPriority() - Gets priority of this category, by which it is sorted among other sibling categories in the tree.
+	\item getPanel(Properties) - Gets the \code{AbstractPropertiesPanel} instance representing the category \emph{properties pane}.
 The only parameter is an instance of \code{Properties} class holding stored properties for the \emph{properties pane}.
-	\item getParent() - getter for a programmatic name of category which is the parent of this category in the tree hierarchy.
-	\item getSubCategories() - gets a \code{List} of \code{VirtualCategoryPanel} instances representing virtual category children.
+	\item getParent() - Getter for a programmatic name of category which is the parent of this category in the tree hierarchy.
+	\item getSubCategories() - Gets a \code{List} of \code{VirtualCategoryPanel} instances representing virtual category children.
 \end{itemize}
@@
 @ServiceProvider(service = PropertiesPanelProvider.class)
-public class PropertiesPanelProviderImpl implements PropertiesPanelProvider{
+public class PropertiesPanelProviderImpl implements PropertiesPanelProvider {
 
   @Override
-  public AbstractPropertiesPanel getPanel(final Properties properties) {
+  public AbstractPropertiesPanel getPanel(Properties properties) {
     return new ExamplePropertiesPanel(properties);
@@
   public List<VirtualCategoryPanel> getSubCategories() {
-    final List<VirtualCategoryPanel> result = new ArrayList<VirtualCategoryPanel>();
+    List<VirtualCategoryPanel> result = new ArrayList<VirtualCategoryPanel>();
     result.add(new VirtualCategoryPanel("virtual_categoryID", "example virtual category",
@@
  \begin{itemize}
-	\item \code{categoryId} - programmatic name of the category or ID.
-	\item \code{categoryName} - human readable name of the category displayed in jInfer properties window.
+	\item \code{categoryId} - Programmatic name of the category or ID.
+	\item \code{categoryName} - Human readable name of the category displayed in jInfer properties window.
 	\item \code{modules} - \code{List<?\ extends NamedModule>} of modules that will be children of this virtual category.
@@
 
-  public ExamplePropertiesPanel(final Properties properties) {
+  public ExamplePropertiesPanel(Properties properties) {
     super(properties);
@@
 \begin{itemize}
-	\item \code{setActiveProject()} - setter of the \code{JInferProject} instance of a running project, which sets the \code{JInferProject} instance if no other inference is running (no other \code{JInferProject} instance is already set).
+	\item \code{setActiveProject()} - Setter of the \code{JInferProject} instance of a running project, which sets the \code{JInferProject} instance if no other inference is running (no other \code{JInferProject} instance is already set).
 Returns \code{boolean} value, based on whether the setting was successful.
-	\item \code{removeActiveProject()} - removes the \code{JInferProject} instance if some inference is running.
-	\item \code{getActiveProject()} - getter of the \code{JInferProject} instance of a running project.
-	\item \code{isActiveProject()} - checks if some inference is running (the \code{JInferProject} instance is set).
-	\item \code{getActiveProjectProps()} - gets the \code{Properties} instance from the actual running inference (\code{Properties} instance is part of the \code{JInferProject} instace).
+	\item \code{removeActiveProject()} - Removes the \code{JInferProject} instance if some inference is running.
+	\item \code{getActiveProject()} - Getter of the \code{JInferProject} instance of a running project.
+	\item \code{isActiveProject()} - Checks if some inference is running (the \code{JInferProject} instance is set).
+	\item \code{getActiveProjectProps()} - Gets the \code{Properties} instance from the actual running inference (\code{Properties} instance is part of the \code{JInferProject} instace).
 If no project is running at the moment, an empty \code{Properties} instance is returned - this allows the JUnit tests to run with default settings.
-	\item \code{getNextModuleCaps()} - getter of capabilities of the next module in the inference chain.
-	\item \code{setNextModuleCaps()} - setter of capabilities of the next module in the inference chain.
+	\item \code{getNextModuleCaps()} - Getter of capabilities of the next module in the inference chain.
+	\item \code{setNextModuleCaps()} - Getter of capabilities of the next module in the inference chain.
 \end{itemize}

### 2011-03-01 doc/jInfer/autoeditor/autoeditor.tex
autoeditor final touch in progress

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Mário Mikula \\ \hline
-Required tokens:       & org.openide.windows.WindowManager \\ \hline
-Provided tokens:       & none \\ \hline
-Module dependencies:   & Base \\ 
+Responsible developer & Mário Mikula \\ \hline
+Required tokens       & org.openide.windows.WindowManager \\ \hline
+Provided tokens       & none \\ \hline
+Module dependencies   & Base \\ 
 					   & JUNG \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.autoeditor \\ 
-					   & cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer \\
-   					   & cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts \\
-   					   & cz.cuni.mff.ksi.jinfer.autoeditor.gui.component \\ \hline
+Public packages       & \code{cz.cuni.mff.ksi.jinfer.autoeditor} \\ 
+					   & \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} \\
+   					   & \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts} \\
+   					   & \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component} \\ \hline
 \end{tabular}
@@
         @Override
-        public String transform(final Step<Regexp<T>> step) {
+        public String transform(Step<Regexp<T>> step) {
           StringBuilder sb = new StringBuilder();

### 2011-03-01 doc/jInfer/autoeditor/autoeditor.tex
3187368 - Review documentation: cleanup - autoeditor done

@@
 Responsible developer & Mário Mikula \\ \hline
-Required tokens       & org.openide.windows.WindowManager \\ \hline
+Required tokens       & \code{org.openide.windows.WindowManager} \\ \hline
 Provided tokens       & none \\ \hline
@@
 \begin{figure}
-\centering\includegraphics[scale=\myscale]{uml_base_classes}
+\centering\includegraphics[width=\textwidth]{uml_base_classes}
 \caption{Class diagram for the base classes.} \label{uml_base_classes}
@@
 
-Source codes of \code{State\.Picking\.Component} resides in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.component}, \code{State\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertex\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
+Source codes of \code{State\.Picking\.Component} resides in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.com\.po\.nent}, \code{State\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Ver\.tex\-Pick\.ing\-Graph\-Mou\.se\-Plug\.in} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
 
@@
     StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
-    Layout<State<T>, Step<T>> layout = LayoutHelperFactory.createUserLayout(automaton, transformer);
+    Layout<State<T>, Step<T>> layout = 
+      LayoutHelperFactory.createUserLayout(automaton, transformer);
     StatePickingVisualizer<Regexp<T>>
@@
 \begin{figure}
-\centering\includegraphics[scale=0.6]{screenshot}
+\centering\includegraphics[width=\textwidth]{screenshot}
 \caption{Screenshot of \jmodule{AutoEditor} GUI} \label{figure_gui}
@@
 
-Source codes of \code{StatesPickingComponent} reside in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.component}, \code{States\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
+Source codes of \code{StatesPickingComponent} reside in package \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.gui.\.com\.po\.nent}, \code{States\-Picking\-Visualizer} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer} and \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} in \code{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.autoeditor.\.automatonvisualizer.\.graphmouseplugins}.\\
 
@@
         layout = LayoutHelperFactory.createUserLayout(inputAutomaton, transformer);
-      StatesPickingVisualizer<T> visualizer = new StatesPickingVisualizer<T>(layout, transformer);
+      StatesPickingVisualizer<T> visualizer = 
+          new StatesPickingVisualizer<T>(layout, transformer);
 
@@
         panel.setLabel("Automaton for element <" + elementName
-          + ">. Please select states to be merged and click continue. Select at least 2 states.");
+          + ">. Please select states to be merged and click continue."
+          + " Select at least 2 states.");
       } else {
@@
 The options panel along with all the logic is in the \code{cz.cuni.mff.ksi.jinfer.autoeditor.options} package.
-Available options include setting the color of the background, colors and shapes of some special types of automaton states.\\
+Available options include setting the color of the background, colors and shapes of some special types of automaton states.
+\\
 
 To have some effect, these settings need to be implemented by extensions of \code{Visualizer} class.
-For examples, see source codes of \code{StatePickingVisualizer} and \code{StatesPickingVisualizer} classes.\\
+For examples, see source codes of \code{StatePickingVisualizer} and \code{StatesPickingVisualizer} classes.
 

### 2011-03-01 doc/jInfer/base/base.tex
fixed overfull hboxes by package names
3187368 - Review documentation: cleanup - base done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & none \\ \hline
-Module dependencies:   & none \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.base.automaton \\ 
-& cz.cuni.mff.ksi.jinfer.base.interfaces \\ 
-& cz.cuni.mff.ksi.jinfer.base.interfaces.inference \\ 
-& cz.cuni.mff.ksi.jinfer.base.interfaces.nodes \\
-& cz.cuni.mff.ksi.jinfer.base.objects \\ 
-& cz.cuni.mff.ksi.jinfer.base.objects.nodes \\ 
-& cz.cuni.mff.ksi.jinfer.base.regexp \\ 
-& cz.cuni.mff.ksi.jinfer.base.utils \\
-& org.apache.log4j.* \\ \hline
+Responsible developer & Matej Vitásek \\ \hline
+Required tokens       & none \\ \hline
+Provided tokens       & none \\ \hline
+Module dependencies   & none \\ \hline
+Public packages       & \code{cz.cuni.mff.ksi.jinfer.base.automaton} \\ 
+& \code{cz.cuni.mff.ksi.jinfer.base.interfaces} \\ 
+& \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} \\ 
+& \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes} \\
+& \code{cz.cuni.mff.ksi.jinfer.base.objects} \\ 
+& \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes} \\ 
+& \code{cz.cuni.mff.ksi.jinfer.base.regexp} \\ 
+& \code{cz.cuni.mff.ksi.jinfer.base.utils} \\
+& \code{org.apache.log4j.*} \\ \hline
 \end{tabular}
@@
 Regular expression representation is contained in \code{cz.cuni.mff.ksi.jinfer.base.regexp} package.
-Most important class is of course \code{Regexp}, assisted by an enum of its type \code{RegexpType} and representation of its interval \code{RegexpInterval}.\\
+Most important class is of course \code{Regexp}, assisted by an enum of its type \code{RegexpType} and representation of its interval \code{RegexpInter\.val}.\\
 XML node representation is described by interfaces in \code{cz.cuni.mff.ksi.jinfer.base.interfaces.nodes} package and more-less concrete implementations in \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes} package.\\

### 2011-03-01 doc/jInfer/basicdtd/basicdtd.tex
3187368 - Review documentation: cleanup - basic{rule,xsd,igg,dtd} done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\ \hline
-Module dependencies:   & Base \\ \hline
-Public packages:       & none \\ \hline
+Responsible developer & Matej Vitásek \\ \hline
+Required tokens       & none \\ \hline
+Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator} \\ \hline
+Module dependencies   & Base \\ \hline
+Public packages       & none \\ \hline
 \end{tabular}

### 2011-03-01 doc/jInfer/basicruledisplayer/basicruledisplayer.tex
regenerating, ahonen cite to make bibtex happy
minor
fixed overfull hboxes by package names
3187368 - Review documentation: cleanup - basic{rule,xsd,igg,dtd} done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
-Module dependencies:   & Base \\ \hline
-Public packages:       & none \\ \hline
+Responsible developer & Matej Vitásek \\ \hline
+Required tokens       & none \\ \hline
+Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} \\ \hline
+Module dependencies   & Base \\ \hline
+Public packages       & none \\ \hline
 \end{tabular}

### 2011-03-01 doc/jInfer/basicigg/basicigg.tex
3187368 - Review documentation: cleanup - basic{rule,xsd,igg,dtd} done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
-Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\ \hline
-Module dependencies:   & Base \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.basicigg.properties \\ \hline
+Responsible developer & Matej Vitásek \\ \hline
+Required tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} \\ \hline
+Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator} \\ \hline
+Module dependencies   & Base \\ \hline
+Public packages       & \code{cz.cuni.mff.ksi.jinfer.basicigg.properties} \\ \hline
 \end{tabular}
@@
 
-One interesting use case for expansion is simple adaptation of legacy inference methods. "Old" methods working only with \emph{simple} grammar can be used unchanged with "new" inputs - schemas.\\
+One interesting use case for expansion is simple adaptation of legacy inference methods. ``Old'' methods working only with \emph{simple} grammar can be used unchanged with ``new'' inputs - schemas.\\
 

### 2011-03-01 doc/jInfer/basicxsd/basicxsd.tex
3187368 - Review documentation: cleanup - basic{rule,xsd,igg,dtd} done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Mário Mikula \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\ \hline
-Module dependencies:   & Base \\ \hline
-Public packages:       & none \\ \hline
+Responsible developer & Mário Mikula \\ \hline
+Required tokens       & none \\ \hline
+Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator} \\ \hline
+Module dependencies   & Base \\ \hline
+Public packages       & none \\ \hline
 \end{tabular}
@@
 
-Named type of an element is defined by XSD element \code{simpleType} (lack of features mentioned above) or \code{complexType} with specified attribute \code{name}.
+Named type of an element is defined by XSD element \code{simpleType} (lack of features mentioned above) or \code{complex\.Ty\.pe} with specified attribute \code{name}.
 Its content is exactly the same as described above.
@@
 
-Default values for \code{minOccurs} and \code{maxOccurs} attributes are "1", if they are not specified. So the example above has the same meaning as the following one.
+Default values for \code{minOccurs} and \code{maxOccurs} attributes are ``1'', if they are not specified. So the example above has the same meaning as the following one.
 
@@
 
-Attributes are defined by XSD element \code{xs:attribute} with attributes \code{name}, \code{type} and optional \code{use}.
-Elements \code{xs:attributes} have to be placed at the end of a \code{complexType} definition.
+Attributes are defined by XSD element \code{xs:\.attri\.bu\.te} with attributes \code{name}, \code{type} and optional \code{use}.
+Elements \code{xs:\.attri\.bu\.tes} have to be placed at the end of a \code{complexType} definition.
 
@@
 There are two classes extending \code{AbstractElementsExporter}.
-\code{GlobalElementsExporter} and \code{RootElementExporter}.
+\code{GlobalElementsExporter} and \code{Root\.Ele\.ment\.Ex\.port\.er}.
 Their constructors has the same signature as constructor of \code{AbstractElementsExporter} and both have \code{run()} method to perform their function.\\
@@
 Name of a global type is derived from name of a corresponding global element by prefixing and suffixing it.
-Default prefix is "T" and default suffix is empty.
+Default prefix is ``T'' and default suffix is empty.
 These values can be changed in project properties.
@@
 \paragraph{Examples of definition of global types}
-Type generated from element named "GlobalElement", containing elements "Text", which is string, and element "AnotherGlobalElement", which is another global element.
+Type generated from element named ``GlobalElement'', containing elements ``Text'', which is string, and element ``AnotherGlobalElement'', which is another global element.
 
@@
 
-Type generated from element named "X", containing element "Y", which is not global element and contains elements "Text" and "GlobalElement" from the previous example.
-Element "X" has one mandatory string attribute named "id".
+Type generated from element named ``X'', containing element ``Y'', which is not global element and contains elements ``Text'' and ``GlobalElement'' from the previous example.
+Element ``X'' has' one mandatory string attribute named ``id''.
 
@@
 This applies also to elements which are not global and their type is one of a built-in types.
-For example, root element "Root" contains one the following two elements.
-Global element "A" and string element "B".
+For example, root element ``Root'' contains one the following two elements.
+Global element ``A'' and string element ``B''.
 
@@
 Elements which are not global and are not of a built-in type are defined recursively, at the place of their occurrence.
-For example, root element "Root" contains element "A", which contains two other string elements.
+For example, root element ``Root'' contains element ``A'', which contains two other string elements.
 
@@
 
-Export of attributes is done alongside with export of elements, as attributes are exported as XSD element \code{xs:attribute} in XSD element \code{xs:complexType}.
+Export of attributes is done alongside with export of elements, as attributes are exported as XSD element \code{xs:\.attri\.bu\.te} in XSD element \code{xs:complexType}.
 
@@
 
-\subsection{"Unique Particle Attribution" problem}
+\subsection{``Unique Particle Attribution'' problem}
 
@@
 	\item For each global element, its type is exported to a XSD representation by \code{GlobalElementsExporter}.
-	\item Root element and recursively all remaining elements are exported to a XSD representation by \code{RootElementExporter}.
-	\item String representation of the schema is returned along with the information that file extension should be "xsd".
+	\item Root element and recursively all remaining elements are exported to a XSD representation by \code{Root\.Element\.Ex\.porter}.
+	\item String representation of the schema is returned along with the information that file extension should be ``xsd''.
 \end{enumerate}

### 2011-03-01 doc/jInfer/projecttype/projecttype.tex
3187368 - Review documentation: cleanup - projecttype mangling after rio

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Michal Švirec \\ \hline
-Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\
- & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\
- & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\
- & org.openide.windows.IOProvider \\ \hline
-Provided tokens:       & none \\ \hline
-Module dependencies:   & Base \\
+Responsible developer & Michal Švirec \\ \hline
+Required tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator} \\
+ & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator} \\
+ & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier} \\
+ & \code{org.openide.windows.IOProvider} \\ \hline
+Provided tokens       & none \\ \hline
+Module dependencies   & Base \\
 	& Runner \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.projecttype.actions \\ \hline
+Public packages       & \code{cz.cuni.mff.ksi.jinfer.projecttype.actions} \\ \hline
 \end{tabular}
@@
 \code{JinferProject} class implements the \code{Project} interface from NBP and represents an in-memory representation of a jInfer project.
-Inner structure of this class is very simple, interface specifies only two methods: \code{getProjectDirectory()} and \code{getLookup()}.
+Inner structure of this class is very simple, interface specifies only two methods: \code{getProject\.Directo\.ry()} and \code{getLookup()}.
 All the extensibility of a project type is done by the \emph{Lookup} functionality of NBP.
@@
 
-A category in the properties window is represented by NBP \code{ProjectCustomizer.Category} class, properties pane is represented by \code{JPanel} class.
-Classes responsible for creation of this properties window are \code{JInferCustomizerProvider} and \code{JInferComponentProvider}.
-From jInfer module point of view the properties category is represented by \code{Properties\-Panel\-Provider} interface and associated properties pane by \code{AbstractPropertiesPanel} abstract class.
+A category in the properties window is represented by NBP \code{Project\.Customizer.\.Category} class, properties pane is represented by \code{JPanel} class.
+Classes responsible for creation of this properties window are \code{JInfer\.Customizer\.Provider} and \code{JInfer\.Component\.Provider}.
+From jInfer module point of view the properties category is represented by \code{Properties\.Panel\.Provider} interface and associated properties pane by \code{Abstract\.Properties\.Panel} abstract class.
 How to use this interface and abstract class to create custom project properties category presented in jInfer project properties window is described in section \ref{properties_reg}.\\
 
-\code{JInferComponentProvider} is a very simple class that extends \code{ProjectCustomizer.CategoryComponentProvider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
-In the \code{JInferCustomizer\-Provider} class (which extends NBP \code{CustomizerProvider} class) is the core functionality of collecting categories (\code{PropertiesPanelProvider} interface) with associated properties panes (\code{AbstractPropertiesPanel} class) from all jInfer modules.\\
+\code{JInferComponentProvider} is a very simple class that extends \code{ProjectCustomizer.Category\.Component\.Provider} class from NBP and its only method is \code{create()} which takes category as a parameter and returns properties pane.
+In the \code{JInfer\.Customizer\.Provider} class (which extends NBP \code{Customizer\.Provider} class) is the core functionality of collecting categories (\code{Properties\.Panel\.Provider} interface) with associated properties panes (\code{Abstract\.Properties\.Panel} class) from all jInfer modules.\\
 
@@
 
-As was written before, each jInfer project properties category is represented in jInfer modules by \code{PropertiesPanelProvider} interface.
+As was written before, each jInfer project properties category is represented in jInfer modules by \code{Properties\.Panel\.Provider} interface.
 To create new category, this interface must be implemented and implementation must be annotated by the following code.

### 2011-03-01 doc/jInfer/xsdimport/xsdimport.tex
XSD Import documentation, unreviewed. Attempt 2

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}
\usepackage{xstring}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\emph{#1}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer XSDImporter Module Description}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}
\maketitle
\noindent Target audience: developers willing to extend jInfer, specifically those creating or extending a parser.

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer: & Robert Smetana \\ \hline
Required tokens:       & none \\ \hline
Provided tokens:       & none \\ \hline
Module dependencies:   & listed for each module separately \\ \hline
Public packages:       & listed for each module separately \\ \hline
\end{tabular}

\section{Introduction}
This document focuses on the description of modules involved in the process of importing of XML Schema Definition (XSD) documents, or Schema documents. 
These modules are responsible for creating the IG rules from input files, therefore they belong in the inference modules category. 
As a whole, they are a specialized extension of \code{BasicIGG} module.

None of the modules provide nor require any tokens. Complete dependency tree is depicted in Figure 9 and described in \cite[section 4]{archdoc}.

\subsection{Naming convention}
Note that there are several types of entities called `element' in the context of the described modules.
To avoid ambiguity, in this document we shall use notation \code{DOM.Element} 
for instances of \code{org.w3c.dom.Element} and notation \code{Element} 
for instances of\code{cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element}.

To add to this complication, XSD Schemas are XML documents which consist of elements and attributes.
We shall refer to these XML elements as \emph{tags} and XML attributes as \emph{tag attributes}.
Moreover, all tags and tag attributes that are known to our parsers are specified in
section \ref{sect:utils}.

For example, the \emph{attribute} tag can have a tag attribute \emph{use}.
While former will match \code{XSDTag.ATTRIBUTE}, the latter will match \code{XSDAttribute.USE}.
Lastly, tag attributes should not be confused with instances of \code{Attribute}, which is a class for storing information about the tag attributes inside jInfer. Every \code{Element} has a list of \code{Attribute}s, as can be seen from Figure 4 in \cite[section 2.3]{archdoc}.

\subsection{Namespaces}
XSD Schemas allow using different namespace prefixes for tag names. Our parser implementations do not recognize namespaces; all prefixes from tag names are trimmed. For example, there is no difference between \code{xs:choice} and \code{xsd:choice} tags, both will match the \code{XSDTag.CHOICE} constant.

\subsection{Rule-trees} \label{sect:ruletree}

Rule-trees are tree data structures, where vertices are instances of \code{Element} class and edges are the relations between them stored in \code{Regexp} instances within vertices. Every vertex is a rule by itself, so in order to obtain IG rules, one must extract all Elements from a rule-tree. Rule-trees are created by \code{DOMHandler}, described in section \ref{sect:domhandler}. The extraction is done by \code{DOMParser}, described in section \ref{sect::domparser}. Or alternatively, when parsing with SAX is selected, \code{SAXHandler} handles both processes.

\section{Structure}
There are three modules that provide support for importing Schema files bundled with jInfer. 
\begin{itemize}
	\item \code{XSDImporter} - Main module providing support for actual parsing modules, as well as common configuration options for the parsers.
	\item \code{XSDImportDOM} - Dependent module implementing DOM parser method.
	\item \code{XSDImportSAX} - Dependent module implementing SAX parser method. 
\end{itemize}
All three modules depend on the \code{Base} module, as they all use the common data structures of jInfer. 
\code{XSDImporter} module also depends on \code{BasicIGG} module.
DOM and SAX parser modules are independent of each other; 
however, they rely on classes from \code{XSDImporter}. 
They provide similar functionality, but use different paradigms.

\subsection{XSD Importer module}

\noindent \begin{tabular}{|l|l|} \hline
Module dependencies:   & Base \\ 
                       & BasicIGG \\ \hline
Public packages:       & cz.cuni.mff.ksi.jinfer.xsdimporter.interfaces \\
                       & cz.cuni.mff.ksi.jinfer.xsdimporter.utils \\ \hline
\end{tabular}
\\

This module has several key roles in the process of importing XSD documents.\\

Firstly, it is an extension of the \code{BasicIGG} module, registering a service provider
of a \code{Processor} class, to handle files with \emph{xsd} extension (see \cite[section 3]{basiciggdoc}). 
It defines a common interface providing the developers with easy integration of their custom parsers to the rest of the inference process.
It also provides a common configuration in the form of a properties panel in project properties for all depending modules.
And finally, this module holds common utility classes related to the task of processing data characteristic to the Schema documents. \\
Functionality is divided between separate packages, in the same order:

\begin{itemize}
	\item \code{cz.cuni.mff.ksi.jinfer.xsdimporter}
	\item \code{cz.cuni.mff.ksi.jinfer.xsdimporter.interfaces}
	\item \code{cz.cuni.mff.ksi.jinfer.xsdimporter.properties}
	\item \code{cz.cuni.mff.ksi.jinfer.xsdimporter.utils}
\end{itemize}

Since some of the packages contain only one important class, we structure the documentation in logical blocks based on significance, not package structure.

\subsubsection{XSD Processor}

Simmilar to the \code{Processor} implementations in the \code{BasicIGG} module, this is the core class of \code{XSDImporter} logic.
XSDProcessor registers itself into \emph{schema} input folder, \emph{xsd} file extension and declares it cannot handle other file extensions.\\
Class depends on NBP lookups (see \cite[section 4.1]{archdoc}) to find available parsers, and uses the selected parser to perform the import operation on defined files.
Initial grammar generated by this processor is considered \emph{complex} and should be followed by expansion (see \cite[section 4]{basiciggdoc}).

\subsubsection{XSD Parser interface}

Key interface which needs to be implemented by any parser handling the import of Schema documents. Lookups of available parsing methods work by searching for service providers based on this interface. In order to support the lookup functionality, this interface extends \code{NamedModule}. It declares method \code{parse()} which takes an \code{InputStream} that contains the input document and returns a complete list of rules from that document. In other words, this method does all the necessary work. Method has the following signature:

\begin{verbatim}
List<Element> parse(final InputStream stream) throws XSDException;
\end{verbatim}

Diagram showing the inheritance of this interface, as well as both parsers bundled with jInfer, which implement it is in Figure \ref{parser_interface}.

\begin{figure}[ht]
\centering\includegraphics[scale=\myscale]{parser_intf}
\caption{Inheritance of XSDParser interface.} \label{parser_interface}
\end{figure}

\subsubsection{Properties}

Settings provided by the package \code{cz.cuni.mff.ksi.jinfer.xsdimporter.properties} are project-wide.
They are meant to be common for all parsers handling the import of Schema documents. Access to these settings is provided by utility class \code{XSDImportSettings} described in section \ref{sect:utils}. For description of each of the settings, please see section \ref{sect:settings}.

\subsubsection{Utils} \label{sect:utils}

Package \code{cz.cuni.mff.ksi.jinfer.xsdimporter.utils} contains classes available for all parsing methods.

- TODO explain the purpose of utility classes \\
- TODO describe every helper class in importer?

\paragraph{XSD Import Settings}

Simple utility class that provides convenient access methods to all the settings, hiding unnecesary implementation details. It holds the following methods.

\begin{itemize}
	\item \code{getParser()} returns selected instance of \code{XSDParser}.
	\item \code{logLevel()} returns current log priority, messages with lower priority are not displayed.
	\item \code{isVerbose()} returns \code{true} or \code{false}, depending on the checkbox.
	\item \code{stopOnError()} returns \code{true} or \code{false}, depending on the checkbox.
\end{itemize}

\paragraph{XSD Occurences}

Helper class for creating intervals from string representation of occurrences found in \code{minOccurs} and \code{maxOccurs} attributes of schema tags. Default values for both attributes are 1. Methods can create all types of intervals. If optional flag \code{minHasPriority} is set, method \code{createInterval} will prefer the value of lower bound, if parsing of upper bound fails, which can result in an interval with equal values of lower and upper limit (\code{min, min}).

\paragraph{XSD Attribude}

Enumeration of tag attributes currently recognized by our parsers. They are the following:
\begin{itemize}
	\item \code{name}
	\item \code{ref}
	\item \code{type}
	\item \code{use}
	\item \code{minOccurs}
	\item \code{maxOccurs}
\end{itemize}
List should be extended according to the parser, to reflect its functionality.

\paragraph{XSD Tag}

Enumeration of tags currently known by our parsers. They are the following:
\begin{itemize}
	\item \code{schema}
	\item \code{element}
	\item \code{complexType}
	\item \code{sequence}
	\item \code{choice}
	\item \code{all}
	\item \code{attribute}
	\item \code{redefine} - this tag is not yet supported, but allows parsing of a certain types of documents
\end{itemize}
This list should also be updated in case the parser functionality is extended.

\paragraph{XSD Utility}

This is basically only a library of short, useful and frequently used functions for parsing, please refer to the JavaDoc.

\subsection{XSD Import DOM module}

\noindent \begin{tabular}{|l|l|} \hline
Module dependencies:   & Base \\ 
                       & XSDImporter \\ \hline
Public packages:       & none \\ \hline
\end{tabular}
\\
This module implements the import functionality using API from Xerces DOM parser. Its structure is very simple, all classes are grouped in one main package, relying on utility classes from \code{XSDImporter} module mentioned before.

\subsubsection{DOM Parser} \label{sect:domparser}
Class implementing \code{XSDParser} interface. It is responsible for creation of DOM tree, extraction of contained rule-trees and aggregation of rules from them. First task is delegated to the underlying Xerces DOM parser API. Conversion between DOM tree and the rule-trees is then passed to \code{DOMHandler}.

\begin{verbatim}
parser = new com.sun.org.apache.xerces.internal.parsers.DOMParser();
parser.parse(new InputSource(stream));
Document doc = parser.getDocument();
DOMHandler handler = new DOMHandler();
List<Element> ruleTrees = handler.createRuleTrees(doc.getDocumentElement());
\end{verbatim}

\subsubsection{DOM Handler} \label{sect:domhandler}
When DOM parsing method is selected, most of the logic is implemented in \code{DOMHandler} class.
This class is responsible for creating so called rule-trees from an existing Document Object Model.

Its only public method iterates through children of the root node, given as parameter, in two passes. This root node should always contain \code{schema} tag, according to XSD specification. Firstly, it registers \code{complexType} and \code{element} tags, as these can be referenced from within the tree. And then launches the algorithm for creating rule-trees for relevant nodes. It can be summarized as follows:

\begin{verbatim}
public List<Element> createRuleTrees(org.w3c.dom.Element root) {
  for each (root.getChildNodes() as tag) {
    if (tag = XSDTag.ELEMENT)     addToReferenced(tag);
    if (tag = XSDTag.COMPLEXTYPE) addNamedCType(tag);
  }
  for each (root.getChildNodes() as tag) {
    subtree = buildRuleSubtree(tag);
    ruleTrees.add(subtree);
  }
  return ruleTrees;
}
\end{verbatim}

BuildRuleSubtree is the most important method. It is divided into four logical parts
\begin{enumerate}
	\item Creates an \code{Element} instance with appropriate parameters (occurences, metadata, etc.).
	\item Inspects the given tag and sets parameters for subnodes \code{Regexp}.
	\item Recursively creates rule-subtrees for its children.
	\item Finalizes the instance by checking for extra conditions.
\end{enumerate}

\subsubsection{DOM Helper}
Helper class for \code{DOMHandler}, containing ulility functions specific for DOM parsing method.
Two most notable of them are \code{createSentinel} and \code{finalizeElement}.

\begin{description}
	\item[CreateSentinel] Creates an instance of \code{Element} class, which is always a leaf in the rule-tree. This is useful when there are cyclic definitions in an input Schema document, or an \code{Element} with this name was defined previously and it is unnecessary to build a rule subtree for it again (it is a reference in this case). More information about sentinels can be found in \cite[section 3.2]{archdoc}, or \cite[section 4]{twostepdoc}.
	\item[FinalizeElement] Checks if \code{Element} is properly defined. Redefine it to \code{LAMBDA} when it was empty, or to token if it only contained a simple data type. This method allows for built-in data types of tags to be retained for export.
\end{description}

\subsection{XSD Import SAX module}

\noindent \begin{tabular}{|l|l|} \hline
Module dependencies:   & Base \\ 
                       & XSDImporter \\ \hline
Public packages:       & none \\ \hline
\end{tabular}
\\
This module implements the import functionality using API of a standard SAX parser. Its structure is divided in two packages. Package \code{cz.cuni.mff.ksi.jinfer.xsdimportsax} holds the main parsing logic, while package \code{cz.cuni.mff.ksi.jinfer.xsdimportsax.utils} contains two utility classes. Both \code{SAXAttributeData} and \code{SAXDocumentElement} serve as wrappers for tag attributes and tags of the input Schema file during parsing, to ease the process of extracting data from them.

Classes \code{SAXParser} and \code{SAXHelper} are very similar to their counterparts in \code{XSDImportDOM} module, for information about their methods please refer to the JavaDoc. Implementation is also relying on utility classes from \code{XSDImporter} module mentioned before.

\subsubsection{SAX Handler}

\section{Data Flow}
Parsing of XSD begins when \code{IGGeneratorImpl} calls \code{XSDProcessor} with an input file.
The overall data flow while parsing a Schema document is simple, it can be summarized into following steps.
\begin{enumerate}
	\item \code{XSDProcessor} finds out the selected \code{XSDParser} from available service providers. This also indicates which module will be handling the actual parsing.
	\item \code{XSDProcessor} passes the input document to a selected module by calling \code{parse()} method on the selected \code{XSDParser}.
	\item Depending on the module, the corresponding API is used to extract IG rules from the input document.
	\item List of rules that were extracted from the document are returned to \code{XSDProcessor}.
	\item If no errors occured during this process, list of extracted rules is returned to \code{IGGeneratorImpl} for expansion. Otherwise, depending on the \emph{Print verbose info} setting, the inference is stops, or an empty list of rules for this input document is returned.
\end{enumerate}

\subsection{Data Flow using DOM parser}
When using the DOM parsing method, extraction of rules from a Schema document, mentioned in step 3 above, proceeds as follows.
\begin{enumerate}
	\item Document Object Model tree data structure is created from the input file by \emph{Xerces DOM parser}.
	\item Root node of the DOM tree is handed over to \code{DOMHandler} class.
	\item \code{DOMHandler} recursively passes the DOM tree by a depth-first algorithm and creates complete subtrees of rules (rule-trees) for each of the direct children of root node.
	\item Rule-trees are returned from \code{DOMHandler}.
	\item \code{DOMParser} traverses the returned rule-trees depth-first to create a list of final rules. Each of the rules now contains a full path, thanks to the way in which they were created.
\end{enumerate}

\subsection{Data Flow using SAX parser}

Due to the nature of SAX parser paradigm, it is not possible to create the tree data structure in the top-to-bottom fashion, as with DOM parser. However, using stack data structures, analogous rule-trees are created from bottom to top, leading to a similar data flow.

\begin{enumerate}
	\item A \code{SAXParserFactory} is created and our implementation of handler class \code{SAXHandler} is passed to it, along with \code{InputStream} containing an input document.
	\item Our handler then sequentially responds to events generated by the SAX parser, creating \code{Element} nodes when appropriate, and keeping them on a stack. This way, a relation can be created by adding an \code{Element} as a child to the \code{Element} on top of the stack.
	\item When parsing is finished, the rule-trees are kept in the data structures of \code{SAXHandler}.
	\item \code{SAXParser} calls \code{getRules()} method of \code{SAXHandler} to traverse the rule-trees to create the list of final rules in a similar manner as in DOM case. These rules also contain a full path.
\end{enumerate}

\section{Known issues and limitations}

In this section we describe known problems that can occur while importing Schema documents. Supported structures are discussed only; tags and tag attributes that are not supported are not in scope of this section.

\subsection{SAX parser limitations}

Due to the inherent difficulty of SAX parsing method to cope with cross-referencing definitions, there is a known issue when schema defines \emph{complexType} tags in the following manner.

\begin{verbatim}
<xsd:complexType name="folderType">
  <xsd:sequence>
    <xsd:element name="file" type="fileType" />
  </xsd:sequence>
</xsd:complexType>

<xsd:complexType name="fileType">
  <xsd:attribute name="ID" type="xsd:ID"/>
</xsd:complexType>

<xsd:element name="folder" type="folderType" />
\end{verbatim}

In this example definition, the \code{folderType} type is dependent on \code{fileType} type, which has not been defined when a new Element, named \code{file}, is trying to resolve its type. This results in the parser assigning empty children to the \code{file} Element, adding the Element to a list of unresolved Elements and continuing with the parsing process. When parser reads the complete definition of \code{folderType} type, the type is only added to a list of known types, as there are no data structures for listing type cross-dependencies.

When the complete definition of \code{fileType} is read, this type is also only added to a list of known types, no dependencies are retrospectively resolved.
The reason for it lies in one-pass approach of SAX parser. In future versions, we are planning to resolve this issue by creating dependency trees of \code{complexType} tag definitions. However, it will make the parsing algorithm more time and memory expensive.

Known fix is to place definitions of dependent types after the types they depend on have been defined, or simply using DOM parser, which is devoid of such complications by design. Using either will result in a desired output:

\begin{verbatim}
<xs:complexType name="Tfile">
  <xs:attribute name="ID" type="xs:ID"/>
</xs:complexType>

<xs:complexType name="Tfolder">
  <xs:sequence>
    <xs:element name="file" type="Tfile"/>
  </xs:sequence>
</xs:complexType>

<xs:element name="folder" type="Tfolder"/>
\end{verbatim}

\section{Settings} \label{sect:settings}

Common settings of modules can be found in project properties under \emph{XSD Import} category. They are project specific. Following list explains their purpose.

\begin{description}
	\item[Parsing method] Setting selects whether SAX or DOM parser is used. 
This selection is based on the number of registered service providers for the \code{XSDParser} class,
so that both the jInfer bundled parsers as well as any custom parsers will be automatically displayed here.
DOM parser is set as the default parsing method. 
	\item[Log level] Setting determines the minimal priority for a message from the selected parser module to be logged. For example, if the module provides extensive debug information about elements being currently parsed, these messages can be filtered by setting the log level higher than \code{DEBUG}.
	\item[Stop on error] Checkbox determines if parsing halts the inference on the first error. If it is not set, then the file that caused the error is simply skipped and not imported. 
	\item[Print verbose info] Setting enables logging of some additional information, for example, complete list of imported rules, or extended warning messages.
\end{description}

\section{Extensibility}

Module \code{XSDImporter} itself is an example of an extension of \code{BasicIGG}, for which it provides a specific funcionality. Other modules that handle different definition languages can be created in a similar fashion.

The modular design of the parser implementations is also an example of extensibility which can be achieved using the NetBeans platform. It allows developers to add, or remove, custom parsers just by creating a module with a new parser implementation; there is no need to change any code within \code{XSDImporter} module. Only exception is the re-definition of the default parser (currently DOM), which is not needed even if the module of default parser is not present.

Parsers with extended or custom functionality, can be created using \code{XSDImportDOM} module as example (recomended).

\nocite{*}
\newpage
\bibliographystyle{alpha}
\bibliography{literature}

\end{document}

### 2011-03-01 doc/jInfer/runner/runner.tex
propset ignore and package names corrected, nocite corrected
3187368 - Review documentation: cleanup - runner done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Matej Vitásek \\ \hline
-Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator \\
- & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator \\
- & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\ 
- & org.openide.windows.IOProvider \\ \hline
-Provided tokens:       & none \\ \hline
-Module dependencies:   & Base \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.runner \\
-  & cz.cuni.mff.ksi.jinfer.runner.properties \\ \hline
+Responsible developer & Matej Vitásek \\ \hline
+Required tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator} \\
+ & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator} \\
+ & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier} \\ 
+ & \code{org.openide.windows.IOProvider} \\ \hline
+Provided tokens       & none \\ \hline
+Module dependencies   & Base \\ \hline
+Public packages       & \code{cz.cuni.mff.ksi.jinfer.runner} \\
+  & \code{cz.cuni.mff.ksi.jinfer.runner.properties} \\ \hline
 \end{tabular}
@@
 
-\jmodule{Runner} it the module responsible for inference process.
-The fact that the inference consists of exactly 3 steps as described in TODO link architecture is hard-wired in jInfer via this module.
+\jmodule{Runner} it the module responsible for inference process. The fact that the inference consists of exactly 3 steps as described in \cite{archdoc}, is hard-wired in jInfer via this module.
 
@@
 The only public method in \code{Runner} is, unsurprisingly, \code{run()}.
-This method will start the first step of the inference process by invoking the selected \jmodule{IGGenerator}'s \code{start()} method, as described in TODO link inference process in architecture.
+This method will start the first step of the inference process by invoking the selected \jmodule{IGGenerator}'s \code{start()} method, as described in \cite[section 3.5.1]{archdoc}.
 Callback methods \code{finishedIGGenerator()}, \code{finishedSimplifier()} and \code{finishedSchemaGenerator()} are responsible for invoking the following stages of inference, or in the latter case for presenting the resulting schema to the user and terminating the inference process.\\

### 2011-03-01 doc/jInfer/treeruledisplayer/treeruledisplayer.tex
Tree RD documentation: fixing a minor problem.
3187368 - Review documentation: cleanup - treeruledisplayer done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Michal Švirec \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
-Module dependencies:   & Base \\
+Responsible developer & Michal Švirec \\ \hline
+Required tokens       & none \\ \hline
+Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} \\ \hline
+Module dependencies   & Base \\
 	& JUNG \\ \hline
-Public packages:       & none \\ \hline
+Public packages       & none \\ \hline
 \end{tabular}
@@
 While the code building tree structure is a nice recursion programming exercise, there is nothing of a special interest in it.
-Finally the method appends forrest into layout and VisualizationViewer which were described above.
+Finally the method appends forrest into layout and \code{VisualizationViewer} which were described above.
 All code responsible for building rule trees is contained in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.logic}.

### 2011-03-01 doc/jInfer/twostep/twostep.tex
3187368 - Review documentation: cleanup - twostep done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Michal Klempa \\ \hline
-Required tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer \\ \hline
-Provided tokens:       & cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier \\ \hline
-Module dependencies:   & AutoEditor, Base, JUNG, Lookup \\ \hline
-Public packages:       & none \\ \hline
+Responsible developer & Michal Klempa \\ \hline
+Required tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} \\ \hline
+Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier} \\ \hline
+Module dependencies   & AutoEditor, Base, JUNG, Lookup \\ \hline
+Public packages       & none \\ \hline
 \end{tabular}
@@
 It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
-type \code{T} (that is representative of some main cluster) has some "attribute" clusters associated with it.
+type \code{T} (that is representative of some main cluster) has some ``attribute'' clusters associated with it.
 Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.
@@
 \subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
-We cluster elements in \code{Iname<AbstractStructuralNode, Attribute>} class (Iname je clusterer!), which takes \code{AbstractStructuralNode} classes to cluster as main, and \code{Attribute} classes as attributes.
-% TODO anti I don't understand the sentence, rewrite.
+We cluster elements in \code{Iname\.<Abstract\.Structural\.Node, Attri\.bu\.te>} class, which takes \code{Abstract\.Structural\.No\.de} classes to cluster as main, and \code{Attribute} classes as attributes.
 Clustering is done based on element's \code{getName()} equality (ignoring case).
@@
 So there can't be cluster with only one sentinel element in it.
-% TODO anti Please rewrite the really long sentence.
   \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule.
-% TODO anti Totally unclear, rewrite
 \end{itemize}
@@
 \subsection{Trie (ClusterProcessor)} \label{section_Trie}
-This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie") of them.
+This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie'') of them.
 More precisely, it takes the first rule and declares it to be a long branch (concatenation of tokens) in a newly created tree.
 After that, it adds the remaining rules one by one as branches like this: as long as it can follow an existing branch, it follows it.
-As soon as the newly added branch starts to differ, it ``branches off" (creates an alternation at that point) the existing tree and hangs the rest of the newly added rule there.
+As soon as the newly added branch starts to differ, it ``branches off'' (creates an alternation at that point) the existing tree and hangs the rest of the newly added rule there.
 Repeating this process creates a prefix tree describing all the rules in the cluster.
@@
 	\item Create prefix-tree automaton (PTA) from positive examples - right sides of rules.
-Call its submodule \jmodule{AutomatonSimplifier} to modify PTA to some generalized automaton by merging states.
+Call its submodule \jmodule{Au\.to\.ma\.ton\.Simpli\.fier} to modify PTA to some generalized automaton by merging states.
 	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.

### 2011-03-01 doc/jInfer/xsdimport/xsdimport.tex
Fixed XSDImport according to first annotations.

@@
 Every vertex is a rule by itself, so in order to obtain IG rules, one must extract all Elements from a rule-tree.
-Rule-trees are created by \code{DOMHandler}, described in section \ref{sect:domhandler}.
-The extraction is done by \code{DOMParser}, described in section \ref{sect::domparser}.
+Rule-trees are created by \code{DOMHandler}, described in subsection \ref{sect:domhandler}.
+The extraction is done by \code{DOMParser}, described in subsection \ref{sect:domparser}.
 Or alternatively, when parsing with SAX is selected, \code{SAXHandler} handles both processes.
+
+Life cycle of rule-trees is apparent from section \ref{sect:dataflow}.
 
@@
 
-This module has several key roles in the process of importing XSD documents.\\
+\noindent This module has several key roles in the process of importing XSD documents.\\
 
@@
 
-Simmilar to the \code{Processor} implementations in the \code{BasicIGG} module, this is the core class of \code{XSDImporter} logic.
-XSDProcessor registers itself into \emph{schema} input folder, \emph{xsd} file extension and declares it cannot handle other file extensions.\\
+Similar to the \code{Processor} implementations in the \code{BasicIGG} module, this is the core class of \code{XSDImporter} logic.
+\code{XSDProcessor} registers itself into \emph{schema} input folder, \emph{xsd} file extension and declares it cannot handle other file extensions.\\
 Class depends on NBP lookups (see \cite[section 4.1]{archdoc}) to find available parsers, and uses the selected parser to perform the import operation on defined files.
@@
 \begin{verbatim}
-List<Element> parse(final InputStream stream) throws XSDException;
+List<Element> parse(InputStream stream) throws XSDException;
 \end{verbatim}
@@
 
-- TODO explain the purpose of utility classes \\
-- TODO describe every helper class in importer?
+\paragraph{Built-in data types}
+
+A special class is designed to contain all built-in data types supported by XSD language.
+Both built-in primitive types and built-in derived types are recognized.
+For a complete list, please see our JavaDoc documentation for \code{XSDBuiltInDataTypes} class.
+These types are contained using instances of \code{SimpleData} class, as described in \cite[section 2.2]{archdoc}.
 
@@
 \\
-This module implements the import functionality using API from Xerces DOM parser.
+
+\noindent This module implements the import functionality using API from Xerces DOM parser.
 Its structure is very simple, all classes are grouped in one main package, relying on utility classes from \code{XSDImporter} module mentioned before.
@@
 
-BuildRuleSubtree is the most important method. It is divided into four logical parts
+The most important method is \code{buildRuleSubtree()}. It is divided into four logical parts.
 \begin{enumerate}
@@
 \end{enumerate}
+
+Signature of the method is:
+\begin{verbatim}
+private Pair<Element, RegexpInterval> buildRuleSubtree(
+	org.w3c.dom.Element currentNode,
+	List<String> context,
+	List<org.w3c.dom.Element> visited) throws XSDException
+\end{verbatim}
+
+Method takes current node, which will be returned as root \code{Element} node (first value of the returned pair) of this subtree, and creates an entire rule-tree below it by recursively calling itself, as mentioned in step 3.
+Context parameter holds the list of names of nodes that create unique path from actual root of the schema document to the current node.
+Visited parameter lists references to instances of \code{Element} that were visited on this path.
+When handling \emph{attribute} tags, context needs to be passed along, but list of visited nodes does not.
 
@@
 \\
-This module implements the import functionality using API of a standard SAX parser.
+
+\noindent This module implements the import functionality using API of a standard SAX parser.
 Its structure is divided in two packages.
@@
 \subsubsection{SAX Handler}
+
+Sax handler implements methods in response to the events generated by the underlying SAX parser API.
+Similar to its counterpart, \code{DOMHandler}, this class holds most on the logic for importing a Schema.
+It works by keeping current tree branch, which is being parsed, on a stack. Adding current node to the parent on top of the stack as current tag finishes.
 
@@
 
-\subsection{Data Flow using SAX parser}
+\subsection{Data Flow using SAX parser} \label{sect:dataflow}
 
@@
 
-\section{Known issues and limitations}
+\section{Known issues and limitations} \label{sect:issues}
 
@@
 In future versions, we are planning to resolve this issue by creating dependency trees of \code{complexType} tag definitions.
-However, it will make the parsing algorithm more time and memory expensive.
+However, it will make the parsing algorithm more time and memory expensive.\\
 
@@
 
-\section{Extensibility}
+\section{Extensibility} \label{sect:extensibility}
 

### 2011-03-01 doc/jInfer/xsdimport/xsdimport.tex
Fixed XSDImport according to second annotations.

@@
 
-None of the modules provide nor require any tokens. Complete dependency tree is depicted in Figure 9 and described in \cite[section 4]{archdoc}.
+None of the modules provide nor require any tokens. Complete dependency tree is described in \cite[section 4]{archdoc} and depicted in Figure 9 of that document.
 
@@
 And finally, this module holds common utility classes related to the task of processing data characteristic to the Schema documents. \\
-Functionality is divided between separate packages, in the same order:
+Functionality is divided between separate packages, in the same order as above:
 
@@
 Class depends on NBP lookups (see \cite[section 4.1]{archdoc}) to find available parsers, and uses the selected parser to perform the import operation on defined files.
-Initial grammar generated by this processor is considered \emph{complex} and should be followed by expansion (see \cite[section 4]{basiciggdoc}).
+Initial grammar generated by this processor is considered \emph{complex} and should be followed by expansion if necessary (see \cite[section 4]{basiciggdoc}).
 
@@
 
-Package \code{cz.cuni.mff.ksi.jinfer.xsdimporter.utils} contains classes available for all parsing methods.
+Package \code{cz.cuni.mff.ksi.jinfer.xsdimporter.utils} contains classes available to all parsing methods.
 
@@
 	\item \code{logLevel()} returns current log priority, messages with lower priority are not displayed.
-	\item \code{isVerbose()} returns \code{true} or \code{false}, depending on the checkbox.
-	\item \code{stopOnError()} returns \code{true} or \code{false}, depending on the checkbox.
+	\item \code{isVerbose()} returns \code{true} or \code{false}, depending on state of the checkbox.
+	\item \code{stopOnError()} returns \code{true} or \code{false}, depending on state of the checkbox.
 \end{itemize}
@@
 
-Enumeration of tag attributes currently recognized by our parsers. They are the following:
+Enumeration of tag attributes currently recognized by our parsers. They are the following.
 \begin{itemize}
@@
 
-Enumeration of tags currently known by our parsers. They are the following:
+Enumeration of tags currently known by our parsers. They are the following.
 \begin{itemize}
@@
 This is basically only a library of short, useful and frequently used functions for parsing, please refer to the JavaDoc.
+
+\paragraph{XSD Exception}
+
+\code{XSDException} is a simple class extending \code{RuntimeException} to announce errors during import operation.
+It is catched by \code{XSDProcessor} which checks if \emph{Stop on error} setting is enabled and stops the inference.
 
@@
 Helper class for \code{DOMHandler}, containing ulility functions specific for DOM parsing method.
-Two most notable of them are \code{createSentinel} and \code{finalizeElement}.
+Two most notable of them are \code{createSentinel()} and \code{finalizeElement()}.
 
 \begin{description}
-	\item[CreateSentinel] Creates an instance of \code{Element} class, which is always a leaf in the rule-tree.
+	\item[CreateSentinel()] Creates an instance of \code{Element} class, which is always a leaf in the rule-tree.
 This is useful when there are cyclic definitions in an input Schema document, or an \code{Element} with this name was defined previously and it is unnecessary to build a rule subtree for it again (it is a reference in this case).
 More information about sentinels can be found in \cite[section 3.2]{archdoc}, or \cite[section 4]{twostepdoc}.
-	\item[FinalizeElement] Checks if \code{Element} is properly defined.
+	\item[FinalizeElement()] Checks if \code{Element} is properly defined.
 Redefine it to \code{LAMBDA} when it was empty, or to token if it only contained a simple data type.
@@
 	\item If no errors occured during this process, list of extracted rules is returned to \code{IGGeneratorImpl} for expansion.
-Otherwise, depending on the \emph{Print verbose info} setting, the inference is stops, or an empty list of rules for this input document is returned.
+Otherwise, depending on the \emph{Stop on error} setting, the inference is stopped, or an empty list of rules for this input document is returned.
 \end{enumerate}
@@
   <xsd:sequence>
-    <xsd:element name="file" type="fileType" />
+    <xsd:element name="file" type="fileType"/>
   </xsd:sequence>
@@
 
-<xsd:element name="folder" type="folderType" />
+<xsd:element name="folder" type="folderType"/>
 \end{verbatim}
@@
 It allows developers to add, or remove, custom parsers just by creating a module with a new parser implementation; there is no need to change any code within \code{XSDImporter} module.
-Only exception is the re-definition of the default parser (currently DOM), which is not needed even if the module of default parser is not present.
-
-Parsers with extended or custom functionality, can be created using \code{XSDImportDOM} module as example (recomended).
+Only exception is the re-definition of the default parser (currently DOM), which is not needed even if the module of default parser is not present.\\
+Parsers with extended or custom functionality can be created using \code{XSDImportDOM} module as example (recomended).
 

### 2011-03-01 doc/jInfer/xsdimport/xsdimport.tex
3187368 - Review documentation: cleanup - xsdimport done

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Responsible developer: & Robert Smetana \\ \hline
-Required tokens:       & none \\ \hline
-Provided tokens:       & none \\ \hline
-Module dependencies:   & listed for each module separately \\ \hline
-Public packages:       & listed for each module separately \\ \hline
+Responsible developer & Robert Smetana \\ \hline
+Required tokens       & none \\ \hline
+Provided tokens       & none \\ \hline
+Module dependencies   & listed for each module separately \\ \hline
+Public packages       & listed for each module separately \\ \hline
 \end{tabular}
@@
 Similar to the \code{Processor} implementations in the \code{BasicIGG} module, this is the core class of \code{XSDImporter} logic.
-\code{XSDProcessor} registers itself into \emph{schema} input folder, \emph{xsd} file extension and declares it cannot handle other file extensions.\\
+\code{XSD\.Pro\.cessor} registers itself into \emph{schema} input folder, \emph{xsd} file extension and declares it cannot handle other file extensions.\\
 Class depends on NBP lookups (see \cite[section 4.1]{archdoc}) to find available parsers, and uses the selected parser to perform the import operation on defined files.
@@
 Package \code{cz.cuni.mff.ksi.jinfer.xsdimportsax} holds the main parsing logic, while package \code{cz.cuni.mff.ksi.jinfer.xsdimportsax.utils} contains two utility classes.
-Both \code{SAXAttributeData} and \code{SAXDocumentElement} serve as wrappers for tag attributes and tags of the input Schema file during parsing, to ease the process of extracting data from them.
+Both \code{SAXAttributeData} and \code{SAX\.Do\.cu\.ment\.Ele\.ment} serve as wrappers for tag attributes and tags of the input Schema file during parsing, to ease the process of extracting data from them.
 

### 2011-03-01 doc/jInfer/xsdimport/xsdimport.tex
"figure"

@@
 
-None of the modules provide nor require any tokens. Complete dependency tree is described in \cite[section 4]{archdoc} and depicted in Figure 9 of that document.
+None of the modules provide nor require any tokens. Complete dependency tree is described in \cite[section 4]{archdoc} and depicted in figure 9 of that document.
 
@@
 Lastly, tag attributes should not be confused with instances of \code{Attribute}, which is a class for storing information about the tag attributes inside jInfer.
-Every \code{Element} has a list of \code{Attribute}s, as can be seen from Figure 4 in \cite[section 2.3]{archdoc}.
+Every \code{Element} has a list of \code{Attribute}s, as can be seen from figure 4 in \cite[section 2.3]{archdoc}.
 
@@
 
-Diagram showing the inheritance of this interface, as well as both parsers bundled with jInfer, which implement it is in Figure \ref{parser_interface}.
+Diagram showing the inheritance of this interface, as well as both parsers bundled with jInfer, which implement it is in figure \ref{parser_interface}.
 

### 2011-03-01 doc/jInfer/twostep/twostep.tex
Fixed some sentences from previous version.

@@
 \section{Introduction}
-In \jmodule{TwoStepSimplifier} we provide user a complex solution to grammar simplification problem.
+In \jmodule{TwoStepSimplifier} we provide the user with a complex solution to grammar simplification problem.
 While this is the only simplifier bundled with jInfer, its strong modularity ensures that it can be extended easily. 
@@
 Strong emphasis is put on generic automata, therefore it is easy to plug-in any automaton merging state algorithm.
-In this document, we will walk over all modules and submodules in simplifier and describe them briefly.
+In this document, we will list all modules and submodules of the simplifier and describe them briefly.
 
@@
 \section{Factory pattern usage} \label{section_factory}
-Lets stay a while at factory pattern usage.
-In \jmodule{TwoStepSimplifier} we employ factory pattern (as described in \cite[section 4.2]{archdoc}) to divide module into submodules.
+In \jmodule{TwoStepSimplifier} we employ a factory pattern (as described in \cite[section 4.2]{archdoc}) to divide the module into submodules.
 Since service providing classes are being kept as singletons in the NB platform, we use them as factories:
@@
 \end{verbatim}
-The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (of which instance is returned by factory).
-For now, it is not important how the worker interface looks like, lets just examine the factory.
+The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (an instance of which is returned by the factory).
+For now, it is not important what the worker interface looks like, lets just examine the factory.
 
-In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
-That means, modules must implement methods:
+In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces 
+so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
+Therefore, the modules must implement following methods:
 \begin{verbatim}
@@
 
-Under certain circumstances it is useful to declare method \code{create()} generic.
-The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic.
-But automaton simplification does not depend on type of symbol of automaton transitions, so interface \code{AutomatonSimplifier<T>}
-generic too.
-Factory interface deals with this by defining the \code{create()} method generic.
+The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic;
+but automaton simplification does not depend on any particular type (of the generic symbol) for automaton transitions, 
+so interface \code{AutomatonSimplifier<T>} is generic too.
+Factory interface deals with this by defining the \code{create()} method also as generic.
 
@@
 
-Lets look at \code{AutomatonMergingStateFactory}. This is factory of module which has \jmodule{AutomatonSimplifier} as a submodule.
+Lets look at \code{AutomatonMergingStateFactory}. This is the factory of a module, which has \jmodule{AutomatonSimplifier} as a submodule.
 Its create method (shortened): 
@@
 \end{verbatim}
-Cluster processor \code{AutomatonMergingState} then receives factories of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
-Cluster processor then may create as many instances of submodule classes as it needs (maybe simplifying more than one automaton).
-Thorough this document, we will mention only worker interface when describing submodule, since all factory interfaces are designed the same way as was just described.
+Cluster processor \code{AutomatonMergingState} then receives factories 
+of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
+Cluster processor then may create as many instances of submodule classes as it needs (it may be simplifying more than one automaton).
+Thorough this document, we will only mention the worker interface when describing a submodule, 
+since all factory interfaces are designed the same way as described previously.
 
@@
 Module is divided into two classes: \code{TwoStepSimplifier} (main logic) and \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
-The latter is registered as service provider (and implements) of \code{Simplifier} interface (defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
+The latter is registered as service provider of (and implements the) \code{Simplifier} interface 
+(defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
 
-Its main method is \code{start()} which receives Initial Grammar in form of
+Its main method \code{start()} receives Initial Grammar in form of
 \begin{verbatim}
@@
 \end{verbatim}
-Each grammar rule is represented as class \code{Element}, where element is left side of rule and its \code{getSubnodes()} method
-returns regular expression representing right side of rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
-Second parameter is 
+Each grammar rule is represented by an instance of \code{Element} class. 
+Element itself represents the left side of the rule, and its \code{getSubnodes()} method returns 
+regular expression representing the right side of the rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
+Its second parameter is 
 \begin{verbatim}
@@
 \end{verbatim}
-Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented same way - list of elements.
+Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented the same way, as a list of elements.
 
@@
 Then it calls \code{simplify(initialGrammar)} method of this instance and its result is passed as parameter to callback function.
-That's all, the magic is in \code{TwoStepSimplifierFactory} class.
+And that's all the magic of \code{TwoStepSimplifierFactory} class.
 
@@
 \end{enumerate}
-Clustering is delegated to \jmodule{Clusterer} submodule, and task of regular expression inference for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
-There is also third submodule called \jmodule{RegularExpressionCleaner} actually, its purpose is just to beautify
- output regular expressions, no inference logic is implemented there. 
-Modules are drawn on fig. \ref{twostep_modules}.
+Clustering is delegated to \jmodule{Clusterer} submodule, and the task of regular expression inference 
+for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
+There is also a third submodule called \jmodule{RegularExpressionCleaner}. 
+Its purpose is just to beautify output regular expressions, no inference logic is implemented there. 
+Modules are drawn on figure \ref{twostep_modules}.
 \begin{figure}[tpb]
@@
 \end{figure}
-We provide one \jmodule{Clusterer}, 4 \jmodule{ClusterProcessor} and 4 \jmodule{RegularExpressionCleaner} implementations.
+We provide one \jmodule{Clusterer}, four \jmodule{ClusterProcessor} and four \jmodule{RegularExpressionCleaner} implementations.
 Each of those will be explained further in this document.
@@
 First, it creates a clusterer, gives it all the rules and performs the clustering.
-Then, empty list of elements is created as the simplified grammar.
+Then, an empty list of elements is created as the simplified grammar.
 For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
-Finally, regular expression cleaning is done in a submodule and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.
+Finally, regular expression cleaning is done by a submodule 
+and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.
 
@@
 There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
-Sentinels can be only on right side of rules and they have to be sent to clusterer, as someone further in the chain can ask for cluster of sentinel element - which appears on right side.
-But sentinels have no content, they must not be taken into account when inferring attribute properties (whether it is required or optional) - since sentinels have no attributes at all.
+Sentinels can be only on the right side of rules. They have to be sent to clusterer, 
+as someone further in the chain can ask for cluster of a sentinel element - which appears on right side.
+However, sentinels have no content, and so they must not be taken into account when inferring attribute properties 
+(whether it is required or optional), since sentinels have no attributes at all.
 
-Now we will examine submodules for clustering, processing and cleaning.
+We will now examine the submodules for clustering, processing and cleaning.
 
@@
 One cluster is represented by the \code{Cluster<T>} class with \code{T} as type parameter of clustered items.
-This class simply holds a \code{Set} of members of this cluster and one of the references as \code{representant} member.
+This class simply holds a \code{Set} of members of the cluster, one of which is chosen as a \code{representant}.
 
-Any clusterer has to comply to \code{Clusterer} interface.
-Its purpose is to cluster bunch of elements (rules) on input, into bunch of \code{Cluster} class instances (clusters) on output.
-It has methods \code{add()} and \code{addAll()} for adding items for clustering.
+Any clusterer has to comply with the \code{Clusterer} interface.
+Its purpose is to cluster a bunch of elements (rules) on input into a bunch of \code{Cluster} class instances (clusters) on output.
+It has methods \code{add()} and \code{addAll()} for adding the input items to be clustered.
 Main part is the method \code{cluster()}, which does the clustering itself.
@@
 Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
-To be able to respond to calls to \code{T getRepresentantForItem(T item)} method, every clusterer implementation has to keep all its clusters even after the clustering is finished.
+To be able to respond to calls to the \code{T getRepresentantForItem(T item)} method, 
+every clusterer implementation has to keep all its clusters even after the clustering had finished. 
 Using this method, given an item the correct cluster representative for this item has to be returned.
@@
 If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
-Missing an item probably indicates error in algorithm rather than normal workflow.
+Missing an item probably indicates error in the algorithm, rather than normal workflow.
 One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
@@
 You might have noticed that the \jmodule{Clusterer} interface and \code{Cluster} class are generic.
-They may be used as design pattern not only for clustering elements in inference process.
+They may be used as design pattern not only for clustering elements in the inference process.
 To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
 It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
-type \code{T} (that is representative of some main cluster) has some ``attribute'' clusters associated with it.
+type \code{T} (that is a representative of some main cluster) has some ``attribute'' clusters associated with it.
 Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.
@@
 \jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression representing that set of elements.
-It returns one rule - element with name set to desired name of element in schema (not all elements in cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
-It processes attributes of all elements in cluster to obtain meaningful schema attribute specification and these attributes have to be attached to the resulting element.
+It returns one rule - element with name set to desired name of element in schema (not all elements in a cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
+It processes attributes of all elements in a cluster to obtain meaningful schema attribute specification. These attributes have to be attached to the resulting element.
 
@@
 Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
-Clusterer can tell you the representative for any element in the whole input (not only those elements in rules, but also those that may be on right side of rules).
+Clusterer can tell you the representative for any arbitrary element from input (not only those elements in rules, but also those that may be on right side of rules).
 Also, clusterer (if it is with attributes) has information about attributes in each cluster.
@@
 \subsection{RegularExpressionCleaner module}
-Finally we examine the \jmodule{RegularExpressionCleaner} interface.
-Purpose of this submodule is only to make output regular expression corrections, to make them nicer.
-For example it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
+In this last subsection, we examine the \jmodule{RegularExpressionCleaner} interface.
+Purpose of this submodule is only to make corrections to regular expression, to make them nicer on output.
+For example, it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
 To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
@@
 \end{verbatim}
-Given regular expression, return regular expression.
+Given a regular expression, return cleaned regular expression.
 Conversion of regular expression is commonly a recursive task and all our implementations work this way.
@@
 \end{figure}
-Process illustrated in fig. \ref{twostep_flow} is as follows.
+Process illustrated in figure \ref{twostep_flow} is as follows.
 \begin{enumerate}
@@
 Each main cluster also has \code{SimpleData} clusterer associated with it.
-Whole scheme is on fig. \ref{clusterer_structure}.
+Whole scheme is on figure \ref{clusterer_structure}.
 \begin{figure}[tpb]
@@
 \end{verbatim}
-The whole submodule structure is in fig. \ref{automaton_merging_state_modules}.
+The whole submodule structure is in figure \ref{automaton_merging_state_modules}.
 
@@
 \end{enumerate}
-The data flow is in fig. \ref{automaton_merging_state_flow}.
+The data flow is depicted in figure \ref{automaton_merging_state_flow}.
 
@@
 That results in cluster-equal behaviour.
-Example of prefix-tree automaton is in fig. \ref{pta}.
+Example of prefix-tree automaton is in figure \ref{pta}.
 
@@
 We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
-This is implemented in \code{StateRemoval} class (on fig. \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
+This is implemented in \code{StateRemoval} class (figure \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
 Before the algorithm starts, states \emph{superInitial} and \emph{superFinal} are added to automaton, former with $\lambda$-transition to initial state.
@@
 \end{enumerate}
-The whole process is step-by-step illustrated in fig. \ref{regaut}.
+The whole process is step-by-step illustrated in figure \ref{regaut}.
 \begin{figure}[pb]
@@
 
-Whole TwoStep submodules structure is in fig. \ref{twostep_modules_complete}.
+Whole TwoStep submodules structure is in figure \ref{twostep_modules_complete}.
 \begin{figure}[tpb]

### 2011-03-01 doc/jInfer/twostep/twostep.tex
Minor change in 2 sentences.

@@
 In \jmodule{TwoStepSimplifier} we employ a factory pattern (as described in \cite[section 4.2]{archdoc}) to divide the module into submodules.
-Since service providing classes are being kept as singletons in the NB platform, we use them as factories:
-For example:
+Since service providing classes are being kept as singletons in the NB platform, we use them as factories. For example:
 \begin{verbatim}
@@
 \subsection{ClusterProcessor module}
-\jmodule{ClusterProcessor} takes rules of one cluster of element and somehow obtains regular expression representing that set of elements.
-It returns one rule - element with name set to desired name of element in schema (not all elements in a cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
+\jmodule{ClusterProcessor} takes rules of one cluster of element and obtains a regular expression representing that set of elements.
+It returns one element, a rule, with name set to desired name of element in schema (not all elements in a cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
 It processes attributes of all elements in a cluster to obtain meaningful schema attribute specification. These attributes have to be attached to the resulting element.

### 2011-03-01 doc/jInfer/architecture/architecture.tex
Removed XSDImport module from list of dependants on RuleDisplayer token.

@@
 Another use of tokens is \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} representing a module providing a rule displayer.
-This token is provided by \jmodule{BasicRuleDisplayer} and \jmodule{TreeRuleDisplayer}, requested by \jmodule{BasicIGG}, \jmodule{XSDImporter} and \jmodule{TwoStepSimplifier}.
+This token is provided by \jmodule{BasicRuleDisplayer} and \jmodule{TreeRuleDisplayer}, requested by \jmodule{BasicIGG} and \jmodule{TwoStepSimplifier}.
 

### 2011-03-01 doc/jInfer/xsdimport/xsdimport.tex
xsdimport blueah

@@
 \noindent \begin{tabular}{|l|l|} \hline
-Module dependencies:   & Base \\ 
+Module dependencies   & Base \\ 
                        & BasicIGG \\ \hline
-Public packages:       & cz.cuni.mff.ksi.jinfer.xsdimporter.interfaces \\
-                       & cz.cuni.mff.ksi.jinfer.xsdimporter.utils \\ \hline
+Public packages       & \code{cz.cuni.mff.ksi.jinfer.xsdimporter.interfaces} \\
+                       & \code{cz.cuni.mff.ksi.jinfer.xsdimporter.utils} \\ \hline
 \end{tabular}
@@
 \noindent \begin{tabular}{|l|l|} \hline
-Module dependencies:   & Base \\ 
+Module dependencies   & Base \\ 
                        & XSDImporter \\ \hline
-Public packages:       & none \\ \hline
+Public packages       & none \\ \hline
 \end{tabular}

### 2011-03-01 doc/jInfer/basicdtd/basicdtd.tex
Minor corrections in some sentences. (1)

@@
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provi\-der is \code{Sche\.ma\.Ge\.ne\.ra\.tor\.Impl}.
-Its \code{start} method first topologically sorts all rules (elements) it got in the simplified grammar on input.
-This toposorting is necessary to avoid using anything not yet defined in the resulting schema.
+The main class implementing \code{SchemaGenerator} interface and simultaneously being registered as its service provi\-der is \code{Sche\.ma\.Ge\.ne\.ra\.tor\.Impl}.
+Its \code{start()} method first topologically sorts all rules (elements) it got in the simplified grammar on input.
+This sorting is necessary to avoid using anything not yet defined in the resulting schema.
 Afterwards, it creates their DTD string represenation.\\
@@
 Export of a single element is handled in the \code{elementToString()} method.
-First the actual \code{<!ELEMENT \ldots>} tag is exported, after that its attributes in a \code{<!ATTLIST \ldots>} tag (if there are any).
+First the actual \code{<!ELEMENT \ldots>} tag is exported, and after that, its attributes in \code{<!ATTLIST \ldots>} tag (if there are any).
 
 \subsection{Element content export}
-Elements are processed \code{elementToString()} method.
-Whole element regexp is sent to method \code{expandIn\.ter\.vals\.Reg\.exp} of class \code{IntervalExpander}.
-Its purpose is to convert intervals on regexp and its children to those representable in DTD.
+Elements are processed by \code{elementToString()} method.
+Whole regexp of an element is sent to method \code{expandIn\.ter\.vals\.Reg\.exp()} of class \code{IntervalExpander}.
+Its purpose is to convert intervals of a regexp and its children to those representable in DTD.
 For example regular expression $(a\{2,5\}, b\{0,2\})$ would be transformed to $(a,a,a?,a?,a?,b?,b?)$.
+
 Class \code{IntervalExpander} works recursively.
-Regexp is passed to method \code{expandIntervalsRegexp} to handle regexp in that element.
+Regexp is passed to method \code{expandIntervalsRegexp()} to handle regexp in that element.
 This method contains a big\code{switch} based on the regexp type.
@@
 For $\lambda$, it simply returns \code{EMPY} as string.
-Tokens are first examined if they are \code{SimpleData}, if so, string \code{\#PCDATA} is returned.
+Tokens are at first examined if they are \code{SimpleData}; if so, string \code{\#PCDATA} is returned.
 If not, element name is returned.
-If interval of this regexp is different from $\{1, 1\}$, the interval \code{toString()} representation is appended.
+If interval of this regexp is different from $\{1, 1\}$, the interval \code{toString()} representation is appended.\\
 

### 2011-03-02 doc/jInfer/basicigg/basicigg.tex
corrected () after methods

@@
 The main class implementing \code{IGGenerator} inference interface and simultaneously registered as its service provider is \code{IGGeneratorImpl}.
-In its \code{start} method it enumerates all files in the input parameter (of type \code{Input}).
+In its \code{start()} method it enumerates all files in the input parameter (of type \code{Input}).
 For each file, based on its extension the correct \emph{processor} is selected, executed and returned rules are aggregated.
 Rules coming from XML documents are verified to be \emph{simple} grammar representation.
-After each file has been processed, the resulting grammar is expanded if necessary and returned by invoking the \code{finished} method of the callback argument.
-This process is illustrated in fig.
-\ref{figure-process}.
+After each file has been processed, the resulting grammar is expanded if necessary and returned by invoking the \code{finished()} method of the callback argument.
+This process is illustrated in figure \ref{figure-process}.
 

### 2011-03-02 doc/jInfer/tutorials/building_jinfer.tex
tutorial pdfs

--- 2011-01-14 doc/jInfer/building-jinfer.tex
@@
 \documentclass[a4paper,10pt,oneside]{article}
-\usepackage{epsfig}
+\usepackage{graphicx}
 \usepackage{color}
 \usepackage{url}
+\usepackage{subfigure}
 \usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{import}
+\usepackage{tgpagella}
+%\usepackage[scale=0.9]{tgcursor}
+%\usepackage[scale=0.9]{tgheros}
+\usepackage{xstring}
+\usepackage{wrapfig}
 
@@
 \newcommand{\vect}[1]{\boldsymbol{#1}}
+\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
+\def\.{\discretionary{}{}{}}
+\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
+
 \setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
@@
 \setlength{\topmargin}{1.5cm}
-\setlength{\textheight}{26cm}
+\setlength{\textheight}{24cm}
 \setlength{\textwidth}{18cm}
 
-\def\mftitle{Building jInfer}
+\def\mftitle{Building jInfer from sources}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \begin{document}
+
 \maketitle
 
-Target audience: jInfer power users. Anyone who needs to build jInfer from sources.\\
+ \section*{Building jInfer from sources}
+\par 
+  Target audience: jInfer power users. Anyone who needs to build jInfer from sources.
 
-\textit{This document assumes that your computer is set to use jInfer, meaning you have correct version of NetBeans installed and you can run jInfer.}
+\par 
+  This document assumes that your computer is set to use jInfer, meaning you
+    have correct version of NetBeans installed and you can run jInfer.
 
-\section{Getting jInfer sources}
-There are two main ways of obtaining jInfer source files: from a version distribution, or from the Subversion repository.
-The first way might be more convenient and secure (version distributions are usually more stable than the bleeding edge code in repository), however you might prefer to build the latest code.\\
+\subsection*{Getting jInfer sources}
+\par 
+There are two main ways of obtaining jInfer source files: from a version
+  distribution, or from the Subversion repository. The first way might be more
+  convenient and secure (version distributions are usually more stable than the
+  bleeding edge code in repository), however you might prefer to build the
+  latest code.
 
+\par 
   The first way, obtaining sources from version distribution:
-\begin{enumerate}
-	\item Go to https://sourceforge.net/projects/jinfer/files/
+
+\begin{enumerate}\item 
+ Go to 
+https://sourceforge.net/projects/jinfer/files/.
   \item Download the latest jInfer sources (jInfer-X.Y-src.zip).
-	\item Unpack the ZIP, you’ll get a few directories. One of them will be \texttt{src}.
+  \item Unpack the ZIP, you'll get a few directories. One of them will be \texttt{src}.
   \item Continue by building jInfer.
   \end{enumerate}
-
-Second way, checking out Subversion repository: you will need a Subversion client, such as the command line client or TortoiseSVN.
-In any case, http://subversion.apache.org/ might be a good place to start.
-To check out jInfer, use the following URL:\\
-
-\texttt{https://jinfer.svn.sourceforge.net/svnroot/jinfer}
-
+\par 
+  Second way, checking out Subversion repository: you will need a
+  Subversion client, such as the command line client or TortoiseSVN. In any case,
+  \url{http://subversion.apache.org/}
+might be a good place to start. To check out
+  jInfer, use the following URL:
+  https://jinfer.svn.sourceforge.net/svnroot/jinfer
   After the checkout, continue by building jInfer.
 
-\section{Building from NetBeans}
-\textit{This is perhaps the easier, more user friendly way of building jInfer.
-If you however wish to automate the building process, you might want to try using Ant - in that case, see the next section.}
+\subsection*{Building from NetBeans}
+\par 
+  This is perhaps the easier, more user friendly way of building jInfer.
+    If you however wish to automate the building process, you might want to
+    try using Ant - in that case, see the next section.
 
-\begin{enumerate}
-	\item Start NetBeans, select \textit{File} > \textit{Open Project...} from the main menu.
-	\item In the dialog window, navigate to the root folder with jInfer sources. Its icon should look like two brown puzzle pieces.
+\begin{enumerate}\item 
+ Start NetBeans, select \textit{File} > \textit{Open Project...} from the main menu.
+  \item In the dialog window, navigate to the root folder with jInfer sources.
+    Its icon should look like two brown puzzle pieces.
   \item You should see the name "jInfer" in the dialog window.
@@
   \item You might want checking the \textit{Open as Main Project} checkbox.
-	\item Click \textit{Open Project}. After a while, all jInfer projects should open and be visible in the \textit{Projects} window.
-\end{enumerate}
+  \item Click \textit{Open Project}. After a while, all jInfer projects should
+    open and be visible in the \textit{Projects} window.
+  \end{enumerate}\par At this moment, you might want to look around, play with the
+  code a bit or apply some patches. If you want to try modified jInfer in a
+  running NetBeans, select \textit{Run} from project jInfer (two brown puzzle
+  pieces icon) context menu. If jInfer is the main project, clicking the
+  green "Play" icon in the toolbar should work too.
+  If you want to build jInfer, for example to test if your changes didn't break
+  something, select \textit{Build} from the context menu. Sometimes it might be
+  necessary to run \textit{Clean \& Build}, try this if it "stops working".
+  If you want to run JUnit test, select \textit{Test} from the same context menu.
+  If you now want to create your own distribution (= NBM modules), select
+  \textit{Create NBMs} from the context menu. They will be created in
+  \textit{build/updates} directory.
+\subsection*{Building with Ant}
+\par 
+  Each jInfer module and jInfer suite itself has
+  its own \texttt{build.xml} Ant build file. The most important one, for the
+  whole jInfer suite, is located in the \texttt{src} folder.
+  It is possible to use vanilla Ant from Apache, but perhaps safer to resort
+  to the Ant bundled with NetBeans. This one should be located in \textit{NetBeans
+    install directory} > \texttt{java/ant/bin/ant}.
+  Whichever Ant is chosen, to build jInfer do the following on command line:
 
-At this moment, you might want to look around, play with the code a bit or apply some patches.
-If you want to try modified jInfer in a running NetBeans, select \textit{Run} from project jInfer (two brown puzzle pieces icon) context menu.
-If jInfer is the main project, clicking the green "Play" icon in the toolbar should work too.\\
-
-If you want to build jInfer, for example to test if your changes didn't break something, select \textit{Build} from the context menu.
-Sometimes it might be necessary to run \textit{Clean \& Build}, try this if it "stops working".
-If you want to run JUnit test, select \textit{Test} from the same context menu.\\
-
-If you now want to create your own distribution (= NBM modules), select \textit{Create NBMs} from the context menu. They will be created in \textit{build/updates} directory.
-
-\section{Building with Ant}
-
-Each jInfer module and jInfer suite itself has its own \texttt{build.xml} Ant build file. The most important one, for the whole jInfer suite, is located in the \texttt{src} folder.\\
-
-It is possible to use vanilla Ant from Apache, but perhaps safer to resort to the Ant bundled with NetBeans.
-This one should be located in \textit{NetBeans install directory} > \texttt{java/ant/bin/ant}.\\
-
-Whichever Ant is chosen, to build jInfer do the following on command line:
-\begin{enumerate}
-	\item Change directory to jInfer's \texttt{src}.
+\begin{enumerate}\item 
+ Change directory to jInfer's \texttt{src}.
   \item Run \texttt{ant [chosen target]}. Interesting targets include:
-	\begin{itemize}
-		\item \texttt{clean}: removes any build artefacts.
+    \begin{itemize}\item 
+        \texttt{clean}: removes any build artefacts.
       \item \texttt{build}: compiles and builds jInfer.
@@
       \end{itemize}
-\end{enumerate}
-
-Chosen target may be empty, which should default to \texttt{build}.
-
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
+  \end{enumerate}\par Chosen target may be empty, which should default to \texttt{build}.
 

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial.tex
tutorial pdfs

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}
\usepackage{xstring}
\usepackage{wrapfig}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{jInfer Tutorial}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}

\maketitle

 \section*{Tutorial}
        \par 
          Target audience: jInfer users. Anyone who needs to use jInfer to create DTD, XSD or another schema from existing XML documents, schemas (again: DTD, XSD, ...) or queries (such as XPath).

        \par 
          Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.

\subsection*{Overview}
\begin{enumerate}
  \item Get \& install NetBeans.
  \item Get \& install jInfer.
  \item Infer schema.
\end{enumerate}
\subsection*{NetBeans}
\par 
  Do this if you don't have any NetBeans installed. If you have an older version than 6.9, jInfer will not run - please update it first.

\begin{enumerate}
  \item Go to \url{http://netbeans.org/}.
  \item Download NetBeans 6.9+, whichever bundle will do.
  \item Follow installation instructions.

\end{enumerate}
\subsection*{jInfer}
\par 
  Do this if you have the correct version of NetBeans installed. Note that jInfer features automatic update - you don't need to follow these steps each time new version comes out.

\begin{enumerate}
  \item Go to https://sourceforge.net/projects/jinfer/files/.
  \item Download the latest jInfer binaries (jInfer-X.Y-bin.zip).
  \item Unpack the ZIP, you'll get a few NBM files.
  \item Start NetBeans, open \textit{Tools} > \textit{Plugins} from the main menu.
  \item Switch to \textit{Downloaded} tab.
  \item Click \textit{Add Plugins...}.
  \item Select all the unpacked NBM files.
  \item Click \textit{Install} and follow the instructions.
  \item Restart NetBeans.
  \end{enumerate}
\subsection*{Basic Inference}
\par 
  This section assumes you have NetBeans and jInfer installed. After the first run, a jInfer Welcome window will appear, guiding you through basically the same steps that will be covered here. You might want to follow them instead.

\par 
  Let's assume that you have one or more XML files, and you want to create XSD schema from them.

\begin{enumerate}\item 
 Create a new jInfer project.
    \begin{enumerate}\item 
        \textit{File} > \textit{New Project} \textbf{or} \textit{New Project} icon from the toolbar \textbf{or} right-click the area in \textit{Projects} window > \textit{New Project}.
      \item Select \textit{jInfer} category, \textit{jInfer Project}.
      \item Click \textit{Next}.
      \item Choose a name (for example MyInference) and location of your project.
      \item Click \textit{Next}.
      \item In the \textit{Schema Generator} combo box select \textit{Basic XSD Exporter}.
      \item Click \textit{Finish}.
      \end{enumerate}
  \item Add input XML files to this project.
    \begin{enumerate}\item 
  Right click the newly created project in \textit{Projects} window.
      \item Select \textit{Add files}.
      \item Select XML files you want to use for schema inference.
      \item If you now open the \textit{XML} folder in your project, you will see the added files.
      \end{enumerate}
  \item Run inference.
    \begin{enumerate}\item 
        \textit{Run Project} in the toolbar (green play icon) \textbf{or} right click the project, select \textit{Run}.
      \item Generated schema will open in a new \textit{Editor} window. It will also appear in the \textit{Output} folder.
      \item You will find the file with the schema in your project (location from step 1d, folder \textit{output}).
      \end{enumerate}
  \end{enumerate}
\subsection*{Project files}\par 
  This section assumes you have successfully inferred at least one schema.

\begin{itemize}\item 
 To add input files to a specific folder (XML documents, schemas, queries) without relying on jInfer to guess the folder from the extension, right click the folder and select \textit{Add XYZ files}.
  \item Note that input folders are virtual - files in them still reside at their original locations. You can't actually delete them from disk from within jInfer, just remove them from these virtual folders by selecting \textit{Delete} from their context menu.
  \item \textit{Output} folder is, on the other hand, a faithful representation of the folder \textit{output} inside your jInfer project (step 1d). If you delete a schema here, it will be deleted from the disk too.
  \item You might want to see a diff between two input documents, or old and inferred schema - just select both files at once, right-click and select \textit{Tools} > \textit{Diff} from their context menu.
  \end{itemize}
\subsection*{Setting up jInfer}
\par 
  This section assumes that you have successfully inferred at least one schema and know how to manipulate input and output files in a project.

\par 
  To change jInfer settings, you have to first realize whether it is a change affecting a single project at a time (for example, changing the output language from XSD to DTD) or whole jInfer at once (whether schema should open in Editor window after the inference finishes).

\begin{enumerate}\item 
 First type of settings is found in so called Project Preferences. You can access them from the respective jInfer project's context menu: right click the project and select \textit{Properties}.
  \item Second type is called jInfer Options and can be found among NetBeans' options: open them (Windows, Linux: \textit{Tools} > \textit{Preferences...}, Mac: \textit{NetBeans} > \textit{Preferences...}) and switch to \textit{jInfer} tab.
  \end{enumerate}
\subsection*{Interactive simplification}
\par 
  This section assumes that you have successfully
inferred a few schemas and would like to play with the process, interfere with it.

\par 
  Default simplifier that jInfer uses is non-interactive: after it is run, it will
work for a while and then display the generated schema. However, jInfer bundles also
a way to guide the inference process interactively, by selecting states in an
automaton that should be merged together and selecting which state should be
removed while converting an automaton to a regular expression. For more
information on this, see the documentation of TwoStep simplifier.
To try out interactive inference, do the following with your jInfer project:

\begin{enumerate}
  \item Open \textit{Project Preferences}.
  \item Navigate to \textit{Simplifiers} > \textit{TwoStep} >
      \textit{Automaton Merging State}.
  \item In the \textit{Automaton simplifier} combobox select
      \textit{User Interactive}.
  \item Navigate to \textit{State Removal}.
  \item In the \textit{State Removal orderer} combobox select
      \textit{User Interactive}.
  \item Save the preferences by clicking \textit{OK}.
  \item Run the inference again.
  \item A new window will appear, which will require you
      to select states to be merged in each step of simplification.
      After selecting these states, proceed by clicking \textit{Continue}.
\end{enumerate}

\end{document}

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial_dev.tex
tutorial pdfs

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}
\usepackage{xstring}
\usepackage{wrapfig}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{Module Developer's Tutorial}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}

\maketitle

\section*{Module Developer's Tutorial}
\par 
  Target audience: jInfer module developers. Anyone who needs to extend jInfer capabilities by writing a new module.

\par

    This tutorial assumes that you are a seasoned Java developer. Having
    experience with programming in some kind of framework (NetBeans Platform
    above all) will help you a lot.
    Make sure you have read the article on
    architecture, data structures and inference process to understand what you
    will be implementing. Having read the documentation for
    remaining modules will help you too.
    Also, before starting this tutorial, make sure you can
    build jInfer from sources.

\subsection*{Overview}
  \begin{enumerate}
    \item Get NetBeans, jInfer sources, try a build.
    \item Decide on the type of module you want to create.
    \item Create new NetBeans module.
    \item Implement jInfer-specific interfaces.
    \item Implement your logic.
  \end{enumerate}
\subsection*{Building jInfer from sources}
  \par 
    Refer to the official instructions.

\subsection*{What type of module?}
  \par 
    First thing you need to realize is what kind of module you will be
    implementing. Is it going to be a part of the inference? If yes, what stage?
    Importing the initial grammar? Simplifying it? Exporting to resulting schema?
    If you are not sure what these terms mean, go read the articles on interence process again.

  \par 
    If your logic doesn't belong to the inference process, it might just extend
    one of the existing modules. Try looking for the code you would like to change.

\subsection*{New NetBeans module}
\par 
  This section assumes you have successfully imported jInfer in NetBeans.

\begin{enumerate}
  \item Expand the jInfer suite (brown puzzles icon).
  \item Right-click \textit{Modules}, select \textit{Add New...}.
  \item On the first page, pick a name for your module. Click \textit{Next}.
  \item On the second page, fill out the following:
    \begin{itemize}
      \item 
        \textit{Code base}: this will be the package structure in which your code is placed.
        If you want, follow our convention: \texttt{cz.cuni.mff.ksi.jinfer.\textit{yourmodule}}.
        Otherwise choose something like \texttt{\textit{tld.company.jinfer.yourmodule}}.

      \item \textit{Module display name}: pretty obvious.
      \item \textit{Generate XML layer}: you may want to check this option, but it doesn't really matter at this stage.
    \end{itemize}
    Rest of the options is uninteresting in most cases.

  \item Click \textit{Finish}.
\end{enumerate}
\subsection*{Implement jInfer's specifics}
\par 
This section will show how to deal with a new inference module - a simplifier
that does no actual simplification, just returns the grammar it got on input.
Your module will need a "main" class implementing the chosen inference
interface, annotated with a \texttt{@ServiceProvider} annotation.

\par 
But first, we need to do some setup.
\begin{enumerate}
  \item Right-click the newly created module, select \textit{Properties}.
  \item In \textit{Libraries > Module Dependencies}, click \textit{Add Dependency...}.
  \item Filter for "base" and select \textit{Base} as a dependency. Click \textit{OK}.
  \item 
      Still in \textit{Properties}, switch to \textit{Display} category. Fill in:
      \begin{itemize}
        \item \textit{Display Name}: Fill in a user-friendly name of your module.
        \item \textit{Display Category}: type in \textit{jInfer}.
        \item \textit{Short \& Long Description}: unleash your inner poet!
      \end{itemize}

  \item Still in \textit{Properties}, switch to \textit{API Versioning}.
  \item Type \texttt{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier} in \textit{Provided Tokens}.
  \item Close \textit{Properties} by clicking \textit{OK}.
\end{enumerate}

\par 
Now to the class itself.
\begin{enumerate}
  \item Add a new Java class: right click \textit{Source Packages} in your new
      module, select \textit{New > Java Class...}.
  \item Fill in class name, for example \textit{MySimplifierImpl}.
  \item Fill in package based on the \textit{Code base} you selected while creating
      the module itself.
  \item Click \textit{Finish}.
  \item In the heading line of the class, add \texttt{implements Simplifier} and
      fix the imports (you need \texttt{Simplifier} from Base module).
  \item Annotate this class with \texttt{@ServiceProvider(service = Simplifier.class)}.
  \item NetBeans will complain about missing method implementations, add them.
  \item 
    Now fill in method bodies:
    \begin{itemize}
      \item \texttt{getName()}: return a string with module unix name, for example \texttt{"mysimplifier"}.
      \item \texttt{getDisplayName()}: return a user friendly module name, for example \texttt{"My First Simplifier"}.
      \item \texttt{getModuleDescription()}: return a short module description. You can also return \texttt{getDisplayName()} in this case.
      \item \texttt{getCapabilities()}: for the moment, it is enough to return \texttt{Collections.emptyList()}.
      \item \texttt{start()}: here is where your main logic belongs. In the case of a simplifier, you would do something with the rules and return a simplified grammar. For now, just do \texttt{callback.finished(initialGrammar);}
    \end{itemize}

\end{enumerate}

\subsection*{Run jInfer}
\par 
  At this moment, run the whole jInfer suite from the NetBeans you imported it into.
  A new, child NetBeans should open with your module correctly installed. You can
  now follow the User tutorial, just select your new
  simplifier while creating a new jInfer project.
  That's it! Start hacking you logic now!

\subsection*{Implement your logic}
\par 
  This is the part where you actually have to do some thinking :-) Implementing
  an importer? Take that \texttt{InputStream} and create some rules of it! Simplifier?
  Take the rules you got and compact them somehow! Exporter? Take those rules and
  write them out as a \texttt{String}! It only depends on which module you are
  in and algorithm you're implementing.

\subsection*{Dealing with jInfer and NetBeans Platform}
\par 
  You will soon get into situation where you need to interact either with
  jInfer or directly NetBeans. There is a
  tutorial for dealing with these cases.

\end{document}

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial_platform.tex
tutorial pdfs

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}
\usepackage{xstring}
\usepackage{wrapfig}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{1.5cm}
\setlength{\evensidemargin}{1.5cm}
\setlength{\topmargin}{1.5cm}
\setlength{\textheight}{24cm}
\setlength{\textwidth}{18cm}

\def\mftitle{Dealing with jInfer and NetBeans Platform}
\def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D., Martin Nečaský, Ph.D.}
\def\mfplacedate{Praha, 2011}
\title{\bf\mftitle}
\author{\mfauthor \\ Advisors: \mfadvisor}
\date{\mfplacedate}

\ifx\pdfoutput\undefined\relax\else\pdfinfo{ /Title (\mftitle) /Author (\mfauthor) /Creator (PDFLaTeX) } \fi

\begin{document}

\maketitle

\section*{Dealing with jInfer and NetBeans Platform}
\par 
  Target audience: jInfer developers. Anyone who needs to interact with jInfer
  or NetBeans while hacking our framework.

\par 
  This tutorial deals with a few important but pretty specific parts of interface
  between your module (or any logic you implement) and either jInfer or NetBeans
  Platform. Not nearly everything regarding interaction with NBP is mentioned here,
  please refer to the relevant FAQ.

\par

    This tutorial assumes that you are a seasoned Java developer. Having
    experience with programming in some kind of framework (NetBeans Platform
    above all) will help you a lot.
    Make sure you have read the article on
    architecture, data structures and inference process to understand what you
    will be implementing. Having read the documentation for
    remaining modules will help you too.
    Also, before starting this tutorial, make sure you can
    build jInfer from sources.

\subsection*{Overview}
\begin{enumerate}
  \item Module visibility
  \item Error handling
  \item Interruptions
  \item Dialogs
  \item Configuration - options, preferences
  \item Rule display
  \item Console output, logging
  \item RunningProject class
  \item Module selection
\end{enumerate}
\subsection*{Module visibility}
\par 
  Each NBP module allows to set some spackages as public, which means their content
  will be available to other modules. That means, if module \textit{A} declares a
  dependency on module \textit{B}, it will be able to use only classes from
  those packages that were set as public in module \textit{B}.

\textit{
  \textbf{Important notice:} Public packages do not affect
  Lookup mechanism, i.e.
  class annotated with \texttt{@ServiceProvider}
  does not need to be in a public package in order to be looked up.
}
\par 
  There are two ways to set packages visible, first is to set it manually in
  module's \texttt{project.xml} file located in
  \textit{ModulePath/nbproject/} folder.

Example:
\begin{verbatim}
// This is a portion of example project.xml file.
<?xml version="1.0" encoding="UTF-8"?>
<project xmlns="\url{http://www.netbeans.org/ns/project/1">}
    <type>org.netbeans.modules.apisupport.project</type>
    <configuration>
        <data xmlns="\url{http://www.netbeans.org/ns/nb-module-project/3">}
            ....
            <public-packages>
                <package>some.example.package.name</package>
            </public-packages>
        </data>
    </configuration>
</project>
\end{verbatim}
\par 
  Another way to set public packages is through GUI:
  \begin{enumerate}
    \item Right-click the module, select \textit{Properties}.
    \item In \textit{API Versioning}, there is a \textit{Public Packages} section.
    \item Check all packages that should be public.
  \end{enumerate}

\subsection*{Error handling}
\par 
  Each run of inference is encapsulated in a \texttt{try-catch} block,
  so it is safe to throw any exception in the inference process.
  Each thrown exception will be caught, get logged, presented to the user and
  inference will stop. However if module uses threads which throw exceptions,
  it is caught by NBP, instead of our code. Because of this its reasonable to
  catch this exception in your module and re-throw it in the right thread.

\subsection*{Interruptions}
\par 
  In jInfer user can stop the inference at any moment of its run. For this
  reason, modules have to check in every time-consuming place (such as long loops)
  whether this is the case. Example below shows how to check the interruption
  from user and what to do to correctly stop the inference.

Example:
\begin{verbatim}
for (forever) {
    if (Thread.interrupted()) {
        throw new InterruptedException();
    }
    doStuff();
}
\end{verbatim}
\subsection*{Dialogs}
\par 
  Creation of open/save dialogs or message windows is done using standard NBP API.
  For dialogs you can use the standard Java \texttt{JFileChooser},
  but NBP \texttt{FileChooserBuilder}
  is more convenient. This class creates dialog which remembers last-used
  directory according to a key passed into its constructor.

Example:
\begin{verbatim}
// The default dir to use if no value is stored
File home = new File(System.getProperty("user.home"));
// Now build a file chooser and invoke the dialog in one line of code
File toAdd = new FileChooserBuilder(CallingClass.class)
                   .setTitle("Some dialog title")
                   .setDefaultWorkingDirectory(home).showOpenDialog();
// Result will be null if the user clicked cancel or closed the dialog w/o OK
if (toAdd != null) {
    //do something
}
\end{verbatim}
\par 
  For message windows the \texttt{DialogDisplayer} NBP API
  class is used. To determine which type of message will be shown
  (Confirmation dialog, message dialog...),
  \texttt{NotifyDescriptor}
  is used. Following is a small example, which creates standard message
  notification. For more information please look at
  NBP Dialogs API FAQ.

Example:
\begin{verbatim}
// Creates standard information message with text "Hello world"
NotifyDescriptor nd = new NotifyDescriptor
                            .Message("Hello world",
                               NotifyDescriptor.INFORMATION_MESSAGE);
DialogDisplayer.getDefault().notify(nd);
\end{verbatim}
\subsection*{Configuration - options, preferences}
\par 
  Option panels located in \textit{Tools > Options} menu are a standard part of
  NBP API. We created a \textit{jInfer} option sub-category to place panels with
  configuration valid across all jInfer projects. Description of the few step
  to create your own option panel follows. For more information, please visit
  NetBeans Options tutorial.

\begin{enumerate}
 \item Create options window:
  \begin{itemize}
   \item Right click \textit{Source Packages} in your module,
                            select \textit{New > Other}.
   \item Under Categories, select Module Development. Under
                            File Types, select Options Panel. Click \textit{Next}.
   \item Keep checked \textit{Create Secondary Panel} and choose
                            jInfer as the \textit{Primary Panel}. Fill in remaining
                            fields. Click \textit{Next}.

   \item Fill in \textit{Class Name Prefix} and \textit{Package}
                            and click \textit{Finish}.
  \end{itemize}

 \item Design newly created \textit{ClassNamePrefix}Panel.java.
 \item Implement \texttt{store()} and \texttt{load()} methods according to comments inside.
\end{enumerate}
\par 
  Both \texttt{store()} and \texttt{load()} methods use
  \texttt{NbPreferences}
  class. Example below shows how to store and load Preferences.

Example:
\begin{verbatim}
public void store() {
  // Saves the state of checkBox into Preferences for ExampleClass
  // class in property with name "PropertyName"
  NbPreferences.forModule(ExampleClass.class)
      .putBoolean("PropertyName", someCheckBox.isSelected());
}
public void load() {
  // Get from ExampleClass class preferences the property with name
  // "PropertyName". Second parameter of method getBoolean is used
  // as default if no property with defined name is saved in preferences.
  someCheckBox.setSelected(NbPreferences.forModule(ExampleClass.class)
                               .getBoolean("PropertyName", true));
}
\end{verbatim}
\par 
  Second type of preferences used in jInfer is Project properties. Each jInfer
  project has its properties panel accessible in its content menu and
  these properties are applied for each project separately. To implement this
  kind of preferences, it is necessary to implement
  \texttt{PropertiesPanelProvider}
  interface and extend the
  \texttt{AbstractPropertiesPanel}
  class. Each Project properties window has category tree, where each category
  represents a separate panel with properties. This category is declared in the
  provider interface, properties panel itself is defined in class extending
  \texttt{AbstractPropertiesPanel}.

\paragraph*{\texttt{PropertiesPanelProvider}}
\par 
  As was mentioned above, \texttt{PropertiesPanelProvider} defines
  category in Project properties windows. In the example below, we'll try to
  explain how to create a simple provider and what each method is responsible for.

Example:
\begin{verbatim}
public class ExampleProvider implemenets PropertiesPanelProvider {
  // Programmatic name of this provider.
  public String getName() {
    return "ExampleProvider";
  }
  // User-friendly name - this name will be displayed in category tree.
  public String getDisplayName() {
    return "Example Category";
  }
  // Priority of category. Higher the number, higher
  // will the category be in the tree. If two categories have
  // same priorities, they are sorted according
  // to their names.
  public int getPriority() {
    return 0;
  }
  // Returns properties panel defined for this category.
  public AbstractPropertiesPanel getPanel(final Properties properties) {
    return new ExamplePropertiesPanel(properties);
  }

  // Defines parent category of this category. If null is returned, this is top
  // level category. In other cases id (programmatic name) of parent category
  // must be returned.
  public String getParent() {
    return null;
  }

  // Optional. For each category, list of virtual categories can be defined.
  // Instead of properties panel, VirtualCategoryPanel only informs of type and
  // installed number of modules of a certain type.
  public List getSubCategories() {
    return null;
  }
}
\end{verbatim}

\paragraph*{\texttt{AbstractPropertiesPanel}}
\par 
  This class represents the visual component of properties category. You can
  design it in whatever way you like. For proper functionality, a few steps must
  be followed.

\begin{enumerate}
  \item 
    Call \texttt{super(Properties)} in constructor: this causes
    \texttt{Properties} instance to be saved into
    \texttt{properties} protected field and you can use it in
    \texttt{load()} and \texttt{store()} methods.

  \item 
    Implement \texttt{load()} method: In this method values previously saved
    into \texttt{Properties} instance provided to this panel are loaded into
    components in this panel.

  \item 
    Implement \texttt{store()} method: Here the values gathered from
    components in this panel are saved into \texttt{Properties}.

\end{enumerate}

Example:
\begin{verbatim}
// This is only a part of the actual class...
public class ExamplePanel extends AbstractPropertiesPanel {

  public ExamplePanel(final Properties properties) {
    super(properties);
  }

  public abstract void store() {
    properties.setProperty("exampleKey", someTextField.getText());
  }

  public abstract void load() {
    String propValue = properties.getProperty("exampleKey", "default value");
    someTextField.setText(propValue);
  }
}
\end{verbatim}

\subsection*{Rule display}

\par 
  \textit{Rule displayer} is a component used for visualization of rules in any
  step of inference process. If rule displayers bundled with jInfer are
  insufficient, you can implement a rule displayer on your own.
  You just need to follow these few easy steps.

\begin{enumerate}
  \item 
    Implement \texttt{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer}
    \begin{itemize}
      \item 
        \texttt{RuleDisplayer} interface extends \texttt{NamedModule},
        so you need to override its methods too.

      \item 
        Implement \texttt{createDisplayer(name, rules)} method which creates
        displayer window responsible for displaying rules.

    \end{itemize}

  \item 
    Annotate this class with
    \texttt{@ServiceProvider(service = RuleDisplayer.class)}.

\end{enumerate}

\par 
  \textbf{Important notice:} never forget to clone the rules you want
  to display. Rule displayer usually runs in its own thread, and accessing the
  rules that are being simplified at the same time might lead to weird results.
  You can use \texttt{CloneHelper} to clone the rules.

Example:
\begin{verbatim}
@ServiceProvider(service = RuleDisplayer.class)
public final class ExampleRuleDisplayer implements RuleDisplayer {

  @Override
  public void createDisplayer(String panelName, List rules) {
    // Creates standard NBP TopComponent in which are rules displayed.
    ExampleRDTopComp topComponent = ExampleRDTopComp.findInstance();
    if (!topComponent.isOpened()) {
      topComponent.open();
    }
    topComponent.createRuleDisplayer(panelName, rules);
  }

  @Override
  public String getName() {
    return "ExampleDisplayer";
  }

  @Override
  public String getDisplayName() {
    return "Example rule displayer";
  }

  @Override
  public String getModuleDescription() {
    return getDisplayName();
  }
}
\end{verbatim}

Example of invocation with cloning:
\begin{verbatim}
RuleDisplayerHelper.showRulesAsync("Panel name",
    new CloneHelper().cloneGrammar(grammar), true);
\end{verbatim}

\subsection*{Console output, logging}

\par 
  To print to console output, NBP provides
  \texttt{IOProvider}
  and \texttt{InputOutput}
  classes. \texttt{IOProvider}'s method \texttt{getIO(String, boolean)}
  returns \texttt{InputOutput} instance, which can be used to obtain
  \texttt{Reader}/\texttt{OutputWriter} to read/write into output window.

Example:
\begin{verbatim}
// Creates new output window, where first parameter is a name and second
// is flag determining whether a new window should be created even if there is
//  already a window with the same name.
InputOutput ioResult = IOProvider.getDefault().getIO("Example", true);
ioResult.getOut().println("Hello world");

// After writing everything into output window, close the output.
ioResult.getOut().close();
\end{verbatim}

\par 
  Logging is done in jInfer using
  Log4j tool.
  If you are not familiar with Log4j, example below shows simple usage. All
  logged messages are saved in a standard log file placed in \textit{UserHome/.jinfer/}
  folder and printed into "jInfer" output window in the application. For each
  of these two places a log level can be set in jInfer options.

Example:
\begin{verbatim}
// Creates Logger instance for this particular class
private static final Logger LOG = Logger.getLogger(Example.class);

    ...

// Log a message with info level
LOG.info("Some message with info log level");
// Log an error message
LOG.error("OMG, error occurred!");
\end{verbatim}

\subsection*{\texttt{RunningProject} class}

\par 
  This class provides access to information regarding currently running inference.
  It is a singleton, corresponding with the fact that at a given time, at most
  one inference (project) can be running in the whole NB. Consequently, all relevant methods
  are synchronized.
  NB project corresponding to the running inference can be retrieved by
  calling \texttt{getActiveProject()}. Its properties can be retrieved by
  calling \texttt{getActiveProjectProps()}. This
  method will work also if there is no running project - it just retrieves an
  empty instance of \texttt{Preferences}, which is useful for example in JUnit
  tests.

Example code retrieving properties for DTD export:
\begin{verbatim}
Properties properties = RunningProject.getActiveProjectProps(
                            DTDExportPropertiesPanel.NAME);
\end{verbatim}

\par 
  Aside from keeping information about the currently running project, this class
  keeps information about the capabilities of the following module in the inference
  chain. This can be retrieved by calling the \texttt{getNextModuleCaps()} method.
  Again, if this information is not available, empty \texttt{Capabilities}
  instance will be returned.

Example code retrieving \textit{Simplifier} capabilities
in \textit{BasicIGG}:
\begin{verbatim}
if (!RunningProject.getNextModuleCaps().getCapabilities().contains(
            Capabilities.CAN_HANDLE_COMPLEX_REGEXPS)) {
  ... expand rules ...
}
\end{verbatim}

\subsection*{Module selection}

\par 
  This part of the tutorial will show the full process of implementing a choice
  between two submodules in an inference module, for example simplifier. Similar
  principles apply one level higher, when dealing with inference modules themselves.

\subsubsection*{Interface}

\par 
  First of all, we need an interface encapsulating the new submodule. It has to
  extend \texttt{NamedModule} interface.

\begin{verbatim}
public interface Foo extends NamedModule {

  String bar();
}
\end{verbatim}

\subsubsection*{Implementations}

\par 
  Second, we need some implementations of this interface. At least two, otherwise
  it's boring.
  They will have to implement all required methods and register themselves as
  service providers for \texttt{Foo}.

Example of the first implementation. The second will be the same (just change "ones" to "twos").
\begin{verbatim}
@ServiceProvider(service = Foo.class)
public class Impl1 implements Foo {

  // NamedModule stuff

  @Override
  public String getName() {
    return "Impl1"; // unix-style name
  }

  @Override
  public String getDisplayName() {
    return "Foo implementation number 1"; // user-friendly name
  }

  @Override
  public String getModuleDescription() {
    return getDisplayName(); // irrelevant now
  }

  // logic from Foo

  @Override
  public String bar() {
    return "Hello world from Impl1!";
  }
}
\end{verbatim}

\subsubsection*{Properties panel}

\par 
  Now we have to create a properties panel. Refer to an
  earlier part of this document for general
  reference. Here we assume a combo box already present on the properties panel,
  with the correct renderer set (\texttt{ProjectPropsComboRenderer}).
  We need to do 2 things: while loading the properties panel, fill it with
  names of all implementations and select the currently chosen one (or default).
  While saving, we have to get the name of
  the selected implementation and save it.

Example of panel load code:
\begin{verbatim}
@Override
public final void load() {
  combo.setModel(new DefaultComboBoxModel(
          ModuleSelectionHelper.lookupImpls(Foo.class).toArray()));

  combo.setSelectedItem(ModuleSelectionHelper.lookupImpl(Foo.class,
                   properties.getProperty(
                      "selected.foo.implementation", // property name
                      "Impl1"))); // default property value
                                             // - same as Impl1.getName()!
}
\end{verbatim}

Example of panel save code:
\begin{verbatim}
@Override
public void store() {
  properties.setProperty("selected.foo.implementation",
          ((NamedModule) combo.getSelectedItem()).getName());
}
\end{verbatim}

Example of combo box renderer setting:
\begin{verbatim}
combo.setRenderer(new ProjectPropsComboRenderer(combo.getRenderer()));
\end{verbatim}

\subsubsection*{Invocation}

\par 
  Finally, there will be a place (in our example simplifier) where we actually
  want to load the properties of the running project and lookup the selected
  implementation of \texttt{Foo}.

Example lookup and usage:
\begin{verbatim}
Properties p = RunningProject.
    getActiveProjectProps("ExampleProvider"); // the name provided in
                                              // properties panel provider

Foo foo = ModuleSelectionHelper.lookupImpl(Foo.class,
    p.getProperty("selected.foo.implementation", "Impl1"));

println(foo.bar());
\end{verbatim}

\par 
  A few things should be noted:
  \begin{itemize}
    \item Module names (\texttt{getName()}) have to be unique.
    \item It is a good practice to store property names and default values as
      constants, for example in super module - do a reference search in our
      code to find out how we do it.
    \item \texttt{getUserDescription()} should work in modules that have
      submodules as follows: it should somehow return the name of the module
      and names of all selected submodules to indicate inner workings of this
      module. For example
      \textit{My Simplifier (Clustering Logic A, Regexper \#3)}.
  \end{itemize}

\end{document}

### 2011-03-02 doc/jInfer/tutorials/building_jinfer.tex
Documentation: Tutorials: changing titles.
propset

@@
 
-\def\mftitle{Building jInfer from sources}
+\def\mftitle{jInfer: Building from sources}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \begin{enumerate}\item 
- Go to 
-https://sourceforge.net/projects/jinfer/files/.
+ Go to \url{https://sourceforge.net/projects/jinfer/files/}.
   \item Download the latest jInfer sources (jInfer-X.Y-src.zip).
@@
   jInfer, use the following URL:
-  https://jinfer.svn.sourceforge.net/svnroot/jinfer
+  \url{https://jinfer.svn.sourceforge.net/svnroot/jinfer}.
   After the checkout, continue by building jInfer.
@@
     Its icon should look like two brown puzzle pieces.
-  \item You should see the name "jInfer" in the dialog window.
+  \item You should see the name ``jInfer'' in the dialog window.
   \item Make sure the \textit{Open Required Projects} check box is checked.
@@
   pieces icon) context menu. If jInfer is the main project, clicking the
-  green "Play" icon in the toolbar should work too.
+  green ``Play'' icon in the toolbar should work too.
   If you want to build jInfer, for example to test if your changes didn't break
   something, select \textit{Build} from the context menu. Sometimes it might be
-  necessary to run \textit{Clean \& Build}, try this if it "stops working".
+  necessary to run \textit{Clean \& Build}, try this if it ``stops working''.
   If you want to run JUnit test, select \textit{Test} from the same context menu.
@@
       \item \texttt{build}: compiles and builds jInfer.
-      \item \texttt{clean build}: performs a clean build, useful if something "stops working".
+      \item \texttt{clean build}: performs a clean build, useful if something ``stops working''.
       \item \texttt{nbms}: creates a distribution by packaging jInfer modules in NBM files.

### 2011-03-02 doc/jInfer/basicigg/basicigg.tex
corrected a lot of : and lowercase letters starting a sentence

@@
 
-XML processor registers itself into \emph{document} input folder, \emph{xml} file extension and declares that it can process other arbitrary file extensions.\\
-SAX traversal is used to collect the rules; the relevant \code{ContentHandler} is the class \code{TrivialHandler}.
-The way it works is following: every time a start of an element is encountered, a new \code{Element} is created along with its attributes.
+XML processor registers itself into \emph{document} input folder, \emph{xml} file extension and declares that it can process other arbitrary file extensions.
+
+SAX traversal is used to collect the rules; the relevant \code{ContentHandler} is the class \code{TrivialHandler}. The way it works is following: \\
+Every time a start of an element is encountered, a new \code{Element} is created along with its attributes.
 This new element representation is then placed on the top of a stack.
@@
 
-Expansion is used to convert complex regular expressions in initial grammar to simple ones (concatenations of tokens) for simplifiers that can handle only the simple form.
+Expansion is used to convert complex regular expressions in initial grammar to simple ones (concatenations of tokens) for simplifiers that can handle the simple form only.
 The process can be seen as breaking down a regular expression into a set of words (positive examples) that are described by this regexp.
@@
 
-\code{ExpanderImpl} works recursively, and it will probably be the best to explain it from bottom to up.\\
+\code{ExpanderImpl} works recursively, and it will probably be best to explain it from bottom to top.\\
 
@@
 Following the same logic, let's denote a set of such words as \code{List<List<T>{}>}.
-Our aim while expanding will be to convert a regular expression representing some node's subnodes to a list of words that are described by this regexp.
+Our aim while expanding will be to convert a regular expression representing some node's subnodes into a list of words that are described by this regexp.
 In this case, \emph{T} will be \code{AbstractStructuralNode}.\\
@@
 The first interesting function is \code{ExpansionHelper.applyInterval()}.
-It takes a list of words extracted from a regexp (thus in the List<List<AbstractStructuralNode>{}> form) and a \code{RegexpInterval} and creates a new list of words: applying the interval on each of the words in a heuristic sense.
+It takes a list of words extracted from a regexp (thus in the List<List<AbstractStructuralNode>{}> form) and a \code{RegexpInterval}, and creates a new list of words.
+Applying the interval on each of the words in a heuristic sense.
 The behaviour depends on the type of the interval.
 \begin{itemize}
-	\item \emph{Once}: the resulting set of words is exactly the original set of words.
-	\item \emph{Optional}: same as above, but an empty word is added to the resulting set (each word from input \emph{may} or \emph{may not} occur - but one empty word is enough).
-	\item \emph{General interval}: has a minimum and a maximum number of occurences.
+	\item \emph{Once}: The resulting set of words is exactly the original set of words.
+	\item \emph{Optional}: Same as above, but an empty word is added to the resulting set (each word from input \emph{may} or \emph{may not} occur - but one empty word is enough).
+	\item \emph{General interval}: Has a minimum and a maximum number of occurences.
 Theoretically, all counts in interval $ [min, max] $ should be considered, but for practical reasons the following heuristic is employed: each word from the input is duplicated\footnote{Duplicating a word $abc$ $N$ times results in a word $\underbrace{abcabc\ldots abc}_{N\times}$.} first $min$, then $max$ times and added to the result.
-	\item \emph{Kleene star}: similar to the general interval and \emph{optional}, the result contains an empty word.
+	\item \emph{Kleene star}: Similar to the general interval and \emph{optional}, the result contains an empty word.
 Then for each word on input: this word and its 3$\times$ duplication is added to the result.
 Obviously, this is a heuristic approach again - theoretically each word should be duplicated between 0 and $\infty$ times in the result.
-	\item \emph{Kleene cross}: except for not including empty word, this is the same case as \emph{Kleene star}: each input word is duplicated 1 and 3 times in the result.
+	\item \emph{Kleene cross}: Except for not including the empty word, this is the same case as \emph{Kleene star}. Each input word is duplicated 1 and 3 times in the result.
 \end{itemize}
 
-Understanding how intervals affect the expansion, it is time to move a level up: how a regular expression is expanded.
+Understanding how intervals affect the expansion, it is time to move up a level, to describe how a regular expression is expanded.
 This is done in \code{ExpanderImpl.unpackRE()} performing a recursive descent into the regexp tree and returning a set of words.
@@
 \begin{itemize}
-	\item \emph{LAMBDA}: resulting word set contains one empty word.
-	\item \emph{TOKEN}: resulting set contains a word with one letter: the content (an \code{AbstractStructuralNode}) of this token; the interval of the regexp is applied on this word before returning it.
-	\item \emph{ALTERNATION}: first, all children of this alternation are recursively evaluated.
+	\item \emph{LAMBDA}: Resulting word set contains one empty word.
+	\item \emph{TOKEN}: Resulting set contains a word with one letter: the content (an \code{AbstractStructuralNode}) of this token; the interval of the regexp is applied on this word before returning it.
+	\item \emph{ALTERNATION}: First, all children of this alternation are recursively evaluated.
 Their respective words are appended in a list, this list has the regexp interval applied and is returned.
@@
 If the regexp is an alternation, any word that can be produced by one of its children can be a word produced by the regexp itself.
-	\item \emph{PERMUTATION}: a heuristic is applied again: because there are $n!$ ways to order children of a permutation, only two are actually picked: the original order in which they appear, and reversal of this order.
+	\item \emph{PERMUTATION}: A heuristic is applied once again.
+Because there are $n!$ ways to order children of a permutation, only two are actually picked: the original order in which they appear, and reversal of this order.
 In both cases the children are treated as if they we concatenation (and thus the following case is invoked) and appended in a list.
 On this list the regexp interval is applied and the list is returned.
-	\item \emph{CONCATENATION}: this is by far the most complicated case.
+	\item \emph{CONCATENATION}: This is by far the most complicated case.
 Let's denote word sets produced by the $N$ children of the concatenation $ w_1 \ldots w_N $.
@@
 
-Now for the next part, description of \code{Element} expansion.
+Now for the next part, a description of \code{Element} expansion.
 This is the content of \code{ExpanderImpl.expandElement()} method.
@@
 \begin{itemize}
-	\item Flag to keep values of attributes found in XML documents - unchecking this saves memory during the inference process, but loses all information about the attribute domain, which can lead to less expressive schema.
+	\item Flag to keep values of attributes found in XML documents - unchecking this saves memory during the inference process, but loses all information about the attribute domain, which can lead to a less expressive schema.
 	\item Flag to keep simple data values from XML documents - again, unchecking saves memory (potentially a lot, with input files containing a lot of text).
-Drawback might again be losing information that could be useful in schema generation.
-	\item Flag to stop on errors: if this flag is set, \jmodule{BasicIGG} will interrupt the inference if it comes across an error while extracting IG.
+A drawback might again be in losing information that could be useful in schema generation.
+	\item Flag to stop on errors - if this flag is set, \jmodule{BasicIGG} will interrupt the inference if it comes across an error while extracting IG.
 Unsetting this flag allows \jmodule{BasicIGG} to ignore these errors and retrieve as much IG as possible.
@@
 
-It is possible to replace the default \code{Expander} implementation: but as mentioned in \ref{section-expansion}, either the old implementation must be removed or module selection must be introduced.
+It is possible to replace the default \code{Expander} implementation, but as mentioned in \ref{section-expansion}, either the old implementation must be removed or module selection must be introduced.
 

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial_dev.tex
Documentation: Tutorials: changing titles.
minor to building, tutorial_dev is OK

@@
 
-\def\mftitle{Module Developer's Tutorial}
+\def\mftitle{jInfer Module Developer's Tutorial}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
         If you want, follow our convention: \texttt{cz.cuni.mff.ksi.jinfer.\textit{yourmodule}}.
-        Otherwise choose something like \texttt{\textit{tld.company.jinfer.yourmodule}}.
+        Otherwise choose something like \texttt{\textit{tld.\.company.\.jinfer.\.yourmodule}}.
 
@@
 that does no actual simplification, just returns the grammar it got on input.
-Your module will need a "main" class implementing the chosen inference
+Your module will need a ``main'' class implementing the chosen inference
 interface, annotated with a \texttt{@ServiceProvider} annotation.
@@
 \par 
-But first, we need to do some setup.
+\noindent But first, we need to do some setup.
 \begin{enumerate}
@@
   \item In \textit{Libraries > Module Dependencies}, click \textit{Add Dependency...}.
-  \item Filter for "base" and select \textit{Base} as a dependency. Click \textit{OK}.
+  \item Filter for ``base'' and select \textit{Base} as a dependency. Click \textit{OK}.
   \item 
@@
 \par 
-Now to the class itself.
+\noindent Now to the class itself.
 \begin{enumerate}
@@
     \begin{itemize}
-      \item \texttt{getName()}: return a string with module unix name, for example \texttt{"mysimplifier"}.
-      \item \texttt{getDisplayName()}: return a user friendly module name, for example \texttt{"My First Simplifier"}.
+      \item \texttt{getName()}: return a string with module unix name, for example \texttt{``mysimplifier''}.
+      \item \texttt{getDisplayName()}: return a user friendly module name, for example \texttt{``My First Simplifier''}.
       \item \texttt{getModuleDescription()}: return a short module description. You can also return \texttt{getDisplayName()} in this case.

### 2011-03-02 doc/jInfer/basicruledisplayer/basicruledisplayer.tex
corrected a few sentences

@@
 
-This is fairly basic rule displayer draws rules as series of nested coloured rectangles.
+This is a fairly basic rule displayer, which draws rules as series of nested coloured rectangles.
 It does so by recursively descending down the regexp tree to draw the innermost nodes first, composing the outer nodes from them.
-User interface-wise, it creates a component with multiple panels: each one for a new grammar to display.
+User interface-wise, it creates a component with multiple tabs - each one for a new grammar to be displayed.
 
@@
 
-The main class implementing \code{RuleDisplayer} inference interface and simultaneously registered as its service provider is \code{BasicRuleDisplayer}.
+The main class implementing \code{RuleDisplayer} inference interface and simultaneously being registered as its service provider is \code{BasicRuleDisplayer}.
 Main method of this class is \code{createDisplayer()}, which looks up the component, adds a new panel to it and renders the specified grammar in it.\\
@@
 The diagram of classes involved in rule display is in figure \ref{figure-class-diagram}.
-The top component of the rule displayer has a few panels in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar once into an internal \code{Image}, then keeping it and drawing it onto panel's \code{Canvas} each time repainting is needed.\\
+The top component of the rule displayer has a few panels in it, each containing its own \code{RuleDisplayer}, responsible for rendering specified grammar into an internal \code{Image} once, and then keeping it and drawing it onto panel's \code{Canvas} each time repainting is needed.\\
 

### 2011-03-02 doc/jInfer/basicdtd/basicdtd.tex
optical separation of paragraphs

@@
 Its purpose is to convert intervals of a regexp and its children to those representable in DTD.
-For example regular expression $(a\{2,5\}, b\{0,2\})$ would be transformed to $(a,a,a?,a?,a?,b?,b?)$.
+For example regular expression $(a\{2,5\}, b\{0,2\})$ would be transformed to $(a,a,a?,a?,a?,b?,b?)$.\\
 
@@
 If it is unbounded and $min$ is zero, attach the regexp once, with interval $\{0, \infty\}$.
-If $min$ is non-zero, last regexp in required string gets interval $\{1, \infty\}$.
+If $min$ is non-zero, last regexp in required string gets interval $\{1, \infty\}$.\\
 
@@
 Type definition of an attribute is generated in the \code{DomainUtils.getAttributeType()} method.
-Based on a user setting, this might decide to enumerate all possible values of this attribute using the \code{(a|b|c)} notation, otherwise it just returns \code{\#CDATA}.\\
+Based on a user setting, this might decide to enumerate all possible values of this attribute using the \code{(a|b|c)} notation, otherwise it just returns \code{\#CDATA}.
+
 Attribute requiredness is assessed based on \code{required} metadata presence.

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial.tex
Documentation: Tutorials: changing titles.
tutorial OK

@@
 
-\def\mftitle{jInfer Tutorial}
+\def\mftitle{jInfer User Tutorial}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
         \par 
-          Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.
+          \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}
 
@@
 \begin{enumerate}
-  \item Go to https://sourceforge.net/projects/jinfer/files/.
+  \item Go to \url{https://sourceforge.net/projects/jinfer/files/}.
   \item Download the latest jInfer binaries (jInfer-X.Y-bin.zip).

### 2011-03-02 doc/jInfer/tutorials/doc_tutorial_platform.tex
Documentation: Tutorials: changing titles.
platform tut OK

@@
 
-\def\mftitle{Dealing with jInfer and NetBeans Platform}
+\def\mftitle{jInfer: Dealing with jInfer and NetBeans Platform}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \par 
-  Each NBP module allows to set some spackages as public, which means their content
+  Each NBP module allows to set some packages as public, which means their content
   will be available to other modules. That means, if module \textit{A} declares a
@@
 
-Example:
+\noindent Example:
 \begin{verbatim}
@@
 <?xml version="1.0" encoding="UTF-8"?>
-<project xmlns="\url{http://www.netbeans.org/ns/project/1">}
+<project xmlns="http://www.netbeans.org/ns/project/1">
     <type>org.netbeans.modules.apisupport.project</type>
     <configuration>
-        <data xmlns="\url{http://www.netbeans.org/ns/nb-module-project/3">}
+        <data xmlns="http://www.netbeans.org/ns/nb-module-project/3">
             ....
@@
 \par 
-  Another way to set public packages is through GUI:
+\noindent  Another way to set public packages is through GUI:
   \begin{enumerate}
@@
 }
+
 public void load() {
@@
   kind of preferences, it is necessary to implement
-  \texttt{PropertiesPanelProvider}
+  \texttt{Properties\.Panel\.Provider}
   interface and extend the
-  \texttt{AbstractPropertiesPanel}
+  \texttt{Abstract\.Properties\.Panel}
   class. Each Project properties window has category tree, where each category
@@
 
-\paragraph*{\texttt{PropertiesPanelProvider}}
+\subsubsection*{PropertiesPanelProvider}
 \par 
@@
 
-Example:
+\noindent Example:	
 \begin{verbatim}
@@
   }
-
   // Defines parent category of this category. If null is returned, this is top
@@
   }
-
   // Optional. For each category, list of virtual categories can be defined.
@@
 
-\paragraph*{\texttt{AbstractPropertiesPanel}}
+\subsubsection*{AbstractPropertiesPanel}
 \par 
@@
 \end{enumerate}
-
-Example:
+\noindent Example:
 \begin{verbatim}
@@
 \end{verbatim}
+\newpage
 
 \subsection*{Rule display}
-
-\par 
   \textit{Rule displayer} is a component used for visualization of rules in any
@@
 
-\par 
+\par \noindent
   \textbf{Important notice:} never forget to clone the rules you want
@@
 \end{verbatim}
-
 Example of invocation with cloning:
@@
 \end{verbatim}
-
-\par 
   Logging is done in jInfer using
@@
   logged messages are saved in a standard log file placed in \textit{UserHome/.jinfer/}
-  folder and printed into "jInfer" output window in the application. For each
+  folder and printed into ``jInfer'' output window in the application. For each
   of these two places a log level can be set in jInfer options.
 
-Example:
+\noindent Example:
 \begin{verbatim}
@@
 
-\subsection*{\texttt{RunningProject} class}
+\subsection*{RunningProject class}
 
@@
 public interface Foo extends NamedModule {
-
   String bar();
@@
 
-Example of the first implementation. The second will be the same (just change "ones" to "twos").
+\noindent Example of the first implementation. The second will be the same (just change ``ones'' to ``twos'').
 \begin{verbatim}
@@
 
-Example of panel load code:
+\noindent Example of panel load code:
 \begin{verbatim}
@@
 
-Example of panel save code:
+\noindent Example of panel save code:
 \begin{verbatim}
@@
 
-Example of combo box renderer setting:
+\noindent Example of combo box renderer setting:
 \begin{verbatim}

### 2011-03-02 doc/jInfer/xsdimport/xsdimport.tex
"others are skipped"

@@
 
-Enumeration of tag attributes currently recognized by our parsers. They are the following.
+Enumeration of tag attributes currently recognized by our parsers. Others are simply skipped. They are the following.
 \begin{itemize}
@@
 
-Enumeration of tags currently known by our parsers. They are the following.
+Enumeration of tags currently known by our parsers. Others are simply skipped. They are the following.
 \begin{itemize}
@@
 It allows developers to add, or remove, custom parsers just by creating a module with a new parser implementation; there is no need to change any code within \code{XSDImporter} module.
-Only exception is the re-definition of the default parser (currently DOM), which is not needed even if the module of default parser is not present.\\
+Only exception is the re-definition of the default parser (currently DOM), which is not needed even if the module of default parser is not present.
+
 Parsers with extended or custom functionality can be created using \code{XSDImportDOM} module as example (recomended).

### 2011-03-02 doc/jInfer/autoeditor/autoeditor.tex
Corrections in AE

@@
 UML diagrams shown in this document are simplified to keep them readable.
-Simplification involves removing not important class members which are not mentioned in the text, removing members of non-jInfer classes and truncating string of inheritance which are not important for this document.\ For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extends \code{JPanel} class directly, though in an UML diagram it has not any methods and extends \code{JPanel} directly, because it is sufficient for understanding of this document.
+Simplification involves removing unimportant class members, which are not mentioned in the text.
+Also, removing members of non-jInfer classes and truncating string of inheritance that are not important for this document.
+For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extend \code{JPanel} class directly.
+However, in the UML diagram, it does not have any methods and is shown to extend \code{JPanel} directly, because it is sufficient for understanding of this document.
 
@@
 Structure of \jmodule{AutoEditor} can be divided into following four main parts.
-\begin{itemize}
-	\item API - API to display automaton in GUI.
-	\item Base classes - Classes providing basic functionality that can be extended and combined to achieve desired visualization of an automaton.
-	\item Derived classes - Classes derived from the base classes that are used in existing modules and simultaneously serve as examples.
-	\item Layout creation - System of creating \code{Layout}s.
-\end{itemize}
-First, \code{Layout}s and use of base classes to create a visualization of automaton will be described.\\
+\begin{description}
+	\item[API] API for displaying automaton in GUI.
+	\item[Base classes] Classes providing basic functionality that can be extended and combined to achieve desired visualization of an automaton.
+	\item[Derived classes] Classes derived from the base classes that are used in existing modules and simultaneously serve as examples.
+	\item[Layout creation] System for creating \code{Layout}s.
+\end{description}
+
+First, \code{Layout}s and use of base classes to create a visualization of an automaton will be described.\\
 
@@
 \subsection{Layout}
-\code{Layout} is a JUNG interface responsible primarily of representation of an automaton and positions of its states.
+\code{Layout} is a JUNG interface responsible primarily for the representation of an automaton and positions of its states.
 JUNG library provides several implementation of \code{Layout} interface.
@@
 It positions automaton states on a square grid.
-This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confusing.\\
+This \code{Layout} gives good results for relatively small automata (about 10 states or less) but for larger ones, the results are quite disarranged and confusing.\\
 
@@
 \subsubsection{Graphviz Layout} \label{subsubsection:graphvizLayout}
-\code{Graphviz Layout} uses \emph{Graphviz}, third-party graph visualization software (see \cite{graphviz}), to create positions of automaton states.
+\code{Graphviz Layout} uses \emph{Graphviz}, a third-party graph visualization software (see \cite{graphviz}), to create positions of automaton states.
 To use this \code{Layout}, \emph{Graphviz} has to be installed and path to \emph{dot} binary has to be set in options.
-This \code{Layout} gives nice results even on large automata.\\
+This \code{Layout} gives nice results even for large automata.\\
 
@@
 \end{verbatim}
-The first argument is an automaton to create layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\
+The first argument is an automaton to create the layout from.
+The second, is a transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\
 
@@
 This section describes classes implementing basic common functionality that are supposed to be extended to create a new suitable visualization of automata for a particular method of inference.
-The new visualization may involve a brand new GUI panel with buttons with various functions, user interaction like selecting states or edges and others.\\
+The new visualization may involve a brand new GUI panel with buttons of various functions.
+For example, user interaction, like selecting states or edges, and others.\\
 
-Main two classes representing visualization of automaton are \code{Visualizer} and \code{AbstractComponent}.
+Two main classes representing the visualization of an automaton are \code{Visualizer} and \code{AbstractComponent}.
 \code{Visualizer} is a graphical representation of automaton and \code{AbstractComponent} is a panel (extends \code{JPanel}) containing \linebreak the \code{Visualizer} which will be displayed in GUI.\\
@@
 Relevant methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}.
-However, to save an image of automaton it is not necessary to call this methods directly.
+However, to save an image of an automaton, it is not necessary to call this methods directly.
 \jmodule{AutoEditor} GUI contains button to save an image of displayed automaton.
@@
 
-\code{PluggableVisualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package is extension of \code{Visualizer} class, which primarily provides an easy way to plug \emph{graph mouse plugins}.\\
+\code{PluggableVisualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package is an extension of \code{Visualizer} class, which primarily provides an easy way to plug \emph{graph mouse plugins}.\\
 
@@
 By default, instance of \code{PluggableVisualizer} is constructed with two plugins enabled.
-They are \code{Scaling\-Graph\-Mouse\-Plugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction.
-In the most cases these plugins are useful, otherwise they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
+They are \code{Scaling\-Graph\-Mouse\-Plugin}, providing zooming functionality, and \code{TranslatingGraphMousePlugin}, providing translation of the displayed automaton in the x and y direction.
+In most cases these plugins are useful, but otherwise, they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\
 
-Public (not inherited) methods of \code{PluggableVisualizer} are the following. Their purpose is clear from their names, for details see their JavaDoc.
+Public (not inherited) methods of \code{PluggableVisualizer} are the following. Their purpose is clear from their names, for details please see their JavaDoc.
 
@@
 	\item \code{guiInterrupt()} - Called when \jmodule{AutoEditor}'s tab is closed to propagate information about terminating of inference to a caller of \jmodule{AutoEditor}.
-There is no need to called this method directly.
+There is no need to call this method directly.
 	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by \code{guiInterrupt()} method or regularly\linebreak(\code{guiDone()} method or GUI was not waiting for user interaction).
@@
 
-Besides those methods, \code{AbstractComponent} has one abstract method, named \code{getAutomatonDrawPanel()}.\\
+Beside those methods, \code{AbstractComponent} has one abstract method, named \code{getAutomatonDrawPanel()}.\\
 
@@
 It is just needed to create \code{JPanel} and define \code{getAutomatonDrawPanel()} method to return this \code{JPanel}.
-\jmodule{AutoEditor} will take care of placing and displaying the \code{Visualizer} in the \code{JPanel}.\\
+\jmodule{AutoEditor} will take care of placing and displaying the \code{Visualizer} in the returned \code{JPanel}.\\
 
 \code{Visualizer} is not set in constructor, because it is often desired to subsequently display several different automata (\code{Visualizer}s) in the same panel.
-In this case it is not needed to create new instance of \code{AbstractComponent} for each \code{Visualizer}, but subsequently call \code{setVisualizer()} method using one instance of \code{AbstractComponent}.
+In this case, it is not needed to create a new instance of \code{AbstractComponent} for each \code{Visualizer}, but subsequently calling \code{setVisualizer()} method using one instance of \code{AbstractComponent}.
 
@@
 
-To display the component in GUI and wait for some user action, method \code{waitForGuiDone()} is used.
+To display the component in GUI and wait for some user action, we use \code{waitForGuiDone()} method.
 After displaying the component, calling this method suspends the running thread, thus code execution of a caller module is stopped at the place of this call.
-However, GUI is ran in another thread, user is able to interact with the panel (component).\\*
-Do not call method \code{waitForGuiDone()} directly.
+However, since GUI is ran in another thread, the user is able to interact with the panel (component).\\*
+Do not call method \code{waitForGuiDone()} directly!
 It is called by \jmodule{AutoEditor} when displaying the component by \jmodule{AutoEditor} \emph{API} method named \code{drawComponentAndWaitForGUI()}.
@@
 
-Method important for a programmer extending \code{AbstractComponent} is called \code{guiDone()}.
+Method important to the programmer extending \code{AbstractComponent} is called \code{guiDone()}.
 This method wakes up the thread suspended in \code{waitForGuiDone()} method and the programmer is responsible for calling it.
-Typically, it is called upon some user action like button click, vertex pick or other GUI event.\\*
-After calling of \code{guiDone()} method, code execution of the caller module is resumed and holding instance of the \code{AbstractComponent} it is able to retrieve results of user interaction, saved in its state.\\
+Typically, it is called upon some user action like a button click, vertex pick or other GUI event.\\*
+After calling of \code{guiDone()} method, code execution of the caller module is resumed, and holding instance of the \code{AbstractComponent} it is able to retrieve results of user interaction, saved in its state.\\
 
@@
 
-\code{StatesPickingVisualizer} is extension of \code{PluggableVisualizer} and upon its construction it adds \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} to the graph mouse plugins.
-With this plugin used, user can pick and unpick automaton states separately or pick several states at once by dragging rectangular selection box over states.\
+\code{StatesPickingVisualizer} is an extension of \code{PluggableVisualizer} and upon its construction it adds \code{Vertices\-Picking\-Graph\-Mouse\-Plugin} to the graph mouse plugins.
+With this plugin used, user can pick and unpick automaton states separately, or pick several states at once by dragging rectangular selection box over states.\
 Main difference compared to the \code{VertexPickingGraphMousePlugin} is that \code{VerticesPickingGraphMousePlugin} does not call component's \code{guiDone()} method.\\
@@
 \code{StatesPickingComponent} is almost the same as \code{StatePickingComponent} but it contains two extra GUI controls.
-'Continue' button and 'Don't ask me anymore to select states to be merger' checkbox.\
-The button is supposed to be clicked when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
+``Continue'' button and ``Don't ask me anymore to select states to be merged'' checkbox.
+The button is supposed to be clicked when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).
 State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.\\
@@
 In the case that none of automaton states was picked, it returns an empty list.
-	\item \code{setLabel()} - Sets text of the component label. The label can be used to communicate some information to user, for example instructions.
+	\item \code{setLabel()} - Sets text of the component label. The label can be used to communicate some information to the user, for example instructions.
 \end{itemize}
@@
 
-As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the editor window.
+As shown in figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab of the editor window.
 It consists of two buttons, horizontal line below them and a panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}).
@@
 
-Buttons have labels 'Save as Image' and 'Show Settings'.
+Buttons have labels ``Save as Image'' and ``Show Settings''.
 Pushing the first one will raise a dialog box to save currently displayed automaton to an image file.
 Set of supported image formats depends on installed JRE.\
-The second one will open \jmodule{AutoEditor} tab in NetBeans options. For description of the settings, see \ref{subsection:Settings}.
+The second one will open an \jmodule{AutoEditor} tab in NetBeans options. For description of the settings, see section \ref{subsection:Settings}.
 
@@
 To have some effect, these settings need to be implemented by extensions of \code{Visualizer} class.
-For examples, see source codes of \code{StatePickingVisualizer} and \code{StatesPickingVisualizer} classes.
+For examples, please see the source codes of \code{StatePickingVisualizer} and \code{StatesPickingVisualizer} classes.
 

### 2011-03-02 doc/jInfer/basicxsd/basicxsd.tex
last minute corrections

@@
 
-The main class implementing \code{SchemaGenerator} inference interface and simultaneously registered as its service provi\-der is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
+The main class implementing \code{SchemaGenerator} inference interface and simultaneously being registered as its service provi\-der is \code{SchemaGeneratorImpl} in package \code{cz.cuni.mff.ksi.jinfer.basicxsd}.
 Process of export consists of two phases described in detail in later sections.
@@
 \end{enumerate}
-Method \code{start()} first creates an instance of \code{Preprocessor} class supplying the rules (elements) it got in the simplified grammar on input.
+Method \code{start()} first creates an instance of \code{Preprocessor} class supplying the rules (elements) it got from the simplified grammar on input.
 Preprocessing itself is done by creating that instance (calling its constructor) and its results can be then retrieved by calling \code{getResult()} method.
@@
 If it is desired to not generate global types, this feature can be turned off in preferences.
-However, turning it off is not recommended and in some cases it may cause invalidity of resulting XSD output.
-For more information, see \ref{section:problems}.\\
+However, turning it off is not recommended and in some cases it may cause invalidity of the resulting XSD output.
+For more information, see section \ref{section:problems}.\\
 
-Minimal number of occurrences of an element to define its type globally can be altered if preferences as well. See \ref{subsection:preferences}.
+Minimal number of occurrences of an element to define its type globally can be altered if preferences as well. See subsection \ref{subsection:preferences}.
 
@@
 This is handled by \code{Indentator} class.
-This classes also serves as a buffer for string representation of XSD that the exporter is creating.\\
+This class also serves as a buffer for string representation of XSD that the exporter is creating.\\
 
@@
 
-Number of spaces characters per one level of indentation can be altered in project properties, see \ref{subsection:preferences}.
+Number of spaces characters per one level of indentation can be altered in project properties, see paragraph \ref{subsection:preferences}.
 
@@
 This subsection covers only XSD features that the exporter supports.
-For list of supported features, see \ref{subsection:supportedXSDFeatures}\\
+For list of supported features, please refer to section \ref{subsection:supportedXSDFeatures}\\
 
@@
 This class is abstract and is supposed to be extended by classes with a particular purpose.
-Its constructor signature is defined as following.
+Its constructor signature is defined as follows.
 
@@
 Parameter \code{indentator} is an instance of \code{Indentator} class to be used to buffer and indent output of exporter.
-This instance doesn't need to be empty .
+This instance doesn't need to be empty.
 Output of exporter is appended at the end of text held by the \code{indentator}.
@@
 
-If a global type contains other elements, these are processed as following.
+If a global type contains other elements, these are processed as follows.
 
@@
 
-This module is just basic implementation, thus many XSD features are not supported. List of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.
+This module is just basic implementation, thus many XSD features are not supported. A list of supported XSD elements and their attributes follows. Let \code{xs} be XMLSchema namespace.
 
@@
 
-This basic implementation of XSD export does not support and does not handle namespaces.
-That means that no namespace definitions are generated in a result, namespace definitions in input are processed as regular elements and attributes and namespace usages (namespace:element) are considered as a part of name of an element.
+This basic implementation of XSD export does not support, nor handle, namespaces; that means that no namespace definitions are generated in the result.
+Namespace definitions in input are processed as regular elements and attributes, and namespace usages (namespace:element) are considered as a part of the name of an element.
 

### 2011-03-02 doc/jInfer/xsdimport/xsdimport.tex
last minute corrections

@@
 \subsection{Naming convention}
-Note that there are several types of entities called `element' in the context of the described modules.
+Note that there are several types of entities called ``element'' in the context of the described modules.
 To avoid ambiguity, in this document we shall use notation \code{DOM.Element} 
@@
 	\item[FinalizeElement()] Checks if \code{Element} is properly defined.
-Redefine it to \code{LAMBDA} when it was empty, or to token if it only contained a simple data type.
+Redefine it to \code{LAMBDA} when it was empty, or to \code{TOKEN} if it only contained a simple data type.
 This method allows for built-in data types of tags to be retained for export.
@@
 Sax handler implements methods in response to the events generated by the underlying SAX parser API.
-Similar to its counterpart, \code{DOMHandler}, this class holds most on the logic for importing a Schema.
-It works by keeping current tree branch, which is being parsed, on a stack. Adding current node to the parent on top of the stack as current tag finishes.
+Similar to its counterpart \code{DOMHandler}, this class holds most on the logic for importing a Schema.
+It works by keeping current tree branch, which is being parsed, on a stack. Adding current node to the parent on top of the stack as current tag closes.
 
@@
 	\item List of rules that were extracted from the document are returned to \code{XSDProcessor}.
-	\item If no errors occured during this process, list of extracted rules is returned to \code{IGGeneratorImpl} for expansion.
+	\item If no errors occured during this process, list of extracted rules is returned to \code{IGGeneratorImpl} for expansion (if needed).
 Otherwise, depending on the \emph{Stop on error} setting, the inference is stopped, or an empty list of rules for this input document is returned.
@@
 	\item Rule-trees are returned from \code{DOMHandler}.
-	\item \code{DOMParser} traverses the returned rule-trees depth-first to create a list of final rules. Each of the rules now contains a full path, thanks to the way in which they were created.
+	\item \code{DOMParser} traverses the returned rule-trees, depth-first, to create a list of final rules. Each of the rules now contains a full path, thanks to the way in which they were created.
 \end{enumerate}
@@
 
-Parsers with extended or custom functionality can be created using \code{XSDImportDOM} module as example (recomended).
+Parsers with extended or custom functionality can be created using \code{XSDImportDOM} module as example (recommended).
 

### 2011-03-02 doc/jInfer/treeruledisplayer/treeruledisplayer.tex
last minute corrections + PDF

@@
 
-The main class implementing \code{RuleDisplayer} inference interface and simultaneously registered as its service provider is \code{TreeRuleDisplayer}.
-Main method of this class is \code{createDisplayer()}, which looks up the component, adds a new panel to it and renders the specified grammar in it.\\
+The main class implementing \code{RuleDisplayer} inference interface and simultaneously being registered as its service provider is \code{TreeRuleDisplayer}.
+Main method of this class is \code{createDisplayer()}, which looks up the component, adds a new panel to it, and renders the specified grammar in it.\\
 
@@
 
-The diagram of classes involved in rule display is in figure \ref{figure-class-diagram}.
+The diagram of classes involved in the process of rule displaying is in figure \ref{figure-class-diagram}.
 Main window of rule displayer may have multiple tabs in it: each contains its own \code{RuleDisplayer} responsible for rendering specified grammar using \jmodule{JUNG} library.
@@
 All the trees representing rules are created in the \code{GraphBuilder} class using the \code{buildGraphPanel()} method.
-This method creates \code{DelegateForrest} of \code{DelegateTree}s where each tree of the forrest represents one rule.
+This method creates \code{DelegateForrest} of \code{DelegateTree}s where each tree of the forest represents one rule.
 Each vertex of the tree is represented by \code{Regexp<?\ extends AbstractNamedNode>} and edge by \code{RegexpInterval}.
 While the code building tree structure is a nice recursion programming exercise, there is nothing of a special interest in it.
-Finally the method appends forrest into layout and \code{VisualizationViewer} which were described above.
-All code responsible for building rule trees is contained in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.logic}.
+Finally, the method appends forrest into layout and \code{VisualizationViewer} which were described above.
+All the code responsible for building rule trees is contained in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.logic}.
 For more details about JUNG library please see \cite{jung}.\\
@@
 All settings provided by \jmodule{TreeRuleDisplayer} are NetBeans-wide.
-The options panel along with all the logic is in the \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.options} package.
+The options panel along with all the logic is in \code{cz.cuni.mff.ksi.jinfer.treeruledisplayer.options} package.
 Available options include setting the color of the background, horizontal and vertical distance between vertices.
-Also for each type of vertex is possible to set the size, the shape and the color.
+Also, for each type of vertex, it is possible to set a size, shape and a color.
 

### 2011-03-02 doc/jInfer/paper/paper.tex
first try

--- 2011-03-01 doc/jInfer/twostep/twostep.tex
@@
-\documentclass[a4paper,10pt,oneside]{article}
+\documentclass[a4paper,10pt,oneside,twocolumn]{article}
 \usepackage{graphicx}
@@
 
-\def\mftitle{jInfer TwoStepSimplifier Module Description}
+\def\mftitle{jInfer XML schema inference framework}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \begin{document}
+\maketitle
 
-\maketitle
-\noindent Target audience: developers willing to extend jInfer, researchers willing to implement own inference methods.
+\abstract
+Today, many algorithms solving XML schema inference problem exist.
+None of them is widely used by public to practically solve the problem.
+This is because the lack of simple user interface to algorithms.
+We present NetBeans (\cite{netbeans}) based pluggable framework to fill this gap.
+Thus enabling people dealing with XML in practice to obtain schema for their documents by following simple steps, without need to read and understand theoretical aspects.
+We hope our solution will boost usage of XML schemas in practice, since creating schema for existing set of documents will be more effordable (one doesn't have to write one by hand).
 
-\noindent \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}
+\section*{Introduction}
+Although many algorithms (\cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska})
+
+\section*{Related work}
+
+%\nocite{*}
+\newpage
+\bibliographystyle{alpha}
+\bibliography{literature}
+\end{document}
 
@@
 
-\nocite{*}
-\newpage
-\bibliographystyle{alpha}
-\bibliography{literature}
-
 \end{document}

### 2011-03-02 doc/jInfer/tutorials/building_jinfer.tex
minor to building, tutorial_dev is OK
last minute corrections + PDF

@@
   This document assumes that your computer is set to use jInfer, meaning you
-    have correct version of NetBeans installed and you can run jInfer.
+    have the correct version of NetBeans installed and you can run jInfer.
 
@@
 \par 
-There are two main ways of obtaining jInfer source files: from a version
+There are two main ways for obtaining jInfer source files: from a version
   distribution, or from the Subversion repository. The first way might be more
@@
 \par 
-  The first way, obtaining sources from version distribution:
-
+\noindent   The first way, obtaining sources from version distribution:
 \begin{enumerate}\item 
@@
 \par 
-  Second way, checking out Subversion repository: you will need a
+  Second way, by checking out Subversion repository: you will need a
   Subversion client, such as the command line client or TortoiseSVN. In any case,
@@
 \par 
-  This is perhaps the easier, more user friendly way of building jInfer.
+  This is perhaps an easier, more user friendly way of building jInfer.
     If you however wish to automate the building process, you might want to

### 2011-03-02 doc/jInfer/projecttype/projecttype.tex
last minute corrections + PDF almost done

@@
 \maketitle
-\noindent Target audience: developers willing to extend jInfer, specifically extend jInfer project structure.
+\noindent Target audience: developers willing to extend jInfer, specifically to extend jInfer project structure.
 
@@
 \begin{figure}
-	\subfigure[Legend: Rectangle is folder, Rounded rectangle is file.] {
+	\subfigure[Legend: Rectangle represents a folder and rounded rectangle a file.] {
 		\centering\includegraphics[scale=\myscale]{project_dir_structure}
@@
 The second file is \emph{input.files} which is an XML file filled with paths to files selected as particular project input (structure of this file is described in section \ref{input.files}).
-\emph{output} subfolder contains schema files which were generated by jInfer inference runs.
+\emph{Output} subfolder contains schema files which were generated by jInfer inference runs.
 
@@
 \code{InputFiles} is a utility class that stores/loads input file paths into/from the \emph{input.files} XML file.
-For this purpose, \code{InputFiles} uses \emph{JAXB} \cite{JAXB} framework to unmarshall/marshall XML file into/from Java content objects.
+For this purpose, \code{InputFiles} uses \emph{JAXB} \cite{JAXB} framework to marshall or unmarshall XML file to, and from, the Java content objects.
 XML structure of \emph{input.files} is very simple, \emph{jinferinput} is the root element under which are \emph{documents}, \emph{schemas} and \emph{queries} elements.
@@
 
-Each jInfer project in NBP is open after its creation in the Projects view and has tree structure where each node is represented by a NBP \code{Node} class.
+Each jInfer project in NBP is open after its creation in the Projects view and has a tree structure where each node is represented by a NBP \code{Node} class.
 This structure in jInfer project type is provided by the \code{JInferLogicalView} class implementing \code{LogicalViewProvider} interface from NBP and is available through the \code{JInferProject} lookup.

### 2011-03-02 doc/jInfer/paper/paper.tex
Project Web: removing empty.html_. Documentation: fixing title in "paper".
minor

@@
 
-\def\mftitle{jInfer XML schema inference framework}
+\def\mftitle{jInfer XML Schema Inference Framework}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 Although many algorithms (\cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska}) 
+% TODO name all papers
+try to solve problem of infering schema for existing set of XML documents, have you ever tried to sovle the problem in practice?
+
+When the problem arises, one wants to find solution as quickly as possible, without need to investigate many scientific papers.
+We focus on this group of potential users - programmers, system administrators, xml coders, xml maintaners in private and scientific sector.
+
+Framework is designed extensible and is implemented as a set of plug-ins for NetBeans platform.
+This brings second group of users (although much smaller in counts) of framework - scientific developers willing to experiment with their new algorithms may easily gain benefits of user interface provided, thus speeding up their development, easing them comparison with other algorithms already developed for framework.
+
+But it is not all about user interface.
+We bring in XML/XSD/DTD import/export modules ready to use.
+We implement sample infering algorithm (see \cite{ahonen}).
+We provide visualization tools to visualize input XML documents as a set of grammar rules, and tools to visualize non-deterministic finite automatons used in many of infering algorithms.
+Thus extending existing NFA base algorithm with user interaction is made simple for scientific developers.
 
 \section*{Related work}
+
+\section*{Short architecture overview}
+
+\section*{More work/Open issues}
+
+\section*{Conclusion?}
 

### 2011-03-02 doc/jInfer/paper/paper.tex
minor
Documentation: "paper" review.

--- 2011-03-01 doc/jInfer/twostep/twostep.tex
@@
 
-\def\mftitle{jInfer TwoStepSimplifier Module Description}
+\def\mftitle{jInfer XML Schema Inference Framework}
 \def\mfauthor{Michal Klempa, Mário Mikula, Robert Smetana, Michal Švirec, Matej Vitásek}
@@
 \begin{document}
+\maketitle
 
-\maketitle
-\noindent Target audience: developers willing to extend jInfer, researchers willing to implement own inference methods.
+\abstract
+At the present day, there are many algorithms solving the XML schema inference problem.
+However, none of them is widely used by the public to practically solve the problem.
+This is partly because of the lack of a simple user interface to these algorithms.
+We present a NetBeans (\cite{netbeans}) based pluggable framework to fill this gap.
+This way we enable people to deal with XML in practice, to obtain schema for their documents by following simple steps, without need to read and understand theoretical aspects.
+We hope our solution will boost usage of XML schemas in practice, since creating schema for existing set of documents will be more affordable (one doesn't have to write one by hand).
 
-\noindent \emph{Note: we use the term \textbf{inference} for the act of creation of schema throughout this and other jInfer documents.}
+\section*{Introduction}
+Although many algorithms (\cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska}) 
+% TODO name all papers
+try to solve the problem of schema inference for an existing set of XML documents, have you ever tried to solve the problem in practice?
+
+When the problem arises, one wants to find a solution as quickly as possible, without the need to investigate many scientific papers.
+We focus on this group of potential users - programmers, system administrators, XML coders, XML maintaners in private and R \& D.
+
+Framework is designed to be extensible and is implemented as a set of modules for NetBeans platform.
+This means a second group of users (although much smaller in counts) of our framework - researchers willing to experiment with their new algorithms may easily gain benefits of user interface provided, thus speeding up their development, easing them comparison with other algorithms already developed for framework.
+
+But it is not all about user interface.
+We provide XML/XSD/DTD import/export modules ready to use.
+We implement a sample inference algorithm (see \cite{ahonen}).
+We provide visualization tools to display input XML documents as a set of grammar rules, and tools to visualize non-deterministic finite automatons used in many of the inference algorithms.
+This way, extending existing NFA base algorithms with user interaction is easier for researchers.
+
+\section*{Related work}
+Most of the algorithms named, doesn't have avalaible implementation, nor are made ready to use.
+Probably best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}.
+It deals with
+
+\section*{Short architecture overview}
+
+\section*{More work/Open issues}
+
+\section*{Conclusion?}
+
+%\nocite{*}
+\newpage
+\bibliographystyle{alpha}
+\bibliography{literature}
+\end{document}
 

### 2011-03-02 doc/jInfer/paper/paper.tex
more and more

@@
 Probably best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}.
-It deals with
+It deals with schema infering, user interaction, schema visualization and user-aided refinement.
+Authors use their own algorithms for schema inference (\cite{Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964}).
+Compared to our software, they are better in user interaction and schema visualization.
+There is no mention of extensibility, however, which we solve as our top priority.
+
+Since both are trying to solve nearly same problem, we predict that we will approach some good ideas from other solution in further work.
+
+Some XML editors also offers XML schema generation, 
+% TODO editors - oxygen, stylus, can they do something? how they do it, they are not free
 
 \section*{Short architecture overview}
+We divide the process of inference into three steps:
+\begin{enumerate}
+	\item 
+	TODO steps, graph from architecture, short mention of internal structure = grammar
+\end{enumerate}	
+\section*{Extensibility}
+TODO short description of how easily can framework be extended ??? maybe not
 
 \section*{More work/Open issues}
+There are still many issues to be worked on.
+First is set of supported input/output formats.
+On input, one can imagine XML Queries, on output Relax NG or Schematron. 
+There is room for implementing schema visualization tool into jInfer, to enable user-aided schema refinement.
+Many algorithms published wait to be implemented as jInfer modules, maybe some of them will be tried in study of our master theses.
 
-\section*{Conclusion?}
+\section*{Conclusion}
+Framework presented aims to be the tool used to obtain schemas from XML/XSD/DTD files (and more inputs after extending).
+While maintaining distributable bundle for users, with best of module selections and algorithms implemented so far, we predict it will be used in future studies of new algorithms as a test environment.
+Best of them will probably migrate into jInfer source tree to aid common users to solve their document set quickly enough, to even bother with schema generation at all.
 
-%\nocite{*}
+Weak side of the solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user.
+Goal was to produce stable framework for future development and we hope to negate this issue by implementing master theses in this environment, providing it with bundle of different algorithms, solving different input and output formats and infering more concise schemas.
+
+\nocite{*}
 \newpage

### 2011-03-02 doc/jInfer/paper/paper.tex
progres

@@
-\documentclass[a4paper,10pt,oneside,twocolumn]{article}
+\documentclass[a4paper,8pt,oneside,twocolumn]{article}
 \usepackage{graphicx}
@@
 \section*{Short architecture overview}
-We divide the process of inference into three steps:
+\begin{figure}
+	\centering\includegraphics[width=\myscale\columnwidth]{inference_process}
+	\caption{High-level view of the inference process} \label{inference_process}
+\end{figure}
+We divide the inference process into three steps (as illustrated on fig. \ref{inference_process}):
 \begin{enumerate}
-	\item 
-	TODO steps, graph from architecture, short mention of internal structure = grammar
+	\item Import of various formats (XML/XSD/DTD/XPath) into internal representation - \emph{Initial Grammar}.
+	\item Simlification (generalization) of \emph{Initial Grammar} with optional help of user interaction.
+	\item Export into various schema formats (XSD/DTD).
 \end{enumerate}	
+Our internal representation consist of regular grammar, that is list of rules.
+We represent right side of grammar rule as regular expression with our own set of classes to represent regular expressions.
+
+On input, all various formats are first converted into internal representation - from XML rules are extracted as positive examples, concatenations, from XSD/DTD rules are extracted as ther are written in schemata.
+From XPath we extract only explicit concatenations for now.
+
+Rules are send to \jmodule{Simplifier} module, which has to do the simplification algorithm and return \emph{Simplified Grammar}, which is then easily exporter by exporter modules.
+
+Simplification can be either automatic, or user-aided.
+We provide visualization of non-deterministic finite automatons.
+This helps algorithms that use merging state algorithm with user interaction.
+Using simple API, algorithm developer can call our library to present automaton to user to make human decisions and continue in simplification process.
+
+We provide visualization of rules itself, which is useful for methods not using automatons at all.
+Algorithm can use API to display rules
+
 \section*{Extensibility}
-TODO short description of how easily can framework be extended ??? maybe not
+Developers willing to experiment with their own algorithms have to take care only of \jmodule{Simplifier} module.
+We provide sample one, which is divided into submodules again, and again and again.
+One can find exact place he wants to replace functionality, and replace only that part of simplification chain.
+It then remains configurable whether to use his new experimental module or the old one, from GUI.
+He doesn't need to code into jInfer sources, keeping his software installed in same NB instance as jInfer is enough.
 

### 2011-03-02 doc/jInfer/paper/paper.tex
progres

@@
 
-\newcommand{\myscale}{0.74}
+\newcommand{\myscale}{0.9}
 \newcommand{\vect}[1]{\boldsymbol{#1}}
@@
 
+\begin{figure}
+	\centering\includegraphics[width=\myscale\columnwidth]{screenshot}
+	\caption{Visualization of automaton in inference process} \label{screenshot}
+\end{figure}
 Simplification can be either automatic, or user-aided.
@@
 Using simple API, algorithm developer can call our library to present automaton to user to make human decisions and continue in simplification process.
+See figure \ref{screenshot} for automaton visualization tool.
 
+\begin{figure}
+	\centering\includegraphics[width=\myscale\columnwidth]{ruledisplayer}
+	\caption{Visualization of grammar rule} \label{ruledisplayer}
+\end{figure}
 We provide visualization of rules itself, which is useful for methods not using automatons at all.
-Algorithm can use API to display rules
+Algorithm can use API to display rules to user in progress (see fig. \ref{ruledisplayer}).
 

### 2011-03-02 doc/jInfer/paper/paper.tex
progres

@@
 We hope our solution will boost usage of XML schemas in practice, since creating schema for existing set of documents will be more affordable (one doesn't have to write one by hand).
+We suggest framework will be used as a experimental sandbox for scientific developers, willing to test their own new algorithms without need to develop complicated GUI.
 
 \section*{Introduction}
-Although many algorithms (\cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska}) 
-% TODO name all papers
-try to solve the problem of schema inference for an existing set of XML documents, have you ever tried to solve the problem in practice?
+Although many algorithms (for example \cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska}) try to solve the problem of schema inference for an existing set of XML documents, have you ever tried to solve the problem in practice?
 
@@
 Probably best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}.
-It deals with schema infering, user interaction, schema visualization and user-aided refinement.
-Authors use their own algorithms for schema inference (\cite{Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964}).
-Compared to our software, they are better in user interaction and schema visualization.
-There is no mention of extensibility, however, which we solve as our top priority.
-
-Since both are trying to solve nearly same problem, we predict that we will approach some good ideas from other solution in further work.
-
-Some XML editors also offers XML schema generation, 
-% TODO editors - oxygen, stylus, can they do something? how they do it, they are not free
+It deals with schema infering and schema evolution as well.
+It employs user interaction, schema visualization and user-aided schema refinement to obtain better results.
+Authors use their own algorithms for automatic schema inference (\cite{Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964}) and then present user generated schema with option of GUI refining it.
+Schema refinement and visualization are strong sides of this solution and also possibility to do schema evolution (given old schema and not-all-valid XML documents, update schema) is unique in the field.
+Since we both, are trying to solve nearly same problem, we predict that we will approach some good ideas from other solution in further work.
+There is no mention of extensibility and pluggable design, however, which we solve as our top priority.
 
@@
 
-\section*{More work/Open issues}
+As we implement \cite{ahonen}, we provide good support for merging state algorithms - automaton visualization for example.
+We divide merging of automaton states in two tasks: merging strategy and merging criterion.
+Former one is responsible of deciding whether to merge two states or not.
+It has to decide where to stop merging and work is done, it uses merging criterion to test two states equivalency and can measure automaton as it wants to decide whether to continue or stop process.
+Merging criterion (we implement $k,h$-context) can be replaced and plugged into merging strategy. One can implement for example $s,k$-strings (\cite{Garofalakis:2000:XSE:335191.335409}).
+
+We provide replacable parts in automaton to regular expression conversion problem, so one can replace our state removal method (with simple weighted heuristic).
+
+Possibilities are open.
+
+\section*{Open issues}
 There are still many issues to be worked on.
@@
 On input, one can imagine XML Queries, on output Relax NG or Schematron. 
-There is room for implementing schema visualization tool into jInfer, to enable user-aided schema refinement.
+There is room for implementing schema visualization tool into jInfer, to enable user-aided schema refinement (to approach \cite{Bex:2008:SSI:1376616.1376750}).
+
+For now, we ignore content models of attribute and text node values, but there is interface available to be implemented.
+So simple data type detection is waiting for its imlementor.
+
 Many algorithms published wait to be implemented as jInfer modules, maybe some of them will be tried in study of our master theses.
@@
 
-Weak side of the solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user.
-Goal was to produce stable framework for future development and we hope to negate this issue by implementing master theses in this environment, providing it with bundle of different algorithms, solving different input and output formats and infering more concise schemas.
+Weak side of the solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user and
+lack of content model inference.
+Goal was to produce stable framework for future development and we hope to negate this issues by implementing master theses in this environment, providing it with bundle of different algorithms, solving different input and output formats and infering more concise schemas.
 

### 2011-03-03 doc/jInfer/paper/paper.tex
Documentation: "paper" - review up to line 100.

@@
 \section*{Related work}
-Most of the algorithms named, doesn't have avalaible implementation, nor are made ready to use.
-Probably best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}.
-It deals with schema infering and schema evolution as well.
+Most of the algorithms mentioned don't have an implementation publicly available, nor are they ready to use.
+Probably best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}, dealing with schema inference and evolution.
 It employs user interaction, schema visualization and user-aided schema refinement to obtain better results.
-Authors use their own algorithms for automatic schema inference (\cite{Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964}) and then present user generated schema with option of GUI refining it.
-Schema refinement and visualization are strong sides of this solution and also possibility to do schema evolution (given old schema and not-all-valid XML documents, update schema) is unique in the field.
-Since we both, are trying to solve nearly same problem, we predict that we will approach some good ideas from other solution in further work.
-There is no mention of extensibility and pluggable design, however, which we solve as our top priority.
+Authors use their own algorithms for automatic schema inference (\cite{Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964}) and then present a user generated schema with optional GUI to refine it.
+Schema refinement and visualization are strong points of this solution.
+Also the possibility to do schema evolution (given old schema and not-all-valid XML documents, update schema) is unique in the field.
+Since we both are trying to solve nearly the same problem, we predict that we will approach some good ideas from the other solution in further work.
+There is no mention of extensibility or pluggable design, however, which we consider our top priority.
 
@@
 \begin{enumerate}
-	\item Import of various formats (XML/XSD/DTD/XPath) into internal representation - \emph{Initial Grammar}.
-	\item Simlification (generalization) of \emph{Initial Grammar} with optional help of user interaction.
-	\item Export into various schema formats (XSD/DTD).
+	\item Import of various formats (XML/XSD/DTD/XPath/\ldots) into an internal representation - \emph{Initial Grammar}.
+	\item Simplification (generalization) of \emph{Initial Grammar} with optional help of user interaction.
+	\item Export into various schema formats (XSD/DTD/\ldots).
 \end{enumerate}	
-Our internal representation consist of regular grammar, that is list of rules.
-We represent right side of grammar rule as regular expression with our own set of classes to represent regular expressions.
+Our internal representation is a regular grammar, i.e. a list of rules.
+We represent the right side of grammar rule as a regular expression with our own set of classes to represent regexps.
 
-On input, all various formats are first converted into internal representation - from XML rules are extracted as positive examples, concatenations, from XSD/DTD rules are extracted as ther are written in schemata.
+On input, all various formats are at first converted into internal representation - from XML, rules are extracted as positive examples, concatenations; from XSD/DTD, rules are extracted as they are written in the schema.
 From XPath we extract only explicit concatenations for now.
 
-Rules are send to \jmodule{Simplifier} module, which has to do the simplification algorithm and return \emph{Simplified Grammar}, which is then easily exporter by exporter modules.
+Rules are sent to \jmodule{Simplifier} module, which has to do the simplification algorithm and return a \emph{Simplified Grammar}, which is then easily exported by \jmodule{Schema Generator} modules.
 
@@
 Simplification can be either automatic, or user-aided.
-We provide visualization of non-deterministic finite automatons.
-This helps algorithms that use merging state algorithm with user interaction.
-Using simple API, algorithm developer can call our library to present automaton to user to make human decisions and continue in simplification process.
-See figure \ref{screenshot} for automaton visualization tool.
+We provide visualization of non-deterministic finite automata.
+This helps algorithms that use merging state algorithms with user interaction.
+Using simple API, algorithm developer can call our library to present the automaton to the user to make human decisions and continue in simplification process.
+See figure \ref{screenshot} for an automaton visualization tool.
 

### 2011-03-03 doc/jInfer/paper/paper.tex
Documentation: "paper" - review round 1.

@@
 \end{figure}
-We divide the inference process into three steps (as illustrated on fig. \ref{inference_process}):
+We divide the inference process into three steps (as illustrated in fig. \ref{inference_process}):
 \begin{enumerate}
@@
 	\centering\includegraphics[width=\myscale\columnwidth]{ruledisplayer}
-	\caption{Visualization of grammar rule} \label{ruledisplayer}
+	\caption{Visualization of a grammar rule} \label{ruledisplayer}
 \end{figure}
-We provide visualization of rules itself, which is useful for methods not using automatons at all.
+We provide visualization of rules themselves, which is useful for methods not using automata at all.
 Algorithm can use API to display rules to user in progress (see fig. \ref{ruledisplayer}).
@@
 \section*{Extensibility}
-Developers willing to experiment with their own algorithms have to take care only of \jmodule{Simplifier} module.
-We provide sample one, which is divided into submodules again, and again and again.
-One can find exact place he wants to replace functionality, and replace only that part of simplification chain.
-It then remains configurable whether to use his new experimental module or the old one, from GUI.
-He doesn't need to code into jInfer sources, keeping his software installed in same NB instance as jInfer is enough.
+Developers willing to experiment with their own algorithms need to take care only of \jmodule{Simplifier} module.
+We provide a sample one, which is recursively divided into submodules.
+One can find the exact place to replace functionality he wants, and replace only that part of the simplification chain.
+It is still optional whether to use this new experimental module or the old, bundled one.
+He doesn't need to code into jInfer sources, keeping his software installed in same NB instance as jInfer is enough. %TODO
 
-As we implement \cite{ahonen}, we provide good support for merging state algorithms - automaton visualization for example.
-We divide merging of automaton states in two tasks: merging strategy and merging criterion.
+As we implement \cite{ahonen}, good support for merging state algorithms is provided - automaton visualization for example.
+We divide automaton state merging in two tasks: merging strategy and merging criterion.
 Former one is responsible of deciding whether to merge two states or not.
 It has to decide where to stop merging and work is done, it uses merging criterion to test two states equivalency and can measure automaton as it wants to decide whether to continue or stop process.
-Merging criterion (we implement $k,h$-context) can be replaced and plugged into merging strategy. One can implement for example $s,k$-strings (\cite{Garofalakis:2000:XSE:335191.335409}).
+Merging criterion (we implement $k,h$-context) can be replaced and plugged into the merging strategy. One can implement for example $s,k$-strings (\cite{Garofalakis:2000:XSE:335191.335409}).
 
-We provide replacable parts in automaton to regular expression conversion problem, so one can replace our state removal method (with simple weighted heuristic).
+We provide replacable parts in automaton-to-regexp conversion problem, so one can replace our state removal method (with simple weighted heuristic).
 
@@
 There are still many issues to be worked on.
-First is set of supported input/output formats.
-On input, one can imagine XML Queries, on output Relax NG or Schematron. 
+First one is the set of supported input/output formats.
+On input, one can imagine support for XML Queries, on output export to languages like Relax NG or Schematron. 
 There is room for implementing schema visualization tool into jInfer, to enable user-aided schema refinement (to approach \cite{Bex:2008:SSI:1376616.1376750}).
 
-For now, we ignore content models of attribute and text node values, but there is interface available to be implemented.
-So simple data type detection is waiting for its imlementor.
+For now we ignore content models of attribute and text node values, but there is an interface waiting to be implemented.
+Thus, simple data type detection is waiting for its implementor.
 
-Many algorithms published wait to be implemented as jInfer modules, maybe some of them will be tried in study of our master theses.
+Many algorithms published wait to be implemented as jInfer modules, perhaps some of them will be implemented in the course of our master theses.
 
 \section*{Conclusion}
-Framework presented aims to be the tool used to obtain schemas from XML/XSD/DTD files (and more inputs after extending).
-While maintaining distributable bundle for users, with best of module selections and algorithms implemented so far, we predict it will be used in future studies of new algorithms as a test environment.
-Best of them will probably migrate into jInfer source tree to aid common users to solve their document set quickly enough, to even bother with schema generation at all.
+The framework presented in this paper aims to be the tool used to obtain schemas from XML/XSD/DTD files (and more after extension).
+While maintaining a distributable bundle for users, with the best modules and algorithms implemented so far, we predict it will be used in future studies of new algorithms as a test environment.
+Best of them will probably merged into jInfer source tree to aid common users to solve their schema inference problems quickly and easily.
 
-Weak side of the solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user and
-lack of content model inference.
-Goal was to produce stable framework for future development and we hope to negate this issues by implementing master theses in this environment, providing it with bundle of different algorithms, solving different input and output formats and infering more concise schemas.
+Weak side of our solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user, as well as the lack of content model inference.
+Our goal was to produce a stable framework for future development and we hope to negate these issues by implementing our master theses in this environment, providing it with a bundle of different algorithms, solving different input and output formats and infering more concise schemas.
 
@@
 \end{document}
-
-\noindent \begin{tabular}{|l|l|} \hline
-Responsible developer & Michal Klempa \\ \hline
-Required tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} \\ \hline
-Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier} \\ \hline
-Module dependencies   & AutoEditor, Base, JUNG, Lookup \\ \hline
-Public packages       & none \\ \hline
-\end{tabular}
-
-%#### INTRO ############################################
-\section{Introduction}
-In \jmodule{TwoStepSimplifier} we provide the user with a complex solution to grammar simplification problem.
-While this is the only simplifier bundled with jInfer, its strong modularity ensures that it can be extended easily. 
-For most of the points of extension we indeed provide multiple implementations. 
-These vary from trivial proof-of-concept to for example $k,h$-context method for grammar rules inference (see \cite{ahonen}). 
-Strong emphasis is put on generic automata, therefore it is easy to plug-in any automaton merging state algorithm.
-In this document, we will list all modules and submodules of the simplifier and describe them briefly.
-
-%#### FACTORY ############################################
-\section{Factory pattern usage} \label{section_factory}
-In \jmodule{TwoStepSimplifier} we employ a factory pattern (as described in \cite[section 4.2]{archdoc}) to divide the module into submodules.
-Since service providing classes are being kept as singletons in the NB platform, we use them as factories. For example:
-\begin{verbatim}
-@ServiceProvider(service = AutomatonSimplifierFactory.class)
-public interface AutomatonSimplifierFactory extends 
-                     NamedModule, Capabilities, UserModuleDescription {
-  <T> AutomatonSimplifier<T> create();
-}
-\end{verbatim}
-The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (an instance of which is returned by the factory).
-For now, it is not important what the worker interface looks like, lets just examine the factory.
-
-In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces 
-so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
-Therefore, the modules must implement following methods:
-\begin{verbatim}
-String getName();
-String getDisplayName();
-String getModuleDescription();
-List<String> getCapabilities();
-String getUserModuleDescription();
-\end{verbatim}
-Nothing non-standard apart from \code{getUserModuleDescription}.
-It returns user-friendly description of the module which is then displayed in properties panels.
-
-The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic;
-but automaton simplification does not depend on any particular type (of the generic symbol) for automaton transitions, 
-so interface \code{AutomatonSimplifier<T>} is generic too.
-Factory interface deals with this by defining the \code{create()} method also as generic.
-
-Our usage of this factory pattern follows the routine:
-\begin{verbatim}
-Properties p = RunningProject.getActiveProjectProps(getName());
-
-AutomatonSimplifierFactory f = ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
-   p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER));
-
-AutomatonSimplifier<AbstractStructuralNode> autSmp = f.<AbstractStructuralNode>create();
-...
-result= autSmp.simplify(something_to_process);
-\end{verbatim}
-If a module has submodules, we implement lookups for submodule implementations in our own factory \code{create()} method.
-Worker class receives factories of all of its submodules as a constructor parameters.
-
-Lets look at \code{AutomatonMergingStateFactory}. This is the factory of a module, which has \jmodule{AutomatonSimplifier} as a submodule.
-Its create method (shortened): 
-\begin{verbatim}
-@Override
-public ClusterProcessor<AbstractStructuralNode> create() {
-  LOG.debug("Creating new ClusterProcessorAutomatonMergingState.");
-  return new AutomatonMergingState(getAutomatonSimplifierFactory(),
-                                   getRegexpAutomatonSimplifierFactory());
-}
-\end{verbatim}
-Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical. Here is the former:
-\begin{verbatim}
-private AutomatonSimplifierFactory getAutomatonSimplifierFactory() {
-  Properties p = RunningProject.getActiveProjectProps(getName());
-
-  return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
-            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER,
-                          PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
-}
-\end{verbatim}
-Cluster processor \code{AutomatonMergingState} then receives factories 
-of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
-Cluster processor then may create as many instances of submodule classes as it needs (it may be simplifying more than one automaton).
-Thorough this document, we will only mention the worker interface when describing a submodule, 
-since all factory interfaces are designed the same way as described previously.
-
-%#### STRUCTURE ############################################
-\section{Structure}
-\jmodule{TwoStepSimplifier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
-Module is divided into two classes: \code{TwoStepSimplifier} (main logic) and \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
-The latter is registered as service provider of (and implements the) \code{Simplifier} interface 
-(defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).
-
-Its main method \code{start()} receives Initial Grammar in form of
-\begin{verbatim}
-List<Element> grammar
-\end{verbatim}
-Each grammar rule is represented by an instance of \code{Element} class. 
-Element itself represents the left side of the rule, and its \code{getSubnodes()} method returns 
-regular expression representing the right side of the rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
-Its second parameter is 
-\begin{verbatim}
-SimplifierCallback callback
-\end{verbatim}
-Callback which should be invoked when the simplification is done.
-On output, simplifier provides Simplified grammar as a parameter of callback function:
-\begin{verbatim}
-void finished(List<Element> grammar);
-\end{verbatim}
-Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented the same way, as a list of elements.
-
-\code{TwoStepSimplifierFactory.start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
-Then it calls \code{simplify(initialGrammar)} method of this instance and its result is passed as parameter to callback function.
-And that's all the magic of \code{TwoStepSimplifierFactory} class.
-
-\section{Modular design}
-\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
-\begin{enumerate}
-	\item Clustering the element instances into clusters of (probably) same elements.
-	\item Inferring regular expression for each element from examples of element contents taken from all elements in a cluster.
-\end{enumerate}
-Clustering is delegated to \jmodule{Clusterer} submodule, and the task of regular expression inference 
-for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
-There is also a third submodule called \jmodule{RegularExpressionCleaner}. 
-Its purpose is just to beautify output regular expressions, no inference logic is implemented there. 
-Modules are drawn on figure \ref{twostep_modules}.
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=\myscale]{twostep_modules}
-	\caption{Submodules of TwoStep simplifier with their implementations} \label{twostep_modules}
-\end{figure}
-We provide one \jmodule{Clusterer}, four \jmodule{ClusterProcessor} and four \jmodule{RegularExpressionCleaner} implementations.
-Each of those will be explained further in this document.
-
-Method \code{TwoStepSimplifier.simplify()} basically does the following.
-\begin{verbatim}
-// 1. cluster elements
-Clusterer<AbstractStructuralNode> clusterer = clustererFactory.create();
-clusterer.addAll(initialGrammar);
-clusterer.cluster();
-
-// 2. prepare emtpy simplified grammar
-List<Element> simplifiedGrammar= new LinkedList<Element>();
-
-// 3. process rules
-ClusterProcessor<AbstractStructuralNode> processor =
-             clusterProcessorFactory.create();
-
-for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
-  AbstractStructuralNode node =  
-                          processor.processCluster(clusterer, cluster.getMembers());
-  RegularExpressionCleaner<AbstractStructuralNode> cleaner = 
-                          regularExpressionCleanerFactory.<AbstractStructuralNode>create();
-  // 4. add to rules
-  simplifiedGrammar.add(
-              new Element(node.getContext(),
-              node.getName(),
-              node.getMetadata(),
-              cleaner.cleanRegularExpression(((Element) node).getSubnodes()),
-              attList));
-}
-
-return simplifiedGrammar;
-\end{verbatim}
-First, it creates a clusterer, gives it all the rules and performs the clustering.
-Then, an empty list of elements is created as the simplified grammar.
-For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
-Finally, regular expression cleaning is done by a submodule 
-and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.
-
-\paragraph{Sentinel processing}
-There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
-Sentinels can be only on the right side of rules. They have to be sent to clusterer, 
-as someone further in the chain can ask for cluster of a sentinel element - which appears on right side.
-However, sentinels have no content, and so they must not be taken into account when inferring attribute properties 
-(whether it is required or optional), since sentinels have no attributes at all.
-
-We will now examine the submodules for clustering, processing and cleaning.
-
-%#### STRUCTURE->Clusterer ############################################
-\subsection{Clusterer submodule}
-Clustering is implemented in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
-One cluster is represented by the \code{Cluster<T>} class with \code{T} as type parameter of clustered items.
-This class simply holds a \code{Set} of members of the cluster, one of which is chosen as a \code{representant}.
-
-Any clusterer has to comply with the \code{Clusterer} interface.
-Its purpose is to cluster a bunch of elements (rules) on input into a bunch of \code{Cluster} class instances (clusters) on output.
-It has methods \code{add()} and \code{addAll()} for adding the input items to be clustered.
-Main part is the method \code{cluster()}, which does the clustering itself.
-
-As this may be a time-consuming operation, method throws \code{InterruptedException}.
-Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
-To be able to respond to calls to the \code{T getRepresentantForItem(T item)} method, 
-every clusterer implementation has to keep all its clusters even after the clustering had finished. 
-Using this method, given an item the correct cluster representative for this item has to be returned.
-
-If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
-Missing an item probably indicates error in the algorithm, rather than normal workflow.
-One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
-Basic usage of the clusterer is:
-\begin{verbatim}
-Clusterer <T> c = new MyContextClusterer<T>();
-c.addAll(initialGrammar);
-c.cluster();
-...
-c.getClusters();
-or
-c.getRepresentantForItem(x);
-\end{verbatim}
-
-%#### STRUCTURE->Clusterer->WithAttributes ############################################
-\subsubsection{ClustererWithAttributes extended interface}
-You might have noticed that the \jmodule{Clusterer} interface and \code{Cluster} class are generic.
-They may be used as design pattern not only for clustering elements in the inference process.
-To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
-It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
-type \code{T} (that is a representative of some main cluster) has some ``attribute'' clusters associated with it.
-Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.
-
-We provide one implementation of this interface described in \ref{section_Iname}.
-
-%#### STRUCTURE->ClusterProcessor ############################################
-\subsection{ClusterProcessor module}
-\jmodule{ClusterProcessor} takes rules of one cluster of element and obtains a regular expression representing that set of elements.
-It returns one element, a rule, with name set to desired name of element in schema (not all elements in a cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
-It processes attributes of all elements in a cluster to obtain meaningful schema attribute specification. These attributes have to be attached to the resulting element.
-
-Interface of submodule is defined as follows.
-\begin{verbatim}
-public interface ClusterProcessor<T> {
-  T processCluster(
-          Clusterer<T> clusterer,
-          List<T> rules
-          ) throws InterruptedException;
-}
-\end{verbatim}
-Why is the cluster processor given a clusterer instance?
-Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
-Clusterer can tell you the representative for any arbitrary element from input (not only those elements in rules, but also those that may be on right side of rules).
-Also, clusterer (if it is with attributes) has information about attributes in each cluster.
-
-We describe each \jmodule{ClusterProcessor} implementation shipped in jInfer in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.
-
-%#### STRUCTURE->RegularExpressionCleaner ############################################
-\subsection{RegularExpressionCleaner module}
-In this last subsection, we examine the \jmodule{RegularExpressionCleaner} interface.
-Purpose of this submodule is only to make corrections to regular expression, to make them nicer on output.
-For example, it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
-To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
-Interface definition is straightforward here.
-\begin{verbatim}
-public interface RegularExpressionCleaner<T> {
-  Regexp<T> cleanRegularExpression(Regexp<T> regexp);
-}
-\end{verbatim}
-Given a regular expression, return cleaned regular expression.
-Conversion of regular expression is commonly a recursive task and all our implementations work this way.
-
-%#### PREFERENCES ############################################
-\section{Preferences}
-All settings provided by \jmodule{TwoStepSimplifier} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.twostep.properties} package.
-It is possible to set the following.
-\begin{itemize}
-	\item Select \jmodule{Clusterer} submodule implementation from those installed.
-	\item Select \jmodule{ClusterProcessor} submodule implementation from those installed.
-	\item Select \jmodule{RegularExpressionCleaner} submodule implementation from those installed.
-\end{itemize}
-
-%#### DATA FLOW ############################################
-\section{Data flow}
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=1]{twostep_flow}
-	\caption{TwoStep data flow} \label{twostep_flow}
-\end{figure}
-Process illustrated in figure \ref{twostep_flow} is as follows.
-\begin{enumerate}
-	\item Initial Grammar on input, which is sent into \jmodule{Clusterer}.
-	\item Each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
-	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
-	\item Regular expressions representing all the processed clusters are added to the list of simplified grammar rules.
-	\item List of simplified rules is returned (Simplified grammar).
-\end{enumerate}
-
-%#### SUBMODULES ############################################
-\section{Submodule implementations}
-%#### SUBMODULES->Iname ############################################
-\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
-We cluster elements in \code{Iname\.<Abstract\.Structural\.Node, Attri\.bu\.te>} class, which takes \code{Abstract\.Structural\.No\.de} classes to cluster as main, and \code{Attribute} classes as attributes.
-Clustering is done based on element's \code{getName()} equality (ignoring case).
-For each rule (element), it is first clustered by finding the cluster which representative has the same name - or create new cluster with this element as representative.
-Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
-Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
-Each node on right side, is examined in the following way.
-\begin{itemize}
-  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster.
-It does nothing more, than holds all \code{SimpleData} instances in one cluster.
-But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
-  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representative of same \code{getName()}, or create new cluster with this sentinel as representative.
-From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future).
-So there can't be cluster with only one sentinel element in it.
-  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule.
-\end{itemize}
-Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
-We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
-This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
-It clusters them by name, case insensitive.
-Each main cluster also has \code{SimpleData} clusterer associated with it.
-Whole scheme is on figure \ref{clusterer_structure}.
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=1]{clusterer_structure}
-	\caption{Iname clusterer structure} \label{clusterer_structure}
-\end{figure}
-
-Once again, there are main clusters for elements, for each main cluster there are two helper clusterers - one for attribute clusters and one for simple data clusters.
-
-%#### SUBMODULES->PassRepresentant ############################################
-\subsection{PassRepresentant (ClusterProcessor)} \label{section_PassRepresentant}
-Trivial ClusterProcessor implementation. For each cluster return its representative as the rule to be in resulting schema.
-This has nothing to do with grammar simplification, it is just a proof-of-concept submodule.
-Input documents are not valid against this odd grammar.
-Do not use this in practice, just read the code to understand the bare minimum needed to implement a submodule.
-
-%#### SUBMODULES->Alternations ############################################
-\subsection{Alternations (ClusterProcessor)} \label{section_Alternations}
-This processor simply gets all right sides from elements in cluster, puts them in one big list and creates alternation regular expression with this list as children.
-That is, it creates one big rule with alternation of every positive example observed.
-No generalization is done at all.
-
-%#### SUBMODULES->Trie ############################################
-\subsection{Trie (ClusterProcessor)} \label{section_Trie}
-This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie'') of them.
-More precisely, it takes the first rule and declares it to be a long branch (concatenation of tokens) in a newly created tree.
-After that, it adds the remaining rules one by one as branches like this: as long as it can follow an existing branch, it follows it.
-As soon as the newly added branch starts to differ, it ``branches off'' (creates an alternation at that point) the existing tree and hangs the rest of the newly added rule there.
-Repeating this process creates a prefix tree describing all the rules in the cluster.
-The tree is the final regular expression.
-
-%#### SUBMODULES->AutomatonMergingState ############################################
-\subsection{AutomatonMergingState (ClusterProcessor)} \label{section_AutomatonMergingState}
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=\myscale]{automaton_merging_state_modules}
-	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_modules}
-\end{figure}
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=1]{automaton_merging_state_flow}
-	\caption{Data flow of AutomatonMergingState cluster processor} \label{automaton_merging_state_flow}
-\end{figure}
-\jmodule{AutomatonMergingState} is an implementation of merging state algorithm on nondeterministic finite automaton (see package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate}).
-\subsubsection{Structure} \label{section_AutomatonMergingState_structure}
-Abridged code of cluster processor operation follows.
-\begin{verbatim}
-// 1. Construct PTA
-Automaton<AbstractStructuralNode> automaton = new Automaton<AbstractStructuralNode>(true);
-
-// Take each rule in cluster and pass right side to automaton to create PTA
-for (AbstractStructuralNode instance : rules) {
-  Element element = (Element) instance;
-  Regexp<AbstractStructuralNode> rightSide = element.getSubnodes();
-
-  List<AbstractStructuralNode> rightSideTokens = rightSide.getTokens();
-
-  List<AbstractStructuralNode> symbolString = new LinkedList<AbstractStructuralNode>();
-  for (AbstractStructuralNode token : rightSideTokens) {
-    symbolString.add(clusterer.getRepresentantForItem(token));
-  }
-  automaton.buildPTAOnSymbol(symbolString);
-}
-
-// 2. Simplify automaton by merging states using automatonSimplifier
-Automaton<AbstractStructuralNode> simplifiedAutomaton =
-    automatonSimplifier.simplify(automaton, elementSymbolToString);
-
-// 3. Convert Automaton<AbstractStructuralNode> to RegexpAutomaton<AbstractStructuralNode>
-RegexpAutomaton<AbstractStructuralNode> regexpAutomaton =
-    new RegexpAutomaton<AbstractStructuralNode>(simplifiedAutomaton);
-
-// 4. Call regexpAutomatonSimplifier to obtain regular expression from regexpAutomaton
-Regexp<AbstractStructuralNode> regexp =
-    regexpAutomatonSimplifier.simplify(regexpAutomaton, regexpAbstractToString);
-
-// 5. Return element with regexp
-return new Element(
-      new ArrayList<String>(),
-      rules.get(0).getName(),
-      new HashMap<String, Object>(),
-      regexp,
-      new ArrayList<Attribute>()
-      );
-\end{verbatim}
-The whole submodule structure is in figure \ref{automaton_merging_state_modules}.
-
-\subsubsection{Data flow}
-\jmodule{AutomatonMergingState} works in following steps.
-\begin{enumerate}
-	\item Create prefix-tree automaton (PTA) from positive examples - right sides of rules.
-Call its submodule \jmodule{Au\.to\.ma\.ton\.Simpli\.fier} to modify PTA to some generalized automaton by merging states.
-	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.
-Regexp automaton is automaton with regular expression as symbol on transitions.
-In automata theory, such automaton is called extended NFA.
-Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
-	\item \code{RegexpAutomaton} is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
-Its job is to derive regular expression from automaton, such that automaton and regular expression represent the same language.
-\end{enumerate}
-The data flow is depicted in figure \ref{automaton_merging_state_flow}.
-
-\subsubsection{Data structures - Automaton} \label{section_Automaton}
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=\myscale]{pta}
-	\caption{PTA created from input strings: $aa, aaaaa, abaa, bab, baa$} \label{pta}
-\end{figure}
-Class \code{Automaton} is shortly described in \cite{archdoc}.
-Let's take a look at method \code{build\.PTA\.On\.Sym\.bol} here.
-Given \code{List<T> symbolString} it traverses automaton - comparing transition symbols with symbols in string.
-It follows transitions in automaton until first difference with string symbol is found (like in prefix trees).
-On different symbol, it creates new branch and all states and transitions on that branch.
-Last state visited gets its \code{finalCount} value incremented, as one more input string ended in this state.
-The code is as follows.
-\begin{verbatim}
-public void buildPTAOnSymbol(List<T> symbolString) {
-  State<T> xState= this.getInitialState();
-  for (T symbol : symbolString) {
-    Step<T> xStep= this.getOutStepOnSymbol(xState, symbol);
-    if (xStep != null) {
-      xStep.incUseCount();
-      xState= xStep.getDestination();
-    } else {
-      State<T> newState= this.createNewState();
-      Step<T> newStep= this.createNewStep(symbol, xState, newState);
-      assert newStep.getDestination().equals(newState);
-      xState= newStep.getDestination();
-    }
-  }
-  xState.incFinalCount();
-}
-\end{verbatim}
-Starting from initial state, traverse the automaton by asking for \code{getOutStepOnSymbol()}.
-If there is such step (transition), follow it to its destination state and move the position in \code{symbolString}.
-If there is no such step, create one and create also the destination state, then follow this newly created step to new state.
-
-As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
-By the way, this is why we cluster all simple data and all sentinels on right sides of rules.	
-While adding a concatenation to the automaton, instead of the instances themselves their \emph{representatives} are used. See \ref{section_AutomatonMergingState_structure} for further reference.
-This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representative is only present in automaton.
-That results in cluster-equal behaviour.
-Example of prefix-tree automaton is in figure \ref{pta}.
-
-\subsubsection{Data structures - RegexpAutomaton} \label{section_RegexpAutomaton}
-Class \code{RegexpAutomaton} extends class \code{Automaton} with generic type \code{T} set to \code{Regexp<T>}.
-That is, you have to instantiate it with same symbol type (e.g. \code{AbstractStructuralNode}) as automaton.
-Implementation is in package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate.regexping}. This is the same package where \jmodule{RegexpAutomatonSimplifier} interface is defined.
-Implementation is nothing special - there is a copy constructor which creates regular expression automaton from ordinary automaton by enclosing each symbol on transition of original automaton in a \code{Regexp.TOKEN}.
-
-\subsubsection{AutomatonSimplifier module}
-\jmodule{AutomatonSimplifier} module has its worker interface defined as following.
-\begin{verbatim}
-public interface AutomatonSimplifier<T> {
-  Automaton<T> simplify(Automaton<T> inputAutomaton, 
-                        SymbolToString<T> symbolToString) throws InterruptedException;
-  Automaton<T> simplify(Automaton<T> inputAutomaton, 
-                        SymbolToString<T> symbolToString,
-                        String elementName) throws InterruptedException;
-}
-\end{verbatim}
-Given input automaton it returns automaton, there is nothing magical to it.
-Interface is defined as generic, implementations don't have to be. However, there is no reason not to make them generic.
-When implementation needs to present an automaton to user, it needs string representation of symbols to display on automaton transitions.
-For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
-It is responsible for converting symbol to string.
-Overloaded version is to inform the user about name of processed element when presenting automaton in \jmodule{AutoEditor}.
-This version of interface will probably change in future.
-
-This generality means that simplifier can not only be used to infer rules for elements, but possibly also to infer patterns strings for content model of attributes using the same algorithms.
-The actual implementation of this still waits for its developer to come.
-Implementations of \jmodule{AutomatonSimplifier} are discussed in \ref{section_Greedy} and \ref{section_UserInteractive}.
-
-\subsubsection{RegexpAutomatonSimplifier submodule}
-Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain a regular expression from a given automaton.
-Interface for doing is is thus simple.
-\begin{verbatim}
-public interface RegexpAutomatonSimplifier<T> {
-  Regexp<T> simplify(RegexpAutomaton<T> inputAutomaton, 
-                     SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
-}
-\end{verbatim}
-Once again we encounter \code{symbolToString} with the same purpose as before, if submodule would need to present the automaton to user, it has to be able to convert symbols of (at runtime) unknown type to strings.
-Our implementation of \jmodule{RegexpAutomatonSimplifier} is discussed in \ref{section_StateRemoval}
-
-\subsubsection{MergeConditionTester submodule}
-\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{Au\.to\.ma\.ton\.Merg\.ing\.Sta\.te}.
-It is at disposal of implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.
-
-One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.
-
-\subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
-It converts nested concatenations into flat ones.
-Example: $(name, (person, id))$ is converted to $(name, person, id)$.
-
-\subsection{Null (RegularExpressionCleaner)} \label{section_Null}
-This cleaner just returns regexp it receives and does nothing.
-Plug this one into chain, if you want to omit this step in inference.
-
-\subsection{EmptyChildren (RegularExpressionCleaner)} \label{section_EmptyChildren}
-Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
-If for example an inference method produces unnecessary empty regexps in regexp tree such as
-$$
-(name, (), (person, id))
-$$
-There is one empty concatenation/alternation/permutation with no sense.
-This is wiped out by this cleaner to produce 
-$$
-(name, (person, id))
-$$
-
-\subsection{Chained (RegularExpressionCleaner)} \label{section_Chained}
-Allows the user to chain more cleaners with output from first one plugged as input into second one and so one.
-Thus regular expression from inference can pass through \jmodule{EmptyChildren} and then through \jmodule{NestedConcatenation} cleaners and then it is returned.
-
-\subsection{Greedy (AutomatonSimplifier)} \label{section_Greedy}
-We implement greedy strategy of automaton simplifying in class \code{Greedy}.
-It simply asks given \code{MergeConditionTester} if it can merge any of automaton states and merges states until there are no states to be merged.
-\begin{verbatim}
-while (there exist pair of states that are equivalent) {
-  merge them
-}
-\end{verbatim}
-We provide one \jmodule{MergeConditionTester} implementation, see \ref{section_KHContext}.
-
-\subsection{UserInteractive (AutomatonSimplifier)} \label{section_UserInteractive}
-We have created automaton simplifier which shows the user the input automaton, asking him to select states to merge.
-Then it merges them and asks user again.
-\begin{verbatim}
-repeat {
-  draw automaton
-  if (user selects more than one state to merge) {
-    merge them
-  }
-} until (user selected at least one state in last attempt)
-\end{verbatim} 
-We use \jmodule{AutoEditor} module described in \cite{autoeditordoc}.
-A screenshot of automaton visualization is in picture \ref{screenshot}.
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=0.5]{screenshot}
-	\caption{Screenshot of UserInteractive (AutomatonSimplifier).} \label{screenshot}
-\end{figure}
-
-\subsection{KHContext (MergeConditionTester)} \label{section_KHContext}
-We have implemented \emph{k,h-context} state equivalence (see \cite{ahonen}) in class \code{KHContext} (implements \code{Mer\.ge\.Con\.di\.tion\.Tes\.ter} interface), which is used by \code{Greedy} to test mergeability of states.
-We don't use $k-grams$ algorithm from \cite{ahonen}, just a simple DFS.
-
-\subsection{StateRemoval (RegexpAutomatonSimplifier)} \label{section_StateRemoval}
-\begin{wrapfigure}{R}{40mm}
-	\centering\includegraphics[scale=\myscale]{state_removal_modules}
-	\caption{Modules of StateRemoval.} \label{state_removal_modules}
-	\vspace{10mm}
-\end{wrapfigure}
-We are using state removal method (see \cite{1224321}) to convert regexp automaton into equivalent regular expression.
-This is implemented in \code{StateRemoval} class (figure \ref{state_removal_modules}) which implements \code{RegexpAutomatonSimplifier} interface.
-Before the algorithm starts, states \emph{superInitial} and \emph{superFinal} are added to automaton, former with $\lambda$-transition to initial state.
-From all final states a $\lambda$ transition to \emph{superFinal} state is added.
-State removal works by removing states of automaton (and redirecting transitions properly) until there are last to two states - \emph{superInitial} and \emph{superFinal}.
-After removing all states, there is only one transition from \emph{superInitial} to \emph{superFinal} state.
-That transition has the final regular expression as symbol on it. This one is returned.
-
-We defined one submodule of this class with \jmodule{Orderer}.
-It has only one method to implement: \code{getStateToRemove}.
-Given automaton, it has to return reference to one state which should be removed from automaton at first.
-State removal calls this submodule and removes state returned.
-
-\subsubsection{Data Structures - StateRemovalRegexpAutomaton}
-We extend \code{RegexpAutomaton} to obtain automaton that fits the state removal procedure.
-This implementation has methods \code{crea\.te\.Su\.per\.In\.it\.ial\.Sta\.te} and \code{crea\.te\.Su\.per\.In\.it\.ial\.Sta\.te}.
-They does exactly what their names stands for.
-\emph{SuperInitial} state has only one $\lambda$-out-transition pointing to original initial state.
-\emph{SuperFinal} state has as many $\lambda$-in-transitions, as there are original final states - from each one it gets one.
-Without going further into many technical details, we just mention method \code{removeState}, which removes given state from automaton.
-Transitions are handled in the following way.
-\begin{enumerate}
-	\item Collapse all loop transitions of removed state to one loop transition with new \code{Regex.ALTERNATION} of all regexps on original loops (with Kleene interval set).
-	\item Collapse all {in | out}-transitions, which same {source | destination} state to one transition with \code{Reg\.exp.AL\.TER\.NA\.TION} of original regexps.
-	\item For each in-transition, loop, and out-transition, add transition from in-transition source to out-transition destination with symbol of \code{Regexp.CONCATENATION} of in-transition regexp, loop regexp and out-transition regexp.
-Thus by-passing the state completely.
-	\item Remove the state and transitions associated with it.
-\end{enumerate}
-The whole process is step-by-step illustrated in figure \ref{regaut}.
-\begin{figure}[pb]
-	\centering
-	\subfigure[RegexpAutomaton after conversion from ordinary automaton.]
-	{
-		\includegraphics[scale=\myscale]{regaut1}
-		\label{regaut1}
-	}
-	\subfigure[First, collapse loop transitions to alternation with Kleene interval set.]
-	{
-		\includegraphics[scale=\myscale]{regaut2}
-		\label{regaut2}
-	}
-	\subfigure[Collapse {in | out}-transitions {from | to} same states.]
-	{
-		\includegraphics[scale=\myscale]{regaut3}
-		\label{regaut3}
-	}
-	\subfigure[Bypass state 3 by adding concatenation transitions from states 1 and 2 to state 4. Remove state 3 and associated transitions.]
-	{
-		\includegraphics[scale=\myscale]{regaut4}
-		\label{regaut4}
-	}
-%	
-%	\subfigure[Finally remove state 3 and associated transitions.]
-%	{
-%		\includegraphics[scale=\myscale]{regaut5}
-%		\label{regaut5}
-%	}
-	\caption{RegexpAutomaton and removing state 3 from it.} \label{regaut}
-\end{figure}
-
-\subsection{Weighted (Orderer)} \label{section_Weighted}
-We implement one orderer, called \code{Weighted}.
-It is simple heuristic - weights all states (weight = sum of {in | out | loop}-transition regular expression lengths) and returns state with lowest weight.
-
-%#### EXTENSIBILITY ############################################
-\section{Extensibility}
-If you are willing to, you can replace each of three submodules of \jmodule{TwoStepSimplifier}.
-It may be useful to replace the default \jmodule{ClustererWithAttributes} implementation with a clever one.
-And regular expression cleaning will be probably the aim of new implementations, too.
-We recommend replacing other parts at lower levels, however.
-You will need to replace \jmodule{ClusterProcessor} only if you want to write inference method that is not based on merging state algorithm (or NFA's at all).
-
-If this is not your case, you can replace module \jmodule{AutomatonSimplifier} to implement ACO heuristics or MDL principle in searching solutions of merging states.
-
-Of course, merge criterion is probably the point to add implementation, implement \jmodule{MergeConditionTester} interface to obtain for example $s,k$-string merge criterion.
-
-Changing ordering of states to remove in \jmodule{StateRemoval} module is easy by replacing \jmodule{Orderer} implementation, for example to one suggested in \cite{1224321} may lead to shorter resulting regular expressions.
-
-Whole TwoStep submodules structure is in figure \ref{twostep_modules_complete}.
-\begin{figure}[tpb]
-	\centering\includegraphics[scale=0.37]{twostep_modules_complete}
-	\caption{Modules of TwoStep simplifier and their submodules. Filled classes are default selection (best of).} \label{twostep_modules_complete}
-\end{figure}
-
-\end{document}

### 2011-03-03 doc/jInfer/paper/paper.tex
Corrected PAPER!

@@
 \abstract
-At the present day, there are many algorithms solving the XML schema inference problem.
-However, none of them is widely used by the public to practically solve the problem.
+At present day, there are many algorithms for solving the problem of XML schema inference.
+However, virtually none of them is widely used by the public to solve this task in practice.
 This is partly because of the lack of a simple user interface to these algorithms.
 We present a NetBeans (\cite{netbeans}) based pluggable framework to fill this gap.
-This way we enable people to deal with XML in practice, to obtain schema for their documents by following simple steps, without need to read and understand theoretical aspects.
-We hope our solution will boost usage of XML schemas in practice, since creating schema for existing set of documents will be more affordable (one doesn't have to write one by hand).
-We suggest framework will be used as a experimental sandbox for scientific developers, willing to test their own new algorithms without need to develop complicated GUI.
+This way we enable people to deal with XML in practice, to obtain schemas for their documents by following a few simple steps, without the need to read and understand theoretical aspects.
+We hope our solution will encourage the usage of XML schemas in practice, since the creation of a schema for existing sets of documents will be more affordable (one doesn't have to write it by hand).
+We suggest that our framework can also be used as an experimental sandbox for scientific developers, willing to test their new algorithms without the need to develop a complicated GUI.
 
 \section*{Introduction}
-Although many algorithms (for example \cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska}) try to solve the problem of schema inference for an existing set of XML documents, have you ever tried to solve the problem in practice?
+Although many algorithms (for example those proposed by \cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, 1802522, vyhnanovska}) try to solve the problem of schema inference for an existing set of XML documents, have you ever tried to solve the problem in practice?
 
 When the problem arises, one wants to find a solution as quickly as possible, without the need to investigate many scientific papers.
-We focus on this group of potential users - programmers, system administrators, XML coders, XML maintaners in private and R \& D.
+We focus on this group of potential users - programmers, system administrators, XML coders, XML maintaners in private and corporate R\&D.
 
-Framework is designed to be extensible and is implemented as a set of modules for NetBeans platform.
-This means a second group of users (although much smaller in counts) of our framework - researchers willing to experiment with their new algorithms may easily gain benefits of user interface provided, thus speeding up their development, easing them comparison with other algorithms already developed for framework.
+Our framework is designed to be extensible and is implemented as a set of modules for NetBeans platform.
+This means that the second group of users of our framework - the researchers willing to experiment with their new algorithms - may easily gain benefits from the user interface we provide.
+Thus, speeding up their development,  making easier the comparison with other algorithms already developed for the framework.
 
-But it is not all about user interface.
-We provide XML/XSD/DTD import/export modules ready to use.
+Nevertheless, it is not all about the user interface.
+We provide XML/XSD/DTD import and export modules that are ready to use.
 We implement a sample inference algorithm (see \cite{ahonen}).
@@
 Most of the algorithms mentioned don't have an implementation publicly available, nor are they ready to use.
-Probably best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}, dealing with schema inference and evolution.
+Probably the best solution nowadays is \cite{Bex:2008:SSI:1376616.1376750}, dealing with schema inference and evolution.
 It employs user interaction, schema visualization and user-aided schema refinement to obtain better results.
 Authors use their own algorithms for automatic schema inference (\cite{Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964}) and then present a user generated schema with optional GUI to refine it.
-Schema refinement and visualization are strong points of this solution.
-Also the possibility to do schema evolution (given old schema and not-all-valid XML documents, update schema) is unique in the field.
-Since we both are trying to solve nearly the same problem, we predict that we will approach some good ideas from the other solution in further work.
-There is no mention of extensibility or pluggable design, however, which we consider our top priority.
+Schema refinement and visualization are strong points of their solution.
+Also, the possibility to do schema evolution (given old schema and not-all-valid XML documents, to update the schema) is unique in the field.
+Since we are both trying to solve (almost) the same problem, we predict that we will approach some good ideas from other known solutions in our further work.
+There is no mention of extensibility or pluggable design, which we, on the other hand, consider our top priority.
 
@@
 \end{figure}
-We divide the inference process into three steps (as illustrated in fig. \ref{inference_process}):
+We divide the inference process into three steps (as illustrated in figure \ref{inference_process}):
 \begin{enumerate}
-	\item Import of various formats (XML/XSD/DTD/XPath/\ldots) into an internal representation - \emph{Initial Grammar}.
+	\item Import of various formats (XML, XSD, DTD, XPath, etc.) into an internal representation - \emph{Initial Grammar}.
 	\item Simplification (generalization) of \emph{Initial Grammar} with optional help of user interaction.
@@
 \section*{Extensibility}
-Developers willing to experiment with their own algorithms need to take care only of \jmodule{Simplifier} module.
+Developers willing to experiment with their own algorithms need to consider the \jmodule{Simplifier} module only.
 We provide a sample one, which is recursively divided into submodules.
-One can find the exact place to replace functionality he wants, and replace only that part of the simplification chain.
-It is still optional whether to use this new experimental module or the old, bundled one.
-He doesn't need to code into jInfer sources, keeping his software installed in same NB instance as jInfer is enough. %TODO
+One can find the exact place where to implement the functionality he wants, and replace only that part of the simplification chain.
+It is still optional, whether to use this new experimental module or the old, bundled one.
+A developer doesn't need to submit modified sources to our repository, just keeping his software (module) installed in the same NetBeans instance as jInfer will suffice.
 
-As we implement \cite{ahonen}, good support for merging state algorithms is provided - automaton visualization for example.
+Due to the fact that we implement \cite{ahonen}, a good support for merging state algorithms is already provided
+(for example automaton visualization).
 We divide automaton state merging in two tasks: merging strategy and merging criterion.
-Former one is responsible of deciding whether to merge two states or not.
+The former one is responsible for deciding whether to merge two states or not.
 It has to decide where to stop merging and work is done, it uses merging criterion to test two states equivalency and can measure automaton as it wants to decide whether to continue or stop process.
@@
 
-We provide replacable parts in automaton-to-regexp conversion problem, so one can replace our state removal method (with simple weighted heuristic).
+We provide replacable parts in automaton-to-regexp conversion problem, so one can replace our state removal method with, for exmple, a simple weighted heuristic.
 
@@
 \section*{Open issues}
-There are still many issues to be worked on.
+There are, however, still many issues to work on.
 First one is the set of supported input/output formats.
@@
 For now we ignore content models of attribute and text node values, but there is an interface waiting to be implemented.
-Thus, simple data type detection is waiting for its implementor.
+Thus, simple data type detection is waiting for its programmer.
 
-Many algorithms published wait to be implemented as jInfer modules, perhaps some of them will be implemented in the course of our master theses.
+Many algorithms published, wait to be implemented as jInfer modules, perhaps some of them will be implemented in the course of our master theses.
 
@@
 While maintaining a distributable bundle for users, with the best modules and algorithms implemented so far, we predict it will be used in future studies of new algorithms as a test environment.
-Best of them will probably merged into jInfer source tree to aid common users to solve their schema inference problems quickly and easily.
+The best of them will probably be merged into jInfer source tree to aid common users to solve their schema inference problems quickly and easily.
 
-Weak side of our solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user, as well as the lack of content model inference.
+Weak side of our solution is the lack of good-quality automatic or semi-automatic algorithms to be used by common user, as well as the lack of content model inference.
 Our goal was to produce a stable framework for future development and we hope to negate these issues by implementing our master theses in this environment, providing it with a bundle of different algorithms, solving different input and output formats and infering more concise schemas.

### 2011-05-09 doc/theses/template/diplomka.tex
theses template

\documentclass[a4paper,12pt,oneside]{report}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{subfigure}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\usepackage{tgpagella}
\usepackage{xstring}
\usepackage{wrapfig}

\newcommand{\myscale}{0.74}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
\def\.{\discretionary{}{}{}}
\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
\linespread{1.5}

\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
\setlength{\voffset}{-1in} %analogous voffset
\setlength{\oddsidemargin}{4cm}
\setlength{\evensidemargin}{4cm}
\setlength{\topmargin}{25mm}
\setlength{\footskip}{1cm}
\setlength{\headheight}{0cm}
\setlength{\headsep}{0cm}
\setlength{\marginparwidth}{0cm}
\setlength{\marginparpush}{0cm}
\setlength{\textheight}{23.7cm}
\setlength{\textwidth}{14.5cm}
\let\openright=\clearpage

\def\mfauthor{Michal Klempa}
\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
\def\mfplacedate{Praha, 2011}

% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
\makeatletter
\def\@makechapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries \thechapter. #1
   \par\nobreak
   \vskip 20\p@
}}
\def\@makeschapterhead#1{
  {\parindent \z@ \raggedright \normalfont
   \Huge\bfseries #1
   \par\nobreak
   \vskip 20\p@
}}
\makeatother

\def\chapwithtoc#1{
\chapter*{#1}
\addcontentsline{toc}{chapter}{#1}
}

\begin{document}

\lefthyphenmin=2
\righthyphenmin=2

%%% Titulní strana práce ======================================================
\pagestyle{empty}
\begin{center}

\large

Charles University in Prague

\medskip

Faculty of Mathematics and Physics

\vfill

{\bf\Large MASTER THESIS}

\vfill

\centerline{\mbox{\includegraphics[width=60mm]{logo.eps}}}

\vfill
\vspace{5mm}

{\LARGE \mfauthor}

\vspace{15mm}

% exactly as assigned
{\LARGE\bfseries Title of the thesis}

\vfill

% Název katedry nebo ústavu, kde byla práce oficiálně zadána
% (dle Organizační struktury MFF UK)
Department of Software Engineering

\vfill

\begin{tabular}{rl}

Supervisor of the master thesis: & 	\mfadvisor \\
\noalign{\vspace{2mm}}
Study programme: & Informatika \\
\noalign{\vspace{2mm}}
Specialization: & ISS \\
\end{tabular}

\vfill

% Zde doplňte rok
\mfplacedate

\end{center}

\newpage % ============================================================
%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
\openright

\noindent
[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
the person who lent the software, literature, etc.)]

\newpage % ============================================================
%%% Strana s čestným prohlášením k diplomové práci
\vglue 0pt plus 1fill
\noindent
I declare that I carried out this master thesis independently, and only with the cited
sources, literature and other professional sources.

\medskip\noindent
I understand that my work relates to the rights and obligations under the Act No.
121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.

\vspace{10mm}

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
signature
\hss}}

\vspace{20mm}

\newpage % ============================================================
%%% Povinná informační strana diplomové práce
\vbox to 0.5\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Název práce: Title [accurate translation of the master thesis title in Czech]

Autor: \mfauthor

Katedra:  Katedra softwarového inženýrství

Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství

Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
master thesis]

Klíčová slova:  [3-5 keywords in Czech]

\vss}\nobreak\vbox to 0.49\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Title:  [the master thesis title in English]

Author: \mfauthor

Department: Department of Software Engineering

Supervisor: \mfadvisor{}, Department of Software Engineering

Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
the master thesis]

Keywords: [3-5 keywords in English]

\vss}

\newpage

%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
%%% na začátku práce, místo na jejím konci.

\openright
\pagestyle{plain}
\setcounter{page}{1}
\tableofcontents

% =============== TEXT ================
\newpage
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec tortor neque, porta a posuere quis, tristique nec diam. Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci. Sed congue vulputate auctor. Morbi laoreet adipiscing fringilla. Aliquam sit amet ante mi. Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus. Nam nec mollis sapien. Sed scelerisque bibendum suscipit. Maecenas ornare est non sapien elementum bibendum a ac diam. Proin nibh arcu, elementum eu lobortis vel, suscipit a augue. Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis. Maecenas consequat tristique pulvinar. Donec porta feugiat neque, ac semper est bibendum vehicula. Vivamus gravida mauris in neque tempus tincidunt. Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
\chapter*{Introduction / Preface}
\addcontentsline{toc}{chapter}{Introduction / Preface}

Aenean in odio eu nisi imperdiet condimentum. Suspendisse id ipsum sed risus pharetra vehicula bibendum sed neque. Nulla id elementum augue. Integer vehicula interdum orci, at gravida ante interdum id. Nunc laoreet orci non nisi cursus vel tincidunt orci venenatis. Quisque blandit odio ac ipsum lacinia a gravida elit mattis. Praesent sagittis, arcu ac accumsan rutrum, tellus est consequat justo, sit amet rutrum sem diam vel ligula. Nulla condimentum, sapien quis hendrerit aliquam, sapien nunc consequat erat, sit amet ultrices est dolor at libero. Duis imperdiet congue elit, at sodales arcu sodales ut. Cras lobortis nisl vel nibh blandit non consectetur tellus suscipit. Proin in massa turpis. Vivamus vulputate, dui eget sagittis dignissim, lacus felis egestas tortor, a dignissim libero diam non lectus. Nam lacinia eros a magna bibendum vehicula. Sed nec nulla quis odio vulputate laoreet. Nulla facilisi.

Phasellus ultrices enim nec velit vehicula malesuada. Ut non orci ipsum. Mauris est nulla, pretium non pellentesque id, tristique vel lorem. Nulla dictum scelerisque ipsum ac sodales. Vestibulum a est vel est vehicula interdum suscipit et augue. In ligula tellus, laoreet id ultrices ac, pretium at ante. Proin augue lorem, blandit et consectetur eu, ultrices nec mi. Maecenas vitae ipsum est, vel venenatis enim. Fusce porttitor adipiscing eros, ac condimentum libero facilisis nec. Fusce commodo viverra odio pretium iaculis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc aliquam, augue ut dignissim scelerisque, mauris tellus lobortis urna, id condimentum nisi felis et velit. Cras tempus dui at lectus fringilla accumsan. Nunc eleifend felis eu sapien ultricies eu laoreet magna porta. Quisque tincidunt nunc in odio vestibulum tristique eget sed lacus. Suspendisse vitae gravida libero.

Donec et nulla risus. Integer nunc ipsum, euismod quis mollis id, pretium quis urna. Aenean ut nibh magna. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Etiam bibendum, dolor eget faucibus sagittis, augue nisi rhoncus diam, at mattis lectus nulla non eros. Donec rutrum, justo sit amet mollis ornare, nulla nulla vehicula velit, sit amet suscipit magna nunc at erat. Maecenas eros quam, adipiscing consectetur imperdiet non, commodo non felis. Duis nec metus mauris, in congue tortor. Duis nibh lectus, pharetra vulputate porttitor id, consectetur vitae nisl. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.

Pellentesque molestie, nisl in accumsan suscipit, neque tortor cursus elit, feugiat sodales tellus mi ac justo. Pellentesque ornare commodo lectus eget laoreet. Ut bibendum, nisl eu pretium ornare, erat diam malesuada dolor, in pellentesque est nibh quis lacus. Morbi bibendum, enim rutrum eleifend pellentesque, risus nisl ultricies justo, ut dignissim magna magna ac libero. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Quisque consequat pulvinar ipsum id congue. Duis porta magna vitae leo rutrum eget commodo ipsum venenatis. Quisque a augue tortor, ut pharetra lacus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Pellentesque id neque id odio pretium cursus quis nec erat.

\chapter*{Epilogue / Conclusion}
\addcontentsline{toc}{chapter}{Epilogue / Conclusion}

%%% Seznam použité literatury
\newpage
\nocite{*}
\bibliographystyle{alpha}
\bibliography{literature}

%%% Tabulky v diplomové práci, existují-li.
\chapwithtoc{List of Tables}
\listoftables

%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
\chapwithtoc{List of Abbreviations}

%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
%%% textu práce. Přílohy se číslují.
\chapwithtoc{Attachments}
\openright
\end{document}

### 2011-06-07 doc/theses/svirec/kap3.tex
First commit of my thesis

\chapter{Analysis of Recent Approaches}

This chapter introduce description and categorization of recent approaches of repairing XML documents with uses of functional dependencies.

The problem of finding optimum repair was primary discussed

\section{este neviem}

\subsection{Repairs and Consistent Answers for XML Data with Functional Dependencies}

A technique for computing repairs which solves the problem of XML data inconsistency with respect to a set of functional dependencies was proposed in \cite{RepAndConsistentAnswer}. In this approach, authors are trying to find minimal set of update operations which makes XML data consistent. This update operations can be divided into two categories i) replacing value associated with element or attribute, and ii) mark particular node information as unreliable.

\subsubsection{XML Tree and Functional Dependency}

To be able to resolve problem of functional dependency violations in XML document, authors tries to introduce concept of functional dependendenies based on those defined for relational databases. In a relational database $D$ a functional dependency $A \rightarrow B$ model correspondence between values $A$ and $B$ in tuples of $D$. Because the standard concent of a tuple is not defined for XML document, authors introduce concept of tree tuple, which corresponds to the concept of tuple in relational database:

\begin{define}[Tree Tuple]
Given an XML tree XT conforming the DTD D, a tree tuple t of XT is a maximal sub-tree of XT such that, for every path $p \in paths(D)$, t.p contains at most one element.\qed
\end{define}

\begin{define}[Functional Dependency]
Given a DTD D, a functional dependency on D is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and p is an element of $paths(D)$.\qed
\end{define}

Given an XML tree $XT$ conforming a DTD $D$ and a functional dependency $F : S_1 \rightarrow S_2$ , we say that $XT$ satisfies $F (XT \models F )$ if for each pair of tree tuples $t_1, t_2$ of $XT$, $t_1.S_1 = t_2.S_1 \land t_1.S_1 = \emptyset \Rightarrow t_1.S_2 = t_2.S_2$ . Given a set of functional dependencies $\mathcal{FD} = \{F_1 , \dots, F_n\}$ over $D$, we say that $XT$ satisfies $\cal FD$ if it satisfies $F_i$ for every $i \in 1..n$.

\subsubsection{Repairing inconsistent XML data}

Authors of this approach chose two kinds of actions to repair inconsistent XML data with regard to functional dependencies. First action is updating value of an attribute or content of an element. As a second action authors choose marking inconsistent element as "unreliable" rather than deleting it, because removing elements from a XML document leads to some undesired drawbacks: it does not always suffice to remove inconsistency and deleting a node can lead to a new document not conforming the given schema.

Depending on XML data and defined functional dependencies, each inconsistency could have many possible strategies to repair it. From all the possible repair strategies, authors prefer those, for which smaller changes are made to the original document.

\begin{example}
Consider XML tree $XT$ of Figure conforming the DTD $D$ defined below, which is representing collection of books, and given the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.
\begin{verbatim}
<!ELEMENT bib (book+)>
<!ELEMENT book (written_by, title,
                pub, year?)>
<!ELEMENT written_by (author+)>
<!ELEMENT author (name)>
<!ATTLIST author ano CDATA>
<!ELEMENT name PCDATA>
<!ELEMENT title PCDATA>
<!ELEMENT pub PCDATA>
<!ELEMENT year PCDATA>
\end{verbatim}
The functional dependency defined above requires that for each book, there is only one element author having a given $@ano$ value, therefore $XT$ does not satisfy given functional dependency, because two author elements of the same book have the same value of attribute $@ano$. To resolve this data inconsistency, we can use two different repair strategies: 1) changing one of the value of attribute $@ano$; 2) marking one of the element author as unreliable. Because the first strategy changes only attribute $@ano$, it is preferred before second strategy, which changes larger portion of document, since it marks whole author element as unreliable.
\qed
\end{example}

\subsubsection{Repair Algorithm}

Before introducing the algorithm to repair inconsistent XML document, lets define reliability of elements in XML tree:

\begin{define}[R-XML Tree]
A R-XML $tree$ is a triplet $RXT = \langle T, \delta, \varrho \rangle$, where $\langle T, \delta \rangle$ is an XML tree and $\varrho$ is a reliability function from $N_T$ to \texttt{\{true, false\}}, such that, for each pair of nodes $n_1 , n_2 \in N_T$ with $n_2$ descendent of $n_1$, it holds that $\varrho(n_1) = false \Rightarrow \varrho(n_2) = false$.\qed
\end{define}

R-XML Tree is used as an input for a main algorithm computing repaired R-XML tree described in Algorithm \ref{repAlgo}. First the algorithm compute all the possible repairs of tuples which doesn't satisfy functional dependecy using the function \texttt{computeRepairs} (lines 2-6). Next all non-minimal repairs are removed from all possible repairs (line 7). In a last step, all the repairs are merged and a unique repaired R-XML tree is returned.

\begin{algorithm}
\caption{XML Repair}
\label{repAlgo}
\begin{algorithmic}[1]
\REQUIRE{\ \\
$RXT = \langle T, \delta, \varrho \rangle$: R-XML tree conforming a DTD $D$\\
$\mathcal{FD} = {F_1, \dots, F_m}$: Set of functional dependencies}
\ENSURE a unique repaired R-XML tree

\STATE $S = \emptyset$ \COMMENT Set of repairs
\FORALL{$(F: S \rightarrow p) \in \mathcal{FD}$ s.t. $RXT \not \models_w F$}
	\FORALL{$t_1, t_2$ tuples of $RXT$ s.t. $t_1, t_2$ do not weakly satisfy $F$}
		\STATE $S = S \cup computeRepairs(F, t_1, t_2, RXT)$
	\ENDFOR
\ENDFOR
\STATE $S = removeNonMinimal(S, RXT)$
\STATE $\langle \delta', \varrho' \rangle = mergeRepairs(S)$
\RETURN $\langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$
\end{algorithmic}
\end{algorithm}

The function \texttt{computeRepairs} gets R-XML tree, functional dependency $F$ and a tuples $t_1, t_2$ of R-XML tree as input and computes repair as follows:
\begin{itemize}
	\item If path $p$ denotes a textual element, one of the two terminal values of $t_1.p$ or $t_2.p$ is changed, so that they become equal (line 3).
	\item Otherwise $p$ denotes a node, so either the node $t_1.p$ or $t_2.p$ is marked as unreliable (line 5).
	\item For each path $p_i$ of the left side of a functional dependency $F$
	\begin{itemize}
		\item If path $p_i$ denotes a textual element, then one of the two terminal values $t_1.p_i$ or $t_2.p_i$ is changed to $\perp$ (line 9).
		\item Otherwise $p_i$ denotes a node, so one of the nodes $t_1.p_i$ or $t_2.p_i$ is marked as unreliable (line 11).
	\end{itemize}
\end{itemize}

\begin{algorithm}[H]
\floatname{algorithm}{Function}
\caption{$computeRepairs(F, t_1, t_2, RXT)$}
\begin{algorithmic}[1]
\REQUIRE{\ \\
$RXT = \langle T, \delta \varrho \rangle$: R-XML tree conforming a DTD $D$\\
$F: X \rightarrow p$ functional dependency\\
$t_1, t_2$ tuples of $RXT$}
\ENSURE $S$: Set of repairs

\STATE $S = \emptyset$
\IF{$p \in StrPaths(D)$}
	\STATE $S = S \cup \{\langle \{\delta(p(t_1)) = t_2.p\}, \varrho \rangle\} \cup \{\langle \{\delta(p(t_2)) = t_1.p\}, \varrho \rangle\} $
\ELSE
	\STATE $S = S \cup \{\langle \emptyset, \varrho_{\{t_1.p\}} \cdot \varrho \rangle\} \cup \{\langle \emptyset, \varrho_{\{t_2.p\}} \cdot \varrho \rangle\}$
\ENDIF
\FORALL{$p_i \in X$}
	\IF{$p_i \in StrPaths(D)$}
		\STATE $S = S \cup \{\langle \{\delta(p_i(t_1)) = \perp_1\}, \varrho \rangle\} \cup \{\langle \{\delta(p_i(t_2)) = \perp_2\}, \varrho \rangle\}$
	\ELSE
		\STATE $S = S \cup \{\langle \emptyset, \varrho_{\{t_1.p_i\}} \cdot \varrho \rangle\} \cup \{\langle \emptyset, \varrho_{\{t_2.p_i\}} \cdot \varrho \rangle\}$
	\ENDIF
\ENDFOR
\RETURN $S$
\end{algorithmic}
\end{algorithm}

\subsubsection{Conclusion}

\todo[inline]{Sem pojde zhrnutie tohto approachu}

\subsection{Querying and Repairing Inconsistent XML Data}

Studying the problem of repairing inconsistent XML document with respect to a set of functional dependencies and investigating the existence of repairs has been introduced in \cite{QueryXML}. Authors introduce two kind of repair primitives, first is deleting "unreliable" portions of document and second is inserting new part. As it is in another approaches, authors prefer minimal set of repair primitives applied to the XML document to form a repair.

Introduced repair primitives authors use in three different repair strategies consisting of:
\begin{enumerate}
	\item \textit{(general) repairs}, where both delete and insert operations are used,
	\item \textit{cleaning repairs}, for documents interpreted as "dirty", only delete operations are used to repair inconsistencies,
	\item \textit{completing repairs}, for documents interpreted as incomplete, insert operations are used.
\end{enumerate}

\subsubsection{General Repair}

With the insert and delete operations as a repair primitives, the structure of the XML document will be updated. The insert operation is represented as $\langle  + [x]a[y], z\rangle$, where i) $x$ is a node identifier, ii) $a$ is a label, iii) $y$ is either a node identifier or a value ($y$ is a value if $a \in Att \cup \{S\}$; otherwise is a node identifier), and iv) $z$ denotes the child of $x$ which must immediately precede $y$ ($\perp$ if $y$ is inserted as first or unique child of $x$). The deletion is represented as $-[x]a[y]$, where $x$ is a node identifier, $a \in Att \cup El \cup \{S\}$, and $y$ is either a node $id$ or a string value, denoting the node to be deleted.\\
With defined basic repair operations for general repair of inconsistent XML document, lets define repair as follows:

\begin{define}[Repair]
Given a XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$, a set of update operations $R$ is said to be a repair of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
\end{define}

With further investigation, authors discover that the problem of deciding whether there exists a repair for a XML document in the presence of DTD with functional dependencies is undecidable. Therefore they consider restricted forms of repairs, more specifically cleaning and completing repairs.

\subsubsection{Conclusion}

\todo[inline]{Sem pojde zhrnutie tohto approachu}

\subsection{Improving XML Data Quality with Functional Dependencies}

The algorithm for repairing XML functional dependency violations which uses modification of node value as the repair primitive has been proposed in \cite{ImprovingXML}. For finding the optimal repair of XML document, authors introduce a cost model, which assign a weight to each leaf node in XML document. The optimal repair is that with lowest repair cost, which is measured by a total weight of the modified nodes.

Authors of this approach found out that by repairing one functional dependency violation can break another, therefore they separate the main algorithm into two phases. In the first phase, the conflict hypergraph capturing the initial functional dependency violations is constructed, and all the violations are fixed by modifying the values of all the nodes on a vertex cover of the conflict hypergraph. In the second phase, remaining violations are resolved by modifying the violating nodes and their core determinants to prevent of introducing new conflicts.

\subsubsection{Cost Model and Repairing Primitive}

For each leaf node $v$ of XML document is associated a weight in the range $[0,1]$, which is denoted $W(v)$. Lets assume that larger the weight of the leaf is, then the leaf is more reliable. The weight may be automatically generated by some statistical methods or can be assigned by a user.

\begin{define}[Functional Dependency]
With a given DTD $D$, a functional dependency is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$. Here $P$ is a  root path, or $P = \epsilon$. Each $P_i (i \in [1,n])$ is a singleton leaf path, and there is a no non-empty common prefix for $P_1, \dots, P_{n+1}$. Given an XML document $T$ conforming to $D$, we say $T$ satisfies $\sigma$:iff $\forall v \in \{[\![P]\!]\}, \forall v_1, v_2 \in \{v[\![P']\!]\}$, if $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, then $v_1[\![P_{n+1}]\!] \equiv v_2[\![p_{n+1}]\!]$.\qed
\end{define}

As was mentioned earlier, repairing primitive is node value modification, where for repairing algorithm is used combination of two rules to resolve violations. Lets have a $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$. If the child nodes of $v_1$ and $v_2$ qualified by paths $P_i$ have equal values of all $i \in [1,n]$ and their child nodes qualified by $P_{n+1}$ have different values, then $v_1$ and $v_2$ violates $\sigma$. The first rule used to repair this violation is to change the value of the node qualified by $P_{n+1}$ from $v_1$ to the value of $v_2$'s child node that matches $P_{n+1}$(or reversely). The second rule is to choose an arbitrary $P_i (i \in [1,n])$, and introduce a fresh new value to the node qualified by $P_i$ from $v_1$ (or $v_2$).

\begin{define}[Optimum repair]
Given an inconsistent XML document $T$ violating a set $\Sigma$ of $FD$s, the repair $T_R$ of $T$ is called optimum repair, if $T_R$ has the minimum cost among all repairs of $T$. The cost $cost(T_R)$ is defined as:
\begin{displaymath}
cost(T_R) = \sum_{v \in T} w(v) \times dist(v, v_R),
\end{displaymath}
where $dist(v, v_R)=1$ if $val(v) \neq val(v_R)$, otherwise $dist(v, v_R)=0$.\qed
\end{define}

\subsubsection{Initial Conflicts Hypergraph}

Weighted Hypergraph is used by the authors in first part of repair algorithm as a tool modeling initial functional dependency violations in XML document. Hypergraph $g$ of XML document $T$ can be defined as a pair $g = (V,E)$, where $V$ stands for a set of elements (called nodes), and $E$ is a set of non-empty subset of $V$ called hyperedges, more accurately each hyperedge indicates a set of value nodes violating FDs. Since hypergraph is weighted and cost model is used in this approach, for each node $v \in V$ of hypergraph is assigned a weight $w(v)$, which is the same as the weight of $v$ in $T$.\\

To actually resolve the problem of repairing FD violations in XML document, authors convert this problem into well-known problem of weighted vertex cover for hypergraph \cite{ApproxAlgo}. Lets have hypergraph $g = (V,E)$ where each hyperedge $e \in E$ is set of value nodes which violates some FD. In a repair of inconsistent XML document, for each hyperedge at least one value node is modified, therefore it is essential to find a vertex cover (VC) for $g$, which is a set $S \subseteq V$, such that for all edges $e \in E$, $S \cap e \neq \emptyset$. Because hypergraph is weighted, we can define weight of VC as a total weight of all vertices in $S$.\\

Algorithm fixing initial FD violations is shown in Algorithm \ref{fixInit}. The algorithm uses approximation algorithm to find VC for the minimum weighted vertex cover proposed in \cite{ApproxAlgo}.

\begin{algorithm}
\caption{Fix-Initial-Conflicts}
\label{fixInit}
\begin{algorithmic}[1]
\REQUIRE An XML document T, a set $\Sigma$ of FDs.
\ENSURE A modified document T.
\STATE Create the initial conflict hypergraph $g$ of $T$ w.r.t $\Sigma$
\STATE Use a known algorithm to find an approxiamtion $VC$ for the minimum weighted vertex cover of $g$
\STATE $remaining$ := VC
\WHILE{There are two target nodes $v_1, v_2 \in T$ violating a FD $\sigma \in \Sigma$, and $v_1[\![P_{n+1}]\!]$ or $v_2[\![P_{n+1}]\!]$ is the only node in VC from the set of nodes $\{v_1[\![P_i]\!](i \in [1, n+1])\} \cup \{v_2[\![P_i]\!](i \in [1, n+1])\}$. (W.l.o.g assume the violation is as follows: $\sigma = (P,P',(P_1,\dots,P_n \rightarrow P_{n+1}))$, $v \in \{[\![P]\!]\}$, $v_1, v_2 \in \{v[\![P']\!]\}$, $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] \not \equiv v_2[\![P_{n+1}]\!]$.)}
\STATE $val(v_1[\![P_{n+1}]\!])$ := $val(v_2[\![P_{n+1}]\!])$ \COMMENT W.l.o.g, we assume $v_1[\![P_{n+1}]\!]$ is in VC
\ENDWHILE
\FORALL{node $u \in remaining$}
\STATE $val(u)$ := $gen\_new\_value()$
\STATE \COMMENT Introduce fresh new values to all the remaining nodes in VC
\ENDFOR
\end{algorithmic}
\end{algorithm}

\subsubsection{Resolving Violations Thoroughly}

After repairing initial FD violations, there is a chance that new violations may be introduced, therefore authors provided a method to do modifications on value nodes without incurring new conflicts (Algorithm \ref{fixRest}). This method uses core determinant $C_u$ of value node $u$ defined as follows:

\begin{define}[Core Determinant]
Given an XML document $T$, a set $\Sigma$ of FDs and a node $u$ in $T$, we say a set of nodes $\{u_1, u_2, \dots, u_n\}$ is a $\sigma\!-\!determinant$ of $u$, if there exists a nontrivial FD $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ logical implied by $\Sigma$, such that $\exists v \in \{[\![P]\!]\}, \exists v_1 \in \{[\![P']\!]\}, v_1[\![P]\!] = u_i$ for $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] = u$.\\
We say that a set $C_u$ of nodes is a core determinant of $u$, if (a) for every nontrivial FD $\sigma$ implied by $\Sigma$ and every set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u \cap W \neq \emptyset$; and (b) for any propert subset $C_u'$ of $C_u$, there exists some nontrivial FD $\sigma$ implied by $\Sigma$, and a set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u' \cap W = \emptyset$.\qed
\end{define}

\begin{algorithm}[H]
\caption{Resolve-Remaining-Violations}
\label{fixRest}
\begin{algorithmic}[1]
\REQUIRE An XML document T, a set $\Sigma$ of FDs.
\ENSURE A modified document T, with all the violations fixed.

\WHILE{there are FD violations in T w.r.t. $\Sigma$}
\STATE pick a violating value node $u$ from T w.r.t. $\Sigma$
\STATE let $C_u$ be a core determinant $u$
\FORALL{node $w \in (C_u \cup \{u\})$}
\STATE $val(w)$ := $gen\_new\_value()$
\STATE \COMMENT it guarantees that no new violations will be introduced
\ENDFOR
\ENDWHILE
\end{algorithmic}
\end{algorithm}

\subsubsection{Conclusion}

\todo[inline]{Sem pojde zhrnutie tohto approachu}

### 2011-06-07 doc/theses/svirec/prace.tex
First commit of my thesis

%%% Hlavní soubor. Zde se definují základní parametry a odkazuje se na ostatní části. %%%

%% Verze pro jednostranný tisk:
% Okraje: levý 40mm, pravý 25mm, horní a dolní 25mm
% (ale pozor, LaTeX si sám přidává 1in)
\documentclass[12pt,a4paper]{report}
\setlength\textwidth{145mm}
\setlength\textheight{247mm}
\setlength\oddsidemargin{15mm}
\setlength\evensidemargin{15mm}
\setlength\topmargin{0mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
% \openright zařídí, aby následující text začínal na pravé straně knihy
\let\openright=\clearpage

%% Pokud tiskneme oboustranně:
% \documentclass[12pt,a4paper,twoside,openright]{report}
% \setlength\textwidth{145mm}
% \setlength\textheight{247mm}
% \setlength\oddsidemargin{15mm}
% \setlength\evensidemargin{0mm}
% \setlength\topmargin{0mm}
% \setlength\headsep{0mm}
% \setlength\headheight{0mm}
% \let\openright=\cleardoublepage

%% Pokud pouľíváte csLaTeX (doporučeno):
%\usepackage{czech}
%% Pokud nikoliv:
%\usepackage[czech]{babel}
%\usepackage[T1]{fontenc}

%% Pouľité kódování znaků: obvykle latin2, cp1250 nebo utf8:
\usepackage[utf8]{inputenc}

%% Ostatní balíčky
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{todonotes}
\usepackage{algorithmic}
\usepackage[chapter]{algorithm}

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

\renewcommand{\algorithmicforall}{\textbf{for each}}

\renewcommand{\algorithmiccomment}[1]{// #1}

% pekne pokope definujeme potrebne udaje
\def\mftitle{Efficient Detection of XML Integrity Constraints}
\def\mfthesistype{Master Thesis}
\def\mfauthor{Michal Švirec}
\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D.}
\def\mfplacedate{Praha, 2010}

%% Balíček hyperref, kterým jdou vyrábět klikací odkazy v PDF,
%% ale hlavně ho pouľíváme k uloľení metadat do PDF (včetně obsahu).
%% POZOR, nezapomeňte vyplnit jméno práce a autora.
\usepackage[ps2pdf,unicode]{hyperref}   % Musí být za vąemi ostatními balíčky
\hypersetup{pdftitle=\mftitle}
\hypersetup{pdfauthor=\mfauthor}

%%% Drobné úpravy stylu

% Tato makra přesvědčují mírně oąklivým trikem LaTeX, aby hlavičky kapitol
% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
%\makeatletter
%\def\@makechapterhead#1{
%  {\parindent \z@ \raggedright \normalfont
%   \Huge\bfseries \thechapter. #1
%   \par\nobreak
%   \vskip 20\p@
%}}
%\def\@makeschapterhead#1{
%  {\parindent \z@ \raggedright \normalfont
%   \Huge\bfseries #1
%   \par\nobreak
%   \vskip 20\p@
%}}
%\makeatother

% Toto makro definuje kapitolu, která není očíslovaná, ale je uvedena v obsahu.
\def\chapwithtoc#1{
\chapter*{#1}
\addcontentsline{toc}{chapter}{#1}
}

\begin{document}

% Trochu volnějąí nastavení dělení slov, neľ je default.
\lefthyphenmin=2
\righthyphenmin=2

%%% Titulní strana práce

\pagestyle{empty}
\begin{center}

\large

Charles University in Prague

\medskip

Faculty of Mathematics and Physics

\vfill

{\bf\Large \mfthesistype}

\vfill

\centerline{\mbox{\includegraphics[width=60mm]{logo.eps}}}

\vfill
\vspace{5mm}

{\LARGE \mfauthor}

\vspace{15mm}

% Název práce přesně podle zadání
{\LARGE\bfseries \mftitle}

\vfill

% Název katedry nebo ústavu, kde byla práce oficiálně zadána
% (dle Organizační struktury MFF UK)
Název katedry nebo ústavu

\vfill

\begin{tabular}{rl}

Supervisor of the master thesis: & \mfadvisor \\
\noalign{\vspace{2mm}}
Study programme: & program \\
\noalign{\vspace{2mm}}
Specialization: & obor \\
\end{tabular}

\vfill

% Zde doplňte rok
\mfplacedate

\end{center}

\newpage

%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.

%%% Na tomto místě mohou být napsána případná poděkování (vedoucímu práce,
%%% konzultantovi, tomu, kdo zapůjčil software, literaturu apod.)

\openright

\noindent
Podakovanie.

\newpage

%%% Strana s čestným prohláąením k diplomové práci

\vglue 0pt plus 1fill

\noindent
I declare that I carried out this master thesis independently, and only with the cited
sources, literature and other professional sources.

\medskip\noindent
I understand that my work relates to the rights and obligations under the Act No.
121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
University in Prague has the right to conclude a license agreement on the use of this
work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.

\vspace{10mm}

\hbox{\hbox to 0.5\hsize{%
In ........ date ............
\hss}\hbox to 0.5\hsize{%
Signature
\hss}}

\vspace{20mm}
\newpage

%%% Povinná informační strana diplomové práce

\vbox to 0.5\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Název práce:
\mftitle
% přesně dle zadání

Autor:
\mfauthor

Katedra:  % Případně Ústav:
Název katedry či ústavu, kde byla práce oficiálně zadána
% dle Organizační struktury MFF UK

Vedoucí diplomové práce:
\mfadvisor, pracoviště
% dle Organizační struktury MFF UK, případně plný název pracoviątě mimo MFF UK

Abstrakt:
% abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce

Klíčová slova:
% 3 aľ 5 klíčových slov

\vss}\nobreak\vbox to 0.49\vsize{
\setlength\parindent{0mm}
\setlength\parskip{5mm}

Title:
% přesný překlad názvu práce v angličtině

Author:
\mfauthor

Department:
Název katedry či ústavu, kde byla práce oficiálně zadána
% dle Organizační struktury MFF UK v angličtině

Supervisor:
\mfadvisor, pracoviště
% dle Organizační struktury MFF UK, případně plný název pracoviątě
% mimo MFF UK v angličtině

Abstract:
% abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
% zadání diplomové práce

Keywords:
% 3 aľ 5 klíčových slov v angličtině

\vss}

\newpage

%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
%%% na začátku práce, místo na jejím konci.

\openright
\pagestyle{plain}
\setcounter{page}{1}
\tableofcontents\thispagestyle{empty}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem*{define}{Definition}	% Definice nečíslujeme, proto "*"
\newtheorem{example}{Example}

%%% Jednotlivé kapitoly práce jsou pro přehlednost uloľeny v samostatných souborech
\include{uvod}
\include{kap1}
\include{kap2}
\include{kap3}

% Ukázka pouľití některých konstrukcí LateXu (odkomentujte, chcete-li)
%\include{ukazka}

\include{zaver}

%%% Seznam pouľité literatury
\nocite{*}
\bibliographystyle{plain}
\bibliography{literatura}

%%% Tabulky v diplomové práci, existují-li.
\chapwithtoc{List of Tables}

%%% Pouľité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
\chapwithtoc{List of Abbreviations}

%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
%%% diagramy apod.). Kaľdá příloha musí být alespoň jednou odkazována z vlastního
%%% textu práce. Přílohy se číslují.
\chapwithtoc{Attachments}

\openright
\end{document}

### 2011-06-07 doc/theses/svirec/kap2.tex
First commit of my thesis

\chapter{Název druhé kapitoly}

\section{Název první podkapitoly v druhé kapitole}

\subsection{Preliminaries}

\subsubsection{XML Trees and DTDs}
\begin{define}[Tree]
{\sl Tree} $T$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is an (acyclic) set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$.
\end{define}

\noindent Let also denote the set of leaf nodes as $Leaves(T)$ and the set of trees defined on an alphabet on node labels $\Sigma$ as $T_\Sigma$.

\begin{define}[XML Tree]
{\sl XML tree} is a pair $XT=\langle T,\delta \rangle$, where:
	\begin{enumerate}
		\item $T = (r, N, E, \lambda)$ is a tree in $T_{\tau\cup\alpha\cup\{S\}}$, where $\tau$ is a tag alphabet, $\alpha$ is an attribute name alphabet and $S$ is a symbol not belonging to $\tau\cup\alpha$ (representing \texttt{\#PCDATA} content of elements)$;$
		\item given a node $n$ of $T$, $\lambda(n) \in \alpha \cup \{S\} \Leftrightarrow n \in Leaves(T);$
		\item $\delta : Leaves(T) \rightarrow Str$, where $Str$ is a string alphabet, is a function associating a (string) value to every leaf of $T$.
	\end{enumerate}\qed
\end{define}

\begin{define}[DTD]
{\sl DTD} is a tuple $D = (\tau, \alpha, P, R, rt)$ where: i) $\tau$ and $\alpha$ are of the same definition as in the \emph{XML tree}; ii) $P$ is the set of \emph{element type definitions}; iii) R is the set of \emph{attribute lists}; iv) $rt \in \tau$ is the tag of the document root element.\qed
\end{define}

\begin{define}[Path]
{\sl Path} p on a DTD $D = (\tau, \alpha, P, R, rt)$ is a sequence $p = s_1, \dots, s_m$ of symbols in $\tau \cup \alpha \cup \{S\}$ such that:
	\begin{enumerate}
		\item $s_1=rt$;
		\item for each $i$ in $2..m-1$, $s_i \in \tau$ and $s_i$ appears in the element type definition of $s_{i-1}$;
		\item $s_m \in \alpha \rightarrow s_m$ appears in the attribute list of $s_{m-1}$;
		\item $s_m \in \tau \cup \{S\} \rightarrow s_m$ appears the element type definition on $s_{m-1}$.
	\end{enumerate}\qed
\end{define}

As we defined \emph{path} on a DTD $D$, let denote as $paths(D)$ the set of paths which can be defined on a DTD $D$. Also another important notion is $p(XT)$, which is the set of nodes from XML tree $XT$ conforming DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$. Moreover, let denote $XT.p$ the \emph{answer} of the path $p$ applied on $XT$ , that is:
\begin{itemize}
	\item[-] if $p \in EPath(D)$, where $EPath(D)$ denotes the set of the paths whose last symbol denotes an element, then $XT.p = p(XT)$
	\item[-] if $p \in StrPath(D)$, where $StrPath(D)$ denotes the set of paths whose last symbol denotes either the textual content of an element or an
attribute, then $XT.p = \{\delta_T(x)|x \in p(XT)\}$.
\end{itemize}

\subsection{XML and Functional Dependencies}

In relational database $D$, correspondence between values $A$ and $B$ in the tuple of $D$, models the functional dependency denoted as $A \rightarrow B$. Because in XML is not such a standard tuple concept, authors introduce the concept of tree tuples, corresponding to relational databases concept of tuples.

\begin{define}[Tree Tuple]
Given an XML tree XT conforming the DTD D, a tree tuple t of XT is a maximal sub-tree of XT such that, for every path $p \in paths(D)$, t.p contains at most one element.\qed
\end{define}

\begin{define}[Functional Dependency]
Given a DTD D, a functional dependency on D is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and p is an element of $paths(D)$.\qed
\end{define}

Given an XML tree $XT$ conforming a DTD $D$ and a functional dependency $F : S_1 \rightarrow S_2$ , we say that $XT$ satisfies $F (XT \models F )$ if for each pair of tree tuples $t_1, t_2$ of $XT$, $t_1.S_1 = t_2.S_1 \land t_1.S_1 = \emptyset \Rightarrow t_1.S_2 = t_2.S_2$ . Given a set of functional dependencies $\mathcal{FD} = \{F_1 , \dots, F_n\}$ over $D$, we say that $XT$ satisfies $\cal FD$ if it satisfies $F_i$ for every $i \in 1..n$.

### 2011-06-30 doc/theses/vitasek/diplomka.tex
Thesis: initial commit of my folder.
Thesis: adding some structure.

--- 2011-05-09 doc/theses/template/diplomka.tex
@@
 
-\def\mfauthor{Michal Klempa}
+\def\mfauthor{Matej Vitásek}
 \def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
@@
 % exactly as assigned
-{\LARGE\bfseries Title of the thesis}
+{\LARGE\bfseries Inference of XML Integrity Constraints}
 
@@
 the person who lent the software, literature, etc.)]
+
+TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
 
@@
 \newpage
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec tortor neque, porta a posuere quis, tristique nec diam.
-Morbi imperdiet, turpis eu pharetra volutpat, felis magna elementum ligula, ut convallis justo orci sed orci.
-Sed congue vulputate auctor.
-Morbi laoreet adipiscing fringilla.
-Aliquam sit amet ante mi.
-Maecenas vehicula felis nec nibh aliquam at convallis erat dapibus.
-Nam nec mollis sapien.
-Sed scelerisque bibendum suscipit.
-Maecenas ornare est non sapien elementum bibendum a ac diam.
-Proin nibh arcu, elementum eu lobortis vel, suscipit a augue.
-Sed adipiscing, velit vel mollis cursus, tellus neque tincidunt augue, at pretium mauris mauris vel felis.
-Maecenas consequat tristique pulvinar.
-Donec porta feugiat neque, ac semper est bibendum vehicula.
-Vivamus gravida mauris in neque tempus tincidunt.
-Nulla blandit, nisi in ullamcorper aliquet, elit tortor viverra velit, id placerat nulla leo eu enim.
-\chapter*{Introduction / Preface}
-\addcontentsline{toc}{chapter}{Introduction / Preface}
 
-Aenean in odio eu nisi imperdiet condimentum.
-Suspendisse id ipsum sed risus pharetra vehicula bibendum sed neque.
-Nulla id elementum augue.
-Integer vehicula interdum orci, at gravida ante interdum id.
-Nunc laoreet orci non nisi cursus vel tincidunt orci venenatis.
-Quisque blandit odio ac ipsum lacinia a gravida elit mattis.
-Praesent sagittis, arcu ac accumsan rutrum, tellus est consequat justo, sit amet rutrum sem diam vel ligula.
-Nulla condimentum, sapien quis hendrerit aliquam, sapien nunc consequat erat, sit amet ultrices est dolor at libero.
-Duis imperdiet congue elit, at sodales arcu sodales ut.
-Cras lobortis nisl vel nibh blandit non consectetur tellus suscipit.
-Proin in massa turpis.
-Vivamus vulputate, dui eget sagittis dignissim, lacus felis egestas tortor, a dignissim libero diam non lectus.
-Nam lacinia eros a magna bibendum vehicula.
-Sed nec nulla quis odio vulputate laoreet.
-Nulla facilisi.
+\chapter*{Preface}
+\addcontentsline{toc}{chapter}{Preface}
 
-Phasellus ultrices enim nec velit vehicula malesuada.
-Ut non orci ipsum.
-Mauris est nulla, pretium non pellentesque id, tristique vel lorem.
-Nulla dictum scelerisque ipsum ac sodales.
-Vestibulum a est vel est vehicula interdum suscipit et augue.
-In ligula tellus, laoreet id ultrices ac, pretium at ante.
-Proin augue lorem, blandit et consectetur eu, ultrices nec mi.
-Maecenas vitae ipsum est, vel venenatis enim.
-Fusce porttitor adipiscing eros, ac condimentum libero facilisis nec.
-Fusce commodo viverra odio pretium iaculis.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nunc aliquam, augue ut dignissim scelerisque, mauris tellus lobortis urna, id condimentum nisi felis et velit.
-Cras tempus dui at lectus fringilla accumsan.
-Nunc eleifend felis eu sapien ultricies eu laoreet magna porta.
-Quisque tincidunt nunc in odio vestibulum tristique eget sed lacus.
-Suspendisse vitae gravida libero.
+TODO why is this important
 
-Donec et nulla risus.
-Integer nunc ipsum, euismod quis mollis id, pretium quis urna.
-Aenean ut nibh magna.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Etiam bibendum, dolor eget faucibus sagittis, augue nisi rhoncus diam, at mattis lectus nulla non eros.
-Donec rutrum, justo sit amet mollis ornare, nulla nulla vehicula velit, sit amet suscipit magna nunc at erat.
-Maecenas eros quam, adipiscing consectetur imperdiet non, commodo non felis.
-Duis nec metus mauris, in congue tortor.
-Duis nibh lectus, pharetra vulputate porttitor id, consectetur vitae nisl.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+\chapter{Definitions}
 
-Pellentesque molestie, nisl in accumsan suscipit, neque tortor cursus elit, feugiat sodales tellus mi ac justo.
-Pellentesque ornare commodo lectus eget laoreet.
-Ut bibendum, nisl eu pretium ornare, erat diam malesuada dolor, in pellentesque est nibh quis lacus.
-Morbi bibendum, enim rutrum eleifend pellentesque, risus nisl ultricies justo, ut dignissim magna magna ac libero.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-Quisque consequat pulvinar ipsum id congue.
-Duis porta magna vitae leo rutrum eget commodo ipsum venenatis.
-Quisque a augue tortor, ut pharetra lacus.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Pellentesque id neque id odio pretium cursus quis nec erat.
+TODO talk about key definition from KfXML
 
-\chapter*{Epilogue / Conclusion}
-\addcontentsline{toc}{chapter}{Epilogue / Conclusion}
+talk about ID, IDREF, ... talk about XML schema keys
+
+\chapter{Research}
+
+TODO talk about FIDAX
+
+\section{Independent Set}
+
+\chapter{MIP Approach}
+
+TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver.
+
+\section{Finding ID sets with GLPK}
+
+TODO how GLPK works, how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
+
+ -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
+
+\section{Heuristics}
+
+TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
+
+\subsection{Constructions Heuristics}
+
+\subsubsection{FIDAX}
+
+TODO this is a heuristic too! We can start with it!
+
+\subsubsection{Random}
+
+TODO this is trivial, but works - link experiments where it bashes FIDAX
+
+\subsubsection{Fuzzy}
+
+TODO kind of trivial
+
+\subsubsection{Incremental}
+
+TODO trivial, hungry
+
+\subsubsection{Removal}
+
+TODO trivial, "hungry"
+
+\subsubsection{Truncated Branch & Bound}
+
+TODO if we limit GLPK's runtime, we get this
+
+\subsection{Improvement Heuristics}
+
+TODO what they are, that they need a pool sometimes, how do we give them pool, ...
+
+\subsubsection{Remove Worst}
+
+TODO trivial
+
+\subsubsection{Remove Random}
+
+TODO trivial, perturbation
+
+\chapter{Experiments}
+
+\section{Experimental Setup}
+
+\section{Experimental Results}
+
+TODO following experiments + more
+
+Input -> How long the search for optimum lasts
+Time limit -> Quality
+Structure of the data -> Which algorithms perform the best
+Alpha, Beta -> Quality
+Which GLPK -> *
+Do I beat FIDAX with random?
+Does Fuzzy beat Random? (Seems that Random is better and even has smaller variance)
+
+\section{The "Best" Algorithm}
+
+\chapter*{Conclusion}
+\addcontentsline{toc}{chapter}{Conclusion}
 
@@
 %%% textu práce. Přílohy se číslují.
+
+%% TODO Appendix - About jInfer
+
 \chapwithtoc{Attachments}

### 2011-06-30 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 \usepackage{wrapfig}
+\usepackage{appendix}
 
@@
 
-Title:  [the master thesis title in English]
+Title:  Inference of XML Integrity Constraints
 
@@
 
-Keywords: [3-5 keywords in English]
+Keywords: XML, ID attributes, inference
 
@@
 
-\chapter*{Preface}
-\addcontentsline{toc}{chapter}{Preface}
+\chapwithtoc{Preface}
 
@@
 
-TODO talk about key definition from KfXML
+TODO talk about key definition from \cite{keX}
 
@@
 
+According to the article \cite{fidax}, ...
+
 TODO talk about FIDAX
+To the best of our knowledge, there are no other articles dealing with this problem
+
+TODO talk about Fajt - but that's different, that's keys
 
 \section{Independent Set}
+
+TODO other approaches to max IS
 
@@
 TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
+TODO mention things like Taboo search
+
+TODO we will be working with heus striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
 
 \subsection{Constructions Heuristics}
+
+TODO construction heuristics are heus that provide us with at least some solution.
 
@@
 
-\subsubsection{Truncated Branch & Bound}
+\subsubsection{Truncated Branch \& Bound}
 
@@
 
+\subsubsection{Identity}
+
+TODO trivial trivial
+
 \subsubsection{Remove Worst}
@@
 
-\subsubsection{Remove Random}
+\subsubsection{Random Remove}
 
@@
 
+\subsubsection{Hungry}
+
+TODO trivial
+
+\subsubsection{Mutation}
+
+TODO
+
+\subsubsection{Crossover}
+
+TODO
+
+\subsubsection{Local Branching}
+
+TODO
+
 \chapter{Experiments}
+
+\section{Experimental Data}
 
@@
 
-\chapter*{Conclusion}
-\addcontentsline{toc}{chapter}{Conclusion}
+\chapwithtoc{Conclusion}
+
+TODO conclusion
 
@@
 %%% Tabulky v diplomové práci, existují-li.
-\chapwithtoc{List of Tables}
+%\chapwithtoc{List of Tables}
 \listoftables
@@
 
-\chapwithtoc{Attachments}
+\appendix
+
 \openright
+\addcontentsline{toc}{chapter}{Appendices}
+
+\chapter{jInfer}
+
 \end{document}

### 2011-07-06 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/gui/StatisticsPanel.java
#3323201 Experimental framework - refactoring, JavaDoc.

@@ -45,7 +45,8 @@ import static cz.cuni.mff.ksi.jinfer.base.utils.AsynchronousUtils.runAsync;
  *
  * @author vektor
  */
-@SuppressWarnings("PMD.SingularField")
+@SuppressWarnings({"PMD.SingularField", "PMD.UnusedFormalParameter",
+  "PMD.MethodArgumentCouldBeFinal", "PMD.TooManyFields", "PMD.TooManyMethods"})
 public class StatisticsPanel extends JPanel {
 
   private static final long serialVersionUID = 5415245241L;

### 2011-07-06 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkInputGenerator.java

@@ -16,15 +16,11 @@
  */
 package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
 
+import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.LocalBranching;
 import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtils;
 import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
 import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
 import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.io.BufferedReader;
-import java.io.IOException;
-import java.io.InputStream;
-import java.io.InputStreamReader;
-import java.util.ArrayList;
 import java.util.Collections;
 import java.util.List;
 import org.apache.log4j.Logger;
@@ -50,20 +46,40 @@ public final class GlpkInputGenerator {
   private static final String FIX_TO_0 = "s.t. f{index}: x['{mapping}'] = 0;";
 
   /**
-   * TODO vektor Comment!
+   * Generates a GLPK MathProg representation of the ID set optimization for the
+   * provided model, in this variant specifically for Local Branching heuristic.
    *
-   * @param model
-   * @param alpha
-   * @param beta
-   * @param incumbentAMs
-   * @param k
-   * @return
+   * <p>
+   * Please see {@link GlpkInputGenerator#generateGlpkInput(AMModel, List, double, double) }
+   * for general info.
+   * </p>
+   *
+   * <p>
+   * This variant will add one more constraint to ensure the
+   * Hamming distance from the incumbent solution in the form:<br/>
+   * <code>SUM<sub>I</sub> (1 - am<sub>i</sub>) + SUM<sub>J</sub> am<sub>j</sub> &lt;= k</code><br/>
+   * where <code>I</code> is the set of AMs in the incumbent solution,
+   * <code>J</code> is the set of remaining AMs in the model.
+   * </p>
+   *
+   * @param model Model from which the problem formulation should be generated.
+   * @param alpha Weight of the attribute mapping <cite>support</cite> in its total weight.
+   * @param beta Weight of the attribute mapping <cite>coverage</cite> in its total weight.
+   * @param incumbentAMs List of AMs contained in the currently incumbent
+   *   solution.
+   * @param k Maximal Hamming distance from the specified incumbent solution.
+   *
[…]

### 2011-07-06 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/sets/RandomLocalBranching.java
#3323201 Experimental framework - refactoring, JavaDoc.

@@ -29,8 +29,6 @@ import java.util.Arrays;
 import java.util.List;
 
 /**
- * TODO vektor Comment!
- *
  * @author vektor
  */
 public class RandomLocalBranching extends AbstractExperimentSet {

### 2011-07-06 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkRunner.java

@@ -18,12 +18,10 @@ package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
 
 import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
 import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import java.io.BufferedReader;
 import java.io.File;
 import java.io.FileReader;
 import java.io.IOException;
 import java.io.InputStreamReader;
-import java.io.PrintWriter;
 import java.util.Calendar;
 import java.util.Collections;
 import java.util.List;
@@ -41,10 +39,6 @@ public final class GlpkRunner {
   }
 
   private static final Logger LOG = Logger.getLogger(GlpkRunner.class);
-
-  private static final String TMP = System.getProperty("java.io.tmpdir");
-  private static final String INPUT = TMP + "/glpk_input.txt";
-  private static final String OUTPUT = TMP + "/glpk_output.txt";
 
   /**
    * @see GlpkRunner#run(cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel, java.util.List, double, double, int)
@@ -78,48 +72,36 @@ public final class GlpkRunner {
   }
 
   /**
-   * TODO vektor Comment!
+   * Runs GLPK optimalization on the specified input (has to be already in
+   * MathProg format) and returns the resulting file in a string. This can
+   * be directly displayed for verification or parsed for actual AMs appearing
+   * in the "optimal" ID set.
    *
-   * @param inputString
-   * @param timeLimit
-   * @return
+   * @param inputString String representation of the GLPK input in MathProg
+   *   format.
+   * @param timeLimit Time limit for this GLPK run in seconds.
+   *
+   * @return String representation of GLPK optimalization output.
    */
   public static String run(final String inputString, final int timeLimit)
         throws InterruptedException {
-    final File inputFile = new File(INPUT);
-    final File outputFile = new File(OUTPUT);
+    final File inputFile = new File(GlpkUtils.INPUT);
+    final File outputFile = new File(GlpkUtils.OUTPUT);
 
     inputFile.delete();
     outputFile.delete();
 
-    writeInput(inputFile, inputString);
+    GlpkUtils.writeInput(inputFile, inputString);
 
     final StringBuilder ret = new StringBuilder();
     try {
       final long startTime = Calendar.getInstance().getTimeInMillis();
 
-      final ProcessBuilder processBuilder = new ProcessBuilder(getParameters(timeLimit));
[…]

### 2011-07-06 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkUtils.java

@@ -19,6 +19,14 @@ package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
 import cz.cuni.mff.ksi.jinfer.attrstats.options.AttrStatsPanel;
 import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
 import cz.cuni.mff.ksi.jinfer.base.utils.FileUtils;
+import java.io.BufferedReader;
+import java.io.File;
+import java.io.IOException;
+import java.io.InputStream;
+import java.io.InputStreamReader;
+import java.io.PrintWriter;
+import java.io.Reader;
+import org.apache.log4j.Logger;
 import org.openide.util.NbPreferences;
 
 /**
@@ -30,6 +38,11 @@ public final class GlpkUtils {
 
   private static final String CMD_OPTS = "-v";
   private static final String EXPECTED = "GLPSOL: GLPK LP/MIP Solver";
+  private static final String TMP = System.getProperty("java.io.tmpdir");
+  public static final String INPUT = TMP + "/glpk_input.txt";
+  public static final String OUTPUT = TMP + "/glpk_output.txt";
+
+  private static final Logger LOG = Logger.getLogger(GlpkUtils.class);
 
   private GlpkUtils() {
 
@@ -89,6 +102,82 @@ public final class GlpkUtils {
   public static String getName(final AttributeMappingId mapping) {
     final String name = mapping.getElement() + "-" +mapping.getAttribute();
     return name.replace(':', '-');
+  }
+
+  /**
+   * Reads the specified {@link Reader} and writes its content to the specified
+   * {@link StringBuilder}.
+   *
+   * @param r Reader to read from.
+   * @param sb String buffer to write to.
+   * @throws InterruptedException If the thread is interrupted.
+   * @throws IOException If there is a problem with reading.
+   */
+  public static void readerToBuilder(final Reader r, final StringBuilder sb)
+          throws InterruptedException, IOException {
+    final BufferedReader reader = new BufferedReader(r);
+
+    String line = reader.readLine();
+    while (line != null) {
+      sb.append(line).append('\n');
+      if (Thread.interrupted()) {
+        throw new InterruptedException();
+      }
+      line = reader.readLine();
+    }
+  }
+
+  /**
+   * Loads the specified resource file and returns it as String.
+   *
+   * @param resource Path to the Java resource.
+   * @return String content of the resource file.
+   */
+  public static String loadTemplate(final String resource) {
+    try {
[…]

### 2011-07-06 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 \usepackage{appendix}
+\usepackage{nomencl}
 
@@
 }
+
+\renewcommand{\nomname}{List of Abbreviations}
+\makenomenclature
 
@@
 
+TODO link to the article mentioning how few XMLs have schemas
+
 \chapter{Definitions}
@@
 
-talk about ID, IDREF, ... talk about XML schema keys
+\begin{itemize}
+	\item XML tree
+	\item Element, Attribute
+	\item ID, IDREF, IDREFS attribute according to specification
+	\item XSD keys (compare them to ID attributes)
+	\item Key according to Keys for XML
+	\item Attribute mapping
+	\item ID set
+\end{itemize}
 
@@
 
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver.
+\nomenclature{MIP}{Mixed Integer Problem}
+
+TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
 
@@
 
+TODO mention intensification, diversification
+
 \subsubsection{Identity}
@@
 
-TODO
+TODO explain how this translates to GLPK input
 
@@
 
+TODO types of data:
+\begin{itemize}
+	\item Realistic - TODO talk about all that are used
+	\item Artificial - random graphs - TODO talk about their generation
+\end{itemize}
+
 \section{Experimental Setup}
+
+TODO we will be using jInfer, see \ref{jInfer}.
 
@@
 
-Input -> How long the search for optimum lasts
-Time limit -> Quality
-Structure of the data -> Which algorithms perform the best
-Alpha, Beta -> Quality
-Which GLPK -> *
-Do I beat FIDAX with random?
-Does Fuzzy beat Random? (Seems that Random is better and even has smaller variance)
+\begin{itemize}
+	\item Input -> How long the search for optimum lasts
+	\item Time limit -> Quality
+	\item Structure of the data -> Which algorithms perform the best
+	\item Alpha, Beta -> Quality
+	\item Which GLPK -> *
+	\item Do I beat FIDAX with random?
+	\item Does Fuzzy beat Random? (Seems that Random is better and even has smaller variance)
+\end{itemize}
 
@@
 \listoftables
+\addcontentsline{toc}{chapter}{List of Tables}
 
 %%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-\chapwithtoc{List of Abbreviations}
+%\chapwithtoc{List of Abbreviations}
+
+\printnomenclature
+\addcontentsline{toc}{chapter}{List of Abbreviations}
 
@@
 %%% textu práce. Přílohy se číslují.
-
-%% TODO Appendix - About jInfer
 
@@
 \chapter{jInfer}
+\label{jInfer}
 

### 2011-07-07 doc/theses/vitasek/diplomka.tex
Thesis: in progress, alpha version of Preface.

@@
 
+% TODO what style of quotation marks to use?
+
 \chapwithtoc{Preface}
 
-TODO why is this important
+TODO why is this important - XMLs are extensively used but lack schemas (link Vosta's Even ant can create...).
+We already have tools to infer the structure (see jInfer), now we would like to go a bit further - to find some integrity constraints to impose on these schemas.
+Integrity schemas can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro?), etc.
+We will focus on the first kind - ID and IDREF attributes.
 
-TODO link to the article mentioning how few XMLs have schemas
+\section{Structure of the thesis}
+
+The thesis will be structured as follows.
+
+First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
+
+Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
+
+This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
+
+We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
+
+Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
+
+An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
 
@@
 
-\nomenclature{MIP}{Mixed Integer Problem}
+TODO rigorously define MIP - optimization problem, some of the variables integral
 
@@
 \subsection{Constructions Heuristics}
+
+\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
 
@@
 \subsection{Improvement Heuristics}
+
+\nomenclature{IH}{Improvement Heuristic}
 
@@
 \begin{itemize}
-	\item Input -> How long the search for optimum lasts
-	\item Time limit -> Quality
+  \item Which construction heuristic is the best - limit GLPK to like 1 second
+  \item With the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
+  \item With the best IH from the previous step, how about using Random as CH?
+	\item GLPK: Input -> How long the search for optimum lasts
+	\item GLPK: Time limit -> Quality
 	\item Structure of the data -> Which algorithms perform the best
@@
 	\item Which GLPK -> *
-	\item Do I beat FIDAX with random?
+	\item Do I beat FIDAX with Random?
 	\item Does Fuzzy beat Random? (Seems that Random is better and even has smaller variance)
@@
 \section{The "Best" Algorithm}
+
+TODO
 

### 2011-07-07 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 
-Název práce: Title [accurate translation of the master thesis title in Czech]
+Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
 
@@
 
-Klíčová slova:  [3-5 keywords in Czech]
+Klíčová slova:  XML, ID atributy, odvozování
 
@@
 
-TODO talk about key definition from \cite{keX}
-
 \begin{itemize}
-	\item XML tree
+	\item XML tree - from FIDAX or wherever
 	\item Element, Attribute
@@
 	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML
+	\item Key according to Keys for XML \cite{keX}
 	\item Attribute mapping
@@
 TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem
+To the best of our knowledge, there are no other articles dealing with this problem.
 
@@
 \section{Independent Set}
+
+TODO define IS rigorously
 
@@
 TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
-TODO mention things like Taboo search
 
-TODO we will be working with heus striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
+TODO mention things like Taboo search (we won't be using them)
+
+TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
 
@@
 
+% TODO can I format things in nomenclature?
 \nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
@@
 \begin{itemize}
-	\item Realistic - TODO talk about all that are used
-	\item Artificial - random graphs - TODO talk about their generation
+	\item Realistic 
+	\item Artificial
 \end{itemize}
+
+\subsection{Realistic data}
+
+TODO talk about all that are used - mention the dataset name, some characteristics (excel sheet)
+
+\subsection{Artificial data}
+
+TODO these are random graphs.
+TODO talk about how they are created, about their structure.
+TODO talk about some "density".
+TODO show a picture, possibly from GraphViz.
 
@@
 
-TODO
+TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
+
+First of all: if we have the time, it is best to just let the GLPK run. 
+We will find the optimum in this way. 
+And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
+
+Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
+For realistic (looks like THIS in graph representation) it is best to do A B C.
+Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
 

### 2011-07-07 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 TODO talk about Fajt - but that's different, that's keys
+TODO add citation for Fajt
 
@@
 
-TODO mention things like Taboo search (we won't be using them)
+TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
 

### 2011-07-09 doc/theses/svirec/kap2.tex
another commit of thesis

@@
-\chapter{Název druhé kapitoly}
+\chapter{Basic Definitions}
 
-\section{Název první podkapitoly v druhé kapitole}
+In this chapter basic definitions necessary in the following chapters are placed. This definitions are cited from \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
 
-\subsection{Preliminaries}
-
-\subsubsection{XML Trees and DTDs}
+\section{XML Trees and DTDs}
 \begin{define}[Tree]
-{\sl Tree} $T$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is an (acyclic) set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$.
+Given an alphabet of nodes $\mathbb{N}$ and an alphabet of node labels $\Sigma$, a {\sl tree} $T$ over $\mathbb{N}$ and $\Sigma$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is an set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$ which are connected and acyclic.
 \end{define}
 
-\noindent Let also denote the set of leaf nodes as $Leaves(T)$ and the set of trees defined on an alphabet on node labels $\Sigma$ as $T_\Sigma$.
+\noindent Let also denote the set of leaf nodes as $Leaves(T)$ and the set of trees defined over an alphabet of node labels $\Sigma$ as $T_\Sigma$.
 
@@
 
-\begin{define}[DTD]
+\begin{example}
+\end{example}
+
+\begin{figure}
+    \centering\includegraphics[scale=\myscale]{example1}
+	\caption{XML Tree} \label{example1}
+\end{figure}
+
+\begin{define}[DTD]\label{dtdDef}
 {\sl DTD} is a tuple $D = (\tau, \alpha, P, R, rt)$ where: i) $\tau$ and $\alpha$ are of the same definition as in the \emph{XML tree}; ii) $P$ is the set of \emph{element type definitions}; iii) R is the set of \emph{attribute lists}; iv) $rt \in \tau$ is the tag of the document root element.\qed
@@
 As we defined \emph{path} on a DTD $D$, let denote as $paths(D)$ the set of paths which can be defined on a DTD $D$.
-Also another important notion is $p(XT)$, which is the set of nodes from XML tree $XT$ conforming DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$.
+Also another important notion is $p(XT)$ (or $\{[\![p]\!]\}$), which is the set of nodes from XML tree $XT$ conforming DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$.
+The set of nodes reachable from a node $v$ following path $p$ is denoted as $\{v[\![p]\!]\}$.
+When there is only one node in $\{v[\![p]\!]\}$, we use $v[\![p]\!]$ to denote this node.
 Moreover, let denote $XT.p$ the \emph{answer} of the path $p$ applied on $XT$ , that is:
@@
 
-\subsection{XML and Functional Dependencies}
+\section{Integrity Constraints and Functional Dependencies}
 
@@
 Given a set of functional dependencies $\mathcal{FD} = \{F_1 , \dots, F_n\}$ over $D$, we say that $XT$ satisfies $\cal FD$ if it satisfies $F_i$ for every $i \in 1..n$.
+
+\todo[inline]{tuto treba doplnit definiciu satisfiability of tree to FD}
+\begin{define}[satisf tree to FD]\label{treeSatisf}
+\end{define}
+
+\todo[inline]{tuto treba doplnit definiciu integrity constraints}
+\begin{define}[Integrity Constraints]\label{integConstr}
+\end{define}

### 2011-07-09 doc/theses/svirec/kap1.tex
another commit of thesis

\chapter{Introduction}

\section{Aim of the Thesis}

\section{Structure of the Thesis}

### 2011-07-09 doc/theses/svirec/kap3.tex
another commit of thesis

@@
 
-This chapter introduce description and categorization of recent approaches of repairing XML documents with uses of functional dependencies.
+This chapter introduces description and categorization of recent approaches of repairing XML documents with usage of functional dependencies.
 
-The problem of finding optimum repair was primary discussed
+All approaches dealing with the problem of finding optimal repair can be divided into categories according to usage of elementary repair primitives.
+This repair primitives are: inserting node, deleting node, updating node and marking node as unreliable of XML document.
 
-\section{este neviem}
-
-\subsection{Repairs and Consistent Answers for XML Data with Functional Dependencies}
+\section[Repairs and Consistent Answers for XML Data]{Repairs and Consistent Answers for XML Data with Functional Dependencies}
 
 A technique for computing repairs which solves the problem of XML data inconsistency with respect to a set of functional dependencies was proposed in \cite{RepAndConsistentAnswer}.
-In this approach, authors are trying to find minimal set of update operations which makes XML data consistent.
-This update operations can be divided into two categories i) replacing value associated with element or attribute, and ii) mark particular node information as unreliable.
+In this approach, the authors are trying to find a minimal set of update operations which makes XML data consistent.
+These update operations can be divided into two categories i) replacing a value associated with an element or an attribute, and ii) marking a particular node information as unreliable.
 
-\subsubsection{XML Tree and Functional Dependency}
+\subsection{XML Tree and Functional Dependency}
 
-To be able to resolve problem of functional dependency violations in XML document, authors tries to introduce concept of functional dependendenies based on those defined for relational databases.
-In a relational database $D$ a functional dependency $A \rightarrow B$ model correspondence between values $A$ and $B$ in tuples of $D$.
-Because the standard concent of a tuple is not defined for XML document, authors introduce concept of tree tuple, which corresponds to the concept of tuple in relational database:
+To be able to resolve problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependendenies based on those defined for relational databases.
+In a relational database $D$ a functional dependency $A \rightarrow B$ models correspondence between values $A$ and $B$ in tuples of $D$.
+Because the standard concent of a tuple is not defined for XML document, authors introduce concept of a tree tuple which corresponds to the concept of tuple in relational database:
 
 \begin{define}[Tree Tuple]
-Given an XML tree XT conforming the DTD D, a tree tuple t of XT is a maximal sub-tree of XT such that, for every path $p \in paths(D)$, t.p contains at most one element.\qed
+Given an XML tree XT conforming a DTD D, a tree tuple t of XT is a maximal sub-tree of XT such that, for every path $p \in paths(D)$, t.p contains at most one element.\qed
 \end{define}
@@
 
-\subsubsection{Repairing inconsistent XML data}
+\subsection{Repairing inconsistent XML data}
 
 Authors of this approach chose two kinds of actions to repair inconsistent XML data with regard to functional dependencies.
-First action is updating value of an attribute or content of an element.
-As a second action authors choose marking inconsistent element as "unreliable" rather than deleting it, because removing elements from a XML document leads to some undesired drawbacks: it does not always suffice to remove inconsistency and deleting a node can lead to a new document not conforming the given schema.
+The first action is updating the value of an attribute or content of an element.
+As the second action authors choose marking inconsistent element as "unreliable" rather than deleting it, because removing elements from an XML document leads to some undesired drawbacks: it does not always suffice to remove inconsistency and deleting a node can lead to a new document not valid against the given schema.
 
 Depending on XML data and defined functional dependencies, each inconsistency could have many possible strategies to repair it.
-From all the possible repair strategies, authors prefer those, for which smaller changes are made to the original document.
+From all the possible repair strategies, the authors prefer those, for which smaller changes are made to the original document.
 
 \begin{example}
-Consider XML tree $XT$ of Figure conforming the DTD $D$ defined below, which is representing collection of books, and given the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.
+Consider XML tree $XT$ conforming the DTD $D$ defined below, which is representing collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.
 \begin{verbatim}
 <!ELEMENT bib (book+)>
-<!ELEMENT book (written_by, title,
-                pub, year?)>
+<!ELEMENT book (written_by, title)>
 <!ELEMENT written_by (author+)>
@@
 <!ELEMENT title PCDATA>
-<!ELEMENT pub PCDATA>
-<!ELEMENT year PCDATA>
 \end{verbatim}
-The functional dependency defined above requires that for each book, there is only one element author having a given $@ano$ value, therefore $XT$ does not satisfy given functional dependency, because two author elements of the same book have the same value of attribute $@ano$.
+The functional dependency defined above requires that for each book, there is only one element author having a given $@ano$ value.
+Therefore $XT$ does not satisfy the given functional dependency, because two author elements of the same book have the same value of attribute $@ano$.
 To resolve this data inconsistency, we can use two different repair strategies: 1) changing one of the value of attribute $@ano$; 2) marking one of the element author as unreliable.
-Because the first strategy changes only attribute $@ano$, it is preferred before second strategy, which changes larger portion of document, since it marks whole author element as unreliable.
+Since the first strategy changes only attribute $@ano$, it is preferred to second strategy, which changes a larger portion of document, since it marks whole author element as unreliable.
 \qed
@@
 
-\subsubsection{Repair Algorithm}
+\subsection{Repair Algorithm}
 
-Before introducing the algorithm to repair inconsistent XML document, lets define reliability of elements in XML tree:
+Before introducing the algorithm to repair inconsistent XML document, let us define reliability of elements in XML tree:
 
 \begin{define}[R-XML Tree]
-A R-XML $tree$ is a triplet $RXT = \langle T, \delta, \varrho \rangle$, where $\langle T, \delta \rangle$ is an XML tree and $\varrho$ is a reliability function from $N_T$ to \texttt{\{true, false\}}, such that, for each pair of nodes $n_1 , n_2 \in N_T$ with $n_2$ descendent of $n_1$, it holds that $\varrho(n_1) = false \Rightarrow \varrho(n_2) = false$.\qed
+A R-XML $tree$ is a triplet $RXT = \langle T, \delta, \varrho \rangle$, where $\langle T, \delta \rangle$ is an XML tree and $\varrho$ is a reliability function from $N_T$ to \texttt{\{true, false\}}, such that, for each pair of nodes $n_1 , n_2 \in N_T$ with $n_2$ descendant of $n_1$, it holds that $\varrho(n_1) = false \Rightarrow \varrho(n_2) = false$.\qed
 \end{define}
 
-R-XML Tree is used as an input for a main algorithm computing repaired R-XML tree described in Algorithm \ref{repAlgo}.
-First the algorithm compute all the possible repairs of tuples which doesn't satisfy functional dependecy using the function \texttt{computeRepairs} (lines 2-6).
+To be able to create a repair, R-XML Tree must not satisfy FD according to definition of weak satisfiability:
+
+\begin{define}[Weak satisfiability]
+Let $RXT = \langle T, \delta, \varrho \rangle$ be an R-XML tree conforming a DTD $D$, and $f: S \rightarrow p$ be a functional dependency.
+We say that $RXT$ weakly satisfies $f$ ($RXT \models_w f$) if one of the following conditions holds:
+\begin{enumerate}
+	\item $\langle T, \delta \rangle \models f$ (definition \ref{treeSatisf});
+    \item for each pair of tuples $t_1$, $t_2$ of $RXT$ one of the following holds:
+    \begin{enumerate}
+    	\item there exists a path $p_i \in S$ such that: \\
+$(\varrho(p_i(t_1)) = false) \lor (\varrho(p_i(t_2)) = false)$;
+        \item $(\varrho(p(t_1)) = false) \lor (\varrho(p(t_2)) = false)$.\qed
+    \end{enumerate}
+\end{enumerate}
+\end{define}
+
+The repair of an R-XML tree which not satisfies $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$).\\
+With a repair $\langle \delta, \varrho \rangle$ of R-XML tree and a set of labelled nodes $N$ of this tree, we denote with $Update_{\delta}(N)$ the set of nodes modified by $\delta$.
+Analogously, we denote $True_{\varrho}(N) = \{n \in N| \varrho(n) = true\}$ and $False_{\varrho}(N) =\{n \in N | \varrho(n) = false\}$.
+
+\begin{define}[Minimal Repair]
+Let $RXT = \langle T, \delta, \varrho \rangle$ be an R-XML tree conforming DTD $D$, $\mathcal{FD}$ a set of functional dependencies and $R_1 = \langle \delta_1, \varrho_1 \rangle$, $R_2 = \langle \delta_2, \varrho_2 \rangle$ two repairs for $RXT$.
+We say that $R_1$ is smaller than $R_2$ ($R_1 \preceq R_2$) if $Updated_{\delta_1}(N_T)\ \cup\ False_{\delta_1}(N_T) \subseteq Updated_{\delta_2}(N_T)\ \cup\ False_{\delta_2}(N_T)$ and $False_{\delta_1}(N_T) \subseteq False_{\delta_2}(N_T)$.
+Repair $R$ is minimal if there is no repair $R' \neq R$ such that $R' \preceq R$.\qed
+\end{define}
+
+An R-XML tree is used as an input for the main algorithm computing repaired R-XML tree described in Algorithm \ref{repAlgo}.
+First the algorithm computes all the possible repairs of tuples which not satisfy a functional dependecy using the function \texttt{computeRepairs} (lines 2-6).
 Next all non-minimal repairs are removed from all possible repairs (line 7).
-In a last step, all the repairs are merged and a unique repaired R-XML tree is returned.
+In the last step, all the repairs are merged and a unique repaired R-XML tree is returned.
 
-\begin{algorithm}
+\begin{algorithm}[H]
 \caption{XML Repair}
@@
 
-The function \texttt{computeRepairs} gets R-XML tree, functional dependency $F$ and a tuples $t_1, t_2$ of R-XML tree as input and computes repair as follows:
+Function \texttt{computeRepairs}(in algorithm \ref{computeRepairs}) gets R-XML tree, a functional dependency $F$ and a tuples $t_1, t_2$ of R-XML tree as input and computes the repair as follows:
 \begin{itemize}
@@
 	\begin{itemize}
-		\item If path $p_i$ denotes a textual element, then one of the two terminal values $t_1.p_i$ or $t_2.p_i$ is changed to $\perp$ (line 9).
+		\item If path $p_i$ denotes a textual element, then one of the two terminal values $t_1.p_i$ or $t_2.p_i$ is changed to the new generated value ($\perp$) (line 9).
 		\item Otherwise $p_i$ denotes a node, so one of the nodes $t_1.p_i$ or $t_2.p_i$ is marked as unreliable (line 11).
@@
 \caption{$computeRepairs(F, t_1, t_2, RXT)$}
-\begin{algorithmic}[1]
+\begin{algorithmic}[1]\label{computeRepairs}
 \REQUIRE{\ \\
@@
 
-\subsubsection{Conclusion}
+\subsection{Conclusion}
 
-\todo[inline]{Sem pojde zhrnutie tohto approachu}
+The authors proposed a technique for repairing XML documents violating functional dependencies based on approaches proposed for relational database repairing.
+The algorithm introduces two possible repair primitives, which creates many possible result from which those with minimal impact on document are chosen.
+However the authors not consider creation of new violations after repairing the initial violations as it is in \cite{ImprovingXML}.
+Another disadvantage of this approach is that a unnecessary repair of some particular violation could be applied to an XML document because another repair could repair that violation before.
 
-\subsection{Querying and Repairing Inconsistent XML Data}
+\section{Querying and Repairing Inconsistent XML Data}
 
-Studying the problem of repairing inconsistent XML document with respect to a set of functional dependencies and investigating the existence of repairs has been introduced in \cite{QueryXML}.
-Authors introduce two kind of repair primitives, first is deleting "unreliable" portions of document and second is inserting new part.
-As it is in another approaches, authors prefer minimal set of repair primitives applied to the XML document to form a repair.
+Studying the problem of repairing inconsistent XML documents with respect to a set of functional dependencies and investigating the existence of repairs has been introduced in \cite{QueryXML}.
+Authors introduce two kind of repair primitives.
+The first is deleting "unreliable" nodes of document and the second is inserting a new nodes.
+Similarly to another approaches, authors prefer minimal set of repair primitives applied to the XML document to form a repair.
 
-Introduced repair primitives authors use in three different repair strategies consisting of:
+The introduced repair primitives the authors use in three different repair strategies consisting of:
 \begin{enumerate}
 	\item \textit{(general) repairs}, where both delete and insert operations are used,
-	\item \textit{cleaning repairs}, for documents interpreted as "dirty", only delete operations are used to repair inconsistencies,
-	\item \textit{completing repairs}, for documents interpreted as incomplete, insert operations are used.
+	\item \textit{cleaning repairs}, where for documents interpreted as "dirty" only delete operations are used to repair inconsistencies,
+	\item \textit{completing repairs}, where for documents interpreted as incomplete, insert operations are used.
 \end{enumerate}
 
-\subsubsection{General Repair}
+\subsection{General Repair}
 
-With the insert and delete operations as a repair primitives, the structure of the XML document will be updated.
-The insert operation is represented as $\langle  + [x]a[y], z\rangle$, where i) $x$ is a node identifier, ii) $a$ is a label, iii) $y$ is either a node identifier or a value ($y$ is a value if $a \in Att \cup \{S\}$; otherwise is a node identifier), and iv) $z$ denotes the child of $x$ which must immediately precede $y$ ($\perp$ if $y$ is inserted as first or unique child of $x$).
-The deletion is represented as $-[x]a[y]$, where $x$ is a node identifier, $a \in Att \cup El \cup \{S\}$, and $y$ is either a node $id$ or a string value, denoting the node to be deleted.\\
-With defined basic repair operations for general repair of inconsistent XML document, lets define repair as follows:
+With the insert and delete operations as repair primitives, is updated the structure of the XML document which conforms DTD $D$ (defined in Definition \ref{dtdDef}) and violated functional dependency.
+The insert operation is represented as $\langle  + [x]a[y], z\rangle$, where i) $x$ is a node identifier, ii) $a$ is a label, iii) $y$ is either a node identifier or a value ($y$ is a value, if $a \in \alpha \cup \{S\}$; otherwise it is a node identifier), and iv) $z$ denotes the child of $x$ which must immediately precede $y$ ($\perp$ if $y$ is inserted as the first or a single child of $x$).
+The deletion is represented as $-[x]a[y]$, where $x$ is a node identifier, $a \in \alpha \cup \tau \cup \{S\}$, and $y$ is either a node $id$ or a string value, denoting the node to be deleted.\\
+The set $R$ of update operations can be divide into two subsets: $R^+$ is a the subset of all the insertion operations in $R$; $R^-$ is the subset of all the deletion operations.
+A set $R$ is said to be consistent if the following conditions are hold:
+\begin{enumerate}
+	\item the deletion of node implies the deletion of all descendant nodes;
+    \item insertions cannot refer to deleted nodes.
+\end{enumerate}
+We say that two sets of update operations $R_1$ and $R_2$ are equivalent ($R_1 \equiv R_2$) if $R_1$ is equal to $R_2$ up to an injective renaming of node identifiers.
+Moreover, we say that $R_1 \preceq R_2$ if $R_1^- \subseteq R_2^- \lor R_1^- = R_2^-$ and $R_1^+ \subseteq R_2^+$.
+We say that $R_1 \sqsubseteq R_2$ if there exists a $R_2' \equiv R_2$ such that  $R_1 \preceq R_2'$.
+At last $R_1 \sqsubset R_2$ if $R_1 \sqsubseteq R_2$ and $R_1 \not \equiv R_2$.\\
+
+With defined basic repair operations for general repair of inconsistent XML document let us define a repair as follows:
 
 \begin{define}[Repair]
-Given a XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$, a set of update operations $R$ is said to be a repair of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
+Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$(definition \ref{integConstr}), a set of update operations $R$ is said to be a repair of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
 \end{define}
 
-With further investigation, authors discover that the problem of deciding whether there exists a repair for a XML document in the presence of DTD with functional dependencies is undecidable.
+With further investigation, the authors discover that the problem of deciding whether there exists a repair for a XML document in the presence of DTD with functional dependencies is undecidable.
 Therefore they consider restricted forms of repairs, more specifically cleaning and completing repairs.
 
-\subsubsection{Conclusion}
+\subsection{Conclusion}
 
-\todo[inline]{Sem pojde zhrnutie tohto approachu}
+In this approach the authors introduce different types of repairs (general, cleaning and completing) and focus on checking whether exists a repair for many classes of integrity constraints (general integrity constraints, inclusion dependencies, functional dependencies, etc.).
+For the functional dependencies the authors found out that the problem of checking whether there exists a general repair for an XML document is $\mathcal{NP}$-complete.
 
-\subsection{Improving XML Data Quality with Functional Dependencies}
+\section{Improving XML Data Quality with Functional Dependencies}
 
 The algorithm for repairing XML functional dependency violations which uses modification of node value as the repair primitive has been proposed in \cite{ImprovingXML}.
-For finding the optimal repair of XML document, authors introduce a cost model, which assign a weight to each leaf node in XML document.
-The optimal repair is that with lowest repair cost, which is measured by a total weight of the modified nodes.
+For finding the optimal repair of an XML document, the authors introduce a cost model, which assign a weight to each leaf node in an XML document.
+The optimal repair is the one with the lowest repair cost which is measured by the total weight of the modified nodes.
 
-Authors of this approach found out that by repairing one functional dependency violation can break another, therefore they separate the main algorithm into two phases.
+The authors of the approach found out that repairing one functional dependency violation can violate another.
+Therefore they divide the main algorithm into two phases.
 In the first phase, the conflict hypergraph capturing the initial functional dependency violations is constructed, and all the violations are fixed by modifying the values of all the nodes on a vertex cover of the conflict hypergraph.
@@
 
-\subsubsection{Cost Model and Repairing Primitive}
+\subsection{Cost Model and Repairing Primitive}
 
-For each leaf node $v$ of XML document is associated a weight in the range $[0,1]$, which is denoted $W(v)$.
-Lets assume that larger the weight of the leaf is, then the leaf is more reliable.
-The weight may be automatically generated by some statistical methods or can be assigned by a user.
+Each leaf node $v$ of XML tree is associated a weight from range $[0,1]$, which is denoted $W(v)$.
+Let us assume that the larger the weight of the leaf the more reliable it is.
+The weight may be automatically generated by statistical methods or it can be assigned by a user.
 
@@
 With a given DTD $D$, a functional dependency is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$.
-Here $P$ is a  root path, or $P = \epsilon$.
+Here $P$ is a  root path (path where first element is a root element of an XML document), or $P = \epsilon$ (empty path).
 Each $P_i (i \in [1,n])$ is a singleton leaf path, and there is a no non-empty common prefix for $P_1, \dots, P_{n+1}$.
@@
 
-As was mentioned earlier, repairing primitive is node value modification, where for repairing algorithm is used combination of two rules to resolve violations.
-Lets have a $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
-If the child nodes of $v_1$ and $v_2$ qualified by paths $P_i$ have equal values of all $i \in [1,n]$ and their child nodes qualified by $P_{n+1}$ have different values, then $v_1$ and $v_2$ violates $\sigma$.
+As was mentioned earlier, a repairing primitive is a node value modification, where for repairing algorithm a combination of two rules to resolve violation is used.
+Let us have a $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
+If the child nodes of $v_1$ and $v_2$ qualified by paths $P_i$ have equal values for all $i \in [1,n]$ and their child nodes qualified by $P_{n+1}$ have different values, then $v_1$ and $v_2$ violates $\sigma$.
 The first rule used to repair this violation is to change the value of the node qualified by $P_{n+1}$ from $v_1$ to the value of $v_2$'s child node that matches $P_{n+1}$(or reversely).
-The second rule is to choose an arbitrary $P_i (i \in [1,n])$, and introduce a fresh new value to the node qualified by $P_i$ from $v_1$ (or $v_2$).
+The second rule is to choose an arbitrary $P_i (i \in [1,n])$, and introduce a new value to the node qualified by $P_i$ from $v_1$ (or $v_2$).
 
-\begin{define}[Optimum repair]
-Given an inconsistent XML document $T$ violating a set $\Sigma$ of $FD$s, the repair $T_R$ of $T$ is called optimum repair, if $T_R$ has the minimum cost among all repairs of $T$.
+\begin{define}[Optimal repair]
+Given an inconsistent XML document $T$ violating a set $\Sigma$ of $FD$s, the repair $T_R$ of $T$ is called optimal repair, if $T_R$ has the minimum cost among all repairs of $T$.
 The cost $cost(T_R)$ is defined as:
@@
 
-\subsubsection{Initial Conflicts Hypergraph}
+\subsection{Initial Conflicts Hypergraph}
 
-Weighted Hypergraph is used by the authors in first part of repair algorithm as a tool modeling initial functional dependency violations in XML document.
+A weighted hypergraph is used in the first part of repair algorithm as a tool modeling initial functional dependency violations in an XML document.
 Hypergraph $g$ of XML document $T$ can be defined as a pair $g = (V,E)$, where $V$ stands for a set of elements (called nodes), and $E$ is a set of non-empty subset of $V$ called hyperedges, more accurately each hyperedge indicates a set of value nodes violating FDs.
-Since hypergraph is weighted and cost model is used in this approach, for each node $v \in V$ of hypergraph is assigned a weight $w(v)$, which is the same as the weight of $v$ in $T$.\\
+Since hypergraph is weighted and a cost model is used in this approach, each node $v \in V$ of the hypergraph is assigned with a weight $w(v)$, which is the same as the weight of $v$ in $T$.\\
 
-To actually resolve the problem of repairing FD violations in XML document, authors convert this problem into well-known problem of weighted vertex cover for hypergraph \cite{ApproxAlgo}.
-Lets have hypergraph $g = (V,E)$ where each hyperedge $e \in E$ is set of value nodes which violates some FD.
-In a repair of inconsistent XML document, for each hyperedge at least one value node is modified, therefore it is essential to find a vertex cover (VC) for $g$, which is a set $S \subseteq V$, such that for all edges $e \in E$, $S \cap e \neq \emptyset$.
-Because hypergraph is weighted, we can define weight of VC as a total weight of all vertices in $S$.\\
+To actually resolve the problem of repairing FD violations in an XML document, the authors convert this problem into well-known problem of weighted vertex cover for hypergraph \cite{ApproxAlgo}.
+Let us have hypergraph $g = (V,E)$, where each hyperedge $e \in E$ is a set of value nodes which violate some FD.
+In a repair of an inconsistent XML document, for each hyperedge at least one value node is modified, therefore it is essential to find a vertex cover (VC) for $g$, which is a set $S \subseteq V$, such that for all edges $e \in E$, $S \cap e \neq \emptyset$.
+Since the hypergraph is weighted, we can define weight of VC as the total weight of all vertices in $S$.\\
 
-Algorithm fixing initial FD violations is shown in Algorithm \ref{fixInit}. The algorithm uses approximation algorithm to find VC for the minimum weighted vertex cover proposed in \cite{ApproxAlgo}.
+The algorithm fixing initial FD violations is shown in Algorithm \ref{fixInit}.
+The algorithm uses approximation algorithm to find VC for the minimum weighted vertex cover proposed in \cite{ApproxAlgo}.
 
@@
 \STATE $val(u)$ := $gen\_new\_value()$
-\STATE \COMMENT Introduce fresh new values to all the remaining nodes in VC
+\STATE \COMMENT Introduce new values to all the remaining nodes in VC
 \ENDFOR
@@
 
-\subsubsection{Resolving Violations Thoroughly}
+\subsection{Resolving Violations Thoroughly}
 
-After repairing initial FD violations, there is a chance that new violations may be introduced, therefore authors provided a method to do modifications on value nodes without incurring new conflicts (Algorithm \ref{fixRest}).
+After repairing initial FD violations, there is a chance that new violations may be introduced, therefore the authors provided a method to do modifications on value nodes without incurring new conflicts (Algorithm \ref{fixRest}).
 This method uses core determinant $C_u$ of value node $u$ defined as follows:
@@
 \begin{define}[Core Determinant]
-Given an XML document $T$, a set $\Sigma$ of FDs and a node $u$ in $T$, we say a set of nodes $\{u_1, u_2, \dots, u_n\}$ is a $\sigma\!-\!determinant$ of $u$, if there exists a nontrivial FD $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ logical implied by $\Sigma$, such that $\exists v \in \{[\![P]\!]\}, \exists v_1 \in \{[\![P']\!]\}, v_1[\![P]\!] = u_i$ for $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] = u$.\\
-We say that a set $C_u$ of nodes is a core determinant of $u$, if (a) for every nontrivial FD $\sigma$ implied by $\Sigma$ and every set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u \cap W \neq \emptyset$; and (b) for any propert subset $C_u'$ of $C_u$, there exists some nontrivial FD $\sigma$ implied by $\Sigma$, and a set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u' \cap W = \emptyset$.\qed
+Given an XML document $T$, a set $\Sigma$ of FDs and a node $u$ in $T$, we say that a set of nodes $\{u_1, u_2, \dots, u_n\}$ is a $\sigma\!-\!determinant$ of $u$, if there exists a nontrivial FD $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ logical implied by $\Sigma$, such that $\exists v \in \{[\![P]\!]\}, \exists v_1 \in \{[\![P']\!]\}, v_1[\![P]\!] = u_i$ for $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] = u$.\\
+We say that a set $C_u$ of nodes is a core determinant of $u$, if (a) for every nontrivial FD $\sigma$ implied by $\Sigma$ and every set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u \cap W \neq \emptyset$; and (b) for any proper subset $C_u'$ of $C_u$, there exists a nontrivial FD $\sigma$ implied by $\Sigma$, and a set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u' \cap W = \emptyset$.\qed
 \end{define}
@@
 
-\subsubsection{Conclusion}
+\subsection{Conclusion}
 
-\todo[inline]{Sem pojde zhrnutie tohto approachu}
+The authors introduced effective two-step heuristic method to solve a problem of finding optimal repair of XML violations against functional dependencies, which is $\mathcal{NP}$-complete.
+Moreover, they experimentally verified the efectivity and scalability of their approach using real-life and synthetic data.

### 2011-07-09 doc/theses/svirec/kap2.tex
added example of XML tree
added some definitions to chapter 2

@@
 
-\begin{example}
+\begin{example}\label{example1ref}
+Consider the following XML document representing a collection of books, and is graphically represented as an XML tree in Fig. \ref{example1}.
+\begin{verbatim}
+<bib>
+  <book>
+    <written_by>
+      <author ano="A1">
+        <name>John Writer</name>
+      </author>
+      <author ano="A1">
+        <name>Eric Seller</name>
+      </author>
+    </written_by>
+    <title>Some title</title>
+  </book>
+  <book>
+    <written_by>
+      <author ano="A2">
+        <name>Adam Publisher</name>
+      </author>
+    </written_by>
+    <title>Some title 2</title>
+  </book>
+</bib>
+\end{verbatim}
+
+The nodes of an XML tree have a label denoting the tag name of the element and unique element identifier in brackets.
+Leaf nodes are either an attribute or the textual content of an element.
+The label of a textual node contains string contained inside of element and unique element identifier.
+The label of an attribute node contains in addition to textual node a name of the attribute.\qed
 \end{example}
 
-\begin{figure}
-    \centering\includegraphics[scale=\myscale]{example1}
-	\caption{XML Tree} \label{example1}
+\begin{figure}[h]
+    \centering\includegraphics[scale=\myscale]{example1-new}
+	\caption{An XML Tree} \label{example1}
 \end{figure}
@@
 {\sl DTD} is a tuple $D = (\tau, \alpha, P, R, rt)$ where: i) $\tau$ and $\alpha$ are of the same definition as in the \emph{XML tree}; ii) $P$ is the set of \emph{element type definitions}; iii) R is the set of \emph{attribute lists}; iv) $rt \in \tau$ is the tag of the document root element.\qed
+\end{define}
+
+\begin{define}[Path Expression]
+A path expression is an expression of the form $p = ('/' | '//')s_1 \dots ('/'|'//')s_m$ where $s_1, \dots, s_{m-1} \in \tau$, and $s_m \in \tau \cup \alpha \cup \{S\}$.
 \end{define}
@@
 		\item for each $i$ in $2..m-1$, $s_i \in \tau$ and $s_i$ appears in the element type definition of $s_{i-1}$;
-		\item $s_m \in \alpha \rightarrow s_m$ appears in the attribute list of $s_{m-1}$;
-		\item $s_m \in \tau \cup \{S\} \rightarrow s_m$ appears the element type definition on $s_{m-1}$.
+		\item $s_m \in \alpha \Rightarrow s_m$ appears in the attribute list of $s_{m-1}$;
+		\item $s_m \in \tau \cup \{S\} \Rightarrow s_m$ appears in the element type definition of $s_{m-1}$.
 	\end{enumerate}\qed
@@
 
-As we defined \emph{path} on a DTD $D$, let denote as $paths(D)$ the set of paths which can be defined on a DTD $D$.
+As we have defined a \emph{path} on a DTD $D$, let us denote as $paths(D)$ the set of paths which can be defined on a DTD $D$.
 Also another important notion is $p(XT)$ (or $\{[\![p]\!]\}$), which is the set of nodes from XML tree $XT$ conforming DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$.
@@
 
-\section{Integrity Constraints and Functional Dependencies}
+\section{Integrity Constraints}
 
-In relational database $D$, correspondence between values $A$ and $B$ in the tuple of $D$, models the functional dependency denoted as $A \rightarrow B$.
-Because in XML is not such a standard tuple concept, authors introduce the concept of tree tuples, corresponding to relational databases concept of tuples.
+Before defining of XML integrity constaint, let us introduce some notation used in the definition.
+A \emph{path atom} is an expression of the form $[x_1]p[x_2]$, where $p$ is a path expression, $x_1$ and $x_2$ are terms, and $x_1 \not \in Str$.\\
+A conjunction of path and built-in atoms $C = [X_1]p_1[Y_1] \cap \dots \cap [X_n]p_n[Y_n] \cap U_1\theta_1 V_1 \cap \dots \cap U_k \theta_k V_k$ is said to be safe if all variables in $C$ are \emph{range restricted}, i.e.
+if
+\begin{itemize}
+ 	\item for every $[X_i]p_i[Y_i]$, either $X_i$ is a constant (node identifier of a string), or there is some $[X_j]p_j[Y_j]$ in $C$ where $X_j$ is range restricted;
+    \item for every built-in term $U_i\theta_i V_i$ occuring in $C$, if $\theta_i$ is equal to $"="$ then at least one of the two terms is range restricted; otherwise both $U_i$ and $V_i$ must be range restricted.
+ \end{itemize}
+A rootless tree formula is an expression of the form $p(\Phi_1 \land \dots \land \Phi_k)$ where $\Phi_i$ is a rootless path formula (expression of the form $p[y]$ where $p$ is a path expression and $y$ is a term) or a rootless tree formula and $p$ is a path expression.
+A tree atom is an expression of the form $[x]T$ where $T$ is a rootless tree formula and $x$ is a term.
 
-\begin{define}[Tree Tuple]
-Given an XML tree XT conforming the DTD D, a tree tuple t of XT is a maximal sub-tree of XT such that, for every path $p \in paths(D)$, t.p contains at most one element.\qed
+\begin{define}[XML Integrity Constraint]\label{integConstr}
+An XML constraint is a formula of the form: $$(\forall X)[\Phi(X)\supset (\exists Y_1)\Psi_1(X,Y_1) \lor \dots \lor (\exists Y_k)(X,Y_k)]$$
+where $X,Y_1,\dots, Y_k$ denote distinct sets of universally and existentially quantified variables, $\Phi(X)$ and $\Phi(X) \land \Psi_i(X, Y_i) (\forall i \in [1..k])$ are safe conjuctions of built-in and tree atoms.\qed
 \end{define}
 
-\begin{define}[Functional Dependency]
-Given a DTD D, a functional dependency on D is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and p is an element of $paths(D)$.\qed
+\section{Functional Dependencies}
+
+Since different approaches uses slighty different representation of an XML document, also definition of functional dependencies differs.
+This is reason we introduce in this section two different definitions.
+
+\subsection{Functional Dependency 1}
+
+In a relational database $DB$, a correspondence between values $A$ and $B$ in the tuple of $D$, models the functional dependency denoted as $A \rightarrow B$.
+Since in XML there is not such a standard tuple concept, the concept of XML tree tuples is introduced, corresponding to relational databases concept of tuples.
+
+\begin{define}[Tree Tuple]\label{treeTuple}
+Given an XML tree XT conforming the DTD $D$, a tree tuple $t$ of $XT$ is a maximal sub-tree of $XT$ such that, for every path $p \in paths(D)$, $t.p$ contains at most one element.\qed
 \end{define}
 
-Given an XML tree $XT$ conforming a DTD $D$ and a functional dependency $F : S_1 \rightarrow S_2$ , we say that $XT$ satisfies $F (XT \models F )$ if for each pair of tree tuples $t_1, t_2$ of $XT$, $t_1.S_1 = t_2.S_1 \land t_1.S_1 = \emptyset \Rightarrow t_1.S_2 = t_2.S_2$ .
-Given a set of functional dependencies $\mathcal{FD} = \{F_1 , \dots, F_n\}$ over $D$, we say that $XT$ satisfies $\cal FD$ if it satisfies $F_i$ for every $i \in 1..n$.
-
-\todo[inline]{tuto treba doplnit definiciu satisfiability of tree to FD}
-\begin{define}[satisf tree to FD]\label{treeSatisf}
+\begin{define}[Functional Dependency]\label{fd1}
+Given a DTD D, a functional dependency on D is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
 \end{define}
 
-\todo[inline]{tuto treba doplnit definiciu integrity constraints}
-\begin{define}[Integrity Constraints]\label{integConstr}
+\begin{example}
+Consider the XML tree of Fig \ref{example1}.
+The following functional dependency express the constraint that two distinct authors of the same book cannot have the same value of attribute \texttt{ano}:$$\{/bib/book, /bib/book/written\_by/author/@ano\} \rightarrow /bib/book/written\_by/author$$\qed
+\end{example}
+
+\subsection{Functional Dependency 2}
+
+\begin{define}[Functional Dependency]\label{fd2}
+With a given DTD $D$, a functional dependency is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$.
+Here $P$ is a  root path (path where first element is a root element of an XML document), or $P = \epsilon$ (empty path).
+Each $P_i (i \in [1,n])$ is a singleton leaf path, and there is a no non-empty common prefix for $P_1, \dots, P_{n+1}$.
+Given an XML document $T$ conforming to $D$, we say $T$ satisfies $\sigma$:iff $\forall v \in \{[\![P]\!]\}, \forall v_1, v_2 \in \{v[\![P']\!]\}$, if $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, then $v_1[\![P_{n+1}]\!] \equiv v_2[\![p_{n+1}]\!]$.\qed
 \end{define}

### 2011-07-09 doc/theses/svirec/kap3.tex
added some definitions to chapter 2

@@
 To be able to resolve problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependendenies based on those defined for relational databases.
-In a relational database $D$ a functional dependency $A \rightarrow B$ models correspondence between values $A$ and $B$ in tuples of $D$.
-Because the standard concent of a tuple is not defined for XML document, authors introduce concept of a tree tuple which corresponds to the concept of tuple in relational database:
-
-\begin{define}[Tree Tuple]
-Given an XML tree XT conforming a DTD D, a tree tuple t of XT is a maximal sub-tree of XT such that, for every path $p \in paths(D)$, t.p contains at most one element.\qed
-\end{define}
-
-\begin{define}[Functional Dependency]
-Given a DTD D, a functional dependency on D is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and p is an element of $paths(D)$.\qed
-\end{define}
+Concept of a tree tuple and functional dependency is defined in Definition \ref{treeTuple} and \ref{fd1}.
 
@@
 \begin{example}
-Consider XML tree $XT$ conforming the DTD $D$ defined below, which is representing collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.
+Consider XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.
 \begin{verbatim}
@@
 \begin{enumerate}
-	\item $\langle T, \delta \rangle \models f$ (definition \ref{treeSatisf});
+	\item $\langle T, \delta \rangle \models f$;
     \item for each pair of tuples $t_1$, $t_2$ of $RXT$ one of the following holds:
@@
 
-\begin{define}[Functional Dependency]
-With a given DTD $D$, a functional dependency is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$.
-Here $P$ is a  root path (path where first element is a root element of an XML document), or $P = \epsilon$ (empty path).
-Each $P_i (i \in [1,n])$ is a singleton leaf path, and there is a no non-empty common prefix for $P_1, \dots, P_{n+1}$.
-Given an XML document $T$ conforming to $D$, we say $T$ satisfies $\sigma$:iff $\forall v \in \{[\![P]\!]\}, \forall v_1, v_2 \in \{v[\![P']\!]\}$, if $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, then $v_1[\![P_{n+1}]\!] \equiv v_2[\![p_{n+1}]\!]$.\qed
-\end{define}
-
 As was mentioned earlier, a repairing primitive is a node value modification, where for repairing algorithm a combination of two rules to resolve violation is used.
-Let us have a $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
+Let us have a $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ (from Definition \ref{fd2}) and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
 If the child nodes of $v_1$ and $v_2$ qualified by paths $P_i$ have equal values for all $i \in [1,n]$ and their child nodes qualified by $P_{n+1}$ have different values, then $v_1$ and $v_2$ violates $\sigma$.

### 2011-07-10 doc/theses/svirec/kap1.tex
another part of thesis

@@
 \chapter{Introduction}
+
+Nowadays, one of the most largely used standard for an information representation and a data exchange on the internet is eXtensible Markup Language (XML) \cite{xml}.
+While XML as a markup language provides syntactic flexibility, the structure of an XML document is described with a so-called XML schema language.
+The two most popular XML schema languages proposed by W3C are Document Type Definition (DTD) \cite{xml} and XML Schema Definition (XSD) \cite{xsd1,xsd2,xsd3}.
+They support some kind of semantic content (e.g., keys and foreign keys), but for improvement of semantic expressiveness of XML integrity constraints for XML \cite{ic} have been defined.
+Several different classes of integrity constraints have been defined for XML, however one most common semantic constraint used in relational databases is functional dependency (FD) \cite{fd}.
+
+With defined set of functional dependencies over an XML document, you can check if an XML document satisfies defined functional dependencies \cite{satifFD}.
+However, data in a real world is typically dirty and often violates (not satisfies) the set of functional dependencies and hence are inconsistent.
+This observation leads to finding an algorithm to compute a repair applied to an XML document such that functional dependencies will be satisfied \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
 
@@
 
+The aim of the thesis is to propose an algorithm repairing XML functional dependency violations for given XML document and set of functional dependencies.
+This thesis analyses recent approaches and discuss their advantages and disadvantages.
+The focus of the thesis is to involve user into process of finding and applying repair and how this interaction helps.
+Part of this thesis is an experimental implementation of the proposed algorithm and its experimental evaluation.
+
 \section{Structure of the Thesis}
+
+Chapter 2 introduces basic definitions which are necessary for further chapters.
+Third Chapter presents recent approaches of repairing XML functional dependency violations.
+In the Chapter 4 is proposed the main algorithm of this thesis.
+Next Chapter contains some details of the experimental implementation.
+Experimental result are shown in Chapter 6.
+Finally, Chapter 7 contains a conclusion and several suggestions for the future work.

### 2011-07-10 doc/theses/svirec/kap2.tex
another part of thesis

@@
 
-Before defining of XML integrity constaint, let us introduce some notation used in the definition.
+Before defining of XML integrity constraint, let us introduce some notation used in the definition.
 A \emph{path atom} is an expression of the form $[x_1]p[x_2]$, where $p$ is a path expression, $x_1$ and $x_2$ are terms, and $x_1 \not \in Str$.\\
@@
  	\item for every $[X_i]p_i[Y_i]$, either $X_i$ is a constant (node identifier of a string), or there is some $[X_j]p_j[Y_j]$ in $C$ where $X_j$ is range restricted;
-    \item for every built-in term $U_i\theta_i V_i$ occuring in $C$, if $\theta_i$ is equal to $"="$ then at least one of the two terms is range restricted; otherwise both $U_i$ and $V_i$ must be range restricted.
+    \item for every built-in term $U_i\theta_i V_i$ occurring in $C$, if $\theta_i$ is equal to $"="$ then at least one of the two terms is range restricted; otherwise both $U_i$ and $V_i$ must be range restricted.
  \end{itemize}
@@
 An XML constraint is a formula of the form: $$(\forall X)[\Phi(X)\supset (\exists Y_1)\Psi_1(X,Y_1) \lor \dots \lor (\exists Y_k)(X,Y_k)]$$
-where $X,Y_1,\dots, Y_k$ denote distinct sets of universally and existentially quantified variables, $\Phi(X)$ and $\Phi(X) \land \Psi_i(X, Y_i) (\forall i \in [1..k])$ are safe conjuctions of built-in and tree atoms.\qed
+where $X,Y_1,\dots, Y_k$ denote distinct sets of universally and existentially quantified variables, $\Phi(X)$ and $\Phi(X) \land \Psi_i(X, Y_i) (\forall i \in [1..k])$ are safe conjunctions of built-in and tree atoms.\qed
 \end{define}
@@
 
-Since different approaches uses slighty different representation of an XML document, also definition of functional dependencies differs.
+Since different approaches uses slightly different representation of an XML document, also definition of functional dependencies differs.
 This is reason we introduce in this section two different definitions.
@@
 \begin{define}[Tree Tuple]\label{treeTuple}
-Given an XML tree XT conforming the DTD $D$, a tree tuple $t$ of $XT$ is a maximal sub-tree of $XT$ such that, for every path $p \in paths(D)$, $t.p$ contains at most one element.\qed
+Given an XML tree $XT$ conforming the DTD $D$, a tree tuple $t$ of $XT$ is a maximal sub-tree of $XT$ such that, for every path $p \in paths(D)$, $t.p$ contains at most one element.\qed
 \end{define}
@@
 \begin{define}[Functional Dependency]\label{fd1}
-Given a DTD D, a functional dependency on D is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
+Given a DTD $D$, a functional dependency on $D$ is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
 \end{define}

### 2011-07-10 doc/theses/svirec/kap3.tex
another part of thesis

@@
 
-To be able to resolve problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependendenies based on those defined for relational databases.
+To be able to resolve problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependencies based on those defined for relational databases.
 Concept of a tree tuple and functional dependency is defined in Definition \ref{treeTuple} and \ref{fd1}.
@@
 An R-XML tree is used as an input for the main algorithm computing repaired R-XML tree described in Algorithm \ref{repAlgo}.
-First the algorithm computes all the possible repairs of tuples which not satisfy a functional dependecy using the function \texttt{computeRepairs} (lines 2-6).
+First the algorithm computes all the possible repairs of tuples which not satisfy a functional dependency using the function \texttt{computeRepairs} (lines 2-6).
 Next all non-minimal repairs are removed from all possible repairs (line 7).
@@
 \STATE Create the initial conflict hypergraph $g$ of $T$ w.r.t $\Sigma$
-\STATE Use a known algorithm to find an approxiamtion $VC$ for the minimum weighted vertex cover of $g$
+\STATE Use a known algorithm to find an approximation $VC$ for the minimum weighted vertex cover of $g$
 \STATE $remaining$ := VC
@@
 The authors introduced effective two-step heuristic method to solve a problem of finding optimal repair of XML violations against functional dependencies, which is $\mathcal{NP}$-complete.
-Moreover, they experimentally verified the efectivity and scalability of their approach using real-life and synthetic data.
+Moreover, they experimentally verified the effectivity and scalability of their approach using real-life and synthetic data.

### 2011-07-11 doc/theses/svirec/kap3.tex
first part of the proposed algorithm

@@
 
-\section[Repairs and Consistent Answers for XML Data]{Repairs and Consistent Answers for XML Data with Functional Dependencies}
+\section[Repairs and Consistent Answers for XML Data]{Repairs and Consistent Answers for XML Data with Functional Dependencies}\label{RepConstAnswers}
 
@@
 
-\begin{define}[R-XML Tree]
+\begin{define}[R-XML Tree]\label{rxmlTree}
 A R-XML $tree$ is a triplet $RXT = \langle T, \delta, \varrho \rangle$, where $\langle T, \delta \rangle$ is an XML tree and $\varrho$ is a reliability function from $N_T$ to \texttt{\{true, false\}}, such that, for each pair of nodes $n_1 , n_2 \in N_T$ with $n_2$ descendant of $n_1$, it holds that $\varrho(n_1) = false \Rightarrow \varrho(n_2) = false$.\qed

### 2011-07-11 doc/theses/svirec/kap4.tex
first part of the proposed algorithm

\chapter{Proposed Algorithm}

\todo[inline]{sem treba dat nejaky zaciatocny obkec}
The main goal of this thesis is to propose an algorithm repairing XML document violating functional dependencies defined for this document. We use an XML tree and tree tuples to represent XML data. To each node of an XML tree is assigned a weight, which is used to measure the cost of a repair.

\section{Repairing Algorithm}

The proposed algorithm is based on the algorithm described in Section \ref{RepConstAnswers} presented in \cite{RepAndConsistentAnswer}. This algorithm was chosen beacause of simple representation of the XML data using a concept which corresponds with concept used in relational databases. Another reason is using besides modification of node value as an update operation also marking particular node information as unreliable, which can reveal forgotten inconsistencies in the data.

The algorithm is splitted into three steps, where second and third step are repeated until all violations are repaired. In the first step, XML document and FDs are loaded, and XML tree with corresponding tree tuples are created. Next step of the algorithm computes the repair candidates for FD violations. In the third step chosen repair candidate is applied to an XML tree.

\subsection{Initial data model}

To represent an XML document we use an extended R-XML tree (Definition \ref{rxmlTree}), called RW-XML tree, which has weights assigned to each node of the tree. The weight of an node indicates correctness of the data the particular node holds, that means the higher the weight is, more correct particular node is. The weights are used to measure cost of repair candidates, where candidate with the lowest cost is picked to be applied to the XML tree. It is important to say that by default to each leaf node is assigned lower weigt than to inner nodes. It is because we prefer the node value modification repair primitive over marking node as unreliable. A definition of an RW-XML tree follows:

\begin{define}[RW-XML tree]
A RW-XML tree is a pair $RWXT = \langle RXT, \omega \rangle$, where $RXT$ is an R-XML tree and $\omega$ is a weight function from $N_T$ to $\mathbb{R}^+_0$.\qed
\end{define}

### 2011-07-12 doc/theses/svirec/kap3.tex
another part of the proposed algorithm

@@
 
-\begin{define}[Weak satisfiability]
+\begin{define}[Weak satisfiability]\label{weakSatisf}
 Let $RXT = \langle T, \delta, \varrho \rangle$ be an R-XML tree conforming a DTD $D$, and $f: S \rightarrow p$ be a functional dependency.

### 2011-07-12 doc/theses/svirec/kap4.tex
another part of the proposed algorithm

@@
 
-\todo[inline]{sem treba dat nejaky zaciatocny obkec}
 The main goal of this thesis is to propose an algorithm repairing XML document violating functional dependencies defined for this document.
@@
 To each node of an XML tree is assigned a weight, which is used to measure the cost of a repair.
+The paths which creates a functional dependency are described with XPath language \cite{xpath}, where only paths with basic construction are allowed (only path constructed with "/", "//" or "@", no wildcards, "[]" or another constructs are allowed).
+We introduce a new concept of the so-called repair group, which clusters repair candidates (repair repairing one FD violation) repairing the same violation, or modifying the same part of the XML tree.
 
@@
 
+\begin{algorithm}
+\caption{Repair RW-XML tree}
+\label{propAlgo}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+$RXT = \langle RT, \omega \rangle$: RW-XML tree conforming a DTD $D$\\
+$\mathcal{FD} = {F_1, \dots, F_m}$: Set of functional dependencies}
+\ENSURE a unique repaired RW-XML tree
+
+\STATE $resultRXT = RXT$
+\WHILE{$resultRXT \not \models_w \mathcal{FD}$}
+    \STATE $S = \emptyset$ \COMMENT Set of repair groups
+    \FORALL{$(F: S \rightarrow p) \in \mathcal{FD}$ s.t. $RXT \not \models_w F$}
+    	\FORALL{$t_1, t_2$ tuples of $RXT$ s.t. $t_1, t_2$ do not weakly satisfy $F$}
+		    \STATE $S = S \cup computeRepairGroup(F, t_1, t_2, RXT, S)$
+	    \ENDFOR
+    \ENDFOR
+
+    \STATE $R = getRepair(S, RXT)$
+    \STATE $resultRXT = applyRepair(R, resultRXT)$
+\ENDWHILE
+
+\end{algorithmic}
+\end{algorithm}
+
 The algorithm is splitted into three steps, where second and third step are repeated until all violations are repaired.
 In the first step, XML document and FDs are loaded, and XML tree with corresponding tree tuples are created.
-Next step of the algorithm computes the repair candidates for FD violations.
+Next step of the algorithm computes repair groups containting repair candidates for FD violations.
 In the third step chosen repair candidate is applied to an XML tree.
+In Algorithm \ref{propAlgo} we can see the simplified process of repairing XML FD violations, which covers second and third step of the whole algorithm.
 
@@
 The weights are used to measure cost of repair candidates, where candidate with the lowest cost is picked to be applied to the XML tree.
+This also first place where interaction of the user can be applied.
+The user could assign weights to the nodes manually (described in \todo{sem doplnit odkaz do implementacnej sekcie}) or can use some sort of statistical methods to generate them automatically.
 It is important to say that by default to each leaf node is assigned lower weigt than to inner nodes.
@@
 \end{define}
+
+After creating RW-XML tree from input XML data, set of tree tuples (defined in Definition \ref{treeTuple}) are constucted.
+Since a set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree.
+Since definition of a tree tuple says that answer of the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on costructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).
+
+Functional dependencies as defined in Definition \ref{fd1} consists of paths described with XPath.
+As was described before, these paths can only have basic construction shown in Example \ref{pathsExample}.
+
+\begin{example}\label{pathsExample}
+Example of paths that can be used in functional dependencies:
+\begin{verbatim}
+//bib/book/author
+/bib/book/author/@ano
+/bib/book/author/name/text()
+\end{verbatim}
+\end{example}
+
+\subsection{Computing Repair Groups}
+
+With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now procceed to compute repair groups.
+First, we need to decide for which FDs violates $RWXT$.
+This can be achieved by finding all tree tuple pairs that not weakly satisfies particular FD (Definition \ref{weakSatisf}).
+For each tuple pair is then computed repair group (line 6 in Algorithm \ref{propAlgo}).
+The function \texttt{computeRepairGroup} is shown in Algorithm \ref{repairGroupAlgo}.
+The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and set of repair groups $RGS$ and computes repair group containig repair candidates as follows:
+
+\begin{enumerate}
+	\item First repair candidates are created using function $computeRepairs$ from Algorithm \ref{repAlgo} (line 1).Since R-XML tree is a special case of RW-XML tree where all weights are equal to zero, we can pass RW-XML tree as a parameter to this function.
+    \item Next is checked if repair candidates intersect another candidades from existing repair groups (line 2)
+    \begin{itemize}
+    	\item If the candidates intersect with some group, they are added to this group (line 3)
+        \item Otherwise new repair group containing repair candidates is created (line 5)
+    \end{itemize}
+\end{enumerate}
+
+\begin{algorithm}
+\floatname{algorithm}{Function}
+\caption{$computeRepairGroup(F, t_1, t_2, RXT, RGS)$}
+\label{repairGroupAlgo}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+$RXT = \langle T, \omega \rangle$: RW-XML tree conforming a DTD $D$\\
+$F: X \rightarrow p$ functional dependency\\
+$t_1, t_2$ tuples of $RXT$\\
+$RGS$ set of repair groups}
+\ENSURE $RG$: repair group
+
+\STATE $S = computeRepairs(F, t_1, t_2, RXT)$ \COMMENT set of repair candidates
+
+\IF{$candidatesIntersectRepairGroups(S, RGS)$}
+\STATE $RG = getIntersectingRepairGroup(S, RGS)$
+\ELSE
+\STATE $RG = createNewRepairGroup(S)$
+\ENDIF
+
+\RETURN $RG$
+\end{algorithmic}
+\end{algorithm}

### 2011-07-13 doc/theses/svirec/kap2.tex
another part of the proposed algorithm

@@
 
-\begin{define}[XML Tree]
+\begin{define}[XML Tree]\label{xmlTree}
 {\sl XML tree} is a pair $XT=\langle T,\delta \rangle$, where:
@@
 \begin{figure}[h]
-    \centering\includegraphics[scale=\myscale]{example1-new}
+    \centering\includegraphics[width=\textwidth]{example1-new}
 	\caption{An XML Tree} \label{example1}
@@
 
+\begin{example}
+Consider the XML tree $XT$ of Fig. \ref{example1}. The subtrees of $XT$ shown in Fig. \ref{tuples} are tree tuples, and the subrees in Fig. \ref{nottuples} are not.
+
+\begin{figure}[h]
+    \centering\includegraphics[width=\textwidth]{tuples}
+	\caption{Two tree tuples of the XML tree of Fig. \ref{example1}} \label{tuples}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \subfloat[]{\label{nottuple1}\includegraphics[scale=0.6]{nottuples1}}
+    \subfloat[]{\label{nottuple2}\includegraphics[scale=0.6]{nottuples2}}
+	\caption{Two subtrees of the XML tree of Fig. \ref{example1} which are not tree tuples}
+    \label{nottuples}
+\end{figure}
+
+In the Fig.
+\ref{nottuple1}, the subtree is not a tree tuple, because the answer of the path $/bib/book/written_by/author$ contains two distinct nodes (i.e.
+$v4$ and $v8$).
+The subtree in the Fig.
+\ref{nottuple2} is not a tree tuble because it is not a maximal subtree (it is a subtree of a tuple from Fig.
+\ref{tuples}).\qed
+
+\end{example}
+
 \begin{define}[Functional Dependency]\label{fd1}

### 2011-07-13 doc/theses/svirec/kap3.tex
another part of the proposed algorithm

@@
 
-\begin{example}
+\begin{example}\label{fdrepairExample}
 Consider XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.

### 2011-07-13 doc/theses/svirec/kap4.tex
another part of the proposed algorithm

@@
 We use an XML tree and tree tuples to represent XML data.
-To each node of an XML tree is assigned a weight, which is used to measure the cost of a repair.
-The paths which creates a functional dependency are described with XPath language \cite{xpath}, where only paths with basic construction are allowed (only path constructed with "/", "//" or "@", no wildcards, "[]" or another constructs are allowed).
-We introduce a new concept of the so-called repair group, which clusters repair candidates (repair repairing one FD violation) repairing the same violation, or modifying the same part of the XML tree.
+Next follows features we want to introduce in our algorithm:
+
+\begin{itemize}
+	\item Incorporation of weight model into the XML data representation to provide the selection of repair candidates with lowest modification cost.
+    \item Involvement of user in the process of finding and applying repair candidates.
+    \item The paths which creates a functional dependency are described with XPath language \cite{xpath}, where only paths with basic construction are allowed (only path constructed with "/", "//" or "@", no wildcards, "[]" or another constructs are allowed).
+    \item We introduce a new concept of the so-called repair group, which clusters repair candidates (repair repairing one FD violation) repairing the same violation, or modifying the same part of the XML tree.
+\end{itemize}
 
@@
 To represent an XML document we use an extended R-XML tree (Definition \ref{rxmlTree}), called RW-XML tree, which has weights assigned to each node of the tree.
-The weight of an node indicates correctness of the data the particular node holds, that means the higher the weight is, more correct particular node is.
+The weight of an node is from interval $[0,1]$ and indicates correctness of the data the particular node holds, that means the higher the weight is, more correct particular node is.
 The weights are used to measure cost of repair candidates, where candidate with the lowest cost is picked to be applied to the XML tree.
-This also first place where interaction of the user can be applied.
+This is also a first place where interaction of the user can be applied.
 The user could assign weights to the nodes manually (described in \todo{sem doplnit odkaz do implementacnej sekcie}) or can use some sort of statistical methods to generate them automatically.
-It is important to say that by default to each leaf node is assigned lower weigt than to inner nodes.
-It is because we prefer the node value modification repair primitive over marking node as unreliable.
+If the user not assign weights nor use statistical methods, default weight is assigned to all nodes.
 A definition of an RW-XML tree follows:
@@
 \begin{define}[RW-XML tree]
-A RW-XML tree is a pair $RWXT = \langle RXT, \omega \rangle$, where $RXT$ is an R-XML tree and $\omega$ is a weight function from $N_T$ to $\mathbb{R}^+_0$.\qed
+A RW-XML tree is a pair $RWXT = \langle RXT, \omega \rangle$, where $RXT$ is an R-XML tree and $\omega$ is a weight function from $N_T$ to $[0,1]$.\qed
 \end{define}
@@
 With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now procceed to compute repair groups.
-First, we need to decide for which FDs violates $RWXT$.
+Repair group is a set of repair candidates, which repairs the same FD or modifies the same part of $RWXT$.
+Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), where either modifies value of an RW-XML tree node ($\delta$ is defined) or marking a node as unreliable ($\varrho$ is definded).
+The $\delta$ function of repair candidate is defined same as in the XML tree (Definition \ref{xmlTree}) and defines new value of the RW-XML tree node.
+Similarly, $\varrho$ function defines the node which is marked as unreliable.
+
+To be able to compute repair groups, we need to decide for which FDs violates $RWXT$.
 This can be achieved by finding all tree tuple pairs that not weakly satisfies particular FD (Definition \ref{weakSatisf}).
 For each tuple pair is then computed repair group (line 6 in Algorithm \ref{propAlgo}).
-The function \texttt{computeRepairGroup} is shown in Algorithm \ref{repairGroupAlgo}.
-The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and set of repair groups $RGS$ and computes repair group containig repair candidates as follows:
+
+The function \texttt{computeRepairGroup}, responsible for creating the repair group, is shown in Algorithm \ref{repairGroupAlgo}.
+The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containig repair candidates as follows:
 
@@
 \end{algorithm}
+
+In Example \ref{repairExample} are shown repair candidates of one repair group modifying node values and marking nodes as unreliable.
+
+\begin{example}\label{repairExample}
+Consider XML data and functional dependency from Example \ref{fdrepairExample}, where XML data is graphically represented as XML tree $XT$ in Fig \ref{example1}.
+The $XT$ violates FD $$\{/bib/book, /bib/book/written\_by/author/@ano\} \rightarrow /bib/book/written\_by/author$$ because two authors of the same book have the same value of attribute $@ano$.
+One repair group with these repair candidates is created: $R_1 = \langle \{\delta(v5) = \perp\},\varrho_{\{\}}(v) \rangle$, $R_2 = \langle \{\delta(v9) = \perp\},\varrho_{\{\}}(v) \rangle$, $R_3 = \langle \{\},\varrho_{\{v4,v5,v6,v7\}}(v) \rangle$, $R_4 = \langle \{\},\varrho_{\{v8,v9,v10,v11\}}(v) \rangle$, $R_5 = \langle \{\},\varrho_{\{v2,v3,\dots,v13\}}(v) \rangle$.
+
+First two repair candidates modifies the value of an attribute $@ano$ by assigning new generated value ($\perp$).
+Next two repair candidates mark each author node with its children nodes that has the same $@ano$ attribute respectively.
+The last candidate marks as unreliable whole subtree with the book node holding authors with the same attribute as root node.\qed
+\end{example}
+
+\subsubsection{Weight model of a Repair Group}
+
+In this section, we introduce a weight model of a repair group, which will be used in the next step of our repair algorithm.
+Before we define a weight of repair group, let us denote some important notations.
+
+Given a repair candidate $R$, we denote with $Modified_\delta(R)$ the set of nodes modified by $\delta$ function from $R$.
+Analogously, we denote $Modified_\varrho(R)$ the set of nodes modified by $\varrho$ function.
+
+Since each repair candidate consists of a set of nodes which are modified (marked as unreliable or has modified value), we can compute cost of each candidate.
+It is important to say, that in out approach, we prefer repair candidates that modify value of the nodes before those marking nodes as unreliable.
+Therefore we added the koeficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying node value will have lower cost.
+The definition of repair candidate cost follows:
+
+\begin{define}[Repair Candidate Cost]
+Given an RW-XML tree $RXT$ and a repair candidate $R$, we define the cost of $R$ as:
+$$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \times k,$$
+where $k$ is koeficient determining a priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
+\end{define}
+
+By default is $k$ set to value that cost of repair candidate marking unreliable node will be higher than the one that modifies node value.
+However this is another place, where can user intervene and change the priority of repair candidates.
+We show in Example \ref{koefK}, how $k$ can change costs of repair candidates.
+
+\begin{example}\label{koefK}
+Consider XML tree $XT$ and repair candidates from Example \ref{repairExample}.
+A default weight of all nodes $v$ of $XT$ is $\omega(v) = 0.5$.
+If the koeficient $k = 1$, cost of repair candidates would be as follows: $cost(R_1) = 0.5$, $cost(R_2) = 0.5$, $cost(R_3) = 2$, $cost(R_4) = 2$ and $cost(R_5) = 6$.
+Let us assume, that a repair candidate with the lowest repair cost will be applied to $XT$.
+If we order costs of repair candidates from the lowest to the highest, you can see that candidates which have marked unreliable nodes are at the end (and will not be applied to $XT$).
+
+Howerver, if we set the koeficient $k = 0.1$, costs of candidates would be sorted as follows: $cost(R_3) = 0.2$, $cost(R_4) = 0.2$, $cost(R_1) = 0.5$, $cost(R_2)  = 0.5$ and $cost(R_5) = 0.6$.
+You can see that order of costs has significantly changed and the cadidate which will be applied is $R_3$ (is marking nodes as unreliable).\qed
+\end{example}
+
+And finally, since a repair group is a set of repair candidates, its weight can be defined as follows:
+
+\begin{define}[Repair Group weight]
+Given an RW-XML tree $RXT$ and a repair group $RG$, we define the weight of $RG$ as the sum of costs of all repair candidates in the $RG$.
+\end{define}
+
+\subsection{Repair Candidate Selection and Application}
+
+The last step of our algorithm can be divided into two steps, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.

### 2011-07-14 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

\chapter{Analysis of Recent Approaches}
\todo[inline]{Translate.}
Existuje mnoho prístupov generovania XML schém. Väčšina z nich spracováva XML data \todo{spomenúť alebo rozpísať tieto prístupy, ako napríklad EXTRACT?}. Podľa mojej znalosti v dobe písania práce existuje len jediný \todo{je pravda?} prístup využívajúci XML operácie\todo{ref do literatury}. Ten sa snaží hľadať keys a foreign keys za pomoci XQuery dotazov.

\section{Discovering XML Keys and Foreign Keys in Queries}
Táto metóda sa snaží vylepšiť automatické odvodzovanie XML schém hľadaním keys a foreign keys zo sady XQuery dotazov. Pre toto hľadanie sú použité len tieto dotazy, nie samotné XML data. Výsledkom je zoznam keys a foreign keys, ktoré môžu byť zachytené pomocou \textbf{key}, \textbf{keyref} \todo{aj unique?} konštrukcií jazyka XML Schema.

\subsection{Keys and foreign keys}
Autori si zavádzajú definíciu pre key a foreign key:

\begin{define}[Key]
Key je trojica $$(C, P, \{L\})$$ kde $C$, $P$, $L$ sú XPath cesty bez predikátov a používajú len osi \texttt{child} a \texttt{descendant}.
$C$ môže byť vynechané. $(P, \{L\})$ je v takom prípade to isté ako $(/, P, \{L\})$ a nazýva sa \emph{global key}, inak sa nazýva \emph{local key}.
\end{define}

The key specifies the following condition. Let $c$ be an element targeted by $C$ and $p$ and $p’$ be two elements targeted by $P$ from $c$. If the value targeted by $L$ from $p$ equals to the value targeted by $L$ from $p’$, then $p$ and $p’$ are the same elements. In other words, no two different elements targeted by $P$ from $c$ can have the same value of $L$.

\begin{define}[Foreign key]
Foreign key je konštrukcia $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ kde $(C, P_2, \{L_2\})$ je key, $P_1$, $L_1$ sú XPath cesty bez predikátov a používajú len osi \texttt{child} a \texttt{descendant}.
$C$ môže byť vynechané podobne ako v prípade key.
\end{define}

Let $c$ be an element targeted by $C$ and $p_1$ be a element targeted by $P_1$ from $c$. The foreign key specifies that there is an element $p_2$ targeted by $P_2$ from $c$ such that the value targeted by $L_1$ from $p_1$ equals to the value targeted by $L_2$ from $p_2$. In other words, each element targeted by $P_1$ from $c$ refers to an element targeted by $P_2$ from $c$ via the pair $L_1$ and $L_2$.

\subsection{Key and foreign key discovery}
\subsubsection{Joins in queries}
Na hľadanie využíva metóda element/element joins v dotazoch.

Assume a query $Q$ that joins a sequence of elements $S_1$ targeted by a path $P_1$ with a sequence of elements $S_2$ targeted by a path $P_2$ on a condition $L_1 = L_2$. It means that $Q$ joins an element $e_1$ from $S_1$ with an element $e_2$ from $S_2$ if $e_1/L_1$ equals to $e_2/L_2$.

\todo[inline]{priklad?}

Metóda predpokladá, že každý join je robený cez pár key/foreign key. On the base of this assumption, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.

Autori uvádzajú dve pozorovania:
\begin{enumerate}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item Assume that an element $e_1$ from $S_1$ can be joined with more elements from $S_2$. It means that there can be more different elements in $S_2$ having their value of $L_2$ equal to $e_1/L_1$. In other words, there can be more different elements in $S_2$ having the same value of $L_2$. Therefore, we infer that $L_2$ can not be a key of the elements in $S_2$. Moreover, because we suppose that one of $L_1$ and $L_2$ is a key and $L_2$ can not be a key, we infer that $L_1$ is a key of the elements in $S_1$. Consequently, we infer that $L_2$ is a foreign key referring $L_1$.
\item Vice versa, assume that $e_1$ can be joined with maximally one element from $S_2$. It means that there is maximally one element in $S_2$ having its value of $L_2$ equal to $e_1/L_1$. In that case we suppose that each element in $S_2$ has a unique value of $L_2$. Therefore, we infer that $L_2$ is a key of the elements in $S_2$ and consequently that $L_1$ is a foreign key referring $L_2$. We cannot infer whether $L_1$ is a key or not.
\end{enumerate}

\subsubsection{Join patterns}
Rozhodnutie pre jeden z prípadov (O1) a (O2) pre konrétny join je robené podľa tvaru tohto joinu. V dotaze sa hľadajú takzvané \emph{join patters}. Tie sú dva, \texttt{for} \emph{join pattern} a \texttt{let} \emph{join pattern} a vyzerajú takto (prvý je \texttt{for} \emph{join pattern}, druhý \texttt{let} \emph{join pattern}):
\todo[inline]{spravne kreslit indexy pri pismenkach}
\begin{verbatim}
01  for e1 in P1               for $e1 in P1
02  return                     return
03    for $e2 in                 let $e2 :=
        P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
04    return CR                  return CR
\end{verbatim}

$P_1$, $P_2$, $L_1$ and $L_2$ in the patterns are XPath paths without predicates. Both patterns differ only at line \texttt{03} where the former applies \texttt{for} while the other applies \texttt{let}. Line \texttt{01} is called \emph{declaration clause}. Line \texttt{03} is called \emph{join clause}. The path $P_1$ is called source path, $P_2$ \emph{target path}, and the condition $L_2 = \$e_1/L_1$ \emph{join condition}. $C_R$ is called \emph{return clause}. All paths $P_1^R$, ..., $P_k^R$ in $C_R$
starting with $\$e2$ are called \emph{target return paths}.

Each pattern occurrence is marked as \emph{repeating} or \emph{non-repeating}. An occurrence marked as \emph{repeating} means that each element targeted by $P_1$ can be joined with more than one elements targeted by $P_2$. In contrary, a pattern occurrence marked as \emph{non-repeating} means that each element targeted by $P_1$ can be joined with zero or one element targeted by $P_2$. In other words, the former means that the observation (O1) can be applied while the other means that (O2) can be applied to infer keys and foreign keys.

Assume a pattern occurrence $\pi$. The decision is made on the base of the following rules (R1) - (R5). Only one rule can be applied. Process of decision starts with (R1). If (Ri) can not be applied, it tries (Ri+1). If any of the following rules (R1) - (R3) can be applied, $\pi$ is marked as \emph{repeating}. Each pattern occurence has also assigned its weight determining how sure the method is about the inferred statement.

\begin{enumerate}
\renewcommand{\theenumi}{(R\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item $\pi$ is an occurrence of the \texttt{for} pattern (weight: 1).
\item Aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path $P_i^R$ in $\pi$ (weight: 1).
\item Aggregation function \texttt{count} is applied on a \emph{return path} $P_i^R$ in $\pi$ (weight: 0.75).
\end{enumerate}

If (R1) - (R3) can not be applied, $\pi$ is marked as \emph{non-repeating}. The weight is assigned according to (R4) and (R5) listed bellow where $k$ denotes the number of target \emph{return paths} in $\pi$.

\begin{enumerate}
\renewcommand{\theenumi}{(R\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\setcounter{enumi}{4}
\item $k > 1$ (weight: 1)
\item$ k <= 1$ (weight 0.5)
\end{enumerate}

\subsubsection{Key and foreign key inference}
Let $w$ be the weight assigned to a pattern occurrence $\pi$. If $\pi$ is marked as \emph{repeating}, (O1) is applied and the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(P_2^{\downarrow}, \{L_2\})$ is not satisfied
\item $(P_1^{\downarrow}, \{L_1\})$ is satisfied
\item $(P_2^{\downarrow}, \{L_2\}) \Rightarrow (P_1^{\downarrow}, \{L_1\})$ is satisfied
\end{itemize}
where for a path $P$, $P^{\downarrow}$ denotes:
\begin{itemize}
\item $P$, if $P$ does not start with a variable
\item $P^{\$e\downarrow}/P'$ (or $P^{\$e\downarrow}//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) where $P^{\$e}$ is a path used for the declaration of the variable $\$e$ in the query
\end{itemize}

If $\pi$ is marked as \emph{non-repeating}, (O2) is applied and the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(P_2^{\downarrow}, \{L_2\})$ is satisfied
\item $(P_1^{\downarrow}, \{L_1\}) \Rightarrow (P_2^{\downarrow}, \{L_2\})$ is satisfied
\end{itemize}

If the \emph{source} and \emph{target path} in $\pi$ have a common context determined by a path $C$, keys and foreign keys are inferred as follows. If $\pi$ is marked as \emph{repeating}, the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(C^{\downarrow}, P_2^{\downarrow C}, \{L_2\})$ is not satisfied
\item $(C^{\downarrow}, P_1^{\downarrow C}, \{L_1\})$ is satisfied
\item $(C^{\downarrow}, (P_2^{\downarrow C}, \{L_2\}) \Rightarrow (P_1^{\downarrow C}, \{L_1\}))$ is satisfied
\end{itemize}
where for paths $C$ and $P$, $P^{\downarrow C}$ denotes:
\begin{itemize}
\item $P$, if $P$ does not start with a variable
\item $P'$ (or $.//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) and $\$e$ is declared by $C$
\item $P^{\$e\downarrow C}/P'$ (or $P^{\$e\downarrow C}//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) and $\$e$ is not declared by $C$
\end{itemize}

If $\pi$ is marked as \emph{non-repeating}, the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(C^{\downarrow}, P_2^{\downarrow C}, \{L_2\})$ is satisfied
\item $(C^{\downarrow}, (P_1^{\downarrow C}, \{L_1\}) \Rightarrow (P_2^{\downarrow C}, \{L_2\}))$ is satisfied
\end{itemize}

\section{Scoring function}
Môže nastať situácia, že nie je splnený nejaký z predpokladov, z ktorých metóda vychádza. Predpokladajme nejakú množinu dotazov. Môže sa stať, že z nejakého dotazu z tejto množiny metóda odvodí key $K$ a z nejakého iného dotazu, že $K$ nie je key. \todo[inline]{priklad?}

Autori preto zavádzajú key scoring function. If a new key $K$ is going to be inferred, it has assigned an initial score 0. Each inferred positive statement about $K$ increases the score of $K$ by the weight of the statement. Respectively, each negative statement about $K$ decreases the score by the respective weight. The resulting score therefore summarizes the weights of all inferred statements about $K$. A positive score means that $K$ is probably satisfied while negative means that $K$ is probably not satisfied. The higher the absolute value of the score is the higher the probability is.

Let $K_1$, ..., $K_n$ be the inferred keys. Let $S_i$ be the score of $K_i$ and $N_i$ be the number of the inferred statements about $K_i$. The precision of the scoring is further enhanced on the base of the following observation. Assume a key $K_i = (C, P, \{L\})$ and $K_j = (C' , P, \{L\})$ where the path $C$ targets ancestors of the elements targeted by $C'$, i.e. the context specified by $C$ covers the context specified by $C'$. In that case it is also said that $K_i$ \emph{covers} $K_j$. It can be easily seen that if $K_i$ is satisfied, $K_j$ must be satisfied as well. On the other hand, if $K_j$ is not satisfied, $K_i$ can not be satisfied too. Therefore, if the score $S_i$ of $K_i$ is positive (i.e. $K_i$ is satisfied), $S_j$ is incremented with $S_i$ and $N_j$ with $N_i$ (i.e. $K_j$ is satisfied as well). Conversely, if the score $S_j$ of $K_j$ is negative (i.e. $K_j$ is not satisfied), $S_i$ is incremented with $S_j$ and $N_i$ with $N_j$ (i.e. $K_i$ can not be satisfied too).

Finally, the scores of the inferred keys are normalized to be comparable with each other. The normalized scores are from the range $\langle -1, 1\rangle$. The normalization takes into account not only the scores summarizing the weights of the statements about keys but also the number of the statements. The normalized score is computed as follows. Let $S^{max}$ be the maximum from $|S_1|$, ..., $|S_n|$ and $N^{max}$ be the maximum from $N_1$, ..., $N_n$. The normalized score $S_i^{norm}$ of $K_i$ is computed as follows:
$$S_i^{norm} = {S_i \over S^{max}} * (1 - {N^{max} - N_i \over \sum _{i=1}^n N_i})$$

\section{Conclusion}
The output of our method is a list of scored keys and for each key a list foreign keys referencing the key. The score of a key can be negative or positive. A negative score means that the key is not specified by the XML documents while positive means that it is satisfied. The absolute value of the score means how sure the method are about it.

Because the method is based on intuition of how XQuery constructs are commonly applied in practice, it can be imprecise in certain cases.

### 2011-07-14 doc/theses/vitasek/diplomka.tex
Thesis: in progress.
Thesis: in progress

@@
 We already have tools to infer the structure (see jInfer), now we would like to go a bit further - to find some integrity constraints to impose on these schemas.
-Integrity schemas can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro?), etc.
+Integrity schemas can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
 We will focus on the first kind - ID and IDREF attributes.
@@
 	\item ID set
+  \item Weight: support, coverage
 \end{itemize}
@@
 
-TODO define IS rigorously
+TODO define IS + MIS rigorously
 
@@
 
-TODO how GLPK works, how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
+TODO how GLPK works (branch&bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
 
@@
 
+TODO describe all the inputs (take it from the mindmap), describe how the output looks (sample trace - link to appendix)
+
 \section{Experimental Data}
@@
 	\item Realistic 
+  \item Realistic with artificial attributes
 	\item Artificial
 \end{itemize}
+
+TODO for all types, show GraphViz pictures.
 
@@
 TODO talk about all that are used - mention the dataset name, some characteristics (excel sheet)
+
+\subsection{Realistic data with artificial attributes}
+
+TODO we have taken some of the text data and converted it to attributes like this, A B C
 
@@
 TODO talk about some "density".
-TODO show a picture, possibly from GraphViz.
 
@@
 
-TODO we will be using jInfer, see \ref{jInfer}.
+TODO we will be using jInfer and its module AttributeStats see \ref{jInfer} and \ref{attrstats}.
 
@@
 	\item Alpha, Beta -> Quality
-	\item Which GLPK -> *
+  \item Alpha, Beta -> How does the solution work?
+	\item Which GLPK -> How does (time->quality) look, how does (input->time) look
 	\item Do I beat FIDAX with Random?
@@
 
+\chapter{Future work}
+
+TODO extension to more documents, IDREFs, ...
+
 \chapwithtoc{Conclusion}
 
-TODO conclusion
+TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
+We discussed the approach chosen in FIDAX.
+Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
+weight) ID set.
+
+However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
+By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
 
@@
 
-\printnomenclature
+\printnomenclature[2cm]
 \addcontentsline{toc}{chapter}{List of Abbreviations}
@@
 
+TODO what it is, who created it and why, what it can do, where to get it.
+
+\chapter{AttributeStats}
+\label{attrstats}
+
+TODO talk a bit about the structure of AttributeStats module.
+
+\chapter{Traces}
+\label{traces}
+
+TODO a few (commented!) traces from experiment runs
+
 \end{document}

### 2011-07-14 doc/theses/svirec/kap4.tex
another part of the proposed algorithm

@@
     \item Involvement of user in the process of finding and applying repair candidates.
+    \item Apply one repair candidate at the time and recalculate repair candidates again to prevent using unnecessary repairs to repair one violation and also to prevent of introducing new violations, which will not be repaired.
     \item The paths which creates a functional dependency are described with XPath language \cite{xpath}, where only paths with basic construction are allowed (only path constructed with "/", "//" or "@", no wildcards, "[]" or another constructs are allowed).
@@
 Analogously, we denote $Modified_\varrho(R)$ the set of nodes modified by $\varrho$ function.
+The count of nodes modified by repair candidate $R$ is denoted with $\lambda(R)$.
+At last, we denote $PS(R)$ a set of paths defining all nodes modified by $R$ (example of the set $PS(R)$ is shown in Example \ref{PSExample}).
+
+\begin{example}\label{PSExample}
+Consider XML tree from Fig. \ref{example1} and a repair candidate, which marks nodes $v12$ and $v13$ as unreliable. Then the set $PS(R)$ contains following paths:
+\begin{verbatim}
+/bib/book/title
+/bib/book/title/text()
+\end{verbatim}\qed
+\end{example}
 
@@
 The last step of our algorithm can be divided into two steps, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.
+
+\subsubsection{Selection of Repair Candidate}
+
+The previous part of the algorithm computes repair groups containing repair candidates, which can repair violations of provided FDs.
+From this candidates, we must choose one that will be applied on RW-XML tree.
+Since we introduced a weight model to the repair group, we can use the weight of repair groups and the cost of repair candidates to choose suitable candidate.
+In our approach, we introduce two distinct algorithms: first that not involve the user to process of selection and the second one that use the user interaction.
+
+The first algorithm simply selects the first repair group with the lowest weight, and from it the repair candidate with lowest cost is selected.
+
+The latter algorithm allows user to choose the repair candidate which is most suitable for his needs.
+Very important aspect of all user interactions in this kind of algorithm is that the user will not be willing to select more than ten repairs.
+We could simply allow user to switch to the first selection algorithm, but that not take in account previous user selections of repair candidates.
+Therefore we introduce in this algorithm a functionality, that is able from selection done by the user before, guess his next selection.
+
+The function $selectRepairByUser(RGS, SR, t)$, responsible for selection of repair candidate involving user interactivity is shown in Algorithm \ref{selectUser}.
+
+First is decided, if the algorithm is in a state that the user is selecting repair candidates (user selection mode), or the user leaves decision making on the algorithm using his previous selections (guess mode) (line 2).
+If we are still in user selection mode, the user choose from repair groups sorted by the weight the most convenient repair group, and from this group the user choose the repair candidate that will be applied on RW-XML tree (line 4).
+Repair candidates in each repair group are also sorted by the cost, working as a hint for the user which repair would be chosen by previous automatic method of selection.
+Last step of user selection mode is saving the information from selected repair candidate (line 5).
+This information consists of the FD which this candidate repairs, nodes the repair changes (their paths) and also if the change was a value modification or marking node as unreliable.
+
+If the user has decided not to selecting repair candidates by himself, the algorithm goes into guess mode (staring on line 6).
+In this mode the algorithm checks all repair groups if some of it containts the repair candidate, that is sufficiently similar with some previously selected repair candidate (line 7-10).
+This similarity is checked by function \texttt{canBeUsedUserSelection} shown in Algorithm \ref{canUserSel}.
+If neither of candidates is sufficiently similar, the algorithm choose repair candidate with the lowest cost from the repair group with the lowest weight (line 15-16).
+
+\begin{algorithm}[H]
+\floatname{algorithm}{Function}
+\caption{$selectRepairByUser(RGS, SR, t)$}
+\label{selectUser}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+$RGS$ set of repair groups\\
+$SR$: set of previously selected repair candidates by the user\\
+$t$: modified nodes count treshold of previously selected repair candidates\\}
+\ENSURE $R$: repair candidate
+
+\STATE $R = \emptyset$
+\IF{$isUserSelection()$}
+    \STATE \COMMENT repair candidate selected by user
+    \STATE $R = getRepairFromUser()$
+    \STATE $saveSelectedRepair(SR, R)$
+\ELSE
+    \FORALL{$RG$ in $RGS$}
+        \STATE \COMMENT for all repair candidates in the repair group
+        \FORALL{$RC$ in $RG$}
+            \IF{$canBeUsedUserSelection(SR, t, RC)$}
+                \RETURN $RC$
+            \ENDIF
+        \ENDFOR
+    \ENDFOR
+    \STATE $RG = getFirstRG(RGS)$ \COMMENT get the repair group with smallest weight
+    \STATE $R = getFirstRepairCandidate(RG)$ \COMMENT get the repair candidate with the lowest cost
+\ENDIF
+
+\RETURN $R$
+\end{algorithmic}
+\end{algorithm}
+
+The function \texttt{canBeUsedUserSelection} gets actual repair candidate $RC$, set of all previously selected repair candidates $SR$, modified nodes count treshold $t$ from interval $(0,1]$ of previously selected repair candidate and determine if $RC$ is sufficiently similar as some repair candidate from $SR$.
+To be similar with some previously selected repair candidate $S$, $RC$ must repair the same $FD$ as $S$, it must use the same update operation as $S$ and $\lambda(RC) \geq \lceil\lambda(S) \times t\rceil$ (line 2).
+Furthermore, if $\lambda(RC) = \lceil\lambda(S) \times t\rceil$, set of paths $PS(RC)$ must be a subset of $PS(S)$ (line 3).
+Otherwise, if $\lambda(RC) > \lceil\lambda(S) \times t\rceil$, there must exists a subset of $PS(S)$ with $\lceil\lambda(S) \times t\rceil$ elements that is subset of $PS(RC)$.
+In other words, without taking into account the treshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is such that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes).
+The treshold $t$ can reduce count of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$.
+In Example \ref{tresholdExample} is shown the usage of the treshold $t$.
+
+\begin{algorithm}
+\floatname{algorithm}{Function}
+\caption{$canBeUsedUserSelection(SR, t, RC)$}
+\label{canUserSel}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+$SR$: set of previously selected repair candidates by the user\\
+$t$: suitability treshold $(0,1]$ of previously selected repair candidates\\
+$RC$: actual repair candidate}
+\ENSURE \TRUE\ if actual repair candidate is similar to some previous one.
+\FORALL{$S$ in $SR$}
+    \IF{$S$ repairs the same FD as $RC$ \AND $S$ use the same update operation as $RC$ \AND $\lceil\lambda(SC) \times t\rceil \leq \lambda(RC)$}
+        \IF{$\lceil\lambda(SC) \times t\rceil = \lambda(RC)$ \AND $PS(RC) \subseteq PS(SC)$}
+            \RETURN \TRUE
+        \ENDIF
+        \IF{$\lceil\lambda(SC) \times t\rceil < \lambda(RC)$ \AND there exists a subset $s$ of $PS(SC)$ with $\lceil\lambda(SC) \times t\rceil$ elements, that $s \subseteq$ $PS(RC)$}
+            \RETURN \TRUE
+        \ENDIF
+   \ENDIF
+\ENDFOR
+\RETURN \FALSE
+\end{algorithmic}
+\end{algorithm}
+
+\begin{example}\label{tresholdExample}
+Consider repair candidate $R_1$ selected by user, that marks as unreliable two nodes and $PS(R_1) = \{/bib/book/title, /bib/book/title/text()\}$.
+Next consider repair candidate $R_2$ with $PS(R_2) = \{/bib/book/title/text()\}$, that the modified node is marked as unreliable and both $R_1$ and $R_2$ repairs the same $FD$.
+If treshold $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$.
+However, if treshold is $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
+\end{example}
+
+\subsubsection{Application of Repair Candidate}

### 2011-07-15 doc/theses/svirec/kap3.tex
chapter 4 finished

@@
 
-The repair of an R-XML tree which not satisfies $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$).\\
+The repair of an R-XML tree which not satisfies $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$).
+The composition of the R-XML tree and the repair (i.e.
+composition of their $\delta$ and $\varrho$ functions) is defined as follows:
+
+\begin{define}[Composition of $\delta$ functions]\label{compDelta}
+The composition of two functions $\delta_1$ and $\delta_2$ asssociating values to leaf nodes is
+\begin{displaymath}
+    \delta_1 \cdot \delta_2(n) =
+        \begin{cases}
+            \delta_1(n) & \text{if $\delta_1(n)$ is defined over $n$,}\\
+            \delta_2(n) & \text{otherwise (i.e. $\delta_1(n)$ is not defined over $n$).}
+        \end{cases}
+\end{displaymath}\qed
+\end{define}
+
+\begin{define}[Composition of $\varrho$ functions]\label{compVarrho}
+The composition of two reliability functions $\varrho_1$ and $\varrho_2$ asssociating boolean value to nodes is $$\varrho_1 \cdot \varrho_2(n) = min(\varrho_1(n), \varrho_2(n)),$$ where to the boolean value $true$ (resp.
+$false$) is associated value 1 (resp.
+0).
+\qed
+\end{define}
+
 With a repair $\langle \delta, \varrho \rangle$ of R-XML tree and a set of labelled nodes $N$ of this tree, we denote with $Update_{\delta}(N)$ the set of nodes modified by $\delta$.

### 2011-07-15 doc/theses/svirec/kap4.tex
chapter 4 finished

@@
 The proposed algorithm is based on the algorithm described in Section \ref{RepConstAnswers} presented in \cite{RepAndConsistentAnswer}.
-This algorithm was chosen beacause of simple representation of the XML data using a concept which corresponds with concept used in relational databases.
+This algorithm was chosen because of simple representation of the XML data using a concept which corresponds with concept used in relational databases.
 Another reason is using besides modification of node value as an update operation also marking particular node information as unreliable, which can reveal forgotten inconsistencies in the data.
@@
 
-The algorithm is splitted into three steps, where second and third step are repeated until all violations are repaired.
+The algorithm is split into three steps, where second and third step are repeated until all violations are repaired.
 In the first step, XML document and FDs are loaded, and XML tree with corresponding tree tuples are created.
-Next step of the algorithm computes repair groups containting repair candidates for FD violations.
+Next step of the algorithm computes repair groups containing repair candidates for FD violations.
 In the third step chosen repair candidate is applied to an XML tree.
@@
 
-After creating RW-XML tree from input XML data, set of tree tuples (defined in Definition \ref{treeTuple}) are constucted.
+After creating RW-XML tree from input XML data, set of tree tuples (defined in Definition \ref{treeTuple}) are constructed.
 Since a set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree.
-Since definition of a tree tuple says that answer of the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on costructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).
+Since definition of a tree tuple says that answer of the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on constructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).
 
@@
 
-With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now procceed to compute repair groups.
+With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now proceed to compute repair groups.
 Repair group is a set of repair candidates, which repairs the same FD or modifies the same part of $RWXT$.
-Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), where either modifies value of an RW-XML tree node ($\delta$ is defined) or marking a node as unreliable ($\varrho$ is definded).
+Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), where either modifies value of an RW-XML tree node ($\delta$ is defined) or marking a node as unreliable ($\varrho$ is defined).
 The $\delta$ function of repair candidate is defined same as in the XML tree (Definition \ref{xmlTree}) and defines new value of the RW-XML tree node.
@@
 The function \texttt{computeRepairGroup}, responsible for creating the repair group, is shown in Algorithm \ref{repairGroupAlgo}.
-The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containig repair candidates as follows:
+The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containing repair candidates as follows:
 
@@
 	\item First repair candidates are created using function $computeRepairs$ from Algorithm \ref{repAlgo} (line 1).Since R-XML tree is a special case of RW-XML tree where all weights are equal to zero, we can pass RW-XML tree as a parameter to this function.
-    \item Next is checked if repair candidates intersect another candidades from existing repair groups (line 2)
+    \item Next is checked if repair candidates intersect another candidates from existing repair groups (line 2)
     \begin{itemize}
@@
 It is important to say, that in out approach, we prefer repair candidates that modify value of the nodes before those marking nodes as unreliable.
-Therefore we added the koeficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying node value will have lower cost.
+Therefore we added the coefficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying node value will have lower cost.
 The definition of repair candidate cost follows:
@@
 $$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \times k,$$
-where $k$ is koeficient determining a priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
+where $k$ is coefficient determining a priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
 \end{define}
@@
 A default weight of all nodes $v$ of $XT$ is $\omega(v) = 0.5$.
-If the koeficient $k = 1$, cost of repair candidates would be as follows: $cost(R_1) = 0.5$, $cost(R_2) = 0.5$, $cost(R_3) = 2$, $cost(R_4) = 2$ and $cost(R_5) = 6$.
+If the coefficient $k = 1$, cost of repair candidates would be as follows: $cost(R_1) = 0.5$, $cost(R_2) = 0.5$, $cost(R_3) = 2$, $cost(R_4) = 2$ and $cost(R_5) = 6$.
 Let us assume, that a repair candidate with the lowest repair cost will be applied to $XT$.
@@
 
-Howerver, if we set the koeficient $k = 0.1$, costs of candidates would be sorted as follows: $cost(R_3) = 0.2$, $cost(R_4) = 0.2$, $cost(R_1) = 0.5$, $cost(R_2)  = 0.5$ and $cost(R_5) = 0.6$.
-You can see that order of costs has significantly changed and the cadidate which will be applied is $R_3$ (is marking nodes as unreliable).\qed
+However, if we set the coefficient $k = 0.1$, costs of candidates would be sorted as follows: $cost(R_3) = 0.2$, $cost(R_4) = 0.2$, $cost(R_1) = 0.5$, $cost(R_2)  = 0.5$ and $cost(R_5) = 0.6$.
+You can see that order of costs has significantly changed and the candidate which will be applied is $R_3$ (is marking nodes as unreliable).\qed
 \end{example}
@@
 
-The last step of our algorithm can be divided into two steps, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.
+The last step of our algorithm can be divided into two parts, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.
 
-\subsubsection{Selection of Repair Candidate}
+\subsubsection{Selection of the Repair Candidate}
 
-The previous part of the algorithm computes repair groups containing repair candidates, which can repair violations of provided FDs.
-From this candidates, we must choose one that will be applied on RW-XML tree.
+The previous step of the algorithm computes repair groups containing repair candidates, which can repair violations of provided FDs.
+From this candidates, we must choose one that is applied on RW-XML tree.
 Since we introduced a weight model to the repair group, we can use the weight of repair groups and the cost of repair candidates to choose suitable candidate.
@@
 If the user has decided not to selecting repair candidates by himself, the algorithm goes into guess mode (staring on line 6).
-In this mode the algorithm checks all repair groups if some of it containts the repair candidate, that is sufficiently similar with some previously selected repair candidate (line 7-10).
+In this mode the algorithm checks all repair groups if some of it contains the repair candidate, that is sufficiently similar with some previously selected repair candidate (line 7-10).
 This similarity is checked by function \texttt{canBeUsedUserSelection} shown in Algorithm \ref{canUserSel}.
@@
 $SR$: set of previously selected repair candidates by the user\\
-$t$: modified nodes count treshold of previously selected repair candidates\\}
+$t$: modified nodes count threshold of previously selected repair candidates\\}
 \ENSURE $R$: repair candidate
@@
 
-The function \texttt{canBeUsedUserSelection} gets actual repair candidate $RC$, set of all previously selected repair candidates $SR$, modified nodes count treshold $t$ from interval $(0,1]$ of previously selected repair candidate and determine if $RC$ is sufficiently similar as some repair candidate from $SR$.
+The function \texttt{canBeUsedUserSelection} gets actual repair candidate $RC$, set of all previously selected repair candidates $SR$, modified nodes count threshold $t$ from interval $(0,1]$ of previously selected repair candidate and determine if $RC$ is sufficiently similar as some repair candidate from $SR$.
 To be similar with some previously selected repair candidate $S$, $RC$ must repair the same $FD$ as $S$, it must use the same update operation as $S$ and $\lambda(RC) \geq \lceil\lambda(S) \times t\rceil$ (line 2).
@@
 Otherwise, if $\lambda(RC) > \lceil\lambda(S) \times t\rceil$, there must exists a subset of $PS(S)$ with $\lceil\lambda(S) \times t\rceil$ elements that is subset of $PS(RC)$.
-In other words, without taking into account the treshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is such that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes).
-The treshold $t$ can reduce count of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$.
-In Example \ref{tresholdExample} is shown the usage of the treshold $t$.
+In other words, without taking into account the threshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is such that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes).
+The threshold $t$ can reduce count of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$.
+In Example \ref{tresholdExample} is shown the usage of the threshold $t$.
 
@@
 $SR$: set of previously selected repair candidates by the user\\
-$t$: suitability treshold $(0,1]$ of previously selected repair candidates\\
+$t$: suitability threshold $(0,1]$ of previously selected repair candidates\\
 $RC$: actual repair candidate}
@@
 Next consider repair candidate $R_2$ with $PS(R_2) = \{/bib/book/title/text()\}$, that the modified node is marked as unreliable and both $R_1$ and $R_2$ repairs the same $FD$.
-If treshold $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$.
-However, if treshold is $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
+If threshold $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$.
+However, if threshold is $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
 \end{example}
 
-\subsubsection{Application of Repair Candidate}
+\subsubsection{Application of the Repair Candidate}
+
+In this part, the selected repair candidate is finally applied to the RW-XML tree.
+If after this part the RW-XML tree not violates any of FDs, the whole repair algorithm ends at this point, otherwise repair groups are regenerated and the selection of the repair candidate part take place.
+
+To apply the selected repair candidate $R$ to the RW-XML tree $RXT$ means, that we compose $\delta$ and $\varrho$ functions of $R$ with the corresponding functions of R-XML tree contained in $RXT$.
+This compositions are defined in Definition \ref{compDelta} and \ref{compVarrho}.
+
+After application of the repair candidate, some of RW-XML tree nodes could become unreliable, which can lead to the situation that some of tree tuples are not anymore considered as a tuple (is no longer a maximal subtree).
+Therefore, we need before regeneration of repair groups also check all tree tuples, if they satisfies definition of tree tuple.

### 2011-07-16 doc/theses/svirec/kap5.tex
added implementation chapter

\chapter{Implementation}

A part of this work is experimental implementation of our proposed algorithm. Besides our approach, we have also implementened algorithm proposed in \cite{RepAndConsistentAnswer}, to be able to compare it with our algorithm. Our algorithm has been implemented in the JAVA language in the JDK version 6.0 as a part of the schema inference framework for XML, called jInfer \cite{jinfer}, developed as a part of a Software Project subject.

\section{Architecture}

The main part of our algorithm is implemented as the \texttt{FDRepairer} extension of the jInfer framework. The core package of the extension is \texttt{cz.cuni.mff.\discretionary{}{}{}ksi.jinfer.functionalDependencies} and contains classes representing base object important for repairing (Tuple, RW-XML tree, Path etc.). This package is further divided into the following subpackages:

\begin{description}
	\item[fd] contains data structure of functional dependencies used as an input to the algorithm.
    \item [interfaces] contains JAVA interfaces necessery for integration to the jInfer framework.
    \item [modelGenerator] contains classes responsible for creating data structures from loaded input files.
    \item [newRepairer] contains class \texttt{NewRepairerImpl}, the class that implements proposed algorithm. Also contains other data structures introduced in our approach
    \item [propeties] contains classes reponsible for showing properties of this extension if the jInfer.
    \item [repairer] contains classes implementing the original algorithm our algorithm is based on.
    \item [weights] contains data structure of node weights used as an input to the algorithm.
\end{description}

\subsection{Input Files Processing}

Files that can be added as an input to our algorithm are of three kind. The XML data, file containing functional dependencies and file containing weights. First, the XML is processed by build-in Java DOM parster. Files with FDs and weights both contains XML data with specific structure defined as XML Schema in Fig. \ref{fdschema} and \ref{weightschema} and are both processed with JAXB \cite{JAXB}.

\section{Restriction of Implementation}

The \texttt{FDRepairer} extension of the jInfer framework is an implementation of our proposed algorithm presented in Chapter \ref{kap4}. However, one part of the algorithm have not been implemented yet, namely the intersection of repair candidates with existing repair group.

### 2011-07-16 doc/theses/svirec/attachments.tex
added implementation chapter

\chapwithtoc{Attachments}

\begin{figure}[h]
\caption{XML Schema for weights}
\label{weightschema}
\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">

<xs:complexType name="Tweight">
  <xs:sequence>
    <xs:element name="path" type="xs:string"/>
    <xs:element name="value" type="xs:decimal"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tweights">
  <xs:sequence>
    <xs:element name="weight" type="Tweight"
        minOccurs="1" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:element name="weights" type="Tweights"/>
</xs:schema>
\end{verbatim}
\end{figure}

\begin{figure}[h]
\caption{XML Schema for functional dependencies}
\label{fdschema}

\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">

<xs:complexType name="TleftSidePaths">
  <xs:sequence>
    <xs:element name="path" type="xs:string"
        minOccurs="1" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="TrightSidePaths">
  <xs:sequence>
    <xs:element name="path" type="xs:string"
        minOccurs="1" maxOccurs="1"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tdependency">
  <xs:sequence>
    <xs:element name="leftSidePaths" type="TleftSidePaths"/>
    <xs:element name="rightSidePaths" type="TrightSidePaths"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tdependencies">
  <xs:sequence>
    <xs:element name="dependency" type="Tdependency"
        minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:element name="dependencies" type="Tdependencies"/>
</xs:schema>
\end{verbatim}
\end{figure}

### 2011-07-16 doc/theses/svirec/kap4.tex
added implementation chapter

@@
-\chapter{Proposed Algorithm}
+\chapter{Proposed Algorithm}\label{kap4}
 
@@
 This is also a first place where interaction of the user can be applied.
-The user could assign weights to the nodes manually (described in \todo{sem doplnit odkaz do implementacnej sekcie}) or can use some sort of statistical methods to generate them automatically.
+The user could assign weights to the nodes manually or can use some sort of statistical methods to generate them automatically.
 If the user not assign weights nor use statistical methods, default weight is assigned to all nodes.

### 2011-07-16 doc/theses/svirec/kap7.tex
added implementation chapter

\chapter{Conclusion}

The aim of this thesis was a proposal and implementation of algorithm repairing XML functional dependencies violations. At first, existing solutions were analyzed and described.

\section{Future Work}

### 2011-07-20 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 % TODO what style of quotation marks to use? 
+% TODO how to format text? is textit / texttt / ... OK?
 
@@
 
-TODO why is this important - XMLs are extensively used but lack schemas (link Vosta's Even ant can create...).
-We already have tools to infer the structure (see jInfer), now we would like to go a bit further - to find some integrity constraints to impose on these schemas.
-Integrity schemas can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
+Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
+However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
+To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
+But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
+One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
+
+This work will be building upon the achievements of jInfer schema inference framework, expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
+
+TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
 We will focus on the first kind - ID and IDREF attributes.
+
+TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
 

### 2011-07-25 doc/theses/svirec/kap2.tex
added implementation chapter
added experiments chapter

@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering\includegraphics[width=\textwidth]{tuples}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \centering
@@
 In the Fig.
-\ref{nottuple1}, the subtree is not a tree tuple, because the answer of the path $/bib/book/written_by/author$ contains two distinct nodes (i.e.
+\ref{nottuple1}, the subtree is not a tree tuple, because the answer of the path $/bib/book/written\_by/author$ contains two distinct nodes (i.e.
 $v4$ and $v8$).
@@
 
-\begin{example}
+\begin{example}\label{fdExample}
 Consider the XML tree of Fig \ref{example1}.

### 2011-07-25 doc/theses/svirec/kap6.tex
added experiments chapter

\chapter{Experimental Results}

In this chapter, experimental results of the \texttt{FDRepairer} extension of the jInfer framework are presented. Set of documents and functional dependencies have been provided as an input and results from the former algorithm have been compared with results from \texttt{FDRepairer}. Documents in provided datasets are split into real-world and synthetic XML documents.

\section{Datasets}

All provided documents with functional dependencies and also repaired documents are placed on CD in the directory \texttt{/datasets/}.

\subsection{Real-World Data}

The first real-world dataset originate from the XML data repository \\(\url{http://www.cs.washington.edu/research/xmldatasets/}), concretely Course data derived from university websites. The course data consists, beside of other information, of day of the week, time and place (building and room), where each course is situated. The consistency of data is evaluated against FD defined as: Two courses starting at the same date and time are each situated in a different place.\\

The second dataset is a set of actors of IMDB database obtained from the Niagara data source (\url{http://www.cs.wisc.edu/niagara/data.html}). The dataset consists of a set of authors, where for each author is listed a set of movies he played in. For this dataset, we defined FD as follows: For each two authors, which played in the movie with the same name, the year of release of this movie must be the same.

\subsection{Synthetic Data}

Two synthetic data sets have been constructed, the first represents data introduced in Example \ref{example1} with FD presented in Example \ref{fdExample}. The latter one is created according to Example 1 introduced in \cite{ImprovingXML}.

\section{Algorithms Comparison}

For each datasets, we first run the former algorithm and collect from a repaired document informations about how many nodes have been modified with particular update operation (we split value modification operation into changing value to new one and copying one value from one node to another).

Thereafter we use our proposed algorithm with minimal repair candidate selection mode, and besides the data we collect for former algorithm, we also gather informations about how many repair groups have been created before first application of repair candidate and the total count of picked repair groups. We run our algorithm multiple times with different value of the coefficient $k$ and assigning of node weights $w$.\\

The first real-world data (course data) is represented by \texttt{reed.xml} and \texttt{wsu.xml}. For each document, we execute our algorithm with three different settings. First two uses default value of $k$ ($k=1.5$) and for the last one, we set $k=0.005$. For the second (resp. third) execution, we set a weight $w$ of nodes represented by XPath \texttt{/root/course/time/start\_time/text()} (resp. \texttt{/root/course/time/\discretionary{}{}{}start/text()}) to 0.1.

The document \texttt{actors.xml} represents the second real-world dataset. Our algorithm is executed two times, where for both executions have been set $k=1.5$. Moreover, the second has weights of nodes represented by XPath \texttt{/W4F\_DOC/\discretionary{}{}{}Actor/Filmography/Movie/Year/text()} set to $w=0.2$.\\

Generated datasets represented by \texttt{bib.xml} and \texttt{customers.xml} documents have been both used as a input for our algorithm two times, where first execution was done with default value of $k$. The second execution on \texttt{bib.xml} uses $k=0.1$ and on \texttt{customers.xml} uses $k=1.5$ and $w=0.2$ is set to all nodes represented by \texttt{/customers/country/c\_list/customer/city/text()}.

\section{Results}

All gathered informations from real-world and synthetic datasets are shown in Tables \ref{realWorldTable} and \ref{syntheticTable}. The columns of both tables are defined as follows: RG is count of repair groups created before application of repair candidate, RG total is total count of picked repair groups, U specifies number of nodes marked as unreliable, NV defines number of nodes changing value to a new one and finally ChV specifies number of nodes with value copied from another node.\\

\begin{table}
    \begin{tabular}{| l | r | *{7}{c|}}
    \hline
    & repairer & $k$ & $w$ & RG & RG total & U & NV & ChV\\ \hline
    \multirow{4}{*}{reed.xml} & old repairer & - & - & - & - & 0 & 780 & 0\\
    & \texttt{FDRepairer} & 1.5 & - & 218 & 195 & 0 & 195 & 0\\
    & \texttt{FDRepairer} & 1.5 & 0.1 & 218 & 195 & 0 & 195 & 0\\
    & \texttt{FDRepairer} & 0.005 & 0.1 & 218 & 195 & 5265 & 0 & 0\\ \hline
    \multirow{4}{*}{wsu.xml} & old repairer & - & - & - & - & 0 & 2612 & 0\\
    & \texttt{FDRepairer} & 1.5 & - & 3520 & 653 & 0 & 653 & 0\\
    & \texttt{FDRepairer} & 1.5 & 0.1 & 3520 & 653 & 0 & 653 & 0\\
    & \texttt{FDRepairer} & 0.005 & 0.1 & 3520 & 653 & 20833 & 0 & 0\\ \hline
    \multirow{3}{*}{actors.xml} & old repairer & - & - & - & - & 0 & 62 & 62\\
    & \texttt{FDRepairer} & 1.5 & - & 69 & 62 & 0 & 32 & 30\\
    & \texttt{FDRepairer} & 1.5 & 0.2 & 69 & 79 & 0 & 0 & 79\\ \hline
    \end{tabular}
\caption{Real-World datasets}
\label{realWorldTable}
\end{table}

From the informations in tables is visible, that our approach found for each dataset the repair, which modifies less nodes than old repair algorithm. One exception, where our algorithm modifies more nodes is the case, when we set value of $k$ near 0, which marks nodes as unreliable. With this setting, we want to demonstrate, that with our approach is possible to choose instead of modifying node values, marking nodes as unreliable. This is not possible with old repairer, since it prefers node value modification before marking node as unreliable.

\begin{table}
    \begin{tabular}{| l | r | *{7}{c|}}
    \hline
    & repairer & $k$ & $w$ & RG & RG total & U & NV & ChV\\ \hline
    \multirow{3}{*}{bib.xml} & old repairer & - & - & - & - & 0 & 913 & 0\\
    & \texttt{FDRepairer} & 1.5 & - & 1444 & 913 & 0 & 913 & 0\\
    & \texttt{FDRepairer} & 0.1 & - & 1444 & 913 & 5478 & 0 & 0\\ \hline
    \multirow{3}{*}{customers.xml} & old repairer & - & - & - & - & 0 & 99 & 99\\
    & \texttt{FDRepairer} & 1.5 & - & 232 & 136 & 0 & 72 & 64\\
    & \texttt{FDRepairer} & 1.5 & 0.2 & 232 & 361 & 0 & 0 & 361\\\hline
    \end{tabular}
\caption{Synthetic datasets}
\label{syntheticTable}
\end{table}

With setting the weight to particular nodes, we want to show that our algorithm is capable of choosing the priority of nodes, which values are modified (\texttt{reed.xml} and \texttt{wsu.xml}), or if the value is changed to new value (\texttt{actors.xml} and \texttt{customers.xml}).

Another interesting observation is that for almost all datasets, our algorithm picks less repair groups (from which picks repair candidate to apply to document) than were created before first pick. It means that some repair candidate repairs more than one FD violation and therefore less value modifications were needed. In the case, where the number of picked repair groups outreach the firstly created ones, is visible that some repair candidate introduces new violation which was consequently repaired in the next iteration of our algorithm.

### 2011-07-25 doc/theses/svirec/prace.tex
another commit of thesis
added example of XML tree
another part of the proposed algorithm
chapter 4 finished
added implementation chapter
added experiments chapter

@@
 \documentclass[12pt,a4paper]{report}
-\setlength\textwidth{145mm}
-\setlength\textheight{247mm}
-\setlength\oddsidemargin{15mm}
-\setlength\evensidemargin{15mm}
-\setlength\topmargin{0mm}
-\setlength\headsep{0mm}
-\setlength\headheight{0mm}
-% \openright zařídí, aby následující text začínal na pravé straně knihy
+
+\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
+\setlength{\voffset}{-1in} %analogous voffset
+\setlength{\oddsidemargin}{4cm}
+\setlength{\evensidemargin}{4cm}
+\setlength{\topmargin}{25mm}
+\setlength{\footskip}{3cm}
+\setlength{\headheight}{0cm}
+\setlength{\headsep}{0cm}
+\setlength{\marginparwidth}{0cm}
+\setlength{\marginparpush}{0cm}
+\setlength{\textheight}{20.7cm}
+\setlength{\textwidth}{14.5cm}
 \let\openright=\clearpage
+
+\linespread{1.3}
+%\renewcommand\baselinestretch{1.3} % riadkovanie jeden a pol
 
@@
 \usepackage{amssymb}
+\usepackage{amsmath}
 \usepackage{todonotes}
@@
 \usepackage[chapter]{algorithm}
+\usepackage{subfig}
+\usepackage{url}
+\usepackage{multirow}
+
+\newcommand{\myscale}{0.5}
 
@@
 \def\mftitle{Efficient Detection of XML Integrity Constraints}
-\def\mfthesistype{Master Thesis}
+\def\mfthesistype{MASTER THESIS}
+\def\mfkatedra{Department of Software Engineering}
+\def\mfpracoviste{}
 \def\mfauthor{Michal Švirec}
 \def\mfadvisor{RNDr. Irena Mlýnková, Ph.D.}
-\def\mfplacedate{Praha, 2010}
+\def\mfplacedate{Praha, 2011}
 
@@
 % Trochu volnějąí nastavení dělení slov, neľ je default.
-\lefthyphenmin=2
-\righthyphenmin=2
+%\lefthyphenmin=2
+%\righthyphenmin=2
 
@@
 % (dle Organizační struktury MFF UK)
-Název katedry nebo ústavu
+\mfkatedra
 
@@
 \noalign{\vspace{2mm}}
-Study programme: & program \\
+Study programme: & Informatika \\
 \noalign{\vspace{2mm}}
-Specialization: & obor \\
+Specialization: & ISS \\
 \end{tabular}
@@
 Katedra:  % Případně Ústav:
-Název katedry či ústavu, kde byla práce oficiálně zadána
+Katedra softwarového inženýrství
 % dle Organizační struktury MFF UK
@@
 Vedoucí diplomové práce:
-\mfadvisor, pracoviště
+\mfadvisor, Katedra softwarového inženýrství
+
 % dle Organizační struktury MFF UK, případně plný název pracoviątě mimo MFF UK
@@
 Department:
-Název katedry či ústavu, kde byla práce oficiálně zadána
+\mfkatedra
 % dle Organizační struktury MFF UK v angličtině
@@
 Supervisor:
-\mfadvisor, pracoviště
+\mfadvisor, \mfkatedra
 % dle Organizační struktury MFF UK, případně plný název pracoviątě
@@
 \pagestyle{plain}
+
+\renewcommand\thepage{}
+\tableofcontents
+\newpage
+\renewcommand\thepage{\arabic{page}}
+
 \setcounter{page}{1}
-\tableofcontents\thispagestyle{empty}
 
@@
 \theoremstyle{definition}
-\newtheorem*{define}{Definition}	% Definice nečíslujeme, proto "*"
+\newtheorem{define}{Definition}[chapter]	% Definice nečíslujeme, proto "*"
 \newtheorem{example}{Example}
@@
 \include{kap3}
+\include{kap4}
+\include{kap5}
+\include{kap6}
+\include{kap7}
 
@@
 
-\include{zaver}
+%\include{zaver}
 
@@
 \nocite{*}
-\bibliographystyle{plain}
+\bibliographystyle{ieeetr}
 \bibliography{literatura}
@@
 
-%%% Pouľité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-\chapwithtoc{List of Abbreviations}
-
 %%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
@@
 %%% textu práce. Přílohy se číslují.
-\chapwithtoc{Attachments}
+\include{attachments}
 

### 2011-07-26 doc/theses/svirec/kap7.tex
added conclusion chapter

@@
 
-The aim of this thesis was a proposal and implementation of algorithm repairing XML functional dependencies violations. At first, existing solutions were analyzed and described.
+The aim of this thesis was a proposal and implementation of algorithm repairing XML functional dependencies violations.
+At first, existing solutions were analyzed and described.
+The result of the analyses was that none of the recent approaches uses, or in only limited form, user interaction in the process of finding suitable repair for the FD violation.
+Consequently, we proposed algorithm based on \cite{RepAndConsistentAnswer}, that uses user interaction in a various ways.\\
+
+First, we incorporated a weight model into the XML data representation, which allows us to measure the modification cost of each possible repair candidate.
+This is also the first place for the user to interact in the repair process.
+Next, we have introduced a new concept of repair candidates clusters called repair groups.
+These clusters groups repair candidates according to the violation of which are repairing or part of the XML data that are modifying.
+
+The main user interaction part of our algorithm is a selection of the repair group and consequently the repair candidate which is applied to the XML document.
+Besides the repair candidate selection, we also introduced a mechanism allowing to guess the next selection from the previous candidates selected by the user, if the user does not want to select the repair candidate anymore.
+
+The experimental implementation of proposed algorithm and also the algorithm we based on was integrated into the jInfer framework.
+Both algorithms have been compared against each other on synthetic as well as real-world datasets.
+From the data gathered from the comparison is clear that our approach have found repairs with less modifications applied on the XML data.
+We also shown that with the user interaction is possible to change the usage of update operations used in repair, which may create more reasonable result for the user.
 
 \section{Future Work}
+
+Although our approach has introduced user in the process of repairing FD violations of the XML data, there is still work to be done.
+The first main task in the further work is to extend paths defining functional dependencies.
+With usage of more constructs that XPath provides, one can define more sophisticated FDs.
+
+Another part of the algorithm that can be improved is clustering repair candidates into repair groups. The user can provide some additional criteria, that can change the resulting repair.
+
+Last but not least, guessing part of the user selection algorithm can be improved by using more sophisticated heuristic algorithm to select proper repair candidate with regard to previous candidates selected by the user.

### 2011-07-27 doc/theses/svirec/kap1.tex
another part of the proposed algorithm
corrections

--- 2011-07-09 doc/theses/svirec/kap1.tex
@@
 \chapter{Introduction}
+
+Nowadays, one of the most largely used standards for information representation and data exchange on the Internet is the eXtensible Markup Language (XML) \cite{xml}.
+While XML as a markup language provides syntactic flexibility, the structure of an XML document is described with a so-called XML schema language.
+The two most popular XML schema languages proposed by W3C are Document Type Definition (DTD) \cite{xml} and XML Schema Definition (XSD) \cite{xsd1,xsd2,xsd3}.
+They support some kind of semantic content (e.g., keys and foreign keys), but for improvement of semantic expressiveness XML integrity constraints for XML \cite{ic} have been defined.
+Several different classes of integrity constraints have been defined for XML, however most common semantic constraint used in relational databases is functional dependency (FD) \cite{fd}.
+
+With a defined set of functional dependencies in an XML document, one can check if an XML document satisfies defined functional dependencies \cite{satifFD}.
+However, data in the real world is typically dirty and often violate (does not satisfy) the set of functional dependencies and hence are inconsistent.
+This observation leads to a need of an algorithm to compute a repair applied to an XML document such that functional dependencies will be satisfied  again \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
 
@@
 
+The aim of the thesis is to propose an algorithm repairing XML functional dependency violations for a given XML document and a set of functional dependencies.
+Analyses recent approaches and discusses their advantages and disadvantages.
+The focus of the thesis is to incorporate the weight model and to involve the user into the process of finding and applying the repair and find out how this interaction helps.
+A part of this thesis is an experimental implementation of the proposed algorithm and its experimental evaluation.
+
 \section{Structure of the Thesis}
+
+Chapter 2 introduces basic definitions which are necessary for further chapters.
+Chapter 3 presents recent approaches of repairing XML functional dependency violations.
+In Chapter 4 is the main algorithm of this thesis proposed.
+Chapter 5 contains details of the experimental implementation.
+Experimental result are presented in Chapter 6.
+Finally, Chapter 7 contains a conclusion and several suggestions for future work.

### 2011-07-27 doc/theses/svirec/kap2.tex
corrections

@@
 
-In this chapter basic definitions necessary in the following chapters are placed. This definitions are cited from \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
+In this chapter basic definitions necessary in the following chapters are placed. The definitions are taken from \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
 
@@
 \begin{define}[Tree]
-Given an alphabet of nodes $\mathbb{N}$ and an alphabet of node labels $\Sigma$, a {\sl tree} $T$ over $\mathbb{N}$ and $\Sigma$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is an set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$ which are connected and acyclic.
+Being given an alphabet of nodes $\mathbb{N}$ and an alphabet of node labels $\Sigma$, a {\sl tree} $T$ over $\mathbb{N}$ and $\Sigma$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is a set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$ which are connected and acyclic.
 \end{define}
 
-\noindent Let also denote the set of leaf nodes as $Leaves(T)$ and the set of trees defined over an alphabet of node labels $\Sigma$ as $T_\Sigma$.
+\noindent Let us also denote the set of leaf nodes as $Leaves(T)$ and the set of trees defined over an alphabet of node labels $\Sigma$ as $T_\Sigma$.
 
@@
 {\sl XML tree} is a pair $XT=\langle T,\delta \rangle$, where:
+\renewcommand{\labelenumi}{\roman{enumi})}
 	\begin{enumerate}
-		\item $T = (r, N, E, \lambda)$ is a tree in $T_{\tau\cup\alpha\cup\{S\}}$, where $\tau$ is a tag alphabet, $\alpha$ is an attribute name alphabet and $S$ is a symbol not belonging to $\tau\cup\alpha$ (representing \texttt{\#PCDATA} content of elements)$;$
+		\item $T = (r, N, E, \lambda)$ is a tree from $T_{\tau\cup\alpha\cup\{S\}}$, where $\tau$ is a tag alphabet, $\alpha$ is an attribute name alphabet and $S$ is a symbol not belonging to $\tau\cup\alpha$ (representing \texttt{\#PCDATA} content of elements)$;$
 		\item given a node $n$ of $T$, $\lambda(n) \in \alpha \cup \{S\} \Leftrightarrow n \in Leaves(T);$
-		\item $\delta : Leaves(T) \rightarrow Str$, where $Str$ is a string alphabet, is a function associating a (string) value to every leaf of $T$.
+		\item $\delta : Leaves(T) \rightarrow Str$ where $Str$ is a string alphabet is a function associating a (string) value to every leaf of $T$.
 	\end{enumerate}\qed
@@
 \begin{example}\label{example1ref}
-Consider the following XML document representing a collection of books, and is graphically represented as an XML tree in Fig. \ref{example1}.
+Consider the following XML document representing a collection of books. Its graphical representation as an XML tree is in Fig. \ref{example1}.
 \begin{verbatim}
@@
 The label of a textual node contains string contained inside of element and unique element identifier.
-The label of an attribute node contains in addition to textual node a name of the attribute.\qed
+The label of an attribute node contains in addition to textual node the name of the attribute.\qed
 \end{example}
@@
 \begin{define}[DTD]\label{dtdDef}
-{\sl DTD} is a tuple $D = (\tau, \alpha, P, R, rt)$ where: i) $\tau$ and $\alpha$ are of the same definition as in the \emph{XML tree}; ii) $P$ is the set of \emph{element type definitions}; iii) R is the set of \emph{attribute lists}; iv) $rt \in \tau$ is the tag of the document root element.\qed
+{\sl DTD} is a tuple $D = (\tau, \alpha, P, R, rt)$ where:
+\renewcommand{\labelenumi}{\roman{enumi})}
+\begin{enumerate}
+	\item $\tau$ and $\alpha$ are of the same definition as in the \emph{XML tree}
+    \item $P$ is the set of \emph{element type definitions}
+    \item $R$ is the set of \emph{attribute lists}
+    \item $rt \in \tau$ is the tag of the document root element.
+\end{enumerate}
+\qed
 \end{define}
@@
 \begin{define}[Path Expression]
-A path expression is an expression of the form $p = ('/' | '//')s_1 \dots ('/'|'//')s_m$ where $s_1, \dots, s_{m-1} \in \tau$, and $s_m \in \tau \cup \alpha \cup \{S\}$.
+A {\sl path expression} is an expression of the form $p = ('/' | '//')s_1 \dots ('/'|'//')s_m$ where $s_1, \dots, s_{m-1} \in \tau$, and $s_m \in \tau \cup \alpha \cup \{S\}$.\qed
 \end{define}
@@
 {\sl Path} p on a DTD $D = (\tau, \alpha, P, R, rt)$ is a sequence $p = s_1, \dots, s_m$ of symbols in $\tau \cup \alpha \cup \{S\}$ such that:
+\renewcommand{\labelenumi}{\roman{enumi})}
 	\begin{enumerate}
@@
 
-As we have defined a \emph{path} on a DTD $D$, let us denote as $paths(D)$ the set of paths which can be defined on a DTD $D$.
-Also another important notion is $p(XT)$ (or $\{[\![p]\!]\}$), which is the set of nodes from XML tree $XT$ conforming DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$.
+As we have defined a \emph{path} on a DTD $D$, let us denote $paths(D)$ as the set of all paths which can be defined on a DTD $D$.
+Also, another important notion is $p(XT)$ (or $\{[\![p]\!]\}$), which is the set of nodes from XML tree $XT$ conforming to DTD $D$, which can be reached by the path $p \in paths(D)$, starting from the root of $XT$.
 The set of nodes reachable from a node $v$ following path $p$ is denoted as $\{v[\![p]\!]\}$.
 When there is only one node in $\{v[\![p]\!]\}$, we use $v[\![p]\!]$ to denote this node.
-Moreover, let denote $XT.p$ the \emph{answer} of the path $p$ applied on $XT$ , that is:
+Moreover, let $XT.p$ denote the \emph{answer} of the path $p$ applied on $XT$ that is:
 \begin{itemize}
-	\item[-] if $p \in EPath(D)$, where $EPath(D)$ denotes the set of the paths whose last symbol denotes an element, then $XT.p = p(XT)$
-	\item[-] if $p \in StrPath(D)$, where $StrPath(D)$ denotes the set of paths whose last symbol denotes either the textual content of an element or an
+	\item if $p \in EPath(D)$, where $EPath(D)$ denotes the set of the paths whose last symbol denotes an element, then $XT.p = p(XT)$
+	\item if $p \in StrPath(D)$, where $StrPath(D)$ denotes the set of paths whose last symbol denotes either the textual content of an element or an
 attribute, then $XT.p = \{\delta_T(x)|x \in p(XT)\}$.
@@
 
-Before defining of XML integrity constraint, let us introduce some notation used in the definition.
+Before defining of the XML integrity constraint, let us introduce some notation used in the definition.
 A \emph{path atom} is an expression of the form $[x_1]p[x_2]$, where $p$ is a path expression, $x_1$ and $x_2$ are terms, and $x_1 \not \in Str$.\\
-A conjunction of path and built-in atoms $C = [X_1]p_1[Y_1] \cap \dots \cap [X_n]p_n[Y_n] \cap U_1\theta_1 V_1 \cap \dots \cap U_k \theta_k V_k$ is said to be safe if all variables in $C$ are \emph{range restricted}, i.e.
+A conjunction of path and built-in atoms $C = [X_1]p_1[Y_1] \cap \dots \cap [X_n]p_n[Y_n] \cap U_1\theta_1 V_1 \cap \dots \cap U_k \theta_k V_k$ is said to be \emph{safe} if all variables in $C$ are \emph{range restricted}, i.e.
 if
@@
 
-Since different approaches uses slightly different representation of an XML document, also definition of functional dependencies differs.
-This is reason we introduce in this section two different definitions.
+Since different approaches uses slightly different representation of an XML document, also the definition of functional dependencies differs, too.
+This is the reason why we introduce two different definitions in this section.
 
@@
 
-In a relational database $DB$, a correspondence between values $A$ and $B$ in the tuple of $D$, models the functional dependency denoted as $A \rightarrow B$.
-Since in XML there is not such a standard tuple concept, the concept of XML tree tuples is introduced, corresponding to relational databases concept of tuples.
+In a relational database, a correspondence between values $A$ and $B$ in the tuple of $D$ models the functional dependency denoted as $A \rightarrow B$.
+Since in XML there is no such standard tuple concept, the concept of XML tree tuples is introduced, corresponding to the concept of tuples in relational databases.
 
 \begin{define}[Tree Tuple]\label{treeTuple}
-Given an XML tree $XT$ conforming the DTD $D$, a tree tuple $t$ of $XT$ is a maximal sub-tree of $XT$ such that, for every path $p \in paths(D)$, $t.p$ contains at most one element.\qed
+Being given an XML tree $XT$ conforming to DTD $D$, a {\sl tree tuple} $t$ of $XT$ is a maximal sub-tree of $XT$ such that for every path $p \in paths(D)$, $t.p$ contains at most one element.\qed
 \end{define}
@@
 \begin{example}
-Consider the XML tree $XT$ of Fig. \ref{example1}. The subtrees of $XT$ shown in Fig. \ref{tuples} are tree tuples, and the subrees in Fig. \ref{nottuples} are not.
+Consider the XML tree $XT$ in Fig. \ref{example1}. The subtrees of $XT$ shown in Fig. \ref{tuples} are tree tuples, and the subrees in Fig. \ref{nottuples} are not.
 
@@
     \centering\includegraphics[width=\textwidth]{tuples}
-	\caption{Two tree tuples of the XML tree of Fig. \ref{example1}} \label{tuples}
+	\caption{Two tree tuples of the XML tree in Fig. \ref{example1}} \label{tuples}
 \end{figure}
@@
     \subfloat[]{\label{nottuple2}\includegraphics[scale=0.6]{nottuples2}}
-	\caption{Two subtrees of the XML tree of Fig. \ref{example1} which are not tree tuples}
+	\caption{Two subtrees of the XML tree in Fig. \ref{example1} which are not tree tuples}
     \label{nottuples}
@@
 
-In the Fig.
+In Fig.
 \ref{nottuple1}, the subtree is not a tree tuple, because the answer of the path $/bib/book/written\_by/author$ contains two distinct nodes (i.e.
@@
 The subtree in the Fig.
-\ref{nottuple2} is not a tree tuble because it is not a maximal subtree (it is a subtree of a tuple from Fig.
+\ref{nottuple2} is not a tree tuple, because it is not a maximal subtree (it is a subtree of a tuple from Fig.
 \ref{tuples}).\qed
@@
 \begin{define}[Functional Dependency]\label{fd1}
-Given a DTD $D$, a functional dependency on $D$ is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
+Being given a DTD $D$, a {\sl functional dependency} on $D$ is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
 \end{define}
@@
 \begin{example}\label{fdExample}
-Consider the XML tree of Fig \ref{example1}.
-The following functional dependency express the constraint that two distinct authors of the same book cannot have the same value of attribute \texttt{ano}:$$\{/bib/book, /bib/book/written\_by/author/@ano\} \rightarrow /bib/book/written\_by/author$$\qed
+Consider the XML tree in Fig \ref{example1}.
+The following functional dependency expresses the constraint that two distinct authors of the same book cannot have the same value of attribute \texttt{ano}:$$\{/bib/book, /bib/book/written\_by/author/@ano\} \rightarrow /bib/book/written\_by/author$$\qed
 \end{example}
@@
 \begin{define}[Functional Dependency]\label{fd2}
-With a given DTD $D$, a functional dependency is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$.
-Here $P$ is a  root path (path where first element is a root element of an XML document), or $P = \epsilon$ (empty path).
+With a given DTD $D$, a {\sl functional dependency} is of the form $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$.
+Here $P$ is a  root path (path where the first element is a root element of an XML document), or $P = \epsilon$ (empty path).
 Each $P_i (i \in [1,n])$ is a singleton leaf path, and there is a no non-empty common prefix for $P_1, \dots, P_{n+1}$.
-Given an XML document $T$ conforming to $D$, we say $T$ satisfies $\sigma$:iff $\forall v \in \{[\![P]\!]\}, \forall v_1, v_2 \in \{v[\![P']\!]\}$, if $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, then $v_1[\![P_{n+1}]\!] \equiv v_2[\![p_{n+1}]\!]$.\qed
+Being given an XML document $T$ conforming to $D$, we say $T$ satisfies $\sigma$:iff $\forall v \in \{[\![P]\!]\}, \forall v_1, v_2 \in \{v[\![P']\!]\}$, if $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, then $v_1[\![P_{n+1}]\!] \equiv v_2[\![p_{n+1}]\!]$.\qed
 \end{define}

### 2011-07-27 doc/theses/svirec/kap4.tex
corrections

@@
 
-The main goal of this thesis is to propose an algorithm repairing XML document violating functional dependencies defined for this document.
+The main goal of this thesis is to propose an algorithm to repair an XML document violating functional dependencies defined for this document.
 We use an XML tree and tree tuples to represent XML data.
-Next follows features we want to introduce in our algorithm:
+We would like our algorithm to have the following features::
 
 \begin{itemize}
-	\item Incorporation of weight model into the XML data representation to provide the selection of repair candidates with lowest modification cost.
-    \item Involvement of user in the process of finding and applying repair candidates.
-    \item Apply one repair candidate at the time and recalculate repair candidates again to prevent using unnecessary repairs to repair one violation and also to prevent of introducing new violations, which will not be repaired.
-    \item The paths which creates a functional dependency are described with XPath language \cite{xpath}, where only paths with basic construction are allowed (only path constructed with "/", "//" or "@", no wildcards, "[]" or another constructs are allowed).
-    \item We introduce a new concept of the so-called repair group, which clusters repair candidates (repair repairing one FD violation) repairing the same violation, or modifying the same part of the XML tree.
+	\item Incorporation of a weight model into the XML data representation to provide the selection of repair candidates with the lowest modification cost.
+    \item Involvement of a user in the process of finding and applying repair candidates.
+    \item Application of a single repair candidate at the time and then recalculate repair candidates again to prevent using unnecessary repairs to repair one violation and also to prevent of introducing new violations, which will not be repaired.
+    \item The paths which form a functional dependency are described with XPath language \cite{xpath}, where only paths with basic construction are allowed (only path constructed with "/", "//" or "@", no wildcards, "[]" or another constructs are allowed).
+    \item We introduce a new concept of the so-called repair group, which clusters repair candidates (repairs for repairing one FD violation) repairing the same violation, or modifying the same part of the XML tree.
 \end{itemize}
@@
 
-The proposed algorithm is based on the algorithm described in Section \ref{RepConstAnswers} presented in \cite{RepAndConsistentAnswer}.
+The proposed algorithm is based on the algorithm described in \ref{RepConstAnswers} presented in \cite{RepAndConsistentAnswer}.
 This algorithm was chosen because of simple representation of the XML data using a concept which corresponds with concept used in relational databases.
-Another reason is using besides modification of node value as an update operation also marking particular node information as unreliable, which can reveal forgotten inconsistencies in the data.
+Another reason, besides modification of node value as an update operation, is using also marking particular node information as unreliable, which can reveal forgotten inconsistencies in the data.
 
@@
 \REQUIRE{\ \\
-$RXT = \langle RT, \omega \rangle$: RW-XML tree conforming a DTD $D$\\
+$RXT = \langle RT, \omega \rangle$: RW-XML tree conforming to DTD $D$\\
 $\mathcal{FD} = {F_1, \dots, F_m}$: Set of functional dependencies}
@@
 
-The algorithm is split into three steps, where second and third step are repeated until all violations are repaired.
+The algorithm is split into three steps, where the second and the third step are repeated until all violations are repaired.
 In the first step, XML document and FDs are loaded, and XML tree with corresponding tree tuples are created.
 Next step of the algorithm computes repair groups containing repair candidates for FD violations.
-In the third step chosen repair candidate is applied to an XML tree.
-In Algorithm \ref{propAlgo} we can see the simplified process of repairing XML FD violations, which covers second and third step of the whole algorithm.
+In the third step the chosen repair candidate is applied to an XML tree.
+In Algorithm \ref{propAlgo} we can see the simplified process of repairing XML FD violations, which covers the second and the third step of the whole algorithm.
 
@@
 To represent an XML document we use an extended R-XML tree (Definition \ref{rxmlTree}), called RW-XML tree, which has weights assigned to each node of the tree.
-The weight of an node is from interval $[0,1]$ and indicates correctness of the data the particular node holds, that means the higher the weight is, more correct particular node is.
-The weights are used to measure cost of repair candidates, where candidate with the lowest cost is picked to be applied to the XML tree.
-This is also a first place where interaction of the user can be applied.
-The user could assign weights to the nodes manually or can use some sort of statistical methods to generate them automatically.
-If the user not assign weights nor use statistical methods, default weight is assigned to all nodes.
+The weight of a node is from interval $[0,1]$ and indicates correctness of the data the particular node holds, meaning the higher the weight is, the more correct that particular node is.
+The weights are used to measure the cost of repair candidates, where candidate with the lowest cost is picked to be applied to the XML tree.
+This is also the first place where some kind of user interaction can be implemented.
+The user could assign the weights to nodes manually or could use some sort of statistical methods to generate them automatically.
+If the user does not assign weights nor does he use statistical methods, a default weight is assigned to all nodes.
 A definition of an RW-XML tree follows:
@@
 
-After creating RW-XML tree from input XML data, set of tree tuples (defined in Definition \ref{treeTuple}) are constructed.
-Since a set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree.
-Since definition of a tree tuple says that answer of the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on constructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).
+After creating the RW-XML tree from input XML data, a set of tree tuples (defined in \ref{treeTuple}) is constructed.
+Since the set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree.
+Since definition of a tree tuple says that answer to the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on constructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).
 
-Functional dependencies as defined in Definition \ref{fd1} consists of paths described with XPath.
-As was described before, these paths can only have basic construction shown in Example \ref{pathsExample}.
+Functional dependencies as defined in Definition \ref{fd1} consist of paths described with XPath.
+As was described before, these paths can only have basic structure shown in Example \ref{pathsExample}.
 
@@
 
-With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now proceed to compute repair groups.
+With RW-XML tree $RWXT$ and corresponding tree tuples from input data created, we can now proceed to compute repair groups.
 Repair group is a set of repair candidates, which repairs the same FD or modifies the same part of $RWXT$.
-Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), where either modifies value of an RW-XML tree node ($\delta$ is defined) or marking a node as unreliable ($\varrho$ is defined).
-The $\delta$ function of repair candidate is defined same as in the XML tree (Definition \ref{xmlTree}) and defines new value of the RW-XML tree node.
-Similarly, $\varrho$ function defines the node which is marked as unreliable.
+Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), which either modifies value of an RW-XML tree node ($\delta$ is defined) or marks a node as unreliable ($\varrho$ is defined).
+The $\delta$ function of a repair candidate is defined the same way as in the XML tree (Definition \ref{xmlTree}) and defines a new value of the RW-XML tree node.
+Similarly, the $\varrho$ function defines the node which is marked as unreliable.
 
-To be able to compute repair groups, we need to decide for which FDs violates $RWXT$.
-This can be achieved by finding all tree tuple pairs that not weakly satisfies particular FD (Definition \ref{weakSatisf}).
-For each tuple pair is then computed repair group (line 6 in Algorithm \ref{propAlgo}).
+To be able to compute repair groups, we need to decide which FDs violate $RWXT$.
+This can be achieved by finding all tree tuple pairs that do not weakly satisfy particular FD (Definition \ref{weakSatisf}).
+A repair group is then computed for each tuple pair (line 6 in Algorithm \ref{propAlgo}).
 
-The function \texttt{computeRepairGroup}, responsible for creating the repair group, is shown in Algorithm \ref{repairGroupAlgo}.
-The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containing repair candidates as follows:
+The function \texttt{computeRepairGroup()} responsible for creating the repair group is shown in Algorithm \ref{repairGroupAlgo}.
+The function gets an RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containing repair candidates as follows:
 
 \begin{enumerate}
-	\item First repair candidates are created using function $computeRepairs$ from Algorithm \ref{repAlgo} (line 1).Since R-XML tree is a special case of RW-XML tree where all weights are equal to zero, we can pass RW-XML tree as a parameter to this function.
-    \item Next is checked if repair candidates intersect another candidates from existing repair groups (line 2)
+	\item First the repair candidates are created using function \texttt{computeRepairs()} from Algorithm \ref{repAlgo} (line 1).
+Since an R-XML tree is a special case of RW-XML tree where all the weights are equal to zero, we can pass RW-XML tree as a parameter to this function.
+    \item Next a check is performed whether the repair candidates intersect other candidates from existing repair groups (line 2)
     \begin{itemize}
     	\item If the candidates intersect with some group, they are added to this group (line 3)
-        \item Otherwise new repair group containing repair candidates is created (line 5)
+        \item Otherwise a new repair group containing repair candidates is created (line 5)
     \end{itemize}
@@
 \REQUIRE{\ \\
-$RXT = \langle T, \omega \rangle$: RW-XML tree conforming a DTD $D$\\
+$RXT = \langle T, \omega \rangle$: RW-XML tree conforming to DTD $D$\\
 $F: X \rightarrow p$ functional dependency\\
@@
 
-In Example \ref{repairExample} are shown repair candidates of one repair group modifying node values and marking nodes as unreliable.
+Example \ref{repairExample} demostrates repair candidates of one repair group modifying node values and marking nodes as unreliable.
 
@@
 The $XT$ violates FD $$\{/bib/book, /bib/book/written\_by/author/@ano\} \rightarrow /bib/book/written\_by/author$$ because two authors of the same book have the same value of attribute $@ano$.
-One repair group with these repair candidates is created: $R_1 = \langle \{\delta(v5) = \perp\},\varrho_{\{\}}(v) \rangle$, $R_2 = \langle \{\delta(v9) = \perp\},\varrho_{\{\}}(v) \rangle$, $R_3 = \langle \{\},\varrho_{\{v4,v5,v6,v7\}}(v) \rangle$, $R_4 = \langle \{\},\varrho_{\{v8,v9,v10,v11\}}(v) \rangle$, $R_5 = \langle \{\},\varrho_{\{v2,v3,\dots,v13\}}(v) \rangle$.
+One repair group with these repair candidates is created:
+\begin{itemize}
+	\item $R_1 = \langle \{\delta(v5) = \perp\},\varrho_{\{\}}(v) \rangle$
+    \item $R_2 = \langle \{\delta(v9) = \perp\},\varrho_{\{\}}(v) \rangle$
+    \item $R_3 = \langle \{\},\varrho_{\{v4,v5,v6,v7\}}(v) \rangle$
+    \item $R_4 = \langle \{\},\varrho_{\{v8,v9,v10,v11\}}(v) \rangle$
+    \item $R_5 = \langle \{\},\varrho_{\{v2,v3,\dots,v13\}}(v) \rangle$
+\end{itemize}
 
-First two repair candidates modifies the value of an attribute $@ano$ by assigning new generated value ($\perp$).
-Next two repair candidates mark each author node with its children nodes that has the same $@ano$ attribute respectively.
-The last candidate marks as unreliable whole subtree with the book node holding authors with the same attribute as root node.\qed
+First two repair candidates modify the value of an attribute $@ano$ by assigning newly generated value ($\perp$).
+Next two repair candidates mark each \emph{author} node with its child nodes that have the same $@ano$ attribute respectively.
+The last candidate marks as unreliable a whole subtree with the \emph{book} node holding \emph{authors} with the same attribute as the root node.\qed
 \end{example}
@@
 
-In this section, we introduce a weight model of a repair group, which will be used in the next step of our repair algorithm.
-Before we define a weight of repair group, let us denote some important notations.
+In this section, we introduce a weight model for a repair group, which will be used in the next step of our repair algorithm.
+Before we define a weight of a repair group, let us introduce some important notations.
 
-Given a repair candidate $R$, we denote with $Modified_\delta(R)$ the set of nodes modified by $\delta$ function from $R$.
+Being given a repair candidate $R$, $Modified_\delta(R)$ denotes the set of nodes modified by $\delta$ function from $R$.
 Analogously, we denote $Modified_\varrho(R)$ the set of nodes modified by $\varrho$ function.
 The count of nodes modified by repair candidate $R$ is denoted with $\lambda(R)$.
-At last, we denote $PS(R)$ a set of paths defining all nodes modified by $R$ (example of the set $PS(R)$ is shown in Example \ref{PSExample}).
+Last, we denote $PS(R)$ a set of paths defining all nodes modified by $R$ (example of the set $PS(R)$ is shown in Example \ref{PSExample}).
 
 \begin{example}\label{PSExample}
-Consider XML tree from Fig. \ref{example1} and a repair candidate, which marks nodes $v12$ and $v13$ as unreliable. Then the set $PS(R)$ contains following paths:
+Consider the XML tree from Fig. \ref{example1} and a repair candidate, which marks nodes $v12$ and $v13$ as unreliable. Then the set $PS(R)$ contains the following paths:
 \begin{verbatim}
@@
 
-Since each repair candidate consists of a set of nodes which are modified (marked as unreliable or has modified value), we can compute cost of each candidate.
-It is important to say, that in out approach, we prefer repair candidates that modify value of the nodes before those marking nodes as unreliable.
-Therefore we added the coefficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying node value will have lower cost.
-The definition of repair candidate cost follows:
+Since each repair candidate consists of a set of nodes which are modified (marked as unreliable or having a modified value), we can compute the cost of each candidate.
+It is important to say that in our approach we prefer repair candidates that modify the value of the nodes to those marking nodes as unreliable.
+Therefore we added a coefficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying the node value will have lower cost.
+The definition of repair candidate cost is as follows:
 
 \begin{define}[Repair Candidate Cost]
-Given an RW-XML tree $RXT$ and a repair candidate $R$, we define the cost of $R$ as:
-$$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \times k,$$
-where $k$ is coefficient determining a priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
+Being given an RW-XML tree $RXT$ and a repair candidate $R$, we define the cost of $R$ as:
+$$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \cdot k,$$
+where $k$ is the coefficient determining the priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
 \end{define}
 
-By default is $k$ set to value that cost of repair candidate marking unreliable node will be higher than the one that modifies node value.
-However this is another place, where can user intervene and change the priority of repair candidates.
-We show in Example \ref{koefK}, how $k$ can change costs of repair candidates.
+By default, $k$ is set to such value that the cost of repair candidate marking as an unreliable node will be higher than the one that modifies node value.
+However, this is another place where user can intervene and change the priority of repair candidates.
+We show in Example \ref{koefK} how $k$ can change the costs of repair candidates.
 
@@
 Consider XML tree $XT$ and repair candidates from Example \ref{repairExample}.
-A default weight of all nodes $v$ of $XT$ is $\omega(v) = 0.5$.
+The default weight of all nodes $v$ of $XT$ is $\omega(v) = 0.5$.
 If the coefficient $k = 1$, cost of repair candidates would be as follows: $cost(R_1) = 0.5$, $cost(R_2) = 0.5$, $cost(R_3) = 2$, $cost(R_4) = 2$ and $cost(R_5) = 6$.
-Let us assume, that a repair candidate with the lowest repair cost will be applied to $XT$.
-If we order costs of repair candidates from the lowest to the highest, you can see that candidates which have marked unreliable nodes are at the end (and will not be applied to $XT$).
+Let us assume that a repair candidate with the lowest repair cost will be applied to $XT$.
+If we order costs of repair candidates from the lowest to the highest, we can see that candidates which have marked unreliable nodes are in the end (and will not be applied to $XT$).
 
 However, if we set the coefficient $k = 0.1$, costs of candidates would be sorted as follows: $cost(R_3) = 0.2$, $cost(R_4) = 0.2$, $cost(R_1) = 0.5$, $cost(R_2)  = 0.5$ and $cost(R_5) = 0.6$.
-You can see that order of costs has significantly changed and the candidate which will be applied is $R_3$ (is marking nodes as unreliable).\qed
+We can see that order of costs has significantly changed and the candidate which will be applied is $R_3$ (marking nodes as unreliable).\qed
 \end{example}
 
-And finally, since a repair group is a set of repair candidates, its weight can be defined as follows:
+And, finally, since a repair group is a set of repair candidates, its weight can be defined as follows:
 
 \begin{define}[Repair Group weight]
-Given an RW-XML tree $RXT$ and a repair group $RG$, we define the weight of $RG$ as the sum of costs of all repair candidates in the $RG$.
+Being given an RW-XML tree $RXT$ and a repair group $RG$, we define the weight of $RG$ as the sum of costs of all repair candidates in the $RG$.
 \end{define}
@@
 
-The last step of our algorithm can be divided into two parts, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.
+The last step of our algorithm can be divided into two parts, namely selection of the repair candidate and the subsequent application of the candidate to RW-XML tree.
 
@@
 The previous step of the algorithm computes repair groups containing repair candidates, which can repair violations of provided FDs.
-From this candidates, we must choose one that is applied on RW-XML tree.
-Since we introduced a weight model to the repair group, we can use the weight of repair groups and the cost of repair candidates to choose suitable candidate.
-In our approach, we introduce two distinct algorithms: first that not involve the user to process of selection and the second one that use the user interaction.
+From these candidates, we must choose the one that is applied to the RW-XML tree.
+Since we introduced a weight model to the repair group, we can use the weight of repair groups and the cost of repair candidates to choose a suitable candidate.
+In our approach, we introduce two distinct algorithms: first that does not involve the user in the process of selection and the second one that uses the user interaction.
 
-The first algorithm simply selects the first repair group with the lowest weight, and from it the repair candidate with lowest cost is selected.
+The former algorithm simply selects the first repair group with the lowest weight, and from it the repair candidate with the lowest cost is selected.
 
-The latter algorithm allows user to choose the repair candidate which is most suitable for his needs.
-Very important aspect of all user interactions in this kind of algorithm is that the user will not be willing to select more than ten repairs.
-We could simply allow user to switch to the first selection algorithm, but that not take in account previous user selections of repair candidates.
-Therefore we introduce in this algorithm a functionality, that is able from selection done by the user before, guess his next selection.
+The latter algorithm allows the user to choose the repair candidate which is the most suitable for his needs.
+Very important aspect of all user interactions in this kind of algorithm is that the user will not be willing to select more than, e.g., ten repairs.
+We could simply allow user to switch to the first selection algorithm, but that does not take in account previous user selections of repair candidates.
+Therefore, we introduce in this algorithm a functionality able to guess his next selection from selection done by the user before.
 
-The function $selectRepairByUser(RGS, SR, t)$, responsible for selection of repair candidate involving user interactivity is shown in Algorithm \ref{selectUser}.
+The function \texttt{selectRepairByUser()} responsible for selection of repair candidate involving user interactivity is shown in Algorithm \ref{selectUser}.
 
-First is decided, if the algorithm is in a state that the user is selecting repair candidates (user selection mode), or the user leaves decision making on the algorithm using his previous selections (guess mode) (line 2).
-If we are still in user selection mode, the user choose from repair groups sorted by the weight the most convenient repair group, and from this group the user choose the repair candidate that will be applied on RW-XML tree (line 4).
-Repair candidates in each repair group are also sorted by the cost, working as a hint for the user which repair would be chosen by previous automatic method of selection.
-Last step of user selection mode is saving the information from selected repair candidate (line 5).
-This information consists of the FD which this candidate repairs, nodes the repair changes (their paths) and also if the change was a value modification or marking node as unreliable.
+First it decides, whether the algorithm is in a state where the user is selecting repair candidates (the user selection mode), or the user leaves decision making on the algorithm using his previous selections (the guess mode) (line 2).
+If we are still in user selection mode, the user chooses from repair groups sorted by the weight the most convenient one, and from this group the user chooses the repair candidate that will be applied on RW-XML tree (line 4).
+Repair candidates in each repair group are also sorted by the cost, working as a hint for the user which repair would be chosen by the previous automatic method of selection.
+The last step of user selection mode is saving the information from selected repair candidate (line 5).
+This information consists of the FD which this candidate repairs, nodes the repair changes (their paths) and also whether the change was a value modification or marking a node as unreliable.
 
-If the user has decided not to selecting repair candidates by himself, the algorithm goes into guess mode (staring on line 6).
-In this mode the algorithm checks all repair groups if some of it contains the repair candidate, that is sufficiently similar with some previously selected repair candidate (line 7-10).
-This similarity is checked by function \texttt{canBeUsedUserSelection} shown in Algorithm \ref{canUserSel}.
-If neither of candidates is sufficiently similar, the algorithm choose repair candidate with the lowest cost from the repair group with the lowest weight (line 15-16).
+If the user has decided not to select repair candidates by himself anymore, the algorithm goes into the guess mode (starting at line 6).
+In this mode the algorithm checks all repair groups whether one of them contains a repair candidate that is sufficiently similar to some previously selected repair candidate (line 7-10).
+This similarity is checked by function \texttt{canBeUsedUserSelection()} shown in Algorithm \ref{canUserSel}.
+If neither of the candidates is sufficiently similar, the algorithm chooses the repair candidate with the lowest cost from the repair group with the lowest weight (line 15-16).
 
@@
 $RGS$ set of repair groups\\
-$SR$: set of previously selected repair candidates by the user\\
+$SR$: the set of repair candidates previously selected by the user\\
 $t$: modified nodes count threshold of previously selected repair candidates\\}
@@
 
-The function \texttt{canBeUsedUserSelection} gets actual repair candidate $RC$, set of all previously selected repair candidates $SR$, modified nodes count threshold $t$ from interval $(0,1]$ of previously selected repair candidate and determine if $RC$ is sufficiently similar as some repair candidate from $SR$.
-To be similar with some previously selected repair candidate $S$, $RC$ must repair the same $FD$ as $S$, it must use the same update operation as $S$ and $\lambda(RC) \geq \lceil\lambda(S) \times t\rceil$ (line 2).
-Furthermore, if $\lambda(RC) = \lceil\lambda(S) \times t\rceil$, set of paths $PS(RC)$ must be a subset of $PS(S)$ (line 3).
-Otherwise, if $\lambda(RC) > \lceil\lambda(S) \times t\rceil$, there must exists a subset of $PS(S)$ with $\lceil\lambda(S) \times t\rceil$ elements that is subset of $PS(RC)$.
-In other words, without taking into account the threshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is such that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes).
-The threshold $t$ can reduce count of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$.
-In Example \ref{tresholdExample} is shown the usage of the threshold $t$.
+The function \texttt{canBeUsedUserSelection()} gets the current repair candidate $RC$, the set of all previously selected repair candidates $SR$, and the modified nodes count threshold $t$ from interval $(0,1]$ of previously selected repair candidate and it determines whether the $RC$ is sufficiently similar to some repair candidate from $SR$.
+To be similar with some previously selected repair candidate $S$, $RC$ must repair the same $FD$ as $S$, it must use the same update operation as $S$ and $\lambda(RC) \geq \lceil\lambda(S) \cdot t\rceil$ (line 2).
+Furthermore, if $\lambda(RC) = \lceil\lambda(S) \cdot t\rceil$, the set of paths $PS(RC)$ must be a subset of $PS(S)$ (line 3).
+Otherwise, if $\lambda(RC) > \lceil\lambda(S) \cdot t\rceil$, there must exist a subset of $PS(S)$ with $\lceil\lambda(S) \cdot t\rceil$ elements that is a subset of $PS(RC)$.
+In other words, without taking into account the threshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is the one that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes).
+The threshold $t$ can reduce the number of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$.
+In Example \ref{tresholdExample} the usage of the threshold $t$ is shown.
 
@@
 \REQUIRE{\ \\
-$SR$: set of previously selected repair candidates by the user\\
-$t$: suitability threshold $(0,1]$ of previously selected repair candidates\\
-$RC$: actual repair candidate}
-\ENSURE \TRUE\ if actual repair candidate is similar to some previous one.
+$SR$: the set of previously selected repair candidates by the user\\
+$t$: a suitability threshold $(0,1]$ of previously selected repair candidates\\
+$RC$: the current repair candidate}
+\ENSURE \TRUE\ if $RC$ is similar to some previous repair candidate.
 \FORALL{$S$ in $SR$}
@@
 Consider repair candidate $R_1$ selected by user, that marks as unreliable two nodes and $PS(R_1) = \{/bib/book/title, /bib/book/title/text()\}$.
-Next consider repair candidate $R_2$ with $PS(R_2) = \{/bib/book/title/text()\}$, that the modified node is marked as unreliable and both $R_1$ and $R_2$ repairs the same $FD$.
-If threshold $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$.
-However, if threshold is $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
+Next, consider repair candidate $R_2$ with $PS(R_2) = \{/bib/book/title/text()\}$, that the modified node is marked as unreliable and both $R_1$ and $R_2$ repairs the same $FD$.
+If $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$.
+However, if $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
 \end{example}
@@
 In this part, the selected repair candidate is finally applied to the RW-XML tree.
-If after this part the RW-XML tree not violates any of FDs, the whole repair algorithm ends at this point, otherwise repair groups are regenerated and the selection of the repair candidate part take place.
+If after this part the RW-XML tree does not violate any FDs, the whole repair algorithm ends at this point, otherwise repair groups are regenerated and the selection of the repair candidate part takes place again.
 
-To apply the selected repair candidate $R$ to the RW-XML tree $RXT$ means, that we compose $\delta$ and $\varrho$ functions of $R$ with the corresponding functions of R-XML tree contained in $RXT$.
-This compositions are defined in Definition \ref{compDelta} and \ref{compVarrho}.
+To apply the selected repair candidate $R$ to the RW-XML tree $RXT$ means to compose $\delta$ and $\varrho$ functions of $R$ with the corresponding functions of R-XML tree contained in $RXT$.
+These compositions are defined in Definition \ref{compDelta} and \ref{compVarrho}.
 
-After application of the repair candidate, some of RW-XML tree nodes could become unreliable, which can lead to the situation that some of tree tuples are not anymore considered as a tuple (is no longer a maximal subtree).
-Therefore, we need before regeneration of repair groups also check all tree tuples, if they satisfies definition of tree tuple.
+After application of the repair candidate, some of RW-XML tree nodes could become unreliable, which can lead to the situation that some of tree tuples are not anymore considered a tuple (it is no longer a maximal subtree).
+Therefore, before regeneration of repair groups we need to check all tree tuples to see whether they satisfy definition of tree tuple.

### 2011-07-27 doc/theses/svirec/kap6.tex
corrections

@@
 
-All provided documents with functional dependencies and also repaired documents are placed on CD in the directory \texttt{/datasets/}.
+All provided documents with functional dependencies and also repaired documents are placed on the enclosed CD in the directory \texttt{/datasets/}.
 
@@
 
-The first real-world dataset originate from the XML data repository \\(\url{http://www.cs.washington.edu/research/xmldatasets/}), concretely Course data derived from university websites.
-The course data consists, beside of other information, of day of the week, time and place (building and room), where each course is situated.
-The consistency of data is evaluated against FD defined as: Two courses starting at the same date and time are each situated in a different place.\\
+The first real-world dataset originates from the XML data repository \\(\url{http://www.cs.washington.edu/research/xmldatasets/}), specifically the Course data derived from university websites.
+The course data consist, along with other information, of day of the week, time and place (building and room), where each course is situated.
+The consistency of data is evaluated against FD defined as: \emph{Two courses starting at the same date and time are each situated in a different place}.\\
 
 The second dataset is a set of actors of IMDB database obtained from the Niagara data source (\url{http://www.cs.wisc.edu/niagara/data.html}).
-The dataset consists of a set of authors, where for each author is listed a set of movies he played in.
+The dataset consists of a set of authors, where for each author a set of movies he played in is listed.
 For this dataset, we defined FD as follows: For each two authors, which played in the movie with the same name, the year of release of this movie must be the same.
@@
 
-Two synthetic data sets have been constructed, the first represents data introduced in Example \ref{example1} with FD presented in Example \ref{fdExample}.
-The latter one is created according to Example 1 introduced in \cite{ImprovingXML}.
+Two synthetic data sets have been constructed, the former represents data introduced in Example \ref{example1} with FD presented in Example \ref{fdExample}.
+The latter is created according to Example 1 introduced in \cite{ImprovingXML}.
 
@@
 
-For each datasets, we first run the former algorithm and collect from a repaired document informations about how many nodes have been modified with particular update operation (we split value modification operation into changing value to new one and copying one value from one node to another).
+For each dataset, we first run the original algorithm and collect from a repaired document informations about how many nodes have been modified with particular update operation (we split value modification operation into changing value to new one and copying one value from one node to another).
 
-Thereafter we use our proposed algorithm with minimal repair candidate selection mode, and besides the data we collect for former algorithm, we also gather informations about how many repair groups have been created before first application of repair candidate and the total count of picked repair groups.
-We run our algorithm multiple times with different value of the coefficient $k$ and assigning of node weights $w$.\\
+After that, we use our proposed algorithm with minimal repair candidate selection mode, and along with the data we collect for original algorithm, we also gather information on how many repair groups have been created before the first application of repair candidate and the total number of picked repair groups.
+We run our algorithm multiple times with different value of the coefficient $k$ and assignment of node weights $w$.\\
 
@@
 For each document, we execute our algorithm with three different settings.
-First two uses default value of $k$ ($k=1.5$) and for the last one, we set $k=0.005$.
-For the second (resp.
-third) execution, we set a weight $w$ of nodes represented by XPath \texttt{/root/course/time/start\_time/text()} (resp.
+First two use the default value of $k$ ($k=1.5$), for the last one, we set $k=0.005$.
+For the second (third) execution, we set the weight $w$ of nodes represented by XPath \texttt{/root/course/time/start\_time/text()} (resp.
 \texttt{/root/course/time/\discretionary{}{}{}start/text()}) to 0.1.
@@
 The document \texttt{actors.xml} represents the second real-world dataset.
-Our algorithm is executed two times, where for both executions have been set $k=1.5$.
+Our algorithm is executed two times, where for both executions $k=1.5$ is set.
 Moreover, the second has weights of nodes represented by XPath \texttt{/W4F\_DOC/\discretionary{}{}{}Actor/Filmography/Movie/Year/text()} set to $w=0.2$.\\
 
-Generated datasets represented by \texttt{bib.xml} and \texttt{customers.xml} documents have been both used as a input for our algorithm two times, where first execution was done with default value of $k$.
+Generated datasets represented by \texttt{bib.xml} and \texttt{customers.xml} documents have been both used as a input for our algorithm two times, where first execution was done with the default value of $k$.
 The second execution on \texttt{bib.xml} uses $k=0.1$ and on \texttt{customers.xml} uses $k=1.5$ and $w=0.2$ is set to all nodes represented by \texttt{/customers/country/c\_list/customer/city/text()}.
@@
 
-All gathered informations from real-world and synthetic datasets are shown in Tables \ref{realWorldTable} and \ref{syntheticTable}.
-The columns of both tables are defined as follows: RG is count of repair groups created before application of repair candidate, RG total is total count of picked repair groups, U specifies number of nodes marked as unreliable, NV defines number of nodes changing value to a new one and finally ChV specifies number of nodes with value copied from another node.\\
+All information gathered from real-world and synthetic datasets is shown in Tables \ref{realWorldTable} and \ref{syntheticTable}.
+The columns of both tables are defined as follows: \emph{RG} is the count of repair groups created before application of repair candidate, \emph{RG total} is the total count of picked repair groups, \emph{U} specifies the number of nodes marked as unreliable, \emph{NV} defines the number of nodes changing value to a new one and finally \emph{ChV} specifies the number of nodes with value copied from another node.\\
 
@@
     \hline
-    & repairer & $k$ & $w$ & RG & RG total & U & NV & ChV\\ \hline
+    dataset & repairer & $k$ & $w$ & RG & RG total & U & NV & ChV\\ \hline
     \multirow{4}{*}{reed.xml} & old repairer & - & - & - & - & 0 & 780 & 0\\
@@
 
-From the informations in tables is visible, that our approach found for each dataset the repair, which modifies less nodes than old repair algorithm.
-One exception, where our algorithm modifies more nodes is the case, when we set value of $k$ near 0, which marks nodes as unreliable.
-With this setting, we want to demonstrate, that with our approach is possible to choose instead of modifying node values, marking nodes as unreliable.
-This is not possible with old repairer, since it prefers node value modification before marking node as unreliable.
+From the information presented in these tables is apparent that our approach found for each dataset a repair that modifies less nodes than the original repair algorithm.
+One exception where our algorithm modifies more nodes is the case when we set value of $k$ near 0, which marks nodes as unreliable.
+With this setting we want to demonstrate that with our approach it is possible to choose instead of modifying node values, marking nodes as unreliable.
+This is not possible with the original repairer, since it prefers node value modification to marking node as unreliable.
 
@@
     \hline
-    & repairer & $k$ & $w$ & RG & RG total & U & NV & ChV\\ \hline
+    dataset & repairer & $k$ & $w$ & RG & RG total & U & NV & ChV\\ \hline
     \multirow{3}{*}{bib.xml} & old repairer & - & - & - & - & 0 & 913 & 0\\
@@
 
-With setting the weight to particular nodes, we want to show that our algorithm is capable of choosing the priority of nodes, which values are modified (\texttt{reed.xml} and \texttt{wsu.xml}), or if the value is changed to new value (\texttt{actors.xml} and \texttt{customers.xml}).
+With setting the weight to particular nodes we want to show that our algorithm is capable of choosing the priority of nodes, which values are modified (\texttt{reed.xml} and \texttt{wsu.xml}), or if the value is changed to new value (\texttt{actors.xml} and \texttt{customers.xml}).
 
 Another interesting observation is that for almost all datasets, our algorithm picks less repair groups (from which picks repair candidate to apply to document) than were created before first pick.
-It means that some repair candidate repairs more than one FD violation and therefore less value modifications were needed.
-In the case, where the number of picked repair groups outreach the firstly created ones, is visible that some repair candidate introduces new violation which was consequently repaired in the next iteration of our algorithm.
+It means that one of the repair candidate repairs more than one FD violation and therefore less value modifications are needed.
+In the case where the number of picked repair groups outreach the firstly created ones, is visible that some repair candidate introduces new violation which was consequently repaired in the next iteration of our algorithm.

### 2011-07-27 doc/theses/svirec/kap7.tex
corrections

--- 2011-07-16 doc/theses/svirec/kap7.tex
@@
 
-The aim of this thesis was a proposal and implementation of algorithm repairing XML functional dependencies violations. At first, existing solutions were analyzed and described.
+The aim of this thesis was to propose and implement an algorithm repairing XML functional dependencies violations.
+At first, existing solutions were analyzed and described.
+The result of the analyses was that none of the recent approaches uses, even in only limited form, user interaction in the process of finding suitable repair for the FD violation.
+Consequently, we proposed algorithm based on \cite{RepAndConsistentAnswer} that uses user interaction in various ways.\\
+
+First, we incorporated a weight model into the XML data representation, which allows us to measure the modification cost of each possible repair candidate.
+This is also the first place for the user to interact with the repair process.
+Next, we have introduced a new concept of repair candidates clusters called repair groups.
+These clusters group repair candidates according to the violation which they are repairing or part of the XML data they are modifying.
+
+The main user interaction part of our algorithm is the selection of the repair group and consequently the repair candidate which is applied to the XML document.
+Besides the repair candidate selection we also introduced a mechanism allowing to guess the next selection based on the previous candidates selected by the user, if the user does not want to select the repair candidate anymore.
+
+The experimental implementation of the original as well as proposed algorithms based on the jInfer framework.
+Both algorithms have been compared against each other on synthetic as well as real-world datasets.
+From the data gathered from the comparison it is clear that our approach has found repairs with less modifications applied on the XML data.
+We have also shown that with the user interaction it is possible to change the usage of update operations used in repair, which may create more reasonable result for the user.
 
 \section{Future Work}
+
+Although our approach has introduced user in the process of repairing FD violations in XML data, there is still work to be done.
+The first main task in the further work is to extend paths defining functional dependencies.
+With usage of more constructs that XPath provides one can define more sophisticated FDs.
+
+Another part of the algorithm to be improved is clustering repair candidates into repair groups. The user can provide some additional criteria, that can change the resulting repair.
+
+Last but not least, guessing part of the user selection algorithm can be improved by using more sophisticated heuristic algorithm to select proper repair candidate with regard to previous candidates selected by the user.

### 2011-07-27 doc/theses/svirec/kap3.tex
corrections

@@
 
-This chapter introduces description and categorization of recent approaches of repairing XML documents with usage of functional dependencies.
+This chapter introduces the description and categorization of recent approaches to repairing XML documents that violate functional dependencies defined in these documents.
 
-All approaches dealing with the problem of finding optimal repair can be divided into categories according to usage of elementary repair primitives.
-This repair primitives are: inserting node, deleting node, updating node and marking node as unreliable of XML document.
+All approaches dealing with the problem of finding optimal repair can be divided into categories according to the usage of elementary repair primitives.
+This repair primitives are: inserting node, deleting node, updating node and marking node as unreliable in XML document.
 
@@
 
-To be able to resolve problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependencies based on those defined for relational databases.
-Concept of a tree tuple and functional dependency is defined in Definition \ref{treeTuple} and \ref{fd1}.
+To be able to resolve the problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependencies based on those defined for relational databases.
+The concept of a tree tuple and functional dependency was introduced in definition \ref{treeTuple} and \ref{fd1}.
 
-Given an XML tree $XT$ conforming a DTD $D$ and a functional dependency $F : S_1 \rightarrow S_2$ , we say that $XT$ satisfies $F (XT \models F )$ if for each pair of tree tuples $t_1, t_2$ of $XT$, $t_1.S_1 = t_2.S_1 \land t_1.S_1 = \emptyset \Rightarrow t_1.S_2 = t_2.S_2$ .
-Given a set of functional dependencies $\mathcal{FD} = \{F_1 , \dots, F_n\}$ over $D$, we say that $XT$ satisfies $\cal FD$ if it satisfies $F_i$ for every $i \in 1..n$.
+Being given an XML tree $XT$ conforming a DTD $D$ and a functional dependency $F : S_1 \rightarrow S_2$ , we say that $XT$ satisfies $F (XT \models F )$ if for each pair of tree tuples $t_1, t_2$ of $XT$, $$t_1.S_1 = t_2.S_1 \land t_1.S_1 = \emptyset \Rightarrow t_1.S_2 = t_2.S_2$$
+Being given a set of functional dependencies $\mathcal{FD} = \{F_1 , \dots, F_n\}$ over $D$, we say that $XT$ satisfies $\cal FD$ if it satisfies $F_i$ for every $i \in 1..n$.
 
@@
 
-Authors of this approach chose two kinds of actions to repair inconsistent XML data with regard to functional dependencies.
-The first action is updating the value of an attribute or content of an element.
-As the second action authors choose marking inconsistent element as "unreliable" rather than deleting it, because removing elements from an XML document leads to some undesired drawbacks: it does not always suffice to remove inconsistency and deleting a node can lead to a new document not valid against the given schema.
+The authors of this approach choose two kinds of actions to repair inconsistent XML data with regard to functional dependencies.
+The first action is updating the value of an attribute or the content of an element.
+As the second action the authors choose marking inconsistent element as ``unreliable" rather than deleting it, because removing elements from an XML document leads to some undesired drawbacks: it does not always suffice to remove inconsistency and deleting a node can lead to a new document not valid against the original schema.
 
 Depending on XML data and defined functional dependencies, each inconsistency could have many possible strategies to repair it.
-From all the possible repair strategies, the authors prefer those, for which smaller changes are made to the original document.
+From all the possible repair strategies the authors prefer those for which smaller changes are made to the original document.
 
 \begin{example}\label{fdrepairExample}
-Consider XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$.
+Consider the XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing a collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$ ($@$ denotes the attribute of an element).
 \begin{verbatim}
@@
 \end{verbatim}
-The functional dependency defined above requires that for each book, there is only one element author having a given $@ano$ value.
+The functional dependency defined above requires that for each book there is only one element author having a given $@ano$ value.
 Therefore $XT$ does not satisfy the given functional dependency, because two author elements of the same book have the same value of attribute $@ano$.
-To resolve this data inconsistency, we can use two different repair strategies: 1) changing one of the value of attribute $@ano$; 2) marking one of the element author as unreliable.
-Since the first strategy changes only attribute $@ano$, it is preferred to second strategy, which changes a larger portion of document, since it marks whole author element as unreliable.
+To resolve this data inconsistency, we can use two different repair strategies: 1) changing one of the value of attribute $@ano$; 2) marking one of the elements author as unreliable.
+Since the first strategy changes only attribute $@ano$, it is preferred to the second strategy, which changes a larger portion of document, since it marks a whole author element as unreliable.
 \qed
@@
 
-Before introducing the algorithm to repair inconsistent XML document, let us define reliability of elements in XML tree:
+Before introducing the algorithm to repair an inconsistent XML document, let us define the notion of reliability of elements in an XML tree:
 
 \begin{define}[R-XML Tree]\label{rxmlTree}
-A R-XML $tree$ is a triplet $RXT = \langle T, \delta, \varrho \rangle$, where $\langle T, \delta \rangle$ is an XML tree and $\varrho$ is a reliability function from $N_T$ to \texttt{\{true, false\}}, such that, for each pair of nodes $n_1 , n_2 \in N_T$ with $n_2$ descendant of $n_1$, it holds that $\varrho(n_1) = false \Rightarrow \varrho(n_2) = false$.\qed
+A R-XML $tree$ is a triplet $RXT = \langle T, \delta, \varrho \rangle$, where $\langle T, \delta \rangle$ is an XML tree and $\varrho$ is a reliability function from $N_T$ to \texttt{\{true, false\}}, such that for each pair of nodes $n_1 , n_2 \in N_T$ with $n_2$ descendant of $n_1$, it holds that $\varrho(n_1) = false \Rightarrow \varrho(n_2) = false$.\qed
 \end{define}
@@
 
-The repair of an R-XML tree which not satisfies $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$).
+The repair of an R-XML tree which does not satisfy $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$).
 The composition of the R-XML tree and the repair (i.e.
@@
 \begin{define}[Composition of $\varrho$ functions]\label{compVarrho}
-The composition of two reliability functions $\varrho_1$ and $\varrho_2$ asssociating boolean value to nodes is $$\varrho_1 \cdot \varrho_2(n) = min(\varrho_1(n), \varrho_2(n)),$$ where to the boolean value $true$ (resp.
-$false$) is associated value 1 (resp.
-0).
-\qed
+The composition of two reliability functions $\varrho_1$ and $\varrho_2$ asssociating a boolean value to nodes is $$\varrho_1 \cdot \varrho_2(n) = \varrho_1(n)\ AND\ \varrho_2(n).$$\qed
 \end{define}
 
-With a repair $\langle \delta, \varrho \rangle$ of R-XML tree and a set of labelled nodes $N$ of this tree, we denote with $Update_{\delta}(N)$ the set of nodes modified by $\delta$.
+With a repair $\langle \delta, \varrho \rangle$ of R-XML tree and a set of labelled nodes $N$ of this tree, we denote  $Updated_{\delta}(N)$ the set of nodes modified by $\delta$.
 Analogously, we denote $True_{\varrho}(N) = \{n \in N| \varrho(n) = true\}$ and $False_{\varrho}(N) =\{n \in N | \varrho(n) = false\}$.
@@
 An R-XML tree is used as an input for the main algorithm computing repaired R-XML tree described in Algorithm \ref{repAlgo}.
-First the algorithm computes all the possible repairs of tuples which not satisfy a functional dependency using the function \texttt{computeRepairs} (lines 2-6).
-Next all non-minimal repairs are removed from all possible repairs (line 7).
+First, the algorithm computes all the possible repairs of tuples which do not satisfy a functional dependency using the function \texttt{computeRepairs()} (lines 2-6).
+Next, all non-minimal repairs are removed from all possible repairs (line 7).
 In the last step, all the repairs are merged and a unique repaired R-XML tree is returned.
@@
 
-Function \texttt{computeRepairs}(in algorithm \ref{computeRepairs}) gets R-XML tree, a functional dependency $F$ and a tuples $t_1, t_2$ of R-XML tree as input and computes the repair as follows:
+Function \texttt{computeRepairs()} (in Algorithm \ref{computeRepairs}) gets an R-XML tree, a functional dependency $F$ and tuples $t_1, t_2$ of the R-XML tree as input and computes the repair as follows:
 \begin{itemize}
@@
 	\item Otherwise $p$ denotes a node, so either the node $t_1.p$ or $t_2.p$ is marked as unreliable (line 5).
-	\item For each path $p_i$ of the left side of a functional dependency $F$
+	\item For each path $p_i$ on the left side of a functional dependency $F$
 	\begin{itemize}
-		\item If path $p_i$ denotes a textual element, then one of the two terminal values $t_1.p_i$ or $t_2.p_i$ is changed to the new generated value ($\perp$) (line 9).
-		\item Otherwise $p_i$ denotes a node, so one of the nodes $t_1.p_i$ or $t_2.p_i$ is marked as unreliable (line 11).
+		\item If path $p_i$ denotes a textual element, then one of the two terminal values $t_1.p_i$ or $t_2.p_i$ is changed to the newly generated value ($\perp$) (line 9).
+		\item Otherwise $p_i$ denotes a node, therefore one of the nodes $t_1.p_i$ or $t_2.p_i$ is marked as unreliable (line 11).
 	\end{itemize}
@@
 \REQUIRE{\ \\
-$RXT = \langle T, \delta \varrho \rangle$: R-XML tree conforming a DTD $D$\\
+$RXT = \langle T, \delta \varrho \rangle$: R-XML tree conforming to DTD $D$\\
 $F: X \rightarrow p$ functional dependency\\
@@
 The authors proposed a technique for repairing XML documents violating functional dependencies based on approaches proposed for relational database repairing.
-The algorithm introduces two possible repair primitives, which creates many possible result from which those with minimal impact on document are chosen.
-However the authors not consider creation of new violations after repairing the initial violations as it is in \cite{ImprovingXML}.
-Another disadvantage of this approach is that a unnecessary repair of some particular violation could be applied to an XML document because another repair could repair that violation before.
+The algorithm introduces two possible repair primitives, which create many possible results from which those with minimal impact on the document are chosen.
+However the authors do not consider creation of new violations after repairing the initial violations as it is in \cite{ImprovingXML}.
+Another disadvantage of this approach is that an unnecessary repair of some particular violation could be applied to an XML document because another repair could repair that violation before.
 
@@
 Studying the problem of repairing inconsistent XML documents with respect to a set of functional dependencies and investigating the existence of repairs has been introduced in \cite{QueryXML}.
-Authors introduce two kind of repair primitives.
-The first is deleting "unreliable" nodes of document and the second is inserting a new nodes.
-Similarly to another approaches, authors prefer minimal set of repair primitives applied to the XML document to form a repair.
+The authors introduce two kinds of repair primitives.
+The first one is deleting ``unreliable" nodes of document, the second one is inserting new nodes.
+Similarly to other approaches, authors prefer minimal set of repair primitives applied to the XML document to form a repair.
 
@@
 
-With the insert and delete operations as repair primitives, is updated the structure of the XML document which conforms DTD $D$ (defined in Definition \ref{dtdDef}) and violated functional dependency.
-The insert operation is represented as $\langle  + [x]a[y], z\rangle$, where i) $x$ is a node identifier, ii) $a$ is a label, iii) $y$ is either a node identifier or a value ($y$ is a value, if $a \in \alpha \cup \{S\}$; otherwise it is a node identifier), and iv) $z$ denotes the child of $x$ which must immediately precede $y$ ($\perp$ if $y$ is inserted as the first or a single child of $x$).
+With the insert and delete operations as repair primitives, the structure of the XML document which conforms DTD $D$ (defined in Definition \ref{dtdDef}) and violated functional dependency is updated.
+The insert operation is denoted as $\langle  + [x]a[y], z\rangle$, where:
+\begin{enumerate}
+\renewcommand{\labelenumi}{\roman{enumi})}
+	\item $x$ is a node identifier
+    \item $a$ is a label
+    \item $y$ is either a node identifier or a value ($y$ is a value, if $a \in \alpha \cup \{S\}$; otherwise it is a node identifier)
+    \item $z$ denotes the child of $x$ which must immediately precede $y$ ($\perp$ if $y$ is inserted as the first or a single child of $x$).
+\end{enumerate}
+
 The deletion is represented as $-[x]a[y]$, where $x$ is a node identifier, $a \in \alpha \cup \tau \cup \{S\}$, and $y$ is either a node $id$ or a string value, denoting the node to be deleted.\\
-The set $R$ of update operations can be divide into two subsets: $R^+$ is a the subset of all the insertion operations in $R$; $R^-$ is the subset of all the deletion operations.
-A set $R$ is said to be consistent if the following conditions are hold:
+The set $R$ of update operations can be divided into two subsets: $R^+$ is the subset of all the insertion operations in $R$; $R^-$ is the subset of all the deletion operations.
+A set $R$ is said to be consistent if the following conditions hold:
 \begin{enumerate}
@@
 
-With defined basic repair operations for general repair of inconsistent XML document let us define a repair as follows:
+With basic repair operations for general repair of inconsistent XML document defined let us define a repair as follows:
 
 \begin{define}[Repair]
-Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$(definition \ref{integConstr}), a set of update operations $R$ is said to be a repair of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
+Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$ (Definition \ref{integConstr}), a set of update operations $R$ is said to be a repair of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
 \end{define}
@@
 
-In this approach the authors introduce different types of repairs (general, cleaning and completing) and focus on checking whether exists a repair for many classes of integrity constraints (general integrity constraints, inclusion dependencies, functional dependencies, etc.).
+In this approach the authors introduce different types of repairs (general, cleaning and completing) and focus on checking whether there exists a repair for many classes of integrity constraints (general integrity constraints, inclusion dependencies, functional dependencies, etc.).
 For the functional dependencies the authors found out that the problem of checking whether there exists a general repair for an XML document is $\mathcal{NP}$-complete.
@@
 
-The algorithm for repairing XML functional dependency violations which uses modification of node value as the repair primitive has been proposed in \cite{ImprovingXML}.
-For finding the optimal repair of an XML document, the authors introduce a cost model, which assign a weight to each leaf node in an XML document.
+The algorithm for repairing XML functional dependency violations which uses a modification of node value as the repair primitive has been proposed in \cite{ImprovingXML}.
+To find the optimal repair of an XML document, the authors introduce a cost model, which assigns a weight to each leaf node in an XML document.
 The optimal repair is the one with the lowest repair cost which is measured by the total weight of the modified nodes.
 
-The authors of the approach found out that repairing one functional dependency violation can violate another.
+The authors of this approach found out that repairing one functional dependency violation can violate another.
 Therefore they divide the main algorithm into two phases.
-In the first phase, the conflict hypergraph capturing the initial functional dependency violations is constructed, and all the violations are fixed by modifying the values of all the nodes on a vertex cover of the conflict hypergraph.
+In the first phase, the conflict hypergraph capturing the initial functional dependency violations is constructed and all the violations are fixed by modifying the values of all the nodes on a vertex cover of the conflict hypergraph.
 In the second phase, remaining violations are resolved by modifying the violating nodes and their core determinants to prevent of introducing new conflicts.
@@
 
-Each leaf node $v$ of XML tree is associated a weight from range $[0,1]$, which is denoted $W(v)$.
-Let us assume that the larger the weight of the leaf the more reliable it is.
-The weight may be automatically generated by statistical methods or it can be assigned by a user.
+Each leaf node $v$ of XML tree is associated with a weight from range $[0,1]$, which is denoted $W(v)$.
+Let us assume that the larger the weight of the leaf, the more reliable it is.
+The weight may be automatically generated by statistical methods or it can be assigned by the user.
 
 As was mentioned earlier, a repairing primitive is a node value modification, where for repairing algorithm a combination of two rules to resolve violation is used.
-Let us have a $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ (from Definition \ref{fd2}) and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
+Let us have an $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ (from definition \ref{fd2}) and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
 If the child nodes of $v_1$ and $v_2$ qualified by paths $P_i$ have equal values for all $i \in [1,n]$ and their child nodes qualified by $P_{n+1}$ have different values, then $v_1$ and $v_2$ violates $\sigma$.
-The first rule used to repair this violation is to change the value of the node qualified by $P_{n+1}$ from $v_1$ to the value of $v_2$'s child node that matches $P_{n+1}$(or reversely).
-The second rule is to choose an arbitrary $P_i (i \in [1,n])$, and introduce a new value to the node qualified by $P_i$ from $v_1$ (or $v_2$).
+The first rule used to repair this violation is to change the value of the node qualified by $P_{n+1}$ from $v_1$ to the value of $v_2$'s child node that matches $P_{n+1}$ (or reversely).
+The second rule is to choose an arbitrary $P_i (i \in [1,n])$ and introduce a new value to the node qualified by $P_i$ from $v_1$ (or $v_2$).
 
 \begin{define}[Optimal repair]
-Given an inconsistent XML document $T$ violating a set $\Sigma$ of $FD$s, the repair $T_R$ of $T$ is called optimal repair, if $T_R$ has the minimum cost among all repairs of $T$.
+Being given an inconsistent XML document $T$ violating a set $\Sigma$ of $FD$s, the repair $T_R$ of $T$ is called {\sl optimal repair}, if $T_R$ has the minimum cost among all repairs of $T$.
 The cost $cost(T_R)$ is defined as:
@@
 
-A weighted hypergraph is used in the first part of repair algorithm as a tool modeling initial functional dependency violations in an XML document.
+A weighted hypergraph is used in the first part of the repair algorithm as a tool modeling initial functional dependency violations in an XML document.
 Hypergraph $g$ of XML document $T$ can be defined as a pair $g = (V,E)$, where $V$ stands for a set of elements (called nodes), and $E$ is a set of non-empty subset of $V$ called hyperedges, more accurately each hyperedge indicates a set of value nodes violating FDs.
@@
 
-The algorithm fixing initial FD violations is shown in Algorithm \ref{fixInit}.
-The algorithm uses approximation algorithm to find VC for the minimum weighted vertex cover proposed in \cite{ApproxAlgo}.
-
-\begin{algorithm}
+\begin{algorithm}[H]
 \caption{Fix-Initial-Conflicts}
@@
 \ENSURE A modified document T.
-\STATE Create the initial conflict hypergraph $g$ of $T$ w.r.t $\Sigma$
-\STATE Use a known algorithm to find an approximation $VC$ for the minimum weighted vertex cover of $g$
+\STATE \COMMENT Create the initial conflict hypergraph $g$ of $T$ w.r.t $\Sigma$
+\STATE \COMMENT Use a known algorithm to find an approximation $VC$ for the minimum weighted vertex cover of $g$
 \STATE $remaining$ := VC
-\WHILE{There are two target nodes $v_1, v_2 \in T$ violating a FD $\sigma \in \Sigma$, and $v_1[\![P_{n+1}]\!]$ or $v_2[\![P_{n+1}]\!]$ is the only node in VC from the set of nodes $\{v_1[\![P_i]\!](i \in [1, n+1])\} \cup \{v_2[\![P_i]\!](i \in [1, n+1])\}$.
+\WHILE{there are two target nodes $v_1, v_2 \in T$ violating a FD $\sigma \in \Sigma$, and $v_1[\![P_{n+1}]\!]$ or $v_2[\![P_{n+1}]\!]$ is the only node in VC from the set of nodes $\{v_1[\![P_i]\!](i \in [1, n+1])\} \cup \{v_2[\![P_i]\!](i \in [1, n+1])\}$.
 (W.l.o.g assume the violation is as follows: $\sigma = (P,P',(P_1,\dots,P_n \rightarrow P_{n+1}))$, $v \in \{[\![P]\!]\}$, $v_1, v_2 \in \{v[\![P']\!]\}$, $v_1[\![P_i]\!] \equiv v_2[\![P_i]\!]$ for all $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] \not \equiv v_2[\![P_{n+1}]\!]$.)}
@@
 
+The algorithm fixing initial FD violations is shown in Algorithm \ref{fixInit}.
+The algorithm uses an approximation algorithm to find VC for the minimum weighted vertex cover proposed in \cite{ApproxAlgo}.
+
 \subsection{Resolving Violations Thoroughly}
@@
 \begin{define}[Core Determinant]
-Given an XML document $T$, a set $\Sigma$ of FDs and a node $u$ in $T$, we say that a set of nodes $\{u_1, u_2, \dots, u_n\}$ is a $\sigma\!-\!determinant$ of $u$, if there exists a nontrivial FD $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ logical implied by $\Sigma$, such that $\exists v \in \{[\![P]\!]\}, \exists v_1 \in \{[\![P']\!]\}, v_1[\![P]\!] = u_i$ for $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] = u$.\\
-We say that a set $C_u$ of nodes is a core determinant of $u$, if (a) for every nontrivial FD $\sigma$ implied by $\Sigma$ and every set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u \cap W \neq \emptyset$; and (b) for any proper subset $C_u'$ of $C_u$, there exists a nontrivial FD $\sigma$ implied by $\Sigma$, and a set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u' \cap W = \emptyset$.\qed
+Being given an XML document $T$, a set $\Sigma$ of FDs and a node $u$ in $T$, we say that a set of nodes $\{u_1, u_2, \dots, u_n\}$ is a $\sigma\!-\!determinant$ of $u$, if there exists a nontrivial FD $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ logicaly implied by $\Sigma$, such that $\exists v \in \{[\![P]\!]\}, \exists v_1 \in \{[\![P']\!]\}, v_1[\![P]\!] = u_i$ for $i \in [1,n]$, and $v_1[\![P_{n+1}]\!] = u$.\\
+We say that a set $C_u$ of nodes is a {\sl core determinant} of $u$, if (a) for every nontrivial FD $\sigma$ implied by $\Sigma$ and every set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u \cap W \neq \emptyset$; and (b) for any proper subset $C_u'$ of $C_u$, there exists a nontrivial FD $\sigma$ implied by $\Sigma$, and a set $W$ that is $\sigma\!-\!determinant$ of $u$, $C_u' \cap W = \emptyset$.\qed
 \end{define}
 
-\begin{algorithm}[H]
+\begin{algorithm}
 \caption{Resolve-Remaining-Violations}
@@
 
-The authors introduced effective two-step heuristic method to solve a problem of finding optimal repair of XML violations against functional dependencies, which is $\mathcal{NP}$-complete.
+The authors introduced an effective two-step heuristic method to solve a problem of finding optimal repair of XML violations against functional dependencies, which is $\mathcal{NP}$-complete.
 Moreover, they experimentally verified the effectivity and scalability of their approach using real-life and synthetic data.

### 2011-07-27 doc/theses/svirec/kap5.tex
corrections

@@
 A part of this work is experimental implementation of our proposed algorithm.
-Besides our approach, we have also implementened algorithm proposed in \cite{RepAndConsistentAnswer}, to be able to compare it with our algorithm.
-Our algorithm has been implemented in the JAVA language in the JDK version 6.0 as a part of the schema inference framework for XML, called jInfer \cite{jinfer}, developed as a part of a Software Project subject.
+Besides our approach, we have also implementened the algorithm proposed in \cite{RepAndConsistentAnswer}, to be able to compare it with our algorithm.
+Our algorithm has been implemented in the Java language in the version 6.0 as a part of the schema inference framework for XML jInfer \cite{jinfer}.
 
@@
 	\item[fd] contains data structure of functional dependencies used as an input to the algorithm.
-    \item [interfaces] contains JAVA interfaces necessery for integration to the jInfer framework.
+    \item [interfaces] contains Java interfaces necessery for integration in the jInfer framework.
     \item [modelGenerator] contains classes responsible for creating data structures from loaded input files.
-    \item [newRepairer] contains class \texttt{NewRepairerImpl}, the class that implements proposed algorithm. Also contains other data structures introduced in our approach
-    \item [propeties] contains classes reponsible for showing properties of this extension if the jInfer.
-    \item [repairer] contains classes implementing the original algorithm our algorithm is based on.
-    \item [weights] contains data structure of node weights used as an input to the algorithm.
+    \item [newRepairer] contains class \texttt{NewRepairerImpl}, the class that implements our proposed algorithm. Also it contains other data structures introduced in our approach
+    \item [properties] contains classes reponsible for showing properties of this extension in the jInfer.
+    \item [repairer] contains classes implementing the original algorithm from \cite{RepAndConsistentAnswer}.
+    \item [weights] contains data structure for node weights used as an input to the algorithm.
 \end{description}
@@
 
-Files that can be added as an input to our algorithm are of three kind.
+Files that can be added as an input to our algorithm are of three types.
 The XML data, file containing functional dependencies and file containing weights.
-First, the XML is processed by build-in Java DOM parster.
-Files with FDs and weights both contains XML data with specific structure defined as XML Schema in Fig.
+First, the XML is processed with build-in Java DOM parster.
+Files with FDs and weights both contain XML data with specific structure defined as XML Schema in Fig.
 \ref{fdschema} and \ref{weightschema} and are both processed with JAXB \cite{JAXB}.

### 2011-07-27 doc/theses/svirec/kap3.tex
added some new examples

@@
 \begin{example}\label{fdrepairExample}
-Consider the XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing a collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$ ($@$ denotes the attribute of an element).
-\begin{verbatim}
-<!ELEMENT bib (book+)>
-<!ELEMENT book (written_by, title)>
-<!ELEMENT written_by (author+)>
-<!ELEMENT author (name)>
-<!ATTLIST author ano CDATA>
-<!ELEMENT name PCDATA>
-<!ELEMENT title PCDATA>
-\end{verbatim}
+Consider the XML tree $XT$ conforming the DTD $D$ from Example \ref{pathExample}, which is representing a collection of books and the following functional dependency:\\ $\{bib.book, bib.book.written\_by.author.@ano\} \rightarrow bib.book.written\_by.author$ ($@$ denotes the attribute of an element).
+
 The functional dependency defined above requires that for each book there is only one element author having a given $@ano$ value.

### 2011-07-27 doc/theses/svirec/kap2.tex
added some new examples

@@
 
+\begin{example}\label{pathExample}
+Consider the XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing a collection of books.
+\begin{verbatim}
+<!ELEMENT bib (book+)>
+<!ELEMENT book (written_by, title)>
+<!ELEMENT written_by (author+)>
+<!ELEMENT author (name)>
+<!ATTLIST author ano CDATA>
+<!ELEMENT name PCDATA>
+<!ELEMENT title PCDATA>
+\end{verbatim}
+
+The set $paths(D)$ contains the following paths:
+\begin{align}
+paths(D) &= \left\{/bib, /bib/book, /bib/book/written\_by,\right.\nonumber\\
+&\qquad \left. /bib/book/written\_by/author, \right.\nonumber\\
+&\qquad \left. /bib/book/written\_by/author/name, \right.\nonumber\\
+&\qquad \left. /bib/book/written\_by/author/name/S, \right.\nonumber\\
+&\qquad \left. /bib/book/written\_by/author/@ano, \right.\nonumber\\
+&\qquad \left. /bib/book/title\nonumber\right\}
+\end{align}\qed
+\end{example}
+
 \section{Integrity Constraints}
@@
  \end{itemize}
+
 A rootless tree formula is an expression of the form $p(\Phi_1 \land \dots \land \Phi_k)$ where $\Phi_i$ is a rootless path formula (expression of the form $p[y]$ where $p$ is a path expression and $y$ is a term) or a rootless tree formula and $p$ is a path expression.
@@
 \end{define}
+
+\begin{example}
+Consider the XML tree $XT$ from Example \ref{example1ref}. Thereafter the integrity constraint that there must exist at least two books differents titles is expressed as
+
+\begin{align*}
+\forall(X)&[[root]/bib[X] \supset\\
+&\exists(Z_1, Z_2)([X]/book/title/S[Z_1] \land [X]/book/title/S[Z_2] \land Z_1 \not = Z_2)]
+\end{align*}\qed
+\end{example}
 
@@
 \begin{define}[Functional Dependency]\label{fd1}
-Being given a DTD $D$, a {\sl functional dependency} on $D$ is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
+Being given a DTD $D$, a {\sl functional dependency} on $D$ is an expression of the form $S \rightarrow p$, where $S$ is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
 \end{define}
@@
 \end{define}
+
+The main difference between two definitions of the functional dependency is that the former definition can use any path from $paths(D)$, whereas the latter considers that each $P_i (i \in [1,n])$ is from $StrPaths(D)$.
+That means the constraint from Example \ref{fdExample} cannot be expressed by functional dependency defined in \ref{fd2}, because of $/bib/book/written\_by/author$ path.
+Nevertheless, modified constraint from Example \ref{fdExample} expressed by FD defined in \ref{fd2} is shown in Example \ref{fd2Example}.
+
+\begin{example}\label{fd2Example}
+Consider the constraint $C$ saying that two distinct authors with different names of the same book cannot have the same value of attribute \texttt{ano}.
+The functional dependency expressing $C$ is defined as follows:
+
+$$(bib/book, written\_by/author, (@ano \rightarrow name))$$\qed
+\end{example}

### 2011-07-28 doc/theses/svirec/kap4.tex
added some new examples
another changes

@@
 If the user does not assign weights nor does he use statistical methods, a default weight is assigned to all nodes.
-A definition of an RW-XML tree follows:
+
+\begin{example}
+Consider the XML document and corresponding XML tree from Example \ref{example1}.
+Let us assign weights to nodes defined by XPath constructs shown in Table \ref{weightTab}.
+The XML tree with assigned custom and also default weights is shown in Fig.
+\ref{weightsFig}.
+
+\begin{table}[h]
+\begin{tabular}{r | c}
+XPath & weight\\\hline
+$/bib/book/title/text()$ & 0.8\\
+$/bib/book/written\_by/author$ & 0.6\\
+$/bib/book/written\_by/author/@ano$ & 0.1\\
+\end{tabular}
+\caption{Table of custom weights assigned to the XML tree.}
+\label{weightTab}
+\end{table}
+
+\begin{figure}[H]
+    \centering\includegraphics[width=\textwidth]{weights}
+	\caption{The XML tree with assigned weights to the nodes.} \label{weightsFig}
+\end{figure}
+\qed
+\end{example}
 
 \begin{define}[RW-XML tree]
-A RW-XML tree is a pair $RWXT = \langle RXT, \omega \rangle$, where $RXT$ is an R-XML tree and $\omega$ is a weight function from $N_T$ to $[0,1]$.\qed
+A {\sl RW-XML tree} is a pair $RWXT = \langle RXT, \omega \rangle$, where $RXT$ is an R-XML tree and $\omega$ is a weight function from $N_T$ to $[0,1]$.\qed
 \end{define}
 
-After creating the RW-XML tree from input XML data, a set of tree tuples (defined in \ref{treeTuple}) is constructed.
+After creating the RW-XML tree from input XML data, a set of tree tuples (defined in Definition r\ref{treeTuple}) is constructed.
 Since the set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree.
@@
 Since each repair candidate consists of a set of nodes which are modified (marked as unreliable or having a modified value), we can compute the cost of each candidate.
-It is important to say that in our approach we prefer repair candidates that modify the value of the nodes to those marking nodes as unreliable.
-Therefore we added a coefficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying the node value will have lower cost.
+It is important to say that unlike in original approach, we do not prefer repair candidates that modify the value of the nodes to those marking nodes as unreliable.
+Therefore we added a coefficient $k$ to the calculation of the repair candidate cost, so that the priority of repair candidate marking node as unreliable can be achieved.
 The definition of repair candidate cost is as follows:
@@
 $$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \cdot k,$$
-where $k$ is the coefficient determining the priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
+where $k$ is the priority of repairing the candidate by modifying the node value in constrast to marking it unreliable.\qed
 \end{define}
@@
 \begin{define}[Repair Group weight]
-Being given an RW-XML tree $RXT$ and a repair group $RG$, we define the weight of $RG$ as the sum of costs of all repair candidates in the $RG$.
+Being given an RW-XML tree $RXT$ and a repair group $RG$, we define the {\sl weight} of $RG$ as the sum of costs of all repair candidates in the $RG$.
 \end{define}
@@
 First it decides, whether the algorithm is in a state where the user is selecting repair candidates (the user selection mode), or the user leaves decision making on the algorithm using his previous selections (the guess mode) (line 2).
-If we are still in user selection mode, the user chooses from repair groups sorted by the weight the most convenient one, and from this group the user chooses the repair candidate that will be applied on RW-XML tree (line 4).
+If we are still in user selection mode, the user chooses from repair groups sorted by the weight the most convenient one, and from this group the user chooses the repair candidate that will be applied on RW-XML tree (line 3).
 Repair candidates in each repair group are also sorted by the cost, working as a hint for the user which repair would be chosen by the previous automatic method of selection.
-The last step of user selection mode is saving the information from selected repair candidate (line 5).
+The last step of user selection mode is saving the information from selected repair candidate (line 4).
 This information consists of the FD which this candidate repairs, nodes the repair changes (their paths) and also whether the change was a value modification or marking a node as unreliable.
 
-If the user has decided not to select repair candidates by himself anymore, the algorithm goes into the guess mode (starting at line 6).
-In this mode the algorithm checks all repair groups whether one of them contains a repair candidate that is sufficiently similar to some previously selected repair candidate (line 7-10).
+If the user has decided not to select repair candidates by himself anymore, the algorithm goes into the guess mode (starting at line 5).
+In this mode the algorithm checks all repair groups whether one of them contains a repair candidate that is sufficiently similar to some previously selected repair candidate (line 6-12).
 This similarity is checked by function \texttt{canBeUsedUserSelection()} shown in Algorithm \ref{canUserSel}.
-If neither of the candidates is sufficiently similar, the algorithm chooses the repair candidate with the lowest cost from the repair group with the lowest weight (line 15-16).
+If neither of the candidates is sufficiently similar, the algorithm chooses the repair candidate with the lowest cost from the repair group with the lowest weight (line 13-14).
 
@@
 \STATE $R = \emptyset$
-\IF{$isUserSelection()$}
-    \STATE \COMMENT repair candidate selected by user
-    \STATE $R = getRepairFromUser()$
+\IF[the user selection mode]{$isUserSelection()$}
+    \STATE $R = getRepairFromUser()$ \COMMENT{repair candidate selected by user}
     \STATE $saveSelectedRepair(SR, R)$
-\ELSE
+\ELSE[the guess mode]
     \FORALL{$RG$ in $RGS$}
-        \STATE \COMMENT for all repair candidates in the repair group
-        \FORALL{$RC$ in $RG$}
+        \FORALL[for all repair candidates in the repair group]{$RC$ in $RG$}
             \IF{$canBeUsedUserSelection(SR, t, RC)$}

### 2011-07-28 doc/theses/svirec/kap6.tex
another changes

@@
 
-For each dataset, we first run the original algorithm and collect from a repaired document informations about how many nodes have been modified with particular update operation (we split value modification operation into changing value to new one and copying one value from one node to another).
+For each dataset we perform the following.
+First, we run the original algorithm and collect from a repaired document informations about how many nodes have been modified with particular update operation.
+Moreover, we split the value modification operation into two types.
+The first is changing value to new one and the second one is copying the value from one node to another.
 
@@
 One exception where our algorithm modifies more nodes is the case when we set value of $k$ near 0, which marks nodes as unreliable.
-With this setting we want to demonstrate that with our approach it is possible to choose instead of modifying node values, marking nodes as unreliable.
+With this setting we want to demonstrate that with our approach it is possible to mark nodes as unreliable instead of modifying their values.
 This is not possible with the original repairer, since it prefers node value modification to marking node as unreliable.
@@
 
-With setting the weight to particular nodes we want to show that our algorithm is capable of choosing the priority of nodes, which values are modified (\texttt{reed.xml} and \texttt{wsu.xml}), or if the value is changed to new value (\texttt{actors.xml} and \texttt{customers.xml}).
+Setting the weight to particular nodes effects the choise of strategies used to repair the violations.
+In datasets \texttt{reed.xml} and \texttt{wsu.xml} setting the custom weight to nodes causes the node values are changed to values of other nodes.
+On the other side, in datasets \texttt{actors.xml} and \texttt{customers.xml} setting the weights causes changing the value of nodes to a newly generated one.
 
@@
 It means that one of the repair candidate repairs more than one FD violation and therefore less value modifications are needed.
-In the case where the number of picked repair groups outreach the firstly created ones, is visible that some repair candidate introduces new violation which was consequently repaired in the next iteration of our algorithm.
+In the case where the number of picked repair groups outreach the number of firstly created ones, it means that some repair candidate introduces new violation which was consequently repaired in the next iteration of our algorithm.

### 2011-07-28 doc/theses/svirec/kap3.tex
another changes

@@
 As was mentioned earlier, a repairing primitive is a node value modification, where for repairing algorithm a combination of two rules to resolve violation is used.
-Let us have an $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ (from definition \ref{fd2}) and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
+Let us have an $FD$ $\sigma = (P, P', (P_1, \dots, P_n \rightarrow P_{n+1}))$ (from Definition \ref{fd2}) and consider two nodes $v_1$ and $v_2$ matching path $P'$ in a subtree rooted at a node in $\{[\![p]\!]\}$.
 If the child nodes of $v_1$ and $v_2$ qualified by paths $P_i$ have equal values for all $i \in [1,n]$ and their child nodes qualified by $P_{n+1}$ have different values, then $v_1$ and $v_2$ violates $\sigma$.

### 2011-07-29 doc/theses/svirec/appendixA.tex
content of CD

\chapter{Content of CD}

The CD is a part of this thesis. It contains text of this thesis, source code of jInfer framework with \texttt{FDRepairer} extension together with documentation and built jInfer as \texttt{nbm} plugins applicable into NetBeans platform. It aslo contains experimental data and results. CD has the following structure:

\begin{itemize}
	\item \textit{content.txt} - A file with this text
    \item \textit{text} - A directory with the PDF file containing text of this thesis
    \item \textit{src} - Source codes of jInfer including \texttt{FDRepairer} extension
    \item \textit{javadoc} - A generated javadoc documentation
    \item \textit{bin} - A directory with \texttt{nbm} plugins of jInfer and FDRepairer
    \item \textit{datasets} - Experimental data and results
\end{itemize}

### 2011-07-30 doc/theses/svirec/appendixC.tex
appendix C - user guide added

\chapter{FDRepairer Guide}

First, before we start repairing XML data violating functional dependencies, new jInfer project needs to be created:

\begin{figure}[H]
    \centering\includegraphics[width=\textwidth]{tutorial_images/new_project}
\end{figure}

Files with XML data, defined functional dependencies and weights assigned to the nodes needs to be added to the newly created project. Files with functional dependencies and weights are xml files with XML Schema shown in Fig. \ref{fdschema} and \ref{weightschema}. These files appears after adding in the \emph{FD} folder of the jInfer project.

\begin{figure}[H]
    \centering\includegraphics[width=\textwidth]{tutorial_images/add_files}
\end{figure}

The settings of the \texttt{FDRepairer} extension is available in the properties window of the jInfer project accessible through project context menu. In these settings, one could choose between original repairer and repairer proposed in this thesis. For the proposed algorithm is then available to choose between \emph{user interactive} repair picker and \emph{Minimal Repair Picker}. In this place is also possible to set coefficient $k$ and threshold $t$.

\begin{figure}[H]
    \centering\includegraphics[width=\textwidth]{tutorial_images/project_properties}
\end{figure}

The repair is executed by the \emph{Run Repair} item from the jInfer project context menu:

\begin{figure}[H]
    \centering\includegraphics[width=\textwidth]{tutorial_images/run_repair}
\end{figure}

If the \emph{user interactive} repair picker has been selected, \emph{RepairPicker Window} is shown to the user. In this window the user picks repair candidate, which will be applied to the repaired XML.

\begin{figure}[H]
    \centering\includegraphics[width=\textwidth]{tutorial_images/user_selection}
\end{figure}

After repairing all the violations, repaired XML document is created in the \emph{Output} subfolder of the jInfer project.

### 2011-07-30 doc/theses/svirec/kap5.tex
appendix C - user guide added

@@
 However, one part of the algorithm have not been implemented yet, namely the intersection of repair candidates with existing repair group.
+
+\section{Building and Executing}
+
+For building jInfer with \texttt{FDRepairer} from sources, please refer to the tutorial at \url{http://jinfer.sourceforge.net/building_jinfer.html}.
+To execute jInfer, install plugins from enclosed CD or built from sources into NetBeans.
+
+Appendix C contains a simple user guide with screenshots about how to use \texttt{FDRepairer}.

### 2011-07-30 doc/theses/svirec/prace.tex
content of CD
appendix C - user guide added

@@
 
-\linespread{1.3}
+%\linespread{1.3}
 %\renewcommand\baselinestretch{1.3} % riadkovanie jeden a pol
@@
 \usepackage{multirow}
+\usepackage{setspace}
+
+\onehalfspacing
 
@@
 \noalign{\vspace{2mm}}
-Study programme: & Informatika \\
+Study programme: & Informatics \\
 \noalign{\vspace{2mm}}
-Specialization: & ISS \\
+Specialization: & Software Systems \\
 \end{tabular}
@@
 Abstrakt:
+Najviac používanými integritnými obmedzeniami v XML sú primárne a cudzie kľúče. Okrem nich však poznáme aj iný typ integritných obmedzení, a to funkčné závislosti. Tak ako pre relačné databázy,
 % abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce
 
-Klíčová slova:
+Klíčová slova: XML, funkčná závislosť, porušenie funkčných závislostí, oprava porušení
 % 3 aľ 5 klíčových slov
@@
 Title:
+\mftitle
 % přesný překlad názvu práce v angličtině
@@
 Abstract:
+
 % abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
@@
 
-Keywords:
+Keywords: XML, functional dependency, functional dependencies violations, violations repair
 % 3 aľ 5 klíčových slov v angličtině
@@
 
+\appendix
+\include{appendixA}
+
 %%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
@@
 \include{attachments}
+\include{appendixC}
 

### 2011-07-30 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 \usepackage{nomencl}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
@@
 \def\.{\discretionary{}{}{}}
+\newcommand{\heu}[1]{\texttt{\textit{#1}}}
 \newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
 \linespread{1.5}
+
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\renewcommand{\algorithmicforall}{\textbf{for each}}
+\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
 
@@
 
-TODO how GLPK works (branch&bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
+TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
 
@@
 
+TODO we will be using a pool - what is a pool
+
 \subsection{Constructions Heuristics}
-
+1
 % TODO can I format things in nomenclature?
@@
 
+See listing \ref{ch-random}.
+
+\begin{algorithm}
+\caption{Random CH}
+\label{ch-random}
+\begin{algorithmic}
+\REQUIRE $N$ required size of pool
+\REQUIRE $C$ list of candidate AMs
+\ENSURE pool of $N$ feasible solutions
+\STATE $r \gets $ empty pool
+\FOR{$i = 1 \to N$} 
+\STATE \COMMENT{create 1 solution}
+\STATE $s \gets $ empty solution
+\WHILE{$s$ is a feasible ID set}
+\STATE $a \gets $ pick at random from $C \textbackslash S$
+\STATE $s \gets s \cup a$
+\ENDWHILE
+\STATE $r \gets r \cup s$
+\ENDFOR
+\RETURN $r$
+\end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Fuzzy}
@@
 
+See listing \ref{ch-fuzzy}.
+
+\begin{algorithm}
+\caption{Fuzzy CH}
+\label{ch-fuzzy}
+\begin{algorithmic}
+\REQUIRE $N$ required size of pool
+\REQUIRE $C$ list of candidate AMs
+\ENSURE pool of $N$ feasible solutions
+\STATE $r \gets $ empty pool
+\FOR{$i = 1 \to N$} 
+\STATE \COMMENT{create 1 solution}
+\STATE $s \gets $ empty solution
+\STATE $C' \gets C$
+
+\WHILE{$C' $ not empty}
+\STATE $a \gets $ pick at weighted random from $C'$
+\IF{$s \cup a$ is a feasible ID set}
+\STATE $s \gets s \cup a$
+\STATE $C' \gets C' \textbackslash a$
+\ENDIF
+\FORALL{$c \in C'$}
+\IF{$s \cup c $ is \textbf{not} a feasible ID set} 
+\STATE \COMMENT {if $c$ cannot be possibly added anymore}
+\STATE $C' \gets C' \textbackslash c$
+\ENDIF
+\ENDFOR
+\ENDWHILE
+
+\STATE $r \gets r + s$
+\ENDFOR
+\RETURN $r$
+\end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Incremental}
 
-TODO trivial, hungry
+TODO trivial, hungry, always 1 solution
+
+See listing \ref{ch-incremental}.
+
+\begin{algorithm}
+\caption{Incremental CH}
+\label{ch-incremental}
+\begin{algorithmic}
+\REQUIRE $C$ list of candidate AMs
+\ENSURE a feasible solution
+\STATE $C' \gets $ sort $C$ by decreasing weight
+\STATE $s \gets $ empty solution
+\FORALL{$c \in C'$}
+\IF{$s \cup c$ is a feasible ID set}
+\STATE $s \gets s + c$
+\ENDIF
+\ENDFOR
+\RETURN $s$
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-TODO trivial, "hungry"
+TODO trivial, "hungry", always 1 solution
+
+See listing \ref{ch-removal}.
+
+\begin{algorithm}
+\caption{Removal CH}
+\label{ch-removal}
+\begin{algorithmic}
+\REQUIRE $C$ list of candidate AMs
+\ENSURE a feasible solution
+\STATE $C' \gets $ sort $C$ by increasing weight
+\STATE $s \gets C'$
+\FORALL{$c \in s$}
+  \IF{$s$ is a feasible ID set}
+    \RETURN $s$
+  \ENDIF
+  \STATE $s \gets s \textbackslash c$
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-TODO what they are, that they need a pool sometimes, how do we give them pool, ...
+TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
 
@@
 
-TODO trivial trivial
+This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{ih-identity}.
+
+\begin{algorithm}
+\caption{Identity IH}
+\label{ih-identity}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\ENSURE the same pool of feasible solutions
+\RETURN $FP$
+\end{algorithmic}
+\end{algorithm}
 
@@
 
+See listing \ref{ih-removeworst}.
+
+\begin{algorithm}
+\caption{Remove Worst IH}
+\label{ih-removeworst}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\ENSURE pool of feasible solutions
+\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
+\RETURN $FP \textbackslash s_{min}$
+\end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Random Remove}
@@
 
+See listing \ref{ih-randomremove}.
+
+\begin{algorithm}
+\caption{Random Remove IH}
+\label{ih-randomremove}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
+\ENSURE pool of feasible solutions
+\FORALL{$s \in FP$}
+  \STATE $K \gets k * |s|$
+  \STATE remove $K$ random AMs from $s$
+\ENDFOR
+\RETURN $FP$
+\end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Hungry}
 
-TODO trivial
+This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
+
+\heu{Hungry} tries to improve each solution in the feasible pool in the following way.
+It orders all candidate AMs not present in the solution by decreasing weight.
+Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition.
+The resulting solution (whether any AMs were added or not) is then returned to the pool.
+Listing \ref{ih-hungry} captures this process.
+
+\begin{algorithm}
+\caption{Hungry IH}
+\label{ih-hungry}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\REQUIRE $C$ list of candidate AMs
+\ENSURE pool of feasible solutions
+\FORALL{$s \in FP$}
+  \STATE \COMMENT {improve a single solution}
+  \STATE $C' \gets C \textbackslash s$
+  \STATE $C' \gets C'$ sorted by decreasing weight
+  \FORALL{$c \in C'$}
+    \IF{$s \cup c$ is a feasible ID set}
+      \STATE $s \gets s \cup c$
+    \ENDIF
+  \ENDFOR
+\ENDFOR
+\RETURN $FP$
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-TODO
+TODO explain how this translates to GLPK input
+
+See listing \ref{ih-mutation}.
+
+\begin{algorithm}
+\caption{Mutation IH}
+\label{ih-mutation}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\REQUIRE $k$ ratio of AMs to fix
+\ENSURE pool of feasible solutions
+\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
+\STATE $K \gets k * |incumbent|$
+\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
+\STATE $improved \gets $ run GLPK
+\RETURN $FP \cup improved$
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-TODO
+TODO explain how this translates to GLPK input
+
+See listing \ref{ih-crossover}.
+
+\begin{algorithm}
+\caption{Crossover IH}
+\label{ih-crossover}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\REQUIRE $k$ ratio of solutions to look for commonalities in
+\ENSURE pool of feasible solutions
+\STATE $K \gets k * |FP|$
+\STATE $FP' \gets K$ random solutions $\in FP$
+\STATE $am \gets$ AMs found in all solutions $\in FP'$
+\STATE fix $am$ in GLPK problem formulation
+\STATE $improved \gets $ run GLPK
+\RETURN $FP \cup improved$
+\end{algorithmic}
+\end{algorithm}
 
@@
 TODO explain how this translates to GLPK input
+
+See listing \ref{ih-localbranching}.
+
+\begin{algorithm}
+\caption{Local Branching IH}
+\label{ih-localbranching}
+\begin{algorithmic}
+\REQUIRE $FP$ pool of feasible solutions
+\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
+\ENSURE pool of feasible solutions
+\STATE $K \gets k * |$total AM count$|$
+\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
+\STATE add max Hamming distance requirement to GLPK problem formulation
+\STATE $improved \gets $ run GLPK
+\RETURN $FP \cup improved$
+\end{algorithmic}
+\end{algorithm}
 
@@
   \item With the best IH from the previous step, how about using Random as CH?
-	\item GLPK: Input -> How long the search for optimum lasts
+	\item GLPK: Input -> How long the search for optimum lasts - this shows we cannot rely on GLPK alone! it just takes too long.
 	\item GLPK: Time limit -> Quality

### 2011-07-31 doc/theses/vitasek/diplomka.tex
Thesis: adding Traces appendix.

@@
 \usepackage[T1]{fontenc}
-%\usepackage{tgpagella}
 \usepackage{xstring}
@@
 \usepackage{algorithmic}
+\usepackage{multicol}
 
@@
 	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping
+	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
 	\item ID set
@@
 %%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-%\chapwithtoc{List of Abbreviations}
 
@@
 
-TODO a few (commented!) traces from experiment runs
+Following is a trace logged from a sample experiment run. It shows all the relevant information related to this instance, any and every piece of information we might be interested in.
+
+To save space, 2-column layout is used. Commentary on the particulars follows right after its end.
+
+\begin{multicols}{2}
+\begin{scriptsize}
+\begin{verbatim}
+CPU info
+  Intel(R) Core(TM)2 Quad CPU Q9550 @ 2.83GHz
+  Cores: 4
+  Clock speed: 2983 MHz
+Memory info
+  Size: 8192 MB
+OS info
+  Name: Windows 7
+  Version: 6.1
+  Architecture: amd64
+Java info
+  Version: 1.6.0_26
+  VM: Java HotSpot(TM) 64-Bit Server VM
+GLPK info
+  GLPSOL: GLPK LP/MIP Solver 4.34
+
+Configuration:
+File name: graph.xml (101599 b)
+  Graph representation: 82 vertices, 1101 edges
+alpha: 1.0, beta: 1.0
+
+Results:
+Start time: Thu Jan 01 01:22:17 CET 1970
+Total time spent: 7754 ms
+Final quality: 0.19951219512195123 (10 AMs)
+Highest quality: 0.23463414634146343 (12 AMs)
+Construction phase: 
+  Algorithm: Random
+    Time taken: 248 ms
+    Time since start: 248 ms
+    Pool size: 10
+    Quality: 0.19975609756097568 (11 AMs)
+Improvement phase: 
+  pass #1: 
+  Algorithm: RandomRemove, ratio = 0.2
+    Time taken: 0 ms
+    Time since start: 841 ms
+    Pool size: 10
+    Quality: 0.15878048780487808 (9 AMs)
+  pass #2: 
+  Algorithm: Mutation, ratio = 0.1, limit = 1 s
+    Time taken: 1512 ms
+    Time since start: 2710 ms
+    Pool size: 11
+    Quality: 0.21975609756097558 (11 AMs)
+
+  <... 7 more passes removed ...>
+
+  pass #10: 
+  Algorithm: Remove Worst
+    Time taken: 80 ms
+    Time since start: 7676 ms
+    Pool size: 12
+    Quality: 0.19951219512195123 (10 AMs)
+Termination reason: Maximum iterations exceeded.
+
+Time,Quality,AMs
+248,0.19975609756097568,11
+841,0.15878048780487808,9
+2710,0.21975609756097558,11
+2927,0.1890243902439024,9
+4421,0.23463414634146343,12
+4703,0.23463414634146343,12
+4896,0.1960975609756098,10
+5793,0.23463414634146337,12
+5972,0.19951219512195123,10
+7433,0.19951219512195123,10
+7676,0.19951219512195123,10
+
+ID
+Element,Attribute,Weight
+vertex0,attr,0.024146341463414635
+vertex2,attr,0.01975609756097561
+vertex33,attr,0.016829268292682928
+vertex34,attr,0.02219512195121951
+vertex4,attr,0.022682926829268292
+vertex41,attr,0.014878048780487804
+vertex7,attr,0.02170731707317073
+vertex70,attr,0.018780487804878048
+vertex76,attr,0.01780487804878049
+vertex8,attr,0.02170731707317073
+vertex80,attr,0.01780487804878049
+vertex97,attr,0.016341463414634147
+
+IDREF
+Element,Attribute
+\end{verbatim}
+\end{scriptsize}
+\end{multicols}
+
+The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
+To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
+It tries to guess whether it's the Cygwin version by looking at the path to the binary.
+
+% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
+The second section states the input file along with its size and graph representation.\\
+Alpha and beta parameters for this instance belong here too.
+
+\begin{footnotesize}
+\begin{verbatim}
+Configuration:
+File name: graph.xml (101599 b)
+  Graph representation: 82 vertices, 1101 edges
+alpha: 1.0, beta: 1.0
+\end{verbatim}
+\end{footnotesize}
+
+Results section opens stating the most important information first: how long did the experiment run and what was the highest and final quality (these two are potentially different).
+Numbers of attribute mappings in the best and final solution respectively are stated as well.
+
+\begin{footnotesize}
+\begin{verbatim}
+Total time spent: 7754 ms
+Final quality: 0.19951219512195123 (10 AMs)
+Highest quality: 0.23463414634146343 (12 AMs)
+\end{verbatim}
+\end{footnotesize}
+
+Construction phase results go next.
+Among reported information are the full identification of the heuristic (possibly along with its parameters), time taken, size of the pool created and the quality of the incumbent solution (again, with the number of its AMs).
+
+\begin{footnotesize}
+\begin{verbatim}
+Algorithm: Random
+  Time taken: 248 ms
+  Time since start: 248 ms
+  Pool size: 10
+  Quality: 0.19975609756097568 (11 AMs)
+\end{verbatim}
+\end{footnotesize}
+
+Now for each of the improvement phases there is one section in output log.
+Information presented here has the same structure as with the construction phase.
+Please note that the \code{Pool size} is always measured \textit{after} the improvement run.
+
+\begin{footnotesize}
+\begin{verbatim}
+Algorithm: Mutation, ratio = 0.1, limit = 1 s
+  Time taken: 1512 ms
+  Time since start: 2710 ms
+  Pool size: 11
+  Quality: 0.21975609756097558 (11 AMs)
+\end{verbatim}
+\end{footnotesize}
+
+After the last improvement phase, the reason why the metaheuristic terminated is stated.
+Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha/beta settings.
+\\
+
+To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
+Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
+
+\begin{footnotesize}
+\begin{verbatim}
+Time,Quality,AMs
+...
+841,0.15878048780487808,9
+2710,0.21975609756097558,11
+...
+\end{verbatim}
+\end{footnotesize}
+
+And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
+Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
+
+\begin{footnotesize}
+\begin{verbatim}
+Element,Attribute,Weight
+vertex0,attr,0.024146341463414635
+...
+\end{verbatim}
+\end{footnotesize}
+
+Note that in this example trace there were no IDREF AMs found.
 

### 2011-07-31 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
+\begin{center}
+  \includegraphics[width=\textwidth]{images/metaheuristic.eps}
+\end{center}
+
 \subsection{Constructions Heuristics}
-1
 % TODO can I format things in nomenclature?
@@
 TODO this is a heuristic too! We can start with it!
+
+The pseudocode is listed (cited from \cite{fidax} with trivial modifications) in \ref{ch-fidax}.
+
+\begin{algorithm}
+\caption{FIDAX CH}
+\label{ch-fidax}
+\begin{algorithmic}
+\REQUIRE $C$ list of candidate AMs
+\STATE $C' \gets C$ sorted by decreasing size
+\STATE Compute the weight $w(m)$ of each $m$ in $C$
+\FORALL{$t$ in $\Sigma^E$}
+  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
+  \STATE Remove from $C'$ all mappings of type $t$ except $m$
+\ENDFOR
+\FORALL{$m$ in $C'$}
+  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
+  \IF{$w(m) > \sum_{p \in S} w(p)$}
+    \STATE remove all $p \in S$ from $C'$
+  \ELSE
+  	\STATE remove $m$ from $C'$
+  \ENDIF
+\ENDFOR
+\RETURN $C'$
+\end{algorithmic}
+\end{algorithm}
 

### 2011-07-31 doc/theses/svirec/kap6.tex
added section about jInfer and some DTDs to the experimental part.

@@
 The course data consist, along with other information, of day of the week, time and place (building and room), where each course is situated.
+DTD of one of the dataset file is shown in Fig.
+\ref{courseDTD}.
 The consistency of data is evaluated against FD defined as: \emph{Two courses starting at the same date and time are each situated in a different place}.\\
+
+\begin{figure}
+    \begin{verbatim}
+        <!ELEMENT root (course*)>
+        <!ELEMENT course (reg_num,subj,crse,sect,title,units,
+                            instructor,days,time,place)>
+        <!ELEMENT reg_num (#PCDATA)>
+        <!ELEMENT subj (#PCDATA)>
+        <!ELEMENT crse (#PCDATA)>
+        <!ELEMENT sect (#PCDATA)>
+        <!ELEMENT title (#PCDATA)>
+        <!ELEMENT units (#PCDATA)>
+        <!ELEMENT instructor (#PCDATA)>
+        <!ELEMENT days (#PCDATA)>
+        <!ELEMENT time (start_time,end_time)>
+        <!ELEMENT start_time (#PCDATA)>
+        <!ELEMENT end_time (#PCDATA)>
+        <!ELEMENT place (building,room)>
+        <!ELEMENT building (#PCDATA)>
+        <!ELEMENT room (#PCDATA)>
+    \end{verbatim}
+    \caption{DTD of reed.xml from course dataset}
+    \label{courseDTD}
+\end{figure}
 
@@
 The dataset consists of a set of authors, where for each author a set of movies he played in is listed.
+DTD of this dataset is shown in Fig.
+\ref{actorsDTD}.
 For this dataset, we defined FD as follows: For each two authors, which played in the movie with the same name, the year of release of this movie must be the same.
+
+\begin{figure}[H]
+    \begin{verbatim}
+        <!ELEMENT W4F_DOC (Actor)>
+        <!ELEMENT Actor (Name,Filmography)>
+        <!ELEMENT Name (FirstName, LastName)>
+        <!ELEMENT FirstName (#PCDATA)>
+        <!ELEMENT LastName (#PCDATA)>
+        <!ELEMENT Filmography (Movie)*>
+        <!ELEMENT Movie (Title,Year)>
+        <!ELEMENT Title (#PCDATA)>
+        <!ELEMENT Year (#PCDATA)>
+    \end{verbatim}
+    \caption{DTD of actors.xml}
+    \label{actorsDTD}
+\end{figure}
 
@@
 
-Two synthetic data sets have been constructed, the former represents data introduced in Example \ref{example1} with FD presented in Example \ref{fdExample}.
+Two synthetic datasets have been constructed, the former represents data introduced in Example \ref{example1} with FD presented in Example \ref{fdExample}.
 The latter is created according to Example 1 introduced in \cite{ImprovingXML}.
+DTD of the former dataset is identical with DTD defined in Example \ref{pathExample}.
+The latter datasets DTD is shown in Fig.
+\ref{synthDTD}.
+
+\begin{figure}[H]
+    \begin{verbatim}
+        <!ELEMENT customers (country)>
+        <!ELEMENT country (name,c_list)>
+        <!ELEMENT name (#PCDATA)>
+        <!ELEMENT c_list (customer)*>
+        <!ELEMENT customer (no,phone,zip,city)>
+        <!ELEMENT no (#PCDATA)>
+        <!ELEMENT phone (#PCDATA)>
+        <!ELEMENT zip (#PCDATA)>
+        <!ELEMENT city (#PCDATA)>
+    \end{verbatim}
+    \caption{DTD of actors.xml}
+    \label{synthDTD}
+\end{figure}
 

### 2011-07-31 doc/theses/svirec/kap5.tex
added section about jInfer and some DTDs to the experimental part.

@@
 
+\section{jInfer Framework}
+
+jInfer framework was developed as a Software Project at the Faculty of Mathematics and Physics, Charles University at Prague.
+The framework is based on NetBeans platform and is mainly used for XML schema inference.
+Although our proposed algorithm is not dealing with the schema inference, jInfer is designed to be extensible and is suitable for incorporation our algorithm into it.
+
+The whole process of repairing inconsistent XML data is divided three steps (illustrated in Fig. \ref{repairProcess}):
+
+\begin{enumerate}
+	\item Import of input data (XML, functional dependencies, weights) into an internal representation - \emph{Initial Model}.
+    \item Repairing of \emph{Initial Model} with optional help of user interaction.
+    \item Export back the repaired data into XML format.
+\end{enumerate}
+
+\begin{figure}[H]
+    \centering\includegraphics[width=0.6\textwidth]{repair_process}
+    \caption{High-level view of the repair process}
+    \label{repairProcess}
+\end{figure}
+
 \section{Architecture}
 
-The main part of our algorithm is implemented as the \texttt{FDRepairer} extension of the jInfer framework.
-The core package of the extension is \texttt{cz.cuni.mff.\discretionary{}{}{}ksi.jinfer.functionalDependencies} and contains classes representing base object important for repairing (Tuple, RW-XML tree, Path etc.).
+The main part of our algorithm is implemented as the \texttt{FDRepairer} module of the jInfer framework.
+The core package of the module is \texttt{cz.cuni.mff.\discretionary{}{}{}ksi.jinfer.functionalDependencies} and contains classes representing base object important for repairing (Tuple, RW-XML tree, Path etc.).
 This package is further divided into the following subpackages:
@@
 
-The \texttt{FDRepairer} extension of the jInfer framework is an implementation of our proposed algorithm presented in Chapter \ref{kap4}.
+The \texttt{FDRepairer} module of the jInfer framework is an implementation of our proposed algorithm presented in Chapter \ref{kap4}.
 However, one part of the algorithm have not been implemented yet, namely the intersection of repair candidates with existing repair group.

### 2011-07-31 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 \todo[inline]{Translate.}
-Existuje mnoho prístupov generovania XML schém.
-Väčšina z nich spracováva XML data \todo{spomenúť alebo rozpísať tieto prístupy, ako napríklad EXTRACT?}.
-Podľa mojej znalosti v dobe písania práce existuje len jediný \todo{je pravda?} prístup využívajúci XML operácie\todo{ref do literatury}.
-Ten sa snaží hľadať keys a foreign keys za pomoci XQuery dotazov.
+Many approaches of XML Schema inference have been proposed. Most of them process XML data as an input for inference. Some of these methods are mentioned in this chapter.
+
+Besides approaches exploiting XML data, approaches that utilize operations over XML data may be developed and they are of the most interest in terms of this work.
+According to my best knowledge at the time of writing there is just one approach of this category.
+It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?}.
 
 \section{Discovering XML Keys and Foreign Keys in Queries}
-Táto metóda sa snaží vylepšiť automatické odvodzovanie XML schém hľadaním keys a foreign keys zo sady XQuery dotazov.
-Pre toto hľadanie sú použité len tieto dotazy, nie samotné XML data.
-Výsledkom je zoznam keys a foreign keys, ktoré môžu byť zachytené pomocou \textbf{key}, \textbf{keyref} \todo{aj unique?} konštrukcií jazyka XML Schema.
+This methods tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
+Just these queries are utilized in inference, no XML data are used.
+The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} \todo{aj unique?} constructs.
 
 \subsection{Keys and foreign keys}
-Autori si zavádzajú definíciu pre key a foreign key:
+Authors introduce formalism that formalizes XML Schema keys.
 
 \begin{define}[Key]
-Key je trojica $$(C, P, \{L\})$$ kde $C$, $P$, $L$ sú XPath cesty bez predikátov a používajú len osi \texttt{child} a \texttt{descendant}.
-$C$ môže byť vynechané. $(P, \{L\})$ je v takom prípade to isté ako $(/, P, \{L\})$ a nazýva sa \emph{global key}, inak sa nazýva \emph{local key}.
+A key is a construct $$(C, P, \{L\})$$ where $C$, $P$, $L$ are XPath paths without predicates thet use only \texttt{child} and \texttt{descendant} axes.
+$C$ can be omitted. $(P, \{L\})$ is then equivalent to $(/, P, \{L\})$ and is called \emph{global key}.Otherwise it is called \emph{local key}.
 \end{define}
@@
 \begin{define}[Foreign key]
-Foreign key je konštrukcia $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ kde $(C, P_2, \{L_2\})$ je key, $P_1$, $L_1$ sú XPath cesty bez predikátov a používajú len osi \texttt{child} a \texttt{descendant}.
-$C$ môže byť vynechané podobne ako v prípade key.
+A foreign key is a construct $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ where $(C, P_2, \{L_2\})$ is a key, $P_1$, $L_1$ are XPath paths without predicates thet use only \texttt{child} and \texttt{descendant} axes.
+$C$ can be omitted as in the case of keys.
 \end{define}
@@
 \subsubsection{Joins in queries}
-Na hľadanie využíva metóda element/element joins v dotazoch.
+To discover keys and foreign keys, method utilizes element/element joins.
 
@@
 
-Metóda predpokladá, že každý join je robený cez pár key/foreign key.
-On the base of this assumption, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.
+Assuming that each join is done via a key/foreign key pair, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.
 
-Autori uvádzajú dve pozorovania:
+Authors introduces two observations:
 \begin{enumerate}
@@
 \subsubsection{Join patterns}
-Rozhodnutie pre jeden z prípadov (O1) a (O2) pre konrétny join je robené podľa tvaru tohto joinu.
-V dotaze sa hľadajú takzvané \emph{join patters}.
-Tie sú dva, \texttt{for} \emph{join pattern} a \texttt{let} \emph{join pattern} a vyzerajú takto (prvý je \texttt{for} \emph{join pattern}, druhý \texttt{let} \emph{join pattern}):
+For a certain join, the decision for one of the cases (O1) and (O2) is made by the form of the join.
+The query is searched for so-called \emph{join patters}.
+These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and look like the following (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
 \todo[inline]{spravne kreslit indexy pri pismenkach}
@@
 \section{Scoring function}
-Môže nastať situácia, že nie je splnený nejaký z predpokladov, z ktorých metóda vychádza.
-Predpokladajme nejakú množinu dotazov.
-Môže sa stať, že z nejakého dotazu z tejto množiny metóda odvodí key $K$ a z nejakého iného dotazu, že $K$ nie je key.
+It can happen that some of the assumptions the method is based on are not met.
+Assuming a set of queries, a key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key.
 \todo[inline]{priklad?}
 
-Autori preto zavádzajú key scoring function.
+Therefore, authors introduce a scoring function.
 If a new key $K$ is going to be inferred, it has assigned an initial score 0.
@@
 A negative score means that the key is not specified by the XML documents while positive means that it is satisfied.
-The absolute value of the score means how sure the method are about it.
+The absolute value of the score means how sure the method is about it.
 

### 2011-07-31 doc/theses/svirec/kap1.tex
introduction changes

@@
 They support some kind of semantic content (e.g., keys and foreign keys), but for improvement of semantic expressiveness XML integrity constraints for XML \cite{ic} have been defined.
-Several different classes of integrity constraints have been defined for XML, however most common semantic constraint used in relational databases is functional dependency (FD) \cite{fd}.
 
-With a defined set of functional dependencies in an XML document, one can check if an XML document satisfies defined functional dependencies \cite{satifFD}.
-However, data in the real world is typically dirty and often violate (does not satisfy) the set of functional dependencies and hence are inconsistent.
-This observation leads to a need of an algorithm to compute a repair applied to an XML document such that functional dependencies will be satisfied  again \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
+However, similarly to the problem of schemas also problem of detection of integrity constrains has two distinct aspects.
+First, if integrity constraints are not explicitly expressed, they need to be detected from the given set of data.
+Second, if integrity constraints are expressed, XML document may not be consistent with respect to them and this needs to be detected (checking the satisfaction of integrity constraint in an XML document).
+Moreover, detected XML data inconsistency needs to be repaired.
+We choose this very aspect of detection of integrity constraints along with the repair of the inconsistent XML document.
+Since several different classes of integrity constraints have been defined for XML, we choose functional dependency (FD) \cite{fd}, which is the most common semantic constraint used in relational databases.
+
+The problem of checking the satisfaction of functional dependencies in an XML document is studied in \cite{satifFD}.
+Also algorithms computing a repair applied to an XML document such that functional dependencies will be satisfied again have been proposed \cite{RepAndConsistentAnswer, QueryXML, ImprovingXML}.
 
@@
 
+As we mention before, we focus on the detection of XML data inconsistency with respect to the functional dependency and consequential repairing this violation.
 The aim of the thesis is to propose an algorithm repairing XML functional dependency violations for a given XML document and a set of functional dependencies.
-Analyses recent approaches and discusses their advantages and disadvantages.
+It analyses recent approaches and discusses their advantages and disadvantages.
 The focus of the thesis is to incorporate the weight model and to involve the user into the process of finding and applying the repair and find out how this interaction helps.

### 2011-08-01 doc/theses/svirec/kap1.tex
fixing

@@
 While XML as a markup language provides syntactic flexibility, the structure of an XML document is described with a so-called XML schema language.
-The two most popular XML schema languages proposed by W3C are Document Type Definition (DTD) \cite{xml} and XML Schema Definition (XSD) \cite{xsd1,xsd2,xsd3}.
-They support some kind of semantic content (e.g., keys and foreign keys), but for improvement of semantic expressiveness XML integrity constraints for XML \cite{ic} have been defined.
+The two most popular XML schema languages proposed by W3C\footnote{The World Wide Web Consorcium, \url{http://www.w3.org}} are Document Type Definition (DTD) \cite{xml} and XML Schema Definition (XSD) \cite{xsd1,xsd2,xsd3}.
+They support some kind of semantic content (e.g., keys and foreign keys), but for improvement of semantic expressiveness integrity constraints for XML \cite{ic} have been defined.
 
-However, similarly to the problem of schemas also problem of detection of integrity constrains has two distinct aspects.
+However, similarly to the problem of schemas, also the problem of detection of integrity constrains has two distinct aspects.
 First, if integrity constraints are not explicitly expressed, they need to be detected from the given set of data.
 Second, if integrity constraints are expressed, XML document may not be consistent with respect to them and this needs to be detected (checking the satisfaction of integrity constraint in an XML document).
-Moreover, detected XML data inconsistency needs to be repaired.
-We choose this very aspect of detection of integrity constraints along with the repair of the inconsistent XML document.
-Since several different classes of integrity constraints have been defined for XML, we choose functional dependency (FD) \cite{fd}, which is the most common semantic constraint used in relational databases.
+Moreover, the detected XML data inconsistency needs to be repaired.
+In this thesis we choose the latter aspect of detection of integrity constraints along with the repair of the inconsistent XML document.
+Since several different classes of integrity constraints have been defined for XML, we choose so-called functional dependency (FD) \cite{fd} which is the most common semantic constraint used in relational databases.
 
@@
 
-As we mention before, we focus on the detection of XML data inconsistency with respect to the functional dependency and consequential repairing this violation.
-The aim of the thesis is to propose an algorithm repairing XML functional dependency violations for a given XML document and a set of functional dependencies.
+The aim of this thesis is to propose an algorithm repairing XML functional dependency violations for a given XML document and a set of functional dependencies.
 It analyses recent approaches and discusses their advantages and disadvantages.
-The focus of the thesis is to incorporate the weight model and to involve the user into the process of finding and applying the repair and find out how this interaction helps.
-A part of this thesis is an experimental implementation of the proposed algorithm and its experimental evaluation.
+The focus of the thesis is to incorporate a weight model, to involve the user the process of finding and applying the repair and to find out how this interaction helps.
+An important part of the thesis is an experimental implementation of the proposed algorithm and its experimental evaluation.
 
@@
 Chapter 3 presents recent approaches of repairing XML functional dependency violations.
-In Chapter 4 is the main algorithm of this thesis proposed.
+In Chapter 4 the main algorithm is proposed.
 Chapter 5 contains details of the experimental implementation.

### 2011-08-01 doc/theses/svirec/kap5.tex
fixing

@@
 
-A part of this work is experimental implementation of our proposed algorithm.
+A part of this work is an experimental implementation of our proposed algorithm.
 Besides our approach, we have also implementened the algorithm proposed in \cite{RepAndConsistentAnswer}, to be able to compare it with our algorithm.
-Our algorithm has been implemented in the Java language in the version 6.0 as a part of the schema inference framework for XML jInfer \cite{jinfer}.
+Our algorithm has been implemented in the Java language, version 6.0 as a part of the schema inference framework for XML called jInfer \cite{jinfer}.
 
@@
 jInfer framework was developed as a Software Project at the Faculty of Mathematics and Physics, Charles University at Prague.
-The framework is based on NetBeans platform and is mainly used for XML schema inference.
-Although our proposed algorithm is not dealing with the schema inference, jInfer is designed to be extensible and is suitable for incorporation our algorithm into it.
+The framework is based on the NetBeans platform and is mainly used for XML schema inference.
+Although our proposed algorithm is not dealing with schema inference, jInfer is designed to be extensible and is suitable for incorporation of our algorithm.
 

### 2011-08-01 doc/theses/svirec/prace.tex
added abstract

@@
 \usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
@@
 \noindent
-Podakovanie.
+I would like to thank my supervisor RNDr. Irena Mlýnková, Ph.D., for her helpful advices, corrections and suggestions.
 
@@
 Název práce:
-\mftitle
+Efektívna detekcia integritných obmedzení v XML
 % přesně dle zadání
@@
 Vedoucí diplomové práce:
-\mfadvisor, Katedra softwarového inženýrství
+\mfadvisor
 
@@
 Abstrakt:
-Najviac používanými integritnými obmedzeniami v XML sú primárne a cudzie kľúče. Okrem nich však poznáme aj iný typ integritných obmedzení, a to funkčné závislosti. Tak ako pre relačné databázy,
+Znalosť integritných obmedzení v XML dátach je jeden z dôležitých aspektov ich spracovania.
+Avšak aj keď tieto integritné obmedzenia pre dané dáta poznáme, je častým javom, že dané dáta sú voči ním nekonzistentné.
+Z tohto dôvodu vznikla snaha detekovať tieto nekonzistentosti dát a následne ich opravovať.
+Táto práca rozširuje a zdokonaľuje doterajšie prístupy opráv XML dokumentov porušujúcich definované integritné obmedzenia, konkrétne takzvané funkčné závislosti.
+Práca prináša algoritmus začleňujúci váhový model a taktiež zapája užívateľa do procesu hľadania a následného aplikovania vhodnej opravy nekonzistentných XML dokumentov.
+Súčasťou práce sú experimentálne výsledky.
+
 % abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce
@@
 
-\vss}\nobreak\vbox to 0.49\vsize{
+\vss}
+
+\newpage
+
+\vbox to 0.5\vsize{
 \setlength\parindent{0mm}
@@
 Supervisor:
-\mfadvisor, \mfkatedra
+\mfadvisor
 % dle Organizační struktury MFF UK, případně plný název pracoviątě
@@
 Abstract:
-
+Knowledge of integrity constraints covered in the XML data is an important aspect of efficient data processing.
+However, although the integrity constraints are defined for the given data, it is a common phenomenon that data violate the predefined set of constraints.
+Therefore detection of these inconsistencies and consecutive repair has emerged.
+This work extends and refines recent approaches to repairing XML documents violating defined set of integrity constraints, specifically so-called functional dependencies.
+The work propose the repair algorithm incorporating the weight model and also involve user into the process of detection and subsequent application of appropriate repair of inconsistent XML documents.
+Experimental results are part of the work.
 % abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad

### 2011-08-01 doc/theses/svirec/prace.tex
fixing abstract

@@
 Abstract:
-Knowledge of integrity constraints covered in the XML data is an important aspect of efficient data processing.
-However, although the integrity constraints are defined for the given data, it is a common phenomenon that data violate the predefined set of constraints.
+Knowledge of integrity constraints covered in XML data is an important aspect of efficient data processing.
+However, although integrity constraints are defined for the given data, it is a common phenomenon that data violate the predefined set of constraints.
 Therefore detection of these inconsistencies and consecutive repair has emerged.
 This work extends and refines recent approaches to repairing XML documents violating defined set of integrity constraints, specifically so-called functional dependencies.
-The work propose the repair algorithm incorporating the weight model and also involve user into the process of detection and subsequent application of appropriate repair of inconsistent XML documents.
+The work proposes the repair algorithm incorporating the weight model and also involve a user into the process of detection and subsequent application of appropriate repair of inconsistent XML documents.
 Experimental results are part of the work.
@@
 
+\clearpage
+\phantomsection
+\addcontentsline{toc}{chapter}{Bibliography}
 %%% Seznam pouľité literatury

### 2011-08-01 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 \chapter{Analysis of Recent Approaches}
-\todo[inline]{Translate.}
-Many approaches of XML Schema inference have been proposed. Most of them process XML data as an input for inference. Some of these methods are mentioned in this chapter.
+\todo[inline]{bibref na prehladovy clanok?}
+Approaches of XML Schema inference can be divided by several criteria. Basic division is by language the resulting schema is written in. Commonly used languages are DTD and XML Schema.
 
-Besides approaches exploiting XML data, approaches that utilize operations over XML data may be developed and they are of the most interest in terms of this work.
-According to my best knowledge at the time of writing there is just one approach of this category.
-It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?}.
+The type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
+Heuristic methods are based on experience with manual construction of schemas and their result commonly does not belong to any class of grammar.
+On the contrary, result of the grammar-inferring methods belongs to a particular class of languages with specific characteristics.
+
+Another important criterion is the type of input data.
+Most of the approaches process XML documents as the input of the inference process and the documents are valid against the resulting schema.
+Besides approaches exploiting XML data, approaches that utilize operations over XML data may be developed and they are the most significant approaches in terms of this work.
+
+According to my best knowledge at the time of writing there is just one approach of this category. It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?}.
+
+\section{XML-Documents-utilizing Approaches}
+The process of inference commonly used by a significant number of approaches can be summarized as the following: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
+e_k$ is constructed.
+The productions form so-called \emph{initial grammar} (IG).
+For each element type the productions are then merged, simplified and generalized using various methods and criteria.
+A common approach is so-called \emph{merging state algorithm}, where a \emph{prefix tree automaton} (PTA) is built from the productions of the same element type and the automaton is generalized via merging of its states.
+Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
+
+\subsection{XTRACT}
+The XTRACT \todo{bibref} system is an example of \emph{heuristic} \emph{merging state algorithm} creating the result in DTD. Its process of inference consists of three steps:
+\begin{enumerate}
+\item Generalization - Generates a set of DTD candidates by searching the input for certain patterns and generalising corresponding parts using regular expressions.
+\item Factoring - Groups of generalized candidate DTDs are factorized to a new ones by finding common sub-expressions to make them more concise.
+\item Minimum Description Length (MDL) Principle - Selecting a near-optimal generalized DTD from the set of all generalized candidate DTDs.
+\end{enumerate}
+
+\todo[inline]{TODO}
 

### 2011-08-01 doc/theses/svirec/kap2.tex
thesis fixing

@@
 \begin{define}[Tree]
-Being given an alphabet of nodes $\mathbb{N}$ and an alphabet of node labels $\Sigma$, a {\sl tree} $T$ over $\mathbb{N}$ and $\Sigma$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is a set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$ which are connected and acyclic.
+Being given an alphabet of nodes $\mathbb{N}$ and an alphabet of node labels $\Sigma$, a {\sl tree} $T$ over $\mathbb{N}$ and $\Sigma$ is a tuple $(r_T, N_T, E_T, \lambda_T)$, where $N_T \subseteq \mathbb{N}$ is the set of nodes, $\lambda_T : N_T \to \Sigma$ is a node labelling function, $r_T \in N_T$ is the distinguished root of $T$, and $E_T \subseteq N_T \times N_T$ is a set of edges such that starting from any node $n_i \in N_T$ it is possible to reach any other node $n_j \in N_T$, walking through a sequence of edges $e_1,\dots,e_k$ which are connected and acyclic.\qed
 \end{define}
@@
 \begin{define}[XML Integrity Constraint]\label{integConstr}
-An XML constraint is a formula of the form: $$(\forall X)[\Phi(X)\supset (\exists Y_1)\Psi_1(X,Y_1) \lor \dots \lor (\exists Y_k)(X,Y_k)]$$
+An {\sl XML constraint} is a formula of the form: $$(\forall X)[\Phi(X)\supset (\exists Y_1)\Psi_1(X,Y_1) \lor \dots \lor (\exists Y_k)(X,Y_k)]$$
 where $X,Y_1,\dots, Y_k$ denote distinct sets of universally and existentially quantified variables, $\Phi(X)$ and $\Phi(X) \land \Psi_i(X, Y_i) (\forall i \in [1..k])$ are safe conjunctions of built-in and tree atoms.\qed
@@
     \centering\includegraphics[width=\textwidth]{tuples}
-	\caption{Two tree tuples of the XML tree in Fig. \ref{example1}} \label{tuples}
+	\caption[Two tree tuples of the XML tree]{Two tree tuples of the XML tree in Fig. \ref{example1}} \label{tuples}
 \end{figure}
@@
     \subfloat[]{\label{nottuple2}\includegraphics[scale=0.6]{nottuples2}}
-	\caption{Two subtrees of the XML tree in Fig. \ref{example1} which are not tree tuples}
+	\caption[Two subtrees of the XML tree]{Two subtrees of the XML tree in Fig. \ref{example1} which are not tree tuples}
     \label{nottuples}

### 2011-08-01 doc/theses/svirec/kap3.tex
thesis fixing

@@
 To be able to resolve the problem of functional dependency violations in XML document, the authors try to introduce the concept of functional dependencies based on those defined for relational databases.
-The concept of a tree tuple and functional dependency was introduced in definition \ref{treeTuple} and \ref{fd1}.
+The concept of a tree tuple and functional dependency was introduced in Definition \ref{treeTuple} and \ref{fd1}.
 
@@
 Let $RXT = \langle T, \delta, \varrho \rangle$ be an R-XML tree conforming a DTD $D$, and $f: S \rightarrow p$ be a functional dependency.
-We say that $RXT$ weakly satisfies $f$ ($RXT \models_w f$) if one of the following conditions holds:
+We say that $RXT$ {\sl weakly satisfies} $f$ ($RXT \models_w f$) if one of the following conditions holds:
 \begin{enumerate}
@@
 \begin{define}[Composition of $\delta$ functions]\label{compDelta}
-The composition of two functions $\delta_1$ and $\delta_2$ asssociating values to leaf nodes is
+The {\sl composition} of two functions $\delta_1$ and $\delta_2$ asssociating values to leaf nodes is
 \begin{displaymath}
@@
 \begin{define}[Composition of $\varrho$ functions]\label{compVarrho}
-The composition of two reliability functions $\varrho_1$ and $\varrho_2$ asssociating a boolean value to nodes is $$\varrho_1 \cdot \varrho_2(n) = \varrho_1(n)\ AND\ \varrho_2(n).$$\qed
+The {\sl composition} of two reliability functions $\varrho_1$ and $\varrho_2$ asssociating a boolean value to nodes is $$\varrho_1 \cdot \varrho_2(n) = \varrho_1(n)\ AND\ \varrho_2(n).$$\qed
 \end{define}
@@
 We say that $R_1$ is smaller than $R_2$ ($R_1 \preceq R_2$) if $Updated_{\delta_1}(N_T)\ \cup\ False_{\delta_1}(N_T) \subseteq Updated_{\delta_2}(N_T)\ \cup\ False_{\delta_2}(N_T)$ and $False_{\delta_1}(N_T) \subseteq False_{\delta_2}(N_T)$.
-Repair $R$ is minimal if there is no repair $R' \neq R$ such that $R' \preceq R$.\qed
+Repair $R$ is {\sl minimal} if there is no repair $R' \neq R$ such that $R' \preceq R$.\qed
 \end{define}
@@
 \begin{define}[Repair]
-Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$ (Definition \ref{integConstr}), a set of update operations $R$ is said to be a repair of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
+Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$ (Definition \ref{integConstr}), a set of update operations $R$ is said to be a {\sl repair} of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
 \end{define}

### 2011-08-01 doc/theses/svirec/kap4.tex
thesis fixing

@@
 
-After creating the RW-XML tree from input XML data, a set of tree tuples (defined in Definition r\ref{treeTuple}) is constructed.
+After creating the RW-XML tree from input XML data, a set of tree tuples (defined in Definition \ref{treeTuple}) is constructed.
 Since the set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree.
@@
 \begin{define}[Repair Candidate Cost]
-Being given an RW-XML tree $RXT$ and a repair candidate $R$, we define the cost of $R$ as:
+Being given an RW-XML tree $RXT$ and a repair candidate $R$, we define the {\sl cost} of $R$ as:
 $$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \cdot k,$$

### 2011-08-02 doc/theses/svirec/appendixA.tex
some fixes

@@
 The CD is a part of this thesis.
-It contains text of this thesis, source code of jInfer framework with \texttt{FDRepairer} extension together with documentation and built jInfer as \texttt{nbm} plugins applicable into NetBeans platform.
+It contains text of this thesis, source code of jInfer framework with \texttt{FDRepairer} module together with documentation and built jInfer as \texttt{nbm} plugins applicable into NetBeans platform.
 It aslo contains experimental data and results.

### 2011-08-02 doc/theses/svirec/kap3.tex
some fixes

@@
 \begin{define}[Repair]
-Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$ (Definition \ref{integConstr}), a set of update operations $R$ is said to be a {\sl repair} of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.
+Given an XML tree $T$, a DTD $D$ and a set of integrity constraints $IC$ (Definition \ref{integConstr}), a set of update operations $R$ is said to be a {\sl repair} of $T$ (with respect to $D$ and $IC$) if $R(T) \models IC$, where $R(T)$ is application of consistent set of updates $R$ to $T$, $R(T)$ conforms $D$ and $\not \exists R' \sqsubset R$ such that $R'(T) \models IC$ and $R'(T)$ conforms $D$.\qed
 \end{define}
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
 \caption{Fix-Initial-Conflicts}

### 2011-08-02 doc/theses/svirec/kap6.tex
thesis fixing
some fixes

@@
 
-In this chapter, experimental results of the \texttt{FDRepairer} extension of the jInfer framework are presented.
+In this chapter, experimental results of the \texttt{FDRepairer} module of the jInfer framework are presented.
 Set of documents and functional dependencies have been provided as an input and results from the former algorithm have been compared with results from \texttt{FDRepairer}.
@@
 \ref{actorsDTD}.
-For this dataset, we defined FD as follows: For each two authors, which played in the movie with the same name, the year of release of this movie must be the same.
+For this dataset, we defined FD as follows: \emph{For each two authors, which played in the movie with the same name, the year of release of this movie must be the same}.
 
@@
     \end{verbatim}
-    \caption{DTD of actors.xml}
+    \caption{DTD of customers.xml}
     \label{synthDTD}

### 2011-08-02 doc/theses/svirec/kap6.tex
fixing experiments

@@
 First two use the default value of $k$ ($k=1.5$), for the last one, we set $k=0.005$.
-For the second (third) execution, we set the weight $w$ of nodes represented by XPath \texttt{/root/course/time/start\_time/text()} (resp.
-\texttt{/root/course/time/\discretionary{}{}{}start/text()}) to 0.1.
+For the last two executions, we set for \texttt{reed.xml} (\texttt{wsu.xml}) the weight $w$ of nodes represented by XPath \texttt{/root/course/time/start\_time/text()} (resp.
+\texttt{/root/\discretionary{}{}{}course/time/\discretionary{}{}{}start/text()}) to 0.1.
 
@@
 
-Setting the weight to particular nodes effects the choise of strategies used to repair the violations.
-In datasets \texttt{reed.xml} and \texttt{wsu.xml} setting the custom weight to nodes causes the node values are changed to values of other nodes.
-On the other side, in datasets \texttt{actors.xml} and \texttt{customers.xml} setting the weights causes changing the value of nodes to a newly generated one.
+Setting the weight to particular nodes causes two different behaviour of the algorithm.
+First, setting the weights effects the choise of strategies used to repair the violations.
+In datasets \texttt{actors.xml} and \texttt{customers.xml} setting the custom weight to nodes causes the node values are changed to values of other nodes.
+Second causes modification of nodes defined by the path specifying the weights.
+In datasets \texttt{reed.xml} and \texttt{wsu.xml} setting the weights still changes the value of nodes to a newly generated one, however for \texttt{reed.xml} only values of nodes defined by \texttt{/root/course/time/start\_time/\discretionary{}{}{}text()} are modified and for \texttt{wsu.xml} are modified values defined by \texttt{root/\discretionary{}{}{}course/time/start/text()}.
 

### 2011-08-02 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 \begin{enumerate}
-\item Generalization - Generates a set of DTD candidates by searching the input for certain patterns and generalising corresponding parts using regular expressions.
+\item Generalization - Generates a set of DTD candidates by searching the input for certain patterns and generalising corresponding fragments using regular expressions.
 \item Factoring - Groups of generalized candidate DTDs are factorized to a new ones by finding common sub-expressions to make them more concise.
-\item Minimum Description Length (MDL) Principle - Selecting a near-optimal generalized DTD from the set of all generalized candidate DTDs.
+\item Minimum Description Length (MDL) Principle - Composing a near-optimal DTD schema from the set of all generalized candidate DTDs.
 \end{enumerate}
 
-\todo[inline]{TODO}
+MDL is the most important step.
+It ranks each DTD candidate using two aspects - conciseness and preciseness.
+Conciseness of a DTD is expressed using the number of bits required to describe the DTD (the smaller, the better).
+Preciseness of a DTD is expressed using the number of bits required for description of the input data using the DTD.
+In other words, the more accurately the structure is described, the fewer bits are required.
+Since the two conditions are contradictory, their balancing brings reasonable and realistic results.
+
+\subsection{SchemaMiner}
+The SchemaMiner system \todo{bibref} is an example of \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
+It focusses on inferring elements with the same name but different structure and unordered sequences.
+
+The elements with the same name but different content are supported only in XSDs and their inference requires exploitation of more sophisticated approaches than combining productions with the same element
+type. On the other hand, although the unordered sequences are a classical example of XSD “syntactic sugar”, their exploitation enables to define simple and, hence, realistic and usable schemas.
+
+\todo[inline]{Moze byt takto? Nejake dalsie pristupy?}
 

### 2011-08-06 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 
-This work will be building upon the achievements of jInfer schema inference framework, expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
+This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
 
@@
 
-The pseudocode is listed (cited from \cite{fidax} with trivial modifications) in \ref{ch-fidax}.
+The pseudocode is listed (cited from \cite{fidax} with trivial modifications without changing the logic) in \ref{ch-fidax}.
 
@@
 
+\section{IDREF}
+
+TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
+
 \chapter{Experiments}
@@
 TODO we will be using jInfer and its module AttributeStats see \ref{jInfer} and \ref{attrstats}.
+TODO describe how to create a model, describe how we get IGG, link BasicIGG
 
@@
 \begin{itemize}
+	\item For each input file: how long does model creation take?
+	\item How long does it take to communicate with GLPK? Creating the input/parsing the output.
 	\item Which construction heuristic is the best - limit GLPK to like 1 second

### 2011-08-06 doc/theses/vitasek/diplomka.tex
Thesis: in progress
Thesis: Appendix describing jInfer.

@@
 
-TODO what it is, who created it and why, what it can do, where to get it.
+This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
+Please see TODO link jInfer project web for complete information, documentation and download options.
+
+jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Robert Smetana, Michal Svirec and Matej Vitasek.
+The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
+The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
+Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
+This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
+
+At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
+Most importantly, the whole process consists of 3 consecutive steps (see TODO link obrazok), responsibility of 3 different services - interchangeable modules.
+
+The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
+This is the representation in which will the structure live until it is used to create the final product - the schema.
+As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
+As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
+(TODO picture?) IG is used to create the AM model used in this thesis, too.
+jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
+
+After leaving the \jmodule{Initial Grammar Generator}, the IG needs to be made more general, shortened, \textit{simplified}.
+This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
+To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
+Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
+
+The last module, \jmodule{Schema Generator} takes the simplified grammar and creates the resulting schema from it.
+This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
+
+TODO elaborate more, if too short:)
 

### 2011-08-08 doc/theses/vitasek/diplomka.tex
Thesis: adding schema of inference.

@@
 At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see TODO link obrazok), responsibility of 3 different services - interchangeable modules.
+Most importantly, the whole process consists of 3 consecutive steps (see TODO link picture - inference process), responsibility of 3 different services - interchangeable modules.
 

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/AttributeStatisticsImpl.java

@@ -1,69 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.gui.StatisticsPanel;
-import cz.cuni.mff.ksi.jinfer.attrstats.gui.AttrStatsTopComponent;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.GraphUtils;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.AttributeStatistics;
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
-import java.util.List;
-import org.openide.util.lookup.ServiceProvider;
-
-/**
- * Implementation of the Attribute Statistics service.
- *
- * @author vektor
- */
-@ServiceProvider(service = AttributeStatistics.class)
-public class AttributeStatisticsImpl implements AttributeStatistics {
-
-  private static final String NAME = "AttributeStatistics";
-  private static final String DISPLAY_NAME = "Attribute Statistics";
-
-  @Override
-  public void showStatistics(final String panelName, final List<Element> grammar) {
-    final AttrStatsTopComponent topComponent = AttrStatsTopComponent.findInstance();
-    if (!topComponent.isOpened()) {
-      topComponent.open();
-    }
-    final StatisticsPanel panel = topComponent.createNewPanel(panelName);
-    panel.setModel(grammar);
-  }
-
-  @Override
-  public String getGraphVizInput(final List<Element> grammar) {
-    return GraphUtils.getGraphVizInput(new AMModel(grammar));
-  }
-
-  @Override
-  public String getName() {
-    return NAME;
-  }
-
-  @Override
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/AbstractExperimentSet.java

@@ -1,101 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ExperimentListener;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Constants;
-import java.io.File;
-import java.util.List;
-import org.apache.log4j.Logger;
-
-/**
- * Abstract class representing an experiment set. After it gets the list of
- * experiments (via {@link ExperimentParameters}) to run, it starts to execute
- * them sequentially.
- *
- * @author vektor
- */
-public abstract class AbstractExperimentSet {
-
-  private final Object monitor = new Object();
-
-  /**
-   * Should return the name of this experimental set.
-   *
-   * @return Name of this set.
-   */
-  public abstract String getName();
-
-  /**
-   * Should return the list of experiment parameters to be run in this set.
-   *
-   * @return List of experiment parameters that will be sequentially run in this
-   * experimental set.
-   */
-  protected abstract List<ExperimentParameters> getExperiments();
-
-  private static final Logger LOG = Logger.getLogger(AbstractExperimentSet.class);
-
-  public void run() throws InterruptedException {
-    int i = 0;
-    for (final ExperimentParameters params : getExperiments()) {
-      final Experiment e = new Experiment(params);
-      final int index = i++;
-      e.addListener(new ExperimentListener() {
-
-        @Override
-        public void experimentFinished(final Experiment e) {
-          logResults(e, index);
-          synchronized (monitor) {
-            monitor.notifyAll();
-          }
-        }
-      });
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/InputFile.java

@@ -1,105 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments;
-
-import java.io.File;
-
-/**
- * Class representing an file being used as an input in an experiment.
- *
- * @author vektor
- */
-public class InputFile {
-
-  /** The file itself. */
-  private final File file;
-
-  /** General characteristic of the file. */
-  private final FileCharacteristics characteristics;
-
-  /**
-   * Full constructor.
-   *
-   * @param file String representing the file name.
-   * @param characteristics General characteristic of the file.
-   */
-  public InputFile(final String file, final FileCharacteristics characteristics) {
-    this(new File(file), characteristics);
-  }
-
-  /**
-   * Full constructor.
-   *
-   * @param file Object representing the file itself.
-   * @param characteristics General characteristic of the file.
-   */
-  public InputFile(final File file, final FileCharacteristics characteristics) {
-    this.file = file;
-    this.characteristics = characteristics;
-  }
-
-  /**
-   * Partial constructor. Assumes the file to be of unknown type
-   * ({@link FileCharacteristics#UNKNOWN}).
-   *
-   * @param file String representing the file name.
-   */
-  public InputFile(final String file) {
-    this(new File(file), FileCharacteristics.UNKNOWN);
-  }
-
-  /**
-   * Partial constructor. Assumes the file to be of unknown type
-   * ({@link FileCharacteristics#UNKNOWN}).
-   *
-   * @param file Object representing the file itself.
-   */
-  public InputFile(final File file) {
-    this(file, FileCharacteristics.UNKNOWN);
-  }
-
-  /**
-   * Returns the object representing this file.
-   */
-  public File getFile() {
-    return file;
-  }
-
-  /**
-   * Returns the general characteristic of this file.
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/SizeTestData.java

@@ -1,84 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Constants;
-import static cz.cuni.mff.ksi.jinfer.attrstats.experiments.FileCharacteristics.*;
-
-/**
- * A set of artificial data, random graphs. They should have approximately the
- * same density (~1/10 of all possible edges).
- *
- * @author vektor
- */
-public enum SizeTestData {
-
-  GRAPH_0_0      ("0-0.xml",     0, 0,    0d),
-  GRAPH_10_5     ("10-5.xml",    10, 5,   0.8500000000000002),
-  GRAPH_20_20    ("20-20.xml",   18, 13,  0.7166666666666669),
-  GRAPH_30_45    ("30-45.xml",   29, 43,  0.7083333333333334),
-  GRAPH_40_80    ("40-80.xml",   39, 72,  0.6950000000000002),
-  GRAPH_50_125   ("50-125.xml",  48, 111, 0.6566666666666666),
-  GRAPH_60_180   ("60-180.xml",  58, 157, 0.6214285714285716),
-  GRAPH_70_245   ("70-245.xml",  67, 205, 0.5982142857142856),
-  GRAPH_80_320   ("80-320.xml",  76, 261, 0.5791666666666667),
-  GRAPH_90_405   ("90-405.xml",  86, 352, 0.528888888888889),
-  GRAPH_100_500  ("100-500.xml", 91, 388, 0.4981818181818182);
-
-  // anything more takes 1800+ seconds...
-
-  private final InputFile file;
-  // Graph representation
-  private final int vertices;
-  private final int edges;
-  // Known optimum for alpha = beta = 1
-  private final Double optimum;
-
-  private SizeTestData(final String fileName,
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/SystemInfo.java

@@ -1,93 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkUtils;
-
-/**
- * Provides the info about the system (CPU, RAM, OS, Java, etc) in textual form.
- *
- * Info that is not easily obtained from within Java has to be set in constants,
- * though.
- *
- * @author vektor
- */
-public final class SystemInfo {
-
-  private SystemInfo() {
-  }
-
-  /** Number of processor cores. */
-  private static final int PROCESSOR_CORES = 4;
-  /** Clock speed in MHz. */
-  private static final int PROCESSOR_CLOCK_SPEED = 2983;
-  /** Processor architecture/family. */
-  private static final String PROCESSOR_ARCH = "Intel(R) Core(TM)2 Quad  CPU   Q9550  @ 2.83GHz";
-  /** RAM size in MB. */
-  private static final int RAM_SIZE = 8192;
-
-  /** Number of processor cores. */
-  //private static final int PROCESSOR_CORES = 2;
-  /** Clock speed in MHz. */
-  //private static final int PROCESSOR_CLOCK_SPEED = 2330;
-  /** Processor architecture/family. */
-  //private static final String PROCESSOR_ARCH = "Intel(R) Core(TM)2 Duo CPU     E6550  @ 2.33GHz";
-  /** RAM size in MB. */
-  //private static final int RAM_SIZE = 4096;
-
-  /** OS name. */
-  private static final String OS_NAME = System.getProperty("os.name");
-  /** OS architecture. */
-  private static final String OS_ARCH = System.getProperty("os.arch");
-  /** OS version. */
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/interfaces/ImprovementHeuristic.java

@@ -1,46 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import java.util.List;
-
-/**
- * Interface representing an improvement heuristic. Its responsibility is to
- * start with a pool (possibly just one) of feasible solutions and improve one
- * or more of them, or add new ones.
- *
- * Note that heuristics use the callback "design pattern", meaning that the
- * result of this heuristic is returned via a callback provided while invoking it.
- *
- * @author vektor
- */
-public interface ImprovementHeuristic extends Heuristics {
-
-  /**
-   * Start the heuristic run. Use the information in the provided experiment to
-   * improve the provided pool (possibly only one) of feasible solutions.
-   *
-   * @param experiment Experiment in context of which to run the heuristic.
-   * @param feasiblePool List of feasible solutions found so far.
-   * @param callback Callback to be invoked when finished.
-   */
-  void start(final Experiment experiment, final List<IdSet> feasiblePool,
-          final HeuristicCallback callback) throws InterruptedException;
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/quality/Weight.java

@@ -1,38 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.QualityMeasurement;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-
-/**
- * Weight-based implementation of the {@link QualityMeasurement} interface.
- *
- * @author vektor
- */
-public class Weight implements QualityMeasurement {
-
-  @Override
-  public Quality measure(final Experiment experiment, final IdSet solution) {
-    return new Quality(
-            experiment.getModel().weight(solution.getMappings(), experiment.getAlpha(), experiment.getBeta()),
-            solution.getMappings().size(),
-            solution.isOptimal());
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/sets/PingAll.java

@@ -1,56 +1,0 @@
-/*
- * Copyright (C) 2011 vitasek
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.sets;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.OfficialTestData;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Weight;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.Glpk;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.Identity;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-
-/**
- *
- * @author vitasek
- */
-public class PingAll extends AbstractExperimentSet {
-
-  @Override
-  public String getName() {
-    return "Ping all";
-  }
-
-  @Override
-  protected List<ExperimentParameters> getExperiments() {
-    final List<ImprovementHeuristic> improvement = Arrays.<ImprovementHeuristic>asList(new Identity());
-
-    final List<ExperimentParameters> ret = new ArrayList<ExperimentParameters>(10);
-
-    for (final OfficialTestData data : OfficialTestData.values()) {
-      ret.add(new ExperimentParameters(data.getFile(), 1, 1, 1, data.getKnownOptimum(),
-              new Glpk(0), improvement, new Weight(), new TimeIterations(1, 10000)));
-    }
-
-    return ret;
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/sets/TenIterations.java

@@ -1,64 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.sets;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Weight;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.Random;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.Crossover;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.Mutation;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.RandomRemove;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.RemoveWorst;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Constants;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-
-/**
- * Simply runs the same experiment ten times.
- *
- * @author vektor
- */
-public class TenIterations extends AbstractExperimentSet {
-
-  @Override
-  public String getName() {
-    return "Ten Iteration";
-  }
-
-  @Override
-  protected List<ExperimentParameters> getExperiments() {
-    final List<ImprovementHeuristic> improvement = Arrays.<ImprovementHeuristic>asList(
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/sets/VariousBetas.java

@@ -1,69 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.sets;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Weight;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.Random;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.Crossover;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.Mutation;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.RandomRemove;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.RemoveWorst;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Constants;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-
-/**
- * Experiment comparing effects of various values of parameters alpha and beta.
- *
- * @author vektor
- */
-public class VariousBetas extends AbstractExperimentSet {
-
-  private static final double[] alphas = new double[] {0.1, 0.5, 1.0};
-  private static final double[] betas = new double[] {0.1, 0.5, 1.0};
-
-  @Override
-  public String getName() {
-    return "Various Betas";
-  }
-
-  @Override
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/termination/TimeIterations.java

@@ -1,91 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.termination;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Quality;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.TerminationCriterion;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.Pair;
-import cz.cuni.mff.ksi.jinfer.base.objects.ImmutablePair;
-import java.util.List;
-
-/**
- * An implementation of the {@link TerminationCriterion} based on the total time
- * spent running the metaheuristics and the number of iterations - improvement
- * heuristics runs. Also, as soon as known optimum is reached, this criterion
- * is met.
- *
- * @author vektor
- */
-public class TimeIterations implements TerminationCriterion {
-
-  private static final double THRESHOLD = 0.0000001;
-  private final int maxIterations;
-  private int iterations = 0;
-  private final long maxTime;
-
-  /**
-   * Constructor. Initializes this criterion to run at most the specified
-   * number of iterations, max time is set to "infinity".
-   *
-   * @param maxIterations Maximum number of iterations - improvement
-   * heuristics runs allowed.
-   */
-  public TimeIterations(final int maxIterations) {
-    this(maxIterations, Long.MAX_VALUE);
-  }
-
-  /**
-   * Full constructor. Initializes this criterion to run at most the specified
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/gui/IdSetPanel.java

@@ -1,186 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.gui;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.idref.IdRefSearch;
-import java.util.List;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.tables.MappingsModel;
-import java.util.Collections;
-import javax.swing.JPanel;
-import javax.swing.table.DefaultTableModel;
-import static cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils.NA;
-
-/**
- * Panel displaying one ID set in a table.
- *
- * @author vektor
- */
-@SuppressWarnings("PMD.SingularField")
-public class IdSetPanel extends JPanel {
-
-  private static final long serialVersionUID = 187541L;
-
-  private AMModel model;
-  private IdSet idSet;
-  private List<AttributeMappingId> idRefCache;
-
-  public IdSetPanel() {
-    super();
-    initComponents();
-    table.setModel(new DefaultTableModel());
-  }
-
-  /**
-   * Sets the model in this panel, thus displaying the list of AMs in the
-   * provided ID set.
-   *
-   * @param idSet ID set - list of {@link AttributeMappingId attribute mappings}
-   * to be displayed.
-   * @param model Model in context of which this ID set was found.
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/gui/StatisticsPanel.java

@@ -1,384 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.gui;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.options.AttrStatsPanel;
-import org.openide.util.NbPreferences;
-import org.netbeans.api.options.OptionsDisplayer;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.JFCWrapper;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkInputGenerator;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkOutputParser;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkRunner;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.fidax.FidaxAlgorithm;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeTreeNode;
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.util.ArrayList;
-import java.util.List;
-import javax.swing.JPanel;
-import javax.swing.tree.DefaultTreeModel;
-import javax.swing.tree.TreePath;
-import org.apache.log4j.Logger;
-import org.openide.DialogDisplayer;
-import org.openide.NotifyDescriptor;
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/Fuzzy.java

@@ -1,102 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ConstructionHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.WeightedRandomGenerator;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.util.ArrayList;
-import java.util.List;
-
-/**
- * This construction heuristics builds up the solution gradually by adding AMs.
- * AMs to be added are picked at weighted random, based on their weight.
- *
- * @author vektor
- */
-public class Fuzzy implements ConstructionHeuristic {
-
-  @Override
-  public void start(final Experiment experiment,
-        final HeuristicCallback callback) throws InterruptedException {
-
-    final List<AttributeMappingId> candidates = MappingUtils.getCandidates(experiment.getModel());
-    final List<Double> weights = new ArrayList<Double>(candidates.size());
-    for (final AttributeMappingId mapping : candidates) {
-      weights.add(Double.valueOf(experiment.getModel().weight(mapping, experiment.getAlpha(), experiment.getBeta())));
-    }
-
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/fidax/Fidax.java

@@ -1,58 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.fidax;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ConstructionHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import java.util.Arrays;
-
-/**
- * Construction heuristic - encapsulation of the heuristic proposed in the
- * "Finding ID Attributes in XML Documents" article.
- * 
- * Note that the pool created by this heuristic contains only one solution
- * ({@link IdSet}).
- *
- * @author vektor
- */
-public class Fidax implements ConstructionHeuristic {
-
-  @Override
-  public void start(final Experiment experiment,
-        final HeuristicCallback callback) throws InterruptedException {
-    final IdSet solution = FidaxAlgorithm.findIDSet(experiment.getModel(), experiment.getAlpha(), experiment.getBeta());
-    callback.finished(Arrays.asList(solution));
-  }
-
-  @Override
-  public String getName() {
-    return "FIDAX";
-  }
-
-  @Override
-  public String getDisplayName() {
-    return getName();
-  }
-
-  @Override
-  public String getModuleDescription() {
-    return getName();
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkOutputParser.java

@@ -1,93 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import java.util.ArrayList;
-import java.util.HashSet;
-import java.util.Scanner;
-import java.util.Set;
-
-/**
- * Utility class for parsing the GLPK optimization result into a set of attribute
- * mappings from the "optimal" ID set.
- *
- * @author vektor
- */
-public final class GlpkOutputParser {
-
-  private GlpkOutputParser() {
-  }
-
-  /**
-   * Parses the string representing the GLPK optimization result and returns
-   * a list of attribute mappings belonging to the "optimal" ID set.
-   *
-   * @param output String representation of the GLPK result. Basically the file
-   * that GLPK produces loaded into a string.
-   * @param model Model to which all the AMs belong.
-   *
-   * @return List of AMs belonging to the "optimal" ID set.
-   */
-  public static IdSet getIDSet(final String output, final AMModel model) {
-    return getIDSet(output, model, false);
-  }
-
-  /**
-   * Parses the string representing the GLPK optimization result and returns
-   * a list of attribute mappings belonging to the "optimal" ID set.
-   *
-   * @param output String representation of the GLPK result. Basically the file
-   * that GLPK produces loaded into a string.
-   * @param model Model to which all the AMs belong.
-   * @param isFixed Flag whether there were any fixed variables. If
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkUtils.java

@@ -1,202 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.options.AttrStatsPanel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.base.utils.FileUtils;
-import java.io.BufferedReader;
-import java.io.File;
-import java.io.IOException;
-import java.io.InputStream;
-import java.io.InputStreamReader;
-import java.io.PrintWriter;
-import java.io.Reader;
-import org.apache.log4j.Logger;
-import org.openide.util.NbPreferences;
-
-/**
- * Utility class for GLPK binding.
- *
- * @author vektor
- */
-public final class GlpkUtils {
-
-  private static final String CMD_OPTS = "-v";
-  private static final String EXPECTED = "GLPSOL: GLPK LP/MIP Solver";
-  private static final String TMP = System.getProperty("java.io.tmpdir");
-  public static final String INPUT = TMP + "/glpk_input.txt";
-  public static final String OUTPUT = TMP + "/glpk_output.txt";
-
-  private static final Logger LOG = Logger.getLogger(GlpkUtils.class);
-
-  private GlpkUtils() {
-
-  }
-
-  /**
-   * Returns the full path to the GLPK Solver binary in a String.
-   *
-   * @return Full path to the GLPK Solver binary.
-   */
-  public static String getPath() {
-    return NbPreferences.forModule(AttrStatsPanel.class).get(AttrStatsPanel.BINARY_PATH_PROP, AttrStatsPanel.BINARY_PATH_DEFAULT);
-  }
-
-  /**
-   * Verifies whether the GLPK binary is valid. See
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/improvement/LocalBranching.java

@@ -1,105 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Quality;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkInputGenerator;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkOutputParser;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkRunner;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.Pair;
-import java.util.ArrayList;
-import java.util.List;
-
-/**
- * This heuristic works by creating a constrained sub-problem and then solving
- * it. The constraint is as follows: the maximum Hamming distance between the
- * old and new solution is limited to a constant <code>k</code>.
- *
- * Some implementation details:
- * <ul>
- *  <li>When working with AMs and ID sets, the Hamming distance is defined
- *    as follows: sum over all AMs of (1 if solutions differ at this AM,
- *    0 otherwise).</li>
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/improvement/RandomRemove.java

@@ -1,72 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.util.ArrayList;
-import java.util.List;
-
-/**
- * A simple improvement heuristic that removes a specified fraction of AMs at
- * random from every solution in the pool.
- *
- * @author vektor
- */
-public class RandomRemove implements ImprovementHeuristic {
-
-  private final double ratio;
-
-  public RandomRemove(final double ratio) {
-    super();
-    if (ratio < 0 || ratio >= 1) {
-      throw new IllegalArgumentException("Invalid ratio: " + ratio);
-    }
-    this.ratio = ratio;
-  }
-
-  @Override
-  public void start(final Experiment experiment, final List<IdSet> feasiblePool,
-        final HeuristicCallback callback) throws InterruptedException {
-
-    final List<IdSet> ret = new ArrayList<IdSet>(feasiblePool.size());
-    for (final IdSet idSet : feasiblePool) {
-      ret.add(new IdSet(BaseUtils.rndSubset(idSet.getMappings(), 1 - ratio)));
-    }
-
-    callback.finished(ret);
-  }
-
-  @Override
-  public String getName() {
-    return "RandomRemove";
-  }
-
-  @Override
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/improvement/RemoveWorst.java

@@ -1,57 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import java.util.List;
-
-/**
- * Simple "improvement" heuristic that works by removing the worst solution
- * ({@link IdSet}) from the feasible pool.
- *
- * @author vektor
- */
-public class RemoveWorst implements ImprovementHeuristic {
-
-  @Override
-  public void start(final Experiment experiment, final List<IdSet> feasiblePool,
-        final HeuristicCallback callback) throws InterruptedException {
-    final IdSet worst = Utils.getWorst(experiment, feasiblePool);
-    feasiblePool.remove(worst);
-    callback.finished(feasiblePool);
-  }
-
-  @Override
-  public String getName() {
-    return "RemoveWorst";
-  }
-
-  @Override
-  public String getDisplayName() {
-    return "Remove Worst";
-  }
-
-  @Override
-  public String getModuleDescription() {
-    return getDisplayName();
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/objects/AttributeMappingId.java

@@ -1,105 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.objects;
-
-/**
- * This is a representation of an attribute mapping identifier, as described in
- * the article "Finding ID Attributes in XML Documents". It contains the
- * element and attribute names. This representation follows the immutable object
- * design pattern.
- *
- * @author vektor
- */
-public class AttributeMappingId implements Comparable<AttributeMappingId> {
-
-  private final String element;
-
-  private final String attribute;
-
-  /**
-   * Full constructor.
-   *
-   * @param element Element name.
-   * @param attribute Attribute name.
-   */
-  public AttributeMappingId(final String element, final String attribute) {
-    this.element = element;
-    this.attribute = attribute;
-  }
-
-  public String getElement() {
-    return element;
-  }
-
-  /**
-   * Checks whether this object represents an attribute mapping of the specified
-   * type (whether the element has the speficied name). This check is
-   * case-sensitive.
-   *
-   * @param type Type to check.
-   * @return <code>true</code> if this mapping is of the specified type,
-   * <code>false</code> otherwise.
-   */
-  public boolean isType(final String type) {
-    return getElement().equals(type);
-  }
-
-  public String getAttribute() {
-    return attribute;
-  }
-
-  /**
-   * Two attribute mapping IDs are equal iff their element and attribute names
-   * are equal (case sensitive).
-   *
-   * @param obj Other object to check equality.
-   * @return <code>true</code> if the other object is {@link AttributeMappingId}
-   * and its element and attribute names are equal to ones in this mapping,
-   * <code>false<code> otherwise.
-   */
-  @Override
-  public boolean equals(final Object obj) {
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/objects/AttributeTreeNode.java

@@ -1,67 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.objects;
-
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Attribute;
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
-import java.util.List;
-import javax.swing.JTree;
-import javax.swing.tree.DefaultMutableTreeNode;
-
-/**
- * Class representing an {@link Attribute} in a {@link JTree}.
- *
- * @author vektor
- */
-public class AttributeTreeNode extends DefaultMutableTreeNode {
-
-  private static final long serialVersionUID = 454563134L;
-
-  private final String elementName;
-
-  private final String attributeName;
-
-  private final List<String> content;
-
-  /**
-   * Full constructor.
-   *
-   * @param elementName Name of the {@link Element} where the represented {@link Attribute} resides.
-   * @param attributeName Name of the represented {@link Attribute}.
-   * @param content Content of the represented {@link Attribute} (see {@link Attribute#content}).
-   */
-  public AttributeTreeNode(final String elementName,
-          final String attributeName, final List<String> content) {
-    super(attributeName);
-    this.elementName = elementName;
-    this.attributeName = attributeName;
-    this.content = content;
-  }
-
-  public String getElementName() {
-    return elementName;
-  }
-
-  public String getAttributeName() {
-    return attributeName;
-  }
-
-  public List<String> getContent() {
-    return content;
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/test/unit/src/cz/cuni/mff/ksi/jinfer/attrstats/objects/AMModelTest.java

@@ -1,256 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.objects;
-
-import java.util.Random;
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
-import cz.cuni.mff.ksi.jinfer.base.utils.TestUtils;
-import java.util.ArrayList;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtilsTest;
-import java.util.Arrays;
-import java.util.HashSet;
-import java.util.List;
-import java.util.Map;
-import java.util.Set;
-import javax.swing.tree.DefaultMutableTreeNode;
-import javax.swing.tree.TreeNode;
-import org.junit.Test;
-import static org.junit.Assert.*;
-
-/**
- * @author vektor
- */
-@SuppressWarnings("PMD.SystemPrintln")
-public class AMModelTest {
-
-  private final static Random RND = new Random();
-  private static final int ITERATIONS = 100;
-  private static final AMModel TRIVIAL_MODEL = new AMModel(Arrays.asList(TestUtils.getElement("e")));
-  private static final AttributeMappingId NULL_MAPPING = new AttributeMappingId(null, null);
-  private static final AttributeMappingId TARGET_MAPPING = new AttributeMappingId("element", "attribute");
-  private static final AttributeMappingId OTHER_MAPPING = new AttributeMappingId("otherElement", "attribute");
-  private static final Element TARGET_ELEMENT = TestUtils.getElementWithAttribute("element", "attribute", "value");
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/test/unit/src/cz/cuni/mff/ksi/jinfer/attrstats/utils/WeightComparatorTest.java

@@ -1,73 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.utils;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
-import cz.cuni.mff.ksi.jinfer.base.utils.TestUtils;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.List;
-import org.junit.Test;
-import static org.junit.Assert.*;
-
-/**
- * @author vektor
- */
-@SuppressWarnings("PMD.SystemPrintln")
-public class WeightComparatorTest {
-
-  /**
-   * Test of compare method, of class WeightComparator.
-   */
-  @Test
-  public void testCompare() {
-    System.out.println("compare");
-
-    final List<Element> grammar = new ArrayList<Element>();
-    grammar.add(TestUtils.getElementWithAttribute("e", "a", "d"));
-    grammar.add(TestUtils.getElementWithAttribute("e1", "a1", "e"));
-    grammar.add(TestUtils.getElementWithAttribute("e1", "a1", "f"));
-    grammar.add(TestUtils.getElementWithAttribute("e2", "a2", "g"));
-    grammar.add(TestUtils.getElementWithAttribute("e2", "a2", "g"));
-    grammar.add(TestUtils.getElementWithAttribute("e2", "a2", "g"));
-
-    final AMModel model = new AMModel(grammar);
-
-    final List<AttributeMappingId> amis = new ArrayList<AttributeMappingId>(2);
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/Experiment.java

@@ -1,309 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ExperimentListener;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Quality;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.QualityMeasurement;
-import cz.cuni.mff.ksi.jinfer.attrstats.idref.IdRefSearch;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.GraphUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.Pair;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGeneratorCallback;
-import cz.cuni.mff.ksi.jinfer.base.objects.Input;
-import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import cz.cuni.mff.ksi.jinfer.base.utils.ModuleSelectionHelper;
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/OfficialTestData.java

@@ -1,84 +1,0 @@
-/*
- * Copyright (C) 2011 vitasek
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Constants;
-import static cz.cuni.mff.ksi.jinfer.attrstats.experiments.FileCharacteristics.*;
-
-/**
- *
- * @author vitasek
- */
-public enum OfficialTestData {
-
-  OVA1           ("OVA1.xml",     REALISTIC, 29, 43,   0.45588235294117635),
-  OVA2           ("OVA2.xml",     REALISTIC, 23, 36,   0.1634615384615385),
-  OVA3           ("OVA3.xml",     REALISTIC, 31, 47,   0.25537156151635415),
-  XMAc           ("XMA-c.xml",    REALISTIC, 1, 0,     0.7546666666666666),
-  XMAp           ("XMA-p.xml",    REALISTIC, 1, 0,     0.2019306150568969),
-  XMD            ("XMD.xml",      REALISTIC, 17, 15,   0.09786094165493507),
-
-  MSH            ("MSH.xml",      CONVERTED, 1, 0,     0.5416472778036296),
-  NTH            ("NTH.xml",      CONVERTED, 5, 7,     0.057918595422124436),
-
-  GRAPH_100_100  ("100-100.xml",  ARTIFICIAL, 99, 95,  0.836666666666667),
-  GRAPH_100_200  ("100-200.xml",  ARTIFICIAL, 96, 174, 0.7260000000000004),
-  GRAPH_100_1000 ("100-1000.xml", ARTIFICIAL, 93, 754, 0.380952380952381);
-
-  private final InputFile file;
-  // Graph representation
-  private final int vertices;
-  private final int edges;
-  // Known optimum for alpha = beta = 1
-  private final Double optimum;
-
-  private OfficialTestData(final String fileName,
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/interfaces/ConstructionHeuristic.java

@@ -1,47 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-
-/**
- * Interface representing a construction heuristic. Its responsibility is to
- * create one or more starting solutions from the model provided in the
- * experiment to be optimized later in improvement heuristics.
- *
- * Note that the solution(s) created by this heuristic might already be optimal.
- * In this case it is up to the experiment to recognize this and stop the
- * optimizations.
- *
- * Note that heuristics use the callback "design pattern", meaning that the
- * result of this heuristic is returned via a callback provided while invoking it.
- *
- * @author vektor
- */
-public interface ConstructionHeuristic extends Heuristics {
-
-  /**
-   * Start the heuristic run. Use the information in the provided experiment to
-   * create a pool (possibly only one) of start solutions.
-   *
-   * @param experiment Experiment in context of which to run the heuristic.
-   * @param callback Callback to be invoked when finished.
-   */
-  void start(final Experiment experiment, final HeuristicCallback callback)
-          throws InterruptedException;
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/interfaces/HeuristicCallback.java

@@ -1,39 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import java.util.List;
-
-/**
- * Callback interface used by heuristics to report their results to
- * their invoker.
- *
- * @author vektor
- */
-public interface HeuristicCallback {
-
-  /**
-   * Report that a heuristic has finished and get the pool of solutions from
-   * the argument.
-   *
-   * @param feasiblePool List of feasible solutions found by this run of the
-   * heuristic.
-   */
-  void finished(final List<IdSet> feasiblePool);
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/interfaces/TerminationCriterion.java

@@ -1,47 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.Pair;
-import java.util.List;
-
-/**
- * Interface encapsulating the decision whether to stop running the metaheuristics.
- *
- * This is based on the total time already spent as well as any properties of
- * the solutions found so far.
- *
- * @author vektor
- */
-public interface TerminationCriterion {
-
-  /**
-   * Returns a flag whether to terminate the iterations of the metaheuristic.
-   *
-   * @param experiment Experiment in context of which to measure the quality.
-   * @param time Total time taken so far.
-   * @param solutions The pool solutions that were produced in the last run.
-   * @return The first item of the pair is <code>true</code> if the
-   * metaheuristic should be terminated, <code>false<code> otherwise. The second
-   * is a string describing why the metaheuristics was terminated, in
-   * user-friendly language.
-   */
-  Pair<Boolean, String> terminate(final Experiment experiment, final long time, final List<IdSet> solutions);
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/quality/Quality.java

@@ -1,58 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality;
-
-/**
- * Class representing quality of a solution. This means some kind of "linear"
- * quality, possibly weight, number of attribute mappings and a flag whether
- * this solution is optimal.
- *
- * @author vektor
- */
-public class Quality {
-
-  private final double scalar;
-
-  private final int amCount;
-
-  private final boolean optimal;
-
-  public static final Quality ZERO = new Quality(0, 0, false);
-
-  public Quality(final double scalar, final int amCount, final boolean optimal) {
-    this.scalar = scalar;
-    this.amCount = amCount;
-    this.optimal = optimal;
-  }
-
-  public String getText() {
-    return getScalar() + " (" + getAmCount() + " AMs)" + (isOptimal() ? ", optimal" : "");
-  }
-
-  public double getScalar() {
-    return scalar;
-  }
-
-  public int getAmCount() {
-    return amCount;
-  }
-
-  public boolean isOptimal() {
-    return optimal;
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/experiments/sets/RandomVsFuzzyStart.java

@@ -1,62 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.experiments.sets;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Weight;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.Fuzzy;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.Random;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.Identity;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Constants;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-
-/**
- * Experiment comparing Random and Fuzzy construction heuristics, only their
- * first steps.
- *
- * @author vektor
- */
-public class RandomVsFuzzyStart extends AbstractExperimentSet {
-
-  @Override
-  public String getName() {
-    return "Random vs Fuzzy first step only";
-  }
-
-  @Override
-  protected List<ExperimentParameters> getExperiments() {
-    final List<ImprovementHeuristic> improvement = Arrays.<ImprovementHeuristic>asList(new Identity());
-
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/Incremental.java

@@ -1,81 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ConstructionHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.WeightComparator;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.Collections;
-import java.util.List;
-
-/**
- * A simple construction heuristic. Orders all AMs according to their weight,
- * and starting from the most valuable tries to insert them in the solution. If
- * the solution remains ID set OK, otherwise the AM is not added.
- *
- * Note that the pool created by this heuristic contains only one solution
- * ({@link IdSet}).
- *
- * @author vektor
- */
-public class Incremental implements ConstructionHeuristic {
-
-  @Override
-  public void start(final Experiment experiment, final HeuristicCallback callback)
-          throws InterruptedException {
-
-    final AMModel model = experiment.getModel();
-
-    final List<AttributeMappingId> candidates = MappingUtils.getCandidates(model);
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/Random.java

@@ -1,79 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ConstructionHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import java.util.ArrayList;
-import java.util.List;
-
-/**
- * A trivial construction heuristics. Incrementally builds the solution by
- * adding random AMs that don't break the ID set property.
- *
- * @author vektor
- */
-public class Random implements ConstructionHeuristic {
-
-  @Override
-  public void start(final Experiment experiment,
-        final HeuristicCallback callback) throws InterruptedException {
-    final List<IdSet> ret = new ArrayList<IdSet>(experiment.getPoolSize());
-    final List<AttributeMappingId> candidates = MappingUtils.getCandidates(experiment.getModel());
-
-    for (int i = 0; i < experiment.getPoolSize(); i++) {
-      final List<AttributeMappingId> idSet = new ArrayList<AttributeMappingId>();
-      final List<AttributeMappingId> cs = new ArrayList<AttributeMappingId>(candidates);
-
-      while (!cs.isEmpty()) {
-        // pick one from cs at random
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/fidax/FidaxAlgorithm.java

@@ -1,146 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.fidax;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.DeletableList;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.ImageSizeComparator;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.HashMap;
-import java.util.HashSet;
-import java.util.List;
-import java.util.Map;
-import java.util.Set;
-
-/**
- * Implementation of the heuristic algorithm for finding ID set as described
- * in the "Finding ID Attributes in XML Documents" article.
- *
- * @author vektor
- */
-public final class FidaxAlgorithm {
-
-  private FidaxAlgorithm() {
-  }
-
-  /**
-   * Finds the "best" ID set in the specified attribute mapping model. For
-   * detailed information on the algorithm itself, please refer to the article.
-   *
-   * @param model Model to work on.
-   * @param alpha Weight of the attribute mapping <cite>support</cite> in its total weight.
-   * @param beta Weight of the attribute mapping <cite>coverage</cite> in its total weight.
-   *
-   * @return List of attribute mappings constituting the ID set found.
-   */
-  public static IdSet findIDSet(final AMModel model,
-          final double alpha, final double beta) {
-
-    // 1.  M := all AMs - hidden here
-
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkInputGenerator.java

@@ -1,244 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement.LocalBranching;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.MappingUtils;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.util.Collections;
-import java.util.List;
-import org.apache.log4j.Logger;
-
-/**
- * Utility class for conversion from {@link AMModel} to GLPK MathProg plaintext
- * format.
- *
- * @author vektor
- */
-public final class GlpkInputGenerator {
-
-  private GlpkInputGenerator() {
-
-  }
-
-  private static final Logger LOG = Logger.getLogger(GlpkInputGenerator.class);
-
-  private static final String TEMPLATE = "cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkInputTemplate.txt";
-  private static final String TEMPLATE_LB = "cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkInputTemplateLB.txt";
-  private static final String CONSTRAINT = "s.t. c{index}: x['{mapping1}'] + x['{mapping2}'] <= 1;";
-  private static final String FIX_TO_1 = "s.t. f{index}: x['{mapping}'] = 1;";
-  private static final String FIX_TO_0 = "s.t. f{index}: x['{mapping}'] = 0;";
-
-  /**
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/construction/glpk/GlpkRunner.java

@@ -1,117 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import java.io.File;
-import java.io.FileReader;
-import java.io.IOException;
-import java.io.InputStreamReader;
-import java.util.Calendar;
-import java.util.Collections;
-import java.util.List;
-import org.apache.log4j.Logger;
-
-/**
- * Utility class for GLPK invocation.
- *
- * @author vektor
- */
-public final class GlpkRunner {
-
-  private GlpkRunner() {
-
-  }
-
-  private static final Logger LOG = Logger.getLogger(GlpkRunner.class);
-
-  /**
-   * @see GlpkRunner#run(cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel, java.util.List, double, double, int)
-   */
-  public static String run(final AMModel model,
-          final double alpha, final double beta, final int timeLimit)
-          throws InterruptedException {
-    return run(model, Collections.<AttributeMappingId>emptyList(), alpha, beta, timeLimit);
-  }
-
-  /**
-   * Constructs the GLPK problem formulation from the provided model, runs
-   * GLPK optimalization and returns the resulting file in a string. This can
-   * be directly displayed for verification or parsed for actual AMs appearing
-   * in the "optimal" ID set.
-   *
-   * @param model Model to find "optimal" ID set in.
-   * @param fixed List of attribute mappings that should be present in the
-   * solution.
-   * @param alpha Weight of the attribute mapping <cite>support</cite> in its total weight.
-   * @param beta Weight of the attribute mapping <cite>coverage</cite> in its total weight.
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/improvement/Identity.java

@@ -1,53 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import java.util.List;
-
-/**
- * This trivial improvement heuristic simply returns the provided solution pool.
- *
- * @author vektor
- */
-public class Identity implements ImprovementHeuristic {
-
-  @Override
-  public void start(final Experiment experiment, final List<IdSet> feasiblePool,
-        final HeuristicCallback callback) throws InterruptedException {
-    callback.finished(feasiblePool);
-  }
-
-  @Override
-  public String getName() {
-    return "Identity";
-  }
-
-  @Override
-  public String getDisplayName() {
-    return getName();
-  }
-
-  @Override
-  public String getModuleDescription() {
-    return getName();
-  }
-
-}

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/heuristics/improvement/Mutation.java

@@ -1,95 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.heuristics.improvement;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.Experiment;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.HeuristicCallback;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.attrstats.experiments.quality.Quality;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkOutputParser;
-import cz.cuni.mff.ksi.jinfer.attrstats.heuristics.construction.glpk.GlpkRunner;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.IdSet;
-import cz.cuni.mff.ksi.jinfer.attrstats.utils.Utils;
-import cz.cuni.mff.ksi.jinfer.base.interfaces.Pair;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.util.ArrayList;
-import java.util.List;
-
-/**
- * A slightly more complex improvement heuristic. Finds the best (incumbent)
- * solution in the solution pool, fixes a fraction of its AMs and runs the
- * GLPK on the resulting sub-problem. Its solution is then added to the pool
- * and returned.
- *
- * @author vektor
- */
-public class Mutation implements ImprovementHeuristic {
-
-  private final double ratio;
-  private final int timeLimit;
-
-  /**
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/objects/AttributeMapping.java

@@ -1,110 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.objects;
-
-import java.util.ArrayList;
-import java.util.List;
-
-/**
- * This is a representation of an attribute mapping, as described in the
- * article "Finding ID Attributes in XML Documents". It contains the mapping
- * itself (element and attribute name), as well as its image (all values the
- * attribute under the element ever has). Note that the image is stored as a
- * list (not a set), thus duplicate values are all stored.
- *
- * @author vektor
- */
-public class AttributeMapping {
-
-  private final AttributeMappingId id;
-
-  private final List<String> image = new ArrayList<String>();
-
-  /**
-   * Minimal constructor. Image of this mapping will be created as empty.
-   *
-   * @param element Element name.
-   * @param attribute Attribute name.
-   */
-  public AttributeMapping(final String element, final String attribute) {
-    this.id = new AttributeMappingId(element, attribute);
-  }
-
-  /**
-   * Minimal constructor. Image of this mapping will be created as empty.
-   *
-   * @param id Identifier of this mapping.
-   */
-  public AttributeMapping(final AttributeMappingId id) {
-    this.id = id;
-  }
-
-  /**
-   * Full constructor.
-   *
-   * @param element Element name.
-   * @param attribute Attribute name.
-   * @param image Attribute mapping image.
-   */
-  public AttributeMapping(final String element, final String attribute, final List<String> image) {
-    this.id = new AttributeMappingId(element, attribute);
-    this.image.addAll(image);
-  }
-
-  /**
-   * Full constructor.
-   *
-   * @param id Identifier of this mapping.
-   * @param image Attribute mapping image.
-   */
-  public AttributeMapping(final AttributeMappingId id, final List<String> image) {
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/utils/MappingUtils.java

@@ -1,191 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.utils;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMapping;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
-import java.util.ArrayList;
-import java.util.HashSet;
-import java.util.List;
-import java.util.Set;
-
-/**
- * Library class containing utility functions for attribute mapping.
- *
- * Please see the article "Finding ID Attributes in XML Documents" for reference.
- *
- * @author vektor
- */
-public final class MappingUtils {
-
-  private MappingUtils() {
-
-  }
-
-  /**
-   * Verifies whether the specified attribute mapping is a candidate mapping.
-   *
-   * <p>An attribute mapping is a <cite>candidate mapping</cite>, iff it is
-   * an injective function: if its every value is unique.</p>
-   *
-   * @param targetMapping Mapping to be verified.
-   * @param allMappings All the attribute mappings of the grammar.
-   * @return True if the specified mapping is a candidate, false otherwise.
-   */
-  public static boolean isCandidateMapping(final AttributeMappingId targetMapping, final AMModel model) {
-    verify(targetMapping, model);
-    final AttributeMapping mapping = model.getAMs().get(targetMapping);
-    final Set<String> domain = new HashSet<String>(mapping.getImage());
-    return domain.size() == mapping.getImage().size();
-  }
-
-  private static void verify(final AttributeMappingId targetMapping, final AMModel model) {
-    if (targetMapping == null || model == null) {
-      throw new IllegalArgumentException("Expecting non-null, non empty parameters");
-    }
[…]

### 2011-08-08 jinfer/trunk/jInfer/AttributeStats/src/cz/cuni/mff/ksi/jinfer/attrstats/utils/WeightComparator.java

@@ -1,47 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.attrstats.utils;
-
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AMModel;
-import cz.cuni.mff.ksi.jinfer.attrstats.objects.AttributeMappingId;
-import java.util.Comparator;
-
-/**
- * Comparator of attribute mappings based on their weights in the context of
- * the provided experiment.
- *
- * @author vektor
- */
-public class WeightComparator implements Comparator<AttributeMappingId> {
-
-  private final AMModel model;
-  private final double alpha;
-  private final double beta;
-
-  public WeightComparator(final AMModel model, final double alpha, final double beta) {
-    this.model = model;
-    this.alpha = alpha;
-    this.beta = beta;
-  }
-
-  @Override
-  public int compare(final AttributeMappingId o1, final AttributeMappingId o2) {
-    final Double weight1 = model.weight(o1, alpha, beta);
-    final Double weight2 = model.weight(o2, alpha, beta);
-    return -weight1.compareTo(weight2);
-  }
-}

### 2011-08-09 doc/theses/vitasek/diplomka.tex
Thesis: in progress
Thesis: improving fonts and literature.
Thesis: in progress.

--- 2011-08-06 doc/theses/vitasek/diplomka.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage{tgpagella}
 \usepackage{xstring}
@@
 \usepackage{multicol}
+\usepackage[pdfborder=0 0 0]{hyperref}
 
@@
 
-\begin{center}
+\begin{figure}
+  \caption{Metaheuristic schema}
+  \label{image-metaheuristic}
+  \centering
     \includegraphics[width=\textwidth]{images/metaheuristic.eps}
-\end{center}
+\end{figure}
 
@@
   \WHILE{$s$ is a feasible ID set}
-    \STATE $a \gets $ pick at random from $C \textbackslash S$
+    \STATE $a \gets $ pick at random from $C \backslash S$
     \STATE $s \gets s \cup a$
@@
       \STATE $s \gets s \cup a$
-      \STATE $C' \gets C' \textbackslash a$
+      \STATE $C' \gets C' \backslash a$
     \ENDIF
@@
         \STATE \COMMENT {if $c$ cannot be possibly added anymore}
-        \STATE $C' \gets C' \textbackslash c$
+        \STATE $C' \gets C' \backslash c$
       \ENDIF
@@
   \ENDIF
-  \STATE $s \gets s \textbackslash c$
+  \STATE $s \gets s \backslash c$
 \ENDFOR
@@
 \STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
-\RETURN $FP \textbackslash s_{min}$
+\RETURN $FP \backslash s_{min}$
 \end{algorithmic}
@@
   \STATE \COMMENT {improve a single solution}
-  \STATE $C' \gets C \textbackslash s$
+  \STATE $C' \gets C \backslash s$
   \STATE $C' \gets C'$ sorted by decreasing weight
@@
 
+\listoffigures
+\addcontentsline{toc}{chapter}{List of Figures}
+
+\listofalgorithms
+\addcontentsline{toc}{chapter}{List of Algorithms}
+
 %%% Tabulky v diplomové práci, existují-li.
@@
 At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see TODO link obrazok), responsibility of 3 different services - interchangeable modules.
+Most importantly, the whole process consists of 3 consecutive steps (see \ref{image-inference-process}), responsibility of 3 different services - interchangeable modules.
+
+\begin{figure}
+  \caption{Inference process in jInfer}
+  \label{image-inference-process}
+  \centering
+    \includegraphics[width=0.5\textwidth]{images/inference-process.eps}
+\end{figure}
 
@@
 
-\chapter{AttributeStats}
-\label{attrstats}
+\chapter{IDSetSearch}
+\label{iss}
 
-TODO talk a bit about the structure of AttributeStats module.
+TODO talk a bit about the structure of ISS module.
 

### 2011-08-10 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 	\item Do I beat FIDAX with Random?
-	\item Does Fuzzy beat Random? (Seems that Random is better and even has smaller variance)
+	\item Does Fuzzy beat Random? (Fuzzy wins for every single case) - take note of quality (mean, variance) but also time (again, mean)
 \end{itemize}
@@
 Results:
-Start time: Thu Jan 01 01:22:17 CET 1970
 Total time spent: 7754 ms

### 2011-08-11 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 	\item Do I beat FIDAX with Random?
-	\item Does Fuzzy beat Random? (Fuzzy wins for every single case) - take note of quality (mean, variance) but also time (again, mean)
+	\item Does Fuzzy beat Random? (Fuzzy wins in every single case) - take note of quality (mean, variance) but also time (again, mean)
+  \item Can FIDAX be improved by Hungry?
 \end{itemize}
@@
 
-TODO extension to more documents, IDREFs, ...
+TODO
+
+\begin{itemize}
+  \item extension to more documents at once
+  \item more heuristics and their chaining
+  \item ACO, genetic, ...
+  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
+  \item jInfer is opensource, thus easy to extend!
+\end{itemize}
 
@@
 \label{jInfer}
+
+TODO to the best of our knowledge, this framework is one-of-a-kind
 

### 2011-08-12 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 \usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
+\usepackage[dvipdfmx,pdfborder=0 0 0]{hyperref}
 
@@
 An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
+
+\section{Conventions}
+
+TODO list conventions - how we write code related stuff, module names, heu names, etc
 
@@
 	\item For each input file: how long does model creation take?
-	\item How long does it take to communicate with GLPK? Creating the input/parsing the output.
+	\item How long does it take to communicate with GLPK? Creating the input / parsing the output.
 	\item Which construction heuristic is the best - limit GLPK to like 1 second
@@
 
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Robert Smetana, Michal Svirec and Matej Vitasek.
+jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
 The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
@@
 TODO talk a bit about the structure of ISS module.
+
+TODO name of the module, its place in jInfer, interface with the framework (service providing)
+
+TODO package name root
+
+TODO "Attribute Stats"
+
+TODO heuristics - interfaces, construction, improvement
+
+TODO experiments - quality, termination, parameters, experiment sets
 
@@
     Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Maximum iterations exceeded.
+Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
 
@@
 After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha/beta settings.
+Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
 \\

### 2011-08-13 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 
+TODO what are trying to establish?
+What is our aim here?
+To get an understanding how the system and its components behave.
+What do the parameters affect.
+To see which types of data are best handled by specific algorithms.
+To see what are the time scales involved in ID search and how various parameters affect the run times of the heuristics.
+
 TODO describe all the inputs (take it from the mindmap), describe how the output looks (sample trace - link to appendix)
@@
 
-TODO types of data:
+Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
+
 \begin{itemize}
 	\item Realistic 
-	\item Realistic with artificial attributes
+	\item Realistic with artificial (converted) attributes
 	\item Artificial
@@
 
-TODO for all types, show GraphViz pictures.
+A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
+The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
+instead of attributes.
+We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
+using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
+And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
+This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.
 
@@
 
+\begin{figure}
+  \caption{Realistic data - first part}
+  \label{image-experiments-data-realistic-1}
+  \centering
+    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1.eps}}
+    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2.eps}}
+    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3.eps}}
+\end{figure}
+
+\begin{figure}
+  \caption{Realistic data - second part}
+  \label{image-experiments-data-realistic-2}
+  \centering
+    \subfigure[XMA-c]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMA-c.eps}}
+    \subfigure[XMA-p]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMA-p.eps}}
+    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD.eps}}
+\end{figure}
+
 \subsection{Realistic data with artificial attributes}
-
-TODO we have taken some of the text data and converted it to attributes like this, A B C
+\label{realistic-converted}
+
+TODO we have taken some of the text data and converted it to attributes like this, A B C.
+
+TODO we can convert it back to keys!
+
+\begin{figure}
+  \caption{Realistic data with converted attributes}
+  \label{image-experiments-data-converted}
+  \centering
+    \subfigure[MSH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/MSH.eps}}
+    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH.eps}}
+\end{figure}
 
@@
 TODO these are random graphs.
+
 TODO talk about how they are created, about their structure.
+
 TODO talk about some "density".
 
+TODO talk about the "size" test data
+
+\begin{figure}
+  \caption{Artificial data}
+  \label{image-experiments-data-artificial}
+  \centering
+%  	\subfigure[100-100]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/100-100.eps}}
+%	\subfigure[100-200]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/100-200.eps}}
+%    \subfigure[100-1000]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/100-1000.eps}}    
+\end{figure}
+
 \section{Experimental Setup}
 
+TODO describe the actual HW/SW setup: two computers, Windows/Linux, versions etc.
+What is running when experimenting, that we are always using more runs to counter for effects like caches and stuff that happens.
+Say many things actually depend on the order in which we run the experiments in one set, and if it is relevant we will talk about it in that particular experiment.
+
 TODO we will be using jInfer and its module AttributeStats see \ref{jInfer} and \ref{attrstats}.
+
 TODO describe how to create a model, describe how we get IGG, link BasicIGG
@@
 
-TODO to the best of our knowledge, this framework is one-of-a-kind
-
 This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
@@
 This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
+
+To the best of our knowledge, at the time of writing this thesis is jInfer the only public, open source and actually working solution for XML schema inference-related tasks.
 
@@
 
-TODO elaborate more, if too short:)
-
 \chapter{IDSetSearch}
@@
 
-\chapter{Traces}
+\chapter{Experimental Trace}
 \label{traces}

### 2011-08-13 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 
+\begin{table}
+  \caption{List of realistic test data files}
+  \label{table-experiments-data-realistic}
+  \centering
+  \begin{tabular}{l | r | c | c | l}
+	Name  & Size [kb] & Vertices & Edges & Optimum \\
+	\hline
+	OVA1  & 4.5 & 29 & 43 & 0.45588235294117635 \\
+	OVA2  & 11.9 & 23 & 36 & 0.1634615384615385  \\ 
+	OVA3  & 237.6 & 31 & 47 & 0.25537156151635415 \\
+	XMA-c & 1 807.7 & 1  & 0  & 0.7546666666666667  \\
+	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
+	XMD   & 1 743.0 & 17 & 15 & 0.09786094165493507 \\
+  \end{tabular}
+\end{table}
+
+TODO explain edges, vertices and say that opt is the known for alpha = beta = 1
+
 \begin{figure}
-  \caption{Realistic data - first part}
-  \label{image-experiments-data-realistic-1}
+  \caption{Realistic data}
+  \label{image-experiments-data-realistic}
   \centering
@@
     \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3.eps}}
-\end{figure}
-
-\begin{figure}
-  \caption{Realistic data - second part}
-  \label{image-experiments-data-realistic-2}
-  \centering
-    \subfigure[XMA-c]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMA-c.eps}}
-    \subfigure[XMA-p]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMA-p.eps}}
+    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c.eps}}
+    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p.eps}}
     \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD.eps}}
@@
 TODO we can convert it back to keys!
+
+\begin{table}
+  \caption{List of realistic test data files with converted attributes}
+  \label{table-experiments-data-converted}
+  \centering
+  \begin{tabular}{l | r | c | c | l}
+	Name  & Size [kb] & Vertices & Edges & Optimum \\
+	\hline
+	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
+	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
+  \end{tabular}
+\end{table}
 
@@
   \centering
-    \subfigure[MSH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/MSH.eps}}
+    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH.eps}}
     \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH.eps}}
@@
 
-TODO talk about some "density".
+TODO talk about some "density". Explain what is the max number of edges
 
 TODO talk about the "size" test data
+
+\begin{table}
+  \caption{List of artificial test data files}
+  \label{table-experiments-data-artificial}
+  \centering
+  \begin{tabular}{l | r | c | c | c | c | l}
+	Name  & Size [kb] & Vertices & Edges & $|E| / |V|$ & $|E| / |E_{max}|$ & Optimum \\
+	\hline
+	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
+	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
+    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{List of ``sized" artificial test data files}
+  \label{table-experiments-data-artificial-size}
+  \centering
+  \begin{tabular}{l | r | c | c | c | c | l}
+	Name  & Size [kb] & Vertices & Edges & $|E| / |V|$ & $|E| / |E_{max}|$ & Optimum \\
+	\hline
+	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
+	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
+    20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
+   	30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
+	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
+	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
+	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
+	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
+	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
+	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
+	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
+  \end{tabular}
+\end{table}
+
+TODO explain why the optimum is getting smaller
 
@@
   \centering
-%  	\subfigure[100-100]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/100-100.eps}}
-%	\subfigure[100-200]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/100-200.eps}}
-%    \subfigure[100-1000]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/100-1000.eps}}    
+    % TODO fix cropping
+  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80.eps}}
+	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed.eps}}
 \end{figure}
+
+TODO why we have less than declared number? because of how we created it
 

### 2011-08-13 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 
+To understand these data sets, we will talk a little about heir origin, and a little more about their \textit{graph representation}. As mentioned earlier, % TODO link
+the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
+Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
+
+The former will be achieved with the help of GraphViz tool, % TODO link
+where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
+Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
+
+The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
+
 \subsection{Realistic data}
 
-TODO talk about all that are used - mention the dataset name, some characteristics (excel sheet)
+From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
+Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
+Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
+Neither will they be included on the DVD distributed with this thesis.
+
+TODO their DTD/XSD - do we have ID attributes?
 
@@
   \caption{List of realistic test data files}
+  \bigskip
   \label{table-experiments-data-realistic}
@@
   \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & Vertices & Edges & Optimum \\
+	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
 	\hline
@@
 \end{table}
-
-TODO explain edges, vertices and say that opt is the known for alpha = beta = 1
 
@@
 
+To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
+
+On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
+We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
+
+Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
+
 \subsection{Realistic data with artificial attributes}
@@
 
-TODO we have taken some of the text data and converted it to attributes like this, A B C.
-
-TODO we can convert it back to keys!
+We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
+Unfortunately, the same problem with disclosure as in the previous case applies here.
+None of these sets had any attributes before the conversion.
+Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
+
+TODO schemas?
+
+To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
+In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
+
+Why is this useful at all?
+Recall TODO link, where we said that ID attributes are a special case of XML keys.
+We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
 
@@
   \caption{List of realistic test data files with converted attributes}
+  \bigskip
   \label{table-experiments-data-converted}
@@
 
+Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
+This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
+
+In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
+Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
+This means we have quite a big set with relatively simple work to be done by the heuristics.
+This is about the change.
+
 \subsection{Artificial data}
 
-TODO these are random graphs.
-
-TODO talk about how they are created, about their structure.
-
-TODO talk about some "density". Explain what is the max number of edges
-
-TODO talk about the "size" test data
+As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
+After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
+We needed to create data that would really put pressure on our heuristics, put them to a real test.
+Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
+Why not create the XML data to contain a more complex graph in the first place?
+What if we could specify how many vertices and edges we want, and get the according XML?
+
+This is actually simple. Consider the following excerpt from an XML file:
+
+\begin{scriptsize}
+\begin{verbatim}
+<graph>
+  <vertex0 attr="-2968876296119015800"/>
+  <vertex1 attr="1729745997570096518"/>
+  <vertex2 attr="-9020549659620928934"/>
+  ...
+  <vertex99 attr="-7545982394508643394"/>
+
+  <vertex82 attr="0"/><vertex21 attr="0"/>
+  <vertex64 attr="1"/><vertex21 attr="1"/>
+  <vertex44 attr="2"/><vertex2 attr="2"/>
+  ...
+  <vertex96 attr="99"/><vertex40 attr="99"/>
+</graph>
+\end{verbatim}
+\end{scriptsize}
+
+We want to create a graph with around $v$ vertices and $e$ edges.
+First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
+To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
+Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
+This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
+
+The pseudocode for this is in listing \ref{listing-random-data}.
+
+\begin{algorithm}
+\caption{Random XML data creation}
+\label{listing-random-data}
+\begin{algorithmic}
+\REQUIRE $v$ requested number of vertices
+\REQUIRE $e$ requested number of edges
+\ENSURE XML file content
+\PRINT \texttt{<graph>}
+\FOR{$i = 1 \to |V|$}
+	\STATE $R \gets RANDOM$
+	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
+\ENDFOR
+\FOR{$i = 1 \to |E|$}
+	\STATE $v1 \gets RANDOM(|V|)$
+	\STATE $v2 \gets RANDOM(|V|)$
+	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
+\ENDFOR
+\PRINT \texttt{</graph>}
+\RETURN
+\end{algorithmic}
+\end{algorithm}
+
+With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
+
+There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
+This can be defined in various ways, we will use two different interpretations for a while.
+The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
+The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
+This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
+
+We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
+Note that the name is always in the form $v-e$.
+
+Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
+For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
+We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
+
+TODO this is on the DVD.
+
+The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
+Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
+
+While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
+This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
+% TODO too much?:)
+Because of the so-called Birthday paradox, % TODO link
+this will happen more with higher $e$.
 
@@
   \caption{List of artificial test data files}
+  \bigskip
   \label{table-experiments-data-artificial}
@@
   \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & Vertices & Edges & $|E| / |V|$ & $|E| / |E_{max}|$ & Optimum \\
+	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
 	\hline
@@
   \caption{List of ``sized" artificial test data files}
+  \bigskip
   \label{table-experiments-data-artificial-size}
@@
 
-TODO explain why the optimum is getting smaller
-
 \begin{figure}
@@
 
-TODO why we have less than declared number? because of how we created it
+To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
+The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
+
+In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
+There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
+This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
+Exploration of this phenomenon is among the possibilities of future work of this thesis.
 

### 2011-08-14 doc/theses/vitasek/diplomka.tex
Reworked makefile and work cycle to pdflatex

@@
 \usepackage{multicol}
-\usepackage[dvipdfmx,pdfborder=0 0 0]{hyperref}
+\usepackage[pdfborder=0 0 0]{hyperref}
 
@@
 
-\centerline{\mbox{\includegraphics[width=60mm]{logo.eps}}}
+\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
 
@@
   \centering
-    \includegraphics[width=\textwidth]{images/metaheuristic.eps}
+    \includegraphics[width=\textwidth]{images/metaheuristic}
 \end{figure}
@@
   \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1.eps}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2.eps}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3.eps}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c.eps}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p.eps}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD.eps}}
+    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
+    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
+    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
+    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
+    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
+    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
 \end{figure}
@@
   \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH.eps}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH.eps}}
+    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
+    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
 \end{figure}
@@
     % TODO fix cropping
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80.eps}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed.eps}}
+  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
+	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
 \end{figure}
@@
   \centering
-    \includegraphics[width=0.5\textwidth]{images/inference-process.eps}
+    \includegraphics[width=0.5\textwidth]{images/inference-process}
 \end{figure}

### 2011-08-14 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 \usepackage[pdfborder=0 0 0]{hyperref}
+\usepackage{amssymb}
 
@@
 
-TODO what are trying to establish?
-What is our aim here?
-To get an understanding how the system and its components behave.
-What do the parameters affect.
-To see which types of data are best handled by specific algorithms.
-To see what are the time scales involved in ID search and how various parameters affect the run times of the heuristics.
-
-TODO describe all the inputs (take it from the mindmap), describe how the output looks (sample trace - link to appendix)
+At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
+
+Now it is the time to move our ideas into reality and test their feasibility.
+But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
+
+First of all, we would like to get an idea of how the whole system and its components behave.
+We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
+Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
+Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
+
+Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
+This will allow us to pick the right heuristic once we see new data.
+
+Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
+
+And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
+
+This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full descriptions and results, and in the end we shall draw some final conclusions.
 
@@
 And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.
-
-To understand these data sets, we will talk a little about heir origin, and a little more about their \textit{graph representation}. As mentioned earlier, % TODO link
+This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
+
+To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
 the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
@@
 
+As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
+Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.
+
+\subsection{Grammar and Model Creation}
+
+TODO link BasicIGG to see how we get IG
+
+TODO describe how we get the model
+
+\subsection{Hardware and Software}
+
 TODO describe the actual HW/SW setup: two computers, Windows/Linux, versions etc.
+
+\subsection{Methodology}
+
 What is running when experimenting, that we are always using more runs to counter for effects like caches and stuff that happens.
@@
 
-TODO we will be using jInfer and its module AttributeStats see \ref{jInfer} and \ref{attrstats}.
-
-TODO describe how to create a model, describe how we get IGG, link BasicIGG
+\subsection{Obtaining the results}
+
+TODO we are logging (link to trace appendix) and creating CSVs.
 
@@
   \item Can FIDAX be improved by Hungry?
+  \item Can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
 \end{itemize}
@@
 \chapter{jInfer}
-\label{jInfer}
+\label{appendix-jInfer}
 
@@
 \chapter{IDSetSearch}
-\label{iss}
+\label{appendix-iss}
 
@@
 TODO experiments - quality, termination, parameters, experiment sets
+
+TODO make sure there was a place where we described the model data model:)
 
@@
 
+\flushright
+$\square$
+
 \end{document}

### 2011-08-14 doc/theses/vitasek/diplomka.tex
Thesis: in progress

@@
 \usepackage{amssymb}
+\usepackage{amsthm}
 
@@
 \makenomenclature
+
+\theoremstyle{definition}
+\newtheorem{define}{Definition}[chapter]
 
@@
 
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full descriptions and results, and in the end we shall draw some final conclusions.
+This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
 
@@
   \centering
-    % TODO fix cropping
   	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
@@
 As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.
+Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
+
+We now have to define a few notions before moving forward to the description of our experiments.
+
+\begin{define}[Experiment parameters]
+	\label{define-experiment-params}
+	\textit{Experiment parameters} are the following.
+	\begin{itemize}
+		\item All the parameters in all the heuristics.
+		\item The specific way in which the heuristics are chained.
+		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
+		\item Initial pool size.
+		\item The termination criteria.
+		\item The input XML file.
+		\item Known optimum for this file and $\alpha$, $\beta$.
+	\end{itemize}
+\end{define}
+
+\begin{define}[Experiment(al) configuration]
+	\label{define-experiment-config}
+	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
+\end{define}
+
+\begin{define}[Experiment(al) set]
+	\label{define-experiment-set}
+	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
+\end{define}
 
@@
 
-TODO describe the actual HW/SW setup: two computers, Windows/Linux, versions etc.
+We will be using two different machines to run the experiments.
+Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
+
+\subsubsection{Linux Machine}
+
+\begin{verbatim}
+Intel Core 2 Quad processor @ 2.83 GHz
+8 GB DDR2 RAM
+x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
+Java HotSpot 64-Bit Server VM (build 20.1-b02)
+GLPK version 4.38
+\end{verbatim}
+
+\subsubsection{Windows Machine}
+
+\begin{verbatim}
+Intel Core 2 Duo processor @ 2.33 GHz
+4 GB DDR2 RAM
+Windows 7 SP1 64bit
+Java HotSpot TODO
+GLPK version TODO
+\end{verbatim}
 
@@
 
-What is running when experimenting, that we are always using more runs to counter for effects like caches and stuff that happens.
-Say many things actually depend on the order in which we run the experiments in one set, and if it is relevant we will talk about it in that particular experiment.
+It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
+First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
+Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
+This improvement is left for the future work.
+
+Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
+OS deciding to run some house cleaning) will be averaged out.
+Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
+% TODO discuss caches and the order in which we run the experiments in one set?
 
@@
 
-TODO we are logging (link to trace appendix) and creating CSVs.
+Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
+We can get all the information relevant to that experiment run from this trace alone.
+An experimental set will produce a number of these traces and store them in plain text files in a folder.
+Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
 
@@
 \chapter{Experimental Trace}
-\label{traces}
+\label{appendix-trace}
 

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
+\renewcommand{\nomname}{List of Abbreviations}
+\makenomenclature
+
+\theoremstyle{definition}
+\newtheorem{define}{Definition}[chapter]
+
 \setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
@@
 }
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
 
@@
 
+\subsection{Measuring the time}
+
+TODO
+
 \subsection{Obtaining the results}
@@
 TODO following experiments + more
+
+\subsection{GLPK: native vs. Cygwin}
+
+In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
+The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
+
+If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
+We might also find that there is no relevant difference, which would be good, too.
+
+\begin{table}
+  \centering
+  \begin{tabular}{| l | l |}
+  \hline
+  Machine & Windows \\
+  Input data set & \texttt{100-500} \\
+  Iterations & 20 \\
+  \hline
+  \end{tabular}
+\end{table}
 

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
-TODO following experiments + more
-
 \subsection{GLPK: native vs. Cygwin}
@@
   Iterations        & 20 \\
+  Pool size         & 1 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
   \hline
@@
 \end{table}
+
+Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
+Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2 (25 settings * 20 iterations = 500 configurations in total).
+There will be no improvement heuristic.
+The only data we gather in the GnuPlot file are the final qualities (weights).\\
+
+TODO input -> time dependency too
+
+\subsection{Still to do}
+
+TODO following experiments + more
 
@@
 	\item Alpha, Beta -> How does the solution work?
-	\item Which GLPK -> How does (time->quality) look, how does (input->time) look
 	\item Do I beat FIDAX with Random?

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 \begin{table}
+  \caption{GLPK: native vs. Cygwin summary}
+  \label{experiment-glpk-native-vs-cygwin-summary}
   \centering
@@
 Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2 (25 settings * 20 iterations = 500 configurations in total).
+Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (pseudocode to be found in \ref{listing-experiment-glpk-native-vs-cygwin}).
 There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).\\
+The only data we gather in the GnuPlot file are the final qualities (weights).
+
+\begin{algorithm}
+\caption{GLPK: native vs. Cygwin generation}
+\label{listing-experiment-glpk-native-vs-cygwin}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$i = 1 \to 20$}
+	\FOR{$time = 1 \to 49 step 2$}
+    \STATE $ES \gets IH = $ \jmodule{Glpk} $, limit = time, IH = \emptyset$  
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
 

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

--- 2011-08-15 doc/theses/vitasek/diplomka.tex
@@
 TODO their DTD/XSD - do we have ID attributes?
+
+TODO verify they satisfy their schemas. Converted too.
 
@@
 x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
+Java SE Runtime Environment (build 1.6.0_26-b03)
 Java HotSpot 64-Bit Server VM (build 20.1-b02)
@@
 Windows 7 SP1 64bit
-Java HotSpot TODO
-GLPK version TODO
+Java SE Runtime Environment (build 1.6.0_26-b03)
+Java HotSpot 32-Bit Client VM (build 20.1-b02)
+GLPK version 4.45 (Cygwin)
+GLPK version 4.34 (native)
 \end{verbatim}
@@
 \begin{table}
+  \caption{GLPK: native vs. Cygwin summary 1}
+  \label{experiment-glpk-native-vs-cygwin-summary-1}
   \centering
@@
 Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2 (25 settings * 20 iterations = 500 configurations in total).
+Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
 There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).\\
-
-TODO input -> time dependency too
+The only data we gather in the GnuPlot file are the final qualities (weights).
+
+\begin{algorithm}
+\caption{GLPK: native vs. Cygwin set generation 1}
+\label{listing-experiment-glpk-native-vs-cygwin-1}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$i = 1 \to 20$}
+	\FOR{$time = 1 \to 49$ step $2$}
+    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$  
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+TODO graph + interpretation
+
+\begin{table}
+  \caption{GLPK: native vs. Cygwin summary 2}
+  \label{experiment-glpk-native-vs-cygwin-summary-2}
+  \centering
+  \begin{tabular}{| l | l |}
+  \hline
+  Machine           & Windows \\
+  Input data set    & sized artificial data \\
+  Iterations        & 20 \\
+  Pool size         & 1 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+  \end{tabular}
+\end{table}
+
+Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
+This experimental set will contain 220 configurations for each version.
+Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
+This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
+There will again be no IH, the only data we will collect are the times of the CH run in each case.
+
+\begin{algorithm}
+\caption{GLPK: native vs. Cygwin set generation 2}
+\label{listing-experiment-glpk-native-vs-cygwin-2}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$i = 1 \to 20$}
+	\FOR{$file \in $ sized artificial data}
+    \STATE $ES \gets ES \cup {file, CH = \jmodule{Glpk}(no limit), IH = \emptyset}$  
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+TODO graph + interpretation
 

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
+\subsection{Grammar and model generation}
+
+%         NB class GrammarModelTiming
+
+TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
+
+The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
+from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
+
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & TODO \\
+  Input data set    & all test data sets \\
+  Iterations        & 20 \\
+  Pool size         & not applicable \\
+  $\alpha$, $\beta$ & not applicable \\
+  \hline
+\end{tabular}
+
+The experimental set will contain $20 * (11 + 11) = 440$ configurations - 20 iterations for 11 test data sets plus 11 sized test data sets.
+There won't be any CHs or IHs.
+We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
+
+TODO graph + interpretation
+
 \subsection{GLPK: native vs. Cygwin}
+
+%         NB class TimeQuality
 
@@
 
-\begin{table}
-  \caption{GLPK: native vs. Cygwin summary 1}
-  \label{experiment-glpk-native-vs-cygwin-summary-1}
-  \centering
 \begin{tabular}{| l | l |}
+  \hline
   \hline
@@
 \end{tabular}
-\end{table}
 
@@
 
-\begin{table}
-  \caption{GLPK: native vs. Cygwin summary 2}
-  \label{experiment-glpk-native-vs-cygwin-summary-2}
-  \centering
 \begin{tabular}{| l | l |}
+  \hline
   \hline
@@
 	\FOR{$file \in $ sized artificial data}
-    \STATE $ES \gets ES \cup {file, CH = \jmodule{Glpk}(no limit), IH = \emptyset}$  
+    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
   \ENDFOR
@@
 \begin{itemize}
-	\item For each input file: how long does model creation take?
   \item How long does it take to communicate with GLPK? Creating the input / parsing the output.
+  \item Big fight Fuzzy vs. Random vs. FIDAX take note of quality (mean, variance) but also time (again, mean)
+  \item Can FIDAX be improved by Hungry?
+
 	\item Which construction heuristic is the best - limit GLPK to like 1 second
+
 	\item With the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
 	\item With the best IH from the previous step, how about using Random as CH?
-	\item GLPK: Input -> How long the search for optimum lasts - this shows we cannot rely on GLPK alone! it just takes too long.
-	\item GLPK: Time limit -> Quality
+
+  \item this can be done inside the GLPK version comparison! GLPK: Input -> How long the search for optimum lasts - this shows we cannot rely on GLPK alone! it just takes too long.
+	\item this can be done inside the GLPK version comparison! GLPK: Time limit -> Quality
+
 	\item Structure of the data -> Which algorithms perform the best
+
   \item Alpha, Beta -> Quality
 	\item Alpha, Beta -> How does the solution work?
-	\item Do I beat FIDAX with Random?
-	\item Does Fuzzy beat Random? (Fuzzy wins in every single case) - take note of quality (mean, variance) but also time (again, mean)
-  \item Can FIDAX be improved by Hungry?
+
   \item Can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
-\subsubsection{FIDAX}
-
-TODO this is a heuristic too! We can start with it!
-
-The pseudocode is listed (cited from \cite{fidax} with trivial modifications without changing the logic) in \ref{ch-fidax}.
+\subsubsection{\jmodule{FIDAX}}
+
+It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
+
+The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{ch-fidax}.
 
@@
 \REQUIRE $C$ list of candidate AMs
+\ENSURE a feasible solution
 \STATE $C' \gets C$ sorted by decreasing size
@@
 
-\subsubsection{Random}
-
-TODO this is trivial, but works - link experiments where it bashes FIDAX
-
-See listing \ref{ch-random}.
+\subsubsection{\jmodule{Random}}
+\label{heu-ch-random}
+
+One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
+This is obviously a hungry heuristic.
+%TODO have I defined ID set condition yet??
+
+The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
+As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
+
+See the Listing \ref{ch-random} for its pseudocode.
 
@@
 
-\subsubsection{Fuzzy}
-
-TODO kind of trivial
-
-See listing \ref{ch-fuzzy}.
+\subsubsection{\jmodule{Fuzzy}}
+\label{heu-ch-fuzzy}
+
+\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
+The weight used here is the usual weight of an AM as defined in TODO.
+Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
+
+Again, this is a hungry heuristic, the Listing \ref{ch-fuzzy} contains its pseudocode.
+
+% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
 
@@
 
-\subsubsection{Incremental}
-
-TODO trivial, hungry, always 1 solution
+\subsubsection{\jmodule{Incremental}}
+
+This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
+This way it can create only one solution, and again, this is a hungry heuristic.
 
@@
 
-\subsubsection{Removal}
-
-TODO trivial, "hungry", always 1 solution
+\subsubsection{\jmodule{Removal}}
+
+This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
+This probably does not satisfy the ID set condition.
+Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
+Again, this is a hungry heuristic returning only one solution.
 
@@
 
+This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
+
 TODO if we limit GLPK's runtime, we get this
@@
 
-\subsubsection{Identity}
+\subsubsection{\jmodule{Identity}}
 
@@
 
-\subsubsection{Remove Worst}
-
-TODO trivial
+\subsubsection{\jmodule{Remove Worst}}
+
+This trivial IH tries to improve the solution pool by removing the worst solution (i.e.
+the one with the lowest quality).
+This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
 
@@
 
-\subsubsection{Random Remove}
-
-TODO trivial, perturbation
+\subsubsection{\jmodule{Random Remove}}
+
+This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
+By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
+
+The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
+For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
+
+This heuristic returns a pool of solutions of the same size as it got on its input.
 
@@
 
-\subsubsection{Hungry}
+\subsubsection{\jmodule{Hungry}}
 
@@
 
-\subsubsection{Mutation}
+\subsubsection{\jmodule{Mutation}}
 
@@
 
-\subsubsection{Crossover}
+\subsubsection{\jmodule{Crossover}}
 
@@
 
-\subsubsection{Local Branching}
+\subsubsection{\jmodule{Local Branching}}
 
@@
 
-The experimental set will contain $20 * (11 + 11) = 440$ configurations - 20 iterations for 11 test data sets plus 11 sized test data sets.
+The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
 There won't be any CHs or IHs.
@@
 
+\subsubsection{GLPK interface}
+
+TODO how long it takes to create GLPK input and parse its output
+
 \subsection{GLPK: native vs. Cygwin}
@@
 %         NB class TimeQuality
+%         NB class TimeTillOptimum
 
@@
 
-\subsection{Still to do}
-
-TODO following experiments + more
-
-\begin{itemize}
-  \item How long does it take to communicate with GLPK? Creating the input / parsing the output.
-  \item Big fight Fuzzy vs. Random vs. FIDAX take note of quality (mean, variance) but also time (again, mean)
-  \item Can FIDAX be improved by Hungry?
-
-	\item Which construction heuristic is the best - limit GLPK to like 1 second
-
-	\item With the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
-	\item With the best IH from the previous step, how about using Random as CH?
-
-  \item this can be done inside the GLPK version comparison! GLPK: Input -> How long the search for optimum lasts - this shows we cannot rely on GLPK alone! it just takes too long.
-	\item this can be done inside the GLPK version comparison! GLPK: Time limit -> Quality
-
-	\item Structure of the data -> Which algorithms perform the best
-
-  \item Alpha, Beta -> Quality
-	\item Alpha, Beta -> How does the solution work?
-
-  \item Can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-\end{itemize}
+TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
+
+\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
+
+TODO take note of quality and time too
+
+\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
+
+TODO
+
+\subsection{Best standalone CH}
+
+TODO compare all CHs - limit GLPK to ~ 1 second
+
+\subsection{Best IH for the best CH}
+
+TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
+
+\subsubsection{\jmodule{Random} as CH}
+
+TODO with the best IH from the previous step, how about using Random as CH?
+
+\subsection{Various $\alpha$, $\beta$}
+
+TODO remind what these parameters mean
+
+\subsubsection{Effects on performance}
+
+TODO
+
+\subsubsection{Effects on the ID set}
+
+TODO
+
+\subsection{Ignoring text data}
+
+TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
+
+\subsection{Chaining the IHs}
+
+TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
+
+\subsection{Algorithms for various data characteristics}
+
+TODO
 

### 2011-08-15 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
-The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{ch-fidax}.
+The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{listing-ch-fidax}.
 
@@
 \caption{FIDAX CH}
-\label{ch-fidax}
+\label{listing-ch-fidax}
 \begin{algorithmic}
@@
 
-See the Listing \ref{ch-random} for its pseudocode.
+See the Listing \ref{listing-ch-random} for its pseudocode.
 
@@
 \caption{Random CH}
-\label{ch-random}
+\label{listing-ch-random}
 \begin{algorithmic}
@@
 
-Again, this is a hungry heuristic, the Listing \ref{ch-fuzzy} contains its pseudocode.
+Again, this is a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
 
@@
 \caption{Fuzzy CH}
-\label{ch-fuzzy}
+\label{listing-ch-fuzzy}
 \begin{algorithmic}
@@
 
-See listing \ref{ch-incremental}.
+See listing \ref{listing-ch-incremental}.
 
@@
 \caption{Incremental CH}
-\label{ch-incremental}
+\label{listing-ch-incremental}
 \begin{algorithmic}
@@
 
-See listing \ref{ch-removal}.
+See listing \ref{listing-ch-removal}.
 
@@
 \caption{Removal CH}
-\label{ch-removal}
+\label{listing-ch-removal}
 \begin{algorithmic}
@@
 
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{ih-identity}.
+This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{listing-ih-identity}.
 
@@
 \caption{Identity IH}
-\label{ih-identity}
+\label{listing-ih-identity}
 \begin{algorithmic}
@@
 
-See listing \ref{ih-removeworst}.
+See listing \ref{listing-ih-removeworst}.
 
@@
 \caption{Remove Worst IH}
-\label{ih-removeworst}
+\label{listing-ih-removeworst}
 \begin{algorithmic}
@@
 
-See listing \ref{ih-randomremove}.
+See listing \ref{listing-ih-randomremove}.
 
@@
 \caption{Random Remove IH}
-\label{ih-randomremove}
+\label{listing-ih-randomremove}
 \begin{algorithmic}
@@
 The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{ih-hungry} captures this process.
+Listing \ref{listing-ih-hungry} captures this process.
 
@@
 \caption{Hungry IH}
-\label{ih-hungry}
+\label{listing-ih-hungry}
 \begin{algorithmic}
@@
 
-See listing \ref{ih-mutation}.
+See listing \ref{listing-ih-mutation}.
 
@@
 \caption{Mutation IH}
-\label{ih-mutation}
+\label{listing-ih-mutation}
 \begin{algorithmic}
@@
 
-See listing \ref{ih-crossover}.
+See listing \ref{listing-ih-crossover}.
 
@@
 \caption{Crossover IH}
-\label{ih-crossover}
+\label{listing-ih-crossover}
 \begin{algorithmic}
@@
 
-See listing \ref{ih-localbranching}.
+See listing \ref{listing-ih-localbranching}.
 
@@
 \caption{Local Branching IH}
-\label{ih-localbranching}
+\label{listing-ih-localbranching}
 \begin{algorithmic}

### 2011-08-16 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
+The latter will come in form of tables containing information regarding the data sets, such their size, known optimum % TODO footnote explaining that we got the optimum by running GLPK w/o time limit
+for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
 
@@
 We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
-Note that the name is always in the form $v-e$.
+Note that the name is always in the form $v-e$.\\
+
+All of the experimental data sets mentioned so far, realistic, converted and artificial alike will be referred to as \textit{official test data (sets)}.\\
 
@@
 For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
+From now on, these will be referred to as \textit{sized test data (sets)}.
 We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
@@
 
+\begin{center}
+\bigskip
 \begin{tabular}{| l | l |}
@@
   Machine           & TODO \\
-  Input data set    & all test data sets \\
+  Input data        & all official and sized test data sets \\
   Iterations        & 20 \\
@@
 \end{tabular}
+\bigskip  
+\end{center}
 
@@
 
+Apart from comparing different versions, we shall see how the pure GLPK approach behaves.
+The first part of this experiment will be limiting the run time, thus making it an instance of Truncated branch \& bound.
+In this case we will see the dependency between the run time and the quality achieved in it.
+In the second time we will let GLPK run until optimum is found.
+We shall see the dependency between input size and run time needed to achieve the optimum.
+
+\begin{center}
+\bigskip
 \begin{tabular}{| l | l |}
@@
   Machine           & Windows \\
-  Input data set    & \texttt{100-500} \\
+  Input data        & \texttt{100-500} \\
   Iterations        & 20 \\
@@
 \end{tabular}
+\bigskip
+\end{center}
 
@@
   Machine           & Windows \\
-  Input data set    & sized artificial data \\
+  Input data        & all sized test data sets \\
   Iterations        & 20 \\
@@
 \FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized artificial data}
+	\FOR{$file \in $ sized test data}
     \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
@@
 
-TODO graph + interpretation
+TODO graph + interpretation. Graph should have its Y axis in log, stress this fact in the text too
 
@@
 \subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
+
+%         NB class RandomVsFuzzyVsFidaxStart
+
+Our investigation into various CH will start by comparing \jmodule{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \jmodule{Random} and \jmodule{Fuzzy}.
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & TODO \\
+  Input data set    & all official test data sets \\
+  Iterations        & 20 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
 

### 2011-08-16 doc/theses/vitasek/appendix-a-jinfer.tex
Thesis: splitting tex files.

--- 2011-08-15 doc/theses/vitasek/diplomka.tex
@@
-\documentclass[a4paper,12pt,oneside]{report}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{url}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{tgpagella}
-\usepackage{xstring}
-\usepackage{wrapfig}
-\usepackage{appendix}
-\usepackage{nomencl}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsthm}
-
-\newcommand{\myscale}{0.74}
-\newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
-\def\.{\discretionary{}{}{}}
-\newcommand{\heu}[1]{\texttt{\textit{#1}}}
-\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
-\linespread{1.5}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
-
-\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
-\setlength{\voffset}{-1in} %analogous voffset
-\setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
-\setlength{\topmargin}{25mm}
-\setlength{\footskip}{1cm}
-\setlength{\headheight}{0cm}
-\setlength{\headsep}{0cm}
-\setlength{\marginparwidth}{0cm}
-\setlength{\marginparpush}{0cm}
-\setlength{\textheight}{23.7cm}
-\setlength{\textwidth}{14.5cm}
-\let\openright=\clearpage
-
-\def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-\begin{document}
-
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce ======================================================
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% exactly as assigned
-{\LARGE\bfseries Inference of XML Integrity Constraints}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & 	\mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage % ============================================================
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-\openright
-
-\noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
-
-\newpage % ============================================================
-%%% Strana s čestným prohlášením k diplomové práci
-\vglue 0pt plus 1fill
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature
-\hss}}
-
-\vspace{20mm}
-
-\newpage % ============================================================
-%%% Povinná informační strana diplomové práce
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
-
-Autor: \mfauthor
-
-Katedra:  Katedra softwarového inženýrství
-
-Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství
-
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
-
-Klíčová slova:  XML, ID atributy, odvozování
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:  Inference of XML Integrity Constraints
-
-Author: \mfauthor
-
-Department: Department of Software Engineering
-
-Supervisor: \mfadvisor{}, Department of Software Engineering
-
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
-
-Keywords: XML, ID attributes, inference
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\setcounter{page}{1}
-\tableofcontents
-
-% =============== TEXT ================
-\newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
-
-\chapwithtoc{Preface}
-
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
-
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
-
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
-
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
-
-\section{Structure of the thesis}
-
-The thesis will be structured as follows.
-
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
-
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
-
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
-
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
-
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
-
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
-
-\section{Conventions}
-
-TODO list conventions - how we write code related stuff, module names, heu names, etc
-
-\chapter{Definitions}
-
-\begin{itemize}
-	\item XML tree - from FIDAX or wherever
-	\item Element, Attribute
-	\item ID, IDREF, IDREFS attribute according to specification
-	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
-	\item ID set
-  \item Weight: support, coverage
-\end{itemize}
-
-\chapter{Research}
-
-According to the article \cite{fidax}, ...
-
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
-
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
-
-TODO other approaches to max IS
-
-\chapter{MIP Approach}
-
-TODO rigorously define MIP - optimization problem, some of the variables integral
-
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
-
-\section{Finding ID sets with GLPK}
-
-TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
-
- -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
-
-\section{Heuristics}
-
-TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
-
-TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
-
-TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
-
-TODO we will be using a pool - what is a pool
-
-\begin{figure}
-  \caption{Metaheuristic schema}
-  \label{image-metaheuristic}
-  \centering
-    \includegraphics[width=\textwidth]{images/metaheuristic}
-\end{figure}
-
-\subsection{Constructions Heuristics}
-% TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
-
-TODO construction heuristics are heus that provide us with at least some solution.
-
-\subsubsection{\jmodule{FIDAX}}
-
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
-
-The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{ch-fidax}.
-
-\begin{algorithm}
-\caption{FIDAX CH}
-\label{ch-fidax}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets C$ sorted by decreasing size
-\STATE Compute the weight $w(m)$ of each $m$ in $C$
-\FORALL{$t$ in $\Sigma^E$}
-  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
-  \STATE Remove from $C'$ all mappings of type $t$ except $m$
-\ENDFOR
-\FORALL{$m$ in $C'$}
-  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
-  \IF{$w(m) > \sum_{p \in S} w(p)$}
-    \STATE remove all $p \in S$ from $C'$
-  \ELSE
-    \STATE remove $m$ from $C'$
-  \ENDIF
-\ENDFOR
-\RETURN $C'$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random}}
-\label{heu-ch-random}
-
-One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a hungry heuristic.
-%TODO have I defined ID set condition yet??
-
-The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
-As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
-
-See the Listing \ref{ch-random} for its pseudocode.
-
-\begin{algorithm}
-\caption{Random CH}
-\label{ch-random}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \WHILE{$s$ is a feasible ID set}
-    \STATE $a \gets $ pick at random from $C \backslash S$
-    \STATE $s \gets s \cup a$
-  \ENDWHILE
-  \STATE $r \gets r \cup s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Fuzzy}}
-\label{heu-ch-fuzzy}
-
-\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in TODO.
-Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
-
-Again, this is a hungry heuristic, the Listing \ref{ch-fuzzy} contains its pseudocode.
-
-% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
-
-\begin{algorithm}
-\caption{Fuzzy CH}
-\label{ch-fuzzy}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \STATE $C' \gets C$
-
-  \WHILE{$C' $ not empty}
-    \STATE $a \gets $ pick at weighted random from $C'$
-    \IF{$s \cup a$ is a feasible ID set}
-      \STATE $s \gets s \cup a$
-      \STATE $C' \gets C' \backslash a$
-    \ENDIF
-    \FORALL{$c \in C'$}
-      \IF{$s \cup c $ is \textbf{not} a feasible ID set}
-        \STATE \COMMENT {if $c$ cannot be possibly added anymore}
-        \STATE $C' \gets C' \backslash c$
-      \ENDIF
-    \ENDFOR
-  \ENDWHILE
-
-  \STATE $r \gets r + s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Incremental}}
-
-This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
-This way it can create only one solution, and again, this is a hungry heuristic.
-
-See listing \ref{ch-incremental}.
-
-\begin{algorithm}
-\caption{Incremental CH}
-\label{ch-incremental}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by decreasing weight
-\STATE $s \gets $ empty solution
-\FORALL{$c \in C'$}
-  \IF{$s \cup c$ is a feasible ID set}
-    \STATE $s \gets s + c$
-  \ENDIF
-\ENDFOR
-\RETURN $s$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Removal}}
-
-This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
-This probably does not satisfy the ID set condition.
-Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
-Again, this is a hungry heuristic returning only one solution.
-
-See listing \ref{ch-removal}.
-
-\begin{algorithm}
-\caption{Removal CH}
-\label{ch-removal}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by increasing weight
-\STATE $s \gets C'$
-\FORALL{$c \in s$}
-  \IF{$s$ is a feasible ID set}
-    \RETURN $s$
-  \ENDIF
-  \STATE $s \gets s \backslash c$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Truncated Branch \& Bound}
-
-This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
-
-TODO if we limit GLPK's runtime, we get this
-
-\subsection{Improvement Heuristics}
-
-\nomenclature{IH}{Improvement Heuristic}
-
-TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
-
-TODO mention intensification, diversification
-
-\subsubsection{\jmodule{Identity}}
-
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{ih-identity}.
-
-\begin{algorithm}
-\caption{Identity IH}
-\label{ih-identity}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE the same pool of feasible solutions
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Remove Worst}}
-
-This trivial IH tries to improve the solution pool by removing the worst solution (i.e.
-the one with the lowest quality).
-This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
-
-See listing \ref{ih-removeworst}.
-
-\begin{algorithm}
-\caption{Remove Worst IH}
-\label{ih-removeworst}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE pool of feasible solutions
-\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
-\RETURN $FP \backslash s_{min}$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random Remove}}
-
-This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
-By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
-
-The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
-For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
-
-This heuristic returns a pool of solutions of the same size as it got on its input.
-
-See listing \ref{ih-randomremove}.
-
-\begin{algorithm}
-\caption{Random Remove IH}
-\label{ih-randomremove}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE $K \gets k * |s|$
-  \STATE remove $K$ random AMs from $s$
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Hungry}}
-
-This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
-
-\heu{Hungry} tries to improve each solution in the feasible pool in the following way.
-It orders all candidate AMs not present in the solution by decreasing weight.
-Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition.
-The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{ih-hungry} captures this process.
-
-\begin{algorithm}
-\caption{Hungry IH}
-\label{ih-hungry}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE \COMMENT {improve a single solution}
-  \STATE $C' \gets C \backslash s$
-  \STATE $C' \gets C'$ sorted by decreasing weight
-  \FORALL{$c \in C'$}
-    \IF{$s \cup c$ is a feasible ID set}
-      \STATE $s \gets s \cup c$
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Mutation}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-mutation}.
-
-\begin{algorithm}
-\caption{Mutation IH}
-\label{ih-mutation}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of AMs to fix
-\ENSURE pool of feasible solutions
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE $K \gets k * |incumbent|$
-\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Crossover}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-crossover}.
-
-\begin{algorithm}
-\caption{Crossover IH}
-\label{ih-crossover}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of solutions to look for commonalities in
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |FP|$
-\STATE $FP' \gets K$ random solutions $\in FP$
-\STATE $am \gets$ AMs found in all solutions $\in FP'$
-\STATE fix $am$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Local Branching}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-localbranching}.
-
-\begin{algorithm}
-\caption{Local Branching IH}
-\label{ih-localbranching}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |$total AM count$|$
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE add max Hamming distance requirement to GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\section{IDREF}
-
-TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
-
-\chapter{Experiments}
-
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
-
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
-
-First of all, we would like to get an idea of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
-
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
-
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
-
-\section{Experimental Data}
-
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
-
-\begin{itemize}
-	\item Realistic 
-	\item Realistic with artificial (converted) attributes
-	\item Artificial
-\end{itemize}
-
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
-
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
-
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
-Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
-
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
-
-\subsection{Realistic data}
-
-From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
-
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
-
-\begin{table}
-  \caption{List of realistic test data files}
-  \bigskip
-  \label{table-experiments-data-realistic}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
-	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data}
-  \label{image-experiments-data-realistic}
-  \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
-\end{figure}
-
-To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
-
-On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
-
-Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
-
-\subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
-
-We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
-None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
-
-TODO schemas?
-
-To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
-
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
-
-\begin{table}
-  \caption{List of realistic test data files with converted attributes}
-  \bigskip
-  \label{table-experiments-data-converted}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data with converted attributes}
-  \label{image-experiments-data-converted}
-  \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
-\end{figure}
-
-Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
-
-In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
-Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
-
-\subsection{Artificial data}
-
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
-
-This is actually simple. Consider the following excerpt from an XML file:
-
-\begin{scriptsize}
-\begin{verbatim}
-<graph>
-  <vertex0 attr="-2968876296119015800"/>
-  <vertex1 attr="1729745997570096518"/>
-  <vertex2 attr="-9020549659620928934"/>
-  ...
-  <vertex99 attr="-7545982394508643394"/>
-
-  <vertex82 attr="0"/><vertex21 attr="0"/>
-  <vertex64 attr="1"/><vertex21 attr="1"/>
-  <vertex44 attr="2"/><vertex2 attr="2"/>
-  ...
-  <vertex96 attr="99"/><vertex40 attr="99"/>
-</graph>
-\end{verbatim}
-\end{scriptsize}
-
-We want to create a graph with around $v$ vertices and $e$ edges.
-First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
-
-The pseudocode for this is in listing \ref{listing-random-data}.
-
-\begin{algorithm}
-\caption{Random XML data creation}
-\label{listing-random-data}
-\begin{algorithmic}
-\REQUIRE $v$ requested number of vertices
-\REQUIRE $e$ requested number of edges
-\ENSURE XML file content
-\PRINT \texttt{<graph>}
-\FOR{$i = 1 \to |V|$}
-	\STATE $R \gets RANDOM$
-	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
-\ENDFOR
-\FOR{$i = 1 \to |E|$}
-	\STATE $v1 \gets RANDOM(|V|)$
-	\STATE $v2 \gets RANDOM(|V|)$
-	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
-\ENDFOR
-\PRINT \texttt{</graph>}
-\RETURN
-\end{algorithmic}
-\end{algorithm}
-
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
-
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
-
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
-Note that the name is always in the form $v-e$.
-
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
-
-TODO this is on the DVD.
-
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
-
-\begin{table}
-  \caption{List of artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
-    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{List of ``sized" artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial-size}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
-  20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-  30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
-	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Artificial data}
-  \label{image-experiments-data-artificial}
-  \centering
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
-\end{figure}
-
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
-
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
-
-\section{Experimental Setup}
-
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
-
-We now have to define a few notions before moving forward to the description of our experiments.
-
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
-	\begin{itemize}
-		\item All the parameters in all the heuristics.
-		\item The specific way in which the heuristics are chained.
-		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
-		\item Initial pool size.
-		\item The termination criteria.
-		\item The input XML file.
-		\item Known optimum for this file and $\alpha$, $\beta$.
-	\end{itemize}
-\end{define}
-
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
-
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
-
-\subsection{Grammar and Model Creation}
-
-TODO link BasicIGG to see how we get IG
-
-TODO describe how we get the model
-
-\subsection{Hardware and Software}
-
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Linux Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 64-Bit Server VM (build 20.1-b02)
-GLPK version 4.38
-\end{verbatim}
-
-\subsubsection{Windows Machine}
-
-\begin{verbatim}
-Intel Core 2 Duo processor @ 2.33 GHz
-4 GB DDR2 RAM
-Windows 7 SP1 64bit
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 32-Bit Client VM (build 20.1-b02)
-GLPK version 4.45 (Cygwin)
-GLPK version 4.34 (native)
-\end{verbatim}
-
-\subsection{Methodology}
-
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
-This improvement is left for the future work.
-
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
-OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
-
-\subsection{Measuring the time}
-
-TODO
-
-\subsection{Obtaining the results}
-
-Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
-We can get all the information relevant to that experiment run from this trace alone.
-An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
-
-\section{Experimental Results}
-
-\subsection{Grammar and model generation}
-
-%         NB class GrammarModelTiming
-
-TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
-
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & TODO \\
-  Input data set    & all test data sets \\
-  Iterations        & 20 \\
-  Pool size         & not applicable \\
-  $\alpha$, $\beta$ & not applicable \\
-  \hline
-\end{tabular}
-
-The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There won't be any CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
-
-TODO graph + interpretation
-
-\subsubsection{GLPK interface}
-
-TODO how long it takes to create GLPK input and parse its output
-
-\subsection{GLPK: native vs. Cygwin}
-
-%         NB class TimeQuality
-%         NB class TimeTillOptimum
-
-In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & \texttt{100-500} \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
-There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
-\label{listing-experiment-glpk-native-vs-cygwin-1}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & sized artificial data \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
-Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
-\label{listing-experiment-glpk-native-vs-cygwin-2}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized artificial data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
-\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
-
-TODO take note of quality and time too
-
-\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
-
-TODO
-
-\subsection{Best standalone CH}
-
-TODO compare all CHs - limit GLPK to ~ 1 second
-
-\subsection{Best IH for the best CH}
-
-TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
-
-\subsubsection{\jmodule{Random} as CH}
-
-TODO with the best IH from the previous step, how about using Random as CH?
-
-\subsection{Various $\alpha$, $\beta$}
-
-TODO remind what these parameters mean
-
-\subsubsection{Effects on performance}
-
-TODO
-
-\subsubsection{Effects on the ID set}
-
-TODO
-
-\subsection{Ignoring text data}
-
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-
-\subsection{Chaining the IHs}
-
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
-
-\subsection{Algorithms for various data characteristics}
-
-TODO
-
-\section{The "Best" Algorithm}
-
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run. 
-We will find the optimum in this way. 
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
-
-\chapter{Future work}
-
-TODO
-
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
-
-\chapwithtoc{Conclusion}
-
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
-
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
-
-%%% Seznam použité literatury
-\newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-
-\listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
-
-\listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
-
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
-\printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
-
-\openright
-\addcontentsline{toc}{chapter}{Appendices}
-
 \chapter{jInfer}
@@
 This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
-
-\chapter{IDSetSearch}
-\label{appendix-iss}
-
-TODO talk a bit about the structure of ISS module.
-
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
-
-TODO package name root
-
-TODO "Attribute Stats"
-
-TODO heuristics - interfaces, construction, improvement
-
-TODO experiments - quality, termination, parameters, experiment sets
-
-TODO make sure there was a place where we described the model data model:)
-
-\chapter{Experimental Trace}
-\label{appendix-trace}
-
-Following is a trace logged from a sample experiment run. It shows all the relevant information related to this instance, any and every piece of information we might be interested in.
-
-To save space, 2-column layout is used. Commentary on the particulars follows right after its end.
-
-\begin{multicols}{2}
-\begin{scriptsize}
-\begin{verbatim}
-CPU info
-  Intel(R) Core(TM)2 Quad CPU Q9550 @ 2.83GHz
-  Cores: 4
-  Clock speed: 2983 MHz
-Memory info
-  Size: 8192 MB
-OS info
-  Name: Windows 7
-  Version: 6.1
-  Architecture: amd64
-Java info
-  Version: 1.6.0_26
-  VM: Java HotSpot(TM) 64-Bit Server VM
-GLPK info
-  GLPSOL: GLPK LP/MIP Solver 4.34
-
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-
-Results:
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-Construction phase: 
-  Algorithm: Random
-    Time taken: 248 ms
-    Time since start: 248 ms
-    Pool size: 10
-    Quality: 0.19975609756097568 (11 AMs)
-Improvement phase: 
-  pass #1: 
-  Algorithm: RandomRemove, ratio = 0.2
-    Time taken: 0 ms
-    Time since start: 841 ms
-    Pool size: 10
-    Quality: 0.15878048780487808 (9 AMs)
-  pass #2: 
-  Algorithm: Mutation, ratio = 0.1, limit = 1 s
-    Time taken: 1512 ms
-    Time since start: 2710 ms
-    Pool size: 11
-    Quality: 0.21975609756097558 (11 AMs)
-
-  <... 7 more passes removed ...>
-
-  pass #10: 
-  Algorithm: Remove Worst
-    Time taken: 80 ms
-    Time since start: 7676 ms
-    Pool size: 12
-    Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
-
-Time,Quality,AMs
-248,0.19975609756097568,11
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-2927,0.1890243902439024,9
-4421,0.23463414634146343,12
-4703,0.23463414634146343,12
-4896,0.1960975609756098,10
-5793,0.23463414634146337,12
-5972,0.19951219512195123,10
-7433,0.19951219512195123,10
-7676,0.19951219512195123,10
-
-ID
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-vertex2,attr,0.01975609756097561
-vertex33,attr,0.016829268292682928
-vertex34,attr,0.02219512195121951
-vertex4,attr,0.022682926829268292
-vertex41,attr,0.014878048780487804
-vertex7,attr,0.02170731707317073
-vertex70,attr,0.018780487804878048
-vertex76,attr,0.01780487804878049
-vertex8,attr,0.02170731707317073
-vertex80,attr,0.01780487804878049
-vertex97,attr,0.016341463414634147
-
-IDREF
-Element,Attribute
-\end{verbatim}
-\end{scriptsize}
-\end{multicols}
-
-The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
-To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
-It tries to guess whether it's the Cygwin version by looking at the path to the binary.
-
-% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
-The second section states the input file along with its size and graph representation.\\
-Alpha and beta parameters for this instance belong here too.
-
-\begin{footnotesize}
-\begin{verbatim}
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-\end{verbatim}
-\end{footnotesize}
-
-Results section opens stating the most important information first: how long did the experiment run and what was the highest and final quality (these two are potentially different).
-Numbers of attribute mappings in the best and final solution respectively are stated as well.
-
-\begin{footnotesize}
-\begin{verbatim}
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Construction phase results go next.
-Among reported information are the full identification of the heuristic (possibly along with its parameters), time taken, size of the pool created and the quality of the incumbent solution (again, with the number of its AMs).
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Random
-  Time taken: 248 ms
-  Time since start: 248 ms
-  Pool size: 10
-  Quality: 0.19975609756097568 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Now for each of the improvement phases there is one section in output log.
-Information presented here has the same structure as with the construction phase.
-Please note that the \code{Pool size} is always measured \textit{after} the improvement run.
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Mutation, ratio = 0.1, limit = 1 s
-  Time taken: 1512 ms
-  Time since start: 2710 ms
-  Pool size: 11
-  Quality: 0.21975609756097558 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
-\\
-
-To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
-Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
-
-\begin{footnotesize}
-\begin{verbatim}
-Time,Quality,AMs
-...
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-...
-\end{verbatim}
-\end{footnotesize}
-
-And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
-Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
-
-\begin{footnotesize}
-\begin{verbatim}
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-...
-\end{verbatim}
-\end{footnotesize}
-
-Note that in this example trace there were no IDREF AMs found.
-
-\flushright
-$\square$
-
-\end{document}

### 2011-08-16 doc/theses/vitasek/appendix-c-trace.tex
Thesis: splitting tex files.

--- 2011-08-15 doc/theses/vitasek/diplomka.tex
@@
-\documentclass[a4paper,12pt,oneside]{report}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{url}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{tgpagella}
-\usepackage{xstring}
-\usepackage{wrapfig}
-\usepackage{appendix}
-\usepackage{nomencl}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsthm}
-
-\newcommand{\myscale}{0.74}
-\newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
-\def\.{\discretionary{}{}{}}
-\newcommand{\heu}[1]{\texttt{\textit{#1}}}
-\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
-\linespread{1.5}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
-
-\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
-\setlength{\voffset}{-1in} %analogous voffset
-\setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
-\setlength{\topmargin}{25mm}
-\setlength{\footskip}{1cm}
-\setlength{\headheight}{0cm}
-\setlength{\headsep}{0cm}
-\setlength{\marginparwidth}{0cm}
-\setlength{\marginparpush}{0cm}
-\setlength{\textheight}{23.7cm}
-\setlength{\textwidth}{14.5cm}
-\let\openright=\clearpage
-
-\def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-\begin{document}
-
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce ======================================================
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% exactly as assigned
-{\LARGE\bfseries Inference of XML Integrity Constraints}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & 	\mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage % ============================================================
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-\openright
-
-\noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
-
-\newpage % ============================================================
-%%% Strana s čestným prohlášením k diplomové práci
-\vglue 0pt plus 1fill
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature
-\hss}}
-
-\vspace{20mm}
-
-\newpage % ============================================================
-%%% Povinná informační strana diplomové práce
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
-
-Autor: \mfauthor
-
-Katedra:  Katedra softwarového inženýrství
-
-Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství
-
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
-
-Klíčová slova:  XML, ID atributy, odvozování
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:  Inference of XML Integrity Constraints
-
-Author: \mfauthor
-
-Department: Department of Software Engineering
-
-Supervisor: \mfadvisor{}, Department of Software Engineering
-
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
-
-Keywords: XML, ID attributes, inference
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\setcounter{page}{1}
-\tableofcontents
-
-% =============== TEXT ================
-\newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
-
-\chapwithtoc{Preface}
-
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
-
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
-
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
-
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
-
-\section{Structure of the thesis}
-
-The thesis will be structured as follows.
-
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
-
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
-
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
-
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
-
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
-
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
-
-\section{Conventions}
-
-TODO list conventions - how we write code related stuff, module names, heu names, etc
-
-\chapter{Definitions}
-
-\begin{itemize}
-	\item XML tree - from FIDAX or wherever
-	\item Element, Attribute
-	\item ID, IDREF, IDREFS attribute according to specification
-	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
-	\item ID set
-  \item Weight: support, coverage
-\end{itemize}
-
-\chapter{Research}
-
-According to the article \cite{fidax}, ...
-
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
-
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
-
-TODO other approaches to max IS
-
-\chapter{MIP Approach}
-
-TODO rigorously define MIP - optimization problem, some of the variables integral
-
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
-
-\section{Finding ID sets with GLPK}
-
-TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
-
- -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
-
-\section{Heuristics}
-
-TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
-
-TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
-
-TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
-
-TODO we will be using a pool - what is a pool
-
-\begin{figure}
-  \caption{Metaheuristic schema}
-  \label{image-metaheuristic}
-  \centering
-    \includegraphics[width=\textwidth]{images/metaheuristic}
-\end{figure}
-
-\subsection{Constructions Heuristics}
-% TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
-
-TODO construction heuristics are heus that provide us with at least some solution.
-
-\subsubsection{\jmodule{FIDAX}}
-
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
-
-The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{ch-fidax}.
-
-\begin{algorithm}
-\caption{FIDAX CH}
-\label{ch-fidax}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets C$ sorted by decreasing size
-\STATE Compute the weight $w(m)$ of each $m$ in $C$
-\FORALL{$t$ in $\Sigma^E$}
-  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
-  \STATE Remove from $C'$ all mappings of type $t$ except $m$
-\ENDFOR
-\FORALL{$m$ in $C'$}
-  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
-  \IF{$w(m) > \sum_{p \in S} w(p)$}
-    \STATE remove all $p \in S$ from $C'$
-  \ELSE
-    \STATE remove $m$ from $C'$
-  \ENDIF
-\ENDFOR
-\RETURN $C'$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random}}
-\label{heu-ch-random}
-
-One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a hungry heuristic.
-%TODO have I defined ID set condition yet??
-
-The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
-As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
-
-See the Listing \ref{ch-random} for its pseudocode.
-
-\begin{algorithm}
-\caption{Random CH}
-\label{ch-random}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \WHILE{$s$ is a feasible ID set}
-    \STATE $a \gets $ pick at random from $C \backslash S$
-    \STATE $s \gets s \cup a$
-  \ENDWHILE
-  \STATE $r \gets r \cup s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Fuzzy}}
-\label{heu-ch-fuzzy}
-
-\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in TODO.
-Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
-
-Again, this is a hungry heuristic, the Listing \ref{ch-fuzzy} contains its pseudocode.
-
-% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
-
-\begin{algorithm}
-\caption{Fuzzy CH}
-\label{ch-fuzzy}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \STATE $C' \gets C$
-
-  \WHILE{$C' $ not empty}
-    \STATE $a \gets $ pick at weighted random from $C'$
-    \IF{$s \cup a$ is a feasible ID set}
-      \STATE $s \gets s \cup a$
-      \STATE $C' \gets C' \backslash a$
-    \ENDIF
-    \FORALL{$c \in C'$}
-      \IF{$s \cup c $ is \textbf{not} a feasible ID set}
-        \STATE \COMMENT {if $c$ cannot be possibly added anymore}
-        \STATE $C' \gets C' \backslash c$
-      \ENDIF
-    \ENDFOR
-  \ENDWHILE
-
-  \STATE $r \gets r + s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Incremental}}
-
-This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
-This way it can create only one solution, and again, this is a hungry heuristic.
-
-See listing \ref{ch-incremental}.
-
-\begin{algorithm}
-\caption{Incremental CH}
-\label{ch-incremental}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by decreasing weight
-\STATE $s \gets $ empty solution
-\FORALL{$c \in C'$}
-  \IF{$s \cup c$ is a feasible ID set}
-    \STATE $s \gets s + c$
-  \ENDIF
-\ENDFOR
-\RETURN $s$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Removal}}
-
-This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
-This probably does not satisfy the ID set condition.
-Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
-Again, this is a hungry heuristic returning only one solution.
-
-See listing \ref{ch-removal}.
-
-\begin{algorithm}
-\caption{Removal CH}
-\label{ch-removal}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by increasing weight
-\STATE $s \gets C'$
-\FORALL{$c \in s$}
-  \IF{$s$ is a feasible ID set}
-    \RETURN $s$
-  \ENDIF
-  \STATE $s \gets s \backslash c$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Truncated Branch \& Bound}
-
-This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
-
-TODO if we limit GLPK's runtime, we get this
-
-\subsection{Improvement Heuristics}
-
-\nomenclature{IH}{Improvement Heuristic}
-
-TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
-
-TODO mention intensification, diversification
-
-\subsubsection{\jmodule{Identity}}
-
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{ih-identity}.
-
-\begin{algorithm}
-\caption{Identity IH}
-\label{ih-identity}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE the same pool of feasible solutions
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Remove Worst}}
-
-This trivial IH tries to improve the solution pool by removing the worst solution (i.e.
-the one with the lowest quality).
-This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
-
-See listing \ref{ih-removeworst}.
-
-\begin{algorithm}
-\caption{Remove Worst IH}
-\label{ih-removeworst}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE pool of feasible solutions
-\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
-\RETURN $FP \backslash s_{min}$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random Remove}}
-
-This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
-By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
-
-The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
-For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
-
-This heuristic returns a pool of solutions of the same size as it got on its input.
-
-See listing \ref{ih-randomremove}.
-
-\begin{algorithm}
-\caption{Random Remove IH}
-\label{ih-randomremove}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE $K \gets k * |s|$
-  \STATE remove $K$ random AMs from $s$
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Hungry}}
-
-This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
-
-\heu{Hungry} tries to improve each solution in the feasible pool in the following way.
-It orders all candidate AMs not present in the solution by decreasing weight.
-Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition.
-The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{ih-hungry} captures this process.
-
-\begin{algorithm}
-\caption{Hungry IH}
-\label{ih-hungry}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE \COMMENT {improve a single solution}
-  \STATE $C' \gets C \backslash s$
-  \STATE $C' \gets C'$ sorted by decreasing weight
-  \FORALL{$c \in C'$}
-    \IF{$s \cup c$ is a feasible ID set}
-      \STATE $s \gets s \cup c$
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Mutation}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-mutation}.
-
-\begin{algorithm}
-\caption{Mutation IH}
-\label{ih-mutation}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of AMs to fix
-\ENSURE pool of feasible solutions
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE $K \gets k * |incumbent|$
-\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Crossover}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-crossover}.
-
-\begin{algorithm}
-\caption{Crossover IH}
-\label{ih-crossover}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of solutions to look for commonalities in
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |FP|$
-\STATE $FP' \gets K$ random solutions $\in FP$
-\STATE $am \gets$ AMs found in all solutions $\in FP'$
-\STATE fix $am$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Local Branching}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-localbranching}.
-
-\begin{algorithm}
-\caption{Local Branching IH}
-\label{ih-localbranching}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |$total AM count$|$
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE add max Hamming distance requirement to GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\section{IDREF}
-
-TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
-
-\chapter{Experiments}
-
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
-
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
-
-First of all, we would like to get an idea of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
-
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
-
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
-
-\section{Experimental Data}
-
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
-
-\begin{itemize}
-	\item Realistic 
-	\item Realistic with artificial (converted) attributes
-	\item Artificial
-\end{itemize}
-
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
-
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
-
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
-Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
-
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
-
-\subsection{Realistic data}
-
-From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
-
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
-
-\begin{table}
-  \caption{List of realistic test data files}
-  \bigskip
-  \label{table-experiments-data-realistic}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
-	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data}
-  \label{image-experiments-data-realistic}
-  \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
-\end{figure}
-
-To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
-
-On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
-
-Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
-
-\subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
-
-We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
-None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
-
-TODO schemas?
-
-To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
-
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
-
-\begin{table}
-  \caption{List of realistic test data files with converted attributes}
-  \bigskip
-  \label{table-experiments-data-converted}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data with converted attributes}
-  \label{image-experiments-data-converted}
-  \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
-\end{figure}
-
-Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
-
-In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
-Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
-
-\subsection{Artificial data}
-
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
-
-This is actually simple. Consider the following excerpt from an XML file:
-
-\begin{scriptsize}
-\begin{verbatim}
-<graph>
-  <vertex0 attr="-2968876296119015800"/>
-  <vertex1 attr="1729745997570096518"/>
-  <vertex2 attr="-9020549659620928934"/>
-  ...
-  <vertex99 attr="-7545982394508643394"/>
-
-  <vertex82 attr="0"/><vertex21 attr="0"/>
-  <vertex64 attr="1"/><vertex21 attr="1"/>
-  <vertex44 attr="2"/><vertex2 attr="2"/>
-  ...
-  <vertex96 attr="99"/><vertex40 attr="99"/>
-</graph>
-\end{verbatim}
-\end{scriptsize}
-
-We want to create a graph with around $v$ vertices and $e$ edges.
-First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
-
-The pseudocode for this is in listing \ref{listing-random-data}.
-
-\begin{algorithm}
-\caption{Random XML data creation}
-\label{listing-random-data}
-\begin{algorithmic}
-\REQUIRE $v$ requested number of vertices
-\REQUIRE $e$ requested number of edges
-\ENSURE XML file content
-\PRINT \texttt{<graph>}
-\FOR{$i = 1 \to |V|$}
-	\STATE $R \gets RANDOM$
-	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
-\ENDFOR
-\FOR{$i = 1 \to |E|$}
-	\STATE $v1 \gets RANDOM(|V|)$
-	\STATE $v2 \gets RANDOM(|V|)$
-	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
-\ENDFOR
-\PRINT \texttt{</graph>}
-\RETURN
-\end{algorithmic}
-\end{algorithm}
-
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
-
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
-
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
-Note that the name is always in the form $v-e$.
-
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
-
-TODO this is on the DVD.
-
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
-
-\begin{table}
-  \caption{List of artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
-    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{List of ``sized" artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial-size}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
-  20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-  30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
-	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Artificial data}
-  \label{image-experiments-data-artificial}
-  \centering
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
-\end{figure}
-
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
-
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
-
-\section{Experimental Setup}
-
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
-
-We now have to define a few notions before moving forward to the description of our experiments.
-
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
-	\begin{itemize}
-		\item All the parameters in all the heuristics.
-		\item The specific way in which the heuristics are chained.
-		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
-		\item Initial pool size.
-		\item The termination criteria.
-		\item The input XML file.
-		\item Known optimum for this file and $\alpha$, $\beta$.
-	\end{itemize}
-\end{define}
-
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
-
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
-
-\subsection{Grammar and Model Creation}
-
-TODO link BasicIGG to see how we get IG
-
-TODO describe how we get the model
-
-\subsection{Hardware and Software}
-
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Linux Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 64-Bit Server VM (build 20.1-b02)
-GLPK version 4.38
-\end{verbatim}
-
-\subsubsection{Windows Machine}
-
-\begin{verbatim}
-Intel Core 2 Duo processor @ 2.33 GHz
-4 GB DDR2 RAM
-Windows 7 SP1 64bit
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 32-Bit Client VM (build 20.1-b02)
-GLPK version 4.45 (Cygwin)
-GLPK version 4.34 (native)
-\end{verbatim}
-
-\subsection{Methodology}
-
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
-This improvement is left for the future work.
-
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
-OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
-
-\subsection{Measuring the time}
-
-TODO
-
-\subsection{Obtaining the results}
-
-Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
-We can get all the information relevant to that experiment run from this trace alone.
-An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
-
-\section{Experimental Results}
-
-\subsection{Grammar and model generation}
-
-%         NB class GrammarModelTiming
-
-TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
-
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & TODO \\
-  Input data set    & all test data sets \\
-  Iterations        & 20 \\
-  Pool size         & not applicable \\
-  $\alpha$, $\beta$ & not applicable \\
-  \hline
-\end{tabular}
-
-The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There won't be any CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
-
-TODO graph + interpretation
-
-\subsubsection{GLPK interface}
-
-TODO how long it takes to create GLPK input and parse its output
-
-\subsection{GLPK: native vs. Cygwin}
-
-%         NB class TimeQuality
-%         NB class TimeTillOptimum
-
-In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & \texttt{100-500} \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
-There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
-\label{listing-experiment-glpk-native-vs-cygwin-1}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & sized artificial data \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
-Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
-\label{listing-experiment-glpk-native-vs-cygwin-2}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized artificial data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
-\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
-
-TODO take note of quality and time too
-
-\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
-
-TODO
-
-\subsection{Best standalone CH}
-
-TODO compare all CHs - limit GLPK to ~ 1 second
-
-\subsection{Best IH for the best CH}
-
-TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
-
-\subsubsection{\jmodule{Random} as CH}
-
-TODO with the best IH from the previous step, how about using Random as CH?
-
-\subsection{Various $\alpha$, $\beta$}
-
-TODO remind what these parameters mean
-
-\subsubsection{Effects on performance}
-
-TODO
-
-\subsubsection{Effects on the ID set}
-
-TODO
-
-\subsection{Ignoring text data}
-
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-
-\subsection{Chaining the IHs}
-
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
-
-\subsection{Algorithms for various data characteristics}
-
-TODO
-
-\section{The "Best" Algorithm}
-
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run. 
-We will find the optimum in this way. 
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
-
-\chapter{Future work}
-
-TODO
-
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
-
-\chapwithtoc{Conclusion}
-
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
-
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
-
-%%% Seznam použité literatury
-\newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-
-\listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
-
-\listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
-
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
-\printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
-
-\openright
-\addcontentsline{toc}{chapter}{Appendices}
-
-\chapter{jInfer}
-\label{appendix-jInfer}
-
-This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
-Please see TODO link jInfer project web for complete information, documentation and download options.
-
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
-The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
-The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
-Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
-This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
-
-To the best of our knowledge, at the time of writing this thesis is jInfer the only public, open source and actually working solution for XML schema inference-related tasks.
-
-At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see \ref{image-inference-process}), responsibility of 3 different services - interchangeable modules.
-
-\begin{figure}
-  \caption{Inference process in jInfer}
-  \label{image-inference-process}
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/inference-process}
-\end{figure}
-
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
-This is the representation in which will the structure live until it is used to create the final product - the schema.
-As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
-As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
-(TODO picture?) IG is used to create the AM model used in this thesis, too.
-jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
-
-After leaving the \jmodule{Initial Grammar Generator}, the IG needs to be made more general, shortened, \textit{simplified}.
-This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
-Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
-
-The last module, \jmodule{Schema Generator} takes the simplified grammar and creates the resulting schema from it.
-This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
-
-\chapter{IDSetSearch}
-\label{appendix-iss}
-
-TODO talk a bit about the structure of ISS module.
-
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
-
-TODO package name root
-
-TODO "Attribute Stats"
-
-TODO heuristics - interfaces, construction, improvement
-
-TODO experiments - quality, termination, parameters, experiment sets
-
-TODO make sure there was a place where we described the model data model:)
-
 \chapter{Experimental Trace}
@@
 Note that in this example trace there were no IDREF AMs found.
-
-\flushright
-$\square$
-
-\end{document}

### 2011-08-16 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: splitting tex files.

--- 2011-08-15 doc/theses/vitasek/diplomka.tex
@@
-\documentclass[a4paper,12pt,oneside]{report}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{url}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{tgpagella}
-\usepackage{xstring}
-\usepackage{wrapfig}
-\usepackage{appendix}
-\usepackage{nomencl}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsthm}
-
-\newcommand{\myscale}{0.74}
-\newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
-\def\.{\discretionary{}{}{}}
-\newcommand{\heu}[1]{\texttt{\textit{#1}}}
-\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
-\linespread{1.5}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
-
-\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
-\setlength{\voffset}{-1in} %analogous voffset
-\setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
-\setlength{\topmargin}{25mm}
-\setlength{\footskip}{1cm}
-\setlength{\headheight}{0cm}
-\setlength{\headsep}{0cm}
-\setlength{\marginparwidth}{0cm}
-\setlength{\marginparpush}{0cm}
-\setlength{\textheight}{23.7cm}
-\setlength{\textwidth}{14.5cm}
-\let\openright=\clearpage
-
-\def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-\begin{document}
-
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce ======================================================
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% exactly as assigned
-{\LARGE\bfseries Inference of XML Integrity Constraints}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & 	\mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage % ============================================================
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-\openright
-
-\noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
-
-\newpage % ============================================================
-%%% Strana s čestným prohlášením k diplomové práci
-\vglue 0pt plus 1fill
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature
-\hss}}
-
-\vspace{20mm}
-
-\newpage % ============================================================
-%%% Povinná informační strana diplomové práce
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
-
-Autor: \mfauthor
-
-Katedra:  Katedra softwarového inženýrství
-
-Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství
-
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
-
-Klíčová slova:  XML, ID atributy, odvozování
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:  Inference of XML Integrity Constraints
-
-Author: \mfauthor
-
-Department: Department of Software Engineering
-
-Supervisor: \mfadvisor{}, Department of Software Engineering
-
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
-
-Keywords: XML, ID attributes, inference
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\setcounter{page}{1}
-\tableofcontents
-
-% =============== TEXT ================
-\newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
-
-\chapwithtoc{Preface}
-
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
-
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
-
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
-
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
-
-\section{Structure of the thesis}
-
-The thesis will be structured as follows.
-
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
-
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
-
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
-
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
-
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
-
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
-
-\section{Conventions}
-
-TODO list conventions - how we write code related stuff, module names, heu names, etc
-
 \chapter{Definitions}
@@
 \end{itemize}
-
-\chapter{Research}
-
-According to the article \cite{fidax}, ...
-
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
-
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
-
-TODO other approaches to max IS
-
-\chapter{MIP Approach}
-
-TODO rigorously define MIP - optimization problem, some of the variables integral
-
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
-
-\section{Finding ID sets with GLPK}
-
-TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
-
- -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
-
-\section{Heuristics}
-
-TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
-
-TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
-
-TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
-
-TODO we will be using a pool - what is a pool
-
-\begin{figure}
-  \caption{Metaheuristic schema}
-  \label{image-metaheuristic}
-  \centering
-    \includegraphics[width=\textwidth]{images/metaheuristic}
-\end{figure}
-
-\subsection{Constructions Heuristics}
-% TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
-
-TODO construction heuristics are heus that provide us with at least some solution.
-
-\subsubsection{\jmodule{FIDAX}}
-
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
-
-The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{ch-fidax}.
-
-\begin{algorithm}
-\caption{FIDAX CH}
-\label{ch-fidax}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets C$ sorted by decreasing size
-\STATE Compute the weight $w(m)$ of each $m$ in $C$
-\FORALL{$t$ in $\Sigma^E$}
-  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
-  \STATE Remove from $C'$ all mappings of type $t$ except $m$
-\ENDFOR
-\FORALL{$m$ in $C'$}
-  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
-  \IF{$w(m) > \sum_{p \in S} w(p)$}
-    \STATE remove all $p \in S$ from $C'$
-  \ELSE
-    \STATE remove $m$ from $C'$
-  \ENDIF
-\ENDFOR
-\RETURN $C'$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random}}
-\label{heu-ch-random}
-
-One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a hungry heuristic.
-%TODO have I defined ID set condition yet??
-
-The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
-As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
-
-See the Listing \ref{ch-random} for its pseudocode.
-
-\begin{algorithm}
-\caption{Random CH}
-\label{ch-random}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \WHILE{$s$ is a feasible ID set}
-    \STATE $a \gets $ pick at random from $C \backslash S$
-    \STATE $s \gets s \cup a$
-  \ENDWHILE
-  \STATE $r \gets r \cup s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Fuzzy}}
-\label{heu-ch-fuzzy}
-
-\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in TODO.
-Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
-
-Again, this is a hungry heuristic, the Listing \ref{ch-fuzzy} contains its pseudocode.
-
-% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
-
-\begin{algorithm}
-\caption{Fuzzy CH}
-\label{ch-fuzzy}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \STATE $C' \gets C$
-
-  \WHILE{$C' $ not empty}
-    \STATE $a \gets $ pick at weighted random from $C'$
-    \IF{$s \cup a$ is a feasible ID set}
-      \STATE $s \gets s \cup a$
-      \STATE $C' \gets C' \backslash a$
-    \ENDIF
-    \FORALL{$c \in C'$}
-      \IF{$s \cup c $ is \textbf{not} a feasible ID set}
-        \STATE \COMMENT {if $c$ cannot be possibly added anymore}
-        \STATE $C' \gets C' \backslash c$
-      \ENDIF
-    \ENDFOR
-  \ENDWHILE
-
-  \STATE $r \gets r + s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Incremental}}
-
-This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
-This way it can create only one solution, and again, this is a hungry heuristic.
-
-See listing \ref{ch-incremental}.
-
-\begin{algorithm}
-\caption{Incremental CH}
-\label{ch-incremental}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by decreasing weight
-\STATE $s \gets $ empty solution
-\FORALL{$c \in C'$}
-  \IF{$s \cup c$ is a feasible ID set}
-    \STATE $s \gets s + c$
-  \ENDIF
-\ENDFOR
-\RETURN $s$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Removal}}
-
-This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
-This probably does not satisfy the ID set condition.
-Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
-Again, this is a hungry heuristic returning only one solution.
-
-See listing \ref{ch-removal}.
-
-\begin{algorithm}
-\caption{Removal CH}
-\label{ch-removal}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by increasing weight
-\STATE $s \gets C'$
-\FORALL{$c \in s$}
-  \IF{$s$ is a feasible ID set}
-    \RETURN $s$
-  \ENDIF
-  \STATE $s \gets s \backslash c$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Truncated Branch \& Bound}
-
-This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
-
-TODO if we limit GLPK's runtime, we get this
-
-\subsection{Improvement Heuristics}
-
-\nomenclature{IH}{Improvement Heuristic}
-
-TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
-
-TODO mention intensification, diversification
-
-\subsubsection{\jmodule{Identity}}
-
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{ih-identity}.
-
-\begin{algorithm}
-\caption{Identity IH}
-\label{ih-identity}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE the same pool of feasible solutions
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Remove Worst}}
-
-This trivial IH tries to improve the solution pool by removing the worst solution (i.e.
-the one with the lowest quality).
-This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
-
-See listing \ref{ih-removeworst}.
-
-\begin{algorithm}
-\caption{Remove Worst IH}
-\label{ih-removeworst}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE pool of feasible solutions
-\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
-\RETURN $FP \backslash s_{min}$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random Remove}}
-
-This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
-By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
-
-The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
-For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
-
-This heuristic returns a pool of solutions of the same size as it got on its input.
-
-See listing \ref{ih-randomremove}.
-
-\begin{algorithm}
-\caption{Random Remove IH}
-\label{ih-randomremove}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE $K \gets k * |s|$
-  \STATE remove $K$ random AMs from $s$
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Hungry}}
-
-This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
-
-\heu{Hungry} tries to improve each solution in the feasible pool in the following way.
-It orders all candidate AMs not present in the solution by decreasing weight.
-Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition.
-The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{ih-hungry} captures this process.
-
-\begin{algorithm}
-\caption{Hungry IH}
-\label{ih-hungry}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE \COMMENT {improve a single solution}
-  \STATE $C' \gets C \backslash s$
-  \STATE $C' \gets C'$ sorted by decreasing weight
-  \FORALL{$c \in C'$}
-    \IF{$s \cup c$ is a feasible ID set}
-      \STATE $s \gets s \cup c$
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Mutation}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-mutation}.
-
-\begin{algorithm}
-\caption{Mutation IH}
-\label{ih-mutation}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of AMs to fix
-\ENSURE pool of feasible solutions
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE $K \gets k * |incumbent|$
-\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Crossover}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-crossover}.
-
-\begin{algorithm}
-\caption{Crossover IH}
-\label{ih-crossover}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of solutions to look for commonalities in
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |FP|$
-\STATE $FP' \gets K$ random solutions $\in FP$
-\STATE $am \gets$ AMs found in all solutions $\in FP'$
-\STATE fix $am$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Local Branching}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-localbranching}.
-
-\begin{algorithm}
-\caption{Local Branching IH}
-\label{ih-localbranching}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |$total AM count$|$
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE add max Hamming distance requirement to GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\section{IDREF}
-
-TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
-
-\chapter{Experiments}
-
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
-
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
-
-First of all, we would like to get an idea of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
-
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
-
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
-
-\section{Experimental Data}
-
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
-
-\begin{itemize}
-	\item Realistic 
-	\item Realistic with artificial (converted) attributes
-	\item Artificial
-\end{itemize}
-
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
-
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
-
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
-Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
-
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
-
-\subsection{Realistic data}
-
-From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
-
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
-
-\begin{table}
-  \caption{List of realistic test data files}
-  \bigskip
-  \label{table-experiments-data-realistic}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
-	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data}
-  \label{image-experiments-data-realistic}
-  \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
-\end{figure}
-
-To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
-
-On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
-
-Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
-
-\subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
-
-We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
-None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
-
-TODO schemas?
-
-To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
-
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
-
-\begin{table}
-  \caption{List of realistic test data files with converted attributes}
-  \bigskip
-  \label{table-experiments-data-converted}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data with converted attributes}
-  \label{image-experiments-data-converted}
-  \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
-\end{figure}
-
-Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
-
-In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
-Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
-
-\subsection{Artificial data}
-
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
-
-This is actually simple. Consider the following excerpt from an XML file:
-
-\begin{scriptsize}
-\begin{verbatim}
-<graph>
-  <vertex0 attr="-2968876296119015800"/>
-  <vertex1 attr="1729745997570096518"/>
-  <vertex2 attr="-9020549659620928934"/>
-  ...
-  <vertex99 attr="-7545982394508643394"/>
-
-  <vertex82 attr="0"/><vertex21 attr="0"/>
-  <vertex64 attr="1"/><vertex21 attr="1"/>
-  <vertex44 attr="2"/><vertex2 attr="2"/>
-  ...
-  <vertex96 attr="99"/><vertex40 attr="99"/>
-</graph>
-\end{verbatim}
-\end{scriptsize}
-
-We want to create a graph with around $v$ vertices and $e$ edges.
-First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
-
-The pseudocode for this is in listing \ref{listing-random-data}.
-
-\begin{algorithm}
-\caption{Random XML data creation}
-\label{listing-random-data}
-\begin{algorithmic}
-\REQUIRE $v$ requested number of vertices
-\REQUIRE $e$ requested number of edges
-\ENSURE XML file content
-\PRINT \texttt{<graph>}
-\FOR{$i = 1 \to |V|$}
-	\STATE $R \gets RANDOM$
-	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
-\ENDFOR
-\FOR{$i = 1 \to |E|$}
-	\STATE $v1 \gets RANDOM(|V|)$
-	\STATE $v2 \gets RANDOM(|V|)$
-	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
-\ENDFOR
-\PRINT \texttt{</graph>}
-\RETURN
-\end{algorithmic}
-\end{algorithm}
-
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
-
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
-
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
-Note that the name is always in the form $v-e$.
-
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
-
-TODO this is on the DVD.
-
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
-
-\begin{table}
-  \caption{List of artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
-    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{List of ``sized" artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial-size}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
-  20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-  30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
-	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Artificial data}
-  \label{image-experiments-data-artificial}
-  \centering
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
-\end{figure}
-
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
-
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
-
-\section{Experimental Setup}
-
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
-
-We now have to define a few notions before moving forward to the description of our experiments.
-
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
-	\begin{itemize}
-		\item All the parameters in all the heuristics.
-		\item The specific way in which the heuristics are chained.
-		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
-		\item Initial pool size.
-		\item The termination criteria.
-		\item The input XML file.
-		\item Known optimum for this file and $\alpha$, $\beta$.
-	\end{itemize}
-\end{define}
-
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
-
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
-
-\subsection{Grammar and Model Creation}
-
-TODO link BasicIGG to see how we get IG
-
-TODO describe how we get the model
-
-\subsection{Hardware and Software}
-
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Linux Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 64-Bit Server VM (build 20.1-b02)
-GLPK version 4.38
-\end{verbatim}
-
-\subsubsection{Windows Machine}
-
-\begin{verbatim}
-Intel Core 2 Duo processor @ 2.33 GHz
-4 GB DDR2 RAM
-Windows 7 SP1 64bit
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 32-Bit Client VM (build 20.1-b02)
-GLPK version 4.45 (Cygwin)
-GLPK version 4.34 (native)
-\end{verbatim}
-
-\subsection{Methodology}
-
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
-This improvement is left for the future work.
-
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
-OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
-
-\subsection{Measuring the time}
-
-TODO
-
-\subsection{Obtaining the results}
-
-Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
-We can get all the information relevant to that experiment run from this trace alone.
-An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
-
-\section{Experimental Results}
-
-\subsection{Grammar and model generation}
-
-%         NB class GrammarModelTiming
-
-TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
-
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & TODO \\
-  Input data set    & all test data sets \\
-  Iterations        & 20 \\
-  Pool size         & not applicable \\
-  $\alpha$, $\beta$ & not applicable \\
-  \hline
-\end{tabular}
-
-The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There won't be any CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
-
-TODO graph + interpretation
-
-\subsubsection{GLPK interface}
-
-TODO how long it takes to create GLPK input and parse its output
-
-\subsection{GLPK: native vs. Cygwin}
-
-%         NB class TimeQuality
-%         NB class TimeTillOptimum
-
-In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & \texttt{100-500} \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
-There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
-\label{listing-experiment-glpk-native-vs-cygwin-1}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & sized artificial data \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
-Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
-\label{listing-experiment-glpk-native-vs-cygwin-2}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized artificial data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
-\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
-
-TODO take note of quality and time too
-
-\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
-
-TODO
-
-\subsection{Best standalone CH}
-
-TODO compare all CHs - limit GLPK to ~ 1 second
-
-\subsection{Best IH for the best CH}
-
-TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
-
-\subsubsection{\jmodule{Random} as CH}
-
-TODO with the best IH from the previous step, how about using Random as CH?
-
-\subsection{Various $\alpha$, $\beta$}
-
-TODO remind what these parameters mean
-
-\subsubsection{Effects on performance}
-
-TODO
-
-\subsubsection{Effects on the ID set}
-
-TODO
-
-\subsection{Ignoring text data}
-
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-
-\subsection{Chaining the IHs}
-
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
-
-\subsection{Algorithms for various data characteristics}
-
-TODO
-
-\section{The "Best" Algorithm}
-
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run. 
-We will find the optimum in this way. 
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
-
-\chapter{Future work}
-
-TODO
-
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
-
-\chapwithtoc{Conclusion}
-
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
-
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
-
-%%% Seznam použité literatury
-\newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-
-\listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
-
-\listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
-
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
-\printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
-
-\openright
-\addcontentsline{toc}{chapter}{Appendices}
-
-\chapter{jInfer}
-\label{appendix-jInfer}
-
-This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
-Please see TODO link jInfer project web for complete information, documentation and download options.
-
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
-The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
-The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
-Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
-This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
-
-To the best of our knowledge, at the time of writing this thesis is jInfer the only public, open source and actually working solution for XML schema inference-related tasks.
-
-At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see \ref{image-inference-process}), responsibility of 3 different services - interchangeable modules.
-
-\begin{figure}
-  \caption{Inference process in jInfer}
-  \label{image-inference-process}
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/inference-process}
-\end{figure}
-
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
-This is the representation in which will the structure live until it is used to create the final product - the schema.
-As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
-As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
-(TODO picture?) IG is used to create the AM model used in this thesis, too.
-jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
-
-After leaving the \jmodule{Initial Grammar Generator}, the IG needs to be made more general, shortened, \textit{simplified}.
-This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
-Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
-
-The last module, \jmodule{Schema Generator} takes the simplified grammar and creates the resulting schema from it.
-This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
-
-\chapter{IDSetSearch}
-\label{appendix-iss}
-
-TODO talk a bit about the structure of ISS module.
-
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
-
-TODO package name root
-
-TODO "Attribute Stats"
-
-TODO heuristics - interfaces, construction, improvement
-
-TODO experiments - quality, termination, parameters, experiment sets
-
-TODO make sure there was a place where we described the model data model:)
-
-\chapter{Experimental Trace}
-\label{appendix-trace}
-
-Following is a trace logged from a sample experiment run. It shows all the relevant information related to this instance, any and every piece of information we might be interested in.
-
-To save space, 2-column layout is used. Commentary on the particulars follows right after its end.
-
-\begin{multicols}{2}
-\begin{scriptsize}
-\begin{verbatim}
-CPU info
-  Intel(R) Core(TM)2 Quad CPU Q9550 @ 2.83GHz
-  Cores: 4
-  Clock speed: 2983 MHz
-Memory info
-  Size: 8192 MB
-OS info
-  Name: Windows 7
-  Version: 6.1
-  Architecture: amd64
-Java info
-  Version: 1.6.0_26
-  VM: Java HotSpot(TM) 64-Bit Server VM
-GLPK info
-  GLPSOL: GLPK LP/MIP Solver 4.34
-
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-
-Results:
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-Construction phase: 
-  Algorithm: Random
-    Time taken: 248 ms
-    Time since start: 248 ms
-    Pool size: 10
-    Quality: 0.19975609756097568 (11 AMs)
-Improvement phase: 
-  pass #1: 
-  Algorithm: RandomRemove, ratio = 0.2
-    Time taken: 0 ms
-    Time since start: 841 ms
-    Pool size: 10
-    Quality: 0.15878048780487808 (9 AMs)
-  pass #2: 
-  Algorithm: Mutation, ratio = 0.1, limit = 1 s
-    Time taken: 1512 ms
-    Time since start: 2710 ms
-    Pool size: 11
-    Quality: 0.21975609756097558 (11 AMs)
-
-  <... 7 more passes removed ...>
-
-  pass #10: 
-  Algorithm: Remove Worst
-    Time taken: 80 ms
-    Time since start: 7676 ms
-    Pool size: 12
-    Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
-
-Time,Quality,AMs
-248,0.19975609756097568,11
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-2927,0.1890243902439024,9
-4421,0.23463414634146343,12
-4703,0.23463414634146343,12
-4896,0.1960975609756098,10
-5793,0.23463414634146337,12
-5972,0.19951219512195123,10
-7433,0.19951219512195123,10
-7676,0.19951219512195123,10
-
-ID
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-vertex2,attr,0.01975609756097561
-vertex33,attr,0.016829268292682928
-vertex34,attr,0.02219512195121951
-vertex4,attr,0.022682926829268292
-vertex41,attr,0.014878048780487804
-vertex7,attr,0.02170731707317073
-vertex70,attr,0.018780487804878048
-vertex76,attr,0.01780487804878049
-vertex8,attr,0.02170731707317073
-vertex80,attr,0.01780487804878049
-vertex97,attr,0.016341463414634147
-
-IDREF
-Element,Attribute
-\end{verbatim}
-\end{scriptsize}
-\end{multicols}
-
-The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
-To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
-It tries to guess whether it's the Cygwin version by looking at the path to the binary.
-
-% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
-The second section states the input file along with its size and graph representation.\\
-Alpha and beta parameters for this instance belong here too.
-
-\begin{footnotesize}
-\begin{verbatim}
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-\end{verbatim}
-\end{footnotesize}
-
-Results section opens stating the most important information first: how long did the experiment run and what was the highest and final quality (these two are potentially different).
-Numbers of attribute mappings in the best and final solution respectively are stated as well.
-
-\begin{footnotesize}
-\begin{verbatim}
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Construction phase results go next.
-Among reported information are the full identification of the heuristic (possibly along with its parameters), time taken, size of the pool created and the quality of the incumbent solution (again, with the number of its AMs).
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Random
-  Time taken: 248 ms
-  Time since start: 248 ms
-  Pool size: 10
-  Quality: 0.19975609756097568 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Now for each of the improvement phases there is one section in output log.
-Information presented here has the same structure as with the construction phase.
-Please note that the \code{Pool size} is always measured \textit{after} the improvement run.
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Mutation, ratio = 0.1, limit = 1 s
-  Time taken: 1512 ms
-  Time since start: 2710 ms
-  Pool size: 11
-  Quality: 0.21975609756097558 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
-\\
-
-To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
-Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
-
-\begin{footnotesize}
-\begin{verbatim}
-Time,Quality,AMs
-...
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-...
-\end{verbatim}
-\end{footnotesize}
-
-And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
-Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
-
-\begin{footnotesize}
-\begin{verbatim}
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-...
-\end{verbatim}
-\end{footnotesize}
-
-Note that in this example trace there were no IDREF AMs found.
-
-\flushright
-$\square$
-
-\end{document}

### 2011-08-16 doc/theses/vitasek/chapter-3-mip.tex
Thesis: splitting tex files.

--- 2011-08-15 doc/theses/vitasek/diplomka.tex
@@
-\documentclass[a4paper,12pt,oneside]{report}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{url}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{tgpagella}
-\usepackage{xstring}
-\usepackage{wrapfig}
-\usepackage{appendix}
-\usepackage{nomencl}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsthm}
-
-\newcommand{\myscale}{0.74}
-\newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
-\def\.{\discretionary{}{}{}}
-\newcommand{\heu}[1]{\texttt{\textit{#1}}}
-\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
-\linespread{1.5}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
-
-\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
-\setlength{\voffset}{-1in} %analogous voffset
-\setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
-\setlength{\topmargin}{25mm}
-\setlength{\footskip}{1cm}
-\setlength{\headheight}{0cm}
-\setlength{\headsep}{0cm}
-\setlength{\marginparwidth}{0cm}
-\setlength{\marginparpush}{0cm}
-\setlength{\textheight}{23.7cm}
-\setlength{\textwidth}{14.5cm}
-\let\openright=\clearpage
-
-\def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-\begin{document}
-
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce ======================================================
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% exactly as assigned
-{\LARGE\bfseries Inference of XML Integrity Constraints}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & 	\mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage % ============================================================
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-\openright
-
-\noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
-
-\newpage % ============================================================
-%%% Strana s čestným prohlášením k diplomové práci
-\vglue 0pt plus 1fill
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature
-\hss}}
-
-\vspace{20mm}
-
-\newpage % ============================================================
-%%% Povinná informační strana diplomové práce
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
-
-Autor: \mfauthor
-
-Katedra:  Katedra softwarového inženýrství
-
-Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství
-
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
-
-Klíčová slova:  XML, ID atributy, odvozování
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:  Inference of XML Integrity Constraints
-
-Author: \mfauthor
-
-Department: Department of Software Engineering
-
-Supervisor: \mfadvisor{}, Department of Software Engineering
-
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
-
-Keywords: XML, ID attributes, inference
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\setcounter{page}{1}
-\tableofcontents
-
-% =============== TEXT ================
-\newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
-
-\chapwithtoc{Preface}
-
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
-
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
-
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
-
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
-
-\section{Structure of the thesis}
-
-The thesis will be structured as follows.
-
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
-
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
-
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
-
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
-
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
-
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
-
-\section{Conventions}
-
-TODO list conventions - how we write code related stuff, module names, heu names, etc
-
-\chapter{Definitions}
-
-\begin{itemize}
-	\item XML tree - from FIDAX or wherever
-	\item Element, Attribute
-	\item ID, IDREF, IDREFS attribute according to specification
-	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
-	\item ID set
-  \item Weight: support, coverage
-\end{itemize}
-
-\chapter{Research}
-
-According to the article \cite{fidax}, ...
-
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
-
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
-
-TODO other approaches to max IS
-
 \chapter{MIP Approach}
@@
 \begin{algorithm}
-\caption{FIDAX CH}
+\caption{\jmodule{FIDAX} CH}
 \label{listing-ch-fidax}
@@
 \begin{algorithm}
-\caption{Random CH}
+\caption{\jmodule{Random} CH}
 \label{listing-ch-random}
@@
 \begin{algorithm}
-\caption{Fuzzy CH}
+\caption{\jmodule{Fuzzy} CH}
 \label{listing-ch-fuzzy}
@@
 \begin{algorithm}
-\caption{Incremental CH}
+\caption{\jmodule{Incremental} CH}
 \label{listing-ch-incremental}
@@
 \begin{algorithm}
-\caption{Removal CH}
+\caption{\jmodule{Removal} CH}
 \label{listing-ch-removal}
@@
 
+TODO we shuffle AMs - we get different runs - pool is possible
+
 \subsection{Improvement Heuristics}
@@
 \begin{algorithm}
-\caption{Identity IH}
+\caption{\jmodule{Identity} IH}
 \label{listing-ih-identity}
@@
 \begin{algorithm}
-\caption{Remove Worst IH}
+\caption{\jmodule{Remove Worst} IH}
 \label{listing-ih-removeworst}
@@
 \begin{algorithm}
-\caption{Random Remove IH}
+\caption{\jmodule{Random Remove} IH}
 \label{listing-ih-randomremove}
@@
 \begin{algorithm}
-\caption{Hungry IH}
+\caption{\jmodule{Hungry} IH}
 \label{listing-ih-hungry}
@@
 \begin{algorithm}
-\caption{Mutation IH}
+\caption{\jmodule{Mutation} IH}
 \label{listing-ih-mutation}
@@
 \begin{algorithm}
-\caption{Crossover IH}
+\caption{\jmodule{Crossover} IH}
 \label{listing-ih-crossover}
@@
 \begin{algorithm}
-\caption{Local Branching IH}
+\caption{\jmodule{Local Branching} IH}
 \label{listing-ih-localbranching}
@@
 TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
-
-\chapter{Experiments}
-
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
-
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
-
-First of all, we would like to get an idea of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
-
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
-
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
-
-\section{Experimental Data}
-
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
-
-\begin{itemize}
-	\item Realistic 
-	\item Realistic with artificial (converted) attributes
-	\item Artificial
-\end{itemize}
-
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
-
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
-
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
-Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
-
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
-
-\subsection{Realistic data}
-
-From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
-
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
-
-\begin{table}
-  \caption{List of realistic test data files}
-  \bigskip
-  \label{table-experiments-data-realistic}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
-	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data}
-  \label{image-experiments-data-realistic}
-  \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
-\end{figure}
-
-To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
-
-On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
-
-Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
-
-\subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
-
-We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
-None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
-
-TODO schemas?
-
-To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
-
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
-
-\begin{table}
-  \caption{List of realistic test data files with converted attributes}
-  \bigskip
-  \label{table-experiments-data-converted}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data with converted attributes}
-  \label{image-experiments-data-converted}
-  \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
-\end{figure}
-
-Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
-
-In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
-Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
-
-\subsection{Artificial data}
-
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
-
-This is actually simple. Consider the following excerpt from an XML file:
-
-\begin{scriptsize}
-\begin{verbatim}
-<graph>
-  <vertex0 attr="-2968876296119015800"/>
-  <vertex1 attr="1729745997570096518"/>
-  <vertex2 attr="-9020549659620928934"/>
-  ...
-  <vertex99 attr="-7545982394508643394"/>
-
-  <vertex82 attr="0"/><vertex21 attr="0"/>
-  <vertex64 attr="1"/><vertex21 attr="1"/>
-  <vertex44 attr="2"/><vertex2 attr="2"/>
-  ...
-  <vertex96 attr="99"/><vertex40 attr="99"/>
-</graph>
-\end{verbatim}
-\end{scriptsize}
-
-We want to create a graph with around $v$ vertices and $e$ edges.
-First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
-
-The pseudocode for this is in listing \ref{listing-random-data}.
-
-\begin{algorithm}
-\caption{Random XML data creation}
-\label{listing-random-data}
-\begin{algorithmic}
-\REQUIRE $v$ requested number of vertices
-\REQUIRE $e$ requested number of edges
-\ENSURE XML file content
-\PRINT \texttt{<graph>}
-\FOR{$i = 1 \to |V|$}
-	\STATE $R \gets RANDOM$
-	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
-\ENDFOR
-\FOR{$i = 1 \to |E|$}
-	\STATE $v1 \gets RANDOM(|V|)$
-	\STATE $v2 \gets RANDOM(|V|)$
-	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
-\ENDFOR
-\PRINT \texttt{</graph>}
-\RETURN
-\end{algorithmic}
-\end{algorithm}
-
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
-
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
-
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
-Note that the name is always in the form $v-e$.
-
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
-
-TODO this is on the DVD.
-
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
-
-\begin{table}
-  \caption{List of artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
-    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{List of ``sized" artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial-size}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
-  20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-  30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
-	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Artificial data}
-  \label{image-experiments-data-artificial}
-  \centering
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
-\end{figure}
-
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
-
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
-
-\section{Experimental Setup}
-
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
-
-We now have to define a few notions before moving forward to the description of our experiments.
-
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
-	\begin{itemize}
-		\item All the parameters in all the heuristics.
-		\item The specific way in which the heuristics are chained.
-		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
-		\item Initial pool size.
-		\item The termination criteria.
-		\item The input XML file.
-		\item Known optimum for this file and $\alpha$, $\beta$.
-	\end{itemize}
-\end{define}
-
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
-
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
-
-\subsection{Grammar and Model Creation}
-
-TODO link BasicIGG to see how we get IG
-
-TODO describe how we get the model
-
-\subsection{Hardware and Software}
-
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Linux Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 64-Bit Server VM (build 20.1-b02)
-GLPK version 4.38
-\end{verbatim}
-
-\subsubsection{Windows Machine}
-
-\begin{verbatim}
-Intel Core 2 Duo processor @ 2.33 GHz
-4 GB DDR2 RAM
-Windows 7 SP1 64bit
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 32-Bit Client VM (build 20.1-b02)
-GLPK version 4.45 (Cygwin)
-GLPK version 4.34 (native)
-\end{verbatim}
-
-\subsection{Methodology}
-
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
-This improvement is left for the future work.
-
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
-OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
-
-\subsection{Measuring the time}
-
-TODO
-
-\subsection{Obtaining the results}
-
-Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
-We can get all the information relevant to that experiment run from this trace alone.
-An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
-
-\section{Experimental Results}
-
-\subsection{Grammar and model generation}
-
-%         NB class GrammarModelTiming
-
-TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
-
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & TODO \\
-  Input data set    & all test data sets \\
-  Iterations        & 20 \\
-  Pool size         & not applicable \\
-  $\alpha$, $\beta$ & not applicable \\
-  \hline
-\end{tabular}
-
-The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There won't be any CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
-
-TODO graph + interpretation
-
-\subsubsection{GLPK interface}
-
-TODO how long it takes to create GLPK input and parse its output
-
-\subsection{GLPK: native vs. Cygwin}
-
-%         NB class TimeQuality
-%         NB class TimeTillOptimum
-
-In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & \texttt{100-500} \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
-There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
-\label{listing-experiment-glpk-native-vs-cygwin-1}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & sized artificial data \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
-Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
-\label{listing-experiment-glpk-native-vs-cygwin-2}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized artificial data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
-\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
-
-TODO take note of quality and time too
-
-\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
-
-TODO
-
-\subsection{Best standalone CH}
-
-TODO compare all CHs - limit GLPK to ~ 1 second
-
-\subsection{Best IH for the best CH}
-
-TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
-
-\subsubsection{\jmodule{Random} as CH}
-
-TODO with the best IH from the previous step, how about using Random as CH?
-
-\subsection{Various $\alpha$, $\beta$}
-
-TODO remind what these parameters mean
-
-\subsubsection{Effects on performance}
-
-TODO
-
-\subsubsection{Effects on the ID set}
-
-TODO
-
-\subsection{Ignoring text data}
-
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-
-\subsection{Chaining the IHs}
-
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
-
-\subsection{Algorithms for various data characteristics}
-
-TODO
-
-\section{The "Best" Algorithm}
-
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run. 
-We will find the optimum in this way. 
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
-
-\chapter{Future work}
-
-TODO
-
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
-
-\chapwithtoc{Conclusion}
-
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
-
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
-
-%%% Seznam použité literatury
-\newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-
-\listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
-
-\listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
-
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
-\printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
-
-\openright
-\addcontentsline{toc}{chapter}{Appendices}
-
-\chapter{jInfer}
-\label{appendix-jInfer}
-
-This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
-Please see TODO link jInfer project web for complete information, documentation and download options.
-
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
-The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
-The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
-Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
-This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
-
-To the best of our knowledge, at the time of writing this thesis is jInfer the only public, open source and actually working solution for XML schema inference-related tasks.
-
-At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see \ref{image-inference-process}), responsibility of 3 different services - interchangeable modules.
-
-\begin{figure}
-  \caption{Inference process in jInfer}
-  \label{image-inference-process}
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/inference-process}
-\end{figure}
-
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
-This is the representation in which will the structure live until it is used to create the final product - the schema.
-As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
-As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
-(TODO picture?) IG is used to create the AM model used in this thesis, too.
-jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
-
-After leaving the \jmodule{Initial Grammar Generator}, the IG needs to be made more general, shortened, \textit{simplified}.
-This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
-Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
-
-The last module, \jmodule{Schema Generator} takes the simplified grammar and creates the resulting schema from it.
-This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
-
-\chapter{IDSetSearch}
-\label{appendix-iss}
-
-TODO talk a bit about the structure of ISS module.
-
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
-
-TODO package name root
-
-TODO "Attribute Stats"
-
-TODO heuristics - interfaces, construction, improvement
-
-TODO experiments - quality, termination, parameters, experiment sets
-
-TODO make sure there was a place where we described the model data model:)
-
-\chapter{Experimental Trace}
-\label{appendix-trace}
-
-Following is a trace logged from a sample experiment run. It shows all the relevant information related to this instance, any and every piece of information we might be interested in.
-
-To save space, 2-column layout is used. Commentary on the particulars follows right after its end.
-
-\begin{multicols}{2}
-\begin{scriptsize}
-\begin{verbatim}
-CPU info
-  Intel(R) Core(TM)2 Quad CPU Q9550 @ 2.83GHz
-  Cores: 4
-  Clock speed: 2983 MHz
-Memory info
-  Size: 8192 MB
-OS info
-  Name: Windows 7
-  Version: 6.1
-  Architecture: amd64
-Java info
-  Version: 1.6.0_26
-  VM: Java HotSpot(TM) 64-Bit Server VM
-GLPK info
-  GLPSOL: GLPK LP/MIP Solver 4.34
-
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-
-Results:
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-Construction phase: 
-  Algorithm: Random
-    Time taken: 248 ms
-    Time since start: 248 ms
-    Pool size: 10
-    Quality: 0.19975609756097568 (11 AMs)
-Improvement phase: 
-  pass #1: 
-  Algorithm: RandomRemove, ratio = 0.2
-    Time taken: 0 ms
-    Time since start: 841 ms
-    Pool size: 10
-    Quality: 0.15878048780487808 (9 AMs)
-  pass #2: 
-  Algorithm: Mutation, ratio = 0.1, limit = 1 s
-    Time taken: 1512 ms
-    Time since start: 2710 ms
-    Pool size: 11
-    Quality: 0.21975609756097558 (11 AMs)
-
-  <... 7 more passes removed ...>
-
-  pass #10: 
-  Algorithm: Remove Worst
-    Time taken: 80 ms
-    Time since start: 7676 ms
-    Pool size: 12
-    Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
-
-Time,Quality,AMs
-248,0.19975609756097568,11
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-2927,0.1890243902439024,9
-4421,0.23463414634146343,12
-4703,0.23463414634146343,12
-4896,0.1960975609756098,10
-5793,0.23463414634146337,12
-5972,0.19951219512195123,10
-7433,0.19951219512195123,10
-7676,0.19951219512195123,10
-
-ID
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-vertex2,attr,0.01975609756097561
-vertex33,attr,0.016829268292682928
-vertex34,attr,0.02219512195121951
-vertex4,attr,0.022682926829268292
-vertex41,attr,0.014878048780487804
-vertex7,attr,0.02170731707317073
-vertex70,attr,0.018780487804878048
-vertex76,attr,0.01780487804878049
-vertex8,attr,0.02170731707317073
-vertex80,attr,0.01780487804878049
-vertex97,attr,0.016341463414634147
-
-IDREF
-Element,Attribute
-\end{verbatim}
-\end{scriptsize}
-\end{multicols}
-
-The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
-To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
-It tries to guess whether it's the Cygwin version by looking at the path to the binary.
-
-% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
-The second section states the input file along with its size and graph representation.\\
-Alpha and beta parameters for this instance belong here too.
-
-\begin{footnotesize}
-\begin{verbatim}
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-\end{verbatim}
-\end{footnotesize}
-
-Results section opens stating the most important information first: how long did the experiment run and what was the highest and final quality (these two are potentially different).
-Numbers of attribute mappings in the best and final solution respectively are stated as well.
-
-\begin{footnotesize}
-\begin{verbatim}
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Construction phase results go next.
-Among reported information are the full identification of the heuristic (possibly along with its parameters), time taken, size of the pool created and the quality of the incumbent solution (again, with the number of its AMs).
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Random
-  Time taken: 248 ms
-  Time since start: 248 ms
-  Pool size: 10
-  Quality: 0.19975609756097568 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Now for each of the improvement phases there is one section in output log.
-Information presented here has the same structure as with the construction phase.
-Please note that the \code{Pool size} is always measured \textit{after} the improvement run.
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Mutation, ratio = 0.1, limit = 1 s
-  Time taken: 1512 ms
-  Time since start: 2710 ms
-  Pool size: 11
-  Quality: 0.21975609756097558 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
-\\
-
-To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
-Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
-
-\begin{footnotesize}
-\begin{verbatim}
-Time,Quality,AMs
-...
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-...
-\end{verbatim}
-\end{footnotesize}
-
-And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
-Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
-
-\begin{footnotesize}
-\begin{verbatim}
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-...
-\end{verbatim}
-\end{footnotesize}
-
-Note that in this example trace there were no IDREF AMs found.
-
-\flushright
-$\square$
-
-\end{document}

### 2011-08-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: splitting tex files.

--- 2011-08-16 doc/theses/vitasek/diplomka.tex
@@
-\documentclass[a4paper,12pt,oneside]{report}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{url}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{tgpagella}
-\usepackage{xstring}
-\usepackage{wrapfig}
-\usepackage{appendix}
-\usepackage{nomencl}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsthm}
-
-\newcommand{\myscale}{0.74}
-\newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
-\def\.{\discretionary{}{}{}}
-\newcommand{\heu}[1]{\texttt{\textit{#1}}}
-\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
-\linespread{1.5}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
-
-\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
-\setlength{\voffset}{-1in} %analogous voffset
-\setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
-\setlength{\topmargin}{25mm}
-\setlength{\footskip}{1cm}
-\setlength{\headheight}{0cm}
-\setlength{\headsep}{0cm}
-\setlength{\marginparwidth}{0cm}
-\setlength{\marginparpush}{0cm}
-\setlength{\textheight}{23.7cm}
-\setlength{\textwidth}{14.5cm}
-\let\openright=\clearpage
-
-\def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-\begin{document}
-
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce ======================================================
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% exactly as assigned
-{\LARGE\bfseries Inference of XML Integrity Constraints}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & 	\mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage % ============================================================
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-\openright
-
-\noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
-
-\newpage % ============================================================
-%%% Strana s čestným prohlášením k diplomové práci
-\vglue 0pt plus 1fill
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature
-\hss}}
-
-\vspace{20mm}
-
-\newpage % ============================================================
-%%% Povinná informační strana diplomové práce
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
-
-Autor: \mfauthor
-
-Katedra:  Katedra softwarového inženýrství
-
-Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství
-
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
-
-Klíčová slova:  XML, ID atributy, odvozování
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:  Inference of XML Integrity Constraints
-
-Author: \mfauthor
-
-Department: Department of Software Engineering
-
-Supervisor: \mfadvisor{}, Department of Software Engineering
-
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
-
-Keywords: XML, ID attributes, inference
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\setcounter{page}{1}
-\tableofcontents
-
-% =============== TEXT ================
-\newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
-
-\chapwithtoc{Preface}
-
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
-
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
-
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
-
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
-
-\section{Structure of the thesis}
-
-The thesis will be structured as follows.
-
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
-
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
-
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
-
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
-
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
-
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
-
-\section{Conventions}
-
-TODO list conventions - how we write code related stuff, module names, heu names, etc
-
-\chapter{Definitions}
-
-\begin{itemize}
-	\item XML tree - from FIDAX or wherever
-	\item Element, Attribute
-	\item ID, IDREF, IDREFS attribute according to specification
-	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
-	\item ID set
-  \item Weight: support, coverage
-\end{itemize}
-
-\chapter{Research}
-
-According to the article \cite{fidax}, ...
-
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
-
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
-
-TODO other approaches to max IS
-
-\chapter{MIP Approach}
-
-TODO rigorously define MIP - optimization problem, some of the variables integral
-
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
-
-\section{Finding ID sets with GLPK}
-
-TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
-
- -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
-
-\section{Heuristics}
-
-TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
-
-TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
-
-TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
-
-TODO we will be using a pool - what is a pool
-
-\begin{figure}
-  \caption{Metaheuristic schema}
-  \label{image-metaheuristic}
-  \centering
-    \includegraphics[width=\textwidth]{images/metaheuristic}
-\end{figure}
-
-\subsection{Constructions Heuristics}
-% TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
-
-TODO construction heuristics are heus that provide us with at least some solution.
-
-\subsubsection{\jmodule{FIDAX}}
-
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
-
-The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{listing-ch-fidax}.
-
-\begin{algorithm}
-\caption{FIDAX CH}
-\label{listing-ch-fidax}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets C$ sorted by decreasing size
-\STATE Compute the weight $w(m)$ of each $m$ in $C$
-\FORALL{$t$ in $\Sigma^E$}
-  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
-  \STATE Remove from $C'$ all mappings of type $t$ except $m$
-\ENDFOR
-\FORALL{$m$ in $C'$}
-  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
-  \IF{$w(m) > \sum_{p \in S} w(p)$}
-    \STATE remove all $p \in S$ from $C'$
-  \ELSE
-    \STATE remove $m$ from $C'$
-  \ENDIF
-\ENDFOR
-\RETURN $C'$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random}}
-\label{heu-ch-random}
-
-One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a hungry heuristic.
-%TODO have I defined ID set condition yet??
-
-The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
-As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
-
-See the Listing \ref{listing-ch-random} for its pseudocode.
-
-\begin{algorithm}
-\caption{Random CH}
-\label{listing-ch-random}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \WHILE{$s$ is a feasible ID set}
-    \STATE $a \gets $ pick at random from $C \backslash S$
-    \STATE $s \gets s \cup a$
-  \ENDWHILE
-  \STATE $r \gets r \cup s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Fuzzy}}
-\label{heu-ch-fuzzy}
-
-\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in TODO.
-Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
-
-Again, this is a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
-
-% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
-
-\begin{algorithm}
-\caption{Fuzzy CH}
-\label{listing-ch-fuzzy}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \STATE $C' \gets C$
-
-  \WHILE{$C' $ not empty}
-    \STATE $a \gets $ pick at weighted random from $C'$
-    \IF{$s \cup a$ is a feasible ID set}
-      \STATE $s \gets s \cup a$
-      \STATE $C' \gets C' \backslash a$
-    \ENDIF
-    \FORALL{$c \in C'$}
-      \IF{$s \cup c $ is \textbf{not} a feasible ID set}
-        \STATE \COMMENT {if $c$ cannot be possibly added anymore}
-        \STATE $C' \gets C' \backslash c$
-      \ENDIF
-    \ENDFOR
-  \ENDWHILE
-
-  \STATE $r \gets r + s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Incremental}}
-
-This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
-This way it can create only one solution, and again, this is a hungry heuristic.
-
-See listing \ref{listing-ch-incremental}.
-
-\begin{algorithm}
-\caption{Incremental CH}
-\label{listing-ch-incremental}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by decreasing weight
-\STATE $s \gets $ empty solution
-\FORALL{$c \in C'$}
-  \IF{$s \cup c$ is a feasible ID set}
-    \STATE $s \gets s + c$
-  \ENDIF
-\ENDFOR
-\RETURN $s$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Removal}}
-
-This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
-This probably does not satisfy the ID set condition.
-Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
-Again, this is a hungry heuristic returning only one solution.
-
-See listing \ref{listing-ch-removal}.
-
-\begin{algorithm}
-\caption{Removal CH}
-\label{listing-ch-removal}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by increasing weight
-\STATE $s \gets C'$
-\FORALL{$c \in s$}
-  \IF{$s$ is a feasible ID set}
-    \RETURN $s$
-  \ENDIF
-  \STATE $s \gets s \backslash c$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Truncated Branch \& Bound}
-
-This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
-
-TODO if we limit GLPK's runtime, we get this
-
-\subsection{Improvement Heuristics}
-
-\nomenclature{IH}{Improvement Heuristic}
-
-TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
-
-TODO mention intensification, diversification
-
-\subsubsection{\jmodule{Identity}}
-
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{listing-ih-identity}.
-
-\begin{algorithm}
-\caption{Identity IH}
-\label{listing-ih-identity}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE the same pool of feasible solutions
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Remove Worst}}
-
-This trivial IH tries to improve the solution pool by removing the worst solution (i.e.
-the one with the lowest quality).
-This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
-
-See listing \ref{listing-ih-removeworst}.
-
-\begin{algorithm}
-\caption{Remove Worst IH}
-\label{listing-ih-removeworst}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE pool of feasible solutions
-\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
-\RETURN $FP \backslash s_{min}$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random Remove}}
-
-This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
-By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
-
-The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
-For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
-
-This heuristic returns a pool of solutions of the same size as it got on its input.
-
-See listing \ref{listing-ih-randomremove}.
-
-\begin{algorithm}
-\caption{Random Remove IH}
-\label{listing-ih-randomremove}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE $K \gets k * |s|$
-  \STATE remove $K$ random AMs from $s$
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Hungry}}
-
-This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
-
-\heu{Hungry} tries to improve each solution in the feasible pool in the following way.
-It orders all candidate AMs not present in the solution by decreasing weight.
-Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition.
-The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{listing-ih-hungry} captures this process.
-
-\begin{algorithm}
-\caption{Hungry IH}
-\label{listing-ih-hungry}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE \COMMENT {improve a single solution}
-  \STATE $C' \gets C \backslash s$
-  \STATE $C' \gets C'$ sorted by decreasing weight
-  \FORALL{$c \in C'$}
-    \IF{$s \cup c$ is a feasible ID set}
-      \STATE $s \gets s \cup c$
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Mutation}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{listing-ih-mutation}.
-
-\begin{algorithm}
-\caption{Mutation IH}
-\label{listing-ih-mutation}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of AMs to fix
-\ENSURE pool of feasible solutions
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE $K \gets k * |incumbent|$
-\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Crossover}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{listing-ih-crossover}.
-
-\begin{algorithm}
-\caption{Crossover IH}
-\label{listing-ih-crossover}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of solutions to look for commonalities in
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |FP|$
-\STATE $FP' \gets K$ random solutions $\in FP$
-\STATE $am \gets$ AMs found in all solutions $\in FP'$
-\STATE fix $am$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Local Branching}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{listing-ih-localbranching}.
-
-\begin{algorithm}
-\caption{Local Branching IH}
-\label{listing-ih-localbranching}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |$total AM count$|$
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE add max Hamming distance requirement to GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\section{IDREF}
-
-TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
-
 \chapter{Experiments}
@@
 
-TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
-
 The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
@@
 The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There won't be any CHs or IHs.
+There will be no CHs or IHs.
 We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
@@
 
-\subsubsection{GLPK interface}
+\subsubsection{GLPK interface timing}
 
-TODO how long it takes to create GLPK input and parse its output
+A related problem is how long it takes to create input for GLPK and then parse its results.
+We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.
+
+TODO graph + interpretation
 
@@
 Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
+This will be repeated in 20 iterations for each of the 11 files from the sized test data set (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
 There will again be no IH, the only data we will collect are the times of the CH run in each case.
@@
 
-TODO take note of quality and time too
+The experimental set will contain 660 configurations in total: 3 different CHs * 11 official test data sets * 20 iterations.
+There will be no improvement heuristics.
+The pool size will be set to 10, even though \jmodule{FIDAX} cannot not profit from this.
+Listing for this can be found in \ref{listing-experiment-random-fuzzy-fidax}.
+
+We will be gathering the running time of the CH itself and incumbent qualities (best in the pool) for GnuPlot.
+
+\begin{algorithm}
+\caption{Random vs. Fuzzy vs. FIDAX set generation}
+\label{listing-experiment-random-fuzzy-fidax}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$i = 1 \to 20$}
+	\FOR{$file \in $ official test data}
+    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Random}(pool = 10), IH = \emptyset\} \cup \{file, CH = \jmodule{Fuzzy}(pool = 10), IH = \emptyset\} \cup \{file, CH = \jmodule{FIDAX}, IH = \emptyset\}$
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+TODO graphs (files with similar behavior together) + interpretation.
 
@@
 
-TODO
+There is a minor question to be answered easily: is \jmodule{FIDAX} ``hungry enough"? Could we improve it using \jmodule{Hungry} as IH? Let us design a short experiment.
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & TODO \\
+  Input data set    & all official test data sets \\
+  Iterations        & 1 \\
+  Pool size         & 1 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+We won't be needing pool bigger than one, nor more iterations - both \jmodule{FIDAX} and \jmodule{Hungry} are deterministic.
+We will try all official data sets, first with empty IH, second with \jmodule{Hungry} as IH.
+We will gather the qualities in each case and see whether there is any improvement.
+
+TODO table + interpretation
 
@@
 Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
-
-\chapter{Future work}
-
-TODO
-
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
-
-\chapwithtoc{Conclusion}
-
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
-
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
-
-%%% Seznam použité literatury
-\newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-
-\listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
-
-\listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
-
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
-\printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
-
-\openright
-\addcontentsline{toc}{chapter}{Appendices}
-
-\chapter{jInfer}
-\label{appendix-jInfer}
-
-This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
-Please see TODO link jInfer project web for complete information, documentation and download options.
-
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
-The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
-The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
-Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
-This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
-
-To the best of our knowledge, at the time of writing this thesis is jInfer the only public, open source and actually working solution for XML schema inference-related tasks.
-
-At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see \ref{image-inference-process}), responsibility of 3 different services - interchangeable modules.
-
-\begin{figure}
-  \caption{Inference process in jInfer}
-  \label{image-inference-process}
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/inference-process}
-\end{figure}
-
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
-This is the representation in which will the structure live until it is used to create the final product - the schema.
-As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
-As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
-(TODO picture?) IG is used to create the AM model used in this thesis, too.
-jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
-
-After leaving the \jmodule{Initial Grammar Generator}, the IG needs to be made more general, shortened, \textit{simplified}.
-This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
-Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
-
-The last module, \jmodule{Schema Generator} takes the simplified grammar and creates the resulting schema from it.
-This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
-
-\chapter{IDSetSearch}
-\label{appendix-iss}
-
-TODO talk a bit about the structure of ISS module.
-
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
-
-TODO package name root
-
-TODO "Attribute Stats"
-
-TODO heuristics - interfaces, construction, improvement
-
-TODO experiments - quality, termination, parameters, experiment sets
-
-TODO make sure there was a place where we described the model data model:)
-
-\chapter{Experimental Trace}
-\label{appendix-trace}
-
-Following is a trace logged from a sample experiment run. It shows all the relevant information related to this instance, any and every piece of information we might be interested in.
-
-To save space, 2-column layout is used. Commentary on the particulars follows right after its end.
-
-\begin{multicols}{2}
-\begin{scriptsize}
-\begin{verbatim}
-CPU info
-  Intel(R) Core(TM)2 Quad CPU Q9550 @ 2.83GHz
-  Cores: 4
-  Clock speed: 2983 MHz
-Memory info
-  Size: 8192 MB
-OS info
-  Name: Windows 7
-  Version: 6.1
-  Architecture: amd64
-Java info
-  Version: 1.6.0_26
-  VM: Java HotSpot(TM) 64-Bit Server VM
-GLPK info
-  GLPSOL: GLPK LP/MIP Solver 4.34
-
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-
-Results:
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-Construction phase: 
-  Algorithm: Random
-    Time taken: 248 ms
-    Time since start: 248 ms
-    Pool size: 10
-    Quality: 0.19975609756097568 (11 AMs)
-Improvement phase: 
-  pass #1: 
-  Algorithm: RandomRemove, ratio = 0.2
-    Time taken: 0 ms
-    Time since start: 841 ms
-    Pool size: 10
-    Quality: 0.15878048780487808 (9 AMs)
-  pass #2: 
-  Algorithm: Mutation, ratio = 0.1, limit = 1 s
-    Time taken: 1512 ms
-    Time since start: 2710 ms
-    Pool size: 11
-    Quality: 0.21975609756097558 (11 AMs)
-
-  <... 7 more passes removed ...>
-
-  pass #10: 
-  Algorithm: Remove Worst
-    Time taken: 80 ms
-    Time since start: 7676 ms
-    Pool size: 12
-    Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
-
-Time,Quality,AMs
-248,0.19975609756097568,11
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-2927,0.1890243902439024,9
-4421,0.23463414634146343,12
-4703,0.23463414634146343,12
-4896,0.1960975609756098,10
-5793,0.23463414634146337,12
-5972,0.19951219512195123,10
-7433,0.19951219512195123,10
-7676,0.19951219512195123,10
-
-ID
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-vertex2,attr,0.01975609756097561
-vertex33,attr,0.016829268292682928
-vertex34,attr,0.02219512195121951
-vertex4,attr,0.022682926829268292
-vertex41,attr,0.014878048780487804
-vertex7,attr,0.02170731707317073
-vertex70,attr,0.018780487804878048
-vertex76,attr,0.01780487804878049
-vertex8,attr,0.02170731707317073
-vertex80,attr,0.01780487804878049
-vertex97,attr,0.016341463414634147
-
-IDREF
-Element,Attribute
-\end{verbatim}
-\end{scriptsize}
-\end{multicols}
-
-The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
-To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
-It tries to guess whether it's the Cygwin version by looking at the path to the binary.
-
-% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
-The second section states the input file along with its size and graph representation.\\
-Alpha and beta parameters for this instance belong here too.
-
-\begin{footnotesize}
-\begin{verbatim}
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-\end{verbatim}
-\end{footnotesize}
-
-Results section opens stating the most important information first: how long did the experiment run and what was the highest and final quality (these two are potentially different).
-Numbers of attribute mappings in the best and final solution respectively are stated as well.
-
-\begin{footnotesize}
-\begin{verbatim}
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Construction phase results go next.
-Among reported information are the full identification of the heuristic (possibly along with its parameters), time taken, size of the pool created and the quality of the incumbent solution (again, with the number of its AMs).
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Random
-  Time taken: 248 ms
-  Time since start: 248 ms
-  Pool size: 10
-  Quality: 0.19975609756097568 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Now for each of the improvement phases there is one section in output log.
-Information presented here has the same structure as with the construction phase.
-Please note that the \code{Pool size} is always measured \textit{after} the improvement run.
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Mutation, ratio = 0.1, limit = 1 s
-  Time taken: 1512 ms
-  Time since start: 2710 ms
-  Pool size: 11
-  Quality: 0.21975609756097558 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
-\\
-
-To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
-Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
-
-\begin{footnotesize}
-\begin{verbatim}
-Time,Quality,AMs
-...
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-...
-\end{verbatim}
-\end{footnotesize}
-
-And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
-Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
-
-\begin{footnotesize}
-\begin{verbatim}
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-...
-\end{verbatim}
-\end{footnotesize}
-
-Note that in this example trace there were no IDREF AMs found.
-
-\flushright
-$\square$
-
-\end{document}

### 2011-08-16 doc/theses/vitasek/diplomka.tex
Thesis: splitting tex files.

--- 2011-08-15 doc/theses/vitasek/diplomka.tex
@@
 
-\chapter{Definitions}
+\include{chapter-1-definitions}
 
-\begin{itemize}
-	\item XML tree - from FIDAX or wherever
-	\item Element, Attribute
-	\item ID, IDREF, IDREFS attribute according to specification
-	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
-	\item ID set
-  \item Weight: support, coverage
-\end{itemize}
+\include{chapter-2-research}
 
-\chapter{Research}
+\include{chapter-3-mip}
 
-According to the article \cite{fidax}, ...
+\include{chapter-4-experiments}
 
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
+\include{chapter-5-future}
 
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
-
-TODO other approaches to max IS
-
-\chapter{MIP Approach}
-
-TODO rigorously define MIP - optimization problem, some of the variables integral
-
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
-
-\section{Finding ID sets with GLPK}
-
-TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
-
- -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
-
-\section{Heuristics}
-
-TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
-
-TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
-
-TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
-
-TODO we will be using a pool - what is a pool
-
-\begin{figure}
-  \caption{Metaheuristic schema}
-  \label{image-metaheuristic}
-  \centering
-    \includegraphics[width=\textwidth]{images/metaheuristic}
-\end{figure}
-
-\subsection{Constructions Heuristics}
-% TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
-
-TODO construction heuristics are heus that provide us with at least some solution.
-
-\subsubsection{\jmodule{FIDAX}}
-
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
-
-The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{ch-fidax}.
-
-\begin{algorithm}
-\caption{FIDAX CH}
-\label{ch-fidax}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets C$ sorted by decreasing size
-\STATE Compute the weight $w(m)$ of each $m$ in $C$
-\FORALL{$t$ in $\Sigma^E$}
-  \STATE Let $m$ be a \textbf{highest-weight} mapping of type $t$ in $C'$
-  \STATE Remove from $C'$ all mappings of type $t$ except $m$
-\ENDFOR
-\FORALL{$m$ in $C'$}
-  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
-  \IF{$w(m) > \sum_{p \in S} w(p)$}
-    \STATE remove all $p \in S$ from $C'$
-  \ELSE
-    \STATE remove $m$ from $C'$
-  \ENDIF
-\ENDFOR
-\RETURN $C'$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random}}
-\label{heu-ch-random}
-
-One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a hungry heuristic.
-%TODO have I defined ID set condition yet??
-
-The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
-As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
-
-See the Listing \ref{ch-random} for its pseudocode.
-
-\begin{algorithm}
-\caption{Random CH}
-\label{ch-random}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \WHILE{$s$ is a feasible ID set}
-    \STATE $a \gets $ pick at random from $C \backslash S$
-    \STATE $s \gets s \cup a$
-  \ENDWHILE
-  \STATE $r \gets r \cup s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Fuzzy}}
-\label{heu-ch-fuzzy}
-
-\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in TODO.
-Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
-
-Again, this is a hungry heuristic, the Listing \ref{ch-fuzzy} contains its pseudocode.
-
-% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
-
-\begin{algorithm}
-\caption{Fuzzy CH}
-\label{ch-fuzzy}
-\begin{algorithmic}
-\REQUIRE $N$ required size of pool
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of $N$ feasible solutions
-\STATE $r \gets $ empty pool
-\FOR{$i = 1 \to N$} 
-  \STATE \COMMENT{create 1 solution}
-  \STATE $s \gets $ empty solution
-  \STATE $C' \gets C$
-
-  \WHILE{$C' $ not empty}
-    \STATE $a \gets $ pick at weighted random from $C'$
-    \IF{$s \cup a$ is a feasible ID set}
-      \STATE $s \gets s \cup a$
-      \STATE $C' \gets C' \backslash a$
-    \ENDIF
-    \FORALL{$c \in C'$}
-      \IF{$s \cup c $ is \textbf{not} a feasible ID set}
-        \STATE \COMMENT {if $c$ cannot be possibly added anymore}
-        \STATE $C' \gets C' \backslash c$
-      \ENDIF
-    \ENDFOR
-  \ENDWHILE
-
-  \STATE $r \gets r + s$
-\ENDFOR
-\RETURN $r$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Incremental}}
-
-This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
-This way it can create only one solution, and again, this is a hungry heuristic.
-
-See listing \ref{ch-incremental}.
-
-\begin{algorithm}
-\caption{Incremental CH}
-\label{ch-incremental}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by decreasing weight
-\STATE $s \gets $ empty solution
-\FORALL{$c \in C'$}
-  \IF{$s \cup c$ is a feasible ID set}
-    \STATE $s \gets s + c$
-  \ENDIF
-\ENDFOR
-\RETURN $s$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Removal}}
-
-This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
-This probably does not satisfy the ID set condition.
-Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
-Again, this is a hungry heuristic returning only one solution.
-
-See listing \ref{ch-removal}.
-
-\begin{algorithm}
-\caption{Removal CH}
-\label{ch-removal}
-\begin{algorithmic}
-\REQUIRE $C$ list of candidate AMs
-\ENSURE a feasible solution
-\STATE $C' \gets $ sort $C$ by increasing weight
-\STATE $s \gets C'$
-\FORALL{$c \in s$}
-  \IF{$s$ is a feasible ID set}
-    \RETURN $s$
-  \ENDIF
-  \STATE $s \gets s \backslash c$
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Truncated Branch \& Bound}
-
-This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
-
-TODO if we limit GLPK's runtime, we get this
-
-\subsection{Improvement Heuristics}
-
-\nomenclature{IH}{Improvement Heuristic}
-
-TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
-
-TODO mention intensification, diversification
-
-\subsubsection{\jmodule{Identity}}
-
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{ih-identity}.
-
-\begin{algorithm}
-\caption{Identity IH}
-\label{ih-identity}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE the same pool of feasible solutions
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Remove Worst}}
-
-This trivial IH tries to improve the solution pool by removing the worst solution (i.e.
-the one with the lowest quality).
-This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
-
-See listing \ref{ih-removeworst}.
-
-\begin{algorithm}
-\caption{Remove Worst IH}
-\label{ih-removeworst}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\ENSURE pool of feasible solutions
-\STATE $s_{min} \gets $ solution with the lowest weight $\in FP$
-\RETURN $FP \backslash s_{min}$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Random Remove}}
-
-This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
-By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
-
-The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
-For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
-
-This heuristic returns a pool of solutions of the same size as it got on its input.
-
-See listing \ref{ih-randomremove}.
-
-\begin{algorithm}
-\caption{Random Remove IH}
-\label{ih-randomremove}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE $K \gets k * |s|$
-  \STATE remove $K$ random AMs from $s$
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Hungry}}
-
-This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
-
-\heu{Hungry} tries to improve each solution in the feasible pool in the following way.
-It orders all candidate AMs not present in the solution by decreasing weight.
-Afterwards, it iteratively tries to extend the solution with these AMs, taking care not to violate the ID set condition.
-The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{ih-hungry} captures this process.
-
-\begin{algorithm}
-\caption{Hungry IH}
-\label{ih-hungry}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $C$ list of candidate AMs
-\ENSURE pool of feasible solutions
-\FORALL{$s \in FP$}
-  \STATE \COMMENT {improve a single solution}
-  \STATE $C' \gets C \backslash s$
-  \STATE $C' \gets C'$ sorted by decreasing weight
-  \FORALL{$c \in C'$}
-    \IF{$s \cup c$ is a feasible ID set}
-      \STATE $s \gets s \cup c$
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\RETURN $FP$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Mutation}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-mutation}.
-
-\begin{algorithm}
-\caption{Mutation IH}
-\label{ih-mutation}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of AMs to fix
-\ENSURE pool of feasible solutions
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE $K \gets k * |incumbent|$
-\STATE fix $K$ random AMs from $incumbent$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Crossover}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-crossover}.
-
-\begin{algorithm}
-\caption{Crossover IH}
-\label{ih-crossover}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of solutions to look for commonalities in
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |FP|$
-\STATE $FP' \gets K$ random solutions $\in FP$
-\STATE $am \gets$ AMs found in all solutions $\in FP'$
-\STATE fix $am$ in GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{\jmodule{Local Branching}}
-
-TODO explain how this translates to GLPK input
-
-See listing \ref{ih-localbranching}.
-
-\begin{algorithm}
-\caption{Local Branching IH}
-\label{ih-localbranching}
-\begin{algorithmic}
-\REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
-\ENSURE pool of feasible solutions
-\STATE $K \gets k * |$total AM count$|$
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE add max Hamming distance requirement to GLPK problem formulation
-\STATE $improved \gets $ run GLPK
-\RETURN $FP \cup improved$
-\end{algorithmic}
-\end{algorithm}
-
-\section{IDREF}
-
-TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
-
-\chapter{Experiments}
-
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
-
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
-
-First of all, we would like to get an idea of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
-
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
-
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
-
-\section{Experimental Data}
-
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
-
-\begin{itemize}
-	\item Realistic 
-	\item Realistic with artificial (converted) attributes
-	\item Artificial
-\end{itemize}
-
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
-
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
-
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
-Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
-
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
-
-\subsection{Realistic data}
-
-From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
-
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
-
-\begin{table}
-  \caption{List of realistic test data files}
-  \bigskip
-  \label{table-experiments-data-realistic}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\ 
-	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data}
-  \label{image-experiments-data-realistic}
-  \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
-\end{figure}
-
-To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
-
-On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
-
-Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
-
-\subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
-
-We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
-None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
-
-TODO schemas?
-
-To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
-
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
-
-\begin{table}
-  \caption{List of realistic test data files with converted attributes}
-  \bigskip
-  \label{table-experiments-data-converted}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-	\hline
-	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data with converted attributes}
-  \label{image-experiments-data-converted}
-  \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
-\end{figure}
-
-Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
-
-In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
-Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
-
-\subsection{Artificial data}
-
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
-
-This is actually simple. Consider the following excerpt from an XML file:
-
-\begin{scriptsize}
-\begin{verbatim}
-<graph>
-  <vertex0 attr="-2968876296119015800"/>
-  <vertex1 attr="1729745997570096518"/>
-  <vertex2 attr="-9020549659620928934"/>
-  ...
-  <vertex99 attr="-7545982394508643394"/>
-
-  <vertex82 attr="0"/><vertex21 attr="0"/>
-  <vertex64 attr="1"/><vertex21 attr="1"/>
-  <vertex44 attr="2"/><vertex2 attr="2"/>
-  ...
-  <vertex96 attr="99"/><vertex40 attr="99"/>
-</graph>
-\end{verbatim}
-\end{scriptsize}
-
-We want to create a graph with around $v$ vertices and $e$ edges.
-First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
-
-The pseudocode for this is in listing \ref{listing-random-data}.
-
-\begin{algorithm}
-\caption{Random XML data creation}
-\label{listing-random-data}
-\begin{algorithmic}
-\REQUIRE $v$ requested number of vertices
-\REQUIRE $e$ requested number of edges
-\ENSURE XML file content
-\PRINT \texttt{<graph>}
-\FOR{$i = 1 \to |V|$}
-	\STATE $R \gets RANDOM$
-	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
-\ENDFOR
-\FOR{$i = 1 \to |E|$}
-	\STATE $v1 \gets RANDOM(|V|)$
-	\STATE $v2 \gets RANDOM(|V|)$
-	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
-\ENDFOR
-\PRINT \texttt{</graph>}
-\RETURN
-\end{algorithmic}
-\end{algorithm}
-
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
-
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
-
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
-Note that the name is always in the form $v-e$.
-
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
-
-TODO this is on the DVD.
-
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
-
-\begin{table}
-  \caption{List of artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\ 
-    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\ 
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{List of ``sized" artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial-size}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-	\hline
-	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\ 
-  20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-  30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\ 
-	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Artificial data}
-  \label{image-experiments-data-artificial}
-  \centering
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
-\end{figure}
-
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
-
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
-
-\section{Experimental Setup}
-
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
-
-We now have to define a few notions before moving forward to the description of our experiments.
-
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
-	\begin{itemize}
-		\item All the parameters in all the heuristics.
-		\item The specific way in which the heuristics are chained.
-		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
-		\item Initial pool size.
-		\item The termination criteria.
-		\item The input XML file.
-		\item Known optimum for this file and $\alpha$, $\beta$.
-	\end{itemize}
-\end{define}
-
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
-
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
-
-\subsection{Grammar and Model Creation}
-
-TODO link BasicIGG to see how we get IG
-
-TODO describe how we get the model
-
-\subsection{Hardware and Software}
-
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Linux Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 64-Bit Server VM (build 20.1-b02)
-GLPK version 4.38
-\end{verbatim}
-
-\subsubsection{Windows Machine}
-
-\begin{verbatim}
-Intel Core 2 Duo processor @ 2.33 GHz
-4 GB DDR2 RAM
-Windows 7 SP1 64bit
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 32-Bit Client VM (build 20.1-b02)
-GLPK version 4.45 (Cygwin)
-GLPK version 4.34 (native)
-\end{verbatim}
-
-\subsection{Methodology}
-
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
-This improvement is left for the future work.
-
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
-OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
-
-\subsection{Measuring the time}
-
-TODO
-
-\subsection{Obtaining the results}
-
-Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
-We can get all the information relevant to that experiment run from this trace alone.
-An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
-
-\section{Experimental Results}
-
-\subsection{Grammar and model generation}
-
-%         NB class GrammarModelTiming
-
-TODO in this experiment we shall see how long it takes to create the IG and the model. So far, this is isolated from heus themselves.
-
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & TODO \\
-  Input data set    & all test data sets \\
-  Iterations        & 20 \\
-  Pool size         & not applicable \\
-  $\alpha$, $\beta$ & not applicable \\
-  \hline
-\end{tabular}
-
-The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There won't be any CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
-
-TODO graph + interpretation
-
-\subsubsection{GLPK interface}
-
-TODO how long it takes to create GLPK input and parse its output
-
-\subsection{GLPK: native vs. Cygwin}
-
-%         NB class TimeQuality
-%         NB class TimeTillOptimum
-
-In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & \texttt{100-500} \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
-There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
-\label{listing-experiment-glpk-native-vs-cygwin-1}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Windows \\
-  Input data set    & sized artificial data \\
-  Iterations        & 20 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
-Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized artificial data set (TODO link) (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
-\label{listing-experiment-glpk-native-vs-cygwin-2}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized artificial data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$  
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO graph + interpretation
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
-\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
-
-TODO take note of quality and time too
-
-\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
-
-TODO
-
-\subsection{Best standalone CH}
-
-TODO compare all CHs - limit GLPK to ~ 1 second
-
-\subsection{Best IH for the best CH}
-
-TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
-
-\subsubsection{\jmodule{Random} as CH}
-
-TODO with the best IH from the previous step, how about using Random as CH?
-
-\subsection{Various $\alpha$, $\beta$}
-
-TODO remind what these parameters mean
-
-\subsubsection{Effects on performance}
-
-TODO
-
-\subsubsection{Effects on the ID set}
-
-TODO
-
-\subsection{Ignoring text data}
-
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-
-\subsection{Chaining the IHs}
-
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
-
-\subsection{Algorithms for various data characteristics}
-
-TODO
-
-\section{The "Best" Algorithm}
-
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run. 
-We will find the optimum in this way. 
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
-
-\chapter{Future work}
-
-TODO
-
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
-
-\chapwithtoc{Conclusion}
-
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
-
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
+\include{chapter-6-conclusion}
 
@@
 
-\chapter{jInfer}
-\label{appendix-jInfer}
+\include{appendix-a-jinfer}
 
-This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
-Please see TODO link jInfer project web for complete information, documentation and download options.
+\include{appendix-b-iss}
 
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
-The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
-The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
-Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
-This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
-
-To the best of our knowledge, at the time of writing this thesis is jInfer the only public, open source and actually working solution for XML schema inference-related tasks.
-
-At heart of jInfer inference process is a modular system provided by NetBeans Platform allowing to define services (interfaces), implement them in any number of ways and then let the user choose which implementation to use.
-Most importantly, the whole process consists of 3 consecutive steps (see \ref{image-inference-process}), responsibility of 3 different services - interchangeable modules.
-
-\begin{figure}
-  \caption{Inference process in jInfer}
-  \label{image-inference-process}
-  \centering
-    \includegraphics[width=0.5\textwidth]{images/inference-process}
-\end{figure}
-
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
-This is the representation in which will the structure live until it is used to create the final product - the schema.
-As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
-As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
-(TODO picture?) IG is used to create the AM model used in this thesis, too.
-jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
-
-After leaving the \jmodule{Initial Grammar Generator}, the IG needs to be made more general, shortened, \textit{simplified}.
-This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
-Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
-
-The last module, \jmodule{Schema Generator} takes the simplified grammar and creates the resulting schema from it.
-This process is not too interesting, but anyone wishing to find out all about it is invited to read the documentation to the two \jmodule{Schema Generator}s bundled with jInfer - the BasicDTD and BasicXSD modules.
-
-\chapter{IDSetSearch}
-\label{appendix-iss}
-
-TODO talk a bit about the structure of ISS module.
-
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
-
-TODO package name root
-
-TODO "Attribute Stats"
-
-TODO heuristics - interfaces, construction, improvement
-
-TODO experiments - quality, termination, parameters, experiment sets
-
-TODO make sure there was a place where we described the model data model:)
-
-\chapter{Experimental Trace}
-\label{appendix-trace}
-
-Following is a trace logged from a sample experiment run. It shows all the relevant information related to this instance, any and every piece of information we might be interested in.
-
-To save space, 2-column layout is used. Commentary on the particulars follows right after its end.
-
-\begin{multicols}{2}
-\begin{scriptsize}
-\begin{verbatim}
-CPU info
-  Intel(R) Core(TM)2 Quad CPU Q9550 @ 2.83GHz
-  Cores: 4
-  Clock speed: 2983 MHz
-Memory info
-  Size: 8192 MB
-OS info
-  Name: Windows 7
-  Version: 6.1
-  Architecture: amd64
-Java info
-  Version: 1.6.0_26
-  VM: Java HotSpot(TM) 64-Bit Server VM
-GLPK info
-  GLPSOL: GLPK LP/MIP Solver 4.34
-
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-
-Results:
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-Construction phase: 
-  Algorithm: Random
-    Time taken: 248 ms
-    Time since start: 248 ms
-    Pool size: 10
-    Quality: 0.19975609756097568 (11 AMs)
-Improvement phase: 
-  pass #1: 
-  Algorithm: RandomRemove, ratio = 0.2
-    Time taken: 0 ms
-    Time since start: 841 ms
-    Pool size: 10
-    Quality: 0.15878048780487808 (9 AMs)
-  pass #2: 
-  Algorithm: Mutation, ratio = 0.1, limit = 1 s
-    Time taken: 1512 ms
-    Time since start: 2710 ms
-    Pool size: 11
-    Quality: 0.21975609756097558 (11 AMs)
-
-  <... 7 more passes removed ...>
-
-  pass #10: 
-  Algorithm: Remove Worst
-    Time taken: 80 ms
-    Time since start: 7676 ms
-    Pool size: 12
-    Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
-
-Time,Quality,AMs
-248,0.19975609756097568,11
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-2927,0.1890243902439024,9
-4421,0.23463414634146343,12
-4703,0.23463414634146343,12
-4896,0.1960975609756098,10
-5793,0.23463414634146337,12
-5972,0.19951219512195123,10
-7433,0.19951219512195123,10
-7676,0.19951219512195123,10
-
-ID
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-vertex2,attr,0.01975609756097561
-vertex33,attr,0.016829268292682928
-vertex34,attr,0.02219512195121951
-vertex4,attr,0.022682926829268292
-vertex41,attr,0.014878048780487804
-vertex7,attr,0.02170731707317073
-vertex70,attr,0.018780487804878048
-vertex76,attr,0.01780487804878049
-vertex8,attr,0.02170731707317073
-vertex80,attr,0.01780487804878049
-vertex97,attr,0.016341463414634147
-
-IDREF
-Element,Attribute
-\end{verbatim}
-\end{scriptsize}
-\end{multicols}
-
-The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
-To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
-It tries to guess whether it's the Cygwin version by looking at the path to the binary.
-
-% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
-The second section states the input file along with its size and graph representation.\\
-Alpha and beta parameters for this instance belong here too.
-
-\begin{footnotesize}
-\begin{verbatim}
-Configuration:
-File name: graph.xml (101599 b)
-  Graph representation: 82 vertices, 1101 edges
-alpha: 1.0, beta: 1.0
-\end{verbatim}
-\end{footnotesize}
-
-Results section opens stating the most important information first: how long did the experiment run and what was the highest and final quality (these two are potentially different).
-Numbers of attribute mappings in the best and final solution respectively are stated as well.
-
-\begin{footnotesize}
-\begin{verbatim}
-Total time spent: 7754 ms
-Final quality: 0.19951219512195123 (10 AMs)
-Highest quality: 0.23463414634146343 (12 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Construction phase results go next.
-Among reported information are the full identification of the heuristic (possibly along with its parameters), time taken, size of the pool created and the quality of the incumbent solution (again, with the number of its AMs).
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Random
-  Time taken: 248 ms
-  Time since start: 248 ms
-  Pool size: 10
-  Quality: 0.19975609756097568 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-Now for each of the improvement phases there is one section in output log.
-Information presented here has the same structure as with the construction phase.
-Please note that the \code{Pool size} is always measured \textit{after} the improvement run.
-
-\begin{footnotesize}
-\begin{verbatim}
-Algorithm: Mutation, ratio = 0.1, limit = 1 s
-  Time taken: 1512 ms
-  Time since start: 2710 ms
-  Pool size: 11
-  Quality: 0.21975609756097558 (11 AMs)
-\end{verbatim}
-\end{footnotesize}
-
-After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
-\\
-
-To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
-Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
-
-\begin{footnotesize}
-\begin{verbatim}
-Time,Quality,AMs
-...
-841,0.15878048780487808,9
-2710,0.21975609756097558,11
-...
-\end{verbatim}
-\end{footnotesize}
-
-And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
-Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
-
-\begin{footnotesize}
-\begin{verbatim}
-Element,Attribute,Weight
-vertex0,attr,0.024146341463414635
-...
-\end{verbatim}
-\end{footnotesize}
-
-Note that in this example trace there were no IDREF AMs found.
+\include{appendix-c-trace}
 

### 2011-08-16 doc/theses/vitasek/appendix-b-iss.tex
Thesis: splitting tex files.

\chapter{IDSetSearch}
\label{appendix-iss}

TODO talk a bit about the structure of ISS module.

TODO name of the module, its place in jInfer, interface with the framework (service providing)

TODO package name root

TODO "Attribute Stats"

TODO heuristics - interfaces, construction, improvement

TODO experiments - quality, termination, parameters, experiment sets

TODO make sure there was a place where we described the model data model:)

### 2011-08-16 doc/theses/vitasek/chapter-2-research.tex
Thesis: splitting tex files.

\chapter{Research}

According to the article \cite{fidax}, ...

TODO talk about FIDAX
To the best of our knowledge, there are no other articles dealing with this problem.

TODO talk about Fajt - but that's different, that's keys
TODO add citation for Fajt

\section{Independent Set}

TODO define IS + MIS rigorously

TODO other approaches to max IS

### 2011-08-16 doc/theses/vitasek/chapter-5-future.tex
Thesis: splitting tex files.

\chapter{Future work}

TODO

\begin{itemize}
  \item extension to more documents at once
  \item more heuristics and their chaining
  \item ACO, genetic, ...
  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
  \item jInfer is opensource, thus easy to extend!
\end{itemize}

### 2011-08-16 doc/theses/vitasek/chapter-6-conclusion.tex
Thesis: splitting tex files.

\chapwithtoc{Conclusion}

TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them. We discussed the approach chosen in FIDAX. Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t. weight) ID set.

However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics. By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.

### 2011-08-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
   	\hline
-	OVA1  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-	OVA2  & 11.9     & 23 & 36 & 0.1634615384615385  \\
-	OVA3  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-	XMA-c & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-	XMA-p & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-	XMD   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
+  	\jmodule{OVA1}  & 4.5      & 29 & 43 & 0.45588235294117635 \\
+  	\jmodule{OVA2}  & 11.9     & 23 & 36 & 0.1634615384615385  \\
+  	\jmodule{OVA3}  & 237.6    & 31 & 47 & 0.25537156151635415 \\
+  	\jmodule{XMA-c} & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
+  	\jmodule{XMA-p} & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
+  	\jmodule{XMD}   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
   \end{tabular}
@@
   \centering
-    \subfigure[OVA1]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[OVA2]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[OVA3]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[XMA-c]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[XMA-p]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[XMD]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
+    \subfigure[\jmodule{OVA1}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
+    \subfigure[\jmodule{OVA2}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
+    \subfigure[\jmodule{OVA3}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
+    \subfigure[\jmodule{XMA-c}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
+    \subfigure[\jmodule{XMA-p}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
+    \subfigure[\jmodule{XMD}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
 \end{figure}
@@
   	\hline
-	MSH  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-	NTH  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\
+  	\jmodule{MSH}  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
+  	\jmodule{NTH}  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\
   \end{tabular}
@@
   \centering
-    \subfigure[MSH]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[NTH]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
+    \subfigure[\jmodule{MSH}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
+    \subfigure[\jmodule{NTH}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
 \end{figure}
@@
   	\hline
-	100-100  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-	100-200  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\
-    100-1000 & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\
+  	\jmodule{100-100}  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
+  	\jmodule{100-200}  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\
+    \jmodule{100-1000} & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\
   \end{tabular}
@@
   	\hline
-	0-0     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-	10-5    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\
-  20-20   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-  30-45   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\
-	40-80   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-	50-125  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-	60-180  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-	70-245  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-	80-320  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-	90-405  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-	100-500 & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
+  	\jmodule{0-0}     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
+  	\jmodule{10-5}    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\
+    \jmodule{20-20}   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
+    \jmodule{30-45}   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\
+  	\jmodule{40-80}   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
+  	\jmodule{50-125}  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
+  	\jmodule{60-180}  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
+  	\jmodule{70-245}  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
+  	\jmodule{80-320}  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
+  	\jmodule{90-405}  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
+  	\jmodule{100-500} & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
   \end{tabular}
@@
   \centering
-  	\subfigure[48-80]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[70-245]{\includegraphics[width=0.45\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
+  	\subfigure[\jmodule{48-80}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/40-80}}
+	\subfigure[\jmodule{70-245}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
 \end{figure}
@@
 \subsubsection{GLPK interface timing}
+
+%         NB class GlpkInterfaceTiming
 
@@
 
+%         NB class FidaxWithHungry
+
 There is a minor question to be answered easily: is \jmodule{FIDAX} ``hungry enough"? Could we improve it using \jmodule{Hungry} as IH? Let us design a short experiment.
@@
 
-TODO table + interpretation
+The experimental results are summarized in the Table \ref{table-experiments-fidax-and-hungry} and are quite surprising.
+As trivial a heuristic \jmodule{Hungry} is, it is still able to improve the ID set found by \jmodule{FIDAX} by as much as almost 50\% (the last row, \jmodule{100-1000}).
+
+Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \jmodule{100-1000}.
+Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute".
+
+\begin{table}
+  \caption{Results of adding \jmodule{Hungry} after \jmodule{FIDAX}}
+  \bigskip
+  \label{table-experiments-fidax-and-hungry}
+  \centering
+  \begin{tabular}{l | l | l}
+    Data set & Quality - \jmodule{FIDAX} & Quality - \jmodule{FIDAX} + \jmodule{Hungry} \\
+    \hline
+    \jmodule{OVA1}     & 0.4411764705882353  & 0.4411764705882353   \\
+    \jmodule{OVA2}     & 0.16346153846153846 & 0.16346153846153846  \\
+    \jmodule{OVA3}     & 0.25482414123443264 & \textbf{0.2553715615163541}   \\
+    \jmodule{XMA-c}    & 0.7546666666666666	 & 0.7546666666666666   \\
+    \jmodule{XMA-p}    & 0.2019306150568969	 & 0.2019306150568969   \\
+    \jmodule{XMD}      & 0.09786094165493509 & 0.09786094165493509  \\
+    \jmodule{MSH}      & 0.5416472778036296	 & 0.5416472778036296   \\
+    \jmodule{NTH}      & 0.05259709474828076 & \textbf{0.057918595422124436} \\
+    \jmodule{100-100}  & 0.56	               & \textbf{0.6766666666666669}   \\
+    \jmodule{100-200}  & 0.44200000000000017 & \textbf{0.5980000000000003}   \\
+    \jmodule{100-1000} & 0.19952380952380955 & \textbf{0.29619047619047617}  \\
+  \end{tabular}
+\end{table}
+
+% TODO for some reason \texttt{\textbf{vertex5}} looks only like \texttt{vertex5}
+\begin{table}
+  \caption{ID sets in \jmodule{FIDAX} versus \jmodule{FIDAX} + \jmodule{Hungry}}
+  \bigskip
+  \label{table-experiments-fidax-and-hungry-idsets}
+  \centering
+  \begin{tabular}{l | l}
+  \jmodule{FIDAX} & \jmodule{FIDAX} + \jmodule{Hungry} \\
+  \hline
+                    & \texttt{\textbf{vertex5}}  \\
+                    & \texttt{\textbf{vertex26}} \\
+  \texttt{vertex30} & \texttt{vertex30} \\
+  \texttt{vertex31} & \texttt{vertex31} \\
+  \texttt{vertex32} & \texttt{vertex32} \\
+  \texttt{vertex34} & \texttt{vertex34} \\
+  \texttt{vertex35} & \texttt{vertex35} \\
+  \texttt{vertex36} & \texttt{vertex36} \\
+  \texttt{vertex37} & \texttt{vertex37} \\
+  \texttt{vertex39} & \texttt{vertex39} \\
+                    & \texttt{\textbf{vertex60}} \\
+                    & \texttt{\textbf{vertex69}} \\
+                    & \texttt{\textbf{vertex70}} \\
+  \texttt{vertex74} & \texttt{vertex74} \\
+  \texttt{vertex75} & \texttt{vertex75} \\
+  \texttt{vertex80} & \texttt{vertex80} \\
+  \end{tabular}
+\end{table}
 

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/BestStandaloneCH.java
#3323201 Experimental framework

@@ -38,13 +38,12 @@ import java.util.List;
 import org.openide.util.lookup.ServiceProvider;
 
 /**
- * Experiment comparing Random, Fuzzy and FIDAX construction heuristics, only
- * their first steps.
+ * Experiment comparing all the CHs, without any IHs.
  *
  * @author vektor
  */
 @ServiceProvider(service = ExperimentSet.class)
-public class RandomVsFuzzyVsFidaxStart extends AbstractExperimentSet {
+public class BestStandaloneCH extends AbstractExperimentSet {
 
   private File finalCsvQuality = new File(Constants.TEST_OUTPUT_ROOT + "/" + getName() + "/result-quality.txt");
   private File finalCsvTime = new File(Constants.TEST_OUTPUT_ROOT + "/" + getName() + "/result-time.txt");

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/FidaxWithHungry.java

@@ -20,55 +20,49 @@ import cz.cuni.mff.ksi.jinfer.base.utils.FileUtils;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.Experiment;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.data.SizeTestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.OfficialTestData;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.TestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.quality.Weight;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.glpk.Glpk;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.fidax.Fidax;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Hungry;
 import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Identity;
-import cz.cuni.mff.ksi.jinfer.iss.options.ISSPanel;
 import cz.cuni.mff.ksi.jinfer.iss.utils.Constants;
 import java.io.File;
 import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.List;
-import org.apache.log4j.Logger;
-import org.openide.util.NbPreferences;
 import org.openide.util.lookup.ServiceProvider;
 
 /**
- * Experiment measuring quality depending on how long we let the GLPK
- * construction run. This can also be used to compare different versions of GLPK.
+ * Experiment determining whether FIDAX can be improved with Hungry.
  *
  * @author vektor
  */
 @ServiceProvider(service = ExperimentSet.class)
-public class TimeQuality extends AbstractExperimentSet {
+public class FidaxWithHungry extends AbstractExperimentSet {
 
-  private static final int BEGIN = 1;
-  private static final int END = 50;
-  private static final int STEP = 2;
-
[…]

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/GlpkInterfaceTiming.java

@@ -16,11 +16,10 @@
  */
 package cz.cuni.mff.ksi.jinfer.iss.experiments.sets;
 
-import cz.cuni.mff.ksi.jinfer.base.utils.FileUtils;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.Experiment;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.OfficialTestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.data.SizeTestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.data.TestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
@@ -28,30 +27,29 @@ import cz.cuni.mff.ksi.jinfer.iss.experiments.quality.Weight;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.termination.TimeIterations;
 import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.glpk.Glpk;
 import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Identity;
-import cz.cuni.mff.ksi.jinfer.iss.options.ISSPanel;
 import cz.cuni.mff.ksi.jinfer.iss.utils.Constants;
-import java.io.File;
+import java.io.IOException;
 import java.util.ArrayList;
 import java.util.Arrays;
 import java.util.List;
+import org.apache.log4j.EnhancedPatternLayout;
+import org.apache.log4j.FileAppender;
 import org.apache.log4j.Logger;
-import org.openide.util.NbPreferences;
+import org.openide.util.Exceptions;
 import org.openide.util.lookup.ServiceProvider;
 
 /**
- * Experiment measuring how long it takes to reach the optimum, based on the
- * input size. This can also be used to compare different versions of GLPK.
+ * Runs all the official test data to see how long it takes them to communicate
+ * with GLPK (which is used as CH, no time limit). No IH is employed thereafter.
  *
  * @author vektor
  */
 @ServiceProvider(service = ExperimentSet.class)
-public class TimeTillOptimum extends AbstractExperimentSet {
-
-  private File finalCsv;
[…]

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/AbstractExperimentSet.java
#3323201 Experimental framework

@@ -47,9 +47,11 @@ public abstract class AbstractExperimentSet implements ExperimentSet {
 
   @Override
   public void run(final int from) throws InterruptedException {
-    LOG.info("Running from index " + from);
     if (from == 0) {
       notifyStart();
+    }
+    else {
+      LOG.info("Running from index " + from);
     }
     final List<ExperimentParameters> experiments = getExperiments();
     final int totalCount = experiments.size();
@@ -70,7 +72,7 @@ public abstract class AbstractExperimentSet implements ExperimentSet {
         }
       }, "Experiment Runner");
 
-      LOG.debug("Running iteration #" + i + "/" + totalCount);
+      LOG.debug("Running iteration #" + (i + 1) + "/" + totalCount);
       t.start();
 
       synchronized (monitor) {
@@ -92,7 +94,7 @@ public abstract class AbstractExperimentSet implements ExperimentSet {
 
     @Override
     public void experimentFinished(final Experiment e) {
-      LOG.debug("Iteration #" + iteration + " finished");
+      LOG.debug("Iteration #" + (iteration + 1) + " finished");
       FileUtils.appendString("\n" + iteration, Constants.RESTART);
       notifyFinished(e, iteration);
       synchronized (monitor) {

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/Experiment.java

@@ -34,6 +34,7 @@ import cz.cuni.mff.ksi.jinfer.base.objects.Input;
 import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
 import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
 import cz.cuni.mff.ksi.jinfer.base.utils.ModuleSelectionHelper;
+import cz.cuni.mff.ksi.jinfer.iss.utils.MappingUtils;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.List;
@@ -208,7 +209,7 @@ public class Experiment {
     igg.start(input, new IGGeneratorCallback() {
 
       @Override
-      public void finished(List<Element> grammar) {
+      public void finished(final List<Element> grammar) {
         grammarTime = delta(grammarStartTime);
         final long modelStartTime = time();
         model = new AMModel(grammar);
@@ -266,6 +267,8 @@ public class Experiment {
       else {
         improvementResults.add(result);
       }
+
+      assert(MappingUtils.isIDset(incumbent.getFirst().getMappings(), model));
 
       final long totTime = delta(startTime);
       final Pair<Boolean, String> termination = params.getTerminationCriterion().terminate(Experiment.this, totTime, feasiblePool);

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/heuristics/construction/glpk/GlpkInputGenerator.java

@@ -21,6 +21,7 @@ import cz.cuni.mff.ksi.jinfer.iss.utils.MappingUtils;
 import cz.cuni.mff.ksi.jinfer.iss.objects.AMModel;
 import cz.cuni.mff.ksi.jinfer.iss.objects.AttributeMappingId;
 import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
+import cz.cuni.mff.ksi.jinfer.iss.utils.Utils;
 import java.util.Collections;
 import java.util.List;
 import org.apache.log4j.Logger;
@@ -78,6 +79,8 @@ public final class GlpkInputGenerator {
           final double alpha, final double beta,
           final List<AttributeMappingId> incumbentAMs, final long k)
           throws InterruptedException {
+    final long start = Utils.time();
+
     final List<AttributeMappingId> candidates = MappingUtils.getCandidates(model);
     Collections.shuffle(candidates);
 
@@ -104,13 +107,17 @@ public final class GlpkInputGenerator {
       }
     }
 
-    return GlpkUtils.loadTemplate(TEMPLATE_LB).toString()
+    final String ret = GlpkUtils.loadTemplate(TEMPLATE_LB).toString()
             .replace("{constraints}", getConstraints(candidates, Collections.<AttributeMappingId>emptyList(), model))
             .replace("{mappings}", mappings)
             .replace("{weights}", weights)
             .replace("{k}", String.valueOf(k))
             .replace("{incumbent}", incumbent)
             .replace("{remaining}", remaining);
+
+    LOG.debug("GLPK input generation took " + Utils.delta(start) + " ms.");
+
+    return ret;
   }
 
   /**
@@ -155,6 +162,8 @@ public final class GlpkInputGenerator {
   public static String generateGlpkInput(final AMModel model,
           final List<AttributeMappingId> fixed,
           final double alpha, final double beta) throws InterruptedException {
+    final long start = Utils.time();
+
     final List<AttributeMappingId> candidates = MappingUtils.getCandidates(model);
     Collections.shuffle(candidates);
 
@@ -173,7 +182,7 @@ public final class GlpkInputGenerator {
               .append('\n');
     }
 
-    final String result = GlpkUtils.loadTemplate(TEMPLATE).toString()
+    final String ret = GlpkUtils.loadTemplate(TEMPLATE).toString()
             .replace("{constraints}", getConstraints(candidates, fixed, model))
             .replace("{mappings}", mappings)
[…]

### 2011-08-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/heuristics/construction/glpk/GlpkOutputParser.java

@@ -19,10 +19,12 @@ package cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.glpk;
 import cz.cuni.mff.ksi.jinfer.iss.objects.AMModel;
 import cz.cuni.mff.ksi.jinfer.iss.objects.AttributeMappingId;
 import cz.cuni.mff.ksi.jinfer.iss.objects.IdSet;
+import cz.cuni.mff.ksi.jinfer.iss.utils.Utils;
 import java.util.ArrayList;
 import java.util.HashSet;
 import java.util.Scanner;
 import java.util.Set;
+import org.apache.log4j.Logger;
 
 /**
  * Utility class for parsing the GLPK optimization result into a set of attribute
@@ -31,6 +33,8 @@ import java.util.Set;
  * @author vektor
  */
 public final class GlpkOutputParser {
+
+  private static final Logger LOG = Logger.getLogger(GlpkOutputParser.class);
 
   private GlpkOutputParser() {
   }
@@ -63,6 +67,8 @@ public final class GlpkOutputParser {
    */
   public static IdSet getIDSet(final String output, final AMModel model,
           final boolean isFixed) {
+    final long start = Utils.time();
+
     final Set<AttributeMappingId> ret = new HashSet<AttributeMappingId>();
 
     final Scanner s = new Scanner(output);
@@ -85,9 +91,13 @@ public final class GlpkOutputParser {
       }
     }
 
-    return new IdSet(
+    final IdSet result = new IdSet(
             new ArrayList<AttributeMappingId>(ret),
             !isFixed && output.contains("INTEGER OPTIMAL SOLUTION FOUND"));
+
+    LOG.debug("GLPK output parsing took " + Utils.delta(start) + " ms.");
+
+    return result;
   }
 
 }

### 2011-08-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
   \hline
-  Machine           & TODO \\
+  Machine           & Windows \\
   Input data set    & all official test data sets \\
@@
 
+%         NB class BestStandaloneCH
+
 TODO compare all CHs - limit GLPK to ~ 1 second
+TODO pool size 10 where applicable
+
+TODO chart for each file: CHs on the X axis, quality Y axis, boxplots
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & TODO \\
+  Input data        & all official test data sets \\
+  Iterations        & 20 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
 

### 2011-08-16 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

\chapter{Analysis of XQuery}
\todo[inline]{bibref na knihu, a W3C XQuery usecases}
This chapter discusses selected constructs of XQuery language and denotes how they could be exploited in the XML schema inference process. It is divided into sections by particular domains of the inference.

\section{Structure of XML documents}
The most queries can be exploited to obtain some information about a structure of respective XML documents. The structure of XML documents captures elements from these documents along with their names, attributes and their organization. What are the outermost elements in these document, which elements can be contained inside a certain element, whether are they optional or mandatory, etc.

Path expressions without predicates which use only child axis are the simpliest example of such queries. Path expression \texttt{/bib/book/author} indicates element \texttt{bib} is the outermost element and contains one or more elements \texttt{book} and these contains one or more elements \texttt{author}.

Additional path expressions \texttt{/bib/book/title}, \texttt{//author/name} indicate that element \texttt{book} also contains element \texttt{title} and element \texttt{author} contains element \texttt{name}. The latter one uses also descendant axis.

Besides elements, attributes could be processed exactly the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} has attribute \texttt{@ISBN}.

However statemets of this kind are necessary for the XML schema inference, their obtaining from queries it not significant because of the following reason. These statement could be determined directly from the XML documents and it could be easily done in a more convenient way. Queries may not cover the whole relevant content of the documents. For example some element may not be queried at all, therefore they mat not be mentioned in any query. In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurence of elements. At last, we cannot be sure that queries target nodes actually presented in the XML documents. Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not. In opposite, even basic methods of XML documents analyzis do not have these inadequacies.

On the other hand, the structure inference could be useful when the whole set of all XML documents is not available and provided XML documents do not cover the structure completly.

\section{Occurrence Count of Elements}
Some XQuery constructs indicate multiplicity of a particular element or limit the element to occur at most once. Consider the following query assuming variable \texttt{\$book1} is bound to a certain \texttt{book} element.

\begin{verbatim}
for $a in $book1/author 
order by $a/last, $a/first
return $a
\end{verbatim}

Apparently, this query expects more than one element \texttt{author} to be children of the element variable \texttt{\$book1} is bound to. Otherwise, any sorting would lack a reason. Although, we cannot be absolutely sure about it again, assuming common sense usage of XQuery it is very likely that element \texttt{author} can occur multiple times as a subelement of element from variable \texttt{\$book1}.

\subsection{Multiple Occurrence}
Similar approach could be applied in many other situations. Another examples are particular usages of function \texttt{count()}, indexation, usage of set operators (\texttt{union}, \texttt{intersect}, \texttt{except}) and usage of function \texttt{one-or-more()}. Queries with description follows.

\begin{verbatim}
<section_count>{ count(/book/section) }</section_count>
\end{verbatim}

Function \texttt{count()} returns number of items in a provided sequence. If the sequence is sequence of elements the count of these elements will probably not be limited to one. The example query indicates that the outermost element \texttt{book} can contain more than one element \texttt{section}.
An exception is usage of function \texttt{count()} in a predicate in expressions where it is used to determine if count of some nodes is greater than zero. Often, this form is used to test presence of a certain node instead of actual counting of its occurrences. In this case, the node could be still limited to occur at most once.

\begin{verbatim}
($s/incision)[2]/instrument
($s/instrument)[position()>=2]
\end{verbatim}

Indexation of nodes and common usage of function \texttt{position()} suggest that writer of such query assumes sequence of respective nodes.

\begin{verbatim}
one-or-more(/catalog/product[@id = 5]/color)
\end{verbatim}

In this query, number of elements \texttt{color} in element \texttt{product} with attribute \texttt{id} equal to 5 has to be at least one otherwise an error is raised upon execution. If we assume that this query is written correctly, with common sense and should not raise the error, we can infer that element \texttt{product} has to containt one or more elements \texttt{color}

\subsection{Occurrence Limited to One}
In opposite to the multiple occurrence, numerous XQuery constructs limit number of occurrences of an element to at most one or exactly one. Some example queries with description follows.

\begin{verbatim}
/catalog/product[1]/number lt 10
\end{verbatim}

\texttt{lt} is a representant of so called \emph{value comparison operators} which operate two sequences of zero or one item. If an operand of a \emph{value comparison operand} is sequence of more than one item \emph{type error} is raised.

\begin{verbatim}
for $item in //item 
order by $item/num 
return $item
\end{verbatim}

Alike the previous example, expression in \texttt{order by} clause can be evaluated to at most one item or the \emph{type error} is raised. Therefore, every element \texttt{item} contains zero or one element \texttt{num} but no more.

Another similar examples are arithmetic expressions and functions accepting a sequence of at most one item. Functions \texttt{zero-or-one()} will raise the \emph{type error} when supplied with an sequence of more than one item.

Those are constructs indicating limitation to zero or one occurrence. Function \texttt{exactly-one} works similar to the function \texttt{zero-or-one} but accepts only sequences of exactly one item (which are in XQuery equal to this item itself).

### 2011-08-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
-From 3 different sources we collected 6 different data sets, called \jmodule{OVA1} - \jmodule{OVA3}, \jmodule{XMA-c}, \jmodule{XMA-p} and \jmodule{XMD}.
+From 3 different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
 Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
@@
   	\hline
-  	\jmodule{OVA1}  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-  	\jmodule{OVA2}  & 11.9     & 23 & 36 & 0.1634615384615385  \\
-  	\jmodule{OVA3}  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-  	\jmodule{XMA-c} & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-  	\jmodule{XMA-p} & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-  	\jmodule{XMD}   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
+  	\dataset{OVA1}  & 4.5      & 29 & 43 & 0.45588235294117635 \\
+  	\dataset{OVA2}  & 11.9     & 23 & 36 & 0.1634615384615385  \\
+  	\dataset{OVA3}  & 237.6    & 31 & 47 & 0.25537156151635415 \\
+  	\dataset{XMA-c} & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
+  	\dataset{XMA-p} & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
+  	\dataset{XMD}   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
   \end{tabular}
@@
   \centering
-    \subfigure[\jmodule{OVA1}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[\jmodule{OVA2}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[\jmodule{OVA3}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[\jmodule{XMA-c}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[\jmodule{XMA-p}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[\jmodule{XMD}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
+    \subfigure[\dataset{OVA1}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
+    \subfigure[\dataset{OVA2}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
+    \subfigure[\dataset{OVA3}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
+    \subfigure[\dataset{XMA-c}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
+    \subfigure[\dataset{XMA-p}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
+    \subfigure[\dataset{XMD}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
 \end{figure}
 
-To interpret the data a little: \jmodule{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
+To interpret the data a little: \dataset{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
 
-On the other hand, the \jmodule{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
+On the other hand, the \dataset{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
 We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
 
-Finally, the \jmodule{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
+Finally, the \dataset{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
 
@@
 
-We found 2 data sets to convert, \jmodule{MSH} and \jmodule{NTH}.
+We found 2 data sets to convert, \dataset{MSH} and \dataset{NTH}.
 Unfortunately, the same problem with disclosure as in the previous case applies here.
@@
 
-To address the conversion: in the case of \jmodule{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \jmodule{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
+To address the conversion: in the case of \dataset{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
+In the case of \dataset{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
 
@@
   	\hline
-  	\jmodule{MSH}  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-  	\jmodule{NTH}  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\
+  	\dataset{MSH}  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
+  	\dataset{NTH}  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\
   \end{tabular}
@@
   \centering
-    \subfigure[\jmodule{MSH}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[\jmodule{NTH}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
+    \subfigure[\dataset{MSH}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
+    \subfigure[\dataset{NTH}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
 \end{figure}
 
-Again to do a little interpretation: in the case of \jmodule{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \jmodule{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
+Again to do a little interpretation: in the case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
+This is then the case similar to \dataset{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
 
-In the case of \jmodule{NTH} we introduced more attributes, 8 to be precise.
+In the case of \dataset{NTH} we introduced more attributes, 8 to be precise.
 Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
@@
 
-The pseudocode for this is in listing \ref{listing-random-data}.
+The pseudocode for this is in Listing \ref{listing-random-data}.
 
@@
 
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \jmodule{100-100}, \jmodule{100-200} and \jmodule{100-1000}.
+We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \dataset{100-100}, \dataset{100-200} and \dataset{100-1000}.
 Note that the name is always in the form $v-e$.\\
@@
 Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \jmodule{0-0} as the trivial one to \jmodule{100-500} as the biggest one.
+For this reason we created 11 more sets, from \dataset{0-0} as the trivial one to \dataset{100-500} as the biggest one.
 From now on, these will be referred to as \textit{sized test data (sets)}.
@@
   	\hline
-  	\jmodule{100-100}  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-  	\jmodule{100-200}  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\
-    \jmodule{100-1000} & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\
+  	\dataset{100-100}  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
+  	\dataset{100-200}  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\
+    \dataset{100-1000} & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\
   \end{tabular}
@@
   	\hline
-  	\jmodule{0-0}     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-  	\jmodule{10-5}    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\
-    \jmodule{20-20}   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-    \jmodule{30-45}   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\
-  	\jmodule{40-80}   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-  	\jmodule{50-125}  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-  	\jmodule{60-180}  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-  	\jmodule{70-245}  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-  	\jmodule{80-320}  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-  	\jmodule{90-405}  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-  	\jmodule{100-500} & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
+  	\dataset{0-0}     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
+  	\dataset{10-5}    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\
+    \dataset{20-20}   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
+    \dataset{30-45}   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\
+  	\dataset{40-80}   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
+  	\dataset{50-125}  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
+  	\dataset{60-180}  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
+  	\dataset{70-245}  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
+  	\dataset{80-320}  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
+  	\dataset{90-405}  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
+  	\dataset{100-500} & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
   \end{tabular}
@@
   \centering
-  	\subfigure[\jmodule{48-80}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[\jmodule{70-245}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
+  	\subfigure[\dataset{48-80}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/40-80}}
+	\subfigure[\dataset{70-245}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
 \end{figure}
@@
 
+\placeholder
+
 TODO graph + interpretation
@@
 We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.
+
+\placeholder
 
@@
 Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \jmodule{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
+Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
 There will be no improvement heuristic.
@@
 	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \jmodule{Glpk}(limit = time), IH = \emptyset}$
+    \STATE $ES \gets ES \cup {CH = \heu{Glpk}(limit = time), IH = \emptyset}$
   \ENDFOR
@@
 \end{algorithm}
+
+\placeholder
 
@@
 This experimental set will contain 220 configurations for each version.
-Every configuration will let \jmodule{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized test data set (see listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
+Every configuration will let \heu{Glpk} CH run for unlimited time, until it finds the optimum.
+This will be repeated in 20 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
 There will again be no IH, the only data we will collect are the times of the CH run in each case.
@@
 	\FOR{$file \in $ sized test data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Glpk}(no\:limit), IH = \emptyset\}$
+    \STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(no\:limit), IH = \emptyset\}$
   \ENDFOR
@@
 
+\placeholder
+
 TODO graph + interpretation. Graph should have its Y axis in log, stress this fact in the text too
@@
 
-\subsection{\jmodule{Random} vs. \jmodule{Fuzzy} vs. \jmodule{FIDAX}}
+\subsection{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX}}
 
@@
 
-Our investigation into various CH will start by comparing \jmodule{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \jmodule{Random} and \jmodule{Fuzzy}.
+Our investigation into various CH will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \heu{Random} and \heu{Fuzzy}.
 
@@
 There will be no improvement heuristics.
-The pool size will be set to 10, even though \jmodule{FIDAX} cannot not profit from this.
+The pool size will be set to 10, even though \heu{FIDAX} cannot not profit from this.
 Listing for this can be found in \ref{listing-experiment-random-fuzzy-fidax}.
@@
 	\FOR{$file \in $ official test data}
-    \STATE $ES \gets ES \cup \{file, CH = \jmodule{Random}(pool = 10), IH = \emptyset\} \cup \{file, CH = \jmodule{Fuzzy}(pool = 10), IH = \emptyset\} \cup \{file, CH = \jmodule{FIDAX}, IH = \emptyset\}$
+    \STATE $ES \gets ES \cup \{file, CH = \heu{Random}(pool = 10), IH = \emptyset\} \cup \{file, CH = \heu{Fuzzy}(pool = 10), IH = \emptyset\} \cup \{file, CH = \heu{FIDAX}, IH = \emptyset\}$
   \ENDFOR
@@
 
+\placeholder
+
 TODO graphs (files with similar behavior together) + interpretation.
 
-\subsubsection{Improving \jmodule{FIDAX} with \jmodule{Hungry}}
+\subsubsection{Improving \heu{FIDAX} with \heu{Hungry}}
 
@@
 
-There is a minor question to be answered easily: is \jmodule{FIDAX} ``hungry enough"? Could we improve it using \jmodule{Hungry} as IH? Let us design a short experiment.
+There is a minor question to be answered easily: is \heu{FIDAX} ``hungry enough"? Could we improve it using \heu{Hungry} as IH? Let us design a short experiment.
 
@@
 
-We won't be needing pool bigger than one, nor more iterations - both \jmodule{FIDAX} and \jmodule{Hungry} are deterministic.
-We will try all official data sets, first with empty IH, second with \jmodule{Hungry} as IH.
+We won't be needing pool bigger than one, nor more iterations - both \heu{FIDAX} and \heu{Hungry} are deterministic.
+We will try all official data sets, first with empty IH, second with \heu{Hungry} as IH.
 We will gather the qualities in each case and see whether there is any improvement.
@@
 The experimental results are summarized in the Table \ref{table-experiments-fidax-and-hungry} and are quite surprising.
-As trivial a heuristic \jmodule{Hungry} is, it is still able to improve the ID set found by \jmodule{FIDAX} by as much as almost 50\% (the last row, \jmodule{100-1000}).
+As trivial a heuristic \heu{Hungry} is, it is still able to improve the ID set found by \heu{FIDAX} by as much as almost 50\% (the last row, \dataset{100-1000}).
 
-Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \jmodule{100-1000}.
+Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \dataset{100-1000}.
 Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute".
@@
 \begin{table}
-  \caption{Results of adding \jmodule{Hungry} after \jmodule{FIDAX}}
+  \caption{Results of adding \heu{Hungry} after \heu{FIDAX}}
   \bigskip
@@
   \begin{tabular}{l | l | l}
-    Data set & Quality - \jmodule{FIDAX} & Quality - \jmodule{FIDAX} + \jmodule{Hungry} \\
+    Data set & Quality - \heu{FIDAX} & Quality - \heu{FIDAX} + \heu{Hungry} \\
     \hline
-    \jmodule{OVA1}     & 0.4411764705882353  & 0.4411764705882353   \\
-    \jmodule{OVA2}     & 0.16346153846153846 & 0.16346153846153846  \\
-    \jmodule{OVA3}     & 0.25482414123443264 & \textbf{0.2553715615163541}   \\
-    \jmodule{XMA-c}    & 0.7546666666666666	 & 0.7546666666666666   \\
-    \jmodule{XMA-p}    & 0.2019306150568969	 & 0.2019306150568969   \\
-    \jmodule{XMD}      & 0.09786094165493509 & 0.09786094165493509  \\
-    \jmodule{MSH}      & 0.5416472778036296	 & 0.5416472778036296   \\
-    \jmodule{NTH}      & 0.05259709474828076 & \textbf{0.057918595422124436} \\
-    \jmodule{100-100}  & 0.56	               & \textbf{0.6766666666666669}   \\
-    \jmodule{100-200}  & 0.44200000000000017 & \textbf{0.5980000000000003}   \\
-    \jmodule{100-1000} & 0.19952380952380955 & \textbf{0.29619047619047617}  \\
+    \dataset{OVA1}     & 0.4411764705882353  & 0.4411764705882353   \\
+    \dataset{OVA2}     & 0.16346153846153846 & 0.16346153846153846  \\
+    \dataset{OVA3}     & 0.25482414123443264 & \textbf{0.2553715615163541}   \\
+    \dataset{XMA-c}    & 0.7546666666666666	 & 0.7546666666666666   \\
+    \dataset{XMA-p}    & 0.2019306150568969	 & 0.2019306150568969   \\
+    \dataset{XMD}      & 0.09786094165493509 & 0.09786094165493509  \\
+    \dataset{MSH}      & 0.5416472778036296	 & 0.5416472778036296   \\
+    \dataset{NTH}      & 0.05259709474828076 & \textbf{0.057918595422124436} \\
+    \dataset{100-100}  & 0.56	               & \textbf{0.6766666666666669}   \\
+    \dataset{100-200}  & 0.44200000000000017 & \textbf{0.5980000000000003}   \\
+    \dataset{100-1000} & 0.19952380952380955 & \textbf{0.29619047619047617}  \\
   \end{tabular}
@@
 \begin{table}
-  \caption{ID sets in \jmodule{FIDAX} versus \jmodule{FIDAX} + \jmodule{Hungry}}
+  \caption{ID sets in \heu{FIDAX} versus \heu{FIDAX} + \heu{Hungry}}
   \bigskip
@@
   \begin{tabular}{l | l}
-  \jmodule{FIDAX} & \jmodule{FIDAX} + \jmodule{Hungry} \\
+  \heu{FIDAX} & \heu{FIDAX} + \heu{Hungry} \\
   \hline
@@
 
+\placeholder
+
+\placeholder
+
+\placeholder
+
 \subsection{Best IH for the best CH}
@@
 
-\subsubsection{\jmodule{Random} as CH}
+\subsubsection{\heu{Random} as CH}
 

### 2011-08-17 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 \subsection{Occurrence Limited to One}
-In opposite to the multiple occurrence, numerous XQuery constructs limit number of occurrences of an element to at most one or exactly one. Some example queries with description follows.
+In opposite to the multiple occurrence, numerous XQuery constructs limit number of occurrences of an element to at most one or exactly one. Some example queries with description follow.
 
@@
 Function \texttt{exactly-one} works similar to the function \texttt{zero-or-one} but accepts only sequences of exactly one item (which are in XQuery equal to this item itself).
+
+\section{Element and Attribute Types}
+
+\subsection{XML Schema Built-in Types}
+XML Schema built-in types of elements and attributes can be inferred from the XML documents by analysing their content.
+Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise especially when a large enough set of XML documents is not available.
+\todo[inline]{priklad?} This is when a set of XQuery queries may be handy.
+
+If a value of an element or attribute is used in an expression then this expression could be often exploited to determine a type of the value.
+Comparing the value to another value of a known type and supplying the value to a function call as an argument of particular type are examples of these expressions as shown below.
+
+\begin{verbatim}
+\\event\date = current-date()
+
+\catalog\product\price < 24.5
+
+declare function local:byteFunction($arg as xs:byte) as xs:byte
+{...};
+\catalog\product\local:byteFunction(@id)
+\end{verbatim}
+\todo[inline]{overit, ci su priklady OK}
+
+Determined types are \texttt{xs:date} for element \texttt{date}, \texttt{xs:decimal} for element \texttt{price} and \texttt{xs:byte} for attribute \texttt{id}.
+
+Alongside common expressions, there are other XQuery constructs indicating types as \emph{type casting}, \emph{constructors} and so called \emph{type declarations}.
+
+Assuming variable \texttt{\$var} bound to a value of some element or attribute, following two fractions of queries indicate its type to be \emph{xs:integer}.
+
+\begin{verbatim}
+$var cast as xs:integer
+xs:integer($var)
+\end{verbatim}
+
+The latter one is usage of \emph{type casting}, the former one is \emph{type constructor}.
+
+Alike, following two examples of \emph{type declarations}. Value of element \texttt{number} is declared to be of type \texttt{xs:integer}.
+
+\begin{verbatim}
+every $number as element(*,xs:integer) in //number satisfies ($number > 0)
+
+declare variable $firstNumber as xs:integer 
+:= data(//product/number[1]); 
+\end{verbatim}
+
+Although, determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues.
+More complex queries are commonly branched by \emph{if-then-else} construct.
+These branches can contain several contradictionary fragments from the view of type inference.
+Therefore, the reliable method would take the query structure into consideration and futher research would be needed.
+
+\subsection{Enumeration}
+In several cases, XQuery construct \emph{if-then-else} may be used to branch the execution of query by all posible values of a certain variable.
+This is equivalent of \emph{switch} construct from other programming languages.
+When the control variable is bound to some element or attribute, type of this node can be inferred as enumeration and its individual values can be determined as well.
+
+\begin{verbatim}
+Query
+let $cat := doc("catalog.xml")/catalog 
+for $dept in distinct-values($cat/product/@dept) 
+return <li>Department: {if ($dept = "ACC") 
+                        then "Accessories" 
+                        else if ($dept = "MEN") 
+                             then "Menswear" 
+                             else if ($dept = "WMN") 
+                                  then "Womens" 
+                                  else () 
+  } ({$dept})</li>
+
+Results 
+<li>Department: Womens (WMN)</li> 
+<li>Department: Accessories (ACC)</li> 
+<li>Department: Menswear (MEN)</li> 
+\end{verbatim}
+
+Finding such patterns in queries could be useful in combination with analysis of XML data this queries are executed upon.
+XML data could help to confirm or disprove this assumption based on the query analysis of vice-versa.

### 2011-08-17 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
-\subsubsection{Linux Machine}
+\subsubsection{Quad Core Machine}
 
@@
 8 GB DDR2 RAM
-x86_64 GNU/Linux 2.6.32-33-generic Ubuntu 10.04
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 64-Bit Server VM (build 20.1-b02)
-GLPK version 4.38
+Windows 7 SP1 64bit
+TODO
+TODO
+GLPK version TODO
 \end{verbatim}
 
-\subsubsection{Windows Machine}
+\subsubsection{Dual Core Machine}
 
@@
   \hline
-  Machine           & Windows \\
+  Machine           & Dual Core \\
   Input data        & \texttt{100-500} \\
@@
   \hline
-  Machine           & Windows \\
+  Machine           & Dual Core \\
   Input data        & all sized test data sets \\
@@
   \hline
-  Machine           & Windows \\
+  Machine           & Dual Core \\
   Input data set    & all official test data sets \\

### 2011-08-18 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 \subsection{Enumeration}
-In several cases, XQuery construct \emph{if-then-else} may be used to branch the execution of query by all posible values of a certain variable.
+In several cases, XQuery \emph{if-then-else} construct may be used to branch the execution of query by all posible values of a certain variable.
 This is equivalent of \emph{switch} construct from other programming languages.
@@
 Finding such patterns in queries could be useful in combination with analysis of XML data this queries are executed upon.
-XML data could help to confirm or disprove this assumption based on the query analysis of vice-versa.
+XML data could help to confirm or disprove this assumption based on the query analysis or vice-versa.
+
+\section{Keys}
+
+\subsection{Approach from Paper Discovering XML Keys and Foreign Keys in Queries}
+Paper \todo{link} introduces a method of discovering keys and foreign keys by investigation of joins in queries.
+Basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form.
+Therefore joins with the same semantics written in different syntax are not taken into consideration.
+
+The following examples are queries that could be processed by a similar method but the actual method will not use them.
+
+\begin{verbatim}
+<result>
+  {
+    for $u in doc("users.xml")//user_tuple
+    for $i in doc("items.xml")//item_tuple
+    where $u/rating > "C" 
+       and $i/reserve_price > 1000 
+       and $i/offered_by = $u/userid
+    return
+        <warning>
+            { $u/name }
+            { $u/rating }
+            { $i/description }
+            { $i/reserve_price }
+        </warning>
+  }
+</result>
+\end{verbatim}
+
+Clause \texttt{where} is used for the join condition \texttt{\$i/offered\_by = \$u/userid instead} of the join condition in predicate of the second \texttt{for} expression.
+
+\begin{verbatim}
+<result>
+  {
+    for $i in doc("items.xml")//item_tuple
+    where empty(doc("bids.xml")//bid_tuple[itemno = $i/itemno])
+    return
+        <no_bid_item>
+            { $i/itemno }
+            { $i/description }
+        </no_bid_item>
+  }
+</result>
+\end{verbatim}
+
+This query illustrates a join where the value of one of the joined elements is not required, the knowledge of its existence is sufficient.
+Therefore the query has not to contain the second \texttt{for} or \texttt{let} keyword and its expression can be moved to \texttt{where} clause.
+
+\begin{verbatim}
+for $item in doc("order.xml")//item, 
+    $product in doc("catalog.xml")//product, 
+    $price in doc("prices.xml")//prices/priceList/prod 
+where $item/@num = $product/number and $product/number = $price/@num 
+return <item num="{$item/@num}" 
+         name="{$product/name}" 
+         price="{$price/price}"/>
+\end{verbatim}
+
+A common three-way join.
+
+\subsection{Join of Recursive Data}
+By the term "recursive data" are meant XML data, for example elements, that somehow reference items from the same set. Example of query that operates upon such data follows.
+
+\begin{verbatim}
+declare function local:one_level($p as element()) as element()
+{
+  <part partid="{ $p/@partid }" name="{ $p/@name }" >
+    {
+      for $s in doc("partlist.xml")//part
+      where $s/@partof = $p/@partid
+      return local:one_level($s)
+    }
+  </part>
+};
+
+<parttree>
+  {
+    for $p in doc("partlist.xml")//part[empty(@partof)]
+    return local:one_level($p)
+  }
+</parttree>
+\end{verbatim}
+
+Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}.
+Elements with not specified attribute \texttt{partof} are at the top of the recursive hierarchy while in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
+
+By using similar approach as described in the paper \todo{ref} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.
+
+\subsection{Negative Statements about Uniqueness}
+In many cases it could be inferred a statement that refuses uniqueness of element's or attribute's values.
+Such statements may prove helpful in combination with other methods in its process of decision making whether a particular element or attribute is unique or not.
+
+Basic representant is a common \emph{FLWOR} query.
+
+\begin{verbatim}
+<bib>
+  {
+    for $b in doc("http://bstore.example.com/bib.xml")//book
+    where $b/publisher = "abcde" and $b/@year > 2000
+    order by $b/title
+    return
+        <book>
+            { $b/@year }
+            { $b/title }
+        </book>
+  }
+</bib>
+\end{verbatim}
+
+Usage of \texttt{for} construct indicates that it is expected a sequence of \texttt{book} elements which satisfy the condition if \emph{where} clause.
+It is a condition composed of two single conditions joined by \emph{and} logical operator.
+Therefore in order to satisfy the whole condition, both two single conditions must be satisfied as well.
+Thus, it is expected that several elements satisfy each of the single conditions.
+
+The first of them is test of equality of book's subelement \texttt{publisher} to a string literal.
+Based on the expectation, element \texttt{publisher} cannot be unique.
+However, the second condition is greater-than comparison of \texttt{year} attribute to an integer literal, it cannot be inferred whether this attribute is unique or not.
+It is because even if it was unique there still might be more than one \texttt{book} element meeting this condition, thus the expectation was fullfiled too.
+Also, any statement, positive nor negative, cannot be inferred about \texttt{title} subelement.
+
+Another simple examples are passing a result of basic path expression to a call of \texttt{distinct-values()} function and usage of \emph{aggregation} functions.
+
+\begin{verbatim}
+<results>
+  {
+    let $doc := doc("prices.xml")
+    for $t in distinct-values($doc//book/title)
+    let $p := $doc//book[title = $t]/price
+    return
+      <minprice title="{ $t }">
+        <price>{ min($p) }</price>
+      </minprice>
+  }
+</results>
+\end{verbatim}
+
+According to the use of \texttt{distinct-values()} function, it can be easily seen that writter of this query assumes possible occurrence of more than one \texttt{book} elements with the same value of their \texttt{title} subelement.
+Thus, \texttt{title} element is not unique.
+
+Alike, variable \texttt{\$p} is bound to a price of each title and then passed to \texttt{min()} function call.
+That indicates that one book title is supposed to have several prices, however, it cannot be said if a certain occurrence of element \emph{book} can contain more than one \texttt{price} subelement.
+
+Also, many other types of queries could be exploited to obtain negative uniqueness statemens like occurrence of element or attribute in \texttt{stable order by} clause or selection of an element set based on a value of particular attribute and consecutive treatment of this set as a sequence of elements.
+
+\begin{verbatim}
+let $prods := doc("catalog.xml")//product 
+for $prod in $prods 
+where $prod << $prods[@dept = $prod/@dept][last( )] 
+return $prod 
+\end{verbatim}
+
+\subsection{Uniqueness}
+TODO

### 2011-08-18 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
   Input data        & all official and sized test data sets \\
-  Iterations        & 20 \\
+  Iterations        & 50 \\
   Pool size         & not applicable \\
@@
 
-The experimental set will contain $20 * (11 + 11) = 440$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
+The experimental set will contain $50 * (11 + 11) = 1100$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
 There will be no CHs or IHs.
@@
 
-\placeholder
-
-TODO graph + interpretation
+TODO just table (mean, variance), no chart + interpretation
 
@@
   Machine           & Dual Core \\
-  Input data        & \texttt{100-500} \\
-  Iterations        & 20 \\
+  Input data        & \dataset{100-500} \\
+  Iterations        & 50 \\
   Pool size         & 1 \\
@@
 Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 49 seconds with increments of 2, meaning 25 settings * 20 iterations = 500 configurations in total (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
+Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
 There will be no improvement heuristic.
@@
 
+TODO try this on XMA-p
+
 \subsection{Chaining the IHs}

### 2011-08-19 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 \subsection{Uniqueness}
-TODO
+In opposite of non-uniqueness discovery, some XQuery constructs can indicate uniqueness of elements or attributes, however, it seems to be more difficult.
+One of approaches of uniqueness discovery could be an investigation of what does the query return.
+Consider the following query.
+
+\begin{verbatim}
+for $product in /catalog/product
+let $number := $product/number
+return <prod xml:id="{concat('prod', number)}"/>
+\end{verbatim}
+
+For each \texttt{product} element, new element \texttt{prod} with attribute \texttt{xml:id} is created.
+Since attribute \texttt{xml:id} is supposed to be unique and there is direct transformation of the values of \texttt{number} elements to the values of attribute \texttt{xml:id}, it is very likely that element \texttt{number} is unique in the source data.
+
+\section{Other}
+
+\subsection{XML Schema \texttt{xs:sequence} and \texttt{xs:all} constructs}
+If an order of appearance of some element set, for example subelements of a certain element, is important this is can be expressed by XML Schema construct \texttt{xs:sequence}.
+On the other hand, \texttt{xs:all} is involved when the elements may occur in any order.
+
+If there is a large enough set of XML documents available it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs.
+However, if the inference is made using a smaller set of XML documents, it may happen that every occurrence of some element set is in the same order but the input data are not representative enough to be sure.
+There comes in handy a comparison of element order in queries, which could help to decide for the use of \texttt{xs:sequence}.
+
+Every two elements in a XML document that are not siblings with the same name and none of them is the outermost element have the nearest common ancestor determining their relative order.
+If there is a lack of evidence to choose between the two constructs in the ancestor queries could be searched for an order comparison of these two elements like in the following query.
+
+\begin{verbatim}
+let $i := //incision)[2]
+for $a in //action)[. >> $i]
+return $a//instrument
+\end{verbatim}
+
+Apparently, the relative order of elements \texttt{incistion} and \texttt{action} is somehow important because the query composes the result using only those \texttt{action} elements that succeed the second \texttt{incision} element.
+Therefore, in a respective type definition part of the their nearest common ancestor it should be used \texttt{xs:sequence} prefered to \texttt{xs:all}.
+
+\subsection{Intermediate XML structure}
+Intermediate XML structure are XML data that are not read from input XML documents nor created as an output of queries but they are used by the queries somehow.
+Objective of this section is intermediate XML structure created directly in the queries.
+It may serve to various purposes a can be created in various ways, hard-coded in queries, computed from the input XML documents to simplify their structure.
+The former is demostrated in the following example.
+
+\begin{verbatim}
+let $deptNames := <deptNames> 
+                    <dept code="ACC" name="Accessories"/> 
+                    <dept code="MEN" name="Menswear"/> 
+                    <dept code="WMN" name="Womens"/> 
+                  </deptNames> 
+let $catalog := doc("catalog.xml")/catalog 
+for $dept in distinct-values($catalog/product/@dept) 
+return <li>Department:
+         {data($deptNames/dept[@code = $dept]/@name)} ({$dept})
+       </li> 
+\end{verbatim}
+
+Intermediate XML structure in this query could be used to identify possible values of attribute \texttt{dept} in \texttt{product} element and thereby determine its type as enumeration.
+
+That is a very simple example of intermediate structure. Since there many admittedly many means of utilization of intermediate structure, further reasearch would be needed.

### 2011-08-19 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
-\placeholder
+\begin{figure}
+  \caption{Time vs. Quality}
+  \label{image-experiment-time-vs-quality}
+  \centering
+    \includegraphics[angle=-90, width=\textwidth]{images/experiments/time-vs-quality}
+\end{figure}
 
-TODO graph + interpretation
+TODO interpretation
 
@@
 
-\placeholder
+\begin{figure}
+  \caption{Time vs. Quality}
+  \label{image-experiment-time-vs-quality}
+  \centering
+    \includegraphics[angle=-90, width=\textwidth]{images/experiments/time-till-optimum}
+\end{figure}
 
-TODO graph + interpretation. Graph should have its Y axis in log, stress this fact in the text too
+TODO interpretation. Graph should has its Y axis in log, stress this fact in the text too
 

### 2011-08-20 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 \section{Structure of XML documents}
-The most queries can be exploited to obtain some information about a structure of respective XML documents.
+Most of queries can be exploited to obtain some information about a structure of respective XML documents.
 The structure of XML documents captures elements from these documents along with their names, attributes and their organization.
-What are the outermost elements in these document, which elements can be contained inside a certain element, whether are they optional or mandatory, etc.
+What the outermost elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.
 
 Path expressions without predicates which use only child axis are the simpliest example of such queries.
-Path expression \texttt{/bib/book/author} indicates element \texttt{bib} is the outermost element and contains one or more elements \texttt{book} and these contains one or more elements \texttt{author}.
+Path expression \texttt{/bib/book/author} indicates that element \texttt{bib} is the outermost element and it contains one or more elements \texttt{book} and these contain one or more elements \texttt{author}.
 
@@
 
-Besides elements, attributes could be processed exactly the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} has attribute \texttt{@ISBN}.
+Besides elements, attributes could be processed exactly in the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} has attribute \texttt{ISBN}.
 
-However statemets of this kind are necessary for the XML schema inference, their obtaining from queries it not significant because of the following reason.
-These statement could be determined directly from the XML documents and it could be easily done in a more convenient way.
+However, statements of this kind are necessary for the XML schema inference, their obtaining from queries is not significant due to the following reason.
+These statements could be determined directly from the XML documents and it could be easily done in a more convenient way.
 Queries may not cover the whole relevant content of the documents.
-For example some element may not be queried at all, therefore they mat not be mentioned in any query.
-In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurence of elements.
+For example some element may not be queried at all, therefore they may not be mentioned in any query.
+In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurrence of elements.
 At last, we cannot be sure that queries target nodes actually presented in the XML documents.
 Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not.
-In opposite, even basic methods of XML documents analyzis do not have these inadequacies.
+In opposite, even basic methods of XML documents analysis do not have these inadequacies.
 
-On the other hand, the structure inference could be useful when the whole set of all XML documents is not available and provided XML documents do not cover the structure completly.
+On the other hand, the structure inference could be useful when the entire set of all XML documents is not available and provided XML documents do not cover the structure completely.
 
@@
 
-Apparently, this query expects more than one element \texttt{author} to be children of the element variable \texttt{\$book1} is bound to.
+Apparently, this query expects more than one \texttt{author} element to be a child of the element the variable \texttt{\$book1} is bound to.
 Otherwise, any sorting would lack a reason.
-Although, we cannot be absolutely sure about it again, assuming common sense usage of XQuery it is very likely that element \texttt{author} can occur multiple times as a subelement of element from variable \texttt{\$book1}.
+Although we cannot be absolutely sure about it again, assuming common-sense usage of XQuery, it is very likely that \texttt{author} element can occur multiple times as a subelement of element from variable \texttt{\$book1}.
 
 \subsection{Multiple Occurrence}
-Similar approach could be applied in many other situations.
+A similar approach could be applied in many other situations.
 Another examples are particular usages of function \texttt{count()}, indexation, usage of set operators (\texttt{union}, \texttt{intersect}, \texttt{except}) and usage of function \texttt{one-or-more()}.
-Queries with description follows.
+Queries with description follow.
 
@@
 Function \texttt{count()} returns number of items in a provided sequence.
-If the sequence is sequence of elements the count of these elements will probably not be limited to one.
+If the sequence is a sequence of elements the count of these elements will probably not be limited to one.
 The example query indicates that the outermost element \texttt{book} can contain more than one element \texttt{section}.
@@
 In this query, number of elements \texttt{color} in element \texttt{product} with attribute \texttt{id} equal to 5 has to be at least one otherwise an error is raised upon execution.
-If we assume that this query is written correctly, with common sense and should not raise the error, we can infer that element \texttt{product} has to containt one or more elements \texttt{color}
+If we assume that this query is written correctly, with common sense and it should not raise the error, we can infer that element \texttt{product} has to contain one or more elements \texttt{color}.
 
@@
 \texttt{lt} is a representant of so called \emph{value comparison operators} which operate two sequences of zero or one item.
-If an operand of a \emph{value comparison operand} is sequence of more than one item \emph{type error} is raised.
+If an operand of a \emph{value comparison operator} is a sequence of more than one item then \emph{type error} is raised.
 
@@
 Another similar examples are arithmetic expressions and functions accepting a sequence of at most one item.
-Functions \texttt{zero-or-one()} will raise the \emph{type error} when supplied with an sequence of more than one item.
+Function \texttt{zero-or-one()} will raise the \emph{type error} when supplied with a sequence of more than one item.
 
 Those are constructs indicating limitation to zero or one occurrence.
-Function \texttt{exactly-one} works similar to the function \texttt{zero-or-one} but accepts only sequences of exactly one item (which are in XQuery equal to this item itself).
+Function \texttt{exactly-one()} works similar to the function \texttt{zero-or-one()} but accepts only sequences of exactly one item (which are in XQuery equal to this item itself).
 
@@
 Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise especially when a large enough set of XML documents is not available.
-\todo[inline]{priklad?} This is when a set of XQuery queries may be handy.
+\todo[inline]{priklad?} This is when a set of XQuery queries may come in handy.
 
@@
 
-Although, determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues.
+Although determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues.
 More complex queries are commonly branched by \emph{if-then-else} construct.
 These branches can contain several contradictionary fragments from the view of type inference.
-Therefore, the reliable method would take the query structure into consideration and futher research would be needed.
+Therefore, the reliable method would take the query structure into consideration and further research would be needed.
 
 \subsection{Enumeration}
-In several cases, XQuery \emph{if-then-else} construct may be used to branch the execution of query by all posible values of a certain variable.
+In several cases, XQuery \emph{if-then-else} construct may be used to branch the execution of query by all possible values of a certain variable.
 This is equivalent of \emph{switch} construct from other programming languages.
@@
 
-Finding such patterns in queries could be useful in combination with analysis of XML data this queries are executed upon.
+Finding such patterns in queries could be useful in combination with analysis of XML data these queries are executed upon.
 XML data could help to confirm or disprove this assumption based on the query analysis or vice-versa.
@@
 
-This query illustrates a join where the value of one of the joined elements is not required, the knowledge of its existence is sufficient.
+This query illustrates a join where the value of one of the joined elements is not required thus the knowledge of its existence is sufficient.
 Therefore the query has not to contain the second \texttt{for} or \texttt{let} keyword and its expression can be moved to \texttt{where} clause.
@@
 \subsection{Join of Recursive Data}
-By the term "recursive data" are meant XML data, for example elements, that somehow reference items from the same set. Example of query that operates upon such data follows.
+By the term "recursive data" are meant XML data, for example elements that somehow reference items from the same set. Example of a query that operates upon such data follows.
 
@@
 Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}.
-Elements with not specified attribute \texttt{partof} are at the top of the recursive hierarchy while in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
+Elements with not specified attribute \texttt{partof} are at the top of the recursive hierarchy while, in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
 
-By using similar approach as described in the paper \todo{ref} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.
+By using a similar approach as described in the paper \todo{ref} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.
 
 \subsection{Negative Statements about Uniqueness}
-In many cases it could be inferred a statement that refuses uniqueness of element's or attribute's values.
+In many cases a statement refusing uniqueness of element's or attribute's values could be inferred.
 Such statements may prove helpful in combination with other methods in its process of decision making whether a particular element or attribute is unique or not.
@@
 It is a condition composed of two single conditions joined by \emph{and} logical operator.
-Therefore in order to satisfy the whole condition, both two single conditions must be satisfied as well.
+Therefore in order to satisfy the whole condition, both single conditions must be satisfied as well.
 Thus, it is expected that several elements satisfy each of the single conditions.
@@
 However, the second condition is greater-than comparison of \texttt{year} attribute to an integer literal, it cannot be inferred whether this attribute is unique or not.
-It is because even if it was unique there still might be more than one \texttt{book} element meeting this condition, thus the expectation was fullfiled too.
+The reason is that even if it was unique there still might be more than one \texttt{book} element meeting this condition.
 Also, any statement, positive nor negative, cannot be inferred about \texttt{title} subelement.
@@
 
-Also, many other types of queries could be exploited to obtain negative uniqueness statemens like occurrence of element or attribute in \texttt{stable order by} clause or selection of an element set based on a value of particular attribute and consecutive treatment of this set as a sequence of elements.
+Also, many other types of queries could be exploited to obtain negative uniqueness statements like occurrence of element or attribute in \texttt{stable order by} clause or selection of an element set based on a value of particular attribute and consecutive treatment of this set as a sequence of elements.
 
@@
 \subsection{Uniqueness}
-In opposite of non-uniqueness discovery, some XQuery constructs can indicate uniqueness of elements or attributes, however, it seems to be more difficult.
-One of approaches of uniqueness discovery could be an investigation of what does the query return.
+In opposite to non-uniqueness discovery, some XQuery constructs can indicate uniqueness of elements or attributes, however, it seems to be more difficult.
+One of approaches of uniqueness discovery could be an investigation of what the query does return.
 Consider the following query.
@@
 \subsection{XML Schema \texttt{xs:sequence} and \texttt{xs:all} constructs}
-If an order of appearance of some element set, for example subelements of a certain element, is important this is can be expressed by XML Schema construct \texttt{xs:sequence}.
+If an order of appearance of some element set, for example subelements of a certain element, is important this can be expressed by XML Schema construct \texttt{xs:sequence}.
 On the other hand, \texttt{xs:all} is involved when the elements may occur in any order.
@@
 Every two elements in a XML document that are not siblings with the same name and none of them is the outermost element have the nearest common ancestor determining their relative order.
-If there is a lack of evidence to choose between the two constructs in the ancestor queries could be searched for an order comparison of these two elements like in the following query.
+If there is a lack of evidence to choose between the two constructs in the ancestor than available queries could be searched for an order comparison of these two elements like in the following query.
 
@@
 
-Apparently, the relative order of elements \texttt{incistion} and \texttt{action} is somehow important because the query composes the result using only those \texttt{action} elements that succeed the second \texttt{incision} element.
-Therefore, in a respective type definition part of the their nearest common ancestor it should be used \texttt{xs:sequence} prefered to \texttt{xs:all}.
+Apparently, the relative order of elements \texttt{incision} and \texttt{action} is somehow important because the query composes the result using only those \texttt{action} elements that succeed the second \texttt{incision} element.
+Therefore, in a respective type definition part of their nearest common ancestor \texttt{xs:sequence} should be used in favour of \texttt{xs:all}.
 
 \subsection{Intermediate XML structure}
-Intermediate XML structure are XML data that are not read from input XML documents nor created as an output of queries but they are used by the queries somehow.
+Intermediate XML structure are XML data that are not read from input XML documents nor created as an output of queries but they are somehow used by the queries.
 Objective of this section is intermediate XML structure created directly in the queries.
-It may serve to various purposes a can be created in various ways, hard-coded in queries, computed from the input XML documents to simplify their structure.
-The former is demostrated in the following example.
+It may serve to various purposes and it can be created in various ways, hard-coded in queries, computed from the input XML documents to simplify their structure.
+The former is demonstrated in the following example.
 
@@
 
-That is a very simple example of intermediate structure. Since there many admittedly many means of utilization of intermediate structure, further reasearch would be needed.
+That is a very simple example of intermediate structure. Since there are admittedly many means of utilization of intermediate structure, further research would be needed.

### 2011-08-20 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 It ranks each DTD candidate using two aspects - conciseness and preciseness.
-Conciseness of a DTD is expressed using the number of bits required to describe the DTD (the smaller, the better).
+Conciseness of a DTD is expressed by using the number of bits required to describe the DTD (the smaller, the better).
 Preciseness of a DTD is expressed using the number of bits required for description of the input data using the DTD.
@@
 The SchemaMiner system \todo{bibref} is an example of \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
-It focusses on inferring elements with the same name but different structure and unordered sequences.
+It focuses on inferring elements with the same name but different structure and unordered sequences.
 
@@
 \section{Discovering XML Keys and Foreign Keys in Queries}
-This methods tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
+This method tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
 Just these queries are utilized in inference, no XML data are used.
@@
 \begin{define}[Key]
-A key is a construct $$(C, P, \{L\})$$ where $C$, $P$, $L$ are XPath paths without predicates thet use only \texttt{child} and \texttt{descendant} axes.
-$C$ can be omitted. $(P, \{L\})$ is then equivalent to $(/, P, \{L\})$ and is called \emph{global key}.Otherwise it is called \emph{local key}.
+A key is a construct $$(C, P, \{L\})$$ where $C$, $P$, $L$ are XPath paths without predicates that use only \texttt{child} and \texttt{descendant} axes.
+$C$ can be omitted. $(P, \{L\})$ is then equivalent to $(/, P, \{L\})$ and is called \emph{global key}. Otherwise it is called \emph{local key}.
 \end{define}
@@
 \begin{define}[Foreign key]
-A foreign key is a construct $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ where $(C, P_2, \{L_2\})$ is a key, $P_1$, $L_1$ are XPath paths without predicates thet use only \texttt{child} and \texttt{descendant} axes.
+A foreign key is a construct $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ where $(C, P_2, \{L_2\})$ is a key, $P_1$, $L_1$ are XPath paths without predicates that use only \texttt{child} and \texttt{descendant} axes.
 $C$ can be omitted as in the case of keys.
@@
 \subsubsection{Joins in queries}
-To discover keys and foreign keys, method utilizes element/element joins.
+To discover keys and foreign keys, the described method utilizes element/element joins.
 
@@
 The query is searched for so-called \emph{join patters}.
-These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and look like the following (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
+These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and look as following (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
 \todo[inline]{spravne kreslit indexy pri pismenkach}
@@
 \section{Conclusion}
-The output of our method is a list of scored keys and for each key a list foreign keys referencing the key.
+The output of the method is a list of scored keys and for each key a list foreign keys referencing the key.
 The score of a key can be negative or positive.

### 2011-08-22 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
+TODO explain boxplots, how to read them
+
 \section{Experimental Results}
@@
   Input data set    & all official test data sets \\
-  Iterations        & 20 \\
+  Iterations        & 50 \\
   Pool size         & 10 \\

### 2011-08-25 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
   \hline
-  Machine           & TODO \\
+  Machine           & Dual Core \\
   Input data        & all official and sized test data sets \\
@@
   \hline
-  Machine           & TODO \\
+  Machine           & Dual Core \\
   Input data set    & all official test data sets \\
@@
   \hline
-  Machine           & TODO \\
+  Machine           & Quad Core \\
   Input data        & all official test data sets \\
@@
 
-TODO with the best construction heuristic, which IH is the best? Note that this needn't be the best combination!
+%         NB class BestIHForGlpk
+
+TODO with the best construction heuristic (GLPK), which IH is the best? Note that this needn't be the best combination!
 

### 2011-08-27 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
-TODO just table (mean, variance), no chart + interpretation
+\begin{table}
+  \caption{Grammar and Model Creation Times}
+  \bigskip
+  \label{table-experiments-grammar-model-timing}
+  \centering
+  \begin{tabular}{l || l | l || l | l || l | l}
+    Data set & GE avg [ms] & GE stdev & MC avg [ms] & MC stdev & Tot avg [ms] & Tot stdev \\
+    \hline
+    \dataset{OVA1}     & < 10 & - & < 10 & - < 10 & - \\
+    \dataset{OVA2}     & < 10 & - & < 10 & - < 10 & - \\
+    \dataset{OVA3}     & 42.94 & 19.8509 & 60.92 & 27.0848 & 103.86	& 31.6911 \\
+    \dataset{XMA-c}    & 140.32 &	33.2618 &	90.24 &	45.8803 & 230.56 & 56.2633 \\
+	\dataset{XMA-p}    & 7518.82 &	922.8882 &	10135.46 &	502.8997 & 17654.28	& 1353.8794 \\
+	\dataset{XMD}      & 979.18 &	307.1760 &	563.04 & 341.4697 & 1542.22	& 134.6883 \\
+	\dataset{MSH}      & 570.24 &	167.1119 &	225.48 &	90.6775 & 795.72 & 161.8340 \\ 
+	\dataset{NTH}      & 328.36 & 118.3766 &	1074.9 &	155.5604 & 1403.26 & 137.8695 \\
+	\dataset{100-100}  & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{100-200}  & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{100-1000} & 18.34 & 10.2372 & 18.84 & 1.0373 & 37.18 & 9.9338 \\
+	\dataset{0-0}      & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{10-5}     & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{20-20}    & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{30-45}    & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{40-80}    & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{50-125}   & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{60-180}   & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{70-245}   & < 10 & - & < 10 & - & < 10 & - \\
+	\dataset{80-320}   & < 10 & - & < 10 & - & 12.48 & 8.3574 \\
+	\dataset{90-405}   & < 10 & - & < 10 & - & 15.88 & 10.3778 \\
+	\dataset{100-500}  & < 10 & - & < 10 & - & 18.74 &	8.8889 \\
+  \end{tabular}
+\end{table}
+
+TODO interpretation
 
@@
 
-\placeholder
+\begin{table}
+  \caption{GLPK Interface Times}
+  \bigskip
+  \label{table-experiments-glpk-timing}
+  \centering
+  \begin{tabular}{l || l | l || l | l || l | l}
+	Data set & IC avg [ms] & IC stdev & OP avg [ms] & OP stdev & Tot avg [ms] & Tot stdev \\
+	\hline
+	\dataset{OVA1} & 36.46 & 66.8517 & 49.8 & 114.0687 & 86.26 & 150.1044 \\
+	\dataset{OVA2} & 39.52 & 75.8210 & 48.8 & 102.4484 & 88.32 & 154.9596 \\
+	\dataset{OVA3} & 34.1 & 74.1838 & 38.62 & 89.3772 & 72.72 & 134.7295 \\
+	\dataset{XMA-c} & 40.88 & 88.6632 & 33.84 & 65.8636 & 74.72 & 127.7338 \\
+	\dataset{XMA-p} & 36.54 & 70.7436 & 49.24 & 101.2412 & 85.78 & 145.2092 \\
+	\dataset{XMD} & 37.98 & 69.2719 & 32.88 & 70.2173 & 70.86 & 114.6692 \\
+	\dataset{MSH} & 40.42 & 91.9885 & 36.52 & 72.1018 & 76.94 & 138.6198 \\
+	\dataset{NTH} & 36.02 & 66.3403 & 38.06 & 88.8244 & 74.08 & 128.9974 \\
+	\dataset{100-100} & 46.5 & 103.3929 & 46.92 & 89.7049 & 93.42 & 158.7267 \\
+	\dataset{100-200} & 42.34 & 96.1204 & 38.22 & 90.0284 & 80.56 & 152.6534 \\
+	\dataset{100-1000} & 32.92 & 64.4534 & 42.1 & 89.4546 & 75.02 & 127.8541 \\
+	\dataset{0-0} & 46.8 & 123.5183 & 46.92 & 102.2601 & 93.72 & 181.5228 \\
+	\dataset{10-5} & 40.06 & 75.7370 & 40.1 & 72.4851 & 80.16 & 126.7135 \\
+	\dataset{20-20} & 33.72 & 70.7263 & 34.1 & 66.2781 & 67.82 & 116.3783 \\
+	\dataset{30-45} & 38.26 & 71.7549 & 45.94 & 110.1284 & 84.2 & 155.7594 \\
+	\dataset{40-80} & 37.06 & 67.0024 & 49.26 & 106.3185 & 86.32 & 144.9918 \\
+	\dataset{50-125} & 50.44 & 101.9162 & 84.76 & 364.7350 & 135.2 & 378.7835 \\
+	\dataset{60-180} & 38.38 & 89.3379 & 42.54 & 94.3742 & 80.92 & 149.6049 \\
+	\dataset{70-245} & 41.5 & 93.2951 & 40.3 & 93.4858 & 81.8 & 149.6797 \\
+	\dataset{80-320} & 51.92 & 121.9812 & 47.98 & 96.0904 & 99.9 & 171.4617 \\
+	\dataset{90-405} & 40.5 & 91.5373 & 36.46 & 88.5099 & 76.96 & 144.2890 \\
+	\dataset{100-500} & 37.82 & 85.7571 & 43.4 & 90.3257 & 81.22 & 141.9103 \\
+  \end{tabular}
+\end{table}
 
-TODO graph + interpretation
+TODO interpretation: it is interesting that even though $|V|$ and $|E|$ are increasing, the times remain almost the same. This probably means that the most of the time is spent in I/O.
 
@@
 
-TODO interpretation. Graph should has its Y axis in log, stress this fact in the text too
+TODO interpretation. Graph has its Y axis in log, stress this fact in the text too
 
 TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
+
+TODO pie chart of all relevant times: grammar, model, GLPK in, GLPK, GLPK out
 
@@
 
-\placeholder
+TODO repair outliers in these 2 graphs
 
-TODO graphs (files with similar behavior together) + interpretation.
+\begin{figure}
+  \caption{Random vs. Fuzzy vs. FIDAX - Quality}
+  \label{image-experiment-random-fuzzy-fidax-quality}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/random-fuzzy-fidax-quality}
+\end{figure}
+
+\begin{figure}
+  \caption{Random vs. Fuzzy vs. FIDAX - Time}
+  \label{image-experiment-random-fuzzy-fidax-time}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/random-fuzzy-fidax-time}
+\end{figure}
+
+TODO interpretation - Fuzzy consistently wins, but takes the longest time by far. Random is better than FIDAX in artificial as well as some real data.
 
@@
 
-TODO compare all CHs - limit GLPK to ~ 1 second
+TODO compare all CHs - limit GLPK to 1 second
 TODO pool size 10 where applicable
@@
   Input data        & all official test data sets \\
-  Iterations        & 20 \\
+  Iterations        & 50 \\
   Pool size         & 10 \\

### 2011-08-27 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
-\placeholder
+TODO XMA-c, XMA-p, MSH and NTH are uninteresting - every CH found the optimum every time
 
-\placeholder
-
-\placeholder
+\begin{figure}
+  \caption{Best Standalone CH}
+  \label{image-experiments-best-standalone-ch}
+  \centering
+  	\subfigure[\dataset{OVA1}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA1}}
+  	\subfigure[\dataset{OVA2}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA2}}
+  	\subfigure[\dataset{OVA3}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA3}}
+  	\subfigure[\dataset{XMD}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-XMD}}
+   	\subfigure[\dataset{100-100}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-100}}
+  	\subfigure[\dataset{100-200}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-200}}
+  	\subfigure[\dataset{100-1000}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-1000}}
+\end{figure}
 

### 2011-08-27 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
+TODO interpretation - GLPK wins all the time, takes 1 second, which is OK.
+
 \subsection{Best IH for the best CH}
@@
 
+TODO count improvement, from there count average + stdev of improvement for every (data set, IH)
+
+TODO we will ignore RandomRemove and RemoveWorst, they cannot help us
+
+\begin{table}
+  \caption{Best IH for \heu{Glpk}}
+  \bigskip
+  \label{table-experiments-best-ih-for-glpk}
+  \centering
+  \begin{tabular}{l || l | l || l | l}
+    Data set & \heu{Hungry} & \heu{Hungry} & \heu{Crossover} & \heu{Crossover} \\
+     & improvement & improvement & improvement & improvement \\
+     & avg & stdev & avg & stdev \\
+    \hline
+    \dataset{80-320} & 0.00017 & 0.00118 & 0.00017 & 0.00118 \\
+    \dataset{90-405} & 0.00502 & 0.00618 & 0.00033 & 0.00165 \\
+    \dataset{100-500} & 0.00664 & 0.00667 & 0.00016 & 0.00081 \\
+    \hline
+    \dataset{100-100} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
+    \dataset{100-200} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
+    \dataset{100-1000} & 0.01630 & 0.01294 & 0.00180 & 0.00506 \\
+  \end{tabular}    
+	% TODO separate somehow
+  \begin{tabular}{l || l | l || l | l}
+    Data set & \heu{LB} & \heu{LB} & \heu{Mutation} & \heu{Mutation} \\
+     & improvement & improvement & improvement & improvement \\
+     & avg & stdev & avg & stdev \\
+    \hline
+    \dataset{80-320} & \textbf{0.00072} & 0.00223 & 0.00064 & 0.00218 \\
+    \dataset{90-405} & 0.00698 & 0.00616 & \textbf{0.00851} & 0.00659 \\
+    \dataset{100-500} & 0.00796 & 0.00797 & \textbf{0.00964} & 0.00804 \\
+    \hline
+    \dataset{100-100} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
+    \dataset{100-200} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
+    \dataset{100-1000} & 0.01710 & 0.01188 & \textbf{0.02337} & 0.01558 \\
+  \end{tabular}
+\end{table}
+
+TODO interpretation: \heu{Mutation} wins.
+
 \subsubsection{\heu{Random} as CH}
 
-TODO with the best IH from the previous step, how about using Random as CH?
+TODO with \heu{Mutation} from the previous step, how about using Random as CH? Do we get an improvement?
 

### 2011-08-27 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 \begin{table}
-  \caption{Grammar and Model Creation Times}
+  \caption{Grammar Extraction and Model Creation Times}
   \bigskip
@@
   \begin{tabular}{l || l | l || l | l || l | l}
-    Data set & GE avg [ms] & GE stdev & MC avg [ms] & MC stdev & Tot avg [ms] & Tot stdev \\
+    Data set & GE & GE & MC & MC & Tot & Tot \\
+     & avg [ms] & stdev & avg [ms] & stdev & avg [ms] & stdev \\
     \hline
@@
   \begin{tabular}{l || l | l || l | l || l | l}
-	Data set & IC avg [ms] & IC stdev & OP avg [ms] & OP stdev & Tot avg [ms] & Tot stdev \\
+	Data set & IC & IC & OP & OP & Tot & Tot \\
+     & avg [ms] & stdev & avg [ms] & stdev & avg [ms] & stdev \\
 	\hline
@@
 \begin{figure}
-  \caption{Time vs. Quality}
-  \label{image-experiment-time-vs-quality}
+  \caption{Time Until Optimum}
+  \label{image-experiment-time-until-optimum}
   \centering
@@
 
-TODO pie chart of all relevant times: grammar, model, GLPK in, GLPK, GLPK out
+\begin{figure}
+  \caption{Timing Summary}
+  \label{image-experiment-timing-summary}
+  \centering
+    \includegraphics[width=.6\textwidth]{images/experiments/timing-pie}
+\end{figure}
+
+TODO interpret
 
@@
 
-TODO count improvement, from there count average + stdev of improvement for every (data set, IH)
+TODO we will ignore RandomRemove and RemoveWorst, they cannot help us
 
-TODO we will ignore RandomRemove and RemoveWorst, they cannot help us
+TODO we will be using only a few sets where we can observe improvement
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & Dual Core \\
+  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
+                    & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
+  Iterations        & 50 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
 
@@
 \subsubsection{\heu{Random} as CH}
+
+%         NB class CHForMutation
 

### 2011-08-27 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/CHForMutation.java

@@ -1,0 +1,77 @@
+/*
+ * Copyright (C) 2011 vektor
+ *
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+package cz.cuni.mff.ksi.jinfer.iss.experiments.sets;
+
+import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.OfficialTestData;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.SizeTestData;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.TestData;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.quality.Weight;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.termination.TimeIterations;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.Random;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.glpk.Glpk;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Mutation;
+import cz.cuni.mff.ksi.jinfer.iss.utils.Utils;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.List;
+import org.openide.util.lookup.ServiceProvider;
+
+/**
+ * This experimental set compares Random and Glpk as CHs for Mutation IH.
+ *
+ * @author vektor
+ */
+@ServiceProvider(service = ExperimentSet.class)
[…]

### 2011-08-28 jinfer/trunk/jInfer/Base/src/cz/cuni/mff/ksi/jinfer/base/utils/RunningProject.java
Thesis & Experimental framework: in progress.

@@ -1,16 +1,16 @@
 /*
  *  Copyright (C) 2010 sviro
- * 
+ *
  *  This program is free software: you can redistribute it and/or modify
  *  it under the terms of the GNU General Public License as published by
  *  the Free Software Foundation, either version 3 of the License, or
  *  (at your option) any later version.
- * 
+ *
  *  This program is distributed in the hope that it will be useful,
  *  but WITHOUT ANY WARRANTY; without even the implied warranty of
  *  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
  *  GNU General Public License for more details.
- * 
+ *
  *  You should have received a copy of the GNU General Public License
  *  along with this program.  If not, see <http://www.gnu.org/licenses/>.
  */
@@ -32,6 +32,8 @@ public final class RunningProject {
   private static Project project = null;
 
   private static Capabilities nextModuleCaps = null;
+
+  private static Properties defaultProperties = new Properties();
 
   private RunningProject() {
   }
@@ -77,16 +79,18 @@ public final class RunningProject {
   }
 
   /**
-   * Gets Properties of active project.
+   * Gets Properties of active project. If there is no project running, default
+   * properties will be returned.
    *
-   * @return propeties of active project.
+   * @return Properties of active project or default properties, if no project
+   * is running.
    */
   public static Properties getActiveProjectProps(final String moduleName) {
     if (isActiveProject()) {
       return new ModuleProperties(moduleName, project.getLookup().lookup(Properties.class));
     }
 
-    return new Properties();
+    return defaultProperties;
   }
 
   /**
@@ -115,5 +119,15 @@ public final class RunningProject {
    */
   public static void setNextModuleCaps(final Capabilities aNextModuleCaps) {
     nextModuleCaps = aNextModuleCaps;
+  }
+
+  /**
+   * Sets the default properties - those will be returned if there is no running
+   * project.
+   *
+   * @param p Properties to be set as default.
+   */
+  public static void setDefaultProperties(final Properties p) {
+    defaultProperties = p;
   }
 }

### 2011-08-28 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/IgnoreTextData.java

@@ -1,0 +1,75 @@
+/*
+ * Copyright (C) 2011 vektor
+ *
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+package cz.cuni.mff.ksi.jinfer.iss.experiments.sets;
+
+import cz.cuni.mff.ksi.jinfer.base.utils.RunningProject;
+import cz.cuni.mff.ksi.jinfer.basicigg.properties.BasicIGGPropertiesPanel;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.Experiment;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.data.OfficialTestData;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.quality.Weight;
+import cz.cuni.mff.ksi.jinfer.iss.experiments.termination.TimeIterations;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.glpk.Glpk;
+import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Identity;
+import cz.cuni.mff.ksi.jinfer.iss.utils.Utils;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.List;
+import java.util.Properties;
+import org.openide.util.lookup.ServiceProvider;
+
+/**
+ * This experiment compares performance when ignoring text data.
+ *
+ * @author vektor
+ */
+@ServiceProvider(service = ExperimentSet.class)
+public class IgnoreTextData extends AbstractExperimentSet {
+
[…]

### 2011-08-28 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/ExperimentAction.java
Thesis & Experimental framework: in progress.

@@ -69,7 +69,7 @@ public final class ExperimentAction implements ActionListener {
             }
           }
           s.close();
-          final int highestNumber = Integer.valueOf(number).intValue();
+          final int highestNumber = number == null ? -1 : Integer.valueOf(number).intValue();
           LOG.info("Restart file: experiment name: " + experimentName + ", highest number: " + highestNumber);
           runExperiment(experimentName, highestNumber + 1);
           return;

### 2011-08-28 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/CHForMutation.java
Thesis & Experimental framework: in progress.

@@ -20,8 +20,6 @@ import cz.cuni.mff.ksi.jinfer.base.utils.FileUtils;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.Experiment;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.data.OfficialTestData;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.data.SizeTestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.data.TestData;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
 import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
@@ -44,11 +42,6 @@ import org.openide.util.lookup.ServiceProvider;
 @ServiceProvider(service = ExperimentSet.class)
 public class CHForMutation extends AbstractExperimentSet {
 
-  private static final List<TestData> TEST_DATA = Arrays.<TestData>asList(
-          SizeTestData.GRAPH_80_320, SizeTestData.GRAPH_90_405,
-          SizeTestData.GRAPH_100_500, OfficialTestData.GRAPH_100_100,
-          OfficialTestData.GRAPH_100_200, OfficialTestData.GRAPH_100_1000);
-
   private static final int POOL_SIZE = 10;
 
   @Override
@@ -63,7 +56,7 @@ public class CHForMutation extends AbstractExperimentSet {
     final List<ExperimentParameters> ret = new ArrayList<ExperimentParameters>(10);
 
     for (int i = 0; i < Utils.getIterations(); i++) {
-      for (final TestData data : TEST_DATA) {
+      for (final TestData data : BestIHForGlpk.TEST_DATA) {
         ret.add(new ExperimentParameters(data.getFile(), POOL_SIZE, 1, 1,
                 data.getKnownOptimum(), new Random(), improvement, new Weight(), new TimeIterations(1)));
         ret.add(new ExperimentParameters(data.getFile(), POOL_SIZE, 1, 1,
@@ -77,7 +70,7 @@ public class CHForMutation extends AbstractExperimentSet {
   @Override
   protected void notifyStart() {
     final StringBuilder sb = new StringBuilder();
-    for (final TestData data : TEST_DATA) {
[…]

### 2011-08-28 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.
Thesis & Experimental framework: in progress.

@@
 \STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
+\FOR{$i = 1 \to 50$}
 	\FOR{$time = 1 \to 49$ step $2$}
@@
   \centering
-    \includegraphics[angle=-90, width=\textwidth]{images/experiments/time-vs-quality}
+    \includegraphics[width=\textwidth]{images/experiments/time-vs-quality}
 \end{figure}
@@
   Input data        & all sized test data sets \\
-  Iterations        & 20 \\
+  Iterations        & 50 \\
   Pool size         & 1 \\
@@
   \centering
-    \includegraphics[angle=-90, width=\textwidth]{images/experiments/time-till-optimum}
+    \includegraphics[width=\textwidth]{images/experiments/time-till-optimum}
 \end{figure}
@@
 \begin{algorithm}
-\caption{Random vs. Fuzzy vs. FIDAX set generation}
+\caption{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX} set generation}
 \label{listing-experiment-random-fuzzy-fidax}
@@
 \STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
+\FOR{$i = 1 \to 50$}
 	\FOR{$file \in $ official test data}
@@
 
+\begin{algorithm}
+\caption{Best Standalone CH set generation}
+\label{listing-experiment-best-standalone-ch}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$file \in $ official test data}
+	\FOR{$i = 1 \to 50$}
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \emptyset\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Fuzzy}, IH = \emptyset\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Incremental}, IH = \emptyset\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Removal}, IH = \emptyset\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{FIDAX}, IH = \emptyset\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
 TODO XMA-c, XMA-p, MSH and NTH are uninteresting - every CH found the optimum every time
@@
 
-\subsection{Best IH for the best CH}
+\subsection{Best IH for Glpk}
 
@@
 
-TODO with the best construction heuristic (GLPK), which IH is the best? Note that this needn't be the best combination!
+TODO with the best construction heuristic (Glpk), which IH is the best? Note that this needn't be the best combination!
 
@@
 \end{center}
+
+\begin{algorithm}
+\caption{Best IH for \heu{Glpk} set generation}
+\label{listing-experiment-best-ih-for-glpk}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
+	\FOR{$i = 1 \to 50$}
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Crossover}(ratio = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Hungry}\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Local Branching}(ratio = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
 
@@
 
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & Dual Core \\
+  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
+                    & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
+  Iterations        & 50 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+\begin{algorithm}
+\caption{\heu{Random} as CH set generation}
+\label{listing-experiment-ch-for-mutation}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
+	\FOR{$i = 1 \to 50$}
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+TODO results
+
+TODO interpretation
+
 \subsection{Various $\alpha$, $\beta$}
@@
 
-TODO
+TODO, we are expecting no influence
+
+TODO summary table
+
+TODO pseudocode
+
+TODO results
+
+TODO interpretation
 
@@
 
+TODO summary table
+
+TODO pseudocode
+
+TODO results
+
+TODO interpretation
+
+\subsection{Various Ratios in IHs}
+
+TODO
+
+TODO summary table
+
+TODO pseudocode
+
+TODO results
+
+TODO interpretation
+
 \subsection{Ignoring text data}
+
+%         NB class IgnoreTextData
 
@@
 
-TODO try this on XMA-p
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & Dual Core \\
+  Input data        & \dataset{XMA-p} \\
+  Iterations        & 50 \\
+  Pool size         & TODO \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+TODO pseudocode
+
+TODO results
+
+TODO interpretation
 

### 2011-08-28 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/BestIHForGlpk.java

@@ -32,8 +32,6 @@ import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Crossover;
 import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Hungry;
 import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.LocalBranching;
 import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Mutation;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.RandomRemove;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.RemoveWorst;
 import cz.cuni.mff.ksi.jinfer.iss.utils.Constants;
 import cz.cuni.mff.ksi.jinfer.iss.utils.Utils;
 import java.io.File;
@@ -50,7 +48,7 @@ import org.openide.util.lookup.ServiceProvider;
 @ServiceProvider(service = ExperimentSet.class)
 public class BestIHForGlpk extends AbstractExperimentSet {
 
-  private static final List<TestData> TEST_DATA = Arrays.<TestData>asList(
+  public static final List<TestData> TEST_DATA = Arrays.<TestData>asList(
           SizeTestData.GRAPH_80_320, SizeTestData.GRAPH_90_405,
           SizeTestData.GRAPH_100_500, OfficialTestData.GRAPH_100_100,
           OfficialTestData.GRAPH_100_200, OfficialTestData.GRAPH_100_1000);
@@ -86,14 +84,6 @@ public class BestIHForGlpk extends AbstractExperimentSet {
                 data.getKnownOptimum(),
                 new Glpk(1), Arrays.<ImprovementHeuristic>asList(new Mutation(0.1, 1)),
                 new Weight(), new TimeIterations(1)));
-        ret.add(new ExperimentParameters(data.getFile(), POOL_SIZE, 1, 1,
-                data.getKnownOptimum(),
-                new Glpk(1), Arrays.<ImprovementHeuristic>asList(new RandomRemove(0.1)),
-                new Weight(), new TimeIterations(1)));
-        ret.add(new ExperimentParameters(data.getFile(), POOL_SIZE, 1, 1,
-                data.getKnownOptimum(),
-                new Glpk(1), Arrays.<ImprovementHeuristic>asList(new RemoveWorst()),
-                new Weight(), new TimeIterations(1)));
       }
     }
 
[…]

### 2011-08-30 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.

@@
 
-TODO results
-
-TODO interpretation
+\begin{figure}
+  \caption{CH for \heu{Mutation}}
+  \label{image-experiment-ch-for-mutation}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/ch-for-mutation}
+\end{figure}
+
+TODO odd are random, even are Glpk
+
+TODO interpretation - Glpk as CH takes always one second, with simpler inputs almost always finds the optimum (thus the capping), with more complex is Random not that bad in the end
 
@@
 
-TODO results
+\begin{figure}
+  \caption{Ignoring Text Data}
+  \label{image-experiment-ignore-text-data}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/ignore-text-data}
+\end{figure}
 

### 2011-09-07 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: Corrections, bibtex.

@@
 \todo[inline]{bibref na prehladovy clanok?}
-Approaches of XML Schema inference can be divided by several criteria. Basic division is by language the resulting schema is written in. Commonly used languages are DTD and XML Schema.
+Approaches of XML schema inference can be divided by several criteria. A basic division is by the language the resulting schema is written in. Commonly used languages are DTD and XML Schema.
 
-The type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
+According to \cite{Mlynkova:2008:AAX:1494650.1495496} The type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
 Heuristic methods are based on experience with manual construction of schemas and their result commonly does not belong to any class of grammar.
@@
 Another important criterion is the type of input data.
-Most of the approaches process XML documents as the input of the inference process and the documents are valid against the resulting schema.
+Most of the approaches process XML documents as the input of the inference process and the documents are supposed to be valid against the resulting schema.
 Besides approaches exploiting XML data, approaches that utilize operations over XML data may be developed and they are the most significant approaches in terms of this work.
 
-According to my best knowledge at the time of writing there is just one approach of this category. It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?}.
+According to my best knowledge at the time of writing there is just one approach of this category.
+It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?
+Ty by tady něškodily všude, u každé té kategorie.}.
 
-\section{XML-Documents-utilizing Approaches}
-The process of inference commonly used by a significant number of approaches can be summarized as the following: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
+\section{Approaches that Utilize XML Documents}
+The process of inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
 e_k$ is constructed.
@@
 
+\todo[inline]{Tady by neškodilo dát odkazy na všechny ex. přístupy a říct, že popíšeme detailněji aspoň dva.}
+
 \subsection{XTRACT}
-The XTRACT \todo{bibref} system is an example of \emph{heuristic} \emph{merging state algorithm} creating the result in DTD. Its process of inference consists of three steps:
+The XTRACT \todo{bibref} system is an example of a \emph{heuristic} \emph{merging state algorithm} creating the result in DTD. Its process of inference consists of three steps:
 \begin{enumerate}
 \item Generalization - Generates a set of DTD candidates by searching the input for certain patterns and generalising corresponding fragments using regular expressions.
-\item Factoring - Groups of generalized candidate DTDs are factorized to a new ones by finding common sub-expressions to make them more concise.
+\item Factoring - Groups of generalized candidate DTDs are factorized to a new ones by finding common sub-expressions to make them more concise. \todo[inline]{priklad}
 \item Minimum Description Length (MDL) Principle - Composing a near-optimal DTD schema from the set of all generalized candidate DTDs.
@@
 \subsection{SchemaMiner}
-The SchemaMiner system \todo{bibref} is an example of \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
+The SchemaMiner system \todo{bibref} is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
+It focuses on inferring elements with the same name but different structure and unordered sequences.is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
 It focuses on inferring elements with the same name but different structure and unordered sequences.
@@
 \todo[inline]{Moze byt takto? Nejake dalsie pristupy?}
+\todo[inline]{U obou by neškodil příklad - tohle je velmi povrchní, pokud někde nečetl i ty články, nepochopí nic.}
 
 \section{Discovering XML Keys and Foreign Keys in Queries}
-This method tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
+This method is described in paper \cite{Necasky:2009:DXK:1529282.1529414}.
+It tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
 Just these queries are utilized in inference, no XML data are used.
@@
 \subsection{Keys and foreign keys}
-Authors introduce formalism that formalizes XML Schema keys.
+The authors introduce a formalism that formalizes XML Schema keys.
+
+\todo[inline]{Definovat, co pouzivame: XPath (aj s verziou), prikady, osy, lomitko, ...}
 
@@
 
+\todo[inline]{Potřebujeme tyto definice k něčemu dále?}
+
 Let $c$ be an element targeted by $C$ and $p_1$ be a element targeted by $P_1$ from $c$.
@@
 
-\subsection{Key and foreign key discovery}
+\subsection{Key and Foreign Key Discovery}
 \subsubsection{Joins in queries}
@@
 \todo[inline]{priklad?}
+\todo[inline]{Určitě. A hlavně je tento popis mnohem detailnější, přesnější a formaálnější než ty předchozí. Mělo by to být všechno v jednodném stylu.}
 
@@
 
-Authors introduces two observations:
+The authors introduce two following observations:
 \begin{enumerate}
@@
 The query is searched for so-called \emph{join patters}.
-These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and look as following (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
+These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and they look as follow (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
+
 \todo[inline]{spravne kreslit indexy pri pismenkach}
@@
 An occurrence marked as \emph{repeating} means that each element targeted by $P_1$ can be joined with more than one elements targeted by $P_2$.
-In contrary, a pattern occurrence marked as \emph{non-repeating} means that each element targeted by $P_1$ can be joined with zero or one element targeted by $P_2$.
-In other words, the former means that the observation (O1) can be applied while the other means that (O2) can be applied to infer keys and foreign keys.
+In contrast, a pattern occurrence marked as \emph{non-repeating} means that each element targeted by $P_1$ can be joined with zero or one element targeted by $P_2$.
+In other words, the former means that the observation (O1) can be applied while the latter means that (O2) can be applied to infer keys and foreign keys.
 
 Assume a pattern occurrence $\pi$.
-The decision is made on the base of the following rules (R1) - (R5).
+The decision is made on the basis of the following rules (R1) - (R5).
 Only one rule can be applied.
@@
 If any of the following rules (R1) - (R3) can be applied, $\pi$ is marked as \emph{repeating}.
-Each pattern occurence has also assigned its weight determining how sure the method is about the inferred statement.
+Each pattern occurence is also assigned with its weight determining how sure the method is about the inferred statement.
 
@@
 
-Therefore, authors introduce a scoring function.
-If a new key $K$ is going to be inferred, it has assigned an initial score 0.
+Therefore, the authors introduce a scoring function.
+If a new key $K$ is going to be inferred, it is assigned with an initial score 0.
 Each inferred positive statement about $K$ increases the score of $K$ by the weight of the statement.
@@
 A positive score means that $K$ is probably satisfied while negative means that $K$ is probably not satisfied.
-The higher the absolute value of the score is the higher the probability is.
+The higher the absolute value of the score is, the higher the probability is.
 
@@
 
-Because the method is based on intuition of how XQuery constructs are commonly applied in practice, it can be imprecise in certain cases.
+Since the method is based on intuition of how XQuery constructs are commonly applied in practice, it can be imprecise in certain cases.

### 2011-09-07 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: Corrections, bibtex.

@@
 \chapter{Analysis of XQuery}
-\todo[inline]{bibref na knihu, a W3C XQuery usecases}
+\todo[inline]{Někde by se mělo říct, kterou verzi uvažujeme, jaké obecně jsou, co je třeba XQuery Core apod.}
 This chapter discusses selected constructs of XQuery language and denotes how they could be exploited in the XML schema inference process.
 It is divided into sections by particular domains of the inference.
+
+Most of sample queries in this chapter are taken from \cite{Walmsley:2007:XQU:1205865} and \cite{Marchiori:07:XQU}, with or without modifications, as these were the main sources for this XQuery analysis.
+\todo[inline]{Overit si, ze mozem pouzivat tieto prevzate priklady.}
 
@@
 
-Besides elements, attributes could be processed exactly in the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} has attribute \texttt{ISBN}.
+Besides elements, attributes can be processed exactly in the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} has attribute \texttt{ISBN}.
 
@@
 In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurrence of elements.
-At last, we cannot be sure that queries target nodes actually presented in the XML documents.
+We cannot be sure that queries target nodes actually presented in the XML documents.
 Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not.
-In opposite, even basic methods of XML documents analysis do not have these inadequacies.
+In contrast, even basic methods of XML documents analysis do not have these inadequacies.
 
-On the other hand, the structure inference could be useful when the entire set of all XML documents is not available and provided XML documents do not cover the structure completely.
+On the other hand, the structural inference could be useful when the entire set of all XML documents is not available and provided XML documents do not cover the structure completely.
 
@@
 Another examples are particular usages of function \texttt{count()}, indexation, usage of set operators (\texttt{union}, \texttt{intersect}, \texttt{except}) and usage of function \texttt{one-or-more()}.
-Queries with description follow.
+Sample queries with a respective description follow.
 
@@
 Function \texttt{count()} returns number of items in a provided sequence.
-If the sequence is a sequence of elements the count of these elements will probably not be limited to one.
-The example query indicates that the outermost element \texttt{book} can contain more than one element \texttt{section}.
+If the sequence is a sequence of elements, the count of these elements will probably not be limited to one.
+The sample query indicates that the root element \texttt{book} can contain more than one element \texttt{section}.
 An exception is usage of function \texttt{count()} in a predicate in expressions where it is used to determine if count of some nodes is greater than zero.
@@
 
-Indexation of nodes and common usage of function \texttt{position()} suggest that writer of such query assumes sequence of respective nodes.
+Indexation of nodes and common usage of function \texttt{position()} suggest that the author of such query assumes a sequence of respective nodes.
 
@@
 
-In this query, number of elements \texttt{color} in element \texttt{product} with attribute \texttt{id} equal to 5 has to be at least one otherwise an error is raised upon execution.
+In this query, number of elements \texttt{color} in element \texttt{product} with attribute \texttt{id} equal to 5 has to be at least one, otherwise an error is raised upon execution.
 If we assume that this query is written correctly, with common sense and it should not raise the error, we can infer that element \texttt{product} has to contain one or more elements \texttt{color}.
@@
 \subsection{Occurrence Limited to One}
-In opposite to the multiple occurrence, numerous XQuery constructs limit number of occurrences of an element to at most one or exactly one. Some example queries with description follow.
+Contrary to the multiple occurrence, numerous XQuery constructs limit number of occurrences of an element to at most one or exactly one. Sample queries with description follow.
 
@@
 
-\texttt{lt} is a representant of so called \emph{value comparison operators} which operate two sequences of zero or one item.
-If an operand of a \emph{value comparison operator} is a sequence of more than one item then \emph{type error} is raised.
+\texttt{lt} is a representant of so called \emph{value comparison operators} \todo{referencia} which operate two sequences of zero or one item.
+If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} \todo{referencia} is raised.
 
@@
 
-Alike the previous example, expression in \texttt{order by} clause can be evaluated to at most one item or the \emph{type error} is raised.
+Alike the previous example, an expression in \texttt{order by} clause can be evaluated to at most one item or the \emph{type error} is raised.
 Therefore, every element \texttt{item} contains zero or one element \texttt{num} but no more.
@@
 Those are constructs indicating limitation to zero or one occurrence.
-Function \texttt{exactly-one()} works similar to the function \texttt{zero-or-one()} but accepts only sequences of exactly one item (which are in XQuery equal to this item itself).
+Function \texttt{exactly-one()} works similarly to the function \texttt{zero-or-one()} but accepts only sequences of exactly one item (which are in XQuery equal to this item itself).
 
@@
 XML Schema built-in types of elements and attributes can be inferred from the XML documents by analysing their content.
-Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise especially when a large enough set of XML documents is not available.
-\todo[inline]{priklad?} This is when a set of XQuery queries may come in handy.
+Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available.
+\todo[inline]{urcite priklad} This is when a set of XQuery queries may come in handy.
 
-If a value of an element or attribute is used in an expression then this expression could be often exploited to determine a type of the value.
+If a value of an element or attribute is used in an expression, then this expression could be often exploited to determine a type of the value.
 Comparing the value to another value of a known type and supplying the value to a function call as an argument of particular type are examples of these expressions as shown below.
@@
 
-Determined types are \texttt{xs:date} for element \texttt{date}, \texttt{xs:decimal} for element \texttt{price} and \texttt{xs:byte} for attribute \texttt{id}.
+The determined types are \texttt{xs:date} for element \texttt{date}, \texttt{xs:decimal} for element \texttt{price} and \texttt{xs:byte} for attribute \texttt{id}.
 
-Alongside common expressions, there are other XQuery constructs indicating types as \emph{type casting}, \emph{constructors} and so called \emph{type declarations}.
+Alongside common expressions, there are other XQuery constructs indicating types, such as \emph{type casting}, \emph{constructors} and so called \emph{type declarations}.
+\todo[inline]{\^ Nevíme co to je. Chce to vysvětlit. Buď v kapitole 2 nebo tady + odkaz do specifikace.}
 
-Assuming variable \texttt{\$var} bound to a value of some element or attribute, following two fractions of queries indicate its type to be \emph{xs:integer}.
+Assuming variable \texttt{\$var} bound to a value of some element or attribute, the following two fractions of queries indicate its type to be \emph{xs:integer}.
 
@@
 
-Alike, following two examples of \emph{type declarations}. Value of element \texttt{number} is declared to be of type \texttt{xs:integer}.
+Similarly, the following two queries are examples of \emph{type declarations}. Value of element \texttt{number} is declared to be of type \texttt{xs:integer}.
 
@@
 
-Finding such patterns in queries could be useful in combination with analysis of XML data these queries are executed upon.
+Finding such patterns in queries could be useful in combination with analysis of respective XML data.
 XML data could help to confirm or disprove this assumption based on the query analysis or vice-versa.
@@
 
-\subsection{Approach from Paper Discovering XML Keys and Foreign Keys in Queries}
+\subsection{Approach from [TODO citace]}
 Paper \todo{link} introduces a method of discovering keys and foreign keys by investigation of joins in queries.
-Basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form.
+The basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form.
 Therefore joins with the same semantics written in different syntax are not taken into consideration.
@@
 
-This query illustrates a join where the value of one of the joined elements is not required thus the knowledge of its existence is sufficient.
-Therefore the query has not to contain the second \texttt{for} or \texttt{let} keyword and its expression can be moved to \texttt{where} clause.
+This query illustrates a join where the value of one of the joined elements is not required, thus the knowledge of its existence is sufficient.
+Therefore the query does not have to contain the second \texttt{for} or \texttt{let} keyword and its expression can be moved to \texttt{where} clause.
 
@@
 \subsection{Join of Recursive Data}
-By the term "recursive data" are meant XML data, for example elements that somehow reference items from the same set. Example of a query that operates upon such data follows.
+By the term 'recursive data' are meant XML elements that somehow reference items from the same set. Example of a query that operates upon such data follows.
 
@@
 Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}.
-Elements with not specified attribute \texttt{partof} are at the top of the recursive hierarchy while, in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
+Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
 
@@
 \subsection{Negative Statements about Uniqueness}
-In many cases a statement refusing uniqueness of element's or attribute's values could be inferred.
-Such statements may prove helpful in combination with other methods in its process of decision making whether a particular element or attribute is unique or not.
+In many cases a statement refusing uniqueness of element or attribute values could be inferred.
+Such statements may be helpful in combination with other methods in its process of making a decision whether a particular element or attribute is unique or not.
 
@@
 
-Usage of \texttt{for} construct indicates that it is expected a sequence of \texttt{book} elements which satisfy the condition if \emph{where} clause.
+Usage of \texttt{for} construct indicates that a sequence of \texttt{book} elements which satisfy the condition in \emph{where} clause is expected.
 It is a condition composed of two single conditions joined by \emph{and} logical operator.
-Therefore in order to satisfy the whole condition, both single conditions must be satisfied as well.
+Therefore, in order to satisfy the whole condition, both single conditions must be satisfied as well.
 Thus, it is expected that several elements satisfy each of the single conditions.
@@
 However, the second condition is greater-than comparison of \texttt{year} attribute to an integer literal, it cannot be inferred whether this attribute is unique or not.
-The reason is that even if it was unique there still might be more than one \texttt{book} element meeting this condition.
+The reason is that even if it was unique, there still might be more than one \texttt{book} element meeting this condition.
 Also, any statement, positive nor negative, cannot be inferred about \texttt{title} subelement.
 
-Another simple examples are passing a result of basic path expression to a call of \texttt{distinct-values()} function and usage of \emph{aggregation} functions.
+Other simple examples are passing a result of basic path expression to a call of \texttt{distinct-values()} function and usage of \emph{aggregation} functions.
 
@@
 
-According to the use of \texttt{distinct-values()} function, it can be easily seen that writter of this query assumes possible occurrence of more than one \texttt{book} elements with the same value of their \texttt{title} subelement.
+According to the use of \texttt{distinct-values()} function, it can be easily seen that the author of this query assumes possible occurrence of more than one \texttt{book} elements with the same value of their \texttt{title} subelement.
 Thus, \texttt{title} element is not unique.
@@
 \subsection{Uniqueness}
-In opposite to non-uniqueness discovery, some XQuery constructs can indicate uniqueness of elements or attributes, however, it seems to be more difficult.
-One of approaches of uniqueness discovery could be an investigation of what the query does return.
+Contrary to non-uniqueness discovery, some XQuery constructs can indicate uniqueness of elements or attributes; however, it seems to be more difficult.
+One of the approaches of uniqueness discovery could be an investigation of what the query does return.
 Consider the following query.
@@
 
-\section{Other}
+\section{Other Constructs}
 
 \subsection{XML Schema \texttt{xs:sequence} and \texttt{xs:all} constructs}
-If an order of appearance of some element set, for example subelements of a certain element, is important this can be expressed by XML Schema construct \texttt{xs:sequence}.
+If an order of appearance of some element set, for example subelements of a certain element, is important, this can be expressed by XML Schema construct \texttt{xs:sequence}.
 On the other hand, \texttt{xs:all} is involved when the elements may occur in any order.
 
-If there is a large enough set of XML documents available it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs.
+If there is a large enough set of XML documents available, it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs.
 However, if the inference is made using a smaller set of XML documents, it may happen that every occurrence of some element set is in the same order but the input data are not representative enough to be sure.
-There comes in handy a comparison of element order in queries, which could help to decide for the use of \texttt{xs:sequence}.
+Comparison of element order in queries can help to decide for the use of \texttt{xs:sequence}.
 
-Every two elements in a XML document that are not siblings with the same name and none of them is the outermost element have the nearest common ancestor determining their relative order.
-If there is a lack of evidence to choose between the two constructs in the ancestor than available queries could be searched for an order comparison of these two elements like in the following query.
+Every two elements in an XML document that are not siblings with the same name and none of them is the root element have the nearest common ancestor determining their relative order.
+If there is a lack of evidence to choose between the two constructs in the ancestor, then available queries can be searched for an order comparison of these two elements like in the following query.
 
@@
 
-Apparently, the relative order of elements \texttt{incision} and \texttt{action} is somehow important because the query composes the result using only those \texttt{action} elements that succeed the second \texttt{incision} element.
+Apparently, the relative order of elements \texttt{incision} and \texttt{action} is important, because the query composes the result using only those \texttt{action} elements that succeed the second \texttt{incision} element.
 Therefore, in a respective type definition part of their nearest common ancestor \texttt{xs:sequence} should be used in favour of \texttt{xs:all}.
@@
 
-Intermediate XML structure in this query could be used to identify possible values of attribute \texttt{dept} in \texttt{product} element and thereby determine its type as enumeration.
+Intermediate XML structure in this query can be used to identify possible values of attribute \texttt{dept} in \texttt{product} element and therefore determine its type as enumeration.
 
-That is a very simple example of intermediate structure. Since there are admittedly many means of utilization of intermediate structure, further research would be needed.
+This is a very simple example of intermediate structure. Since there are admittedly many means of utilization of intermediate structure, further research would be needed.

### 2011-09-08 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: Corrections.

@@
 It focuses on inferring elements with the same name but different structure and unordered sequences.
+\todo[inline]{Definovat skratku XSD a co to je element type.}
 
@@
 It tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
-Just these queries are utilized in inference, no XML data are used.
+Just the queries are utilized in inference, no XML data are used.
 The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} \todo{aj unique?} constructs.
@@
 The query is searched for so-called \emph{join patters}.
-These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and they look as follow (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
+These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and they look as follows (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
 
@@
 
+\todo[inline]{Problém je, že tady ste v podstatě opsal ten článek.
+To je špatně.
+Analýza ex.
+přístupů by měla podat přehled toho, co se dnes v dané oblasti umí.
+Každý přístup by měl být popsán zhruba stejným stylem, stejně podrobně a s jednotným značením.
+Tj.
+ne zkopírovat různé odstavce z různých článků, z něčeho 20 řádků a z něčeho 6 stran.
+To není analýza ale plagiátorství.}
+
 \section{Scoring function}
@@
 Since the method is based on intuition of how XQuery constructs are commonly applied in practice, it can be imprecise in certain cases.
+
+\todo[inline]{Tady by se hodila nějaká srovnávací tabulka metod z různách pohledů (metoda, vstupy, výstupy).
+Dále pak kapitola na téma co dnešní metody neumí a čím se tedy v práci budeme zbývat.
+Tj.
+Nějaká návaznost k další kapitole.}

### 2011-09-08 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: Corrections.

@@
 The structure of XML documents captures elements from these documents along with their names, attributes and their organization.
-What the outermost elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.
+What the \todo{Co to znamená?
+Root element?}outermost elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.
 
@@
 The latter one uses also descendant axis.
+\todo{A k čemu je tato informace při odvozování?}
 
@@
 Queries may not cover the whole relevant content of the documents.
-For example some element may not be queried at all, therefore they may not be mentioned in any query.
-In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurrence of elements.
+For example some element may not be queried at all, therefore they may not be mentioned in any query \todo[inline]{Tady je ale několik případů - chtělo by popsat.
+1) Element není zapsaný v dotazu, ale dostanu ho výsledku - např. rok anrození autora, ptám-li se na autory.
+2) Element není zapsaný v dotazu, ale procházím ho při vyhodnocování -  např. to //autor
+3) Element je v části dokumentu, která se dotazu vlbec netýká.}.
+In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurrence of elements \todo{Tak toto je asi taky diskutabilní, např.
+// by mohlo naznačovat více výskytů.}.
 We cannot be sure that queries target nodes actually presented in the XML documents.
 Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not.
-In contrast, even basic methods of XML documents analysis do not have these inadequacies.
+In contrast, even basic methods of XML schema inference that utilize XML documents do not have these inadequacies.
 
@@
 
+\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}
+
 \begin{verbatim}
@@
 \subsection{Join of Recursive Data}
-By the term 'recursive data' are meant XML elements that somehow reference items from the same set. Example of a query that operates upon such data follows.
+By the term ``recursive data'' are meant XML elements that somehow reference items from the same set.
+\todo[inline]{Tohle je velmi nepřesné.
+Tady asi myslíte reference, ale rekurzivní element je obecně element, kteý obsahuje sám sebe - např.
+kapitola má kapitoly.} Example of a query that operates upon such data follows.
 
@@
 Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
+\todo{Tahle věta je nějaká divná.}
 

### 2011-09-08 doc/theses/mikula/diplomka.tex
Master thesis: In progress.
Master thesis: Corrections, bibtex.
Master thesis: Corrections.

--- 2011-06-07 doc/theses/svirec/prace.tex
@@
 
+\renewcommand\baselinestretch{1.3} % riadkovanie jeden a pol
+
 %% Pokud tiskneme oboustranně:
@@
 %% Pokud nikoliv:
-%\usepackage[czech]{babel}
-%\usepackage[T1]{fontenc}
+\usepackage[czech]{babel}
+\usepackage[T1]{fontenc}
 
@@
 \usepackage{amssymb}
+%\usepackage{alltt}
 \usepackage{todonotes}
@@
 % pekne pokope definujeme potrebne udaje
-\def\mftitle{Efficient Detection of XML Integrity Constraints}
+\def\mftitle{Inference of an XML Schema with the Knowledge of XML Operations}
 \def\mfthesistype{Master Thesis}
-\def\mfauthor{Michal Švirec}
+\def\mfauthor{Mário Mikula}
 \def\mfadvisor{RNDr. Irena Mlýnková, Ph.D.}
-\def\mfplacedate{Praha, 2010}
+\def\mfplacedate{Praha, 2011}
 
@@
 % (dle Organizační struktury MFF UK)
-Název katedry nebo ústavu
+Katedra softwarového inženýrství
 
@@
 \noindent
-Podakovanie.
+\todo[inline]{Poďakovanie.}
 
@@
 \pagestyle{plain}
+\renewcommand\thepage{}
+\tableofcontents
+\newpage
+\renewcommand\thepage{\arabic{page}}
 \setcounter{page}{1}
-\tableofcontents\thispagestyle{empty}
 
@@
 
+%%% Tabulky v diplomové práci, existují-li.
+\chapwithtoc{List of Tables}
+
 %%% Jednotlivé kapitoly práce jsou pro přehlednost uloľeny v samostatných souborech
-\include{uvod}
-\include{kap1}
-\include{kap2}
-\include{kap3}
+\chapter{Úvod}
+\todo[inline]{TODO}
+\chapter{Použité technologie/pojmy}
+\todo[inline]{TODO}
+Krátce popsat co je XML Schema, DTD, XPath, XQuery, možná ňejaký automat apod. Prostě co budete dále v textu potřebovat. Skratky: XSD, ...
+
+\include{analysis_of_recent_approaches}
+\include{analysis_of_xquery}
+%\include{kap2}
+%\include{kap3}
 
@@
 \bibliographystyle{plain}
-\bibliography{literatura}
-
-%%% Tabulky v diplomové práci, existují-li.
-\chapwithtoc{List of Tables}
+\todo[inline]{Literatura tady určitě co nejpodrobněji - odkaz nestačí, za měsíc nemusí existovat. A nemělo by to přetékat okraj.}
+\bibliography{literature}
 
@@
 \chapwithtoc{List of Abbreviations}
+\todo[inline]{Není nutno.}
 

### 2011-09-14 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress.
Thesis: in progress

@@
 
-TODO remind what these parameters mean
-
-\subsubsection{Effects on performance}
-
-TODO, we are expecting no influence
-
-TODO summary table
-
-TODO pseudocode
-
-TODO results
-
-TODO interpretation
-
-\subsubsection{Effects on the ID set}
-
-TODO
-
-TODO summary table
-
-TODO pseudocode
-
-TODO results
-
-TODO interpretation
+TODO remind what these parameters mean, remind that because these are weights, only their ratio matters.
+We have thus a 1-dimensional problem.
+However, for simplicity's sake we will use 25 combinations of alpha and beta and normalize them only during evaluation.
+
+TODO we do not expect any changes in performance and will limit the inquiry to different ID sets produced under different settings
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & Dual Core \\
+  Input data        & all official test data sets \\
+  Iterations        & 1 \\
+  Pool size         & 1 \\
+  $\alpha$, $\beta$ & $\{0.1, 0.25, 0.5, 0.75, 1\} \times \{0.1, 0.25, 0.5, 0.75, 1\}$ \\ % TODO this is not the best way to write it
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+\begin{algorithm}
+\caption{Various values of $\alpha$ and $\beta$}
+\label{listing-experiment-various-betas}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$\alpha \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
+  \FOR{$\beta \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
+    \FOR{$file \in $ official test data}
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1, alpha = \alpha, beta = \beta), IH = \emptyset\}$
+    \ENDFOR
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+Following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{100-1000}, \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.
+
+TODO remaining datasets showed various dependencies on alpha and beta, because this question is not central to our interest we shall mention only 2 examples
+
+TODO in 100-200 we have a clear boundary, in OVA1 (representative of OVAs in general) we have unclear boundary
+
+\subsubsection{Results for \dataset{100-200}}
+
+2 different sets (too long to list here), very nice boundary when we normalize alpha and beta
+
+\subsubsection{Results for \dataset{OVA1}}
+
+TODO found 2 ID sets:
+
+\begin{table}
+  \caption{Different ID sets found for \dataset{OVA1}}
+  \bigskip
+  \label{table-experiments-various-betas-ova1}
+  \centering
+  \begin{tabular}{c || c}
+    ID set 1: element@attribute & ID set 2: element@attribute \\
+    \hline
+    \texttt{aff@fa} & \texttt{aff@fa} \\
+    \texttt{com@ty} & \texttt{com@ty} \\
+    \texttt{cre@da} & \texttt{cre@da} \\
+    \texttt{cri@te} & \texttt{cri@te} \\
+    \texttt{cve@st} & \texttt{cve@st} \\
+    \texttt{\textbf{def@id}} & \texttt{\textbf{def@cl}} \\ % TODO again, my bold typewriter does not work
+    \texttt{fil@co} & \texttt{fil@co} \\
+    \texttt{mod@da} & \texttt{mod@da} \\
+    \texttt{ova@xs} & \texttt{ova@xs} \\
+    \texttt{pat@op} & \texttt{pat@op} \\
+    \texttt{sof@op} & \texttt{sof@op} \\
+    \texttt{sta@da} & \texttt{sta@da} \\
+    \texttt{sub@or} & \texttt{sub@or} \\
+    \texttt{sbt@te} & \texttt{sbt@te} \\
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Effect of $\alpha$, $\beta$ on ID set found for \dataset{OVA1}}
+  \bigskip
+  \label{table-experiments-various-betas-ova1-effect}
+  \centering
+  \begin{tabular}{c | c  c  c  c  c}
+    $\alpha$ \textbackslash $\beta$ & 0.1 & 0.25 & 0.5 & 0.75 & 1 \\
+    \hline
+    0.1  & 1 & 2 & 2 & 1 & 1 \\
+    0.25 & 2 & 2 & 2 & 1 & 2 \\
+    0.5  & 2 & 1 & 1 & 1 & 2 \\
+    0.75 & 1 & 2 & 1 & 2 & 1 \\
+    1    & 1 & 2 & 1 & 1 & 2 \\
+  \end{tabular}
+\end{table}
+
+TODO linear - we have more than 1 combination that has the same ratio (explain what this ratio is)
+
+\begin{table}
+  \caption{Effect of $\alpha$\ ratio on ID set found for \dataset{OVA1}}
+  \bigskip
+  \label{table-experiments-various-betas-ova1-ratio-effect}
+  \centering
+  \begin{tabular}{c | c || c |  c}
+    $\alpha$ ratio & ID set & $\alpha$ ratio & ID set \\
+    \hline
+    0,091	& 1 & 0,500	& 2 \\
+    0,118	& 1 & 0,500	& 2 \\
+    0,167	& 2 & 0,571	& 1 \\
+    0,200	& 2 & 0,600	& 1 \\
+    0,250	& 1 & 0,667	& 1 \\
+    0,286	& 2 & 0,667	& 1 \\
+    0,333	& 2 & 0,714	& 2 \\
+    0,333	& 2 & 0,750	& 2 \\
+    0,400	& 1 & 0,800	& 2 \\
+    0,429	& 1 & 0,833	& 2 \\
+    0,500	& 1 & 0,882	& 1 \\
+    0,500	& 2 & 0,909	& 1 \\
+    0,500	& 1 &       &   \\
+  \end{tabular}
+\end{table}
+
+TODO no clear separation, same ratio producing different sets? Randomness in GLPK input.
+
+\subsubsection{Interpretation}
+
+TODO we just note that there are differences, but will not investigate any further
+
+TODO artificial data always has only one attribute
 
@@
 
-TODO
+TODO play around with ratio in Mutation, but perhaps this is not enough - in that case move this after chained IHs, and try to tweak the best one by fiddling around with ratios.
 
@@
 
+TODO we will try 3 different scenarios, see which performs the best on the data and then try to tweak it a bit.
+
 \subsection{Algorithms for various data characteristics}

### 2011-09-14 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress

@@
 
-\subsection{Various Ratios in IHs}
-
-TODO play around with ratio in Mutation, but perhaps this is not enough - in that case move this after chained IHs, and try to tweak the best one by fiddling around with ratios.
-
-TODO summary table
-
-TODO pseudocode
-
-TODO results
-
-TODO interpretation
-
 \subsection{Ignoring text data}
@@
 
+%         NB class ChainedIHsX
+
 TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
@@
 
-\subsection{Algorithms for various data characteristics}
-
-TODO
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & Dual Core \\
+  Input data        & all official test data sets \\
+  Iterations        & 20 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+\begin{algorithm}
+\caption{Chaining IHs}
+\label{listing-experiment-chaining-ihs}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\STATE $\heu{MUT} \gets \heu{Mutation}(ratio = 0.1, limit = 1)$
+\STATE $\heu{CX} \gets \heu{Crossover}(ratio = 0.1, limit = 1)$
+\STATE $\heu{LB} \gets \heu{LocalBranching}(ratio = 0.1, limit = 1)$
+\STATE $\heu{RR} \gets \heu{RandomRemove}(ratio = 0.1)$
+\STATE $\heu{H} \gets \heu{Hungry}$
+\STATE $\heu{RW} \gets \heu{RemoveWorst}$
+\STATE $IHs \gets \emptyset$
+
+\STATE $IHs \gets IHs \cup (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})$
+
+\STATE $IHs \gets IHs \cup (\heu{CX}, \heu{RW}, \heu{MUT})$
+
+\STATE $IHs \gets IHs \cup (\heu{CX}, \heu{RR}, \heu{MUT}, \heu{RW}, \heu{LB}, \heu{RW}, \heu{H})$
+
+\FOR{$ih \in IHs$}
+  \FOR{$file \in $ official test data}
+    \FOR{$i = 1 \to 20$}
+      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih\}$
+    \ENDFOR
+  \ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+TODO result graphs
+
+MSH, NTH, XMA-c, XMA-p boring - finds the optimum in the first step
+
+TODO interpret
+
+TODO choose the best
+
+TODO experiment to try to improve the best
 

### 2011-09-15 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 The structure of XML documents captures elements from these documents along with their names, attributes and their organization.
-What the \todo{Co to znamená?
-Root element?}outermost elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.
+What the root elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.
 
@@
 The latter one uses also descendant axis.
-\todo{A k čemu je tato informace při odvozování?}
+That means the query considers all elements \texttt{author} in the document, thus it is a hint that maybe there are several elements named \texttt{author} but with different absolute paths and all of them contain an element \texttt{name}.
 
@@
 Queries may not cover the whole relevant content of the documents.
-For example some element may not be queried at all, therefore they may not be mentioned in any query \todo[inline]{Tady je ale několik případů - chtělo by popsat.
+For example some elements may not be queried at all, therefore they may not be mentioned in any query \todo[inline]{Tady je ale několik případů - chtělo by popsat.
 1) Element není zapsaný v dotazu, ale dostanu ho výsledku - např. rok anrození autora, ptám-li se na autory.
@@
 3) Element je v části dokumentu, která se dotazu vlbec netýká.}.
-In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurrence of elements \todo{Tak toto je asi taky diskutabilní, např.
-// by mohlo naznačovat více výskytů.}.
+In addition, these statements do not express any obligation of occurrence of elements and attributes nor clearly determine multiple occurrence of elements.
 We cannot be sure that queries target nodes actually presented in the XML documents.
@@
 
-\texttt{lt} is a representant of so called \emph{value comparison operators} \todo{referencia} which operate two sequences of zero or one item.
-If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} \todo{referencia} is raised.
+\texttt{lt} is a representant of so called \emph{value comparison operators} \todo{referencia na def} which operate two sequences of zero or one item.
+If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} \todo{referencia na def} is raised.
 
@@
 Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available.
-\todo[inline]{urcite priklad} This is when a set of XQuery queries may come in handy.
+This is when a set of XQuery queries may come in handy.
+For example, consider the following occurrences of element \texttt{a}.
+
+\begin{verbatim}
+<a>1</a>
+<a>6</a>
+<a>18</a>
+\end{verbatim}
+
+These three occurrences are not sufficient enough to determine the accurate type of element \texttt{a}, because values \texttt{1}, \texttt{6}, \texttt{18} are valid values of several types: \texttt{decimal}, \texttt{integer}, \texttt{byte}, \texttt{short}, \texttt{unsignedInt}, \texttt{positiveInteger} and many others.
+The following text introduces some possibilities of a type inference from queries.
 
@@
 Therefore, the reliable method would take the query structure into consideration and further research would be needed.
+\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}
 
@@
 When the control variable is bound to some element or attribute, type of this node can be inferred as enumeration and its individual values can be determined as well.
-
-\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}
 
@@
 
-\subsection{Approach from [TODO citace]}
+\subsection{Approach from \cite{Necasky:2009:DXK:1529282.1529414}}
 Paper \todo{link} introduces a method of discovering keys and foreign keys by investigation of joins in queries.
@@
 Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}.
-Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value.
-\todo{Tahle věta je nějaká divná.}
+Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, each element with this attribute specified references an element with attribute \texttt{partid} of the same value.
 
-By using a similar approach as described in the paper \todo{ref} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.
+By using a similar approach as described in \cite{Necasky:2009:DXK:1529282.1529414} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.
 

### 2011-09-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress

@@
 \chapter{Experiments}
+
+% TODO I cannot stress this enough times: make sure there is an explanation of heuristics notation used in pseudocode listing, SOMEWHERE
 
@@
 
-TODO we will try 3 different scenarios, see which performs the best on the data and then try to tweak it a bit.
+TODO we will try 3 different scenarios, see which performs the best on the data and then try to tweak it a bit. First we set all ratios to 0.1, as soon as we find the best scenario, we will tweak it.
 
@@
 
-TODO result graphs
+TODO result graphs - just a \textbf{commented} sample
 
@@
 
-TODO interpret
-
-TODO choose the best
-
-TODO experiment to try to improve the best
+TODO choose the best - see how many times in the 20 runs did they find optimum
+
+             1       2      3
+100-100      20      19     20
+100-200      19      18     17
+100-1000     4       1      5
+OVA1         20      20     20
+OVA2         19      13     18
+OVA3         17      18     20
+
+Strategies 1 and 3 are very similar, we shall choose \textbf{1} for its simplicity to tune the parameters.
+
+\subsubsection{Improving Strategy 1}
+
+TODO remind what Strategy 1 is, list all ratios we will be using, list files we will be using.
+
+TODO we will be gathering for each run the following: what were the parameters, how long did it take and whether it found optimum
+
+TODO summary table
+
+TODO pseudocode
+
+TODO results and interpretation
+
+RR = 0.5 (strong diversification), MUT = 0.05 (only a few fixed), CX = 0.2 (looking for commonalities among 1/5th of the pool)
 

### 2011-09-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress

@@
   Iterations        & 50 \\
-  Pool size         & TODO \\
+  Pool size         & 1 \\
   $\alpha$, $\beta$ & $1$, $1$ \\
@@
 
-TODO pseudocode
+\begin{algorithm}
+\caption{Ignoring text data}
+\label{listing-experiment-ignore-text-data}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+\FOR{$i \in 1 \to 50 $}
+  \STATE $ES \gets ES \cup \{\dataset{XMA-p}, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
+\ENDFOR
+\STATE \textbf{set} ``ignore text data"
+\FOR{$i \in 1 \to 50 $}
+  \STATE $ES \gets ES \cup \{\dataset{XMA-p}, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-TODO interpretation
+TODO interpretation - grammar extraction is very similar if not a bit worse, model creation is about 50\% faster.
 
@@
     \FOR{$i = 1 \to 20$}
-      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih\}$
+      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih, limit = 10\}$
     \ENDFOR
@@
 
-MSH, NTH, XMA-c, XMA-p boring - finds the optimum in the first step
+\dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p} boring - finds the optimum in the first step
 
@@
 
-             1       2      3
-100-100      20      19     20
-100-200      19      18     17
-100-1000     4       1      5
-OVA1         20      20     20
-OVA2         19      13     18
-OVA3         17      18     20
+\begin{table}
+  \caption{Performance of various IH chains}
+  \bigskip
+  \label{table-experiments-chained-ihs-tweaking}
+  \centering
+  \begin{tabular}{l || c | c | c}
+    Dataset & Strategy 1 & Strategy 2 & Strategy 3 \\
+    \hline
+    \dataset{100-100}  & \textbf{20} & 19 & \textbf{20} \\
+    \dataset{100-200}  & \textbf{19} & 18 & 17 \\
+    \dataset{100-1000} & 4  & 1  & \textbf{5}  \\
+    \dataset{OVA1}     & \textbf{20} & \textbf{20} & \textbf{20} \\
+    \dataset{OVA2}     & \textbf{19} & 13 & 18 \\
+    \dataset{OVA3}     & 17 & 18 & \textbf{20} \\
+    \end{tabular}
+  \bigskip
+  Each cell represents the number of times (out of 20) the strategy found the optimum - more is better.
+\end{table}
 
@@
 
-TODO remind what Strategy 1 is, list all ratios we will be using, list files we will be using.
+%         NB class ChainedIHs1Tweak
+
+TODO remind what Strategy 1 is, list all ratios we will be using
 
@@
 
-TODO summary table
-
-TODO pseudocode
-
-TODO results and interpretation
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Machine           & Dual Core \\
+  Input data        & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3} \\
+  Iterations        & 50 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+\begin{algorithm}
+\caption{Chained IHs - Tuning}
+\label{listing-experiment-chained-ihs-tuning}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+
+\STATE $RW \gets \heu{RemoveWorst}$
+
+\FOR{$rrRatio \in \{0, 0.05, 0.1, 0.2, 0.5\}$}
+\FOR{$mutRatio \in \{0.05, 0.1, 0.2\}$}
+\FOR{$cxRatio \in \{0.05, 0.1, 0.2\}$}
+  \STATE $RR \gets \heu{RandomRemoval}(ratio = rrRatio)$
+  \STATE $MUT \gets \heu{Mutation}(ratio = mutRatio, limit = 1)$
+  \STATE $CX \gets \heu{Crossover}(ratio = cxRatio, limit = 1)$
+  \FOR{$file \in \{\dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3}\}$}
+    \FOR{$i = 1 \to 50$}
+      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})\}$
+    \ENDFOR
+  \ENDFOR
+\ENDFOR
+\ENDFOR
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{table}
+  \caption{Performance of Strategy 1 depending on parameters - excerpt}
+  \bigskip
+  \label{table-experiments-chained-ihs-tweaking}
+  \centering
+  \begin{tabular}{c | c | c || c | c | c | c | c | c}
+    \heu{RR} & \heu{MUT} & \heu{CX} & \dataset{100-100} & \dataset{100-1000} & \dataset{100-200} & \dataset{OVA1} & \dataset{OVA2} & \dataset{OVA3} \\
+    \hline
+    0.2	& 0.2	& 0.2	   & 2749.58	& \textbf{8633.76}	  & 3029.82	& 110.82	& 98.68	& 2413.86 \\
+    0.5	& 0.05	& 0.05 & 1031.24	& 10324.46	& 1371.08	& \textbf{71.88}	  & 63.86	& 1282.94 \\
+    0.5	& 0.05	& 0.2	 & \textbf{919.74}	  & 10978.58	& 1322.02	& 74.46	  & 62.42	& \textbf{1217.00} \\
+    0.5	& 0.1	& 0.2	   & 1002.20	& 11030.64	& \textbf{1288.06}	& 85.22	  & 80.96	& 1480.90 \\
+    0.5	& 0.2	& 0.1	   & 1674.78	& 9779.82	  & 2007.42	& 108.14	& \textbf{58.56}	& 1827.46 \\
+    \end{tabular}
+  \bigskip
+  First 3 columns are ratios for specified heuristics. Remaining columns are average run times in ms. Bold are the best times in that column.
+\end{table}
+
+TODO interpretation - we calculated average runtimes for each combination of ratios and each file. Then we picked the ratio combination that yielded the lowest average runtimes.
 
 RR = 0.5 (strong diversification), MUT = 0.05 (only a few fixed), CX = 0.2 (looking for commonalities among 1/5th of the pool)
+
+TODO compare this to pure GLPK run - for example for 100-200, we got from TODO seconds to on average 1.3 seconds.
 

### 2011-09-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/SizeTime.java
#3323201 Experimental framework

@@ -1,59 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.iss.experiments.sets;
-
-import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.data.SizeTestData;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.quality.Weight;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.glpk.Glpk;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Identity;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-import org.openide.util.lookup.ServiceProvider;
-
-/**
- * Dependency of run time (pure GLPK) on input size.
- *
- * @author vektor
- */
-@ServiceProvider(service = ExperimentSet.class)
-public class SizeTime extends AbstractExperimentSet {
-
-  @Override
-  public String getName() {
-    return "Size Time";
-  }
-
-  @Override
-  protected List<ExperimentParameters> getExperiments() {
-    final List<ImprovementHeuristic> improvement = Arrays.<ImprovementHeuristic>asList(new Identity());
-
-    final List<ExperimentParameters> ret = new ArrayList<ExperimentParameters>(10);
-
[…]

### 2011-09-16 jinfer/trunk/jInfer/IDSetSearch/src/cz/cuni/mff/ksi/jinfer/iss/experiments/sets/TenIterations.java

@@ -1,69 +1,0 @@
-/*
- * Copyright (C) 2011 vektor
- *
- * This program is free software: you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation, either version 3 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program.  If not, see <http://www.gnu.org/licenses/>.
- */
-package cz.cuni.mff.ksi.jinfer.iss.experiments.sets;
-
-import cz.cuni.mff.ksi.jinfer.iss.experiments.ExperimentParameters;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.AbstractExperimentSet;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ExperimentSet;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.data.OfficialTestData;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.interfaces.ImprovementHeuristic;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.quality.Weight;
-import cz.cuni.mff.ksi.jinfer.iss.experiments.termination.TimeIterations;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.construction.Random;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Crossover;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.Mutation;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.RandomRemove;
-import cz.cuni.mff.ksi.jinfer.iss.heuristics.improvement.RemoveWorst;
-import java.util.ArrayList;
-import java.util.Arrays;
-import java.util.List;
-import org.openide.util.lookup.ServiceProvider;
-
-/**
- * Simply runs the same experiment ten times.
- *
- * @author vektor
- */
-@ServiceProvider(service = ExperimentSet.class)
-public class TenIterations extends AbstractExperimentSet {
-
-  @Override
-  public String getName() {
-    return "Ten Iterations";
-  }
-
-  @Override
[…]

### 2011-09-16 doc/theses/vitasek/chapter-4-experiments.tex
Thesis: in progress

@@
 
-TODO compare this to pure GLPK run - for example for 100-200, we got from TODO seconds to on average 1.3 seconds.
+TODO compare this to pure GLPK run - for example for \dataset{100-1000}, we got from TODO (90+ on average) seconds to on average TODO seconds (when the 10 second limit is lifted).
+
+TODO this will need 2 boxplots again...
 
@@
 
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.
+Second: if we don't have enough time, we should do Strategy 1. In under 10 seconds we will have very good results (often the optimum).
+
+% TODO if possible, write down in lessons learned:
+% - always allow for interruptions
+% - in experiments, iterations should be the outermost loop
+% - when representing an object, don't be lazy and create a class for it (don't use pairs, lists with specific data at specific locations, etc)

### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

--- 2011-09-16 doc/theses/vitasek/chapter-4-experiments.tex
@@
-\chapter{Experiments}
-
-% TODO I cannot stress this enough times: make sure there is an explanation of heuristics notation used in pseudocode listing, SOMEWHERE
-
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
-
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
-
-First of all, we would like to get an idea of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
-
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
-
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
-
-\section{Experimental Data}
-
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
-
-\begin{itemize}
-	\item Realistic
-	\item Realistic with artificial (converted) attributes
-	\item Artificial
-\end{itemize}
-
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
-
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
-
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
-Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
-
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum % TODO footnote explaining that we got the optimum by running GLPK w/o time limit
-for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
-
-\subsection{Realistic data}
-
-From 3 different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
-
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
-
-\begin{table}
-  \caption{List of realistic test data files}
-  \bigskip
-  \label{table-experiments-data-realistic}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-  	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-  	\hline
-  	\dataset{OVA1}  & 4.5      & 29 & 43 & 0.45588235294117635 \\
-  	\dataset{OVA2}  & 11.9     & 23 & 36 & 0.1634615384615385  \\
-  	\dataset{OVA3}  & 237.6    & 31 & 47 & 0.25537156151635415 \\
-  	\dataset{XMA-c} & 1 807.7  & 1  & 0  & 0.7546666666666667  \\
-  	\dataset{XMA-p} & 13 748.3 & 1  & 0  & 0.2019306150568969  \\
-  	\dataset{XMD}   & 1 743.0  & 17 & 15 & 0.09786094165493507 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data}
-  \label{image-experiments-data-realistic}
-  \centering
-    \subfigure[\dataset{OVA1}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA1}}
-    \subfigure[\dataset{OVA2}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA2}}
-    \subfigure[\dataset{OVA3}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/OVA3}}
-    \subfigure[\dataset{XMA-c}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-c}}
-    \subfigure[\dataset{XMA-p}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/realistic/XMA-p}}
-    \subfigure[\dataset{XMD}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/realistic/XMD}}
-\end{figure}
-
-To interpret the data a little: \dataset{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
-
-On the other hand, the \dataset{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
-
-Finally, the \dataset{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
-
-\subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
-
-We found 2 data sets to convert, \dataset{MSH} and \dataset{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
-None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
-
-TODO schemas?
-
-To address the conversion: in the case of \dataset{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \dataset{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
-
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
-
-\begin{table}
-  \caption{List of realistic test data files with converted attributes}
-  \bigskip
-  \label{table-experiments-data-converted}
-  \centering
-  \begin{tabular}{l | r | c | c | l}
-  	Name  & Size [kb] & $|V|$ & $|E|$ & Optimum \\
-  	\hline
-  	\dataset{MSH}  & 3 100.5 & 1 & 0 & 0.5416472778036296 \\
-  	\dataset{NTH}  & 2 523.5 & 5 & 7 & 0.057918595422124436 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Realistic data with converted attributes}
-  \label{image-experiments-data-converted}
-  \centering
-    \subfigure[\dataset{MSH}]{\includegraphics[width=0.08\textwidth]{images/experiments/data/converted/MSH}}
-    \subfigure[\dataset{NTH}]{\includegraphics[width=0.45\textwidth]{images/experiments/data/converted/NTH}}
-\end{figure}
-
-Again to do a little interpretation: in the case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \dataset{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
-
-In the case of \dataset{NTH} we introduced more attributes, 8 to be precise.
-Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
-
-\subsection{Artificial data}
-
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
-
-This is actually simple. Consider the following excerpt from an XML file:
-
-\begin{scriptsize}
-\begin{verbatim}
-<graph>
-  <vertex0 attr="-2968876296119015800"/>
-  <vertex1 attr="1729745997570096518"/>
-  <vertex2 attr="-9020549659620928934"/>
-  ...
-  <vertex99 attr="-7545982394508643394"/>
-
-  <vertex82 attr="0"/><vertex21 attr="0"/>
-  <vertex64 attr="1"/><vertex21 attr="1"/>
-  <vertex44 attr="2"/><vertex2 attr="2"/>
-  ...
-  <vertex96 attr="99"/><vertex40 attr="99"/>
-</graph>
-\end{verbatim}
-\end{scriptsize}
-
-We want to create a graph with around $v$ vertices and $e$ edges.
-First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
-
-The pseudocode for this is in Listing \ref{listing-random-data}.
-
-\begin{algorithm}
-\caption{Random XML data creation}
-\label{listing-random-data}
-\begin{algorithmic}
-\REQUIRE $v$ requested number of vertices
-\REQUIRE $e$ requested number of edges
-\ENSURE XML file content
-\PRINT \texttt{<graph>}
-\FOR{$i = 1 \to |V|$}
-	\STATE $R \gets RANDOM$
-	\PRINT \texttt{<vertex\textit{i} attr="\textit{R}">}
-\ENDFOR
-\FOR{$i = 1 \to |E|$}
-	\STATE $v1 \gets RANDOM(|V|)$
-	\STATE $v2 \gets RANDOM(|V|)$
-	\PRINT \texttt{<vertex\textit{v1} attr="\textit{i}"> <vertex\textit{v2} attr="\textit{i}">}
-\ENDFOR
-\PRINT \texttt{</graph>}
-\RETURN
-\end{algorithmic}
-\end{algorithm}
-
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
-
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
-
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \dataset{100-100}, \dataset{100-200} and \dataset{100-1000}.
-Note that the name is always in the form $v-e$.\\
-
-All of the experimental data sets mentioned so far, realistic, converted and artificial alike will be referred to as \textit{official test data (sets)}.\\
-
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \dataset{0-0} as the trivial one to \dataset{100-500} as the biggest one.
-From now on, these will be referred to as \textit{sized test data (sets)}.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
-
-TODO this is on the DVD.
-
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
-
-\begin{table}
-  \caption{List of artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-  	Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-  	\hline
-  	\dataset{100-100}  & 8.4  & 99 & 95  & 0.95 & 0.02 & 0.836666666666667 \\
-  	\dataset{100-200}  & 13.0 & 96 & 174 & 1.81 & 0.04 & 0.726000000000000 \\
-    \dataset{100-1000} & 49.5 & 93 & 754 & 8.11 & 0.16 & 0.380952380952381 \\
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{List of ``sized" artificial test data files}
-  \bigskip
-  \label{table-experiments-data-artificial-size}
-  \centering
-  \begin{tabular}{l | r | c | c | c | c | l}
-	  Name  & Size [kb] & $|V|$ & $|E|$ & $\frac{|E|}{|V|}$ & $\frac{|E|}{E_{max}}$ & Optimum \\
-  	\hline
-  	\dataset{0-0}     & 0.2  & 0  & 0   & -    & -    & 0.0                \\
-  	\dataset{10-5}    & 0.6  & 10 & 5   & 0.50 & 0.11 & 0.8500000000000002 \\
-    \dataset{20-20}   & 1.7  & 18 & 13  & 0.72 & 0.08 & 0.7166666666666669 \\
-    \dataset{30-45}   & 3.1  & 29 & 43  & 1.48 & 0.11 & 0.7083333333333334 \\
-  	\dataset{40-80}   & 5.1  & 39 & 72  & 1.85 & 0.10 & 0.6950000000000002 \\
-  	\dataset{50-125}  & 7.5  & 48 & 111 & 2.31 & 0.10 & 0.6566666666666666 \\
-  	\dataset{60-180}  & 10.4 & 58 & 157 & 2.71 & 0.09 & 0.6214285714285716 \\
-  	\dataset{70-245}  & 13.8 & 67 & 205 & 3.06 & 0.09 & 0.5982142857142856 \\
-  	\dataset{80-320}  & 17.6 & 76 & 261 & 3.43 & 0.09 & 0.5791666666666667 \\
-  	\dataset{90-405}  & 21.9 & 86 & 352 & 4.09 & 0.10 & 0.528888888888889  \\
-  	\dataset{100-500} & 26.7 & 91 & 388 & 4.26 & 0.09 & 0.4981818181818182 \\
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \caption{Artificial data}
-  \label{image-experiments-data-artificial}
-  \centering
-  	\subfigure[\dataset{48-80}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[\dataset{70-245}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
-\end{figure}
-
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
-
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
-
-\section{Experimental Setup}
-
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
-
-We now have to define a few notions before moving forward to the description of our experiments.
-
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
-	\begin{itemize}
-		\item All the parameters in all the heuristics.
-		\item The specific way in which the heuristics are chained.
-		\item Parameters $\alpha$ and $\beta$ in the weight (quality) measurement.
-		\item Initial pool size.
-		\item The termination criteria.
-		\item The input XML file.
-		\item Known optimum for this file and $\alpha$, $\beta$.
-	\end{itemize}
-\end{define}
-
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
-
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
-
-\subsection{Grammar and Model Creation}
-
-TODO link BasicIGG to see how we get IG
-
-TODO describe how we get the model
-
-\subsection{Hardware and Software}
-
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Quad Core Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-Windows 7 SP1 64bit
-TODO
-TODO
-GLPK version TODO
-\end{verbatim}
-
-\subsubsection{Dual Core Machine}
-
-\begin{verbatim}
-Intel Core 2 Duo processor @ 2.33 GHz
-4 GB DDR2 RAM
-Windows 7 SP1 64bit
-Java SE Runtime Environment (build 1.6.0_26-b03)
-Java HotSpot 32-Bit Client VM (build 20.1-b02)
-GLPK version 4.45 (Cygwin)
-GLPK version 4.34 (native)
-\end{verbatim}
-
-\subsection{Methodology}
-
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
-This improvement is left for the future work.
-
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
-OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
-
-\subsection{Measuring the time}
-
-TODO
-
-\subsection{Obtaining the results}
-
-Every run of an experiment produces a trace such as the one presented and commented on in Appendix \ref{appendix-trace}.
-We can get all the information relevant to that experiment run from this trace alone.
-An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
-
-TODO explain boxplots, how to read them
-
 \section{Experimental Results}
@@
 
-The experimental set will contain $50 * (11 + 11) = 1100$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
+The experimental set will contain $50 * (11 + 11) = 1100$ configurations: 50 iterations for 11 test data sets plus 11 sized test data sets.
 There will be no CHs or IHs.
@@
     \hline
-    \dataset{OVA1}     & < 10 & - & < 10 & - < 10 & - \\
-    \dataset{OVA2}     & < 10 & - & < 10 & - < 10 & - \\
+    \dataset{OVA1}     & < 10 & - & < 10 & - & < 10 & - \\
+    \dataset{OVA2}     & < 10 & - & < 10 & - & < 10 & - \\
     \dataset{OVA3}     & 42.94 & 19.8509 & 60.92 & 27.0848 & 103.86	& 31.6911 \\
@@
 
-TODO interpretation
+Results are in Table \ref{table-experiments-grammar-model-timing}.
+We are presenting the average grammar extraction (GE) times and their standard deviation, the same for model creation (MC) and total (sum of these two, Tot) times.
+For many data sets the average time is less than 10 ms: this is not enough to be precise and we don't calculate the standard deviation in these cases.
+
+We can see from the results that for most data sets their model can easily be created under around one second, only in case of the biggest set \dataset{XMA-p} (13 MB) this takes some 17 seconds.
+We can conclude that grammar and model creation times are not a bottleneck for now.
+Heuristics run times will be order of magnitude higher.
 
@@
 
-TODO interpretation: it is interesting that even though $|V|$ and $|E|$ are increasing, the times remain almost the same. This probably means that the most of the time is spent in I/O.
+Results are in table \ref{table-experiments-glpk-timing}.
+For each data set there are the times of (GLPK) input creation (IC) - average and standard deviation, then the same for output parsing (OP) and total (Tot).
+
+Interestingly enough, in most cases the times to create an input for GLPK and then to parse its output are very similar.
+Also, for sized test data sets it is interesting to note that even though the $|V|$ and $|E|$ counts are increasing, the times remain almost the same.
+This probably due to the fact that IC and OP times include the I/O when writing to a file for GLPK or reading the file it produced, and these times are probably the most relevant.
 
@@
 In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
+The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution, % TODO link
+or via Cygwin as one of its packages.
+
+If we find out which of these Cygwin version is better, we will be using it exclusively knowing this should not affect any other aspect of our experiments.
+We might also find that there is no relevant difference, which would be and interesting finding, too.
 
 Apart from comparing different versions, we shall see how the pure GLPK approach behaves.
-The first part of this experiment will be limiting the run time, thus making it an instance of Truncated branch \& bound.
+The first part of this experiment will be limiting the run time, thus making it an instance of \heu{Truncated branch \& bound}.
 In this case we will see the dependency between the run time and the quality achieved in it.
@@
 The only data we gather in the GnuPlot file are the final qualities (weights).
+The data set used is \dataset{100-500} as the biggest one in sized test data.
 
@@
 \FOR{$i = 1 \to 50$}
-	\FOR{$time = 1 \to 49$ step $2$}
+	\FOR{$time = 1 \to 46$ step $5$}
     \STATE $ES \gets ES \cup {CH = \heu{Glpk}(limit = time), IH = \emptyset}$
@@
 
-TODO interpretation
+Results are in Figure \ref{image-experiment-time-vs-quality}.
+They should be interpreted as follows: for each time limit from 1 to 46 seconds there are two boxplots next to each other, the left, dashed one is the native GLPK, the right, solid one is the Cygwin GLPK.
+This is reflected in the tics on the X (time) axis, meaning that the axis cannot be interpreted in the usual way.
+
+We can see from the graph that even though for smaller times (1 and 6 seconds, respectively) the Cygwin GLPK is reaching better qualities with smaller variance, starting from 11 seconds the native GLPK is at least as good or better for every following time.
+The results are inconclusive though, it is necessary to wait for confirmation from the second part of this experiment.
 
@@
 
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
+The other way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
+This experimental set will contain 550 configurations for each version.
 Every configuration will let \heu{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
+This will be repeated in 50 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
 There will again be no IH, the only data we will collect are the times of the CH run in each case.
@@
 \STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
+\FOR{$i = 1 \to 50$}
 	\FOR{$file \in $ sized test data}
@@
 
-TODO interpretation. Graph has its Y axis in log, stress this fact in the text too
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
+Results are in Figure \ref{image-experiment-time-until-optimum}, please take a note that the Y axis is in log scale.
+As with the previous case, the X axis cannot be interpreted in the usual way.
+For each data set there are two boxplots next to each other: the left one is the native GLPK, the right one is the Cygwin GLPK.
+
+From these results it becomes clear that the native GLPK has in general shorter running times for each and every input data set than its Cygwin counterpart.
+This becomes less extreme with the increasing input size, which leads us to suspicion that the core parts of computation in both cases are equally powerful.
+Regardless of that, we shall be using the \textbf{native} GLPK for following experiments.
+
+To conclude the first timing experiments we introduce a summary pie chart in Figure \ref{image-experiment-timing-summary}.
+This shows the typical distribution of times needed to find the optimum for the \dataset{OVA3} data set.
+
+These experiments proved that for bigger data sets the times to reach the optimum might become too long. We shall attempt to find heuristics to reach the optimum faster in the following experiments.
+
+% TODO fix this chart
 \begin{figure}
@@
 
-TODO interpret
-
 \subsection{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX}}
@@
 
-Our investigation into various CH will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \heu{Random} and \heu{Fuzzy}.
+Our investigation into various CHs will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \heu{Random} and \heu{Fuzzy}.
 
@@
 
-The experimental set will contain 660 configurations in total: 3 different CHs * 11 official test data sets * 20 iterations.
+The experimental set will contain 1650 configurations in total: 3 different CHs * 11 official test data sets * 50 iterations.
 There will be no improvement heuristics.
@@
 
-We will be gathering the running time of the CH itself and incumbent qualities (best in the pool) for GnuPlot.
+We will be gathering the running time of the CH itself and quality of the best solution found for GnuPlot.
 
@@
 
-TODO repair outliers in these 2 graphs
-
 \begin{figure}
@@
 
-TODO interpretation - Fuzzy consistently wins, but takes the longest time by far. Random is better than FIDAX in artificial as well as some real data.
+Results can be found in Figure \ref{image-experiment-random-fuzzy-fidax-quality} - qualities achieved and Figure \ref{image-experiment-random-fuzzy-fidax-time} - times spent.
+The Y (time) axis in the latter figure is again in log scale.
+For each data set there are 3 boxplots next to each other.
+The first, leftmost, represents \heu{Random}, second \heu{Fuzzy} and finally the third, rightmost is \heu{FIDAX}.
+
+We can draw the following conclusions.
+\heu{Fuzzy} consistently finds the best solution, but it's by far the slowest of these CHs.
+The trivial \heu{Random} is better than \heu{FIDAX} in artificial as well as some real data.
 
@@
 
-There is a minor question to be answered easily: is \heu{FIDAX} ``hungry enough"? Could we improve it using \heu{Hungry} as IH? Let us design a short experiment.
+Now we shall try to answer a minor question, whether it is possible to improve \heu{FIDAX} by using \heu{Hungry} as IH. This short experiment answers that question.
 
@@
 
-We won't be needing pool bigger than one, nor more iterations - both \heu{FIDAX} and \heu{Hungry} are deterministic.
+We need a pool size of one and only a single iteration - both \heu{FIDAX} and \heu{Hungry} are deterministic.
 We will try all official data sets, first with empty IH, second with \heu{Hungry} as IH.
@@
 
-TODO compare all CHs - limit GLPK to 1 second
-TODO pool size 10 where applicable
-
-TODO chart for each file: CHs on the X axis, quality Y axis, boxplots
+We shall now try to find the best standalone CH, that is the CH that finds on average the best solutions when run without any IHs.
+We need to set a time limit for \heu{Glpk} to make it an instance of \heu{Truncated Branch \& Bound}, and we shall use 1 second.
+This is the smallest time limit possible for GLPK and it is still a reasonably short time, fair to other CHs.
 
@@
 \end{center}
+
+We will use all the official data sets, set the pool size to 10 where applicable, $\alpha$ and $\beta$ to 1.
+This experiment will consist of 50 iterations * 11 data sets * 6 CHs = 3300 experimental configurations.
+See the Listing \ref{listing-experiment-best-standalone-ch} for details.
+This time we are not interested in run times, only in qualities which we shall gather in a format for GnuPlot.
 
@@
 
-TODO XMA-c, XMA-p, MSH and NTH are uninteresting - every CH found the optimum every time
+For data sets \heu{XMA-c}, \heu{XMA-p}, \heu{MSH} and \heu{NTH} every CH found the optimum every time.
+Graphs representing the results for remaining data sets can be found in Figure \ref{image-experiments-best-standalone-ch}.
 
@@
 
-TODO interpretation - GLPK wins all the time, takes 1 second, which is OK.
+We can see that \heu{Glpk} wins in every single case. We will start from there and try to build upon this result.
 
@@
 
-TODO with the best construction heuristic (Glpk), which IH is the best? Note that this needn't be the best combination!
-
-TODO we will ignore RandomRemove and RemoveWorst, they cannot help us
-
-TODO we will be using only a few sets where we can observe improvement
+The next logical step will be to try to add one IH after the best CH we have found, \heu{Glpk}.
+We will investigate all IHs except for \heu{RandomRemove} and \heu{RemoveWorst}, which cannot help us at this time.
+
+We should note that the combination best CH - best IH found this way does not necessarily need to be the best one, because we find it using a hungry approach.
 
@@
 \end{center}
+
+This experimental set will contain 6 data sets * 50 iterations * 4 IHs = 1200 experimental configurations.
+Note that we are using only the most challenging data sets, as the combination of \heu{Glpk} as CH and any other IH is already an overkill for easier data sets.
 
@@
 
+The results are listed in Table \ref{table-experiments-best-ih-for-glpk}.
+We shall denote \textit{improvement} the absolute increase in quality after running \heu{Glpk} and after running the IH.
+The table now lists for each data set and each IH the average improvement as well as the standard deviation of the improvement.
+Bold number represents the best IH for that specific data set.
+\heu{Mutation} proves to be the best IH for 3 out of 6 data sets.
+
 \begin{table}
@@
 \end{table}
-
-TODO interpretation: \heu{Mutation} wins.
 
@@
 \begin{algorithm}
-\caption{Various values of $\alpha$ and $\beta$}
+\caption{Various values of $\alpha$ and $\beta$ set generation}
 \label{listing-experiment-various-betas}
@@
 \begin{algorithm}
-\caption{Ignoring text data}
+\caption{Ignoring text data set generation}
 \label{listing-experiment-ignore-text-data}
@@
 \begin{algorithm}
-\caption{Chaining IHs}
+\caption{Chaining IHs set generation}
 \label{listing-experiment-chaining-ihs}
@@
 \begin{algorithm}
-\caption{Chained IHs - Tuning}
+\caption{Chained IHs - Tuning set generation}
 \label{listing-experiment-chained-ihs-tuning}

### 2011-09-17 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress

--- 2011-09-16 doc/theses/vitasek/chapter-4-experiments.tex
@@
 
-Also, every experimental configuration is run 20 times in hope that the effects of any events adversely affecting our results (e.g.
+Also, every experimental configuration is run 50 times in hope that the effects of any events adversely affecting our results (e.g.
 OS deciding to run some house cleaning) will be averaged out.
@@
 TODO explain boxplots, how to read them
-
-\section{Experimental Results}
-
-\subsection{Grammar and model generation}
-
-%         NB class GrammarModelTiming
-
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & all official and sized test data sets \\
-  Iterations        & 50 \\
-  Pool size         & not applicable \\
-  $\alpha$, $\beta$ & not applicable \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-The experimental set will contain $50 * (11 + 11) = 1100$ configurations: 20 iterations for 11 test data sets plus 11 sized test data sets.
-There will be no CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
-
-\begin{table}
-  \caption{Grammar Extraction and Model Creation Times}
-  \bigskip
-  \label{table-experiments-grammar-model-timing}
-  \centering
-  \begin{tabular}{l || l | l || l | l || l | l}
-    Data set & GE & GE & MC & MC & Tot & Tot \\
-     & avg [ms] & stdev & avg [ms] & stdev & avg [ms] & stdev \\
-    \hline
-    \dataset{OVA1}     & < 10 & - & < 10 & - < 10 & - \\
-    \dataset{OVA2}     & < 10 & - & < 10 & - < 10 & - \\
-    \dataset{OVA3}     & 42.94 & 19.8509 & 60.92 & 27.0848 & 103.86	& 31.6911 \\
-    \dataset{XMA-c}    & 140.32 &	33.2618 &	90.24 &	45.8803 & 230.56 & 56.2633 \\
-	\dataset{XMA-p}    & 7518.82 &	922.8882 &	10135.46 &	502.8997 & 17654.28	& 1353.8794 \\
-	\dataset{XMD}      & 979.18 &	307.1760 &	563.04 & 341.4697 & 1542.22	& 134.6883 \\
-	\dataset{MSH}      & 570.24 &	167.1119 &	225.48 &	90.6775 & 795.72 & 161.8340 \\ 
-	\dataset{NTH}      & 328.36 & 118.3766 &	1074.9 &	155.5604 & 1403.26 & 137.8695 \\
-	\dataset{100-100}  & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{100-200}  & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{100-1000} & 18.34 & 10.2372 & 18.84 & 1.0373 & 37.18 & 9.9338 \\
-	\dataset{0-0}      & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{10-5}     & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{20-20}    & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{30-45}    & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{40-80}    & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{50-125}   & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{60-180}   & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{70-245}   & < 10 & - & < 10 & - & < 10 & - \\
-	\dataset{80-320}   & < 10 & - & < 10 & - & 12.48 & 8.3574 \\
-	\dataset{90-405}   & < 10 & - & < 10 & - & 15.88 & 10.3778 \\
-	\dataset{100-500}  & < 10 & - & < 10 & - & 18.74 &	8.8889 \\
-  \end{tabular}
-\end{table}
-
-TODO interpretation
-
-\subsubsection{GLPK interface timing}
-
-%         NB class GlpkInterfaceTiming
-
-A related problem is how long it takes to create input for GLPK and then parse its results.
-We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.
-
-\begin{table}
-  \caption{GLPK Interface Times}
-  \bigskip
-  \label{table-experiments-glpk-timing}
-  \centering
-  \begin{tabular}{l || l | l || l | l || l | l}
-	Data set & IC & IC & OP & OP & Tot & Tot \\
-     & avg [ms] & stdev & avg [ms] & stdev & avg [ms] & stdev \\
-	\hline
-	\dataset{OVA1} & 36.46 & 66.8517 & 49.8 & 114.0687 & 86.26 & 150.1044 \\
-	\dataset{OVA2} & 39.52 & 75.8210 & 48.8 & 102.4484 & 88.32 & 154.9596 \\
-	\dataset{OVA3} & 34.1 & 74.1838 & 38.62 & 89.3772 & 72.72 & 134.7295 \\
-	\dataset{XMA-c} & 40.88 & 88.6632 & 33.84 & 65.8636 & 74.72 & 127.7338 \\
-	\dataset{XMA-p} & 36.54 & 70.7436 & 49.24 & 101.2412 & 85.78 & 145.2092 \\
-	\dataset{XMD} & 37.98 & 69.2719 & 32.88 & 70.2173 & 70.86 & 114.6692 \\
-	\dataset{MSH} & 40.42 & 91.9885 & 36.52 & 72.1018 & 76.94 & 138.6198 \\
-	\dataset{NTH} & 36.02 & 66.3403 & 38.06 & 88.8244 & 74.08 & 128.9974 \\
-	\dataset{100-100} & 46.5 & 103.3929 & 46.92 & 89.7049 & 93.42 & 158.7267 \\
-	\dataset{100-200} & 42.34 & 96.1204 & 38.22 & 90.0284 & 80.56 & 152.6534 \\
-	\dataset{100-1000} & 32.92 & 64.4534 & 42.1 & 89.4546 & 75.02 & 127.8541 \\
-	\dataset{0-0} & 46.8 & 123.5183 & 46.92 & 102.2601 & 93.72 & 181.5228 \\
-	\dataset{10-5} & 40.06 & 75.7370 & 40.1 & 72.4851 & 80.16 & 126.7135 \\
-	\dataset{20-20} & 33.72 & 70.7263 & 34.1 & 66.2781 & 67.82 & 116.3783 \\
-	\dataset{30-45} & 38.26 & 71.7549 & 45.94 & 110.1284 & 84.2 & 155.7594 \\
-	\dataset{40-80} & 37.06 & 67.0024 & 49.26 & 106.3185 & 86.32 & 144.9918 \\
-	\dataset{50-125} & 50.44 & 101.9162 & 84.76 & 364.7350 & 135.2 & 378.7835 \\
-	\dataset{60-180} & 38.38 & 89.3379 & 42.54 & 94.3742 & 80.92 & 149.6049 \\
-	\dataset{70-245} & 41.5 & 93.2951 & 40.3 & 93.4858 & 81.8 & 149.6797 \\
-	\dataset{80-320} & 51.92 & 121.9812 & 47.98 & 96.0904 & 99.9 & 171.4617 \\
-	\dataset{90-405} & 40.5 & 91.5373 & 36.46 & 88.5099 & 76.96 & 144.2890 \\
-	\dataset{100-500} & 37.82 & 85.7571 & 43.4 & 90.3257 & 81.22 & 141.9103 \\
-  \end{tabular}
-\end{table}
-
-TODO interpretation: it is interesting that even though $|V|$ and $|E|$ are increasing, the times remain almost the same. This probably means that the most of the time is spent in I/O.
-
-\subsection{GLPK: native vs. Cygwin}
-
-%         NB class TimeQuality
-%         NB class TimeTillOptimum
-
-In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution TODO link, or via Cygwin as one of its packages.
-
-If we find out which of these Cygwin version is better, we will be using it exclusively without very much fear that this could affect any other aspect of our experiments.
-We might also find that there is no relevant difference, which would be good, too.
-
-Apart from comparing different versions, we shall see how the pure GLPK approach behaves.
-The first part of this experiment will be limiting the run time, thus making it an instance of Truncated branch \& bound.
-In this case we will see the dependency between the run time and the quality achieved in it.
-In the second time we will let GLPK run until optimum is found.
-We shall see the dependency between input size and run time needed to achieve the optimum.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & \dataset{100-500} \\
-  Iterations        & 50 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
-There will be no improvement heuristic.
-The only data we gather in the GnuPlot file are the final qualities (weights).
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
-\label{listing-experiment-glpk-native-vs-cygwin-1}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 50$}
-	\FOR{$time = 1 \to 49$ step $2$}
-    \STATE $ES \gets ES \cup {CH = \heu{Glpk}(limit = time), IH = \emptyset}$
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{figure}
-  \caption{Time vs. Quality}
-  \label{image-experiment-time-vs-quality}
-  \centering
-    \includegraphics[width=\textwidth]{images/experiments/time-vs-quality}
-\end{figure}
-
-TODO interpretation
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & all sized test data sets \\
-  Iterations        & 50 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-Another way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
-This experimental set will contain 220 configurations for each version.
-Every configuration will let \heu{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 20 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
-
-\begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
-\label{listing-experiment-glpk-native-vs-cygwin-2}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 20$}
-	\FOR{$file \in $ sized test data}
-    \STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(no\:limit), IH = \emptyset\}$
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{figure}
-  \caption{Time Until Optimum}
-  \label{image-experiment-time-until-optimum}
-  \centering
-    \includegraphics[width=\textwidth]{images/experiments/time-till-optimum}
-\end{figure}
-
-TODO interpretation. Graph has its Y axis in log, stress this fact in the text too
-
-TODO mention that these times show we cannot rely on GLPK alone! It just takes too long.
-
-\begin{figure}
-  \caption{Timing Summary}
-  \label{image-experiment-timing-summary}
-  \centering
-    \includegraphics[width=.6\textwidth]{images/experiments/timing-pie}
-\end{figure}
-
-TODO interpret
-
-\subsection{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX}}
-
-%         NB class RandomVsFuzzyVsFidaxStart
-
-Our investigation into various CH will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \heu{Random} and \heu{Fuzzy}.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data set    & all official test data sets \\
-  Iterations        & 50 \\
-  Pool size         & 10 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-The experimental set will contain 660 configurations in total: 3 different CHs * 11 official test data sets * 20 iterations.
-There will be no improvement heuristics.
-The pool size will be set to 10, even though \heu{FIDAX} cannot not profit from this.
-Listing for this can be found in \ref{listing-experiment-random-fuzzy-fidax}.
-
-We will be gathering the running time of the CH itself and incumbent qualities (best in the pool) for GnuPlot.
-
-\begin{algorithm}
-\caption{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX} set generation}
-\label{listing-experiment-random-fuzzy-fidax}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$i = 1 \to 50$}
-	\FOR{$file \in $ official test data}
-    \STATE $ES \gets ES \cup \{file, CH = \heu{Random}(pool = 10), IH = \emptyset\} \cup \{file, CH = \heu{Fuzzy}(pool = 10), IH = \emptyset\} \cup \{file, CH = \heu{FIDAX}, IH = \emptyset\}$
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO repair outliers in these 2 graphs
-
-\begin{figure}
-  \caption{Random vs. Fuzzy vs. FIDAX - Quality}
-  \label{image-experiment-random-fuzzy-fidax-quality}
-  \centering
-    \includegraphics[width=\textwidth]{images/experiments/random-fuzzy-fidax-quality}
-\end{figure}
-
-\begin{figure}
-  \caption{Random vs. Fuzzy vs. FIDAX - Time}
-  \label{image-experiment-random-fuzzy-fidax-time}
-  \centering
-    \includegraphics[width=\textwidth]{images/experiments/random-fuzzy-fidax-time}
-\end{figure}
-
-TODO interpretation - Fuzzy consistently wins, but takes the longest time by far. Random is better than FIDAX in artificial as well as some real data.
-
-\subsubsection{Improving \heu{FIDAX} with \heu{Hungry}}
-
-%         NB class FidaxWithHungry
-
-There is a minor question to be answered easily: is \heu{FIDAX} ``hungry enough"? Could we improve it using \heu{Hungry} as IH? Let us design a short experiment.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data set    & all official test data sets \\
-  Iterations        & 1 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-We won't be needing pool bigger than one, nor more iterations - both \heu{FIDAX} and \heu{Hungry} are deterministic.
-We will try all official data sets, first with empty IH, second with \heu{Hungry} as IH.
-We will gather the qualities in each case and see whether there is any improvement.
-
-The experimental results are summarized in the Table \ref{table-experiments-fidax-and-hungry} and are quite surprising.
-As trivial a heuristic \heu{Hungry} is, it is still able to improve the ID set found by \heu{FIDAX} by as much as almost 50\% (the last row, \dataset{100-1000}).
-
-Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \dataset{100-1000}.
-Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute".
-
-\begin{table}
-  \caption{Results of adding \heu{Hungry} after \heu{FIDAX}}
-  \bigskip
-  \label{table-experiments-fidax-and-hungry}
-  \centering
-  \begin{tabular}{l | l | l}
-    Data set & Quality - \heu{FIDAX} & Quality - \heu{FIDAX} + \heu{Hungry} \\
-    \hline
-    \dataset{OVA1}     & 0.4411764705882353  & 0.4411764705882353   \\
-    \dataset{OVA2}     & 0.16346153846153846 & 0.16346153846153846  \\
-    \dataset{OVA3}     & 0.25482414123443264 & \textbf{0.2553715615163541}   \\
-    \dataset{XMA-c}    & 0.7546666666666666	 & 0.7546666666666666   \\
-    \dataset{XMA-p}    & 0.2019306150568969	 & 0.2019306150568969   \\
-    \dataset{XMD}      & 0.09786094165493509 & 0.09786094165493509  \\
-    \dataset{MSH}      & 0.5416472778036296	 & 0.5416472778036296   \\
-    \dataset{NTH}      & 0.05259709474828076 & \textbf{0.057918595422124436} \\
-    \dataset{100-100}  & 0.56	               & \textbf{0.6766666666666669}   \\
-    \dataset{100-200}  & 0.44200000000000017 & \textbf{0.5980000000000003}   \\
-    \dataset{100-1000} & 0.19952380952380955 & \textbf{0.29619047619047617}  \\
-  \end{tabular}
-\end{table}
-
-% TODO for some reason \texttt{\textbf{vertex5}} looks only like \texttt{vertex5}
-\begin{table}
-  \caption{ID sets in \heu{FIDAX} versus \heu{FIDAX} + \heu{Hungry}}
-  \bigskip
-  \label{table-experiments-fidax-and-hungry-idsets}
-  \centering
-  \begin{tabular}{l | l}
-  \heu{FIDAX} & \heu{FIDAX} + \heu{Hungry} \\
-  \hline
-                    & \texttt{\textbf{vertex5}}  \\
-                    & \texttt{\textbf{vertex26}} \\
-  \texttt{vertex30} & \texttt{vertex30} \\
-  \texttt{vertex31} & \texttt{vertex31} \\
-  \texttt{vertex32} & \texttt{vertex32} \\
-  \texttt{vertex34} & \texttt{vertex34} \\
-  \texttt{vertex35} & \texttt{vertex35} \\
-  \texttt{vertex36} & \texttt{vertex36} \\
-  \texttt{vertex37} & \texttt{vertex37} \\
-  \texttt{vertex39} & \texttt{vertex39} \\
-                    & \texttt{\textbf{vertex60}} \\
-                    & \texttt{\textbf{vertex69}} \\
-                    & \texttt{\textbf{vertex70}} \\
-  \texttt{vertex74} & \texttt{vertex74} \\
-  \texttt{vertex75} & \texttt{vertex75} \\
-  \texttt{vertex80} & \texttt{vertex80} \\
-  \end{tabular}
-\end{table}
-
-\subsection{Best standalone CH}
-
-%         NB class BestStandaloneCH
-
-TODO compare all CHs - limit GLPK to 1 second
-TODO pool size 10 where applicable
-
-TODO chart for each file: CHs on the X axis, quality Y axis, boxplots
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Quad Core \\
-  Input data        & all official test data sets \\
-  Iterations        & 50 \\
-  Pool size         & 10 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-\begin{algorithm}
-\caption{Best Standalone CH set generation}
-\label{listing-experiment-best-standalone-ch}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$file \in $ official test data}
-	\FOR{$i = 1 \to 50$}
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \emptyset\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Fuzzy}, IH = \emptyset\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Incremental}, IH = \emptyset\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Removal}, IH = \emptyset\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{FIDAX}, IH = \emptyset\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \emptyset\}$
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO XMA-c, XMA-p, MSH and NTH are uninteresting - every CH found the optimum every time
-
-\begin{figure}
-  \caption{Best Standalone CH}
-  \label{image-experiments-best-standalone-ch}
-  \centering
-  	\subfigure[\dataset{OVA1}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA1}}
-  	\subfigure[\dataset{OVA2}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA2}}
-  	\subfigure[\dataset{OVA3}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA3}}
-  	\subfigure[\dataset{XMD}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-XMD}}
-   	\subfigure[\dataset{100-100}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-100}}
-  	\subfigure[\dataset{100-200}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-200}}
-  	\subfigure[\dataset{100-1000}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-1000}}
-\end{figure}
-
-TODO interpretation - GLPK wins all the time, takes 1 second, which is OK.
-
-\subsection{Best IH for Glpk}
-
-%         NB class BestIHForGlpk
-
-TODO with the best construction heuristic (Glpk), which IH is the best? Note that this needn't be the best combination!
-
-TODO we will ignore RandomRemove and RemoveWorst, they cannot help us
-
-TODO we will be using only a few sets where we can observe improvement
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
-                    & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
-  Iterations        & 50 \\
-  Pool size         & 10 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-\begin{algorithm}
-\caption{Best IH for \heu{Glpk} set generation}
-\label{listing-experiment-best-ih-for-glpk}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
-	\FOR{$i = 1 \to 50$}
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Crossover}(ratio = 0.1, limit = 1)\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Hungry}\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Local Branching}(ratio = 0.1, limit = 1)\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{table}
-  \caption{Best IH for \heu{Glpk}}
-  \bigskip
-  \label{table-experiments-best-ih-for-glpk}
-  \centering
-  \begin{tabular}{l || l | l || l | l}
-    Data set & \heu{Hungry} & \heu{Hungry} & \heu{Crossover} & \heu{Crossover} \\
-     & improvement & improvement & improvement & improvement \\
-     & avg & stdev & avg & stdev \\
-    \hline
-    \dataset{80-320} & 0.00017 & 0.00118 & 0.00017 & 0.00118 \\
-    \dataset{90-405} & 0.00502 & 0.00618 & 0.00033 & 0.00165 \\
-    \dataset{100-500} & 0.00664 & 0.00667 & 0.00016 & 0.00081 \\
-    \hline
-    \dataset{100-100} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
-    \dataset{100-200} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
-    \dataset{100-1000} & 0.01630 & 0.01294 & 0.00180 & 0.00506 \\
-  \end{tabular}    
-	% TODO separate somehow
-  \begin{tabular}{l || l | l || l | l}
-    Data set & \heu{LB} & \heu{LB} & \heu{Mutation} & \heu{Mutation} \\
-     & improvement & improvement & improvement & improvement \\
-     & avg & stdev & avg & stdev \\
-    \hline
-    \dataset{80-320} & \textbf{0.00072} & 0.00223 & 0.00064 & 0.00218 \\
-    \dataset{90-405} & 0.00698 & 0.00616 & \textbf{0.00851} & 0.00659 \\
-    \dataset{100-500} & 0.00796 & 0.00797 & \textbf{0.00964} & 0.00804 \\
-    \hline
-    \dataset{100-100} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
-    \dataset{100-200} & 0.00000 & 0.00000 & 0.00000 & 0.00000 \\
-    \dataset{100-1000} & 0.01710 & 0.01188 & \textbf{0.02337} & 0.01558 \\
-  \end{tabular}
-\end{table}
-
-TODO interpretation: \heu{Mutation} wins.
-
-\subsubsection{\heu{Random} as CH}
-
-%         NB class CHForMutation
-
-TODO with \heu{Mutation} from the previous step, how about using Random as CH? Do we get an improvement?
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
-                    & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
-  Iterations        & 50 \\
-  Pool size         & 10 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-\begin{algorithm}
-\caption{\heu{Random} as CH set generation}
-\label{listing-experiment-ch-for-mutation}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
-	\FOR{$i = 1 \to 50$}
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{figure}
-  \caption{CH for \heu{Mutation}}
-  \label{image-experiment-ch-for-mutation}
-  \centering
-    \includegraphics[width=\textwidth]{images/experiments/ch-for-mutation}
-\end{figure}
-
-TODO odd are random, even are Glpk
-
-TODO interpretation - Glpk as CH takes always one second, with simpler inputs almost always finds the optimum (thus the capping), with more complex is Random not that bad in the end
-
-\subsection{Various $\alpha$, $\beta$}
-
-TODO remind what these parameters mean, remind that because these are weights, only their ratio matters.
-We have thus a 1-dimensional problem.
-However, for simplicity's sake we will use 25 combinations of alpha and beta and normalize them only during evaluation.
-
-TODO we do not expect any changes in performance and will limit the inquiry to different ID sets produced under different settings
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & all official test data sets \\
-  Iterations        & 1 \\
-  Pool size         & 1 \\
-  $\alpha$, $\beta$ & $\{0.1, 0.25, 0.5, 0.75, 1\} \times \{0.1, 0.25, 0.5, 0.75, 1\}$ \\ % TODO this is not the best way to write it
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-\begin{algorithm}
-\caption{Various values of $\alpha$ and $\beta$}
-\label{listing-experiment-various-betas}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\FOR{$\alpha \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
-  \FOR{$\beta \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
-    \FOR{$file \in $ official test data}
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1, alpha = \alpha, beta = \beta), IH = \emptyset\}$
-    \ENDFOR
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-Following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{100-1000}, \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.
-
-TODO remaining datasets showed various dependencies on alpha and beta, because this question is not central to our interest we shall mention only 2 examples
-
-TODO in 100-200 we have a clear boundary, in OVA1 (representative of OVAs in general) we have unclear boundary
-
-\subsubsection{Results for \dataset{100-200}}
-
-2 different sets (too long to list here), very nice boundary when we normalize alpha and beta
-
-\subsubsection{Results for \dataset{OVA1}}
-
-TODO found 2 ID sets:
-
-\begin{table}
-  \caption{Different ID sets found for \dataset{OVA1}}
-  \bigskip
-  \label{table-experiments-various-betas-ova1}
-  \centering
-  \begin{tabular}{c || c}
-    ID set 1: element@attribute & ID set 2: element@attribute \\
-    \hline
-    \texttt{aff@fa} & \texttt{aff@fa} \\
-    \texttt{com@ty} & \texttt{com@ty} \\
-    \texttt{cre@da} & \texttt{cre@da} \\
-    \texttt{cri@te} & \texttt{cri@te} \\
-    \texttt{cve@st} & \texttt{cve@st} \\
-    \texttt{\textbf{def@id}} & \texttt{\textbf{def@cl}} \\ % TODO again, my bold typewriter does not work
-    \texttt{fil@co} & \texttt{fil@co} \\
-    \texttt{mod@da} & \texttt{mod@da} \\
-    \texttt{ova@xs} & \texttt{ova@xs} \\
-    \texttt{pat@op} & \texttt{pat@op} \\
-    \texttt{sof@op} & \texttt{sof@op} \\
-    \texttt{sta@da} & \texttt{sta@da} \\
-    \texttt{sub@or} & \texttt{sub@or} \\
-    \texttt{sbt@te} & \texttt{sbt@te} \\
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{Effect of $\alpha$, $\beta$ on ID set found for \dataset{OVA1}}
-  \bigskip
-  \label{table-experiments-various-betas-ova1-effect}
-  \centering
-  \begin{tabular}{c | c  c  c  c  c}
-    $\alpha$ \textbackslash $\beta$ & 0.1 & 0.25 & 0.5 & 0.75 & 1 \\
-    \hline
-    0.1  & 1 & 2 & 2 & 1 & 1 \\
-    0.25 & 2 & 2 & 2 & 1 & 2 \\
-    0.5  & 2 & 1 & 1 & 1 & 2 \\
-    0.75 & 1 & 2 & 1 & 2 & 1 \\
-    1    & 1 & 2 & 1 & 1 & 2 \\
-  \end{tabular}
-\end{table}
-
-TODO linear - we have more than 1 combination that has the same ratio (explain what this ratio is)
-
-\begin{table}
-  \caption{Effect of $\alpha$\ ratio on ID set found for \dataset{OVA1}}
-  \bigskip
-  \label{table-experiments-various-betas-ova1-ratio-effect}
-  \centering
-  \begin{tabular}{c | c || c |  c}
-    $\alpha$ ratio & ID set & $\alpha$ ratio & ID set \\
-    \hline
-    0,091	& 1 & 0,500	& 2 \\
-    0,118	& 1 & 0,500	& 2 \\
-    0,167	& 2 & 0,571	& 1 \\
-    0,200	& 2 & 0,600	& 1 \\
-    0,250	& 1 & 0,667	& 1 \\
-    0,286	& 2 & 0,667	& 1 \\
-    0,333	& 2 & 0,714	& 2 \\
-    0,333	& 2 & 0,750	& 2 \\
-    0,400	& 1 & 0,800	& 2 \\
-    0,429	& 1 & 0,833	& 2 \\
-    0,500	& 1 & 0,882	& 1 \\
-    0,500	& 2 & 0,909	& 1 \\
-    0,500	& 1 &       &   \\
-  \end{tabular}
-\end{table}
-
-TODO no clear separation, same ratio producing different sets? Randomness in GLPK input.
-
-\subsubsection{Interpretation}
-
-TODO we just note that there are differences, but will not investigate any further
-
-TODO artificial data always has only one attribute
-
-\subsection{Ignoring text data}
-
-%         NB class IgnoreTextData
-
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & \dataset{XMA-p} \\
-  Iterations        & 50 \\
-  Pool size         & TODO \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-TODO pseudocode
-
-\begin{figure}
-  \caption{Ignoring Text Data}
-  \label{image-experiment-ignore-text-data}
-  \centering
-    \includegraphics[width=\textwidth]{images/experiments/ignore-text-data}
-\end{figure}
-
-TODO interpretation
-
-\subsection{Chaining the IHs}
-
-%         NB class ChainedIHsX
-
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
-
-TODO we will try 3 different scenarios, see which performs the best on the data and then try to tweak it a bit. First we set all ratios to 0.1, as soon as we find the best scenario, we will tweak it.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Machine           & Dual Core \\
-  Input data        & all official test data sets \\
-  Iterations        & 20 \\
-  Pool size         & 10 \\
-  $\alpha$, $\beta$ & $1$, $1$ \\
-  \hline
-\end{tabular}
-\bigskip
-\end{center}
-
-\begin{algorithm}
-\caption{Chaining IHs}
-\label{listing-experiment-chaining-ihs}
-\begin{algorithmic}
-\ENSURE experimental set $ES$
-\STATE $ES \gets \emptyset$
-\STATE $\heu{MUT} \gets \heu{Mutation}(ratio = 0.1, limit = 1)$
-\STATE $\heu{CX} \gets \heu{Crossover}(ratio = 0.1, limit = 1)$
-\STATE $\heu{LB} \gets \heu{LocalBranching}(ratio = 0.1, limit = 1)$
-\STATE $\heu{RR} \gets \heu{RandomRemove}(ratio = 0.1)$
-\STATE $\heu{H} \gets \heu{Hungry}$
-\STATE $\heu{RW} \gets \heu{RemoveWorst}$
-\STATE $IHs \gets \emptyset$
-
-\STATE $IHs \gets IHs \cup (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})$
-
-\STATE $IHs \gets IHs \cup (\heu{CX}, \heu{RW}, \heu{MUT})$
-
-\STATE $IHs \gets IHs \cup (\heu{CX}, \heu{RR}, \heu{MUT}, \heu{RW}, \heu{LB}, \heu{RW}, \heu{H})$
-
-\FOR{$ih \in IHs$}
-  \FOR{$file \in $ official test data}
-    \FOR{$i = 1 \to 20$}
-      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih\}$
-    \ENDFOR
-  \ENDFOR
-\ENDFOR
-\RETURN $ES$
-\end{algorithmic}
-\end{algorithm}
-
-TODO result graphs - just a \textbf{commented} sample
-
-MSH, NTH, XMA-c, XMA-p boring - finds the optimum in the first step
-
-TODO choose the best - see how many times in the 20 runs did they find optimum
-
-             1       2      3
-100-100      20      19     20
-100-200      19      18     17
-100-1000     4       1      5
-OVA1         20      20     20
-OVA2         19      13     18
-OVA3         17      18     20
-
-Strategies 1 and 3 are very similar, we shall choose \textbf{1} for its simplicity to tune the parameters.
-
-\subsubsection{Improving Strategy 1}
-
-TODO remind what Strategy 1 is, list all ratios we will be using, list files we will be using.
-
-TODO we will be gathering for each run the following: what were the parameters, how long did it take and whether it found optimum
-
-TODO summary table
-
-TODO pseudocode
-
-TODO results and interpretation
-
-RR = 0.5 (strong diversification), MUT = 0.05 (only a few fixed), CX = 0.2 (looking for commonalities among 1/5th of the pool)
-
-\section{The "Best" Algorithm}
-
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run.
-We will find the optimum in this way.
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do XYZ but it might depend on the characteristics of the data.
-For realistic (looks like THIS in graph representation) it is best to do A B C.
-Whereas, for artificial data (having IJK as their representation) it is best to just do 1 2 3.

### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

@@
   \begin{tabular}{l || l | l || l | l}
-    Data set & \heu{Hungry} & \heu{Hungry} & \heu{Crossover} & \heu{Crossover} \\
-     & improvement & improvement & improvement & improvement \\
-     & avg & stdev & avg & stdev \\
+             & \heu{Hungry} & \heu{Hungry} & \heu{Crossover} & \heu{Crossover} \\
+    Data set & improv - avg & improv - stdev & improv - avg & improv - stdev \\
     \hline
@@
   \begin{tabular}{l || l | l || l | l}
-    Data set & \heu{LB} & \heu{LB} & \heu{Mutation} & \heu{Mutation} \\
-     & improvement & improvement & improvement & improvement \\
-     & avg & stdev & avg & stdev \\
+             & \heu{LB} & \heu{LB} & \heu{Mutation} & \heu{Mutation} \\
+    Data set & improv - avg & improv - stdev & improv - avg & improv - stdev \\
     \hline
@@
 
-TODO with \heu{Mutation} from the previous step, how about using Random as CH? Do we get an improvement?
+As we mentioned before, we chose the combination \heu{Glpk} and \heu{Mutation} in a hungry manner.
+We will now try to take a step back and attempt to replace \heu{Glpk} with \heu{Random}, hoping to get similar qualities in much shorter time (a reminder: \heu{Glpk} always takes 1 second).
 
@@
 
+Setup used will be almost identical to that from the previous experiment.
+Experimental set will consist of 6 data sets * 50 iterations * 2 CHs = 600 experimental configurations, see Listing \ref{listing-experiment-ch-for-mutation}.
+We shall collect the eventual quality after running both the CH and the IH in format suited for GnuPlot.
+
 \begin{algorithm}
@@
 
+Results are summarized in Figure \ref{image-experiment-ch-for-mutation}.
+Again, for each data set there are two boxplots representing \heu{Random} (left one) and \heu{Glpk} (right one).
+The combination \heu{Glpk} + \heu{Mutation} always finds the optimum for the simpler data sets, thus the collapsed boxplots.
+Moreover, it achieves higher quality in each data set.
+On the other hand, combination \heu{Random} + \heu{Mutation} has much shorter running times and in the biggest (and hardest) data set \dataset{100-1000} has almost comparable results.
+This makes it a reasonable choice for big inputs where short time is more important than optimal quality.
+
 \begin{figure}
@@
 
-TODO odd are random, even are Glpk
-
-TODO interpretation - Glpk as CH takes always one second, with simpler inputs almost always finds the optimum (thus the capping), with more complex is Random not that bad in the end
-
 \subsection{Various $\alpha$, $\beta$}
 
-TODO remind what these parameters mean, remind that because these are weights, only their ratio matters.
-We have thus a 1-dimensional problem.
-However, for simplicity's sake we will use 25 combinations of alpha and beta and normalize them only during evaluation.
+After finding the best combination of a CH and IH we turn our attention to some of the parameters. The first ones are the $\alpha$ and $\beta$ from the definition of our weight function. % TODO link
+A short reminder: the weight is defined as TODO copy formula.
+It is thus obvious that only the \textit{ratio} between $\alpha$ and $\beta$ matters, not their actual values.
+This means that investigating effects of these parameters is in fact a 1-dimensional problem.
+However, for simplicity's sake we will use 25 combinations of various $\alpha$ and $\beta$ and normalize them only during evaluation.
 
-TODO we do not expect any changes in performance and will limit the inquiry to different ID sets produced under different settings
+It is worthy noting that we do not expect any changes in performance of heuristics and we will limit the inquiry to different ID sets produced under different settings.
 
@@
   Machine           & Dual Core \\
-  Input data        & all official test data sets \\
+  Input data        & realistic + converted official test data sets \\
   Iterations        & 1 \\
@@
 \end{center}
+
+This experimental set will contain 5 different $\alpha$ settings * 5 $\beta$ settings * 8 data sets = 200 experimental configurations.
+We are not using the artificial data sets, because the way they are generated (attribute values are random numbers) they cannot possibly create different optimal ID sets.
+The pseudocode capturing this is in Listing \ref{listing-experiment-various-betas}.
+We will use \heu{Glpk} constrained to 1 second (thus making it an instance of \heu{Truncated Branch \& Bound}) and no IHs.
+Pool size as well as iteration count will be 1.
+We are noting the actual ID set found by the run of the heuristic.
 
@@
   \FOR{$\beta \in \{0.1, 0.25, 0.5, 0.75, 1\}$}
-    \FOR{$file \in $ official test data}
+    \FOR{$file \in $ realistic of converted official test data}
     	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1, alpha = \alpha, beta = \beta), IH = \emptyset\}$
@@
 
-Following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{100-1000}, \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.
-
-TODO remaining datasets showed various dependencies on alpha and beta, because this question is not central to our interest we shall mention only 2 examples
-
-TODO in 100-200 we have a clear boundary, in OVA1 (representative of OVAs in general) we have unclear boundary
-
-\subsubsection{Results for \dataset{100-200}}
-
-2 different sets (too long to list here), very nice boundary when we normalize alpha and beta
+Following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.
+The \dataset{OVA*} data sets showed various dependencies on $\alpha$ and $\beta$, we shall now describe one representative example.
 
@@
 
-TODO found 2 ID sets:
+The 2 different ID sets found for various $\alpha$ and $\beta$ in \dataset{OVA1} are listed in Table \ref{table-experiments-various-betas-ova1} (note that the actual names had to be anonymized for reasons discussed in Section \ref{section-realistic-data}).
+The differing attribute mapping is highlited.
 
@@
   \begin{tabular}{c || c}
-    ID set 1: element@attribute & ID set 2: element@attribute \\
+    ID set \textbf{1}: element@attribute & ID set \textbf{2}: element@attribute \\
     \hline
@@
     \texttt{cve@st} & \texttt{cve@st} \\
-    \texttt{\textbf{def@id}} & \texttt{\textbf{def@cl}} \\ % TODO again, my bold typewriter does not work
+    \texttt{\textbf{def@id}} & \texttt{\textbf{def@cl}} <- \\ % TODO again, my bold typewriter does not work
     \texttt{fil@co} & \texttt{fil@co} \\
@@
 \end{table}
+
+Table \ref{table-experiments-various-betas-ova1-effect} summarizes the dependency of the ID set found on various values of $\alpha$, $\beta$.
+We than define the $\alpha-ratio$ as $ \dfrac{\alpha}{\alpha + \beta} $ and summarize the findings in a linear manner, sorted by increasing $\alpha-ratio$ in Table \ref{table-experiments-various-betas-ova1-ratio-effect}.
+Note that the $\alpha-ratio$s are not unique due to the way we constructed the experimental configurations here.
 
@@
 
-TODO linear - we have more than 1 combination that has the same ratio (explain what this ratio is)
-
 \begin{table}
-  \caption{Effect of $\alpha$\ ratio on ID set found for \dataset{OVA1}}
+  \caption{Effect of $\alpha-ratio$ on ID set found for \dataset{OVA1}}
   \bigskip
@@
   \begin{tabular}{c | c || c |  c}
-    $\alpha$ ratio & ID set & $\alpha$ ratio & ID set \\
+    $\alpha-ratio$ & ID set & $\alpha-ratio$ & ID set \\
     \hline
@@
 
-TODO no clear separation, same ratio producing different sets? Randomness in GLPK input.
-
-\subsubsection{Interpretation}
-
-TODO we just note that there are differences, but will not investigate any further
-
-TODO artificial data always has only one attribute
+Interestingly enough, there is no clear separation between the two ID sets depending on the $\alpha-ratio$ to be found.
+The very existence of the two sets might be due to the fact that \heu{Glpk} randomizes the order in which AMs are presented to the external GLPK solver.
+However, this question is outside of the scope of this work, and shall be left for future work.
 

### 2011-09-17 doc/theses/vitasek/appendix-c-trace.tex
Thesis: in progress

@@
     Quality: 0.19951219512195123 (10 AMs)
-Termination reason: Max\-imum ite\-rat\-ions ex\-cee\-ded.
+Termination reason: Maximum iterations exceeded.
 

### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

@@
 \subsection{Grammar and model generation}
+\label{section-grammar-model-timing}
 
@@
 
-TODO can we improve performance on huge data by ignoring textual content? Check for building as well as actual heu times.
+When considering data sets such as \dataset{XMA-p}, we notice that they contain a lot of simple text nodes that do not contribute to our search, but possibly slow it down.
+Precisely for this reason the \jmodule{BasicIGG} module in jInfer contains an option to turn off processing of such nodes.
+(It also allows to ignore the content of attributes, but this would be devastating to our cause.) Ignoring the content of text nodes means internally that these are created, but their actual string content is skipped and not saved in the memory structures.
+This means that the whole data model occupies less space on the heap, which can possibly lead to better performance.
+
+We shall now investigate this matter by taking the biggest data set \dataset{XMA-p} containing a lot of text data.
 
@@
 \end{center}
+
+Our experimental set will contain 50 iterations * 2 = 100 experimental configurations as described in Listing \ref{listing-experiment-ignore-text-data}.
+We will be using \heu{Glpk} limited to 1 second with no additional IH and pool size set to 1.
+After the first 50 iterations we will turn on the option to ignore the simple text node data and run the same 50 iterations again.
+We will be collecting the grammar extraction (GE) and model creation (MC) times as in the experiment in Section \ref{section-grammar-model-timing}.
 
@@
 
+Results are summarized in Figure \ref{image-experiment-ignore-text-data}.
+Boxplots drawn in dashed lines represent the original case, not ignoring the text data.
+Solid lines represent the case where we ignore the text data.
+
+Interestingly, the grammar extraction times tend to be shorter in the case when text data is not ignored, although this is inconclusive.
+However, there is a clear improvement of about 50 \% in the case of model creation times.
+The conclusion then is to ignore the simple text node content whenever possible when finding ID attributes.
+
 \begin{figure}
@@
 
-TODO interpretation - grammar extraction is very similar if not a bit worse, model creation is about 50\% faster.
-
 \subsection{Chaining the IHs}
@@
 
-TODO finally we got to the most interesting part, unfortunately this is basically guesswork and intuition.
+In this section we will describe the most interesting experimental area, that is chaining more than one improvement heuristics and running them in a loop.
+Unfortunately, the sheer number of possible combinations in which IHs can be ordered (as well as the number of ways to set their parameters) prohibits us from investigating this in depth.
 
-TODO we will try 3 different scenarios, see which performs the best on the data and then try to tweak it a bit. First we set all ratios to 0.1, as soon as we find the best scenario, we will tweak it.
+We shall then employ a higher-level heuristic: we will choose 3 scenarios (lists of IHs, or metaheuristics), assess their performance to find the best one and then tune its parameters.
+This approach is by no means exhaustive, it is just a probe in the problem space.
+
+The 3 scenarios we will be assessing will be constructed from the following instances of improvement heuristics:
+
+\begin{itemize}
+	\item \heu{RR} is \heu{RandomRemove} with $ratio$ set to $0.1$.
+	\item \heu{MUT} is \heu{Mutation} with $ratio$ set to $0.1$ and time limit set to 1 second.
+	\item \heu{CX} is \heu{Crossover} with $ratio$ set to $0.1$ and time limit set to 1 second.
+	\item \heu{LB} is \heu{LocalBranching} with $ratio$ set to $0.1$ and time limit set to 1 second.
+	\item \heu{RW} is \heu{RemoveWorst}.
+	\item \heu{H} is \heu{Hungry}.
+\end{itemize}
+
+The scenarios themselves shall be the following:
+
+\begin{itemize}
+	\item \textit{\textbf{Scenario 1}}. \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
+	\item \textit{\textbf{Scenario 2}}. \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow$ \heu{MUT} $\rightarrow \ldots$
+	\item \textit{\textbf{Scenario 3}}. \heu{CX} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RW} $\rightarrow$ \heu{LB} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
+\end{itemize}
 
@@
 \end{center}
+
+The experimental set will consist of 3 scenarios * 11 data sets * 20 iterations = 660 experimental configurations.
+Their construction is formalized in the Listing \ref{listing-experiment-chaining-ihs}.
+The construction heuristic will be \heu{Random} with pool size 10.
+All the ratios are set to $0.1$ for the time being.
+The termination criterion is set to limit the total runtime to 10 seconds and (potentialy) infinite iterations.
+
+Data gathered will be traces like the one in Appendix \ref{appendix-trace} - after each iteration, the time taken so far and the quality of incumbent solution is noted.
 
@@
     \FOR{$i = 1 \to 20$}
-      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih, limit = 10\}$
+      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = ih, limit = 10 seconds\}$
     \ENDFOR
@@
 
-TODO result graphs - just a \textbf{commented} sample
+Resulting traces for each data set can be summarized in graphs like the one for \heu{100-100} in Figure \ref{image-experiment-chained-ihs-s1}. This one deserves more explanation than usual.
 
-\dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p} boring - finds the optimum in the first step
+\begin{figure}
+  \caption{Chained IHs - \dataset{100-100} results for Strategy 1}
+  \label{image-experiment-chained-ihs-s1}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/chained-ihs-s1}
+\end{figure}
 
-TODO choose the best - see how many times in the 20 runs did they find optimum
+X and Y axes represent the time and quality, as usual.
+Each line represents one run of the scenario (metaheuristic) in the following way: the N\textsuperscript{th} break in the line (i.e.
+the N\textsuperscript{th} data point) is the partial result after the N\textsuperscript{th} step of the scenario.
+Its X position denotes the absolute time in which this step finished, and its Y position represents the incumbent solution quality after this step.
+Every time a line disappears before reaching 10 seconds it means that this metaheuristic run found the optimum before the 10 second mark.
+There is an obvious repetitive regularity in the shape of each line, this corresponds to the fact that there is a finite number of IHs in this scenario (5 of them in Scenario 1) which repeat over time.
+The obvious similarity between different lines corresponds to the fact that each run is from the same scenario, and over time, they do the same steps.
+
+We can see effects of different IHs from this graph:
+\begin{itemize}
+	\item Every $(1 + 5k)$\textsuperscript{th} and $(3 + 5k)$\textsuperscript{th} step is a \heu{RandomRemove}, and each time this happens there is a rather sharp drop in quality
+	\item Every $(2 + 5k)$\textsuperscript{th} step is a \heu{Mutation}, and there is a consistent increase in quality each time.
+	\item Every $(4 + 5k)$\textsuperscript{th} step is a \heu{Crossover}, and each time it happens there is a consistent increase, yet smaller than with \heu{Mutation}.
+	\item Every $5k$\textsuperscript{th} step is a \heu{RemoveWorst}, and as expected, this removes the worst solution not touching the best ones that decide the incumbent quality.
+The line thus stays flat every time it happens.
+\end{itemize}
+
+In this particular example there is only 1 run out of 10 that does not finish (find optimum) under the 10 second mark.
+
+There are two more graphs like this in Figures \ref{image-experiment-chained-ihs-s2} and \ref{image-experiment-chained-ihs-s3} for comparison, capturing Strategy 2 and Strategy 3 respectively working on the same data set, \dataset{100-100}.
+Describing them in detail is unfortunately outside of the scope of this work.
+
+\begin{figure}
+  \caption{Chained IHs - \dataset{100-100} results for Strategy 2}
+  \label{image-experiment-chained-ihs-s2}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/chained-ihs-s2}
+\end{figure}
+
+\begin{figure}
+  \caption{Chained IHs - \dataset{100-100} results for Strategy 3}
+  \label{image-experiment-chained-ihs-s3}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/chained-ihs-s3}
+\end{figure}
+
+It is now necessary to assess which of the strategies perform the best.
+We shall take a look at the different data sets.
+Easily we can discard \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}, because the optimum is found in the very first step.
+Let us now introduce a metric for assessment of a strategy: namely, how many times of the 20 runs did it find the optimum.
+Results of this are summarized in Table \ref{table-experiments-chained-ihs-tweaking}.
 
@@
 
-Strategies 1 and 3 are very similar, we shall choose \textbf{1} for its simplicity to tune the parameters.
+Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs.
+Highlighted are the strategies that performed best on that data set.
+We see that Strategy 1 and 3 are very similar in performance.
+We shall nonetheless choose Strategy \textbf{1} as the winner for its simplicity.
+Now we can tune its parameters.
 

### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

@@
 Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs.
-Highlighted are the strategies that performed best on that data set.
+Highlited are the strategies that performed best on that data set.
 We see that Strategy 1 and 3 are very similar in performance.
@@
 
-TODO remind what Strategy 1 is, list all ratios we will be using
-
-TODO we will be gathering for each run the following: what were the parameters, how long did it take and whether it found optimum
+A short reminder: Scenario 1 consists of \heu{Random} as the CH and the following IHs: \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
+
+The parameters we can tune in this scenario are the ratios in \heu{RandomRemove} (possibly 2 of them, as there are 2 instances in use), \heu{Mutation} and \heu{Crossover}.
+We shall not tune the time limits in \heu{Mutation} and \heu{Crossover} and leave them set to 1 second.
+This presents us with a 3-dimensional space of parameters, where we want to find a combination best suited for our test data sets.
+We will sample this space by taking a total of 45 configurations of the aforementioned ratios.
 
@@
 \end{center}
+
+This experimental set will consist of 45 ratio combinations * 6 data sets * 50 iterations = 13500 experimental configurations.
+CH will be \heu{Random} with pool size of 10.
+IHs will be the ones from Scenario 1, with their ratios set to one of the 45 combinations produced in the following way.
+\begin{itemize}
+	\item \heu{RandomRemove} ratio will be from $\{0, 0.05, 0.1, 0.2, 0.5\}$
+	\item \heu{Mutatio} ratio will be from $\{0.05, 0.1, 0.2\}$
+	\item \heu{Crossover} ratio will be from $\{0.05, 0.1, 0.2\}$
+\end{itemize}
+The process of creating the configurations is captured in Listing \ref{listing-experiment-chained-ihs-tuning}.
+We will be gathering the following information for each run: what were the parameters, how long did the run take and whether it found optimum.
 
@@
 
+After averaging the data we get a large result table, an excerpt from which is in Table \ref{table-experiments-chained-ihs-tweaking}.
+In the left part are the ratio values, in the right part averaged running times for each data set.
+Highlited are the shortest times for each set.
+Only rows containing at least one such shortest time are presented.
+
 \begin{table}
@@
     \end{tabular}
-  \bigskip
-  First 3 columns are ratios for specified heuristics. Remaining columns are average run times in ms. Bold are the best times in that column.
 \end{table}
 
-TODO interpretation - we calculated average runtimes for each combination of ratios and each file. Then we picked the ratio combination that yielded the lowest average runtimes.
-
-RR = 0.5 (strong diversification), MUT = 0.05 (only a few fixed), CX = 0.2 (looking for commonalities among 1/5th of the pool)
-
-TODO compare this to pure GLPK run - for example for \dataset{100-1000}, we got from TODO (90+ on average) seconds to on average TODO seconds (when the 10 second limit is lifted).
+It is now necessary to pick one ratio combination as the best one.
+It is $(RR = 0.5, MUT = 0.05, CX = 0.2)$ for the following reason: it is the best one for sets \dataset{100-100} and \dataset{OVA3} and second best for \dataset{100-200}, \dataset{OVA1} and \dataset{OVA2}.
+Only the \dataset{100-1000} does not profit from these settings.
+
+Now to interpret ratios in the best combination.
+\heu{RandomRemove} ratio of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
+This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima.
+\heu{Mutation} ratio of 0.05 means only around 5\% of AMs in the incumbent solution will be fixed for the next GLPK optimization.
+\heu{Crossover} ratio of 0.2 means that around 1/5\textsuperscript{th} of ID sets in the pool (randomly chosen) will be scanned for common AMs.
+
+All of the ratios in the best combination are at one end of the range we chose from them.
+As a future work option it is possible to start moving these ratios even more in their preferred way, possibly removing \heu{Mutation} in the process.
+
+TODO compare this to pure GLPK run - for example for \dataset{100-200}, we got from TODO seconds to on average TODO seconds (when the 10 second limit is lifted).
 
@@
 
-TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".
-
-First of all: if we have the time, it is best to just let the GLPK run.
-We will find the optimum in this way.
-And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.
-
-Second: if we don't have enough time, we should do Strategy 1. In under 10 seconds we will have very good results (often the optimum).
-
-% TODO if possible, write down in lessons learned:
+After asking and answering a lot of questions related to the overall system behavior, parameter effects and various heuristic combinations is now the time to summarize our results and draw conclusions.
+
+The first fact is that if we have the time available, it is best to just let the GLPK run.
+It will find the optimum eventually, even though this might take minutes or hours to complete.
+For many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter how long it takes.
+
+Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1.
+In under 10 seconds we will have very good results, often the optimum.
+
+Overall, it is always good to ignore the simple text data nodes, as it will improve the total search time.
+
+%\section{Lessons Learned}
+
+%This short section will list a few lessons that were learned during the development of \jmodule{IDSetSearch} module for jInfer and experimenting with it.
+
+%The first lesson is that when there is an experiment being run in a few nested loops (see any experiment set construction listing), one of which is \textit{iterations}, this one should be the outermost one.
+This is because if the experiment somehow fails, TODO
+
 % - always allow for interruptions
-% - in experiments, iterations should be the outermost loop
 % - when representing an object, don't be lazy and create a class for it (don't use pairs, lists with specific data at specific locations, etc)

### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

@@
   \hline
-  Machine           & Dual Core \\
   Input data        & all official and sized test data sets \\
@@
   $\alpha$, $\beta$ & not applicable \\
+  CH                & not applicable \\
+  IHs               & not applicable \\
   \hline
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & \dataset{100-500} \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Glpk} \\
+  IHs               & $\emptyset$ \\
   \hline
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & all sized test data sets \\
@@
   \hline
-  Machine           & Dual Core \\
   Input data set    & all official test data sets \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Random}, \heu{Fuzzy}, \heu{FIDAX} \\
+  IHs               & $\emptyset$ \\
   \hline
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{FIDAX} \\
+  IHs               & \heu{Hungry} or $\emptyset$ \\
   \hline
@@
 
+% TODO re-run on Dual Core!
+
 We shall now try to find the best standalone CH, that is the CH that finds on average the best solutions when run without any IHs.
@@
   \hline
-  Machine           & Quad Core \\
   Input data        & all official test data sets \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & various \\
+  IHs               & $\emptyset$ \\
   \hline
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Glpk} \\
+  IHs               & \heu{Crossover}, \heu{Hungry}, \heu{Local Branching}, \heu{Mutation} \\
   \hline
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Random} or \heu{Glpk} \\
+  IHs               & \heu{Mutation} \\
   \hline
@@
 
+% TODO this subsection does not play well with the ToC - how to put something else in the ToC than is here?
 \subsection{Various $\alpha$, $\beta$}
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & realistic + converted official test data sets \\
@@
   $\alpha$, $\beta$ & $\{0.1, 0.25, 0.5, 0.75, 1\} \times \{0.1, 0.25, 0.5, 0.75, 1\}$ \\ % TODO this is not the best way to write it
+  CH                & \heu{Glpk} \\
+  IHs               & $\emptyset$ \\
   \hline
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & \dataset{XMA-p} \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Glpk} \\
+  IHs               & not applicable \\
   \hline
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & all official test data sets \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Random} \\
+  IHs               & various \\
   \hline
@@
   \hline
-  Machine           & Dual Core \\
   Input data        & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3} \\
@@
   $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Random} \\
+  IHs               & \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots $ \\
   \hline

### 2011-09-17 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress

@@
 
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do better than just let the solver work: by employing various heuristics.
+At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do more than just let the solver work: by employing various heuristics.
 
-Now it is the time to move our ideas into reality and test their feasibility.
-But before we start talking about the experiments themselves, we should try to formulate our aim, what we will be trying to establish.\\
+Now we shall attempt to implement these ideas. But before we describe the experiments themselves, we should try to formulate our aim.\\
 
-First of all, we would like to get an idea of how the whole system and its components behave.
+First of all, we will try to get understanding of how the whole system and its components behave.
 We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
-Even though we shan't hope for them to be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-Preferably, in the end we should have at least some intuition into what will happen if we try X Y Z, before trying it.
+They probably will not be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
+In the end we should have at some intuition about the system.
 
-Second, as we will be introducing a few broad classes of XML data to run ID set search on, we would like to find out which types are best handled by which configurations and settings.
-This will allow us to pick the right heuristic once we see new data.
+Second, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
 
-Also, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
-
-And in the end, we should be able to formulate some kind of general recommendation in form ``if you see this kind of data, do that".\\
+And in the end, we should be able to formulate some kind of general recommendations regarding the problem of finding ID sets.\\
 
@@
 
-Let us now talk about the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
+The first section will discuss the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
 
@@
 
-A short reasoning for this choice: realistic - of course, we want to see the performance in cases taken from the real world.
-The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as text nodes % TODO is this name consistent with what I've been defining?
-instead of attributes.
-We might try to look at such data, convert some of these ``suspicious" values to attributes (e.g.
+A short reasoning for this choice: realistic - we want to see the performance in cases taken from the real world.
+The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as simple text nodes instead of attributes.
+We will try to convert some of these values to attributes (e.g.
 using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will really put our heuristics to the test.
-This is because the realistic data often prove to be quite easy to solve - the list of candidate AMs is usually too short to be hard to solved to optimality.\\
+And finally, we create completely artificial data to create inputs that will put our heuristics in stress.
+This is because the realistic data often prove to be too easy to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.\\
 
-To understand these data sets, we will talk a little about their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set on a graph.
-Therefore, it might be interesting to actually see the graphs of these data sets and have some kind of numbers associated with them.
+To understand these data sets we will discuss their origin and \textit{graph representation}. As mentioned earlier, % TODO link
+the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set in a graph.
+Therefore it is interesting to actually see the graphs of these data sets and understand some related metrics.
 
@@
 
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum % TODO footnote explaining that we got the optimum by running GLPK w/o time limit
-for $\alpha = \beta = 1$ and the numbers of vertices and edges in aforementioned graphs.
+The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ (found by running the \heu{Glpk} heuristic without a time limit) and the numbers of vertices and edges in aforementioned graphs.
 
 \subsection{Realistic data}
+\label{section-realistic-data}
 
 From 3 different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graphs can be seen in Figure \ref{image-experiments-data-realistic}.
+Their summary is the Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
 Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
@@
 
-TODO their DTD/XSD - do we have ID attributes?
-
-TODO verify they satisfy their schemas. Converted too.
+% TODO their DTD/XSD - do we have ID attributes?
 
@@
 
-To interpret the data a little: \dataset{OVA*} sets have quite interesting and perhaps a bit challenging graphs, but are not too big. We can consider them to be the ``typical" representants.
+To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but are relatively small. We can consider them to be the ``typical" representants.
 
-On the other hand, the \dataset{XMA-*} sets are quite huge, but trivial: their only candidate AM will just get picked and that will be the end of the heuristic.
-We will see the performance of the other components of the whole system, dealing with huge data that are rather simple in the end.
+On the other hand, the \dataset{XMA-*} sets are relatively huge, but trivial: their only candidate AM will just get picked and that the heuristic will end.
+We will see the performance of the other components of the whole system, such as loading the data sets into memory representations.
 
-Finally, the \dataset{XMD} set is quite big and at the same time has non-trivial graph representation. We should see a performance more balanced between processing and finding the ID set here.
+Finally, the \dataset{XMD} set is relatively big and at the same time has non-trivial graph representation.
+In this case we should see a performance more balanced between processing and finding the ID set.
 
@@
 
-We found 2 data sets to convert, \dataset{MSH} and \dataset{NTH}.
+We used 2 data sets to convert, \dataset{MSH} and \dataset{NTH}.
 Unfortunately, the same problem with disclosure as in the previous case applies here.
 None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs can be seen in Figure \ref{image-experiments-data-converted}.
+Their summary is the Table \ref{table-experiments-data-converted}, their graphs are in Figure \ref{image-experiments-data-converted}.
 
-TODO schemas?
-
-To address the conversion: in the case of \dataset{MSH} we found 2 elements with values looking suspiciously like a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
+To address the conversion: in the case of \dataset{MSH} we found 2 elements with values resembling a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
 In the case of \dataset{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
 
-Why is this useful at all?
-Recall TODO link, where we said that ID attributes are a special case of XML keys.
-We can use this approach then to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
+This approach is useful, because as stated in TODO link, ID attributes are a special case of XML keys.
+We can use this approach to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
 
@@
 
-Again to do a little interpretation: in the case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
+Again interpretation: in the case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
 This is then the case similar to \dataset{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
@@
 Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
-This means we have quite a big set with relatively simple work to be done by the heuristics.
-This is about the change.
+This means we have a relatively large set with considerably simple work to be done by the heuristics.
 
@@
 
-As soon as we started experimenting with the data coming from the real world, it was obvious that they present no real challenge to our heuristics.
-After we build the model, we get the most complex graphs of some 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-We needed to create data that would really put pressure on our heuristics, put them to a real test.
-Our solutions is to approach from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
-Why not create the XML data to contain a more complex graph in the first place?
-What if we could specify how many vertices and edges we want, and get the according XML?
+As soon as we started experimenting with the data coming from the real world, it was obvious that they are not complex enough.
+After we build the model, we get the most complex graphs of 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
+Our solution is to approach the problem from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
+We will create the XML data to contain a more complex graph with a specific number of vertices and edges.
 
-This is actually simple. Consider the following excerpt from an XML file:
+Consider the following excerpt from an XML file:
 
@@
 
-We want to create a graph with around $v$ vertices and $e$ edges.
+Our aim is to create a graph with around $v$ vertices and $e$ edges.
 First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values won't conflict with any others.
-Second, for each of the $e$ edges we pick two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
+To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values will not conflict with any others.
+Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
 This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
@@
 
-With this process in place we can create as much data as we wish with any combination of $v$ and $e$ we want.
+With this process it is possible to create as much data as needed, with any combination of $v$ and $e$ requested.
 
-There is one interesting characteristic that can describe random graphs like this, and that is the \textit{density}.
-This can be defined in various ways, we will use two different interpretations for a while.
-The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we'd get the average degree of the vertices).
+There is one characteristic that can describe random graphs like this, and that is the \textit{density}.
+This can be defined in various ways, we will use two different interpretations.
+The first is $\frac{|E|}{|V|}$, that is, how many edges are there for one vertex (multiplied by 2 we would get the average degree of the vertices).
+
 The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
@@
 
-Also, we will need data of comparably similar characteristics but varying size, to study the effects of size on the run times of experiments.
-For this reason we created 11 more sets, from \dataset{0-0} as the trivial one to \dataset{100-500} as the biggest one.
+Also, we will need data of comparably similar characteristics but varying size to study the effects of size on the run times of experiments.
+For this reason we created 11 more sets, from \dataset{0-0} as the trivial one to \dataset{100-500} as the largest one.
 From now on, these will be referred to as \textit{sized test data (sets)}.
@@
 
-TODO this is on the DVD.
+The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
+Some of the graph representations can be seen in Figure \ref{image-experiments-data-artificial}.
 
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
-Some graphs can be seen in Figure \ref{image-experiments-data-artificial}.
-
-While studying the tables you might notice that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set (proof is homework).
-% TODO too much?:)
+While studying the tables it becomes obvious that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
+This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set.
 Because of the so-called Birthday paradox, % TODO link
@@
 
-To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one, which should work very nice with our heuristics.
-The $|V|$ and $|E|$ numbers are orders of magnitude higher than with any realistic (or converted) data set we are using, and this will put real stress on the heuristics, as we shall see in the next section.
+To interpret Tables \ref{table-experiments-data-artificial} and \ref{table-experiments-data-artificial-size}: we get 3 sets of different sizes and densities in the first one.
+The $|V|$ and $|E|$ numbers are orders of magnitude higher than in any realistic (or converted) data set we are using.
 
-In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was quite nicely achieved.
+In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was indeed achieved.
 There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
@@
 Exploration of this phenomenon is among the possibilities of future work of this thesis.
+
+On a final note, the artificial data we used in experiments can be found on the DVD enclosed with this work.
 
@@
 
-We will be using two different machines to run the experiments.
-Every experimental result will have explicitly stated the machine it was performed on, and we shall make sure to run different experiments that can benefit from side-by-side comparison on the same machine.
-
-\subsubsection{Quad Core Machine}
-
-\begin{verbatim}
-Intel Core 2 Quad processor @ 2.83 GHz
-8 GB DDR2 RAM
-Windows 7 SP1 64bit
-TODO
-TODO
-GLPK version TODO
-\end{verbatim}
-
-\subsubsection{Dual Core Machine}
+We will be using the following machine when conducting our experiments.
 
@@
 
-It is impossible to completely shield an experiment from the influence of the environment, but we should try our best.
-First of all, NetBeans running the experiments is the only relevant program running in the system at that time, as far as this is possible.
-Unfortunately, the NetBeans itself is quite a large environment with a life on its own, and we would most certainly get more reliable results, if we could run our experiments outside it.
+We will attempt to shield our experiment from the influence of the environment as much as reasonably possible.
+First of all, NetBeans running the experiments is the only relevant program running in the system while the experiments are performed.
+Unfortunately, NetBeans itself is quite a large environment with a life of its own, and we would most certainly get more reliable results if we could run our experiments outside of it.
 This improvement is left for the future work.
@@
 Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
-% TODO discuss caches and the order in which we run the experiments in one set?
 
@@
 
-TODO
+Whenever it is necessary to measure the duration of an operation, we will use the \texttt{System.nanoTime()} built-in function.
+The result cannot be interpreted in an absolute manner, but by subtracting the time at the start from the time at the end, we can get a reasonably reliable measurement.
+We then have to divide it by one million to obtain the number of milliseconds passed.
 

### 2011-09-18 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress

@@
 
-TODO link BasicIGG to see how we get IG
+This section will briefly describe the process by which an input data set is processed to obtain the AM model as described in TODO link.\\
 
-TODO describe how we get the model
+An input data set is essentialy a single XML file on the filesystem, however there is a straightforward extension to multiple files conforming to the same schema.
+The first step in this process is to use jInfer's module \jmodule{BasicIGG} module (see \cite{basiciggdoc} for details) to obtain a list of rules - an \textit{initial grammar} (IG).
+\nomenclature{IG}{Initial Grammar}
+Please see \cite{archdoc} for detailed specification of IG format.\\
+
+The second step is to convert the grammar into the AM model.
+This is done by scanning it in a linear fashion and retrieving a so-called \textit{flat} representation.
+This consists of a list of tuples in the following format.
+
+\begin{center}
+\textit{(element name, attribute name, attribute value)}
+\end{center}
+
+There is a tuple for every attribute node with a value found in the initial grammar.
+Note that the information about the context in which the element was originally found is lost - but this is not a problem with regard to the definition of XML ID attributes.
+Furthermore, tuples in flat representation do not need to be unique.\\
+
+The model now has to be able to return the list of all attribute mappings and their respective images.
+This is achieved by simply grouping the flat representation by the pair \textit{(element name, attribute name)} and aggregating all attribute values for each such pair.
+Another responsibility of the model is to return the list of \textit{types} - that is simply the list of unique \textit{element name}s.
+
+\subsubsection{Example}
+
+This example will present the whole process, from an input XML file to the model and its AMs and types.
+
+Consider the following XML file fragment.
+\begin{verbatim}
+<x>
+  <y a="1" b="2"/>
+  <y a="3" c="4"/>
+  <y/>
+  <z a="1"/>
+</x>
+\end{verbatim}
+
+Its IG representation is the following set of IG rules.
+
+\begin{eqnarray*}
+	x & \to & y, y, y, z \\
+	y & \to & @a, @b \\
+	y & \to & @a, @c \\
+	y & \to & empty\_concatenation \\
+	z & \to & @a \\
+\end{eqnarray*}
+
+The flat representation will consist of the following set of tuples.
+
+\begin{eqnarray*}
+(y, a, 1) \\
+(y, b, 2) \\
+(y, a, 3) \\
+(y, c, 4) \\
+(z, a, 1) \\
+\end{eqnarray*}
+
+Attribute mappings in this model will be $(y, a)$, $(y, b)$, $(y,c)$ and $(z,a)$. Their images will be $(1,3)$, $(2)$, $(4)$ and $(1)$, respectively. The list of types in this model will be $(y,z)$.
 
@@
 
-\subsection{Measuring the time}
+\subsection{Measuring the Time}
 
@@
 
-\subsection{Obtaining the results}
+\subsection{Obtaining the Results}
 
@@
 An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot, which we use to plot charts found in this work.
+Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot \cite{gnuplot}, which we use to plot charts found in this work.
 
-TODO explain boxplots, how to read them
+\subsection{Reading Boxplots}
+
+To present a set of measurements obtained by iteratively running an experiment we shall prominently use the \textit{boxplot} chart.
+Because we use boxplots produced by GnuPlot, let us quote its manual (\cite{gnuplot-manual}) for the exact definition.
+
+\begin{quote}
+Quartile boundaries are determined such that 1/4 of the points have a value equal or less than the first quartile boundary, 1/2 of the points have a value equal or less than the second quartile (median) value, etc.
+A box is drawn around the region between the first and third quartiles, with a horizontal line at the median value.
+Whiskers extend from the box to user-specified limits.
+Points that lie outside these limits are drawn individually.
+\end{quote}
+
+The ``user-specified limits" of whiskers are set to default value, let us quote again from the manual.
+
+\begin{quote}
+By default the whiskers extend from the ends of the box to the most distant point whose y value lies within 1.5 times the interquartile range.
+\end{quote}

### 2011-09-18 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

@@
 
-\subsection{Grammar and model generation}
+\subsection{Grammar and Model Generation}
 \label{section-grammar-model-timing}
@@
 There will be no CHs or IHs.
-We will be gathering the timing data for IG extraction and model generation in GnuPlot format.
+We will be gathering the timing data for IG extraction and model generation in GnuPlot format.\\
 
@@
 
-\subsubsection{GLPK interface timing}
+\subsubsection{GLPK Interface Timing}
 
@@
 A related problem is how long it takes to create input for GLPK and then parse its results.
-We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.
+We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.\\
 
@@
 
-\subsection{GLPK: native vs. Cygwin}
+\subsection{GLPK: Native vs. Cygwin}
 
@@
 The only data we gather in the GnuPlot file are the final qualities (weights).
-The data set used is \dataset{100-500} as the biggest one in sized test data.
+The data set used is \dataset{100-500} as the biggest one in sized test data.\\
 
 \begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 1}
+\caption{GLPK: Native vs. Cygwin Set Generation 1}
 \label{listing-experiment-glpk-native-vs-cygwin-1}
@@
 This will be repeated in 50 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
-There will again be no IH, the only data we will collect are the times of the CH run in each case.
+There will again be no IH, the only data we will collect are the times of the CH run in each case.\\
 
@@
 \subsection{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX}}
+\label{section-experiments-random-fuzzy-fidax}
 
@@
 
-We will be gathering the running time of the CH itself and quality of the best solution found for GnuPlot.
+We will be gathering the running time of the CH itself and quality of the best solution found for GnuPlot.\\
 
 \begin{algorithm}
-\caption{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX} set generation}
+\caption{\heu{Random} vs. \heu{Fuzzy} vs. \heu{FIDAX} Set Generation}
 \label{listing-experiment-random-fuzzy-fidax}
@@
 We will try all official data sets, first with empty IH, second with \heu{Hungry} as IH.
-We will gather the qualities in each case and see whether there is any improvement.
+We will gather the qualities in each case and see whether there is any improvement.\\
 
@@
 \begin{table}
-  \caption{ID sets in \heu{FIDAX} versus \heu{FIDAX} + \heu{Hungry}}
+  \caption{ID Sets in \heu{FIDAX} Versus \heu{FIDAX} + \heu{Hungry}}
   \bigskip
@@
 
-\subsection{Best standalone CH}
+\subsection{Best Standalone CH}
 
@@
 See the Listing \ref{listing-experiment-best-standalone-ch} for details.
-This time we are not interested in run times, only in qualities which we shall gather in a format for GnuPlot.
+This time we are not interested in run times, only in qualities which we shall gather in a format for GnuPlot.\\
 
 \begin{algorithm}
-\caption{Best Standalone CH set generation}
+\caption{Best Standalone CH Set Generation}
 \label{listing-experiment-best-standalone-ch}
@@
 
-\subsection{Best IH for Glpk}
+\subsection{Best IH for \heu{Glpk}}
 
@@
 This experimental set will contain 6 data sets * 50 iterations * 4 IHs = 1200 experimental configurations.
-Note that we are using only the most challenging data sets, as the combination of \heu{Glpk} as CH and any other IH is already an overkill for easier data sets.
+Note that we are using only the most challenging data sets, as the combination of \heu{Glpk} as CH and any other IH is already an overkill for easier data sets.\\
 
 \begin{algorithm}
-\caption{Best IH for \heu{Glpk} set generation}
+\caption{Best IH for \heu{Glpk} Set Generation}
 \label{listing-experiment-best-ih-for-glpk}
@@
 Experimental set will consist of 6 data sets * 50 iterations * 2 CHs = 600 experimental configurations, see Listing \ref{listing-experiment-ch-for-mutation}.
-We shall collect the eventual quality after running both the CH and the IH in format suited for GnuPlot.
+We shall collect the eventual quality after running both the CH and the IH in format suited for GnuPlot.\\
 
 \begin{algorithm}
-\caption{\heu{Random} as CH set generation}
+\caption{\heu{Random} as CH Set Generation}
 \label{listing-experiment-ch-for-mutation}
@@
 Pool size as well as iteration count will be 1.
-We are noting the actual ID set found by the run of the heuristic.
+We are noting the actual ID set found by the run of the heuristic.\\
 
 \begin{algorithm}
-\caption{Various values of $\alpha$ and $\beta$ set generation}
+\caption{Various Values of $\alpha$ and $\beta$ Set Generation}
 \label{listing-experiment-various-betas}
@@
 \begin{table}
-  \caption{Different ID sets found for \dataset{OVA1}}
+  \caption{Different ID Sets Found for \dataset{OVA1}}
   \bigskip
@@
 \begin{table}
-  \caption{Effect of $\alpha$, $\beta$ on ID set found for \dataset{OVA1}}
+  \caption{Effect of $\alpha$, $\beta$ on ID Set Found for \dataset{OVA1}}
   \bigskip
@@
 \begin{table}
-  \caption{Effect of $\alpha-ratio$ on ID set found for \dataset{OVA1}}
+  \caption{Effect of $\alpha-ratio$ on ID Set Found for \dataset{OVA1}}
   \bigskip
@@
 
-\subsection{Ignoring text data}
+\subsection{Ignoring Text Data}
 
@@
 After the first 50 iterations we will turn on the option to ignore the simple text node data and run the same 50 iterations again.
-We will be collecting the grammar extraction (GE) and model creation (MC) times as in the experiment in Section \ref{section-grammar-model-timing}.
+We will be collecting the grammar extraction (GE) and model creation (MC) times as in the experiment in Section \ref{section-grammar-model-timing}.\\
 
 \begin{algorithm}
-\caption{Ignoring text data set generation}
+\caption{Ignoring Text Data Set Generation}
 \label{listing-experiment-ignore-text-data}
@@
 
-Data gathered will be traces like the one in Appendix \ref{appendix-trace} - after each iteration, the time taken so far and the quality of incumbent solution is noted.
+Data gathered will be traces like the one in Appendix \ref{appendix-trace} - after each iteration, the time taken so far and the quality of incumbent solution is noted.\\
 
 \begin{algorithm}
-\caption{Chaining IHs set generation}
+\caption{Chaining IHs Set Generation}
 \label{listing-experiment-chaining-ihs}
@@
 \begin{figure}
-  \caption{Chained IHs - \dataset{100-100} results for Strategy 1}
+  \caption{Chained IHs - \dataset{100-100} Results for Scenario 1}
   \label{image-experiment-chained-ihs-s1}
@@
 
-In this particular example there is only 1 run out of 10 that does not finish (find optimum) under the 10 second mark.
-
-There are two more graphs like this in Figures \ref{image-experiment-chained-ihs-s2} and \ref{image-experiment-chained-ihs-s3} for comparison, capturing Strategy 2 and Strategy 3 respectively working on the same data set, \dataset{100-100}.
-Describing them in detail is unfortunately outside of the scope of this work.
+In this particular example there is only 1 run out of 10 that does not finish (find optimum) under the 10 second mark.\\
+
+There are two more graphs like this in Figures \ref{image-experiment-chained-ihs-s2} and \ref{image-experiment-chained-ihs-s3} for comparison, capturing Scenario 2 and Scenario 3 respectively working on the same data set, \dataset{100-100}.
+Describing them in detail is unfortunately outside of the scope of this work.\\
 
 \begin{figure}
-  \caption{Chained IHs - \dataset{100-100} results for Strategy 2}
+  \caption{Chained IHs - \dataset{100-100} Results for Scenario 2}
   \label{image-experiment-chained-ihs-s2}
@@
 \begin{figure}
-  \caption{Chained IHs - \dataset{100-100} results for Strategy 3}
+  \caption{Chained IHs - \dataset{100-100} Results for Scenario 3}
   \label{image-experiment-chained-ihs-s3}
@@
 
-It is now necessary to assess which of the strategies perform the best.
+It is now necessary to assess which of the scenarios perform the best.
 We shall take a look at the different data sets.
 Easily we can discard \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}, because the optimum is found in the very first step.
-Let us now introduce a metric for assessment of a strategy: namely, how many times of the 20 runs did it find the optimum.
+Let us now introduce a metric for assessment of a scenario: namely, how many times of the 20 runs did it find the optimum.
 Results of this are summarized in Table \ref{table-experiments-chained-ihs-tweaking}.
@@
 \begin{table}
-  \caption{Performance of various IH chains}
+  \caption{Performance of Various IH Chains}
   \bigskip
@@
   \begin{tabular}{l || c | c | c}
-    Dataset & Strategy 1 & Strategy 2 & Strategy 3 \\
+    Dataset & Scenario 1 & Scenario 2 & Scenario 3 \\
     \hline
@@
     \end{tabular}
-  \bigskip
-  Each cell represents the number of times (out of 20) the strategy found the optimum - more is better.
 \end{table}
 
-Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs.
-Highlited are the strategies that performed best on that data set.
-We see that Strategy 1 and 3 are very similar in performance.
-We shall nonetheless choose Strategy \textbf{1} as the winner for its simplicity.
+Each cell contains the number of times the scenario found optimum in the data set, out of 20 runs.
+Highlited are the scenarios that performed best on that data set.
+We see that Scenario 1 and 3 are very similar in performance.
+We shall nonetheless choose Scenario \textbf{1} as the winner for its simplicity.
 Now we can tune its parameters.
 
-\subsubsection{Improving Strategy 1}
+\subsubsection{Improving Scenario 1}
 
@@
 The process of creating the configurations is captured in Listing \ref{listing-experiment-chained-ihs-tuning}.
-We will be gathering the following information for each run: what were the parameters, how long did the run take and whether it found optimum.
+We will be gathering the following information for each run: what were the parameters, how long did the run take and whether it found optimum.\\
 
 \begin{algorithm}
-\caption{Chained IHs - Tuning set generation}
+\caption{Chained IHs - Improving Scenario 1 Set Generation}
 \label{listing-experiment-chained-ihs-tuning}
@@
 
-After averaging the data we get a large result table, an excerpt from which is in Table \ref{table-experiments-chained-ihs-tweaking}.
+After averaging the data we get a large result table, an excerpt from which is in Table \ref{table-experiments-chained-ihs-tweaking-s1}.
 In the left part are the ratio values, in the right part averaged running times for each data set.
@@
 \begin{table}
-  \caption{Performance of Strategy 1 depending on parameters - excerpt}
-  \bigskip
-  \label{table-experiments-chained-ihs-tweaking}
+  \caption{Performance of Scenario 1 Depending on Parameters - Excerpt}
+  \bigskip
+  \label{table-experiments-chained-ihs-tweaking-s1}
   \centering
@@
 \heu{RandomRemove} ratio of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
-This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima.
+This amounts to a very strong diversification tendency and keeps the scenario from stalling in local optima.
 \heu{Mutation} ratio of 0.05 means only around 5\% of AMs in the incumbent solution will be fixed for the next GLPK optimization.
@@
 
-Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1.
+Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Scenario 1.
 In under 10 seconds we will have very good results, often the optimum.

### 2011-09-18 doc/theses/vitasek/appendix-b-iss.tex
Thesis: in progress

@@
-\chapter{IDSetSearch}
+\chapter{\jmodule{IDSetSearch}}
 \label{appendix-iss}
 
-TODO talk a bit about the structure of ISS module.
+This appendix will shortly describe the \jmodule{IDSetSearch} jInfer module.
+As the name suggests, its main purpose is to find ID and IDREF sets and provide attribute statistics in general for grammars originating from any stage of XML schema inference.
+Virtually every piece of code that was added to jInfer in the course of creating this thesis is contained in this module.\\
 
-TODO name of the module, its place in jInfer, interface with the framework (service providing)
+From jInfer's point of view, this module resides in codebase \texttt{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.iss} and is a service provider for \texttt{cz.\.cuni.\.mff.\.ksi.\.jinfer.\.base.\.inter\.faces.\.ID\.Set\.Search} interface.
+Invoking the \texttt{showIDSetPanel()} method displays a fully-featured window containing all the relevant attribute statistics as well as possibility to find the ID and IDREF sets for a specified grammar.\\
 
-TODO package name root
+Most important packages in \jmodule{IDSetSearch} are the following.
 
-TODO "Attribute Stats"
+\begin{itemize}
+	\item \texttt{objects}, containing the object representation of attribute mappings and AM model.
+	\item \texttt{heuristics.construction}, containing all the CHs hidden behind the \texttt{Con\-struc\-tion\-Heu\-ris\-tic} interface, with sub-packages \texttt{fidax} containing the whole implementation of FIDAX heuristic % TODO link
+	and \texttt{glpk} containing the whole interface to an external GLPK solver. % TODO link
+	\item \texttt{heuristics.improvement}, containing all the IH hidden behind the \texttt{Im\-prove\-ment\-Heu\-ris\-tic} interface.
+	\item \texttt{experiments}, containing everything related to experimenting with these heuristics.
+\end{itemize}
 
-TODO heuristics - interfaces, construction, improvement
+\texttt{Experiment} is a class representing a single experiment with specified input data (encapsulated in \texttt{Test\-Data} interface), settings (encapsulated in \texttt{Ex\-pe\-ri\-ment\-Pa\-ra\-me\-ters}) and a metaheuristic as defined in TODO link.
+Its method \texttt{run()} will launch the metaheuristic, first executing the construction heuristic and then running the specified improvement heuristics in a loop until termination criteria defined in an implementation of \texttt{Ter\-mi\-na\-tion\-Cri\-ter\-ion} are met.
+The quality of a single ID set is measured by an instance of \texttt{Quality\-Measurement}.
+After the experiment finishes, it invokes the \texttt{notify\-Finished()} method.\\
 
-TODO experiments - quality, termination, parameters, experiment sets
+However, experiments are almost never run alone.
+For the purpose of running a whole experimental set there is the \texttt{Experiment\-Set} interface and its abstract implementation \texttt{Abstract\-Experiment\-Set}.
+Its descendants need only to provide a list of \texttt{Experiment\-Parameter}s and looping as well as data collection will be handled for them.
 
-TODO make sure there was a place where we described the model data model:)
+\section{How to Create a New Heuristic}
+
+Decide whether it should be a CH or IH and create a class implementing \texttt{Con\-struc\-tion\-Heu\-ris\-tic} or \texttt{Im\-prove\-ment\-Heu\-ris\-tic}, respectively.
+In each case implement all the \texttt{get*Name()} methods inherited from \texttt{Named\-Module} and then the most important \texttt{start()} method.
+
+In this method use the provided \texttt{Experiment} instance (and \texttt{List<IdSet> feasiblePool} in case of IH) to create a pool of feasible solutions and in the end return it by invoking the \texttt{finished()} method of the provided \texttt{Heuristic\-Callback} parameter.
+
+\section{How to Create a New Experimental Set}
+
+Subclass the \texttt{Abstract\-Experiment\-Set} class, override \texttt{get\-Name()} to provide the name of this set and finally override \texttt{get\-Experiments()} to return the list of \texttt{Experiment\-Parameters} that will constitute this sit.
+
+It is possible to optionaly override any of the following methods: \texttt{notify\-Start()}, \texttt{notify\-Finished()} and \texttt{notify\-Finished\-All()}.
+They will be invoked before running the first experiment, after each experiment run and after all experiments finished, respectively.
+Note that \texttt{notify\-Finished()} already contains logic to output some information regarding the currently finished experiment to a file, but it can be safely overriden without a need to call \texttt{super.\.notify\-Finished()}.

### 2011-09-19 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 Queries may not cover the whole relevant content of the documents.
-For example some elements may not be queried at all, therefore they may not be mentioned in any query \todo[inline]{Tady je ale několik případů - chtělo by popsat.
-1) Element není zapsaný v dotazu, ale dostanu ho výsledku - např. rok anrození autora, ptám-li se na autory.
-2) Element není zapsaný v dotazu, ale procházím ho při vyhodnocování -  např. to //autor
-3) Element je v části dokumentu, která se dotazu vlbec netýká.}.
+Let \texttt{E} be an element occurring in an XML document and let \texttt{Q} be an XQuery query.
+Consider the following cases:
+
+\begin{enumerate}[1)]
+\item \texttt{E} is directly mentioned in \texttt{Q}. For example elements \texttt{bib}, \texttt{book} and \texttt{author} are mentioned in path expression \texttt{/bib/book/author}.
+\item \texttt{E} is not mentioned in \texttt{Q} but it does occur in a result of the evaluation of \texttt{Q}.
+For example path expression \texttt{/bib/book/author} returns elements \texttt{author} along with their subelements \texttt{name}, \texttt{birthdate} and \texttt{nationality}.
+\item \texttt{E} is not mentioned in \texttt{Q} nor it occurs in the result but it is processed by the evaluation of \texttt{Q}.
+For example the evaluation of path expression \texttt{//author} processes amongst others elements \texttt{bib} and \texttt{book}, in a search for elements \texttt{author}.
+\item \texttt{E} occurs in a part of the XML document not related to a query at all.
+\end{enumerate}
+
+How we can obtain any information about element \texttt{E} from query \texttt{Q}?
+In case 1), \texttt{Q} can be directly exploited in the inference process.
+Case 2) requires the result of \texttt{Q} and additional processing of the result.
+Case 3) requires a step-by-step evaluation of \texttt{Q}, so the XML document is also required.
+And in case 4), it is not possible to obtain any information.
+
 In addition, these statements do not express any obligation of occurrence of elements and attributes nor clearly determine multiple occurrence of elements.
-We cannot be sure that queries target nodes actually presented in the XML documents.
+We also cannot be sure that queries target nodes actually presented in the XML documents.
 Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not.
@@
 \subsection{Approach from \cite{Necasky:2009:DXK:1529282.1529414}}
-Paper \todo{link} introduces a method of discovering keys and foreign keys by investigation of joins in queries.
+Paper \cite{Necasky:2009:DXK:1529282.1529414} introduces a method of discovering keys and foreign keys by investigation of joins in queries.
 The basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form.
@@
 
-\subsection{Join of Recursive Data}
-By the term ``recursive data'' are meant XML elements that somehow reference items from the same set.
-\todo[inline]{Tohle je velmi nepřesné.
-Tady asi myslíte reference, ale rekurzivní element je obecně element, kteý obsahuje sám sebe - např.
-kapitola má kapitoly.} Example of a query that operates upon such data follows.
+\subsection{Join of Self-referencing Data}
+By the term ``self-referencing data'' are meant XML elements that somehow reference items from the same set. Example of a query that operates upon such data follows.
 

### 2011-09-19 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 \chapter{Analysis of Recent Approaches}
-\todo[inline]{bibref na prehladovy clanok?}
 Approaches of XML schema inference can be divided by several criteria. A basic division is by the language the resulting schema is written in. Commonly used languages are DTD and XML Schema.
@@
 
-According to my best knowledge at the time of writing there is just one approach of this category.
+According to our best knowledge at the time of writing there is just one approach of this category.
 It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?

### 2011-09-19 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
 
-We should note that the combination best CH - best IH found this way does not necessarily need to be the best one, because we find it using a hungry approach.
+We should note that the combination \textit{best CH - best IH} found this way does not necessarily need to be the best one overall, because we find it using a hungry approach.
 

### 2011-09-19 doc/theses/vitasek/chapter-5-future.tex
Thesis: in progress.

@@
 
-TODO
+A straightforward extension granting the ability to handle more than one input XML file has already been suggested in \cite{fidax}.
+However, it wasn't implemented in this work either, so it still remains an obvious first choice of future work.\\
 
-\begin{itemize}
-  \item extension to more documents at once
-  \item more heuristics and their chaining
-  \item ACO, genetic, ...
-  \item user-friendly experimental interface (choose and setup experiments on the fly in GUI)
-  \item jInfer is opensource, thus easy to extend!
-\end{itemize}
+It is possible (and easy) to add more construction and improvement heuristics, as well as more metaheuristics in which the existing IHs are chained.
+A starting point is in Section TODO link appendix - how to write...\\
+
+As it was mentioned in TODO link, the combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} can be seen as a sort of genetic programming.
+However some modifications would still be necessary to make it a real instance of genetic algorithm metaheuristic.
+
+Likewise it is possible to create an Ant Colony Optimization metaheuristic solving the same problem.
+It would be interesting to see all these metaheuristics compared to each other in a set of comprehensive experiments.\\
+
+The approach used in this work was strictly single-threaded, however there are in principle no limitations to extending this to a parallel, multi-threaded environment.
+For example, creating a pool of initial solutions in \heu{Glpk} construction heuristic can be improved by running several instances of GLPK solver in parallel - as GLPK on its own uses only a single thread to perform the computation.\\
+
+From the point of view of a user - researcher, the current implementation of the experimental framework leaves a lot to be desired.
+As jInfer already contains support for interchangeable and configurable modules, it is possible to create GUI for experiment and experimental set configuration on the fly.\\
+
+jInfer as well as the \jmodule{IDSetSearch} module are open source projects, meaning that anyone wishing to build upon this work can do so easily.

### 2011-09-19 doc/theses/vitasek/appendix-a-jinfer.tex
Thesis: adding the generated PDF to the repository.

@@
 This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
-Please see TODO link jInfer project web for complete information, documentation and download options.
+Please see project web \cite{jinferweb} for complete information, documentation and download options.
 
@@
 The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
-The goal was apparently achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
+The goal was achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
 Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
@@
 This is the representation in which will the structure live until it is used to create the final product - the schema.
-As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (TODO link CFGs).
+As the name suggests, IG is a grammar - an \textit{extended context-free grammar}, to be more precise (see \cite{extendedcfg}).
 As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
 (TODO picture?) IG is used to create the AM model used in this thesis, too.
-jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in (TODO link BasicIGG documentation).
+jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in \cite{basiciggdoc}.
 

### 2011-09-19 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 Ty by tady něškodily všude, u každé té kategorie.}.
+
+\section{On Inference of XML Schema with the Knowledge of an Obsolete One}
+\todo[inline]{spravne zaradit}
+An aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently.
+A XML schema can become obsolete due to changes in a set of XML documents, without capturing these changes in the schema.
+Thus, the schema becames outdated and according to the paper, this case is quite common.
+
+On input, the method is given:
+\begin{itemize}
+\item An original XML schema.
+\item A set of XML documents. Not all have to be valid against the original schema.
+\end{itemize}
+
+And the algorithm consists of two independent steps:
+
+\begin{enumerate}
+\item Correction of the input schema.
+\item Specialization of the input schema.
+\end{enumerate}
+
+In the fist step the input schema is corrected to conform to the whole set of input XML documents.
+This is done by creating a PTA \todo{ref na def} for each production extracted from the input schema and merging it with a respective production from the initial grammar \todo{ref na def}, involving ACO and MDL heuristics.
+
+In the second step, assuming a correct schema it is wanted to specialize the REs involved in the schema with regard to the XML documents, resulting in more precise and readable schema.
+Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.
 

### 2011-09-20 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 Just the queries are utilized in inference, no XML data are used.
-The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} \todo{aj unique?} constructs.
+The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} and \textbf{unique} constructs.
 
-\subsection{Keys and foreign keys}
-The authors introduce a formalism that formalizes XML Schema keys.
+\subsection{Assumptions and Observations}
+To discover keys and foreign keys, the method utilizes element/element joins.
+Assume a query $Q$ that joins a sequence of elements $S_1$ targeted by a path $P_1$ with a sequence of elements $S_2$ targeted by a path $P_2$ on a condition $L_1 = L_2$.
+\todo[inline]{priklad}
+The method is based on assumption that each join is done via key/foreign key pair.
+It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is it's respective foreign key or vice versa.
 
-\todo[inline]{Definovat, co pouzivame: XPath (aj s verziou), prikady, osy, lomitko, ...}
+The author describe two possible cases:
 
-\begin{define}[Key]
-A key is a construct $$(C, P, \{L\})$$ where $C$, $P$, $L$ are XPath paths without predicates that use only \texttt{child} and \texttt{descendant} axes.
-$C$ can be omitted. $(P, \{L\})$ is then equivalent to $(/, P, \{L\})$ and is called \emph{global key}. Otherwise it is called \emph{local key}.
-\end{define}
-
-The key specifies the following condition.
-Let $c$ be an element targeted by $C$ and $p$ and $p’$ be two elements targeted by $P$ from $c$.
-If the value targeted by $L$ from $p$ equals to the value targeted by $L$ from $p’$, then $p$ and $p’$ are the same elements.
-In other words, no two different elements targeted by $P$ from $c$ can have the same value of $L$.
-
-\begin{define}[Foreign key]
-A foreign key is a construct $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ where $(C, P_2, \{L_2\})$ is a key, $P_1$, $L_1$ are XPath paths without predicates that use only \texttt{child} and \texttt{descendant} axes.
-$C$ can be omitted as in the case of keys.
-\end{define}
-
-\todo[inline]{Potřebujeme tyto definice k něčemu dále?}
-
-Let $c$ be an element targeted by $C$ and $p_1$ be a element targeted by $P_1$ from $c$.
-The foreign key specifies that there is an element $p_2$ targeted by $P_2$ from $c$ such that the value targeted by $L_1$ from $p_1$ equals to the value targeted by $L_2$ from $p_2$.
-In other words, each element targeted by $P_1$ from $c$ refers to an element targeted by $P_2$ from $c$ via the pair $L_1$ and $L_2$.
-
-\subsection{Key and Foreign Key Discovery}
-\subsubsection{Joins in queries}
-To discover keys and foreign keys, the described method utilizes element/element joins.
-
-Assume a query $Q$ that joins a sequence of elements $S_1$ targeted by a path $P_1$ with a sequence of elements $S_2$ targeted by a path $P_2$ on a condition $L_1 = L_2$.
-It means that $Q$ joins an element $e_1$ from $S_1$ with an element $e_2$ from $S_2$ if $e_1/L_1$ equals to $e_2/L_2$.
-
-\todo[inline]{priklad?}
-\todo[inline]{Určitě. A hlavně je tento popis mnohem detailnější, přesnější a formaálnější než ty předchozí. Mělo by to být všechno v jednodném stylu.}
-
-Assuming that each join is done via a key/foreign key pair, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.
-
-The authors introduce two following observations:
-\begin{enumerate}
+\begin{itemize}
 \renewcommand{\theenumi}{(O\arabic{enumi})}
 \renewcommand{\labelenumi}{\theenumi}
-\item Assume that an element $e_1$ from $S_1$ can be joined with more elements from $S_2$.
-It means that there can be more different elements in $S_2$ having their value of $L_2$ equal to $e_1/L_1$.
-In other words, there can be more different elements in $S_2$ having the same value of $L_2$.
-Therefore, we infer that $L_2$ can not be a key of the elements in $S_2$.
-Moreover, because we suppose that one of $L_1$ and $L_2$ is a key and $L_2$ can not be a key, we infer that $L_1$ is a key of the elements in $S_1$.
-Consequently, we infer that $L_2$ is a foreign key referring $L_1$.
-\item Vice versa, assume that $e_1$ can be joined with maximally one element from $S_2$.
-It means that there is maximally one element in $S_2$ having its value of $L_2$ equal to $e_1/L_1$.
-In that case we suppose that each element in $S_2$ has a unique value of $L_2$.
-Therefore, we infer that $L_2$ is a key of the elements in $S_2$ and consequently that $L_1$ is a foreign key referring $L_2$.
-We cannot infer whether $L_1$ is a key or not.
-\end{enumerate}
+\item $L_1$ is a key of the elements $S_1$, $L_2$ is a respective foreign key and it itself is not a key of the elements $S_2$.
+\item $L_2$ is a key of the elements $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements $S_1$ or not.
+\end{itemize}
 
-\subsubsection{Join patterns}
+\subsection{Join Patterns and Key Inference}
 For a certain join, the decision for one of the cases (O1) and (O2) is made by the form of the join.
 The query is searched for so-called \emph{join patters}.
-These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and they look as follows (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):
+These are \emph{for join pattern} and \emph{let join pattern} and they are as follows (the first is \emph{for join pattern}, \emph{let join pattern}):
 
@@
 \begin{verbatim}
-01  for e1 in P1               for $e1 in P1
-02  return                     return
-03    for $e2 in                 let $e2 :=
+for $e1 in P1              for $e1 in P1
+return                     return
+    for $e2 in                 let $e2 :=
         P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
-04    return CR                  return CR
+    return CR                  return CR
 \end{verbatim}
 
-$P_1$, $P_2$, $L_1$ and $L_2$ in the patterns are XPath paths without predicates.
-Both patterns differ only at line \texttt{03} where the former applies \texttt{for} while the other applies \texttt{let}.
-Line \texttt{01} is called \emph{declaration clause}.
-Line \texttt{03} is called \emph{join clause}.
-The path $P_1$ is called source path, $P_2$ \emph{target path}, and the condition $L_2 = \$e_1/L_1$ \emph{join condition}.
-$C_R$ is called \emph{return clause}.
-All paths $P_1^R$, ..., $P_k^R$ in $C_R$
-starting with $\$e2$ are called \emph{target return paths}.
+Each occurrence of a join pattern is classified by application of the following rules it this specific order.
+The first satisfied ruled is applied.
+The occurrence is also assigned with a weight determining how sure the method is about the inferred statement.
 
-Each pattern occurrence is marked as \emph{repeating} or \emph{non-repeating}.
-An occurrence marked as \emph{repeating} means that each element targeted by $P_1$ can be joined with more than one elements targeted by $P_2$.
-In contrast, a pattern occurrence marked as \emph{non-repeating} means that each element targeted by $P_1$ can be joined with zero or one element targeted by $P_2$.
-In other words, the former means that the observation (O1) can be applied while the latter means that (O2) can be applied to infer keys and foreign keys.
+The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1), aggregation function \texttt{count} is applied on a return path.
+\todo {definovat return path?}
 
-Assume a pattern occurrence $\pi$.
-The decision is made on the basis of the following rules (R1) - (R5).
-Only one rule can be applied.
-Process of decision starts with (R1).
-If (Ri) can not be applied, it tries (Ri+1).
-If any of the following rules (R1) - (R3) can be applied, $\pi$ is marked as \emph{repeating}.
-Each pattern occurence is also assigned with its weight determining how sure the method is about the inferred statement.
+Else, the pattern occurrence is considered of case (O2) and the assigned weight depends on the number of target return paths.
+If the number is greater than one, the weight is one (weight: 1), else (the number is less or equal one) the weight is one half (weight: 0.5).
 
-\begin{enumerate}
-\renewcommand{\theenumi}{(R\arabic{enumi})}
-\renewcommand{\labelenumi}{\theenumi}
-\item $\pi$ is an occurrence of the \texttt{for} pattern (weight: 1).
-\item Aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path $P_i^R$ in $\pi$ (weight: 1).
-\item Aggregation function \texttt{count} is applied on a \emph{return path} $P_i^R$ in $\pi$ (weight: 0.75).
-\end{enumerate}
+\subsection{Summarization of the Results}
 
-If (R1) - (R3) can not be applied, $\pi$ is marked as \emph{non-repeating}.
-The weight is assigned according to (R4) and (R5) listed bellow where $k$ denotes the number of target \emph{return paths} in $\pi$.
-
-\begin{enumerate}
-\renewcommand{\theenumi}{(R\arabic{enumi})}
-\renewcommand{\labelenumi}{\theenumi}
-\setcounter{enumi}{4}
-\item $k > 1$ (weight: 1)
-\item$ k <= 1$ (weight 0.5)
-\end{enumerate}
-
-\subsubsection{Key and foreign key inference}
-Let $w$ be the weight assigned to a pattern occurrence $\pi$. If $\pi$ is marked as \emph{repeating}, (O1) is applied and the following statements with weight $w$ are inferred:
-\begin{itemize}
-\item $(P_2^{\downarrow}, \{L_2\})$ is not satisfied
-\item $(P_1^{\downarrow}, \{L_1\})$ is satisfied
-\item $(P_2^{\downarrow}, \{L_2\}) \Rightarrow (P_1^{\downarrow}, \{L_1\})$ is satisfied
-\end{itemize}
-where for a path $P$, $P^{\downarrow}$ denotes:
-\begin{itemize}
-\item $P$, if $P$ does not start with a variable
-\item $P^{\$e\downarrow}/P'$ (or $P^{\$e\downarrow}//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) where $P^{\$e}$ is a path used for the declaration of the variable $\$e$ in the query
-\end{itemize}
-
-If $\pi$ is marked as \emph{non-repeating}, (O2) is applied and the following statements with weight $w$ are inferred:
-\begin{itemize}
-\item $(P_2^{\downarrow}, \{L_2\})$ is satisfied
-\item $(P_1^{\downarrow}, \{L_1\}) \Rightarrow (P_2^{\downarrow}, \{L_2\})$ is satisfied
-\end{itemize}
-
-If the \emph{source} and \emph{target path} in $\pi$ have a common context determined by a path $C$, keys and foreign keys are inferred as follows.
-If $\pi$ is marked as \emph{repeating}, the following statements with weight $w$ are inferred:
-\begin{itemize}
-\item $(C^{\downarrow}, P_2^{\downarrow C}, \{L_2\})$ is not satisfied
-\item $(C^{\downarrow}, P_1^{\downarrow C}, \{L_1\})$ is satisfied
-\item $(C^{\downarrow}, (P_2^{\downarrow C}, \{L_2\}) \Rightarrow (P_1^{\downarrow C}, \{L_1\}))$ is satisfied
-\end{itemize}
-where for paths $C$ and $P$, $P^{\downarrow C}$ denotes:
-\begin{itemize}
-\item $P$, if $P$ does not start with a variable
-\item $P'$ (or $.//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) and $\$e$ is declared by $C$
-\item $P^{\$e\downarrow C}/P'$ (or $P^{\$e\downarrow C}//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) and $\$e$ is not declared by $C$
-\end{itemize}
-
-If $\pi$ is marked as \emph{non-repeating}, the following statements with weight $w$ are inferred:
-\begin{itemize}
-\item $(C^{\downarrow}, P_2^{\downarrow C}, \{L_2\})$ is satisfied
-\item $(C^{\downarrow}, (P_1^{\downarrow C}, \{L_1\}) \Rightarrow (P_2^{\downarrow C}, \{L_2\}))$ is satisfied
-\end{itemize}
-
-\todo[inline]{Problém je, že tady ste v podstatě opsal ten článek.
-To je špatně.
-Analýza ex.
-přístupů by měla podat přehled toho, co se dnes v dané oblasti umí.
-Každý přístup by měl být popsán zhruba stejným stylem, stejně podrobně a s jednotným značením.
-Tj.
-ne zkopírovat různé odstavce z různých článků, z něčeho 20 řádků a z něčeho 6 stran.
-To není analýza ale plagiátorství.}
-
-\section{Scoring function}
-It can happen that some of the assumptions the method is based on are not met.
-Assuming a set of queries, a key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key.
+The assumption the method is based on may not be fulfilled for every join in a supplied set of queries.
+A key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key.
 \todo[inline]{priklad?}
 
-Therefore, the authors introduce a scoring function.
-If a new key $K$ is going to be inferred, it is assigned with an initial score 0.
-Each inferred positive statement about $K$ increases the score of $K$ by the weight of the statement.
-Respectively, each negative statement about $K$ decreases the score by the respective weight.
-The resulting score therefore summarizes the weights of all inferred statements about $K$.
-A positive score means that $K$ is probably satisfied while negative means that $K$ is probably not satisfied.
-The higher the absolute value of the score is, the higher the probability is.
+Therefore, the authors introduce a scoring function to summarize the positive and the negative statements about keys using the assigned weights.
+The value of the score expresses the probalitity that a respective key statement is satisfied.
+Finally, the scores of the inferred keys are normalized to be comparable with each other.
 
-Let $K_1$, ..., $K_n$ be the inferred keys.
-Let $S_i$ be the score of $K_i$ and $N_i$ be the number of the inferred statements about $K_i$.
-The precision of the scoring is further enhanced on the base of the following observation.
-Assume a key $K_i = (C, P, \{L\})$ and $K_j = (C' , P, \{L\})$ where the path $C$ targets ancestors of the elements targeted by $C'$, i.e.
-the context specified by $C$ covers the context specified by $C'$.
-In that case it is also said that $K_i$ \emph{covers} $K_j$.
-It can be easily seen that if $K_i$ is satisfied, $K_j$ must be satisfied as well.
-On the other hand, if $K_j$ is not satisfied, $K_i$ can not be satisfied too.
-Therefore, if the score $S_i$ of $K_i$ is positive (i.e.
-$K_i$ is satisfied), $S_j$ is incremented with $S_i$ and $N_j$ with $N_i$ (i.e.
-$K_j$ is satisfied as well).
-Conversely, if the score $S_j$ of $K_j$ is negative (i.e.
-$K_j$ is not satisfied), $S_i$ is incremented with $S_j$ and $N_i$ with $N_j$ (i.e.
-$K_i$ can not be satisfied too).
-
-Finally, the scores of the inferred keys are normalized to be comparable with each other.
-The normalized scores are from the range $\langle -1, 1\rangle$.
-The normalization takes into account not only the scores summarizing the weights of the statements about keys but also the number of the statements.
-The normalized score is computed as follows.
-Let $S^{max}$ be the maximum from $|S_1|$, ..., $|S_n|$ and $N^{max}$ be the maximum from $N_1$, ..., $N_n$.
-The normalized score $S_i^{norm}$ of $K_i$ is computed as follows:
-$$S_i^{norm} = {S_i \over S^{max}} * (1 - {N^{max} - N_i \over \sum _{i=1}^n N_i})$$
-
-\section{Conclusion}
+\subsection{Conclusion}
 The output of the method is a list of scored keys and for each key a list foreign keys referencing the key.
-The score of a key can be negative or positive.
-A negative score means that the key is not specified by the XML documents while positive means that it is satisfied.
-The absolute value of the score means how sure the method is about it.
+Since the method deals only with the inference of keys, it is not an overall method of the XML schema inference.
+It is meant to be used in a collaboration with other schema-inferring methods to refine their results.
 

### 2011-09-20 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
   \hline
-  Input data        & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3} \\
-  Iterations        & 50 \\
+  Input data        & all sized test data sets \\
+  Iterations        & 25 \\
   Pool size         & 10 \\
@@
 
-This experimental set will consist of 45 ratio combinations * 6 data sets * 50 iterations = 13500 experimental configurations.
+This experimental set will consist of 45 ratio combinations * 11 data sets * 25 iterations = 12375 experimental configurations.
 CH will be \heu{Random} with pool size of 10.
@@
   \STATE $CX \gets \heu{Crossover}(ratio = cxRatio, limit = 1)$
-  \FOR{$file \in \{\dataset{100-100}, \dataset{100-200}, \dataset{100-1000}, \dataset{OVA1}, \dataset{OVA2}, \dataset{OVA3}\}$}
+  \FOR{$file \in $ sized test data}
     \FOR{$i = 1 \to 50$}
@@
 \end{algorithm}
+
+% TODO we ignore the time, instead we use the ratio of successful runs
 
@@
 Only rows containing at least one such shortest time are presented.
+
+% TODO replace this table
 
@@
 
+% TODO the ratios actually are 0.5, 0.2 and 0.1, consistently across all all data sets
+
 It is now necessary to pick one ratio combination as the best one.

### 2011-09-21 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
 
-We shall then employ a higher-level heuristic: we will choose 3 scenarios (lists of IHs, or metaheuristics), assess their performance to find the best one and then tune its parameters.
+We shall then employ a higher-level heuristic: we will choose 3 strategies (lists of IHs, or metaheuristics), assess their performance to find the best one and then tune its parameters.
 This approach is by no means exhaustive, it is just a probe in the problem space.
 
-The 3 scenarios we will be assessing will be constructed from the following instances of improvement heuristics:
+The 3 strategies we will be assessing will be constructed from the following instances of improvement heuristics:
 
@@
 
-The scenarios themselves shall be the following:
+The strategies themselves shall be the following:
 
 \begin{itemize}
-	\item \textit{\textbf{Scenario 1}}. \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
-	\item \textit{\textbf{Scenario 2}}. \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow$ \heu{MUT} $\rightarrow \ldots$
-	\item \textit{\textbf{Scenario 3}}. \heu{CX} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RW} $\rightarrow$ \heu{LB} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
+	\item \textit{\textbf{Strategy 1}}. \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
+	\item \textit{\textbf{Strategy 2}}. \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow$ \heu{MUT} $\rightarrow \ldots$
+	\item \textit{\textbf{Strategy 3}}. \heu{CX} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RW} $\rightarrow$ \heu{LB} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
 \end{itemize}
@@
 
-The experimental set will consist of 3 scenarios * 11 data sets * 20 iterations = 660 experimental configurations.
+The experimental set will consist of 3 strategies * 11 data sets * 20 iterations = 660 experimental configurations.
 Their construction is formalized in the Listing \ref{listing-experiment-chaining-ihs}.
@@
 \begin{figure}
-  \caption{Chained IHs - \dataset{100-100} Results for Scenario 1}
+  \caption{Chained IHs - \dataset{100-100} Results for Strategy 1}
   \label{image-experiment-chained-ihs-s1}
@@
 X and Y axes represent the time and quality, as usual.
-Each line represents one run of the scenario (metaheuristic) in the following way: the N\textsuperscript{th} break in the line (i.e.
-the N\textsuperscript{th} data point) is the partial result after the N\textsuperscript{th} step of the scenario.
+Each line represents one run of the strategy (metaheuristic) in the following way: the N\textsuperscript{th} break in the line (i.e.
+the N\textsuperscript{th} data point) is the partial result after the N\textsuperscript{th} step of the strategy.
 Its X position denotes the absolute time in which this step finished, and its Y position represents the incumbent solution quality after this step.
 Every time a line disappears before reaching 10 seconds it means that this metaheuristic run found the optimum before the 10 second mark.
-There is an obvious repetitive regularity in the shape of each line, this corresponds to the fact that there is a finite number of IHs in this scenario (5 of them in Scenario 1) which repeat over time.
-The obvious similarity between different lines corresponds to the fact that each run is from the same scenario, and over time, they do the same steps.
+There is an obvious repetitive regularity in the shape of each line, this corresponds to the fact that there is a finite number of IHs in this strategy (5 of them in Strategy 1) which repeat over time.
+The obvious similarity between different lines corresponds to the fact that each run is from the same strategy, and over time, they do the same steps.
 
@@
 
-There are two more graphs like this in Figures \ref{image-experiment-chained-ihs-s2} and \ref{image-experiment-chained-ihs-s3} for comparison, capturing Scenario 2 and Scenario 3 respectively working on the same data set, \dataset{100-100}.
+There are two more graphs like this in Figures \ref{image-experiment-chained-ihs-s2} and \ref{image-experiment-chained-ihs-s3} for comparison, capturing Strategy 2 and Strategy 3 respectively working on the same data set, \dataset{100-100}.
 Describing them in detail is unfortunately outside of the scope of this work.\\
@@
 \begin{figure}
-  \caption{Chained IHs - \dataset{100-100} Results for Scenario 2}
+  \caption{Chained IHs - \dataset{100-100} Results for Strategy 2}
   \label{image-experiment-chained-ihs-s2}
@@
 \begin{figure}
-  \caption{Chained IHs - \dataset{100-100} Results for Scenario 3}
+  \caption{Chained IHs - \dataset{100-100} Results for Strategy 3}
   \label{image-experiment-chained-ihs-s3}
@@
 
-It is now necessary to assess which of the scenarios perform the best.
+It is now necessary to assess which of the strategies perform the best.
 We shall take a look at the different data sets.
 Easily we can discard \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}, because the optimum is found in the very first step.
-Let us now introduce a metric for assessment of a scenario: namely, how many times of the 20 runs did it find the optimum.
+Let us now introduce a metric for assessment of a strategy: namely, how many times of the 20 runs did it find the optimum.
 Results of this are summarized in Table \ref{table-experiments-chained-ihs-tweaking}.
@@
   \begin{tabular}{l || c | c | c}
-    Dataset & Scenario 1 & Scenario 2 & Scenario 3 \\
+    Dataset & Strategy 1 & Strategy 2 & Strategy 3 \\
     \hline
@@
 
-Each cell contains the number of times the scenario found optimum in the data set, out of 20 runs.
-Highlited are the scenarios that performed best on that data set.
-We see that Scenario 1 and 3 are very similar in performance.
-We shall nonetheless choose Scenario \textbf{1} as the winner for its simplicity.
+Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs.
+Highlited are the strategies that performed best on that data set.
+We see that Strategy 1 and 3 are very similar in performance.
+We shall nonetheless choose Strategy \textbf{1} as the winner for its simplicity.
 Now we can tune its parameters.
 
-\subsubsection{Improving Scenario 1}
+\subsubsection{Improving Strategy 1}
 
@@
 
-A short reminder: Scenario 1 consists of \heu{Random} as the CH and the following IHs: \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
-
-The parameters we can tune in this scenario are the ratios in \heu{RandomRemove} (possibly 2 of them, as there are 2 instances in use), \heu{Mutation} and \heu{Crossover}.
+A short reminder: Strategy 1 consists of \heu{Random} as the CH and the following IHs: \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$
+
+The parameters we can tune in this strategy are the ratios in \heu{RandomRemove} (possibly 2 of them, as there are 2 instances in use), \heu{Mutation} and \heu{Crossover}.
 We shall not tune the time limits in \heu{Mutation} and \heu{Crossover} and leave them set to 1 second.
@@
 CH will be \heu{Random} with pool size of 10.
-IHs will be the ones from Scenario 1, with their ratios set to one of the 45 combinations produced in the following way.
+IHs will be the ones from Strategy 1, with their ratios set to one of the 45 combinations produced in the following way.
 \begin{itemize}
@@
 \begin{algorithm}
-\caption{Chained IHs - Improving Scenario 1 Set Generation}
+\caption{Chained IHs - Improving Strategy 1 Set Generation}
 \label{listing-experiment-chained-ihs-tuning}
@@
 \begin{table}
-  \caption{Performance of Scenario 1 Depending on Parameters - Excerpt}
+  \caption{Performance of Strategy 1 Depending on Parameters - Excerpt}
   \bigskip
@@
 \heu{RandomRemove} ratio of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
-This amounts to a very strong diversification tendency and keeps the scenario from stalling in local optima.
+This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima.
 \heu{Mutation} ratio of 0.05 means only around 5\% of AMs in the incumbent solution will be fixed for the next GLPK optimization.
@@
 TODO this will need 2 boxplots again...
+
+\begin{figure}
+  \caption{Chained IHs - Pure \heu{Glpk} vs. Tuned Strategy 1}
+  \label{image-experiment-chained-ihs-tuned}
+  \centering
+    \includegraphics[width=\textwidth]{images/experiments/chained-ihs-tuned}
+\end{figure}
 
@@
 
-Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Scenario 1.
+Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1.
 In under 10 seconds we will have very good results, often the optimum.

### 2011-09-21 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
 
-The first experiment set will try to establish how long it takes to extract the IG % TODO IG -> nomenclature
-from the input XML file and how long does it take to create the AM model from this IG. For now, we will not be running or measuring any heuristics.
+The first experiment set will try to establish how long it takes to extract the IG from the input XML file and how long does it take to create the AM model from this IG.
+For now, we will not be running or measuring any heuristics.
 
@@
 \subsection{GLPK: Native vs. Cygwin}
+\label{section-glpk-comparison}
 
@@
 In this experiment we will try to remove one of the variables out of the equation: that is the effect of different versions of GLPK on the overall results.
-The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution, % TODO link
-or via Cygwin as one of its packages.
+The rationale is this: on Windows systems, the two most accessible ways to install GLPK are via a binary distribution or via Cygwin as one of its packages.
 
@@
 
-% TODO re-run on Dual Core!
-
 We shall now try to find the best standalone CH, that is the CH that finds on average the best solutions when run without any IHs.
@@
 
-We can see that \heu{Glpk} wins in every single case. We will start from there and try to build upon this result.
+We can see that \heu{Glpk} wins (or is among the best) in every single case. We will start from there and try to build upon this result.
 
@@
 
-\subsubsection{Improving Strategy 1}
+\subsubsection{Tuning Strategy 1}
 
@@
 \end{itemize}
+The total limit will remain at 10 seconds.
 The process of creating the configurations is captured in Listing \ref{listing-experiment-chained-ihs-tuning}.
@@
     \FOR{$i = 1 \to 50$}
-      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})\}$
+      \STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW}), limit = 10 seconds\}$
     \ENDFOR
@@
 
-% TODO we ignore the time, instead we use the ratio of successful runs
-
 After averaging the data we get a large result table, an excerpt from which is in Table \ref{table-experiments-chained-ihs-tweaking-s1}.
-In the left part are the ratio values, in the right part averaged running times for each data set.
-Highlited are the shortest times for each set.
-Only rows containing at least one such shortest time are presented.
-
-% TODO replace this table
+Only the results for the biggest data sets and a few combinations of RR, MUT and CX ratios are presented.
+
+In the left part of the table are the ratio values.
+In the right part are the averaged running times (AT), success ratios (ratio of runs that found the optimum, SR) and average qualities (AQ) for each data set.
+Highlited are the highest success ratios and qualities.\\
 
@@
   \centering
-  \begin{tabular}{c | c | c || c | c | c | c | c | c}
-    \heu{RR} & \heu{MUT} & \heu{CX} & \dataset{100-100} & \dataset{100-1000} & \dataset{100-200} & \dataset{OVA1} & \dataset{OVA2} & \dataset{OVA3} \\
-    \hline
-    0.2	& 0.2	& 0.2	   & 2749.58	       & \textbf{8633.76} & 3029.82	         & 110.82	        & 98.68	         & 2413.86 \\
-    0.5	& 0.05	& 0.05 & 1031.24	       & 10324.46	        & 1371.08	         & \textbf{71.88} & 63.86	         & 1282.94 \\
-    0.5	& 0.05	& 0.2	 & \textbf{919.74} & 10978.58	        & 1322.02	         & 74.46	        & 62.42	         & \textbf{1217.00} \\
-    0.5	& 0.1	& 0.2	   & 1002.20	       & 11030.64	        & \textbf{1288.06} & 85.22	        & 80.96	         & 1480.90 \\
-    0.5	& 0.2	& 0.1	   & 1674.78	       & 9779.82	        & 2007.42	         & 108.14	        & \textbf{58.56} & 1827.46 \\
+  \begin{tabular}{c | c | c | l || c | c | c | c | c}
+    \heu{RR} & \heu{MUT} & \heu{CX} & & \dataset{60-180} & \dataset{70-245} & \dataset{80-320} & \dataset{90-405} & \dataset{100-500} \\
+    \hline
+    \hline
+    $ \ldots $ & $ \ldots $ & $ \ldots $ & $ \ldots $ & $ \ldots $ & $ \ldots $ & $ \ldots $ & $ \ldots $ & $ \ldots $ \\
+    \hline
+    0.5	& 0.1	& 0.1	  & AT  & 1049.72	& 1584.72	& 2349.2	& 7408.28	& 9414.28 \\
+    0.5	& 0.1	& 0.1	  & SR	& 1	& 1	& 1	& 0.52	& 0.28 \\
+    0.5	& 0.1	& 0.1	  & AQ	& \textit{opt} & \textit{opt}	& \textit{opt}	& 0.52533	& 0.49196 \\
+    \hline
+    0.5	& 0.1	& 0.2	  & AT	& 763.44	& 1343.08	& 2599.88	& 9448.12	& 10269.76 \\
+    0.5	& 0.1	& 0.2	  & SR	& 1	& 1	& 1	& 0.24	& 0 \\
+    0.5	& 0.1	& 0.2	  & AQ	& \textit{opt}	& \textit{opt}	& \textit{opt}	& 0.52213	& 0.48462 \\
+    \hline
+    0.5	& 0.2	& 0.05	& AT	& 1438.84	& 1608.32	& 2647.4	& 4954.04	& 7784.88 \\
+    0.5	& 0.2	& 0.05	& SR	& 1	& 1	& 1	& 0.8	& 0.4 \\
+    0.5	& 0.2	& 0.05	& AQ	& \textit{opt}	& \textit{opt}	& \textit{opt}	& 0.52693	& 0.49647 \\
+    \hline
+    0.5	& 0.2	& 0.1	  & AT	& 1333.12	& 1741.08	& 2506.84	& 4720.32	& 8150.6 \\
+    0.5	& 0.2	& 0.1	  & SR	& 1	& 1	& 1	& \textbf{0.84}	& \textbf{0.56} \\
+    0.5	& 0.2	& 0.1	  & AQ	& \textit{opt}	& \textit{opt}	& \textit{opt}	& 0.52733	& \textbf{0.49651} \\
+    \hline
+    0.5	& 0.2	& 0.2	  & AT	& 922.16	& 1353.76	& 2394.48	& 4633.12	& 7424.8 \\
+    0.5	& 0.2	& 0.2	  & SR	& 1	& 1	& 1	& \textbf{0.84}	& 0.44 \\
+    0.5	& 0.2	& 0.2	  & AQ	& \textit{opt}	& \textit{opt}	& \textit{opt}	& \textbf{0.52804}	& 0.49495 \\
     \end{tabular}
@@
 
-% TODO the ratios actually are 0.5, 0.2 and 0.1, consistently across all all data sets
-
-It is now necessary to pick one ratio combination as the best one.
-It is $(RR = 0.5, MUT = 0.05, CX = 0.2)$ for the following reason: it is the best one for sets \dataset{100-100} and \dataset{OVA3} and second best for \dataset{100-200}, \dataset{OVA1} and \dataset{OVA2}.
-Only the \dataset{100-1000} does not profit from these settings.
-
-Now to interpret ratios in the best combination.
+It is now necessary to pick one ratio combination as the best one, and it is $(RR = 0.5, MUT = 0.2, CX = 0.1)$.
+Using this combination for all the data sets from \dataset{10-20} up to \dataset{80-320} the optimum was always found and for \dataset{90-405} and \dataset{100-500} the success ratios were the highest.
+
+Now to interpret the ratios in the best combination.
 \heu{RandomRemove} ratio of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
 This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima.
-\heu{Mutation} ratio of 0.05 means only around 5\% of AMs in the incumbent solution will be fixed for the next GLPK optimization.
-\heu{Crossover} ratio of 0.2 means that around 1/5\textsuperscript{th} of ID sets in the pool (randomly chosen) will be scanned for common AMs.
-
-All of the ratios in the best combination are at one end of the range we chose from them.
-As a future work option it is possible to start moving these ratios even more in their preferred way, possibly removing \heu{Mutation} in the process.
-
-TODO compare this to pure GLPK run - for example for \dataset{100-200}, we got from TODO seconds to on average TODO seconds (when the 10 second limit is lifted).
-
-TODO this will need 2 boxplots again...
+\heu{Mutation} ratio of 0.2 means around 1/5\textsuperscript{th} of AMs in the incumbent solution will be fixed for the next GLPK optimization.
+\heu{Crossover} ratio of 0.1 means that around 10\% of ID sets in the pool (randomly chosen) will be scanned for common AMs.
+
+\heu{RandomRemove} and \heu{Mutation} ratios in the best combination are at the upper bound of the range we chose for them.
+As a future work option it is possible to start moving these ratios even more in their preferred way.
+
+\subsubsection{Final Comparison}
+
+Finally we shall compare the performance of Strategy 1 with tuned parameters (ratios) to the approach we started the experiments with: using the \heu{Glpk} CH with no time limit to find the optimum.
+We will compare them on the biggest of sized test data: \dataset{100-500}.
+
+We already have the running times for pure \heu{Glpk} on \dataset{100-500} from the ``Time Until Optimum" experiment in Section \ref{section-glpk-comparison}.
+The last experiment to find the performance of tuned Strategy 1 without a time limit will have the following parameters.
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Input data        & \dataset{100-500} \\
+  Iterations        & 50 \\
+  Pool size         & 10 \\
+  $\alpha$, $\beta$ & $1$, $1$ \\
+  CH                & \heu{Random} \\
+  IHs               & \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots $ \\
+  \hline
+\end{tabular}
+\bigskip
+\end{center}
+
+The last experimental set will consist of 50 iterations = 50 experimental configurations.
+As with the previous one, \heu{Random} will be the CH, IHs will be from Strategy 1, however this time there will be no time limit.
+The process of generating the experimental set is in Listing \ref{listing-experiment-chained-ihs-tuned}.
+
+\begin{algorithm}
+\caption{Chained IHs - Tuned Strategy 1 Performance Set Generation}
+\label{listing-experiment-chained-ihs-tuned}
+\begin{algorithmic}
+\ENSURE experimental set $ES$
+\STATE $ES \gets \emptyset$
+
+\STATE $RW \gets \heu{RemoveWorst}$
+\STATE $RR \gets \heu{RandomRemoval}(ratio = 0.5)$
+\STATE $MUT \gets \heu{Mutation}(ratio = 0.2, limit = 1)$
+\STATE $CX \gets \heu{Crossover}(ratio = 0.1, limit = 1)$
+\FOR{$i = 1 \to 50$}
+  \STATE $ES \gets ES \cup \{\dataset{100-500}, CH = \heu{Random}, IH = (\heu{RR}, \heu{MUT}, \heu{RR}, \heu{CX}, \heu{RW})\}$
+\ENDFOR
+\RETURN $ES$
+\end{algorithmic}
+\end{algorithm}
+
+Results are summarized in Figure \ref{image-experiment-chained-ihs-tuned}.
+Both boxplots represent run times until the optimum is found.
+It is clear that Strategy 1 is an improvement, achieving on average almost 4x shorter times than pure \heu{Glpk} and finding the optimum under 10 seconds in more than a half of the cases.
 
@@
 
-After asking and answering a lot of questions related to the overall system behavior, parameter effects and various heuristic combinations is now the time to summarize our results and draw conclusions.
+After asking and answering a lot of questions related to the overall system behavior, parameter effects and various heuristic combinations is now the time to summarize our results and draw conclusions.\\
 
@@
 It will find the optimum eventually, even though this might take minutes or hours to complete.
-For many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter how long it takes.
+For many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter how long it takes.\\
 
 Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1.
-In under 10 seconds we will have very good results, often the optimum.
-
-Overall, it is always good to ignore the simple text data nodes, as it will improve the total search time.
-
-%\section{Lessons Learned}
-
-%This short section will list a few lessons that were learned during the development of \jmodule{IDSetSearch} module for jInfer and experimenting with it.
-
-%The first lesson is that when there is an experiment being run in a few nested loops (see any experiment set construction listing), one of which is \textit{iterations}, this one should be the outermost one.
-This is because if the experiment somehow fails, TODO
-
-% - always allow for interruptions
-% - when representing an object, don't be lazy and create a class for it (don't use pairs, lists with specific data at specific locations, etc)
+In all our realistic data sets the optimum was found almost instantly, and the most complex and bigges artificial data sets took only around 1/4 of the time to finish, compared to \heu{Glpk}.\\
+
+Furthermore, it is always good to ignore the simple text data nodes, as it will improve the total run time.

### 2011-09-21 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 And finally, we create completely artificial data to create inputs that will put our heuristics in stress.
-This is because the realistic data often prove to be too easy to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.\\
+This is because the realistic data often prove to be too easy to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.
 
-To understand these data sets we will discuss their origin and \textit{graph representation}. As mentioned earlier, % TODO link
+\begin{define}[Data set]
+	\label{define-data-set}
+	One or more XML files sharing the same schema (even if only an implicit schema) shall be referred to as a \textit{data set}.
+In the scope of this work this will always mean a \textit{single} XML file.
+However, this definition of a data \textit{set} covers also the extension to more XML files as described in \cite{fidax}.
+\end{define}
+
+To understand our test data sets we will discuss their origin and \textit{graph representation}. As mentioned earlier, % TODO link
 the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set in a graph.
@@
 
-The former will be achieved with the help of GraphViz tool, % TODO link
-where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
+The former will be achieved with the help of GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
 Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
@@
 Neither will they be included on the DVD distributed with this thesis.
-
-% TODO their DTD/XSD - do we have ID attributes?
 
@@
 
-Our aim is to create a graph with around $v$ vertices and $e$ edges.
+Our aim is to create a graph with approximately $v$ vertices and $e$ edges.
 First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
@@
 This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set.
-Because of the so-called Birthday paradox, % TODO link
-this will happen more with higher $e$.
+Because of the so-called Birthday paradox (see e.g.
+\cite{birthday}), this will happen more with higher $e$.
 

### 2011-09-21 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 
-According to \cite{Mlynkova:2008:AAX:1494650.1495496} The type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
-Heuristic methods are based on experience with manual construction of schemas and their result commonly does not belong to any class of grammar.
-On the contrary, result of the grammar-inferring methods belongs to a particular class of languages with specific characteristics.
+According to \cite{Mlynkova:2008:AAX:1494650.1495496}, the type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
+Heuristic methods are based on experience with manual construction of schemas, motivated by real-life usages of XML schemata, and their result commonly does not belong to any class of grammar.
+On the contrary, the grammar-inferring methods are based on theoretical knowledge of automata and their results belong to a particular class of languages.
+Thus, these methods guarantee specific characteristics of the results.
+\todo[inline]{Tady by neškodilo dát odkazy na všechny ex.
+přístupy a říct, že popíšeme detailněji aspoň dva.}
 
@@
 Most of the approaches process XML documents as the input of the inference process and the documents are supposed to be valid against the resulting schema.
-Besides approaches exploiting XML data, approaches that utilize operations over XML data may be developed and they are the most significant approaches in terms of this work.
+Besides approaches exploiting XML data, approaches that utilize other or additional sources may be developed.
+An approach utilizing XML data along with an obsolete XML schema is described in \cite{Mlynkova:2009:IXS:1862681.1862693}.
+However, the most significant approaches in terms of this work are those, utilizing operations over XML data.
+According to our best knowledge at the time of writing, there is just one approach of this category, described in \cite{Necasky:2009:DXK:1529282.1529414}.
+It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury - Ty by tady něškodily všude, u každé té kategorie.}.
 
-According to our best knowledge at the time of writing there is just one approach of this category.
-It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury?
-Ty by tady něškodily všude, u každé té kategorie.}.
+\section{Common Caracteristics}
+The process of inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
+e_k$ is constructed.
+The productions form so-called \emph{initial grammar} (IG).
+For each element type the productions are then merged, simplified and generalized using various methods and criteria.
+A common approach is so-called \emph{merging state algorithm}, where a \emph{prefix tree automaton} (PTA) is built from the productions of the same element type and the automaton is generalized via merging of its states.
+Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
+\todo[inline]{Tento odstavec je doslovne prebraty z uvedeneho clanku. Mozem to takto pouzit?}
+
+\section{XTRACT}
+The XTRACT \cite{Garofalakis:2000:XSE:342009.335409} system is an example of a \emph{heuristic} \emph{merging state algorithm} creating the result in DTD.
+Its process of inference consists of three steps:
+\begin{enumerate}
+\item Generalization - Generates a set of DTD candidates by searching the input for certain patterns and generalising corresponding fragments using regular expressions.
+\item Factoring - Groups of generalized candidate DTDs are factorized to a new ones by finding common sub-expressions to make them more concise.
+\item Minimum Description Length (MDL) Principle - Composing a near-optimal DTD schema from the set of all generalized candidate DTDs.
+\end{enumerate}
+
+\subsection{Generalization}
+Purpose of generalization is to create a set of DTD candidates - schemata that cover fractions of the input XML data.
+In the last step, this set will be used to compose a (sub)optimal result with respect to a trade-off between its preciseness and conciseness.
+Therefore, it is desirable to create DTD canditates with various degree of these two characteristics.
+
+Generalization is based on replacing fragmets (sequences of subelements of a given element) from the input XML data by regular expressions, thus, using metacharacters like (*, +, ?).
+To provide a wide set of DTD canditates, each sequence is processed several times using various values of input parameters.
+Due to the very large number of possible DTD canditates, the authors employ certain real-life motivated heuristics.
+
+For instance, the paper (\cite{Garofalakis:2000:XSE:342009.335409}) introduces the following example: Sequences \texttt{abab} and \texttt{bbbe} are generalized to \texttt{(ab)*}, \texttt{(a|b)*}, \texttt{b*e}.
+\todo[inline]{Mozem takto pouzivat prevzate priklady?}
+
+\subsection{Factoring}
+Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decresing their summed size without modifications in their semantics.
+An aim of this step is to decrease a MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
+
+An example introduced it the paper (\cite{Garofalakis:2000:XSE:342009.335409}): DTD candidates \texttt{ac}, \texttt{ad}, \texttt{bc} and \texttt{bd} are factored into \texttt{(a|b)(c|d)}.
+
+Alike the generalization step, also in this step the set of possible factored DTD is huge and the authors propose certain heuristics to make the factorization effective.
+
+\subsection{Minimum Description Length (MDL) Principle}
+This is an important step trying to create the resulting DTD with the best trade-off between its preciseness and conciseness.
+
+Paper \cite{Mlynkova:2008:AAX:1494650.1495496} summarizes this step as follows: It expresses the quality of a DTD candidate using two aspects – conciseness and preciseness.
+Conciseness of a DTD is expressed using the number of bits required to describe the DTD (the smaller, the better).
+Preciseness of a DTD is expressed using the number of bits required for description of the input data using the DTD.
+In other words, the more accurately the structure is described, the fewer bits are required.
+Since the two conditions are contradictory, their balancing brings reasonable and realistic results.
+
+\todo[inline]{Detailnejsie? Priklad?}
+
+\todo[inline]{Popisat nejaky pristup vyuzivajuci ACO}
+
+\section{SchemaMiner}
+\todo[inline]{Popisat detailnejsie}
+The SchemaMiner system \todo{bibref} is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
+It focuses on inferring elements with the same name but different structure and unordered sequences.is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
+It focuses on inferring elements with the same name but different structure and unordered sequences.
+\todo[inline]{Definovat skratku XSD a co to je element type.}
+
+The elements with the same name but different content are supported only in XSDs and their inference requires exploitation of more sophisticated approaches than combining productions with the same element
+type. On the other hand, although the unordered sequences are a classical example of XSD “syntactic sugar”, their exploitation enables to define simple and, hence, realistic and usable schemas.
+
+\todo[inline]{Moze byt takto? Nejake dalsie pristupy?}
+\todo[inline]{U obou by neškodil příklad - tohle je velmi povrchní, pokud někde nečetl i ty články, nepochopí nic.}
 
@@
 
-\section{Approaches that Utilize XML Documents}
-The process of inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
-e_k$ is constructed.
-The productions form so-called \emph{initial grammar} (IG).
-For each element type the productions are then merged, simplified and generalized using various methods and criteria.
-A common approach is so-called \emph{merging state algorithm}, where a \emph{prefix tree automaton} (PTA) is built from the productions of the same element type and the automaton is generalized via merging of its states.
-Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
-
-\todo[inline]{Tady by neškodilo dát odkazy na všechny ex. přístupy a říct, že popíšeme detailněji aspoň dva.}
-
-\subsection{XTRACT}
-The XTRACT \todo{bibref} system is an example of a \emph{heuristic} \emph{merging state algorithm} creating the result in DTD. Its process of inference consists of three steps:
-\begin{enumerate}
-\item Generalization - Generates a set of DTD candidates by searching the input for certain patterns and generalising corresponding fragments using regular expressions.
-\item Factoring - Groups of generalized candidate DTDs are factorized to a new ones by finding common sub-expressions to make them more concise. \todo[inline]{priklad}
-\item Minimum Description Length (MDL) Principle - Composing a near-optimal DTD schema from the set of all generalized candidate DTDs.
-\end{enumerate}
-
-MDL is the most important step.
-It ranks each DTD candidate using two aspects - conciseness and preciseness.
-Conciseness of a DTD is expressed by using the number of bits required to describe the DTD (the smaller, the better).
-Preciseness of a DTD is expressed using the number of bits required for description of the input data using the DTD.
-In other words, the more accurately the structure is described, the fewer bits are required.
-Since the two conditions are contradictory, their balancing brings reasonable and realistic results.
-
-\subsection{SchemaMiner}
-The SchemaMiner system \todo{bibref} is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
-It focuses on inferring elements with the same name but different structure and unordered sequences.is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
-It focuses on inferring elements with the same name but different structure and unordered sequences.
-\todo[inline]{Definovat skratku XSD a co to je element type.}
-
-The elements with the same name but different content are supported only in XSDs and their inference requires exploitation of more sophisticated approaches than combining productions with the same element
-type. On the other hand, although the unordered sequences are a classical example of XSD “syntactic sugar”, their exploitation enables to define simple and, hence, realistic and usable schemas.
-
-\todo[inline]{Moze byt takto? Nejake dalsie pristupy?}
-\todo[inline]{U obou by neškodil příklad - tohle je velmi povrchní, pokud někde nečetl i ty články, nepochopí nic.}
-
 \section{Discovering XML Keys and Foreign Keys in Queries}
@@
 
-\begin{itemize}
+\begin{enumerate}
 \renewcommand{\theenumi}{(O\arabic{enumi})}
@@
 \item $L_2$ is a key of the elements $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements $S_1$ or not.
-\end{itemize}
+\end{enumerate}
 
@@
 
-The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1), aggregation function \texttt{count} is applied on a return path.
+The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1) or if aggregation function \texttt{count} is applied on a return path (weight: 0.75).
 \todo {definovat return path?}

### 2011-09-22 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

@@
 
-\begin{itemize}
-	\item XML tree - from FIDAX or wherever
-	\item Element, Attribute
-	\item ID, IDREF, IDREFS attribute according to specification
-	\item XSD keys (compare them to ID attributes)
-	\item Key according to Keys for XML \cite{keX}
-	\item Attribute mapping \nomenclature{AM}{Attribute Mapping}
-	\item ID set
-  \item Weight: support, coverage
-\end{itemize}
+TODO we shall introduce a number of definitions, some only auxiliary, some necessary in the whole work
+
+\section{XML Tree}
+
+TODO from FIDAX or wherever
+
+\subsection{Element, Attribute}
+
+TODO
+
+\section{ID, IDREF, IDREFS Attributes}
+
+TODO according to specification
+
+\section{XML Keys}
+
+TODO
+
+\subsection{According to Specification}
+
+TODO compare them to ID attributes
+
+\subsection{According to \cite{keX}}
+
+TODO show how a XML key is defined using the keX notation % this is a terrible sentence
+
+\section{Attribute Mappings}
+
+TODO \nomenclature{AM}{Attribute Mapping}
+
+\subsection{Attribute Mapping Model}
+
+TODO
+
+\section{ID Set}
+
+TODO
+
+\section{Attribute Mapping Weight}
+
+TODO
+
+\subsection{Support}
+
+TODO
+
+\subsection{Coverage}
+
+TODO

### 2011-09-22 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress
Thesis: in progress.

@@
 \chapter{MIP Approach}
+\label{chapter-mip}
 
-TODO rigorously define MIP - optimization problem, some of the variables integral
+TODO rigorously define MIP - optimization problem similar to LP, however some of the variables integral or boolean
 
@@
 TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
+
+TODO define \textit{name(AM)}
+
+\begin{verbatim}
+<x>
+  <y a="1" b="2"/>
+  <y a="3" c="4"/>
+  <y/>
+  <z a="1"/>
+</x>
+\end{verbatim}
+
+TODO equations for the MIP formulation
+
+\begin{scriptsize}
+\begin{verbatim}
+set AMs;
+param Weight {i in AMs};
+var x {i in AMs} binary;
+maximize z: sum {i in AMs} x[i] * Weight[i];
+s.t. c1: x['y-b'] + x['y-c'] <= 1;
+s.t. c2: x['y-a'] + x['y-c'] <= 1;
+s.t. c3: x['y-a'] + x['y-b'] <= 1;
+s.t. c4: x['y-a'] + x['z-a'] <= 1;
+data;
+set AMs :=
+y-c
+y-b
+y-a
+z-a;
+param Weight :=
+y-c 0.2
+y-b 0.2
+y-a 0.6
+z-a 0.4;
+end;
+\end{verbatim}
+\end{scriptsize}
 
@@
 
-\subsubsection{\jmodule{FIDAX}}
+\subsubsection{\heu{FIDAX}}
 
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \jmodule{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
+It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \heu{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
 
@@
 \begin{algorithm}
-\caption{\jmodule{FIDAX} CH}
+\caption{\heu{FIDAX} CH}
 \label{listing-ch-fidax}
@@
 
-\subsubsection{\jmodule{Random}}
+\subsubsection{\heu{Random}}
 \label{heu-ch-random}
@@
 The advantages of this trivial heuristic are simplicity, speed and ease with which it can create a pool of variable solutions, almost for free.
-As we will see later in the experiments (TODO link the Random/Fuzzy/FIDAX fight), it performs surprisingly well.
+As we will see later in the experiments (Section \ref{section-experiments-random-fuzzy-fidax}), it performs surprisingly well.
 
@@
 \begin{algorithm}
-\caption{\jmodule{Random} CH}
+\caption{\heu{Random} CH}
 \label{listing-ch-random}
@@
 
-\subsubsection{\jmodule{Fuzzy}}
+\subsubsection{\heu{Fuzzy}}
 \label{heu-ch-fuzzy}
 
-\jmodule{Fuzzy} is an improvement over the \jmodule{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
+\heu{Fuzzy} is an improvement over the \heu{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
 The weight used here is the usual weight of an AM as defined in TODO.
@@
 \begin{algorithm}
-\caption{\jmodule{Fuzzy} CH}
+\caption{\heu{Fuzzy} CH}
 \label{listing-ch-fuzzy}
@@
 
-\subsubsection{\jmodule{Incremental}}
+\subsubsection{\heu{Incremental}}
 
@@
 \begin{algorithm}
-\caption{\jmodule{Incremental} CH}
+\caption{\heu{Incremental} CH}
 \label{listing-ch-incremental}
@@
 
-\subsubsection{\jmodule{Removal}}
+\subsubsection{\heu{Removal}}
 
-This is basically a reversal of the idea from the \jmodule{Incremental} heuristic - start with a solution containing all the candidate AMs.
+This is basically a reversal of the idea from the \heu{Incremental} heuristic - start with a solution containing all the candidate AMs.
 This probably does not satisfy the ID set condition.
@@
 \begin{algorithm}
-\caption{\jmodule{Removal} CH}
+\caption{\heu{Removal} CH}
 \label{listing-ch-removal}
@@
 
-This CH will be called \jmodule{Glpk} from now on, for it is basically a time-constrained run of GLPK.
+This CH will be called \heu{Glpk} from now on, for it is basically a time-constrained run of GLPK.
 
@@
 
-\subsubsection{\jmodule{Identity}}
+\subsubsection{\heu{Identity}}
 
@@
 \begin{algorithm}
-\caption{\jmodule{Identity} IH}
+\caption{\heu{Identity} IH}
 \label{listing-ih-identity}
@@
 
-\subsubsection{\jmodule{Remove Worst}}
+\subsubsection{\heu{Remove Worst}}
 
@@
 \begin{algorithm}
-\caption{\jmodule{Remove Worst} IH}
+\caption{\heu{Remove Worst} IH}
 \label{listing-ih-removeworst}
@@
 
-\subsubsection{\jmodule{Random Remove}}
+\subsubsection{\heu{Random Remove}}
 
@@
 The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
-For example, \jmodule{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
+For example, \heu{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
 
@@
 \begin{algorithm}
-\caption{\jmodule{Random Remove} IH}
+\caption{\heu{Random Remove} IH}
 \label{listing-ih-randomremove}
@@
 
-\subsubsection{\jmodule{Hungry}}
+\subsubsection{\heu{Hungry}}
 
@@
 \begin{algorithm}
-\caption{\jmodule{Hungry} IH}
+\caption{\heu{Hungry} IH}
 \label{listing-ih-hungry}
@@
 
-\subsubsection{\jmodule{Mutation}}
+\subsubsection{\heu{Mutation}}
+
+TODO explain how this works and link wise books
 
 TODO explain how this translates to GLPK input
+
+For every AM $AM_F$ fixed to appear in the solution a following constraint is added to GLPK input:
+\[s.t. f_{index}: x['name(AM_F)'] = 1;\]
+$index$ is a unique integer to number all the constraints.
+
+Additionaly, every other mapping $AM_i$ colliding with $AM_F$ ($\iff \iota(AM_F) \cap \iota(AM_i) \neq \emptyset$) will cause the following constraint to be added:
+\[s.t. f_{index}: x['name(AM_i)'] = 0;\]
+And the original constraint in form:
+\[s.t. c_{index}: x['name(AM_F)'] + x['name(AM_i)'] <= 1;\]
+will not be included.
 
@@
 \begin{algorithm}
-\caption{\jmodule{Mutation} IH}
+\caption{\heu{Mutation} IH}
 \label{listing-ih-mutation}
@@
 
-\subsubsection{\jmodule{Crossover}}
+\subsubsection{\heu{Crossover}}
 
-TODO explain how this translates to GLPK input
+TODO explain how this works and link wise books
+
+TODO explain how this translates to GLPK input - it's again simple fixing to 1, but we get the list of AMs in a different manner.
 
@@
 \begin{algorithm}
-\caption{\jmodule{Crossover} IH}
+\caption{\heu{Crossover} IH}
 \label{listing-ih-crossover}
@@
 
-\subsubsection{\jmodule{Local Branching}}
+\subsubsection{\heu{Local Branching}}
+
+TODO explain how this works and link wise books
 
 TODO explain how this translates to GLPK input
+
+A new constrain describing the maximal allowed distance from the incumbent solution is added to GLPK input.
+
+\[s.t. LB: sum\{i\ in\ INCUMBENT\} (1 - x[i]) + sum\{i\ in\ REMAINING\} x[i] \leq k;\]
+
+Where $INCUMBENT$ is a set of names of AMs in the incumbent solution, $REMAINING$ is a set of all the AMs not included in the incumbent solution and $k$ is the requested maximal distance.
 
@@
 \begin{algorithm}
-\caption{\jmodule{Local Branching} IH}
+\caption{\heu{Local Branching} IH}
 \label{listing-ih-localbranching}
@@
 
-TODO describe the (rather trivial) algorithm of finding IDREF attributes once we have the ID set
+Once an ID set is found, regardless of how exactly, it is easy to find the IDREF set, i.e. the attribute mappings that can be declared as IDREF. % TODO link FIDAX, they did it first.
+
+First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set. This is because the specification % TODO link to the specific anchor in the page
+does not allow an attribute to be ID and IDREF (IDREFS) at the same time.
+Let us denominate these mappings as \textit{IDREF candidates} (obviously different from \textit{candidate AMs}).
+% TODO link candidate AMs
+
+Second, find the image of the ID set as the union of images of all the AMs in this ID set.
+
+\[\iota(ID) = \bigcup_{am \in ID} \iota(am)\]
+
+Now the IDREF set contains all the AMs whose images are a subset of the ID set image.
+
+\[\iota(c) \subset \iota(ID) \Rightarrow c \in IDREF\]
+
+This can be easily determined in a loop over the list of candidates. The process is captured in Listing \ref{listing-idref}.
+
+\begin{algorithm}
+\caption{IDREF Search}
+\label{listing-idref}
+\begin{algorithmic}
+\REQUIRE $AMs$ list of all AMs
+\REQUIRE $ID$ ID set as a list of AMs
+\ENSURE $IDREF$ set as a list of AMs
+\STATE $IDREF \gets \emptyset$
+\STATE $candidates \gets AMs \backslash ID$
+\STATE $\iota(ID) \gets \bigcup_{am \in ID} \iota(am)$
+\FORALL{$c \in candidates$}
+  \IF{$\iota(c) \subset \iota(ID)$}
+    \STATE $IDREF \gets IDREF \cup c$
+  \ENDIF
+\ENDFOR
+\RETURN $IDREF$
+\end{algorithmic}
+\end{algorithm}

### 2011-09-22 doc/theses/vitasek/diplomka.tex
Thesis: in progress.
Thesis & Experimental framework: in progress.
Thesis: in progress

@@
 \usepackage{amsthm}
+\usepackage{amsmath}
 
@@
 \newcommand{\heu}[1]{\texttt{\textit{#1}}}
+\newcommand{\dataset}[1]{\texttt{\textit{#1}}}
 \newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
@@
 \righthyphenmin=2
+
+% -----------------------------------------------------------------------------
+
+%                            FINAL CLEANUP
+
+% - make sure all chapter names are in First Caps
+% - make sure all references have a name, like Table 4.3.4 instead of just 4.3.4
+% - get rid of all short forms: won't -> will not
+% - make sure there are no singletons
+% - make sure LaTeX does not complaint much about under/overfullness
+% - make sure everything is defined and referenced
+% - make sure all the heu/data set names are consistent. This holds especially for GLPK (tool) / Glpk (heu) and FIDAX
+% - make sure that when referring to heus in algorithms, e.g. when describing how experimental sets are constructed, we correctly refer to their parameters as described in their pseudocodes
+% - the official name for things like OVA1 and 100-100 is (test) data set. Make sure I am not calling it files or anything else.
+% - make sure labels have consistent names
 
@@
 
-TODO list conventions - how we write code related stuff, module names, heu names, etc
+As usual, source code excerpts, class, field and method names shall be written in fixed-width font, such as \texttt{getHeuristic()}.
+Names of specific heuristics will be written like this: \heu{Mutation}.
+Name of test data sets will be written like this: \dataset{OVA1}.\\
+
+TODO perhaps explain the way we will be writing pseudocode?\\
+
+There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.
 
@@
 
-\include{chapter-4-experiments}
+\include{chapter-4-1-experiments-setup}
+
+\include{chapter-4-2-experiments-results}
 
@@
 \addcontentsline{toc}{chapter}{List of Abbreviations}
+\label{chapter-list-abbreviations}
 
@@
 
-\flushright
-$\square$
+% TODO a QED square
 

### 2011-09-22 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 According to \cite{Mlynkova:2008:AAX:1494650.1495496}, the type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
-Heuristic methods are based on experience with manual construction of schemas, motivated by real-life usages of XML schemata, and their result commonly does not belong to any class of grammar.
+Heuristic methods are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
 On the contrary, the grammar-inferring methods are based on theoretical knowledge of automata and their results belong to a particular class of languages.
@@
 
-\todo[inline]{Popisat nejaky pristup vyuzivajuci ACO}
+\section{Even an Ant Can Create an XSD}
+This work, described in \cite{Vosta:2008:EAC:1802514.1802522}, combines several previously proposed approaches including the XTRACT system discussed in the previous section.
+Its improvements of the process of XML schema inference include:
 
-\section{SchemaMiner}
-\todo[inline]{Popisat detailnejsie}
-The SchemaMiner system \todo{bibref} is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
-It focuses on inferring elements with the same name but different structure and unordered sequences.is another example of a \emph{heuristic} \emph{merging state algorithm} creating the result in XSD.
-It focuses on inferring elements with the same name but different structure and unordered sequences.
-\todo[inline]{Definovat skratku XSD a co to je element type.}
+\begin{itemize}
+\item Distinguishing elements with the same name but different semantics.
+\item Improvements of algorithms adopted from the previous works.
+\item Incorporating inference of an unordered sequence.
+\item Creating a result in XML Schema language.
+\end{itemize}
 
-The elements with the same name but different content are supported only in XSDs and their inference requires exploitation of more sophisticated approaches than combining productions with the same element
-type. On the other hand, although the unordered sequences are a classical example of XSD “syntactic sugar”, their exploitation enables to define simple and, hence, realistic and usable schemas.
+\subsection{Clustering of Elements}
+This phase clusters elements on a basis of their context and structure.
+It is done by creating tree structures for each input XML documents, where vertices represent elements and attributes and an edge between two vertices expresses the their parent-child relation.
+These trees and their subtrees and then compared, using an imposed tree similarity measure, to find elements with the same semantics.
 
-\todo[inline]{Moze byt takto? Nejake dalsie pristupy?}
-\todo[inline]{U obou by neškodil příklad - tohle je velmi povrchní, pokud někde nečetl i ty články, nepochopí nic.}
+\subsection{Schema Generalization}
+For each cluster, a trivial schema is created, which is then generalized to achieve a reasonable result.
+In a search for the optimal schema, \emph{Ant Colony Optimization (ACO)} heuristic is incorporated.
+The idea behind the ACO heuristic is that a set of artifical ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum.
+An ant
+is performing steps (schema modifications), dying after a predifined number of steps and providing an information - positive feedback - on quality of a found solution.
+The search is performed in a defined count of interations (or it stops if good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
+Every step of an ant represents a modification of a schema, in particular, a merge of states in a corresponding PTA.
+
+One of the improvements of this heuristic is inclusions of a negative feedback after each step of an ant, visible only in a current iteration.
+Due to this improvement, a larger subspace of solutions is searched.
+
+Another improvement lies in a way how an ant decides for a particular step to perform.
+To achieve better results, the authors propose a combinations of several verified approaches.
+A set of all possible steps is created using \emph{k,h-context} and \emph{s,k-string} \todo{citacie} methods.
+The optimal step is then selected employing the MDL principle \todo{citacia}.
+
+\subsection{Result in XSD}
+Unlike the majority of recent approaches, this methods creates its result in XML Schema language (XSD).
+Comparing to DTD, XSD is a newer language and it provides greater expression possibilities.
+Authors of this method focused on inferring elements with the same name but different context and the unordered sequences which can be in XSD expressed by \texttt{xs:all} construct.
+Elements with the same name but different context cannot be expressed in DTD and, however, the unordered sequence can be also expressed in DTD as alternations of ordered sequences, such expression in not practical nor well human-readable.
 

### 2011-09-24 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

@@
 
-TODO from FIDAX or wherever
+We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data.
+This tree shall be called an \textit{XML tree}.
+For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the referenced article does.
 
-\subsection{Element, Attribute}
+Without a loss of generality we ignore the actual ordering of nodes in the tree.
 
-TODO
+\paragraph{Example}
+
+This example introduces an XML file fragment that will be used for demonstration throughout this work.
+XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}, where each node is annotated with a triple $label(v) : id(v) : value(v)$.
+
+\begin{verbatim}
+<x>
+  <y a="1" b="2"/>
+  <y a="3" c="4"/>
+  <y/>
+  <z a="1"/>
+</x>
+\end{verbatim}
+
+\begin{figure}
+  \caption{Example XML Tree}
+  \label{image-definitions-example-xml-tree}
+  \centering
+    \includegraphics[width=\textwidth]{images/definitions/xml-tree}
+\end{figure}
+
+Furthermore, we denote $I$ % TODO fix the capital I
+the set of all ids and $V$ % TODO fix the capital V
+the set of all values in the document. We will need two more definitions from the article.
+
+\begin{define}[Node equality]
+	$v_1$ and $v_2$ are node equal, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
+\end{define}
+
+\begin{define}[Value equality]
+	$v_1$ and $v_2$ are value equal, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
+\end{define}
 
@@
 
-TODO according to specification
+According to \cite{Bray:08:EML}, an XML attribute may have the type \textit{ID}, \textit{IDREF} or \textit{IDREFS} (among others). Following constraints are related to these types.
+
+\begin{quote}
+\textbf{Validity constraint: ID}
+
+Values of type \textbf{ID} \textsc{must} match the Name production.
+A name \textsc{must not} appear more than once in an XML document as a value of this type; i.e., \textbf{ID} values \textsc{must} uniquely identify the elements which bear them.
+
+\textbf{Validity constraint: One ID per Element Type}
+
+An element type \textsc{must not} have more than one \textbf{ID} attribute specified.
+
+\textbf{Validity constraint: ID Attribute Default}
+
+An \textbf{ID} attribute \textsc{must} have a declared default of \textbf{\#IMPLIED} or \textbf{\#REQUIRED}.
+
+\textbf{Validity constraint: IDREF}
+
+Values of type \textbf{IDREF} \textsc{must} match the Name production, and values of type \textbf{IDREFS} \textsc{must} match Names; each Name \textsc{must} match the value of an \textbf{ID} attribute on some element in the XML document; i.e.
+\textbf{IDREF} values \textsc{must} match the value of some ID attribute.
+\end{quote}
 
@@
 
-TODO \nomenclature{AM}{Attribute Mapping}
+Now we return to \cite{fidax} to define the notion of an \textit{attribute mapping} (or AM for short). \nomenclature{AM}{Attribute Mapping}
+We will use a different definition (without introducing keys from \cite{keX} for now) that will however give us the same result.
+
+\begin{define}[$\Sigma^E$, $\Sigma^A$, $\Sigma$]
+	$\Sigma^E$ is the set of all element labels, $\Sigma^A$ is the set of all attribute labels.
+$\Sigma = \Sigma^E \cup \Sigma^A$ is their union and effectively the set of all labels in the document (simple text data nodes do not have a label).
+\end{define}
+
+\begin{define}[Attribute mapping]
+	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the attribute mapping of y over x, denoted $M_{x}^{y}$, the $I \times I$ % TODO fix the capital I
+	relation defined by
+	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}\]
+\end{define}
+
+Thus the relation $M_{x}^{y}$ contains edges in the XML tree connecting element nodes labeled $x$ and attribute nodes labeled $y$.
+
+TODO more from the article?
+
+\paragraph{Example}
+The XML tree from Figure \ref{image-definitions-example-xml-tree} has the following non-empty AMs: $M_{y}^{a} = \{(2,6), (3,8)\}$, $M_{y}^{b} = \{(2,7)\}$, $M_{y}^{c} = \{(3,9)\}$ and $M_{z}^{a} = \{(5,10)\}$.
 

### 2011-09-24 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

@@
   \centering
-    \includegraphics[width=\textwidth]{images/definitions/xml-tree}
+    \subfigure[XML tree]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree}}
+    \subfigure[Attribute mappings]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree-ams}}
 \end{figure}
@@
 \section{ID, IDREF, IDREFS Attributes}
+\label{section-definitions-id-attributes}
 
@@
 \begin{define}[Attribute mapping]
-	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the attribute mapping of y over x, denoted $M_{x}^{y}$, the $I \times I$ % TODO fix the capital I
+	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $I \times I$ % TODO fix the capital I
 	relation defined by
@@
 
-TODO more from the article?
+We can use projection to retrieve all the unique \textit{ids} of either elements or attributes from the relation, with notation $\pi_E(M_{x}^{y})$ and $\pi_A(M_{x}^{y})$.
+
+\begin{define}[Type of the attribute mapping]
+	Attribute mapping $M_{x}^{y}$ is of the \textit{type} $\tau(M_{x}^{y}) = x$.
+\end{define}
 
 \paragraph{Example}
-The XML tree from Figure \ref{image-definitions-example-xml-tree} has the following non-empty AMs: $M_{y}^{a} = \{(2,6), (3,8)\}$, $M_{y}^{b} = \{(2,7)\}$, $M_{y}^{c} = \{(3,9)\}$ and $M_{z}^{a} = \{(5,10)\}$.
+The XML tree from Figure \ref{image-definitions-example-xml-tree} has the following non-empty AMs drawn in bold lines: $M_{y}^{a} = \{(2,6), (3,8)\}$, $M_{y}^{b} = \{(2,7)\}$, $M_{y}^{c} = \{(3,9)\}$ and $M_{z}^{a} = \{(5,10)\}$.
+
+Following example equations hold.
+\begin{eqnarray*}
+\pi_E(M_{y}^{a}) & = & \{2, 3\} \\
+\pi_A(M_{z}^{a}) & = & \{10\} \\
+\tau(M_{y}^{c}) & = & y
+\end{eqnarray*}
+
+\begin{define}[Image of the attribute mapping]
+	\textit{Image} $\iota$ of the attribute mapping $M_{x}^{y}$ is defined as $\iota(M_{x}^{y}) = \{z: z = value(w), w \in \pi_A(M_{x}^{y})\}$
+\end{define}
+
+So the image of an AM is a set of all the values of all the attribute nodes contained in the mapping.
+
+\paragraph{Example}
+Again referring to the XML tree from \ref{image-definitions-example-xml-tree}, we get the following AM images.
+\begin{eqnarray*}
+\iota(M_{y}^{a}) & = & \{1, 3\} \\
+\iota(M_{y}^{b}) & = & \{2\} \\
+\iota(M_{y}^{c}) & = & \{4\} \\
+\iota(M_{z}^{a}) & = & \{1\} \\
+\end{eqnarray*}
 
@@
 
+Based on the requirements for an ID attribute from Section \ref{section-definitions-id-attributes} we will define ID set with the help of the following definition.
+
+\begin{define}[Candidate attribute mapping]
+An attribute mapping $m$ is a \textit{candidate attribute mapping} if it is an injective function, that is,
+\[|m| = |\pi_E(m)| = |\pi_A(m)| = |\iota(m)|\]
+\end{define}
+
+\paragraph{Example}
+In our example all the attribute mappings are candidate AMs.
+
+Now we can proceed to define an ID set.
+
+\begin{define}[ID set]
+A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an ID set iff
+\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset\]
+\end{define}
+
+That is, an ID set has images without repeating values and all the types are unique (an element cannot have more than one ID attribute).
+
+\paragraph{Example}
+Returning to our example, the following are all the possible ID sets: $\{ M_{y}^{a} \}, \{ M_{y}^{b} \}, \{ M_{y}^{b}, M_{z}^{a} \}, \{ M_{y}^{c} \}, \{ M_{y}^{c}, M_{z}^{a} \}$.
+Note that once we select an AM of type $y$ we can never add any other with the same type.
+Note also that $\{ M_{y}^{c}, M_{z}^{a} \}$ is not an ID set, because $\iota(M_{y}^{c}) \cap \iota(M_{z}^{a}) \neq \emptyset$.
+
+\subsection{IDREF and IDREFS Attributes}
+
 TODO
@@
 
-TODO
+TODO this is from FIDAX again
+
+$M = \{m_1, \dots m_i\}$ is the set of all non-empty AMs in the document.
 
@@
 
-TODO
+\begin{define}[Support]
+\textit{Support} of an attribute mapping $m$ is defined as follows.
+\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}\]
+\end{define}
+
+TODO example
 
@@
 
-TODO
+\begin{define}[Coverage]
+\textit{Coverage} of an attribute mapping $m$ is defined as follows.
+\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|\]
+\end{define}
+
+TODO example

### 2011-09-24 doc/theses/vitasek/chapter-5-future.tex
Thesis: in progress.

@@
 It is possible (and easy) to add more construction and improvement heuristics, as well as more metaheuristics in which the existing IHs are chained.
-A starting point is in Section TODO link appendix - how to write...\\
+A starting point is in Section \ref{section-appendix-iss-howto}.\\
 
-As it was mentioned in TODO link, the combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} can be seen as a sort of genetic programming.
+As it was mentioned in Section \ref{section-mip-ihs}, the combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} can be seen as a sort of genetic programming.
 However some modifications would still be necessary to make it a real instance of genetic algorithm metaheuristic.
@@
 The approach used in this work was strictly single-threaded, however there are in principle no limitations to extending this to a parallel, multi-threaded environment.
-For example, creating a pool of initial solutions in \heu{Glpk} construction heuristic can be improved by running several instances of GLPK solver in parallel - as GLPK on its own uses only a single thread to perform the computation.\\
+For example, creating a pool of initial solutions in \heu{Glpk} construction heuristic can be improved by running several instances of GLPK solver in parallel - as GLPK on its own uses only a single thread.\\
 

### 2011-09-25 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.

@@
 
-TODO define \textit{name(AM)}
+\begin{define}[$name()$]
+Given an attribute mapping $m = M_{x}^{y}$, $name(m)$ shall be defined as the string $x\-y$.
+\end{define}
 
@@
 
+For each AM $m$ there will be one binary variable $x_{name(m)}$.
+Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings.
+For each pair of AMs $m_1, m_2$ that cannot share the same ID set (because of conflicting type or image) there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$.
+$\mathbf{b}$ will be a vector of ones with corresponding length.
+
+\[
+\mathbf{x} =
+\begin{bmatrix}
+x_{y-a} \\
+x_{y-b} \\
+x_{y-c} \\
+x_{z-a} \\
+\end{bmatrix},
+\mathbf{c} = 
+\begin{bmatrix}
+weight(M_{y}^{a}) \\
+weight(M_{y}^{b}) \\
+weight(M_{y}^{c}) \\
+weight(M_{z}^{a}) \\
+\end{bmatrix} =
+\begin{bmatrix}
+0.2 \\
+0.2 \\
+0.6 \\
+0.4 \\
+\end{bmatrix},
+\mathbf{b} =
+\begin{bmatrix}
+1 \\
+1 \\
+1 \\
+1 \\
+\end{bmatrix},
+A =
+\begin{pmatrix}
+0 & 1 & 1 & 0 \\
+1 & 0 & 1 & 0 \\
+1 & 1 & 0 & 0 \\
+1 & 0 & 0 & 1 \\
+\end{pmatrix}
+\]
+
+The problem is, recall, to solve the following.
+
+\begin{eqnarray*}
+\max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
+s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
+\end{eqnarray*}
+
+In GLPK \textit{mathprog} language, it translates to the following formulation. % TODO cite mathprog
+
 \begin{scriptsize}
@@
 data;
-set AMs :=
-y-c
-y-b
-y-a
-z-a;
+set AMs := y-c y-b y-a z-a;
 param Weight :=
@@
 \end{scriptsize}
+
+We can use this as an input for the GLPK solver, and we get the solution: TODO GLPK output.
 
@@
 \subsubsection{\heu{FIDAX}}
+\label{section-mip-fidax}
 
@@
 \subsection{Improvement Heuristics}
+\label{section-mip-ihs}
 
@@
 TODO mention intensification, diversification
+
+TODO mention that combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} is a sort of genetic algorithm
 

### 2011-09-25 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

@@
 
-Furthermore, we denote $I$ % TODO fix the capital I
-the set of all ids and $V$ % TODO fix the capital V
-the set of all values in the document. We will need two more definitions from the article.
+Furthermore, we denote $\mathcal{I}$ the set of all ids and $\mathcal{V}$ the set of all values in the document. We will need two more definitions from the article.
 
@@
 
-\section{XML Keys}
+%\section{XML Keys} - is this necessary at all?
 
-TODO
+%TODO
 
-\subsection{According to Specification}
+%\subsection{According to Specification}
 
-TODO compare them to ID attributes
+%TODO compare them to ID attributes
 
-\subsection{According to \cite{keX}}
+%\subsection{According to \cite{keX}}
 
-TODO show how a XML key is defined using the keX notation % this is a terrible sentence
+%TODO show how a XML key is defined using the keX notation % this is a terrible sentence
 
@@
 \begin{define}[Attribute mapping]
-	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $I \times I$ % TODO fix the capital I
-	relation defined by
+	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $\mathcal{I} \times \mathcal{I}$ relation defined by
 	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}\]
@@
 
-\subsection{Attribute Mapping Model}
-
-TODO
+\paragraph{Attribute Mapping Model}
+An attribute mapping model is a data structure containing the information about all the AMs in a document, together with their images. We shall use this notion later in experimental part of this work.
 
 \section{ID Set}
+\label{section-definitions-id-set}
 
@@
 \begin{define}[ID set]
-A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an ID set iff
+A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an \textit{ID set} iff
 \[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset\]
@@
 
-\subsection{IDREF and IDREFS Attributes}
+\subsection{IDREF and IDREFS Condition}
 
-TODO
+Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked IDREF.
+\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i)\]
+Furthermore if $m$ contains multivalued attributes, it is to be marked IDREFS.
 
 \section{Attribute Mapping Weight}
+\label{section-definitions-weight}
 
-TODO this is from FIDAX again
+This definition of weight for AMs or AM sets comes from \cite{fidax} again.
 
-$M = \{m_1, \dots m_i\}$ is the set of all non-empty AMs in the document.
+Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.
 
@@
 
-TODO example
+From the cited article:
+\begin{quote}
+The support of attribute mapping $M_x^y$ is the fraction of edges in the XML tree that connect $x$ elements to $y$ attributes.
+\end{quote}
+
+\paragraph{Example}
+Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of every other mapping is $1 / (2+1+1+1) = 0.2$.
 
@@
 
-TODO example
+Again from the article:
+\begin{quote}
+The coverage of an attribute mapping measures how much of the image of that mapping occurs elsewhere, as a fraction of all mappings images in the document.
+\end{quote}
+
+\paragraph{Example}
+Coverage of $M_{y}^{a}$ in our example is $(0+0+1) / (2+1+1+1) = 0.2$. Coverage of $M_{z}^{a}$ is $0.2$ as well, all the other mappings have coverage $0$.
+\\
+
+\textit{Weight} of an attribute mapping is then defined as a linear combination of its support and coverage.
+
+\begin{define}[Weight]
+For $\alpha, \beta \geq 0$ as relative priorities of support and coverage we define the AM \textit{weight} as follows.
+\[weight(m) = \alpha . \phi(m) + \beta . \chi(m)\]
+For a set of AMs (which may or may not be an ID set) $S = \{m_1, \ldots m_i\}$ we define the weight of this set as the sum of the weights of its AMs.
+\[weight(S) = \sum_{m \in S} weight(m)\]
+\end{define}
+
+\section{Independent Set}
+
+We shall need the notion of an indepentent set of vertices in a graph and its weighted variant.
+
+\begin{define}[Independent set]
+Given an undirected graph $G = (V, E)$, a set of vertices $I \subseteq V$ is an \textit{independent set}, iff the following holds.
+\[\forall v_1, v_2 \in I, v_1 \neq v_2: (v_1, v_2) \notin E \]
+\end{define}
+
+\begin{define}[Maximum weighted independent set]
+Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied.
+\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v in I_{max}} w(v)\]
+\end{define}
+
+TODO finding this is NP-hard - cite
+
+\section{Linear Programming}
+
+TODO define LP rigorously
+
+\begin{eqnarray*}
+\max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
+s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
+\mathbf{x} & \geqslant & 0 \\
+\end{eqnarray*}
+\[
+\mathbf{x} =
+\begin{bmatrix}
+x_1 \\
+x_2 \\
+\ldots \\
+x_n \\
+\end{bmatrix},
+\mathbf{c} = 
+\begin{bmatrix}
+c_1 \\
+c_2 \\
+\ldots \\
+c_n \\
+\end{bmatrix},
+\mathbf{b} =
+\begin{bmatrix}
+b_1 \\
+b_2 \\
+\ldots \\
+b_m \\
+\end{bmatrix}
+\]
+
+Where $\mathbf{x}$ is the vector of variables (to be found by the optimization), $\mathbf{b}$ is the vector and $A$ is the accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function.
+$\mathbf{x}$ and $\mathbf{c}$ have length $n$, $\mathbf{b}$ has length $m$ and $A$ has dimensions $m \times n$.
+
+Another way to write this formulation is in the following.
+
+\begin{eqnarray*}
+\max_{x} z = \sum_{i=1}^{n} c_i x_i \\
+s.t.\, a_{11}x_{1} + a_{12}x_{2} + \ldots + a_{1n}x_{n} & \leqslant & b_{1} \\
+	   a_{21}x_{1} + a_{22}x_{2} + \ldots + a_{2n}x_{n} & \leqslant & b_{2} \\
+	   \ldots \\
+	   a_{m1}x_{1} + a_{m2}x_{2} + \ldots + a_{mn}x_{n} & \leqslant & b_{m} \\
+	   x_i \geqslant 0, i = 1, \ldots n \\
+\end{eqnarray*}
+
+TODO it is possible to formulate minimization function

### 2011-09-25 doc/theses/vitasek/chapter-2-research.tex
Thesis: in progress.

@@
 
-According to the article \cite{fidax}, ...
+According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set (Section \ref{section-definitions-id-set}) with the maximal weight (Section \ref{section-definitions-weight}) is NP-hard.
+The article suggests a heuristic approach described in Section \ref{section-mip-fidax}, which was incorporated into the framework proposed by this work.
 
-TODO talk about FIDAX
 To the best of our knowledge, there are no other articles dealing with this problem.
 
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
-
-\section{Independent Set}
-
-TODO define IS + MIS rigorously
+TODO talk about Fajt \cite{fajt} - but that's different, that's keys. Sometimes, it is translatable, but not always.
 
 TODO other approaches to max IS
+
+TODO LP is simple using Simplex method, however gets hard as soon as we require some of the variables to be integral or boolean.

### 2011-09-25 doc/theses/vitasek/appendix-c-trace.tex
Thesis: in progress.

@@
   Algorithm: Random
-    Time taken: 248 ms
-    Time since start: 248 ms
+    Time taken: 248 ms / Time since start: 248 ms
     Pool size: 10
@@
   Algorithm: RandomRemove, ratio = 0.2
-    Time taken: 0 ms
-    Time since start: 841 ms
+    Time taken: 0 ms / Time since start: 841 ms
     Pool size: 10
@@
   Algorithm: Mutation, ratio = 0.1, limit = 1 s
-    Time taken: 1512 ms
-    Time since start: 2710 ms
+    Time taken: 1512 ms / Time since start: 2710 ms
     Pool size: 11
@@
   Algorithm: Remove Worst
-    Time taken: 80 ms
-    Time since start: 7676 ms
+    Time taken: 80 ms / Time since start: 7676 ms
     Pool size: 12
@@
 
-% TODO make sure G.R. was introduced before! and LINK IT! swedish style!
-The second section states the input file along with its size and graph representation.\\
-Alpha and beta parameters for this instance belong here too.
+The second section states the input file along with its size and graph representation (Section \ref{section-experiments-data}). Alpha and beta parameters for this instance belong here too.
 
@@
 Algorithm: Random
-  Time taken: 248 ms
-  Time since start: 248 ms
+  Time taken: 248 ms / Time since start: 248 ms
   Pool size: 10
@@
 Algorithm: Mutation, ratio = 0.1, limit = 1 s
-  Time taken: 1512 ms
-  Time since start: 2710 ms
+  Time taken: 1512 ms / Time since start: 2710 ms
   Pool size: 11
@@
 
-Note that in this example trace there were no IDREF AMs found.
+Note that in this example trace there were no IDREF AMs found. \qed

### 2011-09-25 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

@@
 \chapter{Definitions}
+\label{chapter-definitions}
 
@@
 Now we return to \cite{fidax} to define the notion of an \textit{attribute mapping} (or AM for short). \nomenclature{AM}{Attribute Mapping}
-We will use a different definition (without introducing keys from \cite{keX} for now) that will however give us the same result.
+We will use a different definition (without introducing keys from \cite{keX}) that will however give us the same result.
 
@@
 \section{Independent Set}
+\label{section-definitions-is}
 

### 2011-09-25 doc/theses/vitasek/chapter-2-research.tex
Thesis: in progress.

@@
 
-According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set (Section \ref{section-definitions-id-set}) with the maximal weight (Section \ref{section-definitions-weight}) is NP-hard.
+According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set (Section \ref{section-definitions-id-set}) with the maximal weight (Section \ref{section-definitions-weight}) can be transformed into the maximum weighted independent set (Section \ref{section-definitions-is}) and is thus NP-hard.
+
+TODO describe the transformation ID -> IS (other way around is in fidax)
+
 The article suggests a heuristic approach described in Section \ref{section-mip-fidax}, which was incorporated into the framework proposed by this work.
 
-To the best of our knowledge, there are no other articles dealing with this problem.
+To the best of our knowledge, there are no other articles dealing with this problem.\\
 
-TODO talk about Fajt \cite{fajt} - but that's different, that's keys. Sometimes, it is translatable, but not always.
+TODO talk about Fajt \cite{fajt} - but that's different, that's keys. Sometimes, it is translatable, but not always.\\
 
-TODO other approaches to max IS
+TODO our problem now is to solve max IS.
 
-TODO LP is simple using Simplex method, however gets hard as soon as we require some of the variables to be integral or boolean.
+TODO other approaches to max IS - approximation methods?

### 2011-09-25 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 \section{Experimental Data}
+\label{section-experiments-data}
 
@@
 
-Consider the following XML file fragment.
+Recall the following XML file fragment from Chapter \ref{chapter-definitions}.
 \begin{verbatim}
@@
 
-The ``user-specified limits" of whiskers are set to default value, let us quote again from the manual.
+The ``user-specified limits" of whiskers are set to default value, let us quote from the manual again.
 

### 2011-09-25 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.

@@
 
-TODO rigorously define MIP - optimization problem similar to LP, however some of the variables integral or boolean
+We have shown that ID set -> IS, now we show the translation to MIP formulation and solve it using GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
 
-TODO FIDAX is cool, but their heuristic might not be the best. It is a IS problem, let's try a multi-purpose solver. Use GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
+TODO MIP formulation: LP is simple using Simplex method, however as soon as we require integrality or boolean values, it is NP-hard.
+
+\begin{define}[Mixed integer problem]
+	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some variables are limited to integral or boolean (0, 1) values.
+\end{define}
+
+TODO show how to translate IS to MIP
 
@@
 
-We can use this as an input for the GLPK solver, and we get the solution: TODO GLPK output.
+We can use this as an input for the GLPK solver, and we get the solution.
 
- -> wow, GLPK works. However, it takes too long to get to optimum (link experiments), so let's try other heuristics.
+\begin{scriptsize}
+\begin{verbatim}
+...
+Problem:    glpk_input
+Rows:       5
+Columns:    4 (4 integer, 4 binary)
+Non-zeros:  12
+Status:     INTEGER OPTIMAL
+Objective:  z = 0.6 (MAXimum)
+
+...
+
+   No. Column name       Activity     Lower bound   Upper bound
+------ ------------    ------------- ------------- -------------
+     1 x[y-a]       *              1             0             1 
+     2 x[y-b]       *              0             0             1 
+     3 x[y-c]       *              0             0             1 
+     4 x[z-a]       *              0             0             1
+
+...
+\end{verbatim}
+\end{scriptsize}
+
+This output tells us that the solution is $x_{y-a} = 1$, $x_{y-b} = 0$, $x_{y-c} = 0$ and $x_{z-a} = 0$.
+This means that the optimal ID set with maximum weight contains only the $M_{y}^{a}$ attribute mapping.\\
+
+It is obvious that this approach works and for any possible input we can let GLPK find the optimal solution.
+However, sometimes it takes too long to find the optimum (see e.g.
+Section \ref{section-glpk-comparison}), we should try to improve this process.
 

### 2011-09-25 doc/theses/vitasek/chapter-6-conclusion.tex
Thesis: in progress.

@@
 
-TODO conclusion: From all integrity constraints in XML we chose the ID attributes and decided to improve the search for them.
-We discussed the approach chosen in FIDAX.
-Based on this article we introduced the MIP approach and demonstrated how to find the optimal (w.r.t.
-weight) ID set.
+From all the integrity constraints in XML we chose the ID / IDREF / IDREFS attributes and decided to improve upon the search for them.
+We discussed the approach from \cite{fidax} and the equivalence of ID set search and maximum weighted independent set.
+Based on this article we introduced the MIP approach and demonstrated how to find the optimal ID set using external GLPK solver in the environment of jInfer framework.\\
 
-However, this approach takes too long for some inputs, so we introduced a whole spectrum of construction as well as improvement heuristics.
-By combining these algorithms and tuning their parameters we were able to find very good solutions while maintaining low running times.
+However, this approach took too long for some inputs, so we introduced a whole range of construction as well as improvement heuristics.
+We combined these algorithms to create a metaheuristic and performed a number of experiments to understand its behavior.
+Finally we selected a promising metaheuristic strategy and tuned its parameters to find very good ID sets while maintaining low running times.\\
+
+To the best of our knowledge, at the time of writing this work is our approach to finding ID attributes the best one known.

### 2011-09-29 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to some \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that we can try to do more than just let the solver work: by employing various heuristics.
+At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to a given \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that there are several possibilities how to optimize the work of the solver by employing various heuristics.
 
-Now we shall attempt to implement these ideas. But before we describe the experiments themselves, we should try to formulate our aim.\\
+We have implemented these ideas and incorporated them in the jInfer framework (see Appendix \ref{appedix-jInfer}).
+But before we describe the experiments themselves, we should try to formulate our aim.\\
 
-First of all, we will try to get understanding of how the whole system and its components behave.
-We would like to see the changes introduced by modifying some of the key parameters, while keeping the others fixed.
+First of all, we describe how the whole system and its components behave.
+We want to see the changes introduced by modifying several parameters, while keeping the others fixed.
 They probably will not be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
-In the end we should have at some intuition about the system.
 
-Second, we will try to evaluate the system performance in terms of the speed of finding good heuristic results. We will try to find tweaks to make the whole process as fast as reasonably possible.
+Second, we evaluate the system performance in terms of the speed of finding good heuristic results. We find tweaks to make the whole process as fast as reasonably possible.
 
-And in the end, we should be able to formulate some kind of general recommendations regarding the problem of finding ID sets.\\
-
-This chapter will be structured in the following way: first we will discuss the experimental data we used, then the methodology used in conducting the experiments, followed by the actual list of experiments with their full description and results, and in the end we shall draw some final conclusions.
+And in the end, we formulate general recommendations regarding the problem of finding ID sets.
 
@@
 
-The first section will discuss the XML data we will be using to conduct out experiments. We are using XML documents of three categories:
+To conduct out experiments, we are using XML documents of three categories:
 
@@
 
-A short reasoning for this choice: realistic - we want to see the performance in cases taken from the real world.
+In case of realistic data we want to see the performance in cases taken from the real world.
 The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as simple text nodes instead of attributes.
 We will try to convert some of these values to attributes (e.g.
-using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see \ref{realistic-converted} for details).
+using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see Section \ref{section-realistic-converted} for details).
 And finally, we create completely artificial data to create inputs that will put our heuristics in stress.
-This is because the realistic data often prove to be too easy to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.
+This is because the realistic data often prove to be too simple to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.
 
@@
 
-To understand our test data sets we will discuss their origin and \textit{graph representation}. As mentioned earlier, % TODO link
+To understand our test data sets we discuss their origin and \textit{graph representation}. As mentioned earlier, % TODO link
 the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set in a graph.
@@
 
-The former will be achieved with the help of GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
+The former will be achieved with the help of the GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
 Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
@@
 From 3 different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
-Their summary is the Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
+Their summary is Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
 Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
@@
 
-To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but are relatively small. We can consider them to be the ``typical" representants.
+To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical" representants.
 
-On the other hand, the \dataset{XMA-*} sets are relatively huge, but trivial: their only candidate AM will just get picked and that the heuristic will end.
-We will see the performance of the other components of the whole system, such as loading the data sets into memory representations.
+On the other hand, the \dataset{XMA-*} sets are relatively huge, but trivial: their only candidate AM will just get picked and the heuristic will end.
+Therefore we will see the performance of the other components of the whole system, such as loading the data sets into memory representations.
 
-Finally, the \dataset{XMD} set is relatively big and at the same time has non-trivial graph representation.
+Finally, the \dataset{XMD} set is relatively big and, at the same time, has non-trivial graph representation.
 In this case we should see a performance more balanced between processing and finding the ID set.
@@
 \subsection{Realistic data with artificial attributes}
-\label{realistic-converted}
+\label{section-realistic-converted}
 
@@
 None of these sets had any attributes before the conversion.
-Their summary is the Table \ref{table-experiments-data-converted}, their graphs are in Figure \ref{image-experiments-data-converted}.
+Their summary is Table \ref{table-experiments-data-converted}, their graphs are in Figure \ref{image-experiments-data-converted}.
 
-To address the conversion: in the case of \dataset{MSH} we found 2 elements with values resembling a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
+To address the conversion: in case of \dataset{MSH} we found 2 elements with values resembling a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
 In the case of \dataset{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
@@
 This approach is useful, because as stated in TODO link, ID attributes are a special case of XML keys.
-We can use this approach to find XML keys: convert some suspicious data into attributes, find the optimal ID set and then create XML key based on this ID set.
+We can use this approach to find XML keys: convert some ``suspicious" data into attributes, find the optimal ID set and then create XML key based on this ID set.
 
@@
 
-Again interpretation: in the case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
+In case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
 This is then the case similar to \dataset{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
 
-In the case of \dataset{NTH} we introduced more attributes, 8 to be precise.
+In case of \dataset{NTH} we introduced 8 attributes.
 Out of them 5 proved to be candidate AMs, with 7 edges constraining them.
@@
 As soon as we started experimenting with the data coming from the real world, it was obvious that they are not complex enough.
-After we build the model, we get the most complex graphs of 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
+After we built the model, we got the most complex graphs of 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
 Our solution is to approach the problem from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
@@
 
-Consider the following excerpt from an XML file:
+To demonstrate this, consider the following excerpt from an XML file:
 
@@
 First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that these values will not conflict with any others.
+To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that they do not conflict with others.
 Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
@@
 
-The pseudocode for this is in Listing \ref{listing-random-data}.
+The respective pseudocode for this is provided in Algorithm \ref{listing-random-data}.
 
@@
 
-We have created 3 sets to use in experiments along with the realistic and converted sets, these are called \dataset{100-100}, \dataset{100-200} and \dataset{100-1000}.
+We have created 3 sets to be used in experiments along with the realistic and converted sets, called \dataset{100-100}, \dataset{100-200} and \dataset{100-1000}.
 Note that the name is always in the form $v-e$.\\
@@
 For this reason we created 11 more sets, from \dataset{0-0} as the trivial one to \dataset{100-500} as the largest one.
-From now on, these will be referred to as \textit{sized test data (sets)}.
-We wanted to keep density the same among these sets, and we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
+These will be referred to as \textit{sized test data (sets)}.
+We wanted to keep the same density among these sets, so we picked the $\frac{|E|}{E_{max}}$ density interpretation for this.
 
-The summary is the Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}, these tables contain 2 new columns: values of density in both interpretations we introduced.
+A summary is provided in Table \ref{table-experiments-data-artificial} and Table \ref{table-experiments-data-artificial-size}; these tables contain 2 new columns: values of density in both interpretations we introduced.
 Some of the graph representations can be seen in Figure \ref{image-experiments-data-artificial}.
@@
 While studying the tables it becomes obvious that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because how the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set.
+This is because of the way the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set.
 Because of the so-called Birthday paradox (see e.g.
@@
 There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximal quality theoretically achievable has to do with the $\frac{|E|}{|V|}$ density, not with the one we fixed.
-Exploration of this phenomenon is among the possibilities of future work of this thesis.
+This intuitively suggests that the maximum quality theoretically achievable is related to the $\frac{|E|}{|V|}$ density, not to the one we fixed.
+Exploration of this phenomenon is beyond the scope of this thesis.
 
-On a final note, the artificial data we used in experiments can be found on the DVD enclosed with this work.
+Note that the artificial data we used in experiments can be found on the DVD enclosed with this work.
 
@@
 
-As was mentioned before, we will be using an extension to the jInfer framework called \jmodule{IDSetSearch}.
-Please see appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
+As was mentioned before, we will use an extension to the jInfer framework called \jmodule{IDSetSearch}.
+Please see Appendices \ref{appendix-jInfer} and \ref{appendix-iss} for more detailed information on these two pieces of software.\\
 
-We now have to define a few notions before moving forward to the description of our experiments.
+We now have to introduce a few notions before moving forward to the description of our experiments.\\
 
-\begin{define}[Experiment parameters]
-	\label{define-experiment-params}
-	\textit{Experiment parameters} are the following.
+\textit{Experiment parameters} are the following ones.
 \begin{itemize}
@@
 \end{itemize}
-\end{define}
 
-\begin{define}[Experiment(al) configuration]
-	\label{define-experiment-config}
-	An \textit{experiment instance}, or \textit{configuration}, is one specific setting of all (experiment) parameters.
-\end{define}
+An \textit{experiment instance}, or \textit{experiment configuration}, is one specific setting of all experiment parameters.\\
 
-\begin{define}[Experiment(al) set]
-	\label{define-experiment-set}
-	One or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experimental set}.
-\end{define}
+And finally, one or more experiment configurations, regardless whether their parameters differ, constitute an \textit{experiment set}.
 
@@
 
-An input data set is essentialy a single XML file on the filesystem, however there is a straightforward extension to multiple files conforming to the same schema.
+An input data set is a single XML file on the filesystem; however, there is a straightforward extension to multiple files conforming to the same schema.
 The first step in this process is to use jInfer's module \jmodule{BasicIGG} module (see \cite{basiciggdoc} for details) to obtain a list of rules - an \textit{initial grammar} (IG).
@@
 The second step is to convert the grammar into the AM model.
-This is done by scanning it in a linear fashion and retrieving a so-called \textit{flat} representation.
+This is done by a linear scan and retrieving a so-called \textit{flat} representation.
 This consists of a list of tuples in the following format.
@@
 \subsubsection{Example}
-
-This example will present the whole process, from an input XML file to the model and its AMs and types.
 
@@
 
-We will be using the following machine when conducting our experiments.
+We will use the following configuration when conducting our experiments.
 
@@
 
-We will attempt to shield our experiment from the influence of the environment as much as reasonably possible.
+We will attempt to protect our experiment from the influence of the environment as much as reasonably possible.
 First of all, NetBeans running the experiments is the only relevant program running in the system while the experiments are performed.
-Unfortunately, NetBeans itself is quite a large environment with a life of its own, and we would most certainly get more reliable results if we could run our experiments outside of it.
+Unfortunately, NetBeans itself is quite a large environment, and we would most certainly get more reliable results if we could run our experiments outside of it.
 This improvement is left for the future work.
 
-Also, every experimental configuration is run 50 times in hope that the effects of any events adversely affecting our results (e.g.
+Also, every experimental configuration is run 50 times so that the effects of any events adversely affecting our results (e.g.
 OS deciding to run some house cleaning) will be averaged out.
-Whenever possible, we will always be using boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
+Whenever possible, we will use boxplots instead of a simple average (or average and variance) to present results of these multiple runs.
 
@@
 The result cannot be interpreted in an absolute manner, but by subtracting the time at the start from the time at the end, we can get a reasonably reliable measurement.
-We then have to divide it by one million to obtain the number of milliseconds passed.
 
@@
 To present a set of measurements obtained by iteratively running an experiment we shall prominently use the \textit{boxplot} chart.
-Because we use boxplots produced by GnuPlot, let us quote its manual (\cite{gnuplot-manual}) for the exact definition.
+Because we use boxplots produced by GnuPlot, let us quote its manual \cite{gnuplot-manual} for the exact definition.
 

### 2011-09-29 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
 
-The first experiment set will try to establish how long it takes to extract the IG from the input XML file and how long does it take to create the AM model from this IG.
+The first experiment set will try to establish how long it takes to extract the IG from the input XML file and to create the AM model from this IG.
 For now, we will not be running or measuring any heuristics.
@@
 
-Results are in Table \ref{table-experiments-grammar-model-timing}.
+The results are captured in Table \ref{table-experiments-grammar-model-timing}.
 We are presenting the average grammar extraction (GE) times and their standard deviation, the same for model creation (MC) and total (sum of these two, Tot) times.
-For many data sets the average time is less than 10 ms: this is not enough to be precise and we don't calculate the standard deviation in these cases.
+For many data sets the average time is less than 10 ms: this is not enough to be precise and we do not calculate the standard deviation in these cases.
 
@@
 
-A related problem is how long it takes to create input for GLPK and then parse its results.
+A related problem is how long it takes to create an input for GLPK and then parse its results.
 We will use the same test data sets as in the previous case, but now we will gather times needed to communicate with GLPK.\\
@@
 
-Results are in table \ref{table-experiments-glpk-timing}.
+The results are captured in Table \ref{table-experiments-glpk-timing}.
 For each data set there are the times of (GLPK) input creation (IC) - average and standard deviation, then the same for output parsing (OP) and total (Tot).
@@
 Interestingly enough, in most cases the times to create an input for GLPK and then to parse its output are very similar.
-Also, for sized test data sets it is interesting to note that even though the $|V|$ and $|E|$ counts are increasing, the times remain almost the same.
-This probably due to the fact that IC and OP times include the I/O when writing to a file for GLPK or reading the file it produced, and these times are probably the most relevant.
+Also, for sized test data sets it is interesting to note that even though the $|V|$ and $|E|$ numbers are increasing, the times remain almost the same.
+This is probably due to the fact that IC and OP times include the I/O when writing to a file for GLPK or reading the file it produced, and these times are probably the most relevant.
 
@@
 In this case we will see the dependency between the run time and the quality achieved in it.
-In the second time we will let GLPK run until optimum is found.
+In the second part we will let GLPK run until the optimum is found.
 We shall see the dependency between input size and run time needed to achieve the optimum.
@@
 Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
-Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-1}).
+Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Algorithm \ref{listing-experiment-glpk-native-vs-cygwin-1}).
 There will be no improvement heuristic.
@@
 
-Results are in Figure \ref{image-experiment-time-vs-quality}.
+The results of this test are captured in Figure \ref{image-experiment-time-vs-quality}.
 They should be interpreted as follows: for each time limit from 1 to 46 seconds there are two boxplots next to each other, the left, dashed one is the native GLPK, the right, solid one is the Cygwin GLPK.
@@
 Every configuration will let \heu{Glpk} CH run for unlimited time, until it finds the optimum.
-This will be repeated in 50 iterations for each of the 11 files from the sized test data set (see Listing \ref{listing-experiment-glpk-native-vs-cygwin-2}).
+This will be repeated in 50 iterations for each of the 11 files from the sized test data set (see Algorithm \ref{listing-experiment-glpk-native-vs-cygwin-2}).
 There will again be no IH, the only data we will collect are the times of the CH run in each case.\\
@@
 \begin{algorithm}
-\caption{GLPK: native vs. Cygwin set generation 2}
+\caption{GLPK: Native vs. Cygwin Set Generation 2}
 \label{listing-experiment-glpk-native-vs-cygwin-2}
@@
 
-Results are in Figure \ref{image-experiment-time-until-optimum}, please take a note that the Y axis is in log scale.
+The results are captured in Figure \ref{image-experiment-time-until-optimum}; please take a note that the Y axis is in log scale.
 As with the previous case, the X axis cannot be interpreted in the usual way.
@@
 This becomes less extreme with the increasing input size, which leads us to suspicion that the core parts of computation in both cases are equally powerful.
-Regardless of that, we shall be using the \textbf{native} GLPK for following experiments.
+Regardless of that, we shall be using the \textbf{native} GLPK for the following experiments.
 
@@
 There will be no improvement heuristics.
-The pool size will be set to 10, even though \heu{FIDAX} cannot not profit from this.
-Listing for this can be found in \ref{listing-experiment-random-fuzzy-fidax}.
+The pool size will be set to 10, even though \heu{FIDAX} cannot profit from this.
+Listing for this can be found in Algorithm \ref{listing-experiment-random-fuzzy-fidax}.
 
@@
 
-We can draw the following conclusions.
-\heu{Fuzzy} consistently finds the best solution, but it's by far the slowest of these CHs.
+We can draw the following conclusions: \heu{Fuzzy} consistently finds the best solution, but it is by far the slowest of these CHs.
 The trivial \heu{Random} is better than \heu{FIDAX} in artificial as well as some real data.
@@
 
-The experimental results are summarized in the Table \ref{table-experiments-fidax-and-hungry} and are quite surprising.
+The experimental results are summarized in Table \ref{table-experiments-fidax-and-hungry} and are quite surprising.
 As trivial a heuristic \heu{Hungry} is, it is still able to improve the ID set found by \heu{FIDAX} by as much as almost 50\% (the last row, \dataset{100-1000}).
@@
 This experiment will consist of 50 iterations * 11 data sets * 6 CHs = 3300 experimental configurations.
-See the Listing \ref{listing-experiment-best-standalone-ch} for details.
+See the Algorithm \ref{listing-experiment-best-standalone-ch} for details.
 This time we are not interested in run times, only in qualities which we shall gather in a format for GnuPlot.\\
@@
 
-The next logical step will be to try to add one IH after the best CH we have found, \heu{Glpk}.
+The next logical step is to try to add one IH after the best CH we have found, \heu{Glpk}.
 We will investigate all IHs except for \heu{RandomRemove} and \heu{RemoveWorst}, which cannot help us at this time.
@@
 Setup used will be almost identical to that from the previous experiment.
-Experimental set will consist of 6 data sets * 50 iterations * 2 CHs = 600 experimental configurations, see Listing \ref{listing-experiment-ch-for-mutation}.
+Experimental set will consist of 6 data sets * 50 iterations * 2 CHs = 600 experimental configurations, see Algorithm \ref{listing-experiment-ch-for-mutation}.
 We shall collect the eventual quality after running both the CH and the IH in format suited for GnuPlot.\\
@@
 
-After finding the best combination of a CH and IH we turn our attention to some of the parameters. The first ones are the $\alpha$ and $\beta$ from the definition of our weight function. % TODO link
+After finding the best combination of a CH and IH we turn our attention to parameters. The first ones are the $\alpha$ and $\beta$ from the definition of our weight function. % TODO link
 A short reminder: the weight is defined as TODO copy formula.
@@
 This means that investigating effects of these parameters is in fact a 1-dimensional problem.
-However, for simplicity's sake we will use 25 combinations of various $\alpha$ and $\beta$ and normalize them only during evaluation.
-
-It is worthy noting that we do not expect any changes in performance of heuristics and we will limit the inquiry to different ID sets produced under different settings.
+However, for the sake of simplicity we will use 25 combinations of various $\alpha$ and $\beta$ and normalize them only during evaluation.
+
+We do not expect any changes in performance of heuristics and we will limit the inquiry to different ID sets produced under different settings.
 
@@
 This experimental set will contain 5 different $\alpha$ settings * 5 $\beta$ settings * 8 data sets = 200 experimental configurations.
-We are not using the artificial data sets, because the way they are generated (attribute values are random numbers) they cannot possibly create different optimal ID sets.
-The pseudocode capturing this is in Listing \ref{listing-experiment-various-betas}.
+We are not using the artificial data sets, because due to the way they are generated (attribute values are random numbers), they cannot possibly create different optimal ID sets.
+The pseudocode capturing this is provided in Algorithm \ref{listing-experiment-various-betas}.
 We will use \heu{Glpk} constrained to 1 second (thus making it an instance of \heu{Truncated Branch \& Bound}) and no IHs.
@@
 
-Following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.
-The \dataset{OVA*} data sets showed various dependencies on $\alpha$ and $\beta$, we shall now describe one representative example.
+The following data sets have the same optimal ID sets regardless of the setting of $\alpha$ and $\beta$: \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}.
+The \dataset{OVA*} data sets showed various dependencies on $\alpha$ and $\beta$; we shall now describe one representative example.
 
@@
 Table \ref{table-experiments-various-betas-ova1-effect} summarizes the dependency of the ID set found on various values of $\alpha$, $\beta$.
-We than define the $\alpha-ratio$ as $ \dfrac{\alpha}{\alpha + \beta} $ and summarize the findings in a linear manner, sorted by increasing $\alpha-ratio$ in Table \ref{table-experiments-various-betas-ova1-ratio-effect}.
+We then define the $\alpha-ratio$ as $ \dfrac{\alpha}{\alpha + \beta} $ and summarize the findings in a linear manner, sorted by increasing $\alpha-ratio$ in Table \ref{table-experiments-various-betas-ova1-ratio-effect}.
 Note that the $\alpha-ratio$s are not unique due to the way we constructed the experimental configurations here.
@@
 Interestingly enough, there is no clear separation between the two ID sets depending on the $\alpha-ratio$ to be found.
-The very existence of the two sets might be due to the fact that \heu{Glpk} randomizes the order in which AMs are presented to the external GLPK solver.
-However, this question is outside of the scope of this work, and shall be left for future work.
+The existence of the two sets might be due to the fact that \heu{Glpk} randomizes the order in which AMs are presented to the external GLPK solver.
+However, this question is beyond the scope of this work, and shall be left for future work.
 
@@
 
-Our experimental set will contain 50 iterations * 2 = 100 experimental configurations as described in Listing \ref{listing-experiment-ignore-text-data}.
+Our experimental set will contain 50 iterations * 2 = 100 experimental configurations as described in Algorithm \ref{listing-experiment-ignore-text-data}.
 We will be using \heu{Glpk} limited to 1 second with no additional IH and pool size set to 1.
@@
 
-Results are summarized in Figure \ref{image-experiment-ignore-text-data}.
+The results are summarized in Figure \ref{image-experiment-ignore-text-data}.
 Boxplots drawn in dashed lines represent the original case, not ignoring the text data.
@@
 Interestingly, the grammar extraction times tend to be shorter in the case when text data is not ignored, although this is inconclusive.
-However, there is a clear improvement of about 50 \% in the case of model creation times.
+However, there is a clear improvement of about 50 \% in case of model creation times.
 The conclusion then is to ignore the simple text node content whenever possible when finding ID attributes.
@@
 
-The 3 strategies we will be assessing will be constructed from the following instances of improvement heuristics:
+The 3 strategies we assess will be constructed from the following instances of improvement heuristics:
 
@@
 The experimental set will consist of 3 strategies * 11 data sets * 20 iterations = 660 experimental configurations.
-Their construction is formalized in the Listing \ref{listing-experiment-chaining-ihs}.
+Their construction is formalized in the Algorithm \ref{listing-experiment-chaining-ihs}.
 The construction heuristic will be \heu{Random} with pool size 10.
@@
 
-Data gathered will be traces like the one in Appendix \ref{appendix-trace} - after each iteration, the time taken so far and the quality of incumbent solution is noted.\\
+Log traces like the one in Appendix \ref{appendix-trace} will be gathered - also, after each iteration, the time taken so far and the quality of incumbent solution is noted.\\
 
@@
 Its X position denotes the absolute time in which this step finished, and its Y position represents the incumbent solution quality after this step.
-Every time a line disappears before reaching 10 seconds it means that this metaheuristic run found the optimum before the 10 second mark.
-There is an obvious repetitive regularity in the shape of each line, this corresponds to the fact that there is a finite number of IHs in this strategy (5 of them in Strategy 1) which repeat over time.
+Every time a line ``disappears" before reaching 10 seconds it means that this metaheuristic run found the optimum before the 10 second mark.
+There is an obvious repetitive regularity in the shape of each line, which corresponds to the fact that there is a finite number of IHs in this strategy (5 of them in Strategy 1) which repeat over time.
 The obvious similarity between different lines corresponds to the fact that each run is from the same strategy, and over time, they do the same steps.
@@
 There are two more graphs like this in Figures \ref{image-experiment-chained-ihs-s2} and \ref{image-experiment-chained-ihs-s3} for comparison, capturing Strategy 2 and Strategy 3 respectively working on the same data set, \dataset{100-100}.
-Describing them in detail is unfortunately outside of the scope of this work.\\
+Describing them in detail is beyond the scope of this work.\\
 
@@
 Easily we can discard \dataset{MSH}, \dataset{NTH}, \dataset{XMA-c}, \dataset{XMA-p}, because the optimum is found in the very first step.
-Let us now introduce a metric for assessment of a strategy: namely, how many times of the 20 runs did it find the optimum.
-Results of this are summarized in Table \ref{table-experiments-chained-ihs-tweaking}.
+Let us now introduce a metric for assessment of a strategy: namely how many times of the 20 runs it found the optimum.
+The respective results are summarized in Table \ref{table-experiments-chained-ihs-tweaking}.
 
@@
 Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs.
-Highlited are the strategies that performed best on that data set.
-We see that Strategy 1 and 3 are very similar in performance.
+The strategies that performed best on that data set are highlited.
+We can see that Strategies 1 and 3 are very similar in performance.
 We shall nonetheless choose Strategy \textbf{1} as the winner for its simplicity.
@@
 The total limit will remain at 10 seconds.
-The process of creating the configurations is captured in Listing \ref{listing-experiment-chained-ihs-tuning}.
-We will be gathering the following information for each run: what were the parameters, how long did the run take and whether it found optimum.\\
+The process of creating the configurations is captured in Algorithm \ref{listing-experiment-chained-ihs-tuning}.
+We will be gathering the following information for each run: what were the parameters, how long did the run take and whether it found the optimum.\\
 
@@
 
-After averaging the data we get a large result table, an excerpt from which is in Table \ref{table-experiments-chained-ihs-tweaking-s1}.
+After averaging the data, we get a large result table; an excerpt of it is in Table \ref{table-experiments-chained-ihs-tweaking-s1}.
 Only the results for the biggest data sets and a few combinations of RR, MUT and CX ratios are presented.
@@
 In the right part are the averaged running times (AT), success ratios (ratio of runs that found the optimum, SR) and average qualities (AQ) for each data set.
-Highlited are the highest success ratios and qualities.\\
+The highest success ratios and qualities are highlited.\\
 
@@
 As with the previous one, \heu{Random} will be the CH, IHs will be from Strategy 1, however this time there will be no time limit.
-The process of generating the experimental set is in Listing \ref{listing-experiment-chained-ihs-tuned}.
+The process of generating the experimental set is in Algorithm \ref{listing-experiment-chained-ihs-tuned}.
 
@@
 
-After asking and answering a lot of questions related to the overall system behavior, parameter effects and various heuristic combinations is now the time to summarize our results and draw conclusions.\\
+After answering a lot of questions related to the overall system behavior, parameter effects and various heuristic combinations we can now summarize our results and draw conclusions.\\
 
@@
 It will find the optimum eventually, even though this might take minutes or hours to complete.
-For many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter how long it takes.\\
-
-Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1.
+For many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it does not matter how long it takes.\\
+
+Secondly, if we do not have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1.
 In all our realistic data sets the optimum was found almost instantly, and the most complex and bigges artificial data sets took only around 1/4 of the time to finish, compared to \heu{Glpk}.\\

### 2011-09-29 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
+% TODO uncomment this to get much shorter ToC
+% \setcounter{tocdepth}{1}
+
 \setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
@@
 \def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Irena Mlýnková Ph.D.}
+\def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá Ph.D.}
 \def\mfplacedate{Praha, 2011}
@@
 
-Vedoucí diplomové práce: \mfadvisor{}, Katedra softwarového inženýrství
+Vedoucí diplomové práce: \mfadvisor{}, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství
 
@@
 
-As usual, source code excerpts, class, field and method names shall be written in fixed-width font, such as \texttt{getHeuristic()}.
+As usual, source code excerpts, class, field and method names shall be written in fixed-width font, such as \texttt{get\-Heu\-ris\-tic()}.
 Names of specific heuristics will be written like this: \heu{Mutation}.

### 2011-09-29 doc/theses/vitasek/list-labels.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+cat *.tex | grep "\\\\label" | sed "s/^.*\\label{\([^}]*\)}.*$/\1/" | sort

### 2011-10-07 doc/theses/mikula/introduction.tex
Master thesis: In progress.

\chapter{Introduction}
Extensible Markup Language (XML) has become a wide-spread standard for data representation and exchange. Its popularity is based on its simplicity and at the same time its flexibility and express power.

With XML used to represent data themselves, it is often needed or convenient to somehow specify a structure of the represented data, their format, internal relations or restrictions, etc. In order to this need, several so-called XML schema languages (XML schemata) have been created. The most used of them are Document Type Definition (DTD) and XML Schema proposed by W3C.

Nevertheless, it is encouraged to use an XML schema along with an XML data representation, in practise it is done sparsely. Commonly, XML documents are not assigned with their respective XML schema at all or the schema is out-dated due to modifications in a structure of data without updating the schema.

Recently in reaction to this situation, significant number of approaches dealing with an automatic construction of an XML schema have been proposed. The aim is that a set of XML documents is available and it is desired to infer a (non-trivial) XML schema, so that the XML documents are valid against it. Most of published approaches are of this type - input is a set of XML documents - and they are based on various ideas and can be classified by several aspects, as discussed in chapter \todo{ref}.

Besides the mentioned type, some approaches that utilize other or additional sources has been developed. If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery queries, etc, these sources can be exploited to refine the inference process. The refinement can be achieved in various aspects such as decreasing the speed of the process, getting a more precise, more concise or more readable result or inference of some statements about the data which cannot be (easily) extracted from the data themselves.

Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (also discussed in chapter \todo{ref}), leaving a wide space for a possible future research.

\todo[inline]{aim of this work}
\todo[inline]{structure of this work}

### 2011-10-09 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress

@@
 \label{chapter-definitions}
-
-TODO we shall introduce a number of definitions, some only auxiliary, some necessary in the whole work
 
@@
 
+\begin{define}[$name()$]
+	Given an attribute mapping $m = M_{x}^{y}$, $name(m)$ shall be defined as the string $x-y$.
+\end{define}
+
 \section{ID Set}
@@
 Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied.
-\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v in I_{max}} w(v)\]
+\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v)\]
 \end{define}
 
-TODO finding this is NP-hard - cite
+It is well known that finding the maximum weighted IS is an NP-hard optimization problem, see e.g. \cite{JM1986425}.
 
@@
 
-TODO define LP rigorously
+The problem of \textit{linear programming} is optimization of a linear function under a set of linear constraints.
+The formulation is usually called a \textit{linear program}.
+It can be written in the following form (minimization version is possible too).
 
@@
 
-TODO it is possible to formulate minimization function
+Solving a linear program is usually possible in polynomial time using the \textit{simplex algorithm} described for example in \cite{dantzig1998linear}.
+
+\section{Mixed Integer Problem}
+
+\begin{define}[Mixed integer problem]
+	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some or all variables are limited to integral or boolean (0, 1) values.
+\end{define}
+
+Solving MIP in general is $\NP$-hard.

### 2011-10-09 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress

@@
 
-We have shown that ID set -> IS, now we show the translation to MIP formulation and solve it using GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
+In this chapter we introduce a new approach to finding maximum ID sets.
+First, we transform the problem formulation to maximum weighted IS problem formulation.
+Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK.
+\nomenclature{GLPK}{GNU Linear Programming Kit}
+We will continue by applying heuristical approaches to improve the performance of the process.
 
-TODO MIP formulation: LP is simple using Simplex method, however as soon as we require integrality or boolean values, it is NP-hard.
+\section{ID Set to IS Formulation}
 
-\begin{define}[Mixed integer problem]
-	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some variables are limited to integral or boolean (0, 1) values.
-\end{define}
+Given $C = \{m_1, \ldots, m_n\}$ a set of all AMs in a document, we construct a graph $G = (V,E)$ as follows.
+For each AM $m_i \in C$ we create a vertex $v_{name(m_i)}$.
+Two vertices $v_{name(m_i)}$ and $v_{name(m_j)}$ shall be connected by an edge iff they cannot share the same ID set, either because they have the same type ($\tau(m_i) = \tau(m_j)$), or their images intersect ($\iota(m_i) \cap \iota(m_j) \neq \emptyset$).
+Weight of a vertex $v_{name(m_i)}$ is the weight of the attribute mapping: $w(v_{name(m_i)}) = weight(m_i)$.
 
-TODO show how to translate IS to MIP
+Now finding the maximum weighted IS in $G$ finds the maximum (optimal) ID set in the original document.
 
-\section{Finding ID sets with GLPK}
+\section{IS to MIP Formulation}
 
-TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...
+Given a graph $G = (V,E)$ with a weight function $w: V \rightarrow \mathbb{R}$, we introduce a binary variable $x_i$ for each vertex $v_i \in V$ and an inequality constraint $x_i + x_j \leq 1$ for each edge $e = (v_i, v_j) \in E$.
+Furthermore we introduce an objective function in form $\sum_{x_i} x_i w(v_i)$.
 
-\begin{define}[$name()$]
-Given an attribute mapping $m = M_{x}^{y}$, $name(m)$ shall be defined as the string $x\-y$.
-\end{define}
+It is obvious that the objective function and all the constraints consitute a MIP instance, and that solving it finds the maximum weigthed IS in $G$.
+
+\section{Finding ID Sets With GLPK}
+
+By chaining these two translations we can create a MIP formulation for a given set of AMs from a document. Solving this MIP instance will give us the optimal ID set for this document.
+
+GLPK \cite{glpk} is a multi-platform, multi-purpose solver well suited for this task.
+It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} (see \cite{land60a} for details) for MIP.
+An advantage of using \textit{Branch \& Bound} is that while traversing the \textit{search tree} it finds intermittent, sub-optimal solutions.
+It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\
+
+We will now demonstrate the full process of finding the optimal ID set of an example XML file using GLPK.
+
+\subsection{Example}
+
+Consider again our XML file fragment.
 
@@
 
-TODO equations for the MIP formulation
+Recall that attribute mappings in this example are $C = \{M_{y}^{a}, M_{y}^{b}, M_{y}^{c}, M_{z}^{a}\}$.
+Corresponding vertices in the IS formulation will be $V = \{v_{y-a}, v_{y-b}, v_{y-c}, v_{z-a}\}$.
+Edges in the IS formulation will be the following.
 
+\begin{eqnarray*}
+(v_{y-a},v_{y-b}) \\
+(v_{y-a},v_{y-c}) \\
+(v_{y-b},v_{y-c}) \\
+(v_{y-a},v_{z-a}) \\
+\end{eqnarray*}
+
+First three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$.
+The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.
+
+\begin{figure}
+  \caption{IS Representation Graph}
+  \label{image-mip-is-graph}
+  \centering
+	\includegraphics[width=.25\textwidth]{images/is-representation}
+\end{figure}
+
+The next step is the MIP formulation.
+We do not need to translate from the IS formulation, as the translation from ID set formulation is straightforward, too.
 For each AM $m$ there will be one binary variable $x_{name(m)}$.
 Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings.
-For each pair of AMs $m_1, m_2$ that cannot share the same ID set (because of conflicting type or image) there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$.
+For each pair of AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$.
 $\mathbf{b}$ will be a vector of ones with corresponding length.
@@
 
-The problem is, recall, to solve the following.
+The problem now is, recall, to solve the following.
 
@@
 
-In GLPK \textit{mathprog} language, it translates to the following formulation. % TODO cite mathprog
+In GLPK \textit{MathProg} language, this translates to the following formulation. % TODO cite mathprog
 
@@
 maximize z: sum {i in AMs} x[i] * Weight[i];
-s.t. c1: x['y-b'] + x['y-c'] <= 1;
+s.t. c1: x['y-a'] + x['y-b'] <= 1;
 s.t. c2: x['y-a'] + x['y-c'] <= 1;
-s.t. c3: x['y-a'] + x['y-b'] <= 1;
+s.t. c3: x['y-b'] + x['y-c'] <= 1;
 s.t. c4: x['y-a'] + x['z-a'] <= 1;
 data;
-set AMs := y-c y-b y-a z-a;
+set AMs := y-a y-b y-c z-a;
 param Weight :=
+y-a 0.6
+y-b 0.2
 y-c 0.2
-y-b 0.2
-y-a 0.6
 z-a 0.4;
@@
 
-This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible.
+This trivial heuristics sorts all candidate AMs by their decreasing weights (Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible.
 This way it can create only one solution, and again, this is a hungry heuristic.

### 2011-10-09 doc/theses/vitasek/chapter-2-research.tex
Thesis: in progress

--- 2011-08-16 doc/theses/vitasek/chapter-2-research.tex
@@
 
-According to the article \cite{fidax}, ...
+According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set with weight more than some $K$ ($K$-\textsc{IDSet}) is in NP.
+Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete.
 
-TODO talk about FIDAX
-To the best of our knowledge, there are no other articles dealing with this problem.
+The transformation from IS problem formulation to $K$-\textsc{IDSet} problem formulation is as follows.
 
-TODO talk about Fajt - but that's different, that's keys
-TODO add citation for Fajt
+\begin{quote}
+Let $G = (V, E)$ be a simple connected graph with vertex set $V = \{v_1, \ldots, v_n\}$, and edge set $E = \{e_1, \ldots, e_m\}$.
+We define the attribute mappings as follows.
+Let $ \mathcal{I} = V \cup E$, and define $value(x) = x, x \in \mathcal{I}$.
+For each vertex $v_i \in V$, we create a mapping $m_i = \{(v_i, e_j): e_j \in E \,\text{is incident on}\, v_i \}$, and define $\tau(m_i) = v_i$; let $C = \{m_1, \ldots, m_n\}$ be set of all such mappings.
+It is clear that G has an independent set of size $K$ iff $C$ has an ID set of size $K$.
+Also, $C$ can be constructed in time polynomial on $n+m$.
+\end{quote}
 
-\section{Independent Set}
+The article continues by proving that the problem of finding maximum weighted IS can be reduced to the problem of finding an ID set with maximum weight (\textsc{Max-IDSet}).
+This again means that \textsc{Max-IDSet} is NP-complete and, furthermore, unless $\P = \NP$, \textsc{Max-IDSet} has no constant factor approximation algorithm.
 
-TODO define IS + MIS rigorously
+The difference in transformation from maximum weighted IS to \textsc{Max-IDSet} is as follows.
 
-TODO other approaches to max IS
+\begin{quote}
+[...] with the added restriction that $w(m_i) = w(v_i), v_i \in V$.
+\end{quote}
+
+Note that the transformation works in both ways: it is equivalently possible to create a maximum weighted IS instance for a given \textsc{Max-IDSet} instance.
+
+The article further suggests a heuristic approach described in Section \ref{section-mip-fidax}, which was incorporated into the framework proposed by this work.
+
+To the best of our knowledge, there are no other articles dealing directly with this problem.\\
+
+Fajt in \cite{fajt} introduces several algorithms to find XML keys (see \cite{keX}) in existing XML documents.
+Simple keys found this way can be translated to an equivalent ID attribute definition.
+This opens a new line of research: finding XML keys using an algorithm modified to look only for useful keys and then converting them to ID attributes.\\
+
+Maximum weigthed IS is a well researched topic with a lot of known direct or approximation algorithms, see e.g. \cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}.\\

### 2011-10-09 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

@@
 \section{Experimental Results}
+
+We will now present the experiments we performed along with their results and conclusions.
+Each experiment will be introduced with a table summarizing the most important parameters of the experiment, such as data used, number of iterations and CH pool size, $\alpha$ and $\beta$ parameters and the actual heuristics used.
 

### 2011-10-09 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress

@@
 
+TODO define incumbent solution
+
 \begin{figure}
@@
 
+We will now introduce several heurisitics we have implemented to aid us in search for optimal ID set.
+
 \subsection{Constructions Heuristics}
 % TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} %TODO decide on how to format CH/IH
+\nomenclature{CH}{Construction Heuristic} % TODO decide on how to format CH/IH
 
-TODO construction heuristics are heus that provide us with at least some solution.
+When we start to solve a problem using a metaheuristic approach, at first we have no solutions at all.
+The purpose of a construction heuristic (CH) is then to provide us with at least some solution.
+This may or may not be already the optimum, in the latter case it will be improved on later using improvement heuristics (IH).
+Some IHs can profit from a pool of several sub-optimal solutions, and some CHs can produce this pool from them.
 
@@
 
-It should be obvious by now that the algorithm described in \cite{fidax} (we shall call it \heu{FIDAX} from now on) can trivially be used as a construction heuristic that will give us one feasible solution.
+The first construction heuristic is the algorithm described in \cite{fidax} (we shall call it \heu{FIDAX} from now on).
+It can trivially be used to give us one feasible solution - it is, after all, a deterministic heuristic algorithm.
 
@@
 
-One of the most natural heuristics when dealing with the IS problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
+One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
 This is obviously a hungry heuristic.
-%TODO have I defined ID set condition yet??
 
@@
 
-\heu{Fuzzy} is an improvement over the \heu{Random} CH: it picks the next AM to try to add based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in TODO.
+\heu{Fuzzy} is an improvement over the \heu{Random} CH: the next AM to be added is selected based on \textit{weighted} instead of uniform random.
+The weight used here is the usual weight of an AM as defined in Section \ref{section-definitions-weight}.
 Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.
 
-Again, this is a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
+This is again a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
 
@@
 
-See listing \ref{listing-ch-incremental}.
+See Listing \ref{listing-ch-incremental} for its pseudocode.
 
@@
 This probably does not satisfy the ID set condition.
-Therefore, order them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
+Therefore, sort them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
 Again, this is a hungry heuristic returning only one solution.
 
-See listing \ref{listing-ch-removal}.
+See Listing \ref{listing-ch-removal} for its pseudocode.
 
@@
 
-\subsubsection{Truncated Branch \& Bound}
+\subsubsection{Truncated Branch \& Bound - \heu{Glpk}}
 
-This CH will be called \heu{Glpk} from now on, for it is basically a time-constrained run of GLPK.
+This construction heuristic will be called \heu{Glpk} from now on, for it is basically a time-constrained run of GLPK.
+Recall that GLPK uses a the \textit{Branch \& Bound} algorithm that produces feasible solutions even before the optimum is found.
+Limiting the run time gives us the best solution found so far, which means this is a construction heuristic.
 
-TODO if we limit GLPK's runtime, we get this
-
-TODO we shuffle AMs - we get different runs - pool is possible
+To be able to create a pool of solutions, the GLPK input in MathProg langauge is always randomly shuffled by changing the order in which variables and constraints appear.
+This is causes the solver to explore the search tree in various orders, producing different solution in each of the time-constrained runs.
 
@@
 
-TODO what they are, that they need a pool sometimes, their input and output is a pool, ...
+Improvement heuristics in general start with a solution pool, attempt to improve one or more solutions in it and then return this improved pool in the end.
+We will need two notions to describe their behavior.
 
-TODO mention intensification, diversification
+\paragraph{Intensification}
 
-TODO mention that combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} is a sort of genetic algorithm
+is the attempt to move the solution towards the nearby local optimum in the solution space.
+
+\paragraph{Diversification}
+
+is the attempt to move the solution away (escape) from from the local optimum, to be able to explore more of the solution space when the metaheuristic starts stagnating.\\
+
+A metaheuristic needs to combine intensification and diversification tendencies to explore the solution space and at the same time arrive at a local optimum.
 
@@
 
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its listing \ref{listing-ih-identity}.
+This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its Listing \ref{listing-ih-identity}.
 
@@
 the one with the lowest quality).
-This might be interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
+This is interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
 
-See listing \ref{listing-ih-removeworst}.
+See Listing \ref{listing-ih-removeworst} for details.
 
@@
 
-This is again a rather trivial IH, something which is usually referred to as a \textit{perturbation} function. % TODO link the notion
-By removing a random subset of specified size from each solution in the pool, it provides variability needed to escape from local optima. % TODO make sure we have talked about local optima before
+This is again a rather trivial diversification improvement heuristic.
+By removing a random subset of specified size from each solution in the pool, it provides the variability needed to escape from local optima.
+% TODO make sure we have talked about local optima before
 
-The number of AMs to remove from each solution is specified as ratio (from the interval $(0, 1)$) of the solution size. % TODO have we defined what a solution size is?
+The number of AMs to remove from each solution is specified as ratio from $(0, 1)$ of the solution size (number of AMs in solution).
 For example, \heu{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
@@
 
-See listing \ref{listing-ih-randomremove}.
+See Listing \ref{listing-ih-randomremove} for pseudocode.
 
@@
 The resulting solution (whether any AMs were added or not) is then returned to the pool.
-Listing \ref{listing-ih-hungry} captures this process.
+This is then intensification, and Listing \ref{listing-ih-hungry} captures the process.
 
@@
 
-TODO explain how this works and link wise books
+\heu{Mutation} is based on the following idea.
+We assume that an incumbent solution may already contain some AMs belonging to the optimal solution.
+We will take a random guess and fix some of these AMs: add new constraints to the MIP formulation fixing values of the respective variables to 1.
 
-TODO explain how this translates to GLPK input
+This new formulation contains less free variables and should be easier to solve, probably even to optimum.
+We run GLPK again using this constrained formulation, enforcing again a time limit.
+Solution found this way is a feasible solution of the original problem, however the optimum is not necessarily the same as in unconstrained formulation.
+It is an intensification approach: we limit the search to the neighborhood of an already found solution.\\
 
-For every AM $AM_F$ fixed to appear in the solution a following constraint is added to GLPK input:
+\heu{Mutation} changes the MIP formulation in following way. For every AM $AM_F$ fixed to appear in the solution a following constraint is added to GLPK input:
 \[s.t. f_{index}: x['name(AM_F)'] = 1;\]
@@
 \[s.t. c_{index}: x['name(AM_F)'] + x['name(AM_i)'] <= 1;\]
-will not be included.
+will not be included.\\
 
-See listing \ref{listing-ih-mutation}.
+Listing \ref{listing-ih-mutation} captures the process of randomly selecting a specified fraction of AMs of the incumbent solution to fix, then running GLPK again.
+\heu{Mutation} requires pool of at least one solution as input, and adds the improved solution to the result pool.
 
@@
 \REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of AMs to fix
+\REQUIRE $k$ fraction of AMs to fix
 \ENSURE pool of feasible solutions
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
+\STATE $incumbent \gets $ incumbent solution in $FP$
 \STATE $K \gets k * |incumbent|$
@@
 
-See listing \ref{listing-ih-localbranching}.
+See Listing \ref{listing-ih-localbranching} for pseudocode.\\
 
@@
 
+TODO mention that combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} is a sort of genetic algorithm
+
 \section{IDREF}
@@
 
-\[\iota(ID) = \bigcup_{am \in ID} \iota(am)\]
+\[\iota(ID) = \bigcup_{m \in ID} \iota(m)\]
 
@@
 \STATE $candidates \gets AMs \backslash ID$
-\STATE $\iota(ID) \gets \bigcup_{am \in ID} \iota(am)$
+\STATE $\iota(ID) \gets \bigcup_{m \in ID} \iota(m)$
 \FORALL{$c \in candidates$}

### 2011-10-09 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress

@@
 
-This section will briefly describe the process by which an input data set is processed to obtain the AM model as described in TODO link.\\
+This section will briefly describe the process by which an input data set is processed to obtain the AM model as described in Section \ref{section-definitions-ams}.\\
 

### 2011-10-09 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress

@@
 
+TODO define incumbent solution, neighborhood, solution space
+
 TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
@@
 TODO we will be using a pool - what is a pool
-
-TODO define incumbent solution
 
@@
 
-TODO explain how this works and link wise books
+This improvement heuristic expands on the idea of \heu{Mutation}.
+But, instead of randomly selecting AMs in the incumbent solution, it looks for commonalities among the solutions in the pool.
+This is based on the hope that if more solutions agree on the same AMs, those are probably included in the optimal solution too.
 
-TODO explain how this translates to GLPK input - it's again simple fixing to 1, but we get the list of AMs in a different manner.
+\heu{Crossover} takes a parameter - fraction of solutions in the pool among which to look for commonalities.
+AMs found in every one of them are fixed in the modified MIP formulation the same way as in \heu{Mutation}.
+This again amounts to an intensification tendency.
 
-See listing \ref{listing-ih-crossover}.
+Listing \ref{listing-ih-crossover} captures the process.
+\heu{Crossover} requires at least one solution in the pool, but to work properly, more are needed.
+Solutions are picked at random from the pool, common AMs found and fixed.
+GLPK is run again (with a time constraint) and the improved solution is added to the result pool.
 
@@
 \REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of solutions to look for commonalities in
+\REQUIRE $k$ fraction of solutions among which to look for commonalities
 \ENSURE pool of feasible solutions
@@
 
-TODO explain how this works and link wise books
+\heu{Local Branching} is another intensification heuristic. This time the neighborhood being searched is defined by edit distance.
 
-TODO explain how this translates to GLPK input
+The incumbent solution is represented by a vector $\mathbf{x}_{INCUMBENT}$ of ones and zeroes.
+Based on it, a new constraint will be added to the MIP formulation.
+For every other solution $\mathbf{x}_i$ the edit distance, i.e.
+number of positions in which $\mathbf{x}_{INCUMBENT}$ and $\mathbf{x}_i$ differ, will have to be lower than some threshold $K$.
+This will be represented in MathProg as follows.
 
-A new constrain describing the maximal allowed distance from the incumbent solution is added to GLPK input.
+\[s.t. LB: sum\{i\ in\ INCUMBENT\} (1 - x[i]) + sum\{i\ in\ REMAINING\} x[i] \leq K;\]
 
-\[s.t. LB: sum\{i\ in\ INCUMBENT\} (1 - x[i]) + sum\{i\ in\ REMAINING\} x[i] \leq k;\]
+Where $INCUMBENT$ is the set of names of AMs in the incumbent solution, $REMAINING$ is the set of all AMs not included in the incumbent solution and $K$ is the maximum edit distance allowed.
+$K$ is determined as a fraction of the count of all AMs, provided as parameter $k$.
 
-Where $INCUMBENT$ is a set of names of AMs in the incumbent solution, $REMAINING$ is a set of all the AMs not included in the incumbent solution and $k$ is the requested maximal distance.
-
-See Listing \ref{listing-ih-localbranching} for pseudocode.\\
+See Listing \ref{listing-ih-localbranching} for pseudocode.
+The heuristic requires a pool containing at least one solution, solves the modified MIP formulation using GLPK limited to some time again and adds the improved solution to the result pool.\\
 
@@
 \REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k$ ratio of total AM count to bound the Hamming distance to
+\REQUIRE $k$ fraction of the total AM count to determine max edit distance
 \ENSURE pool of feasible solutions
 \STATE $K \gets k * |$total AM count$|$
-\STATE $incumbent \gets $ best solution in $FP$ \COMMENT {best = highest weight}
-\STATE add max Hamming distance requirement to GLPK problem formulation
+\STATE $incumbent \gets $ incumbent solution in $FP$
+\STATE add max edit distance requirement to GLPK problem formulation
 \STATE $improved \gets $ run GLPK
@@
 
-TODO mention that combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} is a sort of genetic algorithm
+\paragraph{Genetic algorithms}
+
+It is worth noting that by combining \heu{Mutation}, \heu{Crossover} and \heu{RemoveWorst} we get a very simple genetic algorithm.
 

### 2011-10-10 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 \chapter{Introduction}
-Extensible Markup Language (XML) has become a wide-spread standard for data representation and exchange.
-Its popularity is based on its simplicity and at the same time its flexibility and express power.
+Extensible Markup Language (XML) \cite{Bray:08:EML} has become a popular standard for data representation and exchange.
+Its prevalence is based on its simplicity and at the same time its flexibility and express power.
 
@@
 In order to this need, several so-called XML schema languages (XML schemata) have been created.
-The most used of them are Document Type Definition (DTD) and XML Schema proposed by W3C.
+The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by The World Wide Web Consorcium (W3C).
 
-Nevertheless, it is encouraged to use an XML schema along with an XML data representation, in practise it is done sparsely.
-Commonly, XML documents are not assigned with their respective XML schema at all or the schema is out-dated due to modifications in a structure of data without updating the schema.
+Nevertheless an encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
+Commonly, XML documents are not assigned with their respective XML schema at all or the schema is out-dated due to modifications done to a structure of data without updating the schema.
 
 Recently in reaction to this situation, significant number of approaches dealing with an automatic construction of an XML schema have been proposed.
-The aim is that a set of XML documents is available and it is desired to infer a (non-trivial) XML schema, so that the XML documents are valid against it.
-Most of published approaches are of this type - input is a set of XML documents - and they are based on various ideas and can be classified by several aspects, as discussed in chapter \todo{ref}.
+An aim is to exploit a provided set of XML documents and infer a XML schema, so that the XML documents are valid against it.
+In addition, the inferred schema should be reasonable in views of human-readability, preciseness and conciseness.
+Most of published approaches of an XML schema inference are of this type - input is a set of XML documents.
+They are based on various ideas and can be classified by several aspects as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}.
 
 Besides the mentioned type, some approaches that utilize other or additional sources has been developed.
-If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery queries, etc, these sources can be exploited to refine the inference process.
+If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery \todo{ref} queries, or other, these sources can be exploited to refine the process of inference.
 The refinement can be achieved in various aspects such as decreasing the speed of the process, getting a more precise, more concise or more readable result or inference of some statements about the data which cannot be (easily) extracted from the data themselves.
 
-Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (also discussed in chapter \todo{ref}), leaving a wide space for a possible future research.
+Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (also discussed in Chapter \ref{chapter_analysis_of_recent_approaches}), leaving a wide space for a possible future research and improvements.
 
-\todo[inline]{aim of this work}
-\todo[inline]{structure of this work}
+\section{Aim of this Work}
+\todo[inline]{Toto je skopirovane zo sis-u zo zadania prace, moze to tu byt?}
+The aim of this work is a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts etc.).
+
+Firstly, it is necessary to analyze existing inference solutions in general and to discuss their advantages and disadvantages.
+The core of the work is identification and discussion of information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.
+
+The result of the work will be a proposal of own approach involving the improvements, its implementation and suitable experiments that will show its advantages.
+
+\section{Structure of this Work}
+\todo[inline]{TODO}

### 2011-10-10 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
-\chapter{Analysis of Recent Approaches}
-Approaches of XML schema inference can be divided by several criteria. A basic division is by the language the resulting schema is written in. Commonly used languages are DTD and XML Schema.
+\chapter{Analysis of Recent Approaches} \label{chapter_analysis_of_recent_approaches}
+Existing approaches of XML schema inference can be classified by several criteria.
+A basic division is by the language the resulting schema is written in.
+Commonly used languages are DTD and XML Schema.
 
 According to \cite{Mlynkova:2008:AAX:1494650.1495496}, the type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
-Heuristic methods are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
-On the contrary, the grammar-inferring methods are based on theoretical knowledge of automata and their results belong to a particular class of languages.
+Heuristic methods (\cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522}) are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
+Two of these approaches are described in Section \ref{section_xtract} and Section \ref{section_even_an_ant_can_create_an_xsd}.
+
+On the contrary, the grammar-inferring methods (\cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475}) are based on theoretical knowledge of automata and their results belong to a particular class of languages.
 Thus, these methods guarantee specific characteristics of the results.
-\todo[inline]{Tady by neškodilo dát odkazy na všechny ex.
-přístupy a říct, že popíšeme detailněji aspoň dva.}
 
@@
 According to our best knowledge at the time of writing, there is just one approach of this category, described in \cite{Necasky:2009:DXK:1529282.1529414}.
-It utilizes a set of XQuery queries to discover keys and foreign keys\todo{ref do literatury - Ty by tady něškodily všude, u každé té kategorie.}.
+It utilizes a set of XQuery queries to discover keys and foreign keys.
 
@@
 
-\section{XTRACT}
+\section{XTRACT} \label{section_xtract}
 The XTRACT \cite{Garofalakis:2000:XSE:342009.335409} system is an example of a \emph{heuristic} \emph{merging state algorithm} creating the result in DTD.
@@
 
-For instance, the paper (\cite{Garofalakis:2000:XSE:342009.335409}) introduces the following example: Sequences \texttt{abab} and \texttt{bbbe} are generalized to \texttt{(ab)*}, \texttt{(a|b)*}, \texttt{b*e}.
+For instance, paper \cite{Garofalakis:2000:XSE:342009.335409} introduces the following example: Sequences \texttt{abab} and \texttt{bbbe} are generalized to \texttt{(ab)*}, \texttt{(a|b)*}, \texttt{b*e}.
 \todo[inline]{Mozem takto pouzivat prevzate priklady?}
@@
 Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decresing their summed size without modifications in their semantics.
-An aim of this step is to decrease a MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
+An aim of this step is to decrease an MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
 
-An example introduced it the paper (\cite{Garofalakis:2000:XSE:342009.335409}): DTD candidates \texttt{ac}, \texttt{ad}, \texttt{bc} and \texttt{bd} are factored into \texttt{(a|b)(c|d)}.
+An example introduced it paper \cite{Garofalakis:2000:XSE:342009.335409}: DTD candidates \texttt{ac}, \texttt{ad}, \texttt{bc} and \texttt{bd} are factored into \texttt{(a|b)(c|d)}.
 
@@
 
-\section{Even an Ant Can Create an XSD}
+\section{Even an Ant Can Create an XSD} \label{section_even_an_ant_can_create_an_xsd}
 This work, described in \cite{Vosta:2008:EAC:1802514.1802522}, combines several previously proposed approaches including the XTRACT system discussed in the previous section.
@@
 This phase clusters elements on a basis of their context and structure.
-It is done by creating tree structures for each input XML documents, where vertices represent elements and attributes and an edge between two vertices expresses the their parent-child relation.
-These trees and their subtrees and then compared, using an imposed tree similarity measure, to find elements with the same semantics.
+It is done by creating tree structures for each input XML documents where vertices represent elements and attributes and an edge between two vertices expresses their parent-child relation.
+These trees are their subtrees and then compared using an imposed tree similarity measure to find elements with the same semantics.
 
@@
 An ant
-is performing steps (schema modifications), dying after a predifined number of steps and providing an information - positive feedback - on quality of a found solution.
-The search is performed in a defined count of interations (or it stops if good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
+is performing steps (schema modifications), dying after a predefined number of steps and providing an information - positive feedback - on quality of a solution found.
+The search is performed in a defined count of interations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
 Every step of an ant represents a modification of a schema, in particular, a merge of states in a corresponding PTA.

### 2011-10-11 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 
-On the contrary, the grammar-inferring methods (\cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475}) are based on theoretical knowledge of automata and their results belong to a particular class of languages.
+On the contrary, the grammar-inferring methods (\cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236}) are based on theoretical knowledge of automata and their results belong to a particular class of languages.
 Thus, these methods guarantee specific characteristics of the results.
@@
 It is done by creating tree structures for each input XML documents where vertices represent elements and attributes and an edge between two vertices expresses their parent-child relation.
-These trees are their subtrees and then compared using an imposed tree similarity measure to find elements with the same semantics.
+These trees and their subtrees are then compared using an imposed tree similarity measure to find elements with the same semantics.
 
@@
 
-One of the improvements of this heuristic is inclusions of a negative feedback after each step of an ant, visible only in a current iteration.
+One of the improvements of this heuristic is an inclusion of a negative feedback after each step of an ant, visible only in a current iteration.
 Due to this improvement, a larger subspace of solutions is searched.
@@
 Another improvement lies in a way how an ant decides for a particular step to perform.
-To achieve better results, the authors propose a combinations of several verified approaches.
-A set of all possible steps is created using \emph{k,h-context} and \emph{s,k-string} \todo{citacie} methods.
-The optimal step is then selected employing the MDL principle \todo{citacia}.
+To achieve better results, the authors propose a combination of several verified approaches.
+A set of all possible steps is created using \emph{k,h-context} \cite{Ahonen1996GeneratingGrammars} and \emph{s,k-string} \cite{Raman97thesk-strings, Wong03onstructural} methods.
+The optimal step is then selected employing the MDL principle \cite{Grünwald05atutorial, Garofalakis:2000:XSE:342009.335409}.
 
@@
 \section{On Inference of XML Schema with the Knowledge of an Obsolete One}
-\todo[inline]{spravne zaradit}
 An aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently.
@@
 
-In the second step, assuming a correct schema it is wanted to specialize the REs involved in the schema with regard to the XML documents, resulting in more precise and readable schema.
+In the second step, assuming a correct schema it is wanted to specialize the regular expressions involved in the schema with regard to the XML documents, resulting in more precise and readable schema.
 Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.
@@
 The query is searched for so-called \emph{join patters}.
-These are \emph{for join pattern} and \emph{let join pattern} and they are as follows (the first is \emph{for join pattern}, \emph{let join pattern}):
+These are \emph{for join pattern} and \emph{let join pattern} and they are as follows (the first is \emph{for join pattern}, the latter is \emph{let join pattern}):
 
@@
 
-Else, the pattern occurrence is considered of case (O2) and the assigned weight depends on the number of target return paths.
-If the number is greater than one, the weight is one (weight: 1), else (the number is less or equal one) the weight is one half (weight: 0.5).
+Otherwise, the pattern occurrence is considered of case (O2) and the assigned weight depends on the number of target return paths.
+If the number is greater than one, the weight is one (weight: 1), else (the number equals zero or one) the weight is one half (weight: 0.5).
 
@@
 Nějaká návaznost k další kapitole.}
+
+\section{Summary}
+\begin{table}
+  \footnotesize \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    Method name and/or paper & Input & Output & Year & Standalone \\ \hline \hline
+    \cite{Ahonen1996GeneratingGrammars} & XML documents & DTD & 1996 & Yes \\ \hline
+    XTRACT \cite{Garofalakis:2000:XSE:342009.335409} & XML documents & DTD & 2000 & Yes \\ \hline
+    DTD-miner \cite{Moh:2000:RSW:336597.336638} & XML documents & DTD & 2000 & Yes \\ \hline
+    \cite{Fernau:2001:LXG:645744.667236} & XML documents & DTD & 2001 & Yes \\ \hline
+    ECFG \cite{Chidlovskii:2002:SEX:544220.544288} & XML documents & XSD & 2002 & Yes \\ \hline
+    sk-ANT \cite{Wong03onstructural} & XML documents & DTD & 2003 & Yes \\ \hline
+    \cite {Min:2003:EES:639473.639475} & XML documents & DTD & 2003 & Yes \\ \hline
+    \cite{Bex:2006:ICD:1182635.1164139} & XML documents & DTD & 2006 & Yes \\ \hline
+    XStruct \cite {Hegewald:2006:XES:1129755.1130124} & XML documents & XSD & 2006 & Yes \\ \hline
+    \cite{Bex:2007:IXS:1325851.1325964} & XML documents & XSD & 2007 & Yes \\ \hline
+    SchemaMiner \cite{Vosta:2008:EAC:1802514.1802522} & XML documents & XSD & 2008 & Yes \\ \hline
+    \cite{Mlynkova:2009:IXS:1862681.1862693} & XML documents, XML schema & DTD, XSD* & 2009 & Yes \\ \hline
+    \cite{Necasky:2009:DXK:1529282.1529414} & XQuery queries & XSD* & 2009 & No \\ \hline    
+  \end{tabular}
+  \caption{Summary of Recent Approaches}
+  \label{table_summary_of_recent_approaches}
+\end{table}

### 2011-10-14 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 Assume a query $Q$ that joins a sequence of elements $S_1$ targeted by a path $P_1$ with a sequence of elements $S_2$ targeted by a path $P_2$ on a condition $L_1 = L_2$.
-\todo[inline]{priklad}
+For instance, see Listing \ref{listing_for_join_pattern} and Listing \ref{listing_let_join_pattern}.
+
 The method is based on assumption that each join is done via key/foreign key pair.
@@
 \subsection{Join Patterns and Key Inference}
+\begin{lstlisting}[mathescape, float, caption=For join pattern., frame=single, label=listing_for_join_pattern]
+for $\$e_1$ in $P_1$
+return
+  for $\$e_2$ in
+    $P_2[L_2 = \$e_1/L_1]$
+  return $C_R$
+\end{lstlisting}
+
+\begin{lstlisting}[mathescape, float, caption=Let join pattern., frame=single, label=listing_let_join_pattern]
+for $\$e_1$ in $P_1$
+return
+  let $\$e_2$ :=
+    $P_2[L_2 = \$e_1/L_1]$
+  return $C_R$
+\end{lstlisting}
+
 For a certain join, the decision for one of the cases (O1) and (O2) is made by the form of the join.
 The query is searched for so-called \emph{join patters}.
-These are \emph{for join pattern} and \emph{let join pattern} and they are as follows (the first is \emph{for join pattern}, the latter is \emph{let join pattern}):
-
-\todo[inline]{spravne kreslit indexy pri pismenkach}
-\begin{verbatim}
-for $e1 in P1              for $e1 in P1
-return                     return
-    for $e2 in                 let $e2 :=
-        P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
-    return CR                  return CR
-\end{verbatim}
+These are \emph{for join pattern} and \emph{let join pattern} and they are proclaimed in Listing \ref{listing_for_join_pattern} and Listing \ref{listing_let_join_pattern}.
 
@@
 
-\todo[inline]{Tady by se hodila nějaká srovnávací tabulka metod z různách pohledů (metoda, vstupy, výstupy).
-Dále pak kapitola na téma co dnešní metody neumí a čím se tedy v práci budeme zbývat.
-Tj.
-Nějaká návaznost k další kapitole.}
-
 \section{Summary}
 \begin{table}
-  \footnotesize \begin{tabular}{|c|c|c|c|c|}
+  \begin{threeparttable}
+  \footnotesize \begin{tabular}{|c|c|c|c|}
     \hline
-    Method name and/or paper & Input & Output & Year & Standalone \\ \hline \hline
-    \cite{Ahonen1996GeneratingGrammars} & XML documents & DTD & 1996 & Yes \\ \hline
-    XTRACT \cite{Garofalakis:2000:XSE:342009.335409} & XML documents & DTD & 2000 & Yes \\ \hline
-    DTD-miner \cite{Moh:2000:RSW:336597.336638} & XML documents & DTD & 2000 & Yes \\ \hline
-    \cite{Fernau:2001:LXG:645744.667236} & XML documents & DTD & 2001 & Yes \\ \hline
-    ECFG \cite{Chidlovskii:2002:SEX:544220.544288} & XML documents & XSD & 2002 & Yes \\ \hline
-    sk-ANT \cite{Wong03onstructural} & XML documents & DTD & 2003 & Yes \\ \hline
-    \cite {Min:2003:EES:639473.639475} & XML documents & DTD & 2003 & Yes \\ \hline
-    \cite{Bex:2006:ICD:1182635.1164139} & XML documents & DTD & 2006 & Yes \\ \hline
-    XStruct \cite {Hegewald:2006:XES:1129755.1130124} & XML documents & XSD & 2006 & Yes \\ \hline
-    \cite{Bex:2007:IXS:1325851.1325964} & XML documents & XSD & 2007 & Yes \\ \hline
-    SchemaMiner \cite{Vosta:2008:EAC:1802514.1802522} & XML documents & XSD & 2008 & Yes \\ \hline
-    \cite{Mlynkova:2009:IXS:1862681.1862693} & XML documents, XML schema & DTD, XSD* & 2009 & Yes \\ \hline
-    \cite{Necasky:2009:DXK:1529282.1529414} & XQuery queries & XSD* & 2009 & No \\ \hline    
+    Method name and/or paper & Input & Output & Year \\ \hline \hline
+    \cite{Ahonen1996GeneratingGrammars} & XML documents & DTD & 1996 \\ \hline
+    XTRACT \cite{Garofalakis:2000:XSE:342009.335409} & XML documents & DTD & 2000 \\ \hline
+    DTD-miner \cite{Moh:2000:RSW:336597.336638} & XML documents & DTD & 2000 \\ \hline
+    \cite{Fernau:2001:LXG:645744.667236} & XML documents & DTD & 2001 \\ \hline
+    ECFG \cite{Chidlovskii:2002:SEX:544220.544288} & XML documents & XSD & 2002 \\ \hline
+    sk-ANT \cite{Wong03onstructural} & XML documents & DTD & 2003 \\ \hline
+    \cite {Min:2003:EES:639473.639475} & XML documents & DTD & 2003 \\ \hline
+    \cite{Bex:2006:ICD:1182635.1164139} & XML documents & DTD & 2006 \\ \hline
+    XStruct \cite {Hegewald:2006:XES:1129755.1130124} & XML documents & XSD & 2006 \\ \hline
+    \cite{Bex:2007:IXS:1325851.1325964} & XML documents & XSD & 2007 \\ \hline
+    SchemaMiner \cite{Vosta:2008:EAC:1802514.1802522} & XML documents & XSD & 2008 \\ \hline
+    \cite{Mlynkova:2009:IXS:1862681.1862693} & XML documents, XML schema & DTD, XSD & 2009 \\ \hline
+    \cite{Necasky:2009:DXK:1529282.1529414} & XQuery queries & --\tnote{1} & 2009 \\ \hline    
   \end{tabular}
+  \todo[inline]{doplnit}
+  \begin{tablenotes}
+	\item [1] Result of this method is a list of discovered keys and foreign keys.
+  \end{tablenotes}
   \caption{Summary of Recent Approaches}
   \label{table_summary_of_recent_approaches}
+  \end{threeparttable}
 \end{table}
+
+As shown in Table \ref{table_summary_of_recent_approaches}, most of the recent approaches of XML schema inference are base on utilization of XML documents.
+These incorporates various verified methods and the newer approaches often improve the older ones and/or combine them together to achieve better results.
+
+Lately, several approaches that utilize other input sources have been proposed.
+Paper \cite{Mlynkova:2009:IXS:1862681.1862693} introduces a method that utilizes an XML schema besides XML documents.
+The method described in \cite{Necasky:2009:DXK:1529282.1529414} utilizes XQuery queries, however, the result of this method is a list of discovered keys and foreign keys, not an XML schema.
+Also, to our best knowledge, implementation that combines this method with another approach to get the XML schema has not been proposed yet.
+\todo{Toto je pravda, naozaj ta implementacia neexistuje?}
+
+There are plenty of additional sources that can be exploited in the process of inference such as other XML schema languages, queries, XSLT scripts, negative examples (XML documents that should not conform to the resulting schema).
+In this work we focus on utilizing XQuery queries.

### 2011-10-14 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 Nevertheless an encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
-Commonly, XML documents are not assigned with their respective XML schema at all or the schema is out-dated due to modifications done to a structure of data without updating the schema.
+Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema.
 
 Recently in reaction to this situation, significant number of approaches dealing with an automatic construction of an XML schema have been proposed.
-An aim is to exploit a provided set of XML documents and infer a XML schema, so that the XML documents are valid against it.
+An aim is to exploit a provided set of XML documents and infer an XML schema, so that the XML documents are valid against it.
 In addition, the inferred schema should be reasonable in views of human-readability, preciseness and conciseness.

### 2011-10-27 doc/theses/vitasek/diplomka.tex
Thesis: in progress
Thesis: in progress.

@@
 \usepackage{amsmath}
+\usepackage{complexity}
 
@@
 % - make sure labels have consistent names
+% - ratio -> FRACTION of XYZ to fix/remove/...
+% - what style of quotation marks to use?
 
@@
 \noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
+I would like to thank everyone participating in the jInfer project for creating a~solid foundation for this work, namely Irena Mlýn\-ko\-vá, Mi\-chal Klem\-pa, Má\-rio Mi\-ku\-la, Ro\-bert Sme\-ta\-na and Mi\-chal Švi\-rec.
+The thanks goes again to Ms.
+Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others.
 
@@
 
-Vedoucí diplomové práce: \mfadvisor{}, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství
+Vedoucí diplomové práce: \mfadvisor{} % TODO does this have to be here?, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství
 
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
+Abstrakt: Tato práce navazuje na dřívější pokusy odvodit (inferovat) schému existujících XML dokumentů.
+Jelikož je odvozování \textit{struktury} již relativně dobře popsáno, soustředíme se na integritní omezení.
+Několik jich popisujeme, pozornost pak soustředíme na ID/\.IDREF/\.IDREFS atributy z DTD.
+Na bázi článku od Barbosa a Menelzon stavíme heuristický přístup k problému hledání optimální sady ID atributů, jeho funkčnost a vhodnost pak ověřujeme na škále experimentů.
 
@@
 
-Supervisor: \mfadvisor{}, Department of Software Engineering
+Supervisor: \mfadvisor{} % TODO does this have to be here?, Department of Software Engineering
 
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
+Abstract: In this work we expand upon the previous efforts to extract (infer) schema information from existing XML documents.
+We find the inference of \textit{structure} to be sufficiently researched and focus further on integrity constraints.
+After briefly introducing some of them we turn our attention to ID/\.IDREF/\.IDREFS attributes in DTD.
+Building on the research by Barbosa and Menelzon we introduce a heuristic approach to the problem of finding optimal ID set.
+The approach is evaluated and tuned in a wide range of experiments.
 
@@
 \newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
 
@@
 
-There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.
+There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.\\
+
+TODO we will be ignoring O() complexities
 

### 2011-10-28 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.

@@
 
-TODO what is heuristics (link wise books), what is metaheuristics (we will be using them)
+The definition of \textit{heuristic} or \textit{heuristic algorithm} varies from one source to another. We shall be using it roughly in the following sense.
 
-TODO define incumbent solution, neighborhood, solution space
+\begin{define}[Heuristic]
+	Is an approach to problem solving based on prior experience, educated guess or common knowledge.
+\end{define}
 
-TODO mention things like Taboo search and Genetic Algorithms (we can emulate them with Crossover/Mutation) (we won't be using them)
+\begin{define}[Heuristic Algorithm]
+	Is an algorithm that, in a reasonably short time, will generate a good, maybe even optimal solution to an optimization problem. However, it won’t provide any formal guarantee about its quality.
+\end{define}
 
-TODO we will be working with heuristics striving to do the following: input is a list of AMs, they have their weight, we try to find a non-conflicting subset which maximizes the weight
+This definion of heuristic algorithm coming from \cite{heu-lecture} is rather vague, however, it will be sufficient for us.
 
-TODO we will be using a pool - what is a pool
+An example of a heuristic is the commonly used approach of \textit{trial and error}: after a failed attempt, change something and try again. We will see many more heuristics later in this chapter.
+
+While a heuristic algorithm can be seen as a tool designed to solve one specific problem, the notion of a \textit{metaheuristic} reminds of a recipy to solve a whole family of problems.
+We shall be using the metaheuristic presented in Figure \ref{image-metaheuristic} to find optimal ID sets in this work.
+But before we start describing its structure and components, we need to introduce some more notions.
 
@@
 
-We will now introduce several heurisitics we have implemented to aid us in search for optimal ID set.
+\begin{define}[Solution Space, Solution Quality]
+	Solution space in general is the set of all permissible solutions (not violating any constraints).
+In the specific case of a MIP formulation it is the set of all $\mathbf{x}$ subject to $A\mathbf{x} \leqslant \mathbf{b}$.
+Every solution in the solution space has its \textit{quality}, in case of MIP for solution $\mathbf{x}$ it is the value of the objective function in $\mathbf{x}$.
+\end{define}
+
+\begin{define}[Solution Neighborhood]
+	Neighborhood of a solution $\mathbf{x}$ in the solution space are all the other solutions close to $\mathbf{x}$ according to some metric.
+\end{define}
+
+The precise definition of the neighborhood is always adjusted according to specific needs.
+However, the neighborhood should be defined so that is \textit{continuous} (with respect to quality) to be useful.
+The exact reasons for this requirement are sketched in Section \ref{section-mip-ihs}.
+
+\begin{define}[Solution Pool, Incumbent Solution]
+	\textit{Solution pool} (sometimes called pool of feasible solutions or feasible pool) is simply a set of solutions of different qualities that were found in one of the stages of the metaheuristic.
+The solution(s) with the highest quality is (are) called the \textit{incumbent} solution(s).
+\end{define}
+
+A quick reminder of what we are trying to solve using our metaheuristic: given a list of AMs with weights, find a non-conflicting subset maximizing the sum of weights in the subset.
+We will now describe its structure, please refer again to Figure \ref{image-metaheuristic}.
+
+First we take the list of candidate AMs and ask a \textit{construction heuristic} (see Section \ref{section-mip-chs}) to provide us with a pool of solutions.
+Then, in a loop, we use this pool as input for \textit{improvement heuristics} (see Section \ref{section-mip-ihs}) and in turns ask them to improve it.
+All the time we check whether some \textit{termination criteria} are met, and if it is so, we terminate the metaheuristic.
+The incumbent solution from the last pool is then declared the Output ID set.\\
+
+The notion of metaheuristic covers a wide range of topics in the field of heuristics, such as Tabu Search (see \cite{Glover:1997:TS:549765}), Ant Colony Optimization (see \cite{dorigo2004ant}) or Genetic Algorithms (see e.g.
+\cite{goldberg1989genetic}), to name a few.\\
+
+We will now introduce the heurisitics we have implemented to use in our metaheuristic for finding optimal ID sets.
 
 \subsection{Constructions Heuristics}
+\label{section-mip-chs}
 % TODO can I format things in nomenclature?
-\nomenclature{CH}{Construction Heuristic} % TODO decide on how to format CH/IH
+\nomenclature{CH}{Construction Heuristic}
 
@@
 A metaheuristic needs to combine intensification and diversification tendencies to explore the solution space and at the same time arrive at a local optimum.
+Recall the requirement for the solution space to be continuous in terms of quality: this guarantees that as we approach a solution $\mathbf{x}$, the quality of solutions we encounter approaches the quality of $\mathbf{x}$.
 

### 2011-10-28 doc/theses/vitasek/appendix-b-iss.tex
Thesis: in progress.

@@
 \label{appendix-iss}
+
+\nomenclature{ISS}{ID Set Search}
 
@@
 \begin{itemize}
-	\item \texttt{objects}, containing the object representation of attribute mappings and AM model.
-	\item \texttt{heuristics.construction}, containing all the CHs hidden behind the \texttt{Con\-struc\-tion\-Heu\-ris\-tic} interface, with sub-packages \texttt{fidax} containing the whole implementation of FIDAX heuristic % TODO link
-	and \texttt{glpk} containing the whole interface to an external GLPK solver. % TODO link
+	\item \texttt{objects}, containing classes for attribute mappings and the AM model.
+	\item \texttt{heuristics.construction}, containing all the CHs hidden behind the \texttt{Con\-struc\-tion\-Heu\-ris\-tic} interface, with sub-packages \texttt{fidax} containing the whole implementation of FIDAX heuristic (Section \ref{section-mip-fidax}).
+	and \texttt{glpk} containing the whole interface the external GLPK solver (Section \ref{section-mip-glpk}).
 	\item \texttt{heuristics.improvement}, containing all the IH hidden behind the \texttt{Im\-prove\-ment\-Heu\-ris\-tic} interface.
-	\item \texttt{experiments}, containing everything related to experimenting with these heuristics.
+	\item \texttt{experiments}, containing everything related to experimenting.
 \end{itemize}
@@
 The quality of a single ID set is measured by an instance of \texttt{Quality\-Measurement}.
-After the experiment finishes, it invokes the \texttt{notify\-Finished()} method.\\
+After the experiment finishes, it invokes the \texttt{notify\-Finished()} method.
 
@@
 \section{How to Create a New Heuristic}
+\label{section-appendix-iss-howto}
 
@@
 
-It is possible to optionaly override any of the following methods: \texttt{notify\-Start()}, \texttt{notify\-Finished()} and \texttt{notify\-Finished\-All()}.
+It is possible to override any of the following methods: \texttt{notify\-Start()}, \texttt{notify\-Finished()} and \texttt{notify\-Finished\-All()}.
 They will be invoked before running the first experiment, after each experiment run and after all experiments finished, respectively.
-Note that \texttt{notify\-Finished()} already contains logic to output some information regarding the currently finished experiment to a file, but it can be safely overriden without a need to call \texttt{super.\.notify\-Finished()}.
+Note that \texttt{notify\-Finished()} already can output some information regarding the currently finished experiment to a file, but it can be safely overriden without a need to call \texttt{super.\.notify\-Finished()}.

### 2011-10-28 doc/theses/vitasek/chapter-0-preface.tex
Thesis: in progress.

--- 2011-09-29 doc/theses/vitasek/diplomka.tex
@@
-\documentclass[a4paper,12pt,oneside]{report}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{url}
-\usepackage{subfigure}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{tgpagella}
-\usepackage{xstring}
-\usepackage{wrapfig}
-\usepackage{appendix}
-\usepackage{nomencl}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{amsmath}
-
-\newcommand{\myscale}{0.74}
-\newcommand{\vect}[1]{\boldsymbol{#1}}
-\newcommand{\code}[1]{\texttt{\StrSubstitute{#1}{.}{.\.}}}
-\def\.{\discretionary{}{}{}}
-\newcommand{\heu}[1]{\texttt{\textit{#1}}}
-\newcommand{\dataset}[1]{\texttt{\textit{#1}}}
-\newcommand{\jmodule}[1]{\texttt{\textit{#1}}}
-\linespread{1.5}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\renewcommand{\algorithmiccomment}[1]{\textit{// #1}}
-
-\renewcommand{\nomname}{List of Abbreviations}
-\makenomenclature
-
-\theoremstyle{definition}
-\newtheorem{define}{Definition}[chapter]
-
-% TODO uncomment this to get much shorter ToC
-% \setcounter{tocdepth}{1}
-
-\setlength{\hoffset}{-1in} %left margin will be 0, as hoffset is by default 1inch
-\setlength{\voffset}{-1in} %analogous voffset
-\setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
-\setlength{\topmargin}{25mm}
-\setlength{\footskip}{1cm}
-\setlength{\headheight}{0cm}
-\setlength{\headsep}{0cm}
-\setlength{\marginparwidth}{0cm}
-\setlength{\marginparpush}{0cm}
-\setlength{\textheight}{23.7cm}
-\setlength{\textwidth}{14.5cm}
-\let\openright=\clearpage
-
-\def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-% Tato makra přesvědčují mírně ošklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-\makeatletter
-\def\@makechapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries \thechapter. #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\def\@makeschapterhead#1{
-  {\parindent \z@ \raggedright \normalfont
-   \Huge\bfseries #1
-   \par\nobreak
-   \vskip 20\p@
-}}
-\makeatother
-
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-\begin{document}
-
-\lefthyphenmin=2
-\righthyphenmin=2
-
-% -----------------------------------------------------------------------------
-
-%                            FINAL CLEANUP
-
-% - make sure all chapter names are in First Caps
-% - make sure all references have a name, like Table 4.3.4 instead of just 4.3.4
-% - get rid of all short forms: won't -> will not
-% - make sure there are no singletons
-% - make sure LaTeX does not complaint much about under/overfullness
-% - make sure everything is defined and referenced
-% - make sure all the heu/data set names are consistent. This holds especially for GLPK (tool) / Glpk (heu) and FIDAX
-% - make sure that when referring to heus in algorithms, e.g. when describing how experimental sets are constructed, we correctly refer to their parameters as described in their pseudocodes
-% - the official name for things like OVA1 and 100-100 is (test) data set. Make sure I am not calling it files or anything else.
-% - make sure labels have consistent names
-
-%%% Titulní strana práce ======================================================
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large MASTER THESIS}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% exactly as assigned
-{\LARGE\bfseries Inference of XML Integrity Constraints}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & 	\mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage % ============================================================
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-\openright
-
-\noindent
-[Sample: Here you may thank whoever you wish (the supervisor of the thesis, the consultant,
-the person who lent the software, literature, etc.)]
-
-TODO thank jInfer team, Mlynkova, anyone helping create this, colleagues at work...
-
-\newpage % ============================================================
-%%% Strana s čestným prohlášením k diplomové práci
-\vglue 0pt plus 1fill
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-signature
-\hss}}
-
-\vspace{20mm}
-
-\newpage % ============================================================
-%%% Povinná informační strana diplomové práce
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
-
-Autor: \mfauthor
-
-Katedra:  Katedra softwarového inženýrství
-
-Vedoucí diplomové práce: \mfadvisor{}, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství
-
-Abstrakt: [abstract of 80-200 words in Czech, but not a copy of the assignment of the
-master thesis]
-
-Klíčová slova:  XML, ID atributy, odvozování
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:  Inference of XML Integrity Constraints
-
-Author: \mfauthor
-
-Department: Department of Software Engineering
-
-Supervisor: \mfadvisor{}, Department of Software Engineering
-
-Abstract:  [abstract of 80-200 words in English, but not a copy of the assignment of
-the master thesis]
-
-Keywords: XML, ID attributes, inference
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\setcounter{page}{1}
-\tableofcontents
-
-% =============== TEXT ================
-\newpage
-
-% TODO what style of quotation marks to use? 
-% TODO how to format text? is textit / texttt / ... OK?
-
 \chapwithtoc{Preface}
+\label{chapter-preface}
 
 Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
+However, even though languages such as DTD and XML Schema to describe XML structure exist for a long time, most of the documents use outdated or no schema at all (\cite{1802522}).
 To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
@@
 
-TODO perhaps explain the way we will be writing pseudocode?\\
+Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly presented in the beginning.
 
-There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.
+\begin{algorithm}
+\caption{Example Algorithm}
+\label{listing-example}
+\begin{algorithmic}
+\REQUIRE $I$ input data
+\REQUIRE $n$ maximum number of iterations
+\ENSURE results found
+\FOR{$i = 1 \to n$}
+  \STATE \COMMENT{try to find a solution}
+  \STATE $attempt \gets $ calculate possible solution from $I$
+  \IF{$attempt$ is a valid solution}
+    \RETURN $attempt$
+  \ENDIF
+  \RETURN "solution not found"
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
 
-\include{chapter-1-definitions}
+There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.\\
 
-\include{chapter-2-research}
-
-\include{chapter-3-mip}
-
-\include{chapter-4-1-experiments-setup}
-
-\include{chapter-4-2-experiments-results}
-
-\include{chapter-5-future}
-
-\include{chapter-6-conclusion}
-
-%%% Seznam použité literatury
-\newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-
-\listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
-
-\listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
-
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
-\printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-\label{chapter-list-abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
-
-\openright
-\addcontentsline{toc}{chapter}{Appendices}
-
-\include{appendix-a-jinfer}
-
-\include{appendix-b-iss}
-
-\include{appendix-c-trace}
-
-% TODO a QED square
-
-\end{document}
+Please note that throughout this work we will be explicitly ignoring the $\mathcal{O}()$ complexities of algorithms we use.
+This is because the algorithms we use are by principle strongly stochastic and their performance often depends on behavior of external tools, which we regarded as black boxes and mostly ignored their inner workings.

### 2011-10-28 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 
-\chapwithtoc{Preface}
-
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure already exist since a long time, most of the documents use outdated or no schema at all (link Vosta's Even ant can create...).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
-
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
-
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
-
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
-
-\section{Structure of the thesis}
-
-The thesis will be structured as follows.
-
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
-
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
-
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
-
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
-
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
-
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
-
-\section{Conventions}
-
-As usual, source code excerpts, class, field and method names shall be written in fixed-width font, such as \texttt{get\-Heu\-ris\-tic()}.
-Names of specific heuristics will be written like this: \heu{Mutation}.
-Name of test data sets will be written like this: \dataset{OVA1}.\\
-
-TODO perhaps explain the way we will be writing pseudocode?\\
-
-There is a list of abbreviations following the bibliography in Listing \ref{chapter-list-abbreviations}.\\
-
-TODO we will be ignoring O() complexities
+\include{chapter-0-preface}
 

### 2011-10-28 doc/theses/vitasek/chapter-0-preface.tex
Thesis: in progress.

@@
 
-Along with technologies like JSON, SQL/noSQL databases and bla, XML is one of the leading formats for storing structured data.
+Along with technologies such as SQL/noSQL databases, proprietary binary file formats, plain-text configuration files and JSON, XML is one of the leading formats for storing structured data.
 However, even though languages such as DTD and XML Schema to describe XML structure exist for a long time, most of the documents use outdated or no schema at all (\cite{1802522}).
-To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in A, B, C, jInfer.
-But the schema is not the only constraint that can be imposed on an XML document: the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies here as well.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} (link Sviro).
+To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in \cite{ahonen, bex, vyhnanovska}.
+In particular, \cite{archdoc} introduces the jInfer schema inference framework, dealing primarily with the structural parts of the schema: how all the elements, attributes and text data are to be organized in an XML document conforming to that schema.
+Inference of this kind of structural information was greatly improved in \cite{anti}.\\
 
-This work will be building upon the achievements of jInfer schema inference framework (TODO link Anti's improvements in schema inference), expanding its possibilities in the field of search for \textit{key-} and \textit{foreign key-}like structures in existing XML documents.
+But the schema is not the only constraint that can be imposed on an XML document.
+Any textual or numerical value featured in the document may be subject to type constraints, such as the requirement to conform to a specific regular expression.
+Furthermore, the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies to schemas as well and will be the topic of this work.
+One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} well researched in \cite{sviro}.\\
 
-TODO Integrity constraints can be keys (with ID attributes as a "sub-group"), FKs, functional dependencies (quote Sviro), etc.
-We will focus on the first kind - ID and IDREF attributes.
+From all the constraints that can be applied to an XML document by means of its schema, this work will focus on keys and foreign keys.
+Most important concepts in this field are introduced in \cite{keX} and formalized in the notions of \texttt{ID/IDREF/IDREFS} attributes in DTD and XSD and \texttt{xs:key/xs:keyref} structures in XSD (both in \cite{Bray:08:EML}).\\
 
-TODO argument: test data with DTD (and thus possibly ID/IDREF) is more common, we can have better test sets.
+The scope of this work is finally limited to the inference of \texttt{ID/IDREF/IDREFS} from existing XML documents.
+ID attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with ID attributes are much more common and it is much easier to obtain large data sets for experiments.
+% TODO something like even the search for ID sets is interesting enough?
 
@@
 
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets and keys for XML.
+First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets, linear programming and mixed integer problem.
 
-Secondly, we will review approaches to ID attribute and XML keys search from previous articles on this topic.
+We will then review approaches to ID attribute search from previous articles on this topic and formulate the problem of finding the optimal ID set.
 
-This will lead us to the NP-complete problem of maximal independent set (IS)\nomenclature{IS}{Independent Set}, where we will inspect approaches for solving it.
+This will lead us to the NP-complete problem of Maximum Independent Set, where we will inspect the approaches to solving it.
 
-We will discuss a closely related Mixed Integer Problem (MIP)\nomenclature{MIP}{Mixed Integer Problem} and prove their "equality".
+We will discuss a closely related Mixed Integer Problem and show that by solving MIP we can solve the Maximum Independent Set and thus the original problem of optimal ID set.
 
-Afterwards, we will show how to use an external MIP solver and various heuristics to tackle this problem.
+Afterwards, we will show how to use an external MIP solver, demonstrate that this can take too long and how to use a heuristic approach to find good solutions much faster.
 
@@
 As usual, source code excerpts, class, field and method names shall be written in fixed-width font, such as \texttt{get\-Heu\-ris\-tic()}.
-Names of specific heuristics will be written like this: \heu{Mutation}.
-Name of test data sets will be written like this: \dataset{OVA1}.\\
+Names of specific heuristics will be written like \heu{Mutation}.
+Name of test data sets will be written like \dataset{OVA1}.\\
 
-Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly presented in the beginning.
+Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly marked in the beginning.
 

### 2011-10-28 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress
Thesis: in progress.

@@
 \section{Attribute Mappings}
+\label{section-definitions-ams}
 
@@
 \paragraph{Example}
-Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of every other mapping is $1 / (2+1+1+1) = 0.2$.
+Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of every other mapping is $1 / (2+1+1+1) = 0.2$.\\
+
+Note that this definition of weight is quite arbitrary and all the algorithms mentioned later could easily work with AM weight defined in any other way, even for example defined interactively by the user.
 
@@
 
-We shall need the notion of an indepentent set of vertices in a graph and its weighted variant.
+\nomenclature{IS}{Independent Set}
+
+We shall need the notion of an indepentent set (IS) of vertices in a graph and its weighted variant.
 
@@
 
+\nomenclature{MIP}{Mixed Integer Problem}
+
 \begin{define}[Mixed integer problem]

### 2011-10-30 doc/theses/vitasek/appendix-a-jinfer.tex
Thesis: in progress.

@@
 
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mario Mikula, Ro\-bert Sme\-ta\-na, Michal Svirec and Matej Vitasek.
+jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mário Mikula, Ro\-bert Sme\-ta\-na, Michal Švirec and Matej Vitásek.
 The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.
 The goal was achieved: the SW project was successfuly defended when jInfer was inferring DTD and XSD schemas based on XML documents, old DTD and XSD schemas and XPath queries.
-Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Svirec has extended the framework with capabilities to detect and repair functional dependencies violation and defended his thesis as well.
-This thesis is the third based on this framework, and Mario Mikula's is on its way, too.
+Since then, Michal Klempa has successfuly defended his own thesis improving on the grammar simplification process (see below), Michal Švirec has extended the framework with capabilities to detect and repair functional dependencies violation (see \cite{sviro}) and defended his thesis as well.
+This thesis is the third one based on this framework, and Mário Mikula's is on its way, too.
 
@@
   \caption{Inference process in jInfer}
+  \vspace{10pt}
   \label{image-inference-process}
@@
 
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG, TODO nomenclature).
+The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG).
+\nomenclature{IG}{Initial Grammar}
 This is the representation in which will the structure live until it is used to create the final product - the schema.
@@
 As such, its left hand side is an element, its right hand side is a regular expression representing its content model.
-(TODO picture?) IG is used to create the AM model used in this thesis, too.
+IG is used to create the AM model used in this thesis, too.
 jInfer contains one such module, the \jmodule{BasicIGG}, which is described in detail in \cite{basiciggdoc}.
@@
 This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis (TODO link), which describes this in great detail.
+To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis \cite{anti}, which describes this in great detail.
 Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...

### 2011-10-30 doc/theses/vitasek/appendix-b-iss.tex
Thesis: in progress.

@@
 
-\texttt{Experiment} is a class representing a single experiment with specified input data (encapsulated in \texttt{Test\-Data} interface), settings (encapsulated in \texttt{Ex\-pe\-ri\-ment\-Pa\-ra\-me\-ters}) and a metaheuristic as defined in TODO link.
+\texttt{Experiment} is a class representing a single experiment with specified input data (encapsulated in \texttt{Test\-Data} interface), settings (encapsulated in \texttt{Ex\-pe\-ri\-ment\-Pa\-ra\-me\-ters}) and a metaheuristic as defined in Section \ref{section-mip-heuristics}.
 Its method \texttt{run()} will launch the metaheuristic, first executing the construction heuristic and then running the specified improvement heuristics in a loop until termination criteria defined in an implementation of \texttt{Ter\-mi\-na\-tion\-Cri\-ter\-ion} are met.

### 2011-10-30 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
 
-After finding the best combination of a CH and IH we turn our attention to parameters. The first ones are the $\alpha$ and $\beta$ from the definition of our weight function. % TODO link
-A short reminder: the weight is defined as TODO copy formula.
+After finding the best combination of a CH and IH we turn our attention to parameters.
+The first ones are the $\alpha$ and $\beta$ from the definition of our weight function (Section \ref{section-definitions-weight}).
+A short reminder: the weight is defined as follows.
+\[weight(m) = \alpha . \phi(m) + \beta . \chi(m)\]
+Where $\phi(m)$ is support the attribute mapping $m$ and $\chi(m)$ is its coverage.
+
 It is thus obvious that only the \textit{ratio} between $\alpha$ and $\beta$ matters, not their actual values.

### 2011-10-30 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.

@@
 
-In GLPK \textit{MathProg} language, this translates to the following formulation. % TODO cite mathprog
+In GLPK \textit{MathProg} language (see \cite{mathprog}), this translates to the following formulation.
 
@@
 \section{Heuristics}
+\label{section-mip-heuristics}
 
@@
 This is again a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
-
-% TODO od I need to specify which are my own idea? I guess so, I should link any heu that is not mine. But do I need to stress that for example Fuzzy is mine?
 
@@
 This is again a rather trivial diversification improvement heuristic.
-By removing a random subset of specified size from each solution in the pool, it provides the variability needed to escape from local optima.
-% TODO make sure we have talked about local optima before
+By removing a random subset of specified size from each solution in the pool, it provides the variability needed to escape from local optima in the solution space.
 
@@
 The resulting solution (whether any AMs were added or not) is then returned to the pool.
-This is then intensification, and Listing \ref{listing-ih-hungry} captures the process.
+This is then intensification, and Listing \ref{listing-ih-hungry} captures the process.\\
 
@@
 \end{algorithm}
+
+The following three IHs: \heu{Mutation}, \heu{Crossover} and \heu{Local Branching} are inspired by \cite{heu-lecture}.
 
@@
 
-Once an ID set is found, regardless of how exactly, it is easy to find the IDREF set, i.e. the attribute mappings that can be declared as IDREF. % TODO link FIDAX, they did it first.
+Once an ID set is found, regardless of how exactly, it is easy to find the IDREF set, i.e. the attribute mappings that can be declared as IDREF. This algorithm is adopted from \cite{fidax}.
 
-First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set. This is because the specification % TODO link to the specific anchor in the page
-does not allow an attribute to be ID and IDREF (IDREFS) at the same time.
+First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set.
+This is because the specification of DTD/XSD does not allow an attribute to be ID and IDREF (IDREFS) at the same time.
 Let us denominate these mappings as \textit{IDREF candidates} (obviously different from \textit{candidate AMs}).
-% TODO link candidate AMs
 

### 2011-10-30 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 
-To understand our test data sets we discuss their origin and \textit{graph representation}. As mentioned earlier, % TODO link
-the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set in a graph.
+To understand our test data sets we discuss their origin and \textit{graph representation}.
+As mentioned earlier in Chapter \ref{chapter-mip}, the problem of finding the optimal ID set is in fact the problem of finding the maximum weighted independent set in a graph.
 Therefore it is interesting to actually see the graphs of these data sets and understand some related metrics.
@@
 
-This approach is useful, because as stated in TODO link, ID attributes are a special case of XML keys.
+This approach is useful, because as mentioned in Chapter \ref{chapter-research}, ID attributes are a special case of XML keys.
 We can use this approach to find XML keys: convert some ``suspicious" data into attributes, find the optimal ID set and then create XML key based on this ID set.
@@
   	\subfigure[\dataset{48-80}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[\dataset{70-245}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
+	\subfigure[\dataset{70-245}]{\includegraphics[width=.9\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
 \end{figure}

### 2011-10-30 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 \chapter{Experiments}
-
-% TODO I cannot stress this enough times: make sure there is an explanation of heuristics notation used in pseudocode listing, SOMEWHERE
 
@@
 Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-Neither will they be included on the DVD distributed with this thesis.
+The DVD distributed with this thesis will contain their anonymized version.
 
@@
 We used 2 data sets to convert, \dataset{MSH} and \dataset{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here.
+Unfortunately, the same problem with disclosure as in the previous case applies here, the DVD contains the anonymized version again.
 None of these sets had any attributes before the conversion.
@@
 
-We now have to introduce a few notions before moving forward to the description of our experiments.\\
+We now have to introduce a few notions before moving forward to the description of our experiments.
+In the following text, words \textit{experiment} and \textit{experimental} in various phrases (\textit{experiment X}, \textit{experimental Y}) will be used interchangeably.\\
 

### 2011-10-30 doc/theses/vitasek/chapter-2-research.tex
Thesis: in progress.

@@
 \chapter{Research}
+\label{chapter-research}
 
@@
 Simple keys found this way can be translated to an equivalent ID attribute definition.
-This opens a new line of research: finding XML keys using an algorithm modified to look only for useful keys and then converting them to ID attributes.\\
+The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}.
+This opens a new line of research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to ID attributes.\\
 

### 2011-10-30 doc/theses/vitasek/list-labels.sh

@@ -1,3 +1,7 @@
 #!/bin/bash
 
 cat *.tex | grep "\\\\label" | sed "s/^.*\\label{\([^}]*\)}.*$/\1/" | sort
+
+echo "REFS: "
+
+cat *.tex | grep "\\\\ref" | sed "s/^.*\\ref{\([^}]*\)}.*$/\1/" | sort

### 2011-10-31 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.

@@
 \section{Finding ID Sets With GLPK}
+\label{section-mip-glpk}
 
@@
 \label{section-mip-chs}
-% TODO can I format things in nomenclature?
 \nomenclature{CH}{Construction Heuristic}
@@
 
-The number of AMs to remove from each solution is specified as ratio from $(0, 1)$ of the solution size (number of AMs in solution).
-For example, \heu{Random Remove} with $ratio = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
+The number of AMs to remove from each solution is specified as fraction from $(0, 1)$ of the solution size (number of AMs in solution).
+For example, \heu{Random Remove} with $fraction = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
 
@@
 \REQUIRE $FP$ pool of feasible solutions
-\REQUIRE $k \in (0,1)$ ratio of AMs to remove from each $s \in FP$
+\REQUIRE $k \in (0,1)$ fraction of AMs to remove from each $s \in FP$
 \ENSURE pool of feasible solutions
@@
 
-This simple improvement heuristic assumes that the solutions in the pool are not ``complete", i.e. there are AMs that could be added to them without violating the ID set condition.
+This simple improvement heuristic assumes that the solutions in the pool are not ``complete'', i.e. there are AMs that could be added to them without violating the ID set condition.
 

### 2011-10-31 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 
-We have implemented these ideas and incorporated them in the jInfer framework (see Appendix \ref{appedix-jInfer}).
+We have implemented these ideas and incorporated them in the jInfer framework (see Appendix \ref{appendix-jInfer}).
 But before we describe the experiments themselves, we should try to formulate our aim.\\
@@
 
-To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical" representants.
+To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical'' representants.
 
@@
 This approach is useful, because as mentioned in Chapter \ref{chapter-research}, ID attributes are a special case of XML keys.
-We can use this approach to find XML keys: convert some ``suspicious" data into attributes, find the optimal ID set and then create XML key based on this ID set.
+We can use this approach to find XML keys: convert some ``suspicious'' data into attributes, find the optimal ID set and then create XML key based on this ID set.
 
@@
 The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the ratio of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
+This is the density as the fraction of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
 
@@
 \begin{table}
-  \caption{List of ``sized" artificial test data files}
+  \caption{List of ``sized'' artificial test data files}
   \bigskip
@@
     \subfigure[\dataset{48-80}]{\includegraphics[width=\textwidth]{images/experiments/data/artificial/size/40-80}}
-	\subfigure[\dataset{70-245}]{\includegraphics[width=.9\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
+    \subfigure[\dataset{70-245}]{\includegraphics[width=.8\textwidth]{images/experiments/data/artificial/size/70-245-fixed}}
 \end{figure}
@@
 
-The ``user-specified limits" of whiskers are set to default value, let us quote from the manual again.
+The ``user-specified limits'' of whiskers are set to default value, let us quote from the manual again.
 

### 2011-10-31 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Minor change - add space
Thesis: in progress.

@@
 
-Our experimental set will contain 500 experimental configurations for each of these two GLPK version.
+Our experimental set will contain 500 experimental configurations for each of these two GLPK versions.
 Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Algorithm \ref{listing-experiment-glpk-native-vs-cygwin-1}).
@@
 Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \dataset{100-1000}.
-Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute".
+Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute''.
 
@@
 
-% TODO for some reason \texttt{\textbf{vertex5}} looks only like \texttt{vertex5}
 \begin{table}
@@
   \hline
-                    & \texttt{\textbf{vertex5}}  \\
-                    & \texttt{\textbf{vertex26}} \\
+                    & \texttt{vertex5}  \\
+                    & \texttt{vertex26} \\
   \texttt{vertex30} & \texttt{vertex30} \\
@@
   \texttt{vertex39} & \texttt{vertex39} \\
-                    & \texttt{\textbf{vertex60}} \\
-                    & \texttt{\textbf{vertex69}} \\
-                    & \texttt{\textbf{vertex70}} \\
+                    & \texttt{vertex60} \\
+                    & \texttt{vertex69} \\
+                    & \texttt{vertex70} \\
   \texttt{vertex74} & \texttt{vertex74} \\
@@
   \centering
-  	\subfigure[\dataset{OVA1}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA1}}
-  	\subfigure[\dataset{OVA2}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA2}}
-  	\subfigure[\dataset{OVA3}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-OVA3}}
-  	\subfigure[\dataset{XMD}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-XMD}}
-   	\subfigure[\dataset{100-100}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-100}}
-  	\subfigure[\dataset{100-200}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-200}}
-  	\subfigure[\dataset{100-1000}]{\includegraphics[width=.45\textwidth]{images/experiments/best-ch-100-1000}}
+  	\subfigure[\dataset{OVA1}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-OVA1}}
+  	\subfigure[\dataset{OVA2}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-OVA2}}
+  	\subfigure[\dataset{OVA3}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-OVA3}}
+  	\subfigure[\dataset{XMD}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-XMD}}
+   	\subfigure[\dataset{100-100}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-100-100}}
+  	\subfigure[\dataset{100-200}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-100-200}}
+  	\subfigure[\dataset{100-1000}]{\includegraphics[width=.44\textwidth]{images/experiments/best-ch-100-1000}}
 \end{figure}
@@
 	\FOR{$i = 1 \to 50$}
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Crossover}(ratio = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Crossover}(fraction = 0.1, limit = 1)\}$
     	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Hungry}\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Local Branching}(ratio = 0.1, limit = 1)\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Local Branching}(fraction = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(fraction = 0.1, limit = 1)\}$
   \ENDFOR
@@
   \end{tabular}    
-	% TODO separate somehow
+\\\bigskip
+
   \begin{tabular}{l || l | l || l | l}
@@
 	\FOR{$i = 1 \to 50$}
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
-    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(ratio = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Random}, IH = \heu{Mutation}(fraction = 0.1, limit = 1)\}$
+    	\STATE $ES \gets ES \cup \{file, CH = \heu{Glpk}(limit = 1), IH = \heu{Mutation}(fraction = 0.1, limit = 1)\}$
   \ENDFOR
@@
 
-% TODO this subsection does not play well with the ToC - how to put something else in the ToC than is here?
-\subsection{Various $\alpha$, $\beta$}
+\subsection{\texorpdfstring{Various $\alpha$, $\beta$}{Various alpha, beta}}
 
@@
 The 2 different ID sets found for various $\alpha$ and $\beta$ in \dataset{OVA1} are listed in Table \ref{table-experiments-various-betas-ova1} (note that the actual names had to be anonymized for reasons discussed in Section \ref{section-realistic-data}).
-The differing attribute mapping is highlited.
+The differing attribute mapping is marked.
 
@@
     \texttt{cve@st} & \texttt{cve@st} \\
-    \texttt{\textbf{def@id}} & \texttt{\textbf{def@cl}} <- \\ % TODO again, my bold typewriter does not work
+    \texttt{* def@id} & \texttt{* def@cl} \\
     \texttt{fil@co} & \texttt{fil@co} \\
@@
 \ENDFOR
-\STATE \textbf{set} ``ignore text data"
+\STATE \textbf{set} ``ignore text data''
 \FOR{$i \in 1 \to 50 $}
@@
 \begin{itemize}
-	\item \heu{RR} is \heu{RandomRemove} with $ratio$ set to $0.1$.
-	\item \heu{MUT} is \heu{Mutation} with $ratio$ set to $0.1$ and time limit set to 1 second.
-	\item \heu{CX} is \heu{Crossover} with $ratio$ set to $0.1$ and time limit set to 1 second.
-	\item \heu{LB} is \heu{LocalBranching} with $ratio$ set to $0.1$ and time limit set to 1 second.
+	\item \heu{RR} is \heu{RandomRemove} with $fraction$ set to $0.1$.
+	\item \heu{MUT} is \heu{Mutation} with $fraction$ set to $0.1$ and time limit set to 1 second.
+	\item \heu{CX} is \heu{Crossover} with $fraction$ set to $0.1$ and time limit set to 1 second.
+	\item \heu{LB} is \heu{Local Branching} with $fraction$ set to $0.1$ and time limit set to 1 second.
 	\item \heu{RW} is \heu{RemoveWorst}.
@@
 The construction heuristic will be \heu{Random} with pool size 10.
-All the ratios are set to $0.1$ for the time being.
+All the fractions are set to $0.1$ for the time being.
 The termination criterion is set to limit the total runtime to 10 seconds and (potentialy) infinite iterations.
@@
 \STATE $ES \gets \emptyset$
-\STATE $\heu{MUT} \gets \heu{Mutation}(ratio = 0.1, limit = 1)$
-\STATE $\heu{CX} \gets \heu{Crossover}(ratio = 0.1, limit = 1)$
-\STATE $\heu{LB} \gets \heu{LocalBranching}(ratio = 0.1, limit = 1)$
-\STATE $\heu{RR} \gets \heu{RandomRemove}(ratio = 0.1)$
+\STATE $\heu{MUT} \gets \heu{Mutation}(fraction = 0.1, limit = 1)$
+\STATE $\heu{CX} \gets \heu{Crossover}(fraction = 0.1, limit = 1)$
+\STATE $\heu{LB} \gets \heu{Local Branching}(fraction = 0.1, limit = 1)$
+\STATE $\heu{RR} \gets \heu{RandomRemove}(fraction = 0.1)$
 \STATE $\heu{H} \gets \heu{Hungry}$
@@
 Its X position denotes the absolute time in which this step finished, and its Y position represents the incumbent solution quality after this step.
-Every time a line ``disappears" before reaching 10 seconds it means that this metaheuristic run found the optimum before the 10 second mark.
+Every time a line ``disappears'' before reaching 10 seconds it means that this metaheuristic run found the optimum before the 10 second mark.
 There is an obvious repetitive regularity in the shape of each line, which corresponds to the fact that there is a finite number of IHs in this strategy (5 of them in Strategy 1) which repeat over time.
@@
 
-The parameters we can tune in this strategy are the ratios in \heu{RandomRemove} (possibly 2 of them, as there are 2 instances in use), \heu{Mutation} and \heu{Crossover}.
+The parameters we can tune in this strategy are the fractions in \heu{RandomRemove} (possibly 2 of them, as there are 2 instances in use), \heu{Mutation} and \heu{Crossover}.
 We shall not tune the time limits in \heu{Mutation} and \heu{Crossover} and leave them set to 1 second.
 This presents us with a 3-dimensional space of parameters, where we want to find a combination best suited for our test data sets.
-We will sample this space by taking a total of 45 configurations of the aforementioned ratios.
+We will sample this space by taking a total of 45 configurations of the aforementioned fractions.
 
@@
 
-This experimental set will consist of 45 ratio combinations * 11 data sets * 25 iterations = 12375 experimental configurations.
+This experimental set will consist of 45 fraction combinations * 11 data sets * 25 iterations = 12375 experimental configurations.
 CH will be \heu{Random} with pool size of 10.
-IHs will be the ones from Strategy 1, with their ratios set to one of the 45 combinations produced in the following way.
+IHs will be the ones from Strategy 1, with their fractions set to one of the 45 combinations produced in the following way.
 \begin{itemize}
-	\item \heu{RandomRemove} ratio will be from $\{0, 0.05, 0.1, 0.2, 0.5\}$
-	\item \heu{Mutatio} ratio will be from $\{0.05, 0.1, 0.2\}$
-	\item \heu{Crossover} ratio will be from $\{0.05, 0.1, 0.2\}$
+	\item \heu{RandomRemove} fraction will be from $\{0, 0.05, 0.1, 0.2, 0.5\}$
+	\item \heu{Mutatio} fraction will be from $\{0.05, 0.1, 0.2\}$
+	\item \heu{Crossover} fraction will be from $\{0.05, 0.1, 0.2\}$
 \end{itemize}
@@
 
-\FOR{$rrRatio \in \{0, 0.05, 0.1, 0.2, 0.5\}$}
-\FOR{$mutRatio \in \{0.05, 0.1, 0.2\}$}
-\FOR{$cxRatio \in \{0.05, 0.1, 0.2\}$}
-  \STATE $RR \gets \heu{RandomRemoval}(ratio = rrRatio)$
-  \STATE $MUT \gets \heu{Mutation}(ratio = mutRatio, limit = 1)$
-  \STATE $CX \gets \heu{Crossover}(ratio = cxRatio, limit = 1)$
+\FOR{$rrFraction \in \{0, 0.05, 0.1, 0.2, 0.5\}$}
+\FOR{$mutFraction \in \{0.05, 0.1, 0.2\}$}
+\FOR{$cxFraction \in \{0.05, 0.1, 0.2\}$}
+  \STATE $RR \gets \heu{RandomRemoval}(fraction = rrFraction)$
+  \STATE $MUT \gets \heu{Mutation}(fraction = mutFraction, limit = 1)$
+  \STATE $CX \gets \heu{Crossover}(fraction = cxFraction, limit = 1)$
   \FOR{$file \in $ sized test data}
@@
 After averaging the data, we get a large result table; an excerpt of it is in Table \ref{table-experiments-chained-ihs-tweaking-s1}.
-Only the results for the biggest data sets and a few combinations of RR, MUT and CX ratios are presented.
-
-In the left part of the table are the ratio values.
+Only the results for the biggest data sets and a few combinations of RR, MUT and CX fractions are presented.
+
+In the left part of the table are the fraction values.
 In the right part are the averaged running times (AT), success ratios (ratio of runs that found the optimum, SR) and average qualities (AQ) for each data set.
@@
 
-It is now necessary to pick one ratio combination as the best one, and it is $(RR = 0.5, MUT = 0.2, CX = 0.1)$.
+It is now necessary to pick one fraction combination as the best one, and it is $(RR = 0.5, MUT = 0.2, CX = 0.1)$.
 Using this combination for all the data sets from \dataset{10-20} up to \dataset{80-320} the optimum was always found and for \dataset{90-405} and \dataset{100-500} the success ratios were the highest.
 
-Now to interpret the ratios in the best combination.
-\heu{RandomRemove} ratio of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
+Now to interpret the fractions in the best combination.
+\heu{RandomRemove} fraction of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
 This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima.
-\heu{Mutation} ratio of 0.2 means around 1/5\textsuperscript{th} of AMs in the incumbent solution will be fixed for the next GLPK optimization.
-\heu{Crossover} ratio of 0.1 means that around 10\% of ID sets in the pool (randomly chosen) will be scanned for common AMs.
-
-\heu{RandomRemove} and \heu{Mutation} ratios in the best combination are at the upper bound of the range we chose for them.
-As a future work option it is possible to start moving these ratios even more in their preferred way.
+\heu{Mutation} fraction of 0.2 means around 1/5\textsuperscript{th} of AMs in the incumbent solution will be fixed for the next GLPK optimization.
+\heu{Crossover} fraction of 0.1 means that around 10\% of ID sets in the pool (randomly chosen) will be scanned for common AMs.
+
+\heu{RandomRemove} and \heu{Mutation} fractions in the best combination are at the upper bound of the range we chose for them.
+As a future work option it is possible to start moving these fractions even more in their preferred way.
 
@@
 
-Finally we shall compare the performance of Strategy 1 with tuned parameters (ratios) to the approach we started the experiments with: using the \heu{Glpk} CH with no time limit to find the optimum.
+Finally we shall compare the performance of Strategy 1 with tuned parameters (fractions) to the approach we started the experiments with: using the \heu{Glpk} CH with no time limit to find the optimum.
 We will compare them on the biggest of sized test data: \dataset{100-500}.
 
-We already have the running times for pure \heu{Glpk} on \dataset{100-500} from the ``Time Until Optimum" experiment in Section \ref{section-glpk-comparison}.
+We already have the running times for pure \heu{Glpk} on \dataset{100-500} from the ``Time Until Optimum'' experiment in Section \ref{section-glpk-comparison}.
 The last experiment to find the performance of tuned Strategy 1 without a time limit will have the following parameters.
@@
 \STATE $RW \gets \heu{RemoveWorst}$
-\STATE $RR \gets \heu{RandomRemoval}(ratio = 0.5)$
-\STATE $MUT \gets \heu{Mutation}(ratio = 0.2, limit = 1)$
-\STATE $CX \gets \heu{Crossover}(ratio = 0.1, limit = 1)$
+\STATE $RR \gets \heu{RandomRemoval}(fraction = 0.5)$
+\STATE $MUT \gets \heu{Mutation}(fraction = 0.2, limit = 1)$
+\STATE $CX \gets \heu{Crossover}(fraction = 0.1, limit = 1)$
 \FOR{$i = 1 \to 50$}
@@
 
-\section{The "Best" Algorithm}
+\section{The ``Best'' Algorithm}
 

### 2011-10-31 doc/theses/vitasek/chapter-2-research.tex
Thesis: in progress.

@@
 Fajt in \cite{fajt} introduces several algorithms to find XML keys (see \cite{keX}) in existing XML documents.
-Simple keys found this way can be translated to an equivalent ID attribute definition.
+Simple keys found this way can be translated to an equivalent \texttt{ID} attribute definition.
 The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}.
-This opens a new line of research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to ID attributes.\\
+This opens a new line of research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to \texttt{ID} attributes.\\
 

### 2011-10-31 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 
-\subsection{Realistic data}
+\subsection{Realistic Data}
 \label{section-realistic-data}
@@
 
-\subsection{Realistic data with artificial attributes}
+\subsection{Realistic Data With Artificial Attributes}
 \label{section-realistic-converted}
@@
 
-This approach is useful, because as mentioned in Chapter \ref{chapter-research}, ID attributes are a special case of XML keys.
+This approach is useful, because as mentioned in Chapter \ref{chapter-research}, \texttt{ID} attributes are a special case of XML keys.
 We can use this approach to find XML keys: convert some ``suspicious'' data into attributes, find the optimal ID set and then create XML key based on this ID set.
@@
 In case of \dataset{MSH} we created 2 attributes, of which only one constituted a candidate AM.
-This is then the case similar to \dataset{XMA-*} sets: quite large data, yet only one trivial ID attribute to be found.
+This is then the case similar to \dataset{XMA-*} sets: quite large data, yet only one trivial \texttt{ID} attribute to be found.
 
@@
 
-\subsection{Artificial data}
+\subsection{Artificial Data}
 
@@
 \begin{algorithm}
-\caption{Random XML data creation}
+\caption{Random XML Data Creation}
 \label{listing-random-data}
@@
 There is a tuple for every attribute node with a value found in the initial grammar.
-Note that the information about the context in which the element was originally found is lost - but this is not a problem with regard to the definition of XML ID attributes.
+Note that the information about the context in which the element was originally found is lost - but this is not a problem with regard to the definition of XML \texttt{ID} attributes.
 Furthermore, tuples in flat representation do not need to be unique.\\

### 2011-10-31 doc/theses/vitasek/chapter-6-conclusion.tex
Thesis: in progress.

@@
 
-From all the integrity constraints in XML we chose the ID / IDREF / IDREFS attributes and decided to improve upon the search for them.
+From all the integrity constraints in XML we chose the \texttt{ID}/\.\texttt{IDREF}/\.\texttt{IDREFS} attributes and decided to improve upon the search for them.
 We discussed the approach from \cite{fidax} and the equivalence of ID set search and maximum weighted independent set.
@@
 
-To the best of our knowledge, at the time of writing this work is our approach to finding ID attributes the best one known.
+To the best of our knowledge, at the time of writing this work is our approach to finding \texttt{ID} attributes the best one known.

### 2011-10-31 doc/theses/vitasek/chapter-0-preface.tex
Thesis: in progress.

@@
 Furthermore, the concept of \textit{keys} and \textit{foreign keys}, well known from the relational database world, applies to schemas as well and will be the topic of this work.
-One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} well researched in \cite{sviro}.\\
+One could go even further and try to find even more sophisticated relations in the data, such as \textit{functional dependencies} researched in \cite{sviro}.\\
 
 From all the constraints that can be applied to an XML document by means of its schema, this work will focus on keys and foreign keys.
-Most important concepts in this field are introduced in \cite{keX} and formalized in the notions of \texttt{ID/IDREF/IDREFS} attributes in DTD and XSD and \texttt{xs:key/xs:keyref} structures in XSD (both in \cite{Bray:08:EML}).\\
+Most important concepts in this field are introduced in \cite{keX} and formalized in the notions of \texttt{ID}/\.\texttt{IDREF}/\.\texttt{IDREFS} attributes in DTD and XSD and \texttt{xs:key/xs:keyref} structures in XSD (both in \cite{Bray:08:EML}).\\
 
-The scope of this work is finally limited to the inference of \texttt{ID/IDREF/IDREFS} from existing XML documents.
-ID attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with ID attributes are much more common and it is much easier to obtain large data sets for experiments.
+The scope of this work is finally limited to the inference of \texttt{ID}/\.\texttt{IDREF}/\.\texttt{IDREFS} from existing XML documents.
+\texttt{ID} attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with \texttt{ID} attributes are much more common and it is much easier to obtain large data sets for experiments.
 % TODO something like even the search for ID sets is interesting enough?
@@
 
-First, we will introduce a few notions required throughout the work, such as XML tree, ID attributes, ID sets, linear programming and mixed integer problem.
+First, we will introduce a few notions required throughout the work, such as XML tree, \texttt{ID} attributes, ID sets, linear programming and the mixed integer problem.
 
@@
 
-Afterwards, we will show how to use an external MIP solver, demonstrate that this can take too long and how to use a heuristic approach to find good solutions much faster.
+Afterwards, we will show how to use an external MIP solver and demonstrate that this can take too much time. Next we show how to use a heuristic approach to find good solutions much faster.
 
-An extension to jInfer for finding ID attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
+An extension to jInfer for finding \texttt{ID} attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
 
@@
   \ENDIF
-  \RETURN "solution not found"
+  \RETURN ``solution not found''
 \ENDFOR

### 2011-10-31 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress.

@@
 This tree shall be called an \textit{XML tree}.
-For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the referenced article does.
+For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.
 
@@
 \begin{define}[Node equality]
-	$v_1$ and $v_2$ are node equal, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
+	$v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
 \end{define}
@@
 \begin{define}[Value equality]
-	$v_1$ and $v_2$ are value equal, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
+	$v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
 \end{define}
 
-\section{ID, IDREF, IDREFS Attributes}
+\section{\texttt{ID}, \texttt{IDREF}, \texttt{IDREFS} Attributes}
 \label{section-definitions-id-attributes}
 
-According to \cite{Bray:08:EML}, an XML attribute may have the type \textit{ID}, \textit{IDREF} or \textit{IDREFS} (among others). Following constraints are related to these types.
+According to \cite{Bray:08:EML}, an XML attribute may have the type \texttt{ID}, \texttt{IDREF} or \texttt{IDREFS} (among others). Following constraints are related to these types.
 
 \begin{quote}
-\textbf{Validity constraint: ID}
+\textbf{Validity constraint: \texttt{ID}}
 
-Values of type \textbf{ID} \textsc{must} match the Name production.
-A name \textsc{must not} appear more than once in an XML document as a value of this type; i.e., \textbf{ID} values \textsc{must} uniquely identify the elements which bear them.
+Values of type \texttt{ID} \textsc{must} match the Name production.
+A name \textsc{must not} appear more than once in an XML document as a value of this type; i.e., \texttt{ID} values \textsc{must} uniquely identify the elements which bear them.
 
-\textbf{Validity constraint: One ID per Element Type}
+\textbf{Validity constraint: One \texttt{ID} per Element Type}
 
-An element type \textsc{must not} have more than one \textbf{ID} attribute specified.
+An element type \textsc{must not} have more than one \texttt{ID} attribute specified.
 
-\textbf{Validity constraint: ID Attribute Default}
+\textbf{Validity constraint: \texttt{ID} Attribute Default}
 
-An \textbf{ID} attribute \textsc{must} have a declared default of \textbf{\#IMPLIED} or \textbf{\#REQUIRED}.
+An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED} or \texttt{\#REQUIRED}.
 
-\textbf{Validity constraint: IDREF}
+\textbf{Validity constraint: \texttt{IDREF}}
 
-Values of type \textbf{IDREF} \textsc{must} match the Name production, and values of type \textbf{IDREFS} \textsc{must} match Names; each Name \textsc{must} match the value of an \textbf{ID} attribute on some element in the XML document; i.e.
-\textbf{IDREF} values \textsc{must} match the value of some ID attribute.
+Values of type \texttt{IDREF} \textsc{must} match the Name production, and values of type \texttt{IDREFS} \textsc{must} match Names; each Name \textsc{must} match the value of an \texttt{ID} attribute on some element in the XML document; i.e.
+\texttt{IDREF} values \textsc{must} match the value of some \texttt{ID} attribute.
 \end{quote}
@@
 
-%TODO
-
 %\subsection{According to Specification}
 
-%TODO compare them to ID attributes
+% - compare them to ID attributes
 
@@
 
-%TODO show how a XML key is defined using the keX notation % this is a terrible sentence
+% - show how a XML key is defined using the keX notation % this is a terrible sentence
 
@@
 
-Based on the requirements for an ID attribute from Section \ref{section-definitions-id-attributes} we will define ID set with the help of the following definition.
+Based on the requirements for an \texttt{ID} attribute from Section \ref{section-definitions-id-attributes} we will define ID set with the help of the following definition.
 
@@
 
-That is, an ID set has images without repeating values and all the types are unique (an element cannot have more than one ID attribute).
+That is, an ID set has images without repeating values and all the types are unique (an element cannot have more than one \texttt{ID} attribute).
 
@@
 
-\subsection{IDREF and IDREFS Condition}
+\subsection{\texttt{IDREF} and \texttt{IDREFS} Condition}
 
-Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked IDREF.
+Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}.
 \[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i)\]
-Furthermore if $m$ contains multivalued attributes, it is to be marked IDREFS.
+Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.
 
@@
 Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of every other mapping is $1 / (2+1+1+1) = 0.2$.\\
-
-Note that this definition of weight is quite arbitrary and all the algorithms mentioned later could easily work with AM weight defined in any other way, even for example defined interactively by the user.
 
@@
 \end{define}
+
+Note that this definition of weight is quite arbitrary and all the algorithms mentioned later could easily work with AM weight defined in any other way, even for example defined interactively by the user.
 

### 2011-10-31 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress.

@@
 
-Table \ref{table-experiments-fidax-and-hungry-idsets} lists the ID attributes found in both cases for this most extreme input, \dataset{100-1000}.
-Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as ID attribute''.
+Table \ref{table-experiments-fidax-and-hungry-idsets} lists the \texttt{ID} attributes found in both cases for this most extreme input, \dataset{100-1000}.
+Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as \texttt{ID} attribute''.
 
@@
 However, there is a clear improvement of about 50 \% in case of model creation times.
-The conclusion then is to ignore the simple text node content whenever possible when finding ID attributes.
+The conclusion then is to ignore the simple text node content whenever possible when finding \texttt{ID} attributes.
 

### 2011-11-01 doc/theses/mikula/TODO_NAZOV.tex
Master thesis: In progress.

\chapter{Diskusia k problemom}
\todo[inline]{Vhodny nazov pre tuto kapitolu}

1) odvodzovanie struktury z queries

2) pattern searching vs lexical, syntax and semantics analysis

### 2011-11-01 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

\chapter{Proposed Algorithm}
\todo[inline]{Nejaky uvod}
\todo[inline]{Lexikalna a syntakticka analyza z diplomky J. Schejbala}
\todo[inline]{Staticka analyza typov vyrazov}

\section{Inference of Built-in Types}
Prejst kazdy vyraz a ak z neho vieme nieco odvodit, tak to urobime.

Odvodit nieco vieme ak:
Nejaky podvyraz je path expression (alebo ine urcenie uzlu alebo hodnot uzlov \todo{toto domysliet}
a zaroven cely vyraz
1) je function call a podvyraz je jeho parameter a typ tohoto parametra vieme odvodit z hlavicky funkcie
2) je aritmeticka operacia a podla typov operandov vieme urcit typ podvyrazu (tj napriklad sa uzly porovnavaju s konstantou 5) \todo{co vsetko?}
3) dalsie konstrukcie ako cast as, type constructor, ... \todo{potrebuje riesit? princip je uplne rovnaky}

\begin{algorithm}
\caption{Repair RW-XML tree}
\label{propAlgo}
\begin{algorithmic}[1]
\REQUIRE $queryTrees$: A list of input query trees
\ENSURE Uzly a ich typy nejakym sposobom

\STATE initializacia nodeTypes
\FORALL{$queryTree \in queryTrees$}
    \STATE $partialNodeTypes = processQueryTree(queryTree)$
    \STATE $nodeTypes = mergeNodeTypes(nodeTypes, partialNodeTypes)$
\ENDFOR
\RETURN $nodeTypes$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Process arithmetic comparison}
\label{algorithm_process_arithmetic comparison}
\begin{algorithmic}[1]
\REQUIRE $expr$: An expression which is an arithmetic comparison
\ENSURE Uzly a ich typy nejakym sposobom

\STATE initializacia nodeTypes
\RETURN $nodeTypes$

\WHILE{$resultRXT \not \models_w \mathcal{FD}$}
    \STATE $S = \emptyset$ \COMMENT Set of repair groups
    \FORALL{$(F: S \rightarrow p) \in \mathcal{FD}$ s.t. $RXT \not \models_w F$}
    	\FORALL{$t_1, t_2$ tuples of $RXT$ s.t. $t_1, t_2$ do not weakly satisfy $F$}
		    \STATE $S = S \cup computeRepairGroup(F, t_1, t_2, RXT, S)$
	    \ENDFOR
    \ENDFOR

    \STATE $R = getRepair(S, RXT)$
    \STATE $resultRXT = applyRepair(R, resultRXT)$
\ENDWHILE
\end{algorithmic}
\end{algorithm}

### 2011-11-01 doc/theses/mikula/TODO_NAZOV.tex
Master thesis: In progress.

@@
 \todo[inline]{Vhodny nazov pre tuto kapitolu}
+Based on the previous chapter, we have quite wide space of possible utilization of XQuery queries.
+This chapter discusses some of questions and issues that emerged in an early phase of algorithm fabrication.
+
+\section{Possibilities and Forms of Query Exploitation}
+Queries can be exploited in several ways and levels.
+A basic query utilization can be achieved by analysis of queries without any other input data.
+A more complex method can utilize queries along with XML data.
+Another step can be evaluation of the queries using the XML data and consecutive analysis of the results.
+Even the process of evaluation can be analyzed to obtain some useful information.
+\todo[inline]{Uviest nejake priklady, co by sa dalo v ktorej faze zistovat?}
+
+\todo[inline]{Dopisat, ako sme sa rozhodli.
+Este presne neviem, ale pravdepodobne prva faza bude spracovanie cisto queries a druha by mohla byt kontrola na zaklade XML dat.
+Na dalsie vstupne data asi nebude kapacita}
+
+Another important question is how can be the queries processed.
+Will they be just searched for certain patterns like it is performed in method \cite{Necasky:2009:DXK:1529282.1529414} or will they be processed in a more sophisticated way?
+That could incorporate lexical and syntactic analyses \todo{referencie?} or even a form of an analysis of semantics \todo{referencia?}.
+
+A result of the lexical and syntantic analyses can be some kind of so called syntax tree \todo{referencia}.
+It is a structure representing a word according to a formal grammar of some language \todo{definovat grammar, tree, language}.
+In our case, the language is XQuery, its grammar is defined in \todo{ref} and every query is a word of the XQuery language.
+Leafs of the tree represent terminals of the grammar while inner nodes represent non-terminals.
+\todo{priklad nejakeho stromu?} From point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient.
+For instance, the tree can simplify a search for FLWOR statements.
+It is transitioned and nodes representing FLWORs are found.
+Then each subtree determined by one of the found nodes represents a FLWOR statement and it can be analyzed further.
+
+The syntax tree can be also extended by additional information.
+An example is a static analysis of expression types.
+Types of literal expressions are defined, functions have return types, path expressions return nodes, etc.
+Applying rules defined in \todo{ref} can be determined types of complex expressions.
+These can be helpful for example in the analysis of built-in types of nodes.
+\todo[inline]{priklad}
+\todo[inline]{Pre co sme sa rozhodli?}
+
+\section{Inference of Structure}
 
@@
 2) pattern searching vs lexical, syntax and semantics analysis
+
+3) vstupne data
+
+4) napojenie na existujucu metodu
+
+5) spravavanie dotazov - len ich analyza VS ich spustanie

### 2011-11-02 doc/theses/mikula/TODO_NAZOV.tex
Master thesis: In progress.

@@
 \todo[inline]{Vhodny nazov pre tuto kapitolu}
-Based on the previous chapter, we have quite wide space of possible utilization of XQuery queries.
-This chapter discusses some of questions and issues that emerged in an early phase of algorithm fabrication.
+According to the analysis in the previous chapter, there is quite a wide range of possible utilization of XQuery queries.
+Besides analysis of what information can be extracted from queries, it is needed to devise how the queries will be processed.
+This chapter discusses some questions and issues that emerged in an early phase of the algorithm fabrication.
 
-\section{Possibilities and Forms of Query Exploitation}
-Queries can be exploited in several ways and levels.
+\section{Input Data}
+The first important question is what is the input of the algorithm.
 A basic query utilization can be achieved by analysis of queries without any other input data.
-A more complex method can utilize queries along with XML data.
+The analysis of XQuery in the previous chapter discusses mostly XQuery constructs which can be utilized without respective XML data, for example the inference of built-it types.
+This independence is also the main advantage of this approach for if there are no XML data available this approaches can be still used.
+
+A more complex method can utilize queries along with the respective XML data.
+As discussed in the previous chapter, an element and attribute structure can be inferred from the XML data in a more convenient way.
+Also, the XML data can be used to verify information inferred from the queries or vice-versa.
+For example, utilizing the requeries some attribute is considered as a key of its element.
+But in the data there are elements with the same value of this attribute, and thus, it cannot be the key.
+Vice-versa, we have a notion that the attribute might be the key but we are not sure about that.
+Analysing of the data and finding that values of the attribute are unique can increase our confidence.
+
 Another step can be evaluation of the queries using the XML data and consecutive analysis of the results.
-Even the process of evaluation can be analyzed to obtain some useful information.
-\todo[inline]{Uviest nejake priklady, co by sa dalo v ktorej faze zistovat?}
+And even the process of evaluation can be analyzed to obtain some useful information.
+\todo[inline]{Toto skusit rozpisat nejak podrobnejsie?}
 
-\todo[inline]{Dopisat, ako sme sa rozhodli.
+\todo[inline]{Dopisat, ako sme sa rozhodli - nejaky zaver.
 Este presne neviem, ale pravdepodobne prva faza bude spracovanie cisto queries a druha by mohla byt kontrola na zaklade XML dat.
@@
 
-Another important question is how can be the queries processed.
+\section{Forms of Query Precessing}
+Another important question is how the queries can be processed.
 Will they be just searched for certain patterns like it is performed in method \cite{Necasky:2009:DXK:1529282.1529414} or will they be processed in a more sophisticated way?
-That could incorporate lexical and syntactic analyses \todo{referencie?} or even a form of an analysis of semantics \todo{referencia?}.
+That can mean incorporating lexical and syntactic analyses \todo{referencie?} or even a form of an analysis of semantics \todo{referencia?}.
 
@@
 Applying rules defined in \todo{ref} can be determined types of complex expressions.
-These can be helpful for example in the analysis of built-in types of nodes.
-\todo[inline]{priklad}
+These can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.
+
+Following text is an example of inference of a more complex query processing. Consider the following part of a query.
+\begin{verbatim}
+declare function local:getB($id as xs:string) as element() {
+  //A[@id = $id]/B
+}
+... local:getBs("id") > 10 ...
+\end{verbatim}
+The query consists of a function declaration and an arithmetic comparison.
+The comparison compares a result of the function call and literal value \texttt{10}.
+For the type of \texttt{10} is \texttt{xs:integer} the type of the function call has to be convertible to \texttt{xs:double}.
+It has to be some numeric type.
+The function returns a path expression typed as \texttt{element()}.
+That means the function returns one element \todo{pravda?}.
+In the path expression, the argument \texttt{\$id} can be substituted by the real value specified in the call.
+Thus, the return expression is \texttt{//A[@id = "id"]/B}.
+Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"}, is of some numeric type.
+And, for the function is parametrized there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.
+
+A more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
+But it is also more difficult in terms of implementation.
+On the other hand, simplier approaches of the query processing such as the pattern finding are easier to implement but they also limit possibilities the whole query utilization.
+
 \todo[inline]{Pre co sme sa rozhodli?}
 
-\section{Inference of Structure}
+\section{Inference of XML Structure}
+odvodzovanie struktury z queries
 
-1) odvodzovanie struktury z queries
+\section{Rozsirenie nejakej existujucej metody}
 
-2) pattern searching vs lexical, syntax and semantics analysis
-
-3) vstupne data
-
-4) napojenie na existujucu metodu
-
-5) spravavanie dotazov - len ich analyza VS ich spustanie
+\section{Conclusion}

### 2011-11-04 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

--- 2011-11-02 doc/theses/mikula/TODO_NAZOV.tex
@@
-\chapter{Diskusia k problemom}
+\chapter{Pre-creation of Algorithm}
 \todo[inline]{Vhodny nazov pre tuto kapitolu}
-According to the analysis in the previous chapter, there is quite a wide range of possible utilization of XQuery queries.
+According to the analysis in the previous chapter, there is quite wide range of possible utilizations of XQuery queries.
 Besides analysis of what information can be extracted from queries, it is needed to devise how the queries will be processed.
@@
 Also, the XML data can be used to verify information inferred from the queries or vice-versa.
-For example, utilizing the requeries some attribute is considered as a key of its element.
+For example, utilizing the queries, some attribute is considered a key of its element.
 But in the data there are elements with the same value of this attribute, and thus, it cannot be the key.
@@
 Another step can be evaluation of the queries using the XML data and consecutive analysis of the results.
-And even the process of evaluation can be analyzed to obtain some useful information.
+And even the process of evaluation itself can be analyzed to obtain some useful information.
 \todo[inline]{Toto skusit rozpisat nejak podrobnejsie?}
@@
 \section{Inference of XML Structure}
-odvodzovanie struktury z queries
+The question of the inference of XML structure from queries is partially discussed in the previous chapter.
+We are able to infer XML structure from queries without their evaluation, but in a limited way.
+This inference is based on an analysis of path expressions.
+Its limitation involves the following issues.
+When we infer some subelements of a certain element we often cannot be sure about their number of occurrences.
+Also, we cannot be sure if every occurrence of the certain element contains these subelements and we even do not know if at least one occurrence of the element contains them.
+Thus, the inferred statement is more likely an indication than a fact about the structure.
 
-\section{Rozsirenie nejakej existujucej metody}
+If we want to infer the structure more precisely we need to evaluat the queries, and thus, we need the XML data to evaluate the queries upon.
+But, if we have the XML data we can infer the structure directly from them in an easier and more convenient way.
 
-\section{Conclusion}
+Considering the described issues, we concluded not to deal with the inference of XML structure in this thesis.
+
+\section{Extension of an Existing Approach}
+The existing approaches of XML schema inference deal mainly with the inference of XML structure.
+As stated in the previous section, this work does not deal with the inference of XML structure, and hence, the extension of an existing approach will probably be some kind of an independent addition instead of modification and refinement of its core algorithm.
+
+The existing approaches take the XML data on their input.
+Therefore, the basic idea is that the input will be extended also for XQuery queries and the algorithm will consist of three phases.
+The first phase will be taken from the existing approach and it will process the XML data to infer the XML structure.
+The second phase will process the XQuery queries and it will infer statements that can be inferred indepedently of the XML data.
+The third phase will merge the statements inferred in the second phase into the resulting schema.
+This phase may also infer additional statements from both the XML data and the queries or it may try to verify the statements from the second phase with respect to the XML data.

### 2011-11-06 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

\chapter{Used Technologies and Definitions}

\todo[inline]{Zvyraznovanie pojmov}

\section{XML schema}
An XML schema refers to a description of an XML document in terms of its structure and various constraints. Commonly, the XML schema describes element and attribute names, their parent-child relations, their order and a type of their content. Other constraints often expressed in the XML schema are restrictions on numbers of occurrences of elements, specification of (non-)obligatory attributes, uniqueness and specification of keys.

Various languages have been proposed to express XML schemata. The most known are Document Type Definition (DTD) \todo{link} and XML Schema Definition (XML Schema, XSD) \todo{link} which are briefly described in the following sections. Another examples of the XML schema languages are RELAX NG \todo{link} and  Schematron \todo{link}.

A validity of an XML document againt its XML schema expresses wheather the document is well-formed \todo{link} and at the same time wheather it conforms to the XML schema.

\subsection{DTD}
Document Type Definition (DTD) expresses a structure of XML documents by declarations of elements. An element has its name, content and optional list of attributes.

The content of an element can be denoted by \emph{EMPTY} for an empty element, \emph{ANY} for any content, \emph{(\#PCDATA)} allowing only textual content (without any other subelements), or specified using regular expressions. Names of subelement are combined using the well-known operators (\emph{|}, \emph{+}, \emph{*}, \emph{?} and \emph{,}(comma)). To express the mixed content \emph{\#PCDATA} can be used in an alternation list with the subelement names and this alternation has to be enclosed in \emph{*} operator.

\todo[inline]{dopisat, priklady}

\subsection{XSD}
\todo[inline]{TODO}

\section{XPath}
\todo[inline]{TODO}

\section{XQuery}
\todo[inline]{TODO}

\todo[inline]{možná ňejaký automat apod. Prostě co budete dále v textu potřebovat. Skratky: XSD, XSLT, ...}

### 2011-11-08 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \todo[inline]{Staticka analyza typov vyrazov}
+
+\section{Construction of a Syntax Tree}
+The first step of the algorithm involves lexical and syntactic analyses and produces a syntax tree.
+The analyses are taken from Ji\v r\'{i} Schejbal's master thesis \todo{ref}.
+Since they are much more an issue of implementation than an issue of principle and they are not directly related to the inference, we will not describe them in this thesis.
+
+The mentioned thesis provides us with a helpful processing of XQuery queries, however, it needs to be slightly modified to suit our case.
+The processing writes its result to a file in an XML representation.
+Instead of that, we need to keep the result in memory and pass it to consecutive steps of our algorithm.
+Though, this requirement concerns modifications of implementation while the core of the processing remains untouched.
+Therefore, we also do not describe these modifications.
+
+Since we only modified the implementation of the output representation, the design of the syntax tree remains consequent to the entire design of the processing.
+Each node represents a logical component of a query.
+Inner nodes stand for XQuery constructs that are composed of other constructs and can be further devided.
+Examples of such constructs are FLWORs, if-than-else and function calls.
+Leaf nodes represents elements of XQuery language that can be devided no more.
+Typical representants are literal constants.
+\todo{zistit, ci konstanty nahodou nie su jedine}.
+
+\todo[inline]{priklad stromu}
+\todo[inline]{niekam popisat celu moznu strukturu stromu - najskor do appendixu}
 

### 2011-11-08 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\section{Construction of a Syntax Tree}
+\section{Step 1: Construction of a Syntax Tree}
 The first step of the algorithm involves lexical and syntactic analyses and produces a syntax tree.
@@
 Each node represents a logical component of a query.
-Inner nodes stand for XQuery constructs that are composed of other constructs and can be further devided.
+Inner nodes stand for XQuery constructs that are composed of other constructs and can be further divided.
 Examples of such constructs are FLWORs, if-than-else and function calls.
@@
 Typical representants are literal constants.
-\todo{zistit, ci konstanty nahodou nie su jedine}.
+\todo{zistit, ci konstanty nahodou nie su jedine}
 
@@
 \todo[inline]{niekam popisat celu moznu strukturu stromu - najskor do appendixu}
+
+\section{Step 2: Static Analysis of Expression Types}
+In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree. This process can be divided into two substeps.
+
+\subsection{Determination of Function Return Types}
+Determination of Return Types of Functions is needed because function calls can appear in expressions.
+However a return type of a function can be determined when the analysis of expressions encounters a call of the function, it involves multiple number of transitions of the syntax tree in a search for a definition of a particular function.
+
+Instead, the syntax tree can be searched just once, before the analysis of expressions, and return types of all functions found will be stored within an associative array.
+On request, this array will retrieve the return type of a specified function.
+
+A simple algorithm is presented in \ref{ALG_determination_of_function_return_types} \todo{uvadzat aj taketo jednoduche funkcie?}.
+It uses \texttt{getFunctionDeclarationNodes()} function, which is not presented due to its simplicity.
+It returns a list of all function declaration nodes in the syntax three.
+Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in query prolog.
+
+For simplicity, we will focus on locally defined functions.
+We will not determine types of functions defined in other modules \todo{je to ok?}.
+Return types of built-in functions are fixed, and thus, there is no need to analyze them.
+
+\begin{algorithm}
+\caption{Determination of Function Return Types}
+\label{ALG_determination_of_function_return_types}
+\begin{algorithmic}[1]
+\ENSURE Associative array of function names with their types.
+
+\STATE $functionTypes$ = empty associative array
+\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes()$}
+    \STATE $functionName = functionDeclarationNode.getFunctionName()$
+    \STATE $functionTypes[functionName] = functionDeclarationNode.getReturnType()$
+\ENDFOR
+\RETURN $functionTypes$
+\end{algorithmic}
+\end{algorithm}
+
+For the rest of the thesis, we assume function \texttt{getFunctionReturnType()} which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
+Otherwise, \texttt{null} is returned.
+
+\subsection{Samotna analyza expressions}
+\todo[inline]{Je nutne analyzovat aj typy deklarovanych premennych, kedze tie sa mozu pouzivat vo vyrazoch.}
 

### 2011-11-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \section{Step 2: Static Analysis of Expression Types}
-In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree. This process can be divided into two substeps.
+In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree. This process can be divided into three substeps.
 
@@
 \begin{algorithmic}[1]
-\ENSURE Associative array of function names with their types.
+\ENSURE An associative array of function names with their types.
 
@@
 
-\subsection{Samotna analyza expressions}
+\subsection{Determination of Global Variable Types}
+A similar approaches as in the case of functions can be applied to determine types of global variables.
+Alike the functions, the global variables are defined in the prolog section.
+A type of a variable can be explicitly specified in the definition \todo{je to tak?
+priklad}.
+If it is not it may be often deducible from the binding expression.
+
+\begin{algorithm}
+\caption{Determination of an Expression Type}
+\label{ALG_determination_of_an_expression_type}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $exprNode$: An expression node in the syntax tree. \\
+    $contextVarTypes$: Types of local variables valid in current context.
+}
+\ENSURE The expression type.
+
+\STATE $type = null$
+\IF{$exprNode$ represents a literal value}
+     \STATE $type =$ particular literal value type
+\ELSIF{$exprNode$ represents a call of function $f$}
+     \STATE $type = getFunctionReturnType(f)$
+\ELSIF{$exprNode$ represents variable reference $var$}
+     \STATE $type = getVariableType(contextVarTypes, var)$
+\ELSIF{$exprNode$ represents path expression $path$}
+     \STATE $type = TODO$
+\ELSIF{$exprNode$ represents operator $op$}
+      \STATE TODO
+\ENDIF
+
+\RETURN $type$
+\end{algorithmic}
+\end{algorithm}
+
+The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type than the type is noted.
+Otherwise, an attempt of the type deduction the expression is be made.
+The deduction of type from the expression is presented in \ref{ALG_determination_of_an_expression_type}.
+On input, it takes an expression node and types of local variables that are valid in the expression.
+The presented algorithm is just shortened illustration since the complete version is too long to be presented in this text.
+Nevertheless, the principle is straightforward.
+Depending on the type of the expression, its type can be either determined directly or it depends on its subexpressions which types can be determined recursively.
+In the case of global variables, the argument $contextVarTypes$ is empty because there are no local variables valid in the prolog section.
+
+The types of processed global variables are available trough function $getGlobalVariableType()$.
+If the function gets a variable that is not a global one, it returns $null$.
+Function $getVariableType()$ called in \ref{ALG_determination_of_an_expression_type} is defined in \ref{ALG_function_get_variable_type}.
+It checks if the specified variable is amongst the given local variables.
+If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is $null$.
+
+\begin{algorithm}
+\caption{Function getVariableType}
+\label{ALG_function_get_variable_type}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $contextVarTypes$: Types of local variables valid in current context. \\
+    $var$: A variable to determine the type of.
+}
+\ENSURE The type of variable $var$.
+
+\IF{$contextVarTypes[var]$ is defined}
+     \RETURN $contextVarTypes[var]$
+\ELSE
+     \RETURN $getGlobalVariableType(var)$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Analysis of Expressions}
 \todo[inline]{Je nutne analyzovat aj typy deklarovanych premennych, kedze tie sa mozu pouzivat vo vyrazoch.}

### 2011-11-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\subsection{Analysis of Expressions}
-\todo[inline]{Je nutne analyzovat aj typy deklarovanych premennych, kedze tie sa mozu pouzivat vo vyrazoch.}
+\subsection{Determination of Expression Types}
+\todo[inline]{VYRIESIT: external variables, scope and visibility of variables - ako si udrziavat}
+
+\begin{algorithm}
+\caption{Function analysisOfExpressionTypes}
+\label{ALG_func_analysisOfExpressionTypes}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $startingNode$: A node determinig a subtree to perform the analysis on \\
+    $contextVarTypes$: Types of local variables valid in the current subtree.
+}
+
+\FORALL{$subnode \in$ all subnodes of $startingNode$}
+    \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
+    \IF{$subnode$ defines new variables valid for following subnodes}
+        \STATE $newVarTypes = subnode.variableTypes$
+        \STATE TODO rozsir context
+    \ENDIF
+\ENDFOR
+\IF{$startingNode$ is expression node}
+    \STATE $startingNode.type = determineExpressionType(startingNode)$
+\ENDIF
+\IF{$startingNode$ defines new variables valid in the supernode}
+    \STATE TODO spracuj variables
+    \STATE memorize their types in $startingNode.variableTypes$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Determination of an Expression Type in Query Body}
+\label{ALG_determination_of_an_expression_type_in_query_body}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $exprNode$: An expression node in the syntax tree. \\
+    $contextVarTypes$: Types of local variables valid in the current context.
+}
+\ENSURE The expression type.
+
+\STATE $type = null$
+\IF{$exprNode$ represents a literal value}
+     \STATE $type =$ particular literal value type
+\ELSIF{$exprNode$ represents a call of function $f$}
+     \STATE $type = getFunctionReturnType(f)$
+\ELSIF{$exprNode$ represents variable reference $var$}
+     \STATE $type = getVariableType(contextVarTypes, var)$
+\ELSIF{$exprNode$ represents path expression $path$}
+     \STATE $type = TODO$
+\ELSIF{$exprNode$ represents operator $op$}
+     \STATE TODO
+\ELSIF{$exprNode$ represents $FLWOR$ construct $flwor$}
+     \STATE $contextVarTypes = contextVarTypes + getFLWORVarTypes(flwor)$
+     \STATE
+\ENDIF
+
+\RETURN $type$
+\end{algorithmic}
+\end{algorithm}
 

### 2011-11-10 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 The processing writes its result to a file in an XML representation.
-Instead of that, we need to keep the result in memory and pass it to consecutive steps of our algorithm.
+Instead, we need to keep the result in memory and pass it to consecutive steps of our algorithm.
 Though, this requirement concerns modifications of implementation while the core of the processing remains untouched.
@@
 
-Since we only modified the implementation of the output representation, the design of the syntax tree remains consequent to the entire design of the processing.
+Since we modified only the implementation of the output representation, the design of the syntax tree remains consequent to the entire design of the processing.
 Each node represents a logical component of a query.
@@
 
-\todo[inline]{priklad stromu}
+\todo[inline]{definicia (n-arny, usporiadany z lava do prava) a priklad stromu}
 \todo[inline]{niekam popisat celu moznu strukturu stromu - najskor do appendixu}
+
+\subsection{Vlastnosti stromu}
+\todo[inline]{Napr to, ze pri ceste z vrchola do listu len cez najlavejsie uzly sa nezvacsuje context premennych}
 
@@
 \subsection{Determination of Function Return Types}
-Determination of Return Types of Functions is needed because function calls can appear in expressions.
-However a return type of a function can be determined when the analysis of expressions encounters a call of the function, it involves multiple number of transitions of the syntax tree in a search for a definition of a particular function.
+Determination of return types of functions is needed because function calls can appear in expressions.
+However a return type of a function can be determined at the moment the analysis of expressions encounters a call of the function, it involves multiple number of transitions of the syntax tree in a search for a definition of a particular function.
 
@@
 \begin{algorithm}
-\caption{Determination of an Expression Type}
-\label{ALG_determination_of_an_expression_type}
+\caption{Function determineExpressionType}
+\label{ALG_func_determineExpressionType}
 \begin{algorithmic}[1]
@@
     $exprNode$: An expression node in the syntax tree. \\
-    $contextVarTypes$: Types of local variables valid in current context.
+    $contextVarTypes$: Types of local variables valid in the current subtree.
 }
@@
      \STATE TODO
+\ELSIF{$exprNode$ represents $FLWOR$ construct $flwor$}
+     \STATE  $type =$ type of an expression in the return clause of $flwor$
 \ENDIF
@@
 The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type than the type is noted.
-Otherwise, an attempt of the type deduction the expression is be made.
-The deduction of type from the expression is presented in \ref{ALG_determination_of_an_expression_type}.
+Otherwise, an attempt of the type deduction of the expression is made.
+The deduction of the type from the expression is presented in \ref{ALG_func_determineExpressionType}.
 On input, it takes an expression node and types of local variables that are valid in the expression.
@@
 Nevertheless, the principle is straightforward.
-Depending on the type of the expression, its type can be either determined directly or it depends on its subexpressions which types can be determined recursively.
+Depending on the type of the expression, its type can be either determined directly or it depends on its subexpressions.
+
+\begin{algorithm}
+\caption{Function analysisOfExpressionTypes}
+\label{ALG_func_analysisOfExpressionTypes}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $startingNode$: A node determinig a subtree to perform the analysis on \\
+    $contextVarTypes$: Types of local variables valid in the current subtree.
+}
+
+\FORALL{$subnode \in$ all subnodes of $startingNode$}
+    \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
+    \IF{$subnode$ defines new variables valid for following subnodes}
+        \STATE $contextVarTypes = merge(contextVarTypes, subnode.variableTypes)$
+    \ENDIF
+\ENDFOR
+\IF{$startingNode$ is expression node}
+    \STATE $startingNode.type = determineExpressionType(startingNode)$
+\ENDIF
+\IF{$startingNode$ (or the subtree) defines new variables valid in the supernode}
+    \STATE determine types of the variables
+    \STATE memorize their types in $startingNode.variableTypes$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+Function $determineExpressionType()$ is called within another function presented in \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
+Also, this function handles definitions of local variables, so, on request, their types are provided in the scope of their validity (in the subtree(s) which the variables are valid in).
+The function relies on the fact that definitions of new local variables can make them valid only in nodes with a higher sequence number when numbered in post-order \todo{overit}.
+Therefore, the left-most subnode of a certain node can be processed without an extension of the variable context.
+And, if the left-most subnode extends the variable context for the following nodes in the post-order numbering, this information can be processed as the type of the new variables can be directly determined since the binding expressions are already processed.
+The expressions are processed in post-order because of a possible need of currently processed expresion to know types of its subexpressions to determinde its own type.
+
+Function $analysisOfExpressionTypes$ called upon a binding expression of a global variable determines the type of the binding expression, and hence, the type of the variable.
 In the case of global variables, the argument $contextVarTypes$ is empty because there are no local variables valid in the prolog section.
-
-The types of processed global variables are available trough function $getGlobalVariableType()$.
-If the function gets a variable that is not a global one, it returns $null$.
-Function $getVariableType()$ called in \ref{ALG_determination_of_an_expression_type} is defined in \ref{ALG_function_get_variable_type}.
-It checks if the specified variable is amongst the given local variables.
-If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is $null$.
 
@@
 
+The types of processed global variables are available trough function $getGlobalVariableType()$.
+If the function gets a variable that is not a global one, it returns $null$.
+Function $getVariableType()$ called in \ref{ALG_func_determineExpressionType} is defined in \ref{ALG_function_get_variable_type}.
+It checks if the specified variable is amongst the given local variables.
+If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is $null$.
+
 \subsection{Determination of Expression Types}
 \todo[inline]{VYRIESIT: external variables, scope and visibility of variables - ako si udrziavat}
-
-\begin{algorithm}
-\caption{Function analysisOfExpressionTypes}
-\label{ALG_func_analysisOfExpressionTypes}
-\begin{algorithmic}[1]
-\REQUIRE{\ \\
-    $startingNode$: A node determinig a subtree to perform the analysis on \\
-    $contextVarTypes$: Types of local variables valid in the current subtree.
-}
-
-\FORALL{$subnode \in$ all subnodes of $startingNode$}
-    \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
-    \IF{$subnode$ defines new variables valid for following subnodes}
-        \STATE $newVarTypes = subnode.variableTypes$
-        \STATE TODO rozsir context
-    \ENDIF
-\ENDFOR
-\IF{$startingNode$ is expression node}
-    \STATE $startingNode.type = determineExpressionType(startingNode)$
-\ENDIF
-\IF{$startingNode$ defines new variables valid in the supernode}
-    \STATE TODO spracuj variables
-    \STATE memorize their types in $startingNode.variableTypes$
-\ENDIF
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Determination of an Expression Type in Query Body}
-\label{ALG_determination_of_an_expression_type_in_query_body}
-\begin{algorithmic}[1]
-\REQUIRE{\ \\
-    $exprNode$: An expression node in the syntax tree. \\
-    $contextVarTypes$: Types of local variables valid in the current context.
-}
-\ENSURE The expression type.
-
-\STATE $type = null$
-\IF{$exprNode$ represents a literal value}
-     \STATE $type =$ particular literal value type
-\ELSIF{$exprNode$ represents a call of function $f$}
-     \STATE $type = getFunctionReturnType(f)$
-\ELSIF{$exprNode$ represents variable reference $var$}
-     \STATE $type = getVariableType(contextVarTypes, var)$
-\ELSIF{$exprNode$ represents path expression $path$}
-     \STATE $type = TODO$
-\ELSIF{$exprNode$ represents operator $op$}
-     \STATE TODO
-\ELSIF{$exprNode$ represents $FLWOR$ construct $flwor$}
-     \STATE $contextVarTypes = contextVarTypes + getFLWORVarTypes(flwor)$
-     \STATE
-\ENDIF
-
-\RETURN $type$
-\end{algorithmic}
-\end{algorithm}
 

### 2011-11-11 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \section{Step 2: Static Analysis of Expression Types}
-In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree. This process can be divided into three substeps.
+\todo[inline]{Co su to za typy?}
+\todo[inline]{Preco vlastne tuto analyzu robime?}
+
+In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree.
+This process can be divided into three substeps.
+Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
 
@@
 For simplicity, we will focus on locally defined functions.
-We will not determine types of functions defined in other modules \todo{je to ok?}.
+We will not determine types of functions defined in other modules since the principle is similar but it is somewhat more difficult to implement \todo{je to ok?}.
 Return types of built-in functions are fixed, and thus, there is no need to analyze them.
@@
 If it is not it may be often deducible from the binding expression.
+Again, we do not analyse external variables for the same reason of simplicity as the external functions.
 
@@
 
+Function $determineExpressionType()$ is called within function $analysisOfExpressionTypes$ presented in \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
+Two important questions about this function emerge.
+How (in which order) are the tree nodes recursively processed and how does the function handle definitions (bindings) of new variables which may also appear in the expressions?
+
+The order of the recursion is post-order.
+The reason is that an expression node may need to know types of its subexpressions to determine its own type and we want to determine types of all expressions.
+Therefore, using the post-order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion because the types of the subexpressions are already determined.
+
+The handling of new variable definitions relies on the fact that the definitions make the new variables valid only in nodes with a higher sequence number when numbered in the post-order \todo{overit}.
+Therefore, the left-most subnode of a certain node can be processed without an extension of the variable context.
+And, if the left-most subnode extends the variable context for the following nodes in the post-order numbering, it can be easily handled because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
+Every node "knows" wheather or not it may define new variables for nodes with higher sequence numbers.
+This is expressed in the algorithm by the condition at line 10.
+This condition is double-checked at line 3 where it meaning is that every node even "knows" if its particular subnode may define new variables.
+An example is $FLWOR$ construct.
+The processing of $FLWOR$ knows that the first part (the first subnode) of this node may bind new variables using $for$ and $let$ constructs.
+After the subnode is processed, the processing of the $FLWOR$ node looks if any new variables were bound and if so the context of local variables is extended.
+This is done at line 4.
+Function $mergeContextVarTypes$ presented in \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into ist first argument (the variable context).
+If there records for variables with the same names in the context, they are overwritten to correspond to the variable overlaping.
+It is important to note that we assume the variable context is not passed by refence but by value, and hence, the variable context of a certain node is not affected by the recursive processing of its subnodes.
+
 \begin{algorithm}
@@
     \IF{$subnode$ defines new variables valid for following subnodes}
-        \STATE $contextVarTypes = merge(contextVarTypes, subnode.variableTypes)$
+        \STATE $contextVarTypes = mergeContextVarTypes(contextVarTypes, subnode.variableTypes)$
     \ENDIF
@@
 \ENDIF
-\IF{$startingNode$ (or the subtree) defines new variables valid in the supernode}
+\IF{$startingNode$ defines new variables valid in following nodes}
     \STATE determine types of the variables
@@
 
-Function $determineExpressionType()$ is called within another function presented in \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
-Also, this function handles definitions of local variables, so, on request, their types are provided in the scope of their validity (in the subtree(s) which the variables are valid in).
-The function relies on the fact that definitions of new local variables can make them valid only in nodes with a higher sequence number when numbered in post-order \todo{overit}.
-Therefore, the left-most subnode of a certain node can be processed without an extension of the variable context.
-And, if the left-most subnode extends the variable context for the following nodes in the post-order numbering, this information can be processed as the type of the new variables can be directly determined since the binding expressions are already processed.
-The expressions are processed in post-order because of a possible need of currently processed expresion to know types of its subexpressions to determinde its own type.
+\begin{algorithm}
+\caption{Function mergeContextVarTypes}
+\label{ALG_function_mergeContextVarTypes}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $contextVarTypes$: Types of local variables valid in current context. \\    
+    $extendingVarTypes$: Variable types of the same structure as the first argument and extending it.
+}
+\ENSURE The context variable types extended with the variable types from the second argument.
+
+\FORALL{$varName \in keys(extendingVarTypes)$}
+    \STATE $contextVarTypes[varName] = extendingVarTypes[varName]$
+\ENDFOR
+\RETURN $contextVarTypes$
+\end{algorithmic}
+\end{algorithm}
 
@@
 \subsection{Determination of Expression Types}
-\todo[inline]{VYRIESIT: external variables, scope and visibility of variables - ako si udrziavat}
+To determine types of expressions, we use already introduced function $analysisOfExpressionTypes$.
+The starting node (its first argument) is the node representing the query body and the variable context is empty as there cannot be any local variable valid in the body node.
+
+We can also determine expression types in functions.
+To do this for a certain function, $analysisOfExpressionTypes$ function has to be called with a function declaration node as the starting node.
+In this case, the function declaration node contains a subnode specifying function's formal arguments.
+These argument are set as the variable context for the function body represented by another subnode.
 

### 2011-11-11 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \section{Step 1: Construction of a Syntax Tree}
-The first step of the algorithm involves lexical and syntactic analyses and produces a syntax tree.
+The first step of the algorithm involves lexical and syntactic analyses known from compilers and produces a so-called syntax tree.
 The analyses are taken from Ji\v r\'{i} Schejbal's master thesis \todo{ref}.
-Since they are much more an issue of implementation than an issue of principle and they are not directly related to the inference, we will not describe them in this thesis.
+Since they are not directly related to the inference, and thus, they are not directly related to the topic of this thesis, we will not describe them in this thesis.
+Nevertheless, they provides us with a helpful processing of XQuery queries and we can focus on the inference.
 
-The mentioned thesis provides us with a helpful processing of XQuery queries, however, it needs to be slightly modified to suit our case.
-The processing writes its result to a file in an XML representation.
+The syntactic analysis needs to be slightly modified to suit our case.
+It writes its result into a file in an XML representation.
 Instead, we need to keep the result in memory and pass it to consecutive steps of our algorithm.
@@
 
-Since we modified only the implementation of the output representation, the design of the syntax tree remains consequent to the entire design of the processing.
+\subsection{Description of the Syntax Tree}
+Since we modified only the implementation of the output form, the design of the syntax tree remains consequent to the entire design of the analyses.
 Each node represents a logical component of a query.
 Inner nodes stand for XQuery constructs that are composed of other constructs and can be further divided.
-Examples of such constructs are FLWORs, if-than-else and function calls.
+Examples of such constructs are $FLWOR$, $if-than-else$ and function call.
 Leaf nodes represents elements of XQuery language that can be devided no more.
@@
 
-\todo[inline]{definicia (n-arny, usporiadany z lava do prava) a priklad stromu}
-\todo[inline]{niekam popisat celu moznu strukturu stromu - najskor do appendixu}
+In detail, the syntax tree is a n-ary tree representing syntactic structure of a given query.
+Every node represents a certain XQuery construct and that is a type of the node.
+Some types can be divided into more specific subtypes.
+This is shown \todo{obrazok typov uzlov}.
 
-\subsection{Vlastnosti stromu}
-\todo[inline]{Napr to, ze pri ceste z vrchola do listu len cez najlavejsie uzly sa nezvacsuje context premennych}
+An important characteric of the syntax tree is related to a definition of local variables and ther scope.
+The representation of a definition of a local variable is $VariableBinding$ node.
+Nodes of this type contain only two subnodes; a type of the variable and a binding expression.
+The node and the entire its subtree representing the type do not contain any expressions.
+The binding expression node is an expression node but the represented expression cannot use the variable which is just being bound.
+Therefore the scope of the new variable is not the subtree of the $VariableBinding$ node.
+Assuming the post-order numbering of the nodes, the new variable may be valid in some nodes with a higher post-order sequence number.
+Which nodes that are, it depends on the type of the ancestor node(s).
+
+For example, $FLWORExpr$ node contains four subnodes; $TupleStream$, $WhereClause$, $OrderByClause$ and $ReturnClause$.
+$TupleStream$ node contains a list of $VariableBinding$ node which define variables valid in all other three subnodes of the $FLWORExpr$ node.
+
+\todo[inline]{priklad stromu}
 
@@
 
-\section{Inference of Built-in Types}
+\section{Step 3: Inference of Built-in Types}
 Prejst kazdy vyraz a ak z neho vieme nieco odvodit, tak to urobime.
@@
 
-\begin{algorithm}
-\caption{Repair RW-XML tree}
-\label{propAlgo}
-\begin{algorithmic}[1]
-\REQUIRE $queryTrees$: A list of input query trees
-\ENSURE Uzly a ich typy nejakym sposobom
+In this step, the algorithm transitions through the syntax tree and tries to infer types of elements and attributes from expressions.
+This step could be merged with the step 2 where the syntax tree is being transitioned but for better comprehension we present it separately.
 
-\STATE initializacia nodeTypes
-\FORALL{$queryTree \in queryTrees$}
-    \STATE $partialNodeTypes = processQueryTree(queryTree)$
-    \STATE $nodeTypes = mergeNodeTypes(nodeTypes, partialNodeTypes)$
-\ENDFOR
-\RETURN $nodeTypes$
-\end{algorithmic}
-\end{algorithm}
+How are the types inferred from the expressions?
+Only expressions of a particular type are exploited.
+Specifically, the expression has to contain a subexpression representing an element or an attribute or a set of elements or attributes.
+In addition, we have to known what are the elements or the attributes, what their names are, how their were retrieved.
+Basic example of such expression is a path expression.
+Besides, the expression has to be either a function call or an arithmetic operation \todo{presnejsie} \todo{Nejake ine?}.
 
-\begin{algorithm}
-\caption{Process arithmetic comparison}
-\label{algorithm_process_arithmetic comparison}
-\begin{algorithmic}[1]
-\REQUIRE $expr$: An expression which is an arithmetic comparison
-\ENSURE Uzly a ich typy nejakym sposobom
-
-\STATE initializacia nodeTypes
-\RETURN $nodeTypes$
-
-\WHILE{$resultRXT \not \models_w \mathcal{FD}$}
-    \STATE $S = \emptyset$ \COMMENT Set of repair groups
-    \FORALL{$(F: S \rightarrow p) \in \mathcal{FD}$ s.t. $RXT \not \models_w F$}
-    	\FORALL{$t_1, t_2$ tuples of $RXT$ s.t. $t_1, t_2$ do not weakly satisfy $F$}
-		    \STATE $S = S \cup computeRepairGroup(F, t_1, t_2, RXT, S)$
-	    \ENDFOR
-    \ENDFOR
-
-    \STATE $R = getRepair(S, RXT)$
-    \STATE $resultRXT = applyRepair(R, resultRXT)$
-\ENDWHILE
-\end{algorithmic}
-\end{algorithm}
+\todo[inline]{Popisat odvodzovanie z volania funkcie}
+\todo[inline]{Popisat odvodzovanie z aritmetickej operacie}

### 2011-11-13 doc/theses/vitasek/chapter-2-research.tex
Thesis: ready to send to reviewers.

--- 2011-10-30 doc/theses/vitasek/chapter-2-research.tex
@@
-\chapter{Research}
+\chapter{Related Work}
 \label{chapter-research}
@@
 
-To the best of our knowledge, there are no other articles dealing directly with this problem.\\
+To the best of our knowledge, there are no other articles dealing with this problem.\\
 
-Fajt in \cite{fajt} introduces several algorithms to find XML keys (see \cite{keX}) in existing XML documents.
-Simple keys found this way can be translated to an equivalent ID attribute definition.
+Fajt in \cite{fajt} lists several algorithms to find XML keys (see \cite{keX}) in existing XML documents and introduces some new ones.
+XML keys found this way, being quite different from \texttt{ID} attributes, can under some circumstances (they have to be simple) be translated to an equivalent \texttt{ID} attribute definition.
 The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}.
-This opens a new line of research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to ID attributes.\\
+This opens a new line of possible research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to \texttt{ID} attributes.
+
+In our work we find the \texttt{ID} attributes directly. Even though we can convert them to XML keys by the process mentioned above, we are unable to find more complex keys this way.\\
 

### 2011-11-13 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress.
Thesis: ready to send to reviewers.

@@
 First, we transform the problem formulation to maximum weighted IS problem formulation.
-Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK.
+Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK \cite{glpk}.
 \nomenclature{GLPK}{GNU Linear Programming Kit}
@@
 
-GLPK \cite{glpk} is a multi-platform, multi-purpose solver well suited for this task.
-It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} (see \cite{land60a} for details) for MIP.
-An advantage of using \textit{Branch \& Bound} is that while traversing the \textit{search tree} it finds intermittent, sub-optimal solutions.
+GLPK is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} for MIP.
+
+\paragraph{\textit{Branch \& Bound}} This is an optimization method where the search space is systematically divided into smaller sub-spaces.
+This is the \textit{branching} component, and a so-called \textit{search tree} is built this way.
+Then, the sub-problems are recursively solved and whole branches of the search tree are discarded when it becomes obvious that the solution does not lie there.
+This is the \textit{bounding} component.
+See \cite{land60a} for detailed description.\\
+
+An advantage of using \textit{Branch \& Bound} is that while traversing the search tree it finds intermittent, sub-optimal solutions.
 It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\
@@
 
-\subsection{Example}
+\subsection*{Example}
 
@@
 
-First three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$.
+The first three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$.
 The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.
@@
 However, sometimes it takes too long to find the optimum (see e.g.
-Section \ref{section-glpk-comparison}), we should try to improve this process.
+Section \ref{section-glpk-comparison}), hence, the aim of this work is to improve this process.
 
@@
 \begin{define}[Heuristic]
-	Is an approach to problem solving based on prior experience, educated guess or common knowledge.
+	A \textit{heuristic} is an approach to problem solving based on prior experience, educated guess or common knowledge.
 \end{define}
@@
 \begin{define}[Heuristic Algorithm]
-	Is an algorithm that, in a reasonably short time, will generate a good, maybe even optimal solution to an optimization problem. However, it won’t provide any formal guarantee about its quality.
+	A \textit{heuristic algorithm} is one that, in a reasonably short time, generates a good, maybe even optimal solution to an optimization problem.
+However, it won’t provide any formal guarantee about its quality.
 \end{define}
@@
 
-An example of a heuristic is the commonly used approach of \textit{trial and error}: after a failed attempt, change something and try again. We will see many more heuristics later in this chapter.
+An example of a heuristic is the commonly used approach of \textit{trial and error}: after a failed attempt, change some parameter and try again.
+We will see many more heuristics later in this chapter.
 
@@
 \begin{define}[Solution Pool, Incumbent Solution]
-	\textit{Solution pool} (sometimes called pool of feasible solutions or feasible pool) is simply a set of solutions of different qualities that were found in one of the stages of the metaheuristic.
+	\textit{Solution pool} (sometimes called pool of feasible solutions or feasible pool) is a set of solutions of different qualities that were found in one of the stages of the metaheuristic.
 The solution(s) with the highest quality is (are) called the \textit{incumbent} solution(s).
@@
 Then, in a loop, we use this pool as input for \textit{improvement heuristics} (see Section \ref{section-mip-ihs}) and in turns ask them to improve it.
-All the time we check whether some \textit{termination criteria} are met, and if it is so, we terminate the metaheuristic.
+All the time we check whether \textit{termination criteria} are met, and if so, we terminate the metaheuristic.
 The incumbent solution from the last pool is then declared the Output ID set.\\
@@
 
+The algorithm works in two steps.
+First, all candidate AMs are grouped according to their types, and for each type the AM with the highest weight is selected.
+Second, all these AMs (now called $C'$) are traversed in order of their decreasing size.
+For each AM $m$, a set $S$ of all conflicting AMs from $C'$ is found and weights of both $m$ and $S$ are calculated.
+Then the weights are compared and either $m$ or $S$ is removed from $C'$.
+
 The pseudocode of this CH (taken from the original article with trivial modifications without changing the logic) is in Listing \ref{listing-ch-fidax}.
@@
 \FORALL{$m$ in $C'$}
-  \STATE $S \gets$ all mappings in $C$ whose images intersect $\iota(m)$
+  \STATE $S \gets$ all mappings in $C'$ whose images intersect $\iota(m)$
   \IF{$w(m) > \sum_{p \in S} w(p)$}
@@
 
-This trivial heuristics sorts all candidate AMs by their decreasing weights (Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible.
+This trivial heuristics sorts all candidate AMs by their decreasing weights (see Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible.
 This way it can create only one solution, and again, this is a hungry heuristic.
@@
 
-This construction heuristic will be called \heu{Glpk} from now on, for it is basically a time-constrained run of GLPK.
+This construction heuristic will be called \heu{Glpk} from now on.
+It is basically a time-constrained run of GLPK.
 Recall that GLPK uses a the \textit{Branch \& Bound} algorithm that produces feasible solutions even before the optimum is found.
@@
 We assume that an incumbent solution may already contain some AMs belonging to the optimal solution.
-We will take a random guess and fix some of these AMs: add new constraints to the MIP formulation fixing values of the respective variables to 1.
+We will take a random guess and fix some of these AMs, i.e.
+we add new constraints to the MIP formulation fixing values of the respective variables to 1.
 
@@
 First of all, from the set of all the attribute mappings in the model remove all the AMs contained in the ID set.
-This is because the specification of DTD/XSD does not allow an attribute to be ID and IDREF (IDREFS) at the same time.
+This is because the specification of DTD/XSD does not allow an attribute to be \texttt{ID} and \texttt{IDREF} (\texttt{IDREFS}) at the same time.
 Let us denominate these mappings as \textit{IDREF candidates} (obviously different from \textit{candidate AMs}).

### 2011-11-13 doc/theses/vitasek/appendix-a-jinfer.tex
Minor change - add space
Thesis: ready to send to reviewers.

@@
 
-This appendix will try to describe shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference.
+This appendix describes shortly yet comprehensively \textbf{jInfer} - the Java framework for XML schema inference, in which the algorithms described in this work were implemented.
 Please see project web \cite{jinferweb} for complete information, documentation and download options.
 
-jInfer was developed between 2009 and 2011 at Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mário Mikula, Ro\-bert Sme\-ta\-na, Michal Švirec and Matej Vitásek.
+jInfer was developed between 2009 and 2011 at the Charles University in Prague as a Software Project by team consisting of Michal Klempa, Mário Mikula, Ro\-bert Sme\-ta\-na, Michal Švirec and Matej Vitásek.
 The main idea was to create a structure in which all aspects of XML schema inference can be easily implemented and evaluated.

### 2011-11-13 doc/theses/vitasek/appendix-c-trace.tex
Thesis: in progress.
Thesis: ready to send to reviewers.

@@
 
-And finally, it is important to know what is the ID/IDREF set recommended by this experiment run - the reason why we do all this!
+And finally, it is important to know what is the ID/IDREF set recommended by this experiment run.
 Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
@@
 
-Note that in this example trace there were no IDREF AMs found. \qed
+Note that in this example trace there were no \texttt{IDREF} AMs found. \qed

### 2011-11-13 doc/theses/vitasek/bibliography.tex
Thesis: ready to send to reviewers.

\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{KMS{\etalchar{+}}11b}

\bibitem[Aho96]{ahonen}
H.~Ahonen.
\newblock {\em Generating Grammars for Structured Documents Using Grammatical Inference Methods}.
\newblock PhD thesis, Department of Computer Science, University of Helsinki, Series of Publications A, Report A-1996-4, 1996.

\bibitem[BDF{\etalchar{+}}01]{keX}
Peter Buneman, Susan Davidson, Wenfei Fan, Carmem Hara, and~Wang-Chiew Tan.
\newblock Keys for XML.
\newblock In {\em Proceedings of the 10th international conference on World Wide Web}, WWW '01, pages 201--210, New York, NY, USA, 2001. ACM.

\bibitem[BM03]{fidax}
Denilson Barbosa and Alberto Mendelzon.
\newblock Finding ID Attributes in XML Documents.
\newblock In Zohra Bellahsene, Akmal Chaudhri, Erhard Rahm, Michael Rys, and Rainer Unland, editors, {\em Database and~XML Technologies}, volume 2824 of~{\em Lecture Notes in Computer Science}, pages 180--194. Springer Berlin / Heidelberg, 2003.
\newblock 10.1007/978-3-540-39429-7\_12.

\bibitem[BNV08]{bex}
Geert~Jan Bex, Frank Neven, and Stijn Vansummeren.
\newblock Schemascope: a System for Inferring and Cleaning XML Schemas.
\newblock In {\em Proceedings of the 2008 ACM SIGMOD international conference on Management of data}, SIGMOD '08, pages 1259--1262, New York, NY, USA,
  2008. ACM.

\bibitem[BPM{\etalchar{+}}08]{Bray:08:EML}
Tim Bray, Jean Paoli, Eve Maler, Fran\c{c}ois Yergeau, and C.~M.
  Sperberg-McQueen.
\newblock Extensible Markup Language ({XML}) 1.0 (Fifth Edition).
\newblock {W3C} recommendation, W3C, November 2008.
\newblock \url{http://www.w3.org/TR/2008/REC-xml-20081126/}.

\bibitem[Dan98]{dantzig1998linear}
G.B. Dantzig.
\newblock {\em Linear Programming and Extensions}.
\newblock Landmarks in~Physics and~Mathematics. Princeton University Press,
  1998.

\bibitem[DC10]{heu-lecture}
G.~Di~Caro.
\newblock Heuristics Lecture Notes.
\newblock 2010.

\bibitem[DS04]{dorigo2004ant}
M.~Dorigo and T.~St{\"u}tzle.
\newblock {\em Ant Colony Optimization}.
\newblock Bradford Books. MIT Press, 2004.

\bibitem[Faj10]{fajt}
Stanislav Fajt.
\newblock Mining XML Integrity Constraints.
\newblock Master's thesis, Charles University in Prague, 2010.

\bibitem[FGK09]{Fomin:2009:MCA:1552285.1552286}
Fedor~V. Fomin, Fabrizio Grandoni, and Dieter Kratsch.
\newblock A Measure \& Conquer Approach for the Analysis of Exact Algorithms.
\newblock {\em J. ACM}, 56:25:1--25:32, August 2009.

\bibitem[GL97]{Glover:1997:TS:549765}
Fred Glover and Manuel Laguna.
\newblock {\em Tabu Search}.
\newblock Kluwer Academic Publishers, Norwell, MA, USA, 1997.

\bibitem[glp]{glpk}
GNU Linear Programming Kit.
\newblock \url{http://www.gnu.org/s/glpk/}.

\bibitem[gnu]{gnuplot}
Gnuplot, an Interactive Plotting Program.\\
\newblock \url{http://www.gnuplot.info/}.

\bibitem[Gol89]{goldberg1989genetic}
D.E. Goldberg.
\newblock {\em Genetic Algorithms in Search, Optimization, and Machine Learning}.
\newblock Artificial Intelligence. Addison-Wesley Pub. Co., 1989.

\bibitem[gra]{graphviz}
Graph Visualization Software.
\newblock \url{http://www.graphviz.org/}.

\bibitem[JR86]{JM1986425}
J.M and Robson.
\newblock Algorithms for Maximum Independent Sets.
\newblock {\em Journal of Algorithms}, 7(3):425 -- 440, 1986.

\bibitem[Kel]{gnuplot-manual}
Williams Kelley.
\newblock Gnuplot Manual.
\newblock \url{http://www.gnuplot.info/}.

\bibitem[Kle11]{anti}
Michal Klempa.
\newblock Optimization and Refinement of XML Schema Inference Approaches.
\newblock Master's thesis, Charles University in Prague, 2011.

\bibitem[KMS{\etalchar{+}}11a]{archdoc}
Michal Klempa, M\'{a}rio Mikula, Robert Smetana, Michal \v{S}virec, and Matej Vit\'{a}sek.
\newblock jInfer Architecture.\\
\newblock \url{http://jinfer.sourceforge.net/modules/architecture.pdf}, 2011.

\bibitem[KMS{\etalchar{+}}11b]{basiciggdoc}
Michal Klempa, M\'{a}rio Mikula, Robert Smetana, Michal \v{S}virec, and Matej Vit\'{a}sek.
\newblock jInfer BasicIGG Module Description.\\
\newblock \url{http://jinfer.sourceforge.net/modules/basicigg.pdf}, 2011.

\bibitem[KMS{\etalchar{+}}11c]{jinferweb}
Michal Klempa, M\'{a}rio Mikula, Robert Smetana, Michal \v{S}virec, and Matej Vit\'{a}sek.
\newblock jInfer: Java Framework for XML Schema Inference.\\
\newblock \url{http://jinfer.sourceforge.net}, 2011.

\bibitem[LD60]{land60a}
A.~H. Land and A.~G Doig.
\newblock An Automatic Method of Solving Discrete Programming Problems.
\newblock {\em Econometrica}, 28(3):497--520, 1960.

\bibitem[Mak]{mathprog}
Andrew Makhorin.
\newblock GNU Mathprog.
\newblock \url{http://lpsolve.sourceforge.net/5.5/MathProg.htm}.

\bibitem[McK66]{birthday}
E.~H. McKinney.
\newblock Generalized Birthday Problem.
\newblock {\em The American Mathematical Monthly}, 73(4):pp. 385--387, April
  1966.

\bibitem[Nor]{extendedcfg}
Theodore Norvell.
\newblock A Short Introduction to Regular Expressions and Context-Free Grammars.\\
\newblock \url{http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf}.

\bibitem[Vli02]{vlist2002xml}
E.V. Vlist.
\newblock {\em XML Schema}.
\newblock O'Reilly Series. O'Reilly, 2002.

\bibitem[VMP08]{1802522}
Ond\v{r}ej Vo\v{s}ta, Irena Ml\'{y}nkov\'{a}, and Jaroslav Pokorn\'{y}.
\newblock Even an Ant Can Create An XSD.
\newblock In {\em DASFAA'08: Proceedings of the 13th international conference on Database systems for advanced applications}, pages 35--50, Berlin, Heidelberg, 2008. Springer-Verlag.

\bibitem[\v{S}11]{sviro}
Michal \v{S}virec.
\newblock Efficient Detection of XML Integrity Constraints.
\newblock Master's thesis, Charles University in Prague, 2011.

\bibitem[Vyh09]{vyhnanovska}
Julie Vyhnanovsk\'{a}.
\newblock Automatic Construction of an XML Schema for a Given Set of XML Documents.
\newblock Master's thesis, Charles University in Prague, 2009.

\end{thebibliography}

### 2011-11-13 doc/theses/vitasek/chapter-0-preface.tex
Thesis: ready to send to reviewers.

@@
 
-Along with technologies such as SQL/noSQL databases, proprietary binary file formats, plain-text configuration files and JSON, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema to describe XML structure exist for a long time, most of the documents use outdated or no schema at all (\cite{1802522}).
+Along with technologies such as SQL/noSQL\footnote{noSQL: collection of non-relational database technologies, \url{http://nosql-database.org/}.} databases, proprietary binary file formats, plain-text configuration files and JSON\footnote{JSON: JavaScript Object Notation, lightweight data format, \url{http://www.json.org/}.}, XML is one of the leading formats for storing structured data.
+However, even though languages such as DTD and XML Schema\footnote{DTD and XML Schema: 2 most prominent XML schema languages, \cite{Bray:08:EML}} to describe XML structure exist for a long time, most of the documents use outdated or no schema at all \cite{1802522}.
 To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in \cite{ahonen, bex, vyhnanovska}.
@@
 Inference of this kind of structural information was greatly improved in \cite{anti}.\\
+
+\nomenclature{DTD}{Document Type Definition}
 
@@
 \texttt{ID} attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with \texttt{ID} attributes are much more common and it is much easier to obtain large data sets for experiments.
-% TODO something like even the search for ID sets is interesting enough?
 
-\section{Structure of the thesis}
+\section*{Structure of the thesis}
 
@@
 
-First, we will introduce a few notions required throughout the work, such as XML tree, \texttt{ID} attributes, ID sets, linear programming and the mixed integer problem.
+In Chapter \ref{chapter-definitions} we introduce a few notions required throughout the work, such as XML tree, \texttt{ID} attributes, ID sets, linear programming and the mixed integer problem.
 
-We will then review approaches to ID attribute search from previous articles on this topic and formulate the problem of finding the optimal ID set.
+Then in Chapter \ref{chapter-research}, we review approaches to ID attribute search from previous articles on this topic and formulate the problem of finding the optimal ID set.
 
-This will lead us to the NP-complete problem of Maximum Independent Set, where we will inspect the approaches to solving it.
+This will lead us to the NP-complete problem of Maximum Independent Set, where we will inspect the approaches to solving it in Chapter \ref{chapter-mip}.
 
@@
 
-An extension to jInfer for finding \texttt{ID} attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in the closing chapters.
+An extension to jInfer for finding \texttt{ID} attributes using MIP solver and a combination of heuristics will be presented and experimentally evaluated in Chapter \ref{chapter-experiments}.
 
-\section{Conventions}
+\section*{Conventions}
 

### 2011-11-13 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: ready to send to reviewers.

@@
 We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data.
+Parent nodes are connected to child nodes with the edges.
 This tree shall be called an \textit{XML tree}.
@@
 This example introduces an XML file fragment that will be used for demonstration throughout this work.
-XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}, where each node is annotated with a triple $label(v) : id(v) : value(v)$.
+An XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}(a), where each node is annotated with a triple $label(v) : id(v) : value(v)$.
 
@@
     \subfigure[XML tree]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree}}
-    \subfigure[Attribute mappings]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree-ams}}
+    \subfigure[Attribute Mappings]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree-ams}}
 \end{figure}
@@
 
-According to \cite{Bray:08:EML}, an XML attribute may have the type \texttt{ID}, \texttt{IDREF} or \texttt{IDREFS} (among others). Following constraints are related to these types.
+According to \cite{Bray:08:EML}, an XML attribute may have the type \texttt{ID}, \texttt{IDREF} or \texttt{IDREFS} (among others). The following constraints are related to these types.
 
@@
 
-An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED} or \texttt{\#REQUIRED}.
+An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED}\footnote{\texttt{\#IMPLIED} means that the attribute has a specified default value.} or \texttt{\#REQUIRED}\footnote{\texttt{\#REQUIRED} means that the attribute cannot be empty.}.
 
@@
 \begin{define}[$\Sigma^E$, $\Sigma^A$, $\Sigma$]
-	$\Sigma^E$ is the set of all element labels, $\Sigma^A$ is the set of all attribute labels.
-$\Sigma = \Sigma^E \cup \Sigma^A$ is their union and effectively the set of all labels in the document (simple text data nodes do not have a label).
+	$\Sigma^E$ is the set of all element labels, $\Sigma^A$ is the set of all attribute labels. $\Sigma = \Sigma^E \cup \Sigma^A$ is their union and effectively the set of all labels in the document.
 \end{define}
@@
 \paragraph{Example}
-The XML tree from Figure \ref{image-definitions-example-xml-tree} has the following non-empty AMs drawn in bold lines: $M_{y}^{a} = \{(2,6), (3,8)\}$, $M_{y}^{b} = \{(2,7)\}$, $M_{y}^{c} = \{(3,9)\}$ and $M_{z}^{a} = \{(5,10)\}$.
+The XML tree from Figure \ref{image-definitions-example-xml-tree}(b) has the following non-empty AMs drawn in bold lines: $M_{y}^{a} = \{(2,6), (3,8)\}$, $M_{y}^{b} = \{(2,7)\}$, $M_{y}^{c} = \{(3,9)\}$ and $M_{z}^{a} = \{(5,10)\}$.
 
-Following example equations hold.
+The following example equations hold.
 \begin{eqnarray*}
@@
 \paragraph{Example}
-Again referring to the XML tree from \ref{image-definitions-example-xml-tree}, we get the following AM images.
+Again referring to the XML tree from Figure \ref{image-definitions-example-xml-tree}, we get the following AM images.
 \begin{eqnarray*}
@@
 
-\subsection{\texttt{IDREF} and \texttt{IDREFS} Condition}
+\subsection*{\texttt{IDREF} and \texttt{IDREFS} Condition}
 
@@
 
-From the cited article:
 \begin{quote}
@@
 
-Again from the article:
 \begin{quote}
@@
 
-It is well known that finding the maximum weighted IS is an NP-hard optimization problem, see e.g. \cite{JM1986425}.
+It is well known that finding the maximum weighted IS is an NP-hard optimization problem, \cite{JM1986425}.
 
@@
 
-The problem of \textit{linear programming} is optimization of a linear function under a set of linear constraints.
-The formulation is usually called a \textit{linear program}.
+The problem of \textit{linear programming} is optimization of a linear function under a set of linear constraints. The formulation is usually called a \textit{linear program}.
+
 It can be written in the following form (minimization version is possible too).
@@
 
-Where $\mathbf{x}$ is the vector of variables (to be found by the optimization), $\mathbf{b}$ is the vector and $A$ is the accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function.
+Where $\mathbf{x}$ is the vector of variables (to be found by the optimization), $\mathbf{b}$ is the vector and $A$ its accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function.
 $\mathbf{x}$ and $\mathbf{c}$ have length $n$, $\mathbf{b}$ has length $m$ and $A$ has dimensions $m \times n$.
+Furthermore, \[\max_{x} f(x)\] means maximize the value of $f(x)$ by changing the value of $x$.\\
 
@@
 \max_{x} z = \sum_{i=1}^{n} c_i x_i \\
-s.t.\, a_{11}x_{1} + a_{12}x_{2} + \ldots + a_{1n}x_{n} & \leqslant & b_{1} \\
+s.t. & \\
+	   a_{11}x_{1} + a_{12}x_{2} + \ldots + a_{1n}x_{n} & \leqslant & b_{1} \\
 	   a_{21}x_{1} + a_{22}x_{2} + \ldots + a_{2n}x_{n} & \leqslant & b_{2} \\

### 2011-11-13 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: ready to send to reviewers.

@@
 \chapter{Experiments}
+\label{chapter-experiments}
 
@@
 From 3 different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
-Their summary is Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
+Their summary is in Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
 Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
@@
 None of these sets had any attributes before the conversion.
-Their summary is Table \ref{table-experiments-data-converted}, their graphs are in Figure \ref{image-experiments-data-converted}.
+Their summary is in Table \ref{table-experiments-data-converted}, their graphs are in Figure \ref{image-experiments-data-converted}.
 
@@
 After we built the model, we got the most complex graphs of 31 vertices and 47 edges (see Table \ref{table-experiments-data-realistic}).
-Our solution is to approach the problem from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
+We approach this problem from the other side: in the end, we will be solving the equivalent of IS problem on a graph created from XML data.
 We will create the XML data to contain a more complex graph with a specific number of vertices and edges.
 
-To demonstrate this, consider the following excerpt from an XML file:
+This data will represent very complex real world schemas with a lot of various attributes with similar values.
+
+Our artificial test data will look similar to the following excerpt from an XML file:
 
@@
 Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
-This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.
+This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.\\
+
+This way we can generate data more complex than anything found in the real world. It will however serve us to measure the performance of our algorithms in extreme cases.\\
 

### 2011-11-13 doc/theses/vitasek/diplomka.tex
Thesis: in progress.
Minor change - add space
comment out abbrs
Thesis: ready to send to reviewers.

@@
-\documentclass[a4paper,12pt,oneside]{report}
+\documentclass[a4paper,12pt,twoside]{report}
 \usepackage{graphicx}
@@
 \usepackage{wrapfig}
-\usepackage{appendix}
+\usepackage[title,titletoc]{appendix}
+%\usepackage{textgreek}
 \usepackage{nomencl}
@@
 \usepackage{multicol}
-\usepackage[pdfborder=0 0 0]{hyperref}
 \usepackage{amssymb}
@@
 \usepackage{complexity}
+%\usepackage{refcheck}
+\usepackage[pdfborder=0 0 0,pdftex,unicode]{hyperref}
+\DeclareFontShape{OT1}{cmtt}{bx}{n}{<5><6><7><8><9><10><10.95><12><14.4><17.28><20.74><24.88>cmttb10}{}
 
@@
 \linespread{1.5}
+%\onehalfspacing
 
@@
 
-% - make sure all chapter names are in First Caps
-% - make sure all references have a name, like Table 4.3.4 instead of just 4.3.4
-% - get rid of all short forms: won't -> will not
+% + make sure all chapter names are in First Caps
+% + make sure all references have a name, like Table 4.3.4 instead of just 4.3.4
+% + get rid of all short forms: won't -> will not
 % - make sure there are no singletons
 % - make sure LaTeX does not complaint much about under/overfullness
-% - make sure everything is defined and referenced
-% - make sure all the heu/data set names are consistent. This holds especially for GLPK (tool) / Glpk (heu) and FIDAX
+% + make sure everything is defined and referenced
+% + make sure there is nothing in the bibliography that is unreferenced
+% + make sure all the heu/data set names are consistent. This holds especially for GLPK (tool) / Glpk (heu) and FIDAX
 % - make sure that when referring to heus in algorithms, e.g. when describing how experimental sets are constructed, we correctly refer to their parameters as described in their pseudocodes
@@
 % - make sure labels have consistent names
-% - ratio -> FRACTION of XYZ to fix/remove/...
-% - what style of quotation marks to use?
+% + ratio -> FRACTION of XYZ to fix/remove/...
+% + what style of quotation marks to use? (`` and '')
 
@@
 
-Abstrakt: Tato práce navazuje na dřívější pokusy odvodit (inferovat) schému existujících XML dokumentů.
+Abstrakt: Tato práce navazuje na dřívější pokusy odvodit (inferovat) schéma existujících XML dokumentů.
 Jelikož je odvozování \textit{struktury} již relativně dobře popsáno, soustředíme se na integritní omezení.
 Několik jich popisujeme, pozornost pak soustředíme na ID/\.IDREF/\.IDREFS atributy z DTD.
-Na bázi článku od Barbosa a Menelzon stavíme heuristický přístup k problému hledání optimální sady ID atributů, jeho funkčnost a vhodnost pak ověřujeme na škále experimentů.
+Na bázi článku od Barbosa a Menelzon (2003) stavíme heuristický přístup k problému hledání optimální sady ID atributů, jeho funkčnost a vhodnost pak ověřujeme na škále experimentů.
 
@@
 
-Abstract: In this work we expand upon the previous efforts to extract (infer) schema information from existing XML documents.
-We find the inference of \textit{structure} to be sufficiently researched and focus further on integrity constraints.
-After briefly introducing some of them we turn our attention to ID/\.IDREF/\.IDREFS attributes in DTD.
-Building on the research by Barbosa and Menelzon we introduce a heuristic approach to the problem of finding optimal ID set.
-The approach is evaluated and tuned in a wide range of experiments.
+Abstract: In this work we~expand upon the previous efforts to~extract (infer) schema information from existing XML documents.
+We find the~inference of \textit{structure} to~be~sufficiently researched and focus further on~\textit{integrity constraints}.
+After briefly introducing some of~them we turn our attention to~ID/\.IDREF/\.IDREFS attributes in~DTD.
+Building on~the~research by~Barbosa and Menelzon (2003) we~introduce a~heuristic approach to~the~problem of~finding optimal ID set.
+The approach is~evaluated and tuned in~a~wide range of~experiments.
 
@@
   \include{chapter-0-preface}
-
   \include{chapter-1-definitions}
-
   \include{chapter-2-research}
-
   \include{chapter-3-mip}
-
   \include{chapter-4-1-experiments-setup}
-
   \include{chapter-4-2-experiments-results}
-
   \include{chapter-5-future}
-
   \include{chapter-6-conclusion}
 
-%%% Seznam použité literatury
 \newpage
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
+\phantomsection
+\addcontentsline{toc}{chapter}{Bibliography}
+%\nocite{*}
+%\bibliographystyle{alpha}
+%\bibliography{literature}
+	\include{bibliography}
 
+\newpage
+\phantomsection
+\addcontentsline{toc}{chapter}{Lists}
+
+\clearpage
+\phantomsection
+\addcontentsline{toc}{section}{List of Figures}
 \listoffigures
-\addcontentsline{toc}{chapter}{List of Figures}
 
+\clearpage
+\phantomsection
+\addcontentsline{toc}{section}{List of Algorithms}
 \listofalgorithms
-\addcontentsline{toc}{chapter}{List of Algorithms}
 
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
+\clearpage
+\phantomsection
+\addcontentsline{toc}{section}{List of Tables}
 \listoftables
-\addcontentsline{toc}{chapter}{List of Tables}
 
-%%% Použité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-
+\clearpage
+\phantomsection
+\addcontentsline{toc}{section}{List of Abbreviations}
+\label{chapter-list-abbreviations}
 \printnomenclature[2cm]
-\addcontentsline{toc}{chapter}{List of Abbreviations}
-\label{chapter-list-abbreviations}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Každá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-
-\appendix
 
 \openright
-\addcontentsline{toc}{chapter}{Appendices}
+\clearpage
+\phantomsection
+\addappheadtotoc
+\renewcommand{\appendixname}{Appendix}
 
-\include{appendix-a-jinfer}
-
+\begin{appendices}
+  \input{appendix-a-jinfer} % TODO include?
   \include{appendix-b-iss}
-
   \include{appendix-c-trace}
-
-% TODO a QED square
+\end{appendices}
 

### 2011-11-14 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: in progress.

@@
 Their summary is in Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
-Because the legal status of disclosing these data sets is unclear, we will refrain from identifying them beyond these artificial identifiers.
-The DVD distributed with this thesis will contain their anonymized version.
 
@@
 We used 2 data sets to convert, \dataset{MSH} and \dataset{NTH}.
-Unfortunately, the same problem with disclosure as in the previous case applies here, the DVD contains the anonymized version again.
 None of these sets had any attributes before the conversion.
@@
 This intuitively suggests that the maximum quality theoretically achievable is related to the $\frac{|E|}{|V|}$ density, not to the one we fixed.
-Exploration of this phenomenon is beyond the scope of this thesis.
+Exploration of this phenomenon is beyond the scope of this thesis.\\
 
-Note that the artificial data we used in experiments can be found on the DVD enclosed with this work.
+Note that all the data sets we used in experiments can be found on the DVD enclosed with this work.
 

### 2011-11-14 doc/theses/vitasek/diplomka.tex
Thesis: in progress.

@@
 The thanks goes again to Ms.
-Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others.
+Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HB}.
 
@@
 \hss}\hbox to 0.5\hsize{%
-signature
+
 \hss}}

### 2011-11-17 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \chapter{Proposed Algorithm}
-\todo[inline]{Nejaky uvod}
+\todo[inline]{Nejaky uvod a zkratke hrnutie}
 \todo[inline]{Lexikalna a syntakticka analyza z diplomky J. Schejbala}
 \todo[inline]{Staticka analyza typov vyrazov}
+\todo[inline]{Predpoklad korektnosti dotazov}
 
@@
 The analyses are taken from Ji\v r\'{i} Schejbal's master thesis \todo{ref}.
-Since they are not directly related to the inference, and thus, they are not directly related to the topic of this thesis, we will not describe them in this thesis.
+Since they are not directly related to the inference, and thus, they are not directly related to the topic of this thesis, we will not describe them.
 Nevertheless, they provides us with a helpful processing of XQuery queries and we can focus on the inference.
@@
 
-An important characteric of the syntax tree is related to a definition of local variables and ther scope.
+An important characteric of the syntax tree is related to a definition of local variables and their scope.
 The representation of a definition of a local variable is $VariableBinding$ node.
@@
 \section{Step 2: Static Analysis of Expression Types}
-\todo[inline]{Co su to za typy?}
-\todo[inline]{Preco vlastne tuto analyzu robime?}
+In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree.
+Information on the types of expressions can be used by consecutive steps of the algorithm.
+The consecutive steps will probably not use the determined types of all expressions, however this step may be useful in a future extending.
 
-In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree.
-This process can be divided into three substeps.
-Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
+Specifically, it is desired to determine the XML Schema built-in types \todo{link do appendixu} of expressions, if it is possible.
+Also, for each path expression we want to determine if it selects elements or attributes and we want to store the entire path in the type information.
+This information can be used for example in a $FLWOR$ expression \todo{priklad nejakeho FLWOR a ukazat ako vieme zistit, ktore elementy su prechadzane}.
+\todo[inline]{Uistit sa, ze takto nam to staci.}
+
+To capture sequences, we assign each type with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}. The cardinality expresses one of the following five sequence types.
+
+\begin{itemize}
+\item An empty sequence.
+\item A sequence of exactly one item.
+\item A sequence containing zero or one item.
+\item A sequence containing zero or more items.
+\item A sequence containing one or more items.
+\end{itemize}
+
+\todo[inline]{Popisat co robit a uviest priklad, kedy typ nevieme zistit. Napr porovnavanie hodnot elemetov vybranych z dat.}
+
+This process can be divided into three substeps. Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
 
@@
 Determination of return types of functions is needed because function calls can appear in expressions.
-However a return type of a function can be determined at the moment the analysis of expressions encounters a call of the function, it involves multiple number of transitions of the syntax tree in a search for a definition of a particular function.
+However a return type of a function can be determined at the moment the analysis of expressions encounters a call of the function, it involves multiple transitions of the syntax tree in a search for a definition of a particular function.
 
@@
 It returns a list of all function declaration nodes in the syntax three.
-Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in query prolog.
+Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.
 
@@
 Alike the functions, the global variables are defined in the prolog section.
-A type of a variable can be explicitly specified in the definition \todo{je to tak?
+A type of a variable can be explicitly specified in its definition \todo{je to tak?
 priklad}.
-If it is not it may be often deducible from the binding expression.
+If it is not it may often be deducible from the binding expression.
 Again, we do not analyse external variables for the same reason of simplicity as the external functions.
@@
 Nevertheless, the principle is straightforward.
-Depending on the type of the expression, its type can be either determined directly or it depends on its subexpressions.
+Depending on the class of the expression, its type can be either determined directly or it depends on its subexpressions.
+\todo{link na mozne triedy vyrazov.}
 
@@
 Function $mergeContextVarTypes$ presented in \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into ist first argument (the variable context).
-If there records for variables with the same names in the context, they are overwritten to correspond to the variable overlaping.
+If there are records for variables with the same names in the context, they are overwritten to correspond to the variable overlaping.
 It is important to note that we assume the variable context is not passed by refence but by value, and hence, the variable context of a certain node is not affected by the recursive processing of its subnodes.
@@
 \section{Step 3: Inference of Built-in Types}
-Prejst kazdy vyraz a ak z neho vieme nieco odvodit, tak to urobime.
+In this step, the algorithm goes through the syntax tree and tries to infer types of elements and attributes from expressions using the type information from the previous step.
+These two steps could be merged together but for better comprehension we present it separately.
 
-Odvodit nieco vieme ak:
-Nejaky podvyraz je path expression (alebo ine urcenie uzlu alebo hodnot uzlov \todo{toto domysliet}
-a zaroven cely vyraz
-1) je function call a podvyraz je jeho parameter a typ tohoto parametra vieme odvodit z hlavicky funkcie
-2) je aritmeticka operacia a podla typov operandov vieme urcit typ podvyrazu (tj napriklad sa uzly porovnavaju s konstantou 5) \todo{co vsetko?}
-3) dalsie konstrukcie ako cast as, type constructor, ... \todo{potrebuje riesit? princip je uplne rovnaky}
+How are the types inferred from the expression types?
+We do not exploit all expressions.
+Only expressions of a particular type are exploited.
+Specifically, an expression has to contain a subexpression representing a certain element or attribute or set of elements or attributes.
+This set is denoted $S$ in the following text.
+In other words, the type of the subexpression has to contain a path expression that selects the elements or attributes.
+This subexpression is denoted $E$.
+Another requirement is the expression has to be either a function call or an arithmetic operation \todo{Nejake ine ako napr type constructor, cast as?
+Princip je stale rovnaky}.
 
-In this step, the algorithm transitions through the syntax tree and tries to infer types of elements and attributes from expressions.
-This step could be merged with the step 2 where the syntax tree is being transitioned but for better comprehension we present it separately.
-
-How are the types inferred from the expressions?
-Only expressions of a particular type are exploited.
-Specifically, the expression has to contain a subexpression representing an element or an attribute or a set of elements or attributes.
-In addition, we have to known what are the elements or the attributes, what their names are, how their were retrieved.
-Basic example of such expression is a path expression.
-Besides, the expression has to be either a function call or an arithmetic operation \todo{presnejsie} \todo{Nejake ine?}.
+\todo[inline]{Zlucovanie odvodenych tvrdeni}
+\todo[inline]{Co ked mame path expression s predikatmi?}
 
 \todo[inline]{Popisat odvodzovanie z volania funkcie}
-\todo[inline]{Popisat odvodzovanie z aritmetickej operacie}
+\subsection{Function Calls}
+This case is quite simple.
+The algorithm encounters a function call and one of the arguments is a known set of elements of attributes (subexpression $E$ representing $S$).
+To determine the type of $S$, it is only needed to determine the type of the corresponding formal argument from the definition of the function.
+The function is either a built-in one so its definition is known or it is defined in the prolog section.
+External functions are not processed as was mentioned already.
+
+If the type $T$ of the formal argument is a built-in type or its sequence then $T$ is also the inferred type of $S$. Otherwise, no statement is inferred.
+
+\subsection{Arithmetic Operations}
+If the operator is one of $+$, $-$, $*$, $div$, $mod$ (the class of the expression is $Operator$ and the class of the operator is one of $PLUS$, $MINUS$, $MUL$, $DIV$, $MOD$ \todo{Niekam napisat mozne triedy operatora}), one operand is $E$ and the type $T$ of the other operand is one of numeric built-in types, then $T$ is also the inferred type of $S$.
+
+If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression is $Operator$ and the class of the operator is one of $GEN\_LESS\_THAN$, $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$, $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.

### 2011-11-17 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-Specifically, it is desired to determine the XML Schema built-in types \todo{link do appendixu} of expressions, if it is possible.
-Also, for each path expression we want to determine if it selects elements or attributes and we want to store the entire path in the type information.
-This information can be used for example in a $FLWOR$ expression \todo{priklad nejakeho FLWOR a ukazat ako vieme zistit, ktore elementy su prechadzane}.
-\todo[inline]{Uistit sa, ze takto nam to staci.}
+We want to be able to capture the following types.
 
-To capture sequences, we assign each type with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}. The cardinality expresses one of the following five sequence types.
+\begin{itemize}
+\item XML Schema built-in atomic types \todo{link do appendixu}.
+\item Type $xs:untyped$.
+\item Type representing any item. \todo{potrebujeme?}
+\item Type representing an element, attribute, item, node, text node, comment, processing instruction, document node.
+\item Type representing a set of elements or attributes selected by a certain path expression. The path expression is included in this type.
+\end{itemize}
+
+\todo[inline]{Akeho typu je napr //(elementA * 2)?}
+
+To capture sequences, we assign each type with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}.
+Every type can be perceived as a sequence.
+A type representing one value or one item can be perceived as a sequence of exactly one item.
+The cardinality expresses one of the following five sequence types.
 
@@
 \item A sequence of exactly one item.
-\item A sequence containing zero or one item.
-\item A sequence containing zero or more items.
-\item A sequence containing one or more items.
+\item A sequence containing zero or one item (modifier $?$).
+\item A sequence containing zero or more items (modifier $*$).
+\item A sequence containing one or more items (modifier $+$).
 \end{itemize}
@@
 This process can be divided into three substeps. Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
+
+\todo[inline]{Este by sa dala urobit analyza volania funkcii, do ktorych su predavane nejake elementy/atributy o ktorych vieme ich path expression.
+Vysledkom tejto funkcie by totiz zas mohli byt elementy/atributy o ktorych vieme (pozmenenu) path expression a nie len nezname elementy/atributy.}
 
@@
 
+\todo[inline]{Udrzat si referencie na definicie funkcii pre dalsie pouzitie.}
+
 \subsection{Determination of Global Variable Types}
@@
 \IF{$exprNode$ represents a literal value}
-     \STATE $type =$ particular literal value type
+     \STATE $type =$ particular literal value type, exactly-once cardinality
 \ELSIF{$exprNode$ represents a call of function $f$}
@@
 And, if the left-most subnode extends the variable context for the following nodes in the post-order numbering, it can be easily handled because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
-Every node "knows" wheather or not it may define new variables for nodes with higher sequence numbers.
+Every node ''knows'' wheather or not it may define new variables for nodes with higher sequence numbers.
 This is expressed in the algorithm by the condition at line 10.
-This condition is double-checked at line 3 where it meaning is that every node even "knows" if its particular subnode may define new variables.
+This condition is double-checked at line 3 where it meaning is that every node even ''knows'' if its particular subnode may define new variables.
 An example is $FLWOR$ construct.
@@
 
-\todo[inline]{Popisat odvodzovanie z volania funkcie}
 \subsection{Function Calls}

### 2011-11-19 doc/theses/vitasek/chapter-2-research.tex
Thesis: before merging reviews.

--- 2011-10-30 doc/theses/vitasek/chapter-2-research.tex
@@
 Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete.
-
 The transformation from IS problem formulation to $K$-\textsc{IDSet} problem formulation is as follows.
@@
 
-To the best of our knowledge, there are no other articles dealing directly with this problem.\\
+To the best of our knowledge, there are no other articles dealing with this problem.\\
 
-Fajt in \cite{fajt} introduces several algorithms to find XML keys (see \cite{keX}) in existing XML documents.
-Simple keys found this way can be translated to an equivalent ID attribute definition.
+\section{Finding XML Keys}
+
+XML keys are a structure somewhat similar to \texttt{ID} attributes, but with a much larger expressive strength.
+They have been introduced in \cite{keX} and implemented in XML Schema\footnote{\url{http://www.w3.org/TR/xmlschema11-1/\#Identity-constraint\_Definition\_details}}.
+
+Fajt in \cite{fajt} summarizes several algorithms to help find XML keys in existing data, namely \textit{Gordian}, \textit{XML Primary Keys}, \textit{SPIDER} and \textit{DBA Companion}.
+Except for \textit{XML Primary Keys}, they all are originally purposed to find keys in relational databases.
+We will describe them shortly.
+
+\subsubsection{\textit{Gordian}}
+
+This algorithm from \cite{fajt-41} extracts composite primary keys (PKs) from relational databases.
+
+\begin{quote}
+The idea behind is an observation that a projection of entities corresponds to a key if each counted aggregation for a projection is equal to 1.
+Thus, this method searches for all possible projections of a dataset while computing aggregations on the projected part of the set of entities.
+\end{quote}
+
+This is achieved by constructing a prefix tree from the tuples in the original relation, which is then pruned and traversed depth-first to find non-key attributes from which the primary keys are inferred.
+This algorithm still has to be adapted to search for PKs in XML data.
+
+\subsubsection{\textit{XML Primary Keys}}
+
+This is an algorithm from \cite{fajt-39} capable of finding simple keys and foreign keys directly in XML data.
+This is achieved by building a prefix tree containing all the XML nodes and then evaluating every path in it as a candidate key using metrics called \textit{support} and \textit{confidence}.
+To find more complex keys, the algorithm iteratively constructs candidate keys from simpler ones and evaluates them.\\
+
+The following two algorithms deal with \textit{inclusion dependencies} (INDs), described for example in \cite{fajt-12}.
+
+\subsubsection{\textit{SPIDER}}
+
+The core of this algorithm from \cite{fajt-51, fajt-53} is the following.
+
+\begin{quote}
+The process consists of two steps - sets of values are sorted during the first one and then all the candidates are analyzed in parallel.
+The core of the method is utilizing the data structure called min-heap which synchronizes the processing of all values of all attributes.
+\end{quote}
+
+It is possible to use a number of heuristic pruning strategies to keep the min-heap in a reasonable size.
+This algorithm performs very well for PKs in relational databases, however, it still has to be adapted for XML keys.
+
+\subsubsection{\textit{DBA Companion}}
+
+Like \textit{SPIDER}, this method from \cite{fajt-53} is able to find all the INDs in the database in just one pass.
+However, it uses a different data structure (basically a binary relation between the attributes and their corresponding values) and considers data types.
+Composite INDs are found using the simple ones and pruning the search space.
+According to the authors of \textit{SPIDER}, \textit{DBA Companion} is far inferior in performance.
+This algorithm has yet to be adapted to search for XML keys, too.
+
+\subsubsection{Fajt's Approach - \textit{KeyMiner}}
+
+Fajt introduces a new algorithm based on \textit{Gordian} and \textit{SPIDER} to look for primary and foreign keys in XML data.
+First, relations have to be extracted from the original XML document.
+Then all the primary keys are found using a modified \textit{Gordian} algorithm which can find absolute as well as relative PKs.
+Finally, \textit{SPIDER} is used to compute the foreign keys from the PKs found in the previous step.
+
+\subsection{Relation to \texttt{ID} Attributes}
+
+XML keys found this or any other way can under some circumstances (when they are simple enough) be translated to an equivalent \texttt{ID} attribute definition.
 The process is described in \cite[Ch.\ 9, s.\ 3]{vlist2002xml}.
-This opens a new line of research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to ID attributes.\\
+This opens a new line of possible research: finding XML keys using an algorithm modified to look only for \textit{useful} keys and then converting them to \texttt{ID} attributes.\\
 
-Maximum weigthed IS is a well researched topic with a lot of known direct or approximation algorithms, see e.g. \cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}.\\
+However, in our work we find \texttt{ID} attributes directly. And even though we can always convert them to XML keys by the process mentioned above, we are unable to find more complex keys this way.\\
+
+\section{Maximum Weighted IS}
+
+Maximum weigthed IS is a well researched topic with a lot of known direct or approximation algorithms, see e.g.
+\cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}.
+According to for example \cite{Paschos:1997:SAO:254180.254190}, the best known approximation algorithm for weighted IS to-date achieves an approximation ratio of $3(\Delta + 2)$, where $\Delta$ is the maximum degree of a vertex in the IS graph.
+This article lists several algorithms similar to those we introduce in the following chapters.\\

### 2011-11-19 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: before merging reviews.

@@
 
-Furthermore, it is always good to ignore the simple text data nodes, as it will improve the total run time.
+Furthermore, it is always good to ignore the simple text data nodes, as it will improve the total run time.\\
+
+Now compare this to the performance of the original algorithm from \cite{fidax} that we have implemented as \heu{FIDAX}.
+We cannot use it to find the optimum, even if we do have enough time: it is hungry, deterministic and always gives us only one solution.
+And even if time is constrained, we have shown in Section \ref{section-experiments-random-fuzzy-fidax} that even trivial algorithms such as \heu{Random} or \heu{Fuzzy} very often outperform \heu{FIDAX}.
+And \heu{Glpk} constrained to 1 second run time outperforms \heu{FIDAX} on every data set.\\

### 2011-11-19 doc/theses/vitasek/chapter-6-conclusion.tex
Thesis: before merging reviews.

@@
 
-To the best of our knowledge, at the time of writing this work is our approach to finding \texttt{ID} attributes the best one known.
+To the best of our knowledge, at the time of writing this work is our approach to finding \texttt{ID} attributes the best one known.\\
+
+The wisdom found in the experiments in this work might be the following.
+While it is important to be able to write a heuristic algorithm tailored to the specific problem being solved, such as the authors of \cite{fidax} did, it should be noted that sometimes it is better to solve a more generalized problem.
+In this case the transformation to MIP formulation and using a dedicated solver proved to produce better results in shorter time.

### 2011-11-19 doc/theses/vitasek/diplomka.tex
Thesis: before merging reviews.

@@
 The thanks goes again to Ms.
-Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HB}.
+Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HV}.
 
@@
 \addcontentsline{toc}{chapter}{Bibliography}
-%\nocite{*}
-%\bibliographystyle{alpha}
-%\bibliography{literature}
-	\include{bibliography}
+\nocite{*}
+\bibliographystyle{alpha}
+\bibliography{literature}
+%	\include{bibliography}
 

### 2011-11-19 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \item Type $xs:untyped$.
-\item Type representing any item. \todo{potrebujeme?}
+\item Type representing any item. \todo{je potrebne?}
 \item Type representing an element, attribute, item, node, text node, comment, processing instruction, document node.
-\item Type representing a set of elements or attributes selected by a certain path expression. The path expression is included in this type.
+\item Type representing a set of elements or attributes selected by a certain path expression. The path expression is included in this type. Let this type be identified as $pathType$
 \end{itemize}
 
-\todo[inline]{Akeho typu je napr //(elementA * 2)?}
+\todo[inline]{Akeho typu je napr path expression /root/(elementA * 2)?}
 
@@
 
-\todo[inline]{Popisat co robit a uviest priklad, kedy typ nevieme zistit. Napr porovnavanie hodnot elemetov vybranych z dat.}
+If we cannot determine the type of an expression we express that using $null$. Example of such case is path expression $/rootElement/(elementA * 2)$. \todo{Nejak dopisat}
 
-This process can be divided into three substeps. Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
+The analysis of expression types can be divided into three substeps.
+Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
 
@@
 
-Instead, the syntax tree can be searched just once, before the analysis of expressions, and return types of all functions found will be stored within an associative array.
-On request, this array will retrieve the return type of a specified function.
+Instead, the syntax tree can be searched just once, before the analysis of expressions, and return types of all functions found are stored.
 
@@
 
-\todo[inline]{Udrzat si referencie na definicie funkcii pre dalsie pouzitie.}
+Later phases of the algorithm may need to process the function declarations. Therefore, we also store references to the entire function declaration nodes.
 
@@
 \ELSIF{$exprNode$ represents path expression $path$}
-     \STATE $type = TODO$
+     \IF{$path$ selects an element set}
+         \STATE $type =$ element type, zero-or-more cardianlity, $path$ included
+     \ELSIF{$path$ selects an attribute set}
+         \STATE $type =$ attribute type, zero-or-more cardianlity, $path$ included
+     \ELSE
+         \STATE $type = null$
+     \ENDIF
 \ELSIF{$exprNode$ represents operator $op$}
-     \STATE TODO
+     \IF{$op$ is a comparison operator}
+         \STATE $type = xs:boolean$
+     \ELSIF{$op$ is an additive operator $AND$ one of the operands is of a numeric type}
+         \IF{both operands are of numeric types}
+             \STATE $type =$ the numberic type which is more or equally general than the other
+         \ELSE
+             \STATE $type =$ type of the numeric operand
+         \ENDIF
+     \ENDIF
 \ELSIF{$exprNode$ represents $FLWOR$ construct $flwor$}
@@
 The deduction of the type from the expression is presented in \ref{ALG_func_determineExpressionType}.
-On input, it takes an expression node and types of local variables that are valid in the expression.
+On input, it takes an expression node and types of local variables that are valid in the expression (also called context of variables or variable context).
 The presented algorithm is just shortened illustration since the complete version is too long to be presented in this text.
@@
 Only expressions of a particular type are exploited.
-Specifically, an expression has to contain a subexpression representing a certain element or attribute or set of elements or attributes.
-This set is denoted $S$ in the following text.
-In other words, the type of the subexpression has to contain a path expression that selects the elements or attributes.
-This subexpression is denoted $E$.
+Specifically, an expression has to contain some $pathType$ subexpression $E$ (expression representing a certain element or attribute or set of elements or attributes).
+In the following text, the set represented by expression $E$ is denoted $S$.
 Another requirement is the expression has to be either a function call or an arithmetic operation \todo{Nejake ine ako napr type constructor, cast as?

### 2011-11-19 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: merging Lydka round 1.

@@
 \begin{define}[Node equality]
-	$v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
+	Nodes $v_1$ and $v_2$ are \textit{node equal}, written $v_1 =_n v_2$ iff $id(v_1) = id(v_2)$.
 \end{define}
@@
 \begin{define}[Value equality]
-	$v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
+	Nodes $v_1$ and $v_2$ are \textit{value equal}, written $v_1 =_v v_2$ iff $value(v_1) = value(v_2)$.
 \end{define}
@@
 
-%\section{XML Keys} - is this necessary at all?
-
-%\subsection{According to Specification}
-
-% - compare them to ID attributes
-
-%\subsection{According to \cite{keX}}
-
-% - show how a XML key is defined using the keX notation % this is a terrible sentence
-
 \section{Attribute Mappings}
@@
 	For $x \in \Sigma^E$ and $y \in \Sigma^A$ we define the \textit{attribute mapping} of y over x, denoted $M_{x}^{y}$, the $\mathcal{I} \times \mathcal{I}$ relation defined by
-	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}\]
+	\[M_{x}^{y} = \{ (z,w): label(z) = x, label(w) = y, parent(w) = z \}.\]
 \end{define}
@@
 \begin{define}[Image of the attribute mapping]
-	\textit{Image} $\iota$ of the attribute mapping $M_{x}^{y}$ is defined as $\iota(M_{x}^{y}) = \{z: z = value(w), w \in \pi_A(M_{x}^{y})\}$
+	\textit{Image} $\iota$ of the attribute mapping $M_{x}^{y}$ is defined as $\iota(M_{x}^{y}) = \{z: z = value(w), w \in \pi_A(M_{x}^{y})\}$.
 \end{define}
@@
 \begin{define}[Candidate attribute mapping]
-An attribute mapping $m$ is a \textit{candidate attribute mapping} if it is an injective function, that is,
-\[|m| = |\pi_E(m)| = |\pi_A(m)| = |\iota(m)|\]
+An attribute mapping $m$ is a~\textit{candidate attribute mapping} if it is an injective function, that is,
+\[|m| = |\pi_E(m)| = |\pi_A(m)| = |\iota(m)|.\]
 \end{define}
@@
 A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an \textit{ID set} iff
-\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset\]
+\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset.\]
 \end{define}
@@
 
-Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}.
-\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i)\]
+Given an ID set $I$, the requirements from Section \ref{section-definitions-id-attributes} give us the following condition for an attribute mapping $m$ to be marked \texttt{IDREF}:
+\[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i).\]
 Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.
@@
 
-This definition of weight for AMs or AM sets comes from \cite{fidax} again.
-
-Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.
+This definition of weight for AMs or AM sets comes from \cite{fidax} again. Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.
 
@@
 \begin{define}[Support]
-\textit{Support} of an attribute mapping $m$ is defined as follows.
-\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}\]
+\textit{Support} of an attribute mapping $m$ is defined as follows:
+\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}.\]
 \end{define}
@@
 \begin{define}[Coverage]
-\textit{Coverage} of an attribute mapping $m$ is defined as follows.
-\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|\]
+\textit{Coverage} of an attribute mapping $m$ is defined as follows:
+\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|.\]
 \end{define}
@@
 \begin{define}[Weight]
-For $\alpha, \beta \geq 0$ as relative priorities of support and coverage we define the AM \textit{weight} as follows.
-\[weight(m) = \alpha . \phi(m) + \beta . \chi(m)\]
-For a set of AMs (which may or may not be an ID set) $S = \{m_1, \ldots m_i\}$ we define the weight of this set as the sum of the weights of its AMs.
-\[weight(S) = \sum_{m \in S} weight(m)\]
+For $\alpha, \beta \geq 0$ as relative priorities of support and coverage we define the AM \textit{weight} as follows:
+\[weight(m) = \alpha . \phi(m) + \beta . \chi(m).\]
+For a set of AMs (which may or may not be an ID set) $S = \{m_1, \ldots m_i\}$ we define the weight of this set as the sum of the weights of its AMs:
+\[weight(S) = \sum_{m \in S} weight(m).\]
 \end{define}
@@
 \begin{define}[Independent set]
-Given an undirected graph $G = (V, E)$, a set of vertices $I \subseteq V$ is an \textit{independent set}, iff the following holds.
-\[\forall v_1, v_2 \in I, v_1 \neq v_2: (v_1, v_2) \notin E \]
+Given an undirected graph $G = (V, E)$, a set of vertices $I \subseteq V$ is an \textit{independent set}, iff
+\[\forall v_1, v_2 \in I, v_1 \neq v_2: (v_1, v_2) \notin E.\]
 \end{define}
@@
 \begin{define}[Maximum weighted independent set]
-Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied.
-\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v)\]
+Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied:
+\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v).\]
 \end{define}
@@
 
-It can be written in the following form (minimization version is possible too).
+It can be written in the following form:
 
@@
 b_m \\
-\end{bmatrix}
+\end{bmatrix},
 \]
+
+where a minimization version is possible, too.
 
@@
 
-Another way to write this formulation is in the following.
+Another way to write this formulation is this:
 
@@
 \begin{define}[Mixed integer problem]
-	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some or all variables are limited to integral or boolean (0, 1) values.
+	MIP, or \textit{mixed integer problem}, is an~instance of~linear programming in~which some or~all variables are limited to~integral or~boolean (0, 1) values.
 \end{define}

### 2011-11-19 doc/theses/vitasek/chapter-2-research.tex
Thesis: merging Lydka round 1.

@@
 Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete.
-The transformation from IS problem formulation to $K$-\textsc{IDSet} problem formulation is as follows.
+The transformation from IS problem formulation to~$K$-\textsc{IDSet} problem formulation is as follows.
 
@@
 
-The article continues by proving that the problem of finding maximum weighted IS can be reduced to the problem of finding an ID set with maximum weight (\textsc{Max-IDSet}).
+The article continues by proving that finding the maximum weighted IS can be reduced to the problem of finding an ID set with maximum weight (\textsc{Max-IDSet}).
 This again means that \textsc{Max-IDSet} is NP-complete and, furthermore, unless $\P = \NP$, \textsc{Max-IDSet} has no constant factor approximation algorithm.
@@
 Fajt in \cite{fajt} summarizes several algorithms to help find XML keys in existing data, namely \textit{Gordian}, \textit{XML Primary Keys}, \textit{SPIDER} and \textit{DBA Companion}.
-Except for \textit{XML Primary Keys}, they all are originally purposed to find keys in relational databases.
+Except for \textit{XML Primary Keys}, they all are originally purposed to find keys in~relational databases.
 We will describe them shortly.
@@
 This is achieved by constructing a prefix tree from the tuples in the original relation, which is then pruned and traversed depth-first to find non-key attributes from which the primary keys are inferred.
-This algorithm still has to be adapted to search for PKs in XML data.
+This algorithm still has to~be~adapted to~search for PKs in XML data.
 
@@
 
-It is possible to use a number of heuristic pruning strategies to keep the min-heap in a reasonable size.
-This algorithm performs very well for PKs in relational databases, however, it still has to be adapted for XML keys.
+It is~possible to~use a~number of~heuristic pruning strategies to~keep the~min-heap in~a~reasonable size.
+This algorithm performs very well for PKs in~relational databases, however, it~still has to~be~adapted for~XML keys.
 
@@
 
-Like \textit{SPIDER}, this method from \cite{fajt-53} is able to find all the INDs in the database in just one pass.
-However, it uses a different data structure (basically a binary relation between the attributes and their corresponding values) and considers data types.
-Composite INDs are found using the simple ones and pruning the search space.
-According to the authors of \textit{SPIDER}, \textit{DBA Companion} is far inferior in performance.
-This algorithm has yet to be adapted to search for XML keys, too.
+Like \textit{SPIDER}, this method from \cite{fajt-53} is able to find all the~INDs in~the~da\-ta\-base in~just one pass.
+However, it~uses a~different data structure (basically a~binary relation between the~attributes and~their corresponding values) and~considers data types.
+Composite INDs are~found using the~simple ones and~pruning the~search space.
+According to~the~authors of~\textit{SPIDER}, \textit{DBA Companion} is~far inferior in~performance.
+This algorithm has yet to~be~adapted to~search for XML keys, too.
 

### 2011-11-19 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: merging Lydka round 1.

@@
 
-To interpret the data: \dataset{OVA*} sets have quite interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical'' representants.
+To interpret the data: \dataset{OVA*} sets have interesting and challenging graphs, but they are relatively small. We can consider them to be the ``typical'' representants.
 
@@
 First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an AM, they need an attribute \texttt{attr}, but with large enough random values, so that they do not conflict with others.
-Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of their \texttt{attr}.
+To constitute an~AM, they need an attribute \texttt{attr}, but with large enough random values, so that they do not conflict with others.
+Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of~their \texttt{attr}.
 This will ensure they cannot share the same ID set, thus effectively creating the edge in the graph representation.\\
@@
 The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the fraction of edges that are to all edges that could be in a complete graph with $|V|$ vertices.
+This is the density as the fraction of edges that are to all edges that could be in a~complete graph with $|V|$ vertices.
 
@@
 This is done by a linear scan and retrieving a so-called \textit{flat} representation.
-This consists of a list of tuples in the following format.
+This consists of a~list of tuples in the following format.
 

### 2011-11-19 doc/theses/vitasek/chapter-3-mip.tex
Thesis: merging Lydka round 1.

@@
 
-Given a graph $G = (V,E)$ with a weight function $w: V \rightarrow \mathbb{R}$, we introduce a binary variable $x_i$ for each vertex $v_i \in V$ and an inequality constraint $x_i + x_j \leq 1$ for each edge $e = (v_i, v_j) \in E$.
-Furthermore we introduce an objective function in form $\sum_{x_i} x_i w(v_i)$.
+Given a~graph $G = (V,E)$ with a~weight function $w: V \rightarrow \mathbb{R}$, we introduce a~binary variable $x_i$ for each vertex $v_i \in V$ and an~inequality constraint $x_i + x_j \leq 1$ for each edge $e = (v_i, v_j) \in E$.
+Furthermore we introduce an~objective function in~form $\sum_{x_i} x_i w(v_i)$.
 
-It is obvious that the objective function and all the constraints consitute a MIP instance, and that solving it finds the maximum weigthed IS in $G$.
+It is obvious that the objective function and all the constraints consitute a~MIP instance, and that solving it finds the maximum weigthed IS in $G$.
 
@@
 Corresponding vertices in the IS formulation will be $V = \{v_{y-a}, v_{y-b}, v_{y-c}, v_{z-a}\}$.
-Edges in the IS formulation will be the following.
+Edges in the IS formulation will be the following ones:
 
 \begin{eqnarray*}
-(v_{y-a},v_{y-b}) \\
-(v_{y-a},v_{y-c}) \\
-(v_{y-b},v_{y-c}) \\
-(v_{y-a},v_{z-a}) \\
+(v_{y-a},v_{y-b}), \\
+(v_{y-a},v_{y-c}), \\
+(v_{y-b},v_{y-c}), \\
+(v_{y-a},v_{z-a}). \\
 \end{eqnarray*}
@@
 
-The problem now is, recall, to solve the following.
+The problem now is, recall, to solve the following:
 
@@
 \max_{x} z & = & \mathbf{c}^{\mathrm{T}}\mathbf{x} \\
-s.t.\, A\mathbf{x} & \leqslant & \mathbf{b} \\
+s.t.\, A\mathbf{x} & \leqslant & \mathbf{b}. \\
 \end{eqnarray*}
 
-In GLPK \textit{MathProg} language (see \cite{mathprog}), this translates to the following formulation.
+In GLPK \textit{MathProg} language \cite{mathprog}, this translates to the following.
 
@@
 This construction heuristic will be called \heu{Glpk} from now on.
-It is basically a time-constrained run of GLPK.
+It is basically a~time-constrained run of GLPK.
 Recall that GLPK uses a the \textit{Branch \& Bound} algorithm that produces feasible solutions even before the optimum is found.

### 2011-11-19 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: merging Lydka round 1.

@@
 Our experimental set will contain 500 experimental configurations for each of these two GLPK versions.
-Every configuration will use \heu{Glpk} CH set to a time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Algorithm \ref{listing-experiment-glpk-native-vs-cygwin-1}).
+Every configuration will use \heu{Glpk} CH set to a~time limit from 1 to 46 seconds with increments of 5, meaning 10 settings * 50 iterations = 500 configurations in total (see Algorithm \ref{listing-experiment-glpk-native-vs-cygwin-1}).
 There will be no improvement heuristic.
@@
 The experimental results are summarized in Table \ref{table-experiments-fidax-and-hungry} and are quite surprising.
-As trivial a heuristic \heu{Hungry} is, it is still able to improve the ID set found by \heu{FIDAX} by as much as almost 50\% (the last row, \dataset{100-1000}).
+Bold cells mean that adding \heu{Hungry} helped achieve a better quality.
+However trivial \heu{Hungry} is, it is still able to improve the ID set found by \heu{FIDAX} by as much as almost 50\% (the last row, \dataset{100-1000}).
 
 Table \ref{table-experiments-fidax-and-hungry-idsets} lists the \texttt{ID} attributes found in both cases for this most extreme input, \dataset{100-1000}.
-Note that the content of each cell means ``attribute \texttt{attr} in element \texttt{vertexXY} should be marked as \texttt{ID} attribute''.
+Note that the content of each cell means ``attribute \texttt{attr} in~element \texttt{vertexXY} should be marked as \texttt{ID} attribute''.
 
@@
 See the Algorithm \ref{listing-experiment-best-standalone-ch} for details.
-This time we are not interested in run times, only in qualities which we shall gather in a format for GnuPlot.\\
+This time we are not interested in run times, only in qualities which we shall gather in format for GnuPlot.\\
 
@@
 This experimental set will contain 6 data sets * 50 iterations * 4 IHs = 1200 experimental configurations.
-Note that we are using only the most challenging data sets, as the combination of \heu{Glpk} as CH and any other IH is already an overkill for easier data sets.\\
+Note that we are using only the most challenging data sets, as the combination of \heu{Glpk} as CH and any other IH is already an~overkill for easier data sets.\\
 
@@
 \begin{itemize}
-	\item Every $(1 + 5k)$\textsuperscript{th} and $(3 + 5k)$\textsuperscript{th} step is a \heu{RandomRemove}, and each time this happens there is a rather sharp drop in quality
-	\item Every $(2 + 5k)$\textsuperscript{th} step is a \heu{Mutation}, and there is a consistent increase in quality each time.
-	\item Every $(4 + 5k)$\textsuperscript{th} step is a \heu{Crossover}, and each time it happens there is a consistent increase, yet smaller than with \heu{Mutation}.
-	\item Every $5k$\textsuperscript{th} step is a \heu{RemoveWorst}, and as expected, this removes the worst solution not touching the best ones that decide the incumbent quality.
-The line thus stays flat every time it happens.
+	\item Every $(1 + 5k)$\textsuperscript{th} and $(3 + 5k)$\textsuperscript{th} step is~a~\heu{RandomRemove}, and each time this happens there is~a~rather sharp drop in~quality.
+	\item Every $(2 + 5k)$\textsuperscript{th} step is a~\heu{Mutation}, and there is~a~consistent increase in~quality each time.
+	\item Every $(4 + 5k)$\textsuperscript{th} step is a~\heu{Crossover}, and each time it~happens there is~a~consistent increase, yet smaller than with \heu{Mutation}.
+	\item Every $5k$\textsuperscript{th} step is~a~\heu{RemoveWorst}, and as~expected, this removes the worst solution not touching the best ones that decide the incumbent quality.
+The line thus stays flat every time it~happens.
 \end{itemize}

### 2011-11-19 doc/theses/vitasek/chapter-5-future.tex
Thesis: ready to send to reviewers.
Thesis: merging Lydka round 1.

@@
 A straightforward extension granting the ability to handle more than one input XML file has already been suggested in \cite{fidax}.
-However, it wasn't implemented in this work either, so it still remains an obvious first choice of future work.\\
+However, it was not implemented in this work either, so it remains an obvious first choice of future work.\\
 
-It is possible (and easy) to add more construction and improvement heuristics, as well as more metaheuristics in which the existing IHs are chained.
-A starting point is in Section \ref{section-appendix-iss-howto}.\\
+It is possible (and easy) to add more construction and improvement heuristics, as well as more metaheuristics using the existing IHs. A~starting point is in Section \ref{section-appendix-iss-howto}.\\
 
-As it was mentioned in Section \ref{section-mip-ihs}, the combination of \heu{Crossover}, \heu{Mutation} and \heu{RemoveWorst} can be seen as a sort of genetic programming.
+As it was mentioned in Section \ref{section-mip-ihs}, the combination of \heu{Crossover}, \heu{Mu\-ta\-tion} and \heu{RemoveWorst} can be seen as a sort of genetic programming.
 However some modifications would still be necessary to make it a real instance of genetic algorithm metaheuristic.

### 2011-11-19 doc/theses/vitasek/diplomka.tex
Thesis: merging Lydka round 1.

@@
 \noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
+I declare that I~carried out this master thesis independently, and only with the~cited sources, literature and~other professional sources.
 
 \medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
+I~understand that my~work relates to~the~rights and~obligations under the~Act No.~121/2000 Coll., the~Copyright Act, as~amended, in~particular the~fact that the~Charles University in~Prague has the~right to~conclude a~license agreement on~the~use of this work as~a~school work pursuant to~Section 60 paragraph 1 of~the~Copyright Act.
 
@@
 
-Název práce: Odvozování integritních omezení v XML % TODO how to translate this correctly?
+Název práce: Odvozování integritních omezení v XML
 
@@
 
-Vedoucí diplomové práce: \mfadvisor{} % TODO does this have to be here?, Ka\-ted\-ra soft\-wa\-ro\-vé\-ho in\-že\-nýr\-ství
+Vedoucí diplomové práce: \mfadvisor{}
 
 Abstrakt: Tato práce navazuje na dřívější pokusy odvodit (inferovat) schéma existujících XML dokumentů.
-Jelikož je odvozování \textit{struktury} již relativně dobře popsáno, soustředíme se na integritní omezení.
-Několik jich popisujeme, pozornost pak soustředíme na ID/\.IDREF/\.IDREFS atributy z DTD.
-Na bázi článku od Barbosa a Menelzon (2003) stavíme heuristický přístup k problému hledání optimální sady ID atributů, jeho funkčnost a vhodnost pak ověřujeme na škále experimentů.
+Jelikož je~odvozování \textit{struktury} již relativně dobře popsáno, soustředíme se~na~integritní omezení.
+Několik jich popisujeme, pozornost pak soustředíme na~ID/\.IDREF/\.IDREFS atributy z DTD.
+Na~bázi článku od~Barbosa a~Menelzon (2003) stavíme heuristický přístup k~problému hledání optimální sady ID atributů, jeho funkčnost a~vhodnost pak ověřujeme na~škále experimentů.
 
@@
 
-Supervisor: \mfadvisor{} % TODO does this have to be here?, Department of Software Engineering
+Supervisor: \mfadvisor{}
 
-Abstract: In this work we~expand upon the previous efforts to~extract (infer) schema information from existing XML documents.
-We find the~inference of \textit{structure} to~be~sufficiently researched and focus further on~\textit{integrity constraints}.
+Abstract: In this work we~expand upon the~previous efforts to~infer schema information from existing XML documents.
+We find the~inference of~\textit{structure} to~be~sufficiently researched and focus further on~\textit{integrity constraints}.
 After briefly introducing some of~them we turn our attention to~ID/\.IDREF/\.IDREFS attributes in~DTD.
 Building on~the~research by~Barbosa and Menelzon (2003) we~introduce a~heuristic approach to~the~problem of~finding optimal ID set.
-The approach is~evaluated and tuned in~a~wide range of~experiments.
+The~approach is~evaluated and tuned in~a~wide range of~experiments.
 

### 2011-11-19 doc/theses/vitasek/bibliography.tex
Thesis: fixing the bibliography again.

@@
 
+\bibitem[BDF{\etalchar{+}}00]{fajt-12}
+Peter Buneman, Susan Davidson, Wenfei Fan, Carmem Hara, and Wang chiew Tan.
+\newblock Reasoning About Keys for XML.
+\newblock In {\em Information Systems}, page 2003, 2000.
+
 \bibitem[BDF{\etalchar{+}}01]{keX}
@@
 \newblock In {\em Proceedings of the 10th international conference on World Wide Web}, WWW '01, pages 201--210, New York, NY, USA, 2001. ACM.
+
+\bibitem[BLNT07]{fajt-51}
+Jana Bauckmann, Ulf Leser, Felix Naumann, and Véronique Tietz.
+\newblock Efficiently Detecting Inclusion Dependencies.
+\newblock In {\em In Int. Conf. on Data Engineering (ICDE 07}. Poster, 2007.
 
@@
 \newblock Schemascope: a System for Inferring and Cleaning XML Schemas.
-\newblock In {\em Proceedings of the 2008 ACM SIGMOD international conference on Management of data}, SIGMOD '08, pages 1259--1262, New York, NY, USA,
-  2008. ACM.
+\newblock In {\em Proceedings of the 2008 ACM SIGMOD international conference on Management of data}, SIGMOD '08, pages 1259--1262, New York, NY, USA, 2008. ACM.
 
 \bibitem[BPM{\etalchar{+}}08]{Bray:08:EML}
-Tim Bray, Jean Paoli, Eve Maler, Fran\c{c}ois Yergeau, and C.~M.
-  Sperberg-McQueen.
+Tim Bray, Jean Paoli, Eve Maler, Fran\c{c}ois Yergeau, and C.~M. Sperberg-McQueen.
 \newblock Extensible Markup Language ({XML}) 1.0 (Fifth Edition).
@@
 \newblock {\em Linear Programming and Extensions}.
-\newblock Landmarks in~Physics and~Mathematics. Princeton University Press,
-  1998.
+\newblock Landmarks in~Physics and~Mathematics. Princeton University Press, 1998.
 
@@
 
+\bibitem[GZ02]{fajt-39}
+G\"{o}sta Grahne and Jianfei Zhu.
+\newblock Discovering Approximate Keys in XML Data.
+\newblock In {\em Proceedings of the eleventh international conference on Information and knowledge management}, CIKM '02, pages 453--460, New York, NY, USA, 2002. ACM.
+
 \bibitem[JR86]{JM1986425}
@@
 \newblock {\em Journal of Algorithms}, 7(3):425 -- 440, 1986.
+
+\bibitem[JUF10]{fajt-53}
+B.~Jana, L.~Ulf, and N.~Felix.
+\newblock {\em Efficient and Exact Computation of Inclusion Dependencies for Data Integration}.
+\newblock Technische Berichte des Hasso-Plattner-Instituts f{\"u}r Softwaresystemtechnik an der Universit{\"a}t Potsdam. Universit{\"a}tsverlag Potsdam, 2010.
 
@@
 Andrew Makhorin.
-\newblock GNU Mathprog.
+\newblock GNU MathProg.
 \newblock \url{http://lpsolve.sourceforge.net/5.5/MathProg.htm}.
@@
 \newblock Generalized Birthday Problem.
-\newblock {\em The American Mathematical Monthly}, 73(4):pp. 385--387, April
-  1966.
+\newblock {\em The American Mathematical Monthly}, 73(4):pp. 385--387, April 1966.
 
@@
 \newblock \url{http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf}.
+
+\bibitem[Pas97]{Paschos:1997:SAO:254180.254190}
+Vangelis~T. Paschos.
+\newblock A Survey of Approximately Optimal Solutions to Some Covering and Packing Problems.
+\newblock {\em ACM Comput. Surv.}, 29:171--209, June 1997.
+
+\bibitem[SBHR06]{fajt-41}
+Yannis Sismanis, Paul Brown, Peter~J. Haas, and Berthold Reinwald.
+\newblock Gordian: Efficient and Scalable Discovery of Composite Keys.
+\newblock In {\em In Proc. International Conference on Very Large Data Bases (VLDB}, pages 691--702, 2006.
 
@@
 Ond\v{r}ej Vo\v{s}ta, Irena Ml\'{y}nkov\'{a}, and Jaroslav Pokorn\'{y}.
-\newblock Even an Ant Can Create An XSD.
+\newblock Even an Ant Can Create an XSD.
 \newblock In {\em DASFAA'08: Proceedings of the 13th international conference on Database systems for advanced applications}, pages 35--50, Berlin, Heidelberg, 2008. Springer-Verlag.

### 2011-11-19 doc/theses/vitasek/chapter-0-preface.tex
Thesis: merging Kesy round 1.

@@
 To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in \cite{ahonen, bex, vyhnanovska}.
-In particular, \cite{archdoc} introduces the jInfer schema inference framework, dealing primarily with the structural parts of the schema: how all the elements, attributes and text data are to be organized in an XML document conforming to that schema.
+In particular, \cite{archdoc} introduces the jInfer schema inference framework, dealing primarily with the structural parts of the schema: how all the elements, attributes and text data are to be organized in an XML document conforming to~that schema.
 Inference of this kind of structural information was greatly improved in \cite{anti}.\\
@@
 The scope of this work is finally limited to the inference of \texttt{ID}/\.\texttt{IDREF}/\.\texttt{IDREFS} from existing XML documents.
-\texttt{ID} attributes were chosen over \texttt{xs:key} because the preliminary research found that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with \texttt{ID} attributes are much more common and it is much easier to obtain large data sets for experiments.
+\texttt{ID} attributes were chosen over \texttt{xs:key} because the preliminary research found out that while it is possible to find real-life XML data with schemas containing \texttt{xs:key} structure, schemas with \texttt{ID} attributes are much more common and it is much easier to obtain large data sets for experiments.
 
-\section*{Structure of the thesis}
+\section*{Structure of the Thesis}
 
@@
 
-We will discuss a closely related Mixed Integer Problem and show that by solving MIP we can solve the Maximum Independent Set and thus the original problem of optimal ID set.
+We will discuss a closely related Mixed Integer Problem and show that by~solving MIP we can solve the Maximum Independent Set and thus the original problem of optimal ID set.
 
@@
 
-Please note that throughout this work we will be explicitly ignoring the $\mathcal{O}()$ complexities of algorithms we use.
+Please note that throughout this work we will disregard the $\mathcal{O}()$ complexities of algorithms we use.
 This is because the algorithms we use are by principle strongly stochastic and their performance often depends on behavior of external tools, which we regarded as black boxes and mostly ignored their inner workings.

### 2011-11-19 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: merging Kesy round 1.

@@
 We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data.
-Parent nodes are connected to child nodes with the edges.
+Parent nodes are connected to child nodes with edges.
 This tree shall be called an \textit{XML tree}.
@@
 
-Without a loss of generality we ignore the actual ordering of nodes in the tree.
+Without loss of generality we ignore the actual ordering of nodes in the tree.
 
@@
 This example introduces an XML file fragment that will be used for demonstration throughout this work.
-An XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}(a), where each node is annotated with a triple $label(v) : id(v) : value(v)$.
+An XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}(a), where each node is annotated with a triplet $label(v) : id(v) : value(v)$.
 
@@
 \[\iota(m) \subseteq \bigcup_{m_i \in I} \iota(m_i).\]
-Furthermore if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.
+Furthermore, if $m$ contains multivalued attributes, it is to be marked \texttt{IDREFS}.
 
@@
 \begin{quote}
-The support of attribute mapping $M_x^y$ is the fraction of edges in the XML tree that connect $x$ elements to $y$ attributes.
+The support of attribute mapping $M_x^y$ is the fraction of edges in~the~XML tree that connect $x$ elements to $y$ attributes.
 \end{quote}
@@
 \paragraph{Example}
-Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of every other mapping is $1 / (2+1+1+1) = 0.2$.\\
+Support of $M_{y}^{a}$ in our example is $2 / (2+1+1+1) = 0.4$. Support of~every other mapping is $1 / (2+1+1+1) = 0.2$.\\
 
@@
 
-The problem of \textit{linear programming} is optimization of a linear function under a set of linear constraints. The formulation is usually called a \textit{linear program}.
+The problem of \textit{linear programming} is optimization of a~linear function under a~set of linear constraints. The formulation is usually called a~\textit{linear program}.
 
@@
 
-Solving a linear program is usually possible in polynomial time using the \textit{simplex algorithm} described for example in \cite{dantzig1998linear}.
+Solving a linear program is usually possible in polynomial time using the~\textit{simplex algorithm} described for example in \cite{dantzig1998linear}.
 

### 2011-11-19 doc/theses/vitasek/chapter-2-research.tex
Thesis: merging Kesy round 1.

@@
 
-According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set with weight more than some $K$ ($K$-\textsc{IDSet}) is in NP.
-Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is NP-hard and thus NP-complete.
+According to the article \cite[Chapter~4]{fidax}, the problem of finding an ID set with weight more than some given $K$ ($K$-\textsc{IDSet}) is~in~NP.
+Furthermore, the independent set (IS) problem can be reduced to $K$-\textsc{IDSet}, meaning $K$-\textsc{IDSet} is~NP-hard and thus NP-complete.
 The transformation from IS problem formulation to~$K$-\textsc{IDSet} problem formulation is as follows.
@@
 \begin{quote}
-Let $G = (V, E)$ be a simple connected graph with vertex set $V = \{v_1, \ldots, v_n\}$, and edge set $E = \{e_1, \ldots, e_m\}$.
-We define the attribute mappings as follows.
+Let $G = (V, E)$ be a~simple connected graph with vertex set $V = \{v_1, \ldots, v_n\}$, and edge set $E = \{e_1, \ldots, e_m\}$.
+We define the~attribute mappings as~follows.
 Let $ \mathcal{I} = V \cup E$, and define $value(x) = x, x \in \mathcal{I}$.
-For each vertex $v_i \in V$, we create a mapping $m_i = \{(v_i, e_j): e_j \in E \,\text{is incident on}\, v_i \}$, and define $\tau(m_i) = v_i$; let $C = \{m_1, \ldots, m_n\}$ be set of all such mappings.
-It is clear that G has an independent set of size $K$ iff $C$ has an ID set of size $K$.
-Also, $C$ can be constructed in time polynomial on $n+m$.
+For each vertex $v_i \in V$, we create a~mapping $m_i = \{(v_i, e_j): e_j \in E \,\text{is incident on}\, v_i \}$, and define $\tau(m_i) = v_i$; let $C = \{m_1, \ldots, m_n\}$ be~set of~all such mappings.
+It~is~clear that G has an~independent set of~size $K$ iff $C$ has an~ID set of~size $K$.
+Also, $C$ can be~constructed in~time polynomial on~$n+m$.
 \end{quote}

### 2011-11-19 doc/theses/vitasek/diplomka.tex
Thesis: fixing the bibliography again.
Thesis: merging Kesy round 1.

@@
 \def\mfauthor{Matej Vitásek}
-\def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá Ph.D.}
+\def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá, Ph.D.}
 \def\mfplacedate{Praha, 2011}
@@
 I would like to thank everyone participating in the jInfer project for creating a~solid foundation for this work, namely Irena Mlýn\-ko\-vá, Mi\-chal Klem\-pa, Má\-rio Mi\-ku\-la, Ro\-bert Sme\-ta\-na and Mi\-chal Švi\-rec.
-The thanks goes again to Ms.
-Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HV}.
+The thanks goes again to Irena Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HV}.
 
@@
 After briefly introducing some of~them we turn our attention to~ID/\.IDREF/\.IDREFS attributes in~DTD.
-Building on~the~research by~Barbosa and Menelzon (2003) we~introduce a~heuristic approach to~the~problem of~finding optimal ID set.
+Building on~the~research by~Barbosa and Menelzon (2003) we~introduce a~heuristic approach to~the~problem of~finding an~optimal ID set.
 The~approach is~evaluated and tuned in~a~wide range of~experiments.
@@
 \addcontentsline{toc}{chapter}{Bibliography}
-\nocite{*}
-\bibliographystyle{alpha}
-\bibliography{literature}
-%	\include{bibliography}
+%\nocite{*}
+%\bibliographystyle{alpha}
+%\bibliography{literature}
+	\include{bibliography}
 

### 2011-11-21 doc/theses/vitasek/appendix-a-jinfer.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 
-The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is to parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG).
+The responsibility of the first module, the \jmodule{Initial Grammar Generator}, is~to~parse all input files (documents, schemas and queries) and create a so-called \textit{initial grammar} (IG).
 \nomenclature{IG}{Initial Grammar}
@@
 This is the responsibility of an aptly named module, the \jmodule{Simplifier}.
-To get the full idea about how this can be done it would be probably best to read Michal Klempa's thesis \cite{anti}, which describes this in great detail.
-Whatever happens, there is simplified grammar on the exit of \jmodule{Simplifier}, ready to be processed by...
+To get the full idea about how this can be done it~would be probably best to read Michal Klempa's thesis \cite{anti}, which describes this in great detail.
+Whatever happens, there is simplified grammar on~the exit of \jmodule{Simplifier}, ready to be processed by...
 

### 2011-11-21 doc/theses/vitasek/appendix-b-iss.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 For the purpose of running a whole experimental set there is the \texttt{Experiment\-Set} interface and its abstract implementation \texttt{Abstract\-Experiment\-Set}.
-Its descendants need only to provide a list of \texttt{Experiment\-Parameter}s and looping as well as data collection will be handled for them.
+Its descendants need only to~provide a list of \texttt{Experiment\-Parameter}s and looping as well as data collection will be handled for them.
 
@@
 
-Subclass the \texttt{Abstract\-Experiment\-Set} class, override \texttt{get\-Name()} to provide the name of this set and finally override \texttt{get\-Experiments()} to return the list of \texttt{Experiment\-Parameters} that will constitute this sit.
+Subclass the \texttt{Abstract\-Experiment\-Set} class, override \texttt{get\-Name()} to provide the name of this set and finally override \texttt{get\-Experiments()} to return the list of~\texttt{Ex\-per\-i\-ment\-Parameters} that will constitute this sit.
 
@@
 They will be invoked before running the first experiment, after each experiment run and after all experiments finished, respectively.
-Note that \texttt{notify\-Finished()} already can output some information regarding the currently finished experiment to a file, but it can be safely overriden without a need to call \texttt{super.\.notify\-Finished()}.
+Note that \texttt{notify\-Finished()} already can output some information regarding the currently finished experiment to a file, but it~can be safely overriden without a need to call \texttt{super.\.notify\-Finished()}.

### 2011-11-21 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 
-At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to a given \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that there are several possibilities how to optimize the work of the solver by employing various heuristics.
+At this point of the thesis the reader should be already familiar with the notions we have introduced: the problem of finding the optimal ID set (with respect to~a~given \textit{weight}), that it is directly related to the NP-complete problem of finding the maximal weighted independent set, that this can be solved using the MIP approach, and that there are several possibilities how to optimize the~work of~the~solver by employing various heuristics.
 
@@
 We want to see the changes introduced by modifying several parameters, while keeping the others fixed.
-They probably will not be orthogonal, we might at least isolate some of the parameters that are less important to the overall behavior.
+They probably will not be orthogonal, we might at~least isolate some of the parameters that are less important to the overall behavior.
 
-Second, we evaluate the system performance in terms of the speed of finding good heuristic results. We find tweaks to make the whole process as fast as reasonably possible.
+Second, we evaluate the system performance in terms of the speed of finding good heuristic results. We find tweaks to make the whole process as fast as~reasonably possible.
 
@@
 
-From 3 different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
+From three different sources we collected 6 different data sets, called \dataset{OVA1} - \dataset{OVA3}, \dataset{XMA-c}, \dataset{XMA-p} and \dataset{XMD}.
 Their summary is in Table \ref{table-experiments-data-realistic}, their graph representations can be seen in Figure \ref{image-experiments-data-realistic}.
@@
 On the other hand, the \dataset{XMA-*} sets are relatively huge, but trivial: their only candidate AM will just get picked and the heuristic will end.
-Therefore we will see the performance of the other components of the whole system, such as loading the data sets into memory representations.
+Therefore we will see the performance of the other components of the whole system, such as~loading the data sets into memory representations.
 
@@
 To address the conversion: in case of \dataset{MSH} we found 2 elements with values resembling a key of the records contained in the file, and converted them to be attributes of these records using a simple XSL transformation.
-In the case of \dataset{NTH} we converted all the values in sub-elements of the record elements to be the attributes of the records.
+In the case of \dataset{NTH} we converted all the values in sub-elements of the record elements to~be~the~attributes of the records.
 
@@
 
-Our artificial test data will look similar to the following excerpt from an XML file:
+Our artificial test data will look similar to the following excerpt from an~XML file:
 
@@
 First, we introduce $v$ elements with names \texttt{vertex0} - \texttt{vertex\{V-1\}}.
-To constitute an~AM, they need an attribute \texttt{attr}, but with large enough random values, so that they do not conflict with others.
+To constitute an~AM, they need an attribute \texttt{attr}, but with random values large enough, so~that they do not conflict with others.
 Second, for each of the $e$ edges we choose two \texttt{vertex*} elements at random, and give them the same value of~their \texttt{attr}.
@@
 
-This way we can generate data more complex than anything found in the real world. It will however serve us to measure the performance of our algorithms in extreme cases.\\
+This way we can generate data more complex than anything found in~the~real world. It will however serve us to measure the performance of our algorithms in extreme cases.\\
 
@@
 
-There is one characteristic that can describe random graphs like this, and that is the \textit{density}.
+There is one characteristic that can describe random graphs like this, and~that is the \textit{density}.
 This can be defined in various ways, we will use two different interpretations.
@@
 
-The second, perhaps more interesting is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
-This is the density as the fraction of edges that are to all edges that could be in a~complete graph with $|V|$ vertices.
+The second, perhaps more interesting, is $\frac{|E|}{E_{max}}$, where $E_{max} = \frac{|V|.(|V|-1)}{2}$.
+This is the density as the fraction of edges that are to all edges that could be in~a~complete graph with $|V|$ vertices.
 
 We have created 3 sets to be used in experiments along with the realistic and converted sets, called \dataset{100-100}, \dataset{100-200} and \dataset{100-1000}.
-Note that the name is always in the form $v-e$.\\
+Note that the name is~always in~the~form $v-e$.\\
 
@@
 While studying the tables it becomes obvious that the actual numbers $|V|$ and $|E|$ do not match to the $v$ and $e$ in the names of the sets.
-This is because of the way the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set.
-Because of the so-called Birthday paradox (see e.g.
+This is because of~the~way the random generation algorithm works: it might pick the same edge twice, which will automatically render it unsuitable for the ID set.
+Because of~the~so-called Birthday paradox (see e.g.
 \cite{birthday}), this will happen more with higher $e$.
@@
 In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was indeed achieved.
-There is an interesting observation to be made here: the optimum is steadily decreasing with the increasing overall graph size.
+There is an interesting observation to~be~made here: the optimum is steadily decreasing with the increasing overall graph size.
 This intuitively suggests that the maximum quality theoretically achievable is related to the $\frac{|E|}{|V|}$ density, not to the one we fixed.
@@
 
-An \textit{experiment instance}, or \textit{experiment configuration}, is one specific setting of all experiment parameters.\\
+An \textit{experiment instance}, or \textit{experiment configuration}, is one specific setting of~all experiment parameters.\\
 
@@
 This is done by a linear scan and retrieving a so-called \textit{flat} representation.
-This consists of a~list of tuples in the following format.
+This consists of~a~list of~tuples in the following format.
 
@@
 
-The model now has to be able to return the list of all attribute mappings and their respective images.
+The model now has to be able to return the list of all attribute mappings and~their respective images.
 This is achieved by simply grouping the flat representation by the pair \textit{(element name, attribute name)} and aggregating all attribute values for each such pair.
-Another responsibility of the model is to return the list of \textit{types} - that is simply the list of unique \textit{element name}s.
+Another responsibility of the model is to return the~list of \textit{types} - that is simply the list of unique \textit{element name}s.
 
@@
 Whenever it is necessary to measure the duration of an operation, we will use the \texttt{System.nanoTime()} built-in function.
-The result cannot be interpreted in an absolute manner, but by subtracting the time at the start from the time at the end, we can get a reasonably reliable measurement.
+The result cannot be~interpreted in an absolute manner, but by subtracting the~time at~the~start from the~time at~the~end, we can get a reasonably reliable measurement.
 
@@
 An experimental set will produce a number of these traces and store them in plain text files in a folder.
-Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in format recognized by GnuPlot \cite{gnuplot}, which we use to plot charts found in this work.
+Parsing these files to aggregate and collate them might be a tedious task even using tools like \texttt{sed} and \texttt{grep}, so some of the experiment sets directly output tabular data in~format recognized by GnuPlot \cite{gnuplot}, which we use to plot charts found in~this work.
 
@@
 \begin{quote}
-Quartile boundaries are determined such that 1/4 of the points have a value equal or less than the first quartile boundary, 1/2 of the points have a value equal or less than the second quartile (median) value, etc.
+Quartile boundaries are determined such that 1/4 of the points have a value equal or less than the first quartile boundary, 1/2 of~the~points have a value equal or less than the second quartile (median) value, etc.
 A box is drawn around the region between the first and third quartiles, with a horizontal line at the median value.
@@
 
-The ``user-specified limits'' of whiskers are set to default value, let us quote from the manual again.
+The ``user-specified limits'' of whiskers are set to a default value, let us quote from the manual again.
 

### 2011-11-21 doc/theses/vitasek/chapter-5-future.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 
-It is possible (and easy) to add more construction and improvement heuristics, as well as more metaheuristics using the existing IHs. A~starting point is in Section \ref{section-appendix-iss-howto}.\\
+It is possible (and easy) to add more construction and improvement heuristics, as well as more metaheuristics using the existing IHs. A~starting point is~in~Section \ref{section-appendix-iss-howto}.\\
 
@@
 The approach used in this work was strictly single-threaded, however there are in principle no limitations to extending this to a parallel, multi-threaded environment.
-For example, creating a pool of initial solutions in \heu{Glpk} construction heuristic can be improved by running several instances of GLPK solver in parallel - as GLPK on its own uses only a single thread.\\
+For example, creating a pool of initial solutions in \heu{Glpk} construction heuristic can be improved by running several instances of GLPK solver in~parallel - as GLPK on its own uses only a single thread.\\
 
-From the point of view of a user - researcher, the current implementation of the experimental framework leaves a lot to be desired.
-As jInfer already contains support for interchangeable and configurable modules, it is possible to create GUI for experiment and experimental set configuration on the fly.\\
+From the point of view of a user - researcher, the current implementation of~the experimental framework leaves a lot to be desired.
+As jInfer already contains support for interchangeable and configurable modules, it is possible to~create GUI for experiment and experimental set configuration on the fly.\\
 

### 2011-11-21 doc/theses/vitasek/chapter-6-conclusion.tex
Thesis: merging Lydka round 1.
Thesis: merging Kesy, fixing singletons and duplex.

@@
 
-However, this approach took too long for some inputs, so we introduced a whole range of construction as well as improvement heuristics.
+However, this approach took too long for some inputs, so we introduced a~whole range of construction as well as improvement heuristics.
 We combined these algorithms to create a metaheuristic and performed a number of experiments to understand its behavior.
@@
 While it is important to be able to write a heuristic algorithm tailored to the specific problem being solved, such as the authors of \cite{fidax} did, it should be noted that sometimes it is better to solve a more generalized problem.
-In this case the transformation to MIP formulation and using a dedicated solver proved to produce better results in shorter time.
+In~this case the transformation to MIP formulation and using a dedicated solver proved to~produce better results in shorter time.

### 2011-11-21 doc/theses/vitasek/appendix-c-trace.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 
-The first section deals with system information. Please note that some of these characteristics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
+The first section deals with system information. Note that some of~these char\-ac\-ter\-is\-tics cannot be easily obtained programmatically and are thus stored in the source code as constants.\\
 To obtain GLPK information, the program parses the first line of standard output produced by running \code{glpsol -v}.
@@
 To be able to reconstruct the progress of the metaheuristic, the next section contains CSV \nomenclature{CSV}{Comma Separated Values} formatted data for each iteration.
-Each row contains the time in milliseconds, quality of the incumbent solution and the number of its AMs.
+Each row contains the time in~milliseconds, quality of the incumbent solution and the number of its AMs.
 
@@
 And finally, it is important to know what is the ID/IDREF set recommended by this experiment run.
-Thus the log is concluded by a CSV formatted list of element - attribute name pairs to be included in the ID and IDREF set, respectively.
+Thus the log is concluded by a CSV formatted list of~element - attribute name pairs to be included in the ID and IDREF set, respectively.
 

### 2011-11-21 doc/theses/vitasek/chapter-3-mip.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 First, we transform the problem formulation to maximum weighted IS problem formulation.
-Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK \cite{glpk}.
+Then we transform this into an MIP formulation, and demonstrate how this can be solved using a solver such as GLPK \cite{glpk}.
 \nomenclature{GLPK}{GNU Linear Programming Kit}
-We will continue by applying heuristical approaches to improve the performance of the process.
+We will continue by~applying heuristic approaches to improve the performance of the process.
 
@@
 
-An advantage of using \textit{Branch \& Bound} is that while traversing the search tree it finds intermittent, sub-optimal solutions.
-It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\
+An advantage of using \textit{Branch \& Bound} is~that while traversing the search tree it finds intermittent, sub-optimal solutions.
+It is thus possible to limit the~total search time and instead of the optimum take the best solution found so far.\\
 
-We will now demonstrate the full process of finding the optimal ID set of an example XML file using GLPK.
+We will now demonstrate the full process of~finding the optimal ID set of~an~example XML file using GLPK.
 
@@
 
-The first three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$.
-The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.
+The first three edges are due to the type collision ($y$), the last one is due to~$\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$.
+The graph $G$ constructed in this way is shown in~Figure~\ref{image-mip-is-graph}.
 
 \begin{figure}
+  \centering
+	\includegraphics[width=.25\textwidth]{images/is-representation}
   \caption{IS Representation Graph}
   \label{image-mip-is-graph}
-  \centering
-	\includegraphics[width=.25\textwidth]{images/is-representation}
 \end{figure}
@@
 Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings.
-For each pair of AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$.
-$\mathbf{b}$ will be a vector of ones with corresponding length.
+For each pair of~AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$.
+$\mathbf{b}$ will be a vector of~ones of corresponding length.
 
@@
 	A \textit{heuristic algorithm} is one that, in a reasonably short time, generates a good, maybe even optimal solution to an optimization problem.
-However, it won’t provide any formal guarantee about its quality.
+However, it will not provide any formal guarantee about its quality.
 \end{define}
@@
 
-While a heuristic algorithm can be seen as a tool designed to solve one specific problem, the notion of a \textit{metaheuristic} reminds of a recipy to solve a whole family of problems.
+While a heuristic algorithm can be seen as a tool designed to solve one specific problem, the notion of a \textit{metaheuristic} reminds of a recipe to solve a whole family of problems.
 We shall be using the metaheuristic presented in Figure \ref{image-metaheuristic} to find optimal ID sets in this work.
@@
 \begin{define}[Solution Neighborhood]
-	Neighborhood of a solution $\mathbf{x}$ in the solution space are all the other solutions close to $\mathbf{x}$ according to some metric.
+	Neighborhood of a solution $\mathbf{x}$ in~the~solution space are all the other solutions close to~$\mathbf{x}$ according to~some metric.
 \end{define}
 
-The precise definition of the neighborhood is always adjusted according to specific needs.
+The precise definition of the neighborhood is always adjusted according to~specific needs.
 However, the neighborhood should be defined so that is \textit{continuous} (with respect to quality) to be useful.
@@
 
-The notion of metaheuristic covers a wide range of topics in the field of heuristics, such as Tabu Search (see \cite{Glover:1997:TS:549765}), Ant Colony Optimization (see \cite{dorigo2004ant}) or Genetic Algorithms (see e.g.
+The notion of metaheuristic covers a wide range of topics in the field of~heuristics, such as Tabu Search (see \cite{Glover:1997:TS:549765}), Ant Colony Optimization (see \cite{dorigo2004ant}) or Genetic Algorithms (see e.g.
 \cite{goldberg1989genetic}), to name a few.\\
@@
 The purpose of a construction heuristic (CH) is then to provide us with at least some solution.
-This may or may not be already the optimum, in the latter case it will be improved on later using improvement heuristics (IH).
+This may or may not be already the~optimum, in the latter case it will be improved on later using improvement heuristics (IH).
 Some IHs can profit from a pool of several sub-optimal solutions, and some CHs can produce this pool from them.
@@
 One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a hungry heuristic.
+This is obviously a greedy heuristic.
 
@@
 
-This is again a hungry heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
+This is again a greedy heuristic, the Listing \ref{listing-ch-fuzzy} contains its pseudocode.
 
@@
 
-This trivial heuristics sorts all candidate AMs by their decreasing weights (see Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible.
-This way it can create only one solution, and again, this is a hungry heuristic.
+This trivial heuristic sorts all candidate AMs by their decreasing weights (see Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible.
+This way it can create only one solution, and again, this is a greedy heuristic.
 
@@
 Therefore, sort them by increasing size and start removing them from the solution, until it satisfies the ID set condition.
-Again, this is a hungry heuristic returning only one solution.
+Again, this is a greedy heuristic returning only one solution.
 
@@
 To be able to create a pool of solutions, the GLPK input in MathProg langauge is always randomly shuffled by changing the order in which variables and constraints appear.
-This is causes the solver to explore the search tree in various orders, producing different solution in each of the time-constrained runs.
+This is causes the solver to explore the search tree in~various orders, producing different solution in each of the time-constrained runs.
 
@@
 
-Improvement heuristics in general start with a solution pool, attempt to improve one or more solutions in it and then return this improved pool in the end.
+Improvement heuristics in general start with a solution pool, attempt to improve one or more solutions in it and then return this improved pool in~the~end.
 We will need two notions to describe their behavior.
@@
 
-is the attempt to move the solution towards the nearby local optimum in the solution space.
+is the attempt to move the solution towards the nearby local optimum in~the~solution space.
 
@@
 
-is the attempt to move the solution away (escape) from from the local optimum, to be able to explore more of the solution space when the metaheuristic starts stagnating.\\
+is the attempt to move the solution away (escape) from the~local optimum, to be able to explore more of the solution space when the metaheuristic starts stagnating.\\
 
 A metaheuristic needs to combine intensification and diversification tendencies to explore the solution space and at the same time arrive at a local optimum.
-Recall the requirement for the solution space to be continuous in terms of quality: this guarantees that as we approach a solution $\mathbf{x}$, the quality of solutions we encounter approaches the quality of $\mathbf{x}$.
+Recall the requirement for the solution space to be continuous in~terms of quality: this guarantees that as we approach a solution $\mathbf{x}$, the quality of solutions we encounter approaches the quality of $\mathbf{x}$.
 
@@
 
-This ultimately trivial improvement heuristics does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its Listing \ref{listing-ih-identity}.
+This ultimately trivial improvement heuristic does nothing. It simply returns the feasible pool unchanged. For the sake of completeness, see its Listing \ref{listing-ih-identity}.
 
@@
 the one with the lowest quality).
-This is interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it from growing by pruning inferior solutions.
+This is interesting in cooperation with other improvement heuristics that increase the solution pool size, to keep it~from growing by pruning inferior solutions.
 
@@
 The number of AMs to remove from each solution is specified as fraction from $(0, 1)$ of the solution size (number of AMs in solution).
-For example, \heu{Random Remove} with $fraction = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to rounding).
+For example, \heu{Random Remove} with $fraction = 0.1$ would remove 1 random AM from a solution containing 10 AMs and 2 from a solution containing 17 AMs (due to~rounding).
 
-This heuristic returns a pool of solutions of the same size as it got on its input.
+This heuristic returns a pool of solutions of the same size as it got on input.
 
@@
 We will take a random guess and fix some of these AMs, i.e.
-we add new constraints to the MIP formulation fixing values of the respective variables to 1.
+we add new constraints to~the~MIP formulation fixing values of the respective variables to 1.
 
@@
 This improvement heuristic expands on the idea of \heu{Mutation}.
-But, instead of randomly selecting AMs in the incumbent solution, it looks for commonalities among the solutions in the pool.
+But, instead of~randomly selecting AMs in the incumbent solution, it looks for commonalities among the solutions in the pool.
 This is based on the hope that if more solutions agree on the same AMs, those are probably included in the optimal solution too.
@@
 \heu{Crossover} takes a parameter - fraction of solutions in the pool among which to look for commonalities.
-AMs found in every one of them are fixed in the modified MIP formulation the same way as in \heu{Mutation}.
-This again amounts to an intensification tendency.
+AMs found in every one of them are fixed in the~modified MIP formulation the same way as in \heu{Mutation}.
+This again amounts to~an~intensification tendency.
 
 Listing \ref{listing-ih-crossover} captures the process.
-\heu{Crossover} requires at least one solution in the pool, but to work properly, more are needed.
-Solutions are picked at random from the pool, common AMs found and fixed.
-GLPK is run again (with a time constraint) and the improved solution is added to the result pool.
+\heu{Crossover} requires at least one solution in~the pool, but to work properly, more are needed.
+Solutions are picked at~random from the pool, common AMs found and fixed.
+GLPK is run again (with a~time constraint) and the improved solution is~added to~the~result pool.
 
@@
 See Listing \ref{listing-ih-localbranching} for pseudocode.
-The heuristic requires a pool containing at least one solution, solves the modified MIP formulation using GLPK limited to some time again and adds the improved solution to the result pool.\\
+The heuristic requires a pool containing at~least one solution, solves the modified MIP formulation using GLPK limited to~some time again and adds the improved solution to the result pool.\\
 
@@
 
-Now the IDREF set contains all the AMs whose images are a subset of the ID set image.
+Now the IDREF set contains all the AMs whose images are a subset of the~ID set image.
 

### 2011-11-21 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: merging Kesy, fixing singletons and duplex.

@@
 
-We can see from the results that for most data sets their model can easily be created under around one second, only in case of the biggest set \dataset{XMA-p} (13 MB) this takes some 17 seconds.
-We can conclude that grammar and model creation times are not a bottleneck for now.
-Heuristics run times will be order of magnitude higher.
+We can see from the results that for most data sets their model can easily be created under around one second, only in case of the biggest set \dataset{XMA-p} (13 MB) this takes approximately 17 seconds.
+We can conclude that grammar and~model creation times are not a bottleneck for now.
+Heuristics run times will be~order of magnitude higher.
 
@@
 Interestingly enough, in most cases the times to create an input for GLPK and then to parse its output are very similar.
-Also, for sized test data sets it is interesting to note that even though the $|V|$ and $|E|$ numbers are increasing, the times remain almost the same.
-This is probably due to the fact that IC and OP times include the I/O when writing to a file for GLPK or reading the file it produced, and these times are probably the most relevant.
+Also, for sized test data sets it~is~interesting to note that even though the $|V|$ and $|E|$ numbers are increasing, the~times remain almost the same.
+This is probably due to the fact that IC and OP times include the I/O when writing to a file for GLPK or reading the~file it~produced, and these times are probably the most relevant.
 
@@
 
-If we find out which of these Cygwin version is better, we will be using it exclusively knowing this should not affect any other aspect of our experiments.
+If we find out which of these Cygwin version is better, we will be using it~exclusively knowing this should not affect any other aspect of our experiments.
 We might also find that there is no relevant difference, which would be and interesting finding, too.
@@
 The first part of this experiment will be limiting the run time, thus making it an instance of \heu{Truncated branch \& bound}.
-In this case we will see the dependency between the run time and the quality achieved in it.
+In this case we will see the dependency between the run time and the quality achieved in~it.
 In the second part we will let GLPK run until the optimum is found.
@@
 We can see from the graph that even though for smaller times (1 and 6 seconds, respectively) the Cygwin GLPK is reaching better qualities with smaller variance, starting from 11 seconds the native GLPK is at least as good or better for every following time.
-The results are inconclusive though, it is necessary to wait for confirmation from the second part of this experiment.
+The results are inconclusive though, it is necessary to~wait for confirmation from the second part of this experiment.
 
@@
 
-The other way to compare the performance of these two GLPK versions is to see how long it takes them to find the optimum for a set of data of increasing size.
+The other way to compare the performance of these two GLPK versions is~to~see how long it takes them to find the optimum for a set of data of increasing size.
 This experimental set will contain 550 configurations for each version.
-Every configuration will let \heu{Glpk} CH run for unlimited time, until it finds the optimum.
+Every configuration will let \heu{Glpk} CH run for unlimited time, until it~finds the optimum.
 This will be repeated in 50 iterations for each of the 11 files from the sized test data set (see Algorithm \ref{listing-experiment-glpk-native-vs-cygwin-2}).
@@
 
-The results are captured in Figure \ref{image-experiment-time-until-optimum}; please take a note that the Y axis is in log scale.
-As with the previous case, the X axis cannot be interpreted in the usual way.
+The results are captured in Figure \ref{image-experiment-time-until-optimum}; please take a note that the Y axis is~in~log scale.
+As with the previous case, the X axis cannot be interpreted in~the usual way.
 For each data set there are two boxplots next to each other: the left one is the native GLPK, the right one is the Cygwin GLPK.
@@
 
-Our investigation into various CHs will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized hungry heuristics, \heu{Random} and \heu{Fuzzy}.
+Our investigation into various CHs will start by comparing \heu{FIDAX} from the original article \cite{fidax} to 2 of our trivial randomized greedy heuristics, \heu{Random} and \heu{Fuzzy}.
 
@@
 
-We shall now try to find the best standalone CH, that is the CH that finds on average the best solutions when run without any IHs.
+We shall now try to find the best standalone CH, that is the CH that finds on~average the best solutions when run without any IHs.
 We need to set a time limit for \heu{Glpk} to make it an instance of \heu{Truncated Branch \& Bound}, and we shall use 1 second.
-This is the smallest time limit possible for GLPK and it is still a reasonably short time, fair to other CHs.
+This is the smallest time limit possible for GLPK and it~is~still a~reasonably short time, fair to other CHs.
 
@@
 See the Algorithm \ref{listing-experiment-best-standalone-ch} for details.
-This time we are not interested in run times, only in qualities which we shall gather in format for GnuPlot.\\
+This time we are not interested in run times, only in qualities which we shall gather in~format for GnuPlot.\\
 
@@
 
-We should note that the combination \textit{best CH - best IH} found this way does not necessarily need to be the best one overall, because we find it using a hungry approach.
-
-\begin{center}
-\bigskip
-\begin{tabular}{| l | l |}
-  \hline
-  \hline
-  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
+We should note that the combination \textit{best CH - best IH} found this way does not necessarily need to be the best one overall, because we find it using a greedy approach.
+
+\begin{center}
+\bigskip
+\begin{tabular}{| l | l |}
+  \hline
+  \hline
+  Input data        & \dataset{80-320}, \dataset{90-405}, \dataset{100-500}, \\
                     & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
@@
 \STATE $ES \gets \emptyset$
-\FOR{$file \in \{\dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
+\FOR{$file \in \{\dataset{80-320}, \dataset{90-405}, \dataset{100-500}, \dataset{100-100}, \dataset{100-200}, \dataset{100-1000}\}$}
 	\FOR{$i = 1 \to 50$}
@@
 
-As we mentioned before, we chose the combination \heu{Glpk} and \heu{Mutation} in a hungry manner.
+As mentioned before, we chose the combination \heu{Glpk} and \heu{Mutation} in~a~greedy manner.
 We will now try to take a step back and attempt to replace \heu{Glpk} with \heu{Random}, hoping to get similar qualities in much shorter time (a reminder: \heu{Glpk} always takes 1 second).
@@
   \hline
-  Input data        & \dataset{80-30}, \dataset{90-405}, \dataset{100-500}, \\
+  Input data        & \dataset{80-320}, \dataset{90-405}, \dataset{100-500}, \\
                     & \dataset{100-100}, \dataset{100-200}, \dataset{100-1000} \\
@@
 Moreover, it achieves higher quality in each data set.
-On the other hand, combination \heu{Random} + \heu{Mutation} has much shorter running times and in the biggest (and hardest) data set \dataset{100-1000} has almost comparable results.
+On~the other hand, combination \heu{Random} + \heu{Mutation} has much shorter running times and in the biggest (and hardest) data set \dataset{100-1000} has almost comparable results.
 This makes it a reasonable choice for big inputs where short time is more important than optimal quality.
@@
 
-After finding the best combination of a CH and IH we turn our attention to parameters.
+After finding the best combination of a CH and IH we turn our attention to~parameters.
 The first ones are the $\alpha$ and $\beta$ from the definition of our weight function (Section \ref{section-definitions-weight}).
@@
 The pseudocode capturing this is provided in Algorithm \ref{listing-experiment-various-betas}.
-We will use \heu{Glpk} constrained to 1 second (thus making it an instance of \heu{Truncated Branch \& Bound}) and no IHs.
+We will use \heu{Glpk} constrained to~1~second (thus making it an instance of \heu{Truncated Branch \& Bound}) and no IHs.
 Pool size as well as iteration count will be 1.
@@
 
-The 2 different ID sets found for various $\alpha$ and $\beta$ in \dataset{OVA1} are listed in Table \ref{table-experiments-various-betas-ova1} (note that the actual names had to be anonymized for reasons discussed in Section \ref{section-realistic-data}).
+The 2 different ID sets found for various $\alpha$ and $\beta$ in \dataset{OVA1} are listed in Table \ref{table-experiments-various-betas-ova1} (note that the actual names had to be anonymized for reasons discussed in~Section \ref{section-realistic-data}).
 The differing attribute mapping is marked.
@@
 Interestingly enough, there is no clear separation between the two ID sets depending on the $\alpha-ratio$ to be found.
-The existence of the two sets might be due to the fact that \heu{Glpk} randomizes the order in which AMs are presented to the external GLPK solver.
+The existence of the two sets might be~due to the fact that \heu{Glpk} randomizes the order in which AMs are presented to the external GLPK solver.
 However, this question is beyond the scope of this work, and shall be left for future work.
@@
 
-When considering data sets such as \dataset{XMA-p}, we notice that they contain a lot of simple text nodes that do not contribute to our search, but possibly slow it down.
-Precisely for this reason the \jmodule{BasicIGG} module in jInfer contains an option to turn off processing of such nodes.
-(It also allows to ignore the content of attributes, but this would be devastating to our cause.) Ignoring the content of text nodes means internally that these are created, but their actual string content is skipped and not saved in the memory structures.
-This means that the whole data model occupies less space on the heap, which can possibly lead to better performance.
+When considering data sets such as \dataset{XMA-p}, we notice that they contain a lot of simple text nodes that do not contribute to our search, but possibly slow it~down.
+Precisely for this reason the \jmodule{BasicIGG} module in jInfer contains an~option to turn off processing of such nodes.
+(It also allows to ignore the content of~attributes, but this would be devastating to our cause.) Ignoring the content of~text nodes means internally that these are created, but their actual string content is skipped and not saved in the memory structures.
+This means that the whole data model occupies less space on the heap, which can possibly lead to~better performance.
 
@@
 
-In this section we will describe the most interesting experimental area, that is chaining more than one improvement heuristics and running them in a loop.
+In this section we will describe the most interesting experimental area, that is~chaining more than one improvement heuristics and running them in a loop.
 Unfortunately, the sheer number of possible combinations in which IHs can be ordered (as well as the number of ways to set their parameters) prohibits us from investigating this in depth.
@@
 We shall then employ a higher-level heuristic: we will choose 3 strategies (lists of IHs, or metaheuristics), assess their performance to find the best one and then tune its parameters.
-This approach is by no means exhaustive, it is just a probe in the problem space.
+This approach is by no means exhaustive, it~is~just a probe in the problem space.
 
@@
 All the fractions are set to $0.1$ for the time being.
-The termination criterion is set to limit the total runtime to 10 seconds and (potentialy) infinite iterations.
+The termination criterion is set to~limit the total runtime to 10 seconds and (potentialy) infinite iterations.
 
@@
 X and Y axes represent the time and quality, as usual.
-Each line represents one run of the strategy (metaheuristic) in the following way: the N\textsuperscript{th} break in the line (i.e.
+Each line represents one run of the strategy (metaheuristic) in the following way: the N\textsuperscript{th} break in~the line (i.e.
 the N\textsuperscript{th} data point) is the partial result after the N\textsuperscript{th} step of the strategy.
@@
 CH will be \heu{Random} with pool size of 10.
-IHs will be the ones from Strategy 1, with their fractions set to one of the 45 combinations produced in the following way.
+IHs will be the ones from Strategy 1, with their fractions set to~one of the 45 combinations produced in the following way.
 \begin{itemize}
@@
 
-After averaging the data, we get a large result table; an excerpt of it is in Table \ref{table-experiments-chained-ihs-tweaking-s1}.
-Only the results for the biggest data sets and a few combinations of RR, MUT and CX fractions are presented.
+After averaging the data, we get a large result table; an excerpt of~it~is~in~Table~\ref{table-experiments-chained-ihs-tweaking-s1}.
+Only the results for the biggest data sets and a few combinations of~RR, MUT and CX fractions are presented.
 
@@
 
-It is now necessary to pick one fraction combination as the best one, and it is $(RR = 0.5, MUT = 0.2, CX = 0.1)$.
+It is now necessary to pick one fraction combination as the best one, and it~is~$(RR = 0.5, MUT = 0.2, CX = 0.1)$.
 Using this combination for all the data sets from \dataset{10-20} up to \dataset{80-320} the optimum was always found and for \dataset{90-405} and \dataset{100-500} the success ratios were the highest.
@@
 Now to interpret the fractions in the best combination.
-\heu{RandomRemove} fraction of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded.
+\heu{RandomRemove} fraction of 0.5 means that a randomly chosen half of all AMs from every ID set in~the pool will be discarded.
 This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima.
@@
 \heu{RandomRemove} and \heu{Mutation} fractions in the best combination are at the upper bound of the range we chose for them.
-As a future work option it is possible to start moving these fractions even more in their preferred way.
+As a future work option it~is~possible to start moving these fractions even more in their preferred way.
 
@@
 
-Furthermore, it is always good to ignore the simple text data nodes, as it will improve the total run time.\\
+Furthermore, it is always good to ignore the simple text data nodes, as~it~will improve the total run time.\\
 
 Now compare this to the performance of the original algorithm from \cite{fidax} that we have implemented as \heu{FIDAX}.
-We cannot use it to find the optimum, even if we do have enough time: it is hungry, deterministic and always gives us only one solution.
+We cannot use it to find the optimum, even if we do have enough time: it is greedy, deterministic and always gives us only one solution.
 And even if time is constrained, we have shown in Section \ref{section-experiments-random-fuzzy-fidax} that even trivial algorithms such as \heu{Random} or \heu{Fuzzy} very often outperform \heu{FIDAX}.

### 2011-11-22 doc/theses/vitasek/appendix-b-iss.tex
Thesis: merging Kesy.

@@
 
-Subclass the \texttt{Abstract\-Experiment\-Set} class, override \texttt{get\-Name()} to provide the name of this set and finally override \texttt{get\-Experiments()} to return the list of~\texttt{Ex\-per\-i\-ment\-Parameters} that will constitute this sit.
+Subclass the \texttt{Abstract\-Experiment\-Set} class, override \texttt{get\-Name()} to provide the name of this set and finally override \texttt{get\-Experiments()} to return the list of~\texttt{Ex\-per\-i\-ment\-Parameters} that will constitute this set.
 

### 2011-11-22 doc/theses/vitasek/appendix-c-trace.tex
Thesis: merging Kesy.

@@
 
-The second section states the input file along with its size and graph representation (Section \ref{section-experiments-data}). Alpha and beta parameters for this instance belong here too.
+The second section states the input file along with its size and graph representation (Section \ref{section-experiments-data}). The $\alpha$ and $\beta$ parameters for this instance belong here, too.
 
@@
 After the last improvement phase, the reason why the metaheuristic terminated is stated.
-Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and alpha / beta settings.
+Possible causes are exceeding the maximum time available, maximum iterations or reaching the known optimum for this file and $\alpha$ / $\beta$ settings.
 \\

### 2011-11-22 doc/theses/vitasek/chapter-6-conclusion.tex
Thesis: merging Kesy.

@@
 The wisdom found in the experiments in this work might be the following.
-While it is important to be able to write a heuristic algorithm tailored to the specific problem being solved, such as the authors of \cite{fidax} did, it should be noted that sometimes it is better to solve a more generalized problem.
+While it is important to be able to write a heuristic algorithm tailored to the specific problem being solved, such as the authors of \cite{fidax} did, it should be noted that sometimes it is better to solve a more general problem.
 In~this case the transformation to MIP formulation and using a dedicated solver proved to~produce better results in shorter time.

### 2011-11-22 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: fixing singletons and merging Kesy.

@@
 
-We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data.
+We shall use the representation introduced in \cite{fidax}, where an XML file is~represented by a labeled tree consisting of nodes for elements, attributes and simple text data.
 Parent nodes are connected to child nodes with edges.
@@
 
-We can use projection to retrieve all the unique \textit{ids} of either elements or attributes from the relation, with notation $\pi_E(M_{x}^{y})$ and $\pi_A(M_{x}^{y})$.
+We can use projection to retrieve all the unique \textit{ids} of either elements or~attributes from the relation, with notation $\pi_E(M_{x}^{y})$ and $\pi_A(M_{x}^{y})$.
 
@@
 Returning to our example, the following are all the possible ID sets: $\{ M_{y}^{a} \}, \{ M_{y}^{b} \}, \{ M_{y}^{b}, M_{z}^{a} \}, \{ M_{y}^{c} \}, \{ M_{y}^{c}, M_{z}^{a} \}$.
-Note that once we select an AM of type $y$ we can never add any other with the same type.
-Note also that $\{ M_{y}^{c}, M_{z}^{a} \}$ is not an ID set, because $\iota(M_{y}^{c}) \cap \iota(M_{z}^{a}) \neq \emptyset$.
+Note that once we select an AM of~type $y$ we can never add any other with the same type.
+Note also that $\{ M_{y}^{c}, M_{z}^{a} \}$ is~not an ID set, because $\iota(M_{y}^{c}) \cap \iota(M_{z}^{a}) \neq \emptyset$.
 
@@
 \begin{define}[Support]
-\textit{Support} of an attribute mapping $m$ is defined as follows:
+\textit{Support} of an attribute mapping $m$ is defined as~follows:
 \[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}.\]
@@
 \begin{define}[Coverage]
-\textit{Coverage} of an attribute mapping $m$ is defined as follows:
+\textit{Coverage} of an attribute mapping $m$ is defined as~follows:
 \[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|.\]
@@
 \begin{define}[Maximum weighted independent set]
-Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied:
+Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an~independent set $I_{max}$ is~the \textit{maximum weighted independent set}, iff the following is satisfied:
 \[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v).\]
@@
 
-Where $\mathbf{x}$ is the vector of variables (to be found by the optimization), $\mathbf{b}$ is the vector and $A$ its accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function.
+Where $\mathbf{x}$ is~the vector of variables (to be found by the optimization), $\mathbf{b}$ is~the vector and $A$ its accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function.
 $\mathbf{x}$ and $\mathbf{c}$ have length $n$, $\mathbf{b}$ has length $m$ and $A$ has dimensions $m \times n$.

### 2011-11-22 doc/theses/vitasek/chapter-2-research.tex
Thesis: fixing singletons and merging Kesy.

@@
 \begin{quote}
-The idea behind is an observation that a projection of entities corresponds to a key if each counted aggregation for a projection is equal to 1.
-Thus, this method searches for all possible projections of a dataset while computing aggregations on the projected part of the set of entities.
+The idea behind is~an~observation that a~projection of~entities corresponds to~a~key if~each counted aggregation for a~projection is~equal to 1.
+Thus, this method searches for all possible projections of a~data\-set while computing aggregations on~the projected part of~the set of~entities.
 \end{quote}
@@
 
-Maximum weigthed IS is a well researched topic with a lot of known direct or approximation algorithms, see e.g.
+Maximum weigthed IS is a well researched topic with a lot of~known direct or~approximation algorithms, see e.g.
 \cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}.
-According to for example \cite{Paschos:1997:SAO:254180.254190}, the best known approximation algorithm for weighted IS to-date achieves an approximation ratio of $3(\Delta + 2)$, where $\Delta$ is the maximum degree of a vertex in the IS graph.
-This article lists several algorithms similar to those we introduce in the following chapters.\\
+According to \cite{Paschos:1997:SAO:254180.254190}, the best known approximation algorithm for weighted IS to-date achieves an~approximation ratio of $3(\Delta + 2)$, where $\Delta$ is the maximum degree of a vertex in~the IS graph.
+This article lists several algorithms similar to those we introduce in~the following chapters.\\

### 2011-11-22 doc/theses/vitasek/chapter-3-mip.tex
Thesis: fixing page numbers and a figure.
Thesis: fixing singletons and merging Kesy.

@@
 
-GLPK is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} for MIP.
+GLPK is a multi-platform, multi-purpose solver well suited for this task. It~uses the Simplex method to solve LP problems and \textit{Branch \& Bound} for MIP.
 
-\paragraph{\textit{Branch \& Bound}} This is an optimization method where the search space is systematically divided into smaller sub-spaces.
+\paragraph{\textit{Branch \& Bound}} This is an optimization method where the search space is~systematically divided into smaller sub-spaces.
 This is the \textit{branching} component, and a so-called \textit{search tree} is built this way.
@@
 \begin{figure}
+  \caption{IS Representation Graph}
+  \label{image-mip-is-graph}
   \centering
 	 \includegraphics[width=.25\textwidth]{images/is-representation}
-  \caption{IS Representation Graph}
-  \label{image-mip-is-graph}
 \end{figure}
@@
 \begin{define}[Solution Space, Solution Quality]
-	Solution space in general is the set of all permissible solutions (not violating any constraints).
+	Solution space in general is~the~set of all permissible solutions (not violating any constraints).
 In the specific case of a MIP formulation it is the set of all $\mathbf{x}$ subject to $A\mathbf{x} \leqslant \mathbf{b}$.
-Every solution in the solution space has its \textit{quality}, in case of MIP for solution $\mathbf{x}$ it is the value of the objective function in $\mathbf{x}$.
+Every solution in the solution space has its \textit{quality}, in case of MIP for solution $\mathbf{x}$ it~is~the value of the objective function in $\mathbf{x}$.
 \end{define}
@@
 A quick reminder of what we are trying to solve using our metaheuristic: given a list of AMs with weights, find a non-conflicting subset maximizing the sum of weights in the subset.
+This sum will henceforth be considered the quality of the solution (subset).
 We will now describe its structure, please refer again to Figure \ref{image-metaheuristic}.
@@
 The algorithm works in two steps.
-First, all candidate AMs are grouped according to their types, and for each type the AM with the highest weight is selected.
+First, all candidate AMs are grouped according to their types, and for each type the AM with the highest weight is~selected.
 Second, all these AMs (now called $C'$) are traversed in order of their decreasing size.
@@
 One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution.
-This is obviously a greedy heuristic.
+This is~obviously a greedy heuristic.
 
@@
 
-\heu{Fuzzy} is an improvement over the \heu{Random} CH: the next AM to be added is selected based on \textit{weighted} instead of uniform random.
-The weight used here is the usual weight of an AM as defined in Section \ref{section-definitions-weight}.
+\heu{Fuzzy} is an improvement over the \heu{Random} CH: the next AM to be added is~selected based on \textit{weighted} instead of uniform random.
+The weight used here is~the usual weight of an AM as defined in Section \ref{section-definitions-weight}.
 Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.

### 2011-11-22 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: fixing singletons and merging Kesy.

@@
 using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see Section \ref{section-realistic-converted} for details).
-And finally, we create completely artificial data to create inputs that will put our heuristics in stress.
+And finally, we create completely artificial data to create inputs that will put our heuristics in~stress.
 This is because the realistic data often prove to be too simple to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.
@@
 In the scope of this work this will always mean a \textit{single} XML file.
-However, this definition of a data \textit{set} covers also the extension to more XML files as described in \cite{fidax}.
+However, this definition of~a~data \textit{set} covers also the extension to more XML files as described in \cite{fidax}.
 \end{define}
@@
 
-The former will be achieved with the help of the GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together).
+The former will be achieved with the help of the GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of~their images (and thus cannot be in the same ID set together).
 Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.
 
-The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ (found by running the \heu{Glpk} heuristic without a time limit) and the numbers of vertices and edges in aforementioned graphs.
+The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ (found by running the \heu{Glpk} heuristic without a time limit) and the numbers of vertices and edges in~aforementioned graphs.
 
@@
 There is an interesting observation to~be~made here: the optimum is steadily decreasing with the increasing overall graph size.
-This intuitively suggests that the maximum quality theoretically achievable is related to the $\frac{|E|}{|V|}$ density, not to the one we fixed.
+This intuitively suggests that the maximum quality theoretically achievable is~related to the $\frac{|E|}{|V|}$ density, not to the one we fixed.
 Exploration of this phenomenon is beyond the scope of this thesis.\\
@@
 
-An input data set is a single XML file on the filesystem; however, there is a straightforward extension to multiple files conforming to the same schema.
+An input data set is a single XML file on the filesystem; however, there is~a~straightforward extension to multiple files conforming to the same schema.
 The first step in this process is to use jInfer's module \jmodule{BasicIGG} module (see \cite{basiciggdoc} for details) to obtain a list of rules - an \textit{initial grammar} (IG).

### 2011-11-22 doc/theses/vitasek/chapter-0-preface.tex
Thesis: fixing singletons and merging Kesy.

@@
 Along with technologies such as SQL/noSQL\footnote{noSQL: collection of non-relational database technologies, \url{http://nosql-database.org/}.} databases, proprietary binary file formats, plain-text configuration files and JSON\footnote{JSON: JavaScript Object Notation, lightweight data format, \url{http://www.json.org/}.}, XML is one of the leading formats for storing structured data.
-However, even though languages such as DTD and XML Schema\footnote{DTD and XML Schema: 2 most prominent XML schema languages, \cite{Bray:08:EML}} to describe XML structure exist for a long time, most of the documents use outdated or no schema at all \cite{1802522}.
+However, even though languages such as~DTD and XML Schema\footnote{DTD and XML Schema: 2 most prominent XML schema languages, \cite{Bray:08:EML}} to describe XML structure exist for a long time, most of the documents use outdated or no schema at all \cite{1802522}.
 To tackle this problem one may employ reverse-engineering techniques to infer the schema from existing documents, such as those described in \cite{ahonen, bex, vyhnanovska}.
-In particular, \cite{archdoc} introduces the jInfer schema inference framework, dealing primarily with the structural parts of the schema: how all the elements, attributes and text data are to be organized in an XML document conforming to~that schema.
+In~particular, \cite{archdoc} introduces the jInfer schema inference framework, dealing primarily with the structural parts of the schema: how all the elements, attributes and text data are to be organized in an XML document conforming to~that schema.
 Inference of this kind of structural information was greatly improved in \cite{anti}.\\
@@
 
-Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly marked in the beginning.
+Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly marked in~the~beginning.
 

### 2011-11-23 doc/theses/vitasek/chapter-0-preface.tex
Thesis: finished?

@@
 
-Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly marked in~the~beginning.
+Pseudocode examples such as the one in Listing \ref{listing-example} will always be presented in a functional way, with inputs and outputs of the function clearly marked at~the~beginning.
 

### 2011-11-23 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: finished?

@@
 This tree shall be called an \textit{XML tree}.
-For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.
+For a given node $v$ of an XML tree we define $label(v)$ (name of the node in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.
 
@@
 
-This example introduces an XML file fragment that will be used for demonstration throughout this work.
+This example introduces an XML file fragment that will be used for demonstration throughout the work.
 An XML tree representing it is in Figure \ref{image-definitions-example-xml-tree}(a), where each node is annotated with a triplet $label(v) : id(v) : value(v)$.
@@
   \centering
-    \subfigure[XML tree]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree}}
+    \subfigure[XML Tree]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree}}
     \subfigure[Attribute Mappings]{\includegraphics[width=.45\textwidth]{images/definitions/xml-tree-ams}}
@@
 
-An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED}\footnote{\texttt{\#IMPLIED} means that the attribute has a specified default value.} or \texttt{\#REQUIRED}\footnote{\texttt{\#REQUIRED} means that the attribute cannot be empty.}.
+An \texttt{ID} attribute \textsc{must} have a declared default of \texttt{\#IMPLIED}\footnote{\texttt{\#IMPLIED} means that the attribute has a specified default value.} or \texttt{\#RE\-QUIRED}\footnote{\texttt{\#REQUIRED} means that the attribute cannot be empty.}.
 
@@
 A set of candidate attribute mappings $I = \{m_1, \ldots m_n\}$ is an \textit{ID set} iff
-\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset \wedge \bigcap_{m_i \in I} \iota(m_i) = \emptyset.\]
+\[\bigcap_{m_i \in I} \tau(m_i) = \emptyset\:\text{and} \bigcap_{m_i \in I} \iota(m_i) = \emptyset.\]
 \end{define}
@@
 
-This definition of weight for AMs or AM sets comes from \cite{fidax} again. Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.
+This definition of weight for AMs or AM sets comes from \cite{fidax} again.
+We need the notions of \textit{support} and \textit{coverage} to define it.
+Let $M = \{m_1, \dots m_i\}$ be the set of all non-empty AMs in the document.
 
@@
 \paragraph{Example}
-Coverage of $M_{y}^{a}$ in our example is $(0+0+1) / (2+1+1+1) = 0.2$. Coverage of $M_{z}^{a}$ is $0.2$ as well, all the other mappings have coverage $0$.
+Coverage of $M_{y}^{a}$ in our example is $(0+0+1) / (2+1+1+1) = 0.2$. Coverage of $M_{z}^{a}$ is $0.2$ as well; all the other mappings have coverage $0$.
 \\
@@
 
-Note that this definition of weight is quite arbitrary and all the algorithms mentioned later could easily work with AM weight defined in any other way, even for example defined interactively by the user.
+Note that this definition of weight is quite arbitrary and all the algorithms mentioned later could easily work with AM weight defined in any other way, even, for example, defined interactively by the user.
 

### 2011-11-23 doc/theses/vitasek/diplomka.tex
two side printing margins
Thesis: merging Kesy, fixing singletons and duplex.
Thesis: fixing page numbers and a figure.
Thesis: merging Kesy.
Thesis: finished?

@@
-\documentclass[a4paper,12pt,twoside]{report}
+\documentclass[a4paper,12pt,twoside,openright]{report}
 \usepackage{graphicx}
@@
 \setlength{\oddsidemargin}{4cm}
-\setlength{\evensidemargin}{4cm}
+\setlength{\evensidemargin}{2.5cm}
 \setlength{\topmargin}{25mm}
@@
 \def\mfadvisor{RNDr. Ire\-na Mlýn\-ko\-vá, Ph.D.}
-\def\mfplacedate{Praha, 2011}
+\def\mfplacedate{Prague 2011}
 
@@
 
+\cleardoublepage
+
 \noindent
 I would like to thank everyone participating in the jInfer project for creating a~solid foundation for this work, namely Irena Mlýn\-ko\-vá, Mi\-chal Klem\-pa, Má\-rio Mi\-ku\-la, Ro\-bert Sme\-ta\-na and Mi\-chal Švi\-rec.
-The thanks goes again to Irena Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Mi\-chal Šva\-ger\-ka, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HV}.
+The thanks goes again to Irena Mlýn\-ko\-vá for supervising this work, as well as everyone reviewing it, in no~particular order Lý\-dia Šva\-ger\-ko\-vá, Zu\-za\-na Ma\-sá\-ro\-vá, Mi\-chal Ke\-se\-ly and others\textsuperscript{HV}.
 
@@
 
-\openright
+\cleardoublepage
 \pagestyle{plain}
@@
 % =============== TEXT ================
-\newpage
 
@@
 
-\newpage
+\cleardoublepage
 \phantomsection
@@
 
-\clearpage
+\cleardoublepage
 \phantomsection
@@
 \openright
-\clearpage
+\cleardoublepage
 \phantomsection

### 2011-11-24 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: fixing captions.

@@
 \begin{table}
-  \caption{List of realistic test data files}
+  \caption{List of Realistic Test Data Sets}
   \bigskip
@@
 \begin{figure}
-  \caption{Realistic data}
+  \caption{Realistic Data}
   \label{image-experiments-data-realistic}
@@
 \begin{table}
-  \caption{List of realistic test data files with converted attributes}
+  \caption{List of Realistic Test Data Sets With Converted Attributes}
   \bigskip
@@
 \begin{figure}
-  \caption{Realistic data with converted attributes}
+  \caption{Realistic Data With Converted Attributes}
   \label{image-experiments-data-converted}
@@
 \begin{table}
-  \caption{List of artificial test data files}
+  \caption{List of Artificial Test Data Sets}
   \bigskip
@@
 \begin{table}
-  \caption{List of ``sized'' artificial test data files}
+  \caption{List of ``Sized'' Artificial Test Data Sets}
   \bigskip
@@
 \begin{figure}
-  \caption{Artificial data}
+  \caption{Artificial Data}
   \label{image-experiments-data-artificial}

### 2011-12-08 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 \chapter{Analysis of Recent Approaches} \label{chapter_analysis_of_recent_approaches}
-Existing approaches of XML schema inference can be classified by several criteria.
-A basic division is by the language the resulting schema is written in.
+Existing approaches to XML schema inference can be classified using several criteria.
+A basic classification is based on the language the resulting schema is written in.
 Commonly used languages are DTD and XML Schema.
@@
 According to \cite{Mlynkova:2008:AAX:1494650.1495496}, the type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}.
-Heuristic methods (\cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522}) are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
+Heuristic methods \cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522} are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
 Two of these approaches are described in Section \ref{section_xtract} and Section \ref{section_even_an_ant_can_create_an_xsd}.
+\todo{chce to větu proč zrovna dva}
 
-On the contrary, the grammar-inferring methods (\cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236}) are based on theoretical knowledge of automata and their results belong to a particular class of languages.
+On the contrary, the grammar-inferring methods \cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236} are based on theoretical knowledge of automata and their results belong to a particular class of languages.
 Thus, these methods guarantee specific characteristics of the results.
@@
 An approach utilizing XML data along with an obsolete XML schema is described in \cite{Mlynkova:2009:IXS:1862681.1862693}.
-However, the most significant approaches in terms of this work are those, utilizing operations over XML data.
+However, the most significant approaches in terms of this work are those utilizing operations over XML data.
 According to our best knowledge at the time of writing, there is just one approach of this category, described in \cite{Necasky:2009:DXK:1529282.1529414}.
@@
 \section{Common Caracteristics}
-The process of inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
+The process of XML schema inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ...
 e_k$ is constructed.
@@
 Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
-\todo[inline]{Tento odstavec je doslovne prebraty z uvedeneho clanku. Mozem to takto pouzit?}
+\todo[inline]{Q: Tento odstavec je doslovne prebraty z uvedeneho clanku. Mozem to takto pouzit? A: ano, jen by pak neměla být výrazně jiná úroveň angličtiny}
 
@@
 \subsection{Generalization}
-Purpose of generalization is to create a set of DTD candidates - schemata that cover fractions of the input XML data.
+The purpose of generalization is to create a set of DTD candidates - schemata that cover fractions of the input XML data.
 In the last step, this set will be used to compose a (sub)optimal result with respect to a trade-off between its preciseness and conciseness.
-Therefore, it is desirable to create DTD canditates with various degree of these two characteristics.
+Therefore, it is desirable to create DTD canditates with various degrees of these two characteristics.
 
@@
 For instance, paper \cite{Garofalakis:2000:XSE:342009.335409} introduces the following example: Sequences \texttt{abab} and \texttt{bbbe} are generalized to \texttt{(ab)*}, \texttt{(a|b)*}, \texttt{b*e}.
-\todo[inline]{Mozem takto pouzivat prevzate priklady?}
 
@@
 Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decresing their summed size without modifications in their semantics.
-An aim of this step is to decrease an MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
+The aim of this step is to decrease the MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
 
@@
 
-Alike the generalization step, also in this step the set of possible factored DTD is huge and the authors propose certain heuristics to make the factorization effective.
+Alike the generalization step, also in this step the set of possible factored DTDs is huge and the authors propose certain heuristics to make the factorization effective.
 
@@
 
-\todo[inline]{Detailnejsie? Priklad?}
-
 \section{Even an Ant Can Create an XSD} \label{section_even_an_ant_can_create_an_xsd}
@@
 \begin{itemize}
-\item Distinguishing elements with the same name but different semantics.
+\item Distinguishing elements with the same name but different context.
 \item Improvements of algorithms adopted from the previous works.
 \item Incorporating inference of an unordered sequence.
-\item Creating a result in XML Schema language.
+\item Creating a result in the XML Schema language.
 \end{itemize}
@@
 \subsection{Clustering of Elements}
-This phase clusters elements on a basis of their context and structure.
-It is done by creating tree structures for each input XML documents where vertices represent elements and attributes and an edge between two vertices expresses their parent-child relation.
+This phase clusters elements on the basis of their context and structure.
+It is done by creating tree structures for each input XML document where vertices represent elements and attributes and an edge between two vertices expresses their parent-child relationship.
 These trees and their subtrees are then compared using an imposed tree similarity measure to find elements with the same semantics.
@@
 The idea behind the ACO heuristic is that a set of artifical ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum.
-An ant
-is performing steps (schema modifications), dying after a predefined number of steps and providing an information - positive feedback - on quality of a solution found.
-The search is performed in a defined count of interations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
+An ant is performing steps (schema modifications), dying after a predefined number of steps and providing an information - positive feedback - on the quality of a solution found.
+The search is performed in a defined number of interations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
 Every step of an ant represents a modification of a schema, in particular, a merge of states in a corresponding PTA.
@@
 \subsection{Result in XSD}
-Unlike the majority of recent approaches, this methods creates its result in XML Schema language (XSD).
-Comparing to DTD, XSD is a newer language and it provides greater expression possibilities.
-Authors of this method focused on inferring elements with the same name but different context and the unordered sequences which can be in XSD expressed by \texttt{xs:all} construct.
-Elements with the same name but different context cannot be expressed in DTD and, however, the unordered sequence can be also expressed in DTD as alternations of ordered sequences, such expression in not practical nor well human-readable.
+Unlike the majority of recent approaches, this methods creates its result in XML Schema.
+Contrary to DTD, XSD is a newer language and it provides greater expressive possibilities.
+The authors of this method focused on inferring elements with the same name but different context and the unordered sequences which can be in XSD expressed by \texttt{xs:all} construct.
+Elements with the same name but different context cannot be expressed in DTD and, although, the unordered sequence can be also expressed in DTD as alternations of ordered sequences, such expression in not practical nor well human-readable.
 
 \section{On Inference of XML Schema with the Knowledge of an Obsolete One}
-An aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently.
+The aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently.
 A XML schema can become obsolete due to changes in a set of XML documents, without capturing these changes in the schema.
-Thus, the schema becames outdated and according to the paper, this case is quite common.
+Thus, the schema becomes outdated and according to the paper, this case is quite common.
 
@@
 
-And the algorithm consists of two independent steps:
+The algorithm consists of two independent steps:
 
@@
 In the fist step the input schema is corrected to conform to the whole set of input XML documents.
-This is done by creating a PTA \todo{ref na def} for each production extracted from the input schema and merging it with a respective production from the initial grammar \todo{ref na def}, involving ACO and MDL heuristics.
+This is done by creating a PTA for each production extracted from the input schema and merging it with a respective production from the initial grammar, involving ACO and MDL heuristics.
 
-In the second step, assuming a correct schema it is wanted to specialize the regular expressions involved in the schema with regard to the XML documents, resulting in more precise and readable schema.
+\todo{Tato věta hodně divná.} In the second step, assuming a correct schema it is wanted to specialize the regular expressions involved in the schema with regard to the XML documents, resulting in more precise and readable schema.
 Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.
@@
 \section{Discovering XML Keys and Foreign Keys in Queries}
-This method is described in paper \cite{Necasky:2009:DXK:1529282.1529414}.
-It tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
+The method described in paper \cite{Necasky:2009:DXK:1529282.1529414} improves automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.
 Just the queries are utilized in inference, no XML data are used.
@@
 The method is based on assumption that each join is done via key/foreign key pair.
-It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is it's respective foreign key or vice versa.
+It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is its respective foreign key or vice versa.
 
-The author describe two possible cases:
+The authors describe two possible cases:
 
@@
 \renewcommand{\labelenumi}{\theenumi}
-\item $L_1$ is a key of the elements $S_1$, $L_2$ is a respective foreign key and it itself is not a key of the elements $S_2$.
-\item $L_2$ is a key of the elements $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements $S_1$ or not.
+\item $L_1$ is a key of the elements in $S_1$, $L_2$ is a respective foreign key and it itself is not a key of the elements in $S_2$.
+\item $L_2$ is a key of the elements in $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements in $S_1$ or not.
 \end{enumerate}
@@
 The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1) or if aggregation function \texttt{count} is applied on a return path (weight: 0.75).
-\todo {definovat return path?}
+\todo {Q: definovat return path?
+A: no, možná spíš jednu větu co to zhruba je}
 
@@
 A key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key.
-\todo[inline]{priklad?}
 
@@
 Since the method deals only with the inference of keys, it is not an overall method of the XML schema inference.
-It is meant to be used in a collaboration with other schema-inferring methods to refine their results.
+It is meant to be used in collaboration with other schema-inferring methods to refine their results.
 
@@
   \begin{tablenotes}
-	\item [1] Result of this method is a list of discovered keys and foreign keys.
+	\item [1] The result of this method is a list of discovered keys and foreign keys.
   \end{tablenotes}
@@
 
-As shown in Table \ref{table_summary_of_recent_approaches}, most of the recent approaches of XML schema inference are base on utilization of XML documents.
-These incorporates various verified methods and the newer approaches often improve the older ones and/or combine them together to achieve better results.
+As shown in Table \ref{table_summary_of_recent_approaches}, most of the recent approaches of XML schema inference are based on utilization of XML documents.
+These incorporate various verified methods and the newer approaches often improve the older ones and/or combine them together to achieve better results.
 
@@
 Also, to our best knowledge, implementation that combines this method with another approach to get the XML schema has not been proposed yet.
-\todo{Toto je pravda, naozaj ta implementacia neexistuje?}
 

### 2011-12-08 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 \chapter{Introduction}
-Extensible Markup Language (XML) \cite{Bray:08:EML} has become a popular standard for data representation and exchange.
-Its prevalence is based on its simplicity and at the same time its flexibility and express power.
+The Extensible Markup Language (XML) \cite{Bray:08:EML} has become a popular standard for data representation and exchange.
+Its prevalence is based on its simplicity and, at the same time, its flexibility and expressive power.
 
@@
 In order to this need, several so-called XML schema languages (XML schemata) have been created.
-The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by The World Wide Web Consorcium (W3C).
+The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by the World Wide Web Consorcium (W3C).
 
-Nevertheless an encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
+Despite an conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
 Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema.
+\todo{ref na clanok s tvrdenim}
 
-Recently in reaction to this situation, significant number of approaches dealing with an automatic construction of an XML schema have been proposed.
+Recently in reaction to this situation, a significant number of approaches dealing with an automatic construction of an XML schema has been proposed.
 An aim is to exploit a provided set of XML documents and infer an XML schema, so that the XML documents are valid against it.
 In addition, the inferred schema should be reasonable in views of human-readability, preciseness and conciseness.
-Most of published approaches of an XML schema inference are of this type - input is a set of XML documents.
+Most of published approaches to XML schema inference are of this type - the input is a set of XML documents.
 They are based on various ideas and can be classified by several aspects as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}.
 
-Besides the mentioned type, some approaches that utilize other or additional sources has been developed.
+Besides the mentioned type, some approaches that utilize other or additional sources have been developed \todo{reference, např.
+DP Klempa}.
 If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery \todo{ref} queries, or other, these sources can be exploited to refine the process of inference.
-The refinement can be achieved in various aspects such as decreasing the speed of the process, getting a more precise, more concise or more readable result or inference of some statements about the data which cannot be (easily) extracted from the data themselves.
+The refinement can be achieved in various aspects such as increasing the speed of the process, getting a more precise, more concise or more readable result or inference of statements about the data which cannot be (easily) extracted from the data themselves.
 
-Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (also discussed in Chapter \ref{chapter_analysis_of_recent_approaches}), leaving a wide space for a possible future research and improvements.
+Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (as also discussed in Chapter \ref{chapter_analysis_of_recent_approaches}), leaving a wide space for a possible future research and improvements.
 
 \section{Aim of this Work}
-\todo[inline]{Toto je skopirovane zo sis-u zo zadania prace, moze to tu byt?}
-The aim of this work is a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts etc.).
+Our aim is to perform a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts \todo{ref} etc.).
 
-Firstly, it is necessary to analyze existing inference solutions in general and to discuss their advantages and disadvantages.
-The core of the work is identification and discussion of information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.
+Firstly, we indend to analyze existing inference solutions in general and to discuss their advantages and disadvantages.
+Then, we try to identificate and discuss information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.
 
-The result of the work will be a proposal of own approach involving the improvements, its implementation and suitable experiments that will show its advantages.
+The work should result into a proposal of own approach involving the improvements, its implementation and suitable experiments that will show its advantages.
 
-\section{Structure of this Work}
+\section{Structure of the Work}
 \todo[inline]{TODO}

### 2011-12-08 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 Only expressions of a particular type are exploited.
-Specifically, an expression has to contain some $pathType$ subexpression $E$ (expression representing a certain element or attribute or set of elements or attributes).
+Specifically, an expression has to contain an subexpression $E$ of $pathType$ type (expression representing a certain element or attribute or set of elements or attributes).
 In the following text, the set represented by expression $E$ is denoted $S$.
@@
 
-\todo[inline]{Zlucovanie odvodenych tvrdeni}
-\todo[inline]{Co ked mame path expression s predikatmi?}
+The output of this step is a set of couples ($pathType$, XML Schema built-in type).
+
+\todo[inline]{Zlucovanie odvodenych tvrdeni.}
+\todo[inline]{Path expressions s predikatmi.}
 

### 2011-12-08 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

@@
 
-\section{XML schema}
+\section{XML Schema}
 An XML schema refers to a description of an XML document in terms of its structure and various constraints.
@@
 
-A validity of an XML document againt its XML schema expresses wheather the document is well-formed \todo{link} and at the same time wheather it conforms to the XML schema.
+A validity of an XML document againt its XML schema expresses wheather the document is well-formed \todo{link}, and at the same time, wheather it conforms to the XML schema.
 
@@
 The content of an element can be denoted by \emph{EMPTY} for an empty element, \emph{ANY} for any content, \emph{(\#PCDATA)} allowing only textual content (without any other subelements), or specified using regular expressions.
-Names of subelement are combined using the well-known operators (\emph{|}, \emph{+}, \emph{*}, \emph{?} and \emph{,}(comma)).
+Names of subelements are combined using operators (\emph{|}, \emph{+}, \emph{*}, \emph{?} and \emph{,}(comma)).
 To express the mixed content \emph{\#PCDATA} can be used in an alternation list with the subelement names and this alternation has to be enclosed in \emph{*} operator.

### 2011-12-08 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

@@
 \chapter{Pre-creation of Algorithm}
-\todo[inline]{Vhodny nazov pre tuto kapitolu}
-According to the analysis in the previous chapter, there is quite wide range of possible utilizations of XQuery queries.
+According to the analysis in the previous chapter, there is quite a wide range of possible utilizations of XQuery queries.
 Besides analysis of what information can be extracted from queries, it is needed to devise how the queries will be processed.
@@
 The analysis of XQuery in the previous chapter discusses mostly XQuery constructs which can be utilized without respective XML data, for example the inference of built-it types.
-This independence is also the main advantage of this approach for if there are no XML data available this approaches can be still used.
+This independence is also the main advantage of this approach, if there are no XML data available, this approaches can be still used.
 
@@
 And even the process of evaluation itself can be analyzed to obtain some useful information.
-\todo[inline]{Toto skusit rozpisat nejak podrobnejsie?}
+\todo[inline]{Q: Toto skusit rozpisat nejak podrobnejsie?
+A: Ty useful information by neškodilo popsat, jinak OK.}
 
@@
 Will they be just searched for certain patterns like it is performed in method \cite{Necasky:2009:DXK:1529282.1529414} or will they be processed in a more sophisticated way?
-That can mean incorporating lexical and syntactic analyses \todo{referencie?} or even a form of an analysis of semantics \todo{referencia?}.
+That can mean incorporating lexical and syntactic analyses \todo{Q: referencie?}, known from creation of compilers, or even a form of an analysis of semantics \todo{referencia?}.
 
-A result of the lexical and syntantic analyses can be some kind of so called syntax tree \todo{referencia}.
-It is a structure representing a word according to a formal grammar of some language \todo{definovat grammar, tree, language}.
+\todo[inline]{Q: referencie? A: Možná jednu společnou na nějakou základní klinu o překadačích. Toto asi nebude jádro práce, takže není třeba přesně referencovat každý pojem.}
+
+The result of lexical and syntantic analyses can be some kind of so-called syntax tree \todo{referencia}.
+It is a structure representing a word according to a formal grammar of some language \todo{Q: definovat grammar, tree, language?
+A: Pokud bude použito dále, tak ano.}.
 In our case, the language is XQuery, its grammar is defined in \todo{ref} and every query is a word of the XQuery language.
-Leafs of the tree represent terminals of the grammar while inner nodes represent non-terminals.
-\todo{priklad nejakeho stromu?} From point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient.
+Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals.
+\todo{priklad nejakeho stromu?} From the point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient.
 For instance, the tree can simplify a search for FLWOR statements.
@@
 An example is a static analysis of expression types.
-Types of literal expressions are defined, functions have return types, path expressions return nodes, etc.
-Applying rules defined in \todo{ref} can be determined types of complex expressions.
+Types of literal expressions are defined, functions have return types, path expressions can return nodes, etc.
+Applying \todo{divná věta} rules defined in TODO \todo{ref} can be determined types of complex expressions.
 These can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.
 
-Following text is an example of inference of a more complex query processing. Consider the following part of a query.
+The following text is an example of inference of a more complex query processing. Consider the following part of a query.
 \begin{verbatim}
@@
 For the type of \texttt{10} is \texttt{xs:integer} the type of the function call has to be convertible to \texttt{xs:double}.
-It has to be some numeric type.
+Thus, it has to be a numeric type.
 The function returns a path expression typed as \texttt{element()}.
@@
 Thus, the return expression is \texttt{//A[@id = "id"]/B}.
-Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"}, is of some numeric type.
+Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"} is of some numeric type.
 And, for the function is parametrized there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.
@@
 But it is also more difficult in terms of implementation.
-On the other hand, simplier approaches of the query processing such as the pattern finding are easier to implement but they also limit possibilities the whole query utilization.
+On the other hand, simplier approaches of the query processing such as the pattern finding are easier to implement but they also limit possibilities of the query utilization.
 
-\todo[inline]{Pre co sme sa rozhodli?}
+\todo[inline]{Tak náročnost implementace rozhodně není kritérium.
+Ta se udělá jednou.
+Navíc ve vědě (tj.
+i DP) nemá opodstatnění.
+Pro DP navíc platí, že nemusíte implementovat vše, co navrhnete.
+Implementace je pomocná věc, která slouží k tomu, abychom experimentálně ukázali vlastnosti a chování navrženého algoritmu.}
+
+\todo[inline]{Pre co sme sa rozhodli? Odvozovat strukturu čistě z dotazů je nesmysl.}
 
 \section{Inference of XML Structure}
-The question of the inference of XML structure from queries is partially discussed in the previous chapter.
+The question of inference of XML structure from queries is partially discussed in the previous chapter.
 We are able to infer XML structure from queries without their evaluation, but in a limited way.
@@
 
-If we want to infer the structure more precisely we need to evaluat the queries, and thus, we need the XML data to evaluate the queries upon.
-But, if we have the XML data we can infer the structure directly from them in an easier and more convenient way.
+If we want to infer the structure more precisely we need to evaluate the queries, and thus, we need the XML data to evaluate the queries upon.
+But, if we have the XML data, we can infer the structure directly from them in an easier and more convenient way.
+
+\todo[inline]{Ono je to spíš naopak.
+To, že známe dotazy, nám může omezit prostor možných XML schémat, která popisují strukturu XML dat.
+Jinak řečeno, může nám to zrychlit odvozování struktury, protože z dotazu víme, že některé možnosti můžeme vyřadit.}
 
@@
 The existing approaches of XML schema inference deal mainly with the inference of XML structure.
-As stated in the previous section, this work does not deal with the inference of XML structure, and hence, the extension of an existing approach will probably be some kind of an independent addition instead of modification and refinement of its core algorithm.
+As stated in the previous section, this work does not deal with the inference of XML structure, and hence, the extension of an existing approach will be a kind of an independent addition instead of modification and refinement of its core algorithm.
 
@@
 Therefore, the basic idea is that the input will be extended also for XQuery queries and the algorithm will consist of three phases.
-The first phase will be taken from the existing approach and it will process the XML data to infer the XML structure.
+The first phase will be taken from an existing approach and it will process the XML data to infer the XML structure.
 The second phase will process the XQuery queries and it will infer statements that can be inferred indepedently of the XML data.

### 2011-12-14 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

@@
 Various languages have been proposed to express XML schemata.
-The most known are Document Type Definition (DTD) \todo{link} and XML Schema Definition (XML Schema, XSD) \todo{link} which are briefly described in the following sections.
-Another examples of the XML schema languages are RELAX NG \todo{link} and  Schematron \todo{link}.
+The most known are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema Definition (XML Schema, XSD) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} which are briefly described in the following sections.
+Another examples of the XML schema languages are RELAX NG \cite{relaxng} and  Schematron \cite{schematron}.
 
-A validity of an XML document againt its XML schema expresses wheather the document is well-formed \todo{link}, and at the same time, wheather it conforms to the XML schema.
+A validity of an XML document againt its XML schema expresses wheather the document is well-formed \cite{Bray:08:EML}, and at the same time, wheather it conforms to the XML schema.
 

### 2011-12-14 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 Two of these approaches are described in Section \ref{section_xtract} and Section \ref{section_even_an_ant_can_create_an_xsd}.
-\todo{chce to větu proč zrovna dva}
+The former one incorporates so-called MDL principle to create the (sub)optimal result.
+The latter one combines several verified methods together to improve the quality of the result.
 
@@
 
-\todo{Tato věta hodně divná.} In the second step, assuming a correct schema it is wanted to specialize the regular expressions involved in the schema with regard to the XML documents, resulting in more precise and readable schema.
+In the second step, regular expressions from the corrected schema are specialized with regard to the XML documents, resulting in a more precise and readable schema.
 Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.
@@
 
-The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1) or if aggregation function \texttt{count} is applied on a return path (weight: 0.75).
-\todo {Q: definovat return path?
-A: no, možná spíš jednu větu co to zhruba je}
+The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a target return path (weight: 1) or if aggregation function \texttt{count} is applied on a target return path (weight: 0.75), where target return paths are paths in $C_R$ starting with $\$e_2$.
 

### 2011-12-14 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 Despite an conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
-Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema.
-\todo{ref na clanok s tvrdenim}
+Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema \cite{Mlynkova:2008:AAX:1494650.1495496}.
 
@@
 \section{Aim of this Work}
-Our aim is to perform a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts \todo{ref} etc.).
+Our aim is to perform a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts \cite{Clark:99:XTV} etc.).
 

### 2011-12-14 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 Instead, we need to keep the result in memory and pass it to consecutive steps of our algorithm.
-Though, this requirement concerns modifications of implementation while the core of the processing remains untouched.
+Though this requirement concerns modifications of implementation, the core of the processing remains untouched.
 Therefore, we also do not describe these modifications.
+
+\todo[inline]{Toto vše je OK, ale asi budeme potřebovat vědět, co nám z té analýzy leze za informace se kterými dál pracujeme.
+Čili vstupy našeho algoritmu je potřeba popsat přesně, ideálně formálně, tj.
+je to asi nějaký strom T= (V, E), kde V je...
+atd.
+Jeslti už je to pak implementováno jako objekty v paměti nebo XML data je celkem jedno a bude popsáno v kapitola Implementation.}
 
@@
 Examples of such constructs are $FLWOR$, $if-than-else$ and function call.
-Leaf nodes represents elements of XQuery language that can be devided no more.
+Leaf nodes represent elements of XQuery language that cannot be further devided.
 Typical representants are literal constants.
 \todo{zistit, ci konstanty nahodou nie su jedine}
+
+\todo[inline]{\^ Toto bude potřeba popsat přesně, ne jen příkladem. Klidně převezměnte z diplomky JS.}
 
@@
 
+\todo[inline]{\^ Toto je skoro stejný odstavec ne?}
+
 An important characteric of the syntax tree is related to a definition of local variables and their scope.
@@
 Nodes of this type contain only two subnodes; a type of the variable and a binding expression.
-The node and the entire its subtree representing the type do not contain any expressions.
+The node and its subtree representing the type do not contain any expressions.
 The binding expression node is an expression node but the represented expression cannot use the variable which is just being bound.
 Therefore the scope of the new variable is not the subtree of the $VariableBinding$ node.
-Assuming the post-order numbering of the nodes, the new variable may be valid in some nodes with a higher post-order sequence number.
-Which nodes that are, it depends on the type of the ancestor node(s).
+Assuming \todo{Toto je velmi high-level popis na to, jak málo o syntaktickém stromě a dalších položkách víme.} the post-order numbering of the nodes, the new variable may be valid in some nodes with a higher post-order sequence number.
+Which \todo{Toto není anglický slovosled.} nodes that are, it depends on the type of the ancestor node(s).
 
@@
 \section{Step 2: Static Analysis of Expression Types}
-In the second step, we try to statically (without evaluation of the queries) determine types of expressions in the syntax tree.
+In the second step, we statically (i.e.
+without evaluation of the queries) determine types of expressions in the syntax tree.
 Information on the types of expressions can be used by consecutive steps of the algorithm.
-The consecutive steps will probably not use the determined types of all expressions, however this step may be useful in a future extending.
+The consecutive steps will not use the determined types of all expressions, however this step may be useful in a future extending.
 
-We want to be able to capture the following types.
+\todo[inline]{Tady z těch try a probably je jasně vidět, že netušíte co vlastně přesně algoritmus bude dělat.
+Nejdříve si to rozmyslete, pak pište text.
+Rozhodně by se takové výrazy neměly v DP dobjevovat, působít o velmi nedůvěryhodně.
+Měl by to být strohý popis toho, co bylo uděláno.}
+
+We want to capture the following types.
 
@@
 \item XML Schema built-in atomic types \todo{link do appendixu}.
-\item Type $xs:untyped$.
+\item Type $xs:untyped$. \todo{to je co?}
 \item Type representing any item. \todo{je potrebne?}
 \item Type representing an element, attribute, item, node, text node, comment, processing instruction, document node.
-\item Type representing a set of elements or attributes selected by a certain path expression. The path expression is included in this type. Let this type be identified as $pathType$
+\item Type representing a set of elements or attributes selected by a certain path expression. The path expression is included in this type. Let this type be identified as $pathType$.
 \end{itemize}
 
-\todo[inline]{Akeho typu je napr path expression /root/(elementA * 2)?}
+\todo[inline]{Q: Akeho typu je napr path expression /root/(elementA * 2)? A: Datový typ je daný operátorem. Hodnota elementu se přetypuje.}
 
@@
 
-If we cannot determine the type of an expression we express that using $null$. Example of such case is path expression $/rootElement/(elementA * 2)$. \todo{Nejak dopisat}
+If we cannot determine the type of an expression, we express that using $null$. An example of such case is path expression $/rootElement/(elementA * 2)$. \todo{Nejak dopisat}
 
@@
 \todo[inline]{Este by sa dala urobit analyza volania funkcii, do ktorych su predavane nejake elementy/atributy o ktorych vieme ich path expression.
-Vysledkom tejto funkcie by totiz zas mohli byt elementy/atributy o ktorych vieme (pozmenenu) path expression a nie len nezname elementy/atributy.}
+Vysledkom tejto funkcie by totiz zas mohli byt elementy/atributy o ktorych vieme (pozmenenu) path expression a nie len nezname elementy/atributy.
+A: Přesně tak. A jak říkám - není třeba vše co navrhnete implementovat.}
 
@@
 Determination of return types of functions is needed because function calls can appear in expressions.
-However a return type of a function can be determined at the moment the analysis of expressions encounters a call of the function, it involves multiple transitions of the syntax tree in a search for a definition of a particular function.
+A return type of a function can be determined at the moment the analysis of expressions encounters a call of the function, however, it involves multiple transitions of the syntax tree in a search for a definition of a particular function.
 
@@
 
-A simple algorithm is presented in \ref{ALG_determination_of_function_return_types} \todo{uvadzat aj taketo jednoduche funkcie?}.
-It uses \texttt{getFunctionDeclarationNodes()} function, which is not presented due to its simplicity.
-It returns a list of all function declaration nodes in the syntax three.
+A simple algorithm is presented in Algorithm \ref{ALG_determination_of_function_return_types} \todo{Q: uvadzat aj taketo jednoduche funkcie?
+A: Obecně by mělo být popsáno vše.
+Ale pokud se nějaká jednoduchá věc např.
+opakuje, tak stačí napsat, že je algoritmus podobný jako v předchozím případě nebo tak něco.}.
+It uses \texttt{getFunctionDeclarationNodes()} function, which returns a list of all function declaration nodes in the syntax three.
 Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.
@@
 For simplicity, we will focus on locally defined functions.
-We will not determine types of functions defined in other modules since the principle is similar but it is somewhat more difficult to implement \todo{je to ok?}.
-Return types of built-in functions are fixed, and thus, there is no need to analyze them.
+We will not determine types of functions defined in other modules since the principle is similar but it is somewhat more difficult to implement \todo{Q: je to ok?
+A: No spíš napište něco jako without loss of generality.}.
+Return types of built-in functions are fixed and, thus, there is no need to analyze them.
 
@@
 
-For the rest of the thesis, we assume function \texttt{getFunctionReturnType()} which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
+For the rest of the thesis, we assume function \todo{Většinou se na začátek sekce s algoritmem dává seznam používaných  funkcí a jejich význam.} \texttt{getFunctionReturnType()} which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
 Otherwise, \texttt{null} is returned.
@@
 \subsection{Determination of Global Variable Types}
-A similar approaches as in the case of functions can be applied to determine types of global variables.
+A similar approach as in the case of functions can be applied to determine types of global variables.
 Alike the functions, the global variables are defined in the prolog section.
@@
 priklad}.
-If it is not it may often be deducible from the binding expression.
-Again, we do not analyse external variables for the same reason of simplicity as the external functions.
+If it is not, it may often be deducible from the binding expression.
+Again, we do not analyse external variables for the same reason as in case of external functions.
 
@@
 
-The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type than the type is noted.
+\todo[inline]{algoritmus 6.2, riadok 2, slovo "represents" - Tady by neškodilo definovat nějakou funkci a její možné návratové hodnoty a toto používat.}
+\todo[inline]{algoritmus 6.2, riadok 19 - Totéž, neškodila by nějaká funkce, tj. trochu formálnější zápis. Jednak se pravděpodobně bude hodit dál a druhak to zpřehlední popis.}
+
+The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted.
 Otherwise, an attempt of the type deduction of the expression is made.
-The deduction of the type from the expression is presented in \ref{ALG_func_determineExpressionType}.
+The deduction of the type from the expression is presented in Algorithm \ref{ALG_func_determineExpressionType}.
 On input, it takes an expression node and types of local variables that are valid in the expression (also called context of variables or variable context).
@@
 
-Function $determineExpressionType()$ is called within function $analysisOfExpressionTypes$ presented in \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
+Function $determineExpressionType()$ is called within function $analysisOfExpressionTypes$ presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
 Two important questions about this function emerge.
@@
 The reason is that an expression node may need to know types of its subexpressions to determine its own type and we want to determine types of all expressions.
-Therefore, using the post-order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion because the types of the subexpressions are already determined.
+Therefore, using the post-order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion, because the types of the subexpressions are already determined.
 
@@
 This is expressed in the algorithm by the condition at line 10.
-This condition is double-checked at line 3 where it meaning is that every node even ''knows'' if its particular subnode may define new variables.
-An example is $FLWOR$ construct.
+This condition is double-checked at line 3 where its meaning is that every node even ''knows'' if its particular subnode may define new variables.
+An example is a $FLWOR$ construct.
 The processing of $FLWOR$ knows that the first part (the first subnode) of this node may bind new variables using $for$ and $let$ constructs.
@@
 This is done at line 4.
-Function $mergeContextVarTypes$ presented in \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into ist first argument (the variable context).
+Function $mergeContextVarTypes$ presented in Algorithm \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into ist first argument (the variable context).
 If there are records for variables with the same names in the context, they are overwritten to correspond to the variable overlaping.
@@
 
+\todo[inline]{Algoritmus analysisOfEXpressionTypes, line 10 - Tady tyhle věty jsou v pseudoalgoritmech divné.
+Měy by se vyskytovat minimálně.
+JInak je to pak v podstatě jen odřádkovaná věta, ne formální popis algoritmu bez technických detailů.}
+
+\todo[inline]{Pozor na to přetékání textu - na spoustě míst.}
+
 \begin{algorithm}
@@
 Function $analysisOfExpressionTypes$ called upon a binding expression of a global variable determines the type of the binding expression, and hence, the type of the variable.
-In the case of global variables, the argument $contextVarTypes$ is empty because there are no local variables valid in the prolog section.
+In case of global variables, the argument $contextVarTypes$ is empty, because there are no local variables valid in the prolog section.
 
@@
 If the function gets a variable that is not a global one, it returns $null$.
-Function $getVariableType()$ called in \ref{ALG_func_determineExpressionType} is defined in \ref{ALG_function_get_variable_type}.
+Function $getVariableType()$ called in Algorithm \ref{ALG_func_determineExpressionType} is defined in Algorithm \ref{ALG_function_get_variable_type}.
 It checks if the specified variable is amongst the given local variables.
@@
 \section{Step 3: Inference of Built-in Types}
-In this step, the algorithm goes through the syntax tree and tries to infer types of elements and attributes from expressions using the type information from the previous step.
+In this step, the algorithm goes through the syntax tree to infer types of elements and attributes from expressions using the type information from the previous step.
 These two steps could be merged together but for better comprehension we present it separately.
@@
 In the following text, the set represented by expression $E$ is denoted $S$.
-Another requirement is the expression has to be either a function call or an arithmetic operation \todo{Nejake ine ako napr type constructor, cast as?
-Princip je stale rovnaky}.
+Another requirement is that the expression has to be either a function call or an arithmetic operation \todo{Q: Nejake ine ako napr type constructor, cast as?
+Princip je stale rovnaky.
+A: ak to jen popište.
+To je naprosto OK.}.
 
@@
 \subsection{Function Calls}
-This case is quite simple.
+This case is quite straightforward.
 The algorithm encounters a function call and one of the arguments is a known set of elements of attributes (subexpression $E$ representing $S$).
@@
 
-If the type $T$ of the formal argument is a built-in type or its sequence then $T$ is also the inferred type of $S$. Otherwise, no statement is inferred.
+If the type $T$ of the formal argument is a built-in type or its sequence, then $T$ is also the inferred type of $S$. Otherwise, no statement is inferred.
 
@@
 If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression is $Operator$ and the class of the operator is one of $GEN\_LESS\_THAN$, $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$, $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.
+
+\todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme.}

### 2011-12-14 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
+\subsection{Definition of the Syntax Tree}
+\begin{definition}
+Syntax tree of a XQuery query $Q$ is a tuple $T = (V, E, o)$ where $V$ is a set of nodes representing XQuery language constructs (TODO\todo{kompletny zoznam}), $E$ is a set of edges - couples $(a,b):a,b \in V$ and $o:V \rightarrow \mathbb{N}$ is an ordering function and $T$ satisfies the following conditions:
+\begin{itemize}
+\item $(V,E)$ is a tree.
+\item The root node is $ModuleNode$ representing the entire query Q.
+\item $(a,b)$ is an element of $E$ if and only if $b$ is a direct component of construct $a$ in $Q$.
+\item For every node $a \in V$ the ordering function $o$ is defined as follows \\* $o(a) = \left\{
+	\begin{array}{ll}
+		1  & \mbox{if }a\mbox{ is the root node} \\
+		seq(a) & \mbox{else}
+	\end{array}
+\right.$ \\* where $seq(a)$ is a sequential number of appearence of construct $a$ in $Q$.
+\end{itemize}
+\end{definition}
+
 \subsection{Description of the Syntax Tree}
@@
 Examples of such constructs are $FLWOR$, $if-than-else$ and function call.
-Leaf nodes represent elements of XQuery language that cannot be further devided.
+Leaf nodes represent elements of XQuery language that cannot be further divided.
 Typical representants are literal constants.

### 2011-12-14 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

@@
 And even the process of evaluation itself can be analyzed to obtain some useful information.
-\todo[inline]{Q: Toto skusit rozpisat nejak podrobnejsie?
-A: Ty useful information by neškodilo popsat, jinak OK.}
+For instance, these are partial results of evaluating of expressions (elements selected by each path expression, real arguments in fuction calls, etc).
 
@@
 Will they be just searched for certain patterns like it is performed in method \cite{Necasky:2009:DXK:1529282.1529414} or will they be processed in a more sophisticated way?
-That can mean incorporating lexical and syntactic analyses \todo{Q: referencie?}, known from creation of compilers, or even a form of an analysis of semantics \todo{referencia?}.
+That can mean incorporating lexical and syntax analyses, known from creation of compilers, or even a form of an analysis of semantics \cite{compilers}.
 
-\todo[inline]{Q: referencie? A: Možná jednu společnou na nějakou základní klinu o překadačích. Toto asi nebude jádro práce, takže není třeba přesně referencovat každý pojem.}
-
-The result of lexical and syntantic analyses can be some kind of so-called syntax tree \todo{referencia}.
-It is a structure representing a word according to a formal grammar of some language \todo{Q: definovat grammar, tree, language?
-A: Pokud bude použito dále, tak ano.}.
-In our case, the language is XQuery, its grammar is defined in \todo{ref} and every query is a word of the XQuery language.
+The result of lexical and syntax analyses can be some kind of so-called syntax tree \cite{compilers}.
+It is a structure representing a word according to a formal grammar of some language.
+In our case, the language is XQuery, its grammar is defined in REF \todo{ref} and every query is a word of the XQuery language.
 Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals.
@@
 
+\todo[inline]{Q: definovat grammar, tree, language? A: Pokud bude použito dále, tak ano.}
+
 The syntax tree can be also extended by additional information.
@@
 Types of literal expressions are defined, functions have return types, path expressions can return nodes, etc.
-Applying \todo{divná věta} rules defined in TODO \todo{ref} can be determined types of complex expressions.
-These can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.
+Types of complex expressions can be determined applying the rules defined in REF \todo{ref}.
+The inferred expression types can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.
 
@@
 
-A more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
-But it is also more difficult in terms of implementation.
-On the other hand, simplier approaches of the query processing such as the pattern finding are easier to implement but they also limit possibilities of the query utilization.
+While simplier approaches of the query processing such as the pattern finding limit possibilities of the query utilization, a more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
+Therefore, we decided to incorporate query processing using the lexical and syntax analyses.
 
@@
 Implementace je pomocná věc, která slouží k tomu, abychom experimentálně ukázali vlastnosti a chování navrženého algoritmu.}
-
-\todo[inline]{Pre co sme sa rozhodli? Odvozovat strukturu čistě z dotazů je nesmysl.}
 

### 2011-12-16 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \end{definition}
+
+\subsection{Syntax Tree Nodes}
+\todo[inline]{Dat do appendixu?}
+Nodes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where a simple node can be of several types.
+Nodes with emphasised names express a common type of a group of nodes, their descendants, and these type nodes cannot be directly used in the syntax tree.
+The remaining nodes represent particular constructs of the XQuery language and their instances can be used as nodes of the syntax tree.
+Each of the instances additionaly contains details about the particular form of the respective XQuery construct used in the query.
+
+For example, an instance of the syntax tree cannot directly contain nodes $Node$ and $ExprNode$, but it can contain instances of nodes $AttributeNode$ and $LiteralNode$.
+Regarding the types, an instance of $LiteralNode$ node is of three types: $LiteralNode$, $ExprNode$ and $Node$.
+It carries information on a type of the literal and its value, for example $xs:integer$ of value $5$.
+
+\begin{description}
+\item \textbf{Node}
+\begin{description}
+\item AttributeNode
+\item AttrListNode
+\item AxisNode
+\item CaseClauseNode
+\item CaseClausesNode
+\item CDataSectionNode
+\item ContentNode
+\item DefaultCaseNode
+\item EntityRefNode
+\item \textbf{ExprHolderNode}
+  \begin{description}
+  \item BindingSequenceNode
+  \item DefaultValueNode
+  \item ElseExpressionNode
+  \item OperandExpressionNode
+  \item ReturnClauseNode
+  \item TestExpressionNode
+  \item ThenExpressionNode
+  \item WhereClauseNode
+  \end{description}
+\item \textbf{ExprNode}
+  \begin{description}
+  \item CommaOperatorNode
+  \item ConstructorNode
+  \item ContextItemExprNode
+  \item EmptySequenceNode
+  \item ExtensionExprNode
+  \item FLWORExprNode
+  \item FunctionCallNode
+  \item IfExprNode
+  \item LiteralNode
+  \item OperatorNode
+  \item OrderedExprNode
+  \item PathExprNode
+  \item QuantifiedExprNode
+  \item TypeswitchExprNode
+  \item UnorderedExprNode
+  \item ValidateExprNode
+  \item VarRefNode
+  \end{description}
+\item FunctionBodyNode
+\item CharRefNode
+\item InClausesNode
+\item \textbf{ItemTypeNode}
+  \begin{description}
+  \item AnyItemNode
+  \item AtomicTypeNode
+  \item KindTestNode
+  \item NameTestNode
+  \end{description}
+\item LocationHintNode
+\item LocationHintsNode
+\item \textbf{ModuleChildNode}
+  \begin{description}
+  \item ModuleDeclNode
+  \item PrologNode
+  \item QueryBodyNode
+  \end{description}
+\item ModuleNode
+\item NameNode
+\item OrderByClauseNode
+\item OrderSpecNode
+\item ParamListNode
+\item ParamNode
+\item PITargetNode
+\item PragmaListNode
+\item PragmaNode
+\item PredicateListNode
+\item \textbf{PrologChildNode}
+  \begin{description}
+  \item BaseURIDeclNode
+  \item BoundarySpaceDeclNode
+  \item ConstructionDeclNode
+  \item CopyNamespacesDeclNode
+  \item DefaultCollationDeclNode
+  \item DefaultNamespaceDeclNode
+  \item EmptyOrderDeclNode
+  \item FunctionDeclNode
+  \item \textbf{ImportNode}
+    \begin{description}
+    \item ModuleImportNode
+    \item SchemaImportNode
+    \end{description}
+  \item NamespaceDeclNode
+  \item OptionDeclNode
+  \item OrderingModeDeclNode
+  \item VarDeclNode
+  \end{description}
+\item \textbf{StepExprNode}
+  \begin{description}
+  \item SelfOrDescendantStepNode
+  \end{description}
+\item StringNode
+\item TupleStreamNode
+\item TypeNode
+\item \textbf{VariableBindingNode}
+  \begin{description}
+  \item ForClauseNode
+  \item InClauseNode
+  \item LetClauseNode
+  \end{description}
+\item VarValueNode
+\end{description}
+\end{description}
 

### 2011-12-16 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 Each of the instances additionaly contains details about the particular form of the respective XQuery construct used in the query.
+
+\todo[inline]{V ExprNode si pamatame aj typ vyrazu}
 
@@
 \end{description}
+
+Some pseudo code algorithms in the following sections need to determine the type of a node instance.
+For this purpose, we define function $isA(nodeInstance, nodeType)$.
+The first argument is an instance of any node and the second is a node type itself.
+The function returns boolean value $true$ if the instance is of the specified node type, according to the described principle.
+Otherwise, it returns $false$.
+
+For instance, assuming variable $literal$ is an instance of node $LiteralNode$, calls $isA(literal, LiteralNode)$, $isA(literal, ExprNode)$ return $true$, while call $isA(literal, ContentNode)$ returns $false$.
 
@@
 \STATE $type = null$
-\IF{$exprNode$ represents a literal value}
-     \STATE $type =$ particular literal value type, exactly-once cardinality
-\ELSIF{$exprNode$ represents a call of function $f$}
-     \STATE $type = getFunctionReturnType(f)$
-\ELSIF{$exprNode$ represents variable reference $var$}
-     \STATE $type = getVariableType(contextVarTypes, var)$
-\ELSIF{$exprNode$ represents path expression $path$}
+\IF{$isA(exprNode, LiteralNode)$}
+     \STATE $type =$ ((LiteralNode)exprNode).type, exactly-once cardinality
+\ELSIF{$isA(exprNode, FunctionCallNode)$}
+     \STATE $type = getFunctionReturnType(((FunctionCallNode)exprNode).functionName)$
+\ELSIF{$isA(exprNode, VarRefNode)$}
+     \STATE $type = getVariableType(contextVarTypes, ((VarRefNode)exprNode).variableName)$
+\ELSIF{$isA(exprNode, PathExprNode)$}
+     \STATE $path = ((PathExprNode)exprNode).path$
      \IF{$path$ selects an element set}
@@
      \ENDIF
-\ELSIF{$exprNode$ represents operator $op$}
-     \IF{$op$ is a comparison operator}
-         \STATE $type = xs:boolean$
-     \ELSIF{$op$ is an additive operator $AND$ one of the operands is of a numeric type}
-         \IF{both operands are of numeric types}
-             \STATE $type =$ the numberic type which is more or equally general than the other
-         \ELSE
-             \STATE $type =$ type of the numeric operand
-         \ENDIF
-     \ENDIF
-\ELSIF{$exprNode$ represents $FLWOR$ construct $flwor$}
-     \STATE  $type =$ type of an expression in the return clause of $flwor$
+\ELSIF{$isA(exprNode, OperatorNode)$}
+     \STATE $type = determineOperatorType(exprNode, contextVarTypes)$
+\ELSIF{$isA(exprNode, FLWORExprNode)$}
+     \STATE $returnClauseNode = ((FLWORExprNode)exprNode).returnClauseNode$
+     \STATE $type = returnClauseNode.type$
 \ENDIF
@@
 
-\todo[inline]{algoritmus 6.2, riadok 2, slovo "represents" - Tady by neškodilo definovat nějakou funkci a její možné návratové hodnoty a toto používat.}
-\todo[inline]{algoritmus 6.2, riadok 19 - Totéž, neškodila by nějaká funkce, tj. trochu formálnější zápis. Jednak se pravděpodobně bude hodit dál a druhak to zpřehlední popis.}
+\todo[inline]{Napisat, ze determineOperatorType() tiez nie je kompletna}
+\todo[inline]{definovat isOperatorClassAddition()}
+\todo[inline]{definovat isNumericType()}
+\todo[inline]{definovat isOperatorClassComparison()}
+
+\begin{algorithm}
+\caption{Function determineOperatorType}
+\label{ALG_func_determineOperatorType}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $operatorNode$: An operator expression node in the syntax tree. \\
+    $contextVarTypes$: Types of local variables valid in the current subtree.
+}
+\ENSURE The operator expression type.
+
+\STATE $type = null$
+\STATE $operator = operatorNode.operator$
+\IF{$isOperatorClassComparison(operator)$}
+    \STATE $type = xs:boolean$
+\ELSIF{$isOperatorClassAddition(operator) AND$ one of the operands is of a numeric type}
+    \STATE $leftOperand = ((OperatorNode)exprNode).leftOperand$
+    \STATE $rightOperand = ((OperatorNode).exprNode).rightOperand$
+    \IF{$isNumericType(leftOperand.type) AND isNumericType(rightOperand.type)$}
+        \STATE $type = selectMoreGeneralType(leftOperand.type, rightOperand.type)$
+    \ELSIF{$isNumericType(leftOperand.type)$}
+        \STATE $type = leftOperand.type$
+    \ELSIF{$isNumericType(rightOperand.type)$}
+        \STATE $type = rightOperand.type$
+    \ENDIF
+\ENDIF
+
+\RETURN $type$
+\end{algorithmic}
+\end{algorithm}
 

### 2011-12-17 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 Nodes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where a simple node can be of several types.
-Nodes with emphasised names express a common type of a group of nodes, their descendants, and these type nodes cannot be directly used in the syntax tree.
+Nodes with emphasised names express a common type of a group of nodes - their descendants, and these type nodes cannot be directly used in the syntax tree.
 The remaining nodes represent particular constructs of the XQuery language and their instances can be used as nodes of the syntax tree.
@@
 
-\todo[inline]{V ExprNode si pamatame aj typ vyrazu}
+\todo[inline]{V ExprNode s ExprHolderNode si pamatame aj typ vyrazu}
 
@@
 
+Nodes can be classified into three groups: inner nodes, leaf nodes and nodes that can be both inner and leaf.
+Inner nodes (marked $I$) stand for XQuery constructs that are composed of other constructs and can be further divided.
+An example of such node is $FLWORExprNode$ which is composed of nodes $TupleStreamNode$, $WhereClauseNode$, $OrderByClauseNode$ and $ReturnClauseNode$.
+Leaf nodes (marked $L$) represent elements of XQuery language that cannot be further divided.
+For example, $LiteralNode$.
+Node $FunctionCallNode$ is an example of the third group (marked $IL$).
+Function call of a function without arguments is represented by a leaf while function call with arguments is represented by an inner node and its subnodes are nodes representing those arguments.
+
 \begin{description}
@@
 \begin{description}
-\item AttributeNode
+\item AttributeNode (I)
 \item AttrListNode
-\item AxisNode
-\item CaseClauseNode
-\item CaseClausesNode
+\item AxisNode (I)
+\item CaseClauseNode (I)
+\item CaseClausesNode (I)
 \item CDataSectionNode
 \item ContentNode
-\item DefaultCaseNode
+\item DefaultCaseNode (I)
 \item EntityRefNode
@@
   \begin{description}
-  \item BindingSequenceNode
-  \item DefaultValueNode
-  \item ElseExpressionNode
-  \item OperandExpressionNode
-  \item ReturnClauseNode
-  \item TestExpressionNode
-  \item ThenExpressionNode
-  \item WhereClauseNode
+  \item BindingSequenceNode (I)
+  \item DefaultValueNode (I)
+  \item ElseExpressionNode (I)
+  \item OperandExpressionNode (I)
+  \item ReturnClauseNode (I)
+  \item TestExpressionNode (I)
+  \item ThenExpressionNode (I)
+  \item WhereClauseNode (I)
   \end{description}
@@
   \begin{description}
-  \item CommaOperatorNode
-  \item ConstructorNode
+  \item CommaOperatorNode (I)
+  \item ConstructorNode (I)
   \item ContextItemExprNode
   \item EmptySequenceNode
-  \item ExtensionExprNode
-  \item FLWORExprNode
+  \item ExtensionExprNode (I)
+  \item FLWORExprNode (I)
   \item FunctionCallNode
-  \item IfExprNode
+  \item IfExprNode (I)
   \item LiteralNode
-  \item OperatorNode
-  \item OrderedExprNode
-  \item PathExprNode
-  \item QuantifiedExprNode
-  \item TypeswitchExprNode
-  \item UnorderedExprNode
-  \item ValidateExprNode
+  \item OperatorNode (I)
+  \item OrderedExprNode (I)
+  \item PathExprNode (I)
+  \item QuantifiedExprNode (I)
+  \item TypeswitchExprNode (I)
+  \item UnorderedExprNode (I)
+  \item ValidateExprNode (I)
   \item VarRefNode
@@
   \end{description}
-\item \textbf{StepExprNode}
+\item StepExprNode
   \begin{description}
@@
 Each node represents a logical component of a query.
-Inner nodes stand for XQuery constructs that are composed of other constructs and can be further divided.
-Examples of such constructs are $FLWOR$, $if-than-else$ and function call.
-Leaf nodes represent elements of XQuery language that cannot be further divided.
-Typical representants are literal constants.
-\todo{zistit, ci konstanty nahodou nie su jedine}
+
+Inner nodes are: $AttributeNode$, $AxisNode$, $CaseClauseNode$, $CaseClausesNode$, $DefaultCaseNode$, $BindingSequenceNode$, $DefaultValueNode$, $ElseExpressionNode$, $OperandExpressionNode$, $ReturnClauseNode$, $TestExpressionNode$, $ThenExpressionNode$, $WhereClauseNode$, $CommaOperatorNode$, $ConstructorNode$, $ExtensionExprNode$, $FLWORExprNode$, $IfExprNode$, $OperatorNode$, $OrderedExprNode$, $PathExprNode$, $QuantifiedExprNode$, $TypeswitchExprNode$, $UnorderedExprNode$, $ValidateExprNode$, $InClausesNode$, $QueryBodyNode$, $ModuleNode$, $OrderByClauseNode$, $OrderSpecNode$, $ParamListNode$, $ParamNode$, $PragmaListNode$, $PredicateListNode$, $FunctionDeclNode$, $ModuleImportNode$, $SchemaImportNode$, $VarDeclNode$, $StepExprNode$, $TupleStreamNode$, $ForClauseNode$, $InClauseNode$, $LetClauseNode$.
+
+Nodes that can be both inner and leaf are: $AttrListNode$, $ContentNode$, $FunctionCallNode$, $FunctionBodyNode$, $KindTestNode$, $LocationHintsNode$, $NameNode$, $PrologNode$, $PITargetNode$, $VarValueNode$.
+
+Leaf nodes are: $CDataSectionNode$, $EntityRefNode$, $ContextItemExprNode$, $EmptySequenceNode$, $VarRefNode$, $PragmaNode$, $BaseURIDeclNode$, $BoundarySpaceDeclNode$, $ConstructionDeclNode$, $CopyNamespacesDeclNode$, $DefaultCollationDeclNode$, $DefaultNamespaceDeclNode$, $EmptyOrderDeclNode$, $LocationHintNode$, $NamespaceDeclNode$, $OptionDeclNode$, $OrderingModeDeclNode$, $SelfOrDescendantStepNode$, $StringNode$, $TypeNode$.
+
+\todo{zistit, ci konstanty nahodou nie su jedine.
+NIE: CDataSectionNode, ContentNode(ale nemusi byt list), EntityRefNode, ContextItemExprNode, EmptySequenceNode, FunctionCallNode(bez parametrov), VarRefNode, FunctionBodyNode(prazdna funkcia), CharRefNode, AnyItemNode, AtomicTypeNode, KindTestNode(nemusi), NameTestNode, LocationHintNode, ModuleDeclNode, PrologNode, NameNode(nemusi), PITargetNode(nemusi), PragmaNode, BaseURIDeclNode, BoundarySpaceDeclNode, ConstructionDeclNode, CopyNamespacesDeclNode, DefaultCollationDeclNode, DefaultNamespaceDeclNode, EmptyOrderDeclNode, NamespaceDeclNode, OptionDeclNode, OrderingModeDeclNode, SelfOrDescendantStepNode, StringNode, TypeNode, VarValueNode}
 

### 2011-12-17 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\subsection{Syntax Tree Nodes}
-\todo[inline]{Dat do appendixu?}
-Nodes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where a simple node can be of several types.
-Nodes with emphasised names express a common type of a group of nodes - their descendants, and these type nodes cannot be directly used in the syntax tree.
+\subsection{Syntax Tree Node Types}
+\todo[inline]{Dat figures do appendixu?}
+The node types (or classes) of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where a node can be of several types.
+This hierarchy is shown in figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}.
+Nodes with emphasised names express a common type of a group of nodes - their descendants, and these type nodes cannot be directly used in the syntax tree (abstract classes).
 The remaining nodes represent particular constructs of the XQuery language and their instances can be used as nodes of the syntax tree.
@@
 
-For example, an instance of the syntax tree cannot directly contain nodes $Node$ and $ExprNode$, but it can contain instances of nodes $AttributeNode$ and $LiteralNode$.
+For example, an instance of the syntax tree cannot directly contain instances of nodes $Node$ and $ExprNode$, but it can contain instances of nodes $AttributeNode$ and $LiteralNode$.
 Regarding the types, an instance of $LiteralNode$ node is of three types: $LiteralNode$, $ExprNode$ and $Node$.
 It carries information on a type of the literal and its value, for example $xs:integer$ of value $5$.
+
+\todo[inline]{Asi bude potrebne nakreslit pre kazdy node, co presne obsahuje. Alebo staci pre tie, ktore budeme my pouzivat?}
 
@@
 
+\begin{figure}
+\caption{Syntax tree node types part 1.}
+\label{FIG_syntax_tree_node_types_part_1}
 \begin{description}
@@
 \item AttributeNode (I)
-\item AttrListNode
+\item AttrListNode (IL)
 \item AxisNode (I)
@@
 \item CaseClausesNode (I)
-\item CDataSectionNode
-\item ContentNode
+\item CDataSectionNode (L)
+\item ContentNode (IL)
 \item DefaultCaseNode (I)
-\item EntityRefNode
+\item EntityRefNode (L)
+\item \textbf{ExprHolderNode}
+\item \textbf{ExprNode}
+\item FunctionBodyNode (IL)
+\item CharRefNode (L)
+\item InClausesNode (I)
+\item \textbf{ItemTypeNode}
+\item LocationHintNode (L)
+\end{description}
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 2.}
+\label{FIG_syntax_tree_node_types_part_2}
+\begin{description}
+\item \textbf{Node}
+\begin{description}
+\item LocationHintsNode (IL)
+\item \textbf{ModuleChildNode}
+\item ModuleNode (I)
+\item NameNode (IL)
+\item OrderByClauseNode (I)
+\item OrderSpecNode (I)
+\item ParamListNode (I)
+\item ParamNode (I)
+\item PITargetNode (IL)
+\item PragmaListNode (I)
+\item PragmaNode (L)
+\item PredicateListNode (I)
+\item \textbf{PrologChildNode}
+\item StepExprNode (I)
+\item StringNode (L)
+\item TupleStreamNode (I)
+\item TypeNode (L)
+\item \textbf{VariableBindingNode}
+\item VarValueNode (IL)
+\end{description}
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 3.}
+\label{FIG_syntax_tree_node_types_part_3}
+\begin{description}
 \item \textbf{ExprHolderNode}
@@
   \end{description}
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 4.}
+\label{FIG_syntax_tree_node_types_part_4}
+\begin{description}
 \item \textbf{ExprNode}
@@
   \item ConstructorNode (I)
-  \item ContextItemExprNode
-  \item EmptySequenceNode
+  \item ContextItemExprNode (L)
+  \item EmptySequenceNode (L)
   \item ExtensionExprNode (I)
   \item FLWORExprNode (I)
-  \item FunctionCallNode
+  \item FunctionCallNode (IL)
   \item IfExprNode (I)
-  \item LiteralNode
+  \item LiteralNode (L)
   \item OperatorNode (I)
@@
   \item ValidateExprNode (I)
-  \item VarRefNode
+  \item VarRefNode (L)
   \end{description}
-\item FunctionBodyNode
-\item CharRefNode
-\item InClausesNode
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 5.}
+\label{FIG_syntax_tree_node_types_part_5}
+\begin{description}
 \item \textbf{ItemTypeNode}
   \begin{description}
-  \item AnyItemNode
-  \item AtomicTypeNode
-  \item KindTestNode
-  \item NameTestNode
+  \item AnyItemNode (L)
+  \item AtomicTypeNode (L)
+  \item KindTestNode (IL)
+  \item NameTestNode (L)
   \end{description}
-\item LocationHintNode
-\item LocationHintsNode
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 6.}
+\label{FIG_syntax_tree_node_types_part_6}
+\begin{description}
 \item \textbf{ModuleChildNode}
   \begin{description}
-  \item ModuleDeclNode
-  \item PrologNode
-  \item QueryBodyNode
+  \item ModuleDeclNode (L)
+  \item PrologNode (IL)
+  \item QueryBodyNode (I)
   \end{description}
-\item ModuleNode
-\item NameNode
-\item OrderByClauseNode
-\item OrderSpecNode
-\item ParamListNode
-\item ParamNode
-\item PITargetNode
-\item PragmaListNode
-\item PragmaNode
-\item PredicateListNode
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 7.}
+\label{FIG_syntax_tree_node_types_part_7}
+\begin{description}
 \item \textbf{PrologChildNode}
   \begin{description}
-  \item BaseURIDeclNode
-  \item BoundarySpaceDeclNode
-  \item ConstructionDeclNode
-  \item CopyNamespacesDeclNode
-  \item DefaultCollationDeclNode
-  \item DefaultNamespaceDeclNode
-  \item EmptyOrderDeclNode
-  \item FunctionDeclNode
+  \item BaseURIDeclNode (L)
+  \item BoundarySpaceDeclNode (L)
+  \item ConstructionDeclNode (L)
+  \item CopyNamespacesDeclNode (L)
+  \item DefaultCollationDeclNode (L)
+  \item DefaultNamespaceDeclNode (L)
+  \item EmptyOrderDeclNode (L)
+  \item FunctionDeclNode (I)
   \item \textbf{ImportNode}
     \begin{description}
-    \item ModuleImportNode
-    \item SchemaImportNode
+    \item ModuleImportNode (I)
+    \item SchemaImportNode (I)
     \end{description}
-  \item NamespaceDeclNode
-  \item OptionDeclNode
-  \item OrderingModeDeclNode
-  \item VarDeclNode
+  \item NamespaceDeclNode (L)
+  \item OptionDeclNode (L)
+  \item OrderingModeDeclNode (L)
+  \item VarDeclNode (I)
   \end{description}
-\item StepExprNode
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 8.}
+\label{FIG_syntax_tree_node_types_part_8}
 \begin{description}
-  \item SelfOrDescendantStepNode
+\item StepExprNode (I)
+  \begin{description}
+  \item SelfOrDescendantStepNode (L)
   \end{description}
-\item StringNode
-\item TupleStreamNode
-\item TypeNode
+\end{description}
+\end{figure}
+
+\begin{figure}
+\caption{Syntax tree node types part 9.}
+\label{FIG_syntax_tree_node_types_part_9}
+\begin{description}
 \item \textbf{VariableBindingNode}
   \begin{description}
-  \item ForClauseNode
-  \item InClauseNode
-  \item LetClauseNode
-  \end{description}
-\item VarValueNode
+  \item ForClauseNode (I)
+  \item InClauseNode (I)
+  \item LetClauseNode (I)
   \end{description}
 \end{description}
+\end{figure}
 
@@
 \subsection{Description of the Syntax Tree}
-Since we modified only the implementation of the output form, the design of the syntax tree remains consequent to the entire design of the analyses.
-Each node represents a logical component of a query.
-
-Inner nodes are: $AttributeNode$, $AxisNode$, $CaseClauseNode$, $CaseClausesNode$, $DefaultCaseNode$, $BindingSequenceNode$, $DefaultValueNode$, $ElseExpressionNode$, $OperandExpressionNode$, $ReturnClauseNode$, $TestExpressionNode$, $ThenExpressionNode$, $WhereClauseNode$, $CommaOperatorNode$, $ConstructorNode$, $ExtensionExprNode$, $FLWORExprNode$, $IfExprNode$, $OperatorNode$, $OrderedExprNode$, $PathExprNode$, $QuantifiedExprNode$, $TypeswitchExprNode$, $UnorderedExprNode$, $ValidateExprNode$, $InClausesNode$, $QueryBodyNode$, $ModuleNode$, $OrderByClauseNode$, $OrderSpecNode$, $ParamListNode$, $ParamNode$, $PragmaListNode$, $PredicateListNode$, $FunctionDeclNode$, $ModuleImportNode$, $SchemaImportNode$, $VarDeclNode$, $StepExprNode$, $TupleStreamNode$, $ForClauseNode$, $InClauseNode$, $LetClauseNode$.
-
-Nodes that can be both inner and leaf are: $AttrListNode$, $ContentNode$, $FunctionCallNode$, $FunctionBodyNode$, $KindTestNode$, $LocationHintsNode$, $NameNode$, $PrologNode$, $PITargetNode$, $VarValueNode$.
-
-Leaf nodes are: $CDataSectionNode$, $EntityRefNode$, $ContextItemExprNode$, $EmptySequenceNode$, $VarRefNode$, $PragmaNode$, $BaseURIDeclNode$, $BoundarySpaceDeclNode$, $ConstructionDeclNode$, $CopyNamespacesDeclNode$, $DefaultCollationDeclNode$, $DefaultNamespaceDeclNode$, $EmptyOrderDeclNode$, $LocationHintNode$, $NamespaceDeclNode$, $OptionDeclNode$, $OrderingModeDeclNode$, $SelfOrDescendantStepNode$, $StringNode$, $TypeNode$.
-
-\todo{zistit, ci konstanty nahodou nie su jedine.
-NIE: CDataSectionNode, ContentNode(ale nemusi byt list), EntityRefNode, ContextItemExprNode, EmptySequenceNode, FunctionCallNode(bez parametrov), VarRefNode, FunctionBodyNode(prazdna funkcia), CharRefNode, AnyItemNode, AtomicTypeNode, KindTestNode(nemusi), NameTestNode, LocationHintNode, ModuleDeclNode, PrologNode, NameNode(nemusi), PITargetNode(nemusi), PragmaNode, BaseURIDeclNode, BoundarySpaceDeclNode, ConstructionDeclNode, CopyNamespacesDeclNode, DefaultCollationDeclNode, DefaultNamespaceDeclNode, EmptyOrderDeclNode, NamespaceDeclNode, OptionDeclNode, OrderingModeDeclNode, SelfOrDescendantStepNode, StringNode, TypeNode, VarValueNode}
-
-\todo[inline]{\^ Toto bude potřeba popsat přesně, ne jen příkladem. Klidně převezměnte z diplomky JS.}
-
-In detail, the syntax tree is a n-ary tree representing syntactic structure of a given query.
-Every node represents a certain XQuery construct and that is a type of the node.
-Some types can be divided into more specific subtypes.
-This is shown \todo{obrazok typov uzlov}.
-
-\todo[inline]{\^ Toto je skoro stejný odstavec ne?}
-
 An important characteric of the syntax tree is related to a definition of local variables and their scope.
@@
 Assuming \todo{Toto je velmi high-level popis na to, jak málo o syntaktickém stromě a dalších položkách víme.} the post-order numbering of the nodes, the new variable may be valid in some nodes with a higher post-order sequence number.
-Which \todo{Toto není anglický slovosled.} nodes that are, it depends on the type of the ancestor node(s).
+It depends on the type of the ancestor node(s) which nodes that are.
 

### 2011-12-18 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

@@
 
+\subsection{An XML Sample}
+To demonstrate the described technologies, we introduce a part of a simple XML document \ref{FIG_a_simple_xml_sample}.
+It represents books in a bookstore.
+Each book has a mandatory id, a title, a list of authors and an optional ISBN.
+
+\begin{figure}
+\caption{A simple XML sample.}
+\label{FIG_a_simple_xml_sample}
+\begin{verbatim}
+<bookstore>
+  <book id="b1">
+    <title>Compilers: Principles, Techniques, and Tools (2nd Edition)</title>
+    <authors>
+      <author>Alfred V. Aho</author>
+      <author>Monica S. Lam</author>
+      <author>Ravi Sethi</author>
+      <author>Jeffrey D. Ullman</author>
+    </authors>
+  </book>
+  <book id="b2">
+    <title>XQuery</title>
+    <authors>
+      <author>Priscilla Walmsley</author>
+    </authors>
+    <isbn>0596006349</isbn>
+  </book>
+</bookstore>
+\end{verbatim}
+\end{figure}
+
 \subsection{DTD}
-Document Type Definition (DTD) expresses a structure of XML documents by declarations of elements. An element has its name, content and optional list of attributes.
+Document Type Definition (DTD) expresses a structure of XML documents by declarations of elements. An element has its name and content is declared using syntax \texttt{<!ELEMENT $name$ $content$>}.
 
-The content of an element can be denoted by \emph{EMPTY} for an empty element, \emph{ANY} for any content, \emph{(\#PCDATA)} allowing only textual content (without any other subelements), or specified using regular expressions.
-Names of subelements are combined using operators (\emph{|}, \emph{+}, \emph{*}, \emph{?} and \emph{,}(comma)).
-To express the mixed content \emph{\#PCDATA} can be used in an alternation list with the subelement names and this alternation has to be enclosed in \emph{*} operator.
+The content of an element can be denoted by \texttt{EMPTY} for an empty element, \texttt{ANY} for any content, \texttt{(\#PCDATA)} allowing only textual content (without any other subelements), or specified using regular expressions.
+Names of subelements are combined using operators (\texttt{|}, \texttt{+}, \texttt{*}, \texttt{?} and \texttt{,}(comma)).
+To express the mixed content \texttt{\#PCDATA} can be used in an alternation list with the subelement names and this alternation has to be enclosed in \texttt{*} operator.
 
-\todo[inline]{dopisat, priklady}
+Attributes of an element are specified in an attribute list \texttt{<!ATTLIST $element\_name$ $attribute\_name$ $type$ $default\_value$>}.
+Each attribute has its name, its type and its default value or definition of obligation of occurrence.
+The type can be an enumeration of values \texttt{($value_1$ | $value_2$ | $\dots$ | $value_n$)} or one of the following values.
+
+\begin{itemize}
+\item \texttt{CDATA} Character data - any string.
+\item \texttt{ID} An unique identifier.
+\item \texttt{IDREF} An ID reference - a value of the ID.
+\item \texttt{IDREFS} A space-separated list of ID references. 
+\item \texttt{NMTOKEN}	A valid XML name.
+\item \texttt{NMTOKENS} A space-separated list of valid XML names.
+\item \texttt{ENTITY} An entity.
+\item \texttt{ENTITIES} A space-separated list of entities.
+\item \texttt{NOTATION} A name of a notation.
+\end{itemize}
+
+The dafault value is either a literal value or one of the following specifiers.
+
+\begin{itemize}
+\item \texttt{\#REQUIRED} The attribute is mandatory.
+\item \texttt{\#IMPLIED} The attribute is optional.
+\item \texttt{\#FIXED $value$} The attribute value is constant $value$.
+\end{itemize}
+
+The DTD also provides other constructs such as declaration of entities and notations not mentioned in this work.
+
+An example of DTD describing the book element from the XML sample \ref{FIG_a_simple_xml_sample} follows.
+
+\begin{verbatim}
+<!ELEMENT book (title, authors, isbn?)>
+<!ATTRLIST book id ID #REQUIRED>
+
+<!ELEMENT title #PCDATA>
+
+<!ELEMENT authors (author+)>
+
+<!ELEMENT author #PCDATA>
+
+<!ELEMENT isbn #PCDATA>
+\end{verbatim}
+
+\todo[inline]{overit, zvalidovat}
 
 \subsection{XSD}
-\todo[inline]{TODO}
+Since the XSD language, containing many constructs and features, is quite comprehensive, we will describe just its basic principles.
+An important fact is that each XSD instance is a valid XML document.
+Its root element is always \texttt{<schema>} and the the XSD instance may (and it often does) begins with the following two lines.
+
+\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">
+\end{verbatim}
+
+Definition of an element in the XSD is
+\begin{alltt}
+<xs:element name="\(name\)" type="\(type\)"/>
+\end{alltt}
+where $type$ is either one of the built-in types \todo{ref?} or a user-defined type. Definition of an attribute is similar.
+\begin{alltt}
+<xs:attribute name="\(name\)" type="\(type\)"/>
+\end{alltt}
+Attributes are optional by default. If an attribute is mandatory, it is expressed by adding \texttt{use} attribute to its definition.
+\begin{alltt}
+<xs:attribute name="\(name\)" type="\(type\)" use="required"/>
+\end{alltt}
+
+The user-defined type in the XSD language is either a simple type, if it does not contain other elements and attributes, or a complex type otherwise.
+Attributes are only allowed to be of the simple types.
+Definition of the a simple type is
+\begin{alltt}
+<xs:simpleType name="\(name\)">
+   \(type details\)
+</xs:simpleType>
+\end{alltt}
+The user-defined simple types often serve to restrict the built-in types in various ways such as limiting lengths of strings, allowing only certain values and thus creating an enumeration type, and other.
+
+A complex type is defined in the same way.
+\begin{alltt}
+<xs:complexType name="\(name\)">
+  \(type details\)
+</xs:complexType>
+\end{alltt}
+The complex types can contain many constructs.
+Subelements are declared using \texttt{<xs:sequence>}, \texttt{<xs:choice>} and \texttt{<xs:all>} schema elements.
+If the order of the subelements is significant, we use \texttt{<xs:sequence>}, where the subelements in an XML must occur in the same order as they are defined in the sequence.
+If the order is not significant, we use \texttt{<xs:all>}, where the subelements must only occur but their order is not relevant.
+Construct \texttt{<xs:choice>} is equivalent to the alternation of several elements in the DTD.
+
+Moreover, these three construct can be nested and combined, assuming the combination is not ambiguous.
+
+Many schema elements (including \texttt{<xs:element>}, \texttt{<xs:sequence>}, \texttt{<xs:choice>}, \texttt{<xs:all>}) can be assigned with an occurrence interval.
+The occurrence interval is expressed using \texttt{minOccurs} and \texttt{maxOccurs} attributes.
+For instance, if an element is optional, it definition can be
+\begin{alltt}
+<xs:element name="\(name\)" type="\(type\)" minOccurs="0" maxOccurs="1"/>
+\end{alltt}
+
+An element can has a mixed content (can contain text and other elements at the same time).
+Such element has to be of a complex type and definition of the type has to contain attribute \texttt{mixed} with value \texttt{true}.
+The following example demonstrates the mixed content along with the definition of the element type inside \texttt{<xs:element>} schema element.
+\begin{alltt}
+<xs:element name="\(name\)">
+  <xs:complexType mixed="true">
+    \(complex type details\)
+  </xs:complexType>
+</xs:element>
+\end{alltt}
+
+The XSD language consists of many more constructs we do not mention such as substitution groups, type extensions, integrity constranints and other.
+
+The book element from the sample XML \ref{FIG_a_simple_xml_sample} can be described using the XSD as follows.
+\begin{alltt}
+<xs:element name="book">
+  <xs:complexType>
+    <xs:sequence>
+      <xs:element name="title" type="xs:string"/>
+      <xs:element name="authors>
+        <xs:complexType>
+          <xs:element name="author" type="xs:string" maxOccurs="unbound"/>
+        </xs:complexType>
+      </xs:element>
+      <xs:element name="isbn" type="xs:string" minOccurs="0"/>
+    </xs:sequence>
+    <xs:attribute name="id" type="xs:string" use="required"/>
+  </xs:complexContent>
+</xs:element>
+\end{alltt}
+
+\todo[inline]{overit, zvalidovat}
 
 \section{XPath}
-\todo[inline]{TODO}
+XML Path Language (XPath, version 1.0) \cite{xpath} is a language to select fractions of XML documents using so-called path expressions.
+XPath considers an XML document as a tree of nodes.
+It recognizes seven types of nodes: document node, element node, attribute node, text node, namespace node, processing instruction node, and comment node.
+The root of the tree is the document node representing the entire XML document.
+
+\subsection{Path Expressions}
+A path expression is composed of individual path steps separated by \texttt{/} and can be absolute or relative.
+The absolute path begins with \texttt{/} representing the document node.
+The relative path needs a non-empty starting node-set to be evaluated.
+A path step is \\ \texttt{$axis$::$node\_test$[$predicate_1$]$\dots$[$predicate_n$]} \\
+where $axis$ and all predicates are optional.
+
+\subsection{Axes}
+An axis specifies a node-set relative to the current node.
+The default axis is \texttt{child} selecting all children nodes of the current node.
+Another important axes are \texttt{attribute}, selecting attributes of the current node, and \texttt{descendant-or-self} selecting the current node and all its descendants in the tree.
+The remaining axes are: \texttt{ancestor}, \texttt{ancestor-or-self}, \texttt{descendant}, \texttt{following}, \texttt{following-sibling}, \texttt{namespace}, \texttt{parent}, \texttt{preceding}, \texttt{preceding-sibling}, and \texttt{self}.
+
+\subsection{Node Tests}
+A node test identifies node(s) within all nodes selected by an axis. It can be a node type and/or node name. Examples of the node tests follow.
+\begin{itemize}
+\item \texttt{node()} All nodes selected by an axis.
+\item \texttt{text()} Text nodes.
+\item \texttt{*} All nodes assigned with their name (element and attributes).
+\end{itemize}
+
+\subsection{Abbrevations}
+Abbrevations for several most widely used construct are defined.
+\begin{itemize}
+\item \texttt{$P$/@$identifier$} stands for \texttt{$P$/atribute::$identifier$}
+\item \texttt{$P$/../$identifier$} stands for \texttt{$P$/parent::*/$identifier$}
+\item \texttt{$P$//$identifier$} stands for \texttt{$P$/descendant-or-self::node()/$identifier$} 
+\end{itemize}
+
+\subsection{Predicates}
+Predicates places additional conditions upon nodes that passed the node test.
+They can be either relative XPath paths or comparison expressions.
+The path predicates evaluate to true if they select a not empty set of nodes.
+Operators in the comparison expressions are \texttt{=}, \texttt{!=}, \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{\&eq;}, \texttt{\&ne;}, \texttt{\&lt;}, \texttt{\&gt;}, \texttt{\&le;}, and \texttt{\&ge;}.
+Operands are any XPath expressions (paths, literal values, etc).
+
+\subsection{Built-in Functions}
+XPath also provides a set of built-in functions.
+Few examples are \texttt{count($path$)}, returning a number of nodes selected by $path$, \texttt{position()}, returning the position of the current node in the current node-set, and \texttt{sum($path$)}, returning the sum of all nodes selected by $path$.
+
+\subsection{Usage Examples}
+Finally, we introduce several examples of XPath expressions, with their description, using the sample XML \ref{FIG_a_simple_xml_sample}.
+
+Path that selects \texttt{author} elements of the book with id $b1$:
+\begin{alltt}
+/bookstore/book[@id = "b1"]/authors/author
+\end{alltt}
+
+Expression returning a number of all books:
+\begin{alltt}
+count(//book)
+\end{alltt}
+
+Path selecting the ISBN element of the book titles $XQuery$:
+\begin{alltt}
+/bookstore/book[title = "XQuery"]/isbn
+\end{alltt}
+
+Path returning the titles of the books written by more than one author:
+\begin{alltt}
+/bookstore/book[count(authors/author) > 1]/title/text()
+\end{alltt}
+
+\todo[inline]{overit, otestovat}
 
 \section{XQuery}
-\todo[inline]{TODO}
+\todo[inline]{Tuto sekciu som vypracovaval primarne podla vasich slidov na prednasku Technologie XML. Mam to niekde uviest?}
+
+An XML Query Language (XQuery, version 1.0) \todo{ref} is a language designed to query XML data.
+It is based on XPath 2.0.
+XPath 2.0 is an extension of XPath 1.0 (but not entirely compatible), adding ordered sequences, their iterations, set operations, conditional expressions, quantified expressions, and XML Schema types.
+
+\subsection{Sequences}
+A sequence is an ordered set of items.
+The result of each XPath 2.0 (and also XQuery 1.0) path is a sequence.
+An item is either an atomic value or a node.
+An atomic value is a value of a XML Schema simple type.
+A node is an instance of one of the node types.
+
+Each node has its identity, data type (simple or complex, according to the XML Schema types), typed value (can be retrieved by function \texttt{fn:data()}, and string value (retrieved by function \texttt{fn:string()}.
+
+\subsection{FLWOR Expressions}
+A basic construct of the XQuery language is $FLWOR$. It is an abbrevation of its five clauses: for, let, where, order by, return. For clause is \\
+\texttt{for $var$ in $expr$} \\
+Expression $expr$ is evaluated and its result is a sequence. Items of the sequence are iteratively assigned to $var$ variable, which is valid also in the following clauses.
+
+Let clause is \\
+\texttt{let $var$ := $expr$} \\
+which evaluates $expr$ expression and its result is assigned to $var$ variable.
+
+Where clause is \\
+\texttt{where $expr$} \\
+Expression $expr$ can (and usually should) contain the variables from the for clause(s) and the remaining clauses are executed for only those tuples of values of the for variables, that the $expr$ evaluates to true.
+
+Order by clause is \\
+\texttt{order by $expr$)} \\
+and it orders the tuples of values that passed the where clause by the specified criterion.
+
+Return clause is \\
+\texttt{return $expr$} \\
+The result of the whole $FLWOR$ expression is $expr$. It can (and usually is) be constructed using the tuples of the for variable values and the let variables.
+
+\subsection{Conditional Expressions}
+
+\subsection{Quantified Expressions}
+
+\subsection{Built-in Functions}
+
+\subsection{User-defined Functions}
+
+\subsection{Value Comparison Expressions}
+
+\subsection{General Comparison Expressions}
+
+\subsection{Node Comparison Expressions}
+
+\subsection{Usage Examples}
+
+\todo[inline]{TODO XQuery Core?}
 

### 2011-12-19 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

@@
 
+In the remainder of this section, basic features of the XQuery languages are briefly described. Many query samples can be found in  chapter \ref{CHAPTER_analysis_of_xquery}.
+
 \subsection{Sequences}
@@
 \subsection{Conditional Expressions}
+XQuery provides common \texttt{if-than-else} conditional expressions with common syntax.
+\begin{alltt}
+if (\(condition\))
+than \(expr1\)
+else \(expr2\)
+\end{alltt}
+If $condition$ expression evaluates to true, $expr1$ is evaluated, else $expr2$ is evaluated. The else branch is optional.
 
 \subsection{Quantified Expressions}
+XQuery provides two types of quantified expressions, \texttt{every} and \texttt{some}. Their syntax follows.
+\begin{alltt}
+every \(var\) in \(expr1\)
+satisfies \(expr2\)
+\end{alltt}
+\begin{alltt}
+some \(var\) in \(expr1\)
+satisfies \(expr2\)
+\end{alltt}
 
-\subsection{Built-in Functions}
+Firstly, $expr1$ is evaluated, and then the result of the quantified expression is true if $expr2$ is true for every (some) item, represented by $var$, of the result of $expr1$ (which is a sequence).
 
-\subsection{User-defined Functions}
+\subsection{Functions}
+XQuery provides a wide set of built-in functions. \todo{ref?}
 
-\subsection{Value Comparison Expressions}
+And also new functions can defined using the following syntax.
+\begin{alltt}
+define function \(name\)(\(parameters\)) as \(type\)
+\{
+  \(expr\)
+\}
+\end{alltt}
+where $name$ is the function's name, $parameters$ is a list of parameters (with or without specification of their types), and $type$ is a type of a return value, which is the result of $expr$.
 
-\subsection{General Comparison Expressions}
+\subsection{Comparison Expressions}
+Comparison expressions are of three types: $value$ comparison expressions, $general$ comparison expressions, and $node$ comparison expressions.
+A very brief insight follows.
+More details can be found in REF\todo{ref, moze to takto ostat?}.
 
-\subsection{Node Comparison Expressions}
+The $value$ comparison expressions use operators \texttt{lt}, \texttt{gt}, \texttt{le}, \texttt{ge}, \texttt{eq}, and \texttt{ne} and they compare operands that are or can be converted to atomic values.
 
-\subsection{Usage Examples}
+The $general$ comparison expressions use operators \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, and \texttt{!=} and they can also compare sequences.
+
+And the $node$ comparison expressions use operators \texttt{is}, \texttt{<<}, and \texttt{>>} and they compare nodes, their identities and order.
 
@@
 
+\section{Other Technologies}
+
+\begin{description}
+\item[XSLT] e\textbf{X}tensible \textbf{S}tylesheet \textbf{L}anguage \textbf{T}ransformations \cite{Clark:99:XTV} is a language to transform input XML document(s) into other (not necessarily XML) document(s).
+\end{description}
+
 \todo[inline]{možná ňejaký automat apod. Prostě co budete dále v textu potřebovat. Skratky: XSD, XSLT, ...}

### 2011-12-20 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \begin{definition}
-Syntax tree of a XQuery query $Q$ is a tuple $T = (V, E, o)$ where $V$ is a set of nodes representing XQuery language constructs (TODO\todo{kompletny zoznam}), $E$ is a set of edges - couples $(a,b):a,b \in V$ and $o:V \rightarrow \mathbb{N}$ is an ordering function and $T$ satisfies the following conditions:
-\begin{itemize}
-\item $(V,E)$ is a tree.
-\item The root node is $ModuleNode$ representing the entire query Q.
-\item $(a,b)$ is an element of $E$ if and only if $b$ is a direct component of construct $a$ in $Q$.
-\item For every node $a \in V$ the ordering function $o$ is defined as follows \\* $o(a) = \left\{
-	\begin{array}{ll}
-		1  & \mbox{if }a\mbox{ is the root node} \\
-		seq(a) & \mbox{else}
-	\end{array}
-\right.$ \\* where $seq(a)$ is a sequential number of appearence of construct $a$ in $Q$.
-\end{itemize}
+Syntax tree of XQuery query $Q$ is tuple $T = (V, E, \mathcal{F})$ where $V$ is a set of nodes REF\todo{ref} representing the XQuery syntax constructs, $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ in query $Q$, and $\mathcal{F}$ is a set of functions specifying additional properties of the nodes.
+
+Since two different literal values in $Q$ are represented by the same item from $LiteralNode \in V$, it is needed to distinguish them using $\mathcal{F}$.
+Each literal value has a type and value.
+Therefore $\mathcal{F}$ contains functions $type_{LiteralNode}: V_{LiteralNode} \rightarrow Types_{literal}$ and $value_{LiteralNode}: V_{LiteralNode} \rightarrow Values_{literal}$, where $V_{LiteralNode}$ is set $\{v | v \in V, v = LiteralNode\}$, $Types_{literal}$ is set of all types of literal values $\{DECIMAL, INTEGER, DOUBLE, STRING\}$, and $Values_{literal}$ is a set of all literal values (all valid XQuery decimal numbers, integers, double numbers and strings).
+
+\todo[inline]{Podobne funkcie pre dalsie clenske premenne dalsich tried}
+Set $\mathcal{F}$ also contains functions that return children of particular nodes.
+For $FLWORExprNode$ it is function $whereClauseNode_{FLWORNodeExpr} : V_{FLWORNodeExpr} \rightarrow V$ where $V_{FLWORNodeExpr}$ is set $\{v | v \in V, v = FLWORNodeExpr\}$.
+For every $v \in V_{FLWORNodeExpr}: whereClauseNode_{FLWORNodeExpr}(v) = w$ so that $(v, w) \in E$ and $w = WhereClauseNode$.
+
+For $ParamListNode$, which can has multiple $ParamNode$ children, set $\mathcal{F}$ contains $paramOrder_{ParamListNode} : V_{ParamListNode} \rightarrow \mathcal{PO}$ where $V_{ParamListNode}$ is set $\{v | v \in V, v = ParamListNode\}$ and $\mathcal{PO}$ is set of functions $\{paramOrder_v : E_v \rightarrow \mathbb{N} | v \in V_{ParamListNode}, E_v = \{(v,w) | (v,w) \in E, w = ParamNode\}\}$.
+For every $v \in V_{ParamListNode}: paramOrder_{ParamListNode}(v) = paramOrder_v$ and for every $e = (v, w) \in E_v: paramOrder_v(e)$ is a sequential number of $w$ in children of $v$ in query $Q$.
 \end{definition}

### 2011-12-21 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

--- 2011-12-17 doc/theses/mikula/proposed_algorithm.tex
@@
 \section{Step 1: Construction of a Syntax Tree}
-The first step of the algorithm involves lexical and syntactic analyses known from compilers and produces a so-called syntax tree.
+The first step of the algorithm involves lexical and syntax analyses known from the construction of compilers and produces a so-called syntax tree.
 The analyses are taken from Ji\v r\'{i} Schejbal's master thesis \todo{ref}.
@@
 \subsection{Definition of the Syntax Tree}
-\begin{definition}
-Syntax tree of a XQuery query $Q$ is a tuple $T = (V, E, o)$ where $V$ is a set of nodes representing XQuery language constructs (TODO\todo{kompletny zoznam}), $E$ is a set of edges - couples $(a,b):a,b \in V$ and $o:V \rightarrow \mathbb{N}$ is an ordering function and $T$ satisfies the following conditions:
-\begin{itemize}
-\item $(V,E)$ is a tree.
-\item The root node is $ModuleNode$ representing the entire query Q.
-\item $(a,b)$ is an element of $E$ if and only if $b$ is a direct component of construct $a$ in $Q$.
-\item For every node $a \in V$ the ordering function $o$ is defined as follows \\* $o(a) = \left\{
-	\begin{array}{ll}
-		1  & \mbox{if }a\mbox{ is the root node} \\
-		seq(a) & \mbox{else}
-	\end{array}
-\right.$ \\* where $seq(a)$ is a sequential number of appearence of construct $a$ in $Q$.
-\end{itemize}
-\end{definition}
+Firstly, we formally define the syntax tree.
 
-\subsection{Syntax Tree Node Types}
+\begin{define}[Syntax tree]
+Syntax tree of XQuery query $Q$ is tuple $T = (V, E, c, \mathcal{P}, \mathcal{O})$ where $V \subset \mathbb{N}$ is a set of nodes, each node representing a particular XQuery construct in query $Q$, $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ in query $Q$, $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9} (those with non-bolded names), $\mathcal{P}$ is a set of functions specifying additional properties of the nodes, and $\mathcal{O}$ is a set of functions specifying an order of children of particular nodes.
+\end{define}
+
+Regarding the additional properties, two constructs in $Q$ represented by two nodes of the same class from $C$ may differ in certain ways, and therefore, it is needed to distinguish them.
+For instance, two different literal values in $Q$ are represented by nodes $l_1, l_2 \in V$ and $c(l_1) = c(l_2) = LiteralNode$ but each has a different value and type.
+Therefore $\mathcal{P}$ contains functions \\
+$type_{LiteralNode}: V_{LiteralNode} \rightarrow Types_{literal}$ \\
+$value_{LiteralNode}: V_{LiteralNode} \rightarrow Values_{literal}$ \\
+where $V_{LiteralNode}$ is set $\{v | v \in V, c(v) = LiteralNode\}$, $Types_{literal}$ is set of all types of literal values $\{DECIMAL, INTEGER, DOUBLE, STRING\}$, and $Values_{literal}$ is a set of all literal values (all valid XQuery decimal numbers, integers, double numbers and strings).
+
+Set $\mathcal{P}$ contains other similar functions but due to their large count, we do not define them formally.
+\todo[inline]{Moze takto ostat, alebo definovat aspon tie co pouzivame? Alebo ich aspon vypisat?}
+
+Set $\mathcal{P}$ also contains functions that return children of particular nodes.
+For instance, in every syntax tree, for every $v \in V, c(v) = FLWORExprNode$ it exists $w = \in V, c(w) = WhereClauseNode$ so that $(v,w) \in E$.
+In other words, every $FLWORExprNode$ node has a $WhereClauseNode$ node amongst its children.
+Set $\mathcal{P}$ contains function \\
+$whereClauseNode_{FLWORExprNode} : V_{FLWORExprNode} \rightarrow V$ \\ 
+where $V_{FLWORExprNode}$ is set $\{v | v \in V, c(v) = FLWORExprNode\}$.
+For every $v \in V_{FLWORExprNode}: whereClauseNode_{FLWORExprNode}(v) = w$ so that $(v, w) \in E$ and $c(w) = WhereClauseNode$.
+Function $whereClauseNode_{FLWORExprNode}$ is a convenient function to specify a $WhereClauseNode$ child of a given $FLWORExprNode$ node.
+
+Set $\mathcal{O}$ contains functions specifying an order of children of a some nodes.
+Those nodes are nodes with all children of the same class from $C$.
+For instance, $ParamListNode$ can has multiple $ParamNode$ children.
+Set $\mathcal{O}$ contains function \\
+$paramOrder_{ParamListNode} : V_{ParamListNode} \rightarrow \mathcal{O}_P$ \\
+where $V_{ParamListNode}$ is set $\{v | v \in V, c(v) = ParamListNode\}$ and $\mathcal{O}_P$ is set of functions $\{paramOrder_v : E_v \rightarrow \mathbb{N} | v \in V_{ParamListNode}, E_v = \{(v,w) | (v,w) \in E, c(w) = ParamNode\}\}$.
+For every $v \in V_{ParamListNode}: paramOrder_{ParamListNode}(v) = paramOrder_v$ and for every $e = (v, w) \in E_v: paramOrder_v(e)$ is a sequential number of $w$ amongst children of $v$ in query $Q$.
+In other words, it represents the order of the children as the respective constructs appear in the query.
+
+\subsection{Syntactic Abbrevations}
+Since we need to use the syntax tree in pseudo-algorithms, we define several abbrevations to make its usage more simplier.
+
+Assuming, $v \in V$
+$v.X$ is abbrevation for $X(v)$
+
+\subsection{Closer Look at the Nodes of the Syntax Tree}
 \todo[inline]{Dat figures do appendixu?}
 The node types (or classes) of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where a node can be of several types.
-This hierarchy is shown in figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}.
+This hierarchy is shown in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}.
 Nodes with emphasised names express a common type of a group of nodes - their descendants, and these type nodes cannot be directly used in the syntax tree (abstract classes).

### 2011-12-25 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\todo[inline]{Toto vše je OK, ale asi budeme potřebovat vědět, co nám z té analýzy leze za informace se kterými dál pracujeme.
-Čili vstupy našeho algoritmu je potřeba popsat přesně, ideálně formálně, tj.
-je to asi nějaký strom T= (V, E), kde V je...
-atd.
-Jeslti už je to pak implementováno jako objekty v paměti nebo XML data je celkem jedno a bude popsáno v kapitola Implementation.}
-
 \subsection{Definition of the Syntax Tree}
@@
 \begin{define}[Syntax tree]
-Syntax tree of XQuery query $Q$ is tuple $T = (V, E, c, \mathcal{P}, \mathcal{O})$ where $V \subset \mathbb{N}$ is a set of nodes, each node representing a particular XQuery construct in query $Q$, $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ in query $Q$, $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9} (those with non-bolded names), $\mathcal{P}$ is a set of functions specifying additional properties of the nodes, and $\mathcal{O}$ is a set of functions specifying an order of children of particular nodes.
+\label{DEF_syntax_tree}
+Syntax tree of XQuery query $Q$ is tuple $T = (V, E, c, \mathcal{P}, o)$ where $V \subset \mathbb{N}$ is a set of nodes, each node representing a particular XQuery construct in query $Q$, $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ ($b$ is a child of $a$) in query $Q$, $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}, $\mathcal{P}$ is a set of functions specifying additional properties of the nodes and distinguishing the nodes of the same classes, and $o : V \rightarrow \mathcal{O}$ is a function specifying an order of children of the nodes where $\mathcal{O} = \{o_v : E_v \rightarrow \mathbb{N} | v \in V, E_v = \{(v, w) | w \in V, (v,w) \in E\}\}$ is set of functions specifying the children order for each node.
+For every $v \in V, o(v) = o_v$ so that $o_v(v,w)$ is a sequential number of a construct represented by $w$ amongst constructs represented by children of $v$ in query $Q$.
 \end{define}
@@
 
-Set $\mathcal{P}$ also contains functions that return children of particular nodes.
-For instance, in every syntax tree, for every $v \in V, c(v) = FLWORExprNode$ it exists $w = \in V, c(w) = WhereClauseNode$ so that $(v,w) \in E$.
-In other words, every $FLWORExprNode$ node has a $WhereClauseNode$ node amongst its children.
-Set $\mathcal{P}$ contains function \\
-$whereClauseNode_{FLWORExprNode} : V_{FLWORExprNode} \rightarrow V$ \\ 
-where $V_{FLWORExprNode}$ is set $\{v | v \in V, c(v) = FLWORExprNode\}$.
-For every $v \in V_{FLWORExprNode}: whereClauseNode_{FLWORExprNode}(v) = w$ so that $(v, w) \in E$ and $c(w) = WhereClauseNode$.
-Function $whereClauseNode_{FLWORExprNode}$ is a convenient function to specify a $WhereClauseNode$ child of a given $FLWORExprNode$ node.
+\subsection{Syntactic Abbreviations}
+Since we need to use the syntax tree in pseudo-algorithms, we define the following abbreviations to make its usage more simple.
 
-Set $\mathcal{O}$ contains functions specifying an order of children of a some nodes.
-Those nodes are nodes with all children of the same class from $C$.
-For instance, $ParamListNode$ can has multiple $ParamNode$ children.
-Set $\mathcal{O}$ contains function \\
-$paramOrder_{ParamListNode} : V_{ParamListNode} \rightarrow \mathcal{O}_P$ \\
-where $V_{ParamListNode}$ is set $\{v | v \in V, c(v) = ParamListNode\}$ and $\mathcal{O}_P$ is set of functions $\{paramOrder_v : E_v \rightarrow \mathbb{N} | v \in V_{ParamListNode}, E_v = \{(v,w) | (v,w) \in E, c(w) = ParamNode\}\}$.
-For every $v \in V_{ParamListNode}: paramOrder_{ParamListNode}(v) = paramOrder_v$ and for every $e = (v, w) \in E_v: paramOrder_v(e)$ is a sequential number of $w$ amongst children of $v$ in query $Q$.
-In other words, it represents the order of the children as the respective constructs appear in the query.
+For every $v \in V$, abbreviation
+\begin{itemize}
+\item $v.p$ stands for $p_{c(v)}(v)$ if $p_{c(v)} \in \mathcal{P}$.
+\item $v.getChildren()$ stands for $\{w | w \in V, (v,w) \in E\}$ which is a set of all children of $v$.
+\item $v.getChild(i)$, $i \in \mathbb{N} \cup \{0\}$, stands for $w \in V$ so that $o(v)(v,w) = i + 1$.
+\item $v.getChild(class), class \in C$, stands for $w \in V$ so that $(v,w) \in E$ and $c(w) = class$, if $\Bigl | \{u | u \in V, (v,u) \in E, c(u) = class\}\Bigr | = 1$.
+\end{itemize}
 
-\subsection{Syntactic Abbrevations}
-Since we need to use the syntax tree in pseudo-algorithms, we define several abbrevations to make its usage more simplier.
-
-Assuming, $v \in V$
-$v.X$ is abbrevation for $X(v)$
+\todo[inline]{Je nutne presnejsie popisat kedy a preco tieto skratky maju zmysel? V pseudo-kodoch budu samozrejme pouzivane len v pripadoch, ked su dobre definovane a maju zmysel.}
 
@@
 \todo[inline]{Dat figures do appendixu?}
-The node types (or classes) of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where a node can be of several types.
+The node classes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where an object can be of several types.
 This hierarchy is shown in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}.
-Nodes with emphasised names express a common type of a group of nodes - their descendants, and these type nodes cannot be directly used in the syntax tree (abstract classes).
-The remaining nodes represent particular constructs of the XQuery language and their instances can be used as nodes of the syntax tree.
-Each of the instances additionaly contains details about the particular form of the respective XQuery construct used in the query.
+The classes with emphasised names represent a common class of a group of subclasses, and these classes cannot be directly used in the syntax tree (abstract classes).
+The remaining nodes represent particular constructs of the XQuery language and nodes of these classes can be used in the syntax tree.
 
-\todo[inline]{V ExprNode s ExprHolderNode si pamatame aj typ vyrazu}
-
-For example, an instance of the syntax tree cannot directly contain instances of nodes $Node$ and $ExprNode$, but it can contain instances of nodes $AttributeNode$ and $LiteralNode$.
-Regarding the types, an instance of $LiteralNode$ node is of three types: $LiteralNode$, $ExprNode$ and $Node$.
-It carries information on a type of the literal and its value, for example $xs:integer$ of value $5$.
+For example, an instance of the syntax tree cannot directly contain nodes of $Node$ and $ExprNode$ classes (for every $v \in V, c(v) \neq Node, c(v) \neq ExprNode$), but it can contain nodes of $AttributeNode$ and $LiteralNode$ classes.
+Regarding the multiplicity of types, a node of $LiteralNode$ class is also considered to be of two indirect types: $ExprNode$ and $Node$.
 
@@
 
-Nodes can be classified into three groups: inner nodes, leaf nodes and nodes that can be both inner and leaf.
-Inner nodes (marked $I$) stand for XQuery constructs that are composed of other constructs and can be further divided.
-An example of such node is $FLWORExprNode$ which is composed of nodes $TupleStreamNode$, $WhereClauseNode$, $OrderByClauseNode$ and $ReturnClauseNode$.
-Leaf nodes (marked $L$) represent elements of XQuery language that cannot be further divided.
+The node classes can be classified into three groups: inner node classes, leaf node classes and node classes that can be both inner and leaf.
+Inner node classes (marked $I$) stand for XQuery constructs that are composed of other constructs and can be further divided.
+An example of such class is $FLWORExprNode$ which is composed of $TupleStreamNode$, $WhereClauseNode$, $OrderByClauseNode$ and $ReturnClauseNode$ classes.
+Leaf node classes (marked $L$) represent elements of XQuery language that cannot be further divided.
 For example, $LiteralNode$.
 Node $FunctionCallNode$ is an example of the third group (marked $IL$).
-Function call of a function without arguments is represented by a leaf while function call with arguments is represented by an inner node and its subnodes are nodes representing those arguments.
+Function call of a function without arguments is represented by a leaf node while function call with arguments is represented by an inner node and its subnodes are nodes representing those arguments.
 
@@
 
-Some pseudo code algorithms in the following sections need to determine the type of a node instance.
-For this purpose, we define function $isA(nodeInstance, nodeType)$.
-The first argument is an instance of any node and the second is a node type itself.
-The function returns boolean value $true$ if the instance is of the specified node type, according to the described principle.
+Some pseudo code algorithms in the following sections need to determine a class of a node.
+The node's direct class can be determined by function $c$ from the definition of the syntax tree \ref{DEF_syntax_tree}, and hence, we also can determine its indirect classes.
+For the purpose of pseudo code, for every $v \in V$ and every $class \in C$, we define function $is(v, class)$.
+It returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
 Otherwise, it returns $false$.
 
-For instance, assuming variable $literal$ is an instance of node $LiteralNode$, calls $isA(literal, LiteralNode)$, $isA(literal, ExprNode)$ return $true$, while call $isA(literal, ContentNode)$ returns $false$.
+For instance, assuming $v \in V, c(v) = LiteralNode$, calls $is(v, LiteralNode)$, $is(v, ExprNode)$ return $true$, while call $is(v, ContentNode)$ returns $false$.
 
-\subsection{Description of the Syntax Tree}
-An important characteric of the syntax tree is related to a definition of local variables and their scope.
-The representation of a definition of a local variable is $VariableBinding$ node.
-Nodes of this type contain only two subnodes; a type of the variable and a binding expression.
-The node and its subtree representing the type do not contain any expressions.
-The binding expression node is an expression node but the represented expression cannot use the variable which is just being bound.
-Therefore the scope of the new variable is not the subtree of the $VariableBinding$ node.
-Assuming \todo{Toto je velmi high-level popis na to, jak málo o syntaktickém stromě a dalších položkách víme.} the post-order numbering of the nodes, the new variable may be valid in some nodes with a higher post-order sequence number.
-It depends on the type of the ancestor node(s) which nodes that are.
+\subsection{Characteristics of the Syntax Tree}
+An important characteristic of the syntax tree is related to a definition of local variables and their scope in the XQuery language.
+The representation of a definition of a local variable in the syntax tree is a node of $VariableBinding$ class.
+Nodes of that class contain have only two children; a node representing the type of the variable and a node representing the binding expression (expression defining the value of the variable, and thus, it cannot use the variable).
+Hence, the entire subtree doesn't contain any expressions that use the variable.
+Therefore the scope of the new variable is not the subtree of the $VariableBinding$ class node.
+It depends on the type of XQuery construct that the variable binding is a (indirect) component of.
 
-For example, $FLWORExpr$ node contains four subnodes; $TupleStream$, $WhereClause$, $OrderByClause$ and $ReturnClause$.
-$TupleStream$ node contains a list of $VariableBinding$ node which define variables valid in all other three subnodes of the $FLWORExpr$ node.
+For example, a node of $FLWORExpr$ class contains four subnodes of $TupleStream$, $WhereClause$, $OrderByClause$ and $ReturnClause$ classes.
+The $TupleStream$ class node contains a list of node of $variableBinding$ class which define variables valid in all other three subnodes of the FLWOR node.
+
+\todo[inline]{Nejak sa z tohoto vytratilo to, preco to je pre nas dolezita charakteristika}
 
@@
 
-For simplicity, we will focus on locally defined functions.
-We will not determine types of functions defined in other modules since the principle is similar but it is somewhat more difficult to implement \todo{Q: je to ok?
-A: No spíš napište něco jako without loss of generality.}.
-Return types of built-in functions are fixed and, thus, there is no need to analyze them.
+Without loss of generality, we will focus on locally defined functions.
+We will not determine types of functions defined in other modules since the principle is similar but it is needed to look up the definitions in syntax trees of other queries.
+Return types of built-in functions are fixed and, thus, there is no need to analyse them.
 

### 2011-12-25 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
+\begin{algorithm}
+\caption{Retrieval of Function Declaration Nodes $getFunctionDeclarationNodes(syntaxTree)$}
+\label{ALG_retrieval_of_function_declaration_nodes}
+\begin{algorithmic}[1]
+\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
+\ENSURE A list of syntax tree nodes representing function declarations.
+
+\STATE $prologNode = null$
+\FORALL{$moduleChildNode \in syntaxTree.getChildren()$}
+    \IF{$is(moduleChildNode, PrologNode)$}
+        \STATE $prologNode = moduleChildNode$
+    \ENDIF
+\ENDFOR
+
+\STATE $functionDeclarationNodes =$ an empty list
+\IF{$prologNode \neq null$}
+    \FORALL{$prologChildNode \in prologNode.getChildren()$}
+        \IF{$is(prologChildNode, FunctionDeclNode)$}
+            \STATE add $prologChildNode$ to $functionDeclarationNodes$
+        \ENDIF
+    \ENDFOR
+\ENDIF
+
+\RETURN $functionDeclarationNodes$
+\end{algorithmic}
+\end{algorithm}
+
 A simple algorithm is presented in Algorithm \ref{ALG_determination_of_function_return_types} \todo{Q: uvadzat aj taketo jednoduche funkcie?
@@
 opakuje, tak stačí napsat, že je algoritmus podobný jako v předchozím případě nebo tak něco.}.
-It uses \texttt{getFunctionDeclarationNodes()} function, which returns a list of all function declaration nodes in the syntax three.
+It uses \texttt{getFunctionDeclarationNodes()} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three.
 Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.
@@
 \begin{algorithmic}[1]
-\ENSURE An associative array of function names with their types.
+\REQUIRE{\ \\
+    $getFunctionDeclarationNodes()$: Function defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}. \\
+    $syntaxTree$: A reference to the root node of a syntax tree.
+}
+\ENSURE An associative array of function names with their types and references to the their declaration nodes.
 
-\STATE $functionTypes$ = empty associative array
-\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes()$}
-    \STATE $functionName = functionDeclarationNode.getFunctionName()$
-    \STATE $functionTypes[functionName] = functionDeclarationNode.getReturnType()$
+\STATE $functionArray$ = an empty associative array
+\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes(syntaxTree)$}
+    \STATE $functionName = functionDeclarationNode.funcName$
+    \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
+    \STATE $functionArray[functionName] = (PROCESS\_TYPE(typeNode), functionDeclarationNode)$
 \ENDFOR
-\RETURN $functionTypes$
+\RETURN $functionArray$
 \end{algorithmic}
@@
 
-For the rest of the thesis, we assume function \todo{Většinou se na začátek sekce s algoritmem dává seznam používaných funkcí a jejich význam.} \texttt{getFunctionReturnType()} which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
+\todo[inline]{Definovat a premenovat PROCESS\_TYPE() v Determination of Function Return Types. V skratke robi to, ze instanciu TypeNode prevediu na nasu reprezentaciu typu.}
+
+For the rest of the thesis, we assume function \todo{Většinou se na začátek sekce s algoritmem dává seznam používaných funkcí a jejich význam.} $getFunctionReturnType()$ which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
 Otherwise, \texttt{null} is returned.
 
-Later phases of the algorithm may need to process the function declarations. Therefore, we also store references to the entire function declaration nodes.
+Later phases of the algorithm may need to process the function declarations.
+Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.
 
@@
 \STATE $type = null$
-\IF{$isA(exprNode, LiteralNode)$}
-     \STATE $type =$ ((LiteralNode)exprNode).type, exactly-once cardinality
-\ELSIF{$isA(exprNode, FunctionCallNode)$}
-     \STATE $type = getFunctionReturnType(((FunctionCallNode)exprNode).functionName)$
-\ELSIF{$isA(exprNode, VarRefNode)$}
-     \STATE $type = getVariableType(contextVarTypes, ((VarRefNode)exprNode).variableName)$
-\ELSIF{$isA(exprNode, PathExprNode)$}
-     \STATE $path = ((PathExprNode)exprNode).path$
+\IF{$is(exprNode, LiteralNode)$}
+     \STATE $type = exprNode.type$, exactly-once cardinality
+\ELSIF{$is(exprNode, FunctionCallNode)$}
+     \STATE $type = getFunctionReturnType(exprNode.functionName)$
+\ELSIF{$is(exprNode, VarRefNode)$}
+     \STATE $type = getVariableType(contextVarTypes, exprNode.variableName)$
+\ELSIF{$is(exprNode, PathExprNode)$}
+     \STATE $path = exprNode.path$
      \IF{$path$ selects an element set}
-         \STATE $type =$ element type, zero-or-more cardianlity, $path$ included
+         \STATE $type =$ element type, zero-or-more cardinality, $path$ included
      \ELSIF{$path$ selects an attribute set}
-         \STATE $type =$ attribute type, zero-or-more cardianlity, $path$ included
+         \STATE $type =$ attribute type, zero-or-more cardinality, $path$ included
      \ELSE
@@
      \ENDIF
-\ELSIF{$isA(exprNode, OperatorNode)$}
+\ELSIF{$is(exprNode, OperatorNode)$}
      \STATE $type = determineOperatorType(exprNode, contextVarTypes)$
-\ELSIF{$isA(exprNode, FLWORExprNode)$}
-     \STATE $returnClauseNode = ((FLWORExprNode)exprNode).returnClauseNode$
+\ELSIF{$is(exprNode, FLWORExprNode)$}
+     \STATE $returnClauseNode = exprNode.returnClauseNode$
      \STATE $type = returnClauseNode.type$

### 2011-12-26 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 The node's direct class can be determined by function $c$ from the definition of the syntax tree \ref{DEF_syntax_tree}, and hence, we also can determine its indirect classes.
-For the purpose of pseudo code, for every $v \in V$ and every $class \in C$, we define function $is(v, class)$.
-It returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
+For the purpose of pseudo code, we define the following function.
+For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
 Otherwise, it returns $false$.
@@
 \item XML Schema built-in atomic types \todo{link do appendixu}.
-\item Type $xs:untyped$. \todo{to je co?}
-\item Type representing any item. \todo{je potrebne?}
-\item Type representing an element, attribute, item, node, text node, comment, processing instruction, document node.
-\item Type representing a set of elements or attributes selected by a certain path expression. The path expression is included in this type. Let this type be identified as $pathType$.
+\item Types $ElementType$, $AttributeType$, $NodeType$, $TextNodeType$, $CommentType$, $ProcessingInstructionType$, $DocumentType$ representing an element, attribute, node, text node, comment, processing instruction, document node.
+\item Type representing a node selected by a certain path expression. The path expression is included in this type. Let this type be identified as $PathType$.
 \end{itemize}
-
-\todo[inline]{Q: Akeho typu je napr path expression /root/(elementA * 2)? A: Datový typ je daný operátorem. Hodnota elementu se přetypuje.}
 
@@
 \end{itemize}
+
+$PathType$ has always cardinality of zero-or-more, because the result of a path expression is a general sequence and since we do not evaluate the queries we cannot determine its actual cardinality.
 
@@
 
-A simple algorithm is presented in Algorithm \ref{ALG_determination_of_function_return_types} \todo{Q: uvadzat aj taketo jednoduche funkcie?
+A simple algorithm is presented in Algorithm \ref{ALG_processing_of_functions} \todo{Q: uvadzat aj taketo jednoduche funkcie?
 A: Obecně by mělo být popsáno vše.
@@
 \begin{algorithm}
-\caption{Determination of Function Return Types}
-\label{ALG_determination_of_function_return_types}
+\caption{Processing of Functions}
+\label{ALG_processing_of_functions}
 \begin{algorithmic}[1]
@@
     $getFunctionDeclarationNodes()$: Function defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}. \\
+    $getTypeTN()$: Function defined in Algorithm \ref{ALG_extraction_of_a_type_from_typenode}. \\
     $syntaxTree$: A reference to the root node of a syntax tree.
@@
     \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
-    \STATE $functionArray[functionName] = (PROCESS\_TYPE(typeNode), functionDeclarationNode)$
+    \STATE $functionArray[functionName] = (getTypeTN(typeNode), functionDeclarationNode)$
 \ENDFOR
@@
 
-\todo[inline]{Definovat a premenovat PROCESS\_TYPE() v Determination of Function Return Types. V skratke robi to, ze instanciu TypeNode prevediu na nasu reprezentaciu typu.}
+\begin{algorithm}
+\caption{Extraction of a Type from TypeNode getTypeTN()}
+\label{ALG_extraction_of_a_type_from_typenode}
+\begin{algorithmic}[1]
+\REQUIRE $typeNode$: Syntax tree node of TypeNode class.
+\ENSURE Type extracted from $typeNode$.
 
-For the rest of the thesis, we assume function \todo{Většinou se na začátek sekce s algoritmem dává seznam používaných funkcí a jejich význam.} $getFunctionReturnType()$ which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
+\STATE $type = null$
+\STATE $cardinality = typeNode.cardinality$
+\STATE $itemTypeNode = typeNode.getChild(ItemTypeNode)$
+\IF{$is(itemTypeNode, AtomicTypeNode)$}
+    \STATE $type = itemTypeNode.typeName$
+\ELSIF{$is(itemTypeNode, KindTestNode)$}
+    \STATE $type = itemTypeNode.nodeKind$
+\ENDIF
+
+\RETURN $type$, $cardinality$
+\end{algorithmic}
+\end{algorithm}
+
+\todo[inline]{ALG Extraction of a Type from TypeNode nie uplne sedi s definiciou typov. DP JS ma nieco navyse.}
+
+For the rest of the thesis, we assume function $getFunctionReturnType()$ which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
 Otherwise, \texttt{null} is returned.
 
-Later phases of the algorithm may need to process the function declarations.
+Later phases of the algorithm may need to process the declarations of certain functions.
 Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.
@@
 
+\todo[inline]{ALG \ref{ALG_func_determineExpressionType}: Vymysliet a dopisat ako ziskavat typ z PathExprNode. Bude potrebne vyskusat implementaciu, ako sa chova.}
+
 \begin{algorithm}
-\caption{Function determineExpressionType}
+\caption{Function determineExpressionType()}
 \label{ALG_func_determineExpressionType}
@@
 \REQUIRE{\ \\
+    $getFunctionReturnType()$: Function returning a type of a function. \\
+    $getVariableType()$: Function returning a type of a variable. \\
+    $determineOperatorType()$: Function defined in \ref{ALG_func_determineOperatorType}. \\
+    $getType()$: Function returning a type of already processed expression. \\
     $exprNode$: An expression node in the syntax tree. \\
@@
 \IF{$is(exprNode, LiteralNode)$}
-     \STATE $type = exprNode.type$, exactly-once cardinality
+     \STATE $type =$ XSD atomic type $exprNode.type$, exactly-once cardinality
 \ELSIF{$is(exprNode, FunctionCallNode)$}
-     \STATE $type = getFunctionReturnType(exprNode.functionName)$
+     \STATE $type = getFunctionReturnType(exprNode.fncName)$
 \ELSIF{$is(exprNode, VarRefNode)$}
-     \STATE $type = getVariableType(contextVarTypes, exprNode.variableName)$
+     \STATE $type = getVariableType(contextVarTypes, exprNode.varName)$
 \ELSIF{$is(exprNode, PathExprNode)$}
-     \STATE $path = exprNode.path$
-     \IF{$path$ selects an element set}
-         \STATE $type =$ element type, zero-or-more cardinality, $path$ included
-     \ELSIF{$path$ selects an attribute set}
-         \STATE $type =$ attribute type, zero-or-more cardinality, $path$ included
-     \ELSE
-         \STATE $type = null$
+     \IF{$exprNode$ selects an element set}
+         \STATE $type = ElementType$, zero-or-more cardinality, $exprNode$ included
+     \ELSIF{$exprNode$ selects an attribute set}
+         \STATE $type = AttributeType$, zero-or-more cardinality, $exprNode$ included
      \ENDIF
@@
 \ELSIF{$is(exprNode, FLWORExprNode)$}
-     \STATE $returnClauseNode = exprNode.returnClauseNode$
-     \STATE $type = returnClauseNode.type$
+     \STATE $returnClauseNode = exprNode.getChild(ReturnClauseNode)$
+     \STATE $type = getType(returnClauseNode)$
 \ENDIF
@@
 
-\todo[inline]{Napisat, ze determineOperatorType() tiez nie je kompletna}
-\todo[inline]{definovat isOperatorClassAddition()}
-\todo[inline]{definovat isNumericType()}
-\todo[inline]{definovat isOperatorClassComparison()}
-
 \begin{algorithm}
-\caption{Function determineOperatorType}
+\caption{Function determineOperatorType()}
 \label{ALG_func_determineOperatorType}
@@
 \REQUIRE{\ \\ 
+    $getType()$: Function returning a type of already processed expression. \\
+    $isOperatorClassComparison()$: Function defined in \ref{ALG_func_isOperatorClassComparison}. \\
+    $isOperatorClassAddition()$: Function defined in \ref{ALG_func_isOperatorClassAddition}. \\
+    $isNumericType()$: Function defined in \ref{ALG_func_isNumericType}. \\
+    $selectMoreGenericNumericType()$: Function defined in \ref{ALG_func_selectMoreGeneralNumericType}. \\
     $operatorNode$: An operator expression node in the syntax tree. \\
@@
 \IF{$isOperatorClassComparison(operator)$}
-    \STATE $type = xs:boolean$
-\ELSIF{$isOperatorClassAddition(operator) AND$ one of the operands is of a numeric type}
-    \STATE $leftOperand = ((OperatorNode)exprNode).leftOperand$
-    \STATE $rightOperand = ((OperatorNode).exprNode).rightOperand$
-    \IF{$isNumericType(leftOperand.type) AND isNumericType(rightOperand.type)$}
-        \STATE $type = selectMoreGeneralType(leftOperand.type, rightOperand.type)$
-    \ELSIF{$isNumericType(leftOperand.type)$}
-        \STATE $type = leftOperand.type$
-    \ELSIF{$isNumericType(rightOperand.type)$}
-        \STATE $type = rightOperand.type$
+    \STATE $type = xs:boolean$, exactly-one cardinality
+\ELSIF{$isOperatorClassAddition(operator)$}
+    \STATE $leftOperandType = getType(exprNode.leftSide)$
+    \STATE $rightOperandType = getType(exprNode.rightSide)$
+    \IF{$isNumericType(leftOperandType)) AND isNumericType(rightOperandType)$}
+        \STATE $type = selectMoreGeneralNumericType(leftOperandType, rightOperandType)$
+    \ELSIF{$isNumericType(leftOperandType)$}
+        \STATE $type = leftOperandType$
+    \ELSIF{$isNumericType(rightOperandType)$}
+        \STATE $type = rightOperandType$
     \ENDIF
@@
 
+\begin{algorithm}
+\caption{Function isOperatorClassComparison()}
+\label{ALG_func_isOperatorClassComparison}
+\begin{algorithmic}[1]
+\REQUIRE $operator$: Representation of a XQuery operator.
+\ENSURE $true$ if $operator$ is a comparison operator, $false$ otherwise.
+
+\IF{$operator$ equals one of $GEN\_EQUALS$, $GEN\_GREATER\_THAN$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_LESS\_THAN$, $GEN\_LESS_THAN\_EQUALS$, $GEN\_NOT\_EQUALS$, $VAL\_EQUALS$, $VAL\_GREATER\_THAN$, $VAL\_GREATER\_THAN\_EQUALS$, $VAL\_LESS\_THAN$, $VAL\_LESS\_THAN\_EQUALS$, $VAL\_NOT\_EQUALS$}
+    \RETURN $true$
+\ELSE
+    \RETURN $false$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function isOperatorClassAddition()}
+\label{ALG_func_isOperatorClassAddition}
+\begin{algorithmic}[1]
+\REQUIRE $operator$: Representation of a XQuery operator.
+\ENSURE $true$ if $operator$ is an addition operator, $false$ otherwise.
+
+\IF{$operator$ equals one of $PLUS$, $MINUS$}
+    \RETURN $true$
+\ELSE
+    \RETURN $false$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function isNumericType()}
+\label{ALG_func_isNumericType}
+\begin{algorithmic}[1]
+\REQUIRE $type$: Representation of a type.
+\ENSURE $true$ if $type$ represents one of the XSD atomic built-in numeric types, $false$ otherwise.
+
+\IF{$type$ represents one of $float$, $double$, $decimal$, $integer$, $long$, $int$, $short$, $byte$, $nonPositiveInteger$, $negativeInteger$, $nonNegativeInteger$, $positiveInteger$, $unsignedLong$, $unsignedInt$, $unsignedShort$, $unsignedByte$}
+    \RETURN $true$
+\ELSE
+    \RETURN $false$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function selectMoreGeneralNumericType()}
+\label{ALG_func_selectMoreGeneralNumericType}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $numericType1$: Numeric type one.
+    $numericType2$: Numeric type two.
+}
+\ENSURE Returns the more generic numeric type of the two specified.
+\STATE TODO
+\end{algorithmic}
+\end{algorithm}
+
 The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted.
-Otherwise, an attempt of the type deduction of the expression is made.
+Otherwise, an attempt of the type deduction of the binding expression is made.
 The deduction of the type from the expression is presented in Algorithm \ref{ALG_func_determineExpressionType}.
@@
 Nevertheless, the principle is straightforward.
-Depending on the class of the expression, its type can be either determined directly or it depends on its subexpressions.
-\todo{link na mozne triedy vyrazov.}
+Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_4}), its type can be either determined directly or it depends on its subexpressions.
+A part of the function's semantic is separated in \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
 
@@
 
-The order of the recursion is post-order.
+The order of the recursion is to firstly process children of a node and then the node itself.
+The order of children processing is specified by function $o$ from the definition of the syntax tree \ref{DEF_syntax_tree}.
 The reason is that an expression node may need to know types of its subexpressions to determine its own type and we want to determine types of all expressions.
-Therefore, using the post-order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion, because the types of the subexpressions are already determined.
+Therefore, using the described order, subexpressions of an expression are processed first, and then the expression itself without any need for further recursion, because the types of the subexpressions are already determined.
 
-The handling of new variable definitions relies on the fact that the definitions make the new variables valid only in nodes with a higher sequence number when numbered in the post-order \todo{overit}.
-Therefore, the left-most subnode of a certain node can be processed without an extension of the variable context.
-And, if the left-most subnode extends the variable context for the following nodes in the post-order numbering, it can be easily handled because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
+The handling of new variable definitions relies on the fact that the definitions make the new variables valid only in nodes with a higher sequence number when numbered in the order described in in the previous paragraph.
+Therefore, the left-most (the first amongst the ordered subnodes) subnode of a certain node can be processed without an extension of the variable context.
+And, if the left-most subnode extends the variable context for the following nodes in the numbering, it can be easily handled because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
 Every node ''knows'' wheather or not it may define new variables for nodes with higher sequence numbers.
@@
 This is done at line 4.
-Function $mergeContextVarTypes$ presented in Algorithm \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into ist first argument (the variable context).
-If there are records for variables with the same names in the context, they are overwritten to correspond to the variable overlaping.
-It is important to note that we assume the variable context is not passed by refence but by value, and hence, the variable context of a certain node is not affected by the recursive processing of its subnodes.
+Function $mergeContextVarTypes$ presented in Algorithm \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into its first argument (the variable context).
+If there are records for variables with the same names in the context, they are overwritten to correspond to the variable overlapping.
+It is important to note that we assume the variable context is not passed by reference but by value, and hence, the variable context of a certain node is not affected by the recursive processing of its subnodes.
 

### 2011-12-28 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
     $isNumericType()$: Function defined in \ref{ALG_func_isNumericType}. \\
-    $selectMoreGenericNumericType()$: Function defined in \ref{ALG_func_selectMoreGeneralNumericType}. \\
+    $selectMoreGenericNumericType()$: Function that select a more generic numeric XSD type from two given types according to the type promotions rules (TODO-REF). \\
     $operatorNode$: An operator expression node in the syntax tree. \\
@@
 
-\begin{algorithm}
-\caption{Function selectMoreGeneralNumericType()}
-\label{ALG_func_selectMoreGeneralNumericType}
-\begin{algorithmic}[1]
-\REQUIRE{\ \\
-    $numericType1$: Numeric type one.
-    $numericType2$: Numeric type two.
-}
-\ENSURE Returns the more generic numeric type of the two specified.
-\STATE TODO
-\end{algorithmic}
-\end{algorithm}
-
 The algorithm iterates through the variable declaration nodes from the prolog and if the variable is explicitly assigned with its type, the type is noted.
@@
 Every node ''knows'' wheather or not it may define new variables for nodes with higher sequence numbers.
-This is expressed in the algorithm by the condition at line 10.
-This condition is double-checked at line 3 where its meaning is that every node even ''knows'' if its particular subnode may define new variables.
-An example is a $FLWOR$ construct.
-The processing of $FLWOR$ knows that the first part (the first subnode) of this node may bind new variables using $for$ and $let$ constructs.
-After the subnode is processed, the processing of the $FLWOR$ node looks if any new variables were bound and if so the context of local variables is extended.
-This is done at line 4.
+This is expressed in the algorithm by the condition stating at line 10.
+If it does define new variables, they are memorized.
+Later, upon processing of its parent, they are retrieved and the variable context is extended.
+This is done at lines 4 and 5.
 Function $mergeContextVarTypes$ presented in Algorithm \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into its first argument (the variable context).
@@
 \REQUIRE{\ \\
-    $startingNode$: A node determinig a subtree to perform the analysis on \\
+    $memorizeType()$: Function memorizing a determined type of a specified node. \\
+    $set()$: Function memorizing a specified property of a specified node. \\
+    $add()$: Function adding a value to a list of specified properties of a specified node. \\
+    $get()$: Function returning a specified property (or list) set by functions $set()$ and $add()$. \\
+    $mergeContextVarTypes()$: Function defined in Algorithm \ref{ALG_function_mergeContextVarTypes}. \\
+    $startingNode$: A node determining a subtree to perform the analysis on. \\
     $contextVarTypes$: Types of local variables valid in the current subtree.
@@
 
-\FORALL{$subnode \in$ all subnodes of $startingNode$}
+\FORALL{$i \in \{1, \dots, |startingNode.getChildren()|\}$ ordered from the lowest to the highest}
+    \STATE $subnode = startingNode.getChild(i)$
     \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
-    \IF{$subnode$ defines new variables valid for following subnodes}
-        \STATE $contextVarTypes = mergeContextVarTypes(contextVarTypes, subnode.variableTypes)$
+    \STATE $newVars = get(subnode, newVars)$    
+    \STATE $contextVarTypes = mergeContextVarTypes(contextVarTypes, newVars)$
+\ENDFOR
+\IF{$is(startingNode, ExprNode)$}
+    \STATE $memorizeType(startingNode, determineExpressionType(startingNode))$
 \ENDIF
+\IF{$is(startingNode, VariableBindingNode)$}
+    \STATE $type = null$
+    \STATE $typeNode = startingNode.getChild(TypeNode)$
+    \IF{$typeNode \neq null$}
+        \STATE $type = getTypeTN(typeNode)$
+    \ELSE
+        \STATE $type = determineExpressionType(startingNode.getChild(BindingSequenceNode).getChild(ExprNode), contextVarTypes)$
+    \ENDIF
+    \STATE $set(startingNode, newVars, (startingNode.varName, type))$
+\ELSIF{$is(startingNode, TupleStreamNode)$}
+    \FORALL{$varBindingNode \in startingNode.getChildren()$}
+        \STATE $add(startingNode, newVars, get(varBindingNode, newVars))$
     \ENDFOR
-\IF{$startingNode$ is expression node}
-    \STATE $startingNode.type = determineExpressionType(startingNode)$
-\ENDIF
-\IF{$startingNode$ defines new variables valid in following nodes}
-    \STATE determine types of the variables
-    \STATE memorize their types in $startingNode.variableTypes$
 \ENDIF
@@
 
-\todo[inline]{Algoritmus analysisOfEXpressionTypes, line 10 - Tady tyhle věty jsou v pseudoalgoritmech divné.
-Měy by se vyskytovat minimálně.
-JInak je to pak v podstatě jen odřádkovaná věta, ne formální popis algoritmu bez technických detailů.}
+\todo[inline]{Popisat dvojicu funkcii memorizeType() a getType(). Taktiez trojicu set(), get(), add().}
+
+\todo[inline]{Overit si, ci ALG \ref{ALG_func_analysisOfExpressionTypes} pokryva vsetky pripady, co sa tyka definovania novych premennych a ich propagacie vyssim uzlom.}
 
@@
 \REQUIRE{\ \\
+    $keys()$: Function retrieving index values of an indexed structure. \\
     $contextVarTypes$: Types of local variables valid in current context. \\    
-    $extendingVarTypes$: Variable types of the same structure as the first argument and extending it.
+    $extendingVarTypes$: New variables in the same structure as the first argument.
 }
@@
 \REQUIRE{\ \\
+    $getGlobalVariableType()$: Function returning a determined type of a specified global variable. \\
     $contextVarTypes$: Types of local variables valid in current context. \\
@@
 Only expressions of a particular type are exploited.
-Specifically, an expression has to contain an subexpression $E$ of $pathType$ type (expression representing a certain element or attribute or set of elements or attributes).
+Specifically, an expression has to contain an subexpression $E$ of $PathType$ type (expression representing a certain element or attribute or set of elements or attributes).
 In the following text, the set represented by expression $E$ is denoted $S$.
@@
 \subsection{Arithmetic Operations}
-If the operator is one of $+$, $-$, $*$, $div$, $mod$ (the class of the expression is $Operator$ and the class of the operator is one of $PLUS$, $MINUS$, $MUL$, $DIV$, $MOD$ \todo{Niekam napisat mozne triedy operatora}), one operand is $E$ and the type $T$ of the other operand is one of numeric built-in types, then $T$ is also the inferred type of $S$.
+If the operator is one of $+$, $-$, $*$, $div$, $mod$ (the class of the expression node is $Operator$ and it represents one of $PLUS$, $MINUS$, $MUL$, $DIV$, $MOD$ \todo{Niekam napisat mozne triedy operatora}), one operand is $E$ and the type $T$ of the other operand is one of numeric built-in types, then $T$ is also the inferred type of $S$.
 
-If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression is $Operator$ and the class of the operator is one of $GEN\_LESS\_THAN$, $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$, $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.
+If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression node is $Operator$ and it represents one of $GEN\_LESS\_THAN$, $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$, $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.
 

### 2012-01-02 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \todo[inline]{Nejaky uvod a zkratke hrnutie}
-\todo[inline]{Lexikalna a syntakticka analyza z diplomky J. Schejbala}
-\todo[inline]{Staticka analyza typov vyrazov}
 \todo[inline]{Predpoklad korektnosti dotazov}
@@
 
-\todo[inline]{Tady z těch try a probably je jasně vidět, že netušíte co vlastně přesně algoritmus bude dělat.
-Nejdříve si to rozmyslete, pak pište text.
-Rozhodně by se takové výrazy neměly v DP dobjevovat, působít o velmi nedůvěryhodně.
-Měl by to být strohý popis toho, co bylo uděláno.}
-
 We want to capture the following types.
@@
 
-A simple algorithm is presented in Algorithm \ref{ALG_processing_of_functions} \todo{Q: uvadzat aj taketo jednoduche funkcie?
-A: Obecně by mělo být popsáno vše.
-Ale pokud se nějaká jednoduchá věc např.
-opakuje, tak stačí napsat, že je algoritmus podobný jako v předchozím případě nebo tak něco.}.
+A simple algorithm is presented in Algorithm \ref{ALG_processing_of_functions}.
 It uses \texttt{getFunctionDeclarationNodes()} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three.
@@
 Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.
+
+\subsection{Auxiliary Functions}
+Before proceeding to the next phase of the algorithm, we introduce auxiliary functions used in pseudo algorithms.
+
+Function $memorizeType()$ takes two arguments $node$ and $type$ and it memorizes the given type of the specified expression node.
+This information can be then retrieved by function $getType()$, specifying the particular node as its argument.
+
+Another group consists of functions $set()$, $add()$, and $get()$.
+Function $set()$ memorizes a given value of a specified property of a specified node.
+It is used to assign a node with a named value, for example, $set(someNode, color, 'red')$ will assign $someNode$ with string value $'red'$ wich can be then retrieved by function $get()$, specifying the node and the property name.
+For example, $get(someNode, color)$ returns $'red'$.
+A subsequent call of $set()$ assigning a node with a value of already assigned property will overwrite it, leaving the property with the newer value.
+However, a property can has several values and this can be achieved using function $add()$ with the same syntax as function $set()$ but instead of overwriting the existing property, $add()$ will append the new value to the existing ones.
+Then, a call of $get()$ on this node and property returns a list of all its values, preserving the order of their addition.
 
@@
 
-\todo[inline]{Popisat dvojicu funkcii memorizeType() a getType(). Taktiez trojicu set(), get(), add().}
-
 \todo[inline]{Overit si, ci ALG \ref{ALG_func_analysisOfExpressionTypes} pokryva vsetky pripady, co sa tyka definovania novych premennych a ich propagacie vyssim uzlom.}
@@
 \todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme.}
+
+\section{Step 4: Key Discovery}

### 2012-01-03 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 
-\section{Discovering XML Keys and Foreign Keys in Queries}
+\section{Discovering XML Keys and Foreign Keys in Queries}\label{Keys}
 The method described in paper \cite{Necasky:2009:DXK:1529282.1529414} improves automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries.

### 2012-01-03 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme.}
+\todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme. -- To urobim pocas implementacie pre nejaky vzorovy dotaz.}
 
 \section{Step 4: Key Discovery}
+In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in section \ref{Keys}, and extending it.
+Firstly, the syntax tree is searched for certain forms of FLWOR expressions to infer the keys from.
+Then, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.
+
+\subsection{Join Patterns}
+The two join patterns from the original approach are shown in Listings \ref{listing_for_join_pattern} and \ref{listing_let_join_pattern}.
+Additional join patterns are introduced in Listings \ref{listing_join_pattern_3}, \ref{listing_join_pattern_4} and \ref{listing_join_pattern_5}.
+In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes.
+
+\begin{lstlisting}[mathescape, float, caption=Join pattern 3, frame=single, label=listing_join_pattern_3]
+for $\$e_1$ in $P_1$
+for $\$e_2$ in
+    $P_2[L_2 = \$e_1/L_1]$
+return $C_R$
+\end{lstlisting}
+
+\begin{lstlisting}[mathescape, float, caption=Join pattern 4, frame=single, label=listing_join_pattern_4]
+for $\$e_1$ in $P_1$
+let $\$e_2$ :=
+    $P_2[L_2 = \$e_1/L_1]$
+return $C_R$
+\end{lstlisting}
+
+\begin{lstlisting}[mathescape, float, caption=Join pattern 5, frame=single, label=listing_join_pattern_5]
+for $\$e_1$ in $P_1$
+for $\$e_2$ in $P_2$
+where $\$e_2/L_2 = \$e_1/L_1$
+return $C_R$
+\end{lstlisting}

### 2012-01-05 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

--- 2012-01-02 doc/theses/mikula/proposed_algorithm.tex
@@
 
-\todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme.}
+\todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme. -- To urobim pocas implementacie pre nejaky vzorovy dotaz.}
 
 \section{Step 4: Key Discovery}
+In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in section \ref{Keys}, and extending it.
+Firstly, the syntax tree is searched for certain forms of FLWOR expressions to infer the keys from.
+Then, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.
+
+\subsection{Join Patterns}
+
+\begin{lstlisting}[mathescape, float, caption=Other form of the for join pattern., frame=single, label=listing_other_for_join_pattern]
+for $\$e_1$ in $P_1$
+for $\$e_2$ in
+    $P_2[L_2 = \$e_1/L_1]$
+return $C_R$
+\end{lstlisting}
+
+\begin{lstlisting}[mathescape, float, caption=Other form of the let join pattern., frame=single, label=listing_other_let_join_pattern]
+for $\$e_1$ in $P_1$
+let $\$e_2$ :=
+    $P_2[L_2 = \$e_1/L_1]$
+return $C_R$
+\end{lstlisting}
+
+\begin{lstlisting}[mathescape, float, caption=Join pattern 3., frame=single, label=listing_join_pattern_3]
+for $\$e_1$ in $P_1$
+for $\$e_2$ in $P_2$
+where $\$e_2/L_2 = \$e_1/L_1$
+return $C_R$
+\end{lstlisting}
+
+The two join patterns from the original approach are shown in Listings \ref{listing_for_join_pattern} and \ref{listing_let_join_pattern}.
+Additional join patterns are introduced in Listings \ref{listing_other_for_join_pattern}, \ref{listing_other_let_join_pattern} and \ref{listing_join_pattern_3}.
+In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes.
+Actually, the join patterns from \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly because their structure in the syntax tree is different.
+
+Since, in this step, we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of $PathType$ without predicates, using only child and descendant axes.
+
+As described in the original method and summarized in section \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns and rules to classify each occurrence into one of these cases.
+For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
+\todo{Popisat dovod nizsej vahy (nie je take jasne, ze je to par key-foreign key), pripadne zmenit vahu.}
+
+\begin{algorithm}
+\caption{Processing of FLWOR expressions}
+\label{ALG_processing_of_flwor_expressions}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $flworNode$: A node representing a FLWOR expression. \\
+    $forVars$: An array of couples of for variables and their for clause nodes.\\
+}
+
+\ENSURE Updated variable $forVars$.
+
+\STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).getChildren()$
+\FORALL{$bindingNode \in bindingNodes$}
+    \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).getChild(ExprNode)$
+    \IF{$getType(BindingExpr) == PathType$ using only child and descendant axes, ending with exactly one predicate $P$}
+      \FORALL{$(var, node) \in forVars$}
+            \IF{$P$ is of form $bindingNode.varName/L_1 = var/L_2$}
+                \IF{$is(bindingExpr, ForClauseNode)$}
+                    \STATE memorize $(node, bindingExpr)$ as an occurrence of the for join pattern
+                \ELSIF{$is(bindingExpr, LetClauseNode)$}
+                    \STATE memorize $(node, bindingExpr)$ as an occurrence of the for let pattern                
+                \ENDIF
+            \ENDIF
+        \ENDFOR
+    \ELSIF{$getType(BindingExpr) == PathType$ without predicates using only child and descendant axes}
+        \IF{$is(bindingNode, ForClauseNode$}
+            \STATE add $(bindingNode.varName, bindingNode)$ to forVars
+        \ENDIF
+    \ENDIF
+\ENDFOR
+
+\RETURN $forVars$
+\end{algorithmic}
+\end{algorithm}
+
+\todo[inline]{Processing function pre join pattern 3.}
+
+To find the occurrences of the join patterns, the algorithm recursively, in pre-order, searches the syntax tree and every node representing a FLWOR expression is processed.
+The FLWOR processing is shown in Algorithm \ref{ALG_processing_of_flwor_expressions}.
+Its input is an array of couples containing names of variables which has been bound in for clauses in FLWOR expressions represented by ancestor nodes, and references to those respective for clause nodes.
+The processing iterates trough subnodes of a current node and the iteration consists of two logical parts.
+
+In the first one, a subnode is checked if it is a for or let clause that forms a join pattern occurrence with one of clauses from the input array.
+If it does, the occurrence is noted.
+Otherwise, if it is a for clause binding a new variable satisfying the conditions to be the first for clause of a join pattern, it is added to the array along with the variable name.
+It can be then processed in the following clauses of the currect FLWOR node or in its descendant FLWOR nodes, later in the recursion.
+
+\subsection{Analysis of the Join Pattern Occurrences}

### 2012-01-08 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
-\chapter{Analysis of XQuery}
+\chapter{Analysis of XQuery} \label{CHAPTER_analysis_of_xquery}
 \todo[inline]{Někde by se mělo říct, kterou verzi uvažujeme, jaké obecně jsou, co je třeba XQuery Core apod.}
@@
 Case 2) requires the result of \texttt{Q} and additional processing of the result.
-Case 3) requires a step-by-step evaluation of \texttt{Q}, so the XML document is also required.
+Case 3) requires a step-by-step evaluation of \texttt{Q}, hence, the XML document is also required.
 And in case 4), it is not possible to obtain any information.
@@
 
-\subsection{Negative Statements about Uniqueness}
+\subsection{Negative Statements about Uniqueness} \label{Negative_uniqueness}
 In many cases a statement refusing uniqueness of element or attribute values could be inferred.

### 2012-01-08 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-As described in the original method and summarized in section \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns and rules to classify each occurrence into one of these cases.
+As described in the original method and summarized in section \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases.
 For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
@@
     $flworNode$: A node representing a FLWOR expression. \\
-    $forVars$: An array of couples of for variables and their for clause nodes.\\
+    $forVars$: An array of couples of for variables and their for clause nodes.
 }
@@
 \STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).getChildren()$
+
+\STATE $whereClause = flworNode.getChild(WhereClauseNode)$
+\STATE $checkJoinPattern3 = false$
+\STATE $whereExpr = null$
+\IF{$whereClause != null$}
+    \STATE $whereExpr = whereClause.getChild(ExprNode)$
+    \IF{$is(whereExpr, OperatorNode)$ AND $whereExpr.operator == GEN\_EQUALS$}
+        \STATE $checkJoinPattern3 = true$
+    \ENDIF
+\ENDIF
+
 \FORALL{$bindingNode \in bindingNodes$}
@@
         \FORALL{$(var, node) \in forVars$}
-            \IF{$P$ is of form $bindingNode.varName/L_1 = var/L_2$}
-                \IF{$is(bindingExpr, ForClauseNode)$}
-                    \STATE memorize $(node, bindingExpr)$ as an occurrence of the for join pattern
-                \ELSIF{$is(bindingExpr, LetClauseNode)$}
-                    \STATE memorize $(node, bindingExpr)$ as an occurrence of the for let pattern                
-                \ENDIF
-            \ENDIF
+            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var, forVars)$
         \ENDFOR
@@
 
+\begin{algorithm}
+\caption{Function determinateJoinPatterns}
+\label{ALG_func_determinateJoinPatterns}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+    $curBindingNode$: A variable binding node in the current FLWOR node. \\
+    $P$: A predicate from the binding expression. \\
+    $forBindingNode$: A for binding node from an ancestor FLWOR node. \\
+    $forVar$: A variable from the ancestor for binding node. \\
+    $forVars$: An array of couples of for variables and their for clause nodes.
+}
+
+\ENSURE Updated variable $forVars$.
+
+\IF{$P$ is of form $curBindingNode.varName/L_1 = var/L_2$}
+    \IF{$is(curBindingNode, ForClauseNode)$}
+        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for join pattern
+        \STATE remove $(forVar, forBindingNode)$ from $forVars$
+    \ELSIF{$is(curBindingNode, LetClauseNode)$}
+        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for let pattern                
+        \STATE remove $(forVar, forBindingNode)$ from $forVars$
+    \ENDIF
+\ELSIF{$checkJoinPattern3$ AND $is(curBindingNode, ForClauseNode)$}
+    \IF{$whereExpr$ is of form $bindingNode.varName/L_1 = var/L_2$}
+        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the join pattern 3
+        \STATE remove $(forVar, forBindingNode)$ from $forVars$
+    \ENDIF
+\ENDIF
+
+\RETURN $forVars$
+\end{algorithmic}
+\end{algorithm}
+
 \todo[inline]{Processing function pre join pattern 3.}
@@
 Its input is an array of couples containing names of variables which has been bound in for clauses in FLWOR expressions represented by ancestor nodes, and references to those respective for clause nodes.
-The processing iterates trough subnodes of a current node and the iteration consists of two logical parts.
+The processing iterates through subnodes of a current node and the iteration consists of two logical parts.
 
@@
 If it does, the occurrence is noted.
-Otherwise, if it is a for clause binding a new variable satisfying the conditions to be the first for clause of a join pattern, it is added to the array along with the variable name.
+The responsible code is partially moved to function $determineJoinPatterns()$ in Algorithm \ref{ALG_func_determinateJoinPatterns}.
+If the subnode does not form an occurrence, and if it is a for clause binding a new variable satisfying the conditions to be the first for clause of a join pattern, it is added to the array along with the variable name.
 It can be then processed in the following clauses of the currect FLWOR node or in its descendant FLWOR nodes, later in the recursion.
@@
 \subsection{Analysis of the Join Pattern Occurrences}
+For each found join pattern occurrence, the algorithm displayed in Algorithm \ref{ALG_classification_of_join_pattern_occurrences} decides whether it is (O1) or (O2) case, using the rules from the original method and the rule for the join pattern 3 considering it of case (O1) and assigning it with weight of 0.5.
+
+\begin{algorithm}
+\caption{Classification of join pattern occurrences}
+\label{ALG_classification_of_join_pattern_occurrences}
+\begin{algorithmic}[1]
+\REQUIRE $\mathcal{O}$: A set of all found join pattern occurrences.
+
+\ENSURE $\mathcal{O'}$: A set of the classified join pattern occurrences from $\mathcal{O}$ assigned with their respective weights.
+
+\FORALL{$jpOccurrence \in \mathcal{O}$}
+    \IF{$jpOccurrence$ is the for join pattern}
+        \STATE mark $jpOccurrence$ as case (O1), weight 1
+    \ELSIF{$jpOccurrence$ is the join pattern 3}
+        \STATE mark $jpOccurrence$ as case (O1), weight 0.5
+    \ELSE
+        \STATE TODO (R2,R3,R4,R5)
+    \ENDIF
+    \STATE update $jpOccurrence$ in $\mathcal{O}$
+\ENDFOR
+
+\RETURN $\mathcal{O}$
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Rejection of Uniqueness}
+While the previous steps of the key discovery produce positive statements about the keys, in this step, the algorithm searches for evidence on non-uniqueness of elements and attributes.
+An aim of this step is to eliminate or decrease the number of falsely inferred keys.
+\todo[inline]{Napisat niekam, ze odvodzovanie kludov predpoklada, ze kazdy element/element join je robeny cez key-foreign key pair.}
+When the algorithm discovers that a certain element is probably not unique, it assign the element with a negative weight that will decrease the sum of its weights in the summarizing step.
+
+The first group of XQuery constructs utilized by this step is a group of $distinct-values()$ function and aggregation functions $min()$, $max()$, TODO \todo{Doplnit} as discussed in Section \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
+The algorithm searches for calls of the mentioned functions with a $PathType$ on their input.
+Elements and attributes selected by those $PathTypes$ are then considered to be not unique, and consequently, they can not be keys.
+
+Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates.
+Predicates in a path restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates.
+\todo[inline]{Toto asi nebude uplne pravda pre nejake exoticke predikaty.
+Overit pri implementacii.}
+
+\begin{algorithm}
+\caption{Rejection of uniqueness - aggregation functions}
+\label{ALG_rejection_of_uniqueness_aggregation_functions}
+\begin{algorithmic}[1]
+\REQUIRE $node$: A node of the syntax tree.
+
+\IF{$is(node, FunctionCallNode)$}
+    \IF{$node.fncName$ is one of $distinct-values$, $min$, $max$, TODO}
+        \STATE $argument = node.getChild(0)$
+        \IF{$getType(argument) == PathType$ using only child and descendant axes}
+            \STATE memorize $argument$ as the negative key statement, weight 1
+        \ENDIF
+    \ENDIF
+\ENDIF
+
+\FORALL{$subnode \in node.getChildren()$}
+    \STATE call this function recursively with $subnode$
+\ENDFOR
+
+\end{algorithmic}
+\end{algorithm}

### 2012-01-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \end{algorithm}
+
+Besides the aggregation functions, a certain form of FLWOR expressions is utilized in this step.
+They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (lieral constant, reference of a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
+
+As discussed in Section \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the count of the items may be more than one.
+And, together with that, the set is restricted to contain only items which satisfy a condition that a value of one of their subnodes (or descendant nodes) equals a certain value.
+That implies that the value in those nodes is not unique, and, alike the utilization of the aggregation functions, non-unique nodes cannot be keys.
+
+\begin{algorithm}
+\caption{Rejection of uniqueness - comparison with a constant}
+\label{ALG_rejection_of_uniqueness_comparison_with_a_constant}
+\begin{algorithmic}[1]
+\REQUIRE $node$: A node of the syntax tree.
+
+\STATE $forVars =$ an empty array
+
+\IF{$is(node, FLWORExprNode)$}
+    \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
+        \IF{$is(bindingNode, ForClauseNode)$}
+            \STATE $expr = bindingNode.getChild(BindingSequenceNode).getChild(ExprNode)$
+            \IF{$getType(expr) == PathExpr$ without predicates, using only child and descendant axes}
+                \STATE add $bindingNode.varName$ to $forVars$
+            \ENDIF
+        \ENDIF
+    \ENDFOR
+\ENDIF
+
+\STATE $whereClause = node.getChild(WhereClauseNode)$
+\IF{$whereClause != null$}
+    \STATE $processWhere(forVars, whereClause.getChild(ExprNode))$
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Rejection of uniqueness - function processWhere}
+\label{ALG_rejection_of_uniqueness_function_process_where}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+	$forVars$: Names of variables to look for in the where expression. \\
+	$exprNode$: The where expression node.
+}
+
+\IF{$is(exprNode, OperatorNode$)}
+    \IF{$exprNode.operator == GEN\_EQUALS$}
+        \FORALL{$var \in forVars$}
+            \IF{$exprNode$ is of a form $var/L = C$ where $L$ is a path without predicates using only child and descendant axes, and getType(C) is a constant}
+                \STATE memorize $var/L$ as the negative key statement, weight 0.9
+            \ENDIF
+        \ENDFOR
+    \ELSIF{$exprNode.operator == AND$}
+        \STATE processWhere(forVars, exprNode.leftSide)
+        \STATE processWhere(forVars, exprNode.rightSide)
+    \ENDIF
+\ENDIF
+
+\end{algorithmic}
+\end{algorithm}
+
+\todo[inline]{Doplnit algoritmus comparison with a constant pre predikaty.}

### 2012-02-12 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 
-Recently in reaction to this situation, a significant number of approaches dealing with an automatic construction of an XML schema has been proposed.
-An aim is to exploit a provided set of XML documents and infer an XML schema, so that the XML documents are valid against it.
+Recently in reaction to this situation, a significant number of approaches dealing with automatic construction of an XML schema has been proposed.
+The aim is to exploit a provided set of XML documents and infer an XML schema, so that the XML documents are valid against it.
 In addition, the inferred schema should be reasonable in views of human-readability, preciseness and conciseness.
@@
 
-Firstly, we indend to analyze existing inference solutions in general and to discuss their advantages and disadvantages.
-Then, we try to identificate and discuss information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.
+Firstly, we analyze existing inference solutions in general and discuss their advantages and disadvantages.
+Then, we identify and discuss information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.
 
-The work should result into a proposal of own approach involving the improvements, its implementation and suitable experiments that will show its advantages.
+The work results in a proposal of own approach involving the improvements, its implementation and suitable experiments that show its advantages.
 

### 2012-02-12 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

--- 2011-12-18 doc/theses/mikula/used_technologies_and_definitions.tex
@@
 
-A validity of an XML document againt its XML schema expresses wheather the document is well-formed \cite{Bray:08:EML}, and at the same time, wheather it conforms to the XML schema.
+Validity of an XML document against its XML schema expresses whether the document is well-formed \cite{Bray:08:EML} and, at the same time, whether it conforms to the XML schema.
 
-\subsection{An XML Sample}
-To demonstrate the described technologies, we introduce a part of a simple XML document \ref{FIG_a_simple_xml_sample}.
+\subsection{An XML Example}
+To demonstrate the described technologies, we introduce a part of a simple XML document (see Figure \ref{FIG_a_simple_xml_example}).
 It represents books in a bookstore.
@@
 \begin{figure}
-\caption{A simple XML sample.}
-\label{FIG_a_simple_xml_sample}
+\caption{A simple XML example.}
+\label{FIG_a_simple_xml_example}
 \begin{verbatim}
@@
   <book id="b1">
-    <title>Compilers: Principles, Techniques, and Tools (2nd Edition)</title>
+    <title>
+      Compilers: Principles, Techniques, and Tools
+      (2nd Edition)
+    </title>
     <authors>
@@
 \subsection{DTD}
-Document Type Definition (DTD) expresses a structure of XML documents by declarations of elements. An element has its name and content is declared using syntax \texttt{<!ELEMENT $name$ $content$>}.
+Document Type Definition (DTD) expresses a structure of XML documents by declarations of elements. An element has its name and a content declared using syntax \texttt{<!ELEMENT $name$ $content$>}.
 
@@
 
-Attributes of an element are specified in an attribute list \texttt{<!ATTLIST $element\_name$ $attribute\_name$ $type$ $default\_value$>}.
+Attributes of an element are specified in an attribute list \texttt{<!ATTLIST\linebreak $element\_name$ $attribute\_name$ $type$ $default\_value$>}.
 Each attribute has its name, its type and its default value or definition of obligation of occurrence.
@@
 
-An example of DTD describing the book element from the XML sample \ref{FIG_a_simple_xml_sample} follows.
+An example of DTD describing the book element from the XML example in Figure \ref{FIG_a_simple_xml_example} follows.
 
@@
 An important fact is that each XSD instance is a valid XML document.
-Its root element is always \texttt{<schema>} and the the XSD instance may (and it often does) begins with the following two lines.
+Its root element is always \texttt{<schema>} and the XSD instance begins with the following two lines.
 
@@
 \end{alltt}
-where $type$ is either one of the built-in types \todo{ref?} or a user-defined type. Definition of an attribute is similar.
+where $type$ is either one of the built-in types or a user-defined type. Definition of an attribute is similar.
 \begin{alltt}
@@
 Attributes are only allowed to be of the simple types.
-Definition of the a simple type is
+Definition of a simple type is
 \begin{alltt}
@@
 Subelements are declared using \texttt{<xs:sequence>}, \texttt{<xs:choice>} and \texttt{<xs:all>} schema elements.
-If the order of the subelements is significant, we use \texttt{<xs:sequence>}, where the subelements in an XML must occur in the same order as they are defined in the sequence.
-If the order is not significant, we use \texttt{<xs:all>}, where the subelements must only occur but their order is not relevant.
+If the order of the subelements is significant, we use \texttt{<xs:sequence>}, where the subelements in an XML instance must occur in the same order as they are defined in the sequence.
+If the order is not significant, we use \texttt{<xs:all>}.
 Construct \texttt{<xs:choice>} is equivalent to the alternation of several elements in the DTD.
@@
 
-An element can has a mixed content (can contain text and other elements at the same time).
+An element can have a mixed content (can contain text and other elements at the same time).
 Such element has to be of a complex type and definition of the type has to contain attribute \texttt{mixed} with value \texttt{true}.
@@
 
-The book element from the sample XML \ref{FIG_a_simple_xml_sample} can be described using the XSD as follows.
+The book element from the sample XML document in Figure \ref{FIG_a_simple_xml_example} can be described using the XSD as follows.
 \begin{alltt}
@@
 A path step is \\ \texttt{$axis$::$node\_test$[$predicate_1$]$\dots$[$predicate_n$]} \\
-where $axis$ and all predicates are optional.
+where all predicates are optional.
 
@@
 \subsection{Abbrevations}
-Abbrevations for several most widely used construct are defined.
+Abbrevations for several most widely used constructs are defined as follows.
 \begin{itemize}
+\item \texttt{$P$/$identifier$} stands for \texttt{$P$/child::$identifier$} (child is a default axis)
 \item \texttt{$P$/@$identifier$} stands for \texttt{$P$/atribute::$identifier$}
@@
 \subsection{Predicates}
-Predicates places additional conditions upon nodes that passed the node test.
+Predicates place additional conditions upon nodes that passed the node test.
 They can be either relative XPath paths or comparison expressions.
@@
 \subsection{Usage Examples}
-Finally, we introduce several examples of XPath expressions, with their description, using the sample XML \ref{FIG_a_simple_xml_sample}.
+Finally, we introduce several examples of XPath expressions, with their description, using the sample XML in Figure \ref{FIG_a_simple_xml_example}.
 
-Path that selects \texttt{author} elements of the book with id $b1$:
+A path that selects \texttt{author} elements of the book with id $b1$:
 \begin{alltt}
@@
 
-Expression returning a number of all books:
+An expression returning the number of all books:
 \begin{alltt}
@@
 
-Path selecting the ISBN element of the book titles $XQuery$:
+A path selecting the ISBN element of the book entitled $XQuery$:
 \begin{alltt}
@@
 
-Path returning the titles of the books written by more than one author:
+A path returning the titles of the books written by more than one author:
 \begin{alltt}
@@
 \section{XQuery}
-\todo[inline]{Tuto sekciu som vypracovaval primarne podla vasich slidov na prednasku Technologie XML. Mam to niekde uviest?}
-
 An XML Query Language (XQuery, version 1.0) \todo{ref} is a language designed to query XML data.
@@
 XPath 2.0 is an extension of XPath 1.0 (but not entirely compatible), adding ordered sequences, their iterations, set operations, conditional expressions, quantified expressions, and XML Schema types.
+
+In the remainder of this section, basic features of the XQuery languages are briefly described. Many query samples can be found in Chapter \ref{CHAPTER_analysis_of_xquery}.
 
@@
 An item is either an atomic value or a node.
-An atomic value is a value of a XML Schema simple type.
+An atomic value is a value of an XML Schema simple type.
 A node is an instance of one of the node types.
 
-Each node has its identity, data type (simple or complex, according to the XML Schema types), typed value (can be retrieved by function \texttt{fn:data()}, and string value (retrieved by function \texttt{fn:string()}.
+Each node has its identity, data type (simple or complex, according to the XML Schema types), typed value (which can be retrieved by function \texttt{fn:data()}), and string value (which can be retrieved by function \texttt{fn:string()}).
 
@@
 \texttt{return $expr$} \\
-The result of the whole $FLWOR$ expression is $expr$. It can (and usually is) be constructed using the tuples of the for variable values and the let variables.
+The result of the whole $FLWOR$ expression is $expr$. It is constructed using the tuples of the for variable values and the let variables.
 
 \subsection{Conditional Expressions}
+XQuery provides common \texttt{if-then-else} conditional expressions with common syntax.
+\begin{alltt}
+if (\(condition\))
+then \(expr1\)
+else \(expr2\)
+\end{alltt}
+If $condition$ expression evaluates to true, $expr1$ is evaluated, else $expr2$ is evaluated. The else branch is optional.
 
 \subsection{Quantified Expressions}
+XQuery provides two types of quantified expressions, \texttt{every} and \texttt{some}. Their syntax follows.
+\begin{alltt}
+every \(var\) in \(expr1\)
+satisfies \(expr2\)
+\end{alltt}
+\begin{alltt}
+some \(var\) in \(expr1\)
+satisfies \(expr2\)
+\end{alltt}
 
-\subsection{Built-in Functions}
+Firstly, $expr1$ is evaluated, and then the result of the quantified expression is true if $expr2$ is true for every (some) item, represented by $var$, of the result of $expr1$ (which is a sequence).
 
-\subsection{User-defined Functions}
+\subsection{Functions}
+XQuery provides a wide set of built-in functions.
 
-\subsection{Value Comparison Expressions}
+And also new functions can be defined using the following syntax.
+\begin{alltt}
+define function \(name\)(\(parameters\)) as \(type\)
+\{
+  \(expr\)
+\}
+\end{alltt}
+where $name$ is the function's name, $parameters$ is a list of parameters (with or without specification of their types), and $type$ is a type of a return value, which is the result of $expr$.
 
-\subsection{General Comparison Expressions}
+%\subsection{Comparison Expressions}
+%Comparison expressions are of three types: $value$ comparison expressions, $general$ comparison expressions, and %$node$ comparison expressions.
+A very brief insight follows.
+More details can be found in REF\todo{ref, moze to takto ostat?}.
 
-\subsection{Node Comparison Expressions}
+%The $value$ comparison expressions use operators \texttt{lt}, \texttt{gt}, \texttt{le}, \texttt{ge}, %\texttt{eq}, and \texttt{ne} and they compare operands that are or can be converted to atomic values.
 
-\subsection{Usage Examples}
+%The $general$ comparison expressions use operators \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, %and \texttt{!=} and they can also compare sequences.
+
+%And the $node$ comparison expressions use operators \texttt{is}, \texttt{<<}, and \texttt{>>} and they compare %nodes, their identities and order.
 
@@
 
+%\section{Other Technologies}
+
+%\begin{description}
+%\item[XSLT] e\textbf{X}tensible \textbf{S}tylesheet \textbf{L}anguage \textbf{T}ransformations %\cite{Clark:99:XTV} is a language to transform input XML document(s) into other (not necessarily XML) document(s).
+%\end{description}
+
 \todo[inline]{možná ňejaký automat apod. Prostě co budete dále v textu potřebovat. Skratky: XSD, XSLT, ...}

### 2012-02-12 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 Heuristic methods \cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522} are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
-Two of these approaches are described in Section \ref{section_xtract} and Section \ref{section_even_an_ant_can_create_an_xsd}.
+Two of these approaches are described in Chapter \ref{section_xtract} and Chapter \ref{section_even_an_ant_can_create_an_xsd} \todo{někde Chapter, někde Section - sjednotit a dodržovat}.
 The former one incorporates so-called MDL principle to create the (sub)optimal result.
@@
 \subsection{Factoring}
-Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decresing their summed size without modifications in their semantics.
+Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decreasing their summed size without modifications in their semantics.
 The aim of this step is to decrease the MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
 
-An example introduced it paper \cite{Garofalakis:2000:XSE:342009.335409}: DTD candidates \texttt{ac}, \texttt{ad}, \texttt{bc} and \texttt{bd} are factored into \texttt{(a|b)(c|d)}.
+An example is introduced it paper \cite{Garofalakis:2000:XSE:342009.335409}: DTD candidates \texttt{ac}, \texttt{ad}, \texttt{bc} and \texttt{bd} are factored into \texttt{(a|b)(c|d)}.
 
@@
 Unlike the majority of recent approaches, this methods creates its result in XML Schema.
-Contrary to DTD, XSD is a newer language and it provides greater expressive possibilities.
 The authors of this method focused on inferring elements with the same name but different context and the unordered sequences which can be in XSD expressed by \texttt{xs:all} construct.
@@
 
-In the fist step the input schema is corrected to conform to the whole set of input XML documents.
+In the first step the input schema is corrected to conform to the whole set of input XML documents.
 This is done by creating a PTA for each production extracted from the input schema and merging it with a respective production from the initial grammar, involving ACO and MDL heuristics.
@@
 
-The method is based on assumption that each join is done via key/foreign key pair.
-It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is its respective foreign key or vice versa.
+The method is based on an assumption that each join is done via key/foreign key pair.
+It means it is supposed that $L_1$ is a key of the elements in $S_1$ and $L_2$ is its respective foreign key or vice versa.
 
@@
 
-The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a target return path (weight: 1) or if aggregation function \texttt{count} is applied on a target return path (weight: 0.75), where target return paths are paths in $C_R$ starting with $\$e_2$.
+The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a target return path (weight: 1) or if aggregation function \texttt{count} is applied on a target return path (weight: 0.75), where target return paths are paths in $C_R$ starting with $\$e_2$ (see Listing \ref{listing_for_join_pattern}).
 
@@
 The output of the method is a list of scored keys and for each key a list foreign keys referencing the key.
-Since the method deals only with the inference of keys, it is not an overall method of the XML schema inference.
+Since the method deals only with the inference of keys, it is not a complete method of the XML schema inference.
 It is meant to be used in collaboration with other schema-inferring methods to refine their results.
@@
 Paper \cite{Mlynkova:2009:IXS:1862681.1862693} introduces a method that utilizes an XML schema besides XML documents.
-The method described in \cite{Necasky:2009:DXK:1529282.1529414} utilizes XQuery queries, however, the result of this method is a list of discovered keys and foreign keys, not an XML schema.
+The method described in \cite{Necasky:2009:DXK:1529282.1529414} utilizes XQuery queries; however, the result of this method is a list of discovered keys and foreign keys, not an XML schema.
 Also, to our best knowledge, implementation that combines this method with another approach to get the XML schema has not been proposed yet.

### 2012-02-13 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
-\chapter{Analysis of XQuery} \label{CHAPTER_analysis_of_xquery}
+svn diff\chapter{Analysis of XQuery} \label{CHAPTER_analysis_of_xquery}
 \todo[inline]{Někde by se mělo říct, kterou verzi uvažujeme, jaké obecně jsou, co je třeba XQuery Core apod.}
@@
 Most of sample queries in this chapter are taken from \cite{Walmsley:2007:XQU:1205865} and \cite{Marchiori:07:XQU}, with or without modifications, as these were the main sources for this XQuery analysis.
-\todo[inline]{Overit si, ze mozem pouzivat tieto prevzate priklady.}
 
 \section{Structure of XML documents}
-Most of queries can be exploited to obtain some information about a structure of respective XML documents.
+Most of queries can be exploited to obtain some information about the structure of respective XML documents.
 The structure of XML documents captures elements from these documents along with their names, attributes and their organization.
@@
 Path expressions without predicates which use only child axis are the simpliest example of such queries.
-Path expression \texttt{/bib/book/author} indicates that element \texttt{bib} is the outermost element and it contains one or more elements \texttt{book} and these contain one or more elements \texttt{author}.
+For example, path expression \texttt{/bib/book/author} indicates that element \texttt{bib} is the root element and it contains one or more elements \texttt{book} and these contain one or more elements \texttt{author}.
 
@@
 The latter one uses also descendant axis.
-That means the query considers all elements \texttt{author} in the document, thus it is a hint that maybe there are several elements named \texttt{author} but with different absolute paths and all of them contain an element \texttt{name}.
+That means the query considers all elements \texttt{author} in the document, thus it is a hint that maybe there are several elements named \texttt{author} but with different absolute paths and some of them contain an element \texttt{name}.
 
-Besides elements, attributes can be processed exactly in the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} has attribute \texttt{ISBN}.
+Besides elements, attributes can be processed exactly in the same way. Path expression \texttt{/bib/book/@ISBN} indicates that element \texttt{book} can have attribute \texttt{ISBN}.
 
-However, statements of this kind are necessary for the XML schema inference, their obtaining from queries is not significant due to the following reason.
+Statements of this kind are necessary for the XML schema inference, however their obtaining from queries is not significant due to the following reason.
 These statements could be determined directly from the XML documents and it could be easily done in a more convenient way.
-Queries may not cover the whole relevant content of the documents.
+In addition, queries may not cover the whole relevant content of the documents.
 Let \texttt{E} be an element occurring in an XML document and let \texttt{Q} be an XQuery query.
@@
 For example path expression \texttt{/bib/book/author} returns elements \texttt{author} along with their subelements \texttt{name}, \texttt{birthdate} and \texttt{nationality}.
-\item \texttt{E} is not mentioned in \texttt{Q} nor it occurs in the result but it is processed by the evaluation of \texttt{Q}.
+\item \texttt{E} is neither mentioned in \texttt{Q}, nor it occurs in the result, but it is processed by the evaluation of \texttt{Q}.
 For example the evaluation of path expression \texttt{//author} processes amongst others elements \texttt{bib} and \texttt{book}, in a search for elements \texttt{author}.
-\item \texttt{E} occurs in a part of the XML document not related to a query at all.
+\item \texttt{E} occurs in a part of the XML document not related to \texttt{Q} at all.
 \end{enumerate}
@@
 
-\section{Occurrence Count of Elements}
+\section{Number of Occurrences of Elements}
 Some XQuery constructs indicate multiplicity of a particular element or limit the element to occur at most once.
@@
 
-Function \texttt{count()} returns number of items in a provided sequence.
-If the sequence is a sequence of elements, the count of these elements will probably not be limited to one.
+Function \texttt{count()} returns the number of items in a provided sequence.
+If the sequence is a sequence of elements, the number of these elements will probably not be limited to one.
 The sample query indicates that the root element \texttt{book} can contain more than one element \texttt{section}.
@@
 
-In this query, number of elements \texttt{color} in element \texttt{product} with attribute \texttt{id} equal to 5 has to be at least one, otherwise an error is raised upon execution.
+In this query, the number of elements \texttt{color} in element \texttt{product} with attribute \texttt{id} equal to 5 has to be at least one, otherwise an error is raised upon execution.
 If we assume that this query is written correctly, with common sense and it should not raise the error, we can infer that element \texttt{product} has to contain one or more elements \texttt{color}.
@@
 Alike the previous example, an expression in \texttt{order by} clause can be evaluated to at most one item or the \emph{type error} is raised.
-Therefore, every element \texttt{item} contains zero or one element \texttt{num} but no more.
+Therefore, every element \texttt{item} contains zero or one element \texttt{num} but not more.
 
@@
 XML Schema built-in types of elements and attributes can be inferred from the XML documents by analysing their content.
-Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available.
-This is when a set of XQuery queries may come in handy.
+Since the number of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available.
+This is a case when a set of XQuery queries may bring optimizations.
 For example, consider the following occurrences of element \texttt{a}.
@@
 These three occurrences are not sufficient enough to determine the accurate type of element \texttt{a}, because values \texttt{1}, \texttt{6}, \texttt{18} are valid values of several types: \texttt{decimal}, \texttt{integer}, \texttt{byte}, \texttt{short}, \texttt{unsignedInt}, \texttt{positiveInteger} and many others.
-The following text introduces some possibilities of a type inference from queries.
 
 If a value of an element or attribute is used in an expression, then this expression could be often exploited to determine a type of the value.
-Comparing the value to another value of a known type and supplying the value to a function call as an argument of particular type are examples of these expressions as shown below.
+Comparing the value to another value of a known type and supplying the value to a function call as an argument of a particular type are examples of these expressions as shown below.
 
@@
 \begin{verbatim}
-every $number as element(*,xs:integer) in //number satisfies ($number > 0)
+every $number as element(*,xs:integer) in //number
+  satisfies ($number > 0)
 
@@
 In several cases, XQuery \emph{if-then-else} construct may be used to branch the execution of query by all possible values of a certain variable.
-This is equivalent of \emph{switch} construct from other programming languages.
+This is equivalent to \emph{switch} construct from other programming languages.
 When the control variable is bound to some element or attribute, type of this node can be inferred as enumeration and its individual values can be determined as well.
@@
 
-The following examples are queries that could be processed by a similar method but the actual method will not use them.
+The following examples are queries that could be processed by a similar method, but the actual method will not use them.
 
@@
 
-A common three-way join.
+A common three-way join. \todo{Tady asi něco chybí.}
 
@@
 \subsection{Negative Statements about Uniqueness} \label{Negative_uniqueness}
-In many cases a statement refusing uniqueness of element or attribute values could be inferred.
+In many cases a statement refusing uniqueness of element or attribute values can be inferred.
 Such statements may be helpful in combination with other methods in its process of making a decision whether a particular element or attribute is unique or not.
@@
 
-Usage of \texttt{for} construct indicates that a sequence of \texttt{book} elements which satisfy the condition in \emph{where} clause is expected.
+Usage of \texttt{for} construct indicates that a sequence of \texttt{book} elements which satisfy the condition in \emph{where} \todo{tady se nám nějak změnil styl} clause is expected.
 It is a condition composed of two single conditions joined by \emph{and} logical operator.
@@
 
-The first of them is test of equality of book's subelement \texttt{publisher} to a string literal.
+The first of them is a test of equality of book's subelement \texttt{publisher} to a string literal.
 Based on the expectation, element \texttt{publisher} cannot be unique.
@@
 
+\todo[inline]{Tady to chce ějaký komentář.}
+
 \subsection{Uniqueness}
@@
 For each \texttt{product} element, new element \texttt{prod} with attribute \texttt{xml:id} is created.
-Since attribute \texttt{xml:id} is supposed to be unique and there is direct transformation of the values of \texttt{number} elements to the values of attribute \texttt{xml:id}, it is very likely that element \texttt{number} is unique in the source data.
+Since attribute \texttt{xml:id} is supposed to be unique and there is a direct transformation of the values of \texttt{number} elements to the values of attribute \texttt{xml:id}, it is very likely that element \texttt{number} is unique in the source data.
 
@@
 \begin{verbatim}
-let $i := //incision)[2]
-for $a in //action)[. >> $i]
+let $i := //incision[2]
+for $a in //action[. >> $i]
 return $a//instrument
@@
 \subsection{Intermediate XML structure}
-Intermediate XML structure are XML data that are not read from input XML documents nor created as an output of queries but they are somehow used by the queries.
+Intermediate XML structure represents XML data that are neither read from input XML documents nor created as an output of queries but they are somehow used by the queries.
 Objective of this section is intermediate XML structure created directly in the queries.

### 2012-02-13 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

@@
 
+\todo[inline]{Určitě to chce kombinaci dotazů a dat.
+Na začátku textu říkáme, že samotné dotazy nestačí, resp.
+neříkají všechno.
+Ale klidně můžete v textu popsat oboje a pak implementovat jen část z obojího a porovnat výsledky.
+Nebo vzít nějakou metodu, která je v jInfer (DP p.
+Klempy), rozšířit ji o zpracování dotazů a na nějakých pěkných datech ukázat, že je to rychlejší/dává lepší (přesnější) výsledky.}
+
 \section{Forms of Query Precessing}
@@
 
-The result of lexical and syntax analyses can be some kind of so-called syntax tree \cite{compilers}.
-It is a structure representing a word according to a formal grammar of some language.
+The result of lexical and syntax analyses can be a kind of so-called syntax tree \cite{compilers}.
+It is a structure representing a word according to a formal grammar of a language.
 In our case, the language is XQuery, its grammar is defined in REF \todo{ref} and every query is a word of the XQuery language.
@@
 Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"} is of some numeric type.
-And, for the function is parametrized there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.
+And, for the function is parametrized \todo{divná věta} there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.
 
@@
 Therefore, we decided to incorporate query processing using the lexical and syntax analyses.
-
-\todo[inline]{Tak náročnost implementace rozhodně není kritérium.
-Ta se udělá jednou.
-Navíc ve vědě (tj.
-i DP) nemá opodstatnění.
-Pro DP navíc platí, že nemusíte implementovat vše, co navrhnete.
-Implementace je pomocná věc, která slouží k tomu, abychom experimentálně ukázali vlastnosti a chování navrženého algoritmu.}
 
@@
 When we infer some subelements of a certain element we often cannot be sure about their number of occurrences.
-Also, we cannot be sure if every occurrence of the certain element contains these subelements and we even do not know if at least one occurrence of the element contains them.
+Also, we cannot be sure if every occurrence of a certain element contains the subelements and we even do not know if at least one occurrence of the element contains them.
 Thus, the inferred statement is more likely an indication than a fact about the structure.
+
+\todo[inline]{S proč potřebuju vyhonotit dotazy?
+Dostanu nějaký výsledek, ale ten může obsahovat třeba úplně jiná data ne jsou v dokumentech - použiju konstruktory, něco sečtu a původní idokumenty můžou být úplně jiné.}
 
@@
 
+\todo[inline]{Ta věta je nejasná.
+Spíš pro to odvození struktury použijeme některý standardní přístup analyzující data a dotazy použijeme pro další upřesnění, ne?
+Nebo strukturu nebudeme řešit vůbec?
+To je asi divné.}
+
 \section{Extension of an Existing Approach}
-The existing approaches of XML schema inference deal mainly with the inference of XML structure.
-As stated in the previous section, this work does not deal with the inference of XML structure, and hence, the extension of an existing approach will be a kind of an independent addition instead of modification and refinement of its core algorithm.
+The existing approaches of XML schema inference deal mainly with inference of XML structure.
+Hence, the extension of an existing approach will be a kind of an independent addition instead of modification and refinement of its core algorithm.
 
@@
 This phase may also infer additional statements from both the XML data and the queries or it may try to verify the statements from the second phase with respect to the XML data.
+
+\todo[inline]{Také bychom mohli vzít přístup odvozující strukturu a informace z dotazů využít v  jeho průběhu.
+Obecně tyto metody vytvoří prefixový strom z iniciální gramatiky a pak se ho snaží zobecňovat sléváním stavů a zobecňováním automatu.
+Např.
+můžeme mít pravidlo, že pokud se element A vyskytuje alespoň 5x, pravděpodobně se může vyskytovat libovolněkrát.
+My ale v datech najdeme element jenom 4x.
+Pak nevím jeslti je pravidlo rozumné.
+Dotazy nám to mohou potvrdit/vyvrátit.
+Čili můžeme zpřesnit proces odvozování.
+Nebo ho můžeme zoptimalizovat. Např. Vošta ve své diplomce poměrně složitě analyzuje automat aby našek konstrukt xs:all. My bychom to pomocí analýzy dotazů mohli umět jednodušeji = efektivněji.
+Nebo jsme schopni pomocí dotazů přesněji určit kontext elementu (opět řeší Vošta v diplomce).
+Např.
+mám element jméno, který se vyskytuje v různých kontextech (absolutních cestách).
+Já ale nevím, jeslti ten element má stený význam a tudíž by pro něj měl být odvozen stejný model obsahu.
+Např.
+jméno autora může být složeno z podelementu křestní a příjmení, zatímco jméno knihy bude prostý string.
+Vošta používá podobnostní funkci, ale tam je potřeba nastavit threshold.
+Pokud ale budu mít dotaz, který mi tyto elementy dává do jedné proměnné (např.
+jméno zákazníka i jméno autora), zřejmě jsou sémanticky podobné a mohou mít stejný model obsahu.
+Stejný dotaz pro knihy i autory by byl asi divný.
+Tohle by asi nebylo špatné minimálně odiskutovat.
+Mrkněte se např.
+na diplomku Vošty (popř.
+Klempy).
+Jsou tam popsané typické postupy pro odvozování a základní pravidla slévání.
+U nich by také dotazy mohly pomoci.}

### 2012-02-13 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 It writes its result into a file in an XML representation.
-Instead, we need to keep the result in memory and pass it to consecutive steps of our algorithm.
+Instead, we need to keep the result in the main memory and pass it to consecutive steps of our algorithm.
 Though this requirement concerns modifications of implementation, the core of the processing remains untouched.
@@
 \label{DEF_syntax_tree}
-Syntax tree of XQuery query $Q$ is tuple $T = (V, E, c, \mathcal{P}, o)$ where $V \subset \mathbb{N}$ is a set of nodes, each node representing a particular XQuery construct in query $Q$, $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ ($b$ is a child of $a$) in query $Q$, $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}, $\mathcal{P}$ is a set of functions specifying additional properties of the nodes and distinguishing the nodes of the same classes, and $o : V \rightarrow \mathcal{O}$ is a function specifying an order of children of the nodes where $\mathcal{O} = \{o_v : E_v \rightarrow \mathbb{N} | v \in V, E_v = \{(v, w) | w \in V, (v,w) \in E\}\}$ is set of functions specifying the children order for each node.
+Syntax tree of XQuery query $Q$ is tuple $T = (V, E, c, \mathcal{P}, o)$ where
+\begin{itemize}
+\item $V \subset \mathbb{N}$ is a set of nodes, each node representing a particular XQuery construct in query $Q$,
+\item $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ ($b$ is a child of $a$) in query $Q$,
+\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}, \todo{Tohle je hrozné peklo.
+Potřebujeme to takto?
+Nedalo by se to nacpat do 1-2 tabulek/obrázků?
+Je nutné aby každá položky byla na samostatném řádku.
+Trochu to vypadá jako umělé natahování diplomky.}
+\item $\mathcal{P}$ is a set of functions specifying additional properties of the nodes and distinguishing the nodes of the same classes,
+\item and $o : V \rightarrow \mathcal{O}$ is a function specifying an order of children of the nodes, where $\mathcal{O} = \{o_v : E_v \rightarrow \mathbb{N} | v \in V, E_v = \{(v, w) | w \in V, (v,w) \in E\}\}$ is set of functions specifying the children order for each node.
 For every $v \in V, o(v) = o_v$ so that $o_v(v,w)$ is a sequential number of a construct represented by $w$ amongst constructs represented by children of $v$ in query $Q$.
+\end{itemize}
 \end{define}
@@
 
-Set $\mathcal{P}$ contains other similar functions but due to their large count, we do not define them formally.
-\todo[inline]{Moze takto ostat, alebo definovat aspon tie co pouzivame? Alebo ich aspon vypisat?}
+Set $\mathcal{P}$ contains other similar functions but due to their large number, we do not define them formally.
+\todo[inline]{Q: Moze takto ostat, alebo definovat aspon tie co pouzivame? Alebo ich aspon vypisat? A: aspoň příklad, lépe vypsat ty, co použijeme}
+
+\todo[inline]{Hodil by se příklad stromu - jeslti nebude dál.}
 
@@
 
-\todo[inline]{Je nutne presnejsie popisat kedy a preco tieto skratky maju zmysel? V pseudo-kodoch budu samozrejme pouzivane len v pripadoch, ked su dobre definovane a maju zmysel.}
+\todo[inline]{Q: Je nutne presnejsie popisat kedy a preco tieto skratky maju zmysel?
+V pseudo-kodoch budu samozrejme pouzivane len v pripadoch, ked su dobre definovane a maju zmysel.
+A: Ne, ale hodilo by se slovně popsat co znamenají.}
 
@@
 This hierarchy is shown in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}.
-The classes with emphasised names represent a common class of a group of subclasses, and these classes cannot be directly used in the syntax tree (abstract classes).
+The classes with names in bold represent a common class of a group of subclasses, and these classes cannot be directly used in the syntax tree (abstract classes).
 The remaining nodes represent particular constructs of the XQuery language and nodes of these classes can be used in the syntax tree.
@@
 
-\todo[inline]{Asi bude potrebne nakreslit pre kazdy node, co presne obsahuje. Alebo staci pre tie, ktore budeme my pouzivat?}
+\todo[inline]{Q: Asi bude potrebne nakreslit pre kazdy node, co presne obsahuje.
+Alebo staci pre tie, ktore budeme my pouzivat?
+A: Určitě všude stačí popsat jen to, co použijeme, s poznámkou, že zbytek je např.
+v diplomce, ze které vycházíte.}
 
@@
 Some pseudo code algorithms in the following sections need to determine a class of a node.
-The node's direct class can be determined by function $c$ from the definition of the syntax tree \ref{DEF_syntax_tree}, and hence, we also can determine its indirect classes.
+The node's direct class can be determined by function $c$ from the definition of the syntax tree in Definition \ref{DEF_syntax_tree}, and hence, we also can determine its indirect classes.
 For the purpose of pseudo code, we define the following function.
-For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
+
+\begin{define}[Function is(v, class)]For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
 Otherwise, it returns $false$.
+\end{define}
 
@@
 The representation of a definition of a local variable in the syntax tree is a node of $VariableBinding$ class.
-Nodes of that class contain have only two children; a node representing the type of the variable and a node representing the binding expression (expression defining the value of the variable, and thus, it cannot use the variable).
-Hence, the entire subtree doesn't contain any expressions that use the variable.
+Nodes of that class have only two children; a node representing the type of the variable and a node representing the binding expression (expression defining the value of the variable, and thus, it cannot use the variable).
+Hence, the entire subtree does not contain any expressions that use the variable.
 Therefore the scope of the new variable is not the subtree of the $VariableBinding$ class node.
@@
 
-For example, a node of $FLWORExpr$ class contains four subnodes of $TupleStream$, $WhereClause$, $OrderByClause$ and $ReturnClause$ classes.
+For example, a node of $FLWORExpr$ class contains four subnodes of\linebreak $TupleStream$, $WhereClause$, $OrderByClause$ and $ReturnClause$ classes.
 The $TupleStream$ class node contains a list of node of $variableBinding$ class which define variables valid in all other three subnodes of the FLWOR node.
@@
 \item XML Schema built-in atomic types \todo{link do appendixu}.
-\item Types $ElementType$, $AttributeType$, $NodeType$, $TextNodeType$, $CommentType$, $ProcessingInstructionType$, $DocumentType$ representing an element, attribute, node, text node, comment, processing instruction, document node.
+\item Types $ElementType$, $AttributeType$, $NodeType$, $TextNodeType$,\linebreak $CommentType$, $ProcessingInstructionType$, $DocumentType$ representing an element, attribute, node, text node, comment, processing instruction, document node.
 \item Type representing a node selected by a certain path expression. The path expression is included in this type. Let this type be identified as $PathType$.
@@
 
-$PathType$ has always cardinality of zero-or-more, because the result of a path expression is a general sequence and since we do not evaluate the queries we cannot determine its actual cardinality.
+$PathType$ has always cardinality of zero-or-more, because the result of a path expression is a general sequence and since we do not evaluate the queries, we cannot determine its actual cardinality.
 
@@
 Vysledkom tejto funkcie by totiz zas mohli byt elementy/atributy o ktorych vieme (pozmenenu) path expression a nie len nezname elementy/atributy.
-A: Přesně tak. A jak říkám - není třeba vše co navrhnete implementovat.}
+A: Přesně tak. A jak říkám - není třeba vše co navrhnete implementovat.
+A2: Stačí to popsat a v kapitole Implementace říct, které ásti algoritmu implementujeme. Stačí takové, které už ukáží zajímavé chování a dokážou přínosy návrhu. Přesně k tomu implmentace je.}
 
@@
 Determination of return types of functions is needed because function calls can appear in expressions.
-A return type of a function can be determined at the moment the analysis of expressions encounters a call of the function, however, it involves multiple transitions of the syntax tree in a search for a definition of a particular function.
+A return type of a function can be determined at the moment the analysis of expressions encounters a call of the function; however, it involves multiple transitions of the syntax tree in a search for a definition of a particular function.
 
@@
 A simple algorithm is presented in Algorithm \ref{ALG_processing_of_functions}.
-It uses \texttt{getFunctionDeclarationNodes()} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three.
+It uses \texttt{getFunction\-DeclarationNodes()} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three.
 Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.
@@
 We will not determine types of functions defined in other modules since the principle is similar but it is needed to look up the definitions in syntax trees of other queries.
-Return types of built-in functions are fixed and, thus, there is no need to analyse them.
+Return types of built-in functions are fixed and, thus, there is no need to analyze them.
 

### 2012-02-13 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \begin{algorithmic}[1]
-\REQUIRE{\ \\
-    $getFunctionDeclarationNodes()$: Function defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}. \\
-    $getTypeTN()$: Function defined in Algorithm \ref{ALG_extraction_of_a_type_from_typenode}. \\
-    $syntaxTree$: A reference to the root node of a syntax tree.
-}
+\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
 \ENSURE An associative array of function names with their types and references to the their declaration nodes.
@@
 \STATE $functionArray$ = an empty associative array
-\FORALL{$functionDeclarationNode \in getFunctionDeclarationNodes(syntaxTree)$}
+\FORALL{$functionDeclarationNode \in$ \\ $getFunctionDeclarationNodes(syntaxTree)$}
     \STATE $functionName = functionDeclarationNode.funcName$
     \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
-    \STATE $functionArray[functionName] = (getTypeTN(typeNode), functionDeclarationNode)$
+    \STATE $functionArray[functionName] =$ \\ $(getTypeTN(typeNode), functionDeclarationNode)$
 \ENDFOR
@@
 
-Later phases of the algorithm may need to process the declarations of certain functions.
+Later phases of the algorithm need to process the declarations of certain functions.
 Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.
@@
 It is used to assign a node with a named value, for example, $set(someNode, color, 'red')$ will assign $someNode$ with string value $'red'$ wich can be then retrieved by function $get()$, specifying the node and the property name.
-For example, $get(someNode, color)$ returns $'red'$.
+For example,\linebreak $get(someNode, color)$ returns $'red'$.
 A subsequent call of $set()$ assigning a node with a value of already assigned property will overwrite it, leaving the property with the newer value.
-However, a property can has several values and this can be achieved using function $add()$ with the same syntax as function $set()$ but instead of overwriting the existing property, $add()$ will append the new value to the existing ones.
+However, a property can have several values and this can be achieved using function $add()$ with the same syntax as function $set()$ but instead of overwriting the existing property, $add()$ will append the new value to the existing ones.
 Then, a call of $get()$ on this node and property returns a list of all its values, preserving the order of their addition.
@@
 \REQUIRE{\ \\
-    $getFunctionReturnType()$: Function returning a type of a function. \\
-    $getVariableType()$: Function returning a type of a variable. \\
-    $determineOperatorType()$: Function defined in \ref{ALG_func_determineOperatorType}. \\
-    $getType()$: Function returning a type of already processed expression. \\
     $exprNode$: An expression node in the syntax tree. \\
@@
 \REQUIRE{\ \\ 
-    $getType()$: Function returning a type of already processed expression. \\
-    $isOperatorClassComparison()$: Function defined in \ref{ALG_func_isOperatorClassComparison}. \\
-    $isOperatorClassAddition()$: Function defined in \ref{ALG_func_isOperatorClassAddition}. \\
-    $isNumericType()$: Function defined in \ref{ALG_func_isNumericType}. \\
-    $selectMoreGenericNumericType()$: Function that select a more generic numeric XSD type from two given types according to the type promotions rules (TODO-REF). \\
     $operatorNode$: An operator expression node in the syntax tree. \\
@@
     \STATE $rightOperandType = getType(exprNode.rightSide)$
-    \IF{$isNumericType(leftOperandType)) AND isNumericType(rightOperandType)$}
-        \STATE $type = selectMoreGeneralNumericType(leftOperandType, rightOperandType)$
+    \IF{$isNumericType(leftOperandType)$ $AND$ $isNumericType(rightOperandType)$}
+        \STATE $type = selectMoreGeneralNumericType(leftOperandType,$ $rightOperandType)$
     \ELSIF{$isNumericType(leftOperandType)$}
@@
 Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_4}), its type can be either determined directly or it depends on its subexpressions.
-A part of the function's semantic is separated in \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
+A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
 
-Function $determineExpressionType()$ is called within function $analysisOfExpressionTypes$ presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
+Function $determineExpressionType$ is called within function $analysis\-Of\-Expression\-Types$ presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
 Two important questions about this function emerge.
@@
 The order of the recursion is to firstly process children of a node and then the node itself.
-The order of children processing is specified by function $o$ from the definition of the syntax tree \ref{DEF_syntax_tree}.
+The order of child processing is specified by function $o$ from the definition of the syntax tree in Definition \ref{DEF_syntax_tree}.
 The reason is that an expression node may need to know types of its subexpressions to determine its own type and we want to determine types of all expressions.
@@
 \REQUIRE{\ \\
-    $memorizeType()$: Function memorizing a determined type of a specified node. \\
-    $set()$: Function memorizing a specified property of a specified node. \\
-    $add()$: Function adding a value to a list of specified properties of a specified node. \\
-    $get()$: Function returning a specified property (or list) set by functions $set()$ and $add()$. \\
-    $mergeContextVarTypes()$: Function defined in Algorithm \ref{ALG_function_mergeContextVarTypes}. \\
     $startingNode$: A node determining a subtree to perform the analysis on. \\
@@
     \ELSE
-        \STATE $type = determineExpressionType(startingNode.getChild(BindingSequenceNode).getChild(ExprNode), contextVarTypes)$
+        \STATE $type = determineExpressionType(startingNode.$ $getChild(BindingSequenceNode).getChild(ExprNode),$ $contextVarTypes)$
     \ENDIF
@@
 \REQUIRE{\ \\
-    $getGlobalVariableType()$: Function returning a determined type of a specified global variable. \\
     $contextVarTypes$: Types of local variables valid in current context. \\
@@
 
-The types of processed global variables are available trough function $getGlobalVariableType()$.
+The types of processed global variables are available trough function \linebreak $getGlobalVariableType()$.
 If the function gets a variable that is not a global one, it returns $null$.
@@
 \subsection{Determination of Expression Types}
-To determine types of expressions, we use already introduced function $analysisOfExpressionTypes$.
+To determine types of expressions, we use already introduced function \linebreak $analysisOfExpressionTypes$.
 The starting node (its first argument) is the node representing the query body and the variable context is empty as there cannot be any local variable valid in the body node.
@@
 
-If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression node is $Operator$ and it represents one of $GEN\_LESS\_THAN$, $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$, $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.
+If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression node is $Operator$ and it represents one of $GEN\_LESS\_THAN$, \linebreak $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$,\linebreak $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$,\linebreak $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.
 
@@
 \section{Step 4: Key Discovery}
-In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in section \ref{Keys}, and extending it.
-Firstly, the syntax tree is searched for certain forms of FLWOR expressions to infer the keys from.
-Then, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.
+In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it.
+Firstly, the syntax tree is searched for forms of FLWOR expressions to infer the keys from.
+Then, as described later, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.
 
@@
 In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes.
-Actually, the join patterns from \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly because their structure in the syntax tree is different.
+Actually, the join patterns from Listings \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly because their structure in the syntax tree is different.
 
@@
 
-As described in the original method and summarized in section \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases.
+As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases.
 For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
@@
 
-\STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).getChildren()$
+\STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).$ $getChildren()$
 
@@
 \FORALL{$bindingNode \in bindingNodes$}
-    \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).getChild(ExprNode)$
+    \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
     \IF{$getType(BindingExpr) == PathType$ using only child and descendant axes, ending with exactly one predicate $P$}
         \FORALL{$(var, node) \in forVars$}
-            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var, forVars)$
+            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var,$ $forVars)$
         \ENDFOR
@@
 \end{algorithm}
+
+\todo[inline]{Alg 6.12 FLWOR expressions - pre podmienky na riadkoch 13 a 17 definovat nejake funkcie}
 
@@
 While the previous steps of the key discovery produce positive statements about the keys, in this step, the algorithm searches for evidence on non-uniqueness of elements and attributes.
-An aim of this step is to eliminate or decrease the number of falsely inferred keys.
+The aim of this step is to eliminate or decrease the number of falsely inferred keys.
 \todo[inline]{Napisat niekam, ze odvodzovanie kludov predpoklada, ze kazdy element/element join je robeny cez key-foreign key pair.}
@@
 \todo[inline]{Toto asi nebude uplne pravda pre nejake exoticke predikaty.
-Overit pri implementacii.}
+Overit pri implementacii.
+A: Kdyžtak popište.}
 
@@
         \IF{$is(bindingNode, ForClauseNode)$}
-            \STATE $expr = bindingNode.getChild(BindingSequenceNode).getChild(ExprNode)$
+            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
             \IF{$getType(expr) == PathExpr$ without predicates, using only child and descendant axes}

### 2012-02-20 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \item $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ ($b$ is a child of $a$) in query $Q$,
-\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}, \todo{Tohle je hrozné peklo.
+\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \todo{Tohle je hrozné peklo.
 Potřebujeme to takto?
@@
 The node classes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where an object can be of several types.
-This hierarchy is shown in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \ref{FIG_syntax_tree_node_types_part_3}, \ref{FIG_syntax_tree_node_types_part_4}, \ref{FIG_syntax_tree_node_types_part_5}, \ref{FIG_syntax_tree_node_types_part_6}, \ref{FIG_syntax_tree_node_types_part_7}, \ref{FIG_syntax_tree_node_types_part_8}, \ref{FIG_syntax_tree_node_types_part_9}.
+This hierarchy is shown in tables in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}.
+The tables are composed of classes in italic and their subclasses.
 The classes with names in bold represent a common class of a group of subclasses, and these classes cannot be directly used in the syntax tree (abstract classes).
-The remaining nodes represent particular constructs of the XQuery language and nodes of these classes can be used in the syntax tree.
+The remaining non-bold nodes represent particular constructs of the XQuery language and nodes of these classes can be used in the syntax tree.
 
@@
 \label{FIG_syntax_tree_node_types_part_1}
-\begin{description}
-\item \textbf{Node}
-\begin{description}
-\item AttributeNode (I)
-\item AttrListNode (IL)
-\item AxisNode (I)
-\item CaseClauseNode (I)
-\item CaseClausesNode (I)
-\item CDataSectionNode (L)
-\item ContentNode (IL)
-\item DefaultCaseNode (I)
-\item EntityRefNode (L)
-\item \textbf{ExprHolderNode}
-\item \textbf{ExprNode}
-\item FunctionBodyNode (IL)
-\item CharRefNode (L)
-\item InClausesNode (I)
-\item \textbf{ItemTypeNode}
-\item LocationHintNode (L)
-\end{description}
-\end{description}
+\begin{tabular}{l l | l}
+\multicolumn{2}{l|}{\textbf{\emph{Node}}} & \textbf{\emph{ExprNode}} \\
+\hline
+AttributeNode (I) & ModuleNode (I) & CommaOperatorNode (I) \\
+AttrListNode (IL) & NameNode (IL) & ConstructorNode (I) \\
+AxisNode (I) & OrderByClauseNode (I) & ContextItemExprNode (L) \\
+CaseClauseNode (I) & OrderSpecNode (I) & EmptySequenceNode (L) \\
+CaseClausesNode (I) & ParamListNode (I) & ExtensionExprNode (I) \\
+CDataSectionNode (L) & ParamNode (I) & FLWORExprNode (I) \\
+ContentNode (IL) & PITargetNode (IL) & FunctionCallNode (IL) \\
+DefaultCaseNode (I) & PragmaListNode (I) & IfExprNode (I) \\
+EntityRefNode (L) & PragmaNode (L) & LiteralNode (L) \\
+\textbf{ExprHolderNode} & PredicateListNode (I) & OperatorNode (I) \\
+\textbf{ExprNode} & \textbf{PrologChildNode} & OrderedExprNode (I) \\
+FunctionBodyNode (IL) & StepExprNode (I) & PathExprNode (I) \\
+CharRefNode (L) & StringNode (L) & QuantifiedExprNode (I) \\
+InClausesNode (I) & TupleStreamNode (I) & TypeswitchExprNode (I) \\
+\textbf{ItemTypeNode} & TypeNode (L) & UnorderedExprNode (I) \\
+LocationHintNode (L) & \textbf{VariableBindingNode} & ValidateExprNode (I) \\
+LocationHintsNode (IL) & VarValueNode (IL) & VarRefNode (L) \\
+\textbf{ModuleChildNode} & & \\
+\end{tabular}
 \end{figure}
@@
 \label{FIG_syntax_tree_node_types_part_2}
-\begin{description}
-\item \textbf{Node}
-\begin{description}
-\item LocationHintsNode (IL)
-\item \textbf{ModuleChildNode}
-\item ModuleNode (I)
-\item NameNode (IL)
-\item OrderByClauseNode (I)
-\item OrderSpecNode (I)
-\item ParamListNode (I)
-\item ParamNode (I)
-\item PITargetNode (IL)
-\item PragmaListNode (I)
-\item PragmaNode (L)
-\item PredicateListNode (I)
-\item \textbf{PrologChildNode}
-\item StepExprNode (I)
-\item StringNode (L)
-\item TupleStreamNode (I)
-\item TypeNode (L)
-\item \textbf{VariableBindingNode}
-\item VarValueNode (IL)
-\end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 3.}
-\label{FIG_syntax_tree_node_types_part_3}
-\begin{description}
-\item \textbf{ExprHolderNode}
-  \begin{description}
-  \item BindingSequenceNode (I)
-  \item DefaultValueNode (I)
-  \item ElseExpressionNode (I)
-  \item OperandExpressionNode (I)
-  \item ReturnClauseNode (I)
-  \item TestExpressionNode (I)
-  \item ThenExpressionNode (I)
-  \item WhereClauseNode (I)
-  \end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 4.}
-\label{FIG_syntax_tree_node_types_part_4}
-\begin{description}
-\item \textbf{ExprNode}
-  \begin{description}
-  \item CommaOperatorNode (I)
-  \item ConstructorNode (I)
-  \item ContextItemExprNode (L)
-  \item EmptySequenceNode (L)
-  \item ExtensionExprNode (I)
-  \item FLWORExprNode (I)
-  \item FunctionCallNode (IL)
-  \item IfExprNode (I)
-  \item LiteralNode (L)
-  \item OperatorNode (I)
-  \item OrderedExprNode (I)
-  \item PathExprNode (I)
-  \item QuantifiedExprNode (I)
-  \item TypeswitchExprNode (I)
-  \item UnorderedExprNode (I)
-  \item ValidateExprNode (I)
-  \item VarRefNode (L)
-  \end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 5.}
-\label{FIG_syntax_tree_node_types_part_5}
-\begin{description}
-\item \textbf{ItemTypeNode}
-  \begin{description}
-  \item AnyItemNode (L)
-  \item AtomicTypeNode (L)
-  \item KindTestNode (IL)
-  \item NameTestNode (L)
-  \end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 6.}
-\label{FIG_syntax_tree_node_types_part_6}
-\begin{description}
-\item \textbf{ModuleChildNode}
-  \begin{description}
-  \item ModuleDeclNode (L)
-  \item PrologNode (IL)
-  \item QueryBodyNode (I)
-  \end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 7.}
-\label{FIG_syntax_tree_node_types_part_7}
-\begin{description}
-\item \textbf{PrologChildNode}
-  \begin{description}
-  \item BaseURIDeclNode (L)
-  \item BoundarySpaceDeclNode (L)
-  \item ConstructionDeclNode (L)
-  \item CopyNamespacesDeclNode (L)
-  \item DefaultCollationDeclNode (L)
-  \item DefaultNamespaceDeclNode (L)
-  \item EmptyOrderDeclNode (L)
-  \item FunctionDeclNode (I)
-  \item \textbf{ImportNode}
-    \begin{description}
-    \item ModuleImportNode (I)
-    \item SchemaImportNode (I)
-    \end{description}
-  \item NamespaceDeclNode (L)
-  \item OptionDeclNode (L)
-  \item OrderingModeDeclNode (L)
-  \item VarDeclNode (I)
-  \end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 8.}
-\label{FIG_syntax_tree_node_types_part_8}
-\begin{description}
-\item StepExprNode (I)
-  \begin{description}
-  \item SelfOrDescendantStepNode (L)
-  \end{description}
-\end{description}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 9.}
-\label{FIG_syntax_tree_node_types_part_9}
-\begin{description}
-\item \textbf{VariableBindingNode}
-  \begin{description}
-  \item ForClauseNode (I)
-  \item InClauseNode (I)
-  \item LetClauseNode (I)
-  \end{description}
-\end{description}
+\begin{tabular}{l | l}
+\textbf{\emph{PrologChildNode}} & \textbf{\emph{ExprHolderNode}} \\
+\hline
+BaseURIDeclNode (L) & BindingSequenceNode (I) \\
+BoundarySpaceDeclNode (L) & DefaultValueNode (I) \\
+ConstructionDeclNode (L) & ElseExpressionNode (I) \\
+CopyNamespacesDeclNode (L) & OperandExpressionNode (I) \\
+DefaultCollationDeclNode (L) & ReturnClauseNode (I) \\
+DefaultNamespaceDeclNode (L) & TestExpressionNode (I) \\
+EmptyOrderDeclNode (L) & ThenExpressionNode (I) \\
+FunctionDeclNode (I) & WhereClauseNode (I) \\
+\cline{2-2} \textbf{ImportNode} & \textbf{\emph{ItemTypeNode}} \\
+\cline{2-2} NamespaceDeclNode (L) & AnyItemNode (L) \\
+OptionDeclNode (L) & AtomicTypeNode (L) \\
+OrderingModeDeclNode (L) & KindTestNode (IL) \\
+VarDeclNode (I) & NameTestNode (L) \\
+\cline{1-2} \textbf{\emph{ImportNode}} & \textbf{\emph{VariableBindingNode}} \\
+\cline{1-2} ModuleImportNode (I) & ForClauseNode (I) \\
+SchemaImportNode (I) & InClauseNode (I) \\
+\cline{1-1} \textbf{\emph{ModuleChildNode}} & LetClauseNode (I) \\
+\cline{1-2} ModuleDeclNode (L) & \emph{StepExprNode} (I) \\
+\cline{2-2}PrologNode (IL) & SelfOrDescendantStepNode (L) \\
+QueryBodyNode (I) & 
+\end{tabular}
 \end{figure}
@@
 \todo[inline]{Napisat niekam, ze odvodzovanie kludov predpoklada, ze kazdy element/element join je robeny cez key-foreign key pair.}
-When the algorithm discovers that a certain element is probably not unique, it assign the element with a negative weight that will decrease the sum of its weights in the summarizing step.
+When the algorithm discovers that a certain element is probably not unique, it assigns the element with a negative weight that will decrease the sum of its weights in the summarizing step.
 
@@
 The algorithm searches for calls of the mentioned functions with a $PathType$ on their input.
-Elements and attributes selected by those $PathTypes$ are then considered to be not unique, and consequently, they can not be keys.
+Elements and attributes selected by those $PathTypes$ are then considered to be not unique, and consequently, they cannot be keys.
 
@@
 \ENDIF
-
-\FORALL{$subnode \in node.getChildren()$}
-    \STATE call this function recursively with $subnode$
-\ENDFOR
-
 \end{algorithmic}
@@
 
+\IF{$is(node, FLWORExprNode)$}
 	\STATE $forVars =$ an empty array
 
-\IF{$is(node, FLWORExprNode)$}
     \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
@@
     \ENDFOR
-\ENDIF
 
@@
     	\STATE $processWhere(forVars, whereClause.getChild(ExprNode))$
+	\ENDIF
 \ENDIF

### 2012-02-21 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \item $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ ($b$ is a child of $a$) in query $Q$,
-\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}, \todo{Tohle je hrozné peklo.
-Potřebujeme to takto?
-Nedalo by se to nacpat do 1-2 tabulek/obrázků?
-Je nutné aby každá položky byla na samostatném řádku.
-Trochu to vypadá jako umělé natahování diplomky.}
+\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2},
 \item $\mathcal{P}$ is a set of functions specifying additional properties of the nodes and distinguishing the nodes of the same classes,
@@
 \subsection{Closer Look at the Nodes of the Syntax Tree}
-\todo[inline]{Dat figures do appendixu?}
 The node classes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where an object can be of several types.
@@
 \item Types $ElementType$, $AttributeType$, $NodeType$, $TextNodeType$,\linebreak $CommentType$, $ProcessingInstructionType$, $DocumentType$ representing an element, attribute, node, text node, comment, processing instruction, document node.
-\item Type representing a node selected by a certain path expression. The path expression is included in this type. Let this type be identified as $PathType$.
+\item Type representing a node or a set of nodes selected by a certain path expression. The path expression is included in this type. Let this type be identified as $PathType$.
+\item $UnknownType$ representing a type without known details, which does not suit one of the three previous types. An example is XSD type $anyType$.
 \end{itemize}
 
-To capture sequences, we assign each type with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}.
-Every type can be perceived as a sequence.
+\todo[inline]{Doplnit PathType o boolean for bound, volania zoznam funkcii volanych na tuto path.}
+
+To capture sequences, we assign the first two categories of types (all types except of $PathType$ and $UnknownType$) with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}.
+Each of those types can be perceived as a sequence.
 A type representing one value or one item can be perceived as a sequence of exactly one item.
@@
 
-$PathType$ has always cardinality of zero-or-more, because the result of a path expression is a general sequence and since we do not evaluate the queries, we cannot determine its actual cardinality.
-
-If we cannot determine the type of an expression, we express that using $null$. An example of such case is path expression $/rootElement/(elementA * 2)$. \todo{Nejak dopisat}
+$PathType$ is not assigned with the cardinality since we do not evaluate the queries, and therefore, we cannot determine if a certain XQuery path targets zero, one or more nodes.
+Alike, $UnknownType$ is neither assigned with the cardinality.
+Expressions of $UnknownType$ are not utilized it the inference, therefore, their cardinality is not needed.
 
@@
 \ELSIF{$is(exprNode, PathExprNode)$}
-     \IF{$exprNode$ selects an element set}
-         \STATE $type = ElementType$, zero-or-more cardinality, $exprNode$ included
-     \ELSIF{$exprNode$ selects an attribute set}
-         \STATE $type = AttributeType$, zero-or-more cardinality, $exprNode$ included
-     \ENDIF
+     \STATE $type = createPathType(exprNode)$
 \ELSIF{$is(exprNode, OperatorNode)$}
@@
      \STATE $returnClauseNode = exprNode.getChild(ReturnClauseNode)$
-     \STATE $type = getType(returnClauseNode)$
+     \STATE $type = createForUnboundType(getType(returnClauseNode))$
 \ENDIF
@@
         \STATE $type = determineExpressionType(startingNode.$ $getChild(BindingSequenceNode).getChild(ExprNode),$ $contextVarTypes)$
+        \IF{$is(startingNode, ForClauseNode)$}
+            \STATE $type = createForBoundType(type)$
+        \ENDIF
     \ENDIF
@@
 
+\begin{algorithm}
+\caption{Function createPathType}
+\label{ALG_function_create_path_type}
+\begin{algorithmic}[1]
+\REQUIRE $pathExprNode$: A reference to a PathExprNode to create the PathType from.
+\ENSURE The PathType of $pathExprNode$ expression.
+
+\STATE TODO
+\end{algorithmic}
+\end{algorithm}
+
 \subsection{Determination of Expression Types}
@@
 The starting node (its first argument) is the node representing the query body and the variable context is empty as there cannot be any local variable valid in the body node.
+
+\begin{algorithm}
+\caption{Function createForBoundType}
+\label{ALG_function_create_for_bound_type}
+\begin{algorithmic}[1]
+\REQUIRE $type$: A type to create the for bound type from.
+\ENSURE The for bound type from $type$.
+
+\STATE TODO
+\end{algorithmic}
+\end{algorithm}
 
@@
 These argument are set as the variable context for the function body represented by another subnode.
+
+\begin{algorithm}
+\caption{Function createForUnboundType}
+\label{ALG_function_create_for_unbound_type}
+\begin{algorithmic}[1]
+\REQUIRE $type$: A type to create the for unbound type from.
+\ENSURE The for unbound type from $type$.
+
+\STATE TODO
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-The first group of XQuery constructs utilized by this step is a group of $distinct-values()$ function and aggregation functions $min()$, $max()$, TODO \todo{Doplnit} as discussed in Section \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
+The first group of XQuery constructs utilized by this step is a group of $distinct-values$ function and aggregation functions $min$, $max$, $sum$ as discussed in Section \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
 The algorithm searches for calls of the mentioned functions with a $PathType$ on their input.
@@
 
-Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates.
-Predicates in a path restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates.
-\todo[inline]{Toto asi nebude uplne pravda pre nejake exoticke predikaty.
-Overit pri implementacii.
-A: Kdyžtak popište.}
+Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates in their lasts steps.
+Predicates in a path's last step restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates.
 
@@
 \IF{$is(node, FunctionCallNode)$}
-    \IF{$node.fncName$ is one of $distinct-values$, $min$, $max$, TODO}
+    \IF{$node.fncName$ is one of $distinct-values$, $min$, $max$, $sum$}
         \STATE $argument = node.getChild(0)$
-        \IF{$getType(argument) == PathType$ using only child and descendant axes}
+        \IF{$getType(argument) == PathType$ using only child and descendant axes, without predicates except the last step}
             \STATE memorize $argument$ as the negative key statement, weight 1
@@
             \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-            \IF{$getType(expr) == PathExpr$ without predicates, using only child and descendant axes}
+            \IF{$getType(expr) == PathExpr$ without predicates except the last step, using only child and descendant axes}
                 \STATE add $bindingNode.varName$ to $forVars$
@@
         \FORALL{$var \in forVars$}
-            \IF{$exprNode$ is of a form $var/L = C$ where $L$ is a path without predicates using only child and descendant axes, and getType(C) is a constant}
+            \IF{$exprNode$ is of a form $var/L = C$ where $L$ is a path without predicates, except the last step, using only child and descendant axes, and C is a literal constant}
                 \STATE memorize $var/L$ as the negative key statement, weight 0.9
@@
 \end{algorithm}
-
-\todo[inline]{Doplnit algoritmus comparison with a constant pre predikaty.}

### 2012-02-22 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-We want to capture the following types.
+The analysis of expression types can be divided into three substeps.
+Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
+
+But firstly, we describe types of expressions we want to capture and their features.
+
+\subsection{Expression Types}
 
@@
 
-\todo[inline]{Doplnit PathType o boolean for bound, volania zoznam funkcii volanych na tuto path.}
+\subsection{PathType}
+$PathType$ contains additional information.
+The represented path is contained by a list of its steps, in particular instances of $StepExprNode$.
+If a step is a reference to a variable which type is $PathType$, we also want to include this information.
+Therefore, $PathType$ contain association between the steps and other $PathTypes$ and this association is defined for the $PathType$ variable steps.
+
+To differ between a common $PathType$ selecting a set of nodes and a $PathType$ bound to a for variable in a FLWOR expression, $PathType$ structure contains a flag $isForBound$.
+
+Additionaly, $PathType$ contains a list of special functions that were called with an argument of $PathType$ type.
+The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of $PathType$ because then, we can determine a type of the function call more precisely.
+Those special functions are built-in functions $data$, $min$, $max$ \todo{Doplnit}.
+
+In summary, we represent $PathType$ as a structure with the following member variables. \todo{Radsej UML?}
+\begin{itemize}
+\item $stepNodes$ - A list of references to $StepExprNode$ instances.
+\item $forBoundSubsteps$ - An association between references from $stepNodes$ and instances of $PathType$ type.
+\item $isForBound$ - Boolean flag determining if the type was bound to a for variable in a for clause.
+\item $specialFunctionCalls$ - List of special functions called with this instance as an argument.
+\end{itemize}
+
+\subsection{Cardinality of Types}
 
@@
 Expressions of $UnknownType$ are not utilized it the inference, therefore, their cardinality is not needed.
-
-The analysis of expression types can be divided into three substeps.
-Determination of return types of functions, determination of types of global variables, and finally determination of types of expressions.
 
@@
 
-\todo[inline]{ALG Extraction of a Type from TypeNode nie uplne sedi s definiciou typov. DP JS ma nieco navyse.}
+\todo[inline]{ALG Extraction of a Type from TypeNode nie uplne sedi s definiciou typov. DP JS ma nieco navyse. A tento alg je nedokonceny!!}
 
@@
 Again, we do not analyse external variables for the same reason as in case of external functions.
-
-\todo[inline]{ALG \ref{ALG_func_determineExpressionType}: Vymysliet a dopisat ako ziskavat typ z PathExprNode. Bude potrebne vyskusat implementaciu, ako sa chova.}
 
@@
 
-\STATE TODO
+\STATE $stepNodes =$ an empty array
+\STATE $forBoundSubsteps =$ an empty associative array
+\FORALL{$stepNode \in pathExprNode.getChildren()$}
+	\STATE $detailNode = stepNode.getChild(DetailNode)$
+	\IF{$detailNode != null$}
+		\IF{$is(detailNode, VarRefNode)$}
+			\STATE $type = getType(detailNode)$ // $PathType$
+			\IF{$type.isForBound$}
+				\STATE $forBoundSubsteps[stepNode] = type$
+				\STATE add $stepNode$ to $stepNodes$
+			\ELSE
+				\STATE add all $type.stepNodes$ to $stepNodes$
+			\ENDIF
+		\ELSE
+			\STATE add $stepNode$ to $stepNodes$
+		\ENDIF
+	\ELSE
+		\STATE add $stepNode$ to $stepNodes$
+	\ENDIF
+\ENDFOR
+
+\RETURN $PathType$ with $stepNodes$, $forBoundSubsteps$, and $isForBound$ set to $false$
 \end{algorithmic}
@@
 
-\STATE TODO
+\IF{$type$ is $UnknownType$}
+	\RETURN $UnknownType$
+\ELSIF{$type$ is $NodeType$ or XSD atomic type}
+	\RETURN $type$ with cardinality set to exactly-one
+\ELSE
+	\STATE // $PathType$
+	\RETURN $type$ with $isForBound$ flag set to $true$
+\ENDIF
 \end{algorithmic}
@@
 
-\STATE TODO
+\IF{$type$ is $UnknownType$}
+	\RETURN $UnknownType$
+\ELSIF{$type$ is $NodeType$ or XSD atomic type}
+	\RETURN $type$ with cardinality set to zero-or-more
+\ELSE
+	\STATE // $PathType$
+	\RETURN $type$ with $isForBound$ flag set to $false$
+\ENDIF
 \end{algorithmic}
@@
             \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-            \IF{$getType(expr) == PathExpr$ without predicates except the last step, using only child and descendant axes}
+            \IF{$getType(expr) == PathType$ using only child and descendant axes}
                 \STATE add $bindingNode.varName$ to $forVars$

### 2012-02-22 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \FORALL{$stepNode \in pathExprNode.getChildren()$}
-	\STATE $detailNode = stepNode.getChild(DetailNode)$
+	\STATE $detailNode = stepNode.getChild(ExprNode)$
 	\IF{$detailNode != null$}
@@
         \FORALL{$(var, node) \in forVars$}
-            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var,$ $forVars)$
+            \STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var,$ $forVars,$ $checkJoinPattern3,$ $whereExpr)$
         \ENDFOR
@@
     $forVar$: A variable from the ancestor for binding node. \\
-    $forVars$: An array of couples of for variables and their for clause nodes.
+    $forVars$: An array of couples of for variables and their for clause nodes. \\
+    $checkJoinPattern3$: A flag determining whether to analyze the join pattern as possible join pattern 3. \\
+    $whereExpr$: An expression from a where clause of the current FLWOR node.
 }
@@
     \ELSIF{$exprNode.operator == AND$}
-        \STATE processWhere(forVars, exprNode.leftSide)
-        \STATE processWhere(forVars, exprNode.rightSide)
+        \STATE $processWhere(forVars, exprNode.leftSide)$
+        \STATE $processWhere(forVars, exprNode.rightSide)$
+    \ENDIF
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function usesOnlyChildAndDescendantAxes}
+\label{ALG_function_uses_only_child_and_descendant_axes}
+\begin{algorithmic}[1]
+\REQUIRE $pathType$: A $PathType$ instance.
+\ENSURE A boolean result.
+
+\todo[inline]{Dalsie pomocne funkcie a zaradit ich do nejakej kapitoly o pom. funkciach.}
+\FORALL{$stepNode \in pathType.stepNodes$}
+	\IF{$stepNode.isAxisStep()$}
+		\STATE $axisKind = stepNode.getAxisNode().getAxisKind()$
+		\IF{$axisKind$ $!=$ $CHILD$ $AND$ $axisKind$ $!=$ $DESCENDANT$}
+			\RETURN false
+		\ENDIF
+	\ENDIF
+\ENDFOR
+
+\STATE $detailNode = stepNode.getChild(ExprNode)$
+\IF{$detailNode != null$}
+	\IF{$is(detailNode, VarRefNode)$}
+		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) ==$ false}
+			\RETURN false
+		\ENDIF
 	\ENDIF
@@
 
+\RETURN true
 \end{algorithmic}

### 2012-02-22 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
+\subsection{Auxiliary Functions}
+\todo[inline]{Dalsie pomocne funkcie a okomentovat.}
+\todo[inline]{Algoritmu pouzivaju nejake nedefinovane funkcie, napr stepNode.getAxisNode().}
+
+\begin{algorithm}
+\caption{Function usesOnlyChildAndDescendantAxes}
+\label{ALG_function_uses_only_child_and_descendant_axes}
+\begin{algorithmic}[1]
+\REQUIRE $pathType$: A $PathType$ instance.
+\ENSURE A boolean result.
+
+\FORALL{$stepNode \in pathType.stepNodes$}
+	\IF{$stepNode.isAxisStep()$}
+		\STATE $axisKind = stepNode.getAxisNode().getAxisKind()$
+		\IF{$axisKind$ $!=$ $CHILD$ $AND$ $axisKind$ $!=$ $DESCENDANT$}
+			\RETURN false
+		\ENDIF
+	\ENDIF
+\ENDFOR
+
+\STATE $detailNode = stepNode.getChild(ExprNode)$
+\IF{$detailNode != null$}
+	\IF{$is(detailNode, VarRefNode)$}
+		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) ==$ false}
+			\RETURN false
+		\ENDIF
+	\ENDIF
+\ENDIF
+
+\RETURN true
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function isWithoutPredicates}
+\label{ALG_function_is_without_predicates}
+\begin{algorithmic}[1]
+\REQUIRE $pathType$: A $PathType$ instance.
+\ENSURE A boolean result.
+
+\FORALL{$stepNode \in pathType.stepNodes$}
+	\IF{$stepNode.hasPredicates()$}
+		\RETURN false
+	\ENDIF
+	\IF{$pathType.forBoundSubsteps[stepNode]$ $!=$ $null$}
+		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode]$ $==$ false}
+			\RETURN false
+		\ENDIF
+	\ENDIF
+\ENDFOR
+
+\RETURN true
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function endsWithExactlyOnePredicate}
+\label{ALG_function_ends_with_exactly_one_predicate}
+\begin{algorithmic}[1]
+\REQUIRE $pathType$: A $PathType$ instance.
+\ENSURE If $pathType$ ends with exactly one predicate, return value is the predicate, otherwise return value is false.
+
+\STATE $lastStepNode =$ last item from $pathType.stepNodes$
+\IF{$lastStepNode.hasPredicates()$ $==$ false}
+	\RETURN false
+\ENDIF
+\IF{number of item in list $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
+	\RETURN false
+\ENDIF
+
+\RETURN the only item in $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$
+\end{algorithmic}
+\end{algorithm}
+
 \subsection{Join Patterns}
@@
 \end{algorithm}
-
-\begin{algorithm}
-\caption{Function usesOnlyChildAndDescendantAxes}
-\label{ALG_function_uses_only_child_and_descendant_axes}
-\begin{algorithmic}[1]
-\REQUIRE $pathType$: A $PathType$ instance.
-\ENSURE A boolean result.
-
-\todo[inline]{Dalsie pomocne funkcie a zaradit ich do nejakej kapitoly o pom. funkciach.}
-\FORALL{$stepNode \in pathType.stepNodes$}
-	\IF{$stepNode.isAxisStep()$}
-		\STATE $axisKind = stepNode.getAxisNode().getAxisKind()$
-		\IF{$axisKind$ $!=$ $CHILD$ $AND$ $axisKind$ $!=$ $DESCENDANT$}
-			\RETURN false
-		\ENDIF
-	\ENDIF
-\ENDFOR
-
-\STATE $detailNode = stepNode.getChild(ExprNode)$
-\IF{$detailNode != null$}
-	\IF{$is(detailNode, VarRefNode)$}
-		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) ==$ false}
-			\RETURN false
-		\ENDIF
-	\ENDIF
-\ENDIF
-
-\RETURN true
-\end{algorithmic}
-\end{algorithm}

### 2012-02-23 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \subsection{Auxiliary Functions}
-\todo[inline]{Dalsie pomocne funkcie a okomentovat.}
-\todo[inline]{Algoritmu pouzivaju nejake nedefinovane funkcie, napr stepNode.getAxisNode().}
+Firstly, we define some auxiliary functions used by algorithms in this chapter.
+
+Function $usesOnlyChildAndDescendantAxes$ in Algorithm \ref{ALG_function_uses_only_child_and_descendant_axes} takes an instance of $PathType$ as its arguments and returns true if the path uses only child and descendant axes.
+Otherwise, it returns false.
 
@@
 
+Function $isWithoutPredicates$ in Algorithm \ref{ALG_function_is_without_predicates} checks whether a path represented by a given $PathType$ instance does not contain predicates.
+A Similar function is  $isWithoutPredicatesExceptLastStep$.
+This function is not explicitly defined, because its definition is the same as the definition of function $isWithoutPredicates$, except it does not check the last step of the path.
+So the last step may contain predicates and the function will return true.
+
 \begin{algorithm}
@@
 
+Function $endsWithExactlyOnePredicate$ checks if a path represented by a given $PathType$ instance contains exactly one predicate in its last step and if so, the predicate is returned.
+Otherwise, it returns false.
+
 \begin{algorithm}
@@
 \end{algorithm}
+
+\todo[inline]{Algoritmy pouzivaju nejake nedefinovane funkcie, napr stepNode.getAxisNode().}
 
@@
     \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-    \IF{$getType(BindingExpr) == PathType$ using only child and descendant axes, ending with exactly one predicate $P$}
+    \STATE $type = getType(bindingExpr)$
+    \IF{$type$ is $PathType$ $AND$ $usesOnlyChildAndDescendantAxes(type)$ $AND$ $isWithoutPredicatesExceptLastStep(type)$}
+    	\STATE $P = endsWithExactlyOnePredicate(type)$
+    	\IF{$P$}
         	\FORALL{$(var, node) \in forVars$}
@@
         	\ENDFOR
-    \ELSIF{$getType(BindingExpr) == PathType$ without predicates using only child and descendant axes}
+        \ELSIF{$isWithoutPredicates(type)$}
         	\IF{$is(bindingNode, ForClauseNode$}
             	\STATE add $(bindingNode.varName, bindingNode)$ to forVars
+        	\ENDIF
         \ENDIF
@@
 
+\todo[inline]{Definovat getParent() na Node.}
+\todo[inline]{Popisat preco raz bereme target return path types a raz priamo paths.}
+
 \FORALL{$jpOccurrence \in \mathcal{O}$}
@@
     \ELSE
-        \STATE TODO (R2,R3,R4,R5)
+        \STATE $secondBindingNode =$ the second binding node from $jpOccurrence$
+        \STATE $flworNode = secondBindingNode.getParent().getParent()$
+        \STATE $secondVarName = secondBindingNode.varName$
+        \STATE $returnPathTypes = getReturnPathTypes(flworNode, secondVarName)$
+        \FORALL{$returnPathType \in returnPathTypes$}
+        	\IF{$returnPathType.specialFunctionCalls$ contains one of $min$, $max$, $sum$, $avg$}
+        		\STATE mark $jpOccurrence$ as case (O1), weight 1 // R2
+        		\STATE continue the outmost for cycle with next $jpOccurrence$
+        	\ENDIF
+        \ENDFOR
+        \FORALL{$returnPathType \in returnPathTypes$}
+        	\IF{$returnPathType.specialFunctionCalls$ contains $count$}
+        		\STATE mark $jpOccurrence$ as case (O1), weight 0.75 // R3
+        		\STATE continue the outmost for cycle with next $jpOccurrence$
+        	\ENDIF
+        \ENDFOR
+        \STATE $targetReturnPaths = getTargetReturnPaths(flworNode, secontVarName)$
+        \STATE $retPathsNumber = $ number of paths in $targetReturnPaths$
+        \IF{$retPathsNumber > 1$}
+        	\STATE mark $jpOccurrence$ as case (O2), weight 1 // R4
+        \ELSE
+            \STATE mark $jpOccurrence$ as case (O2), weight 0.5 // R5
+        \ENDIF
     \ENDIF

### 2012-02-23 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/KeysInferrer.java
Master thesis implementation: In progress.

@@ -108,9 +108,8 @@ public class KeysInferrer {
   private static boolean usesOnlyChildAndDescendantAxes(final PathType pathType) {
     for (final StepExprNode stepNode : pathType.getStepNodes()) {
       if (stepNode.isAxisStep()) {
-        final AxisNode axisNode = stepNode.getAxisNode();
-        if (axisNode.getAxisKind() != AxisKind.CHILD
-                && axisNode.getAxisKind() != AxisKind.DESCENDANT_OR_SELF) {
+        final AxisKind axisKind = stepNode.getAxisNode().getAxisKind();
+        if (axisKind != AxisKind.CHILD && axisKind != AxisKind.DESCENDANT) {
           return false;
         }
       }
@@ -223,9 +222,6 @@ public class KeysInferrer {
         List<PathType> returnPathTypes = getTargetReturnPathTypes(flworNode, joinPattern.getSecondVariableBindingNode().getVarName());
         boolean cont = true;
         for (final PathType pathType : returnPathTypes) {
-          if (!cont) {
-            break;
-          }
           for (final String functionName : pathType.getSpecialFunctionCalls()) {
             if (functionName.equals("min")
                     || functionName.equals("max")
@@ -236,14 +232,17 @@ public class KeysInferrer {
               break;
             }
           }
-        }
-        
-        // R3
-        cont = true;
-        for (final PathType pathType : returnPathTypes) {
           if (!cont) {
             break;
           }
+        }
+        
+        if (!cont) {
+          continue;
+        }
+        
+        // R3
+        for (final PathType pathType : returnPathTypes) {
           for (final String functionName : pathType.getSpecialFunctionCalls()) {
             if (functionName.equals("count")) { // TODO rio pozor na nazvy funkcii, mozu byt aj prefixovane.
               classifiedJoinPatterns.add(new ClassifiedJoinPattern(joinPattern, ClassifiedJoinPattern.Type.O1, 75));
@@ -251,8 +250,15 @@ public class KeysInferrer {
               break;
             }
           }
+          if (!cont) {
+            break;
+          }
         }
         
+        if (!cont) {
+          continue;
+        }
+
         // R4 R5
         final List<PathExprNode> returnPaths = getTargetReturnPaths(flworNode, joinPattern.getSecondVariableBindingNode().getVarName());
         final int returnPathsNumber = returnPaths.size();

### 2012-02-23 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 	\IF{$stepNode.isAxisStep()$}
-		\STATE $axisKind = stepNode.getAxisNode().getAxisKind()$
-		\IF{$axisKind$ $!=$ $CHILD$ $AND$ $axisKind$ $!=$ $DESCENDANT$}
+		\STATE $axisKind$ = $stepNode.getAxisNode().getAxisKind()$
+		\IF{$axisKind$ $\neq$ $CHILD$ AND $axisKind$ $\neq$ $DESCENDANT$}
 			\RETURN false
@@
 
-\STATE $detailNode = stepNode.getChild(ExprNode)$
-\IF{$detailNode != null$}
+\STATE $detailNode$ = $stepNode.getChild(ExprNode)$
+\IF{$detailNode \neq$ null}
 	\IF{$is(detailNode, VarRefNode)$}
-		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) ==$ false}
+		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode])$ == false}
 			\RETURN false
@@
 \REQUIRE $pathType$: A $PathType$ instance.
-\ENSURE A boolean result.
+\ENSURE A boolean result.getTargetReturnPathTypesRecursive
 
@@
 	\ENDIF
-	\IF{$pathType.forBoundSubsteps[stepNode]$ $!=$ $null$}
-		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode]$ $==$ false}
+	\IF{$pathType.forBoundSubsteps[stepNode]$ $\neq$ null}
+		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode]$ == false}
 			\RETURN false
@@
 
-Function $endsWithExactlyOnePredicate$ checks if a path represented by a given $PathType$ instance contains exactly one predicate in its last step and if so, the predicate is returned.
+Function $endsWithExactlyOnePredicate$ in Algorithm \ref{ALG_function_ends_with_exactly_one_predicate} checks if a path represented by a given $PathType$ instance contains exactly one predicate in its last step and if so, the predicate is returned.
 Otherwise, it returns false.
@@
 
-\STATE $lastStepNode =$ last item from $pathType.stepNodes$
+\STATE $lastStepNode$ = last item from $pathType.stepNodes$
 \IF{$lastStepNode.hasPredicates()$ $==$ false}
@@
 \RETURN the only item in $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$
+\end{algorithmic}
+\end{algorithm}
+
+Function $isTargetPath$ in Algorithm \ref{ALG_function_is_target_path} checks if a path represented by a given $PathType$ instance is so-called target path for the specified variable.
+A target path for a variable is a path where its first step is a reference to the variable.
+
+\begin{algorithm}
+\caption{Function isTargetPath}
+\label{ALG_function_is_target_path}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $pathType$: A $PathType$ instance. \\
+  $varName$: A variable name.
+}
+\ENSURE Returns true, if the first step of path represented by the $PathType$ instance is a reference to the specified variable.
+
+\todo[inline]{Upravit styl vsetkych algoritmov (null, ==, !=, true/false, ...) tak ako je tu.}
+
+\STATE $firstStep$ = first item from  $pathType.stepNodes$
+\STATE $detailNode$ = $firstStep.detailNode$
+\IF{$detailNode$ == null}
+	\RETURN false
+\ENDIF
+\IF{$is(detailNode, VarRefNode)$ == false}
+	\RETURN false
+\ENDIF
+\IF{$detailNode.varName$ == $varName$}
+	\RETURN true
+\ELSE
+	\RETURN false
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+Function $getTargetReturnPathTypes$ in Algorithm \ref{ALG_function_get_target_return_path_types} searches the given FLWOR for $PathTypes$ representing so-called target return paths.
+A target return path is a return path in a where clause of a FLWOR.
+The FLWOR is processed recursively using function $getTargetReturnPathTypesRecursive$ in Algorithm \ref{ALG_function_get_target_return_path_types_recursive}.
+
+\begin{algorithm}
+\caption{Function getTargetReturnPathTypes}
+\label{ALG_function_get_target_return_path_types}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $flworNode$: A $FLWORExprNode$ instance to search for target return paths. \\
+  $varName$: A variable name.
+}
+\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by $PathType$ instances.
+
+\STATE $pathTypes$ = an empty array
+\STATE $getTargetReturnPathTypesRecursive(flworNode.$ $getChild(ReturnClauseNode), varName, pathTypes)$
+\RETURN $pathTypes$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function getTargetReturnPathTypesRecursive}
+\label{ALG_function_get_target_return_path_types_recursive}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $node$: A node to search for target return paths. \\
+  $varName$: A variable name. \\
+  $pathTypes$ An output list to add found target return path types to.
+}
+
+\IF{$is(node, ExprNode)$}
+	\STATE $type$ = $getType(node)$
+	\IF{$type$ is $PathType$ AND $isTargetPath(type, varName)$}
+		\STATE add $type$ to $pathTypes$
+	\ENDIF
+\ENDIF
+\FORALL{$child \in node.getChildren()$}
+	\STATE $getTargetReturnPathTypesRecursive(child, varName, pathTypes)$
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+A similar couple of functions is function $getTargetReturnPaths$ and function $getTargetReturnPathsRecursive$.
+Only difference between those two couples of functions is that the former one searches for all expressions which are of $PathType$ representing a target return path, while the later one searches only for instances of $PathExprNode$.
+
+To ilustrate this difference, we introduce the following example.
+Assuming $P$ is a return target path expression for variable $var$ and $flwor$ is a FLWOR expression with return clause $data(P)$, function call $getTargetReturnPathTypes(flwor, var)$ returns two $PathTypes$.
+One is the type of expression $P$ and the other is the type of the function call $data(P)$ which is also $PathType$.
+On the other hand, function call $getTargetReturnPaths(flwor, var)$ returns just one $PathType$ representing $P$.
+
+\begin{algorithm}
+\caption{Function getTargetReturnPaths}
+\label{ALG_function_get_target_return_paths}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $flworNode$: A $FLWORExprNode$ instance to search for target return paths. \\
+  $varName$: A variable name.
+}
+\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by $PartExprNode$ instances.
+
+\STATE $paths$ = an empty array
+\STATE $getTargetReturnPathsRecursive(flworNode.$ $getChild(ReturnClauseNode), varName, paths)$
+\RETURN $paths$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function getTargetReturnPathsRecursive}
+\label{ALG_function_get_target_return_paths_recursive}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $node$: A node to search for target return paths. \\
+  $varName$: A variable name. \\
+  $paths$ An output list to add found target return path types to.
+}
+
+\IF{$is(node, PathExprNode)$}
+	\STATE $type$ = $getType(node)$
+	\IF{$type$ is $PathType$ AND $isTargetPath(type, varName)$}
+		\STATE add $type$ to $paths$
+	\ENDIF
+\ENDIF
+\FORALL{$child \in node.getChildren()$}
+	\STATE $getTargetReturnPathsRecursive(child, varName, paths)$
+\ENDFOR
 \end{algorithmic}
@@
         \STATE $argument = node.getChild(0)$
-        \IF{$getType(argument) == PathType$ using only child and descendant axes, without predicates except the last step}
+        \STATE $type = getType(argument)$
+        \IF{$type$ is $PathType$ AND $usesOnlyChildAndDescendantAxes(type)$ AND $isWithoutPredicatesExceptLastStep(type)$}
             \STATE memorize $argument$ as the negative key statement, weight 1
@@
             \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-            \IF{$getType(expr) == PathType$ using only child and descendant axes}
+            \IF{$getType(expr)$ is $PathType$ AND $usesOnlyChildAndDescendantAxes(getType(expr))$}
                 \STATE add $bindingNode.varName$ to $forVars$
@@
         \FORALL{$var \in forVars$}
-            \IF{$exprNode$ is of a form $var/L = C$ where $L$ is a path without predicates, except the last step, using only child and descendant axes, and C is a literal constant}
+            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is $PathType$ AND $isWithoutPredicatesExceptLastStep(getType(var/L))$ AND $usesOnlyChildAndDescendantAxes(getType(var/L))$ AND $C$ is a literal constant}
                 \STATE memorize $var/L$ as the negative key statement, weight 0.9

### 2012-02-24 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \item $v.getChild(class), class \in C$, stands for $w \in V$ so that $(v,w) \in E$ and $c(w) = class$, if $\Bigl | \{u | u \in V, (v,u) \in E, c(u) = class\}\Bigr | = 1$.
+\item If exists $u \in V$ so that $(u,v) \in E$, $v.getParent()$ stands for that $u$.
 \end{itemize}
@@
 
-For instance, assuming $v \in V, c(v) = LiteralNode$, calls $is(v, LiteralNode)$, $is(v, ExprNode)$ return $true$, while call $is(v, ContentNode)$ returns $false$.
+For instance, assuming $v \in V, c(v) = LiteralNode$, calls $is(v, LiteralNode)$, $is(v, ExprNode)$ returns $true$, while call $is(v, ContentNode)$ returns $false$.
 
@@
 Nevertheless, the principle is straightforward.
-Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_4}), its type can be either determined directly or it depends on its subexpressions.
+Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_1}), its type can be either determined directly or it depends on its subexpressions.
 A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
@@
 
-\subsection{Auxiliary Functions}
+\subsection{Auxiliary Functions} \label{step_4_auxiliary_functions}
 Firstly, we define some auxiliary functions used by algorithms in this chapter.
@@
 \FORALL{$stepNode \in pathType.stepNodes$}
-	\IF{$stepNode.isAxisStep()$}
-		\STATE $axisKind$ = $stepNode.getAxisNode().getAxisKind()$
+	\IF{$stepNode.isAxisStep$}
+		\STATE $axisKind$ = $stepNode.axisNode.getAxisKind()$
 		\IF{$axisKind$ $\neq$ $CHILD$ AND $axisKind$ $\neq$ $DESCENDANT$}
@@
 \end{algorithm}
-
-\todo[inline]{Algoritmy pouzivaju nejake nedefinovane funkcie, napr stepNode.getAxisNode().}
 
@@
 
-As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules to classify each occurrence into one of these cases.
+As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules (R1 - R5) to classify each occurrence into one of these cases.
 For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
@@
 
-\todo[inline]{Alg 6.12 FLWOR expressions - pre podmienky na riadkoch 13 a 17 definovat nejake funkcie}
-
 \begin{algorithm}
@@
 
-\todo[inline]{Processing function pre join pattern 3.}
-
 To find the occurrences of the join patterns, the algorithm recursively, in pre-order, searches the syntax tree and every node representing a FLWOR expression is processed.
@@
 \subsection{Analysis of the Join Pattern Occurrences}
-For each found join pattern occurrence, the algorithm displayed in Algorithm \ref{ALG_classification_of_join_pattern_occurrences} decides whether it is (O1) or (O2) case, using the rules from the original method and the rule for the join pattern 3 considering it of case (O1) and assigning it with weight of 0.5.
+For each found join pattern occurrence, the algorithm displayed in Algorithm \ref{ALG_classification_of_join_pattern_occurrences} and Algorithms \ref{ALG_function_check_r2}, \ref{ALG_function_check_r3}, \ref{ALG_function_use_r4_r5} decides whether it is (O1) or (O2) case, using the rules from the original method and the rule for the join pattern 3 considering it of case (O1) and assigning it with weight of 0.5.
+
+As it can be seen in Algorithm \ref{ALG_classification_of_join_pattern_occurrences}, rules R2 and R3 are applied using instances of $PathType$ while, rules R4 and R5 use instances of $PathExprNode$.
+This difference is partially explained in definition of the auxiliary functions in Chapter \ref{step_4_auxiliary_functions}.
+Rules R4 and R5 count target return paths.
+If they use instances of $PathType$, they can count one path more times (as is described in the mentioned chapter), and thus, give a wrong result.
 
@@
 \ENSURE $\mathcal{O'}$: A set of the classified join pattern occurrences from $\mathcal{O}$ assigned with their respective weights.
-
-\todo[inline]{Definovat getParent() na Node.}
-\todo[inline]{Popisat preco raz bereme target return path types a raz priamo paths.}
 
@@
         \STATE $secondVarName = secondBindingNode.varName$
-        \STATE $returnPathTypes = getReturnPathTypes(flworNode, secondVarName)$
+        \STATE $returnPathTypes = getReturnPathTypes(flworNode,$ $secondVarName)$
+		\IF{$checkR2(returnPathTypes, jpOccurrence)$}
+			\STATE continue
+		\ENDIF
+		\IF{$checkR3(returnPathTypes, jpOccurrence)$}
+			\STATE continue
+		\ENDIF
+		\STATE $targetReturnPaths = getTargetReturnPaths(flworNode,$ $secontVarName)$
+		\STATE $useR4R5(returnPaths, jpOccurrence)$
+    \ENDIF
+    \STATE update $jpOccurrence$ in $\mathcal{O}$
+\ENDFOR
+
+\RETURN $\mathcal{O}$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function checkR2}
+\label{ALG_function_check_r2}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $returnPathTypes$: A list of target return path types. \\
+  $jpOccurrence$: An occurrence of a join pattern to classify.
+}
+
+\ENSURE True if the $jpOccurrence$ was classified as case (O1) using the R2 rule. False otherwise.
+
 \FORALL{$returnPathType \in returnPathTypes$}
@@
     	\STATE mark $jpOccurrence$ as case (O1), weight 1 // R2
-        		\STATE continue the outmost for cycle with next $jpOccurrence$
-        	\ENDIF
-        \ENDFOR
+  		\RETURN true
+  	\ENDIF
+\ENDFOR
+
+\RETURN false
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function checkR3}
+\label{ALG_function_check_r3}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $returnPathTypes$: A list of target return path types. \\
+  $jpOccurrence$: An occurrence of a join pattern to classify.
+}
+
+\ENSURE True if the $jpOccurrence$ was classified as case (O1) using the R3 rule. False otherwise.
+
 \FORALL{$returnPathType \in returnPathTypes$}
@@
         \STATE mark $jpOccurrence$ as case (O1), weight 0.75 // R3
-        		\STATE continue the outmost for cycle with next $jpOccurrence$
-        	\ENDIF
-        \ENDFOR
-        \STATE $targetReturnPaths = getTargetReturnPaths(flworNode, secontVarName)$
+        \RETURN true
+    \ENDIF
+\ENDFOR
+
+\RETURN false
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function useR4R5}
+\label{ALG_function_use_r4_r5}
+\begin{algorithmic}[1]
+\REQUIRE{\ \\
+  $targetReturnPaths$: A list of target return paths. \\
+  $jpOccurrence$: An occurrence of a join pattern to classify.
+}
+
 \STATE $retPathsNumber = $ number of paths in $targetReturnPaths$
@@
 \ENDIF
-    \ENDIF
-    \STATE update $jpOccurrence$ in $\mathcal{O}$
-\ENDFOR
-
-\RETURN $\mathcal{O}$
 \end{algorithmic}

### 2012-02-26 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/BuiltinFunctions.java
Master thesis implementation: In progress.

@@ -21,6 +21,7 @@ import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.NodeType;
 import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.PathType;
 import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.Type;
 import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.UnknownType;
+import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.XSDType;
 import java.util.ArrayList;
 import java.util.HashMap;
 import java.util.List;
@@ -37,6 +38,8 @@ public class BuiltinFunctions {
     builtinFunctionTypes.put("data", new UnknownType());
     builtinFunctionTypes.put("doc", new NodeType(NodeType.NType.DOCUMENT, Cardinality.ONE));
     builtinFunctionTypes.put("min", new UnknownType());
+    builtinFunctionTypes.put("distinct-values", new UnknownType());
+    builtinFunctionTypes.put("count", new XSDType(XSDType.XSDAtomicType.INTEGER, Cardinality.ONE));
   }
   
   private final static List<String> DEFAULT_BUILTIN_FUNCTIONS_NAMESPACES = new ArrayList<String>();
@@ -72,23 +75,23 @@ public class BuiltinFunctions {
   }
   
   public static Type getFunctionCallType(final FunctionCallNode functionCallNode) {
-    String functionName = functionCallNode.getFuncName();
-    functionName = functionName.substring(functionName.lastIndexOf(':') + 1);
+    final String functionName = functionCallNode.getFuncName();
+    final String builtinFuncName = BuiltinFunctions.isBuiltinFunction(functionName);
     
-    if (functionName.equals("data")) {
-      final Type type = functionCallNode.getParams().get(0).getType();
-      if (PathType.class.isInstance(type)) {
-        ((PathType)type).addSpecialFunctionCall("data");
+    assert (builtinFuncName != null);
+    
+    for (final String specFuncName : PathType.SPECIAL_FUNCTION_NAMES) {
+      if (builtinFuncName.equals(specFuncName)) {
+        final Type type = functionCallNode.getParams().get(0).getType();
+        if (PathType.class.isInstance(type)) {
+          ((PathType)type).addSpecialFunctionCall(builtinFuncName);
+        }
+        return type;
       }
-      return type;
-    } else if (functionName.equals("min")) {
[…]

### 2012-02-26 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/FunctionsProcessor.java

@@ -23,7 +23,7 @@ import java.util.HashMap;
 import java.util.Map;
 
 /**
- * TODO rio
+ * TODO rio comment
  * Analyza globalnych funkcii pre jeden syntakticky strom.
  * @author rio
  */
@@ -86,7 +86,7 @@ public class FunctionsProcessor {
     return fTypes;
   }
   
-  // TODO rio Dopisat do DP, ze pocitame s tym, ze builtin funkcie su bud bez rpefixu alebo s prefixom fn:
+  // TODO rio Dopisat do DP, ze pocitame s tym, ze builtin funkcie su bud bez prefixu alebo s prefixom fn:
   public ParamListNode getParamListNode(final String functionName) {
     final String builtinFunctionName = BuiltinFunctions.isBuiltinFunction(functionName);
             

### 2012-02-26 jinfer/branches/thesis-mikula/Base/src/cz/cuni/mff/ksi/jinfer/base/xqueryanalyzer/types/PathType.java
Master thesis implementation: In progress.

@@ -27,6 +27,16 @@ import java.util.Map;
  */
 public class PathType implements Type {
   
+  public final static List<String> SPECIAL_FUNCTION_NAMES = new ArrayList<String>();
+  static {
+    SPECIAL_FUNCTION_NAMES.add("data");
+    SPECIAL_FUNCTION_NAMES.add("distinc-values");
+    SPECIAL_FUNCTION_NAMES.add("min");
+    SPECIAL_FUNCTION_NAMES.add("max");
+    SPECIAL_FUNCTION_NAMES.add("avg");
+    SPECIAL_FUNCTION_NAMES.add("sum");
+  }
+  
   private final List<StepExprNode> stepNodes;
   private final Map<StepExprNode, PathType> forBoundSubsteps; // StepExprNode -> ForBoundPathType
   private final List<String> specialFunctionCalls = new ArrayList<String>(); // An ordered list of functions that have been applied to this path.

### 2012-02-26 jinfer/branches/thesis-mikula/BasicIGG/src/cz/cuni/mff/ksi/jinfer/basicigg/xquery/XQueryProcessor.java

@@ -74,15 +74,6 @@ public class XQueryProcessor implements Processor<ModuleNode> {
       converter.convert();
     } catch (final XQParseException e) {
       LOG.error("Error parsing XQuery file, ignoring and going on.", e);
-      
-      ByteArrayOutputStream baos = new ByteArrayOutputStream();
-      PrintStream ps = new PrintStream(baos);
-      e.printStackTrace(ps);
-      String content = baos.toString();
-      LOG.error(content);
-      
-      
-      
       return Collections.EMPTY_LIST;
     }
     final List<ModuleNode> result = new ArrayList<ModuleNode>();

### 2012-02-26 jinfer/branches/thesis-mikula/Runner/src/cz/cuni/mff/ksi/jinfer/runner/Runner.java

@@ -98,7 +98,7 @@ public class Runner {
     schemaGenerator = ModuleSelectionHelper.lookupImpl(SchemaGenerator.class,
             projectProperties.getProperty(ModuleSelectionPropertiesPanel.SCHEMAGEN_PROP, ModuleSelectionPropertiesPanel.SCHEMAGEN_DEFAULT));
     // TODO rio Selection in properties panel is not implemented yet.
-    xqueryAnalyzer = ModuleSelectionHelper.lookupImpl(XQueryAnalyzer.class, "XQueryAnalyzer");
+    xqueryAnalyzer = ModuleSelectionHelper.lookupImpl(XQueryAnalyzer.class, "XQuery_Analyzer");
   }
 
   /**

### 2012-02-26 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/BuiltinTypesInferrer.java

@@ -17,6 +17,7 @@
 package cz.cuni.mff.ksi.jinfer.xqueryanalyzer;
 
 import cz.cuni.mff.ksi.jinfer.base.objects.nodes.xqanalyser.*;
+import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.PathType;
 import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.Type;
 import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.TypeFactory;
 import java.util.HashMap;
@@ -31,7 +32,7 @@ public class BuiltinTypesInferrer {
   
   private final ModuleNode root;
   private final FunctionsProcessor fp;
-  private final Map<Type, Type> inferredTypes = new HashMap<Type, Type>(); // TODO rio Type->Type je blbost. Treba nejak vymyslet path->Type.
+  private final Map<PathType, Type> inferredTypes = new HashMap<PathType, Type>();
   
   public BuiltinTypesInferrer(final ModuleNode root, final FunctionsProcessor fp) {
     this.root = root;
@@ -76,7 +77,7 @@ public class BuiltinTypesInferrer {
         if (paramTypeDeclarationNode == null) {
           continue;
         }
-        inferredTypes.put(paramType, TypeFactory.createType(paramTypeDeclarationNode));
+        inferredTypes.put((PathType)paramType, TypeFactory.createType(paramTypeDeclarationNode));
       }
     }
   }
@@ -84,7 +85,7 @@ public class BuiltinTypesInferrer {
   public void processOperator(final OperatorNode opNode) {
     final Operator op = opNode.getOperator();
     
-    Type pathType = null;
+    PathType pathType = null;
     Type type = null;
     
     switch (op) {
@@ -103,12 +104,12 @@ public class BuiltinTypesInferrer {
         final Type rightType = rightOperand.getType();
         if (leftType.getCategory() == Type.Category.PATH) {
           if (rightType.isNumeric() && rightType.getCategory() == Type.Category.BUILT_IN) {
-            pathType = leftType;
+            pathType = (PathType)leftType;
             type = rightType;
           }
         } else if (rightType.getCategory() == Type.Category.PATH) {
           if (leftType.isNumeric() && leftType.getCategory() == Type.Category.BUILT_IN) {
-            pathType = rightType;
+            pathType = (PathType)rightType;
             type = leftType;
           }
         }
@@ -127,12 +128,12 @@ public class BuiltinTypesInferrer {
[…]

### 2012-02-26 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/ExpressionsProcessor.java

@@ -34,7 +34,6 @@ import java.util.Map;
  * vyrazy v tele dotazu.
  * 
  * TODO rio Treba dorobit analyzu volani funkcii s tym, ze sa bude zistovat typ tak, ze sa vyuzije tele funkcie.
- * TODO rio Skompletovat rozne pomocne metody.
  * TODO rio Porozmyslat, ci nie je porebne zlucit zistovanie typov user-defined funkcii a globalnych premennych, kedze glob. premenna moze pouzivat funkcie a opacne.
  * 
  * @author rio
@@ -150,6 +149,10 @@ public class ExpressionsProcessor {
       return TypeFactory.createForUnboundType(((FLWORExprNode) expressionNode).getReturnClauseNode().getExprNode().getType());
     } else if (PathExprNode.class.isInstance(expressionNode)) {
       return TypeFactory.createPathType((PathExprNode)expressionNode);
+    } else if (ConstructorNode.class.isInstance(expressionNode)) {
+      //final ConstructorNode constructorNode = (ConstructorNode)expressionNode;
+      // TODO rio Aky typ ma constructor?
+      return new UnknownType();
     }
     assert (false); // TODO rio dorobit pre ostatne typy
     return null;
@@ -168,21 +171,76 @@ public class ExpressionsProcessor {
 
     if (isOperatorClassAddition(operator)) {
       final Type leftType = operatorNode.getOperand().getType();
+      
+      // Moze byt aj unarny operator, v takom pripade je typ jasny.
+      if (operatorNode.getRightSide() == null) {
+        return leftType;
+      }
+      
       final Type rightType = operatorNode.getRightSide().getType();
+      
+      // TODO rio Co so sekvenciami? Plati aj pre ne?
       if (leftType.isNumeric() && rightType.isNumeric()) {
-        return leftType; // TODO rio vybrat obecnejsi z typov
+        return selectCommonType(leftType, rightType); 
       } else if (leftType.isNumeric()) {
         return leftType;
       } else if (rightType.isNumeric()) {
         return rightType;
-      } // TODO rio else?
+      }
     }
 
     if (operator == Operator.TO) {
       return new XSDType(XSDType.XSDAtomicType.INTEGER, Cardinality.ZERO_OR_MORE);
     }
 
-    return null;
+    return new UnknownType();
+  }
+  
[…]

### 2012-02-26 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/KeysInferrer.java

@@ -377,9 +377,12 @@ public class KeysInferrer {
     if (FunctionCallNode.class.isInstance(node)) {
       final FunctionCallNode funcCallNode = (FunctionCallNode)node;
       final String funcName = funcCallNode.getFuncName();
-      if (funcName.equals("distinct-values")
-              || funcName.equals("min")
-              || funcName.equals("max")) { // TODO dalsie funkcie, aj do DP, riesit aj prefixy
+      final String builtinFuncName = BuiltinFunctions.isBuiltinFunction(funcName);
+      if (builtinFuncName != null
+              && builtinFuncName.equals("distinct-values")
+              || builtinFuncName.equals("min")
+              || builtinFuncName.equals("max")
+              || builtinFuncName.equals("sum")) {
         final ExprNode argument = funcCallNode.getParams().get(0);
         assert(argument != null);
         final Type type = argument.getType();

### 2012-02-26 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/XQueryAnalyzerImpl.java

@@ -16,7 +16,6 @@
  */
 package cz.cuni.mff.ksi.jinfer.xqueryanalyzer;
 
-import cz.cuni.mff.ksi.jinfer.base.xqueryanalyzer.types.Type;
 import cz.cuni.mff.ksi.jinfer.base.interfaces.inference.XQueryAnalyzer;
 import cz.cuni.mff.ksi.jinfer.base.interfaces.inference.XQueryAnalyzerCallback;
 import cz.cuni.mff.ksi.jinfer.base.objects.InferenceDataHolder;
@@ -26,7 +25,7 @@ import org.apache.log4j.Logger;
 import org.openide.util.lookup.ServiceProvider;
 
 /**
- * TODO rio
+ * TODO rio comment
  *
  * Hlavna trieda XQuery modulu. Postupne pusta jednotlive kroky algoritmu
  * na zistovanie informacii z dotazov.
@@ -37,6 +36,9 @@ import org.openide.util.lookup.ServiceProvider;
 public class XQueryAnalyzerImpl implements XQueryAnalyzer {
   
   private final static Logger LOG = Logger.getLogger(XQueryAnalyzerImpl.class);
+  
+  private static final String NAME = "XQuery_Analyzer";
+  private static final String DISPLAY_NAME = "XQuery analyzer";
 
   @Override
   public void start(InferenceDataHolder idh, XQueryAnalyzerCallback callback) throws InterruptedException {
@@ -62,26 +64,22 @@ public class XQueryAnalyzerImpl implements XQueryAnalyzer {
   
   @Override
   public String getDisplayName() {
-    // TODO rio
-    return "XQueryAnalyzer";
+    return DISPLAY_NAME;
   }
 
   @Override
   public String getModuleDescription() {
-    // TODO rio
-    return "XQueryAnalyzer";
+    return getDisplayName();
   }
 
   @Override
   public String getName() {
-    // TODO rio
-    return "XQueryAnalyzer";
+    return NAME;
   }
 
   @Override
   public List<String> getCapabilities() {
-    // TODO rio
-    return new LinkedList<String>();
+    return Collections.emptyList();
   }
   
   private void init(final XQNode node) {

### 2012-02-26 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
+\todo[inline]{Niektore typy mozu byt take, ze pri prvom prechode ich nevieme zistit, ale ked uz mame zistene ine typy, tak dokazeme zistit aj tie, co sme pred tym nevedeli.
+Takze by bolo treba robit viac priechodov, ale skor nieco ako dopologicke triedenie, co zavisi na com, a potom typy zistovat v tomto poradi (mozno aj v druhom kole).}
+
 The analysis of expression types can be divided into three substeps.
@@
 
-\STATE $type = null$
+\STATE $type = UnknownType$
 \IF{$is(exprNode, LiteralNode)$}
@@
 
-\STATE $type = null$
+\STATE $type = UnknownType$
 \STATE $operator = operatorNode.operator$
@@
 
-\IF{$operator$ equals one of $PLUS$, $MINUS$}
+\IF{$operator$ equals one of $PLUS$, $MINUS$, $UNARY\_PLUS$, $UNARY\_MINUS$}
     \RETURN $true$

### 2012-03-03 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/FunctionsProcessor.java
Master thesis implementation: In progress.

@@ -86,7 +86,6 @@ public class FunctionsProcessor {
     return fTypes;
   }
   
-  // TODO rio Dopisat do DP, ze pocitame s tym, ze builtin funkcie su bud bez prefixu alebo s prefixom fn:
   public ParamListNode getParamListNode(final String functionName) {
     final String builtinFunctionName = BuiltinFunctions.isBuiltinFunction(functionName);
             

### 2012-03-03 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/JoinPattern.java

@@ -35,8 +35,6 @@ public class JoinPattern {
   private final JoinPatternType type;
   private final VariableBindingNode firstVariableBindingNode;
   private final VariableBindingNode secondVariableBindingNode;
-  private final PathType pathType1;
-  private final PathType pathType2;
   private final PathExprNode P1;
   private final PathExprNode P2;
   private final PathExprNode L1;
@@ -47,8 +45,6 @@ public class JoinPattern {
     this.type = type;
     this.firstVariableBindingNode = firstVariableBindingNode;
     this.secondVariableBindingNode = secondVariableBindingNode;
-    pathType1 = (PathType)firstVariableBindingNode.getBindingSequenceNode().getExprNode().getType();
-    pathType2 = (PathType)secondVariableBindingNode.getBindingSequenceNode().getExprNode().getType();
     this.P1 = P1;
     this.P2 = P2;
     this.L1 = L1;

### 2012-03-03 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/ExpressionsProcessor.java
Master thesis implementation: In progress.

@@ -36,7 +36,7 @@ import java.util.Map;
  * vyrazy v tele dotazu.
  * 
  * TODO rio Treba dorobit analyzu volani funkcii s tym, ze sa bude zistovat typ tak, ze sa vyuzije tele funkcie.
- * TODO rio Porozmyslat, ci nie je porebne zlucit zistovanie typov user-defined funkcii a globalnych premennych, kedze glob. premenna moze pouzivat funkcie a opacne.
+ * TODO rio Porozmyslat, ci nie je potrebne zlucit zistovanie typov user-defined funkcii a globalnych premennych, kedze glob. premenna moze pouzivat funkcie a opacne.
  * 
  * @author rio
  */
@@ -146,7 +146,7 @@ public class ExpressionsProcessor {
     } else if (OperatorNode.class.isInstance(expressionNode)) {
       return determineOperatorType((OperatorNode) expressionNode);
     } else if (VarRefNode.class.isInstance(expressionNode)) {
-      // TODO rio Toto dopisat do DP. Je to kvoli tomu, aby vo FLWORoch bola premenna obsahujuca cestu sama cesta zacinajuce touto premennou, aby bolo mozne ju povazovat za target return path.
+      // Je to kvoli tomu, aby vo FLWORoch bola premenna obsahujuca cestu sama cesta zacinajuce touto premennou, aby bolo mozne ju povazovat za target return path.
       final Type varType = contextVarTypes.get(((VarRefNode) expressionNode).getVarName());
       if (varType.getCategory() == Type.Category.PATH) {
         final List<StepExprNode> steps = new ArrayList<StepExprNode>();

### 2012-03-03 jinfer/branches/thesis-mikula/XQueryAnalyzer/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/KeysInferrer.java

@@ -214,7 +214,6 @@ public class KeysInferrer {
 
     rejectionOfUniqueness_aggregationFunctions(node);
     rejectionOfUniqueness_comparisonWithAConstant(node);
-    // TODO rio zlucovanie vah
   }
 
   public Map<String, ForClauseNode> processFLWORExpr(final FLWORExprNode flworExpr, Map<String, ForClauseNode> forVariables) {
@@ -349,7 +348,7 @@ public class KeysInferrer {
             if (functionName.equals("min")
                     || functionName.equals("max")
                     || functionName.equals("avg")
-                    || functionName.equals("sum")) { // TODO rio pozor na nazvy funkcii, mozu byt aj prefixovane.
+                    || functionName.equals("sum")) {
               classifiedJoinPatterns.add(new ClassifiedJoinPattern(joinPattern, ClassifiedJoinPattern.Type.O1, 100));
               cont = false;
               break;
@@ -450,16 +449,14 @@ public class KeysInferrer {
     } else if (VarRefNode.class.isInstance(node)) {
       final VarRefNode varRefNode = (VarRefNode)node;
       if (varRefNode.getVarName().equals(varName)) {
-        // TODO rio Toto dopisat do DP. Je to kvoli tomu, aby vo FLWORoch bola premenna obsahujuca cestu sama cesta zacinajuce touto premennou, aby bolo mozne ju povazovat za target return path.
+        // Toto je to kvoli tomu, aby vo FLWORoch bola premenna obsahujuca cestu sama cesta zacinajuce touto premennou, aby bolo mozne ju povazovat za target return path.
         assert(varRefNode.getType().getCategory() == Type.Category.PATH);
-        final List<StepExprNode> steps = new ArrayList<StepExprNode>();
-        final StepExprNode step = new StepExprNode(varRefNode, null);
-        steps.add(step);
-        final Map<StepExprNode, PathType> substeps = new HashMap<StepExprNode, PathType>();
-        substeps.put(step, (PathType)varRefNode.getType());
-        final PathExprNode path = new PathExprNode(steps, InitialStep.CONTEXT);
-        path.setParentNode(varRefNode.getParentNode());
-        paths.add(path);
+        final PathType pathType = (PathType)varRefNode.getType();
+        if (isTargetPath(pathType, varName)) {
+          final PathExprNode path = pathType.getPathExprNode();
+          path.setParentNode(varRefNode.getParentNode());
+          paths.add(path);
+        }
       }
     }
 
[…]

### 2012-03-03 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 $PathType$ contains additional information.
-The represented path is contained by a list of its steps, in particular instances of $StepExprNode$.
+The represented path is contained by an instance of $PathExprNode$ containing list of its steps, in particular instances of $StepExprNode$.
 If a step is a reference to a variable which type is $PathType$, we also want to include this information.
@@
 \begin{itemize}
-\item $stepNodes$ - A list of references to $StepExprNode$ instances.
-\item $forBoundSubsteps$ - An association between references from $stepNodes$ and instances of $PathType$ type.
+\item $pathExprNode$ - An instance of $PathExprNode$.
+\item $substeps$ - An association between references from $stepNodes$ and instances of $PathType$ type.
 \item $isForBound$ - Boolean flag determining if the type was bound to a for variable in a for clause.
@@
 Return types of built-in functions are fixed and, thus, there is no need to analyze them.
+Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by $fn$ or not prefixed.
 
@@
      \STATE $type = getVariableType(contextVarTypes, exprNode.varName)$
+     \IF{$type$ is $PathType$}
+         \STATE $step$ = a path step (new instance of $StepExprNode$) representing the variable reference
+		 \STATE $substeps$ = empty associative array         
+         \STATE $substeps[step]$ = $type$
+         \STATE $pathExprNode$ = new instance of $PathExprNode$ containing one step $step$
+         \STATE $type$ = $PathType$ containing $pathExprNode$ and $substeps$
+     \ENDIF
 \ELSIF{$is(exprNode, PathExprNode)$}
@@
 Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_1}), its type can be either determined directly or it depends on its subexpressions.
+
+The meaning of the code at lines 6-14 is the following.
+If a variable is bound to an expression which type is $PathType$, we want to assign the variable reference expression with $PathType$ which represents path containing exactly one step, the variable reference.
+Some algorithm in later sections search for paths starting with a certain variable and we want them to include also that types of expressions.
+
 A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
@@
 	\ENDIF
+\ELSIF{$is(node, VarRefNode)$}
+	\IF{$node.varName$ == $varName$}
+		\STATE $type$ = $getType(node)$
+		\IF {$type$ is $PathType$ AND $isTargetPath(type, varName)$}
+			\STATE add $type$ to $paths$
+		\ENDIF
+	\ENDIF
 \ENDIF
@@
 
-The first group of XQuery constructs utilized by this step is a group of $distinct-values$ function and aggregation functions $min$, $max$, $sum$ as discussed in Section \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
+The first XQuery construct utilized by this step is a function call of $distinct-values$ function as discussed in Section \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
 The algorithm searches for calls of the mentioned functions with a $PathType$ on their input.
@@
 
+Aggregation functions $min$, $max$, $sum$ are not utilized by the function because they were utilized by the search for join patterns, and in the process of classification of found join patterns, they were included in the computation of the key weights.
+
 Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates in their lasts steps.
@@
 \IF{$is(node, FunctionCallNode)$}
-    \IF{$node.fncName$ is one of $distinct-values$, $min$, $max$, $sum$}
+    \IF{$node.fncName$ == $distinct-values$}
         \STATE $argument = node.getChild(0)$

### 2012-03-06 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

\chapter{Combination with Existing Methods of Inference}
In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference. We focus on a class of methods which are based on a creation and subsequent simlification of the initial grammar (as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}).

The initial grammar contains rules of form $e \rightarrow e_1e_2\dots e_k$, where $e$ is an element and $e_1, \dots , e_k$ are its subelements. After the simplification, the simplified grammar contain rules of form $e \rightarrow E$, where $E$ is a regular expression composed of subelements of element $e$, describing its content.

Attributes of elements are not contained in the grammar directly, but every element carries an information on its attributes.

The combination with such methods of inference is straightforward. The rules of the grammar describe a XML structure using the most general aspect of elements and their subelements. Since the statements inferred by our method do not involve the XML structure by defining a subelement structure of elements, there are no conflicts between the grammar rules and our statements that need to be resolved.

Our statements are of the three following forms. The first one is $p \rightarrow t$, where $p$ is a $PathType$ representing an XQuery path selecting a set of elements or attributes \todo{plati nejake obmedzenie na tuto cestu?}
, and $t$ is an XSD built-in atomic type.

The second and third forms are $k = (C,P,\{L\})$ and $fk = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key, where $C,P,P_1,P_2,L,L_1,L_2$ are $PathTypes$ without predicates using only child and descendant axes.

In all tree cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form). This is done in two steps. The first one is normalization of a particular $PathType$ and the second one is selection of the targeted elements.

A path represented by $PathType$ can contain variable references and association of these references with other paths. That is the reason why the normalization is convenient. It simply finds the variable references in the path and replaces them with steps from the associated paths, which are normalized recursively.

The selection of elements is a simplified XPath evaluation. The simplification involves the aspects; the evaluation is not performed upon XML data, but the simplified grammar containing rules with a regular expression on their right side, and, partially related, predicates in path steps are not evaluated, they are ignored.

It iterate through steps of a path, maintaining a so-called context set, which is a set of elements to evaluate the current step upon. The evaluation of one step is shown in Algorithm \ref{FUNC_evaluate_step}. For the retaining of readability of the code, we present an evaluation of self or descendant, child, and attribute axes, and nodes specified by name.

At first, the algorithm determine the axis of the step. If it is self or descendant axis, it returns the result of $evaluateStep\_selfOrDescendant$ function. This function returns the given context set extended by all descendants of the elements from the context set in the given grammar.

If the axis is child axis, for each element in the context set, the algorithm determines its subelement using the grammar, and, if those with the name equal to the name specified by to step are added to the resulting context set. Function $getTokens$ at line 10 retrieves all elements from a regular expression. A specific form of the regular expression is not important, because we only need to know which elements are possible subelements of the particular element.

And, at last, if the axis is attribute axis, elements from the context set are searched to contain an attribute with the specified name and those found attributes are added to the result.

\begin{algorithm}
\caption{Function evaluateStep}
\label{FUNC_evaluate_step}
\begin{algorithmic}[1]
\REQUIRE{\ \\
	$step$: An instance of $StepExprNode$ representing a step of the path to evaluate.
	$contextSet$: A set of elements and attributes to evaluate $step$ upon.
	$grammar$: The grammar.
}

\ENSURE A context set after the step evaluation.

\IF{$is(step, SelfOrDescendantStep)$}
	\RETURN $evaluateStep\_selfOrDescendant(contextSet, grammar)$
\ENDIF

\STATE $newContextSet$ = empty set

\STATE $axisKind$ = $step.getChild(axisNode).axisKind$
\STATE $nodeName$ = $step.getChild(axisNode).getChild(nameTestNode).name$
\IF{$axisKind$ == $CHILD$}
	\FORALL{$node \in contextSet$}
		\IF{$node$ is element}
			\STATE $subelements$ = $getTokens(grammar[node])$
			\FORALL{$subelement \in subelements$}
				\IF{$subelement.name$ == $nodeName$}
					\STATE add $subelement$ to $newContextSet$
				\ENDIF
			\ENDFOR
		\ENDIF
	\ENDFOR
\ELSIF{$axisKind$ == $ATTRIBUTE$}
	\FORALL{$node \in contextSet$}
		\IF{$node$ is element}
			\STATE $attributes$ = $node.attributes$
			\FORALL{$attribute \in attributes$}
				\IF{$attribute.name$ == $nodeName$}
					\STATE add $attribute$ to $newContextSet$
				\ENDIF
			\ENDFOR
		\ENDIF
	\ENDFOR
\ENDIF

\RETURN $newContextSet$
\end{algorithmic}
\end{algorithm}

### 2012-03-06 doc/theses/mikula/diplomka.tex
Master thesis: In progress.

@@
 %% Pokud nikoliv:
-\usepackage[czech]{babel}
-\usepackage[T1]{fontenc}
+%\usepackage[czech]{babel}
+%\usepackage[T1]{fontenc}
 
@@
 \usepackage{amssymb}
-%\usepackage{alltt}
+\usepackage{alltt}
 \usepackage{todonotes}
+\usepackage{algorithmic}
+\usepackage[chapter]{algorithm}
+\usepackage{enumerate}
+\usepackage{multirow}
+\usepackage{threeparttable}
+\usepackage{listings}
 \usepackage{algorithmic}
@@
 
+% Theorem environment
+\newenvironment{definition}[1][Definition]{\begin{trivlist}
+\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+
 \begin{document}
@@
 % (dle Organizační struktury MFF UK)
-Katedra softwarového inženýrství
+Department of Software Engineering
 
@@
 \tableofcontents
-\newpage
-\renewcommand\thepage{\arabic{page}}
-\setcounter{page}{1}
 
@@
 \theoremstyle{definition}
-\newtheorem*{define}{Definition}	% Definice nečíslujeme, proto "*"
+%\newtheorem*{define}{Definition}	% Definice nečíslujeme, proto "*"
+\newtheorem{define}{Definition}[chapter]
 \newtheorem{example}{Example}
@@
 
+\newpage
+\renewcommand\thepage{\arabic{page}}
+\setcounter{page}{1}
+
 %%% Jednotlivé kapitoly práce jsou pro přehlednost uloľeny v samostatných souborech
-\chapter{Úvod}
-\todo[inline]{TODO}
-\chapter{Použité technologie/pojmy}
-\todo[inline]{TODO}
-Krátce popsat co je XML Schema, DTD, XPath, XQuery, možná ňejaký automat apod. Prostě co budete dále v textu potřebovat. Skratky: XSD, ...
-
+\include{introduction}
+\include{used_technologies_and_definitions}
 \include{analysis_of_recent_approaches}
 \include{analysis_of_xquery}
-%\include{kap2}
-%\include{kap3}
-
-% Ukázka pouľití některých konstrukcí LateXu (odkomentujte, chcete-li)
-%\include{ukazka}
-
-%\include{zaver}
+\include{pre_creation_of_algorithm}
+\include{proposed_algorithm}
+\include{combination_with_existing_methods_of_inference}
 
@@
 \todo[inline]{Literatura tady určitě co nejpodrobněji - odkaz nestačí, za měsíc nemusí existovat. A nemělo by to přetékat okraj.}
+\todo[inline]{Literatura - jInfer modules documentation - Tady by neškodily odkazy na jednotlivé dokumenty.}
 \bibliography{literature}
@@
 %%% Pouľité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-\chapwithtoc{List of Abbreviations}
-\todo[inline]{Není nutno.}
+%\chapwithtoc{List of Abbreviations}
+%\todo[inline]{Není nutno.}
 

### 2012-03-06 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
+\subsection{Inferrence of Keys from Join Pattern Occurrences}
+Now, when the join patterns occurrences are classified into (O1) and (O2) cases, we infer the key statements according to the original method \cite{Necasky:2009:DXK:1529282.1529414}.
+\todo[inline]{Popisat? To by som musel asi jedine skopirovat z clanku, to vlastnymi slovami nepopisem.}
+
 \subsection{Rejection of Uniqueness}
-While the previous steps of the key discovery produce positive statements about the keys, in this step, the algorithm searches for evidence on non-uniqueness of elements and attributes.
+While the previous steps of the key discovery produce mostly positive statements about the keys, in this step, the algorithm searches for evidence on non-uniqueness of elements and attributes.
 The aim of this step is to eliminate or decrease the number of falsely inferred keys.
@@
 \end{algorithm}
+
+\subsection{Normalization of Key Statements}
+\todo[inline]{Znovu pouzitie originalnej metody. Popisat?}

### 2012-03-07 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

@@
 \chapter{Combination with Existing Methods of Inference}
+\todo[inline]{Kontrola pomocou dat.}
+
 In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference.
@@
 
+\section{Evaluation of Paths}
 A path represented by $PathType$ can contain variable references and association of these references with other paths.
@@
 For the retaining of readability of the code, we present an evaluation of self or descendant, child, and attribute axes, and nodes specified by name.
-
-At first, the algorithm determine the axis of the step.
-If it is self or descendant axis, it returns the result of $evaluateStep\_selfOrDescendant$ function.
-This function returns the given context set extended by all descendants of the elements from the context set in the given grammar.
-
-If the axis is child axis, for each element in the context set, the algorithm determines its subelement using the grammar, and, if those with the name equal to the name specified by to step are added to the resulting context set.
-Function $getTokens$ at line 10 retrieves all elements from a regular expression.
-A specific form of the regular expression is not important, because we only need to know which elements are possible subelements of the particular element.
-
-And, at last, if the axis is attribute axis, elements from the context set are searched to contain an attribute with the specified name and those found attributes are added to the result.
 
@@
 \end{algorithm}
+
+At first, the algorithm determine the axis of the step.
+If it is self or descendant axis, it returns the result of $evaluateStep\_selfOrDescendant$ function.
+This function returns the given context set extended by all descendants of the elements from the context set in the given grammar.
+
+If the axis is child axis, for each element in the context set, the algorithm determines its subelement using the grammar, and, if those with the name equal to the name specified by to step are added to the resulting context set.
+Function $getTokens$ at line 10 retrieves all elements from a regular expression.
+A specific form of the regular expression is not important, because we only need to know which elements are possible subelements of the particular element.
+
+And, at last, if the axis is attribute axis, elements from the context set are searched to contain an attribute with the specified name and those found attributes are added to the result.
+
+\section{Saving the Inferred Statements}
+In the case of keys, the algorithm iterates through the inferred key statements.
+A key's context path $C$ is evaluated and the key is assigned to the target elements selected by the context path, one element can be assigned with multiple keys.
+Additionally, each key is assigned with a list of foreign keys that are referencing it.
+
+In the case of inferred types, the situation is not that simple, because there can occur conflicting statements. In particular, two kinds of conflicts can emerge between two statements.
+
+\begin{enumerate}
+\item The two types do not equal, but one is castable to the other.
+\item The two types do not equal and none od them is castable to the other.
+\end{enumerate}
+
+The example of conflict 1 for path $p$ without predicates is $p_1 \rightarrow date$, $p_2 \rightarrow string$.
+And another one is $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.
+There are two reasons of this kind of conflicts.
+
+The first reason is that both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a more convenient expression and is more precise.
+Consider these two expressions; $PathType$ $p$ is compared to an integer literal constant, and, $PathType$ $p$ is an argument of function where a formal type of the argument is byte.
+The first expression is utilized to infer statement $p \rightarrow integer$, and the second one to infer statement $p \rightarrow byte$.
+Both of the types are correct, but the second one is more accurate.

### 2012-03-08 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
-svn diff\chapter{Analysis of XQuery} \label{CHAPTER_analysis_of_xquery}
-\todo[inline]{Někde by se mělo říct, kterou verzi uvažujeme, jaké obecně jsou, co je třeba XQuery Core apod.}
+\chapter{Analysis of XQuery} \label{CHAPTER_analysis_of_xquery}
 This chapter discusses selected constructs of XQuery language and denotes how they could be exploited in the XML schema inference process.
@@
 Paper \cite{Necasky:2009:DXK:1529282.1529414} introduces a method of discovering keys and foreign keys by investigation of joins in queries.
-The basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form.
+The basis of this discovery is a search for particular forms of joins, so called \emph{join patterns}, but the joins are processed only if they are found in a particular syntactic form.
 Therefore joins with the same semantics written in different syntax are not taken into consideration.
@@
 
-Usage of \texttt{for} construct indicates that a sequence of \texttt{book} elements which satisfy the condition in \emph{where} \todo{tady se nám nějak změnil styl} clause is expected.
-It is a condition composed of two single conditions joined by \emph{and} logical operator.
+Usage of \texttt{for} construct indicates that a sequence of \texttt{book} elements which satisfy the condition in \texttt{where} clause is expected.
+It is a condition composed of two single conditions joined by \texttt{and} logical operator.
 Therefore, in order to satisfy the whole condition, both single conditions must be satisfied as well.
@@
 Alike, variable \texttt{\$p} is bound to a price of each title and then passed to \texttt{min()} function call.
-That indicates that one book title is supposed to have several prices, however, it cannot be said if a certain occurrence of element \emph{book} can contain more than one \texttt{price} subelement.
+That indicates that one book title is supposed to have several prices, however, it cannot be said if a certain occurrence of element \texttt{book} can contain more than one \texttt{price} subelement.
 

### 2012-03-08 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

@@
 In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference.
-We focus on a class of methods which are based on a creation and subsequent simlification of the initial grammar (as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}).
+We focus on a class of methods which are based on a creation and subsequent simplification of the initial grammar (as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}).
 
@@
 Our statements are of the three following forms.
-The first one is $p \rightarrow t$, where $p$ is a $PathType$ representing an XQuery path selecting a set of elements or attributes \todo{plati nejake obmedzenie na tuto cestu?}
-, and $t$ is an XSD built-in atomic type.
+The first one is $p \rightarrow t$, where $p$ is a $PathType$ representing an XQuery path selecting a set of elements or attributes, and $t$ is an XSD built-in atomic type.
 
@@
 
-In the case of inferred types, the situation is not that simple, because there can occur conflicting statements. In particular, two kinds of conflicts can emerge between two statements.
+In the case of inferred types, the situation is slightly less simple, because there can occur conflicting statements.
+The examples of such conflicts for path $p$ without predicates are $p_1 \rightarrow date$, $p_2 \rightarrow string$, and, $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.
 
-\begin{enumerate}
-\item The two types do not equal, but one is castable to the other.
-\item The two types do not equal and none od them is castable to the other.
-\end{enumerate}
+Note that, in both examples, one type is castable to the other (date to string, and byte to int).
+Both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a more convenient expression and is more precise.
 
-The example of conflict 1 for path $p$ without predicates is $p_1 \rightarrow date$, $p_2 \rightarrow string$.
-And another one is $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.
-There are two reasons of this kind of conflicts.
-
-The first reason is that both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a more convenient expression and is more precise.
 Consider these two expressions; $PathType$ $p$ is compared to an integer literal constant, and, $PathType$ $p$ is an argument of function where a formal type of the argument is byte.
@@
 Both of the types are correct, but the second one is more accurate.
+
+%In the case of inferred types, the situation is not that simple, because there can occur conflicting statements. In %particular, two kinds of conflicts can emerge between two statements.
+%
+%\begin{enumerate}
+%\item The two types do not equal, but one is castable to the other.
+%\item The two types do not equal and none of them is castable to the other.
+%\end{enumerate}
+%
+%The example of conflict of the first kind for path $p$ without predicates is $p_1 \rightarrow date$, $p_2 \rightarrow %string$.
+And another one is $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when %stripped of predicates, they equal $p$.
+There are two reasons of this kind of conflicts.
+%
+%The first reason is that both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a %more convenient expression and is more precise.
+Consider these two expressions; $PathType$ $p$ is compared to an integer %literal constant, and, $PathType$ $p$ is an argument of function where a formal type of the argument is byte.
+The first %expression is utilized to infer statement $p \rightarrow integer$, and the second one to infer statement $p \rightarrow %byte$.
+Both of the types are correct, but the second one is more accurate.
+
+From all inferred types for a particular path, the most specific (the most accurate) is selected and assigned to respective nodes from the grammar targeted by the path.

### 2012-03-08 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 
-Besides the mentioned type, some approaches that utilize other or additional sources have been developed \todo{reference, např.
-DP Klempa}.
-If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery \todo{ref} queries, or other, these sources can be exploited to refine the process of inference.
+Besides the mentioned type, some approaches that utilize other or additional sources have been developed, for example \cite{Mlynkova:2009:IXS:1862681.1862693} and \cite{thesis_klempa}.
+If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery \cite{w3c_xquery} queries, or other, these sources can be exploited to refine the process of inference.
 The refinement can be achieved in various aspects such as increasing the speed of the process, getting a more precise, more concise or more readable result or inference of statements about the data which cannot be (easily) extracted from the data themselves.

### 2012-03-08 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

@@
 It is a structure representing a word according to a formal grammar of a language.
-In our case, the language is XQuery, its grammar is defined in REF \todo{ref} and every query is a word of the XQuery language.
+In our case, the language is XQuery, its grammar is defined in \cite{w3c_xquery} and every query is a word of the XQuery language.
 Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals.
@@
 It is transitioned and nodes representing FLWORs are found.
-Then each subtree determined by one of the found nodes represents a FLWOR statement and it can be analyzed further.
+Then each subtree determined by one of the found nodes represents a FLWOR statement and it can be analysed further.
 
@@
 Types of literal expressions are defined, functions have return types, path expressions can return nodes, etc.
-Types of complex expressions can be determined applying the rules defined in REF \todo{ref}.
+Types of complex expressions can be determined applying the rules defined in \cite{w3c_xquery}.
 The inferred expression types can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.

### 2012-03-08 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 Heuristic methods \cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522} are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar.
-Two of these approaches are described in Chapter \ref{section_xtract} and Chapter \ref{section_even_an_ant_can_create_an_xsd} \todo{někde Chapter, někde Section - sjednotit a dodržovat}.
+Two of these approaches are described in Chapter \ref{section_xtract} and Chapter \ref{section_even_an_ant_can_create_an_xsd}.
 The former one incorporates so-called MDL principle to create the (sub)optimal result.
@@
 Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
-\todo[inline]{Q: Tento odstavec je doslovne prebraty z uvedeneho clanku. Mozem to takto pouzit? A: ano, jen by pak neměla být výrazně jiná úroveň angličtiny}
 

### 2012-03-08 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\subsection{Definition of the Syntax Tree}
+\section{Definition of the Syntax Tree}
 Firstly, we formally define the syntax tree.
@@
 Regarding the additional properties, two constructs in $Q$ represented by two nodes of the same class from $C$ may differ in certain ways, and therefore, it is needed to distinguish them.
-For instance, two different literal values in $Q$ are represented by nodes $l_1, l_2 \in V$ and $c(l_1) = c(l_2) = LiteralNode$ but each has a different value and type.
+For instance, two different literal values in $Q$ are represented by nodes $l_1, l_2 \in V$ and $c(l_1) = c(l_2) =$ \texttt{LiteralNode} but each has a different value and type.
 Therefore $\mathcal{P}$ contains functions \\
@@
 $value_{LiteralNode}: V_{LiteralNode} \rightarrow Values_{literal}$ \\
-where $V_{LiteralNode}$ is set $\{v | v \in V, c(v) = LiteralNode\}$, $Types_{literal}$ is set of all types of literal values $\{DECIMAL, INTEGER, DOUBLE, STRING\}$, and $Values_{literal}$ is a set of all literal values (all valid XQuery decimal numbers, integers, double numbers and strings).
+where $V_{LiteralNode}$ is set $\{v | v \in V, c(v) =$ \texttt{LiteralNode}$\}$, $Types_{literal}$ is set of all types of literal values \{\texttt{DECIMAL}, \texttt{INTEGER}, \texttt{DOUBLE}, \texttt{STRING}\}, and $Values_{literal}$ is a set of all literal values (all valid XQuery decimal numbers, integers, double numbers and strings).
 
@@
 
-For example, an instance of the syntax tree cannot directly contain nodes of $Node$ and $ExprNode$ classes (for every $v \in V, c(v) \neq Node, c(v) \neq ExprNode$), but it can contain nodes of $AttributeNode$ and $LiteralNode$ classes.
-Regarding the multiplicity of types, a node of $LiteralNode$ class is also considered to be of two indirect types: $ExprNode$ and $Node$.
+For example, an instance of the syntax tree cannot directly contain nodes of \texttt{Node} and \texttt{ExprNode} classes (for every $v \in V, c(v) \neq$ \texttt{Node}$, c(v) \neq$ \texttt{ExprNode}), but it can contain nodes of \texttt{AttributeNode} and \texttt{LiteralNode} classes.
+Regarding the multiplicity of types, a node of \texttt{LiteralNode} class is also considered to be of two indirect types: \texttt{ExprNode} and \texttt{Node}.
 
@@
 The node classes can be classified into three groups: inner node classes, leaf node classes and node classes that can be both inner and leaf.
-Inner node classes (marked $I$) stand for XQuery constructs that are composed of other constructs and can be further divided.
-An example of such class is $FLWORExprNode$ which is composed of $TupleStreamNode$, $WhereClauseNode$, $OrderByClauseNode$ and $ReturnClauseNode$ classes.
-Leaf node classes (marked $L$) represent elements of XQuery language that cannot be further divided.
-For example, $LiteralNode$.
-Node $FunctionCallNode$ is an example of the third group (marked $IL$).
+Inner node classes (marked \texttt{I}) stand for XQuery constructs that are composed of other constructs and can be further divided.
+An example of such class is \texttt{FLWORExprNode} which is composed of \texttt{TupleStreamNode}, \texttt{WhereClauseNode}, \texttt{OrderByClauseNode} and \texttt{ReturnClauseNode} classes.
+Leaf node classes (marked \texttt{L}) represent elements of XQuery language that cannot be further divided.
+For example, \texttt{LiteralNode}.
+Node \texttt{FunctionCallNode} is an example of the third group (marked \texttt{IL}).
 Function call of a function without arguments is represented by a leaf node while function call with arguments is represented by an inner node and its subnodes are nodes representing those arguments.
@@
 \hline
-AttributeNode (I) & ModuleNode (I) & CommaOperatorNode (I) \\
-AttrListNode (IL) & NameNode (IL) & ConstructorNode (I) \\
-AxisNode (I) & OrderByClauseNode (I) & ContextItemExprNode (L) \\
-CaseClauseNode (I) & OrderSpecNode (I) & EmptySequenceNode (L) \\
-CaseClausesNode (I) & ParamListNode (I) & ExtensionExprNode (I) \\
-CDataSectionNode (L) & ParamNode (I) & FLWORExprNode (I) \\
-ContentNode (IL) & PITargetNode (IL) & FunctionCallNode (IL) \\
-DefaultCaseNode (I) & PragmaListNode (I) & IfExprNode (I) \\
-EntityRefNode (L) & PragmaNode (L) & LiteralNode (L) \\
-\textbf{ExprHolderNode} & PredicateListNode (I) & OperatorNode (I) \\
-\textbf{ExprNode} & \textbf{PrologChildNode} & OrderedExprNode (I) \\
-FunctionBodyNode (IL) & StepExprNode (I) & PathExprNode (I) \\
-CharRefNode (L) & StringNode (L) & QuantifiedExprNode (I) \\
-InClausesNode (I) & TupleStreamNode (I) & TypeswitchExprNode (I) \\
-\textbf{ItemTypeNode} & TypeNode (L) & UnorderedExprNode (I) \\
-LocationHintNode (L) & \textbf{VariableBindingNode} & ValidateExprNode (I) \\
-LocationHintsNode (IL) & VarValueNode (IL) & VarRefNode (L) \\
+\texttt{AttributeNode (I)} & \texttt{ModuleNode (I)} & \texttt{CommaOperatorNode (I)} \\
+\texttt{AttrListNode (IL)} & \texttt{NameNode (IL)} & \texttt{ConstructorNode (I)} \\
+\texttt{AxisNode (I)} & \texttt{OrderByClauseNode (I)} & \texttt{ContextItemExprNode (L)} \\
+\texttt{CaseClauseNode (I)} & \texttt{OrderSpecNode (I)} & \texttt{EmptySequenceNode (L)} \\
+\texttt{CaseClausesNode (I)} & \texttt{ParamListNode (I)} & \texttt{ExtensionExprNode (I)} \\
+\texttt{CDataSectionNode (L)} & \texttt{ParamNode (I)} & \texttt{FLWORExprNode (I)} \\
+\texttt{ContentNode (IL)} & \texttt{PITargetNode (IL)} & \texttt{FunctionCallNode (IL)} \\
+\texttt{DefaultCaseNode (I)} & \texttt{PragmaListNode (I)} & \texttt{IfExprNode (I)} \\
+\texttt{EntityRefNode (L)} & \texttt{PragmaNode (L)} & \texttt{LiteralNode (L)} \\
+\textbf{ExprHolderNode} & \texttt{PredicateListNode (I)} & \texttt{OperatorNode (I)} \\
+\textbf{ExprNode} & \textbf{PrologChildNode} & \texttt{OrderedExprNode (I)} \\
+\texttt{FunctionBodyNode (IL)} & \texttt{StepExprNode (I)} & \texttt{PathExprNode (I)} \\
+\texttt{CharRefNode (L)} & \texttt{StringNode (L)} & \texttt{QuantifiedExprNode (I)} \\
+\texttt{InClausesNode (I)} & \texttt{TupleStreamNode (I)} & \texttt{TypeswitchExprNode (I)} \\
+\textbf{ItemTypeNode} & \texttt{TypeNode (L)} & \texttt{UnorderedExprNode (I)} \\
+\texttt{LocationHintNode (L)} & \textbf{VariableBindingNode} & \texttt{ValidateExprNode (I)} \\
+\texttt{LocationHintsNode (IL)} & \texttt{VarValueNode (IL)} & \texttt{VarRefNode (L)} \\
 \textbf{ModuleChildNode} & & \\
@@
 \hline
-BaseURIDeclNode (L) & BindingSequenceNode (I) \\
-BoundarySpaceDeclNode (L) & DefaultValueNode (I) \\
-ConstructionDeclNode (L) & ElseExpressionNode (I) \\
-CopyNamespacesDeclNode (L) & OperandExpressionNode (I) \\
-DefaultCollationDeclNode (L) & ReturnClauseNode (I) \\
-DefaultNamespaceDeclNode (L) & TestExpressionNode (I) \\
-EmptyOrderDeclNode (L) & ThenExpressionNode (I) \\
-FunctionDeclNode (I) & WhereClauseNode (I) \\
+\texttt{BaseURIDeclNode (L)} & \texttt{BindingSequenceNode (I)} \\
+\texttt{BoundarySpaceDeclNode (L)} & \texttt{DefaultValueNode (I)} \\
+\texttt{ConstructionDeclNode (L)} & \texttt{ElseExpressionNode (I)} \\
+\texttt{CopyNamespacesDeclNode (L)} & \texttt{OperandExpressionNode (I)} \\
+\texttt{DefaultCollationDeclNode (L)} & \texttt{ReturnClauseNode (I)} \\
+\texttt{DefaultNamespaceDeclNode (L)} & \texttt{TestExpressionNode (I)} \\
+\texttt{EmptyOrderDeclNode (L)} & \texttt{ThenExpressionNode (I)} \\
+\texttt{FunctionDeclNode (I)} & \texttt{WhereClauseNode (I)} \\
 \cline{2-2} \textbf{ImportNode} & \textbf{\emph{ItemTypeNode}} \\
-\cline{2-2} NamespaceDeclNode (L) & AnyItemNode (L) \\
-OptionDeclNode (L) & AtomicTypeNode (L) \\
-OrderingModeDeclNode (L) & KindTestNode (IL) \\
-VarDeclNode (I) & NameTestNode (L) \\
+\cline{2-2} \texttt{NamespaceDeclNode (L)} & \texttt{AnyItemNode (L)} \\
+\texttt{OptionDeclNode (L)} & \texttt{AtomicTypeNode (L)} \\
+\texttt{OrderingModeDeclNode (L)} & \texttt{KindTestNode (IL)} \\
+\texttt{VarDeclNode (I)} & \texttt{NameTestNode (L)} \\
 \cline{1-2} \textbf{\emph{ImportNode}} & \textbf{\emph{VariableBindingNode}} \\
-\cline{1-2} ModuleImportNode (I) & ForClauseNode (I) \\
-SchemaImportNode (I) & InClauseNode (I) \\
-\cline{1-1} \textbf{\emph{ModuleChildNode}} & LetClauseNode (I) \\
-\cline{1-2} ModuleDeclNode (L) & \emph{StepExprNode} (I) \\
-\cline{2-2}PrologNode (IL) & SelfOrDescendantStepNode (L) \\
-QueryBodyNode (I) & 
+\cline{1-2} \texttt{ModuleImportNode (I)} & \texttt{ForClauseNode (I)} \\
+\texttt{SchemaImportNode (I)} & \texttt{InClauseNode (I)} \\
+\cline{1-1} \textbf{\emph{ModuleChildNode}} & \texttt{LetClauseNode (I)} \\
+\cline{1-2} \texttt{ModuleDeclNode (L)} & \texttt{\emph{StepExprNode} (I)} \\
+\cline{2-2} \texttt{PrologNode (IL)} & \texttt{SelfOrDescendantStepNode (L)} \\
+\texttt{QueryBodyNode (I)} & 
 \end{tabular}
@@
 
-\begin{define}[Function is(v, class)]For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
+\begin{define}[Function $is(v, class)$]For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
 Otherwise, it returns $false$.
@@
 
-For instance, assuming $v \in V, c(v) = LiteralNode$, calls $is(v, LiteralNode)$, $is(v, ExprNode)$ returns $true$, while call $is(v, ContentNode)$ returns $false$.
+For instance, assuming $v \in V, c(v) =$ \texttt{LiteralNode}, calls $is(v,$ \texttt{LiteralNode}$)$ and $is(v,$ \texttt{ExprNode}$)$ return $true$, while call $is(v,$ \texttt{ContentNode}$)$ returns $false$.
 
@@
 An important characteristic of the syntax tree is related to a definition of local variables and their scope in the XQuery language.
-The representation of a definition of a local variable in the syntax tree is a node of $VariableBinding$ class.
+The representation of a definition of a local variable in the syntax tree is a node of \texttt{VariableBindingNode} class.
 Nodes of that class have only two children; a node representing the type of the variable and a node representing the binding expression (expression defining the value of the variable, and thus, it cannot use the variable).
 Hence, the entire subtree does not contain any expressions that use the variable.
-Therefore the scope of the new variable is not the subtree of the $VariableBinding$ class node.
+Therefore the scope of the new variable is not the subtree of the \texttt{VariableBindingNode} class node.
 It depends on the type of XQuery construct that the variable binding is a (indirect) component of.
 
-For example, a node of $FLWORExpr$ class contains four subnodes of\linebreak $TupleStream$, $WhereClause$, $OrderByClause$ and $ReturnClause$ classes.
-The $TupleStream$ class node contains a list of node of $variableBinding$ class which define variables valid in all other three subnodes of the FLWOR node.
+For example, a node of \texttt{FLWORExprNode} class contains four subnodes of\linebreak \texttt{TupleStreamNode}, \texttt{WhereClauseNode}, \texttt{OrderByClauseNode} and\linebreak \texttt{ReturnClauseNode} classes.
+The \texttt{TupleStreamNode} class node contains a list of nodes of \texttt{VariableBindingNode} class which define variables valid in all other three subnodes of the FLWOR node.
 
@@
 \item XML Schema built-in atomic types \todo{link do appendixu}.
-\item Types $ElementType$, $AttributeType$, $NodeType$, $TextNodeType$,\linebreak $CommentType$, $ProcessingInstructionType$, $DocumentType$ representing an element, attribute, node, text node, comment, processing instruction, document node.
-\item Type representing a node or a set of nodes selected by a certain path expression. The path expression is included in this type. Let this type be identified as $PathType$.
-\item $UnknownType$ representing a type without known details, which does not suit one of the three previous types. An example is XSD type $anyType$.
+\item Types \texttt{ElementType}, \texttt{AttributeType}, \texttt{NodeType}, \texttt{TextNodeType},\linebreak \texttt{CommentType}, \texttt{ProcessingInstructionType}, \texttt{DocumentType} representing an element, attribute, node, text node, comment, processing instruction, document node.
+\item Type representing a node or a set of nodes selected by a certain path expression. The path expression is included in this type. Let this type be identified as \texttt{PathType}.
+\item \texttt{UnknownType} representing a type without known details, which does not suit one of the three previous types. An example is XSD type \texttt{anyType}.
 \end{itemize}
@@
 \subsection{PathType}
-$PathType$ contains additional information.
-The represented path is contained by an instance of $PathExprNode$ containing list of its steps, in particular instances of $StepExprNode$.
-If a step is a reference to a variable which type is $PathType$, we also want to include this information.
-Therefore, $PathType$ contain association between the steps and other $PathTypes$ and this association is defined for the $PathType$ variable steps.
-
-To differ between a common $PathType$ selecting a set of nodes and a $PathType$ bound to a for variable in a FLWOR expression, $PathType$ structure contains a flag $isForBound$.
-
-Additionaly, $PathType$ contains a list of special functions that were called with an argument of $PathType$ type.
-The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of $PathType$ because then, we can determine a type of the function call more precisely.
-Those special functions are built-in functions $data$, $min$, $max$ \todo{Doplnit}.
-
-In summary, we represent $PathType$ as a structure with the following member variables. \todo{Radsej UML?}
+\texttt{PathType} contains additional information.
+The represented path is contained by an instance of \texttt{PathExprNode} containing list of its steps, in particular instances of \texttt{StepExprNode}.
+If a step is a reference to a variable which type is \texttt{PathType}, we also want to include this information.
+Therefore, \texttt{PathType} contain association between the steps and other \texttt{PathType}s and this association is defined for the \texttt{PathType} variable steps.
+
+To differ between a common \texttt{PathType} selecting a set of nodes and a \texttt{PathType} bound to a for variable in a FLWOR expression, \texttt{PathType} structure contains a flag \texttt{isForBound}.
+
+Additionally, \texttt{PathType} contains a list of special functions that were called with an argument of \texttt{PathType} type.
+The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of \texttt{PathType} because then, we can determine a type of the function call more precisely.
+Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max} \todo{Doplnit}.
+
+In summary, we represent \texttt{PathType} as a structure with the following member variables.
 \begin{itemize}
-\item $pathExprNode$ - An instance of $PathExprNode$.
-\item $substeps$ - An association between references from $stepNodes$ and instances of $PathType$ type.
-\item $isForBound$ - Boolean flag determining if the type was bound to a for variable in a for clause.
-\item $specialFunctionCalls$ - List of special functions called with this instance as an argument.
+\item \texttt{pathExprNode} - An instance of \texttt{PathExprNode}.
+\item \texttt{substeps} - An association between steps from \texttt{pathExprNode} and instances of \texttt{PathType} type.
+\item \texttt{isForBound} - Boolean flag determining if the type was bound to a for variable in a for clause.
+\item \texttt{specialFunctionCalls} - List of special functions called with this instance as an argument.
 \end{itemize}
@@
 
-To capture sequences, we assign the first two categories of types (all types except of $PathType$ and $UnknownType$) with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}.
+To capture sequences, we assign the first two categories of types (all types except of \texttt{PathType} and \texttt{UnknownType}) with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}.
 Each of those types can be perceived as a sequence.
@@
 \item A sequence of exactly one item.
-\item A sequence containing zero or one item (modifier $?$).
-\item A sequence containing zero or more items (modifier $*$).
-\item A sequence containing one or more items (modifier $+$).
+\item A sequence containing zero or one item (modifier \texttt{?}).
+\item A sequence containing zero or more items (modifier \texttt{*}).
+\item A sequence containing one or more items (modifier \texttt{+}).
 \end{itemize}
 
-$PathType$ is not assigned with the cardinality since we do not evaluate the queries, and therefore, we cannot determine if a certain XQuery path targets zero, one or more nodes.
-Alike, $UnknownType$ is neither assigned with the cardinality.
-Expressions of $UnknownType$ are not utilized it the inference, therefore, their cardinality is not needed.
+\texttt{PathType} is not assigned with the cardinality since we do not evaluate the queries, and therefore, we cannot determine if a certain XQuery path targets zero, one or more nodes.
+Alike, \texttt{UnknownType} is neither assigned with the cardinality.
+Expressions of \texttt{UnknownType} are not utilized it the inference, therefore, their cardinality is not needed.
 
@@
 \begin{algorithm}
-\caption{Retrieval of Function Declaration Nodes $getFunctionDeclarationNodes(syntaxTree)$}
+\caption{Function \texttt{getFunctionDeclarationNodes}: Retrieval of Function Declaration Nodes}
 \label{ALG_retrieval_of_function_declaration_nodes}
 \begin{algorithmic}[1]
-\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
+\REQUIRE syntaxTree: A reference to the root node of a syntax tree.
 \ENSURE A list of syntax tree nodes representing function declarations.
 
-\STATE $prologNode = null$
-\FORALL{$moduleChildNode \in syntaxTree.getChildren()$}
-    \IF{$is(moduleChildNode, PrologNode)$}
-        \STATE $prologNode = moduleChildNode$
-    \ENDIF
-\ENDFOR
-
-\STATE $functionDeclarationNodes =$ an empty list
-\IF{$prologNode \neq null$}
-    \FORALL{$prologChildNode \in prologNode.getChildren()$}
-        \IF{$is(prologChildNode, FunctionDeclNode)$}
-            \STATE add $prologChildNode$ to $functionDeclarationNodes$
-        \ENDIF
-    \ENDFOR
-\ENDIF
-
-\RETURN $functionDeclarationNodes$
+\STATE prologNode $=$ null
+\FORALL{moduleChildNode $\in$ syntaxTree.getChildren()}
+    \IF{is(moduleChildNode, PrologNode)}
+        \STATE prologNode $=$ moduleChildNode
+    \ENDIF
+\ENDFOR
+
+\STATE functionDeclarationNodes $=$ an empty list
+\IF{prologNode $\neq$ null}
+    \FORALL{prologChildNode $\in$ prologNode.getChildren()}
+        \IF{is(prologChildNode, FunctionDeclNode)}
+            \STATE add prologChildNode to functionDeclarationNodes
+        \ENDIF
+    \ENDFOR
+\ENDIF
+
+\RETURN functionDeclarationNodes
 \end{algorithmic}
@@
 A simple algorithm is presented in Algorithm \ref{ALG_processing_of_functions}.
-It uses \texttt{getFunction\-DeclarationNodes()} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three.
+It uses\linebreak \texttt{getFunctionDeclarationNodes} function, defined in Algorithm \ref{ALG_retrieval_of_function_declaration_nodes}, which returns a list of all function declaration nodes in the syntax three.
 Actually, it does not have to search the whole syntax tree as the function declaration nodes can be present only in the query prolog section.
@@
 We will not determine types of functions defined in other modules since the principle is similar but it is needed to look up the definitions in syntax trees of other queries.
-Return types of built-in functions are fixed and, thus, there is no need to analyze them.
+Return types of built-in functions are fixed and, thus, there is no need to analyse them.
 Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by $fn$ or not prefixed.
@@
 \begin{algorithmic}[1]
-\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
+\REQUIRE syntaxTree: A reference to the root node of a syntax tree.
 \ENSURE An associative array of function names with their types and references to the their declaration nodes.
 
-\STATE $functionArray$ = an empty associative array
-\FORALL{$functionDeclarationNode \in$ \\ $getFunctionDeclarationNodes(syntaxTree)$}
-    \STATE $functionName = functionDeclarationNode.funcName$
-    \STATE $typeNode = functionDeclarationNode.getChild(TypeNode)$
-    \STATE $functionArray[functionName] =$ \\ $(getTypeTN(typeNode), functionDeclarationNode)$
-\ENDFOR
-\RETURN $functionArray$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Extraction of a Type from TypeNode getTypeTN()}
+\STATE functionArray $=$ an empty associative array
+\FORALL{functionDeclarationNode $\in$ \\ getFunctionDeclarationNodes(syntaxTree)}
+    \STATE functionName $=$ functionDeclarationNode.funcName
+    \STATE typeNode $=$ functionDeclarationNode.getChild(TypeNode)
+    \STATE functionArray[functionName] $=$ \\ (getTypeTN(typeNode), functionDeclarationNode)
+\ENDFOR
+\RETURN functionArray
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function \texttt{getTypeTN}: Extraction of a Type from TypeNode}
 \label{ALG_extraction_of_a_type_from_typenode}
 \begin{algorithmic}[1]
-\REQUIRE $typeNode$: Syntax tree node of TypeNode class.
-\ENSURE Type extracted from $typeNode$.
-
-\STATE $type = null$
-\STATE $cardinality = typeNode.cardinality$
-\STATE $itemTypeNode = typeNode.getChild(ItemTypeNode)$
-\IF{$is(itemTypeNode, AtomicTypeNode)$}
-    \STATE $type = itemTypeNode.typeName$
-\ELSIF{$is(itemTypeNode, KindTestNode)$}
-    \STATE $type = itemTypeNode.nodeKind$
-\ENDIF
-
-\RETURN $type$, $cardinality$
+\REQUIRE typeNode: Syntax tree node of TypeNode class.
+\ENSURE Type extracted from typeNode.
+
+\STATE type = null
+\STATE cardinality = typeNode.cardinality
+\STATE itemTypeNode = typeNode.getChild(ItemTypeNode)
+\IF{is(itemTypeNode, AtomicTypeNode)}
+    \STATE type = itemTypeNode.typeName
+\ELSIF{is(itemTypeNode, KindTestNode)}
+    \STATE type = itemTypeNode.nodeKind
+\ENDIF
+
+\RETURN type, cardinality
 \end{algorithmic}
@@
 
-For the rest of the thesis, we assume function $getFunctionReturnType()$ which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
+For the rest of the thesis, we assume function \texttt{getFunctionReturnType} which takes a function name and returns the return type of the function if the function is either built-in or it is recorded in the associative array.
 Otherwise, \texttt{null} is returned.
@@
 Later phases of the algorithm need to process the declarations of certain functions.
-Therefore, we also store references to the entire function declaration nodes, accessible using function $getFunctionDeclNode()$.
+Therefore, we also store references to the entire function declaration nodes, accessible using function \texttt{getFunctionDeclNode}.
 
@@
 
-Function $memorizeType()$ takes two arguments $node$ and $type$ and it memorizes the given type of the specified expression node.
-This information can be then retrieved by function $getType()$, specifying the particular node as its argument.
-
-Another group consists of functions $set()$, $add()$, and $get()$.
-Function $set()$ memorizes a given value of a specified property of a specified node.
-It is used to assign a node with a named value, for example, $set(someNode, color, 'red')$ will assign $someNode$ with string value $'red'$ wich can be then retrieved by function $get()$, specifying the node and the property name.
-For example,\linebreak $get(someNode, color)$ returns $'red'$.
-A subsequent call of $set()$ assigning a node with a value of already assigned property will overwrite it, leaving the property with the newer value.
-However, a property can have several values and this can be achieved using function $add()$ with the same syntax as function $set()$ but instead of overwriting the existing property, $add()$ will append the new value to the existing ones.
-Then, a call of $get()$ on this node and property returns a list of all its values, preserving the order of their addition.
+Function \texttt{memorizeType} takes two arguments \texttt{node} and \texttt{type} and it memorizes the given type of the specified expression node.
+This information can be then retrieved by function \texttt{getType}, specifying the particular node as its argument.
+
+Another group consists of functions \texttt{set}, \texttt{add}, and \texttt{get}.
+Function \texttt{set} memorizes a given value of a specified property of a specified node.
+It is used to assign a node with a named value, for example, \texttt{set(someNode, color, "red")} will assign \texttt{someNode} with string value \texttt{"red"} which can be then retrieved by function \texttt{get}, specifying the node and the property name.
+For example,\linebreak \texttt{get(someNode, color)} returns \texttt{"red"}.
+A subsequent call of \texttt{set} assigning a node with a value of already assigned property will overwrite it, leaving the property with the newer value.
+However, a property can have several values and this can be achieved using function \texttt{add} with the same syntax as function \texttt{set} but instead of overwriting the existing property, \texttt{add} will append the new value to the existing ones.
+Then, a call of \texttt{get} on this node and property returns a list of all its values, preserving the order of their addition.
 
@@
 \begin{algorithm}
-\caption{Function determineExpressionType()}
+\caption{Function \texttt{determineExpressionType}}
 \label{ALG_func_determineExpressionType}
@@
 \REQUIRE{\ \\
-    $exprNode$: An expression node in the syntax tree. \\
-    $contextVarTypes$: Types of local variables valid in the current subtree.
+    exprNode: An expression node in the syntax tree. \\
+    contextVarTypes: Types of local variables valid in the current subtree.
 }
@@
 
-\STATE $type = UnknownType$
-\IF{$is(exprNode, LiteralNode)$}
-     \STATE $type =$ XSD atomic type $exprNode.type$, exactly-once cardinality
-\ELSIF{$is(exprNode, FunctionCallNode)$}
-     \STATE $type = getFunctionReturnType(exprNode.fncName)$
-\ELSIF{$is(exprNode, VarRefNode)$}
-     \STATE $type = getVariableType(contextVarTypes, exprNode.varName)$
-     \IF{$type$ is $PathType$}
-         \STATE $step$ = a path step (new instance of $StepExprNode$) representing the variable reference
-		 \STATE $substeps$ = empty associative array         
-         \STATE $substeps[step]$ = $type$
-         \STATE $pathExprNode$ = new instance of $PathExprNode$ containing one step $step$
-         \STATE $type$ = $PathType$ containing $pathExprNode$ and $substeps$
-     \ENDIF
-\ELSIF{$is(exprNode, PathExprNode)$}
-     \STATE $type = createPathType(exprNode)$
-\ELSIF{$is(exprNode, OperatorNode)$}
-     \STATE $type = determineOperatorType(exprNode, contextVarTypes)$
-\ELSIF{$is(exprNode, FLWORExprNode)$}
-     \STATE $returnClauseNode = exprNode.getChild(ReturnClauseNode)$
-     \STATE $type = createForUnboundType(getType(returnClauseNode))$
-\ENDIF
-
-\RETURN $type$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Function determineOperatorType()}
+\STATE type $=$ UnknownType
+\IF{is(exprNode, LiteralNode)}
+     \STATE type $=$ XSD atomic type exprNode.type, exactly-once cardinality
+\ELSIF{is(exprNode, FunctionCallNode)}
+     \STATE type $=$ getFunctionReturnType(exprNode.fncName)
+\ELSIF{is(exprNode, VarRefNode)}
+     \STATE type = getVariableType(contextVarTypes, exprNode.varName)
+     \IF{type is PathType}
+         \STATE step $=$ a path step (new instance of StepExprNode) representing the variable reference
+		 \STATE substeps = empty associative array         
+         \STATE substeps[step] $=$ type
+         \STATE pathExprNode $=$ new instance of PathExprNode containing one step step
+         \STATE type $=$ PathType containing pathExprNode and substeps
+     \ENDIF
+\ELSIF{is(exprNode, PathExprNode)}
+     \STATE type $=$ createPathType(exprNode)
+\ELSIF{is(exprNode, OperatorNode)}
+     \STATE type $=$ determineOperatorType(exprNode, contextVarTypes)
+\ELSIF{is(exprNode, FLWORExprNode)}
+     \STATE returnClauseNode $=$ exprNode.getChild(ReturnClauseNode)
+     \STATE type $=$ createForUnboundType(getType(returnClauseNode))
+\ENDIF
+
+\RETURN type
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function \texttt{determineOperatorType}}
 \label{ALG_func_determineOperatorType}
@@
 
-\STATE $type = UnknownType$
-\STATE $operator = operatorNode.operator$
-\IF{$isOperatorClassComparison(operator)$}
-    \STATE $type = xs:boolean$, exactly-one cardinality
-\ELSIF{$isOperatorClassAddition(operator)$}
-    \STATE $leftOperandType = getType(exprNode.leftSide)$
-    \STATE $rightOperandType = getType(exprNode.rightSide)$
-    \IF{$isNumericType(leftOperandType)$ $AND$ $isNumericType(rightOperandType)$}
-        \STATE $type = selectMoreGeneralNumericType(leftOperandType,$ $rightOperandType)$
-    \ELSIF{$isNumericType(leftOperandType)$}
+\STATE $type =$ UnknownType
+\STATE $operator =$ $operatorNode.operator$
+\IF{isOperatorClassComparison($operator$)}
+    \STATE $type =$ boolean, exactly-one cardinality
+\ELSIF{isOperatorClassAddition($operator$)}
+    \STATE $leftOperandType =$ getType($exprNode.leftSide$)
+    \STATE $rightOperandType =$ getType($exprNode.rightSide$)
+    \IF{isNumericType($leftOperandType$) $\wedge$ isNumericType($rightOperandType$)}
+        \STATE $type =$ selectMoreGeneralNumericType($leftOperandType$, $rightOperandType$)
+    \ELSIF{isNumericType($leftOperandType$)}
         \STATE $type = leftOperandType$
-    \ELSIF{$isNumericType(rightOperandType)$}
+    \ELSIF{isNumericType($rightOperandType$)}
         \STATE $type = rightOperandType$
@@
 
-\RETURN $type$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Function isOperatorClassComparison()}
+\RETURN type
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function \texttt{isOperatorClassComparison}}
 \label{ALG_func_isOperatorClassComparison}
@@
 \REQUIRE $operator$: Representation of a XQuery operator.
-\ENSURE $true$ if $operator$ is a comparison operator, $false$ otherwise.
-
-\IF{$operator$ equals one of $GEN\_EQUALS$, $GEN\_GREATER\_THAN$, $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_LESS\_THAN$, $GEN\_LESS_THAN\_EQUALS$, $GEN\_NOT\_EQUALS$, $VAL\_EQUALS$, $VAL\_GREATER\_THAN$, $VAL\_GREATER\_THAN\_EQUALS$, $VAL\_LESS\_THAN$, $VAL\_LESS\_THAN\_EQUALS$, $VAL\_NOT\_EQUALS$}
-    \RETURN $true$
+\ENSURE true if operator is a comparison operator, false otherwise.
+
+\IF{$operator$ equals one of GEN\_EQUALS, GEN\_GREATER\_THAN, GEN\_GREATER\_THAN\_EQUALS, GEN\_LESS\_THAN, GEN\_LESS\_THAN\_EQUALS, GEN\_NOT\_EQUALS, VAL\_EQUALS, VAL\_GREATER\_THAN, VAL\_GREATER\_THAN\_EQUALS, VAL\_LESS\_THAN, VAL\_LESS\_THAN\_EQUALS, VAL\_NOT\_EQUALS}
+    \RETURN true
 \ELSE
-    \RETURN $false$
+    \RETURN false
 \ENDIF
@@
 
-The first XQuery construct utilized by this step is a function call of $distinct-values$ function as discussed in Section \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
+The first XQuery construct utilized by this step is a function call of $distinct-values$ function as discussed in Chapter \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
 The algorithm searches for calls of the mentioned functions with a $PathType$ on their input.
@@
 
-As discussed in Section \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the count of the items may be more than one.
+As discussed in Chapter \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the count of the items may be more than one.
 And, together with that, the set is restricted to contain only items which satisfy a condition that a value of one of their subnodes (or descendant nodes) equals a certain value.

### 2012-03-08 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 \begin{verbatim}
-\\event\date = current-date()
+//event/date = current-date()
 
-\catalog\product\price < 24.5
+/catalog/product/price < 24.5
 
@@
 {...};
-\catalog\product\local:byteFunction(@id)
+/catalog/product/local:byteFunction(@id)
 \end{verbatim}
-\todo[inline]{overit, ci su priklady OK}
 
@@
 
-Alongside common expressions, there are other XQuery constructs indicating types, such as \emph{type casting}, \emph{constructors} and so called \emph{type declarations}.
-\todo[inline]{\^ Nevíme co to je. Chce to vysvětlit. Buď v kapitole 2 nebo tady + odkaz do specifikace.}
+Alongside common expressions, there are other XQuery constructs indicating types, such as \emph{type casting}, \emph{type constructors} and so called \emph{type declarations} (see \cite{w3c_xquery} for specification), demonstrated in the following examples.
 
@@
 
-Although determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues.
+%Although determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues.
 More complex queries are commonly branched by \emph{if-then-else} construct.
@@
 Therefore, the reliable method would take the query structure into consideration and further research would be needed.
-\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}
+%\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}
 
@@
 
-A common three-way join. \todo{Tady asi něco chybí.}
+A common three-way join can be also utilize to infer keys and foreign keys, however, the respective XML data and their analysis would be needed to tell what is a key and what is a foreign key.
 
@@
 for $prod in $prods 
-where $prod << $prods[@dept = $prod/@dept][last( )] 
+where $prod << $prods[@dept = $prod/@dept][last()] 
 return $prod 
@@
 
-\todo[inline]{Tady to chce ějaký komentář.}
+Usage of predicate \texttt{[last()]} indicates that \texttt{\$prod} is a sequence of elements.
 

### 2012-03-08 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

--- 2011-12-14 doc/theses/mikula/pre_creation_of_algorithm.tex
@@
 Another step can be evaluation of the queries using the XML data and consecutive analysis of the results.
-And even the process of evaluation itself can be analyzed to obtain some useful information.
-For instance, these are partial results of evaluating of expressions (elements selected by each path expression, real arguments in fuction calls, etc).
-
-\todo[inline]{Dopisat, ako sme sa rozhodli - nejaky zaver.
-Este presne neviem, ale pravdepodobne prva faza bude spracovanie cisto queries a druha by mohla byt kontrola na zaklade XML dat.
-Na dalsie vstupne data asi nebude kapacita}
+And even the process of evaluation itself can be analysed to obtain some useful information.
+For instance, these are partial results of evaluating of expressions (elements selected by each path expression, real arguments in function calls, etc).
 
@@
 
-The result of lexical and syntax analyses can be some kind of so-called syntax tree \cite{compilers}.
-It is a structure representing a word according to a formal grammar of some language.
-In our case, the language is XQuery, its grammar is defined in REF \todo{ref} and every query is a word of the XQuery language.
+The result of lexical and syntax analyses can be a kind of so-called syntax tree \cite{compilers}.
+It is a structure representing a word according to a formal grammar of a language.
+In our case, the language is XQuery, its grammar is defined in \cite{w3c_xquery} and every query is a word of the XQuery language.
 Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals.
@@
 The function returns a path expression typed as \texttt{element()}.
-That means the function returns one element \todo{pravda?}.
+That means the function returns one element.
 In the path expression, the argument \texttt{\$id} can be substituted by the real value specified in the call.
@@
 Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"} is of some numeric type.
-And, for the function is parametrized there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.
+And, since the function is parametrized, there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.
 
-While simplier approaches of the query processing such as the pattern finding limit possibilities of the query utilization, a more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
+While simplifier approaches of the query processing such as the pattern finding limit possibilities of the query utilization, a more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
 Therefore, we decided to incorporate query processing using the lexical and syntax analyses.
-
-\todo[inline]{Tak náročnost implementace rozhodně není kritérium.
-Ta se udělá jednou.
-Navíc ve vědě (tj.
-i DP) nemá opodstatnění.
-Pro DP navíc platí, že nemusíte implementovat vše, co navrhnete.
-Implementace je pomocná věc, která slouží k tomu, abychom experimentálně ukázali vlastnosti a chování navrženého algoritmu.}
 
@@
 When we infer some subelements of a certain element we often cannot be sure about their number of occurrences.
-Also, we cannot be sure if every occurrence of the certain element contains these subelements and we even do not know if at least one occurrence of the element contains them.
+Also, we cannot be sure if every occurrence of a certain element contains the subelements and we even do not know if at least one occurrence of the element contains them.
 Thus, the inferred statement is more likely an indication than a fact about the structure.
 
-If we want to infer the structure more precisely we need to evaluate the queries, and thus, we need the XML data to evaluate the queries upon.
-But, if we have the XML data, we can infer the structure directly from them in an easier and more convenient way.
-
-\todo[inline]{Ono je to spíš naopak.
-To, že známe dotazy, nám může omezit prostor možných XML schémat, která popisují strukturu XML dat.
-Jinak řečeno, může nám to zrychlit odvozování struktury, protože z dotazu víme, že některé možnosti můžeme vyřadit.}
-
-Considering the described issues, we concluded not to deal with the inference of XML structure in this thesis.
+Since the inference of XML structure utilizing only queries is not clear, we need the XML data, if we want to infer the structure more precisely.
+And, if we have the XML data, we can infer the structure directly from them using an existing approach and utilize queries to refine its result.
 
 \section{Extension of an Existing Approach}
-The existing approaches of XML schema inference deal mainly with the inference of XML structure.
-As stated in the previous section, this work does not deal with the inference of XML structure, and hence, the extension of an existing approach will be a kind of an independent addition instead of modification and refinement of its core algorithm.
+The existing approaches of XML schema inference deal mainly with inference of XML structure.
+Hence, the extension of an existing approach will be a kind of an independent addition instead of modification and refinement of its core algorithm.
 
@@
 The first phase will be taken from an existing approach and it will process the XML data to infer the XML structure.
-The second phase will process the XQuery queries and it will infer statements that can be inferred indepedently of the XML data.
+The second phase will process the XQuery queries and it will infer statements that can be inferred independently of the XML data.
 The third phase will merge the statements inferred in the second phase into the resulting schema.
 This phase may also infer additional statements from both the XML data and the queries or it may try to verify the statements from the second phase with respect to the XML data.
+
+A more advanced method can exploit queries to refine a core algorithm (e.g.
+merging of PTA) of an existing approach.
+The approach described in \cite{Vosta:2008:EAC:1802514.1802522} distinguishes elements with the same name, but a different content model and context.
+Some information from queries may help to improve this algorithm.
+For example, consider an XML representation of company data containing names of employees, costumers, and products represented by element \texttt{name}.
+The names of employees and costumers consists of two subelements for first name and surname.
+The names of products are atomic strings.
+During analysis of available queries, we may find that elements representing the names of employees and elements representing the names of costumers are processed in the same way (e.g.
+they are mixed in one sequence), while elements representing the names of products are processed separately.
+This suggests that the elements representing the names of employees and costumers have the same semantic and the same content model which is different from the content model of the elements representing the names of products.

### 2012-03-09 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

@@
 
-\texttt{lt} is a representant of so called \emph{value comparison operators} \todo{referencia na def} which operate two sequences of zero or one item.
-If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} \todo{referencia na def} is raised.
+\texttt{lt} is a representant of so called \emph{value comparison operators} (\texttt{eq}, \texttt{ne}, \texttt{lt}, \texttt{le}, \texttt{gt}, \texttt{ge}, see \cite{w3c_xquery}) which operate two sequences of zero or one item.
+If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} (see \cite{w3c_xquery}) is raised.
 

### 2012-03-09 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

@@
 Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals.
-\todo{priklad nejakeho stromu?} From the point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient.
+From the point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient.
 For instance, the tree can simplify a search for FLWOR statements.
@@
 Then each subtree determined by one of the found nodes represents a FLWOR statement and it can be analysed further.
-
-\todo[inline]{Q: definovat grammar, tree, language? A: Pokud bude použito dále, tak ano.}
 

### 2012-03-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 The first step of the algorithm involves lexical and syntax analyses known from the construction of compilers and produces a so-called syntax tree.
-The analyses are taken from Ji\v r\'{i} Schejbal's master thesis \todo{ref}.
+The analyses are taken from Ji\v r\'{i} Schejbal's master thesis \cite{thesis_schejbal}.
 Since they are not directly related to the inference, and thus, they are not directly related to the topic of this thesis, we will not describe them.
@@
 Set $\mathcal{P}$ contains other similar functions but due to their large number, we do not define them formally.
-\todo[inline]{Q: Moze takto ostat, alebo definovat aspon tie co pouzivame? Alebo ich aspon vypisat? A: aspoň příklad, lépe vypsat ty, co použijeme}
+Functions $varName_{VarRefNode}$, $axisKind_{AxisNode}$, $operator_{OperatorNode}$ are examples of commonly used functions from $\mathcal{P}$.
+Their meaning will be explained in a place of their use.
+For detail, see \cite{thesis_schejbal}.
 
@@
 \item $v.p$ stands for $p_{c(v)}(v)$ if $p_{c(v)} \in \mathcal{P}$.
+This is a shortened syntax for functions from $\mathcal{P}$.
+Assuming $v \in V$, $c(v) =$ \texttt{VarRefNode}, and $varName_{VarRefNode} \in \mathcal{P}$; expression $varName_{VarRefNode}(v)$ can be abbreviated to $v.varName$.
 \item $v.getChildren()$ stands for $\{w | w \in V, (v,w) \in E\}$ which is a set of all children of $v$.
 \item $v.getChild(i)$, $i \in \mathbb{N} \cup \{0\}$, stands for $w \in V$ so that $o(v)(v,w) = i + 1$.
+In other words, it is a child of $v$ with sequential number $i$ in order specified by $o$, starting with 0.
 \item $v.getChild(class), class \in C$, stands for $w \in V$ so that $(v,w) \in E$ and $c(w) = class$, if $\Bigl | \{u | u \in V, (v,u) \in E, c(u) = class\}\Bigr | = 1$.
-\item If exists $u \in V$ so that $(u,v) \in E$, $v.getParent()$ stands for that $u$.
+It is a node satisfying two conditions; it is a child of $v$ and its class is $class$.
+The function is defined when there is exactly one child of $v$ of $class$ class.
+\item If exists $u \in V$ so that $(u,v) \in E$, $v.getParent()$ stands for that $u$, which is a parent of $v$.
 \end{itemize}
-
-\todo[inline]{Q: Je nutne presnejsie popisat kedy a preco tieto skratky maju zmysel?
-V pseudo-kodoch budu samozrejme pouzivane len v pripadoch, ked su dobre definovane a maju zmysel.
-A: Ne, ale hodilo by se slovně popsat co znamenají.}
 
@@
 Regarding the multiplicity of types, a node of \texttt{LiteralNode} class is also considered to be of two indirect types: \texttt{ExprNode} and \texttt{Node}.
-
-\todo[inline]{Q: Asi bude potrebne nakreslit pre kazdy node, co presne obsahuje.
-Alebo staci pre tie, ktore budeme my pouzivat?
-A: Určitě všude stačí popsat jen to, co použijeme, s poznámkou, že zbytek je např.
-v diplomce, ze které vycházíte.}
 
@@
 \begin{figure}
-\caption{Syntax tree node types part 1.}
+\caption{Syntax tree node types part 1. For details, see \cite{thesis_schejbal}.}
 \label{FIG_syntax_tree_node_types_part_1}
@@
 \begin{figure}
-\caption{Syntax tree node types part 2.}
+\caption{Syntax tree node types part 2. For details, see \cite{thesis_schejbal}.}
 \label{FIG_syntax_tree_node_types_part_2}
@@
 
-\todo[inline]{Nejak sa z tohoto vytratilo to, preco to je pre nas dolezita charakteristika}
+This characteristic is explicitly described, because several algorithms later in this chapter rely on it.
 
@@
 \subsection{Expression Types}
-
 \begin{itemize}
-\item XML Schema built-in atomic types \todo{link do appendixu}.
+\item XML Schema built-in atomic types. See Attachment \ref{FIG_xsd_built_in_atomic_types}.
 \item Types \texttt{ElementType}, \texttt{AttributeType}, \texttt{NodeType}, \texttt{TextNodeType},\linebreak \texttt{CommentType}, \texttt{ProcessingInstructionType}, \texttt{DocumentType} representing an element, attribute, node, text node, comment, processing instruction, document node.
@@
 
-To capture sequences, we assign the first two categories of types (all types except of \texttt{PathType} and \texttt{UnknownType}) with its cardinality as proposed in \todo{ref na Schajbalovu diplomku}.
+To capture sequences, we assign the first two categories of types (all types except of \texttt{PathType} and \texttt{UnknownType}) with its cardinality as proposed in \cite{thesis_schejbal}.
 Each of those types can be perceived as a sequence.
@@
 
-To ilustrate this difference, we introduce the following example.
+To illustrate this difference, we introduce the following example.
 Assuming $P$ is a return target path expression for variable $var$ and $flwor$ is a FLWOR expression with return clause $data(P)$, function call $getTargetReturnPathTypes(flwor, var)$ returns two $PathTypes$.

### 2012-03-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of \texttt{PathType} because then, we can determine a type of the function call more precisely.
-Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max} \todo{Doplnit}.
+Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max}, \texttt{avg}, \texttt{sum}, \texttt{distinct-values}.
 
@@
 Alike the functions, the global variables are defined in the prolog section.
-A type of a variable can be explicitly specified in its definition \todo{je to tak?
-priklad}.
+A type of a variable can be explicitly specified in its definition, for instance \texttt{declare variable \$x as xs:byte := 12;}.
 If it is not, it may often be deducible from the binding expression.
@@
 
-\todo[inline]{Overit si, ci ALG \ref{ALG_func_analysisOfExpressionTypes} pokryva vsetky pripady, co sa tyka definovania novych premennych a ich propagacie vyssim uzlom.}
-
-\todo[inline]{Pozor na to přetékání textu - na spoustě míst.}
-
 \begin{algorithm}
@@
 
-The output of this step is a set of couples ($pathType$, XML Schema built-in type).
-
-\todo[inline]{Zlucovanie odvodenych tvrdeni.}
-\todo[inline]{Path expressions s predikatmi.}
+The output of this step is a set of statements of form \\ an instance of \texttt{PathType} $\rightarrow$ XML Schema built-in atomic type.
 

### 2012-03-10 doc/theses/mikula/diplomka.tex
Master thesis: In progress.

@@
 \noindent
-\todo[inline]{Poďakovanie.}
+I would like to thank my supervisor RNDr. Irena Mlýnková, Ph.D., for her helpful advices, corrections and suggestions.
 
@@
 \chapwithtoc{Attachments}
+\begin{figure}
+\caption{XSD built-in atomic types.}
+\label{FIG_xsd_built_in_atomic_types}
+\includegraphics[scale=0.65]{xsd-types.eps}
+\end{figure}
+
+\begin{figure}
+\caption{All possible values representing an operator in an instance of \texttt{OperatorNode}}
+\label{FIG_operators}
+Logical: \texttt{AND}, \texttt{OR}.
+
+Comparison: \texttt{GEN\_EQUALS}, \texttt{GEN\_NOT\_EQUALS}, \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{VAL\_EQUALS}, \texttt{VAL\_NOT\_EQUALS}, \texttt{VAL\_LESS\_THAN}, \texttt{VAL\_LESS\_THAN\_EQUALS}, \texttt{VAL\_GREATER\_THAN}, \texttt{VAL\_GREATER\_THAN\_EQUALS}, \texttt{NOD\_IS}, \texttt{NOD\_PRECEDES}, \texttt{NOD\_FOLLOWS}.
+
+Range: \texttt{TO}.
+
+Additive: \texttt{PLUS}, \texttt{MINUS}.
+
+Multiplicative: \texttt{MUL}, \texttt{DIV}, \texttt{IDIV}, \texttt{MOD}.
+
+Set: \texttt{UNION}, \texttt{INTERSECTION}, \texttt{DIFFERENCE}.
+
+Type test: \texttt{INSTANCE\_OF}, \texttt{CASTABLE\_AS}.
+
+Type conversion: \texttt{TREAT\_AS}, \texttt{CAST\_AS}.
+
+Unary: \texttt{UNARY\_PLUS}, \texttt{UNARY\_MINUS}.
+\end{figure}
 

### 2012-03-10 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 In the following text, the set represented by expression $E$ is denoted $S$.
-Another requirement is that the expression has to be either a function call or an arithmetic operation \todo{Q: Nejake ine ako napr type constructor, cast as?
-Princip je stale rovnaky.
-A: ak to jen popište.
-To je naprosto OK.}.
+Another requirement is that the expression has to be either a function call or an arithmetic operation.
+As discussed in Chapter \ref{CHAPTER_analysis_of_xquery}, also other XQuery constructs can be utilized to infer built-in types, but, since the principle is similar, we focus on the two mentioned.
 
@@
 \subsection{Arithmetic Operations}
-If the operator is one of $+$, $-$, $*$, $div$, $mod$ (the class of the expression node is $Operator$ and it represents one of $PLUS$, $MINUS$, $MUL$, $DIV$, $MOD$ \todo{Niekam napisat mozne triedy operatora}), one operand is $E$ and the type $T$ of the other operand is one of numeric built-in types, then $T$ is also the inferred type of $S$.
+If the operator is one of $+$, $-$, $*$, $div$, $mod$ (the class of the expression node is $Operator$ and it represents one of $PLUS$, $MINUS$, $MUL$, $DIV$, $MOD$ (for all operators, see Attachment \ref{FIG_operators})), one operand is $E$ and the type $T$ of the other operand is one of numeric built-in types, then $T$ is also the inferred type of $S$.
 
@@
 Firstly, the syntax tree is searched for forms of FLWOR expressions to infer the keys from.
-Then, as described later, certain costructs are used to support or degrade the evidence of inferred statements in a final summarization.
+Then, as described later, certain constructs are used to support or degrade the evidence of inferred statements in a final summary.
 
@@
 \end{algorithm}
+
+\subsection{Definition of Keys}
+To define a key and a foreign key, we adopt the definitions introduced in the original approach \cite{Necasky:2009:DXK:1529282.1529414} with a modification that XPath paths are replaced by \texttt{PathType} instances.
+
+\begin{define}[Key]
+A key is a construct $(C, P, \{L\})$ where $C$, $P$ and $L$ are \texttt{PathType} instances without predicates and without special function calls that use only child and descendant axes.
+$C$ is called \emph{context path}, $P$ \emph{target path} and $L$ \emph{key path}.
+$C$ can be omitted, i.e.
+we can write $(P, \{L\})$.
+This is equivalent to $(/, P, \{L\})$.
+If $C$ is omitted we call the key \emph{global key}.
+Otherwise, it is called \emph{relative key}.
+\end{define}
+
+\begin{define}[Foreign key]
+A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$ where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are \texttt{PathType} instances without predicates and without special function calls that use only child and descendant axes.
+$C$ can be omitted as in the case of keys.
+\end{define}
 
@@
 
-As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inferrence of keys from occurrences of the join patterns, and rules (R1 - R5) to classify each occurrence into one of these cases.
+As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inference of keys from occurrences of the join patterns, and rules (R1 - R5) to classify each occurrence into one of these cases.
 For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
-\todo{Popisat dovod nizsej vahy (nie je take jasne, ze je to par key-foreign key), pripadne zmenit vahu.}
+The lesser weight is chosen because there is a lower probability that join of the join pattern 3 type is done via a key/foreign key pair.
 
@@
 
-\subsection{Inferrence of Keys from Join Pattern Occurrences}
-Now, when the join patterns occurrences are classified into (O1) and (O2) cases, we infer the key statements according to the original method \cite{Necasky:2009:DXK:1529282.1529414}.
-\todo[inline]{Popisat? To by som musel asi jedine skopirovat z clanku, to vlastnymi slovami nepopisem.}
+\subsection{Inference of Keys from Join Pattern Occurrences}
+Now, when the join patterns occurrences are classified into (O1) and (O2) cases, we infer the key statements according to the original approach \cite{Necasky:2009:DXK:1529282.1529414}.
+
+Let $w$ be the weight assigned to a pattern occurrence $\pi$. If $\pi$ is marked as (O1), the following statements with weight $w$ are inferred:
+\begin{itemize}
+\item $(P_2, \{L_2\})$ is not satisfied
+\item $(P_1, \{L_1\})$ is satisfied
+\item $(P_2, \{L_2\}) \rightarrow (P_1, \{L_1\})$ is satisfied
+\end{itemize}
+
+If $\pi$ is marked as (O2), the following statements with weight $w$ are inferred:
+\begin{itemize}
+\item $(P_2, \{L_2\})$ is satisfied
+\item $(P_1, \{L_1\}) \rightarrow (P_2, \{L_2\})$ is satisfied
+\end{itemize}
+
+The original approach \cite{Necasky:2009:DXK:1529282.1529414} describes also inference of relative keys. Refer to it for details.
 
@@
 The aim of this step is to eliminate or decrease the number of falsely inferred keys.
-\todo[inline]{Napisat niekam, ze odvodzovanie kludov predpoklada, ze kazdy element/element join je robeny cez key-foreign key pair.}
+A key can be inferred falsely when the assumption that every join is done via a key/foreign key pair is not correct for a particular join.
+
 When the algorithm discovers that a certain element is probably not unique, it assigns the element with a negative weight that will decrease the sum of its weights in the summarizing step.
@@
 \subsection{Normalization of Key Statements}
-\todo[inline]{Znovu pouzitie originalnej metody. Popisat?}
+The original approach \cite{Necasky:2009:DXK:1529282.1529414} employs the following normalization which we adopt. Is also incorporates a heuristic for precision enhancement. For details, refer to it.
+
+Let $K_1, \dots, K_n$ be the inferred keys.
+Let $S_i$ be the score of $K_i$ and $N_i$ be the number of the inferred statements about $K_i$.
+Let $S^{max}$ be the maximum from $|S_1|, \dots , |S_n|$ and $N^{max}$ be the maximum norm
+from $N_1, \dots , N_n$. The normalized score $S_i$ of $K_i$ is computed as $S^{norm}_i = \frac{S_i}{S^{max}} * (1 - \frac{N^{max} - N_i}{\sum_{i=1}^{n}N_i})$

### 2012-03-12 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

@@
 \chapter{Combination with Existing Methods of Inference}
-\todo[inline]{Kontrola pomocou dat.}
-
 In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference.

### 2012-03-12 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \chapter{Proposed Algorithm}
-\todo[inline]{Nejaky uvod a zkratke hrnutie}
-\todo[inline]{Predpoklad korektnosti dotazov}
+In this thesis we are developing an approach dealing with inference of XML schema, which input consists of two components; a set of XML documents and a set of XQuery queries related to the documents.
+The proposed algorithm introduced in this chapter describes the processing of XQuery queries only and a combination of its output with an existing method of XML schema inference (to produce a complete XML schema) is discussed in the next chapter.
+
+There is an important assumption placed upon the input XQuery queries saying that each query must be syntactically and semantically correct.
+That means that if a query was evaluated (using a XQuery processor meeting the XQuery specification), the evaluation would not raise any error, neither static nor dynamic.
+In other words, each query is syntactically correct and it correctly queries the data in the XML documents.
+
+The proposed algorithm consists of the following four main steps, described in detail in the rest of the chapter.
+
+\begin{enumerate}
+\item Construction of a syntax tree. We use lexical and syntax analyses proposed in \cite{thesis_schejbal} and for each XQuery on input, we construct a data structure defined in \ref{DEF_syntax_tree}.
+\item Static Analysis of expression types. The algorithm searches for expressions in the syntax trees and statically (without evaluation) determines their types.
+\item Inference of built-in types. When the types of expressions are determined, selected forms of expression are utilized to infer types of elements and attributes.
+\item Key discovery. The final step is an extension of approach \cite{Necasky:2009:DXK:1529282.1529414} inferring keys and foreign keys.
+\end{enumerate}
 
@@
 
-\todo[inline]{Niektore typy mozu byt take, ze pri prvom prechode ich nevieme zistit, ale ked uz mame zistene ine typy, tak dokazeme zistit aj tie, co sme pred tym nevedeli.
-Takze by bolo treba robit viac priechodov, ale skor nieco ako dopologicke triedenie, co zavisi na com, a potom typy zistovat v tomto poradi (mozno aj v druhom kole).}
-
 The analysis of expression types can be divided into three substeps.
@@
 Expressions of \texttt{UnknownType} are not utilized it the inference, therefore, their cardinality is not needed.
-
-\todo[inline]{Este by sa dala urobit analyza volania funkcii, do ktorych su predavane nejake elementy/atributy o ktorych vieme ich path expression.
-Vysledkom tejto funkcie by totiz zas mohli byt elementy/atributy o ktorych vieme (pozmenenu) path expression a nie len nezname elementy/atributy.
-A: Přesně tak. A jak říkám - není třeba vše co navrhnete implementovat.
-A2: Stačí to popsat a v kapitole Implementace říct, které ásti algoritmu implementujeme. Stačí takové, které už ukáží zajímavé chování a dokážou přínosy návrhu. Přesně k tomu implmentace je.}
 
@@
 And, if the left-most subnode extends the variable context for the following nodes in the numbering, it can be easily handled because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
-Every node ''knows'' wheather or not it may define new variables for nodes with higher sequence numbers.
+Every node ''knows'' whether or not it may define new variables for nodes with higher sequence numbers.
 This is expressed in the algorithm by the condition stating at line 10.
@@
 
+\subsection{Possible Extensions}
+As mentioned, the proposed algorithm utilizes only a small portion of XQuery constructs which can be be possibly utilized.
+Also, the algorithm does not perform the inference of types inside user-defined functions (it just uses the return types determined in earlier phases).
+That can be easily done as we know the definitions of functions and types of their real arguments.
+An algorithm performing the analysis in the user-defined functions can works as follows.
+
+If a currently processed expression is a function call of an user-defined function and at least one of its real arguments is an instance of \texttt{PathType}, get the function declaration node.
+In the function body, replace the formal arguments by the real arguments and set them as the local variables, so the proposed algorithms can be applied to process the function's body.
+Analyse the types of expressions in the body with the new information on the types of arguments by incorporating of the proposed algorithms and replace the function's return type by the one currently determined.
+
+A little complication is a recursion.
+A simple algorithm of that kind may never end because of the infinite analysis of the same recursive function(s).
+Therefore, we shall keep track of the currently analysed functions (e.g.
+in a stack of function calls) and do not process the recursive calls.
+
 \todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme. -- To urobim pocas implementacie pre nejaky vzorovy dotaz.}

### 2012-03-12 doc/theses/mikula/appendix_a.tex
Master thesis: In progress.

--- 2012-03-10 doc/theses/mikula/diplomka.tex
@@
-%%% Hlavní soubor. Zde se definují základní parametry a odkazuje se na ostatní části. %%%
-
-%% Verze pro jednostranný tisk:
-% Okraje: levý 40mm, pravý 25mm, horní a dolní 25mm
-% (ale pozor, LaTeX si sám přidává 1in)
-\documentclass[12pt,a4paper]{report}
-\setlength\textwidth{145mm}
-\setlength\textheight{247mm}
-\setlength\oddsidemargin{15mm}
-\setlength\evensidemargin{15mm}
-\setlength\topmargin{0mm}
-\setlength\headsep{0mm}
-\setlength\headheight{0mm}
-% \openright zařídí, aby následující text začínal na pravé straně knihy
-\let\openright=\clearpage
-
-\renewcommand\baselinestretch{1.3} % riadkovanie jeden a pol
-
-%% Pokud tiskneme oboustranně:
-% \documentclass[12pt,a4paper,twoside,openright]{report}
-% \setlength\textwidth{145mm}
-% \setlength\textheight{247mm}
-% \setlength\oddsidemargin{15mm}
-% \setlength\evensidemargin{0mm}
-% \setlength\topmargin{0mm}
-% \setlength\headsep{0mm}
-% \setlength\headheight{0mm}
-% \let\openright=\cleardoublepage
-
-%% Pokud pouľíváte csLaTeX (doporučeno):
-%\usepackage{czech}
-%% Pokud nikoliv:
-%\usepackage[czech]{babel}
-%\usepackage[T1]{fontenc}
-
-%% Pouľité kódování znaků: obvykle latin2, cp1250 nebo utf8:
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-
-%% Ostatní balíčky
-\usepackage{graphicx}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{alltt}
-\usepackage{todonotes}
-\usepackage{algorithmic}
-\usepackage[chapter]{algorithm}
-\usepackage{enumerate}
-\usepackage{multirow}
-\usepackage{threeparttable}
-\usepackage{listings}
-\usepackage{algorithmic}
-\usepackage[chapter]{algorithm}
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-
-\renewcommand{\algorithmiccomment}[1]{// #1}
-
-% pekne pokope definujeme potrebne udaje
-\def\mftitle{Inference of an XML Schema with the Knowledge of XML Operations}
-\def\mfthesistype{Master Thesis}
-\def\mfauthor{Mário Mikula}
-\def\mfadvisor{RNDr. Irena Mlýnková, Ph.D.}
-\def\mfplacedate{Praha, 2011}
-
-%% Balíček hyperref, kterým jdou vyrábět klikací odkazy v PDF,
-%% ale hlavně ho pouľíváme k uloľení metadat do PDF (včetně obsahu).
-%% POZOR, nezapomeňte vyplnit jméno práce a autora.
-\usepackage[ps2pdf,unicode]{hyperref}   % Musí být za vąemi ostatními balíčky
-\hypersetup{pdftitle=\mftitle}
-\hypersetup{pdfauthor=\mfauthor}
-
-%%% Drobné úpravy stylu
-
-% Tato makra přesvědčují mírně oąklivým trikem LaTeX, aby hlavičky kapitol
-% sázel příčetněji a nevynechával nad nimi spoustu místa. Směle ignorujte.
-%\makeatletter
-%\def\@makechapterhead#1{
-%  {\parindent \z@ \raggedright \normalfont
-%   \Huge\bfseries \thechapter. #1
-%   \par\nobreak
-%   \vskip 20\p@
-%}}
-%\def\@makeschapterhead#1{
-%  {\parindent \z@ \raggedright \normalfont
-%   \Huge\bfseries #1
-%   \par\nobreak
-%   \vskip 20\p@
-%}}
-%\makeatother
-
-% Toto makro definuje kapitolu, která není očíslovaná, ale je uvedena v obsahu.
-\def\chapwithtoc#1{
-\chapter*{#1}
-\addcontentsline{toc}{chapter}{#1}
-}
-
-% Theorem environment
-\newenvironment{definition}[1][Definition]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-
-\begin{document}
-
-% Trochu volnějąí nastavení dělení slov, neľ je default.
-\lefthyphenmin=2
-\righthyphenmin=2
-
-%%% Titulní strana práce
-
-\pagestyle{empty}
-\begin{center}
-
-\large
-
-Charles University in Prague
-
-\medskip
-
-Faculty of Mathematics and Physics
-
-\vfill
-
-{\bf\Large \mfthesistype}
-
-\vfill
-
-\centerline{\mbox{\includegraphics[width=60mm]{logo.eps}}}
-
-\vfill
-\vspace{5mm}
-
-{\LARGE \mfauthor}
-
-\vspace{15mm}
-
-% Název práce přesně podle zadání
-{\LARGE\bfseries \mftitle}
-
-\vfill
-
-% Název katedry nebo ústavu, kde byla práce oficiálně zadána
-% (dle Organizační struktury MFF UK)
-Department of Software Engineering
-
-\vfill
-
-\begin{tabular}{rl}
-
-Supervisor of the master thesis: & \mfadvisor \\
-\noalign{\vspace{2mm}}
-Study programme: & Informatika \\
-\noalign{\vspace{2mm}}
-Specialization: & ISS \\
-\end{tabular}
-
-\vfill
-
-% Zde doplňte rok
-\mfplacedate
-
-\end{center}
-
-\newpage
-
-%%% Následuje vevázaný list -- kopie podepsaného "Zadání diplomové práce".
-%%% Toto zadání NENÍ součástí elektronické verze práce, nescanovat.
-
-%%% Na tomto místě mohou být napsána případná poděkování (vedoucímu práce,
-%%% konzultantovi, tomu, kdo zapůjčil software, literaturu apod.)
-
-\openright
-
-\noindent
-I would like to thank my supervisor RNDr. Irena Mlýnková, Ph.D., for her helpful advices, corrections and suggestions.
-
-\newpage
-
-%%% Strana s čestným prohláąením k diplomové práci
-
-\vglue 0pt plus 1fill
-
-\noindent
-I declare that I carried out this master thesis independently, and only with the cited
-sources, literature and other professional sources.
-
-\medskip\noindent
-I understand that my work relates to the rights and obligations under the Act No.
-121/2000 Coll., the Copyright Act, as amended, in particular the fact that the Charles
-University in Prague has the right to conclude a license agreement on the use of this
-work as a school work pursuant to Section 60 paragraph 1 of the Copyright Act.
-
-\vspace{10mm}
-
-\hbox{\hbox to 0.5\hsize{%
-In ........ date ............
-\hss}\hbox to 0.5\hsize{%
-Signature
-\hss}}
-
-\vspace{20mm}
-\newpage
-
-%%% Povinná informační strana diplomové práce
-
-\vbox to 0.5\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Název práce:
-\mftitle
-% přesně dle zadání
-
-Autor:
-\mfauthor
-
-Katedra:  % Případně Ústav:
-Název katedry či ústavu, kde byla práce oficiálně zadána
-% dle Organizační struktury MFF UK
-
-Vedoucí diplomové práce:
-\mfadvisor, pracoviště
-% dle Organizační struktury MFF UK, případně plný název pracoviątě mimo MFF UK
-
-Abstrakt:
-% abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce
-
-Klíčová slova:
-% 3 aľ 5 klíčových slov
-
-\vss}\nobreak\vbox to 0.49\vsize{
-\setlength\parindent{0mm}
-\setlength\parskip{5mm}
-
-Title:
-% přesný překlad názvu práce v angličtině
-
-Author:
-\mfauthor
-
-Department:
-Název katedry či ústavu, kde byla práce oficiálně zadána
-% dle Organizační struktury MFF UK v angličtině
-
-Supervisor:
-\mfadvisor, pracoviště
-% dle Organizační struktury MFF UK, případně plný název pracoviątě
-% mimo MFF UK v angličtině
-
-Abstract:
-% abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
-% zadání diplomové práce
-
-Keywords:
-% 3 aľ 5 klíčových slov v angličtině
-
-\vss}
-
-\newpage
-
-%%% Strana s automaticky generovaným obsahem diplomové práce. U matematických
-%%% prací je přípustné, aby seznam tabulek a zkratek, existují-li, byl umístěn
-%%% na začátku práce, místo na jejím konci.
-
-\openright
-\pagestyle{plain}
-\renewcommand\thepage{}
-\tableofcontents
-
-\newtheorem{theorem}{Theorem}
-\theoremstyle{definition}
-%\newtheorem*{define}{Definition}	% Definice nečíslujeme, proto "*"
-\newtheorem{define}{Definition}[chapter]
-\newtheorem{example}{Example}
-
-%%% Tabulky v diplomové práci, existují-li.
-\chapwithtoc{List of Tables}
-
-\newpage
-\renewcommand\thepage{\arabic{page}}
-\setcounter{page}{1}
-
-%%% Jednotlivé kapitoly práce jsou pro přehlednost uloľeny v samostatných souborech
-\include{introduction}
-\include{used_technologies_and_definitions}
-\include{analysis_of_recent_approaches}
-\include{analysis_of_xquery}
-\include{pre_creation_of_algorithm}
-\include{proposed_algorithm}
-\include{combination_with_existing_methods_of_inference}
-
-%%% Seznam pouľité literatury
-\nocite{*}
-\bibliographystyle{plain}
-\todo[inline]{Literatura tady určitě co nejpodrobněji - odkaz nestačí, za měsíc nemusí existovat. A nemělo by to přetékat okraj.}
-\todo[inline]{Literatura - jInfer modules documentation - Tady by neškodily odkazy na jednotlivé dokumenty.}
-\bibliography{literature}
-
-%%% Pouľité zkratky v diplomové práci, existují-li, včetně jejich vysvětlení.
-%\chapwithtoc{List of Abbreviations}
-%\todo[inline]{Není nutno.}
-
-%%% Přílohy k diplomové práci, existují-li (různé dodatky jako výpisy programů,
-%%% diagramy apod.). Kaľdá příloha musí být alespoň jednou odkazována z vlastního
-%%% textu práce. Přílohy se číslují.
-\chapwithtoc{Attachments}
-\begin{figure}
-\caption{XSD built-in atomic types.}
-\label{FIG_xsd_built_in_atomic_types}
-\includegraphics[scale=0.65]{xsd-types.eps}
-\end{figure}
-
+\chapter{Attachments}
 \begin{figure}
@@
 
-\openright
-\end{document}
+\begin{figure}
+\caption{XSD built-in atomic types.}
+\label{FIG_xsd_built_in_atomic_types}
+\includegraphics[scale=0.65]{xsd-types.eps}
+\end{figure}

### 2012-03-14 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 
-On the contrary, the grammar-inferring methods \cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236} are based on theoretical knowledge of automata and their results belong to a particular class of languages.
+On the contrary, the grammar-inferring methods \cite{ahonen, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236} are based on theoretical knowledge of automata and their results belong to a particular class of languages.
 Thus, these methods guarantee specific characteristics of the results.
@@
 To achieve better results, the authors propose a combination of several verified approaches.
-A set of all possible steps is created using \emph{k,h-context} \cite{Ahonen1996GeneratingGrammars} and \emph{s,k-string} \cite{Raman97thesk-strings, Wong03onstructural} methods.
+A set of all possible steps is created using \emph{k,h-context} \cite{ahonen} and \emph{s,k-string} \cite{Raman97thesk-strings, Wong03onstructural} methods.
 The optimal step is then selected employing the MDL principle \cite{Grünwald05atutorial, Garofalakis:2000:XSE:342009.335409}.
@@
     Method name and/or paper & Input & Output & Year \\ \hline \hline
-    \cite{Ahonen1996GeneratingGrammars} & XML documents & DTD & 1996 \\ \hline
+    \cite{ahonen} & XML documents & DTD & 1996 \\ \hline
     XTRACT \cite{Garofalakis:2000:XSE:342009.335409} & XML documents & DTD & 2000 \\ \hline
@@
     \cite{Necasky:2009:DXK:1529282.1529414} & XQuery queries & --\tnote{1} & 2009 \\ \hline
+    \cite{thesis_klempa} & XML documents, XML schema & DTD, XSD & 2011 \\ \hline
   \end{tabular}
-  \todo[inline]{doplnit}
   \begin{tablenotes}

### 2012-03-14 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \begin{algorithmic}[1]
-\REQUIRE syntaxTree: A reference to the root node of a syntax tree.
+\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
 \ENSURE A list of syntax tree nodes representing function declarations.
 
-\STATE prologNode $=$ null
-\FORALL{moduleChildNode $\in$ syntaxTree.getChildren()}
-    \IF{is(moduleChildNode, PrologNode)}
-        \STATE prologNode $=$ moduleChildNode
-    \ENDIF
-\ENDFOR
-
-\STATE functionDeclarationNodes $=$ an empty list
-\IF{prologNode $\neq$ null}
-    \FORALL{prologChildNode $\in$ prologNode.getChildren()}
-        \IF{is(prologChildNode, FunctionDeclNode)}
-            \STATE add prologChildNode to functionDeclarationNodes
+\STATE $prologNode :=$ null
+\FORALL{$moduleChildNode \in syntaxTree.getChildren()$}
+    \IF{$is(moduleChildNode,$ PrologNode$)$}
+        \STATE $prologNode := moduleChildNode$
+    \ENDIF
+\ENDFOR
+
+\STATE $functionDeclarationNodes :=$ an empty list
+\IF{$prologNode \neq$ null}
+    \FORALL{$prologChildNode \in prologNode.getChildren()$}
+        \IF{$is(prologChildNode,$ FunctionDeclNode$)$}
+            \STATE add $prologChildNode$ to $functionDeclarationNodes$
         \ENDIF
@@
 \begin{algorithmic}[1]
-\REQUIRE syntaxTree: A reference to the root node of a syntax tree.
+\REQUIRE $syntaxTree$: A reference to the root node of a syntax tree.
 \ENSURE An associative array of function names with their types and references to the their declaration nodes.
 
-\STATE functionArray $=$ an empty associative array
-\FORALL{functionDeclarationNode $\in$ \\ getFunctionDeclarationNodes(syntaxTree)}
-    \STATE functionName $=$ functionDeclarationNode.funcName
-    \STATE typeNode $=$ functionDeclarationNode.getChild(TypeNode)
-    \STATE functionArray[functionName] $=$ \\ (getTypeTN(typeNode), functionDeclarationNode)
-\ENDFOR
-\RETURN functionArray
+\STATE $functionArray :=$ an empty associative array
+\FORALL{$functionDeclarationNode \in$ \\ $getFunctionDeclarationNodes(syntaxTree)$}
+    \STATE $functionName := functionDeclarationNode.funcName$
+    \STATE $typeNode := functionDeclarationNode.getChild(TypeNode)$
+    \STATE $functionArray[functionName] :=$ \\ $(getTypeTN(typeNode), functionDeclarationNode)$
+\ENDFOR
+\RETURN $functionArray$
 \end{algorithmic}
@@
 \begin{algorithmic}[1]
-\REQUIRE typeNode: Syntax tree node of TypeNode class.
-\ENSURE Type extracted from typeNode.
-
-\STATE type = null
-\STATE cardinality = typeNode.cardinality
-\STATE itemTypeNode = typeNode.getChild(ItemTypeNode)
-\IF{is(itemTypeNode, AtomicTypeNode)}
-    \STATE type = itemTypeNode.typeName
-\ELSIF{is(itemTypeNode, KindTestNode)}
-    \STATE type = itemTypeNode.nodeKind
-\ENDIF
-
-\RETURN type, cardinality
+\REQUIRE $typeNode$: Syntax tree node of TypeNode class.
+\ENSURE Type extracted from $typeNode$.
+
+\STATE $type :=$ null
+\STATE $cardinality := typeNode.cardinality$
+\STATE $itemTypeNode := typeNode.getChild(ItemTypeNode)$
+\IF{$is(itemTypeNode,$ AtomicTypeNode$)$}
+    \STATE $type := itemTypeNode.typeName$
+\ELSIF{$is(itemTypeNode,$ KindTestNode$)$}
+    \STATE $type := itemTypeNode.nodeKind$
+\ENDIF
+
+\RETURN $type, cardinality$
 \end{algorithmic}
@@
 \REQUIRE{\ \\
-    exprNode: An expression node in the syntax tree. \\
-    contextVarTypes: Types of local variables valid in the current subtree.
+    $exprNode$: An expression node in the syntax tree. \\
+    $contextVarTypes$: Types of local variables valid in the current subtree.
 }
@@
 
-\STATE type $=$ UnknownType
-\IF{is(exprNode, LiteralNode)}
-     \STATE type $=$ XSD atomic type exprNode.type, exactly-once cardinality
-\ELSIF{is(exprNode, FunctionCallNode)}
-     \STATE type $=$ getFunctionReturnType(exprNode.fncName)
-\ELSIF{is(exprNode, VarRefNode)}
-     \STATE type = getVariableType(contextVarTypes, exprNode.varName)
-     \IF{type is PathType}
-         \STATE step $=$ a path step (new instance of StepExprNode) representing the variable reference
-		 \STATE substeps = empty associative array         
-         \STATE substeps[step] $=$ type
-         \STATE pathExprNode $=$ new instance of PathExprNode containing one step step
-         \STATE type $=$ PathType containing pathExprNode and substeps
-     \ENDIF
-\ELSIF{is(exprNode, PathExprNode)}
-     \STATE type $=$ createPathType(exprNode)
-\ELSIF{is(exprNode, OperatorNode)}
-     \STATE type $=$ determineOperatorType(exprNode, contextVarTypes)
-\ELSIF{is(exprNode, FLWORExprNode)}
-     \STATE returnClauseNode $=$ exprNode.getChild(ReturnClauseNode)
-     \STATE type $=$ createForUnboundType(getType(returnClauseNode))
+\STATE $type :=$ UnknownType
+\IF{$is(exprNode,$ LiteralNode$)$}
+     \STATE $type :=$ XSD atomic type $exprNode.type$, exactly-once cardinality
+\ELSIF{$is(exprNode,$ FunctionCallNode$)$}
+     \STATE $type := getFunctionReturnType(exprNode.fncName)$
+\ELSIF{$is(exprNode,$ VarRefNode$)$}
+     \STATE $type := getVariableType(contextVarTypes, exprNode.varName)$
+     \IF{$type$ is PathType}
+         \STATE $step :=$ a path step (new instance of StepExprNode) representing the variable reference
+		 \STATE $substeps :=$ an empty associative array         
+         \STATE $substeps[step] := type$
+         \STATE $pathExprNode :=$ new instance of PathExprNode containing one step step
+         \STATE $type :=$ PathType containing $pathExprNode$ and $substeps$
+     \ENDIF
+\ELSIF{$is(exprNode,$ PathExprNode$)$}
+     \STATE $type := createPathType(exprNode)$
+\ELSIF{$is(exprNode,$ OperatorNode$)$}
+     \STATE $type := determineOperatorType(exprNode, contextVarTypes)$
+\ELSIF{$is(exprNode,$ FLWORExprNode$)$}
+     \STATE $returnClauseNode := exprNode.getChild(ReturnClauseNode)$
+     \STATE $type := createForUnboundType(getType(returnClauseNode))$
 \ENDIF
@@
 
-\STATE $type =$ UnknownType
-\STATE $operator =$ $operatorNode.operator$
-\IF{isOperatorClassComparison($operator$)}
-    \STATE $type =$ boolean, exactly-one cardinality
-\ELSIF{isOperatorClassAddition($operator$)}
-    \STATE $leftOperandType =$ getType($exprNode.leftSide$)
-    \STATE $rightOperandType =$ getType($exprNode.rightSide$)
-    \IF{isNumericType($leftOperandType$) $\wedge$ isNumericType($rightOperandType$)}
-        \STATE $type =$ selectMoreGeneralNumericType($leftOperandType$, $rightOperandType$)
-    \ELSIF{isNumericType($leftOperandType$)}
-        \STATE $type = leftOperandType$
-    \ELSIF{isNumericType($rightOperandType$)}
-        \STATE $type = rightOperandType$
+\STATE $type :=$ UnknownType
+\STATE $operator := operatorNode.operator$
+\IF{$isOperatorClassComparison(operator)$}
+    \STATE $type :=$ boolean, exactly-one cardinality
+\ELSIF{$isOperatorClassAddition(operator)$}
+    \STATE $leftOperandType := getType(exprNode.leftSide)$
+    \STATE $rightOperandType := getType(exprNode.rightSide)$
+    \IF{$isNumericType(leftOperandType)$ \\ $\wedge$ $isNumericType(rightOperandType)$}
+        \STATE $type := selectMoreGeneralNumericType(leftOperandType,$ $rightOperandType)$
+    \ELSIF{$isNumericType(leftOperandType)$}
+        \STATE $type := leftOperandType$
+    \ELSIF{$isNumericType(rightOperandType)$}
+        \STATE $type := rightOperandType$
     \ENDIF
@@
 \begin{algorithmic}[1]
-\REQUIRE $operator$: Representation of a XQuery operator.
-\ENSURE true if operator is a comparison operator, false otherwise.
+\REQUIRE $operator$: A representation of a XQuery operator.
+\ENSURE true if $operator$ is a comparison operator, false otherwise.
 
@@
 \begin{algorithm}
-\caption{Function isOperatorClassAddition()}
+\caption{Function \texttt{isOperatorClassAddition}}
 \label{ALG_func_isOperatorClassAddition}
 \begin{algorithmic}[1]
-\REQUIRE $operator$: Representation of a XQuery operator.
-\ENSURE $true$ if $operator$ is an addition operator, $false$ otherwise.
-
-\IF{$operator$ equals one of $PLUS$, $MINUS$, $UNARY\_PLUS$, $UNARY\_MINUS$}
-    \RETURN $true$
+\REQUIRE $operator$: A representation of a XQuery operator.
+\ENSURE true if $operator$ is an addition operator, false otherwise.
+
+\IF{$operator$ equals one of PLUS, MINUS, UNARY\_PLUS, UNARY\_MINUS}
+    \RETURN true
 \ELSE
-    \RETURN $false$
-\ENDIF
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Function isNumericType()}
+    \RETURN false
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function \texttt{isNumericType}}
 \label{ALG_func_isNumericType}
 \begin{algorithmic}[1]
-\REQUIRE $type$: Representation of a type.
-\ENSURE $true$ if $type$ represents one of the XSD atomic built-in numeric types, $false$ otherwise.
-
-\IF{$type$ represents one of $float$, $double$, $decimal$, $integer$, $long$, $int$, $short$, $byte$, $nonPositiveInteger$, $negativeInteger$, $nonNegativeInteger$, $positiveInteger$, $unsignedLong$, $unsignedInt$, $unsignedShort$, $unsignedByte$}
-    \RETURN $true$
+\REQUIRE $type$: A representation of a type.
+\ENSURE true if $type$ represents one of the XSD atomic built-in numeric types, false otherwise.
+
+\IF{$type$ represents one of float, double, decimal, integer, long, int, short, byte, nonPositiveInteger, negativeInteger, nonNegativeInteger, positiveInteger, unsignedLong, unsignedInt, unsignedShort, unsignedByte}
+    \RETURN true
 \ELSE
-    \RETURN $false$
+    \RETURN false
 \ENDIF
@@
 \begin{algorithm}
-\caption{Function analysisOfExpressionTypes}
+\caption{Function \texttt{analysisOfExpressionTypes}}
 \label{ALG_func_analysisOfExpressionTypes}
@@
 \FORALL{$i \in \{1, \dots, |startingNode.getChildren()|\}$ ordered from the lowest to the highest}
-    \STATE $subnode = startingNode.getChild(i)$
+    \STATE $subnode := startingNode.getChild(i)$
     \STATE $analysisOfExpressionTypes(subnode, contextVarTypes)$ \COMMENT{recursion}
-    \STATE $newVars = get(subnode, newVars)$    
-    \STATE $contextVarTypes = mergeContextVarTypes(contextVarTypes, newVars)$
-\ENDFOR
-\IF{$is(startingNode, ExprNode)$}
+    \STATE $newVars := get(subnode, newVars)$    
+    \STATE $contextVarTypes := mergeContextVarTypes(contextVarTypes, newVars)$
+\ENDFOR
+\IF{$is(startingNode,$ ExprNode$)$}
     \STATE $memorizeType(startingNode, determineExpressionType(startingNode))$
 \ENDIF
-\IF{$is(startingNode, VariableBindingNode)$}
-    \STATE $type = null$
-    \STATE $typeNode = startingNode.getChild(TypeNode)$
+\IF{$is(startingNode,$ VariableBindingNode$)$}
+    \STATE $type :=$ null
+    \STATE $typeNode := startingNode.getChild(TypeNode)$
     \IF{$typeNode \neq null$}
-        \STATE $type = getTypeTN(typeNode)$
+        \STATE $type := getTypeTN(typeNode)$
     \ELSE
-        \STATE $type = determineExpressionType(startingNode.$ $getChild(BindingSequenceNode).getChild(ExprNode),$ $contextVarTypes)$
-        \IF{$is(startingNode, ForClauseNode)$}
-            \STATE $type = createForBoundType(type)$
+        \STATE $type := determineExpressionType(startingNode.$ $getChild(BindingSequenceNode).getChild(ExprNode),$ $contextVarTypes)$
+        \IF{$is(startingNode,$ ForClauseNode$)$}
+            \STATE $type := createForBoundType(type)$
         \ENDIF
@@
     \STATE $set(startingNode, newVars, (startingNode.varName, type))$
-\ELSIF{$is(startingNode, TupleStreamNode)$}
+\ELSIF{$is(startingNode,$ TupleStreamNode$)$}
     \FORALL{$varBindingNode \in startingNode.getChildren()$}
@@
 \begin{algorithm}
-\caption{Function mergeContextVarTypes}
+\caption{Function \texttt{mergeContextVarTypes}}
 \label{ALG_function_mergeContextVarTypes}
@@
 \REQUIRE{\ \\
-    $keys()$: Function retrieving index values of an indexed structure. \\
     $contextVarTypes$: Types of local variables valid in current context. \\    
@@
 \FORALL{$varName \in keys(extendingVarTypes)$}
-    \STATE $contextVarTypes[varName] = extendingVarTypes[varName]$
+    \STATE $contextVarTypes[varName] := extendingVarTypes[varName]$
 \ENDFOR
@@
 \begin{algorithm}
-\caption{Function getVariableType}
+\caption{Function \texttt{getVariableType}}
 \label{ALG_function_get_variable_type}
@@
 \begin{algorithm}
-\caption{Function createPathType}
+\caption{Function \texttt{createPathType}}
 \label{ALG_function_create_path_type}
@@
 
-\STATE $stepNodes =$ an empty array
-\STATE $forBoundSubsteps =$ an empty associative array
+\STATE $stepNodes :=$ an empty array
+\STATE $forBoundSubsteps :=$ an empty associative array
 \FORALL{$stepNode \in pathExprNode.getChildren()$}
-	\STATE $detailNode = stepNode.getChild(ExprNode)$
-	\IF{$detailNode != null$}
-		\IF{$is(detailNode, VarRefNode)$}
-			\STATE $type = getType(detailNode)$ // $PathType$
+	\STATE $detailNode := stepNode.getChild(ExprNode)$
+	\IF{$detailNode \neq$ null}
+		\IF{$is(detailNode,$ VarRefNode$)$}
+			\STATE $type := getType(detailNode)$ \COMMENT{PathType}
 			\IF{$type.isForBound$}
-				\STATE $forBoundSubsteps[stepNode] = type$
+				\STATE $forBoundSubsteps[stepNode] := type$
 				\STATE add $stepNode$ to $stepNodes$
@@
 
-\RETURN $PathType$ with $stepNodes$, $forBoundSubsteps$, and $isForBound$ set to $false$
+\RETURN $PathType$ with $stepNodes$, $forBoundSubsteps$, and $isForBound$ set to false
 \end{algorithmic}
@@
 \begin{algorithm}
-\caption{Function createForBoundType}
+\caption{Function \texttt{createForBoundType}}
 \label{ALG_function_create_for_bound_type}
@@
 
-\IF{$type$ is $UnknownType$}
-	\RETURN $UnknownType$
-\ELSIF{$type$ is $NodeType$ or XSD atomic type}
+\IF{$type$ is UnknownType}
+	\RETURN UnknownType
+\ELSIF{$type$ is NodeType or XSD atomic type}
 	\RETURN $type$ with cardinality set to exactly-one
 \ELSE
-	\STATE // $PathType$
-	\RETURN $type$ with $isForBound$ flag set to $true$
+	\STATE \COMMENT {It is PathType}
+	\RETURN $type$ with $isForBound$ flag set to true
 \ENDIF
@@
 \begin{algorithm}
-\caption{Function createForUnboundType}
+\caption{Function \texttt{createForUnboundType}}
 \label{ALG_function_create_for_unbound_type}
@@
 
-\IF{$type$ is $UnknownType$}
-	\RETURN $UnknownType$
-\ELSIF{$type$ is $NodeType$ or XSD atomic type}
+\IF{$type$ is UnknownType}
+	\RETURN UnknownType
+\ELSIF{$type$ is NodeType or XSD atomic type}
 	\RETURN $type$ with cardinality set to zero-or-more
 \ELSE
-	\STATE // $PathType$
-	\RETURN $type$ with $isForBound$ flag set to $false$
+	\STATE \COMMENT {It is PathType}
+	\RETURN $type$ with $isForBound$ flag set to false
 \ENDIF
@@
 \begin{algorithm}
-\caption{Function usesOnlyChildAndDescendantAxes}
+\caption{Function \texttt{usesOnlyChildAndDescendantAxes}}
 \label{ALG_function_uses_only_child_and_descendant_axes}
 \begin{algorithmic}[1]
-\REQUIRE $pathType$: A $PathType$ instance.
+\REQUIRE $pathType$: A PathType instance.
 \ENSURE A boolean result.
@@
 	\IF{$stepNode.isAxisStep$}
-		\STATE $axisKind$ = $stepNode.axisNode.getAxisKind()$
-		\IF{$axisKind$ $\neq$ $CHILD$ AND $axisKind$ $\neq$ $DESCENDANT$}
-			\RETURN false
-		\ENDIF
-	\ENDIF
-\ENDFOR
-
-\STATE $detailNode$ = $stepNode.getChild(ExprNode)$
+		\STATE $axisKind := stepNode.axisNode.getAxisKind()$
+		\IF{$axisKind \neq$ CHILD $\wedge$ $axisKind \neq$ DESCENDANT}
+			\RETURN false
+		\ENDIF
+	\ENDIF
+\ENDFOR
+
+\STATE $detailNode := stepNode.getChild(ExprNode)$
 \IF{$detailNode \neq$ null}
-	\IF{$is(detailNode, VarRefNode)$}
-		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode])$ == false}
+	\IF{$is(detailNode,$ VarRefNode$)$}
+		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) =$ false}
 			\RETURN false
@@
 \begin{algorithm}
-\caption{Function isWithoutPredicates}
+\caption{Function \texttt{isWithoutPredicates}}
 \label{ALG_function_is_without_predicates}
 \begin{algorithmic}[1]
-\REQUIRE $pathType$: A $PathType$ instance.
-\ENSURE A boolean result.getTargetReturnPathTypesRecursive
+\REQUIRE $pathType$: A PathType instance.
+\ENSURE A boolean result.
 
 \FORALL{$stepNode \in pathType.stepNodes$}
-	\IF{$stepNode.hasPredicates()$}
-		\RETURN false
-	\ENDIF
-	\IF{$pathType.forBoundSubsteps[stepNode]$ $\neq$ null}
-		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode]$ == false}
+	\IF{$stepNode.hasPredicates$}
+		\RETURN false
+	\ENDIF
+	\IF{$pathType.forBoundSubsteps[stepNode] \neq$ null}
+		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode] =$ false}
 			\RETURN false
@@
 \begin{algorithm}
-\caption{Function endsWithExactlyOnePredicate}
+\caption{Function \texttt{endsWithExactlyOnePredicate}}
 \label{ALG_function_ends_with_exactly_one_predicate}
 \begin{algorithmic}[1]
-\REQUIRE $pathType$: A $PathType$ instance.
+\REQUIRE $pathType$: A PathType instance.
 \ENSURE If $pathType$ ends with exactly one predicate, return value is the predicate, otherwise return value is false.
 
-\STATE $lastStepNode$ = last item from $pathType.stepNodes$
-\IF{$lastStepNode.hasPredicates()$ $==$ false}
-	\RETURN false
-\ENDIF
-\IF{number of item in list $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
+\STATE $lastStepNode :=$ the last item from $pathType.stepNodes$
+\IF{$lastStepNode.hasPredicates() =$ false}
+	\RETURN false
+\ENDIF
+\IF{number of items in list $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
 	\RETURN false
@@
 \begin{algorithm}
-\caption{Function isTargetPath}
+\caption{Function \texttt{isTargetPath}}
 \label{ALG_function_is_target_path}
@@
 \REQUIRE{\ \\
-  $pathType$: A $PathType$ instance. \\
+  $pathType$: A PathType instance. \\
   $varName$: A variable name.
 }
-\ENSURE Returns true, if the first step of path represented by the $PathType$ instance is a reference to the specified variable.
-
-\todo[inline]{Upravit styl vsetkych algoritmov (null, ==, !=, true/false, ...) tak ako je tu.}
-
-\STATE $firstStep$ = first item from  $pathType.stepNodes$
-\STATE $detailNode$ = $firstStep.detailNode$
-\IF{$detailNode$ == null}
-	\RETURN false
-\ENDIF
-\IF{$is(detailNode, VarRefNode)$ == false}
-	\RETURN false
-\ENDIF
-\IF{$detailNode.varName$ == $varName$}
+\ENSURE Returns true, if the first step of path represented by the PathType instance is a reference to the specified variable.
+
+\STATE $firstStep :=$ first item from  $pathType.stepNodes$
+\STATE $detailNode := firstStep.detailNode$
+\IF{$detailNode =$ null}
+	\RETURN false
+\ENDIF
+\IF{$is(detailNode, VarRefNode) =$ false}
+	\RETURN false
+\ENDIF
+\IF{$detailNode.varName = varName$}
 	\RETURN true
@@
 \begin{algorithm}
-\caption{Function getTargetReturnPathTypes}
+\caption{Function \texttt{getTargetReturnPathTypes}}
 \label{ALG_function_get_target_return_path_types}
@@
 \REQUIRE{\ \\
-  $flworNode$: A $FLWORExprNode$ instance to search for target return paths. \\
+  $flworNode$: A FLWORExprNode instance to search for target return paths. \\
   $varName$: A variable name.
 }
-\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by $PathType$ instances.
-
-\STATE $pathTypes$ = an empty array
+\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by PathType instances.
+
+\STATE $pathTypes :=$ an empty array
 \STATE $getTargetReturnPathTypesRecursive(flworNode.$ $getChild(ReturnClauseNode), varName, pathTypes)$
@@
 \begin{algorithm}
-\caption{Function getTargetReturnPathTypesRecursive}
+\caption{Function \texttt{getTargetReturnPathTypesRecursive}}
 \label{ALG_function_get_target_return_path_types_recursive}
@@
 
-\IF{$is(node, ExprNode)$}
-	\STATE $type$ = $getType(node)$
-	\IF{$type$ is $PathType$ AND $isTargetPath(type, varName)$}
+\IF{$is(node,$ ExprNode$)$}
+	\STATE $type := getType(node)$
+	\IF{$type$ is $PathType$ $\wedge$ $isTargetPath(type, varName)$}
 		\STATE add $type$ to $pathTypes$
@@
 \begin{algorithm}
-\caption{Function getTargetReturnPaths}
+\caption{Function \texttt{getTargetReturnPaths}}
 \label{ALG_function_get_target_return_paths}
@@
 \REQUIRE{\ \\
-  $flworNode$: A $FLWORExprNode$ instance to search for target return paths. \\
+  $flworNode$: A FLWORExprNode instance to search for target return paths. \\
   $varName$: A variable name.
 }
-\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by $PartExprNode$ instances.
-
-\STATE $paths$ = an empty array
+\ENSURE A list of target return paths for the specified variable, in the given FLWOR, represented by PathExprNode instances.
+
+\STATE $paths :=$ an empty array
 \STATE $getTargetReturnPathsRecursive(flworNode.$ $getChild(ReturnClauseNode), varName, paths)$
@@
 \begin{algorithm}
-\caption{Function getTargetReturnPathsRecursive}
+\caption{Function \texttt{getTargetReturnPathsRecursive}}
 \label{ALG_function_get_target_return_paths_recursive}
@@
 
-\IF{$is(node, PathExprNode)$}
-	\STATE $type$ = $getType(node)$
-	\IF{$type$ is $PathType$ AND $isTargetPath(type, varName)$}
+\IF{$is(node,$ PathExprNode$)$}
+	\STATE $type := getType(node)$
+	\IF{$type$ is PathType $\wedge$ $isTargetPath(type, varName)$}
 		\STATE add $type$ to $paths$
 	\ENDIF
-\ELSIF{$is(node, VarRefNode)$}
-	\IF{$node.varName$ == $varName$}
-		\STATE $type$ = $getType(node)$
-		\IF {$type$ is $PathType$ AND $isTargetPath(type, varName)$}
+\ELSIF{$is(node,$ VarRefNode$)$}
+	\IF{$node.varName = varName$}
+		\STATE $type := getType(node)$
+		\IF {$type$ is PathType $\wedge$ $isTargetPath(type, varName)$}
 			\STATE add $type$ to $paths$
@@
 
-\STATE $bindingNodes = bindingNode \in flworNode.getChild(TupleStreamNode).$ $getChildren()$
-
-\STATE $whereClause = flworNode.getChild(WhereClauseNode)$
-\STATE $checkJoinPattern3 = false$
-\STATE $whereExpr = null$
-\IF{$whereClause != null$}
-    \STATE $whereExpr = whereClause.getChild(ExprNode)$
-    \IF{$is(whereExpr, OperatorNode)$ AND $whereExpr.operator == GEN\_EQUALS$}
-        \STATE $checkJoinPattern3 = true$
+\STATE $bindingNodes := bindingNode \in flworNode.getChild(TupleStreamNode).$ $getChildren()$
+
+\STATE $whereClause := flworNode.getChild(WhereClauseNode)$
+\STATE $checkJoinPattern3 :=$ false
+\STATE $whereExpr :=$ null
+\IF{$whereClause \neq$ null}
+    \STATE $whereExpr := whereClause.getChild(ExprNode)$
+    \IF{$is(whereExpr,$ OperatorNode$)$ $\wedge$ $whereExpr.operator =$ GEN\_EQUALS}
+        \STATE $checkJoinPattern3 :=$ true
     \ENDIF
@@
 \FORALL{$bindingNode \in bindingNodes$}
-    \STATE $bindingExpr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-    \STATE $type = getType(bindingExpr)$
-    \IF{$type$ is $PathType$ $AND$ $usesOnlyChildAndDescendantAxes(type)$ $AND$ $isWithoutPredicatesExceptLastStep(type)$}
-    	\STATE $P = endsWithExactlyOnePredicate(type)$
+    \STATE $bindingExpr := bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
+    \STATE $type := getType(bindingExpr)$
+    \IF{$type$ is PathType $\wedge$ $usesOnlyChildAndDescendantAxes(type)$ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
+    	\STATE $P := endsWithExactlyOnePredicate(type)$
     	\IF{$P$}
         	\FORALL{$(var, node) \in forVars$}
-            	\STATE $forVars = determinateJoinPatterns(bindingNode, P, node, var,$ $forVars,$ $checkJoinPattern3,$ $whereExpr)$
+            	\STATE $forVars := determinateJoinPatterns(bindingNode, P, node, var,$ $forVars,$ $checkJoinPattern3,$ $whereExpr)$
         	\ENDFOR
         \ELSIF{$isWithoutPredicates(type)$}
-        	\IF{$is(bindingNode, ForClauseNode$}
-            	\STATE add $(bindingNode.varName, bindingNode)$ to forVars
+        	\IF{$is(bindingNode,$ ForClauseNode$)$}
+            	\STATE add $(bindingNode.varName, bindingNode)$ to $forVars$
   	       	\ENDIF
@@
 \begin{algorithm}
-\caption{Function determinateJoinPatterns}
+\caption{Function \texttt{determinateJoinPatterns}}
 \label{ALG_func_determinateJoinPatterns}
@@
 \IF{$P$ is of form $curBindingNode.varName/L_1 = var/L_2$}
-    \IF{$is(curBindingNode, ForClauseNode)$}
+    \IF{$is(curBindingNode,$ ForClauseNode$)$}
         \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for join pattern
         \STATE remove $(forVar, forBindingNode)$ from $forVars$
-    \ELSIF{$is(curBindingNode, LetClauseNode)$}
+    \ELSIF{$is(curBindingNode,$ LetClauseNode$)$}
         \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for let pattern                
@@
     \ENDIF
-\ELSIF{$checkJoinPattern3$ AND $is(curBindingNode, ForClauseNode)$}
+\ELSIF{$checkJoinPattern3$ AND $is(curBindingNode,$ ForClauseNode$)$}
     \IF{$whereExpr$ is of form $bindingNode.varName/L_1 = var/L_2$}
@@
     \ELSE
-        \STATE $secondBindingNode =$ the second binding node from $jpOccurrence$
-        \STATE $flworNode = secondBindingNode.getParent().getParent()$
-        \STATE $secondVarName = secondBindingNode.varName$
-        \STATE $returnPathTypes = getReturnPathTypes(flworNode,$ $secondVarName)$
+        \STATE $secondBindingNode :=$ the second binding node from $jpOccurrence$
+        \STATE $flworNode := secondBindingNode.getParent().getParent()$
+        \STATE $secondVarName := secondBindingNode.varName$
+        \STATE $returnPathTypes := getReturnPathTypes(flworNode,$ $secondVarName)$
 		\IF{$checkR2(returnPathTypes, jpOccurrence)$}
@@
 		\ENDIF
-		\STATE $targetReturnPaths = getTargetReturnPaths(flworNode,$ $secontVarName)$
+		\STATE $targetReturnPaths := getTargetReturnPaths(flworNode,$ $secontVarName)$
 		\STATE $useR4R5(returnPaths, jpOccurrence)$
@@
 \begin{algorithm}
-\caption{Function checkR2}
+\caption{Function \texttt{checkR2}}
 \label{ALG_function_check_r2}
@@
 \FORALL{$returnPathType \in returnPathTypes$}
-	\IF{$returnPathType.specialFunctionCalls$ contains one of $min$, $max$, $sum$, $avg$}
-    	\STATE mark $jpOccurrence$ as case (O1), weight 1 // R2
+	\IF{$returnPathType.specialFunctionCalls$ contains one of 'min', 'max', 'sum', 'avg'}
+    	\STATE mark $jpOccurrence$ as case (O1), weight 1 \COMMENT {R2}
   		\RETURN true
@@
 \begin{algorithm}
-\caption{Function checkR3}
+\caption{Function \texttt{checkR3}}
 \label{ALG_function_check_r3}
@@
     \IF{$returnPathType.specialFunctionCalls$ contains $count$}
-        \STATE mark $jpOccurrence$ as case (O1), weight 0.75 // R3
+        \STATE mark $jpOccurrence$ as case (O1), weight 0.75 \COMMENT {R3}
         \RETURN true
@@
 \begin{algorithm}
-\caption{Function useR4R5}
+\caption{Function \texttt{useR4R5}}
 \label{ALG_function_use_r4_r5}
@@
 
-\STATE $retPathsNumber = $ number of paths in $targetReturnPaths$
+\STATE $retPathsNumber :=$ a number of paths in $targetReturnPaths$
 \IF{$retPathsNumber > 1$}
-	\STATE mark $jpOccurrence$ as case (O2), weight 1 // R4
+	\STATE mark $jpOccurrence$ as case (O2), weight 1 \COMMENT {R4}
 \ELSE
-	\STATE mark $jpOccurrence$ as case (O2), weight 0.5 // R5
+	\STATE mark $jpOccurrence$ as case (O2), weight 0.5 // \COMMENT {R5}
 \ENDIF
@@
 
-\IF{$is(node, FunctionCallNode)$}
-    \IF{$node.fncName$ == $distinct-values$}
-        \STATE $argument = node.getChild(0)$
-        \STATE $type = getType(argument)$
-        \IF{$type$ is $PathType$ AND $usesOnlyChildAndDescendantAxes(type)$ AND $isWithoutPredicatesExceptLastStep(type)$}
+\IF{$is(node$, FunctionCallNode$)$}
+    \IF{$node.fncName =$ 'distinct-values'}
+        \STATE $argument := node.getChild(0)$
+        \STATE $type := getType(argument)$
+        \IF{$type$ is PathType $\wedge$ $usesOnlyChildAndDescendantAxes(type)$ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
             \STATE memorize $argument$ as the negative key statement, weight 1
@@
 
-\IF{$is(node, FLWORExprNode)$}
-	\STATE $forVars =$ an empty array
+\IF{$is(node,$ FLWORExprNode$)$}
+	\STATE $forVars :=$ an empty array
 
     \FORALL{$bindingNode \in node.getChild(TupleStreamNode).getChildren()$}
-        \IF{$is(bindingNode, ForClauseNode)$}
-            \STATE $expr = bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-            \IF{$getType(expr)$ is $PathType$ AND $usesOnlyChildAndDescendantAxes(getType(expr))$}
+        \IF{$is(bindingNode,$ ForClauseNode$)$}
+            \STATE $expr := bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
+            \IF{$getType(expr)$ is PathType $\wedge$ $usesOnlyChildAndDescendantAxes(getType(expr))$}
                 \STATE add $bindingNode.varName$ to $forVars$
@@
 
-    \STATE $whereClause = node.getChild(WhereClauseNode)$
-	\IF{$whereClause != null$}
+    \STATE $whereClause := node.getChild(WhereClauseNode)$
+	\IF{$whereClause \neq$ null}
     	\STATE $processWhere(forVars, whereClause.getChild(ExprNode))$
@@
 
-\IF{$is(exprNode, OperatorNode$)}
-    \IF{$exprNode.operator == GEN\_EQUALS$}
+\IF{$is(exprNode,$ OperatorNode$)$}
+    \IF{$exprNode.operator =$ GEN\_EQUALS}
         \FORALL{$var \in forVars$}
-            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is $PathType$ AND $isWithoutPredicatesExceptLastStep(getType(var/L))$ AND $usesOnlyChildAndDescendantAxes(getType(var/L))$ AND $C$ is a literal constant}
+            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is PathType $\wedge$ $isWithoutPredicatesExceptLastStep(getType(var/L))$ $\wedge$ $usesOnlyChildAndDescendantAxes(getType(var/L))$ $\wedge$ $C$ is a literal constant}
                 \STATE memorize $var/L$ as the negative key statement, weight 0.9
@@
         \ENDFOR
-    \ELSIF{$exprNode.operator == AND$}
+    \ELSIF{$exprNode.operator =$ AND}
         \STATE $processWhere(forVars, exprNode.leftSide)$

### 2012-03-15 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\begin{define}[Function $is(v, class)$]For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value $true$ if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
-Otherwise, it returns $false$.
+\begin{define}[Function $is(v, class)$]For every $v \in V$ and every $class \in C$, function $is(v, class)$ returns boolean value \texttt{true} if $c(v) = class$ or $c(v)$ is a (direct or indirect) subclass of $class$, according to the described principle.
+Otherwise, it returns \texttt{false}.
 \end{define}
 
-For instance, assuming $v \in V, c(v) =$ \texttt{LiteralNode}, calls $is(v,$ \texttt{LiteralNode}$)$ and $is(v,$ \texttt{ExprNode}$)$ return $true$, while call $is(v,$ \texttt{ContentNode}$)$ returns $false$.
+For instance, assuming $v \in V, c(v) =$ \texttt{LiteralNode}, calls $is(v,$ \texttt{LiteralNode}$)$ and $is(v,$ \texttt{ExprNode}$)$ return \texttt{true}, while call $is(v,$ \texttt{ContentNode}$)$ returns \texttt{false}.
 
@@
 Return types of built-in functions are fixed and, thus, there is no need to analyse them.
-Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by $fn$ or not prefixed.
+Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by \texttt{fn} or not prefixed.
 
@@
 The meaning of the code at lines 6-14 is the following.
-If a variable is bound to an expression which type is $PathType$, we want to assign the variable reference expression with $PathType$ which represents path containing exactly one step, the variable reference.
+If a variable is bound to an expression which type is PathType, we want to assign the variable reference expression with PathType which represents path containing exactly one step, the variable reference.
 Some algorithm in later sections search for paths starting with a certain variable and we want them to include also that types of expressions.
@@
 
-Function $determineExpressionType$ is called within function $analysis\-Of\-Expression\-Types$ presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
+Function \texttt{determineExpressionType} is called within function \texttt{analysis\-Of\-Expression\-Types} presented in Algorithm \ref{ALG_func_analysisOfExpressionTypes}, which recursively determines types of all subexpressions.
 Two important questions about this function emerge.
@@
 This is done at lines 4 and 5.
-Function $mergeContextVarTypes$ presented in Algorithm \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into its first argument (the variable context).
+Function \texttt{mergeContextVarTypes} presented in Algorithm \ref{ALG_function_mergeContextVarTypes} writes every record from its second argument (the new variables) into its first argument (the variable context).
 If there are records for variables with the same names in the context, they are overwritten to correspond to the variable overlapping.
@@
 
-Function $analysisOfExpressionTypes$ called upon a binding expression of a global variable determines the type of the binding expression, and hence, the type of the variable.
-In case of global variables, the argument $contextVarTypes$ is empty, because there are no local variables valid in the prolog section.
+Function \texttt{analysisOfExpressionTypes} called upon a binding expression of a global variable determines the type of the binding expression, and hence, the type of the variable.
+In case of global variables, the argument \texttt{contextVarTypes} is empty, because there are no local variables valid in the prolog section.
 
@@
 
-The types of processed global variables are available trough function \linebreak $getGlobalVariableType()$.
-If the function gets a variable that is not a global one, it returns $null$.
-Function $getVariableType()$ called in Algorithm \ref{ALG_func_determineExpressionType} is defined in Algorithm \ref{ALG_function_get_variable_type}.
+The types of processed global variables are available trough function \linebreak \texttt{getGlobalVariableType}.
+If the function gets a variable that is not a global one, it returns \texttt{null}.
+Function \texttt{getVariableType} called in Algorithm \ref{ALG_func_determineExpressionType} is defined in Algorithm \ref{ALG_function_get_variable_type}.
 It checks if the specified variable is amongst the given local variables.
-If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is $null$.
+If so, it returns its type, else it handles the variable as a global one and if such global variable does not exist the result is \texttt{null}.
 
@@
 \subsection{Determination of Expression Types}
-To determine types of expressions, we use already introduced function \linebreak $analysisOfExpressionTypes$.
+To determine types of expressions, we use already introduced function \\ \texttt{analysisOfExpressionTypes}.
 The starting node (its first argument) is the node representing the query body and the variable context is empty as there cannot be any local variable valid in the body node.
@@
 We can also determine expression types in functions.
-To do this for a certain function, $analysisOfExpressionTypes$ function has to be called with a function declaration node as the starting node.
+To do this for a certain function, \texttt{analysisOfExpressionTypes} function has to be called with a function declaration node as the starting node.
 In this case, the function declaration node contains a subnode specifying function's formal arguments.
@@
 Only expressions of a particular type are exploited.
-Specifically, an expression has to contain an subexpression $E$ of $PathType$ type (expression representing a certain element or attribute or set of elements or attributes).
+Specifically, an expression has to contain an subexpression $E$ of PathType type (expression representing a certain element or attribute or set of elements or attributes).
 In the following text, the set represented by expression $E$ is denoted $S$.
@@
 
-The output of this step is a set of statements of form \\ an instance of \texttt{PathType} $\rightarrow$ XML Schema built-in atomic type.
+The output of this step is a set of statements of form $P \rightarrow T$, where $P$ is an instance of PathType and $T$ is an XML Schema built-in atomic type.
 
@@
 This case is quite straightforward.
-The algorithm encounters a function call and one of the arguments is a known set of elements of attributes (subexpression $E$ representing $S$).
+The algorithm encounters a function call and one of the arguments is a set of elements of attributes (subexpression $E$ representing $S$) represented by PathType $P$.
 To determine the type of $S$, it is only needed to determine the type of the corresponding formal argument from the definition of the function.
@@
 
-If the type $T$ of the formal argument is a built-in type or its sequence, then $T$ is also the inferred type of $S$. Otherwise, no statement is inferred.
+If the type $T$ of the formal argument is a built-in type or its sequence, then $T$ is also the inferred type of $S$. The inferred statement is $P \rightarrow T$. Otherwise, no statement is inferred.
 
 \subsection{Arithmetic Operations}
-If the operator is one of $+$, $-$, $*$, $div$, $mod$ (the class of the expression node is $Operator$ and it represents one of $PLUS$, $MINUS$, $MUL$, $DIV$, $MOD$ (for all operators, see Attachment \ref{FIG_operators})), one operand is $E$ and the type $T$ of the other operand is one of numeric built-in types, then $T$ is also the inferred type of $S$.
-
-If the operator is one of $<$, $>$, $<=$, $>=$, $=$, $!=$ (the class of the expression node is $Operator$ and it represents one of $GEN\_LESS\_THAN$, \linebreak $GEN\_GREATER\_THAN$, $GEN\_LESS\_THAN\_EQUALS$,\linebreak $GEN\_GREATER\_THAN\_EQUALS$, $GEN\_EQUALS$,\linebreak $GEN\_NOT\_EQUALS$), one operand is $E$ and the type $T$ of the other operand is one of built-in types, then $T$ is also the inferred type of $S$.
+If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{*}, \texttt{div}, \texttt{mod} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{MUL}, \texttt{DIV}, \texttt{MOD} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
+
+\todo[inline]{Toto je zlozitejsie.}
+
+If the operator is one of \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, \texttt{!=} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{GEN\_EQUALS}, \\ \texttt{GEN\_NOT\_EQUALS}), one operand is of PathType $P$ and the type $T$ of the other operand is one of built-in types, then the inferred statement is $P \rightarrow T$.
 
@@
 
-If a currently processed expression is a function call of an user-defined function and at least one of its real arguments is an instance of \texttt{PathType}, get the function declaration node.
+If a currently processed expression is a function call of an user-defined function and at least one of its real arguments is an instance of PathType, get the function declaration node.
 In the function body, replace the formal arguments by the real arguments and set them as the local variables, so the proposed algorithms can be applied to process the function's body.
@@
 
-Function $usesOnlyChildAndDescendantAxes$ in Algorithm \ref{ALG_function_uses_only_child_and_descendant_axes} takes an instance of $PathType$ as its arguments and returns true if the path uses only child and descendant axes.
+Function \texttt{usesOnlyChildAndDescendantAxes} in Algorithm \ref{ALG_function_uses_only_child_and_descendant_axes} takes an instance of PathType as its arguments and returns true if the path uses only child and descendant axes.
 Otherwise, it returns false.
@@
 
-Function $isWithoutPredicates$ in Algorithm \ref{ALG_function_is_without_predicates} checks whether a path represented by a given $PathType$ instance does not contain predicates.
-A Similar function is  $isWithoutPredicatesExceptLastStep$.
-This function is not explicitly defined, because its definition is the same as the definition of function $isWithoutPredicates$, except it does not check the last step of the path.
+Function \texttt{isWithoutPredicates} in Algorithm \ref{ALG_function_is_without_predicates} checks whether a path represented by a given PathType instance does not contain predicates.
+A Similar function is  \texttt{isWithoutPredicatesExceptLastStep}.
+This function is not explicitly defined, because its definition is the same as the definition of function \texttt{isWithoutPredicates}, except it does not check the last step of the path.
 So the last step may contain predicates and the function will return true.
@@
 
-Function $endsWithExactlyOnePredicate$ in Algorithm \ref{ALG_function_ends_with_exactly_one_predicate} checks if a path represented by a given $PathType$ instance contains exactly one predicate in its last step and if so, the predicate is returned.
+Function \texttt{endsWithExactlyOnePredicate} in Algorithm \ref{ALG_function_ends_with_exactly_one_predicate} checks if a path represented by a given PathType instance contains exactly one predicate in its last step and if so, the predicate is returned.
 Otherwise, it returns false.
@@
 
-Function $isTargetPath$ in Algorithm \ref{ALG_function_is_target_path} checks if a path represented by a given $PathType$ instance is so-called target path for the specified variable.
+Function \texttt{isTargetPath} in Algorithm \ref{ALG_function_is_target_path} checks if a path represented by a given PathType instance is so-called target path for the specified variable.
 A target path for a variable is a path where its first step is a reference to the variable.
@@
 \ENDIF
-\IF{$is(detailNode, VarRefNode) =$ false}
+\IF{$is(detailNode,$ VarRefNode$) =$ false}
 	\RETURN false
@@
 
-Function $getTargetReturnPathTypes$ in Algorithm \ref{ALG_function_get_target_return_path_types} searches the given FLWOR for $PathTypes$ representing so-called target return paths.
+Function \texttt{getTargetReturnPathTypes} in Algorithm \ref{ALG_function_get_target_return_path_types} searches the given FLWOR for PathTypes representing so-called target return paths.
 A target return path is a return path in a where clause of a FLWOR.
-The FLWOR is processed recursively using function $getTargetReturnPathTypesRecursive$ in Algorithm \ref{ALG_function_get_target_return_path_types_recursive}.
+The FLWOR is processed recursively using function \texttt{getTargetReturnPathTypesRecursive} in Algorithm \ref{ALG_function_get_target_return_path_types_recursive}.
 
@@
 
-A similar couple of functions is function $getTargetReturnPaths$ and function $getTargetReturnPathsRecursive$.
-Only difference between those two couples of functions is that the former one searches for all expressions which are of $PathType$ representing a target return path, while the later one searches only for instances of $PathExprNode$.
+A similar couple of functions is function \texttt{getTargetReturnPaths} and function \texttt{getTargetReturnPathsRecursive}.
+Only difference between those two couples of functions is that the former one searches for all expressions which are of PathType representing a target return path, while the later one searches only for instances of \texttt{PathExprNode}.
 
 To illustrate this difference, we introduce the following example.
-Assuming $P$ is a return target path expression for variable $var$ and $flwor$ is a FLWOR expression with return clause $data(P)$, function call $getTargetReturnPathTypes(flwor, var)$ returns two $PathTypes$.
-One is the type of expression $P$ and the other is the type of the function call $data(P)$ which is also $PathType$.
-On the other hand, function call $getTargetReturnPaths(flwor, var)$ returns just one $PathType$ representing $P$.
+Assuming $P$ is a return target path expression for variable $var$ and $flwor$ is a FLWOR expression with return clause \texttt{data(}$P$\texttt{)}, function call \texttt{getTargetReturnPathTypes(} $flwor$, $var$\texttt{)} returns two PathTypes.
+One is the type of expression $P$ and the other is the type of the function call \texttt{data(}$P$\texttt{)} which is also PathType.
+On the other hand, function call \texttt{getTargetReturnPaths(}$flwor$, $var$\texttt{)} returns just one PathType representing $P$.
 
@@
 \subsection{Definition of Keys}
-To define a key and a foreign key, we adopt the definitions introduced in the original approach \cite{Necasky:2009:DXK:1529282.1529414} with a modification that XPath paths are replaced by \texttt{PathType} instances.
+To define a key and a foreign key, we adopt the definitions introduced in the original approach \cite{Necasky:2009:DXK:1529282.1529414} with a modification that XPath paths are replaced by PathType instances.
 
 \begin{define}[Key]
-A key is a construct $(C, P, \{L\})$ where $C$, $P$ and $L$ are \texttt{PathType} instances without predicates and without special function calls that use only child and descendant axes.
+A key is a construct $(C, P, \{L\})$ where $C$, $P$ and $L$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
 $C$ is called \emph{context path}, $P$ \emph{target path} and $L$ \emph{key path}.
@@
 \begin{define}[Foreign key]
-A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$ where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are \texttt{PathType} instances without predicates and without special function calls that use only child and descendant axes.
+A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$ where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
 $C$ can be omitted as in the case of keys.
@@
 
-Since, in this step, we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of $PathType$ without predicates, using only child and descendant axes.
+Since, in this step, we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of PathType without predicates, using only child and descendant axes.
 
@@
 If it does, the occurrence is noted.
-The responsible code is partially moved to function $determineJoinPatterns()$ in Algorithm \ref{ALG_func_determinateJoinPatterns}.
+The responsible code is partially moved to function \texttt{determineJoinPatterns} in Algorithm \ref{ALG_func_determinateJoinPatterns}.
 If the subnode does not form an occurrence, and if it is a for clause binding a new variable satisfying the conditions to be the first for clause of a join pattern, it is added to the array along with the variable name.
-It can be then processed in the following clauses of the currect FLWOR node or in its descendant FLWOR nodes, later in the recursion.
+It can be then processed in the following clauses of the current FLWOR node or in its descendant FLWOR nodes, later in the recursion.
 
@@
 
-As it can be seen in Algorithm \ref{ALG_classification_of_join_pattern_occurrences}, rules R2 and R3 are applied using instances of $PathType$ while, rules R4 and R5 use instances of $PathExprNode$.
+As it can be seen in Algorithm \ref{ALG_classification_of_join_pattern_occurrences}, rules R2 and R3 are applied using instances of PathType while, rules R4 and R5 use instances of \texttt{PathExprNode}.
 This difference is partially explained in definition of the auxiliary functions in Chapter \ref{step_4_auxiliary_functions}.
 Rules R4 and R5 count target return paths.
-If they use instances of $PathType$, they can count one path more times (as is described in the mentioned chapter), and thus, give a wrong result.
+If they use instances of PathType, they can count one path more times (as is described in the mentioned chapter), and thus, give a wrong result.
 
@@
 
-The first XQuery construct utilized by this step is a function call of $distinct-values$ function as discussed in Chapter \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
-The algorithm searches for calls of the mentioned functions with a $PathType$ on their input.
-Elements and attributes selected by those $PathTypes$ are then considered to be not unique, and consequently, they cannot be keys.
-
-Aggregation functions $min$, $max$, $sum$ are not utilized by the function because they were utilized by the search for join patterns, and in the process of classification of found join patterns, they were included in the computation of the key weights.
-
-Contrary the key discovery, $PathTypes$ utilized in this step may contain predicates in their lasts steps.
+The first XQuery construct utilized by this step is a function call of \\ \texttt{distinct-values} function as discussed in Chapter \ref{Negative_uniqueness} and shown in Algorithm \ref{ALG_rejection_of_uniqueness_aggregation_functions}.
+The algorithm searches for calls of the mentioned functions with a PathType on their input.
+Elements and attributes selected by those PathTypes are then considered to be not unique, and consequently, they cannot be keys.
+
+Aggregation functions \texttt{min}, \texttt{max}, \texttt{sum} are not utilized by the function because they were utilized by the search for join patterns, and in the process of classification of found join patterns, they were included in the computation of the key weights.
+
+Contrary the key discovery, PathTypes utilized in this step may contain predicates in their lasts steps.
 Predicates in a path's last step restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates.
@@
 Besides the aggregation functions, a certain form of FLWOR expressions is utilized in this step.
-They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (lieral constant, reference of a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
+They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference of a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
 
@@
             \STATE $expr := bindingNode.getChild(BindingSequenceNode).$ $getChild(ExprNode)$
-            \IF{$getType(expr)$ is PathType $\wedge$ $usesOnlyChildAndDescendantAxes(getType(expr))$}
+            \IF{$getType(expr)$ is PathType $\wedge$ $usesOnlyChildAndDescendantAxes($ $getType(expr))$}
                 \STATE add $bindingNode.varName$ to $forVars$

### 2012-03-15 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

@@
 Our statements are of the three following forms.
-The first one is $p \rightarrow t$, where $p$ is a $PathType$ representing an XQuery path selecting a set of elements or attributes, and $t$ is an XSD built-in atomic type.
+The first one is $p \rightarrow t$, where $p$ is a PathType representing an XQuery path selecting a set of elements or attributes, and $t$ is an XSD built-in atomic type.
 
-The second and third forms are $k = (C,P,\{L\})$ and $fk = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key, where $C,P,P_1,P_2,L,L_1,L_2$ are $PathTypes$ without predicates using only child and descendant axes.
+The second and third forms are $k = (C,P,\{L\})$ and $fk = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key, where $C,P,P_1,P_2,L,L_1,L_2$ are PathTypes without predicates using only child and descendant axes.
 
@@
 This is done in two steps.
-The first one is normalization of a particular $PathType$ and the second one is selection of the targeted elements.
+The first one is normalization of a particular PathType and the second one is selection of the targeted elements.
 
 \section{Evaluation of Paths}
-A path represented by $PathType$ can contain variable references and association of these references with other paths.
+A path represented by PathType can contain variable references and association of these references with other paths.
 That is the reason why the normalization is convenient.
@@
 \REQUIRE{\ \\
-	$step$: An instance of $StepExprNode$ representing a step of the path to evaluate.
+	$step$: An instance of StepExprNode representing a step of the path to evaluate.
 	$contextSet$: A set of elements and attributes to evaluate $step$ upon.
@@
 
-\IF{$is(step, SelfOrDescendantStep)$}
+\IF{$is(step,$ SelfOrDescendantStep$)$}
 	\RETURN $evaluateStep\_selfOrDescendant(contextSet, grammar)$
@@
 
-\STATE $newContextSet$ = empty set
+\STATE $newContextSet :=$ an empty set
 
-\STATE $axisKind$ = $step.getChild(axisNode).axisKind$
-\STATE $nodeName$ = $step.getChild(axisNode).getChild(nameTestNode).name$
-\IF{$axisKind$ == $CHILD$}
+\STATE $axisKind := step.getChild(axisNode).axisKind$
+\STATE $nodeName := step.getChild(axisNode).getChild(nameTestNode).name$
+\IF{$axisKind =$ CHILD}
 	\FORALL{$node \in contextSet$}
 		\IF{$node$ is element}
-			\STATE $subelements$ = $getTokens(grammar[node])$
+			\STATE $subelements := getTokens(grammar[node])$
 			\FORALL{$subelement \in subelements$}
-				\IF{$subelement.name$ == $nodeName$}
+				\IF{$subelement.name = nodeName$}
 					\STATE add $subelement$ to $newContextSet$
@@
 	\ENDFOR
-\ELSIF{$axisKind$ == $ATTRIBUTE$}
+\ELSIF{$axisKind =$ ATTRIBUTE}
 	\FORALL{$node \in contextSet$}
 		\IF{$node$ is element}
-			\STATE $attributes$ = $node.attributes$
+			\STATE $attributes := node.attributes$
 			\FORALL{$attribute \in attributes$}
-				\IF{$attribute.name$ == $nodeName$}
+				\IF{$attribute.name = nodeName$}
 					\STATE add $attribute$ to $newContextSet$
@@
 At first, the algorithm determine the axis of the step.
-If it is self or descendant axis, it returns the result of $evaluateStep\_selfOrDescendant$ function.
+If it is self or descendant axis, it returns the result of \texttt{evaluateStep\_selfOrDescendant} function.
 This function returns the given context set extended by all descendants of the elements from the context set in the given grammar.
@@
 If the axis is child axis, for each element in the context set, the algorithm determines its subelement using the grammar, and, if those with the name equal to the name specified by to step are added to the resulting context set.
-Function $getTokens$ at line 10 retrieves all elements from a regular expression.
+Function \texttt{getTokens} at line 10 retrieves all elements from a regular expression.
 A specific form of the regular expression is not important, because we only need to know which elements are possible subelements of the particular element.
@@
 
-Consider these two expressions; $PathType$ $p$ is compared to an integer literal constant, and, $PathType$ $p$ is an argument of function where a formal type of the argument is byte.
+Consider these two expressions; PathType $p$ is compared to an integer literal constant, and, PathType $p$ is an argument of function where a formal type of the argument is byte.
 The first expression is utilized to infer statement $p \rightarrow integer$, and the second one to infer statement $p \rightarrow byte$.

### 2012-03-15 doc/theses/mikula/implementation.tex
Master thesis: In progress.

\chapter{Implementation}
The solution proposed in Chapters \ref{CHAPTER_proposed_algorithm} and \ref{CHAPTER_combination_with_existings_methods_of_inference} was implemented using the jInfer framework \cite{jinfer}. It is a framework for implementing methods of XML schema inferring created as a software project at Faculty of Mathematics and Physics, Charles University in Prague. It is written in Java as a plugin for NetBeans platform.

The framework consists of modules representing logical parts of a process of inference. The main idea behind the modules is that they can be replaced by other modules with the same interface but different implementation and new modules can be connected to extend functionality.

\begin{figure}
\label{FIG_jinfer}
\caption{jInfer process of inference.}
\includegraphics[scale=1]{jinfer.eps}
\end{figure}

Figure \ref{FIG_jinfer} shows the process of inference in jInfer. The grey rectangles represent steps of the inference, the white boxes represent input and output. Originally, the inference was composed of \textbf{Initial Grammar Generator}, \textbf{Simplifier}, and \textbf{Schema Generator} steps. Steps \textbf{XQAnalyzer}, \textbf{XQuery Processor}, and \textbf{Merger} are implementations of main parts of our proposed solution.

\begin{itemize}
\item \textbf{XQAnalyzer} - An implementation of the lexical and syntax analyses proposed in \cite{thesis_schejbal} and modified to create syntax trees. It is partially incorporated in \textbf{XQueryImporter} module (lexer, parser and syntax tree creation) which handles input XQuery files, and \textbf{Base} module (syntax tree data structures) which contains classes shared by several modules.
\item \textbf{XQuery Processor}
\item \textbf{Merger}
\end{itemize}

### 2012-03-15 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

@@
-\chapter{Combination with Existing Methods of Inference}
+\chapter{Combination with Existing Methods of Inference} \label{CHAPTER_combination_with_existings_methods_of_inference}
 In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference.

### 2012-03-15 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
-\chapter{Proposed Algorithm}
+\chapter{Proposed Algorithm} \label{CHAPTER_proposed_algorithm}
 In this thesis we are developing an approach dealing with inference of XML schema, which input consists of two components; a set of XML documents and a set of XQuery queries related to the documents.

### 2012-03-15 doc/theses/mikula/implementation.tex
Master thesis: In progress.

@@
 
+\section{jInfer Process of Inference}
 \begin{figure}
-\label{FIG_jinfer}
+\label{FIG_jinfer_steps}
 \caption{jInfer process of inference.}
-\includegraphics[scale=1]{jinfer.eps}
+\includegraphics[scale=0.9]{jinfer_steps.eps}
 \end{figure}
 
-Figure \ref{FIG_jinfer} shows the process of inference in jInfer.
-The grey rectangles represent steps of the inference, the white boxes represent input and output.
-Originally, the inference was composed of \textbf{Initial Grammar Generator}, \textbf{Simplifier}, and \textbf{Schema Generator} steps.
-Steps \textbf{XQAnalyzer}, \textbf{XQuery Processor}, and \textbf{Merger} are implementations of main parts of our proposed solution.
+Figure \ref{FIG_jinfer_steps} shows the process of inference in jInfer.
+The grey and yellow rectangles represent steps of the inference, the white boxes represent input and output.
+Originally, the inference was composed of \textbf{Initial Grammar Generator}, \textbf{Simplifier}, and \textbf{Schema Generator} steps (grey).
+Steps (yellow) \textbf{XQAnalyzer}, \textbf{XQuery Processor}, and \textbf{Merger} are implementations of main parts of our proposed solution.
 
@@
 \item \textbf{XQAnalyzer} - An implementation of the lexical and syntax analyses proposed in \cite{thesis_schejbal} and modified to create syntax trees.
-It is partially incorporated in \textbf{XQueryImporter} module (lexer, parser and syntax tree creation) which handles input XQuery files, and \textbf{Base} module (syntax tree data structures) which contains classes shared by several modules.
-\item \textbf{XQuery Processor}
-\item \textbf{Merger}
+It parses input XQuery files and outputs their syntax trees.
+\item \textbf{XQuery Processor} - An implementation of algorithms proposed in Chapter \ref{CHAPTER_proposed_algorithm}.
+In particular, construction of a syntax tree, static analysis of expression types, inference of built-in types, and key discovery.
+On input, it takes the syntax trees, and its output are statements inferred from the queries represented by the syntax tress.
+\item \textbf{Merger} - A step of inference responsible for combining the simplified grammar with the statements inferred by \textbf{XQuery Processor} as described in Chapter \ref{CHAPTER_combination_with_existings_methods_of_inference}.
+Its result is the simplified grammar extended by the statements from \textbf{XQuery Processor} in a way that each statement is analysed and its information is assigned to concerned elements and attributes from the grammar.
 \end{itemize}
+
+\section{jInfer Modules}
+\begin{figure}
+\label{FIG_jinfer_modules}
+\caption{jInfer modules.}
+\includegraphics[scale=0.8]{jinfer_modules.eps}
+\end{figure}
+
+The previous section describes a high level schematic view of the process of inference.
+But, from a more technical view, jInfer modules do not utterly correspond with the steps of the presented process of inference.
+
+The first difference is that none of the modules runs in parallel (in several threads) with other, as it is schematically shown in Figure \ref{FIG_jinfer_steps}.
+The modules run in a serial order as shown in Figure \ref{FIG_jinfer_modules}.
+The second difference is that one step is not always represented by one module, or one module is not always representing one step.
+
+Actually, the blue modules shown in \ref{FIG_jinfer_modules} are module classes with specified interface.
+Actual modules are then implementations complying the interface.
+Since we provide only one implementation of modules for the proposed solution, this is not very important for this work and we do not describe the principle in detail.
+
+As in the previous section, newly added modules are those shown in yellow.
+They are \textbf{XQuery Processor} and \textbf{XQuery Importer}.
+We also modified modules \textbf{Basic XSD Exporter} (implementation of module class \textbf{Schema Generator}) and \textbf{Base} to extend their functionality.
+
+\begin{itemize}
+\item \textbf{Base} - This is module containing common classes used by other modules and defining interfaces.
+We added two packages to this module.
+Package \texttt{cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes} implements the structure of syntax trees.
+And package \texttt{cz.cuni.mff.ksi\-.jinfer.base.objects.xquery.xqueryprocessor.types} contains data \\ structures and utilities for types defined and used by the proposed solution.
+\item \textbf{XQuery Importer} - This module represents \textbf{XQAnalyzer} step in Figure \ref{FIG_jinfer_steps}.
+\item \textbf{XQuery Processor} - The main module consisting of steps \textbf{XQuery Processor} and \textbf{Merger} in Figure \ref{FIG_jinfer_steps}.
+Package \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer} contains the main class \texttt{XQueryAnalyzerImpl} and classes implementing the steps of the proposed algorithm.
+Packages \texttt{cz.cuni.mff\-.ksi.jinfer\-.xqueryanalyzer.keys} and \texttt{cz.cuni.mff.ksi.jinfer-\\.xqueryanalyzer.joinpatterns} contain representation of keys and join patterns.
+And package \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer.utils} contains various utilities.
+\item \textbf{Basic XSD Exporter} - This is the module responsible for generating the resulting schema in XSD.
+It was modified to process the additional information assigned to the grammar in \textbf{XQuery Processor} module.
+\end{itemize}

### 2012-03-16 doc/theses/mikula/conclusion.tex
Master thesis: In progress.

\chapter{Conclusion}
The aim of this thesis was to employ XML operations in the XML schema inference. We analysed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations, only to found one, which infers keys from XQuery queries.

Since the notion XML operations is very general and the range of XML technologies is very large, for purpose of this work, we decided to focus on XQuery technology. We made the overview of possible utilization of XQuery queries in the process of XML schema inference.

Before a creation of the algorithm itself, we had to take several decisions in questions that emerged. Since there is a lack of the methods dealing with the utilisation of XML operation, there is also a lack of practically proven solutions that can help in such decision making.

In the proposed solution, we decided to incorporate lexical and syntax analyses of XQuery queries, because it is more general and more extensible than a pattern searching. To achieve that, we adopted the algorithm from a recent master thesis dealing with an analysis of XQuery queries.

We also implemented several ideas from the overview to infer XSD built-in types of elements and attributes. And we slightly extended the one existing method dealing with the inference of keys. \todo{Ako dopadli testy.}

At last, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using jInfer framework. Thus, we created the first complete, ready-to-use, and extensible implementation of the XML xchema inference exploiting XQuery queries besides XML data.

\section{Future Work}

### 2012-03-16 doc/theses/mikula/diplomka.tex
Master thesis: In progress.

--- 2012-03-06 doc/theses/mikula/diplomka.tex
@@
 \include{combination_with_existing_methods_of_inference}
+\include{implementation}
+\include{experiments}
+\include{conclusion}
 
@@
 \bibliographystyle{plain}
-\todo[inline]{Literatura tady určitě co nejpodrobněji - odkaz nestačí, za měsíc nemusí existovat. A nemělo by to přetékat okraj.}
-\todo[inline]{Literatura - jInfer modules documentation - Tady by neškodily odkazy na jednotlivé dokumenty.}
 \bibliography{literature}
@@
 %%% textu práce. Přílohy se číslují.
-\chapwithtoc{Attachments}
+%\chapwithtoc{Attachments}
+\appendix
+\include{appendix_a}
 

### 2012-03-16 doc/theses/mikula/conclusion.tex
Master thesis: In progress.

@@
 \section{Future Work}
+This work implements only some ideas discussed in the overview in Chapter \ref{CHAPTER_analysis_of_xquery}, leaving most of them for a future work.
+Also, the implemented algorithms can be further refined as was already mentioned in Chapter \ref{CHAPTER_proposed_algorithm} with the presentation of the algorithms.
+For example thorough processing of user-defined function calls.
+
+Chapter \ref{CHAPTER_precreation_of_algorithm} discusses some possible future enhancements, as well.
+A research on possibilities of modifying the grammar rules based on information extracted from XQuery queries may bring interesting results.
+We analyse the queries statically only, we do not evaluate them.
+An analysis of queries together with their results can be a topic of some future research.
+
+The utilization of XQuery queries certainly provides a space for incorporating an interaction with user. For example, a user may influence the scoring of inferred keys to get a more precise results.
+
+And, a very large space for a possible future research is provided by utilisation of other XML operations. The main representant is XSLT.

### 2012-03-16 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

@@
-\chapter{Pre-creation of Algorithm}
+\chapter{Pre-creation of Algorithm} \label{CHAPTER_precreation_of_algorithm}
 According to the analysis in the previous chapter, there is quite a wide range of possible utilizations of XQuery queries.

### 2012-03-19 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/expressiontypesanalysis/ExpressionTypesAnalyser.java
Master thesis implementation: In progress.

@@ -142,7 +142,7 @@ public class ExpressionTypesAnalyser {
       }
     } else if (TupleStreamNode.class.isInstance(root)) {
       newVars = newContextVarTypes;
-    } else if (InClausesNode.class.isInstance(root)) { // TODO rio do diplomky!!
+    } else if (InClausesNode.class.isInstance(root)) {
       newVars = newContextVarTypes;
     }
 
@@ -192,7 +192,7 @@ public class ExpressionTypesAnalyser {
     } else if (IfExprNode.class.isInstance(expressionNode)) {
       final IfExprNode ifExprNode = (IfExprNode)expressionNode;
       return ifExprNode.getThenExpressionNode().getExprNode().getType();
-    } else if (QuantifiedExprNode.class.isInstance(expressionNode)) { // TODO rio do diplomky!!
+    } else if (QuantifiedExprNode.class.isInstance(expressionNode)) {
       return new XSDType(XSDBuiltinAtomicType.BOOLEAN, Cardinality.ONE);
     }
     assert (false); // TODO rio Implement other expressions.

### 2012-03-19 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/expressiontypesanalysis/FunctionsAnalyser.java

@@ -20,8 +20,10 @@ import cz.cuni.mff.ksi.jinfer.xqueryanalyzer.utils.BuiltinFunctionsUtils;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.*;
 import cz.cuni.mff.ksi.jinfer.base.interfaces.xquery.Type;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.AbstractType;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.UnknownType;
 import java.util.HashMap;
 import java.util.Map;
+import org.apache.log4j.Logger;
 import sun.reflect.generics.reflectiveObjects.NotImplementedException;
 
 /**
@@ -37,6 +39,7 @@ import sun.reflect.generics.reflectiveObjects.NotImplementedException;
  */
 public class FunctionsAnalyser {
   
+  private static final Logger LOG = Logger.getLogger(FunctionsAnalyser.class);
   private final Map<String, FunctionDeclNode> functionDeclarationNodes;
   private final Map<String, Type> functionTypes;
   
@@ -73,7 +76,8 @@ public class FunctionsAnalyser {
       }
     }
     
-    throw new NotImplementedException();
+    LOG.warn("getFunctionType: Function " + qualifiedName + " is not implemented yet. Returning UnknownType.");
+    return new UnknownType();
   }
   
   /**
@@ -92,7 +96,8 @@ public class FunctionsAnalyser {
       }
     }
     
-    throw new NotImplementedException();
+    LOG.warn("getParamListNode: Function " + functionName + " is not implemented yet.");
+    return null;
   }
   
   private Map<String, FunctionDeclNode> getFunctionDeclarationNodes(final ModuleNode root) {

### 2012-03-19 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/keydiscovery/joinpatterns/JoinPatternsFinder.java

@@ -114,7 +114,7 @@ public class JoinPatternsFinder {
             for (final Entry<String, ForClauseNode> entry : forVariables.entrySet()) {
               determineJoinPattern(bindingNode, predicate, entry.getValue(), entry.getKey(), checkJoinPattern3, whereExpr, foundJoinPatterns);
             }
-          } else if (checkJoinPattern3 && PathTypeUtils.isWithoutPredicates((PathType) bindingExprType)) { // TODO rio do diplomky!!
+          } else if (checkJoinPattern3 && PathTypeUtils.isWithoutPredicates((PathType) bindingExprType)) {
             for (final Entry<String, ForClauseNode> entry : forVariables.entrySet()) {
               determineJoinPattern(bindingNode, predicate, entry.getValue(), entry.getKey(), checkJoinPattern3, whereExpr, foundJoinPatterns);
             }
@@ -140,7 +140,7 @@ public class JoinPatternsFinder {
         final JoinWhereFormExprParser exprParser = new JoinWhereFormExprParser(whereExpr, forVar, bindingNode.getVarName());
         if (exprParser.isExprWhereForm()) {
           final PathType P1 = (PathType) forClauseNode.getBindingSequenceNode().getExprNode().getType(); // TODO rio I'm not sure is this is correct - if these expressions are always path types.
-          final PathType P2 = (PathType) bindingNode.getBindingSequenceNode().getExprNode().getType(); // TODO rio do DP!!
+          final PathType P2 = (PathType) bindingNode.getBindingSequenceNode().getExprNode().getType();
           foundJoinPatterns.add(new JoinPattern(JoinPattern.JoinPatternType.JP3, forClauseNode, bindingNode, P1, P2, exprParser.getL1(), exprParser.getL2()));
         }
       }
@@ -151,7 +151,7 @@ public class JoinPatternsFinder {
     final JoinPredicateFormExprParser exprParser = new JoinPredicateFormExprParser(predicate, forVar);
     if (exprParser.isExprPredicateForm()) {
       final PathType P1 = (PathType) forClauseNode.getBindingSequenceNode().getExprNode().getType(); // TODO rio I'm not sure is this is correct - if these expressions are always path types.
-      final PathType P2WithPredicate = (PathType) bindingNode.getBindingSequenceNode().getExprNode().getType(); // TODO rio do DP!!
[…]

### 2012-03-19 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/utils/BuiltinFunctionsUtils.java

@@ -27,6 +27,7 @@ import java.util.ArrayList;
 import java.util.HashMap;
 import java.util.List;
 import java.util.Map;
+import org.apache.log4j.Logger;
 import sun.reflect.generics.reflectiveObjects.NotImplementedException;
 
 /**
@@ -58,6 +59,8 @@ public class BuiltinFunctionsUtils {
     DEFAULT_BUILTIN_FUNCTIONS_NAMESPACES.add("");
     DEFAULT_BUILTIN_FUNCTIONS_NAMESPACES.add("fn");
   }
+  
+  private static final Logger LOG = Logger.getLogger(BuiltinFunctionsUtils.class);
   
   /**
    * If a specified function is built-in returns its name without prefix, else
@@ -111,7 +114,8 @@ public class BuiltinFunctionsUtils {
     if (builtinFunctionTypes.containsKey(builtinFuncName)) {
       return builtinFunctionTypes.get(builtinFuncName);
     } else {
-      throw new NotImplementedException();
+      LOG.warn("getFunctioncCallType: Built-in function " + builtinFuncName + " is not implemented yet. Returning UnknownType.");
+      return new UnknownType(); // TODO rio Implement completely if needed.
     }
   }
   
@@ -160,7 +164,8 @@ public class BuiltinFunctionsUtils {
       paramNodes.add(paramNode2);
       paramNodes.add(paramNode3);
     } else {
-      throw new NotImplementedException(); // TODO rio Implement completely if needed.
+      LOG.warn("getParamListNode: Built-in function " + functionName + " is not implemented yet.");
+      return null; // TODO rio Implement completely if needed.
     }
     
     return new ParamListNode(paramNodes);

### 2012-03-19 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/builtintypeinference/BuiltinTypesInferrer.java
Master thesis implementation: In progress.

@@ -78,7 +78,7 @@ public class BuiltinTypesInferrer {
         if (paramTypeDeclarationNode == null) {
           continue;
         }
-        final Type formalParamType = AbstractType.createType(paramTypeDeclarationNode); // TODO rio do diplomky!
+        final Type formalParamType = AbstractType.createType(paramTypeDeclarationNode);
         if (formalParamType.getCategory() != Type.Category.UNKNOWN) {
           inferredTypes.add(new InferredTypeStatement((PathType)paramType, formalParamType));
         }

### 2012-03-19 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/merger/Merger.java

@@ -47,7 +47,7 @@ import java.util.LinkedHashSet;
 import java.util.List;
 import java.util.Map;
 import java.util.Set;
-import sun.reflect.generics.reflectiveObjects.NotImplementedException;
+import org.apache.log4j.Logger;
 
 /**
  * Class responsible for merging the inferred statements with the grammar.
@@ -60,6 +60,8 @@ import sun.reflect.generics.reflectiveObjects.NotImplementedException;
  * @author rio
  */
 public class Merger {
+  
+  private final static Logger LOG = Logger.getLogger(Merger.class);
 
   /* Grammar metadata keys. Inferred statements will be written to the metadata
    * under these keys. Any exporter which want to use statements inferred by
@@ -285,7 +287,7 @@ public class Merger {
                 }
               }
             }
-          } else if (KindTestNode.class.isInstance(itemTypeNode)) { // TODO rio do diplomky!!
+          } else if (KindTestNode.class.isInstance(itemTypeNode)) {
             final KindTestNode ktn = (KindTestNode) itemTypeNode;
             final NodeKind nk = ktn.getNodeKind();
             if (nk == NodeKind.TEXT) {
@@ -295,16 +297,19 @@ public class Merger {
                 }
               }
             } else {
-              throw new NotImplementedException();
+              LOG.warn("evaluateStep: NodeKind " + nk.toString() + " is not implemented yet, the result of evaluation is very likely wrong.");
+              return contextSet;
             }
           } else {
-            throw new NotImplementedException(); // Not completely finished, but probably sufficient for our purpose.
+            LOG.warn("evaluateStep: This subclass of ItemTypeNode is not implemented yet, the result of evaluation is very likely wrong.");
+            return contextSet; // Not completely finished, but probably sufficient for our purpose.
           }
           break;
         }
 
         default:
-          throw new NotImplementedException(); // Not completely finished, but probably sufficient for our purpose.
+         LOG.warn("evaluateStep: AxisKind " + axisNode.getAxisKind().toString() + " is not implemented yet, the result of evaluation is very likely wrong.");
+         return contextSet; // Not completely finished, but probably sufficient for our purpose.
       }
     } else {
       final ExprNode detailNode = step.getDetailNode();
@@ -316,7 +321,8 @@ public class Merger {
         }
       }
 
-      throw new NotImplementedException(); // Not completely finished, but probably sufficient for our purpose.
[…]

### 2012-03-19 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \item XML Schema built-in atomic types. See Attachment \ref{FIG_xsd_built_in_atomic_types}.
-\item Types \texttt{ElementType}, \texttt{AttributeType}, \texttt{NodeType}, \texttt{TextNodeType},\linebreak \texttt{CommentType}, \texttt{ProcessingInstructionType}, \texttt{DocumentType} representing an element, attribute, node, text node, comment, processing instruction, document node.
-\item Type representing a node or a set of nodes selected by a certain path expression. The path expression is included in this type. Let this type be identified as \texttt{PathType}.
-\item \texttt{UnknownType} representing a type without known details, which does not suit one of the three previous types. An example is XSD type \texttt{anyType}.
+\item Types \emph{ElementType}, \emph{AttributeType}, \emph{NodeType}, \emph{TextNodeType},\linebreak \emph{CommentType}, \emph{ProcessingInstructionType}, \emph{DocumentType} representing an element, attribute, node, text node, comment, processing instruction, document node.
+\item Type representing a node or a set of nodes selected by a certain path expression. The path expression is included in this type. Let this type be identified as \emph{PathType}.
+\item \emph{UnknownType} representing a type without known details, which does not suit one of the three previous types. An example is XSD type \texttt{anyType}.
 \end{itemize}
@@
 \subsection{PathType}
-\texttt{PathType} contains additional information.
+PathType contains additional information.
 The represented path is contained by an instance of \texttt{PathExprNode} containing list of its steps, in particular instances of \texttt{StepExprNode}.
-If a step is a reference to a variable which type is \texttt{PathType}, we also want to include this information.
-Therefore, \texttt{PathType} contain association between the steps and other \texttt{PathType}s and this association is defined for the \texttt{PathType} variable steps.
-
-To differ between a common \texttt{PathType} selecting a set of nodes and a \texttt{PathType} bound to a for variable in a FLWOR expression, \texttt{PathType} structure contains a flag \texttt{isForBound}.
-
-Additionally, \texttt{PathType} contains a list of special functions that were called with an argument of \texttt{PathType} type.
-The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of \texttt{PathType} because then, we can determine a type of the function call more precisely.
+If a step is a reference to a variable which type is PathType, we also want to include this information.
+Therefore, PathType contain association between the steps and other PathTypes and this association is defined for the PathType variable steps.
+
+To differ between a common PathType selecting a set of nodes and a PathType bound to a for variable in a FLWOR expression, PathType structure contains a flag \texttt{isForBound}.
+
+Additionally, PathType contains a list of special functions that were called with an argument of PathType type.
+The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of PathType because then, we can determine a type of the function call more precisely.
 Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max}, \texttt{avg}, \texttt{sum}, \texttt{distinct-values}.
 
-In summary, we represent \texttt{PathType} as a structure with the following member variables.
+In summary, we represent PathType as a structure with the following member variables.
 \begin{itemize}
 \item \texttt{pathExprNode} - An instance of \texttt{PathExprNode}.
-\item \texttt{substeps} - An association between steps from \texttt{pathExprNode} and instances of \texttt{PathType} type.
+\item \texttt{substeps} - An association between steps from \texttt{pathExprNode} and instances of PathType type.
 \item \texttt{isForBound} - Boolean flag determining if the type was bound to a for variable in a for clause.
@@
 
-To capture sequences, we assign the first two categories of types (all types except of \texttt{PathType} and \texttt{UnknownType}) with its cardinality as proposed in \cite{thesis_schejbal}.
+To capture sequences, we assign the first two categories of types (all types except of PathType and UnknownType) with its cardinality as proposed in \cite{thesis_schejbal}.
 Each of those types can be perceived as a sequence.
@@
 
-\texttt{PathType} is not assigned with the cardinality since we do not evaluate the queries, and therefore, we cannot determine if a certain XQuery path targets zero, one or more nodes.
-Alike, \texttt{UnknownType} is neither assigned with the cardinality.
-Expressions of \texttt{UnknownType} are not utilized it the inference, therefore, their cardinality is not needed.
+PathType is not assigned with the cardinality since we do not evaluate the queries, and therefore, we cannot determine if a certain XQuery path targets zero, one or more nodes.
+Alike, UnknownType is neither assigned with the cardinality.
+Expressions of UnknownType are not utilized it the inference, therefore, their cardinality is not needed.
 
@@
 
-The order of the recursion is to firstly process children of a node and then the node itself.
+The order of the recursion is firstly to process children of a node and then the node itself.
 The order of child processing is specified by function $o$ from the definition of the syntax tree in Definition \ref{DEF_syntax_tree}.
@@
     \STATE $set(startingNode, newVars, (startingNode.varName, type))$
-\ELSIF{$is(startingNode,$ TupleStreamNode$)$}
+\ELSIF{$is(startingNode,$ TupleStreamNode$) \vee is(startingNode,$ InClausesNode$)$}
     \FORALL{$varBindingNode \in startingNode.getChildren()$}
@@
     	\STATE $P := endsWithExactlyOnePredicate(type)$
-    	\IF{$P$}
+    	\IF{$P \vee (checkJoinPattern3 \wedge isWithoutPredicates(type))$}
         	\FORALL{$(var, node) \in forVars$}
-            	\STATE $forVars := determinateJoinPatterns(bindingNode, P, node, var,$ $forVars,$ $checkJoinPattern3,$ $whereExpr)$
-        	\ENDFOR
-        \ELSIF{$isWithoutPredicates(type)$}
+            	\STATE $forVars := determineJoinPattern(bindingNode, P, node, var,$ $forVars,$ $checkJoinPattern3,$ $whereExpr)$
+        	\ENDFOR
+        \ENDIF
+        \IF{$isWithoutPredicates(type)$}
         	\IF{$is(bindingNode,$ ForClauseNode$)$}
@@
 \begin{algorithm}
-\caption{Function \texttt{determinateJoinPatterns}}
-\label{ALG_func_determinateJoinPatterns}
+\caption{Function \texttt{determineJoinPattern}}
+\label{ALG_func_determineJoinPattern}
 \begin{algorithmic}[1]
@@
 If it does, the occurrence is noted.
-The responsible code is partially moved to function \texttt{determineJoinPatterns} in Algorithm \ref{ALG_func_determinateJoinPatterns}.
+The responsible code is partially moved to function \texttt{determineJoinPattern} in Algorithm \ref{ALG_func_determineJoinPattern}.
 If the subnode does not form an occurrence, and if it is a for clause binding a new variable satisfying the conditions to be the first for clause of a join pattern, it is added to the array along with the variable name.

### 2012-03-19 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

--- 2012-03-15 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
@@
-\chapter{Combination with Existing Methods of Inference}
+\chapter{Combination with Existing Methods of Inference} \label{CHAPTER_combination_with_existings_methods_of_inference}
+\todo[inline]{Kontrola na zaklade dat.}
 In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference.

### 2012-03-19 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\STATE $type :=$ null
+\STATE $type :=$ UnknownType
 \STATE $cardinality := typeNode.cardinality$
@@
 \IF{$is(itemTypeNode,$ AtomicTypeNode$)$}
-    \STATE $type := itemTypeNode.typeName$
+    \STATE $type :=$ XSD atomic built-in type $itemTypeNode.typeName$, $cardinality$
 \ELSIF{$is(itemTypeNode,$ KindTestNode$)$}
-    \STATE $type := itemTypeNode.nodeKind$
-\ENDIF
-
-\RETURN $type, cardinality$
-\end{algorithmic}
-\end{algorithm}
-
-\todo[inline]{ALG Extraction of a Type from TypeNode nie uplne sedi s definiciou typov. DP JS ma nieco navyse. A tento alg je nedokonceny!!}
+    \STATE $type := itemTypeNode.nodeKind$, $cardinality$
+\ENDIF
+
+\RETURN $type$
+\end{algorithmic}
+\end{algorithm}
 
@@
 \subsection{Arithmetic Operations}
-If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{*}, \texttt{div}, \texttt{mod} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{MUL}, \texttt{DIV}, \texttt{MOD} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
-
-\todo[inline]{Toto je zlozitejsie.}
+If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
+
+If the operator in an arithmetic operation is one of \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{MUL}, \texttt{DIV}, one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow $ \texttt{double}.
 

### 2012-03-19 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \subsection{Arithmetic Operations}
-If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
-
-If the operator in an arithmetic operation is one of \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{MUL}, \texttt{DIV}, one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow $ \texttt{double}.
+If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod}, \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD}, \texttt{MUL}, \texttt{DIV} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
 

### 2012-03-19 jinfer/branches/thesis-mikula/BasicXSDExporter/src/cz/cuni/mff/ksi/jinfer/basicxsd/elementsexporters/AbstractElementsExporter.java
Master thesis implementation: In progress.

@@ -111,7 +111,7 @@ public abstract class AbstractElementsExporter {
       return;
     }
 
-    // If element's type is global set it and finish.
+    // If element's type is global set it, set keys and finish.
     if (preprocessingResult.isElementGlobal(element.getName())) {
       indentator.append(" type=\"");
       indentator.append(typenamePrefix);
@@ -119,7 +119,16 @@ public abstract class AbstractElementsExporter {
       indentator.append(typenamePostfix);
       indentator.append("\"");
       indentator.append(OccurencesProcessor.processOccurrences(interval));
-      indentator.append("/>\n");
+      if (hasElementKeys(element)) {
+        indentator.append(">\n");
+        indentator.increaseIndentation();
+        processElementKeys(element);
+        processElementForeignKeys(element);
+        indentator.decreaseIndentation();
+        indentator.indent("</xs:element>\n");
+      } else {
+        indentator.append("/>\n");
+      }
       return;
     }
 
@@ -185,6 +194,16 @@ public abstract class AbstractElementsExporter {
     processElementAttributes(element);
   }
   
+  protected boolean hasElementKeys(final Element element) {
+    final List<Key> keys = (List<Key>)element.getMetadata().get("xquery_processor_keys");
+    
+    if (BaseUtils.isEmpty(keys)) {
+      return false;
+    } else {    
+      return true;
+    }
+  }
+  
   protected void processElementKeys(final Element element) {
     final List<Key> keys = (List<Key>)element.getMetadata().get("xquery_processor_keys");
     
@@ -222,7 +241,7 @@ public abstract class AbstractElementsExporter {
       final String keyName = keyNames.get(fKey.getKey());
       assert(keyName != null);
 
-      indentator.indent("<xs:keyref name=\"" + keyName + "Ref\">\n");
+      indentator.indent("<xs:keyref name=\"" + keyName + "Ref\" refer=\"" + keyName + "\">\n");
       indentator.increaseIndentation();
       indentator.indent("<xs:selector xpath=\"" + targetPath.toString() + "\"/>\n");
       indentator.indent("<xs:field xpath=\"" + keyPath.toString() + "\"/>\n");

### 2012-03-19 jinfer/branches/thesis-mikula/BasicXSDExporter/src/cz/cuni/mff/ksi/jinfer/basicxsd/elementsexporters/GlobalElementsExporter.java

@@ -89,9 +89,6 @@ public final class GlobalElementsExporter extends AbstractElementsExporter {
     indentator.increaseIndentation();
 
     processElementContent(element);
-    
-    processElementKeys(element);
-    processElementForeignKeys(element);
 
     indentator.decreaseIndentation();
 

### 2012-03-20 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.

--- 2012-03-15 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
@@
-\chapter{Combination with Existing Methods of Inference}
-In this chapter, we describe how to incorporate the inferred statements with existing methods of XML schema inference.
+\chapter{Combination with Existing Methods of Inference} \label{CHAPTER_combination_with_existings_methods_of_inference}
+In this chapter, we describe how to incorporate the inferred statements to existing methods of XML schema inference.
 We focus on a class of methods which are based on a creation and subsequent simplification of the initial grammar (as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}).
@@
 The selection of elements is a simplified XPath evaluation.
-The simplification involves the aspects; the evaluation is not performed upon XML data, but the simplified grammar containing rules with a regular expression on their right side, and, partially related, predicates in path steps are not evaluated, they are ignored.
+The simplification involves two aspects; the evaluation is not performed upon XML data, but the simplified grammar containing rules with a regular expression on their right side, and, partially related, predicates in path steps are not evaluated, they are ignored.
 
@@
 
-%In the case of inferred types, the situation is not that simple, because there can occur conflicting statements. In %particular, two kinds of conflicts can emerge between two statements.
-%
-%\begin{enumerate}
-%\item The two types do not equal, but one is castable to the other.
-%\item The two types do not equal and none of them is castable to the other.
-%\end{enumerate}
-%
-%The example of conflict of the first kind for path $p$ without predicates is $p_1 \rightarrow date$, $p_2 \rightarrow %string$.
-And another one is $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when %stripped of predicates, they equal $p$.
-There are two reasons of this kind of conflicts.
-%
-%The first reason is that both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a %more convenient expression and is more precise.
-Consider these two expressions; $PathType$ $p$ is compared to an integer %literal constant, and, $PathType$ $p$ is an argument of function where a formal type of the argument is byte.
-The first %expression is utilized to infer statement $p \rightarrow integer$, and the second one to infer statement $p \rightarrow %byte$.
-Both of the types are correct, but the second one is more accurate.
+A problem emerges for example if PathType $p$ is compared to an integer constant, and the real type of elements (or attributes) selected by $p$ is \texttt{double}.
+In that case, statement $p \rightarrow$ \texttt{integer} is inferred, but it is not correct.
 
-From all inferred types for a particular path, the most specific (the most accurate) is selected and assigned to respective nodes from the grammar targeted by the path.
+\subsection{Verification using XML data}
+To solves those problems, we propose a simple verification using XML data.
+
+For each PathType $p$ that occurs in the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types.
+$T_p = \{t|(p \rightarrow t) \in \mathcal{S}_t\}$.
+Then, we create sequence $T_p'$ by ordering the set $T_p$ from the most specific type to the most general one.
+For example, if $T_p = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $T_p' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.
+
+Since we have the XML data and path $p$ is XQuery path, we can use an XQuery process (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$.
+The verification algorithms iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$.
+If so, $t'$ is the inferred type for nodes $N$ (and PathType $p$), else the inferred type is \texttt{string}.

### 2012-03-20 doc/theses/mikula/implementation.tex
Master thesis: In progres.

@@
 \chapter{Implementation}
-The solution proposed in Chapters \ref{CHAPTER_proposed_algorithm} and \ref{CHAPTER_combination_with_existings_methods_of_inference} was implemented using the jInfer framework \cite{jinfer}.
-It is a framework for implementing methods of XML schema inferring created as a software project at Faculty of Mathematics and Physics, Charles University in Prague.
+The solution proposed in Chapters \ref{CHAPTER_proposed_algorithm} and \ref{CHAPTER_combination_with_existings_methods_of_inference}, except the verification of types, was implemented using the jInfer framework \cite{jinfer}.
+It is a framework for implementing methods of XML schema inference created as a software project at Faculty of Mathematics and Physics, Charles University in Prague.
 It is written in Java as a plugin for NetBeans platform.
@@
 In particular, construction of a syntax tree, static analysis of expression types, inference of built-in types, and key discovery.
-On input, it takes the syntax trees, and its output are statements inferred from the queries represented by the syntax tress.
-\item \textbf{Merger} - A step of inference responsible for combining the simplified grammar with the statements inferred by \textbf{XQuery Processor} as described in Chapter \ref{CHAPTER_combination_with_existings_methods_of_inference}.
+On input, it takes the syntax trees, and its output are statements inferred from the syntax tress.
+\item \textbf{Merger} - A step of inference responsible for combining the simplified grammar with the statements inferred by \textbf{XQuery Processor} as described in Chapter \ref{CHAPTER_combination_with_existings_methods_of_inference}, except the verification of types.
 Its result is the simplified grammar extended by the statements from \textbf{XQuery Processor} in a way that each statement is analysed and its information is assigned to concerned elements and attributes from the grammar.
@@
 The modules run in a serial order as shown in Figure \ref{FIG_jinfer_modules}.
-The second difference is that one step is not always represented by one module, or one module is not always representing one step.
+The second difference is that one step is not always represented by one module, or one module is not always representing just one step.
 
-Actually, the blue modules shown in \ref{FIG_jinfer_modules} are module classes with specified interface.
+Actually, the blue modules shown in \ref{FIG_jinfer_modules} are module abstractions with specified interface.
 Actual modules are then implementations complying the interface.
@@
 
-As in the previous section, newly added modules are those shown in yellow.
+As in the previous section, newly added modules are those shown in yellow boxes.
 They are \textbf{XQuery Processor} and \textbf{XQuery Importer}.
-We also modified modules \textbf{Basic XSD Exporter} (implementation of module class \textbf{Schema Generator}) and \textbf{Base} to extend their functionality.
+We also modified modules \textbf{Basic XSD Exporter} (an implementation of \textbf{Schema Generator} interface) and \textbf{Base} to extend their functionality.
 
@@
 \item \textbf{Base} - This is module containing common classes used by other modules and defining interfaces.
-We added two packages to this module.
+We added five packages to this module.
 Package \texttt{cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes} implements the structure of syntax trees.
-And package \texttt{cz.cuni.mff.ksi\-.jinfer.base.objects.xquery.xqueryprocessor.types} contains data \\ structures and utilities for types defined and used by the proposed solution.
-\item \textbf{XQuery Importer} - This module represents \textbf{XQAnalyzer} step in Figure \ref{FIG_jinfer_steps}.
+Package \texttt{cz.cuni.mff.ksi.jinfer\-.base.interfaces.xquery} contains \texttt{Type} interface which is an interface for types used in the type analysis of syntax trees.
+Package \texttt{cz.cuni.mff.ksi\-.jinfer.base\-.objects.xquery.types} contains implementations of \texttt{Type} interface and type utility classes.
+Package \texttt{cz.cuni.mff.ksi.jinfer\-.base\-.objects.xquery.keys} provides representations of keys and foreign keys, and package \texttt{cz.cuni.mff.ksi\-.jinfer.base.objects.xsd} provides representation of XSD built-in atomic types.
+\item \textbf{XQuery Importer} - This module represents \textbf{XQAnalyzer} step in Figure \ref{FIG_jinfer_steps}. It is responsible for creating syntax trees from input XQuery files (step 1).
 \item \textbf{XQuery Processor} - The main module consisting of steps \textbf{XQuery Processor} and \textbf{Merger} in Figure \ref{FIG_jinfer_steps}.
-Package \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer} contains the main class \texttt{XQueryAnalyzerImpl} and classes implementing the steps of the proposed algorithm.
-Packages \texttt{cz.cuni.mff\-.ksi.jinfer\-.xqueryanalyzer.keys} and \texttt{cz.cuni.mff.ksi.jinfer-\\.xqueryanalyzer.joinpatterns} contain representation of keys and join patterns.
-And package \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer.utils} contains various utilities.
+Package \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer} contains the main module class \texttt{XQueryAnalyzerImpl}.
+Other packages are \texttt{cz.cuni.mff.ksi.jinfer.xqueryanalyzer.-\\expressiontypesanalysis} implementing the static analysis of expression types (step 2), \texttt{cz.cuni.mff.ksi.jinfer.xqueryanalyzer.builtintype\-inference} implementing the inference of XSD built-in types (step 3), \texttt{cz\-.cuni.mff\-.ksi.jinfer\-.xqueryanalyzer.keydiscovery} and their subpackages implementing the key discovery (step 4), \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer.merger} implementing the merging with the grammar, and \texttt{cz.cuni.mff.ksi.jinfer.xqueryanalyzer.utils} containing various utilities.
 \item \textbf{Basic XSD Exporter} - This is the module responsible for generating the resulting schema in XSD.

### 2012-03-20 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-The proposed algorithm consists of the following four main steps, described in detail in the rest of the chapter.
+\begin{figure}
+\label{FIG_steps_algorithm}
+\caption{Steps of the proposed algorithm.}
+\includegraphics[scale=1]{algorithm.eps}
+\end{figure}
+
+The proposed algorithm consists of the following four main steps, show in Figure \ref{FIG_steps_algorithm} and described in detail in the rest of the chapter.
 
 \begin{enumerate}
-\item Construction of a syntax tree. We use lexical and syntax analyses proposed in \cite{thesis_schejbal} and for each XQuery on input, we construct a data structure defined in \ref{DEF_syntax_tree}.
-\item Static Analysis of expression types. The algorithm searches for expressions in the syntax trees and statically (without evaluation) determines their types.
-\item Inference of built-in types. When the types of expressions are determined, selected forms of expression are utilized to infer types of elements and attributes.
-\item Key discovery. The final step is an extension of approach \cite{Necasky:2009:DXK:1529282.1529414} inferring keys and foreign keys.
+\item \textbf{Construction of a syntax tree}.
+We use lexical and syntax analyses proposed in \cite{thesis_schejbal} and for each XQuery on input, we construct a data structure defined in \ref{DEF_syntax_tree}.
+\item \textbf{Static Analysis of expression types}. The algorithm searches for expressions in the syntax trees and statically (without evaluation) determines their types.
+\item \textbf{Inference of built-in types}. When the types of expressions are determined, selected forms of expression are utilized to infer types of elements and attributes.
+\item \textbf{Key discovery}. The final step is an extension of approach \cite{Necasky:2009:DXK:1529282.1529414} inferring keys and foreign keys.
 \end{enumerate}
+
+As can be seen in Figure \ref{FIG_steps_algorithm}, step 4 is independent on step 3. They both depend on step 2.
 

### 2012-03-20 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-Some pseudo code algorithms in the following sections need to determine a class of a node.
+Some pseudo-code algorithms in the following sections need to determine a class of a node.
 The node's direct class can be determined by function $c$ from the definition of the syntax tree in Definition \ref{DEF_syntax_tree}, and hence, we also can determine its indirect classes.
@@
 \todo[inline]{priklad stromu}
+
+\subsection{Syntax Tree Example}
+
+\begin{lstlisting}[float, caption=Sample query 1., frame=single, label=listing_sample_query_1]
+declare namespace local = "http://www.foobar.org";
+declare function local:convert($v as xs:decimal?)
+  as xs:decimal?
+{
+  2.20371 * $v (: convert Dfl to Euro :)
+};
+
+for $i in /site/open_auctions/open_auction
+return local:convert(zero-or-one($i/reserve))
+\end{lstlisting}
+
+\begin{figure}
+\label{FIG_syntax_tree_1}
+\caption{Sample syntax tree 1.}
+\includegraphics[scale=0.6]{syntax_tree_1.eps}
+\end{figure}
+
+A syntax tree constructed from sample query in Listing \ref{listing_sample_query_1} is shown in Figure \ref{FIG_syntax_tree_1}.
 

### 2012-03-20 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max}, \texttt{avg}, \texttt{sum}, \texttt{distinct-values}.
+\todo{DOPLNIT}
 
@@
 
+\begin{figure}
+\label{FIG_syntax_tree_1_types}
+\caption{Sample syntax tree 1 after the static analysis of types.}
+\includegraphics[scale=0.6]{syntax_tree_1_types.eps}
+\end{figure}
+
 We can also determine expression types in functions.
@@
 These argument are set as the variable context for the function body represented by another subnode.
+
+Figure \ref{FIG_syntax_tree_1_types} shows the syntax tree from Figure \ref{FIG_syntax_tree_1} after the static analysis of expression types.
+The types are shown in red color.
+Note that the node representing \texttt{zero-or-one} function call is of a PathType type, as well as its argument.
+It is so, because the function returns its argument unchanged, and thus, we included the function in the special functions list in the PathType definition.
+
+Also note, that the function call node of \texttt{local:convert} in the FLWOR return clause has \texttt{zero-or-one} cardinality and the FLWOR node has \texttt{zero-or-more} cardinality.
+The change of cardinality is a result of the for unbinding shown in Algorithm \ref{ALG_function_create_for_unbound_type}.
 

### 2012-03-20 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 The motivation is that in some cases of function calls, we want to know that the function call is performed with an instance of PathType because then, we can determine a type of the function call more precisely.
-Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max}, \texttt{avg}, \texttt{sum}, \texttt{distinct-values}.
-\todo{DOPLNIT}
+Those special functions are built-in functions \texttt{data}, \texttt{min}, \texttt{max}, \texttt{avg}, \texttt{sum}, \texttt{distinct-values}, \texttt{zero-or-one}, \texttt{exactly-one}.
+And other may be added, when needed.
 
@@
 \caption{Sample syntax tree 1 after the static analysis of types.}
-\includegraphics[scale=0.6]{syntax_tree_1_types.eps}
+\includegraphics[scale=0.5]{syntax_tree_1_types.eps}
 \end{figure}
@@
 
-\subsection{Function Calls}
+\subsection{Function Calls}\label{SUBSECTION_function_calls}
 This case is quite straightforward.
-The algorithm encounters a function call and one of the arguments is a set of elements of attributes (subexpression $E$ representing $S$) represented by PathType $P$.
+The algorithm encounters a function call and one of the arguments is a set of elements of attributes (subexpression $E$ representing $S$) represented by PathType $p$.
 To determine the type of $S$, it is only needed to determine the type of the corresponding formal argument from the definition of the function.
@@
 
-If the type $T$ of the formal argument is a built-in type or its sequence, then $T$ is also the inferred type of $S$. The inferred statement is $P \rightarrow T$. Otherwise, no statement is inferred.
+If the type $t$ of the formal argument is a built-in type or its sequence, then $t$ is also the inferred type of $S$. The inferred statement is $p \rightarrow t$. Otherwise, no statement is inferred.
 
 \subsection{Arithmetic Operations}
-If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod}, \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD}, \texttt{MUL}, \texttt{DIV} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
-
-If the operator is one of \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, \texttt{!=} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{GEN\_EQUALS}, \\ \texttt{GEN\_NOT\_EQUALS}), one operand is of PathType $P$ and the type $T$ of the other operand is one of built-in types, then the inferred statement is $P \rightarrow T$.
+If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod}, \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD}, \texttt{MUL}, \texttt{DIV} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $p$ and the type $t$ of the other operand is one of numeric built-in types, then the inferred statement is $p \rightarrow t$.
+
+If the operator is one of \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, \texttt{!=} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{GEN\_EQUALS}, \\ \texttt{GEN\_NOT\_EQUALS}), one operand is of PathType $p$ and the type $t$ of the other operand is one of built-in types, then the inferred statement is $p \rightarrow t$.
+
+\subsection{Example}
+
+\begin{figure}
+\label{FIG_syntax_tree_1_types_type_inference}
+\caption{Example of an inference of a type from a function call.}
+\includegraphics[scale=1]{syntax_tree_1_types_type_inference.eps}
+\end{figure}
+
+Figure \ref{FIG_syntax_tree_1_types_type_inference} shows a fraction of the syntax tree from Figure \ref{FIG_syntax_tree_1_types}.
+When the inference of built-in types encounters the node marked by blue color, a statement will be inferred using the principle described in Chapter \ref{SUBSECTION_function_calls}.
+The node represents a call of \texttt{local:convert} function.
+
+This function has one formal argument of type $t =$ \texttt{xs:decimal}, \texttt{zero-or-one} cardinality.
+The real argument is of a PathType $p$.
+In particular, it is a PathType representing path \texttt{\$i/reserve}, where the \texttt{\$i} variable refers to a PathType representing for-bound \texttt{/site/open\_auctions/open\_auction} path, and one special function call of \texttt{zero-or-one} function is noted.
+
+Since the criteria for the inference from a function call are met, statement $p \rightarrow t$ is inferred.
 
@@
 in a stack of function calls) and do not process the recursive calls.
-
-\todo[inline]{Pěkné, neškodilo by proložit příklady, tj. dotaz, strom, co z něho odvodíme. -- To urobim pocas implementacie pre nejaky vzorovy dotaz.}
 

### 2012-03-21 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 from $N_1, \dots , N_n$. The normalized score $S_i$ of $K_i$ is computed as $S^{norm}_i = \frac{S_i}{S^{max}} * (1 - \frac{N^{max} - N_i}{\sum_{i=1}^{n}N_i})$
+
+\subsection{Example}
+\begin{lstlisting}[float, caption=Sample query 2., frame=single, label=listing_sample_query_2]
+for $p in /site/people/person
+let $a :=
+  for $t in /site/closed_auctions/closed_auction
+  where $t/buyer/@person = $p/@id
+  return $t
+return <item person="{$p/name/text()}">{count($a)}</item>
+\end{lstlisting}
+
+\begin{figure}
+\label{FIG_syntax_tree_2}
+\caption{Sample syntax tree 2.}
+\includegraphics[scale=0.7]{syntax_tree_2.eps}
+\end{figure}
+
+Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_sample_query_2}.
+In this tree, the algorithm founds one occurrence of the join pattern 3.
+The two for clauses and one where clause which form the occurrence are shown in the red boxes.
+
+The next phase marks this occurrence as (O1) case with weight 0.5 and the following statements are inferred:
+
+\begin{itemize}
+\item Key $k = (/site/people/person, \{@id\})$ is satisfied.
+\item Foreign key $((/site/closed\_auctions/closed\_auction, \{buyer/@person\}) \rightarrow k)$ is satisfied.
+\item Key $(/site/closed\_auctions/closed\_auction, \{buyer/@person\})$ is not satisfied.
+\end{itemize}

### 2012-03-21 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-The next phase marks this occurrence as (O1) case with weight 0.5 and the following statements are inferred:
-
 \begin{itemize}
-\item Key $k = (/site/people/person, \{@id\})$ is satisfied.
-\item Foreign key $((/site/closed\_auctions/closed\_auction, \{buyer/@person\}) \rightarrow k)$ is satisfied.
-\item Key $(/site/closed\_auctions/closed\_auction, \{buyer/@person\})$ is not satisfied.
+\item $P_1 =$ \texttt{/site/people/person}
+\item $P_2 =$ \texttt{/site/closed\_auctions/closed\_auction}
+\item $L_1 =$ \texttt{@id}
+\item $L_2 =$ \texttt{buyer/@person}
 \end{itemize}
+
+The next phase marks this occurrence as (O1) case with weight 0.5. Paths $P_1$ and $P_2$ does not have a common context, and thus, the following statements are inferred:
+
+\begin{itemize}
+\item Key $k = ($\texttt{/site/people/person}$, \{$\texttt{@id}$\})$ is satisfied.
+\item Foreign key $(($\texttt{/site/closed\_auctions/closed\_auction}$, \{$\texttt{buyer/@person}$\}) \rightarrow k)$ is satisfied.
+\item Key $($\texttt{/site/closed\_auctions/closed\_auction}$, \{$\texttt{buyer/@person}$\})$ is not satisfied.
+\end{itemize}

### 2012-03-22 doc/theses/mikula/experiments.tex
Master thesis: In progress.

\chapter{Experiments}
In this chapter we describe how we performed experiments with the implementation and what problems we encountered.

\section{Testing Data}
To get meaningful results of the experiments, testing data should be composed of XML documents, which are instances of a certain, possibly not known, XML schema, and a set of XQuery queries which query the XML documents. The amount of the XML data does not have to be large. On the other hand, the set of queries should be large (at least hundreds of queries) and the queries should be real, not artificially made.

In a search for such testing data, we have not succeeded. Large sets of XML data are available, but large sets of XQuery queries are not or it is not a simple task to obtain them.

If we cannot obtain an ideal set of testing data, we can at least try to find the most suitable one from available non-ideal sets.

Sets of XML data and XQuery queries can be found in W3C XML Query Use Cases \cite{Marchiori:07:XQU}. However, those are very small sets of queries and the analysis of XQuery in Chapter \ref{CHAPTER_analysis_of_xquery} was worked out using those queries, and thus, the relevancy of such testing data is questionable.

Another considered possibility was to obtain some set of XML data and create queries to it. This notion was rejected because such set would have all of the negative characteristics; it would be small, artificially made, and it would not be independent, as well.

At last, we concluded to use data provided by the XMark project \cite{xmark}. It consists of automatically generated XML data and a set of twenty XQuery queries related to the data. Although, this set is also very small, it is more or less real and we did not known the set in the process of developing the algorithm.

### 2012-03-22 doc/theses/mikula/diplomka.tex
Master thesis: In progress.

@@
 \usepackage[chapter]{algorithm}
+
+\usepackage{color}
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+\newcommand{\lstfont}{\small\ttfamily}
+\lstset{ %
+  language=,                % the language of the code
+  basicstyle=\lstfont,           % the size of the fonts that are used for the code
+  %numbers=left,                   % where to put the line-numbers
+  %numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
+  %stepnumber=2,                   % the step between two line-numbers. If it's 1, each line 
+                                  % will be numbered
+  numbersep=5pt,                  % how far the line-numbers are from the code
+  backgroundcolor=\color{white},      % choose the background color. You must add \usepackage{color}
+  showspaces=false,               % show spaces adding particular underscores
+  showstringspaces=false,         % underline spaces within strings
+  showtabs=false,                 % show tabs within strings adding particular underscores
+  frame=single,                   % adds a frame around the code
+  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
+  tabsize=2,                      % sets default tabsize to 2 spaces
+  captionpos=t,                   % sets the caption-position to bottom
+  breaklines=true,                % sets automatic line breaking
+  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
+  title=\lstname,                   % show the filename of files included with \lstinputlisting;
+                                  % also try caption instead of title
+  keywordstyle=\color{blue},          % keyword style
+  commentstyle=\color{dkgreen},       % comment style
+  stringstyle=\color{mauve},         % string literal style
+  escapeinside={\%*}{*)},            % if you want to add a comment within your code
+  morekeywords={*,...},              % if you want to add more keywords to the set
+}
 
@@
 \include{appendix_a}
+\include{appendix_b}
 

### 2012-03-22 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 
-\begin{lstlisting}[float, caption=Sample query 1., frame=single, label=listing_sample_query_1]
-declare namespace local = "http://www.foobar.org";
-declare function local:convert($v as xs:decimal?)
-  as xs:decimal?
-{
-  2.20371 * $v (: convert Dfl to Euro :)
-};
-
-for $i in /site/open_auctions/open_auction
-return local:convert(zero-or-one($i/reserve))
-\end{lstlisting}
-
 \begin{figure}
@@
 
-A syntax tree constructed from sample query in Listing \ref{listing_sample_query_1} is shown in Figure \ref{FIG_syntax_tree_1}.
+A syntax tree constructed from sample query in Listing \ref{listing_testing_query_18} is shown in Figure \ref{FIG_syntax_tree_1}.
 
@@
 \subsection{Example}
-\begin{lstlisting}[float, caption=Sample query 2., frame=single, label=listing_sample_query_2]
-for $p in /site/people/person
-let $a :=
-  for $t in /site/closed_auctions/closed_auction
-  where $t/buyer/@person = $p/@id
-  return $t
-return <item person="{$p/name/text()}">{count($a)}</item>
-\end{lstlisting}
-
 \begin{figure}
@@
 
-Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_sample_query_2}.
+Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_testing_query_8} (subtree of ConstructorNode is omitted because of lack of space).
 In this tree, the algorithm founds one occurrence of the join pattern 3.

### 2012-03-22 doc/theses/mikula/appendix_b.tex
Master thesis: In progress.

\chapter{Test Data} \label{APPENDIX_test_data}

Test data from the XMark project \cite{xmark}. Using the provided XML generator, XML document of size approximately 1.5 MB was generated. Figure \ref{listing_dtd} is its DTD and other figures in this appendix are XQuery queries that query the XML document.

The queries was slightly modified by replacing calls of \texttt{doc} function in paths by \texttt{/} (document node).

\begin{lstlisting}[caption=DTD of the test XML data, frame=single, label=listing_dtd]
<!-- DTD for auction database -->
<!-- $Id: auction.dtd,v 1.15 2001/01/29 21:42:35 albrecht Exp $ -->

<!ELEMENT site            (regions, categories, catgraph, people, open_auctions, closed_auctions)>

<!ELEMENT categories      (category+)>
<!ELEMENT category        (name, description)>
<!ATTLIST category        id ID #REQUIRED>
<!ELEMENT name            (#PCDATA)>
<!ELEMENT description     (text | parlist)>
<!ELEMENT text            (#PCDATA | bold | keyword | emph)*>
<!ELEMENT bold            (#PCDATA | bold | keyword | emph)*>
<!ELEMENT keyword         (#PCDATA | bold | keyword | emph)*>
<!ELEMENT emph            (#PCDATA | bold | keyword | emph)*>
<!ELEMENT parlist         (listitem)*>
<!ELEMENT listitem        (text | parlist)*>

<!ELEMENT catgraph        (edge*)>
<!ELEMENT edge            EMPTY>
<!ATTLIST edge            from IDREF #REQUIRED to IDREF #REQUIRED>

<!ELEMENT regions         (africa, asia, australia, europe, namerica, samerica)>
<!ELEMENT africa          (item*)>
<!ELEMENT asia            (item*)>
<!ELEMENT australia       (item*)>
<!ELEMENT namerica        (item*)>
<!ELEMENT samerica        (item*)>
<!ELEMENT europe          (item*)>
<!ELEMENT item            (location, quantity, name, payment, description, shipping, incategory+, mailbox)>
<!ATTLIST item            id ID #REQUIRED
                          featured CDATA #IMPLIED>
<!ELEMENT location        (#PCDATA)>
<!ELEMENT quantity        (#PCDATA)>
<!ELEMENT payment         (#PCDATA)>
<!ELEMENT shipping        (#PCDATA)>
<!ELEMENT reserve         (#PCDATA)>
<!ELEMENT incategory      EMPTY>
<!ATTLIST incategory      category IDREF #REQUIRED>
<!ELEMENT mailbox         (mail*)>
<!ELEMENT mail            (from, to, date, text)>
<!ELEMENT from            (#PCDATA)>
<!ELEMENT to              (#PCDATA)>
<!ELEMENT date            (#PCDATA)>
<!ELEMENT itemref         EMPTY>
<!ATTLIST itemref         item IDREF #REQUIRED>
<!ELEMENT personref       EMPTY>
<!ATTLIST personref       person IDREF #REQUIRED>

<!ELEMENT people          (person*)>
<!ELEMENT person          (name, emailaddress, phone?, address?, homepage?, creditcard?, profile?, watches?)>
<!ATTLIST person          id ID #REQUIRED>
<!ELEMENT emailaddress    (#PCDATA)>
<!ELEMENT phone           (#PCDATA)>
<!ELEMENT address         (street, city, country, province?, zipcode)>
<!ELEMENT street          (#PCDATA)>
<!ELEMENT city            (#PCDATA)>
<!ELEMENT province        (#PCDATA)>
<!ELEMENT zipcode         (#PCDATA)>
<!ELEMENT country         (#PCDATA)>
<!ELEMENT homepage        (#PCDATA)>
<!ELEMENT creditcard      (#PCDATA)>
<!ELEMENT profile         (interest*, education?, gender?, business, age?)>
<!ATTLIST profile         income CDATA #IMPLIED>
<!ELEMENT interest        EMPTY>
<!ATTLIST interest        category IDREF #REQUIRED>
<!ELEMENT education       (#PCDATA)>
<!ELEMENT income          (#PCDATA)>
<!ELEMENT gender          (#PCDATA)>
<!ELEMENT business        (#PCDATA)>
<!ELEMENT age             (#PCDATA)>
<!ELEMENT watches         (watch*)>
<!ELEMENT watch           EMPTY>
<!ATTLIST watch           open_auction IDREF #REQUIRED>

<!ELEMENT open_auctions   (open_auction*)>
<!ELEMENT open_auction    (initial, reserve?, bidder*, current, privacy?, itemref, seller, annotation, quantity, type, interval)>
<!ATTLIST open_auction    id ID #REQUIRED>
<!ELEMENT privacy         (#PCDATA)>
<!ELEMENT initial         (#PCDATA)>
<!ELEMENT bidder          (date, time, personref, increase)>
<!ELEMENT seller          EMPTY>
<!ATTLIST seller          person IDREF #REQUIRED>
<!ELEMENT current         (#PCDATA)>
<!ELEMENT increase        (#PCDATA)>
<!ELEMENT type            (#PCDATA)>
<!ELEMENT interval        (start, end)>
<!ELEMENT start           (#PCDATA)>
<!ELEMENT end             (#PCDATA)>
<!ELEMENT time            (#PCDATA)>
<!ELEMENT status          (#PCDATA)>
<!ELEMENT amount          (#PCDATA)>

<!ELEMENT closed_auctions (closed_auction*)>
<!ELEMENT closed_auction  (seller, buyer, itemref, price, date, quantity, type, annotation?)>
<!ELEMENT buyer           EMPTY>
<!ATTLIST buyer           person IDREF #REQUIRED>
<!ELEMENT price           (#PCDATA)>
<!ELEMENT annotation      (author, description?, happiness)>

<!ELEMENT author          EMPTY>
<!ATTLIST author          person IDREF #REQUIRED>
<!ELEMENT happiness       (#PCDATA)>
\end{lstlisting}

\begin{lstlisting}[caption=Test query 1., frame=single, label=listing_test_query_1]
for $b in /site/people/person[@id = "person0"] return $b/name/text()
\end{lstlisting}

\begin{lstlisting}[caption=Test query 2., frame=single, label=listing_test_query_2]
for $b in /site/open_auctions/open_auction
return <increase>{$b/bidder[1]/increase/text()}</increase>
\end{lstlisting}

\begin{lstlisting}[caption=Test query 3., frame=single, label=listing_test_query_3]
for $b in /site/open_auctions/open_auction
where zero-or-one($b/bidder[1]/increase/text()) * 2 <= $b/bidder[last()]/increase/text()
return
  <increase
  first="{$b/bidder[1]/increase/text()}"
  last="{$b/bidder[last()]/increase/text()}"/>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 4., frame=single, label=listing_test_query_4]
for $b in /site/open_auctions/open_auction
where
  some $pr1 in $b/bidder/personref[@person = "person20"],
       $pr2 in $b/bidder/personref[@person = "person51"]
  satisfies $pr1 << $pr2
return <history>{$b/reserve/text()}</history>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 5., frame=single, label=listing_test_query_5]
count(
  for $i in /site/closed_auctions/closed_auction
  where $i/price/text() >= 40
  return $i/price
)
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 6., frame=single, label=listing_test_query_6]
for $b in //site/regions return count($b//item)
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 7., frame=single, label=listing_test_query_7]
for $p in /site
return
  count($p//description) + count($p//annotation) + count($p//emailaddress)
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 8., frame=single, label=listing_test_query_8]
for $p in /site/people/person
let $a :=
  for $t in /site/closed_auctions/closed_auction
  where $t/buyer/@person = $p/@id
  return $t
return <item person="{$p/name/text()}">{count($a)}</item>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 9., frame=single, label=listing_test_query_9]
let $ca := /site/closed_auctions/closed_auction return
let
    $ei := /site/regions/europe/item
for $p in /site/people/person
let $a :=
  for $t in $ca
  where $p/@id = $t/buyer/@person
  return
    let $n := for $t2 in $ei where $t/itemref/@item = $t2/@id return $t2
    return <item>{$n/name/text()}</item>
return <person name="{$p/name/text()}">{$a}</person>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 10., frame=single, label=listing_test_query_10]
for $i in
  distinct-values(/site/people/person/profile/interest/@category)
let $p :=
  for $t in /site/people/person
  where $t/profile/interest/@category = $i
  return
    <personne>
      <statistiques>
        <sexe>{$t/profile/gender/text()}</sexe>
        <age>{$t/profile/age/text()}</age>
        <education>{$t/profile/education/text()}</education>
        <revenu>{fn:data($t/profile/@income)}</revenu>
      </statistiques>
      <coordonnees>
        <nom>{$t/name/text()}</nom>
        <rue>{$t/address/street/text()}</rue>
        <ville>{$t/address/city/text()}</ville>
        <pays>{$t/address/country/text()}</pays>
        <reseau>
          <courrier>{$t/emailaddress/text()}</courrier>
          <pagePerso>{$t/homepage/text()}</pagePerso>
        </reseau>
      </coordonnees>
      <cartePaiement>{$t/creditcard/text()}</cartePaiement>
    </personne>
return <categorie>{<id>{$i}</id>, $p}</categorie>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 11., frame=single, label=listing_test_query_11]
for $p in /site/people/person
let $l :=
  for $i in /site/open_auctions/open_auction/initial
  where $p/profile/@income > 5000 * exactly-one($i/text())
  return $i
return <items name="{$p/name/text()}">{count($l)}</items>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 12., frame=single, label=listing_test_query_12]
for $p in /site/people/person
let $l :=
  for $i in /site/open_auctions/open_auction/initial
  where $p/profile/@income > 5000 * exactly-one($i/text())
  return $i
where $p/profile/@income > 50000
return <items person="{$p/profile/@income}">{count($l)}</items>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 13., frame=single, label=listing_test_query_13]
for $i in /site/regions/australia/item
return <item name="{$i/name/text()}">{$i/description}</item>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Testing query 14., frame=single, label=listing_test_query_14]
for $i in /site//item
where contains(string(exactly-one($i/description)), "gold")
return $i/name/text()
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 15., frame=single, label=listing_test_query_15]
for $a in
  /site/closed_auctions/closed_auction/annotation/description/parlist/
   listitem/
   parlist/
   listitem/
   text/
   emph/
   keyword/
   text()
return <text>{$a}</text>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 16., frame=single, label=listing_test_query_16]
for $a in /site/closed_auctions/closed_auction
where
  not(
    empty(
      $a/annotation/description/parlist/listitem/parlist/listitem/text/emph/
       keyword/
       text()
    )
  )
return <person id="{$a/seller/@person}"/>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 17., frame=single, label=listing_test_query_17]
for $p in /site/people/person
where empty($p/homepage/text())
return <person name="{$p/name/text()}"/>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 18., frame=single, label=listing_test_query_18]
declare namespace local = "http://www.foobar.org";
declare function local:convert($v as xs:decimal?) as xs:decimal?
{
  2.20371 * $v (: convert Dfl to Euro :)
};

for $i in /site/open_auctions/open_auction
return local:convert(zero-or-one($i/reserve))
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 19., frame=single, label=listing_test_query_19]
for $b in /site/regions//item
let $k := $b/name/text()
order by zero-or-one($b/location) ascending empty greatest
return <item name="{$k}">{$b/location/text()}</item>
\end{lstlisting}

\begin{lstlisting}[float=htpb, caption=Test query 20., frame=single, label=listing_test_query_20]
<result>
  <preferred>
    {count(/site/people/person/profile[@income >= 100000])}
  </preferred>
  <standard>
    {
      count(
        /site/people/person/
         profile[@income < 100000 and @income >= 30000]
      )
    }
  </standard>
  <challenge>
    {count(/site/people/person/profile[@income < 30000])}
  </challenge>
  <na>
    {
      count(
        for $p in /site/people/person
        where empty($p/profile/@income)
        return $p
      )
    }
  </na>
</result>
\end{lstlisting}

### 2012-03-22 doc/theses/mikula/experiments.tex
Master thesis: In progress.

@@
 
-\section{Testing Data}
-To get meaningful results of the experiments, testing data should be composed of XML documents, which are instances of a certain, possibly not known, XML schema, and a set of XQuery queries which query the XML documents.
+\section{Test Data}
+To get meaningful results of the experiments, test data should be composed of XML documents, which are instances of a certain, possibly not known, XML schema, and a set of XQuery queries which query the XML documents.
 The amount of the XML data does not have to be large.
@@
 
-In a search for such testing data, we have not succeeded. Large sets of XML data are available, but large sets of XQuery queries are not or it is not a simple task to obtain them.
+In a search for such test data, we have not succeeded. Large sets of XML data are available, but large sets of XQuery queries are not or it is not a simple task to obtain them.
 
-If we cannot obtain an ideal set of testing data, we can at least try to find the most suitable one from available non-ideal sets.
+If we cannot obtain an ideal set of test data, we can at least try to find the most suitable one from available non-ideal sets.
 
 Sets of XML data and XQuery queries can be found in W3C XML Query Use Cases \cite{Marchiori:07:XQU}.
-However, those are very small sets of queries and the analysis of XQuery in Chapter \ref{CHAPTER_analysis_of_xquery} was worked out using those queries, and thus, the relevancy of such testing data is questionable.
+However, those are very small sets of queries and the analysis of XQuery in Chapter \ref{CHAPTER_analysis_of_xquery} was worked out using those queries, and thus, the relevancy of such test data is questionable.
 
@@
 At last, we concluded to use data provided by the XMark project \cite{xmark}.
-It consists of automatically generated XML data and a set of twenty XQuery queries related to the data.
+They are attached in Appendix \ref{APPENDIX_test_data} and they consists of automatically generated XML data and a set of twenty XQuery queries related to the data.
 Although, this set is also very small, it is more or less real and we did not known the set in the process of developing the algorithm.
+
+\section{Results}
+
+\subsection{Type Statements Inference}
+The following six type statements were inferred.\\
+\texttt{/site/open\_auctions/open\_auction/bidder/increase/text()} $\rightarrow$ \texttt{integer}\\
+\texttt{/site/closed\_auctions/closed\_auction/price/text()} $\rightarrow$ \texttt{integer}\\
+\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer}\\
+\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer}\\
+\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{integer}\\
+\texttt{/site/open\_auctions/open\_auction/reserve} $\rightarrow$ \texttt{decimal}
+
+Only the sixth statement is correct. Other five are incorrect, because, comparing to the data, real type of nodes selected by the paths is \texttt{decimal}, as well.
+
+To reveal the cause of the incorrect type inference, see, for example, query in Listing \ref{listing_test_query_12}.
+In the where clause, values of \texttt{/site/people/person/profile/@income} are compared to the integer literal constant \texttt{50000}.
+From this expression, it is not possible to infer the type correctly.
+Problem is that this is the only inferred statement.
+Better results can be achieved by providing a larger set of input queries, containing also expressions that can be exploited to infer correct statements.
+Then the verification with data can be incorporated to choose the correct statements.
+
+\subsection{Key Discovery}

### 2012-03-22 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/keydiscovery/negativeuniqueness/NegativeUniquenessFinder.java
Master thesis implementation: In progress.

@@ -104,7 +104,7 @@ public class NegativeUniquenessFinder {
         }
 
         final PathType pathType = (PathType) type;
-        if (PathTypeUtils.usesOnlyChildAndDescendantAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) {
+        if (PathTypeUtils.usesOnlyChildAndDescendantAndAttributeAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) { // TODO rio pridana attr axe, do DP!!
           memorizeNegativeWeight(pathType, 100);
         }
       }

### 2012-03-22 jinfer/branches/thesis-mikula/Base/src/cz/cuni/mff/ksi/jinfer/base/objects/xquery/types/NormalizedPathType.java
Master thesis implementation: In progress.

@@ -21,12 +21,15 @@ import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.AxisNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.ExprNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.InitialStep;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.ItemTypeNode;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.KindTestNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.NameTestNode;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.NodeKind;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.SelfOrDescendantStepNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.StepExprNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.VarRefNode;
 import java.util.ArrayList;
 import java.util.List;
+import org.apache.log4j.Logger;
 
 /**
  * A representation of a path type in a normalized form. It means that all
@@ -45,6 +48,8 @@ import java.util.List;
  * @author rio
  */
 public class NormalizedPathType {
+  
+  private final static Logger LOG = Logger.getLogger(NormalizedPathType.class);
 
   private final List<StepExprNode> steps;
   private InitialStep initialStep;
@@ -173,6 +178,10 @@ public class NormalizedPathType {
     boolean isFirstNode = true;
     boolean printChildAxis = true;
     
+    if (getInitialStep() == InitialStep.ROOT) {
+      stringBuilder.append("/");
+    }
+    
     for (final StepExprNode step : steps) {
             
       if (step instanceof SelfOrDescendantStepNode) {
@@ -204,8 +213,18 @@ public class NormalizedPathType {
         }
         
         final ItemTypeNode itemTypeNode = axisNode.getNodeTestNode();
-        assert(NameTestNode.class.isInstance(itemTypeNode));
-        stringBuilder.append(((NameTestNode)itemTypeNode).getName());
[…]

### 2012-03-22 jinfer/branches/thesis-mikula/Base/src/cz/cuni/mff/ksi/jinfer/base/objects/xquery/types/PathType.java

@@ -17,7 +17,6 @@
 package cz.cuni.mff.ksi.jinfer.base.objects.xquery.types;
 
 import cz.cuni.mff.ksi.jinfer.base.interfaces.xquery.Type;
-import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.AbstractType;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.ExprNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.InitialStep;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.PathExprNode;

### 2012-03-22 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/XQueryAnalyzerImpl.java

@@ -25,9 +25,13 @@ import cz.cuni.mff.ksi.jinfer.base.interfaces.inference.XQueryProcessor;
 import cz.cuni.mff.ksi.jinfer.base.interfaces.inference.XQueryProcessorCallback;
 import cz.cuni.mff.ksi.jinfer.base.objects.Input;
 import cz.cuni.mff.ksi.jinfer.base.objects.nodes.Element;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.InitialStep;
 import cz.cuni.mff.ksi.jinfer.base.utils.BaseUtils;
 import cz.cuni.mff.ksi.jinfer.base.utils.FileUtils;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.ModuleNode;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.NormalizedPathType;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.XSDType;
+import cz.cuni.mff.ksi.jinfer.base.objects.xsd.XSDBuiltinAtomicType;
 import cz.cuni.mff.ksi.jinfer.xqueryanalyzer.builtintypeinference.InferredTypeStatement;
 import cz.cuni.mff.ksi.jinfer.xqueryanalyzer.keydiscovery.summary.SummarizedKey;
 import java.io.File;
@@ -77,7 +81,19 @@ public class XQueryAnalyzerImpl implements XQueryProcessor {
       keysInferrer.addNegativeUniquenessStatements(syntaxTreeProcessor.getNegativeUniquenessStatements());
     }
     
-    LOG.info("Total Number of inferred type statements: " + inferredTypes.size());
+    // TODO rio do DP, okomentovat!!
+    final List<InferredTypeStatement> inferredTypeAbsolutePaths = new ArrayList<InferredTypeStatement>();
+    for (final InferredTypeStatement inferredTypeStatement : inferredTypes) {
+      final NormalizedPathType path = new NormalizedPathType(inferredTypeStatement.getPathType());
+      final XSDBuiltinAtomicType inferredType = ((XSDType)inferredTypeStatement.getType()).getAtomicType();
+      if (path.getInitialStep() == InitialStep.ROOT) {
+        inferredTypeAbsolutePaths.add(inferredTypeStatement);
+        LOG.info("Inferred type statement: " + path.toString() + " -> " + inferredType);
+      } else {
[…]

### 2012-03-22 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/keydiscovery/summary/KeySummarizer.java

@@ -19,6 +19,7 @@ package cz.cuni.mff.ksi.jinfer.xqueryanalyzer.keydiscovery.summary;
 import cz.cuni.mff.ksi.jinfer.xqueryanalyzer.keydiscovery.negativeuniqueness.NegativeUniquenessStatement;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.keys.Key;
 import cz.cuni.mff.ksi.jinfer.xqueryanalyzer.keydiscovery.weightedkeys.WeightedKey;
+import cz.cuni.mff.ksi.jinfer.xqueryanalyzer.utils.PathTypeUtils;
 import java.util.Collection;
 import java.util.HashMap;
 import java.util.List;
@@ -83,7 +84,7 @@ public class KeySummarizer {
     
     for (final NegativeUniquenessStatement nus : negativeUniquenessStatements) {
       if ((nus.getContextPath() == null && key.getContextPath() == null || nus.getContextPath().equals(key.getContextPath()))
-              && nus.getTargetPath().equals(key.getTargetPath())) {
+              && nus.getTargetPath().equals(PathTypeUtils.join(key.getTargetPath(), key.getKeyPath()))) { // TODO rio do DP!!
         weight -= nus.getWeight();
       }
     }

### 2012-03-22 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/utils/PathTypeUtils.java

@@ -18,11 +18,15 @@ package cz.cuni.mff.ksi.jinfer.xqueryanalyzer.utils;
 
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.AxisKind;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.ExprNode;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.InitialStep;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.StepExprNode;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.syntaxtree.nodes.VarRefNode;
+import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.NormalizedPathType;
 import cz.cuni.mff.ksi.jinfer.base.objects.xquery.types.PathType;
+import java.util.ArrayList;
 import java.util.List;
 import java.util.Map;
+import sun.swing.SwingUtilities2.AATextInfo;
 
 /**
  * An utility class providing PathType functions.
@@ -97,6 +101,31 @@ public class PathTypeUtils {
 
     return true;
   }
+  
+  /**
+   * Determines if a path uses only axis child, descendant, descendant-or-self, and attribute.
+   */
+  public static boolean usesOnlyChildAndDescendantAndAttributeAxes(final PathType pathType) { // TODO do DP!!
+    for (final StepExprNode stepNode : pathType.getSteps()) {
+      if (stepNode.isAxisStep()) {
+        final AxisKind axisKind = stepNode.getAxisNode().getAxisKind();
+        if (axisKind != AxisKind.CHILD && axisKind != AxisKind.DESCENDANT && axisKind != AxisKind.DESCENDANT_OR_SELF && axisKind != AxisKind.ATTRIBUTE) {
+          return false;
+        }
+      }
+
+      final ExprNode detailNode = stepNode.getDetailNode();
+      if (detailNode != null) {
+        if (VarRefNode.class.isInstance(detailNode)) {
+          if (usesOnlyChildAndDescendantAndAttributeAxes(pathType.getSubpaths().get(stepNode)) == false) {
+            return false;
+          }
+        }
+      }
+    }
+
+    return true;
+  }
 
   /**
    * If a path has just one predicate and it is in its last step, returns this
@@ -115,5 +144,16 @@ public class PathTypeUtils {
     }
 
[…]

### 2012-03-23 doc/theses/mikula/experiments.tex
Master thesis: In progress.

@@
 
-\subsection{Type Statements Inference}
-The following six type statements were inferred.\\
-\texttt{/site/open\_auctions/open\_auction/bidder/increase/text()} $\rightarrow$ \texttt{integer}\\
-\texttt{/site/closed\_auctions/closed\_auction/price/text()} $\rightarrow$ \texttt{integer}\\
-\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer}\\
-\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer}\\
-\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{integer}\\
-\texttt{/site/open\_auctions/open\_auction/reserve} $\rightarrow$ \texttt{decimal}
+\subsection{Type Inference}
+Six type statements shown in Table \ref{TAB_inferred_types} were inferred.
 
-Only the sixth statement is correct. Other five are incorrect, because, comparing to the data, real type of nodes selected by the paths is \texttt{decimal}, as well.
+\begin{table}
+\label{TAB_inferred_types}
+\caption{Inferred type statements.}
+\begin{tabular}{|l|}
+\hline
+\texttt{/site/open\_auctions/open\_auction/bidder/increase/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/closed\_auctions/closed\_auction/price/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/open\_auctions/open\_auction/reserve} $\rightarrow$ \texttt{decimal} \\ \hline
+\end{tabular}
+\end{table}
+
+Only the last inferred statement is correct. Others are incorrect, because, comparing to the data, real type of nodes selected by the paths is \texttt{decimal}, as well.
 
 To reveal the cause of the incorrect type inference, see, for example, query in Listing \ref{listing_test_query_12}.
-In the where clause, values of \texttt{/site/people/person/profile/@income} are compared to the integer literal constant \texttt{50000}.
+In the where clause, values of \texttt{/site/people/person/profile/\-@income} are compared to the integer literal constant \texttt{50000}.
 From this expression, it is not possible to infer the type correctly.
-Problem is that this is the only inferred statement.
+The problem is that this is the only inferred statement.
 Better results can be achieved by providing a larger set of input queries, containing also expressions that can be exploited to infer correct statements.
@@
 \subsection{Key Discovery}
+Four key statements and their normalized weights shown in Table \ref{TAB_inferred_keys} were inferred.
+
+\begin{table}
+\label{TAB_inferred_keys}
+\caption{Inferred key statements.}
+\begin{tabular}{|l|c|}
+\hline
+\textbf{Key} & \textbf{Weight} \\ \hline \hline
+\texttt{(/site/closed\_auctions/closed\_auction, \{buyer/@person\})} & -1.0 \\ \hline
+\texttt{(/site/regions/europe/item, \{@id\})} & -0.417 \\ \hline
+\texttt{(/site/people/person, \{@id\})} & 1.0 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction, \{itemref/@item\})} & 0.417 \\ \hline
+\end{tabular}
+\end{table}
+
+The first one is correct, a buyer is not a key of closed auctions.
+The second one is not correct, because \texttt{id} attribute is a key of \texttt{item} elements.
+The third one is correct and the forth one declares \texttt{itemref} element to be a key of closed auctions, which is not true, but only with weight 0.417.
+
+Closer analysis of input queries reveals that all of the statements were inferred from occurrences of the join pattern 3. That knowledge leads to the two following observations.
+
+\begin{itemize}
+\item The original method of key discovery would not infer anything on this input data.
+\item The cause of the incorrectly inferred statements are not that they were inferred from join patterns occurrences, where a join is not done by a key/foreign key pair.
+It is that the for clauses (definitions of paths $P_1$ and $P_2$) in the join pattern 3 occurrence (in query in Listing \ref{listing_test_query_9}) are swapped, so the real key is considered as a foreign key and vice-versa.
+\end{itemize}
+
+A partial solution is an extension of the test data by queries containing expression that can be exploited to infer negative uniqueness statements.
+Such expression is, for example,\\ \texttt{distinct-values(/site/closed\_auctions/closed\_auction/itemref/@item)}\\ to get unique ids of items that was sold in some auction.
+Since, one item may be sold several times in auctions organized in different time periods, \texttt{distinct-values} function is applied.
+
+From the original data, one negative uniqueness statement was inferred:\\
+\texttt{/site/people/person/profile/interest/@category} is not unique with weight 1, and it it correct.
+Since, there is not such key statement inferred, the negative uniqueness statement is not used to any correction of weight.
+
+\begin{table}
+\label{TAB_inferred_keys_2}
+\caption{Key statements inferred from the extended test data.}
+\begin{tabular}{|l|c|}
+\hline
+\textbf{Key} & \textbf{Weight} \\ \hline \hline
+\texttt{(/site/closed\_auctions/closed\_auction, \{buyer/@person\})} & -1.0 \\ \hline
+\texttt{(/site/regions/europe/item, \{@id\})} & -0.417 \\ \hline
+\texttt{(/site/people/person, \{@id\})} & 1.0 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction, \{itemref/@item\})} & -0.417 \\ \hline
+\end{tabular}
+\end{table}
+
+From data extended by the mentioned expression, another negative uniqueness was inferred:\\
+\texttt{/site/closed\_auctions/closed\_auction/itemref/@item} with weight 1.
+This one decreases the weight of the falsely inferred key.
+Key statements inferred from the extended data are shown in Table \ref{TAB_inferred_keys_2}.
+
+As was demonstrated, larger sets of input data may lead to better results.
+However, the problem with the falsely rejected key still remains.
+It may be solved by modification of the statements inferred from occurrences of the join pattern 3.
+If we omit the negative statement of a key from the second for clause ($(P_2, \{L_2\})$ is not satisfied), we get better results.
+The modified statements inferred from an occurrence of the join pattern 3 are the following (assigned weight remains 0.5):
+
+\begin{itemize}
+\item $(P_1, \{L_1\})$ is satisfied
+\item $(P_2, \{L_2\}) \rightarrow (P_1, \{L_1\})$ is satisfied
+\end{itemize}
+
+From the extended test data, using the modified join pattern 3 statements, key statements in Table \ref{TAB_inferred_keys_3} were inferred.
+Those are the best results from all test runs, though it is not clear that the modified statements will produce the best results also on other larger input data.
+Therefore, further tests with large sets of queries are required to determine the best settings for the algorithm.
+
+\begin{table}
+\label{TAB_inferred_keys_3}
+\caption{Key statements inferred from the extended test data using the modified JP3 statements.}
+\begin{tabular}{|l|c|}
+\hline
+\textbf{Key} & \textbf{Weight} \\ \hline \hline
+\texttt{(/site/people/person, \{@id\})} & 1.0 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction, \{itemref/@item\})} & -0.666 \\ \hline
+\end{tabular}
+\end{table}

### 2012-03-24 doc/theses/mikula/conclusion.tex
Master thesis: In progress.

@@
 And we slightly extended the one existing method dealing with the inference of keys.
-\todo{Ako dopadli testy.}
+We experimentally demonstrated that on some input files, results of the extended method are better than results of the original method.
+However, we did not succeed in a search of an ideal, large enough set of test data.
+The testing was performed on a small set of input queries and further testing and algorithm tuning is required.
 
@@
 Thus, we created the first complete, ready-to-use, and extensible implementation of the XML xchema inference exploiting XQuery queries besides XML data.
+
+\todo[inline]{Nejake prehladne zhrnutie plusov a minusov?}
 
@@
 And, a very large space for a possible future research is provided by utilisation of other XML operations. The main representant is XSLT.
+
+Besides the mentioned, our opinion is that the most urgent future work is obtaining a large enough test data set, performing proper experiments, and refining the algorithm by modification of its settings (statements inferred from join pattern occurrences, weights, etc.) according to experimental results.
+This process was slightly suggested in Chapter \ref{CHAPTER_experiments}.

### 2012-03-24 doc/theses/mikula/introduction.tex
Master thesis: In progress.

@@
 
-\section{Structure of the Work}
-\todo[inline]{TODO}
+\section{Structure of the Thesis}
+The thesis is structured as follows.
+
+In Chapter 2, we introduce XML technologies used in this thesis and we briefly explain their basic principles.
+
+Chapter 3 contains summaries of several existing approaches of XML schema inference. One of them proposes an utilisation of XQuery queries, which interests us the most.
+
+We made an analysis of XQuery technology from a view of XML schema inference, where we suggest several ideas on how XQuery queries can be exploited in the inference process.
+This is a content of Chapter 4.
+
+In Chapter 5, we discusses some essential questions that we needed to decide before a development of an algorithm.
+
+The algorithm itself is proposed in Chapter 6. And Chapter 7 describes how to combine results of the algorithm with existing grammar inferring methods of XML schema inference.
+
+An implementation of the proposed algorithm using a jInfer framework is described in Chapter 8. Chapter 9 then discusses performed experiments and their outcomes.

### 2012-03-24 doc/theses/mikula/experiments.tex
Master thesis: In progress.

@@
-\chapter{Experiments}
-In this chapter we describe how we performed experiments with the implementation and what problems we encountered.
+\chapter{Experiments} \label{CHAPTER_experiments}
+In this chapter we describe how we performed experiments with the implementation and what problems we faced.
 
@@
 \texttt{(/site/people/person, \{@id\})} & 1.0 \\ \hline
-\texttt{(/site/closed\_auctions/closed\_auction, \{itemref/@item\})} & -0.666 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction, \{itemref/@item\})} & -0.333 \\ \hline
 \end{tabular}
 \end{table}
+
+\todo[inline]{Pripojit vysledne XSD? Je dost velke. Alebo aspon relevantnu cast?}

### 2012-03-24 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 For detail, see \cite{thesis_schejbal}.
-
-\todo[inline]{Hodil by se příklad stromu - jeslti nebude dál.}
 
@@
 
-\todo[inline]{priklad stromu}
-
 \subsection{Syntax Tree Example}
@@
 
-A syntax tree constructed from sample query in Listing \ref{listing_testing_query_18} is shown in Figure \ref{FIG_syntax_tree_1}.
+A syntax tree constructed from sample query in Listing \ref{listing_test_query_18} is shown in Figure \ref{FIG_syntax_tree_1}.
 
@@
 
-Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_testing_query_8} (subtree of ConstructorNode is omitted because of lack of space).
+Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_test_query_8} (subtree of ConstructorNode is omitted because of lack of space).
 In this tree, the algorithm founds one occurrence of the join pattern 3.

### 2012-03-24 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/builtintypeinference/BuiltinTypesInferrer.java
Master thesis implementation: In progress.

@@ -53,6 +53,11 @@ public class BuiltinTypesInferrer {
       // To process function body nodes, we need know the real arguments.
       // It is not implemented yet.
       return;
+    } else if (node instanceof PathExprNode) {
+      // We do not want to dive to PathExprNode instances, because it results
+      // in inference of types for parts of paths. For example /n[@a = 5] infers
+      // @a -> integer.
+      return;
     }
     
     final List<XQNode> subnodes = node.getSubnodes();

### 2012-03-24 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/keydiscovery/summary/KeySummarizer.java

@@ -84,7 +84,7 @@ public class KeySummarizer {
     
     for (final NegativeUniquenessStatement nus : negativeUniquenessStatements) {
       if ((nus.getContextPath() == null && key.getContextPath() == null || nus.getContextPath().equals(key.getContextPath()))
-              && nus.getTargetPath().equals(PathTypeUtils.join(key.getTargetPath(), key.getKeyPath()))) { // TODO rio do DP!!
+              && nus.getTargetPath().equals(PathTypeUtils.join(key.getTargetPath(), key.getKeyPath()))) {
         weight -= nus.getWeight();
       }
     }

### 2012-03-24 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/XQueryAnalyzerImpl.java
Master thesis implementation: In progress.

@@ -86,19 +86,15 @@ public class XQueryAnalyzerImpl implements XQueryProcessor {
       }
     }
     
-    // TODO rio do DP, okomentovat!!
-    final List<InferredTypeStatement> inferredTypeAbsolutePaths = new ArrayList<InferredTypeStatement>();
     for (final InferredTypeStatement inferredTypeStatement : inferredTypes) {
       final NormalizedPathType path = new NormalizedPathType(inferredTypeStatement.getPathType());
       final XSDBuiltinAtomicType inferredType = ((XSDType)inferredTypeStatement.getType()).getAtomicType();
-      if (path.getInitialStep() == InitialStep.ROOT) {
-        inferredTypeAbsolutePaths.add(inferredTypeStatement);
-        LOG.info("Inferred type statement: " + path.toString() + " -> " + inferredType);
-      } else {
-        LOG.debug("Inferred type statement (skipped because the path is not absolute): " + path.toString() + " -> " + inferredType);
+      if (path.getInitialStep() != InitialStep.ROOT) {
+        LOG.error("Inferred type statement should not contain a relative path. Probably a bug.");
       }
+      LOG.info("Inferred type statement: " + path.toString() + " -> " + inferredType);
     }
-    LOG.info("Total number of inferred type statements: " + inferredTypeAbsolutePaths.size());
+    LOG.info("Total number of inferred type statements: " + inferredTypes.size());
     
     keysInferrer.summarize();
     final Collection<SummarizedKey> keys = keysInferrer.getKeys();
@@ -115,7 +111,7 @@ public class XQueryAnalyzerImpl implements XQueryProcessor {
     LOG.info("Total number of inferred key statements: " + keys.size());
     
     final Merger merger = new Merger(grammar);
-    merger.mergeInferredTypes(inferredTypeAbsolutePaths);
+    merger.mergeInferredTypes(inferredTypes);
     merger.mergeInferredKeys(keys, foreignKeys);
     
     callback.finished(grammar);

### 2012-03-24 jinfer/branches/thesis-mikula/XQueryProcessor/src/cz/cuni/mff/ksi/jinfer/xqueryanalyzer/keydiscovery/negativeuniqueness/NegativeUniquenessFinder.java

@@ -104,7 +104,7 @@ public class NegativeUniquenessFinder {
         }
 
         final PathType pathType = (PathType) type;
-        if (PathTypeUtils.usesOnlyChildAndDescendantAndAttributeAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) { // TODO rio pridana attr axe, do DP!!
+        if (PathTypeUtils.usesOnlyChildAndDescendantAndAttributeAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) {
           memorizeNegativeWeight(pathType, 100);
         }
       }
@@ -149,7 +149,7 @@ public class NegativeUniquenessFinder {
       if (leftOperand.getType().getCategory() == Type.Category.PATH
               && LiteralNode.class.isInstance(rightOperand)) {
         final PathType pathType = (PathType) leftOperand.getType();
-        if (PathTypeUtils.usesOnlyChildAndDescendantAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) {
+        if (PathTypeUtils.usesOnlyChildAndDescendantAndAttributeAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) {
           for (final String var : forVars) {
             final ExprNode detailNode = pathType.getSteps().get(0).getDetailNode();
             if (detailNode != null) {
@@ -164,7 +164,7 @@ public class NegativeUniquenessFinder {
       } else if (rightOperand.getType().getCategory() == Type.Category.PATH
               && LiteralNode.class.isInstance(leftOperand)) {
         final PathType pathType = (PathType) rightOperand.getType();
-        if (PathTypeUtils.usesOnlyChildAndDescendantAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) {
+        if (PathTypeUtils.usesOnlyChildAndDescendantAndAttributeAxes(pathType) && PathTypeUtils.isWithoutPredicatesExceptLastStep(pathType)) {
           for (final String var : forVars) {
             final ExprNode detailNode = pathType.getSteps().get(0).getDetailNode();
             if (detailNode != null) {

### 2012-03-24 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \section{Step 3: Inference of Built-in Types}
+\todo[inline]{Rekurzia sa nevnara do FunctionBodyNodes a PathExprNodes - napisat, alebo je to detail?}
 In this step, the algorithm goes through the syntax tree to infer types of elements and attributes from expressions using the type information from the previous step.
@@
 \subsection{Auxiliary Functions} \label{step_4_auxiliary_functions}
+\todo[inline]{Ako je to s osou descendant or self pri hladani join patternov? Orig. metoda o nej explicitne nehovori, ale zrejme pouziva}
+\todo[inline]{Ako je to s attribute osou v L cestach join patternov? Atribut predsa moze byt kluc.}
 Firstly, we define some auxiliary functions used by algorithms in this chapter.
 
-Function \texttt{usesOnlyChildAndDescendantAxes} in Algorithm \ref{ALG_function_uses_only_child_and_descendant_axes} takes an instance of PathType as its arguments and returns true if the path uses only child and descendant axes.
+Function \texttt{usesOnlyChildAndDescendantAxes} in Algorithm \ref{ALG_function_uses_only_child_and_descendant_axes} takes an instance of PathType as its arguments and returns true if the path uses only child, descendant and descendant or self axes.
 Otherwise, it returns false.
+Function \texttt{usesOnlyChildAnd\-DescendantAndAttributeAxes} is the same except it allows also attributes axis.
 
@@
 		\STATE $axisKind := stepNode.axisNode.getAxisKind()$
-		\IF{$axisKind \neq$ CHILD $\wedge$ $axisKind \neq$ DESCENDANT}
+		\IF{$axisKind \neq$ CHILD $\wedge$ $axisKind \neq$ DESCENDANT $\wedge$ $axisKind \neq$ DESCENDANT\_OR\_SELF}
 			\RETURN false
@@
         \STATE $type := getType(argument)$
-        \IF{$type$ is PathType $\wedge$ $usesOnlyChildAndDescendantAxes(type)$ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
+        \IF{$type$ is PathType \\ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(type)$ \\ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
             \STATE memorize $argument$ as the negative key statement, weight 1
@@
 That implies that the value in those nodes is not unique, and, alike the utilization of the aggregation functions, non-unique nodes cannot be keys.
+
+\todo[inline]{Zlucovanie vah s klucmi.}
 
@@
         \FORALL{$var \in forVars$}
-            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is PathType $\wedge$ $isWithoutPredicatesExceptLastStep(getType(var/L))$ $\wedge$ $usesOnlyChildAndDescendantAxes(getType(var/L))$ $\wedge$ $C$ is a literal constant}
+            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is PathType $\wedge$ $isWithoutPredicatesExceptLastStep(getType(var/L))$ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(getType(var/L))$ $\wedge$ $C$ is a literal constant}
                 \STATE memorize $var/L$ as the negative key statement, weight 0.9

### 2012-03-27 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.
Master thesis: Corrections.

@@
 The combination with such methods of inference is straightforward.
-The rules of the grammar describe a XML structure using the most general aspect of elements and their subelements.
+The rules of the grammar describe the XML structure using the most general aspect of elements and their subelements.
 Since the statements inferred by our method do not involve the XML structure by defining a subelement structure of elements, there are no conflicts between the grammar rules and our statements that need to be resolved.
@@
 
-In all tree cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form).
+In all three cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form).
 This is done in two steps.
@@
 
-It iterate through steps of a path, maintaining a so-called context set, which is a set of elements to evaluate the current step upon.
+It iterates through steps of a path, maintaining a so-called \emph{context set}, which is a set of elements to evaluate the current step upon.
 The evaluation of one step is shown in Algorithm \ref{FUNC_evaluate_step}.
@@
 
-At first, the algorithm determine the axis of the step.
+At first, the algorithm determines the axis of the step.
 If it is self or descendant axis, it returns the result of \texttt{evaluateStep\_selfOrDescendant} function.
@@
 \section{Saving the Inferred Statements}
-In the case of keys, the algorithm iterates through the inferred key statements.
+In case of keys, the algorithm iterates through the inferred key statements.
 A key's context path $C$ is evaluated and the key is assigned to the target elements selected by the context path, one element can be assigned with multiple keys.
@@
 
-In the case of inferred types, the situation is slightly less simple, because there can occur conflicting statements.
+In case of inferred types, the situation is slightly less simple, because there can occur conflicting statements.
 The examples of such conflicts for path $p$ without predicates are $p_1 \rightarrow date$, $p_2 \rightarrow string$, and, $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.
@@
 
-For each PathType $p$ that occurs in the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types.
+For each normalized PathType $p$ from the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types.
 $T_p = \{t|(p \rightarrow t) \in \mathcal{S}_t\}$.
@@
 
-Since we have the XML data and path $p$ is XQuery path, we can use an XQuery process (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$.
-The verification algorithms iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$.
+Since we have the XML data and path $p$ is an XQuery path, we can use an XQuery processor (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$.
+The verification algorithm iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$.
 If so, $t'$ is the inferred type for nodes $N$ (and PathType $p$), else the inferred type is \texttt{string}.

### 2012-03-27 doc/theses/mikula/conclusion.tex
Master thesis: Corrections.

@@
 The aim of this thesis was to employ XML operations in the XML schema inference.
-We analysed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations, only to found one, which infers keys from XQuery queries.
+We analyzed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations, only to found one\todo{Jiny vyraz}, which infers keys from XQuery queries.
 
-Since the notion XML operations is very general and the range of XML technologies is very large, for purpose of this work, we decided to focus on XQuery technology.
+Since the notion of XML operations is very general and the range of XML technologies is very large, for the purpose of this work, we decided to focus on XQuery technology.
 We made the overview of possible utilization of XQuery queries in the process of XML schema inference.
@@
 We also implemented several ideas from the overview to infer XSD built-in types of elements and attributes.
-And we slightly extended the one existing method dealing with the inference of keys.
+And we extended the one existing method dealing with the inference of keys.
 We experimentally demonstrated that on some input files, results of the extended method are better than results of the original method.
@@
 
-At last, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using jInfer framework.
+Finally, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using the jInfer framework.
 Thus, we created the first complete, ready-to-use, and extensible implementation of the XML xchema inference exploiting XQuery queries besides XML data.
@@
 
-Chapter \ref{CHAPTER_precreation_of_algorithm} discusses some possible future enhancements, as well.
+Chapter \ref{CHAPTER_precreation_of_algorithm} discusses some possible future enhancements as well.
 A research on possibilities of modifying the grammar rules based on information extracted from XQuery queries may bring interesting results.
 We analyse the queries statically only, we do not evaluate them.
-An analysis of queries together with their results can be a topic of some future research.
+An analysis of queries together with their results can be a topic of another future research direction.
 
-The utilization of XQuery queries certainly provides a space for incorporating an interaction with user. For example, a user may influence the scoring of inferred keys to get a more precise results.
+The utilization of XQuery queries certainly provides a space for incorporating an interaction with user. For example, a user may influence the scoring of inferred keys to get more precise results.
 
@@
 Besides the mentioned, our opinion is that the most urgent future work is obtaining a large enough test data set, performing proper experiments, and refining the algorithm by modification of its settings (statements inferred from join pattern occurrences, weights, etc.) according to experimental results.
-This process was slightly suggested in Chapter \ref{CHAPTER_experiments}.
+This process was suggested in Chapter \ref{CHAPTER_experiments}.

### 2012-03-27 doc/theses/mikula/introduction.tex
Master thesis: Corrections.

@@
 
-With XML used to represent data themselves, it is often needed or convenient to somehow specify a structure of the represented data, their format, internal relations or restrictions, etc.
+With XML used to represent data themselves, it is often needed or convenient to somehow specify the structure of the represented data, their format, internal relations or restrictions, etc.
 In order to this need, several so-called XML schema languages (XML schemata) have been created.
@@
 
-Despite an conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
+Despite a conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely.
 Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema \cite{Mlynkova:2008:AAX:1494650.1495496}.

### 2012-03-27 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.
Master thesis: Corrections.

@@
 For each cluster, a trivial schema is created, which is then generalized to achieve a reasonable result.
-In a search for the optimal schema, \emph{Ant Colony Optimization (ACO)} heuristic is incorporated.
+In search for the optimal schema, \emph{Ant Colony Optimization (ACO)} heuristic is incorporated.
 The idea behind the ACO heuristic is that a set of artifical ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum.
@@
 
-One of the improvements of this heuristic is an inclusion of a negative feedback after each step of an ant, visible only in a current iteration.
+One of the improvements of this heuristic is an inclusion of a negative feedback after each step of an ant, visible only in the current iteration.
 Due to this improvement, a larger subspace of solutions is searched.
@@
 The aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently.
-A XML schema can become obsolete due to changes in a set of XML documents, without capturing these changes in the schema.
+An XML schema can become obsolete due to changes in a set of XML documents, without capturing these changes in the schema.
 Thus, the schema becomes outdated and according to the paper, this case is quite common.
@@
     \hline
-    Method name and/or paper & Input & Output & Year \\ \hline \hline
+    \textbf{Method name and/or paper} & \textbf{Input} & \textbf{Output} & \textbf{Year} \\ \hline \hline
     \cite{ahonen} & XML documents & DTD & 1996 \\ \hline

### 2012-03-27 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: Corrections.

@@
 
-On the other hand, the structural inference could be useful when the entire set of all XML documents is not available and provided XML documents do not cover the structure completely.
+On the other hand, the structural inference could be useful when the entire set of all XML documents is not available and the provided XML documents do not cover the structure completely.
 
@@
 The sample query indicates that the root element \texttt{book} can contain more than one element \texttt{section}.
-An exception is usage of function \texttt{count()} in a predicate in expressions where it is used to determine if count of some nodes is greater than zero.
+An exception is usage of function \texttt{count()} in a predicate in expressions where it is used to determine if the number of some nodes is greater than zero.
 Often, this form is used to test presence of a certain node instead of actual counting of its occurrences.
@@
 
-Another similar examples are arithmetic expressions and functions accepting a sequence of at most one item.
+Other similar examples are arithmetic expressions and functions accepting a sequence of at most one item.
 Function \texttt{zero-or-one()} will raise the \emph{type error} when supplied with a sequence of more than one item.
@@
 Since the number of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available.
-This is a case when a set of XQuery queries may bring optimizations.
+This is the case when a set of XQuery queries may bring optimizations.
 For example, consider the following occurrences of element \texttt{a}.
@@
 
-A common three-way join can be also utilize to infer keys and foreign keys, however, the respective XML data and their analysis would be needed to tell what is a key and what is a foreign key.
+A common three-way join can be also utilized to infer keys and foreign keys, however, the respective XML data and their analysis would be needed to tell what is a key and what is a foreign key.
 
@@
 Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}.
-Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, each element with this attribute specified references an element with attribute \texttt{partid} of the same value.
+Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy, while each element with this attribute specified references an element with attribute \texttt{partid} of the same value.
 
@@
 
-According to the use of \texttt{distinct-values()} function, it can be easily seen that the author of this query assumes possible occurrence of more than one \texttt{book} elements with the same value of their \texttt{title} subelement.
+According to the use of \texttt{distinct-values()} function, it can be easily seen that the author of this query assumes possible occurrence of more than one \texttt{book} element with the same value of their \texttt{title} subelement.
 Thus, \texttt{title} element is not unique.
@@
 
-If there is a large enough set of XML documents available, it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs.
+If there is a large enough set of XML documents available, it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs \todo{Reference na Vostuv clanek}.
 However, if the inference is made using a smaller set of XML documents, it may happen that every occurrence of some element set is in the same order but the input data are not representative enough to be sure.

### 2012-03-27 doc/theses/mikula/implementation.tex
Master thesis: Corrections.

@@
 The solution proposed in Chapters \ref{CHAPTER_proposed_algorithm} and \ref{CHAPTER_combination_with_existings_methods_of_inference}, except the verification of types, was implemented using the jInfer framework \cite{jinfer}.
-It is a framework for implementing methods of XML schema inference created as a software project at Faculty of Mathematics and Physics, Charles University in Prague.
+It is a framework for implementing methods of XML schema inference created as a software project at the Faculty of Mathematics and Physics, Charles University in Prague.
 It is written in Java as a plugin for NetBeans platform.
@@
 \section{jInfer Modules}
+\todo[inline]{Labely obrazku dat pod.}
 \begin{figure}
@@
 
-Actually, the blue modules shown in \ref{FIG_jinfer_modules} are module abstractions with specified interface.
+Actually, the blue modules shown in Figure \ref{FIG_jinfer_modules} are module abstractions with specified interface.
 Actual modules are then implementations complying the interface.

### 2012-03-27 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: Corrections.

@@
 A basic query utilization can be achieved by analysis of queries without any other input data.
-The analysis of XQuery in the previous chapter discusses mostly XQuery constructs which can be utilized without respective XML data, for example the inference of built-it types.
-This independence is also the main advantage of this approach, if there are no XML data available, this approaches can be still used.
+The analysis of XQuery in the previous chapter discusses mostly XQuery constructs which can be utilized without respective XML data, for example the inference of built-in types.
+This independence is also the main advantage of this approach, if there are no XML data available, this approach can be still used.
 
@@
 The query consists of a function declaration and an arithmetic comparison.
-The comparison compares a result of the function call and literal value \texttt{10}.
+The comparison compares the result of the function call and literal value \texttt{10}.
 For the type of \texttt{10} is \texttt{xs:integer} the type of the function call has to be convertible to \texttt{xs:double}.
@@
 
-While simplifier approaches of the query processing such as the pattern finding limit possibilities of the query utilization, a more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
+While simpler approaches of the query processing such as the pattern finding limit possibilities of the query utilization, a more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions.
 Therefore, we decided to incorporate query processing using the lexical and syntax analyses.

### 2012-03-27 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

@@
 \chapter{Proposed Algorithm} \label{CHAPTER_proposed_algorithm}
-In this thesis we are developing an approach dealing with inference of XML schema, which input consists of two components; a set of XML documents and a set of XQuery queries related to the documents.
+In this thesis we are developing an approach dealing with inference of XML schema, whose input consists of two components; a set of XML documents and a set of XQuery queries related to the documents.
 The proposed algorithm introduced in this chapter describes the processing of XQuery queries only and a combination of its output with an existing method of XML schema inference (to produce a complete XML schema) is discussed in the next chapter.
@@
 \item \textbf{Construction of a syntax tree}.
-We use lexical and syntax analyses proposed in \cite{thesis_schejbal} and for each XQuery on input, we construct a data structure defined in \ref{DEF_syntax_tree}.
-\item \textbf{Static Analysis of expression types}. The algorithm searches for expressions in the syntax trees and statically (without evaluation) determines their types.
+We use lexical and syntax analyses proposed in \cite{thesis_schejbal} and for each XQuery on input, we construct a data structure defined in Definition \ref{DEF_syntax_tree}.
+\item \textbf{Static Analysis of expression types}.
+The algorithm searches for expressions in the syntax trees and statically (without evaluation) determines their types.
+\todo{Co se mysli?
+Datove typy?}
 \item \textbf{Inference of built-in types}. When the types of expressions are determined, selected forms of expression are utilized to infer types of elements and attributes.
@@
 Since they are not directly related to the inference, and thus, they are not directly related to the topic of this thesis, we will not describe them.
-Nevertheless, they provides us with a helpful processing of XQuery queries and we can focus on the inference.
+Nevertheless, they provide us with a helpful processing of XQuery queries and we can focus on the inference.
 
@@
 
-Regarding the additional properties, two constructs in $Q$ represented by two nodes of the same class from $C$ may differ in certain ways, and therefore, it is needed to distinguish them.
+Regarding the additional properties, two constructs in $Q$ represented by two nodes of the same class from $C$ may differ in certain ways, and, therefore, it is needed to distinguish them.
 For instance, two different literal values in $Q$ are represented by nodes $l_1, l_2 \in V$ and $c(l_1) = c(l_2) =$ \texttt{LiteralNode} but each has a different value and type.
@@
 Functions $varName_{VarRefNode}$, $axisKind_{AxisNode}$, $operator_{OperatorNode}$ are examples of commonly used functions from $\mathcal{P}$.
-Their meaning will be explained in a place of their use.
-For detail, see \cite{thesis_schejbal}.
+Their meaning will be explained in a place of their usage.
+For details, see \cite{thesis_schejbal}.
 
@@
 Function call of a function without arguments is represented by a leaf node while function call with arguments is represented by an inner node and its subnodes are nodes representing those arguments.
+
+\todo[inline]{Triedy hodit do tabulky, nie do figure.}
 
@@
 Therefore the scope of the new variable is not the subtree of the \texttt{VariableBindingNode} class node.
-It depends on the type of XQuery construct that the variable binding is a (indirect) component of.
+It depends on the type of XQuery construct that the variable binding is an (indirect) component of.
 
@@
 \begin{itemize}
-\item XML Schema built-in atomic types. See Attachment \ref{FIG_xsd_built_in_atomic_types}.
+\item XML Schema built-in atomic types. See Attachment \ref{FIG_xsd_built_in_atomic_types}. \todo{Kludne do textu}
 \item Types \emph{ElementType}, \emph{AttributeType}, \emph{NodeType}, \emph{TextNodeType},\linebreak \emph{CommentType}, \emph{ProcessingInstructionType}, \emph{DocumentType} representing an element, attribute, node, text node, comment, processing instruction, document node.
@@
 The represented path is contained by an instance of \texttt{PathExprNode} containing list of its steps, in particular instances of \texttt{StepExprNode}.
-If a step is a reference to a variable which type is PathType, we also want to include this information.
-Therefore, PathType contain association between the steps and other PathTypes and this association is defined for the PathType variable steps.
-
-To differ between a common PathType selecting a set of nodes and a PathType bound to a for variable in a FLWOR expression, PathType structure contains a flag \texttt{isForBound}.
+If a step is a reference to a variable whose type is PathType, we also want to include this information.
+Therefore, PathType contains association between the steps and other PathTypes and this association is defined for the PathType variable steps.
+
+To distinguish between a common PathType selecting a set of nodes and a PathType bound to a for variable in a FLWOR expression, PathType structure contains a flag \texttt{isForBound}.
 
@@
 
-To capture sequences, we assign the first two categories of types (all types except of PathType and UnknownType) with its cardinality as proposed in \cite{thesis_schejbal}.
+To capture sequences, we assign the first two categories of types (all types except for PathType and UnknownType) with its cardinality as proposed in \cite{thesis_schejbal}.
 Each of those types can be perceived as a sequence.
@@
 The meaning of the code at lines 6-14 is the following.
-If a variable is bound to an expression which type is PathType, we want to assign the variable reference expression with PathType which represents path containing exactly one step, the variable reference.
-Some algorithm in later sections search for paths starting with a certain variable and we want them to include also that types of expressions.
+If a variable is bound to an expression whose type is PathType, we want to assign the variable reference expression with PathType which represents path containing exactly one step, the variable reference.
+Some algorithm in later sections searches for paths starting with a certain variable and we want them to include also that types of expressions.
 
@@
 
-The handling of new variable definitions relies on the fact that the definitions make the new variables valid only in nodes with a higher sequence number when numbered in the order described in in the previous paragraph.
+The handling of new variable definitions relies on the fact that the definitions make the new variables valid only in nodes with a higher sequence number when numbered in the order described in the previous paragraph.
 Therefore, the left-most (the first amongst the ordered subnodes) subnode of a certain node can be processed without an extension of the variable context.
-And, if the left-most subnode extends the variable context for the following nodes in the numbering, it can be easily handled because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
+And, if the left-most subnode extends the variable context for the following nodes in the numbering, it can be easily handled, because the types of the new variables can be directly determined since the binding expressions are already processed (thus, their types are known).
 Every node ''knows'' whether or not it may define new variables for nodes with higher sequence numbers.
@@
 In this case, the function declaration node contains a subnode specifying function's formal arguments.
-These argument are set as the variable context for the function body represented by another subnode.
+These arguments are set as the variable context for the function body represented by another subnode.
 
@@
 Only expressions of a particular type are exploited.
-Specifically, an expression has to contain an subexpression $E$ of PathType type (expression representing a certain element or attribute or set of elements or attributes).
+Specifically, an expression has to contain an subexpression $E$ of PathType type (expression representing a certain element or attribute or a set of elements or attributes).
 In the following text, the set represented by expression $E$ is denoted $S$.
@@
 
-The output of this step is a set of statements of form $P \rightarrow T$, where $P$ is an instance of PathType and $T$ is an XML Schema built-in atomic type.
+The output of this step is a set of statements of the form $P \rightarrow T$, where $P$ is an instance of PathType and $T$ is an XML Schema built-in atomic type.
 
@@
 \section{Step 4: Key Discovery}
+\todo[inline]{Tu by neskodilo napisat, ze sice preberame alg, ale ho aj implementujeme a overime, co v povodnom clanku nebolo.}
 In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it.
@@
 A similar couple of functions is function \texttt{getTargetReturnPaths} and function \texttt{getTargetReturnPathsRecursive}.
-Only difference between those two couples of functions is that the former one searches for all expressions which are of PathType representing a target return path, while the later one searches only for instances of \texttt{PathExprNode}.
+The only difference between those two couples of functions is that the former one searches for all expressions which are of PathType representing a target return path, while the later one searches only for instances of \texttt{PathExprNode}.
 
@@
 \begin{define}[Key]
-A key is a construct $(C, P, \{L\})$ where $C$, $P$ and $L$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
+A key is a construct $(C, P, \{L\})$, where $C$, $P$ and $L$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
 $C$ is called \emph{context path}, $P$ \emph{target path} and $L$ \emph{key path}.
@@
 \begin{define}[Foreign key]
-A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$ where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
+A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$, where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
 $C$ can be omitted as in the case of keys.
 \end{define}
+
+\todo[inline]{priklad}
 
@@
 In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes.
-Actually, the join patterns from Listings \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly because their structure in the syntax tree is different.
-
-Since, in this step, we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of PathType without predicates, using only child and descendant axes.
-
-As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inference of keys from occurrences of the join patterns, and rules (R1 - R5) to classify each occurrence into one of these cases.
+Actually, the join patterns from Listings \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly, because their structure in the syntax tree is different.
+
+Since in this step we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of PathType without predicates, using only child and descendant axes.
+
+As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inference of keys from occurrences of the join patterns, and rules (R1 - R5) \todo{V sekci 3.5 zadna takova nevidim} to classify each occurrence into one of these cases.
 For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
@@
 
-Aggregation functions \texttt{min}, \texttt{max}, \texttt{sum} are not utilized by the function because they were utilized by the search for join patterns, and in the process of classification of found join patterns, they were included in the computation of the key weights.
-
-Contrary the key discovery, PathTypes utilized in this step may contain predicates in their lasts steps.
-Predicates in a path's last step restrict its result, and therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is a set of elements or attributes selected by the same path stripped of the predicates.
+Aggregation functions \texttt{min}, \texttt{max}, \texttt{sum} are not utilized by the function, because they were utilized by the search for join patterns, and in the process of classification of found join patterns, they were included in the computation of the key weights.
+
+Contrary to the key discovery, PathTypes utilized in this step may contain predicates in their lasts steps.
+Predicates in a path's last step restrict its result, and, therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is the set of elements or attributes selected by the same path stripped of the predicates.
 
@@
 
-As discussed in Chapter \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the count of the items may be more than one.
+As discussed in Chapter \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the number of the items may be more than one.
 And, together with that, the set is restricted to contain only items which satisfy a condition that a value of one of their subnodes (or descendant nodes) equals a certain value.

### 2012-03-27 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.
Master thesis: Corrections.

@@
 \chapter{Used Technologies and Definitions}
-
-\todo[inline]{Zvyraznovanie pojmov}
 
@@
 An XML schema refers to a description of an XML document in terms of its structure and various constraints.
-Commonly, the XML schema describes element and attribute names, their parent-child relations, their order and a type of their content.
+Commonly, the XML schema describes element and attribute names, their parent-child relations, their order and type of their content.
 Other constraints often expressed in the XML schema are restrictions on numbers of occurrences of elements, specification of (non-)obligatory attributes, uniqueness and specification of keys.
@@
 \begin{figure}
-\caption{A simple XML example.}
+\caption{A simple XML example}
 \label{FIG_a_simple_xml_example}
@@
 \end{figure}
+\todo[inline]{Zmensit Figure \ref{FIG_a_simple_xml_example}?}
 
 \subsection{DTD}
-Document Type Definition (DTD) expresses a structure of XML documents by declarations of elements. An element has its name and a content declared using syntax \texttt{<!ELEMENT $name$ $content$>}.
+Document Type Definition (DTD) expresses the structure of XML documents by declarations of elements. An element has its name and a content declared using syntax \texttt{<!ELEMENT $name$ $content$>}.
 
@@
 \item \texttt{CDATA} Character data - any string.
-\item \texttt{ID} An unique identifier.
+\item \texttt{ID} A unique identifier.
 \item \texttt{IDREF} An ID reference - a value of the ID.
@@
 
+\todo[inline]{Toto do obrazku.}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\todo[inline]{overit, zvalidovat}
 
@@
 
-Moreover, these three construct can be nested and combined, assuming the combination is not ambiguous.
+Moreover, these three constructs can be nested and combined, assuming the combination is not ambiguous.
 
@@
 The book element from the sample XML document in Figure \ref{FIG_a_simple_xml_example} can be described using the XSD as follows.
+
+\todo[inline]{Do obrazku.}
 \begin{alltt}
@@
 
-\todo[inline]{overit, zvalidovat}
-
 \section{XPath}
@@
 An axis specifies a node-set relative to the current node.
-The default axis is \texttt{child} selecting all children nodes of the current node.
+The default axis is \texttt{child} selecting all child nodes of the current node.
 Another important axes are \texttt{attribute}, selecting attributes of the current node, and \texttt{descendant-or-self} selecting the current node and all its descendants in the tree.
@@
 \item \texttt{text()} Text nodes.
-\item \texttt{*} All nodes assigned with their name (element and attributes).
+\item \texttt{*} All nodes assigned with their name (elements and attributes).
 \end{itemize}
 
-\subsection{Abbrevations}
-Abbrevations for several most widely used constructs are defined as follows.
+\subsection{Abbreviations}
+Abbreviations for several most widely used constructs are defined as follows.
 \begin{itemize}
@@
 
-\todo[inline]{overit, otestovat}
-
 \section{XQuery}
-An XML Query Language (XQuery, version 1.0) \todo{ref} is a language designed to query XML data.
+An XML Query Language (XQuery, version 1.0) \cite{w3c_xquery} is a language designed to query XML data.
 It is based on XPath 2.0.
@@
 
-In the remainder of this section, basic features of the XQuery languages are briefly described. Many query samples can be found in Chapter \ref{CHAPTER_analysis_of_xquery}.
+In the remainder of this section, basic features of the XQuery languages are briefly described. Many query examples can be found in Chapter \ref{CHAPTER_analysis_of_xquery}.
 
@@
 Order by clause is \\
-\texttt{order by $expr$)} \\
+\texttt{order by $expr$} \\
 and it orders the tuples of values that passed the where clause by the specified criterion.
@@
 
-\todo[inline]{TODO XQuery Core?}
-
 %\section{Other Technologies}
@@
 %\end{description}
-
-\todo[inline]{možná ňejaký automat apod. Prostě co budete dále v textu potřebovat. Skratky: XSD, XSLT, ...}

### 2012-03-27 doc/theses/mikula/experiments.tex
Master thesis: Corrections.

@@
 \begin{table}
-\label{TAB_inferred_types}
-\caption{Inferred type statements.}
 \begin{tabular}{|l|}
@@
 \end{tabular}
+\caption{Inferred type statements.}
+\label{TAB_inferred_types}
 \end{table}
@@
 \begin{table}
-\label{TAB_inferred_keys}
-\caption{Inferred key statements.}
 \begin{tabular}{|l|c|}
@@
 \end{tabular}
+\caption{Inferred key statements.}
+\label{TAB_inferred_keys}
 \end{table}
@@
 The second one is not correct, because \texttt{id} attribute is a key of \texttt{item} elements.
-The third one is correct and the forth one declares \texttt{itemref} element to be a key of closed auctions, which is not true, but only with weight 0.417.
+The third one is correct and the fourth one declares \texttt{itemref} element to be a key of closed auctions, which is not true, but only with weight 0.417.
 
@@
 \begin{table}
-\label{TAB_inferred_keys_2}
-\caption{Key statements inferred from the extended test data.}
 \begin{tabular}{|l|c|}
@@
 \end{tabular}
+\caption{Key statements inferred from the extended test data.}
+\label{TAB_inferred_keys_2}
 \end{table}
@@
 \begin{table}
-\label{TAB_inferred_keys_3}
-\caption{Key statements inferred from the extended test data using the modified JP3 statements.}
 \begin{tabular}{|l|c|}
@@
 \end{tabular}
+\caption[Key statements inferred using the modified JP3 statements.]{Key statements inferred from the extended test data using the modified JP3 statements.}
+\label{TAB_inferred_keys_3}
 \end{table}
 
-\todo[inline]{Pripojit vysledne XSD? Je dost velke. Alebo aspon relevantnu cast?}
+\todo[inline]{Pripojit vysledne XSD? Je dost velke. Alebo aspon relevantnu cast? A: Neni treba, kdyz tak do appendixu.}

### 2012-03-27 doc/theses/mikula/diplomka.tex
Master thesis: Corrections.

@@
   tabsize=2,                      % sets default tabsize to 2 spaces
-  captionpos=t,                   % sets the caption-position to bottom
+  captionpos=b,                   % sets the caption-position to bottom
   breaklines=true,                % sets automatic line breaking
@@
 %%% Tabulky v diplomové práci, existují-li.
-\chapwithtoc{List of Tables}
+%\chapwithtoc{List of Tables}
+\listoftables
+\todo[inline]{Je List of Tables potrebny? Co figures?}
 

### 2012-03-27 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

@@
 \begin{figure}
+\includegraphics[scale=1]{algorithm.eps}
+\caption{Steps of the proposed algorithm.}
 \label{FIG_steps_algorithm}
-\caption{Steps of the proposed algorithm.}
-\includegraphics[scale=1]{algorithm.eps}
 \end{figure}
@@
 \item $E$ is a set of pairs $(v, w)$ where $v,w \in V$ and for every $a, b \in V, a \neq b: (a, b) \in E$ if and only if a construct represented by $b$ is a direct component of a construct represented by $a$ ($b$ is a child of $a$) in query $Q$,
-\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2},
+\item $c: V \rightarrow C$ is function assigning each node with its class from set $C$ of all XQuery language constructs listed in Tables \ref{TAB_syntax_tree_node_types_part_1}, \ref{TAB_syntax_tree_node_types_part_2},
 \item $\mathcal{P}$ is a set of functions specifying additional properties of the nodes and distinguishing the nodes of the same classes,
@@
 The node classes of the syntax are organized in an is-a hierarchical structure, commonly used in the object oriented programming languages, where an object can be of several types.
-This hierarchy is shown in tables in Figures \ref{FIG_syntax_tree_node_types_part_1}, \ref{FIG_syntax_tree_node_types_part_2}.
+This hierarchy is shown in tables in Tables \ref{TAB_syntax_tree_node_types_part_1}, \ref{TAB_syntax_tree_node_types_part_2}.
 The tables are composed of classes in italic and their subclasses.
@@
 
-\todo[inline]{Triedy hodit do tabulky, nie do figure.}
-
-\begin{figure}
-\caption{Syntax tree node types part 1. For details, see \cite{thesis_schejbal}.}
-\label{FIG_syntax_tree_node_types_part_1}
+\begin{table}
 \begin{tabular}{l l | l}
@@
 \end{tabular}
-\end{figure}
-
-\begin{figure}
-\caption{Syntax tree node types part 2. For details, see \cite{thesis_schejbal}.}
-\label{FIG_syntax_tree_node_types_part_2}
+\caption{Syntax tree node types part 1. For details, see \cite{thesis_schejbal}.}
+\label{TAB_syntax_tree_node_types_part_1}
+\end{table}
+
+\begin{table}
 \begin{tabular}{l | l}
@@
 \end{tabular}
-\end{figure}
+\caption{Syntax tree node types part 2. For details, see \cite{thesis_schejbal}.}
+\label{TAB_syntax_tree_node_types_part_2}
+\end{table}
 
@@
 \begin{figure}
+\includegraphics[scale=0.6]{syntax_tree_1.eps}
+\caption{Sample syntax tree 1.}
 \label{FIG_syntax_tree_1}
-\caption{Sample syntax tree 1.}
-\includegraphics[scale=0.6]{syntax_tree_1.eps}
 \end{figure}
@@
 Nevertheless, the principle is straightforward.
-Depending on the class of the expression node (show in Figure \ref{FIG_syntax_tree_node_types_part_1}), its type can be either determined directly or it depends on its subexpressions.
+Depending on the class of the expression node (show in Table \ref{TAB_syntax_tree_node_types_part_1}), its type can be either determined directly or it depends on its subexpressions.
 
@@
 \begin{figure}
+\includegraphics[scale=0.5]{syntax_tree_1_types.eps}
+\caption{Sample syntax tree 1 after the static analysis of types.}
 \label{FIG_syntax_tree_1_types}
-\caption{Sample syntax tree 1 after the static analysis of types.}
-\includegraphics[scale=0.5]{syntax_tree_1_types.eps}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[scale=1]{syntax_tree_1_types_type_inference.eps}
+\caption{Example of an inference of a type from a function call.}
 \label{FIG_syntax_tree_1_types_type_inference}
-\caption{Example of an inference of a type from a function call.}
-\includegraphics[scale=1]{syntax_tree_1_types_type_inference.eps}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[scale=0.7]{syntax_tree_2.eps}
+\caption{Sample syntax tree 2.}
 \label{FIG_syntax_tree_2}
-\caption{Sample syntax tree 2.}
-\includegraphics[scale=0.7]{syntax_tree_2.eps}
 \end{figure}

### 2012-03-27 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: Corrections.

@@
 
-If there is a large enough set of XML documents available, it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs \todo{Reference na Vostuv clanek}.
+If there is a large enough set of XML documents available, it should be possible to correctly detect where to use \texttt{xs:sequence} and \texttt{xs:all} constructs (paper \cite{Vosta:2008:EAC:1802514.1802522} deals with this problem).
 However, if the inference is made using a smaller set of XML documents, it may happen that every occurrence of some element set is in the same order but the input data are not representative enough to be sure.

### 2012-03-27 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: Corrections.

@@
 \begin{figure}
-\caption{A simple XML example}
-\label{FIG_a_simple_xml_example}
 \begin{verbatim}
@@
 \end{verbatim}
+\caption{A simple XML example.}
+\label{FIG_a_simple_xml_example}
 \end{figure}
-\todo[inline]{Zmensit Figure \ref{FIG_a_simple_xml_example}?}
 
@@
 
-An example of DTD describing the book element from the XML example in Figure \ref{FIG_a_simple_xml_example} follows.
+An example of DTD describing the book element from the XML example in Figure \ref{FIG_a_simple_xml_example} is shown in Figure \ref{FIG_a_simple_dtd_example}.
 
-\todo[inline]{Toto do obrazku.}
+\begin{figure}
 \begin{verbatim}
@@
 \end{verbatim}
+\caption{A simple DTD example.}
+\label{FIG_a_simple_dtd_example}
+\end{figure}
 
@@
 
-The book element from the sample XML document in Figure \ref{FIG_a_simple_xml_example} can be described using the XSD as follows.
+The book element from the sample XML document in Figure \ref{FIG_a_simple_xml_example} can be described using the XSD as shown in Figure \ref{FIG_a_simple_xsd_example}.
 
-\todo[inline]{Do obrazku.}
+\begin{figure}
 \begin{alltt}
@@
         <xs:complexType>
-          <xs:element name="author" type="xs:string" maxOccurs="unbound"/>
+          <xs:element name="author" type="xs:string"
+           maxOccurs="unbound"/>
         </xs:complexType>
@@
 \end{alltt}
+\caption{A simple XSD example.}
+\label{FIG_a_simple_xsd_example}
+\end{figure}
 

### 2012-03-29 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.
Master thesis: In progress.

@@
 
+\section{Motivation}
+There is a large space of possible XQuery utilizations, but obviously, they all cannot be covered by one thesis.
+Hence, we had to choose only some of them.
+We decided to focus on an inference of XSD built-in atomic types of elements and attributes, and a key discovery.
+
+The inference of types is included in several recent works (\cite{Chidlovskii:2002:SEX:544220.544288, Hegewald:2006:XES:1129755.1130124}), but generally, it is considered a side problem and it is omitted by most of the works.
+Since a presence of XSD simple types in an inferred schema may be often desired or demanded, and an information on types of elements and attributes can be conveniently extracted from several XQuery constructs, we decided to incorporate the inference of simple types in our proposed algorithm and to implement it.
+
+As a second part of the algorithm, we decided to extend the only method \cite{Necasky:2009:DXK:1529282.1529414} which utilizes XQuery queries and infers keys and foreign keys.
+The extension is based on a more general and precise processing of queries, addition of a new case of query patterns that are used in the process, and a more accurate summary of key statements by incorporating a search for statements that reject a uniqueness of certain elements and attributes.
+
+Moreover, the original method has not been implemented and experimentally proven yet. We will do that as well.
+
+\section{Overview}
 \begin{figure}
@@
 \ELSE
-	\STATE mark $jpOccurrence$ as case (O2), weight 0.5 // \COMMENT {R5}
+	\STATE mark $jpOccurrence$ as case (O2), weight 0.5 \COMMENT {R5}
 \ENDIF

### 2012-03-29 doc/theses/mikula/implementation.tex
Master thesis: Corrections.
Master thesis: In progress.

@@
 \begin{figure}
+\includegraphics[scale=0.9]{jinfer_steps.eps}
+\caption{jInfer process of inference.}
 \label{FIG_jinfer_steps}
-\caption{jInfer process of inference.}
-\includegraphics[scale=0.9]{jinfer_steps.eps}
 \end{figure}
@@
 \section{jInfer Modules}
-\todo[inline]{Labely obrazku dat pod.}
 \begin{figure}
+\includegraphics[scale=0.8]{jinfer_modules.eps}
+\caption{jInfer modules.}
 \label{FIG_jinfer_modules}
-\caption{jInfer modules.}
-\includegraphics[scale=0.8]{jinfer_modules.eps}
 \end{figure}
@@
 Since we provide only one implementation of modules for the proposed solution, this is not very important for this work and we do not describe the principle in detail.
+We only note that \textbf{Basic XQuery Processor} module is an implementation of \textbf{Non-grammatical Input Processor} module abstraction.
 
 As in the previous section, newly added modules are those shown in yellow boxes.
-They are \textbf{XQuery Processor} and \textbf{XQuery Importer}.
+They are \textbf{Basic XQuery Processor} and \textbf{XQuery Importer}.
 We also modified modules \textbf{Basic XSD Exporter} (an implementation of \textbf{Schema Generator} interface) and \textbf{Base} to extend their functionality.
@@
 \item \textbf{XQuery Importer} - This module represents \textbf{XQAnalyzer} step in Figure \ref{FIG_jinfer_steps}. It is responsible for creating syntax trees from input XQuery files (step 1).
-\item \textbf{XQuery Processor} - The main module consisting of steps \textbf{XQuery Processor} and \textbf{Merger} in Figure \ref{FIG_jinfer_steps}.
-Package \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer} contains the main module class \texttt{XQueryAnalyzerImpl}.
-Other packages are \texttt{cz.cuni.mff.ksi.jinfer.xqueryanalyzer.-\\expressiontypesanalysis} implementing the static analysis of expression types (step 2), \texttt{cz.cuni.mff.ksi.jinfer.xqueryanalyzer.builtintype\-inference} implementing the inference of XSD built-in types (step 3), \texttt{cz\-.cuni.mff\-.ksi.jinfer\-.xqueryanalyzer.keydiscovery} and their subpackages implementing the key discovery (step 4), \texttt{cz.cuni.mff.ksi.jinfer\-.xqueryanalyzer.merger} implementing the merging with the grammar, and \texttt{cz.cuni.mff.ksi.jinfer.xqueryanalyzer.utils} containing various utilities.
+\item \textbf{Basic XQuery Processor} - The main module consisting of steps \textbf{XQuery Processor} and \textbf{Merger} in Figure \ref{FIG_jinfer_steps}.
+Package \texttt{cz.cuni.mff.ksi.jinfer\-.basicxqueryprocessor} contains the main module class \texttt{NonGrammatical\-InputProcessorImpl}.
+Other packages are \texttt{cz.cuni.mff.ksi.jinfer.\-basicxqueryprocessor.expressiontypesanalysis} implementing the static analysis of expression types (step 2), \texttt{cz.cuni.mff.ksi.jinfer.basic-\\xqueryprocessor.builtintype\-inference} implementing the inference of XSD built-in types (step 3), \texttt{cz\-.cuni.mff\-.ksi.jinfer\-.basicxquery-\\processor.keydiscovery} and their subpackages implementing the key discovery (step 4), \texttt{cz.cuni.mff.ksi.jinfer.basicxqueryprocessor.merger} implementing the merging with the grammar, and \texttt{cz.cuni.mff.ksi.jinfer\-.basicxqueryprocessor.utils} containing various utilities.
 \item \textbf{Basic XSD Exporter} - This is the module responsible for generating the resulting schema in XSD.
-It was modified to process the additional information assigned to the grammar in \textbf{XQuery Processor} module.
+It was modified to process the additional information assigned to the grammar in \textbf{Basic XQuery Processor} module.
 \end{itemize}

### 2012-03-29 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

@@
 
-Each occurrence of a join pattern is classified by application of the following rules it this specific order.
-The first satisfied ruled is applied.
+Each occurrence of a join pattern is classified by application of the following rules R1 - R5 in this specific order.
+The first satisfied rule is applied.
 The occurrence is also assigned with a weight determining how sure the method is about the inferred statement.
 
-The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a target return path (weight: 1) or if aggregation function \texttt{count} is applied on a target return path (weight: 0.75), where target return paths are paths in $C_R$ starting with $\$e_2$ (see Listing \ref{listing_for_join_pattern}).
+The pattern occurrence is considered of case (O1) if it is the for join pattern (R1, weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a target return path (R2, weight: 1) or if aggregation function \texttt{count} is applied on a target return path (R3, weight: 0.75), where target return paths are paths in $C_R$ starting with $\$e_2$ (see Listing \ref{listing_for_join_pattern}).
 
 Otherwise, the pattern occurrence is considered of case (O2) and the assigned weight depends on the number of target return paths.
-If the number is greater than one, the weight is one (weight: 1), else (the number equals zero or one) the weight is one half (weight: 0.5).
+If the number is greater than one, the weight is one (R4, weight: 1), else (the number equals zero or one) the weight is one half (R5, weight: 0.5).
 

### 2012-03-29 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \subsection{Expression Types}
+\begin{figure}
+\caption{XSD built-in atomic types.}
+\label{FIG_xsd_built_in_atomic_types}
+\includegraphics[scale=0.65]{xsd-types.eps}
+\end{figure}
+
 \begin{itemize}
-\item XML Schema built-in atomic types. See Attachment \ref{FIG_xsd_built_in_atomic_types}. \todo{Kludne do textu}
+\item XML Schema built-in atomic types. See Figure \ref{FIG_xsd_built_in_atomic_types}.
 \item Types \emph{ElementType}, \emph{AttributeType}, \emph{NodeType}, \emph{TextNodeType},\linebreak \emph{CommentType}, \emph{ProcessingInstructionType}, \emph{DocumentType} representing an element, attribute, node, text node, comment, processing instruction, document node.
@@
 \subsection{Definition of Keys}
-To define a key and a foreign key, we adopt the definitions introduced in the original approach \cite{Necasky:2009:DXK:1529282.1529414} with a modification that XPath paths are replaced by PathType instances.
+To define a key and a foreign key, we adopt the definitions introduced in the original approach \cite{Necasky:2009:DXK:1529282.1529414} with the following modifications.
+
+\begin{itemize}
+\item XPath paths are replaced by PathType instances.
+\item All paths can use also descendant or self axis.
+\item \emph{Key paths} can use also attribute axis.
+\end{itemize}
+
+The modified definitions are as follows.
 
 \begin{define}[Key]
-A key is a construct $(C, P, \{L\})$, where $C$, $P$ and $L$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
+A key is a construct $(C, P, \{L\})$, where $C$, $P$ and $L$ are PathType instances without predicates and without special function calls that use only child, descendant, and descendant or self ($L$ also attribute) axes.
 $C$ is called \emph{context path}, $P$ \emph{target path} and $L$ \emph{key path}.
@@
 \begin{define}[Foreign key]
-A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$, where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are PathType instances without predicates and without special function calls that use only child and descendant axes.
+A foreign key is a construct $(C, (P_1, \{L_1\}) \rightarrow (P_2, \{L_2\}))$, where $(C, P_2, \{L_2\})$ is a key and $P_1$ and $L_1$ are PathType instances without predicates and without special function calls that use only child, descendant, and descendant or self ($L$ also attribute) axes.
 $C$ can be omitted as in the case of keys.
@@
 
-\todo[inline]{priklad}
+An example of a global key is \texttt{(/site/people/person, \{@id\})}.
+An example of a foreign key to that key is  \texttt{((/site/closed\_auctions/closed\_auction, \{buyer/@person\}) $\rightarrow$ (/site/people/person, \{@id\}))}.
 
@@
 
-As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inference of keys from occurrences of the join patterns, and rules (R1 - R5) \todo{V sekci 3.5 zadna takova nevidim} to classify each occurrence into one of these cases.
+As described in the original method and summarized in Chapter \ref{Keys} we recognize two cases (O1) and (O2) of inference of keys from occurrences of the join patterns, and rules (R1 - R5) to classify each occurrence into one of these cases.
 For the join pattern 3, we introduce a new rule, considering the join pattern 3 of case (O1), assigned with weight of 0.5.
@@
 
-\todo[inline]{Zlucovanie vah s klucmi.}
-
 \begin{algorithm}
@@
 
+\todo[inline]{Definicia NUS a context Path, podobne ako v klucoch.}
+
 \subsection{Normalization of Key Statements}
@@
 
-Let $K_1, \dots, K_n$ be the inferred keys.
-Let $S_i$ be the score of $K_i$ and $N_i$ be the number of the inferred statements about $K_i$.
+Let $K_1, \dots, K_n$ be the inferred keys. Let $S_i$ be the score (the sum of weights) of $K_i$  and $N_i$ be the number of the inferred statements about $K_i$.
+
+Let $NU_1, \dots, NU_m$ be the discovered negative uniqueness statements. Let $W_j$ be the weight of $NU_j$.
+
+The negative uniqueness statements affect the scores as follows.
+For each negative uniqueness statement $NU_j = (C^{NU}_j, P^{NU}_j)$, find $\mathcal{K}_j$ set of keys which are rejected by the statement.
+\begin{center}
+$\mathcal{K}_j = \Bigl \{K_i = (C^K_i, P^K_i, \{L^K_i\}) \Big | C^{NU}_j = C^K_i, P^{NU}_j = P^K_i/L^K_i\Bigr \}$.
+\end{center}
+Then, for each key $K_i \in \mathcal{K}_j$ decrement its score $S_i$ with weight of the negative uniqueness statement $W_j$.
+
 Let $S^{max}$ be the maximum from $|S_1|, \dots , |S_n|$ and $N^{max}$ be the maximum norm
-from $N_1, \dots , N_n$. The normalized score $S_i$ of $K_i$ is computed as $S^{norm}_i = \frac{S_i}{S^{max}} * (1 - \frac{N^{max} - N_i}{\sum_{i=1}^{n}N_i})$
+from $N_1, \dots , N_n$. The normalized score $S_i$ of $K_i$ is computed as
+\begin{center}
+$S^{norm}_i = \frac{S_i}{S^{max}} * (1 - \frac{N^{max} - N_i}{\sum_{i=1}^{n}N_i})$
+\end{center}
 

### 2012-03-30 doc/jInfer/architecture-trunk/architecture.tex
Architecture documentation renamed to architecture-1.0. It was also copied and modified to create architecture-trunk documenting architecture in trunk.

--- 2011-03-01 doc/jInfer/architecture/architecture.tex
@@
 \ref{inference_process}.
-From the high-level viewpoint, it consists of three consecutive steps carried out by three different modules:
+From the high-level viewpoint, it consists of four consecutive steps carried out by four different modules:
 \begin{enumerate}
-	\item Initial Grammar (IG) generation: done by the \jmodule{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation.
-All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
+	\item Initial Grammar (IG) generation: done by the \jmodule{Initial Grammar Generator (IGG)} module, this is the process of converting of the inputs to IG representation.
+Documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.
+Not all types of inputs have to be used to create the initial grammar, as will be described in step 3.\\
 	For example, a trivial XML document
@@
 User interaction might be used in this step to help achieve better simplification.
-At the end of this step, all rules are sent to the export step.\\
+At the end of this step, all rules are sent to the next step.\\
 	For example, previous rules for element person could be simplified to a single rule
@@
 In Initial Grammar, all regexps are concatenations (even empty), but in simplified grammar, if element have to be empty in schema, it has to have lambda regular expresion as subnodes.
+	\item Non-grammatical input processing: done by the \jmodule{Non-grammatical Input Processor}, in this step, certain inputs are exploited to extract any non-grammatical information that can help in the inference.
+The information is then merged into the grammar which is sent to the last step.
+Inputs processed in this step can be inputs that were used to create the initial grammar, and, as well inputs that were not touched by the \jmodule{IGG} module.
+This step is not a crucial part of the inference, it's purpose is to provide a possibility to easily extend the inference process by incorporating other non-grammatical inputs.
+Therefore, the inputs and the way of merging the extracted information with the grammar are not specified in detail and they depend on an implementation.
 	\item Schema export: done by the \jmodule{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema file from the simplified rules.
@@
 	\end{verbatim}
-	For element person, even when simplified grammar specifies its occurence to at least once, at most 3 times, as DTD has no such construct, export module have to do some magic.
+	For element person, even when simplified grammar specifies its occurrence to at least once, at most 3 times, as DTD has no such construct, export module have to do some magic.
 Situation is even worse for elements that contain simple data inside simplified rule.
@@
 	\item \code{Simplifier}
+	\item \code{NonGrammaticalInputProcessor}
 	\item \code{SchemaGenerator}
@@
 Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.
+
+\jmodule{Non-grammatical Input Processor} takes both grammar and \code{Input} object.
 
@@
 	\item \code{SimplifierCallback}
+	\item \code{NonGrammaticalInputProcessor}
 	\item \code{SchemaGeneratorCallback}
@@
 \code{IGGeneratorCallback.finished()} and \code{SimplifierCallback.finished()} have a grammar (Initial Grammar in case of IGG) as their only argument.
+\code{NonGrammaticalInputProcessorCallback.finished()} has that argument, as well.
 
@@
 	\item When Simplifier finishes, it invokes callback's \code{finish} method, passing the simplified grammar as parameter.
-	\item This again causes \code{Runner} to encapsulate and execute SchemaGen's \code{start}, passing the simplified grammar from the second callback and the third callback as parameters.
+	\item This again causes \code{Runner} to encapsulate and execute NonGrammaticalInputProcessor's \code{start}, passing the simplified from the second callback and the third callback as parameters.
+	\item When NonGrammaticalInputProcessor finishes, it invokes callback's \code{finish} method, passing the merged simplified grammar as parameter.
+	\item This once again causes \code{Runner} to encapsulate and execute SchemaGen's \code{start}, passing the grammar from the third callback and the fourth callback as parameters.
 	\item SchemaGen finishes and invokes the last callback's \code{finish}, passing the resulting schema and its extension as parameters.
@@
 \begin{itemize}
-	\item \jmodule{Runner}: as described in the section \ref{section_runner}, this module encapsulates the inference process (including the fact that inference consists of 3 phases as described before).
+	\item \jmodule{Runner}: as described in the section \ref{section_runner}, this module encapsulates the inference process (including the fact that inference consists of 4 phases as described before).
 Most important class: \code{Runner}.
@@
 	\item \jmodule{XSDImportDOM}: XSD input handling using DOM parser.
+	\item \jmodule{XQueryImporter}: XQuery input handling incorporating lexical and syntax analyses as proposed in \cite{mikula}.
 	\item \jmodule{TwoStepSimplifier}: a \jmodule{Simplifier} implementation.
@@
 Read the separated documentation of module.
+	\item \jmodule{BasicXQueryProcessor}: a \jmodule{Non-grammatical Input Processor} implementation.
+Implements algorithm proposed in \cite{mikula} processing XQuery queries as an non-grammatical input and refining the inference process by information extracted from the queries.
 	\item \jmodule{BasicDTDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to DTD.
@@
 The dependencies between all the modules are summarized in figure \ref{module_deps}.
-Figure \ref{modules_inference} shows modules implementing the 3 main inference interfaces described in \ref{section_inference_process}.
+Figure \ref{modules_inference} shows modules implementing the 4 main inference interfaces described in \ref{section_inference_process}.
 
@@
 
-jInfer uses tokens in the following way: \jmodule{Runner}, as the module encapsulating the inference process, and \jmodule{ProjectType} as an encapsulation of jInfer projects both require the following 3 tokens.
+jInfer uses tokens in the following way: \jmodule{Runner}, as the module encapsulating the inference process, and \jmodule{ProjectType} as an encapsulation of jInfer projects both require the following 4 tokens.
 \begin{itemize}
@@
 	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator}
+	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.NonGrammaticalInputProcessor}
 	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier}

### 2012-03-31 doc/theses/mikula/diplomka.tex
Master thesis: In progress.
Master thesis: Corrections.

@@
 Název práce:
-\mftitle
+Odvozování XML schémy se znalostí XML operací
 % přesně dle zadání
@@
 Vedoucí diplomové práce:
-\mfadvisor, pracoviště
+\mfadvisor
 % dle Organizační struktury MFF UK, případně plný název pracoviątě mimo MFF UK
 
-Abstrakt:
+Abstrakt: \todo[inline]{TODO cca 10-15 riadkov}
 % abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce
 
-Klíčová slova:
+Klíčová slova: \todo[inline]{TODO 5-8 slov}
 % 3 aľ 5 klíčových slov
@@
 Department:
-Název katedry či ústavu, kde byla práce oficiálně zadána
+Department of Software Engineering
 % dle Organizační struktury MFF UK v angličtině
@@
 Supervisor:
-\mfadvisor, pracoviště
+\mfadvisor
 % dle Organizační struktury MFF UK, případně plný název pracoviątě
@@
 
-Abstract:
+Abstract: \todo[inline]{TODO}
 % abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
@@
 
-Keywords:
+Keywords: \todo[inline]{TODO}
 % 3 aľ 5 klíčových slov v angličtině
@@
 \listoftables
-\todo[inline]{Je List of Tables potrebny? Co figures?}
 

### 2012-03-31 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

@@
 Hence, we had to choose only some of them.
-We decided to focus on an inference of XSD built-in atomic types of elements and attributes, and a key discovery.
+We decided to focus on inference of XSD built-in atomic types of elements and attributes, and key discovery.
 
@@
 As a second part of the algorithm, we decided to extend the only method \cite{Necasky:2009:DXK:1529282.1529414} which utilizes XQuery queries and infers keys and foreign keys.
-The extension is based on a more general and precise processing of queries, addition of a new case of query patterns that are used in the process, and a more accurate summary of key statements by incorporating a search for statements that reject a uniqueness of certain elements and attributes.
+The extension is based on a more general and precise processing of queries, addition of a new case of query patterns that are used in the process, and a more accurate summary of key statements by incorporating searching for statements that reject uniqueness of certain elements and attributes.
 
@@
 \subsection{Normalization of Key Statements}
-The original approach \cite{Necasky:2009:DXK:1529282.1529414} employs the following normalization which we adopt. Is also incorporates a heuristic for precision enhancement. For details, refer to it.
+The original approach \cite{Necasky:2009:DXK:1529282.1529414} employs the following normalization which we adopt.
+Is also incorporates a heuristic for precision enhancement.
+For details, we refer to it.
 

### 2012-03-31 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: Corrections.

@@
 \subsection{Usage Examples}
-Finally, we introduce several examples of XPath expressions, with their description, using the sample XML in Figure \ref{FIG_a_simple_xml_example}.
+Finally, we introduce several examples of XPath expressions, with their description, using the sample XML document in Figure \ref{FIG_a_simple_xml_example}.
 

### 2012-03-31 doc/theses/mikula/appendix_a.tex
Master thesis: In progress.

@@
 \chapter{Attachments}
-\begin{figure}
-\caption{All possible values representing an operator in an instance of \texttt{OperatorNode}}
-\label{FIG_operators}
+\begin{figure}[float=htpb]
 Logical: \texttt{AND}, \texttt{OR}.
@@
 Unary: \texttt{UNARY\_PLUS}, \texttt{UNARY\_MINUS}.
+\caption{All possible values representing an operator in an instance of \texttt{OperatorNode}}
+\label{FIG_operators}
 \end{figure}
 
-\begin{figure}
-\caption{XSD built-in atomic types.}
-\label{FIG_xsd_built_in_atomic_types}
-\includegraphics[scale=0.65]{xsd-types.eps}
-\end{figure}
+\begin{lstlisting}[caption=Resulting XSD of Test Scenario B, frame=single, label=listing_xsd]
+<?xml version="1.0" encoding="UTF-8"?>
+<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">
+<!-- Inferred on Sat Mar 31 21:28:32 CEST 2012 by Basic IG generator, TwoStep(Iname (with attributes), Automaton Merging State(GreedyMDL(Combined(k,h-context, s,k-strings, Null, Null),Naive Alphabet), State Removal Ordered(Weighted)), Chained(Empty Children, Nested Concatenation, Null)), Basic XSD exporter -->
+
+<!-- global types -->
+<xs:complexType name="Temph" mixed="true">
+  <xs:sequence>
+    <xs:sequence minOccurs="0" maxOccurs="unbounded">
+      <xs:choice>
+        <xs:element name="keyword" type="Tkeyword"/>
+        <xs:element name="bold" type="Tbold"/>
+      </xs:choice>
+    </xs:sequence>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tbold" mixed="true">
+  <xs:sequence>
+    <xs:choice>
+      <xs:choice>
+        <xs:sequence>
+          <xs:element name="emph" type="Temph"/>
+        </xs:sequence>
+      </xs:choice>
+      <xs:sequence>
+        <xs:element name="keyword" type="Tkeyword"/>
+      </xs:sequence>
+    </xs:choice>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tkeyword" mixed="true">
+  <xs:sequence>
+    <xs:choice>
+      <xs:sequence minOccurs="0">
+        <xs:element name="bold" type="Tbold"/>
+      </xs:sequence>
+      <xs:sequence>
+        <xs:element name="emph" type="Temph"/>
+      </xs:sequence>
+    </xs:choice>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Ttext" mixed="true">
+  <xs:choice minOccurs="0" maxOccurs="unbounded">
+    <xs:element name="bold" type="Tbold"/>
+    <xs:element name="keyword" type="Tkeyword"/>
+    <xs:element name="emph" type="Temph"/>
+  </xs:choice>
+</xs:complexType>
+
+<xs:complexType name="Tparlist">
+  <xs:sequence>
+    <xs:element name="listitem" type="Tlistitem" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tlistitem">
+  <xs:choice>
+    <xs:element name="text" type="Ttext"/>
+    <xs:element name="parlist" type="Tparlist"/>
+  </xs:choice>
+</xs:complexType>
+
+<xs:complexType name="Tdescription">
+  <xs:choice>
+    <xs:element name="parlist" type="Tparlist"/>
+    <xs:element name="text" type="Ttext"/>
+  </xs:choice>
+</xs:complexType>
+
+<xs:complexType name="Tincategory">
+  <xs:attribute name="category" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tmail">
+  <xs:sequence>
+    <xs:element name="from" type="xs:string"/>
+    <xs:element name="to" type="xs:string"/>
+    <xs:element name="date" type="xs:string"/>
+    <xs:element name="text" type="Ttext"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tmailbox">
+  <xs:sequence minOccurs="0">
+    <xs:element name="mail" type="Tmail"/>
+    <xs:element name="mail" type="Tmail" minOccurs="0"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Titem">
+  <xs:sequence>
+    <xs:element name="location" type="xs:string"/>
+    <xs:element name="quantity" type="xs:string"/>
+    <xs:element name="name" type="xs:string"/>
+    <xs:element name="payment" type="xs:string"/>
+    <xs:element name="description" type="Tdescription"/>
+    <xs:element name="shipping" type="xs:string"/>
+    <xs:element name="incategory" type="Tincategory" minOccurs="0" maxOccurs="unbounded"/>
+    <xs:element name="mailbox" type="Tmailbox"/>
+  </xs:sequence>
+  <xs:attribute name="id" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tafrica">
+  <xs:sequence>
+    <xs:element name="item" type="Titem"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tasia">
+  <xs:sequence>
+    <xs:element name="item" type="Titem"/>
+    <xs:element name="item" type="Titem"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Taustralia">
+  <xs:sequence>
+    <xs:element name="item" type="Titem"/>
+    <xs:element name="item" type="Titem"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Teurope">
+  <xs:sequence>
+    <xs:element name="item" type="Titem" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tnamerica">
+  <xs:sequence>
+    <xs:element name="item" type="Titem" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tsamerica">
+  <xs:sequence>
+    <xs:element name="item" type="Titem"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tregions">
+  <xs:sequence>
+    <xs:element name="africa" type="Tafrica"/>
+    <xs:element name="asia" type="Tasia"/>
+    <xs:element name="australia" type="Taustralia"/>
+    <xs:element name="europe" type="Teurope"/>
+    <xs:element name="namerica" type="Tnamerica"/>
+    <xs:element name="samerica" type="Tsamerica"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tcategory">
+  <xs:sequence>
+    <xs:element name="name" type="xs:string"/>
+    <xs:element name="description" type="Tdescription"/>
+  </xs:sequence>
+  <xs:attribute name="id" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tcategories">
+  <xs:sequence>
+    <xs:element name="category" type="Tcategory"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tedge">
+  <xs:attribute name="from" type="xs:string" use="required"/>
+  <xs:attribute name="to" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tcatgraph">
+  <xs:sequence>
+    <xs:element name="edge" type="Tedge"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Taddress">
+  <xs:sequence>
+    <xs:element name="street" type="xs:string"/>
+    <xs:element name="city" type="xs:string"/>
+    <xs:element name="country" type="xs:string"/>
+    <xs:choice>
+      <xs:element name="zipcode" type="xs:string"/>
+      <xs:sequence>
+        <xs:element name="province" type="xs:string"/>
+        <xs:element name="zipcode" type="xs:string"/>
+      </xs:sequence>
+    </xs:choice>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tinterest">
+  <xs:attribute name="category" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tprofile">
+  <xs:sequence>
+    <xs:choice minOccurs="0" maxOccurs="unbounded">
+      <xs:element name="education" type="xs:string"/>
+      <xs:element name="interest" type="Tinterest"/>
+    </xs:choice>
+    <xs:choice>
+      <xs:element name="business" type="xs:string"/>
+      <xs:sequence>
+        <xs:element name="gender" type="xs:string"/>
+        <xs:element name="business" type="xs:string"/>
+      </xs:sequence>
+    </xs:choice>
+    <xs:element name="age" type="xs:string" minOccurs="0"/>
+  </xs:sequence>
+  <xs:attribute name="income" type="xs:integer" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Twatch">
+  <xs:attribute name="open_auction" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Twatches">
+  <xs:sequence>
+    <xs:element name="watch" type="Twatch" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tperson">
+  <xs:sequence>
+    <xs:element name="name" type="xs:string"/>
+    <xs:element name="emailaddress" type="xs:string"/>
+    <xs:choice minOccurs="0" maxOccurs="unbounded">
+      <xs:element name="address" type="Taddress"/>
+      <xs:element name="phone" type="xs:string"/>
+      <xs:element name="homepage" type="xs:string"/>
+      <xs:element name="creditcard" type="xs:string"/>
+    </xs:choice>
+    <xs:choice>
+      <xs:sequence minOccurs="0">
+        <xs:element name="profile" type="Tprofile"/>
+        <xs:element name="watches" type="Twatches" minOccurs="0"/>
+      </xs:sequence>
+      <xs:element name="watches" type="Twatches"/>
+    </xs:choice>
+  </xs:sequence>
+  <xs:attribute name="id" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tpeople">
+  <xs:sequence>
+    <xs:element name="person" type="Tperson" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tpersonref">
+  <xs:attribute name="person" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tbidder">
+  <xs:sequence>
+    <xs:element name="date" type="xs:string"/>
+    <xs:element name="time" type="xs:string"/>
+    <xs:element name="personref" type="Tpersonref"/>
+    <xs:element name="increase" type="xs:integer"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Titemref">
+  <xs:attribute name="item" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tseller">
+  <xs:attribute name="person" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tauthor">
+  <xs:attribute name="person" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tannotation">
+  <xs:sequence>
+    <xs:element name="author" type="Tauthor"/>
+    <xs:element name="description" type="Tdescription"/>
+    <xs:element name="happiness" type="xs:string"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tinterval">
+  <xs:sequence>
+    <xs:element name="start" type="xs:string"/>
+    <xs:element name="end" type="xs:string"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Topen_auction">
+  <xs:sequence>
+    <xs:element name="initial" type="xs:integer"/>
+    <xs:choice minOccurs="0" maxOccurs="unbounded">
+      <xs:element name="reserve" type="xs:decimal"/>
+      <xs:element name="bidder" type="Tbidder"/>
+    </xs:choice>
+    <xs:element name="current" type="xs:string"/>
+    <xs:element name="privacy" type="xs:string" minOccurs="0" maxOccurs="unbounded"/>
+    <xs:element name="itemref" type="Titemref"/>
+    <xs:element name="seller" type="Tseller"/>
+    <xs:element name="annotation" type="Tannotation"/>
+    <xs:element name="quantity" type="xs:string"/>
+    <xs:element name="type" type="xs:string"/>
+    <xs:element name="interval" type="Tinterval"/>
+  </xs:sequence>
+  <xs:attribute name="id" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Topen_auctions">
+  <xs:sequence>
+    <xs:element name="open_auction" type="Topen_auction" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tbuyer">
+  <xs:attribute name="person" type="xs:string" use="required"/>
+</xs:complexType>
+
+<xs:complexType name="Tclosed_auction">
+  <xs:sequence>
+    <xs:element name="seller" type="Tseller"/>
+    <xs:element name="buyer" type="Tbuyer"/>
+    <xs:element name="itemref" type="Titemref"/>
+    <xs:element name="price" type="xs:integer"/>
+    <xs:element name="date" type="xs:string"/>
+    <xs:element name="quantity" type="xs:string"/>
+    <xs:element name="type" type="xs:string"/>
+    <xs:element name="annotation" type="Tannotation"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tclosed_auctions">
+  <xs:sequence>
+    <xs:element name="closed_auction" type="Tclosed_auction" minOccurs="0" maxOccurs="unbounded"/>
+  </xs:sequence>
+</xs:complexType>
+
+<xs:complexType name="Tsite">
+  <xs:sequence>
+    <xs:element name="regions" type="Tregions"/>
+    <xs:element name="categories" type="Tcategories"/>
+    <xs:element name="catgraph" type="Tcatgraph"/>
+    <xs:element name="people" type="Tpeople"/>
+    <xs:element name="open_auctions" type="Topen_auctions"/>
+    <xs:element name="closed_auctions" type="Tclosed_auctions"/>
+  </xs:sequence>
+</xs:complexType>
+
+<!-- top level element -->
+<xs:element name="site" type="Tsite">
+  <xs:key name="key1">
+    <xs:selector xpath="people/person"/>
+    <xs:field xpath="@id"/>
+  </xs:key>
+  <xs:key name="key2">
+    <xs:selector xpath=".//item"/>
+    <xs:field xpath="@id"/>
+  </xs:key>
+  <xs:key name="key3">
+    <xs:selector xpath="regions/europe/item"/>
+    <xs:field xpath="@id"/>
+  </xs:key>
+  <xs:keyref name="keyRef1_key1" refer="key1">
+    <xs:selector xpath="closed_auctions/closed_auction"/>
+    <xs:field xpath="buyer/@person"/>
+  </xs:keyref>
+  <xs:keyref name="keyRef2_key2" refer="key2">
+    <xs:selector xpath="closed_auctions/closed_auction"/>
+    <xs:field xpath="itemref/@item"/>
+  </xs:keyref>
+  <xs:keyref name="keyRef3_key3" refer="key3">
+    <xs:selector xpath="closed_auctions/closed_auction"/>
+    <xs:field xpath="itemref/@item"/>
+  </xs:keyref>
+  <xs:keyref name="keyRef4_key3" refer="key3">
+    <xs:selector xpath="open_auctions/open_auction"/>
+    <xs:field xpath="itemref/@item"/>
+  </xs:keyref>
+</xs:element>
+</xs:schema>
+\end{lstlisting}

### 2012-03-31 doc/theses/mikula/experiments.tex
Master thesis: In progress.

@@
 
-\section{Test Data}
+We made two test scenarios; one dealing with input data that were not made for purposes of this experiments, and the other extending the first dataset by data created for better test coverage.
+
+\section{Test Scenario A}
+
+\subsection{Test Data}
 To get meaningful results of the experiments, test data should be composed of XML documents, which are instances of a certain, possibly not known, XML schema, and a set of XQuery queries which query the XML documents.
@@
 
-\section{Results}
+\subsection{Results}
 
-\subsection{Type Inference}
+\subsubsection{Type Inference}
 Six type statements shown in Table \ref{TAB_inferred_types} were inferred.
@@
 
-\subsection{Key Discovery}
+\subsubsection{Key Discovery}
 Four key statements and their normalized weights shown in Table \ref{TAB_inferred_keys} were inferred.
@@
 
-\todo[inline]{Pripojit vysledne XSD? Je dost velke. Alebo aspon relevantnu cast? A: Neni treba, kdyz tak do appendixu.}
+\section{Test Scenario B}
+As mentioned before, the previous test scenario involves only data with JP3 occurrences, and therefore, it only tests the extension of the original method, while the original method itself remains untested.
+To correct it, we add several new test queries created by ourselves.
+
+\subsection{Test Data}
+
+\begin{lstlisting}[caption=Test query B1 containing a for join pattern occurrence., frame=single, label=listing_test_query_b1]
+for $item in /site/regions/europe/item
+let $closed_auctions := /site/closed_auctions/closed_auction[itemref/@item = $item/@id]
+return <item><id>{$item/@id}</id><max-price>{max($closed_auctions/price)}</max-price></item>
+\end{lstlisting}
+
+\begin{lstlisting}[caption=Test query B2 containing a let join pattern occurrence., frame=single, label=listing_test_query_b2]
+for $open_auction in /site/open_auctions/open_auction
+let $item := /site/regions/europe/item[@id = $open_auction/itemref/@item]
+return $item/personref/@person
+\end{lstlisting}
+
+\begin{lstlisting}[caption=Test query B3 containing let join pattern and JP3 occurrences., frame=single, label=listing_test_query_b3, numbers=left, numberstyle=\tiny]
+for $person in /site/people/person
+where $person/profile/@income < 100.00 and $person/profile/gender = "m"
+return
+    <list>{
+        for $auction in /site/closed_auctions/closed_auction
+        let $item := /site//item[@id = $auction/itemref/@item]
+        let $price := $auction/price
+        where $auction/buyer/@person = $person/@id
+        return
+            <record><person>{$person/@id}</person><item>{$item/@id}</item><price>{$price}</price></record>
+    }</list>
+\end{lstlisting}
+
+In this scenario, we use test data from the Test Scenario A extended by queries in Listings \ref{listing_test_query_b1}, \ref{listing_test_query_b2}, and \ref{listing_test_query_b3}.
+
+\todo[inline]{Mozno popisat aj vyznam tychto dotazov?}
+
+The first two ones are simple queries containing one join pattern occurrence each.
+The third one contains occurrences of let join pattern (binding clauses at lines 5 and 6) and JP3 (binding clauses at lines 1 and 5, and where clause at line 8), two expressions exploitable by the inference of types (both at line 2) and one negative uniqueness statement of the comparison-with-a-constant form (comparison with \texttt{"m"} at line 2).
+Thus, the third query contains instances of all types of the constructs utilized by our method.
+
+\subsubsection{Results}
+
+\begin{table}
+\begin{tabular}{|l|}
+\hline
+\texttt{/site/open\_auctions/open\_auction/bidder/increase/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/closed\_auctions/closed\_auction/price/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{integer} \\ \hline
+\texttt{/site/open\_auctions/open\_auction/reserve} $\rightarrow$ \texttt{decimal} \\ \hline
+\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{decimal} \\ \hline
+\texttt{/site/people/person/profile/gender} $\rightarrow$ \texttt{string} \\ \hline
+\end{tabular}
+\caption{Inferred type statements B.}
+\label{TAB_inferred_types_b}
+\end{table}
+
+\begin{table}
+\begin{tabular}{|l|c|}
+\hline
+\textbf{Key} & \textbf{Weight} \\ \hline \hline
+\texttt{(/site//item, {@id})} & 0.333 \\ \hline
+\texttt{(/site/people/person, {@id})} & 0.6 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction, {buyer/@person})} & -0.6 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction, {itemref/@item})} & -0.917 \\ \hline
+\texttt{(/site/regions/europe/item, {@id})} & 0.6 \\ \hline
+\end{tabular}
+\caption{Inferred key statements B.}
+\label{TAB_inferred_keys_b}
+\end{table}
+
+\begin{table}
+\begin{tabular}{|l|c|}
+\hline
+\textbf{Node} & \textbf{Weight} \\ \hline \hline
+\texttt{(/site/people/person/profile/interest/@category)} & 1 \\ \hline
+\texttt{(/site/closed\_auctions/closed\_auction/itemref/@item)} & 1 \\ \hline
+\texttt{(/site/people/person/profile/gender)} & 0.9 \\ \hline
+\end{tabular}
+\caption{Inferred negative uniqueness statements B.}
+\label{TAB_inferred_negative_uniqueness_statements_b}
+\end{table}
+
+Results of this test scenario are shown in Tables \ref{TAB_inferred_types_b}, \ref{TAB_inferred_keys_b}, and \ref{TAB_inferred_negative_uniqueness_statements_b}. All seams to be as we expected.
+
+Note that all inferred type statements are correct, however, the weight of the \texttt{(/site//item, {@id})} key is only 0.333.
+The reason is that there is only one join pattern occurrence resulting to this key, and therefore in summary, its normalized weight is low and it is correct.
+
+We demonstrated that also the original method works and we can get better results with little larger input dataset.
+Though, this set is still very small and we do not show how the methods works with large real-world data, possibly containing FLWOR constructs not satisfying our assumption that each join is done by a key/foreign key pair.
+
+We attach the resulting XML Schema in Appendix \ref{listing_xsd}.
+It was made using threshold 0.3.
+It means that all key statements with normalized weight equal or higher than 0.3 are considered correct and are included in the schema.

### 2012-03-31 doc/theses/mikula/conclusion.tex
Master thesis: In progress.

@@
 The aim of this thesis was to employ XML operations in the XML schema inference.
-We analyzed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations, only to found one\todo{Jiny vyraz}, which infers keys from XQuery queries.
+We analyzed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations.
+We found only one method, inferring keys from XQuery queries.
 
@@
 \todo[inline]{Nejake prehladne zhrnutie plusov a minusov?}
+\todo[inline]{testy}
 

### 2012-03-31 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

@@
 \subsection{Auxiliary Functions} \label{step_4_auxiliary_functions}
-\todo[inline]{Ako je to s osou descendant or self pri hladani join patternov? Orig. metoda o nej explicitne nehovori, ale zrejme pouziva}
-\todo[inline]{Ako je to s attribute osou v L cestach join patternov? Atribut predsa moze byt kluc.}
 Firstly, we define some auxiliary functions used by algorithms in this chapter.

### 2012-03-31 jinfer/trunk/jInfer/BasicXQueryProcessor/src/cz/cuni/mff/ksi/jinfer/basicxqueryprocessor/keydiscovery/joinpatterns/JoinPatternsFinder.java
Master thesis implementation: Fixes.

@@ -146,7 +146,6 @@ public class JoinPatternsFinder {
           }
         }
       }
-      return;
     }
 
     // Check for and let join patterns.

### 2012-03-31 jinfer/trunk/jInfer/BasicXQueryProcessor/src/cz/cuni/mff/ksi/jinfer/basicxqueryprocessor/keydiscovery/negativeuniqueness/NegativeUniquenessFinder.java

@@ -74,15 +74,15 @@ public class NegativeUniquenessFinder {
   }
   
   private void processRecursive(final XQNode node) {
+    rejectionOfUniqueness_aggregationFunctions(node);
+    rejectionOfUniqueness_comparisonWithAConstant(node);
+    
     final List<XQNode> subnodes = node.getSubnodes();
     if (subnodes != null) {
       for (final XQNode subnode : node.getSubnodes()) {
         processRecursive(subnode);
       }
     }
-
-    rejectionOfUniqueness_aggregationFunctions(node);
-    rejectionOfUniqueness_comparisonWithAConstant(node);
   }
 
   private void rejectionOfUniqueness_aggregationFunctions(final XQNode node) {
@@ -131,7 +131,7 @@ public class NegativeUniquenessFinder {
 
     final WhereClauseNode whereClauseNode = flworNode.getWhereClauseNode();
     if (whereClauseNode != null) {
-      rejectionOfUniqueness_processWhere(flworNode, forVars);
+      rejectionOfUniqueness_processWhere(whereClauseNode.getExprNode(), forVars);
     }
   }
 
@@ -141,7 +141,10 @@ public class NegativeUniquenessFinder {
     }
 
     final Operator op = ((OperatorNode) whereExprNode).getOperator();
-    if (op == Operator.GEN_EQUALS) {
+    if (op == Operator.AND) {
+      rejectionOfUniqueness_processWhere(((OperatorNode)whereExprNode).getOperand(), forVars);
+      rejectionOfUniqueness_processWhere(((OperatorNode)whereExprNode).getRightSide(), forVars);
+    } else if (op == Operator.GEN_EQUALS) {
       final ExprNode leftOperand = ((OperatorNode) whereExprNode).getOperand();
       final ExprNode rightOperand = ((OperatorNode) whereExprNode).getRightSide();
 

### 2012-03-31 jinfer/trunk/jInfer/BasicXQueryProcessor/src/cz/cuni/mff/ksi/jinfer/basicxqueryprocessor/utils/BuiltinFunctionsUtils.java

@@ -40,6 +40,7 @@ public class BuiltinFunctionsUtils {
     builtinFunctionTypes.put("data", new UnknownType());
     builtinFunctionTypes.put("doc", new NodeType(NodeType.NodeTypeCategory.DOCUMENT, Cardinality.ONE));
     builtinFunctionTypes.put("min", new UnknownType());
+    builtinFunctionTypes.put("max", new UnknownType());
     builtinFunctionTypes.put("distinct-values", new UnknownType());
     builtinFunctionTypes.put("count", new XSDType(XSDBuiltinAtomicType.INTEGER, Cardinality.ONE));
     builtinFunctionTypes.put("exists", new XSDType(XSDBuiltinAtomicType.BOOLEAN, Cardinality.ONE));
@@ -136,7 +137,7 @@ public class BuiltinFunctionsUtils {
       final ItemTypeNode itemTypeNode = new AtomicTypeNode("xs:string");
       final ParamNode paramNode = new ParamNode("uri", new TypeNode(Cardinality.ZERO_OR_ONE, itemTypeNode));
       paramNodes.add(paramNode);
-    } else if (functionName.equals("min") || functionName.equals("distinct-values")) {
+    } else if (functionName.equals("min") || functionName.equals("max") || functionName.equals("distinct-values")) {
       final ParamNode paramNode1 = new ParamNode("arg", new TypeNode(Cardinality.ZERO_OR_MORE));
       final ItemTypeNode itemTypeNode2 = new AtomicTypeNode("xs:string");
       final ParamNode paramNode2 = new ParamNode("collation", new TypeNode(Cardinality.ONE, itemTypeNode2));

### 2012-03-31 jinfer/trunk/jInfer/BasicXSDExporter/src/cz/cuni/mff/ksi/jinfer/basicxsd/elementsexporters/AbstractElementsExporter.java

@@ -56,6 +56,7 @@ public abstract class AbstractElementsExporter {
   protected final String typenamePrefix;
   protected final String typenamePostfix;
   private int keyNumber = 1;
+  private int foreignKeyNumber = 1;
   private final Map<Key, String> keyNames = new HashMap<Key, String>();
   
   private static final String XQUERY_PROCESSOR_METADATA_TYPE = "xquery_processor_type";
@@ -245,7 +246,8 @@ public abstract class AbstractElementsExporter {
       final String keyName = keyNames.get(fKey.getKey());
       assert(keyName != null);
 
-      indentator.indent("<xs:keyref name=\"" + keyName + "Ref\" refer=\"" + keyName + "\">\n");
+      indentator.indent("<xs:keyref name=\"" + "keyRef" + foreignKeyNumber + '_' + keyName + "\" refer=\"" + keyName + "\">\n");
+      ++foreignKeyNumber;
       indentator.increaseIndentation();
       indentator.indent("<xs:selector xpath=\"" + targetPath.toString() + "\"/>\n");
       indentator.indent("<xs:field xpath=\"" + keyPath.toString() + "\"/>\n");

### 2012-03-31 jinfer/trunk/jInfer/BasicXQueryProcessor/src/cz/cuni/mff/ksi/jinfer/basicxqueryprocessor/merger/Merger.java
Master thesis implementation: Fixes.

@@ -92,7 +92,7 @@ public class Merger {
     writeInferredTypesToGrammar(topologicallySortedGrammar, inferredTypes);
   }
 
-  /**
+  /**<person>{$person/@id}</person><item>{$item/@id}</item><price>{$price}</price>
    * Merges specified inferred key statements with the grammar.
    * @param keys
    * @param foreignKeys
@@ -136,7 +136,7 @@ public class Merger {
       for (final AbstractStructuralNode node : contextSet.getNodes()) {
         final Map<String, Object> metadata = node.getMetadata();
         final XSDBuiltinAtomicType xsdType = (XSDBuiltinAtomicType) metadata.get(METADATA_KEY_TYPE);
-        if (type == null) {
+        if (xsdType == null) {
           metadata.put(METADATA_KEY_TYPE, inferredAtomicType);
           metadata.put(METADATA_KEY_HAS_PREDICATES, hasPredicates);
         } else {

### 2012-04-02 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.
Master thesis: In progress.

@@
 In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it.
+Despite the approach was proposed, it has not been implemented yet.
+Therefore we will be the first to implement it and perform consecutive testing.
+
 Firstly, the syntax tree is searched for forms of FLWOR expressions to infer the keys from.
@@
     $forVar$: A variable from the ancestor for binding node. \\
-    $forVars$: An array of couples of for variables and their for clause nodes. \\
     $checkJoinPattern3$: A flag determining whether to analyze the join pattern as possible join pattern 3. \\
@@
 
-\ENSURE Updated variable $forVars$.
-
-\IF{$P$ is of form $curBindingNode.varName/L_1 = var/L_2$}
+\IF{$P$ is of form $forVar/L_1 = curBindingNode.varName/L_2 \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_1) \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_2)$}
     \IF{$is(curBindingNode,$ ForClauseNode$)$}
         \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for join pattern
-        \STATE remove $(forVar, forBindingNode)$ from $forVars$
     \ELSIF{$is(curBindingNode,$ LetClauseNode$)$}
         \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for let pattern                
-        \STATE remove $(forVar, forBindingNode)$ from $forVars$
-    \ENDIF
-\ELSIF{$checkJoinPattern3$ AND $is(curBindingNode,$ ForClauseNode$)$}
-    \IF{$whereExpr$ is of form $bindingNode.varName/L_1 = var/L_2$}
+    \ENDIF
+\ENDIF
+\IF{$checkJoinPattern3 \wedge is(curBindingNode,$ ForClauseNode$) \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_1) \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_2)$}
+    \IF{$whereExpr$ is of form $forVar/L_1 = bindingNode.varName/L_2$}
         \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the join pattern 3
-        \STATE remove $(forVar, forBindingNode)$ from $forVars$
-    \ENDIF
-\ENDIF
-
-\RETURN $forVars$
+    \ENDIF
+\ENDIF
 \end{algorithmic}
@@
 
-When the algorithm discovers that a certain element is probably not unique, it assigns the element with a negative weight that will decrease the sum of its weights in the summarizing step.
+When the algorithm discovers that a certain element (or attribute) is probably not unique, it assigns the element (or attribute) with a negative weight that will decrease the sum of its weights in the summarizing step.
+
+We can formally define negative uniqueness statements as follows:
+
+\begin{define}[Negative uniqueness Statement]
+A negative uniqueness statement is a construct $(C, P)$, where $C$ and $P$ are PathType instances that use only child, descendant, and descendant or self ($P$ also attribute) axes.
+$C$ is called \emph{context path}, $P$ \emph{target path}.
+$C$ can be omitted as in the case of keys.
+\end{define}
+
+Note that the restrictions placed on the paths $C$ and $P$ are less restrictive as in the case of keys.
+If we found a negative uniqueness statement ($C$, $P$), where $P$ or $C$ contains predicates, then it implies that also $(C', P')$ is a negative uniqueness statement, where $C'$ and $P'$ are $C$ and $P$ stripped of predicates.
+In the case of special function calls, the explanation is the same.
 
@@
         \IF{$type$ is PathType \\ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(type)$ \\ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
-            \STATE memorize $argument$ as the negative key statement, weight 1
+            \STATE memorize $(argument)$ as the negative key statement, weight 1
         \ENDIF
@@
 Besides the aggregation functions, a certain form of FLWOR expressions is utilized in this step.
-They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference of a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
+They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference to a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
 
@@
 That implies that the value in those nodes is not unique, and, alike the utilization of the aggregation functions, non-unique nodes cannot be keys.
+This situation is handled in Algorithms \ref{ALG_rejection_of_uniqueness_comparison_with_a_constant} and \ref{ALG_rejection_of_uniqueness_function_process_where}.
 
@@
             \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is PathType $\wedge$ $isWithoutPredicatesExceptLastStep(getType(var/L))$ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(getType(var/L))$ $\wedge$ $C$ is a literal constant}
-                \STATE memorize $var/L$ as the negative key statement, weight 0.9
+                \STATE memorize $(var/L)$ as the negative key statement, weight 0.9
             \ENDIF
@@
 
-\todo[inline]{Definicia NUS a context Path, podobne ako v klucoch.}
+In spite of the presented algorithms search for global negative uniqueness statements ($C$ path is omitted) only, we included the context path in the definition for possible future extensions.
 

### 2012-04-02 doc/theses/mikula/diplomka.tex
Master thesis: In progress.

@@
 
-Abstrakt: \todo[inline]{TODO cca 10-15 riadkov}
+Abstrakt: V poslední době bylo vyvinuto množství metod zabívajících se automatickým odvozováním XML schémy, většina z nich ale používá XML dokumenty ako jediný vstup.
+My se v tehle práci zaměřujeme na rozšíření odvozování začleněním XML operací, konkrétně XQuery dotazů.
+Diskutujeme jak je možné využít XQuery dotazy na vylepšení procesu odvozování a navrhujeme algoritmus založený na vybraných vylepšeních, který rozšiřuje existujíci metodu hledání klíčů a který může být začleněn do tzv.
+\emph{metod odvozujících gramatiku}.
+Implementováním algoritmu sme vytvořili první řešení odvozování XML schémy, ktoré využívá XML dokumenty spolu s XML operacemi.
 % abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce
 
-Klíčová slova: \todo[inline]{TODO 5-8 slov}
+Klíčová slova: XML, XML schéma, odvozování, XQuery, dotazy, XML operace
 % 3 aľ 5 klíčových slov
@@
 
-Abstract: \todo[inline]{TODO}
+Abstract: Recently, plenty of methods dealing with an automatic inference of XML schema have been developed, however, most of them utilize XML documents as their only input.
+In this thesis we focus on extending the inference by incorporating XML operations, in particular XQuery queries.
+We discuss on how can XQuery queries help in improving the inference process and we propose an algorithm based on chosen improvements, extending an existing method of a key discovery, that can be integrated to so-called \emph{grammar-inferring} methods.
+By implementing it, we created the first solution of XML schema inference using XML documents along with XML operations.
 % abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad
@@
 
-Keywords: \todo[inline]{TODO}
+Keywords: XML, XML schema, inference, XQuery, queries, XML operations
 % 3 aľ 5 klíčových slov v angličtině

### 2012-04-06 doc/theses/mikula/conclusion.tex
Master thesis: In progress.
Master thesis: Corrections.

@@
 \chapter{Conclusion}
-The aim of this thesis was to employ XML operations in the XML schema inference.
-We analyzed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations.
+The aim of this thesis was to employ XML operations in the XML schema inference process.
+We analyzed several existing methods of the XML schema inference and we searched for methods that utilize selected XML operations.
 We found only one method, inferring keys from XQuery queries.
@@
 
-Before a creation of the algorithm itself, we had to take several decisions in questions that emerged.
+Before creation of the algorithm itself, we had to take several decisions in questions that emerged.
 Since there is a lack of the methods dealing with the utilisation of XML operation, there is also a lack of practically proven solutions that can help in such decision making.
@@
 We also implemented several ideas from the overview to infer XSD built-in types of elements and attributes.
-And we extended the one existing method dealing with the inference of keys.
+And we extended and implemented the one existing method dealing with the inference of keys.
 We experimentally demonstrated that on some input files, results of the extended method are better than results of the original method.
-However, we did not succeed in a search of an ideal, large enough set of test data.
+However, we did not succeed in the search of an ideal, large enough set of test data.
 The testing was performed on a small set of input queries and further testing and algorithm tuning is required.
@@
 
-\todo[inline]{Nejake prehladne zhrnutie plusov a minusov?}
-\todo[inline]{testy}
+\todo[inline]{Nejake prehladne zhrnutie plusov a minusov? ANO}
 
 \section{Future Work}
-This work implements only some ideas discussed in the overview in Chapter \ref{CHAPTER_analysis_of_xquery}, leaving most of them for a future work.
+This work implements only some ideas discussed in the overview in Chapter \ref{CHAPTER_analysis_of_xquery}, leaving most of them for future work.
 Also, the implemented algorithms can be further refined as was already mentioned in Chapter \ref{CHAPTER_proposed_algorithm} with the presentation of the algorithms.
@@
 
-The utilization of XQuery queries certainly provides a space for incorporating an interaction with user. For example, a user may influence the scoring of inferred keys to get more precise results.
+The utilization of XQuery queries certainly provides a space for incorporating interaction with user. For example, a user may influence the scoring of inferred keys to get more precise results.
 

### 2012-04-06 doc/theses/mikula/diplomka.tex
Master thesis: Corrections.

--- 2012-03-31 doc/theses/mikula/diplomka.tex
@@
 
-Abstrakt: \todo[inline]{TODO cca 10-15 riadkov}
+Abstrakt: V poslednej dobe bolo vyvinutých množstvo metód zaoberajúcich sa automatickým odvodzovaním XML schémy.
+Väčšina z nich ale používa XML dokumenty ako jediný vstup.
+My sa v tejto práci zameriavame na rozšírenie odvo\-dzo\-vania zahrnutím XML operácií, konkrétne XQuery dotazov.
+Diskutujeme ako je možné využiť XQuery dotazy na vylepšenie procesu odvodzovania a navrhujeme algoritmus založený na vybraných vylepšeniach, ktorý rozširuje existujúcu metódu hľadania kľúčov a ktorý môže byť začlenený do tzv.
+\emph{metód odvodzujúcich gramatiku}.
+Implementovaním algoritmu sme vytvorili prvé kompletné riešenie odvodzovania XML schémy, ktoré využíva XML dokumenty spolu s XML operáciami.
 % abstrakt v rozsahu 80-200 slov; nejedná se vąak o opis zadání diplomové práce
 
-Klíčová slova: \todo[inline]{TODO 5-8 slov}
+Klíčová slova: XML, XML schéma, odvozování, XQuery, dotazy, XML operace
 % 3 aľ 5 klíčových slov
@@
 
-Abstract: \todo[inline]{TODO}
+Abstract: Recently, plenty of methods dealing with automatic inference of XML schema have been developed, however, most of them utilize XML documents as their only input.
+In this thesis we focus on extending inference by incorporating XML operations, in particular XQuery queries.
+We discuss how can XQuery queries help in improving the inference process and we propose an algorithm based on chosen improvements, extending an existing method of a key discovery, that can be integrated to so-called \emph{grammar-inferring} methods.
+By implementing it, we created the first solution of XML schema inference using XML documents along with XML operations.
 % abstrakt v rozsahu 80-200 slov v angličtině; nejedná se vąak o překlad

### 2012-04-06 doc/theses/mikula/experiments.tex
Master thesis: Corrections.

@@
 
-\todo[inline]{Mozno popisat aj vyznam tychto dotazov?}
-
 The first two ones are simple queries containing one join pattern occurrence each.
@@
 
-Results of this test scenario are shown in Tables \ref{TAB_inferred_types_b}, \ref{TAB_inferred_keys_b}, and \ref{TAB_inferred_negative_uniqueness_statements_b}. All seams to be as we expected.
+Results of this test scenario are shown in Tables \ref{TAB_inferred_types_b}, \ref{TAB_inferred_keys_b}, and \ref{TAB_inferred_negative_uniqueness_statements_b}. All seem to be as we expected.
 
@@
 We demonstrated that also the original method works and we can get better results with little larger input dataset.
-Though, this set is still very small and we do not show how the methods works with large real-world data, possibly containing FLWOR constructs not satisfying our assumption that each join is done by a key/foreign key pair.
+Though, this set is still very small and we do not show how the methods work with large real-world data, possibly containing FLWOR constructs not satisfying our assumption that each join is done by a key/foreign key pair.
 
-We attach the resulting XML Schema in Appendix \ref{listing_xsd}.
+We attach the resulting XSD in Appendix \ref{listing_xsd}.
 It was made using threshold 0.3.

### 2012-04-06 doc/theses/mikula/introduction.tex
Master thesis: Corrections.

@@
 
-In Chapter 5, we discusses some essential questions that we needed to decide before a development of an algorithm.
+In Chapter 5, we discuss some essential questions that we needed to decide before a development of an algorithm.
 
@@
 
-An implementation of the proposed algorithm using a jInfer framework is described in Chapter 8. Chapter 9 then discusses performed experiments and their outcomes.
+An implementation of the proposed algorithm using the jInfer framework is described in Chapter 8. Chapter 9 then discusses performed experiments and their outcomes. And, finally, Chapter 10 concludes.

### 2012-04-06 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

@@
          \STATE $substeps[step] := type$
-         \STATE $pathExprNode :=$ new instance of PathExprNode containing one step step
+         \STATE $pathExprNode :=$ new instance of PathExprNode containing one step
          \STATE $type :=$ PathType containing $pathExprNode$ and $substeps$
@@
 \section{Step 4: Key Discovery}
-\todo[inline]{Tu by neskodilo napisat, ze sice preberame alg, ale ho aj implementujeme a overime, co v povodnom clanku nebolo.}
 In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it.
@@
 
-Note that the restrictions placed on the paths $C$ and $P$ are less restrictive as in the case of keys.
-If we found a negative uniqueness statement ($C$, $P$), where $P$ or $C$ contains predicates, then it implies that also $(C', P')$ is a negative uniqueness statement, where $C'$ and $P'$ are $C$ and $P$ stripped of predicates.
+Note that the restrictions placed on the paths $C$ and $P$ are less restrictive than in the case of keys.
+If we find a negative uniqueness statement ($C$, $P$), where $P$ or $C$ contains predicates, then it implies that also $(C', P')$ is a negative uniqueness statement, where $C'$ and $P'$ are $C$ and $P$ stripped of predicates.
 In the case of special function calls, the explanation is the same.
@@
 The original approach \cite{Necasky:2009:DXK:1529282.1529414} employs the following normalization which we adopt.
-Is also incorporates a heuristic for precision enhancement.
+It also incorporates a heuristic for precision enhancement.
 For details, we refer to it.

### 2012-04-06 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: Corrections and final(?) additions.

@@
 
-\section{On Inference of XML Schema with the Knowledge of an Obsolete One}
+\section[On Inference of XML Schema with the Knowledge of an Obs. One]{On Inference of XML Schema with the Knowledge of an Obsolete One}
 The aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently.
@@
 In the second step, regular expressions from the corrected schema are specialized with regard to the XML documents, resulting in a more precise and readable schema.
-Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.
+Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simpler form, if this transcription preserves the validity, and refactorization to improve readability.
 

### 2012-04-06 doc/theses/mikula/experiments.tex
Master thesis: Corrections and final(?) additions.

@@
 \end{tabular}
-\caption{Inferred type statements.}
+\caption{Inferred type statements A}
 \label{TAB_inferred_types}
@@
 \end{tabular}
-\caption{Inferred key statements.}
+\caption{Inferred key statements A}
 \label{TAB_inferred_keys}
@@
 \end{tabular}
-\caption{Key statements inferred from the extended test data.}
+\caption{Key statements inferred from the extended test data A}
 \label{TAB_inferred_keys_2}
@@
 \end{tabular}
-\caption[Key statements inferred using the modified JP3 statements.]{Key statements inferred from the extended test data using the modified JP3 statements.}
+\caption[Key statements inferred using the modified JP3 statements]{Key statements inferred from the extended test data using the modified JP3 statements}
 \label{TAB_inferred_keys_3}
@@
 \end{tabular}
-\caption{Inferred type statements B.}
+\caption{Inferred type statements B}
 \label{TAB_inferred_types_b}
@@
 \end{tabular}
-\caption{Inferred key statements B.}
+\caption{Inferred key statements B}
 \label{TAB_inferred_keys_b}
@@
 \end{tabular}
-\caption{Inferred negative uniqueness statements B.}
+\caption{Inferred negative uniqueness statements B}
 \label{TAB_inferred_negative_uniqueness_statements_b}

### 2012-04-06 doc/theses/mikula/conclusion.tex
Master thesis: Corrections and final(?) additions.

@@
 Finally, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using the jInfer framework.
-Thus, we created the first complete, ready-to-use, and extensible implementation of the XML xchema inference exploiting XQuery queries besides XML data.
+Thus, we created the first complete, ready-to-use, and extensible implementation of the XML schema inference exploiting XQuery queries besides XML data.
 
-\todo[inline]{Nejake prehladne zhrnutie plusov a minusov? ANO}
+As the main advantages and disadvantages of the work we list the following.
+\begin{itemize}
+\renewcommand{\labelitemi}{$+$}
+\item Elaboration of the overview how XQuery queries can help in the process of XML schema inference.
+\item Incorporation of the lexical and syntax analyses of input queries creating their syntax trees.
+\item Development of the static type analysis including PathTypes, which can be easily extended and use in possible future extensions.
+\item Extension of the existing method of key discovery, achieving better results.
+\item Our implementation using the jInfer framework is the first implementation of XML schema inference method utilizing XML operations.
+\renewcommand{\labelitemi}{$-$}
+\item In the proposed algorithm we dealt with a small part of possible refinements discussed in the overview.
+\item The combination with existing methods of inference is done in the simplest way, not modifying the grammar rules.
+But, from the view of recent XML schema inference research, the modification of the grammar rules according to the statements inferred from XML operations is vastly interesting.
+\item Since large real-world sets of XQuery queries are not available, we did not perform experiments using such sets.
+We only tested our method using a relatively small number input of queries, decreasing the level of reliability of the experimental results.
+\end{itemize}
 
 \section{Future Work}
-This work implements only some ideas discussed in the overview in Chapter \ref{CHAPTER_analysis_of_xquery}, leaving most of them for future work.
+As already mentioned, this work implements only some ideas discussed in the overview in Chapter \ref{CHAPTER_analysis_of_xquery}, leaving most of them for future work.
 Also, the implemented algorithms can be further refined as was already mentioned in Chapter \ref{CHAPTER_proposed_algorithm} with the presentation of the algorithms.

### 2012-04-06 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections and final(?) additions.

@@
 \includegraphics[scale=1]{algorithm.eps}
-\caption{Steps of the proposed algorithm.}
+\caption{Steps of the proposed algorithm}
 \label{FIG_steps_algorithm}
@@
 The algorithm searches for expressions in the syntax trees and statically (without evaluation) determines their types.
-\todo{Co se mysli?
-Datove typy?}
+See Section \ref{SECTION_expression_types}.
 \item \textbf{Inference of built-in types}. When the types of expressions are determined, selected forms of expression are utilized to infer types of elements and attributes.
@@
 \includegraphics[scale=0.6]{syntax_tree_1.eps}
-\caption{Sample syntax tree 1.}
+\caption{Sample syntax tree 1}
 \label{FIG_syntax_tree_1}
@@
 
-\subsection{Expression Types}
+\subsection{Expression Types} \label{SECTION_expression_types}
 \begin{figure}
-\caption{XSD built-in atomic types.}
+\caption{XSD built-in atomic types}
 \label{FIG_xsd_built_in_atomic_types}
@@
 \includegraphics[scale=0.5]{syntax_tree_1_types.eps}
-\caption{Sample syntax tree 1 after the static analysis of types.}
+\caption{Sample syntax tree 1 after the static analysis of types}
 \label{FIG_syntax_tree_1_types}
@@
 \section{Step 3: Inference of Built-in Types}
-\todo[inline]{Rekurzia sa nevnara do FunctionBodyNodes a PathExprNodes - napisat, alebo je to detail?}
 In this step, the algorithm goes through the syntax tree to infer types of elements and attributes from expressions using the type information from the previous step.
@@
 
+Likewise the previous step, the syntax tree is recursively searched for expressions meeting the conditions for the type inference.
+A little difference is that the recursion stops at \texttt{FunctionsBodyNode} and \texttt{PathExprNode} nodes, because the processing of these nodes requires a different approach, which we do not deal with in this thesis.
+
 The output of this step is a set of statements of the form $P \rightarrow T$, where $P$ is an instance of PathType and $T$ is an XML Schema built-in atomic type.
@@
 \includegraphics[scale=1]{syntax_tree_1_types_type_inference.eps}
-\caption{Example of an inference of a type from a function call.}
+\caption{Example of an inference of a type from a function call}
 \label{FIG_syntax_tree_1_types_type_inference}
@@
 \includegraphics[scale=0.7]{syntax_tree_2.eps}
-\caption{Sample syntax tree 2.}
+\caption{Sample syntax tree 2}
 \label{FIG_syntax_tree_2}

### 2012-04-09 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: Corrections.

@@
 In the last step, this set will be used to compose a (sub)optimal result with respect to a trade-off between its preciseness and conciseness.
-Therefore, it is desirable to create DTD canditates with various degrees of these two characteristics.
+Therefore, it is desirable to create DTD candidates with various degrees of these two characteristics.
 
-Generalization is based on replacing fragmets (sequences of subelements of a given element) from the input XML data by regular expressions, thus, using metacharacters like (*, +, ?).
-To provide a wide set of DTD canditates, each sequence is processed several times using various values of input parameters.
-Due to the very large number of possible DTD canditates, the authors employ certain real-life motivated heuristics.
+Generalization is based on replacing fragments (sequences of subelements of a given element) from the input XML data by regular expressions, thus, using metacharacters like (*, +, ?).
+To provide a wide set of DTD candidates, each sequence is processed several times using various values of input parameters.
+Due to the very large number of possible DTD candidates, the authors employ certain real-life motivated heuristics.
 
@@
 \subsection{Factoring}
-Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decreasing their summed size without modifications in their semantics.
+Factoring is a process of creating a new DTD candidate from two or more DTD candidates, decreasing their summed size without modifications in their semantics.
 The aim of this step is to decrease the MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.
@@
 In search for the optimal schema, \emph{Ant Colony Optimization (ACO)} heuristic is incorporated.
-The idea behind the ACO heuristic is that a set of artifical ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum.
+The idea behind the ACO heuristic is that a set of artificial ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum.
 An ant is performing steps (schema modifications), dying after a predefined number of steps and providing an information - positive feedback - on the quality of a solution found.
-The search is performed in a defined number of interations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
+The search is performed in a defined number of iterations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations.
 Every step of an ant represents a modification of a schema, in particular, a merge of states in a corresponding PTA.
@@
 Therefore, the authors introduce a scoring function to summarize the positive and the negative statements about keys using the assigned weights.
-The value of the score expresses the probalitity that a respective key statement is satisfied.
+The value of the score expresses the probability that a respective key statement is satisfied.
 Finally, the scores of the inferred keys are normalized to be comparable with each other.

### 2012-04-09 doc/theses/mikula/introduction.tex
Master thesis: Corrections.

@@
 In order to this need, several so-called XML schema languages (XML schemata) have been created.
-The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by the World Wide Web Consorcium (W3C).
+The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by the World Wide Web Consortium (W3C).
 

### 2012-04-09 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: Corrections.

@@
 Thus, the return expression is \texttt{//A[@id = "id"]/B}.
-Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equaling string value \texttt{"id"} is of some numeric type.
+Therefore, we can infer that element \texttt{B} in element \texttt{A} with attribute \texttt{id} equalling string value \texttt{"id"} is of some numeric type.
 And, since the function is parametrized, there is a notion that this statement may be correct also for other elements \texttt{A} and \texttt{B}.

### 2012-04-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

@@
 We will not determine types of functions defined in other modules since the principle is similar but it is needed to look up the definitions in syntax trees of other queries.
-Return types of built-in functions are fixed and, thus, there is no need to analyse them.
+Return types of built-in functions are fixed and, thus, there is no need to analyze them.
 Also, since a determination of a prefix for built-in functions is a technical issue, we assume that built-in functions are either prefixed by \texttt{fn} or not prefixed.
@@
 If it is not, it may often be deducible from the binding expression.
-Again, we do not analyse external variables for the same reason as in case of external functions.
+Again, we do not analyze external variables for the same reason as in case of external functions.
 
@@
 In the function body, replace the formal arguments by the real arguments and set them as the local variables, so the proposed algorithms can be applied to process the function's body.
-Analyse the types of expressions in the body with the new information on the types of arguments by incorporating of the proposed algorithms and replace the function's return type by the one currently determined.
+Analyze the types of expressions in the body with the new information on the types of arguments by incorporating of the proposed algorithms and replace the function's return type by the one currently determined.
 

### 2012-04-09 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: Corrections and final(?) additions.
Master thesis: Corrections.

@@
 \end{verbatim}
-\caption{A simple XML example.}
+\caption{A simple XML example}
 \label{FIG_a_simple_xml_example}
@@
 
-The dafault value is either a literal value or one of the following specifiers.
+The default value is either a literal value or one of the following specifiers.
 
@@
 \end{verbatim}
-\caption{A simple DTD example.}
+\caption{A simple DTD example}
 \label{FIG_a_simple_dtd_example}
@@
 
-The XSD language consists of many more constructs we do not mention such as substitution groups, type extensions, integrity constranints and other.
+The XSD language consists of many more constructs we do not mention such as substitution groups, type extensions, integrity constraints and other.
 
@@
 \end{alltt}
-\caption{A simple XSD example.}
+\caption{A simple XSD example}
 \label{FIG_a_simple_xsd_example}
@@
 \item \texttt{$P$/$identifier$} stands for \texttt{$P$/child::$identifier$} (child is a default axis)
-\item \texttt{$P$/@$identifier$} stands for \texttt{$P$/atribute::$identifier$}
+\item \texttt{$P$/@$identifier$} stands for \texttt{$P$/attribute::$identifier$}
 \item \texttt{$P$/../$identifier$} stands for \texttt{$P$/parent::*/$identifier$}
@@
 \subsection{FLWOR Expressions}
-A basic construct of the XQuery language is $FLWOR$. It is an abbrevation of its five clauses: for, let, where, order by, return. For clause is \\
+A basic construct of the XQuery language is $FLWOR$. It is an abbreviation of its five clauses: for, let, where, order by, return. For clause is \\
 \texttt{for $var$ in $expr$} \\

### 2012-04-09 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: Corrections.

@@
 
-Path expressions without predicates which use only child axis are the simpliest example of such queries.
+Path expressions without predicates which use only child axis are the simplest example of such queries.
 For example, path expression \texttt{/bib/book/author} indicates that element \texttt{bib} is the root element and it contains one or more elements \texttt{book} and these contain one or more elements \texttt{author}.
@@
 In addition, queries may not cover the whole relevant content of the documents.
-Let \texttt{E} be an element occurring in an XML document and let \texttt{Q} be an XQuery query.
+Let $E$ be an element occurring in an XML document and let $Q$ be an XQuery query.
 Consider the following cases:
@@
 \begin{enumerate}[1)]
-\item \texttt{E} is directly mentioned in \texttt{Q}. For example elements \texttt{bib}, \texttt{book} and \texttt{author} are mentioned in path expression \texttt{/bib/book/author}.
-\item \texttt{E} is not mentioned in \texttt{Q} but it does occur in a result of the evaluation of \texttt{Q}.
+\item $E$ is directly mentioned in $Q$. For example elements \texttt{bib}, \texttt{book} and \texttt{author} are mentioned in path expression \texttt{/bib/book/author}.
+\item $E$ is not mentioned in $Q$ but it does occur in a result of the evaluation of $Q$.
 For example path expression \texttt{/bib/book/author} returns elements \texttt{author} along with their subelements \texttt{name}, \texttt{birthdate} and \texttt{nationality}.
-\item \texttt{E} is neither mentioned in \texttt{Q}, nor it occurs in the result, but it is processed by the evaluation of \texttt{Q}.
+\item $E$ is neither mentioned in $Q$, nor it occurs in the result, but it is processed by the evaluation of $Q$.
 For example the evaluation of path expression \texttt{//author} processes amongst others elements \texttt{bib} and \texttt{book}, in a search for elements \texttt{author}.
-\item \texttt{E} occurs in a part of the XML document not related to \texttt{Q} at all.
+\item $E$ occurs in a part of the XML document not related to $Q$ at all.
 \end{enumerate}
 
-How we can obtain any information about element \texttt{E} from query \texttt{Q}?
-In case 1), \texttt{Q} can be directly exploited in the inference process.
-Case 2) requires the result of \texttt{Q} and additional processing of the result.
-Case 3) requires a step-by-step evaluation of \texttt{Q}, hence, the XML document is also required.
+How we can obtain any information about element $E$ from query $Q$?
+In case 1), $Q$ can be directly exploited in the inference process.
+Case 2) requires the result of $Q$ and additional processing of the result.
+Case 3) requires a step-by-step evaluation of $Q$, hence, the XML document is also required.
 And in case 4), it is not possible to obtain any information.
@@
 
-Basic representant is a common \emph{FLWOR} query.
+Basic representant is a common FLWOR query.
 

### 2012-04-09 doc/theses/mikula/diplomka.tex
Master thesis: Corrections and final(?) additions.
Master thesis: Corrections.

@@
 \hss}\hbox to 0.5\hsize{%
-Signature
+%Signature
 \hss}}
@@
 My sa v tejto práci zameriavame na rozšírenie odvo\-dzo\-vania zahrnutím XML operácií, konkrétne XQuery dotazov.
-Diskutujeme ako je možné využiť XQuery dotazy na vylepšenie procesu odvodzovania a navrhujeme algoritmus založený na vybraných vylepšeniach, ktorý rozširuje existujúcu metódu hľadania kľúčov a ktorý môže byť začlenený do tzv.
-\emph{metód odvodzujúcich gramatiku}.
+Diskutujeme ako je možné využiť XQuery dotazy na vylepšenie procesu odvodzovania a navrhujeme algoritmus založený na vybraných vylepšeniach, ktorý rozširuje existujúcu metódu hľadania kľúčov a ktorý môže byť začlenený do metód odvodzujúcich tzv.
+\emph{počiatočnú gramatiku}.
 Implementovaním algoritmu sme vytvorili prvé kompletné riešenie odvodzovania XML schémy, ktoré využíva XML dokumenty spolu s XML operáciami.
@@
 In this thesis we focus on extending inference by incorporating XML operations, in particular XQuery queries.
-We discuss how can XQuery queries help in improving the inference process and we propose an algorithm based on chosen improvements, extending an existing method of a key discovery, that can be integrated to so-called \emph{grammar-inferring} methods.
+We discuss how can XQuery queries help in improving the inference process and we propose an algorithm based on chosen improvements, extending an existing method of a key discovery, that can be integrated to methods inferring so-called \emph{initial grammar}.
 By implementing it, we created the first solution of XML schema inference using XML documents along with XML operations.
@@
 \newtheorem{example}{Example}
-
-%%% Tabulky v diplomové práci, existují-li.
-%\chapwithtoc{List of Tables}
-\listoftables
 
@@
 \include{appendix_b}
+\include{appendix_c}
+
+%%% Tabulky v diplomové práci, existují-li.
+%\chapwithtoc{List of Tables}
+\listoftables
+\listoffigures
 

### 2012-04-09 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: Corrections.

@@
       <xs:element name="title" type="xs:string"/>
-      <xs:element name="authors>
+      <xs:element name="authors">
         <xs:complexType>
+          <xs:sequence>
             <xs:element name="author" type="xs:string"
-           maxOccurs="unbound"/>
+             maxOccurs="unbounded"/>
+          </xs:sequnce>
         </xs:complexType>
@@
     <xs:attribute name="id" type="xs:string" use="required"/>
-  </xs:complexContent>
+  </xs:complexType>
 </xs:element>

### 2012-04-09 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: Corrections.

@@
 
-Generalization is based on replacing fragments (sequences of subelements of a given element) from the input XML data by regular expressions, thus, using metacharacters like (*, +, ?).
+Generalization is based on replacing fragments (sequences of subelements of a given element) from the input XML data by regular expressions, thus, using metacharacters like \texttt{*}, \texttt{+}, \texttt{?}.
 To provide a wide set of DTD candidates, each sequence is processed several times using various values of input parameters.

### 2012-04-09 doc/theses/mikula/conclusion.tex
Master thesis: Corrections.

@@
 
-Finally, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using the jInfer framework.
+Finally, we proposed a simple way how to combine the inferred statements with existing methods of the XML schema inference inferring initial grammar and we implemented it using the jInfer framework.
 Thus, we created the first complete, ready-to-use, and extensible implementation of the XML schema inference exploiting XQuery queries besides XML data.

### 2012-04-09 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: Corrections.

@@
 Our statements are of the three following forms.
-The first one is $p \rightarrow t$, where $p$ is a PathType representing an XQuery path selecting a set of elements or attributes, and $t$ is an XSD built-in atomic type.
+The first one is $P \rightarrow T$, where $P$ is a PathType representing an XQuery path selecting a set of elements or attributes, and $T$ is an XSD built-in atomic type.
 
-The second and third forms are $k = (C,P,\{L\})$ and $fk = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key, where $C,P,P_1,P_2,L,L_1,L_2$ are PathTypes without predicates using only child and descendant axes.
+The second and third forms are $K = (C,P,\{L\})$ and $K_f = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key.
 
-In all three cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form).
+In all three cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($P$ from the first form and $C$ from the second and third form).
 This is done in two steps.
@@
 In case of inferred types, the situation is slightly less simple, because there can occur conflicting statements.
-The examples of such conflicts for path $p$ without predicates are $p_1 \rightarrow date$, $p_2 \rightarrow string$, and, $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.
+The examples of such conflicts for path $P$ without predicates are $P_1 \rightarrow$ \texttt{date}, $P_2 \rightarrow$ \texttt{string}, and, $P_1 \rightarrow$ \texttt{byte}, $P_2 \rightarrow$ \texttt{int}, where $P_1$, $P_2$ are paths that when stripped of predicates, they equal $P$.
 
-Note that, in both examples, one type is castable to the other (date to string, and byte to int).
-Both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a more convenient expression and is more precise.
+Note that, in both examples, one type is castable to the other (\texttt{date} to \texttt{string}, and \texttt{byte} to \texttt{int}).
+Both types are inferred correctly for nodes targeted by $P$, but one of them was inferred from a more convenient expression and is more precise.
 
-Consider these two expressions; PathType $p$ is compared to an integer literal constant, and, PathType $p$ is an argument of function where a formal type of the argument is byte.
-The first expression is utilized to infer statement $p \rightarrow integer$, and the second one to infer statement $p \rightarrow byte$.
+Consider these two expressions; PathType $P$ is compared to an integral literal constant, and, PathType $P$ is an argument of a function where a formal type of the argument is \texttt{byte}.
+The first expression is utilized to infer statement $P \rightarrow$ \texttt{integer}, and the second one to infer statement $P \rightarrow$ \texttt{byte}.
 Both of the types are correct, but the second one is more accurate.
 
-A problem emerges for example if PathType $p$ is compared to an integer constant, and the real type of elements (or attributes) selected by $p$ is \texttt{double}.
-In that case, statement $p \rightarrow$ \texttt{integer} is inferred, but it is not correct.
+A problem emerges for example if PathType $P$ is compared to an integral constant, and the real type of elements (or attributes) selected by $P$ is \texttt{double}.
+In that case, statement $P \rightarrow$ \texttt{integer} is inferred, but it is not correct.
 
 \subsection{Verification using XML data}
-To solves those problems, we propose a simple verification using XML data.
+To solve the problems, we propose a simple verification using XML data.
 
-For each normalized PathType $p$ from the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types.
-$T_p = \{t|(p \rightarrow t) \in \mathcal{S}_t\}$.
-Then, we create sequence $T_p'$ by ordering the set $T_p$ from the most specific type to the most general one.
-For example, if $T_p = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $T_p' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.
+For each normalized PathType $P$ from the inferred type statements $\mathcal{S}_t$, we find set $\mathcal{T}_P$ of all inferred types.
+$\mathcal{T}_P = \{T|(P \rightarrow T) \in \mathcal{S}_t\}$.
+Then, we create sequence $\mathcal{T}_P'$ by ordering the set $\mathcal{T}_P$ from the most specific type to the most general one.
+For example, if $\mathcal{T}_P = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $\mathcal{T}_P' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.
 
-Since we have the XML data and path $p$ is an XQuery path, we can use an XQuery processor (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$.
-The verification algorithm iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$.
-If so, $t'$ is the inferred type for nodes $N$ (and PathType $p$), else the inferred type is \texttt{string}.
+Since we have the XML data and path $P$ is an XQuery path, we can use an XQuery processor (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $P$.
+The verification algorithm iterates through $\mathcal{T}_P'$ and for each $T' \in \mathcal{T}_P'$ it checks if every node in $N$ conforms to $T'$.
+If so, $T'$ is the inferred type for nodes $N$ (and PathType $P$), else the inferred type is \texttt{string}.

### 2012-04-09 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.

@@
 PathType contains additional information.
-The represented path is contained by an instance of \texttt{PathExprNode} containing list of its steps, in particular instances of \texttt{StepExprNode}.
+The represented path is contained by a list of its steps, in particular instances of \texttt{StepExprNode}.
 If a step is a reference to a variable whose type is PathType, we also want to include this information.
@@
 \begin{itemize}
-\item \texttt{pathExprNode} - An instance of \texttt{PathExprNode}.
-\item \texttt{substeps} - An association between steps from \texttt{pathExprNode} and instances of PathType type.
+\item \texttt{steps} - A list of \texttt{PathExprNode} instances.
+\item \texttt{substeps} - An association between variable-referencing steps and instances of PathType type.
 \item \texttt{isForBound} - Boolean flag determining if the type was bound to a for variable in a for clause.
@@
          \STATE $substeps[step] := type$
-         \STATE $pathExprNode :=$ new instance of PathExprNode containing one step
-         \STATE $type :=$ PathType containing $pathExprNode$ and $substeps$
+         \STATE $steps :=$ an array containing one item which is $step$
+         \STATE $type :=$ PathType containing $steps$ and $substeps$
      \ENDIF
@@
 
-A part of the function's semantic is separated in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
+A parts of the function's semantic are separated in other functions, like the one in Algorithm \ref{ALG_func_determineOperatorType}, which is not presented completely because of the same reason.
 
@@
 
-\STATE $stepNodes :=$ an empty array
-\STATE $forBoundSubsteps :=$ an empty associative array
-\FORALL{$stepNode \in pathExprNode.getChildren()$}
-	\STATE $detailNode := stepNode.getChild(ExprNode)$
+\STATE $steps :=$ an empty array
+\STATE $substeps :=$ an empty associative array
+\FORALL{$step \in pathExprNode.getChildren()$}
+	\STATE $detailNode := step.getChild(ExprNode)$
 	\IF{$detailNode \neq$ null}
@@
 			\IF{$type.isForBound$}
-				\STATE $forBoundSubsteps[stepNode] := type$
-				\STATE add $stepNode$ to $stepNodes$
+				\STATE $substeps[stepNode] := type$
+				\STATE add $step$ to $steps$
 			\ELSE
-				\STATE add all $type.stepNodes$ to $stepNodes$
+				\STATE add all $type.steps$ to $steps$
 			\ENDIF
 		\ELSE
-			\STATE add $stepNode$ to $stepNodes$
+			\STATE add $step$ to $steps$
 		\ENDIF
 	\ELSE
-		\STATE add $stepNode$ to $stepNodes$
-	\ENDIF
-\ENDFOR
-
-\RETURN $PathType$ with $stepNodes$, $forBoundSubsteps$, and $isForBound$ set to false
+		\STATE add $step$ to $steps$
+	\ENDIF
+\ENDFOR
+
+\RETURN $PathType$ with $steps$, $substeps$, and $isForBound$ set to false
 \end{algorithmic}
@@
 Only expressions of a particular type are exploited.
-Specifically, an expression has to contain an subexpression $E$ of PathType type (expression representing a certain element or attribute or a set of elements or attributes).
+Specifically, an expression has to contain a subexpression $E$ of PathType type (expression representing a certain element or attribute or a set of elements or attributes).
 In the following text, the set represented by expression $E$ is denoted $S$.
@@
 This case is quite straightforward.
-The algorithm encounters a function call and one of the arguments is a set of elements of attributes (subexpression $E$ representing $S$) represented by PathType $p$.
+The algorithm encounters a function call and one of the arguments is a set of elements of attributes (subexpression $E$ representing $S$) represented by PathType $P$.
 To determine the type of $S$, it is only needed to determine the type of the corresponding formal argument from the definition of the function.
@@
 
-If the type $t$ of the formal argument is a built-in type or its sequence, then $t$ is also the inferred type of $S$. The inferred statement is $p \rightarrow t$. Otherwise, no statement is inferred.
+If the type $T$ of the formal argument is a built-in type or its sequence, then $T$ is also the inferred type of $S$. The inferred statement is $P \rightarrow T$. Otherwise, no statement is inferred.
 
 \subsection{Arithmetic Operations}
-If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod}, \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD}, \texttt{MUL}, \texttt{DIV} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $p$ and the type $t$ of the other operand is one of numeric built-in types, then the inferred statement is $p \rightarrow t$.
-
-If the operator is one of \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, \texttt{!=} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{GEN\_EQUALS}, \\ \texttt{GEN\_NOT\_EQUALS}), one operand is of PathType $p$ and the type $t$ of the other operand is one of built-in types, then the inferred statement is $p \rightarrow t$.
+If the operator in an arithmetic operation is one of \texttt{+}, \texttt{-}, \texttt{div}, \texttt{mod}, \texttt{*}, \texttt{/} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{PLUS}, \texttt{MINUS}, \texttt{IDIV}, \texttt{MOD}, \texttt{MUL}, \texttt{DIV} (for all operators, see Attachment \ref{FIG_operators})), one operand is of PathType $P$ and the type $T$ of the other operand is one of numeric built-in types, then the inferred statement is $P \rightarrow T$.
+
+If the operator is one of \texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=}, \texttt{=}, \texttt{!=} (the class of the expression node is \texttt{Operator} and it represents one of \texttt{GEN\_LESS\_THAN}, \texttt{GEN\_GREATER\_THAN}, \texttt{GEN\_LESS\_THAN\_EQUALS}, \texttt{GEN\_GREATER\_THAN\_EQUALS}, \texttt{GEN\_EQUALS}, \\ \texttt{GEN\_NOT\_EQUALS}), one operand is of PathType $P$ and the type $T$ of the other operand is one of built-in types, then the inferred statement is $P \rightarrow T$.
 
@@
 
-This function has one formal argument of type $t =$ \texttt{xs:decimal}, \texttt{zero-or-one} cardinality.
-The real argument is of a PathType $p$.
+This function has one formal argument of type $T =$ \texttt{xs:decimal}, \texttt{zero-or-one} cardinality.
+The real argument is of a PathType $P$.
 In particular, it is a PathType representing path \texttt{\$i/reserve}, where the \texttt{\$i} variable refers to a PathType representing for-bound \texttt{/site/open\_auctions/open\_auction} path, and one special function call of \texttt{zero-or-one} function is noted.
 
-Since the criteria for the inference from a function call are met, statement $p \rightarrow t$ is inferred.
+Since the criteria for the inference from a function call are met, statement $P \rightarrow T$ is inferred.
 
@@
 
-\FORALL{$stepNode \in pathType.stepNodes$}
-	\IF{$stepNode.isAxisStep$}
-		\STATE $axisKind := stepNode.axisNode.getAxisKind()$
+\FORALL{$step \in pathType.steps$}
+	\IF{$step.isAxisStep$}
+		\STATE $axisKind := step.axisNode.getAxisKind()$
 		\IF{$axisKind \neq$ CHILD $\wedge$ $axisKind \neq$ DESCENDANT $\wedge$ $axisKind \neq$ DESCENDANT\_OR\_SELF}
@@
 
-\STATE $detailNode := stepNode.getChild(ExprNode)$
+\STATE $detailNode := step.getChild(ExprNode)$
 \IF{$detailNode \neq$ null}
 	\IF{$is(detailNode,$ VarRefNode$)$}
-		\IF{$usesOnlyChildAndDescendantAxes($ $pathType.forBoundSubsteps[stepNode]) =$ false}
+		\IF{$usesOnlyChildAndDescendantAxes(pathType.substeps[step]) =$ false}
 			\RETURN false
@@
 
-\FORALL{$stepNode \in pathType.stepNodes$}
-	\IF{$stepNode.hasPredicates$}
+\FORALL{$step \in pathType.steps$}
+	\IF{$step.hasPredicates$}
 		\RETURN false
 	\ENDIF
-	\IF{$pathType.forBoundSubsteps[stepNode] \neq$ null}
-		\IF{$isWithoutPredicates(pathType.forBoundSubsteps[stepNode] =$ false}
+	\IF{$pathType.substeps[step] \neq$ null}
+		\IF{$isWithoutPredicates(pathType.substeps[step] =$ false}
 			\RETURN false
@@
 
-\STATE $lastStepNode :=$ the last item from $pathType.stepNodes$
-\IF{$lastStepNode.hasPredicates() =$ false}
+\STATE $lastStep :=$ the last item from $pathType.steps$
+\IF{$lastStep.hasPredicates() =$ false}
 	\RETURN false
 \ENDIF
-\IF{number of items in list $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
+\IF{number of items in list $lastStep.getChild(PredicateListNode)$ $.getChildren()$ is higher than 1}
 	\RETURN false
@@
 
-\RETURN the only item in $lastStepNode.getChild(PredicateListNode)$ $.getChildren()$
+\RETURN the only item in $lastStep.getChild(PredicateListNode)$ $.getChild(0)$
 \end{algorithmic}
@@
 
-\STATE $firstStep :=$ first item from  $pathType.stepNodes$
+\STATE $firstStep :=$ first item from  $pathType.steps$
 \STATE $detailNode := firstStep.detailNode$
@@
 Additional join patterns are introduced in Listings \ref{listing_other_for_join_pattern}, \ref{listing_other_let_join_pattern} and \ref{listing_join_pattern_3}.
-In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child and descendant axes.
+In all join patterns $P_1$, $P_2$, $L_1$, $L_2$ are XQuery paths without predicates, using only child, descendant, and descendant or self ($L_1$, $L_2$ also attribute) axes.
 Actually, the join patterns from Listings \ref{listing_other_for_join_pattern} and \ref{listing_other_let_join_pattern} are covered by the for and let join pattern in the original approach, but we introduce them explicitly, because their structure in the syntax tree is different.
 
-Since in this step we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of PathType without predicates, using only child and descendant axes.
+Since in this step we already know the types of all expressions in the syntax tree, the requirement that $P_1$, $P_2$, $L_1$, $L_2$ are paths of the described form can be replaced by a more general requirement that $P_1$, $P_2$, $L_1$, $L_2$ are expressions of PathType satisfying the same requirements.
 
@@
 \FORALL{$returnPathType \in returnPathTypes$}
-    \IF{$returnPathType.specialFunctionCalls$ contains $count$}
+    \IF{$returnPathType.specialFunctionCalls$ contains 'count'}
         \STATE mark $jpOccurrence$ as case (O1), weight 0.75 \COMMENT {R3}
@@
 
-Contrary to the key discovery, PathTypes utilized in this step may contain predicates in their lasts steps.
-Predicates in a path's last step restrict its result, and, therefore, if a set of elements or attributes selected by a path with predicates is not unique, neither is the set of elements or attributes selected by the same path stripped of the predicates.
-
 \begin{algorithm}
@@
 
-Besides the aggregation functions, a certain form of FLWOR expressions is utilized in this step.
-They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference to a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
+Besides the \texttt{distinct-values} function, a certain form of FLWOR expressions is utilized in this step.
+They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child, descendant, and descendant or self axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference to a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.
 
@@
 
-Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_test_query_8} (subtree of ConstructorNode is omitted because of lack of space).
+Figure \ref{FIG_syntax_tree_2} shows the syntax tree of the query in Listing \ref{listing_test_query_8} (subtree of \texttt{Construc\-torNode} is omitted because of lack of space).
 In this tree, the algorithm founds one occurrence of the join pattern 3.
@@
 \begin{itemize}
-\item Key $k = ($\texttt{/site/people/person}$, \{$\texttt{@id}$\})$ is satisfied.
-\item Foreign key $(($\texttt{/site/closed\_auctions/closed\_auction}$, \{$\texttt{buyer/@person}$\}) \rightarrow k)$ is satisfied.
+\item Key $K = ($\texttt{/site/people/person}$, \{$\texttt{@id}$\})$ is satisfied.
+\item Foreign key $(($\texttt{/site/closed\_auctions/closed\_auction}$, \{$\texttt{buyer/@person}$\}) \rightarrow K)$ is satisfied.
 \item Key $($\texttt{/site/closed\_auctions/closed\_auction}$, \{$\texttt{buyer/@person}$\})$ is not satisfied.

### 2012-04-10 doc/theses/mikula/conclusion.tex
Master thesis: Final corrections.

@@
 \item Incorporation of the lexical and syntax analyses of input queries creating their syntax trees.
-\item Development of the static type analysis including PathTypes, which can be easily extended and use in possible future extensions.
+\item Development of the static type analysis including PathTypes, which can be easily extended and used in possible future extensions.
 \item Extension of the existing method of key discovery, achieving better results.
@@
 \renewcommand{\labelitemi}{$-$}
-\item In the proposed algorithm we dealt with a small part of possible refinements discussed in the overview.
-\item The combination with existing methods of inference is done in the simplest way, not modifying the grammar rules.
-But, from the view of recent XML schema inference research, the modification of the grammar rules according to the statements inferred from XML operations is vastly interesting.
-\item Since large real-world sets of XQuery queries are not available, we did not perform experiments using such sets.
-We only tested our method using a relatively small number input of queries, decreasing the level of reliability of the experimental results.
+\item In the proposed algorithm we dealt with only a small part of possible refinements discussed in the overview.
+The reason for this is that inclusion of more of them would exceed scope of one thesis.
+\item The combination with existing methods of inference is done in the simplest possible way - by not modifying the grammar rules.
+However, in the view of recent XML schema inference research, the modification of the grammar rules according to the statements inferred from XML operations is of considerable interest.
+The reason for the simplest combination is related with the previous point and it is that we did not deal with refinements affecting the grammar rules.
+\item Since large real-world sets of XQuery queries are not available, we did not perform tests using such sets.
+We only tested our method using a relatively small number of input queries, decreasing the level of reliability of the experimental results.
 \end{itemize}

### 2012-04-12 doc/theses/mikula/appendix_c.tex
Master thesis: Adding forgotten appendix.

\chapter{Content of CD}
The CD attached to this thesis has the following structure.

\begin{itemize}
\item \texttt{content.txt} - A file with this text.
\item \texttt{text/} - A PDF version of the thesis.
\item \texttt{src/} - Source codes of the jInfer framework including implementation of our method. The same source codes can be also obtained from public Subversion repository by issuing command:\\
\texttt{svn co -r 2155 \\ https://jinfer.svn.sourceforge.net/svnroot/jinfer/jinfer/trunk/}
\item \texttt{bin/} - jInfer plugins for NetBeans 7.0.1. See jInfer Tutorial \cite{jinfer_tutorial} for step-by-step instructions, but use NetBeans 7.0.1. Since the tutorial is for the last official jInfer release 1.0, it says the required version of NetBeans is at least 6.9. But this is not true in our case, because we use the development version of jInfer and it requires NetBeans 7.0.1.
\item \texttt{testing/} - A directory containing sets of test data and test results. Again, see the Jinfer Tutorial \cite{jinfer_tutorial} for instructions how to run the inference. 
\end{itemize}
